### 2019-10-09 README.Rmd
Initial commit of cleaned up regional effects repo

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# The Regional Effects of Marine Protected Areas

This repo contains code and data needed to reproduce "The Regional Effects of Marine Protected Areas" by Ovando et al. In Prep.

Users will need to install the simulation model that can be found [here](https://github.com/DanOvando/spasm), and installed using `remotes::install_github("danovando/spasm")`

>> Marine Protected Areas (MPAs) cover between 3-7% of the world's oceans, up from less than 1% in the year 2000. The Convention on Biological Diversity calls for 10% of coastal waters to be protected inside MPAs by 2020, while the International Union for Conservation of Nature calls for 30% protection by 2030. It is often clear that MPAs produce conservation benefits inside their borders, but many MPAs are also justified on the grounds that they also benefit the broader region outside their borders. Despite the expanding role of MPAs in marine resource management, we lack a clear understanding of what the regional effects of MPAs might be, and how we might detect them in scientific studies. We used a bio-economic simulation model to demonstrate how the regional effects of MPAs are affected by a suite of economic and environmental drivers and show how commonly used methods to measure inside-MPA effects are likely to fail to capture the regional impacts of protection. We present an empirical strategy for estimating regional MPA effects on biomass density using 16 years of data from inside and outside of MPAs within the Channel Islands, California, USA. We find no statistically clear effect of MPAs on aggregate mean biomass densities of targeted finfish throughout the study region. To make meaningful progress on MPA design and evaluation, the next generation of MPA science must address the challenge of understanding and estimating the regional effects of MPAs across the diversity of species in the ecosystem.

### 2019-10-09 documents/PNAS-template-main.tex
Initial commit of cleaned up regional effects repo

\documentclass[9pt,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnassupportinginfo}
% \readytosubmit %% Uncomment this line before submitting, so that the instruction page is removed.

\title{Your Main Manuscript Title}
\author{Author1, Author2 and Author3 (Complete author list)}
\correspondingauthor{Corresponding Author Name.\\E-mail: author.two@email.com}

\begin{document}

%% Comment/remove this line before generating final copy for submission
%%\instructionspage

\maketitle

%% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
\SItext

\subsection*{Subhead}
Type or paste text here. This should be additional explanatory text such as an extended technical description of results, full details of mathematical models, extended lists of acknowledgments, etc.

\section*{Heading}
\subsection*{Subhead}
Type or paste text here. You may break this section up into subheads as needed (e.g., one section on ``Materials'' and one on ``Methods'').

\subsection*{Materials}
Add a Materials subsection if you need to.

\subsection*{Methods}
Add a Methods subsection if you need to.

%%% Each figure should be on its own page
\begin{figure}
\centering
\includegraphics[width=\textwidth]{example-image}
\caption{First figure}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{frog}
\caption{Second figure}
\end{figure}

\begin{table}\centering
\caption{This is a table}

\begin{tabular}{lrrr}
Species & CBS & CV & G3 \\
\midrule
1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
\bottomrule
\end{tabular}
\end{table}

%%% Add this line AFTER all your figures and tables
\FloatBarrier

\movie{Type caption for the movie here.}

\movie{Type caption for the other movie here. Adding longer text to show what happens, to decide on alignment and/or indentations.}

\movie{A third movie, just for kicks.}

\dataset{dataset_one.txt}{Type or paste caption here.}

\dataset{dataset_two.txt}{Type or paste caption here. Adding longer text to show what happens, to decide on alignment and/or indentations for multi-line or paragraph captions.}

\bibliography{pnas-sample}

\end{document}

### 2019-10-09 documents/ovando-regional-effects-of-mpas-si.Rmd
Initial commit of cleaned up regional effects repo

---
title: "Supporting Information: The Regional Effects of Marine Protected Areas"
author:  Daniel Ovando,Jennifer E. Caselle,Christopher Costello,Olivier Deschenes,Steven D. Gaines,Ray Hilborn,Owen Liu
output: 
  bookdown::pdf_document2:
    keep_tex: true
  pdf_document: default
bibliography: My Library.bib
header-includes:
    - \usepackage{longtable}
    - \usepackage{booktabs}
    - \usepackage{eso-pic}
    - \usepackage{graphicx}
    - \usepackage[left]{lineno}
    - \linenumbers
params:
  run_name: ["v5.0"]
  sim_years: [50]
  burn_years: [20]
  num_patches: [25]
always_allow_html: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE)

library(sf)
library(ggmap)
library(viridis)
library(hrbrthemes)
library(scales)
library(patchwork)
library(rpart)
library(extrafont)
library(caret)
library(ggsci)
library(rEDM)
library(tidyverse)
library(Synth)
library(spasm)
extrafont::loadfonts()
functions <- list.files(here::here("functions"))

walk(functions, ~ here::here("functions", .x) %>% source()) # load local functions
save_plots <- TRUE

# sim_years <- 50
# 
# burn_years <- 20
# 
# num_patches <- 25
# 
# run_name <- "v5.0"

run_name <- params$run_name

sim_years <- params$sim_years

burn_years <- params$burn_years

num_patches <- params$num_patches

run_dir <- here::here("results", run_name)

experiment_dir <- here::here("results",run_name, "experiments")

load(file = here::here("results",run_name, "rawish_zissou_data.Rdata"))

load(file = here::here("results",run_name, "model_runs.Rdata"))

load(file = here::here("results",run_name,"processed_grid.Rdata"))

load(file.path(run_dir,"abundance_data.Rdata"))

density_ratios <- readRDS(file.path(run_dir,"density_ratios.rds"))

outcomes <- readRDS(file.path(run_dir,"outcomes.rds"))

   eqo <- outcomes %>%
     filter(year == max(year))

facet_labels <- c(
  mpa_size = "Range in MPA",
  depletion = "Without-MPA Depletion"
)

channel_islands <- readRDS(here::here("data","channel_islands_map.RDS"))

ca_mpas <- sf::st_read(here::here("data","MPA_CA_Existing_160301")) %>%
  rmapshaper::ms_simplify() %>%
  sf::st_transform(crs = 4326)

zissou_theme <-
  theme_ipsum(
    base_size = 12,
    axis_title_size = 12,
    strip_text_size = 12,
    base_family = "Helvetica"
  )

zissou_theme <- zissou_theme +
  theme(   panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank())

testplot <- ggplot()

panel_height = unit(1,"npc") - sum(ggplotGrob(testplot)[["heights"]][-3]) - unit(4,"line")

gc <- guide_colorbar(frame.colour = "black",
                     ticks.colour = "white",
                     barheight = panel_height)

  hgc <- guide_colorbar(frame.colour = "black",
                       ticks.colour = "black",
                       barwidth = 10)

    wide_hgc <- guide_colorbar(frame.colour = "black",
                       ticks.colour = "black",
                       barwidth = 25)

theme_set(zissou_theme)

models_worked <- model_runs$tmb_fit %>% map("error") %>% map_lgl(is_null)

model_runs <- model_runs %>%
  filter(models_worked) %>%
  mutate(tmb_fit = map(tmb_fit,"result")) %>%
  mutate(processed_fits = map(tmb_fit, process_fits)) %>%
  mutate(did_plot = map(processed_fits, "did_plot"))

base_run <- model_runs %>%
filter(var_names == "pisco_a", mpa_only == FALSE, center_scale == TRUE)

mpa_run <- model_runs %>%
  filter(data_source == "pisco",
         var_names == "pisco_a",
         mpa_only == TRUE)

kfm_run <- model_runs %>%
  filter(data_source == "kfm")

fitted_data <- base_run$data[[1]]

zissou_fit <- base_run$tmb_fit[[1]]

report <- base_run$tmb_fit[[1]]

site_data <- read_csv(here::here("data",'Final_Site_Table_UCSB.csv')) %>%
  magrittr::set_colnames(., tolower(colnames(.))) %>%
  select(
    site,
    side,
    mpagroup,
    mpa_status,
    reserve,
    region,
    year_mpa,
    mpaareanm2,
    lat_wgs84,
    lon_wgs84
  ) %>%
  rename(lat_wgs84 = lon_wgs84,
         lon_wgs84 = lat_wgs84) %>%
  unique() %>%
  mutate(eventual_mpa = (year_mpa > 0))

zissou_fit$zissou_estimates <- zissou_fit$zissou_estimates %>%
  mutate(variable = str_replace_all(variable, "(\\.).*$",""))

report$zissou_estimates <- report$zissou_estimates %>% 
    mutate(variable = str_replace_all(variable, "(\\.).*$",""))

seen_non_nested_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seen_non_nested_betas")) %>%
  rename(group = variable) %>%
  mutate(variable  = zissou_fit$zissou_data$x_seen_non_nested %>% colnames())

seeing_non_nested_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seeing_non_nested_betas")) %>%
  rename(group = variable) %>%
  mutate(variable  = zissou_fit$zissou_data$x_seen_non_nested %>% colnames())

seen_year_species_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seen_year_species_betas")) %>%
  rename(group = variable) %>%
  mutate(variable = zissou_fit$zissou_data$x_seen_year_species %>% colnames())

seeing_year_species_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seeing_year_species_betas")) %>%
  # filter(variable == "seeing_year_species_betas") %>%
  rename(group = variable) %>%
  mutate(variable = zissou_fit$zissou_data$x_seen_year_species %>% colnames())

seen_region_cluster_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seen_region_cluster_betas")) %>%
  # filter(variable == "seen_region_cluster_betas") %>%
  rename(group = variable) %>%
  mutate(variable = zissou_fit$zissou_data$x_seen_region_cluster %>% colnames())

seeing_region_cluster_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seeing_region_cluster_betas")) %>%
  # filter(variable == "seeing_region_cluster_betas") %>%
  rename(group = variable) %>%
  mutate(variable = zissou_fit$zissou_data$x_seen_region_cluster %>% colnames())

did_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "mpa_effect")) %>%
  # filter(variable == "mpa_effect") %>%
  mutate(group = variable) %>%
  mutate(year = zissou_fit$did_data$year %>% unique())

 betas <- bind_rows(
   seen_non_nested_betas,
   seeing_non_nested_betas,
   seen_year_species_betas,
   seeing_year_species_betas,
   seen_region_cluster_betas,
   seeing_region_cluster_betas,
   did_betas %>% select(-year)
 ) %>%
   as_data_frame()

 non_nested_beta_plot <- betas %>%
   filter(str_detect(group, "non_nested")) %>%
   ggplot() +
   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
   geom_pointrange(aes(x = variable,
                       y = estimate,
                       ymin = lower,
                       ymax = upper)) +
   facet_wrap(~group) +
   coord_flip() +
   theme(axis.text.y = element_text(size = 10))

 year_species_effects_plots <- betas %>%
   filter(str_detect(group, "year_species_betas")) %>%
   mutate(year = str_replace_all(variable,"\\D","") %>% as.numeric()) %>%
   mutate(classcode = str_split(variable,'-', simplify = T)[,2]) %>%
   ggplot() +
   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
   geom_ribbon(aes(x = year, ymin = lower, ymax = upper, fill = classcode), alpha = 0.25) +
   geom_line(aes(x = year, y = estimate, color = classcode)) +
   facet_wrap(~group)

  region_cluster_plots <- betas %>%
   filter(str_detect(group, "region_cluster_betas")) %>%
   mutate(cluster = str_replace_all(variable,"\\D","")) %>%
   mutate(region = str_split(variable,'-', simplify = T)[,3]) %>%
   ggplot() +
    geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
    geom_pointrange(aes(x = region,
                        y = estimate,
                        ymin = lower,
                        ymax = upper, color = cluster)) +
    facet_wrap(~group)

  sim_plot <- outcomes %>%
    mutate(rsize = plyr::round_any(mpa_size, 0.25),
           reff = plyr::round_any(mpa_effect,0.1)) %>%
    filter(rsize == 0.25) %>%
    mutate(year = years_protected + 2001) %>%
    filter(year %in% did_betas$year) %>% 
    group_by(year, reff) %>% 
    count() %>% 
    group_by(year) %>% 
    mutate(sc = n / sum(n))

  did_plot <- did_betas %>%
    ggplot() +
    geom_tile(data = sim_plot, aes(year, reff, fill = sc), alpha = 0.75) +
    geom_vline(aes(xintercept = 2003), color = 'red', linetype = 2, size = 2) +
    geom_hline(aes(yintercept = 0)) +
    geom_pointrange(aes(
      year,
      y = exp(estimate) - 1,
      ymin = exp(lower) - 1,
      ymax = exp(upper) - 1
    ),
    size = 1.5
    ) +
    labs(x = "Year", y = "Targeted Trend Divergence") +
    scale_fill_viridis(guide = hgc,
                       name = "% of Annual Sims",
                       labels = scales::percent_format(accuracy = 1),
                       option = "plasma") + 
    theme(legend.position = "top")

top_species <- base_run$data[[1]]$classcode %>% unique()

cip_data <- pisco_data %>%
  left_join(site_data, by = c("site","side")) %>%
  filter(is.na(eventual_mpa) == F) %>%
  filter(region %in% c("ANA", "SCI","SRI",'SMI'),
         classcode %in% top_species)

# process outcomes

filter_summary <- fitted_data %>% 
  select(classcode, targeted) %>% 
  unique() %>% 
  group_by(targeted) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(targeted = ifelse(targeted == 1,"yes","no")) %>% 
  spread(targeted, n)

```

# Supporting Information (SI)

# SI Text {#si-text .unnumbered}

## Computing environment

The computing environment needed to reproduce the results presented here is stored in a Docker container available [here]()XX. Interested users can load the container on their system to reproduce the complete analysis and manuscript. Analysis were performed in `r R.version.string`. Package versions are shown in Table.S\@ref(tab:pkg-info).

```{r pkg-info, results = "asis"}
table <- devtools::package_info() %>% 
  select(package, loadedversion, date,source)

knitr::kable(table,
             digits = 2,
             caption = "Package versions and sources used in this paper",
             longtable = TRUE,
             booktabs = TRUE,
             format = "latex",
             col.names = colnames(table)) %>% 
  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))

```

## Operating Model

The operating model is a spatial single-species age-structured bio-economic model. The operating model itself is organized as an R package, which can be found and installed [here](https://github.com/DanOvando/spasm). Users can explore the functionality of the operating mode through an interactive web application [here](https://danovando.shinyapps.io/simmpa/).

For the population model, numbers *n* at time *t* for age *a* are given by

\begin{equation}
n_{t,a}=\begin{cases}
      = BH(ssb_{t-1}) & \text{if $a = 1$}\\
     = n_{t-1,a-1}e^{-(m + qE_{t-1}{\times}s_{a-1})}, & \text{if $1< a < max(age)$}\\
     =  n_{t-1,a}e^{-(m + qE_{t-1}{\times}s_a)} + n_{t-1,a-1}e^{-(m + qE_{t-1}{\times}s_{a-1})}, & \text{if $a = max(a)$}
  \end{cases}
  (\#eq:pop)
\end{equation}

where *BH* is the Beverton-Holt recruitment function, *ssb* is spawning-stock-biomass, *m* is natural mortality, *q* is catchability,*E* is fishing effort at time *t*, and *s* is selectivity at age *a*.

Selectivity is modeled through a logistic form per

\begin{equation}
s_a=\frac{1}{(1 + e^{-log(19)\times\frac{l_a - l_{sel}}{\delta_{sel}}})}
  (\#eq:sel)
\end{equation}

where $l_a$ is the mean length at at age, $l_sel$ is the length at which on average 50\% of individuals are selected by the fishery, and $\delta_{sel}$ are the additional units of length at which on average 95\% of fish are selected by the fishery.

*ssb* is calculated by converting age to mean length, calculating weight at age, maturity at age, and then calculating spawning stock biomass as the sum of spawning potential at age in a given time step.

\begin{equation}
  l_{a} = l_{\infty}\left(1 - e^{-k(a - a_0)}\right)
  (\#eq:len)
\end{equation}

Weight at age is then given by

\begin{equation}
  b_{a} = w_a \times l_{a}^{w_b}
  (\#eq:weight)
\end{equation}

and maturity *mat* is calculated as

\begin{equation}
  \frac{1}{(1 + e^{-log(19)\times\frac{l_a - l_{mat}}{\delta_{mat}}})}
  (\#eq:mat)
\end{equation}

where $l_{mat}$ is the length at which on average 50\% of individuals are sexual maturity, and $\delta_{mat}$ is the units of length beyond $l_{mat}$ at which on average 95\% of fish are sexually mature.

Spawning stock biomass at time *t* is then calculated as

\begin{equation}
  ssb_t = \sum_{a=1}^Aw_{a,t}mat_{a,t}n_{a,t}
    (\#eq:ssb)
\end{equation}

### Recruitment

Recruitment follows Beverton-Holt dynamics. We do however allow for three variants in the timing of density dependence:

  1. Local density dependence: Density dependence occurs independently in each patch, and recruits then disperse to nearby patches

\begin{equation}
  n_{t,a = 1,p} = \left(\frac{0.8{\times}r0_{p}\times{h}\times{ssb_{t-1,p}}}{0.2\times{ssb0_p}\times(1 - h)+(h - 0.2)\times{ssb_{t-1,p}}}\right)\times \boldsymbol{d^l} \times \epsilon_t
  (\#eq:dd1)
\end{equation}

where **d^l^** is the larval movement matrix, *h* is Beverton-Holt steepness (constrained between 0.6 and 0.99), *r0* is unfished recruitment, and *ssb0* is unfished spawning stock biomass.

  2. Global density dependence: Density dependence is a function of the sum of spawning biomass across all patches, and recruits are then distributed according to habitat quality

\begin{equation}
n_{t,a = 1,p} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p}}\times{h}\times\sum_{p=1}^P{ssb_{t-1,p}}}{0.2\times{\sum_{p=1}^Pssb0_p}\times(1 - h)+(h - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p}}}\right)\times {hab_p} \times \epsilon_t
  (\#eq:dd2)
\end{equation}

where *hab* is a vector of habitat quality by patch that sums to 1.

3. Post-dispersal density dependence: Larvae are distributed throughout the system, and then density dependence occurs based on the density of adult biomass at the destination patch.

\begin{equation}
  larv_{t,p} = ssb_{t-1} \times\boldsymbol{d^l}
  (\#eq:larvmove)
\end{equation}

\begin{equation}
n_{t,a = 1,p} = \left(\frac{0.8{\times}r0_{p}\times{h}\times{larv_{t,p}}}{0.2\times{ssb0_p}\times(1 - h)+(h - 0.2)\times{larv_{t,p}}}\right) \times \epsilon_t
  (\#eq:dd3)
\end{equation}

$\epsilon$ represents multiplicative recruitment deviates. Deviates are calculated as

$$\epsilon_t = e^{recdev_t}$$

`$recdev$ are the log-normal recruitment deviates in time *t*.

The stochastic component of the deviate is

$$\gamma_t \sim norm(-\sigma_r^2/2,\sigma_r)$$

and the final multiplicative recruitment deviate in time *t* is then

$$recdev_t = \gamma_t\sqrt{1 - ac_r^2} + recdev_{t-1}ac_r $$

where *ac* is the autocorrelation of the recruitment function (between 0 and 1).

### Dispersal

Dispersal in the model is broken into two components: adult and larval. Both assume a Gaussian dispersal kernel of the form

\begin{equation}
m_{s,p_i,p_j}=\frac{1}{\sqrt{2\pi\sigma_{s,p_i}^2}}e^{-\frac{d_{p_i,p_j}^2}{2\sigma_{s,p_i}^2}}
  (\#eq:move)
\end{equation}

where *i* is the source patch, *j* is the destination patch, *d* is the distance between patches *i* and *j* (where distance is measured with wrapped edges, such that if there are 50 patches, patch 1 and patch 50 have a distance of 1), and $\sigma_s*$is the movement rate, in units of patches, for life stage *s* (adult or larval).

We allow the adult dispersal matrix to be affected by adult density dependence. The idea behind this is that adult fish will move more as densities increase, and become more sedentary as densities decrease (as habitat and food become more available for example). This allows us to simulate a scenario where as MPAs build up density they begin to export more adults to the surrounding waters, and if densities are lower in the fished areas these fish will actually become more sedentary.

Under these conditions, the adult movement rate is a linear function of depletion (measured as $ssb/ssb_0$)

\begin{equation}
\sigma_{s=a,p}^* = max(slope\times{d _p+ \sigma_{s=a} \times dmod},0)
  (\#eq:ddmove1)
\end{equation}

where

\begin{equation}
slope = \sigma_{s=a} - ( \sigma_{s=a} \times dmod)
  (\#eq:ddmove2)
\end{equation}

Under these conditions, when depletion *d = 1* (meaning the stock is unfished) the adult movement rate equals the max adult movement rate ($\sigma_{s=a}^* = \sigma_{s=a}$). When $d=0$ $\sigma_{s=a}^* = \sigma_{s=a} * dmod$. The greater *dmod* is then, the more movement rates from a patch decline as density declines.

We also allow for a "sprinkler" condition in which MPAs are placed in locations that disperse larvae to a much wider area than non-MPA locations. In this world, we simply multiply $\sigma_{s=l,p}$ by a sprinkler factor (by default 4) for any patch *p* that would eventually become an MPA (whether or not MPAs are ever introduced). In other words, when we compare two scenarios, one with MPA and one without, the "without" scenario still has higher larval movement rates in patches that become MPAs in the "with" scenario.

### Fleet Dynamics

We allow for three fleet models: constant effort, constant catch, and open-access. Constant effort means that total effort across all patches is equal in all time steps (unless MPAs force exit of effort as discussed below). Under constant catch, we set a target catch volume (in biomass, summed across all patches). Each time step, we calculate the fishing mortality rate that, given the fishable biomass in that time step, would produce the target catch. If there is insufficient fishable biomass available to support the target catch, we mark the population as crashed and stop the simulation (these crashed simulations are not included in the final analysis).

Under open-access, fishing effort expands in proportion to a weighted mean of profit-per-unit effort over the last *t* time steps.

\begin{equation}
profit_{t} = price\times{catch_t} - cost\times{E_t}^2
  (\#eq:profit)
\end{equation}

From there, we determine the new effort as

\begin{equation}
E_t = E_{t-1} + \theta\times\sum_{i=t-1-l}^{t-1}w_i\frac{profit_{i}}{E^i}
  (\#eq:openaccess)
\end{equation}

where *w* is a weighting function which is just a linear function of time

\begin{equation}
  w_i = \frac{i}{\sum_{i=1}^li}
  (\#eq:oaweight)
\end{equation}

and *l* is the number of lagged time steps over which to calculate the weighted mean PPUE.

The open-access model can enter chaotic dynamics if the model parameters are not properly tuned. To address this, we first set price at 1, and set a $\theta$ such that when profits are about as large as they might conceivably be the fishery doubles in size. We then estimate reference points for the simulated fishery (Bmsy, Fmsy, MSY), and set a target bionomic equilibrium B/Bmsy. Holding the other parameters constant, we thing find a cost coefficient that produces the desired bionomic equilibrium.

### Spatial Fleet Distribution

Given a total amount of effort, we then need to distribute that effort in space. In the simplest form, effort is evenly distributed throughout the available patches.

\begin{equation}
E_{t,p} = E_t \times \frac{open_p}{\sum_{p=1}^Popen_p}
  (\#eq:simpleeffort)
\end{equation}

where *open* indicates whether patch *p* is open to fishing or not.

Effort can also be distributed according to spawning stock biomass in fishable patches

\begin{equation}
E_{t,p} = E_t \times \frac{open_{t,p}ssb_{t,p}}{\sum_{p=1}^Popen_{t,p}ssb_{t,p}}
  (\#eq:biomasseffort)
\end{equation}

And lastly effort can be distributed according to profit-per-unit-effort

\begin{equation}
E_{t,p} = E_t \times \frac{open_{t,p}ppue_{t,p}}{\sum_{p=1}^Popen_{t,p}ppue_{t,p}}
  (\#eq:ppueeffort)
\end{equation}

Under the constant effort or open access scenarios, effort can immediately respond to MPA placement in one of two ways. Effort can concentrate outside the MPAs (such that the sum of effort before and after MPA placement stays constant), or effort can leave the MPAs, such that the total effort in the fishery is reduced by the amount of effort that occurred inside the MPAs immediately before MPA placement. This is intended to simulate a scenario where fishers that used to use the MPA simply leave the fishery rather than redistribute outside the MPA, due for example to costs or lack of location specific knowledge to fish outside the MPA.

### MPA Design

MPA design is relatively straightforward. We set a percentage of patches that are to be placed inside no-take MPAs. MPAs can either be placed continuously (e.g if there are 100 patches and 25% are in MPAs, patches 1 to 25 are in MPAs) or randomly. If the MPAs are placed randomly, we can also set a minimum MPA size. This controls the patchiness of the MPAs. As the "patchiness" factor approaches zero, the behavior equals that of random placement. As it approaches 1, the behavior approaches that of continuous placement. In between, the greater the patchiness, the more clustered together MPAs become.

## Simulations

We use this our operating model to simulate 10,000 different fisheries, where each fishery is a random combination of variables, described below

Table.S1 - Range of simulated variables

| Variable | Distribution|
|---------:|:-----------:|
| Scientific Name | Drawn from all possible species in `FishLife` (@thorson2017c)|
|steepness (h) | ~uniform(0.6,0.95)|
|Adult movement ($\sigma_{s=a}$) | ~uniform(0,0.5 * P)|
|Larval movement ($\sigma_{s=l}$) | ~uniform(0,0.5 * P)|
|Recruitment variation ($\sigma_{r}$) |  $\in\{0,0.05,.1\}$|
|Recruitment autocorrelation ($ac_{r}$) |  $\in\{0,0.05,.1\}$|
|DD adult movement (dmod) | $\in\{0.25,1\}$|
|Density-dependence timing | $\in\{local, global,post-dispersal\}$|
|\% Patches in MPA | ~uniform(0.01,1)|
|Initial fishing relative to natural mortality | ~uniform(0.01,4)|
|Selectivity as a multiple of maturity length| ~uniform(0.1,1.25)|
|Fleet model| $\in\{open-access, constant-effort,constant-catch\}$|
|Spatial effort model| $\in\{uniform, biomass,profits\}$|
|Years into simulation to start MPA| ~round(uniform(5,0.66 * T))|
|MPA is sprinkler?| $\in\{TRUE,FALSE\}$|
|Randomly place MPA?| $\in\{TRUE,FALSE\}$|
|Fleet reaction to MPA| $\in\{concentrate, abandon-ship\}$|
|Patchiness | ~uniform(0.01,0.75)|
|MPA habitat factor| $\in\{1, 4\}$|

One thing to note here is the random sampling of species' scientific names. The effect of MPAs, especially over the short term, will clearly be affected by factors such as the growth rate, the mortality rate, and the maturity schedule. These life history traits are related through a variety of biological processes, as such randomly sampling these parameters can lead to biologically nonsensical "frankenfish". We resolve this by using the `FishLife` package (@thorson2017c) instead. `FishLife` builds off of [FishBase](https://fishbase.org/), and provides estimate of key life history traits taking into account the relationships across these variables. For simulations then, we randomly pull a species from `FishLife`, and then pull the available life history information from that species for use in the operating model. This allows us to simulate a wide range of life history types in a realistic manner.

We ran 20,000 simulations from these distributions. Each simulation runs for 50 years in 50 patches (with a 25 year unfished burn-in period for conditions in which initial conditions cannot be solved analytically, for example when MPAs have better habitat than non-MPAs). For each simulation, we run one scenario without MPAs, though taking note of where the MPA would be as needed. For the second scenario, we hold everything constant except we now add in the MPAs as dictated by the particular simulation.

##### Filtering Simulations

After the 20,000 simulations have run, we perform a series of filtering steps to remove runs that either a) produced chaotic dynamics during the open-access scenario; b) did not converge to the correct bionomic equilibrium in the open-access scenario; or c) crashed the population before the MPAs went into place (population falls below 5% of unfished biomass). These filtering steps left us with `r nrow(processed_grid) ` viable simulations.

#### Additional Simulation Results

Simulation results are presented as percent differences in biomass densities with and without MPAs, in order to be comparable to the estimates that the regression model produces. However, this metric presents some problems as a measure of how "detectable" an effect size is. As depletion increases, relatively small changes in total biomass (relative to the variance in the observation process) can translate into large percent changes in biomass. For example, moving from a density of .2kg/m^2^ to .4kg/m^2^ translates to a 100% percent increase, but only a .2kg/m^2^ absolute increase, a small value to detect with a real observation program.

To illustrate these, we present an alternative to our simulation results in which changes in biomass caused by MPAs is scaled by the unfished biomass in the system.

```{r pop-conservation-effect, include = TRUE, fig.cap="Median (A) and range (B) of equilibrium regional MPA conservation effects (change in total biomass with MPAs relative to without MPAs as a percentage of unfished biomass) across a range of depletion and MPA sizes (and incorporating the full range of scenarios included in our study). 'Range in MPA' is the percent of patches covered by an MPA, 'Depletion' is the depletion that would have occurred in equilibrium without the MPA"}

pop_time_plot <- outcomes %>%
  filter(years_protected > 0) %>%
  ggplot(aes(years_protected, pop_effect)) +
  geom_bin2d() +
  scale_fill_viridis(guide = gc)

ylabs <- c(expression(""<= "-50%"), paste0(seq(-25,75, by = 25),"%"),expression("">= "100%") )

ybreaks = seq(-.50,1, by = .25)

pop_depletion_plot <- outcomes %>%
  filter(year == max(year)) %>%
  ggplot() +
  geom_bin2d(aes(depletion, pop_effect), binwidth = c(.1, .1),
             show.legend = FALSE) +
  scale_fill_viridis(
  option = "A",
  # trans = "log10",
  guide = guide_colorbar(frame.colour = "black",
  frame.linewidth = 1,),
  name = "Median Effect"
  ) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1), name = "Depletion") +
  scale_y_continuous(labels = ylabs, breaks = ybreaks, name = "")

pop_depletion_plot <- outcomes %>%
    filter(year == max(year)) %>%
    mutate(
      rdep = cut(
        depletion,
        breaks = seq(0, 1, by = .2),
        labels = seq(0, .8, by = .2),
        include.lowest = TRUE
      )
      ,
      reff = cut(
        pop_effect,
        breaks = seq(-.6, 1, by = .2),
        labels = seq(-.6, .8, by = .2),
        include.lowest = TRUE
      )
    ) %>%
    group_by(rdep, reff) %>%
    count() %>%
    group_by(rdep) %>%
    mutate(pn = n / sum(n)) %>%
    ungroup() %>%
    mutate(rdep = as.numeric(as.character(rdep)),
           reff = as.numeric(as.character(reff))) %>%
    ggplot(aes(rdep, reff, fill = pn)) +
    geom_tile(show.legend = FALSE) +
    scale_fill_viridis(
      option = "A",
      # trans = "log10",
      guide = hgc,
      name = "% of Sims",
      labels = scales::percent_format(accuracy = 1)
    ) +
    scale_x_continuous(labels = scales::percent_format(accuracy = 1),
                       name = "Depletion",
                       expand = expand_scale(add = c(0,0))) +
    scale_y_continuous(labels = ylabs,
                       breaks = ybreaks,
                       name = "")

  pop_size_plot <- outcomes %>%
    filter(year == max(year)) %>%
    mutate(
      rsize = cut(
        mpa_size,
        breaks = seq(0, 1, by = .2),
        labels = seq(0, .8, by = .2)
      )
      ,
      reff = cut(
        pop_effect,
        breaks = seq(-.6, 1, by = .2),
        labels = seq(-.6, .8, by = .2)
      )
    ) %>%
    group_by(rsize, reff) %>%
    count() %>%
    group_by(rsize) %>%
    mutate(pn = n / sum(n)) %>%
    ungroup() %>%
    mutate(rsize = as.numeric(as.character(rsize)),
           reff = as.numeric(as.character(reff))) %>%
    ggplot(aes(rsize, reff, fill = pn)) +
    geom_tile() +
    scale_fill_viridis(
      option = "A",
      # trans = "log10",
      guide = hgc,
      name = "% of Sims",
      labels = scales::percent_format(accuracy = 1)
    ) +
    scale_x_continuous(labels = scales::percent_format(accuracy = 1),
                       name = "Range in MPA",
                       expand = expand_scale(add = c(0,0))) +
    scale_y_continuous(labels = ylabs,
                       breaks = ybreaks,
                       name = "Conservation Effect") +
    theme(legend.position = "top") +
    labs(title = 'B)')

 pop_facet_effect_plot <- outcomes %>%
  filter(year == max(year)) %>%
   select(pop_effect, mpa_size, depletion) %>%
   gather(measure, value, -pop_effect) %>%
   ggplot() +
   geom_bin2d(aes(value, pop_effect), binwidth = c(.1, .1), show.legend = TRUE) +
  scale_fill_viridis(
  option = "A",
  # trans = "log10",
  guide = hgc,
  name = "# of Sims"
  ) +
   facet_wrap(~measure, labeller = labeller(measure = facet_labels), strip.position = "bottom") +
   scale_x_percent(name = "") +
   scale_y_percent(name = "Pop. Conservation Effect") + 
   theme(legend.position = "top")

  pop_depletion_and_size_plot <- outcomes %>%
    filter(years_protected >= 0) %>%
    group_by(experiment) %>%
    mutate(depletion = plyr::round_any(depletion[years_protected == 0], .1),
           mpa_size = plyr::round_any(mpa_size, .1)) %>%
    filter(year == max(year)) %>%
    group_by(depletion, mpa_size) %>%
    summarise(median_mpa_effect = median(pop_effect, na.rm = TRUE)) %>%
    ggplot(aes(depletion, mpa_size, fill = median_mpa_effect)) +
    geom_tile() +
    # geom_contour(aes(z = median_mpa_effect, color = ..level..), show.legend = FALSE) +
    scale_fill_viridis(labels = scales::percent_format(accuracy = 1),
                       guide = hgc,
                       name = "Median Pop. Effect") +
  labs(x = "Depletion", y = "Range in MPA", title = "A)") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                       expand = c(0,0)) +
    scale_x_continuous(labels = scales::percent_format(accuracy = 1), expand = c(0,0),
                       limits = c(0,1)) + 
    theme(legend.position = "top")

  # pop_depletion_and_size_plot <-  directlabels::direct.label(pop_depletion_and_size_plot, list("angled.boxes", box.alpha = 0.5))

  pop_combo_plot <-
    (pop_depletion_and_size_plot) + ((pop_size_plot) / pop_depletion_plot)  + plot_layout(widths = c(1.5, 1)) & theme(
    plot.margin = unit(c(0, 1, 0, 1), units = "lines"),
    axis.text.x = element_text(size = 8),
    legend.box.margin = unit(c(0, 0, 0, 0), units = "lines"),
    axis.text.y = element_text(size = 10)
    )

pop_combo_plot

```

<!-- ##### BACI -->

<!-- # ```{r baci-vs-mpa-effect-plot, include = FALSE, eval = FALSE} -->
<!-- #  -->
<!-- # baci_plot <- baci_outcomes %>% -->
<!-- #   filter(sprinkler == FALSE, mpa_habfactor == 1) %>% -->
<!-- #   # select(experiment, -->
<!-- #   #        years_protected, -->
<!-- #   #        mpa_effect, -->
<!-- #   #        adult_movement, -->
<!-- #   #        mpa_size, -->
<!-- #   #        contains("baci")) %>% -->
<!-- #   group_by(experiment) %>%  -->
<!-- #   mutate(b0 = unique(`no-mpa`[year == 1])) %>%  -->
<!-- #   ungroup() %>%  -->
<!-- #   mutate(mpa_effect = (`with-mpa` - `no-mpa`) / b0, -->
<!-- #          baci = baci / b0) %>%  -->
<!-- #   mutate(adult_movement = adult_movement / max(adult_movement)) %>% -->
<!-- #   group_by(experiment) %>%  -->
<!-- #   filter(years_protected == max(years_protected)) %>% -->
<!-- #   ggplot(aes(y = (mpa_effect), x = (baci / b0))) + -->
<!-- #     # geom_hline(aes(yintercept = 0)) +  -->
<!-- #   # geom_vline(aes(xintercept = 0)) +  -->
<!-- #   geom_smooth(method = "lm", se = FALSE) + -->
<!-- #   # geom_abline(aes(slope = 1, intercept = 0), linetype = 2, size = 1.5, color = "red") + -->
<!-- #   geom_point(aes(color = adult_movement, size = mpa_size), alpha = 0.5, show.legend = FALSE)  + -->
<!-- #   # labs(y = "True % Increase in Biomass", x = "% BACI") + -->
<!-- #   scale_color_viridis(guide = wide_hgc, labels = scales::percent_format(accuracy = 1), name = "Fish Movement", -->
<!-- #                       option = "cividis") +  -->
<!-- #   theme(legend.position = "top") -->
<!-- #  -->
<!-- # pdf(file =  "figs/baci.pdf",width = 8, height = 5, useDingbats = TRUE) -->
<!-- # print(baci_plot) -->
<!-- # dev.off() -->
<!-- #  -->
<!-- # ``` -->

<!-- At the extreme sizes of MPAs there is clearly a tension between conservation and fishery goals. However, most MPAs are relatively small (XX), and therefore likely to cover lower percentages of a population as simulated here. A key question then is how do tradeoffs   -->

```{r mpa-tradeoffs, eval = FALSE}

tradeoffs <- fishery_outcomes %>%
  select(experiment, year, years_protected,mpa_size, fishery_effect) %>%
  left_join(outcomes %>%
  select(experiment,year, mpa_effect), by = c("experiment","year")
)

vplot <- tibble(fishery_effect = -1:1, mpa_effect = -1:1) %>%
  mutate(slope = ifelse(fishery_effect < 0, -1, 1)) %>%
  mutate(vline = slope * fishery_effect)

tradeoff_kobe_plot = tradeoffs %>%
  group_by(experiment) %>%
  filter(years_protected == max(years_protected)) %>%
  ungroup() %>%
  mutate(mpa_bin = cut(mpa_size, breaks = seq(0, 1, by = 0.25))) %>%
  ggplot(aes(fishery_effect, mpa_effect)) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_vline(aes(xintercept = 0), linetype = 2) +
  geom_hex(bins = 5,    aes(fill = ..density..),color = "black") +
    geom_line(data = vplot,aes(fishery_effect, vline), color = "red",
              linetype = 2, size = 2) +

  # geom_abline(aes(slope = -1, intercept = 0), linetype = 2, color = "red", size = 2) +
  #   geom_abline(aes(slope = 1, intercept = 0), linetype = 2, color = "black", size = 2) +
  scale_fill_distiller(
    # option = "plasma",
    labels = scales::percent_format(accuracy = 1),
    name = "% of Sims",
    guide = hgc
  ) +
  # geom_point(size = 2, alpha = 0.2, aes(color = factor(mpa_bin))) +
  # scale_color_viridis_d() +
  facet_wrap( ~ mpa_bin) +
  theme_minimal() +
  theme(legend.position = "top") +
    scale_x_continuous(labels = scales::percent_format(accuracy = 1),
                     name = "% Change Fishery") +
      scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     name = "% Change Conservation") + 
  labs(caption = "facets are percent of range in MPA. Red dashed line shows points where fisheries losses/gains are equal to conservation gains")

# suppose that utility is the sum of percentage gain in conservation and fishery

tradeoff_bins <- tradeoffs %>%
  group_by(experiment) %>%
  filter(years_protected == max(years_protected)) %>%
  mutate(tradeoff = mpa_effect + fishery_effect) %>%
  mutate(tradeoff_bin = plyr::round_any(tradeoff,.25),
         mpa_bin = plyr::round_any(mpa_size,.1)) %>%
  group_by(tradeoff_bin, mpa_bin) %>%
  count() %>%
  group_by(mpa_bin) %>%
  mutate(n = n / sum(n)) %>%
  ungroup()

tradeoff_summary <- tradeoffs %>%
  group_by(experiment) %>%
  filter(years_protected == max(years_protected)) %>%
  mutate(tradeoff = mpa_effect + fishery_effect) %>%
  mutate(mpa_bin = plyr::round_any(mpa_size,.1)) %>%
  group_by(mpa_bin) %>%
  summarise(net_tradeoff = mean(tradeoff > 0))

tradeoff_density_plot <- tradeoff_bins %>%
  ggplot(aes(mpa_bin, tradeoff_bin, fill = n)) +
  # geom_hline(aes(yintercept = 0), linetype = 2, size = 2) +
  geom_tile(alpha = 0.9) +
  scale_fill_viridis(
    option = "plasma",
    labels = scales::percent_format(accuracy = 1),
    name = "% of Sims",
    guide = hgc
  ) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1),
                     name = "% of Range in MPA") +
  scale_y_continuous(name = "Net Utility") +
  theme(legend.position = "top")

tradeoff_summary_plot <- tradeoff_summary %>%
  ggplot(aes(mpa_bin, net_tradeoff)) +
  geom_col() +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1),
                     name = "% of Range in MPA") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     name = "% Positive Utility")

# tradeoff_density_plot + tradeoff_summary_plot & theme(
#     plot.margin = unit(c(0, 1, 0, 1), units = "lines"),
#     axis.text.x = element_text(size = 8),
#     legend.box.margin = unit(c(0, 0, 0, 0), units = "lines"),
#     axis.text.y = element_text(size = 10)
#     )

```

<!-- why is it going to be hard ot detect regional effects of MPAs few if any empirical studies at net regional effects. does this mean that i doesn't exist, or impossible tot est, why? three fundamental stumbling blocks -->

## Estimation Model

The goal of the estimation model is to estimate the difference in the mean densities of targeted and non-targeted species pre-and-post MPA implementation.

### Data

All observation data used in the estimation model were collected by PISCO. PISCO staff also compiled allometric information used to convert lengths to expected weights, and hence biomass densities. We do not account for error in this translation step Table.S\@ref(tab:classcode) lists the species included in the model. In order to be included in the estimation model, a species must have been observed at least twice a year every year for at least 14 years. We also omitted all observations of "young of the year" fish due to inability to identify these observations to the nearest species level in many cases. We omitted data from 1999 due to changes in the sampling procedures that occurred after 1999. Per recommendations from PISCO staff we omit observations from the canopy level of the transects (leaving the middle, bottom, and middle canopy levels).

PISCO data report positive observations of fish. In order to use these data in our model we need to add in zeros for any transect that could have observed a given species of fish but did not. We assume that a fish could have been observed on a given transect if that species has ever been observed at that site in any time period in the data (PISCO data are organized by sites, with multiple transect at different locations within the borders of a site). If a species has never been observed at a site we assume that it does not occur at that site.

Interested readers can examine the mean biomass densities of all species in Fig.\@ref(fig:targ-trend-plot)-\@ref(fig:nontarg-trend-plot).

We include several additional sources of data in our regression analysis. Temperature readings are included from the PISCO data for each transect. We also include PISCO data on the estimated surge and visibility. We augmented these data with information on kelp cover over time from the Santa Barbara Channel Long Term Ecological Research Network [@lter2017]. We used a k-nearest neighbors algorithm to fill in missing kelp observations, and matched the interpolated kelp data to the PISCO data at the resolution of year-month-site (Fig.\@ref(fig:kelp-plot)).

Temperature data were augmented with data from `FishLife` [@thorson2017c] to include the estimated preferred temperature for a given species, so that we can include deviations from the preferred temperature envelope as a predictor in the model. This allows different temperatures to have different effects on each species (and is less computationally intensive than estimating species-temperature slopes) (Fig.\@ref(fig:tempdev-plot)).

We also included lagged catch totals in the Santa Barbara region for the commercially harvest species in the database, in an effort to control for changes in density caused by changes in fishing pressure. Catches were pulled from the CDFW website (https://www.wildlife.ca.gov/Fishing/Commercial/Landings), and extracted using the `tabulizer` package in R [@leeper2018] (Fig.\@ref(fig:catches-plot)).

```{r tempdev-plot, fig.cap = "Mean deviations from preferred temperature by species and year"}

abundance_data$data[[1]] %>% 
  group_by(commonname, year) %>% 
  summarise(mean_tempdev = mean(temp_deviation)) %>% 
  ggplot(aes(year, mean_tempdev)) + 
  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
  geom_line() + 
  facet_wrap(~commonname, scales = "free_y") + 
  theme_minimal() + 
  labs(x = "Year", y = "Deviation from mean preferred temperature")

```

```{r catches-plot, fig.cap = "Total CDFW reported commercial catches in the Santa Barbara region"}

abundance_data$data[[1]] %>% 
  filter(targeted == 1) %>% 
  group_by(commonname, year) %>% 
  summarise(catches = unique(lag_catch)) %>% 
  group_by(commonname) %>% 
  mutate(total_catches = sum(catches)) %>% 
  ungroup() %>% 
  filter(total_catches > 0) %>% 
  ggplot(aes(year, catches)) + 
  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
  geom_line() + 
  facet_wrap(~commonname, scales = "free_y") + 
  theme_minimal() + 
  labs(x = "Year", y = "Catch (lbs)")

```

```{r kelp-plot, fig.cap="Mean kelp biomass by island over time from SBC LTER"}

abundance_data$data[[1]] %>% 
  group_by(region,year) %>% 
  summarise(mean_kelp = mean(interp_kelp, na.rm = TRUE)) %>% 
  ggplot(aes(year, (mean_kelp), color = region)) + 
  geom_line() + 
  labs(x = "Year", y = "Mean Kelp Biomass")

```

```{r targ-trend-plot, fig.cap="Centered and scaled mean biomass densities of all targeted finfish in analysis before filtering"}

targeted_trend_plot <- pisco_data %>% 
  filter(targeted == 1) %>% 
  group_by(year, classcode, targeted, commonname) %>% 
  summarise(mean_biomass_density = mean(density_g_m2, na.rm = TRUE)) %>% 
  group_by(classcode) %>% 
  mutate(mean_biomass_density = scale(mean_biomass_density)) %>% 
  ggplot(aes(year,mean_biomass_density)) + 
  geom_hline(aes(yintercept = 0), linetype = 2) + 
  geom_vline(aes(xintercept = 2003), color = "red") +
  geom_line() + 
  facet_wrap(~commonname) +
  theme_minimal() + 
  theme(strip.text = element_text(size = 6),
        axis.text.x = element_text(size = 8)) + 
  labs(x = "Year", y = "Centered and Scaled Mean Biomass Density")

targeted_trend_plot
```

```{r nontarg-trend-plot, fig.cap="Centered and scaled mean biomass densities of all non-targeted finfish in analysis before filtering"}

nontargeted_trend_plot <- pisco_data %>% 
  filter(targeted == 0) %>% 
  group_by(year, classcode, targeted, commonname) %>% 
  summarise(mean_biomass_density = mean(density_g_m2, na.rm = TRUE)) %>% 
  group_by(classcode) %>% 
  mutate(mean_biomass_density = scale(mean_biomass_density)) %>% 
  ggplot(aes(year,mean_biomass_density)) + 
  geom_hline(aes(yintercept = 0), linetype = 2) + 
  geom_vline(aes(xintercept = 2003), color = "red") +
  geom_line() + 
  facet_wrap(~commonname) +
  theme_minimal() + 
  theme(strip.text = element_text(size = 6),
        axis.text.x = element_text(size = 8)) + 
  labs(x = "Year", y = "Centered and Scaled Mean Biomass Density")

nontargeted_trend_plot
```

```{r classcode}

species_data <- cip_data %>% 
  select(classcode, commonname, taxa, targeted) %>% 
  unique() %>% 
  arrange(desc(targeted)) %>% 
  mutate(targeted = targeted == 1) %>% 
  rename(`Common Name` = commonname, 
         `Scientific Name` = taxa,
         `Targeted?` = targeted)

status <- read_csv(here::here("data","stock-status-summaries.csv")) %>% 
  select(classcode, `Stock Status`)

species_data <- species_data %>% 
  left_join(status, by = "classcode")

knitr::kable(species_data,
             digits = 2,
             caption = "Species included in estimation model",
             longtable = TRUE,
             booktabs = TRUE,
             col.names = colnames(species_data)) %>%
  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))

```

### Model

The regression analysis uses a mixed-effects hierarchical model. The raw data are estimated length compositions by fish species along a survey transect at a site. Lengths are converted to biomass per allometric relationships supplied by PISCO and supplemented by the `FishLife` [@thorson2017c] package in R where needed. We performed some minimal data filtering to reduce noise in the data. We only include species that were observed at least twice in each year of the dataset (2000-2017) somewhere in the core Channel Islands (Anacapa, Santa Cruz, Santa Rosa, San Miguel). While some data are available from 1999, per consultation with PISCO we omit those data due to changes in survey protocols. We assign species to targeted and non-targeted groups per the PISCO classifications. This filtering process results in `r filter_summary$no` non-targeted species and `r filter_summary$yes` targeted species remaining in the analysis.

The simplified explanation of the estimation is a hierarchical model in which we first standardize the observed biomass densities into an abundance index of each species over time. The abundance indices in each year are assumed to be log-normally distributed with means and standard deviations for the targeted and non-targeted groups, giving an estimate of the mean densities of targeted and non-targeted species over time. We then calculate the difference between mean density of targeted species and the mean density of non-targeted species in each year.

\newpage
```{r cartoon, fig.cap= "Cartoon illustration of the hierarchichal difference-in-difference estimator", fig.width=3}
knitr::include_graphics("figs/estimation-cartoon.png")
```

\pagebreak

The first stage of the regression is a log-normal delta model. The model estimates two regressions, the first is a binomial generalized linear model (GLM) with a logit link estimating the probability of observing a given fish species at a observation *i* (transect at time *t*). The probability that a given species was observed *o* at a given observation is distributed

\begin{equation}
o_{s,i} \sim binomial(\frac{1}{1 +e^{-\beta^{o}{X}}})
  (\#eq:binom)
\end{equation}

 where $\beta^{o}$ are the estimated coefficients for the observation model and *X* is a matrix of covariates that include random effects for each year in the data (2000 to 2017).

The expected density *d* of positive observations is modeled per a log-normal distribution

 \begin{equation}
  log(d_{s,i}) \sim normal(\beta^{d}X, \sigma_s)
   (\#eq:logd)
 \end{equation}

where $\beta^{d}$ are the estimated coefficients for the expected density model and *X* is the same matrix of covariates as used in the observation portion of the model and $\sigma_s$ allows for each species *s* to have different standard deviations.

Our covariate matrix *X* contains both fixed and random effects. Fixed effects include the depth level of the transect, the sampling site, the month of the observation, the estimated surge at the transect, visibility, the depth of the transect, and the experience (and experience squared) of the diver conducting the transect. We classify each species into one of two clusters based on the mean longitude the species was encountered at, breaking the species into two groups: those primarily found in the western end of the Channel Islands those found more in the eastern end.  We then estimate random effects for each island for each cluster

\begin{equation}
\beta_{island,cluster} \sim normal(0,\sigma_{cluster})
  (\#eq:island)
\end{equation}

This allows the mean effect of each island to differ for each cluster, e.g. allowing San Miguel, the easternmost island, to have a higher mean density for eastern species than for more western species (if the data suggest it).

The second critical component of the covariate matrix *X* are random effects for each year for each species

\begin{equation}
  \beta_{year,species} \sim normal(0,\sigma_{species})
  (\#eq:species)
\end{equation}

These $\beta_{year,species}$ represent our "standardized" estimate of observed abundance of each species in each time step, controlling for the included covariates.

However, we still need to account for changes in the probability of detection over time. For that, we create a standard matrix of with rows equal to the number of years and columns corresponding to each of the columns in *X*, holding everything fixed at mean (or most frequently observed level for factors) levels for all variables in X except for the year and species interaction indices. Calling this standardized matrix $X^{standard}$, the probability of observing a given species in year *y* is then

\begin{equation}
p_{s,y} = (\frac{1}{1 +e^{-\beta^{o}{X^{standard}}}})
  (\#eq:pdetect)
\end{equation}

In the same manner as described by @punt2000, The standardized index of abundance for species *s* in year *y* then is

\begin{equation}
I_{species,year} = p_{species,year}e^{\beta_{species,year}}
  (\#eq:index)
\end{equation}

The next phase of the model requires us to estimate the mean abundance of targeted and non-targeted species over time. The concept here is that each $I_{species,year}$ can be modeled by a regression that contains random effects for each year for targeted and non-targeted fishes, the assumption then being that there is a mean density for targeted and non-target species, and $I_{species,year}$ represent deviations from that mean.

\begin{equation}
log(I_{species,year}) \sim normal(\beta^{effect}X^{effect}, \sigma_I)
  (\#eq:indexprior)
\end{equation}

$X^{effect}$ contains both fixed and random effects. The fixed effects include an intercept and the temperature deviation for a given species in a year, where temperature deviation is

\begin{equation}
t_{s,y} = (t^{pref}_{s} -  \bar{t_{y}})^2
  (\#eq:tempdev)
\end{equation}

where $t^{pref}$ is the preferred temperature for species *s* (drawn form `FishLife`, @thorson2017c), and $\bar{t_{y}}$ is the mean temperature encountered by that species in year *y*. We also include as variables in the model the mean kelp cover experienced by a given species in a given year, as well as the total fishery catches reported in the previous year for that species in the Santa Barbara region [drawn from the California Department of Fish and Wildlife database].  We also include random intercepts for each species in $X^{effect}$. The most important random effects are year effects for targeted and non-targeted species

\begin{equation}
\beta_{year,targeted} \sim normal(0,\sigma_{targeted})
  (\#eq:betatarg)
\end{equation}

$\beta_{year,targeted}$ is the mean log density of targeted species in year *y*, controlling for included covariates. Therefore, the final step in the model, the divergence in the standardized abundance trends of targeted and non-targeted species is

\begin{equation}
divergence_{year} =  \beta_{year,targeted = 1} - \beta_{year,targeted = 0}
  (\#eq:mpaeffect)
\end{equation}

The model is fit in TMB to integrate the uncertainty across all levels of the model, with standard errors for each coefficient in the model estimated through the Laplace approximation.

A complete table of estimated coefficients can be seen in Table.S\@ref(tab:betatable).

<!-- Stock status estimates are not available for many of the targeted species in the database. The California Department of Fish and Wildlife compiles status estimates for many nearshore finfish (https://www.wildlife.ca.gov/Conservation/Marine/Groundfish/Nearshore-Finfish), including recent stock assessments.  For species not listed here we searched for available information on recent stock status from relevant stock assessments.  -->

```{r betatable}

table <- betas %>% select(-std_error,-group)

knitr::kable(table,
             digits = 2,
             caption = "Complete table of estimated coefficients from hierarchical difference-in-difference model",
             longtable = TRUE,
             booktabs = TRUE,
             col.names = colnames(table)) %>% 
  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))

```

Figures S\@ref(fig:fe-plot):S\@ref(fig:targ-plot) present estimated effects for covariates included in the model, along with the raw estimated mean trends of the targeted and non-targeted species (while the difference between these trends is presented in our main results).

\newpage

```{r fe-plot, fig.cap = "Estimated coefficients for non-spatial fixed effects in observation model (seeing) and observed model (seen)"}

 non_nested_beta_plot <- betas %>%
   filter(str_detect(group, "non_nested") & !str_detect(variable, "site_side")) %>%
   ggplot() +
   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
   geom_pointrange(aes(x =variable,
                       y = estimate,
                       ymin = lower,
                       ymax = upper)) +
   facet_wrap(~group, labeller=labeller(group=c(
     seeing_non_nested_betas="Observation Model\n(seeing)",
     seen_non_nested_betas = "Observed Model\n(seen)"
   ))) +
   coord_flip() +
   theme(axis.text.y = element_text(size = 10))

non_nested_beta_plot

```

\pagebreak

```{r site-plot, fig.cap = "Estimated coefficients for spatial random effects in observation model (seeing) and observed model (seen)", fig.height=10}

 non_nested_site_beta_plot <- betas %>%
   filter(str_detect(group, "non_nested") & str_detect(variable, "site_side")) %>%
  mutate(variable = tolower(variable)) %>% 
  mutate(variable = str_replace(variable, "site_side","")) %>% 
  mutate(island = str_replace(str_split(variable, '_', simplify = TRUE)[,1],'-','')
 ) %>% 
   ggplot() +
   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
   geom_pointrange(aes(x = variable,
                       y = estimate,
                       ymin = lower,
                       ymax = upper)) +
   facet_grid(island~group, scales = "free",labeller=labeller(group=c(
     seeing_non_nested_betas="Observation Model\n(seeing)",
     seen_non_nested_betas = "Observed Model\n(seen)"),
     island=c(anacapa="ANA",sci="SCI",smi="SMI",sri="SRI"))) +
   coord_flip() +
   theme(axis.text.y = element_text(size = 10)) + 
  theme_minimal()

non_nested_site_beta_plot + 
  theme(axis.text.y = element_text(size = 6))

```

\pagebreak

```{r region-plot, fig.cap="Random effects for region by cluster", eval = FALSE}

  region_cluster_plots <- betas %>%
   filter(str_detect(group, "region_cluster_betas")) %>%
   mutate(cluster = str_replace_all(variable,"\\D","")) %>%
   mutate(region = str_split(variable,'-', simplify = T)[,3]) %>%
   ggplot() +
    geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
    geom_pointrange(aes(x = region,
                        y = estimate,
                        ymin = lower,
                        ymax = upper, color = cluster)) +
    facet_wrap(~group,labeller=labeller(group=c(
     seeing_region_cluster_betas="Observation Model\n(seeing)",
     seen_region_cluster_betas = "Observed Model\n(seen)")))

region_cluster_plots

```

\pagebreak

```{r targ-plot, fig.cap = "Trends in standardized mean abundance of targeted and non-targeted species"}

targeted_trends <- report$zissou_estimates %>% 
  filter(str_detect(variable, "targeted_did")) %>% 
  group_by(variable) %>% 
  mutate(year =zissou_fit$did_data$year %>% unique(),
         labvar=ifelse(variable=="nontargeted_did_betas","Nontargeted","Targeted")) %>% 
  ungroup() %>% 
  ggplot() + 
  geom_line(aes(year, estimate, color = labvar),alpha = 0.5) +
  geom_pointrange(aes(year, estimate, ymin = lower, ymax = upper, color = labvar)) + 
  scale_color_d3(name = "") +
  theme(legend.position = c(0.8,0.9))

targeted_trends

```

\pagebreak

### Regression Diagnostics

We include visual diagnostics of our estimation model. All coefficients passed convergence criteria for TMB.

Looking first at the predictions of the model for the positive observations in the data (i.e. using the full model to predict biomass densities, and then comparing those predictions to cases where some positive biomass densities were observed), the model diagnostics show no clear problems. The $R^2$ of the model is 0.43. Residuals do not exhibit trends, though some grouping the residuals is evident. The quantile-quantile plot suggest that on the assumption of log-normal errors on the observed densities is reasonable, though the model appears to have some slight problems estimating the highest observed densities (Fig.S\@ref(fig:obs-v-pred-plot)).

```{r}

fit_report <- base_run$tmb_fit[[1]]$zissou_report

observed_seen <- base_run$tmb_fit[[1]]$zissou_data$log_density

observed_seeing <- base_run$tmb_fit[[1]]$zissou_data$any_seen

seen_diagnostics <- data_frame(observed = observed_seen, predicted = as.numeric(fit_report$log_density_hat)) %>% 
  mutate(residuals = predicted - observed)

seeing_diagnostics <- data_frame(observed = observed_seeing, predicted = as.numeric(fit_report$prob_seeing))

```

```{r, eval = FALSE}

a = base_run$tmb_fit[[1]]$zissou_data$x_seen_non_nested %>% 
  as_data_frame() %>% 
  select(contains("mean_")) %>% 
  mutate(resid = seen_diagnostics$residuals) %>% 
  gather(variable, value, -resid) %>% 
  ggplot(aes(value, resid)) + 
  geom_point() + 
  facet_wrap(~variable)

a

```

```{r obs-v-pred-plot, fig.cap="High level diagnostics for observed compontent of Delta-GLM: Observed vs predicted log densities (A), predicted log density vs residuals (B), and a normal qq-plot of the residuals (C)"}

obs_v_pred_plot <- seen_diagnostics %>% 
  ggplot(aes(observed, predicted)) + 
  geom_point(alpha = 0.5) + 
  geom_abline(aes(intercept = 0, slope = 1), linetype = 2, color = "red") + 
  labs(caption = glue::glue("R^2 is {round(yardstick::rsq(data = seen_diagnostics, truth = observed, estimate = predicted)$.estimate,2)}"), title= "A") + 
  theme(plot.margin = unit(rep(.1,4), units = "lines"))

fit_v_resid_plot <- seen_diagnostics %>% 
  ggplot(aes(predicted, residuals)) + 
  geom_point() + 
  geom_hline(aes(yintercept = 0), linetype = 2, color = "red") + 
  labs(title = "B") +
    theme(plot.margin = unit(rep(.1,4), units = "lines"))

qq_plot <- ggplot(seen_diagnostics, aes(sample = residuals)) +
  stat_qq() +
  stat_qq_line() + 
  labs(title = "C") + 
    theme(plot.margin = unit(rep(.1,4), units = "lines"))

obs_v_pred_plot/(fit_v_resid_plot + qq_plot)

```

In order to evaluate the ability of the model to estimate positive observations, we can compare the the binned predicted probability of a positive observation to the proportion of observations in that bin that recorded positive observations. If our model is working well, we would expect a group of fisheries that our model estimates on average should have a 50% probability of a positive observation, then we should expect about 50% of those observations to have positive observations. This is indeed what we see from the model (Fig.S\@ref(fig:obs-prob-plot)).

```{r obs-prob-plot, fig.cap = "Binned mean predicted probability of detection provided by the first stage of the hurdle model vs observed proportion of positive detections"}

seeing_diag_plot <- seeing_diagnostics %>% 
  ungroup() %>% 
  mutate(binned_prob = ntile(predicted, 20)) %>% 
  arrange(binned_prob) %>% 
  group_by(binned_prob) %>% 
  summarise(mean_prob_seen = mean(predicted),
            percent_observed = mean(observed))  %>% 
  ggplot() +
  geom_point(aes(mean_prob_seen, percent_observed)) + 
  geom_abline(aes(intercept = 0, slope = 1), color = "red", linetype = 2) + 
  labs(x = "Binned Predicted Probability of Detection", y = "Proportion of Positive Observations")

seeing_diag_plot
```

We can also examine the receiver-operator-curve (ROC) to assess the performance of the observation component of the model. The area under the curve (AUC) value for the model is 0.84 (on a scale of 0.5 to 1), indicating the model is an overall good predictor of whether or not a given observation of biomass densities will be positive or not.

```{r roc-plot, fig.cap = "Receiver operating characteristic curve of predictions of positive biomass densities"}
library(plotROC)
rocplot <- ggplot(seeing_diagnostics, aes(m = predicted, d = observed))+ geom_roc(n.cuts=20,labels=FALSE) + 
  geom_abline(aes(slope = 1, intercept = 0))

roc_plot <- rocplot + style_roc(theme = zissou_theme)

roc_plot
```

### Standardized Abundance Indices

Overall most species showed consistent trends in biomass densities across the different islands at which they have been observed (Fig.S\@ref(fig:reg-pop)).

```{r reg-pop, fig.cap = "Mean density by island by year for each fish species included in the analysis"}

reg_pop_plot <- fitted_data %>% 
  group_by(region, year, classcode) %>% 
  summarise(mean_density = mean(exp(log_density))) %>% 
  mutate(log_mean_density = log(mean_density)) %>% 
  ungroup() %>% 
  group_by(classcode, region) %>% 
  mutate(cs_mean_density = (mean_density - mean(mean_density)) / sd(mean_density)) %>% 
  ungroup() %>% 
  ggplot(aes(year, cs_mean_density, color = region)) + 
  scale_color_npg(name="Region")+
  geom_line() + 
  labs(y="Centered and Scaled Density",x="Year")+
  facet_wrap(~classcode, scales = "free_y") + 
  theme_minimal()+
  theme(axis.text=element_text(size=6))

reg_pop_plot

```

The standardized indices of abundance generally did not very substantially from the raw mean densities by species over time. However, for some species, such as blue rockfish, the standardized abundance index suggests much higher biomass densities in the pre-MPA period than those reported in the raw data. We suspect this is largely a function of changes in sampling sites over time, that the standardization is better able to account for (Fig.S\@ref(fig:raw-v-stand-plot)).

```{r raw-v-stand-plot, fig.cap = "Raw (points) and standardized (lines) indices of abundance for each of the fishes included in the analysis"}

life_history <- fitted_data %>%
  select(classcode) %>%
  unique() %>%
  mutate(numeric_classcode = as.numeric(as.factor(classcode)))

raw <- fitted_data %>%
  group_by(year, classcode, targeted) %>%
  summarise(md = mean(exp(log_density)),
            mt = mean(temp_deviation)) %>%
  group_by(classcode) %>%
  mutate(smd = (md - mean(md))/(sd(md))) %>%
  ungroup() %>%
  left_join(life_history, by = "classcode")

years <- unique(fitted_data$year)

years <- zissou_fit$did_data$year %>% unique()

abundance_trends <- data_frame(abundance_hat = fit_report$abundance_hat,
                               classcode = rep(1:n_distinct(fitted_data$classcode), each = length(years))) %>%
  mutate(log_abundance_hat = log(abundance_hat)) %>%
  group_by(classcode) %>%
  mutate(year = years) %>%
  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
  ungroup() %>%
  rename(numeric_classcode = classcode) %>%
  left_join(life_history, by = "numeric_classcode" )

# 
# abundance_trends %>%
#   ggplot(aes(year, log(abundance_hat), color = factor(classcode))) +
#   geom_line(show.legend = F) +
#   geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
#   facet_wrap(~classcode) +
#   theme_minimal()

raw_v_std_plot <- abundance_trends %>%
  filter(year %in% years) %>% 
  ggplot(aes(year, scaled_abundance_hat, color = factor(classcode))) +
  geom_line(show.legend = F) +
  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
  geom_point(data = raw,aes(year, smd, color = factor(classcode)), show.legend = F)+
  facet_wrap(~classcode) +
  labs(y="Raw (dots) and Standardized (lines) Abundance",x="Year")+
  theme_minimal()+
  theme(axis.text=element_text(size=6))

raw_v_std_plot

```

We include a variety of environmental, observation, and temporal indicators in our model. Inclusion of highly co-linear variables in a model can inflate standard errors and obscure "true" effects. To account for this we calculated the Pearson's correlation coefficients for all of the continuous data included in our model to ensure that none of the included variables had correlation coefficients greater than 0.7, a general rule of thumb for co-inclusion of variables. We did not find problematic levels of correlation among any of our included continuous variables.

```{r numeric-cor, fig.cap="Pearson correlation coefficients of continuos data included in the regression model"}

numeric_cor_plot <- base_run$data[[1]] %>%
  select_if(is.numeric) %>%
  select(-log_density, -targeted) %>%
  na.omit() %>%
  corrr::correlate() %>%
  gather("colname", "correlation", -rowname) %>%
  ggplot(aes(rowname, colname, fill = correlation)) +
  geom_tile() +
  geom_text(aes(rowname, colname, label = round(correlation, 2))) +
  scale_fill_gradient2(
    low = "tomato",
    mid = "white",
    high = "steelblue",
    midpoint = 0,
    limits = c(-1, 1)
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      vjust = 1,
      hjust = 1,
    size = 6),
    axis.text.y = element_text(size = 6),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )

numeric_cor_plot
```

## Alternative estimation models

Our proposed identification strategy attempts to control for non-MPA (and not directly modeled) related changes in abundances through the trend in the non-targeted species. However, a simpler alternative would be to simply compare densities before-and-after MPA implementation, while explicitly controlling for non-MPA related factors that we believe may have some effect on densities (a "selection on observables" strategy). To that end, we fit a mixed-effects regression that models log densities of targeted species only (positive observations only, for the sake of simplicity) as a function of temperature deviations, kelp cover, observer experience, random effects for species and region, and fixed effects for each year in the data (omitting the year 2000). The hypothesis here is that any non-MPA related factors that affect densities are accounted for in the observed variables included in the model.

Using this model, densities of targeted species appear to have been declining steadily since 2000, and appear to have plateaued off since the implementation of MPAs in 2003. Without an identification strategy such as the one employed in this study then, all we could conclude is that densities appear to be lower post-MPA, and have not increased substantially over time (Fig.S\@ref(fig:obs-plot)).

```{r obs-plot, fig.cap = "Selection on observables identification strategy. Plotted estimates are fixed effects of year on log-density (relative to the year 2000), controlling for observer experience, temperature deviations, and kelp cover, with random effects for species and region"}

test <- lme4::lmer(log_density~ factor_year + cumulative_n_obs  + temp_deviation + interp_kelp + (1|classcode) + (1|region), data = fitted_data %>% filter(targeted == 1, any_seen == T))

sel_on_obs_plot <- broom::tidy(test) %>% 
  filter(str_detect(term, "factor_year")) %>% 
  mutate(year = str_replace_all(term,"\\D","") %>% as.numeric()) %>% 
  ggplot() + 
  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
  geom_pointrange(aes(year, estimate, ymin = estimate - 1.96*std.error, ymax = estimate + 1.96*std.error)) + 
  labs(y = "Estimated Difference", x = "Year")

sel_on_obs_plot
```

```{r, include=F}
raw <- fitted_data %>%
  group_by(year, classcode, region, targeted) %>%
  summarise(md = mean(exp(log_density)),
            mt = mean(temp_deviation)) %>%
  group_by(classcode) %>%
  mutate(smd = (md - mean(md))/(sd(md))) %>%
  ungroup() %>%
  left_join(life_history, by = "classcode")

raw_region <- fitted_data %>%
  group_by(year, region,classcode, targeted) %>%
  summarise(md = mean(exp(log_density)),
            mt = mean(temp_deviation)) %>%
  group_by(classcode) %>%
  mutate(smd = (md - mean(md))/(sd(md))) %>%
  ungroup() %>%
  left_join(life_history, by = "classcode")

raw_mpa_only <- fitted_data %>%
  filter(eventual_mpa == T) %>% 
  group_by(year, classcode, targeted) %>%
  summarise(md = mean(exp(log_density)),
            mt = mean(temp_deviation)) %>%
  group_by(classcode) %>%
  mutate(smd = (md - mean(md))/(sd(md))) %>%
  ungroup() %>%
  left_join(life_history, by = "classcode")

raw_nompa_only <- fitted_data %>%
  filter(eventual_mpa == F) %>% 
  group_by(year, classcode, targeted) %>%
  summarise(md = mean(exp(log_density)),
            mt = mean(temp_deviation)) %>%
  group_by(classcode) %>%
  mutate(smd = (md - mean(md))/(sd(md))) %>%
  ungroup() %>%
  left_join(life_history, by = "classcode")

# basic_model <- rstanarm::stan_glm(log(md + 1e-3) ~ factor(year) + targeted + factor(year):targeted, data = raw)

```

The estimation model used for our main results is complicated. We feel this complexity is justified in order to best capture the uncertainty inherent in the challenging task of conducting underwater visual surveys, as well as the spatio-temporal nature of the underlying data. However, we also ran several simpler models in order to examine the sensitivity of our results to the model structure selected here.

Much of the complexity of our model comes from the integration of a standardized abundance index for each of the species, which are then compiled into a standardized index of abundance index for targeted species as a whole.

In one simpler approach, we aggregated the PISCO data into mean biomass densities by species and year across all the Channel Islands. Since all included species have at least some positive observations in each year, this removes the complication of dealing with the probability of detection problem in the raw data, while of course assuming that the net effects of all of the other factors included in our main model are on average zero (e.g. observer skill, visibility, etc.). A difference-in-difference model fit to these simplified data show qualitatively similar (though much more uncertain) results to our full model. In particular, the model shows the same gradual increase in targeted densities relative to non-targeted until 2013, followed by an attenuation of this trend. Importantly, though, this simplified model fails to correct for some pre-MPA differences in the two groups, leading to negative "effects" of the MPAs being estimated before the MPAs themselves went in place. While these sort of anticipatory effects are certainly possible [@mcdermott2019], in this case we would suggest the likelier explanation---given that these anticipatory effects disappear in the full model---is that our full model provides important controls for pre-MPA characteristics (Fig.S\@ref(fig:sortasimple-did))

```{r, results = "hide"}

year_mpa <- 2002

  data <- abundance_data$data[abundance_data$data_source == "pisco"][[1]] %>%
    mutate(density = exp(log_density) * any_seen) %>%
    mutate(targ = case_when(targeted == 1 ~ "Targeted",  TRUE ~ "Non-Targeted"))

  sortasimple <- data %>%
    group_by(year, classcode) %>%
    summarise(mean_density = mean(density),
              targ = unique(targ)) %>%
    mutate(mpa = year > year_mpa) %>%
    ungroup() %>%
    mutate(fyear = factor(year)) %>%
    mutate(fyear = relevel(fyear, ref = "2003"))

  sortasimple_reg <-
    rstanarm::stan_glmer(
      log(mean_density + 1e-3) ~ targ * fyear + (1 |
                                                   classcode),
      data = sortasimple,
      cores = 4
    )

  sortasimple_plot <- bayesplot::mcmc_areas(as.matrix(sortasimple_reg),
                                            regex_pars = ":fyear") +
    geom_vline(aes(xintercept = 0), linetype = 2, color = "red") +
    scale_y_discrete(labels = c(2000:2001, 2003:2017)) +
    coord_flip() +
    scale_x_percent() + 
    zissou_theme + 
    labs(x="Estimated Difference",y="Year")+
    theme(axis.text.x = element_text(size = 8))

```
```{r sortasimple-did, fig.cap = "Results of simplified difference-in-difference regression. The model estimates the difference in the mean trend of densities of targeted and non-targeted species over time, controlling for the mean densities of each individual species group. Data are initially aggregated to the level of species-by-year."}
sortasimple_plot
```

```{r species-by-species-did, results = "hide"}

year_mpa <- 2002

  data <- abundance_data$data[abundance_data$data_source == "pisco"][[1]] %>%
    mutate(density = exp(log_density) * any_seen) %>%
    mutate(targ = case_when(targeted == 1 ~ "Targeted",  TRUE ~ "Non-Targeted"))

  sortasimple <- data %>%
    group_by(year, classcode) %>%
    summarise(mean_density = mean(density),
              targ = unique(targ)) %>%
    mutate(mpa = year > year_mpa) %>%
    ungroup() %>%
    mutate(fyear = factor(year)) %>%
    mutate(fyear = relevel(fyear, ref = "2003"))

  sortasimple_reg <-
    rstanarm::stan_glmer(
      log(mean_density + 1e-3) ~ targ * fyear + (1 |
                                                   classcode),
      data = sortasimple,
      cores = 4
    )

  sortasimple_plot <- bayesplot::mcmc_areas(as.matrix(sortasimple_reg),
                                            regex_pars = ":fyear") +
    geom_vline(aes(xintercept = 0), linetype = 2, color = "red") +
    scale_y_discrete(labels = c(2000:2001, 2003:2017)) +
    coord_flip() +
    scale_x_percent() + 
    zissou_theme + 
    labs(x="Estimated Difference",y="Year")+
    theme(axis.text.x = element_text(size = 8))

```

As an even simpler approach, we aggregated the data to the level of targeted and non-targeted species, and estimated the divergence in their trends over time.

```{r, results = "hide"}

  data <- abundance_data$data[abundance_data$data_source == "pisco"][[1]] %>%
    mutate(density = exp(log_density) * any_seen) %>%
    mutate(targ = case_when(targeted == 1 ~ "Targeted",  TRUE ~ "Non-Targeted"))

  supersimple <- data %>%
    group_by(year, targ) %>%
    summarise(mean_density = mean(density)) %>% 
    mutate(mpa = year > year_mpa) %>%
    ungroup() %>%
    mutate(fyear = factor(year)) %>%
    mutate(fyear = relevel(fyear, ref = "2002"))

  supersimple_reg <-
    rstanarm::stan_glm(
      log(mean_density) ~ targ + fyear + targ:fyear,
      data = supersimple,
      cores = 4,
      iter = 2000,
      adapt_delta = 0.95
    )

    # supersimple_reg <-
    # rstanarm::stan_glm(
    #   log(mean_density + 1e-3) ~ targ * fyear,
    #   data = supersimple,
    #   cores = 4,
    #   iter = 10000
    # )

  supersimple_plot <- bayesplot::mcmc_areas(as.matrix(supersimple_reg),
                                            regex_pars = ":fyear") +
    geom_vline(aes(xintercept = 0), linetype = 2, color = "red") +
    scale_y_discrete(labels = c(2000:2001, 2003:2017)) +
    coord_flip() +
    scale_x_percent() + 
    zissou_theme + 
    labs(y="Year",x="Estimated Difference")+
    theme(axis.text.x = element_text(size = 8))

```

```{r}
supersimple_plot
```

We can also explore the effects of our species filtering by running a simplified regression but now including all the species in the database, no matter of how infrequently they are observed.

```{r, results= "hide"}
year_mpa <- 2002

  sortasimple <- pisco_data %>%
    group_by(year, classcode) %>%
    summarise(mean_density = mean(density_g_m2, na.rm = TRUE),
              targ = unique(targeted)) %>%
    mutate(mpa = year > year_mpa) %>%
    ungroup() %>%
    mutate(fyear = factor(year)) %>%
    mutate(fyear = relevel(fyear, ref = "2003"))

  sortasimple_reg <-
    rstanarm::stan_glmer(
      log(mean_density + 1e-3) ~ targ * fyear + (1 |
                                                   classcode),
      data = sortasimple,
      cores = 4
    )

  all_species_sortasimple_plot <- bayesplot::mcmc_areas(as.matrix(sortasimple_reg),
                                            regex_pars = ":fyear") +
    geom_vline(aes(xintercept = 0), linetype = 2, color = "red") +
    scale_y_discrete(labels = c(1999:2001, 2003:2017)) +
    coord_flip() +
    scale_x_percent() + 
    zissou_theme + 
    labs(x="Estimated Difference",y="Year")+
    theme(axis.text.x = element_text(size = 8))
```

```{r sortasimple-noilter-did, fig.cap = "Results of simplified difference-in-difference regression including all observed finfish. The model estimates the difference in the mean trend of densities of targeted and non-targeted species over time, controlling for the mean densities of each individual species group. Data are initially aggregated to the level of species-by-year."}
all_species_sortasimple_plot
```

### Synthetic controls

Synthetic controls are an alternative method for attempting to estimate the causal effect of a policy intervention [@abadie2010]. A difference-in-difference approach assumes that some observable group serves as an adequate control for the state of the treated group in an untreated world. In our default case, we assume that the mean standardized index of non-targeted species are our control for the targeted species. Alternatively, synthetic controls use timeseries of treated and non-treated groups before and after treatment to construct a new "control" group built by weighting the pre-treatment timeseries of un-treated observations (together with covariates) such that the synthetic control group matches the trends in the treated group pre-treatment.

We chose to present difference-in-difference as our main result since it better allows us to capture the uncertainty in the data generating process through our hierarchical model. However, we felt that it was worth exploring whether synthetic controls provided substantially different results than our default model.

For the first synthetic control, we pulled our standardized mean index of abundance for targeted species as a whole from our difference-in-difference model as our treated group. We then pull the standardized indices of abundance for each of the non-targeted groups from the difference-in-difference to use as the candidate untreated components for the synthetic controls. A complete synthetic control analysis would require more extensive validation of the methods, but we use this approach simply to explore whether we observe substantially divergent results in the synthetic control versus the difference-in-difference model.

We centered and scaled the candidate abundance indices to facilitate model convergence given the very few number of pre-treatment years available. The results of a synthetic control model are presented as the difference between the observed treatment outcome and the synthetic control (the difference in this case being in units of standard deviations). The model was not able to construct an adequate synthetic control at this level, as shown by the differences between the treated group and the synthetic control pre-treatment. However, we would note that the post-treatment results do show similarities with our main results, namely a lack of a clear divergence between the treatment and the control, and an upwards trend up through the earl 2010s followed by a decline (Fig.S\@ref(fig:total-synth)).

```{r, include = FALSE}

nontargeted_index <- data_frame(abundance_hat = fit_report$abundance_hat,
                               classcode = rep(1:n_distinct(fitted_data$classcode), each = length(years))) %>%
  mutate(log_abundance_hat = log(abundance_hat)) %>%
  group_by(classcode) %>%
  mutate(year =years) %>%
  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
  ungroup() %>%
  rename(numeric_classcode = classcode) %>%
  left_join(life_history, by = "numeric_classcode" ) %>% 
  left_join(unique(abundance_data$data[[1]] %>% select(classcode, targeted)), by = "classcode") %>% 
  filter(targeted == 0) %>% 
  select(log_abundance_hat, year, classcode, targeted)

targeted_index <- zissou_fit$zissou_estimates %>%
  filter(
    stringr::str_detect(variable, "(targeted_did_betas)"),!str_detect(variable, "non")
  ) %>%
  mutate(
    year = years,
    classcode = "targeted",
    targeted = 1
  ) %>%
  rename(log_abundance_hat = estimate) %>% 
  select(colnames(nontargeted_index))

  # filter(variable == "seeing_year_species_betas") %>%

nontargeted_covariates <- pisco_data %>% 
  select(classcode, contains("mean"),year, targeted) %>% 
  filter(targeted == 0) %>% 
  select(-mean_length) %>% 
  gather(variable, value, -classcode,-year,-targeted) %>% 
  group_by(classcode, variable, year) %>% 
  summarise(mean_value = mean(value, na.rm = TRUE)) %>% 
  spread(variable, mean_value)

targeted_covariates <- pisco_data %>% 
  select(classcode, contains("mean"),year, targeted) %>% 
  filter(targeted == 1) %>% 
  mutate(classcode = "targeted") %>% 
  select(-mean_length) %>% 
  gather(variable, value, -classcode,-year,-targeted) %>% 
  group_by(classcode, variable, year) %>% 
  summarise(mean_value = mean(value, na.rm = TRUE)) %>% 
  spread(variable, mean_value)

synth_covariates <- nontargeted_covariates %>% 
  bind_rows(targeted_covariates)

synth_data <- targeted_index %>% 
  bind_rows(nontargeted_index) %>% 
  left_join(synth_covariates, by = c("classcode","year")) %>% 
    mutate(numeric_classcode = as.numeric(factor(classcode))) %>% 
  group_by(classcode) %>% 
  mutate(log_abundance_hat = scale(exp(log_abundance_hat))) %>% 
  ungroup() %>% 
  as.data.frame()

# synth_data %>%
#   group_by(classcode) %>%
#   mutate(index = (log_abundance_hat)) %>%
#   ggplot(aes(
#     year,
#     index,
#     color = factor(targeted),
#     group = interaction(targeted, classcode)
#   )) +
#   geom_line()

prepped_synth_data <-
  Synth::dataprep(
    foo = synth_data,
    predictors = colnames(synth_data)[str_detect(colnames(synth_data),"mean")],
    predictors.op = "mean",
    dependent = "log_abundance_hat",
    unit.variable = "numeric_classcode",
    unit.names.variable = "classcode",
    time.variable = "year",
    treatment.identifier = "targeted",
    controls.identifier = unique(synth_data$classcode)[unique(synth_data$classcode) != "targeted"],
    time.predictors.prior = years[years <= 2003],
    time.optimize.ssr = years[years <= 2003],
    time.plot = years
  )

synth_fit <- Synth::synth(prepped_synth_data)

gaps <-
  rownames_to_column(
    prepped_synth_data$Y1plot - (prepped_synth_data$Y0plot %*% synth_fit$solution.w) %>%
      as.data.frame()
  ) %>%
  rename(gap = w.weight,
         year = rowname) %>%
  mutate(year = as.numeric(year)) %>%
  mutate(synthetic = as.numeric(prepped_synth_data$Y0plot %*% synth_fit$solution.w)) %>% 
  mutate(treatment = gap + synthetic)

# path.plot(synth.res = synth_fit,
#           dataprep.res = prepped_synth_data
# )

total_gap_plot <- gaps %>% 
  ggplot(aes(year, gap)) + 
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_vline(aes(xintercept = 2003), color = "red")+
  geom_line() + 
    labs(x = "Year", y = "Standard deviations from synthetic control")

total_synth_plot <- gaps %>% 
  select(year, synthetic, treatment) %>% 
  gather(source, value, -year) %>% 
  group_by(source) %>% 
  # mutate(value = scale(value)) %>% 
  ggplot(aes(year, value, color = source, linetype = source)) + 
  geom_vline(aes(xintercept = 2003), color = "red")+
  geom_line() + 
    labs(x = "Year", y = "Standard deviations from synthetic control")

```

```{r  total-synth, fig.cap = "Difference in centered and scaled standardized targeted abundance and synthetic standardized targeted abundance"}
total_gap_plot
```

As an extension, we repeated this process, but now treating each targeted species individually as the treated group, and the non-targeted species as the non-targeted. This is intended to explore whether we see clearer signals for individual species than we do for the targeted class as a whole.

Overall we see similarly unclear results as the aggregate targeted synthetic control (and our main results). The synthetic control was better constructed for some individual species, but not clearly for any one, and most species showed some evidence of the upward-then-downward trend seen throughout our results (Fig.S\@ref(fig:synth-classcode-fits)).

```{r, include = FALSE}

synth_foo <-
  function(treated_classcode,
           synth_data,
           targeted_classcodes) {
    # treated_classcode <- targeted_classcodes[1]

    temp_synth_data <- synth_data %>%
      filter(classcode == treated_classcode |
               !(classcode %in% targeted_classcodes))

    prepped_synth_data <-
      Synth::dataprep(
        foo = temp_synth_data,
        predictors = colnames(temp_synth_data)[str_detect(colnames(temp_synth_data), "mean")],
        predictors.op = "mean",
        dependent = "log_abundance_hat",
        unit.variable = "numeric_classcode",
        unit.names.variable = "classcode",
        time.variable = "year",
        treatment.identifier = treated_classcode,
        controls.identifier = unique(temp_synth_data$classcode)[unique(temp_synth_data$classcode) != treated_classcode],
        time.predictors.prior = c(2000:2003),
        time.optimize.ssr = c(2000:2003),
        time.plot = years
      )

    synth_fit <- Synth::synth(prepped_synth_data)

    gaps <-
      rownames_to_column(
        prepped_synth_data$Y1plot - (prepped_synth_data$Y0plot %*% synth_fit$solution.w) %>%
          as.data.frame()
      ) %>%
      rename(gap = w.weight,
             year = rowname) %>%
      mutate(year = as.numeric(year)) %>%
      mutate(synthetic = as.numeric(prepped_synth_data$Y0plot %*% synth_fit$solution.w)) %>%
      mutate(treatment = gap + synthetic)
  }

nontargeted_index <- data_frame(abundance_hat = fit_report$abundance_hat,
                               classcode = rep(1:n_distinct(fitted_data$classcode), each = length(years))) %>%
  mutate(log_abundance_hat = log(abundance_hat)) %>%
  group_by(classcode) %>%
  mutate(year = 1999 + 1:length(abundance_hat)) %>%
  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
  ungroup() %>%
  rename(numeric_classcode = classcode) %>%
  left_join(life_history, by = "numeric_classcode" ) %>% 
  left_join(unique(abundance_data$data[[1]] %>% select(classcode, targeted)), by = "classcode") %>% 
  select(log_abundance_hat, year, classcode, targeted)

synth_covariates <- pisco_data %>% 
  select(classcode, contains("mean"),year, targeted) %>% 
  select(-mean_length) %>% 
  gather(variable, value, -classcode,-year,-targeted) %>% 
  group_by(classcode, variable, year) %>% 
  summarise(mean_value = mean(value, na.rm = TRUE)) %>% 
  spread(variable, mean_value)

synth_data <- nontargeted_index %>% 
  left_join(synth_covariates, by = c("classcode","year")) %>% 
    mutate(numeric_classcode = as.numeric(factor(classcode))) %>% 
  group_by(classcode) %>% 
  mutate(log_abundance_hat = scale(exp(log_abundance_hat))) %>% 
  ungroup() %>% 
  as.data.frame()

targeted_classcodes <- unique(synth_data$classcode[synth_data$targeted == 1])

synth_classcode_fits <-
  tibble(targeted_classcodes = targeted_classcodes) %>%
  mutate(
    synthetic_fit = map(
      targeted_classcodes,
      synth_foo,
      synth_data = synth_data,
      targeted_classcodes = targeted_classcodes
    )
  )

synth_classcode_fits <- synth_classcode_fits %>% 
  unnest()

synth_classcode_fits_plot <- 
synth_classcode_fits %>% 
  ggplot(aes(year, gap)) + 
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_vline(aes(xintercept = 2003), color = "red")+
  geom_line() + 
  facet_wrap(~targeted_classcodes) + 
  labs(x = "Year", y = "Standard deviations from synthetic control")

```

```{r synth-classcode-fits, fig.cap="Synthetic control gaps for each targeted species"}
synth_classcode_fits_plot + 
  theme_minimal()
```

## Testing Model Assumptions

### Simulation testing

We state that a difference-in-difference model using targeted and non-targeted species is capable (conditional on assumptions) of estimating the causal effect of MPAs. We simulated MPA outcomes to test this claim. We first test our estimation strategy under idealized circumstances, where recruitment is deterministic and PISCO divers all have constant and perfect observer skills. We simulate five species that vary only in their maximum size and length at maturity. For each of these species, we set one version that is targeted by fishing and one that is not. We set a constant fishing mortality rate for each simulated targeted species, and then ran two matched simulations, one with MPAs and one without. We then have our simulated divers sample data from each of these scenarios, and then pass the sampled biomass densities to a simplified version of our difference-in-difference model (omitting the probability of detection step). We can then compare the difference-in-difference estimates of the MPA effect to the true simulated effect. The difference in difference model is able to capture the simulated MPA effect under these circumstances (Fig.S\@ref(fig:simple-did-test))

```{r simple-did-test, fig.cap = "Simulated mean (red dashed line) and individual species (solid lines) MPA effects over time, along with difference-in-difference estimated MPA effects (mean with 95% confidence intervals)"}

load(file = file.path(run_dir,'simulated_did.Rdata'))

simple_performance$mixed_effect_did + 
  labs(title = '', x = "Years with MPA protection",
       y = "MPA Effect")
```

We then simulated a more complex example. We use the actual targeted and non-targeted species from our model. We assign species predominately seen in the western Channel Islands as "cold water" and those in the eastern Channel Islands as "warm water". We allow for stochasticity in recruitment. We use El Niño data as a simulated environmental recruitment driver, where we assume that El Niño events produce negative recruitment shocks for cold water species and *vice versa* for warm water species. We simulate three different divers each with different base skill levels, visual selectivities, and an evolving skill rate (such that observers get better over time). We hold fishing mortality rates constant for each species, although that fishing mortality affects each species differently because of intrinsic biological differences in maturity-at-age and steepness. We then test the ability of the difference-in-difference model to isolate the mean MPA effect across all of these targeted species, which our results show it is capable of doing (Fig.S\@ref(fig:complex-did-test)).

```{r complex-did-test,  fig.cap = "Simulated mean (red dashed line) and individual species (solid lines) MPA effects over time, along with difference-in-difference estimated MPA effects (mean with 95% confidence intervals)"}

pisco_performance$mixed_effect_did + 
    labs(title = '', x = "Years with MPA protection",
       y = "MPA Effect")

```

### Sensitivity to "missing" observations

Appropriately addressing the problem of "missing" observations is a critical challenge in any field observation study. If no observations of a given fish species were recorded on a given transect, should the density of that species on that transect be marked as zero, and influence the estimate of the overall mean density accordingly? The obvious answer seems to be yes, but what if that species simply does not live in the environment covered by a particular transect, or was not present during the particular time of the diver's observation? For our base runs, we assign a value of zero density on a given transect for any fish species that has been observed at least once at a given site at any time in our data but was not observed on that particular transect. If that species was never observed at that site, we do not include a zero for that species. Our rationale for this is that given the shifting nature of the sampled sites, and the intensity of sampling at those sites, we do not want to skew density trends by changes in the amount of suitable habitat for a given species sampled. However, this is clearly a strong assumption. For example, perhaps the decreasing trend in mean densities from 2000 to 2004 is due to increased number of sites (and therefore zeros) included in the data. To assess the potential importance of this choice, we can compare the mean densities of targeted and non-targeted species over time with the added zeros to the mean densities using only positive observations (i.e. not including any zeros in the data, (Fig.S\@ref(fig:nozero-raw-trend)). The trends in the raw densities, and most importantly the mean trends of targeted and non-targeted fishes, are nearly identical whether or not zeros are added, providing strong evidence that our choice of how to incorporate missing observations into the data are not strongly influencing our overall results.

```{r nozero-raw-trend, fig.cap = "Centered and scaled mean annual density, excluding zeros, of included fishes (points) and smoothed means of targeted and non-targeted groups (line) over time" }

nozero_trend_plot <- base_run$data[[1]] %>% 
  filter(any_seen == T) %>% 
  select(year, targeted,classcode, log_density, any_seen) %>% 
  mutate(density = exp(log_density) * any_seen) %>% 
  group_by(classcode,targeted, year) %>% 
  summarise(md = mean(density)) %>% 
  group_by(classcode,targeted) %>% 
  mutate(smd = (md - mean(md)) / sd(md)) %>% 
  ggplot(aes(year, smd, color = targeted == 1, fill = targeted == 1)) + 
  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
  geom_line(aes(group = interaction(targeted, classcode)), alpha = 0.25) +
  geom_smooth() +
  scale_color_d3(name = "Targeted?") + 
  scale_fill_d3(name = "Targeted?") +
  labs(y = "Centered and Scaled Mean Density", x = "Year")

nozero_trend_plot
```

### Testing SUTVA with Convergent Cross Mapping

The difference-in-difference model also assumes that the targeted and non-targeted fishes do not directly or indirectly affect each other.This assumption is clearly violated on some level: all the fishes in this analysis are part of the same ecosystem and therefore interact to some degree. For example, if the protection of targeted predatory fishes results in increased mortality of non-targeted fishes, the model would attribute that as an increased regional effect (greater divergence between the abundance of targeted and non-targeted species). Given the time scale of analysis (15 years of protection), we do not feel that massive trophic cascades are likely to have developed yet, given both the pace and complexity of trophic cascade development [@babcock2010; @pershing2015a]. A complete assessment of evidence for trophic cascades in the Channel Islands is beyond the scope of this study, but to address this question somewhat we utilized convergent cross mapping *sensu* @sugihara2012 to test for a significant causal signal between different broad trophic groups in the data, implemented in the `rEDM` package in R.

Convergent cross mapping is a nonlinear forecasting method that uses observed time series data to test for significant causal links between variables. Following methods laid out in @clark2015 and @sugihara2012, we pool the abundance of each broad trophic group by region (Fig.S\@ref(fig:trophic-plot). This uses the data from the islands as "replicates", requiring the assumption that the islands are all part of the same dynamic system, but allowing us to take advantage of the extra information provided by each island to further resolve the reconstructed manifolds. Using these aggregations, we then test whether the variables can be properly embedded, i.e., if they have predictable manifold dynamics. We do this through a simplex forecasting test, using an individual timeseries' own lags to build a manifold. For each timeseries, the "best embedding dimension" is an approximation of the dimensionality of the dynamic system, in other words, the number of dimensions that define and predict the evolving states of the timeseries. This analysis shows that only the carnivore, piscivore, and planktivores groups show evidence of significant predictability (that is, that past dynamics of these species groups can predict future dynamics, Fig.S\@ref(fig:embed-plot)).

```{r trophic-plot, fig.cap="Centered and scaled densities by broad trophic group and island over time"}

trophic_groups <- cip_data %>% 
  select(classcode, broadtrophic) %>% 
  unique()

# dat <- abundance_trends %>%
#   left_join(trophic_groups, by = "classcode") %>%
#   group_by(broadtrophic, year) %>%
#   summarise(density = mean(scaled_abundance_hat)) %>%
#   group_by(broadtrophic) %>%
#   mutate(density = scale(density)) %>%
#   ungroup() %>% 
#   spread(broadtrophic, density) %>% 
#   arrange(year) %>% 
#   mutate(lag4_carnivore = lag(carnivore,4),
#          lag4_piscovore = lag(piscivore,4)) %>% 
#   na.omit() %>% 
#   ungroup() %>% 
#   filter(year > 1999) 
# dat %>%
#   ggplot(aes(year, density, color = broadtrophic)) +
#   geom_line()

dat <- cip_data %>%
  group_by(broadtrophic, region, site, side, transect, year, eventual_mpa) %>%
  summarise(density = sum(density_g_m2)) %>%
  group_by(broadtrophic, region, site, year) %>%
  summarise(mean_density = mean(density, na.rm = T)) %>%
  group_by(broadtrophic,region, year) %>%
  summarise(regional_density = mean(mean_density, na.rm = T)) %>% 
  ungroup() %>% 
  spread(broadtrophic, regional_density) %>% 
  group_by(region) %>% 
  arrange(year) %>% 
  mutate(lag4_carnivore = lag(carnivore,4),
         lag4_piscovore = lag(piscivore,4)) %>% 
  na.omit() %>% 
  ungroup() %>% 
  filter(year > 1999)

 dat <- dat %>%
  select(-lag4_carnivore,-lag4_piscovore)%>%
  group_by(region)%>%
  mutate_at(vars(carnivore:planktivore),funs(norm=(.-mean(.))/sd(.)))->dat

group_trends_plot <- dat %>%
  select(region,year,contains("norm"))%>%
  gather("group","density",-region,-year)%>%
  ggplot(aes(year,density,col=group))+
  geom_line()+
  geom_point() +
  scale_color_locuszoom(name="Trophic Group",labels=c("carnivore","herbivore","piscivore","planktivore"))+
  labs(x="year",y="Centered and Scaled Density")+
  facet_wrap(~region)

group_trends_plot
```

```{r embed-plot, fig.cap = "Predictive skill as a function of embedding dimensions"}

## split data by region to analyze timeseries from each island separately
ana.dat <- dat %>% filter(region=="ANA")
sci.dat <- dat %>% filter(region=="SCI")
smi.dat <- dat %>% filter(region=="SMI")
sri.dat <- dat %>% filter(region=="SRI")

datnest <- dat %>% group_by(region) %>% nest()

datnorm <- dat %>% select(region,year,contains('norm')) %>% ungroup()

# have to record the segments corresponding to each "replicate" so simplex algorithm does not try to make predictions crossing time barriers
segs <- datnorm %>% mutate(ind=row_number()) %>% group_by(region) %>% summarise(first=first(ind),last=last(ind)) %>%
  select(-region)

var_names <- c("carnivore_norm","herbivore_norm","piscivore_norm","planktivore_norm")

regions.combined.simp.list <- map(var_names,function(x){
  temp <- datnorm %>% ungroup() %>% select(matches(x)) %>% as.data.frame()
  out <- simplex(as.numeric(temp[,1]),E=1:10,lib=as.matrix(segs),silent=T) %>%
    mutate(trophic=str_replace(x,"_norm",""))
  out
})

embed_plot <- bind_rows(regions.combined.simp.list) %>%
  ggplot(aes(E,rho,color=trophic))+
  geom_line(size=2)+
  facet_wrap(~trophic,nrow=2,scales="free_y")+
  geom_hline(yintercept = 0,color="black")+
  labs(x="Embedding Dimension (E)",y=expression(paste("Predictive Skill, ",rho)))+
  scale_x_continuous(breaks=seq(0,12,by=2))+
  scale_color_locuszoom()+
  guides(color=F)

embed_plot

```

Focusing on just these three groups then (removing herbivores), we can test for causal relationships between groups using convergent cross mapping and the logic of Takens' theorem of dynamic systems. Generalizations of Takens' theorem indicate that if two variables (in our case, species or physical variables) are part of the same dynamic system, their individual dynamics should reflect their relative causal influence [@sugihara2012]. In other words, if one variable is causally forced by another, that forcing should leave a signature on the first time series. Convergent cross mapping (CCM) tests for causation by using the attractor/manifold built from the time series of one variable to predict another (hence the "cross-mapping"). In simple terms, the *causal effect of A on B is determined by how well B cross-maps A*.

There are two criteria for CCM to establish causality:  First, and most obviously, predictive cross-map skill using all available data should be significantly greater than zero. Second, that predictability should be convergent.  Convergence means that cross-mapped estimates improve with library length (the number of state-space vectors used to build the attractor), because the attractor is more fully resolved and therefore estimation error should decline. Convergence is key to distinguishing causation from simple or spurious correlation. If two variables are spuriously correlated and not causally linked, CCM should fail to satisfy this second criterion. Based on these criteria, there is little evidence of significant dynamic interactions between trophic groups (Fig.S\@ref(fig:cross-map-cp)-\@ref(fig:cross-map-pis)). Cross-mapping produced positive predictive skill, but was non-convergent for all cross-mappings with the exception of carnivores cross-mapping planktivores (providing some evidence that planktivore dynamics may be a driver of carnivore dynamics). This analysis provides evidence that trophic cascades are unlikely to be a significant driver of our results. It is important to note though that this analysis does not mean that trophic cascades could not emerge in this system, rather that we do not detect them with these data at this time.

```{r cross-map-fxn}
# plots a cross mapping between two variables in the data, given their variable names (grp), their optimal embedding dimensions (from simplex forecasting), and their preferred names for plotting
xmap <- function(grp1,grp2,E1,E2,name1,name2){
    # cross-map group 1 to group 2

  ccm1 <- ccm(datnorm,lib=as.matrix(segs),pred=as.matrix(segs),E=E1,lib_column= grp1,target_column = grp2,lib_sizes = c(10,25,50,75),num_samples=100,replace=T,silent=T,RNGseed = 41389)

  # test for convergence (is predictability at max library size significantly greater than at minimum library size?)

  rhomeans1 <- ccm1 %>% ccm_means()

  conv_test1 <- t.test(ccm1$rho[ccm1$lib_size==10],ccm1$rho[ccm1$lib_size==75])$p.value
  if((rhomeans1$rho[rhomeans1$lib_size==75]-rhomeans1$rho[rhomeans1$lib_size==10])<0.05) conv_test1 <- 1
  conv_test1 <- ifelse(conv_test1<0.05,"Convergent","Non-convergent")

  xmap1 <- ccm1 %>%
    group_by(lib_size)%>%
    summarise(rhomean=mean(rho,na.rm=T),upper=quantile(rho, 0.975),lower=quantile(rho, 0.025))%>%
    ungroup()%>%
    ggplot(aes(lib_size,rhomean))+
    geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.3,fill="red")+
    geom_line(color="darkorchid3")+
    geom_hline(aes(yintercept = 0), linetype = 2) +
    labs(x="Library Size",y=expression(paste(rho, " (predictive skill)")),title=paste0(name1," xmap ",name2,"\n","(",conv_test1,")"))+
    theme(plot.title = element_text(size=6))

  # cross map group 2 to group 1
  # inspect the output of simplex from the previous step and use the best embedding dimension (highest rho) for the carnivore time series

  ccm2 <- ccm(datnorm,lib=as.matrix(segs),pred=as.matrix(segs),E=E2,lib_column= grp2,target_column = grp1,lib_sizes = c(10,25,50,75),num_samples=100,replace=T,silent=T,RNGseed = 41389)

  rhomeans2 <- ccm2 %>% ccm_means()

  conv_test2 <- t.test(ccm2$rho[ccm2$lib_size==10],ccm2$rho[ccm2$lib_size==75])$p.value
  if((rhomeans2$rho[rhomeans2$lib_size==75]-rhomeans2$rho[rhomeans2$lib_size==10])<0.05) conv_test2 <- 1
  conv_test2 <- ifelse(conv_test2<0.05,"Convergent","Non-convergent")

  xmap2 <- ccm2 %>%
    group_by(lib_size)%>%
    summarise(rhomean=mean(rho,na.rm=T),upper=quantile(rho, 0.975),lower=quantile(rho, 0.025))%>%
    ungroup()%>%
    ggplot(aes(lib_size,rhomean))+
    geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.3,fill="red")+
    geom_line(color="darkorchid3")+
    geom_hline(aes(yintercept = 0), linetype = 2) +
    labs(x="Library Size",y=expression(paste(rho, " (predictive skill)")),title=paste0(name2," xmap ",name1,"\n","(",conv_test2,")"))+
    theme(plot.title = element_text(size=6))

  cross_map_plot <- xmap1 + xmap2

cross_map_plot
}
```

```{r cross-map-cp, fig.cap = "Cross mapping of effect of piscivores on carnivores (A) and carnivores on piscivores (B) in the PISCO data from 2000 to 2017. Shaded region show 95% confidence interval"}

car_pis<-xmap(grp1="carnivore_norm",grp2="piscivore_norm",E1=4,E2=4,name1="Carnivores",name2="Piscivores")
car_pl <- xmap(grp1="carnivore_norm",grp2="planktivore_norm",E1=4,E2=4,name1="Carnivores",name2="Planktivores")
pl_pis <- xmap(grp1="planktivore_norm",grp2="piscivore_norm",E1=4,E2=4,name1="Planktivores",name2="Piscivores")

car_pis
```

```{r cross-map-pl, fig.cap = "Cross mapping of effect of planktivores on carnivores (A) and carnivores on planktivores (B) in the PISCO data from 2000 to 2017. Shaded region show 95% confidence interval"}

car_pl

```

```{r cross-map-pis, fig.cap = "Cross mapping of effect of piscivores on carnivores (A) and carnivores on planktivores (B) in the PISCO data from 2000 to 2017. Shaded region show 95% confidence interval"}

 pl_pis 
```

```{r trophic-cor-plot, fig.cap="Estimated correlation coefficients between broad trophic groups at the island by year resolution", eval = FALSE}

trophic_effects_plot <- cip_data %>%
  group_by(broadtrophic, region, site, side, transect, year, eventual_mpa) %>%
  summarise(density = sum(density_g_m2)) %>%
  group_by(broadtrophic, region, site, year) %>%
  summarise(mean_density = mean(density, na.rm = T)) %>%
  group_by(broadtrophic,region, year) %>%
  summarise(regional_density = mean(mean_density, na.rm = T)) %>% 
  ungroup() %>% 
  spread(broadtrophic, regional_density) %>% 
  group_by(region) %>% 
  arrange(year) %>% 
  mutate(lag4_carnivore = lag(carnivore,4),
         lag4_piscovore = lag(piscivore,4)) %>% 
  na.omit() %>% 
  ungroup() %>% 
  filter(year > 1999) %>% 
  select(-year,-region) %>% 
  corrr::correlate() %>% 
   gather("colname","correlation",-rowname) %>% 
  ggplot(aes(rowname, colname,fill = correlation)) + 
  geom_tile() + 
  geom_text(aes(rowname, colname, label = round(correlation,2))) + 
  scale_fill_gradient2(low = "tomato", mid = "white", high = "steelblue", midpoint = 0, limits = c(-1,1),
                       guide = guide_colorbar(frame.colour = "black",barheight = 13)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1,
                                   hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

trophic_effects_plot

```

## Repeat Analysis with Kelp Forest Monitoring Data

As a robustness check to our main results, we repeated our analysis utilizing data provided by the [Kelp Forest Monitoring Program (KFM)](https://science.nature.nps.gov/im/units/medn/monitor/kelpforest.cfm) conducted in the Channel Islands. Despite having similar but different methods and survey locations, we find almost identical estimated trends in divergences between targeted and non-targeted species using the KFM data (Fig.S\@ref(fig:kfm-did)).

```{r kfm-did,fig.cap = "Estimated divergence in biomass densities of targeted and non-targeted fishes in the Channel Islands (i.e. integrated across inside and outside of MPAs) using the KFM data . MPAs are implemented in 2003 (red dashed line). Estimates are from a regression on log(abundance index), and so estimated effects roughly correspond to percentage changes"}

kfm_did_plot <- kfm_run$did_plot[[1]] + 
  scale_y_continuous(name = "Divergence from Non-Targeted", labels = percent) + 
  labs(caption = "")

kfm_did_plot

```

## MPA Only Analysis

Given trends in mean densities observed in the raw data, the "regional conservation effect" estimated by our model---defined as the divergence in trends between the targeted and non-targeted species across the Channel Islands region---is not surprising. By jumping through countless statistical hoops we reach a similar conclusion that we would just by looking at the divergences in the mean trends. The integration of data from inside and outside of MPAs is a possible explanation for this lack of a clear regional effect. If spillover is limited or has simply not developed yet, especially relative to the effect of fishing outside of MPAs, then it is possible that there is a clear positive effect inside the MPAs, a clear negative effect outside, and when we look across both types of sites we get an unclear average of the two.

To address this, we can first repeat some exploratory data analysis of trends in densities inside and outside the MPAs for targeted and non-targeted species. @caselle2015 provides a thorough look at this question of differences inside and outside of MPAs. We update that analysis here to account for our specific questions of trend divergence, potential differences in filtering methods, to include data up through 2017, and to utilize our estimation method on just the inside-MPA data. For all exploratory analyses, we consider the same top `r length(top_species)` consistently observed species. Looking first at simple trends in total mean biomass density across these species inside and outside of MPAs, we find evidence that biomass densities inside the MPAs is increasing faster (and is higher inside) than outside (Fig.S\@ref(fig:bio-mpa)).

```{r bio-mpa, fig.cap= "Annual mean aggregate biomass density (summed across all fishes) inside and outside of eventual MPA locations over time. Red dashed line indicates MPA implementation year"}

abundance_index_plot <- cip_data %>% 
  group_by(targeted, region,site,side, transect, year, eventual_mpa, classcode) %>% 
  summarise(density = sum(density_g_m2)) %>% 
  group_by(targeted,classcode, region,site, year, eventual_mpa) %>% 
  summarise(mean_density = mean(density, na.rm = T)) %>% 
  group_by(targeted, classcode, year, eventual_mpa) %>% 
  summarise(mmd = mean(mean_density, na.rm = T)) %>% 
  group_by(year, eventual_mpa) %>% 
  summarise(abundance_index = sum(mmd, na.rm = T)) %>% 
  ggplot(aes(year, abundance_index, color = eventual_mpa)) + 
  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
  geom_line(size = 2) + 
  scale_color_manual(values = c("#009ACD", "#548B54"), name = "" , labels = c("Non-MPA",'MPA')) + 
  scale_y_continuous(name = "Total Abundance Index")+
  labs(x="Year")

abundance_index_plot

```

Our proposed identification strategy here though is not that total biomass density should be different inside and outside, but that the non-targeted species should serve as the control to the targeted. If we believe that the MPA effects are greater inside the MPA, then we would expect to see stronger divergences in biomass densities between these two targeted and non-targeted fishes inside the MPAs than outside

```{r div-mpa, fig.cap="Trends in total biomass density inside and outside of eventual MPAs for targeted and non-targeted fishes. Red dashed line indicates MPA implementation year"}

facet_labels <- c(
  `TRUE` = "MPA",
  `FALSE` = "Non-MPA"
)

mpa_abundance_plot <- cip_data %>% 
  mutate(targeted = targeted == 1) %>% 
  group_by(targeted, region,site,side, transect, year, eventual_mpa, classcode) %>% 
  summarise(density = sum(density_g_m2)) %>% 
  group_by(targeted,classcode, region,site, year, eventual_mpa) %>% 
  summarise(mean_density = mean(density, na.rm = T)) %>% 
  group_by(targeted, classcode, year, eventual_mpa) %>% 
  summarise(mmd = mean(mean_density, na.rm = T)) %>% 
  group_by(year, targeted, eventual_mpa) %>% 
  summarise(abundance_index = sum(mmd, na.rm = T)) %>% 
  ungroup() %>% 
  group_by(targeted, eventual_mpa) %>% 
  mutate(abundance_index = (abundance_index - mean(abundance_index)) / sd(abundance_index)) %>% 
  ungroup() %>% 
  rename(`Eventual MPA?` = eventual_mpa) %>% 
  ggplot(aes(year, abundance_index, color = targeted)) + 
  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
  geom_line(size = 2) + 
  scale_color_manual(values = c("steelblue", "tomato") ,name = "",
                     labels = c("Non-Targeted","Targeted")) + 
  facet_wrap(~`Eventual MPA?`, labeller = labeller(`Eventual MPA?` = facet_labels)) + 
    scale_y_continuous(name = "Total Abundance Index")+
  labs(x='Year')

mpa_abundance_plot

```

Here we see a different picture. While there is some visual evidence that the targeted species were diverging from the non-targeted faster inside the MPAs than outside, both inside and outside we see that the trend in total biomass density of targeted species is trending downward, relative to the trend in the non-targeted species in recent years. This analysis is of total biomass density. However, our model estimates the mean difference in targeted and non-targeted species. Both have their advantages, but we chose the mean to reflect a hypothesis that the MPAs would provide positive benefits across all targeted species. The total biomass density could be strongly affected by a sharp increase or decrease in one or two species, even if the mean trend is different. Examining the mean trends though, we see the same results (Fig.S\@ref(fig:mean-div-mpa)).

```{r mean-div-mpa, fig.cap = "Trends in mean total biomass density inside and outside of eventual MPAs for targeted and non-targeted fishes. Red dashed line indicates MPA implementation year"}

targeted_mpa_plot <- cip_data %>% 
  mutate(targeted = targeted == 1) %>% 
  group_by(targeted, region,site,side, transect, year, eventual_mpa, classcode) %>% 
  summarise(density = sum(density_g_m2)) %>% 
  group_by(targeted,classcode, region,site, year, eventual_mpa) %>% 
  summarise(mean_density = mean(density, na.rm = T)) %>% 
  group_by(targeted, classcode, year, eventual_mpa) %>% 
  summarise(mmd = mean(mean_density, na.rm = T)) %>% 
  group_by(year, targeted, eventual_mpa) %>% 
  summarise(abundance_index = mean(mmd, na.rm = T)) %>% 
  ungroup() %>% 
    group_by(targeted, eventual_mpa) %>% 
  mutate(abundance_index = (abundance_index - mean(abundance_index)) / sd(abundance_index)) %>% 
  ungroup() %>% 
  rename(`Eventual MPA?` = eventual_mpa) %>% 
  ggplot(aes(year, abundance_index, color = targeted)) + 
  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
  geom_line(size = 2) + 
  scale_color_manual(values = c("steelblue", "tomato") ,name = "Targeted?") + 
  facet_wrap(~`Eventual MPA?`, labeller = label_both) + 
    scale_y_continuous(name = "Mean Abundance Index")+
  labs(x="Year")

targeted_mpa_plot
```

Lastly, we can examine both the mean and individual trend to check clear species-by-species outliers in the overall biomass density trends. This analysis shows noise, but overall the targeted and non-targeted species seem to be following similar trends within their respective groups

```{r cs-trends, fig.cap = "Centered and scaled biomass density trends for each fish grouped by targeted and non targeted (pale lines) and fitted LOESS smoother (with 95% confidence intervals around mean) and mean by targeted and non-targeted groups, inside and outside od MPAs. Red dashed line indicates MPA implementation year"}

species_mpa_trend_plot <- cip_data %>% 
  group_by(year, classcode, eventual_mpa, targeted) %>% 
  summarise(mean_density = mean(density_g_m2, na.rm = T)) %>% 
  ungroup() %>% 
  mutate(targeted = targeted == 1) %>% 
  group_by(classcode, eventual_mpa) %>% 
  mutate(cs_density = (mean_density - mean(mean_density)) / sd(mean_density)) %>% 
  ungroup() %>% 
    rename(`Eventual MPA?` = eventual_mpa) %>% 
  ggplot(aes(year, cs_density, color = targeted, fill = targeted)) + 
    geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
  geom_line(aes(group = interaction(classcode, targeted)),alpha = 0.5) + 
  geom_smooth() +
    scale_color_manual(values = c("steelblue", "tomato") ,name = "", labels = c("Non-Targeted","Targeted")) +
      scale_fill_manual(values = c("steelblue", "tomato") ,name = "",labels = c("Non-Targeted","Targeted")) +

    facet_wrap(~`Eventual MPA?`, labeller = labeller(`Eventual MPA?` = facet_labels)) + 
    labs(y = "Centered and Scaled Mean Density", x = "Year")

species_mpa_trend_plot

```

These visual assessments suggest that similar to our results looking both inside and outside of MPAs, we would expect that our estimation model fitted only on data from inside eventual MPAs would reach similar conclusions as our results fitted to data from both inside and outside MPAs. To test this, we re-ran our analysis, but only using data from sites that are eventually placed inside MPAs. Our results reflect the same trends as displayed in the raw data and the statistical region-wide analysis, providing robust statistical support to the conclusions we would reach from visually examining the raw data (Fig.S \@ref(fig:mpa-did)).

```{r mpa-did,fig.cap = "Estimated divergence in biomass densities of targeted and non-targeted fishes inside eventual Channel Islands MPAs. MPAs are implement in 2003 (red dashed line). Estimates are from a regression on log(abundance index), and so estimated effects roughly correspond to percentage changes" }

mpa_did_plot <- mpa_run$did_plot[[1]] + 
       labs(x = "Year", y = "Estimate of Regional Effect",
          caption = "")

mpa_did_plot

```

## References

### 2019-10-09 documents/ovando-regional-effects-of-mpas-si.tex
Initial commit of cleaned up regional effects repo

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Supporting Information: The Regional Effects of Marine Protected Areas},
            pdfauthor={Daniel Ovando,Jennifer E. Caselle,Christopher Costello,Olivier Deschenes,Steven D. Gaines,Ray Hilborn,Owen Liu},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Supporting Information: The Regional Effects of Marine Protected Areas}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Daniel Ovando,Jennifer E. Caselle,Christopher Costello,Olivier Deschenes,Steven D. Gaines,Ray Hilborn,Owen Liu}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

\usepackage{longtable}
\usepackage{booktabs}
\usepackage{eso-pic}
\usepackage{graphicx}
\usepackage[left]{lineno}
\linenumbers

\begin{document}
\maketitle

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\hypertarget{supporting-information-si}{%
\section{Supporting Information (SI)}\label{supporting-information-si}}

\hypertarget{si-text}{%
\section*{SI Text}\label{si-text}}
\addcontentsline{toc}{section}{SI Text}

\hypertarget{computing-environment}{%
\subsection{Computing environment}\label{computing-environment}}

The computing environment needed to reproduce the results presented here is stored in a Docker container available \href{}{here}XX. Interested users can load the container on their system to reproduce the complete analysis and manuscript. Analysis were performed in R version 3.6.1 (2019-07-05). Package versions are shown in Table.S\ref{tab:pkg-info}.

\begin{longtable}[t]{lllll}
\caption{\label{tab:pkg-info}Package versions and sources used in this paper}\\
\toprule
  & package & loadedversion & date & source\\
\midrule
\endfirsthead
\caption[]{\label{tab:pkg-info}Package versions and sources used in this paper \textit{(continued)}}\\
\toprule
  & package & loadedversion & date & source\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
assertthat & assertthat & 0.2.1 & 2019-03-21 & CRAN (R 3.6.0)\\
backports & backports & 1.1.5 & 2019-10-02 & CRAN (R 3.6.1)\\
bitops & bitops & 1.0-6 & 2013-08-17 & CRAN (R 3.6.0)\\
bookdown & bookdown & 0.13 & 2019-08-21 & CRAN (R 3.6.0)\\
broom & broom & 0.5.2 & 2019-04-07 & CRAN (R 3.6.0)\\
\addlinespace
callr & callr & 3.3.2 & 2019-09-22 & CRAN (R 3.6.0)\\
caret & caret & 6.0-84 & 2019-04-27 & CRAN (R 3.6.0)\\
cellranger & cellranger & 1.1.0 & 2016-07-27 & CRAN (R 3.6.0)\\
class & class & 7.3-15 & 2019-01-01 & CRAN (R 3.6.1)\\
classInt & classInt & 0.4-1 & 2019-08-06 & CRAN (R 3.6.0)\\
\addlinespace
cli & cli & 1.1.0 & 2019-03-19 & CRAN (R 3.6.0)\\
codetools & codetools & 0.2-16 & 2018-12-24 & CRAN (R 3.6.1)\\
colorspace & colorspace & 1.4-1 & 2019-03-18 & CRAN (R 3.6.0)\\
crayon & crayon & 1.3.4 & 2017-09-16 & CRAN (R 3.6.0)\\
crul & crul & 0.8.4 & 2019-08-02 & CRAN (R 3.6.0)\\
\addlinespace
curl & curl & 4.2 & 2019-09-24 & CRAN (R 3.6.0)\\
data.table & data.table & 1.12.4 & 2019-10-03 & CRAN (R 3.6.1)\\
DBI & DBI & 1.0.0 & 2018-05-02 & CRAN (R 3.6.0)\\
desc & desc & 1.2.0 & 2018-05-01 & CRAN (R 3.6.0)\\
devtools & devtools & 2.2.0 & 2019-09-07 & CRAN (R 3.6.0)\\
\addlinespace
digest & digest & 0.6.21 & 2019-09-20 & CRAN (R 3.6.0)\\
dplyr & dplyr & 0.8.3 & 2019-07-04 & CRAN (R 3.6.0)\\
DT & DT & 0.9 & 2019-09-17 & CRAN (R 3.6.1)\\
e1071 & e1071 & 1.7-2 & 2019-06-05 & CRAN (R 3.6.0)\\
ellipsis & ellipsis & 0.3.0 & 2019-09-20 & CRAN (R 3.6.0)\\
\addlinespace
evaluate & evaluate & 0.14 & 2019-05-28 & CRAN (R 3.6.0)\\
extrafont & extrafont & 0.17 & 2014-12-08 & CRAN (R 3.6.0)\\
extrafontdb & extrafontdb & 1.0 & 2012-06-11 & CRAN (R 3.6.0)\\
forcats & forcats & 0.4.0 & 2019-02-17 & CRAN (R 3.6.0)\\
foreach & foreach & 1.4.7 & 2019-07-27 & CRAN (R 3.6.0)\\
\addlinespace
foreign & foreign & 0.8-72 & 2019-08-02 & CRAN (R 3.6.0)\\
fs & fs & 1.3.1 & 2019-05-06 & CRAN (R 3.6.0)\\
gdtools & gdtools & 0.2.0 & 2019-09-03 & CRAN (R 3.6.0)\\
generics & generics & 0.0.2 & 2018-11-29 & CRAN (R 3.6.0)\\
geojson & geojson & 0.3.2 & 2019-01-31 & CRAN (R 3.6.0)\\
\addlinespace
geojsonio & geojsonio & 0.7.0 & 2019-04-25 & CRAN (R 3.6.0)\\
geojsonlint & geojsonlint & 0.3.0 & 2019-02-08 & CRAN (R 3.6.0)\\
ggmap & ggmap & 3.0.0.901 & 2019-10-05 & Github (dkahle/ggmap@37a8672)\\
ggplot2 & ggplot2 & 3.2.1 & 2019-08-10 & CRAN (R 3.6.0)\\
ggsci & ggsci & 2.9 & 2018-05-14 & CRAN (R 3.6.0)\\
\addlinespace
ggspatial & ggspatial & 1.0.3 & 2018-12-14 & CRAN (R 3.6.0)\\
glue & glue & 1.3.1 & 2019-03-12 & CRAN (R 3.6.0)\\
gower & gower & 0.2.1 & 2019-05-14 & CRAN (R 3.6.0)\\
gridExtra & gridExtra & 2.3 & 2017-09-09 & CRAN (R 3.6.0)\\
gtable & gtable & 0.3.0 & 2019-03-25 & CRAN (R 3.6.0)\\
\addlinespace
haven & haven & 2.1.1 & 2019-07-04 & CRAN (R 3.6.0)\\
here & here & 0.1 & 2017-05-28 & CRAN (R 3.6.0)\\
hms & hms & 0.5.1 & 2019-08-23 & CRAN (R 3.6.0)\\
hrbrthemes & hrbrthemes & 0.6.0 & 2019-01-21 & CRAN (R 3.6.0)\\
htmltools & htmltools & 0.4.0 & 2019-10-04 & CRAN (R 3.6.1)\\
\addlinespace
htmlwidgets & htmlwidgets & 1.3 & 2018-09-30 & CRAN (R 3.6.0)\\
httpcode & httpcode & 0.2.0 & 2016-11-14 & CRAN (R 3.6.0)\\
httr & httr & 1.4.1 & 2019-08-05 & CRAN (R 3.6.1)\\
ipred & ipred & 0.9-9 & 2019-04-28 & CRAN (R 3.6.0)\\
iterators & iterators & 1.0.12 & 2019-07-26 & CRAN (R 3.6.0)\\
\addlinespace
jpeg & jpeg & 0.1-8 & 2014-01-23 & CRAN (R 3.6.0)\\
jqr & jqr & 1.1.0 & 2018-10-22 & CRAN (R 3.6.0)\\
jsonlite & jsonlite & 1.6 & 2018-12-07 & CRAN (R 3.6.0)\\
jsonvalidate & jsonvalidate & 1.1.0 & 2019-06-25 & CRAN (R 3.6.0)\\
kernlab & kernlab & 0.9-27 & 2018-08-10 & CRAN (R 3.6.0)\\
\addlinespace
KernSmooth & KernSmooth & 2.23-15 & 2015-06-29 & CRAN (R 3.6.1)\\
knitr & knitr & 1.25 & 2019-09-18 & CRAN (R 3.6.1)\\
labeling & labeling & 0.3 & 2014-08-23 & CRAN (R 3.6.0)\\
lattice & lattice & 0.20-38 & 2018-11-04 & CRAN (R 3.6.1)\\
lava & lava & 1.6.6 & 2019-08-01 & CRAN (R 3.6.0)\\
\addlinespace
lazyeval & lazyeval & 0.2.2 & 2019-03-15 & CRAN (R 3.6.0)\\
lifecycle & lifecycle & 0.1.0 & 2019-08-01 & CRAN (R 3.6.0)\\
lubridate & lubridate & 1.7.4 & 2018-04-11 & CRAN (R 3.6.0)\\
magrittr & magrittr & 1.5 & 2014-11-22 & CRAN (R 3.6.0)\\
maptools & maptools & 0.9-5 & 2019-02-18 & CRAN (R 3.6.0)\\
\addlinespace
MASS & MASS & 7.3-51.4 & 2019-03-31 & CRAN (R 3.6.1)\\
Matrix & Matrix & 1.2-17 & 2019-03-22 & CRAN (R 3.6.1)\\
memoise & memoise & 1.1.0 & 2017-04-21 & CRAN (R 3.6.0)\\
ModelMetrics & ModelMetrics & 1.2.2 & 2018-11-03 & CRAN (R 3.6.0)\\
modelr & modelr & 0.1.5 & 2019-08-08 & CRAN (R 3.6.0)\\
\addlinespace
munsell & munsell & 0.5.0 & 2018-06-12 & CRAN (R 3.6.0)\\
nlme & nlme & 3.1-141 & 2019-08-01 & CRAN (R 3.6.0)\\
nnet & nnet & 7.3-12 & 2016-02-02 & CRAN (R 3.6.1)\\
numDeriv & numDeriv & 2016.8-1.1 & 2019-06-06 & CRAN (R 3.6.0)\\
optimx & optimx & 2018-7.10 & 2018-09-30 & CRAN (R 3.6.0)\\
\addlinespace
packrat & packrat & 0.5.0 & 2018-11-14 & CRAN (R 3.6.0)\\
patchwork & patchwork & 0.0.1 & 2019-10-03 & Github (thomasp85/patchwork@36b4918)\\
pillar & pillar & 1.4.2 & 2019-06-29 & CRAN (R 3.6.0)\\
pkgbuild & pkgbuild & 1.0.5 & 2019-08-26 & CRAN (R 3.6.0)\\
pkgconfig & pkgconfig & 2.0.3 & 2019-09-22 & CRAN (R 3.6.0)\\
\addlinespace
pkgload & pkgload & 1.0.2 & 2018-10-29 & CRAN (R 3.6.0)\\
plyr & plyr & 1.8.4 & 2016-06-08 & CRAN (R 3.6.0)\\
png & png & 0.1-7 & 2013-12-03 & CRAN (R 3.6.0)\\
prettyunits & prettyunits & 1.0.2 & 2015-07-13 & CRAN (R 3.6.0)\\
processx & processx & 3.4.1 & 2019-07-18 & CRAN (R 3.6.0)\\
\addlinespace
prodlim & prodlim & 2018.04.18 & 2018-04-18 & CRAN (R 3.6.0)\\
ps & ps & 1.3.0 & 2018-12-21 & CRAN (R 3.6.0)\\
purrr & purrr & 0.3.2 & 2019-03-15 & CRAN (R 3.6.0)\\
R6 & R6 & 2.4.0 & 2019-02-14 & CRAN (R 3.6.0)\\
Rcpp & Rcpp & 1.0.2 & 2019-07-25 & CRAN (R 3.6.0)\\
\addlinespace
readr & readr & 1.3.1 & 2018-12-21 & CRAN (R 3.6.0)\\
readxl & readxl & 1.3.1 & 2019-03-13 & CRAN (R 3.6.0)\\
recipes & recipes & 0.1.7 & 2019-09-15 & CRAN (R 3.6.0)\\
rEDM & rEDM & 0.7.4 & 2019-08-19 & Github (ha0ye/rEDM@88554a4)\\
remotes & remotes & 2.1.0 & 2019-06-24 & CRAN (R 3.6.0)\\
\addlinespace
reshape2 & reshape2 & 1.4.3 & 2017-12-11 & CRAN (R 3.6.0)\\
rgdal & rgdal & 1.4-4 & 2019-05-29 & CRAN (R 3.6.0)\\
rgeos & rgeos & 0.5-1 & 2019-08-05 & CRAN (R 3.6.1)\\
RgoogleMaps & RgoogleMaps & 1.4.4 & 2019-08-20 & CRAN (R 3.6.0)\\
rjson & rjson & 0.2.20 & 2018-06-08 & CRAN (R 3.6.0)\\
\addlinespace
rlang & rlang & 0.4.0 & 2019-06-25 & CRAN (R 3.6.0)\\
rmapshaper & rmapshaper & 0.4.1 & 2018-10-16 & CRAN (R 3.6.0)\\
rmarkdown & rmarkdown & 1.16 & 2019-10-01 & CRAN (R 3.6.1)\\
rpart & rpart & 4.1-15 & 2019-04-12 & CRAN (R 3.6.1)\\
rprojroot & rprojroot & 1.3-2 & 2018-01-03 & CRAN (R 3.6.0)\\
\addlinespace
rstudioapi & rstudioapi & 0.10 & 2019-03-19 & CRAN (R 3.6.0)\\
rticles & rticles & 0.10 & 2019-08-21 & CRAN (R 3.6.0)\\
Rttf2pt1 & Rttf2pt1 & 1.3.7 & 2018-06-29 & CRAN (R 3.6.0)\\
rvest & rvest & 0.3.4 & 2019-05-15 & CRAN (R 3.6.0)\\
scales & scales & 1.0.0 & 2018-08-09 & CRAN (R 3.6.0)\\
\addlinespace
sessioninfo & sessioninfo & 1.1.1 & 2018-11-05 & CRAN (R 3.6.0)\\
sf & sf & 0.8-0 & 2019-09-17 & CRAN (R 3.6.1)\\
sp & sp & 1.3-1 & 2018-06-05 & CRAN (R 3.6.0)\\
spasm & spasm & 0.1.0.9000 & 2019-10-09 & local\\
stringi & stringi & 1.4.3 & 2019-03-12 & CRAN (R 3.6.0)\\
\addlinespace
stringr & stringr & 1.4.0 & 2019-02-10 & CRAN (R 3.6.0)\\
survival & survival & 2.44-1.1 & 2019-04-01 & CRAN (R 3.6.1)\\
Synth & Synth & 1.1-5 & 2014-01-27 & CRAN (R 3.6.0)\\
systemfonts & systemfonts & 0.1.1 & 2019-07-01 & CRAN (R 3.6.0)\\
testthat & testthat & 2.2.1 & 2019-07-25 & CRAN (R 3.6.0)\\
\addlinespace
tibble & tibble & 2.1.3 & 2019-06-06 & CRAN (R 3.6.0)\\
tidyr & tidyr & 1.0.0 & 2019-09-11 & CRAN (R 3.6.0)\\
tidyselect & tidyselect & 0.2.5 & 2018-10-11 & CRAN (R 3.6.0)\\
tidyverse & tidyverse & 1.2.1 & 2017-11-14 & CRAN (R 3.6.0)\\
timeDate & timeDate & 3043.102 & 2018-02-21 & CRAN (R 3.6.0)\\
\addlinespace
tinytex & tinytex & 0.16 & 2019-09-17 & CRAN (R 3.6.1)\\
units & units & 0.6-4 & 2019-08-22 & CRAN (R 3.6.0)\\
usethis & usethis & 1.5.1 & 2019-07-04 & CRAN (R 3.6.0)\\
V8 & V8 & 2.3 & 2019-07-02 & CRAN (R 3.6.0)\\
vctrs & vctrs & 0.2.0 & 2019-07-05 & CRAN (R 3.6.0)\\
\addlinespace
viridis & viridis & 0.5.1 & 2018-03-29 & CRAN (R 3.6.0)\\
viridisLite & viridisLite & 0.3.0 & 2018-02-01 & CRAN (R 3.6.0)\\
withr & withr & 2.1.2 & 2018-03-15 & CRAN (R 3.6.0)\\
xfun & xfun & 0.10 & 2019-10-01 & CRAN (R 3.6.1)\\
xml2 & xml2 & 1.2.2 & 2019-08-09 & CRAN (R 3.6.0)\\
\addlinespace
yaml & yaml & 2.2.0 & 2018-07-25 & CRAN (R 3.6.0)\\
zeallot & zeallot & 0.1.0 & 2018-01-28 & CRAN (R 3.6.0)\\*
\end{longtable}

\hypertarget{operating-model}{%
\subsection{Operating Model}\label{operating-model}}

The operating model is a spatial single-species age-structured bio-economic model. The operating model itself is organized as an R package, which can be found and installed \href{https://github.com/DanOvando/spasm}{here}. Users can explore the functionality of the operating mode through an interactive web application \href{https://danovando.shinyapps.io/simmpa/}{here}.

For the population model, numbers \emph{n} at time \emph{t} for age \emph{a} are given by

\begin{equation}
n_{t,a}=\begin{cases}
      = BH(ssb_{t-1}) & \text{if $a = 1$}\\
     = n_{t-1,a-1}e^{-(m + qE_{t-1}{\times}s_{a-1})}, & \text{if $1< a < max(age)$}\\
     =  n_{t-1,a}e^{-(m + qE_{t-1}{\times}s_a)} + n_{t-1,a-1}e^{-(m + qE_{t-1}{\times}s_{a-1})}, & \text{if $a = max(a)$}
  \end{cases}
  \label{eq:pop}
\end{equation}

where \emph{BH} is the Beverton-Holt recruitment function, \emph{ssb} is spawning-stock-biomass, \emph{m} is natural mortality, \emph{q} is catchability,\emph{E} is fishing effort at time \emph{t}, and \emph{s} is selectivity at age \emph{a}.

Selectivity is modeled through a logistic form per

\begin{equation}
s_a=\frac{1}{(1 + e^{-log(19)\times\frac{l_a - l_{sel}}{\delta_{sel}}})}
  \label{eq:sel}
\end{equation}

where \(l_a\) is the mean length at at age, \(l_sel\) is the length at which on average 50\% of individuals are selected by the fishery, and \(\delta_{sel}\) are the additional units of length at which on average 95\% of fish are selected by the fishery.

\emph{ssb} is calculated by converting age to mean length, calculating weight at age, maturity at age, and then calculating spawning stock biomass as the sum of spawning potential at age in a given time step.

\begin{equation}
  l_{a} = l_{\infty}\left(1 - e^{-k(a - a_0)}\right)
  \label{eq:len}
\end{equation}

Weight at age is then given by

\begin{equation}
  b_{a} = w_a \times l_{a}^{w_b}
  \label{eq:weight}
\end{equation}

and maturity \emph{mat} is calculated as

\begin{equation}
  \frac{1}{(1 + e^{-log(19)\times\frac{l_a - l_{mat}}{\delta_{mat}}})}
  \label{eq:mat}
\end{equation}

where \(l_{mat}\) is the length at which on average 50\% of individuals are sexual maturity, and \(\delta_{mat}\) is the units of length beyond \(l_{mat}\) at which on average 95\% of fish are sexually mature.

Spawning stock biomass at time \emph{t} is then calculated as

\begin{equation}
  ssb_t = \sum_{a=1}^Aw_{a,t}mat_{a,t}n_{a,t}
    \label{eq:ssb}
\end{equation}

\hypertarget{recruitment}{%
\subsubsection{Recruitment}\label{recruitment}}

Recruitment follows Beverton-Holt dynamics. We do however allow for three variants in the timing of density dependence:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Local density dependence: Density dependence occurs independently in each patch, and recruits then disperse to nearby patches
\end{enumerate}

\begin{equation}
  n_{t,a = 1,p} = \left(\frac{0.8{\times}r0_{p}\times{h}\times{ssb_{t-1,p}}}{0.2\times{ssb0_p}\times(1 - h)+(h - 0.2)\times{ssb_{t-1,p}}}\right)\times \boldsymbol{d^l} \times \epsilon_t
  \label{eq:dd1}
\end{equation}

where \textbf{d\textsuperscript{l}} is the larval movement matrix, \emph{h} is Beverton-Holt steepness (constrained between 0.6 and 0.99), \emph{r0} is unfished recruitment, and \emph{ssb0} is unfished spawning stock biomass.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  Global density dependence: Density dependence is a function of the sum of spawning biomass across all patches, and recruits are then distributed according to habitat quality
\end{enumerate}

\begin{equation}
n_{t,a = 1,p} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p}}\times{h}\times\sum_{p=1}^P{ssb_{t-1,p}}}{0.2\times{\sum_{p=1}^Pssb0_p}\times(1 - h)+(h - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p}}}\right)\times {hab_p} \times \epsilon_t
  \label{eq:dd2}
\end{equation}

where \emph{hab} is a vector of habitat quality by patch that sums to 1.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Post-dispersal density dependence: Larvae are distributed throughout the system, and then density dependence occurs based on the density of adult biomass at the destination patch.
\end{enumerate}

\begin{equation}
  larv_{t,p} = ssb_{t-1} \times\boldsymbol{d^l}
  \label{eq:larvmove}
\end{equation}

\begin{equation}
n_{t,a = 1,p} = \left(\frac{0.8{\times}r0_{p}\times{h}\times{larv_{t,p}}}{0.2\times{ssb0_p}\times(1 - h)+(h - 0.2)\times{larv_{t,p}}}\right) \times \epsilon_t
  \label{eq:dd3}
\end{equation}

\(\epsilon\) represents multiplicative recruitment deviates. Deviates are calculated as

\[\epsilon_t = e^{recdev_t}\]

`\(recdev\) are the log-normal recruitment deviates in time \emph{t}.

The stochastic component of the deviate is

\[\gamma_t \sim norm(-\sigma_r^2/2,\sigma_r)\]

and the final multiplicative recruitment deviate in time \emph{t} is then

\[recdev_t = \gamma_t\sqrt{1 - ac_r^2} + recdev_{t-1}ac_r \]

where \emph{ac} is the autocorrelation of the recruitment function (between 0 and 1).

\hypertarget{dispersal}{%
\subsubsection{Dispersal}\label{dispersal}}

Dispersal in the model is broken into two components: adult and larval. Both assume a Gaussian dispersal kernel of the form

\begin{equation}
m_{s,p_i,p_j}=\frac{1}{\sqrt{2\pi\sigma_{s,p_i}^2}}e^{-\frac{d_{p_i,p_j}^2}{2\sigma_{s,p_i}^2}}
  \label{eq:move}
\end{equation}

where \emph{i} is the source patch, \emph{j} is the destination patch, \emph{d} is the distance between patches \emph{i} and \emph{j} (where distance is measured with wrapped edges, such that if there are 50 patches, patch 1 and patch 50 have a distance of 1), and \(\sigma_s*\)is the movement rate, in units of patches, for life stage \emph{s} (adult or larval).

We allow the adult dispersal matrix to be affected by adult density dependence. The idea behind this is that adult fish will move more as densities increase, and become more sedentary as densities decrease (as habitat and food become more available for example). This allows us to simulate a scenario where as MPAs build up density they begin to export more adults to the surrounding waters, and if densities are lower in the fished areas these fish will actually become more sedentary.

Under these conditions, the adult movement rate is a linear function of depletion (measured as \(ssb/ssb_0\))

\begin{equation}
\sigma_{s=a,p}^* = max(slope\times{d _p+ \sigma_{s=a} \times dmod},0)
  \label{eq:ddmove1}
\end{equation}

where

\begin{equation}
slope = \sigma_{s=a} - ( \sigma_{s=a} \times dmod)
  \label{eq:ddmove2}
\end{equation}

Under these conditions, when depletion \emph{d = 1} (meaning the stock is unfished) the adult movement rate equals the max adult movement rate (\(\sigma_{s=a}^* = \sigma_{s=a}\)). When \(d=0\) \(\sigma_{s=a}^* = \sigma_{s=a} * dmod\). The greater \emph{dmod} is then, the more movement rates from a patch decline as density declines.

We also allow for a ``sprinkler'' condition in which MPAs are placed in locations that disperse larvae to a much wider area than non-MPA locations. In this world, we simply multiply \(\sigma_{s=l,p}\) by a sprinkler factor (by default 4) for any patch \emph{p} that would eventually become an MPA (whether or not MPAs are ever introduced). In other words, when we compare two scenarios, one with MPA and one without, the ``without'' scenario still has higher larval movement rates in patches that become MPAs in the ``with'' scenario.

\hypertarget{fleet-dynamics}{%
\subsubsection{Fleet Dynamics}\label{fleet-dynamics}}

We allow for three fleet models: constant effort, constant catch, and open-access. Constant effort means that total effort across all patches is equal in all time steps (unless MPAs force exit of effort as discussed below). Under constant catch, we set a target catch volume (in biomass, summed across all patches). Each time step, we calculate the fishing mortality rate that, given the fishable biomass in that time step, would produce the target catch. If there is insufficient fishable biomass available to support the target catch, we mark the population as crashed and stop the simulation (these crashed simulations are not included in the final analysis).

Under open-access, fishing effort expands in proportion to a weighted mean of profit-per-unit effort over the last \emph{t} time steps.

\begin{equation}
profit_{t} = price\times{catch_t} - cost\times{E_t}^2
  \label{eq:profit}
\end{equation}

From there, we determine the new effort as

\begin{equation}
E_t = E_{t-1} + \theta\times\sum_{i=t-1-l}^{t-1}w_i\frac{profit_{i}}{E^i}
  \label{eq:openaccess}
\end{equation}

where \emph{w} is a weighting function which is just a linear function of time

\begin{equation}
  w_i = \frac{i}{\sum_{i=1}^li}
  \label{eq:oaweight}
\end{equation}

and \emph{l} is the number of lagged time steps over which to calculate the weighted mean PPUE.

The open-access model can enter chaotic dynamics if the model parameters are not properly tuned. To address this, we first set price at 1, and set a \(\theta\) such that when profits are about as large as they might conceivably be the fishery doubles in size. We then estimate reference points for the simulated fishery (Bmsy, Fmsy, MSY), and set a target bionomic equilibrium B/Bmsy. Holding the other parameters constant, we thing find a cost coefficient that produces the desired bionomic equilibrium.

\hypertarget{spatial-fleet-distribution}{%
\subsubsection{Spatial Fleet Distribution}\label{spatial-fleet-distribution}}

Given a total amount of effort, we then need to distribute that effort in space. In the simplest form, effort is evenly distributed throughout the available patches.

\begin{equation}
E_{t,p} = E_t \times \frac{open_p}{\sum_{p=1}^Popen_p}
  \label{eq:simpleeffort}
\end{equation}

where \emph{open} indicates whether patch \emph{p} is open to fishing or not.

Effort can also be distributed according to spawning stock biomass in fishable patches

\begin{equation}
E_{t,p} = E_t \times \frac{open_{t,p}ssb_{t,p}}{\sum_{p=1}^Popen_{t,p}ssb_{t,p}}
  \label{eq:biomasseffort}
\end{equation}

And lastly effort can be distributed according to profit-per-unit-effort

\begin{equation}
E_{t,p} = E_t \times \frac{open_{t,p}ppue_{t,p}}{\sum_{p=1}^Popen_{t,p}ppue_{t,p}}
  \label{eq:ppueeffort}
\end{equation}

Under the constant effort or open access scenarios, effort can immediately respond to MPA placement in one of two ways. Effort can concentrate outside the MPAs (such that the sum of effort before and after MPA placement stays constant), or effort can leave the MPAs, such that the total effort in the fishery is reduced by the amount of effort that occurred inside the MPAs immediately before MPA placement. This is intended to simulate a scenario where fishers that used to use the MPA simply leave the fishery rather than redistribute outside the MPA, due for example to costs or lack of location specific knowledge to fish outside the MPA.

\hypertarget{mpa-design}{%
\subsubsection{MPA Design}\label{mpa-design}}

MPA design is relatively straightforward. We set a percentage of patches that are to be placed inside no-take MPAs. MPAs can either be placed continuously (e.g if there are 100 patches and 25\% are in MPAs, patches 1 to 25 are in MPAs) or randomly. If the MPAs are placed randomly, we can also set a minimum MPA size. This controls the patchiness of the MPAs. As the ``patchiness'' factor approaches zero, the behavior equals that of random placement. As it approaches 1, the behavior approaches that of continuous placement. In between, the greater the patchiness, the more clustered together MPAs become.

\hypertarget{simulations}{%
\subsection{Simulations}\label{simulations}}

We use this our operating model to simulate 10,000 different fisheries, where each fishery is a random combination of variables, described below

Table.S1 - Range of simulated variables

\begin{longtable}[]{@{}rc@{}}
\toprule
\begin{minipage}[b]{0.41\columnwidth}\raggedleft
Variable\strut
\end{minipage} & \begin{minipage}[b]{0.53\columnwidth}\centering
Distribution\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.41\columnwidth}\raggedleft
Scientific Name\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
Drawn from all possible species in \texttt{FishLife} (Thorson et al. (2017))\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.41\columnwidth}\raggedleft
steepness (h)\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
\textasciitilde{}uniform(0.6,0.95)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.41\columnwidth}\raggedleft
Adult movement (\(\sigma_{s=a}\))\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
\textasciitilde{}uniform(0,0.5 * P)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.41\columnwidth}\raggedleft
Larval movement (\(\sigma_{s=l}\))\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
\textasciitilde{}uniform(0,0.5 * P)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.41\columnwidth}\raggedleft
Recruitment variation (\(\sigma_{r}\))\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
\(\in\{0,0.05,.1\}\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.41\columnwidth}\raggedleft
Recruitment autocorrelation (\(ac_{r}\))\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
\(\in\{0,0.05,.1\}\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.41\columnwidth}\raggedleft
DD adult movement (dmod)\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
\(\in\{0.25,1\}\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.41\columnwidth}\raggedleft
Density-dependence timing\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
\(\in\{local, global,post-dispersal\}\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.41\columnwidth}\raggedleft
\% Patches in MPA\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
\textasciitilde{}uniform(0.01,1)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.41\columnwidth}\raggedleft
Initial fishing relative to natural mortality\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
\textasciitilde{}uniform(0.01,4)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.41\columnwidth}\raggedleft
Selectivity as a multiple of maturity length\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
\textasciitilde{}uniform(0.1,1.25)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.41\columnwidth}\raggedleft
Fleet model\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
\(\in\{open-access, constant-effort,constant-catch\}\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.41\columnwidth}\raggedleft
Spatial effort model\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
\(\in\{uniform, biomass,profits\}\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.41\columnwidth}\raggedleft
Years into simulation to start MPA\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
\textasciitilde{}round(uniform(5,0.66 * T))\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.41\columnwidth}\raggedleft
MPA is sprinkler?\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
\(\in\{TRUE,FALSE\}\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.41\columnwidth}\raggedleft
Randomly place MPA?\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
\(\in\{TRUE,FALSE\}\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.41\columnwidth}\raggedleft
Fleet reaction to MPA\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
\(\in\{concentrate, abandon-ship\}\)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.41\columnwidth}\raggedleft
Patchiness\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
\textasciitilde{}uniform(0.01,0.75)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.41\columnwidth}\raggedleft
MPA habitat factor\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
\(\in\{1, 4\}\)\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

One thing to note here is the random sampling of species' scientific names. The effect of MPAs, especially over the short term, will clearly be affected by factors such as the growth rate, the mortality rate, and the maturity schedule. These life history traits are related through a variety of biological processes, as such randomly sampling these parameters can lead to biologically nonsensical ``frankenfish''. We resolve this by using the \texttt{FishLife} package (Thorson et al. (2017)) instead. \texttt{FishLife} builds off of \href{https://fishbase.org/}{FishBase}, and provides estimate of key life history traits taking into account the relationships across these variables. For simulations then, we randomly pull a species from \texttt{FishLife}, and then pull the available life history information from that species for use in the operating model. This allows us to simulate a wide range of life history types in a realistic manner.

We ran 20,000 simulations from these distributions. Each simulation runs for 50 years in 50 patches (with a 25 year unfished burn-in period for conditions in which initial conditions cannot be solved analytically, for example when MPAs have better habitat than non-MPAs). For each simulation, we run one scenario without MPAs, though taking note of where the MPA would be as needed. For the second scenario, we hold everything constant except we now add in the MPAs as dictated by the particular simulation.

\hypertarget{filtering-simulations}{%
\subparagraph{Filtering Simulations}\label{filtering-simulations}}

After the 20,000 simulations have run, we perform a series of filtering steps to remove runs that either a) produced chaotic dynamics during the open-access scenario; b) did not converge to the correct bionomic equilibrium in the open-access scenario; or c) crashed the population before the MPAs went into place (population falls below 5\% of unfished biomass). These filtering steps left us with 9672 viable simulations.

\hypertarget{additional-simulation-results}{%
\paragraph{Additional Simulation Results}\label{additional-simulation-results}}

Simulation results are presented as percent differences in biomass densities with and without MPAs, in order to be comparable to the estimates that the regression model produces. However, this metric presents some problems as a measure of how ``detectable'' an effect size is. As depletion increases, relatively small changes in total biomass (relative to the variance in the observation process) can translate into large percent changes in biomass. For example, moving from a density of .2kg/m\textsuperscript{2} to .4kg/m\textsuperscript{2} translates to a 100\% percent increase, but only a .2kg/m\textsuperscript{2} absolute increase, a small value to detect with a real observation program.

To illustrate these, we present an alternative to our simulation results in which changes in biomass caused by MPAs is scaled by the unfished biomass in the system.

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/pop-conservation-effect-1.pdf}
\caption{\label{fig:pop-conservation-effect}Median (A) and range (B) of equilibrium regional MPA conservation effects (change in total biomass with MPAs relative to without MPAs as a percentage of unfished biomass) across a range of depletion and MPA sizes (and incorporating the full range of scenarios included in our study). `Range in MPA' is the percent of patches covered by an MPA, `Depletion' is the depletion that would have occurred in equilibrium without the MPA}
\end{figure}

\hypertarget{estimation-model}{%
\subsection{Estimation Model}\label{estimation-model}}

The goal of the estimation model is to estimate the difference in the mean densities of targeted and non-targeted species pre-and-post MPA implementation.

\hypertarget{data}{%
\subsubsection{Data}\label{data}}

All observation data used in the estimation model were collected by PISCO. PISCO staff also compiled allometric information used to convert lengths to expected weights, and hence biomass densities. We do not account for error in this translation step Table.S\ref{tab:classcode} lists the species included in the model. In order to be included in the estimation model, a species must have been observed at least twice a year every year for at least 14 years. We also omitted all observations of ``young of the year'' fish due to inability to identify these observations to the nearest species level in many cases. We omitted data from 1999 due to changes in the sampling procedures that occurred after 1999. Per recommendations from PISCO staff we omit observations from the canopy level of the transects (leaving the middle, bottom, and middle canopy levels).

PISCO data report positive observations of fish. In order to use these data in our model we need to add in zeros for any transect that could have observed a given species of fish but did not. We assume that a fish could have been observed on a given transect if that species has ever been observed at that site in any time period in the data (PISCO data are organized by sites, with multiple transect at different locations within the borders of a site). If a species has never been observed at a site we assume that it does not occur at that site.

Interested readers can examine the mean biomass densities of all species in Fig.\ref{fig:targ-trend-plot}-\ref{fig:nontarg-trend-plot}.

We include several additional sources of data in our regression analysis. Temperature readings are included from the PISCO data for each transect. We also include PISCO data on the estimated surge and visibility. We augmented these data with information on kelp cover over time from the Santa Barbara Channel Long Term Ecological Research Network (LTER et al. 2017). We used a k-nearest neighbors algorithm to fill in missing kelp observations, and matched the interpolated kelp data to the PISCO data at the resolution of year-month-site (Fig.\ref{fig:kelp-plot}).

Temperature data were augmented with data from \texttt{FishLife} (Thorson et al. 2017) to include the estimated preferred temperature for a given species, so that we can include deviations from the preferred temperature envelope as a predictor in the model. This allows different temperatures to have different effects on each species (and is less computationally intensive than estimating species-temperature slopes) (Fig.\ref{fig:tempdev-plot}).

We also included lagged catch totals in the Santa Barbara region for the commercially harvest species in the database, in an effort to control for changes in density caused by changes in fishing pressure. Catches were pulled from the CDFW website (\url{https://www.wildlife.ca.gov/Fishing/Commercial/Landings}), and extracted using the \texttt{tabulizer} package in R (Leeper 2018) (Fig.\ref{fig:catches-plot}).

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/tempdev-plot-1.pdf}
\caption{\label{fig:tempdev-plot}Mean deviations from preferred temperature by species and year}
\end{figure}

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/catches-plot-1.pdf}
\caption{\label{fig:catches-plot}Total CDFW reported commercial catches in the Santa Barbara region}
\end{figure}

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/kelp-plot-1.pdf}
\caption{\label{fig:kelp-plot}Mean kelp biomass by island over time from SBC LTER}
\end{figure}

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/targ-trend-plot-1.pdf}
\caption{\label{fig:targ-trend-plot}Centered and scaled mean biomass densities of all targeted finfish in analysis before filtering}
\end{figure}

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/nontarg-trend-plot-1.pdf}
\caption{\label{fig:nontarg-trend-plot}Centered and scaled mean biomass densities of all non-targeted finfish in analysis before filtering}
\end{figure}

\begin{longtable}[t]{lllll}
\caption{\label{tab:classcode}Species included in estimation model}\\
\toprule
classcode & Common Name & Scientific Name & Targeted? & Stock Status\\
\midrule
\endfirsthead
\caption[]{\label{tab:classcode}Species included in estimation model \textit{(continued)}}\\
\toprule
classcode & Common Name & Scientific Name & Targeted? & Stock Status\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
pcla & kelp bass & Paralabrax clathratus & TRUE & unknown - increasing CPUE\\
smys & blue rockfish & Sebastes mystinus & TRUE & Near minimum stock threshold - increasing\\
ejac & black surfperch & Embiotoca jacksoni & TRUE & unknown\\
spul & California sheephead & Semicossyphus pulcher & TRUE & Below target levels\\
rvac & pile perch & Rhacochilus vacca & TRUE & unknown\\
\addlinespace
satr & kelp rockfish & Sebastes atrovirens & TRUE & unknown\\
elat & striped seaperch & Embiotoca lateralis & TRUE & unknown\\
saur & brown rockfish & Sebastes auriculatus & TRUE & unknown\\
cpri & ocean whitefish & Caulolatilus princeps & TRUE & unknown\\
rtox & rubberlip seaperch & Rhacochilus toxotes & TRUE & unknown\\
\addlinespace
schr & black-and-yellow rockfish & Sebastes chrysomelas & TRUE & unknown\\
scau & copper rockfish & Sebastes caurinus & TRUE & Signs of overfishing\\
cpun & blacksmith & Chromis punctipinnis & FALSE & NA\\
gnig & opaleye & Girella nigricans & FALSE & NA\\
hrub & garibaldi & Hypsypops rubicundus & FALSE & NA\\
\addlinespace
hsem & rock wrasse & Halichoeres semicinctus & FALSE & NA\\
ocal & senorita & Oxyjulis californica & FALSE & NA\\
hros & giant kelpfish & Heterostichus rostratus & FALSE & NA\\
mcal & halfmoon & Medialuna californiensis & FALSE & NA\\
opic & painted greenling & Oxylebius pictus & FALSE & NA\\
\addlinespace
bfre & kelp surfperch & Brachyistius frenatus & FALSE & NA\\
hcar & rainbow seaperch & Hypsurus caryi & FALSE & NA\\
pfur & white seaperch & Phanerodon furcatus & FALSE & NA\\*
\end{longtable}

\hypertarget{model}{%
\subsubsection{Model}\label{model}}

The regression analysis uses a mixed-effects hierarchical model. The raw data are estimated length compositions by fish species along a survey transect at a site. Lengths are converted to biomass per allometric relationships supplied by PISCO and supplemented by the \texttt{FishLife} (Thorson et al. 2017) package in R where needed. We performed some minimal data filtering to reduce noise in the data. We only include species that were observed at least twice in each year of the dataset (2000-2017) somewhere in the core Channel Islands (Anacapa, Santa Cruz, Santa Rosa, San Miguel). While some data are available from 1999, per consultation with PISCO we omit those data due to changes in survey protocols. We assign species to targeted and non-targeted groups per the PISCO classifications. This filtering process results in 11 non-targeted species and 12 targeted species remaining in the analysis.

The simplified explanation of the estimation is a hierarchical model in which we first standardize the observed biomass densities into an abundance index of each species over time. The abundance indices in each year are assumed to be log-normally distributed with means and standard deviations for the targeted and non-targeted groups, giving an estimate of the mean densities of targeted and non-targeted species over time. We then calculate the difference between mean density of targeted species and the mean density of non-targeted species in each year.

\newpage

\begin{figure}
\centering
\includegraphics{figs/estimation-cartoon.png}
\caption{\label{fig:cartoon}Cartoon illustration of the hierarchichal difference-in-difference estimator}
\end{figure}

\pagebreak

The first stage of the regression is a log-normal delta model. The model estimates two regressions, the first is a binomial generalized linear model (GLM) with a logit link estimating the probability of observing a given fish species at a observation \emph{i} (transect at time \emph{t}). The probability that a given species was observed \emph{o} at a given observation is distributed

\begin{equation}
o_{s,i} \sim binomial(\frac{1}{1 +e^{-\beta^{o}{X}}})
  \label{eq:binom}
\end{equation}

where \(\beta^{o}\) are the estimated coefficients for the observation model and \emph{X} is a matrix of covariates that include random effects for each year in the data (2000 to 2017).

The expected density \emph{d} of positive observations is modeled per a log-normal distribution

\begin{equation}
  log(d_{s,i}) \sim normal(\beta^{d}X, \sigma_s)
   \label{eq:logd}
 \end{equation}

where \(\beta^{d}\) are the estimated coefficients for the expected density model and \emph{X} is the same matrix of covariates as used in the observation portion of the model and \(\sigma_s\) allows for each species \emph{s} to have different standard deviations.

Our covariate matrix \emph{X} contains both fixed and random effects. Fixed effects include the depth level of the transect, the sampling site, the month of the observation, the estimated surge at the transect, visibility, the depth of the transect, and the experience (and experience squared) of the diver conducting the transect. We classify each species into one of two clusters based on the mean longitude the species was encountered at, breaking the species into two groups: those primarily found in the western end of the Channel Islands those found more in the eastern end. We then estimate random effects for each island for each cluster

\begin{equation}
\beta_{island,cluster} \sim normal(0,\sigma_{cluster})
  \label{eq:island}
\end{equation}

This allows the mean effect of each island to differ for each cluster, e.g.~allowing San Miguel, the easternmost island, to have a higher mean density for eastern species than for more western species (if the data suggest it).

The second critical component of the covariate matrix \emph{X} are random effects for each year for each species

\begin{equation}
  \beta_{year,species} \sim normal(0,\sigma_{species})
  \label{eq:species}
\end{equation}

These \(\beta_{year,species}\) represent our ``standardized'' estimate of observed abundance of each species in each time step, controlling for the included covariates.

However, we still need to account for changes in the probability of detection over time. For that, we create a standard matrix of with rows equal to the number of years and columns corresponding to each of the columns in \emph{X}, holding everything fixed at mean (or most frequently observed level for factors) levels for all variables in X except for the year and species interaction indices. Calling this standardized matrix \(X^{standard}\), the probability of observing a given species in year \emph{y} is then

\begin{equation}
p_{s,y} = (\frac{1}{1 +e^{-\beta^{o}{X^{standard}}}})
  \label{eq:pdetect}
\end{equation}

In the same manner as described by Punt et al. (2000), The standardized index of abundance for species \emph{s} in year \emph{y} then is

\begin{equation}
I_{species,year} = p_{species,year}e^{\beta_{species,year}}
  \label{eq:index}
\end{equation}

The next phase of the model requires us to estimate the mean abundance of targeted and non-targeted species over time. The concept here is that each \(I_{species,year}\) can be modeled by a regression that contains random effects for each year for targeted and non-targeted fishes, the assumption then being that there is a mean density for targeted and non-target species, and \(I_{species,year}\) represent deviations from that mean.

\begin{equation}
log(I_{species,year}) \sim normal(\beta^{effect}X^{effect}, \sigma_I)
  \label{eq:indexprior}
\end{equation}

\(X^{effect}\) contains both fixed and random effects. The fixed effects include an intercept and the temperature deviation for a given species in a year, where temperature deviation is

\begin{equation}
t_{s,y} = (t^{pref}_{s} -  \bar{t_{y}})^2
  \label{eq:tempdev}
\end{equation}

where \(t^{pref}\) is the preferred temperature for species \emph{s} (drawn form \texttt{FishLife}, Thorson et al. (2017)), and \(\bar{t_{y}}\) is the mean temperature encountered by that species in year \emph{y}. We also include as variables in the model the mean kelp cover experienced by a given species in a given year, as well as the total fishery catches reported in the previous year for that species in the Santa Barbara region {[}drawn from the California Department of Fish and Wildlife database{]}. We also include random intercepts for each species in \(X^{effect}\). The most important random effects are year effects for targeted and non-targeted species

\begin{equation}
\beta_{year,targeted} \sim normal(0,\sigma_{targeted})
  \label{eq:betatarg}
\end{equation}

\(\beta_{year,targeted}\) is the mean log density of targeted species in year \emph{y}, controlling for included covariates. Therefore, the final step in the model, the divergence in the standardized abundance trends of targeted and non-targeted species is

\begin{equation}
divergence_{year} =  \beta_{year,targeted = 1} - \beta_{year,targeted = 0}
  \label{eq:mpaeffect}
\end{equation}

The model is fit in TMB to integrate the uncertainty across all levels of the model, with standard errors for each coefficient in the model estimated through the Laplace approximation.

A complete table of estimated coefficients can be seen in Table.S\ref{tab:betatable}.

\begin{longtable}[t]{rrrl}
\caption{\label{tab:betatable}Complete table of estimated coefficients from hierarchical difference-in-difference model}\\
\toprule
estimate & lower & upper & variable\\
\midrule
\endfirsthead
\caption[]{\label{tab:betatable}Complete table of estimated coefficients from hierarchical difference-in-difference model \textit{(continued)}}\\
\toprule
estimate & lower & upper & variable\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
0.20 & 0.15 & 0.25 & cumulative\_n\_obs\\
0.00 & -0.04 & 0.04 & surge\\
0.09 & 0.07 & 0.12 & mean\_depth\\
-0.05 & -0.08 & -0.03 & mean\_vis\\
1.47 & 1.25 & 1.69 & intercept\\
\addlinespace
-0.08 & -0.28 & 0.13 & site\_side-ANACAPA\_ADMIRALS-E\\
-0.04 & -0.23 & 0.15 & site\_side-ANACAPA\_ADMIRALS-W\\
0.10 & -0.08 & 0.27 & site\_side-ANACAPA\_BLACK\_SEA\_BASS-CEN\\
0.08 & -0.13 & 0.29 & site\_side-ANACAPA\_EAST\_FISH\_CAMP-CEN\\
-0.19 & -0.40 & 0.03 & site\_side-ANACAPA\_EAST\_FISH\_CAMP-E\\
\addlinespace
-0.34 & -0.53 & -0.14 & site\_side-ANACAPA\_EAST\_FISH\_CAMP-W\\
-0.11 & -0.26 & 0.04 & site\_side-ANACAPA\_EAST\_ISLE-CEN\\
-0.10 & -0.25 & 0.04 & site\_side-ANACAPA\_EAST\_ISLE-E\\
-0.33 & -0.47 & -0.18 & site\_side-ANACAPA\_EAST\_ISLE-W\\
0.11 & -0.05 & 0.27 & site\_side-ANACAPA\_LIGHTHOUSE\_REEF-CEN\\
\addlinespace
0.05 & -0.11 & 0.20 & site\_side-ANACAPA\_LIGHTHOUSE\_REEF-E\\
0.05 & -0.11 & 0.21 & site\_side-ANACAPA\_LIGHTHOUSE\_REEF-W\\
-0.10 & -0.25 & 0.05 & site\_side-ANACAPA\_MIDDLE\_ISLE-CEN\\
-0.27 & -0.43 & -0.12 & site\_side-ANACAPA\_MIDDLE\_ISLE-E\\
0.03 & -0.12 & 0.19 & site\_side-ANACAPA\_MIDDLE\_ISLE-W\\
\addlinespace
0.01 & -0.14 & 0.16 & site\_side-ANACAPA\_WEST\_ISLE-CEN\\
0.07 & -0.08 & 0.22 & site\_side-ANACAPA\_WEST\_ISLE-E\\
0.06 & -0.09 & 0.22 & site\_side-ANACAPA\_WEST\_ISLE-W\\
0.04 & -0.11 & 0.18 & site\_side-SCI\_CAVERN\_POINT-E\\
0.17 & 0.03 & 0.32 & site\_side-SCI\_CAVERN\_POINT-W\\
\addlinespace
0.10 & -0.05 & 0.24 & site\_side-SCI\_COCHE\_POINT-E\\
0.14 & 0.00 & 0.29 & site\_side-SCI\_COCHE\_POINT-W\\
0.09 & -0.06 & 0.24 & site\_side-SCI\_FORNEY-E\\
0.39 & 0.24 & 0.54 & site\_side-SCI\_FORNEY-W\\
0.45 & 0.30 & 0.59 & site\_side-SCI\_GULL\_ISLE-E\\
\addlinespace
0.02 & -0.13 & 0.17 & site\_side-SCI\_GULL\_ISLE-W\\
0.10 & -0.05 & 0.25 & site\_side-SCI\_HAZARDS-CEN\\
-0.23 & -0.38 & -0.08 & site\_side-SCI\_HAZARDS-E\\
0.04 & -0.11 & 0.19 & site\_side-SCI\_HAZARDS-W\\
0.25 & 0.08 & 0.42 & site\_side-SCI\_LITTLE\_SCORPION-E\\
\addlinespace
0.08 & -0.08 & 0.25 & site\_side-SCI\_LITTLE\_SCORPION-W\\
0.20 & 0.04 & 0.35 & site\_side-SCI\_PAINTED\_CAVE-CEN\\
0.30 & 0.15 & 0.45 & site\_side-SCI\_PAINTED\_CAVE-E\\
0.14 & -0.01 & 0.30 & site\_side-SCI\_PAINTED\_CAVE-W\\
0.04 & -0.10 & 0.19 & site\_side-SCI\_PELICAN-CEN\\
\addlinespace
0.28 & 0.13 & 0.43 & site\_side-SCI\_PELICAN-E\\
0.46 & 0.09 & 0.84 & site\_side-SCI\_PELICAN-FAR\_WEST\\
0.09 & -0.06 & 0.24 & site\_side-SCI\_PELICAN-W\\
0.06 & -0.11 & 0.24 & site\_side-SCI\_POTATO\_PASTURE-E\\
0.19 & 0.01 & 0.37 & site\_side-SCI\_POTATO\_PASTURE-W\\
\addlinespace
0.14 & -0.03 & 0.31 & site\_side-SCI\_SAN\_PEDRO\_POINT-E\\
0.30 & 0.14 & 0.46 & site\_side-SCI\_SAN\_PEDRO\_POINT-W\\
0.13 & -0.03 & 0.30 & site\_side-SCI\_SCORPION\_ANCHORAGE-CEN\\
0.31 & 0.16 & 0.45 & site\_side-SCI\_SCORPION-E\\
0.16 & 0.01 & 0.31 & site\_side-SCI\_SCORPION-W\\
\addlinespace
0.11 & -0.04 & 0.27 & site\_side-SCI\_VALLEY-CEN\\
0.03 & -0.15 & 0.21 & site\_side-SCI\_VALLEY-E\\
-0.09 & -0.25 & 0.07 & site\_side-SCI\_VALLEY-W\\
0.14 & -0.01 & 0.30 & site\_side-SCI\_YELLOWBANKS-CEN\\
0.14 & -0.16 & 0.44 & site\_side-SCI\_YELLOWBANKS-E\\
\addlinespace
-0.14 & -0.29 & 0.01 & site\_side-SCI\_YELLOWBANKS-W\\
0.21 & -0.19 & 0.61 & site\_side-SMI\_BAY\_POINT-CEN\\
-0.41 & -0.60 & -0.22 & site\_side-SMI\_CROOK\_POINT-E\\
0.06 & -0.12 & 0.25 & site\_side-SMI\_CROOK\_POINT-W\\
-0.15 & -0.33 & 0.04 & site\_side-SMI\_CUYLER-E\\
\addlinespace
-0.11 & -0.27 & 0.05 & site\_side-SMI\_CUYLER-W\\
0.27 & 0.11 & 0.43 & site\_side-SMI\_HARRIS\_PT\_RESERVE-E\\
0.04 & -0.13 & 0.22 & site\_side-SMI\_HARRIS\_PT\_RESERVE-W\\
0.46 & 0.29 & 0.63 & site\_side-SMI\_TYLER\_BIGHT-E\\
0.45 & 0.28 & 0.63 & site\_side-SMI\_TYLER\_BIGHT-W\\
\addlinespace
0.10 & -0.21 & 0.42 & site\_side-SRI\_BEACON\_REEF-E\\
-0.24 & -0.61 & 0.12 & site\_side-SRI\_BEACON\_REEF-W\\
-0.08 & -0.38 & 0.22 & site\_side-SRI\_BEE\_ROCK-E\\
-0.46 & -0.79 & -0.12 & site\_side-SRI\_BEE\_ROCK-W\\
-0.42 & -0.92 & 0.08 & site\_side-SRI\_CARRINGTON-CEN\\
\addlinespace
0.29 & -0.22 & 0.80 & site\_side-SRI\_CARRINGTON-E\\
0.27 & -0.12 & 0.67 & site\_side-SRI\_CARRINGTON-W\\
0.24 & 0.07 & 0.41 & site\_side-SRI\_CHICKASAW-E\\
0.26 & 0.08 & 0.43 & site\_side-SRI\_CHICKASAW-W\\
-0.10 & -0.26 & 0.06 & site\_side-SRI\_CLUSTER\_POINT-N\\
\addlinespace
-0.05 & -0.22 & 0.13 & site\_side-SRI\_CLUSTER\_POINT-S\\
0.10 & -0.26 & 0.47 & site\_side-SRI\_FORD\_POINT-CEN\\
0.02 & -0.16 & 0.20 & site\_side-SRI\_JOHNSONS\_LEE\_NORTH-E\\
-0.09 & -0.28 & 0.11 & site\_side-SRI\_JOHNSONS\_LEE\_NORTH-W\\
0.21 & 0.05 & 0.36 & site\_side-SRI\_JOHNSONS\_LEE\_SOUTH-E\\
\addlinespace
0.00 & -0.16 & 0.16 & site\_side-SRI\_JOHNSONS\_LEE\_SOUTH-W\\
-0.36 & -0.60 & -0.12 & site\_side-SRI\_JOLLA\_VIEJA-E\\
-0.28 & -0.56 & 0.00 & site\_side-SRI\_JOLLA\_VIEJA-W\\
0.24 & -0.09 & 0.58 & site\_side-SRI\_MONACOS-E\\
0.16 & -0.17 & 0.48 & site\_side-SRI\_MONACOS-W\\
\addlinespace
0.06 & -0.20 & 0.33 & site\_side-SRI\_RODES\_REEF-E\\
0.21 & -0.04 & 0.45 & site\_side-SRI\_RODES\_REEF-W\\
0.09 & -0.06 & 0.24 & site\_side-SRI\_SOUTH\_POINT-E\\
0.19 & 0.04 & 0.35 & site\_side-SRI\_SOUTH\_POINT-W\\
0.07 & -0.12 & 0.26 & site\_side-SRI\_TRANCION\_CANYON-E\\
\addlinespace
0.14 & -0.06 & 0.33 & site\_side-SRI\_TRANCION\_CANYON-W\\
-0.03 & -0.09 & 0.02 & level-CNMD\\
0.03 & -0.01 & 0.06 & level-MID\\
0.00 & -0.05 & 0.05 & factor\_month-11\\
0.14 & 0.07 & 0.21 & factor\_month-12\\
\addlinespace
-0.28 & -0.50 & -0.07 & factor\_month-7\\
-0.02 & -0.06 & 0.02 & factor\_month-8\\
-0.02 & -0.05 & 0.01 & factor\_month-9\\
-0.05 & -0.11 & 0.01 & cumulative\_n\_obs\_2\\
0.42 & 0.36 & 0.48 & cumulative\_n\_obs\\
\addlinespace
-0.04 & -0.09 & 0.00 & surge\\
-0.26 & -0.29 & -0.23 & mean\_depth\\
0.09 & 0.06 & 0.11 & mean\_vis\\
-0.59 & -0.76 & -0.42 & intercept\\
-0.38 & -0.60 & -0.16 & site\_side-ANACAPA\_ADMIRALS-E\\
\addlinespace
-0.16 & -0.37 & 0.06 & site\_side-ANACAPA\_ADMIRALS-W\\
-0.39 & -0.58 & -0.19 & site\_side-ANACAPA\_BLACK\_SEA\_BASS-CEN\\
-0.46 & -0.70 & -0.23 & site\_side-ANACAPA\_EAST\_FISH\_CAMP-CEN\\
-0.52 & -0.75 & -0.28 & site\_side-ANACAPA\_EAST\_FISH\_CAMP-E\\
-0.27 & -0.49 & -0.05 & site\_side-ANACAPA\_EAST\_FISH\_CAMP-W\\
\addlinespace
0.27 & 0.10 & 0.44 & site\_side-ANACAPA\_EAST\_ISLE-CEN\\
0.34 & 0.18 & 0.51 & site\_side-ANACAPA\_EAST\_ISLE-E\\
-0.05 & -0.22 & 0.12 & site\_side-ANACAPA\_EAST\_ISLE-W\\
-0.27 & -0.45 & -0.09 & site\_side-ANACAPA\_LIGHTHOUSE\_REEF-CEN\\
-0.03 & -0.21 & 0.15 & site\_side-ANACAPA\_LIGHTHOUSE\_REEF-E\\
\addlinespace
-0.21 & -0.39 & -0.03 & site\_side-ANACAPA\_LIGHTHOUSE\_REEF-W\\
-0.18 & -0.35 & -0.01 & site\_side-ANACAPA\_MIDDLE\_ISLE-CEN\\
-0.48 & -0.65 & -0.31 & site\_side-ANACAPA\_MIDDLE\_ISLE-E\\
-0.49 & -0.67 & -0.32 & site\_side-ANACAPA\_MIDDLE\_ISLE-W\\
0.07 & -0.10 & 0.24 & site\_side-ANACAPA\_WEST\_ISLE-CEN\\
\addlinespace
-0.06 & -0.23 & 0.11 & site\_side-ANACAPA\_WEST\_ISLE-E\\
-0.34 & -0.51 & -0.16 & site\_side-ANACAPA\_WEST\_ISLE-W\\
0.11 & -0.06 & 0.27 & site\_side-SCI\_CAVERN\_POINT-E\\
0.12 & -0.04 & 0.29 & site\_side-SCI\_CAVERN\_POINT-W\\
0.20 & 0.03 & 0.36 & site\_side-SCI\_COCHE\_POINT-E\\
\addlinespace
-0.01 & -0.17 & 0.16 & site\_side-SCI\_COCHE\_POINT-W\\
-0.50 & -0.67 & -0.34 & site\_side-SCI\_FORNEY-E\\
-0.36 & -0.52 & -0.19 & site\_side-SCI\_FORNEY-W\\
0.19 & 0.03 & 0.35 & site\_side-SCI\_GULL\_ISLE-E\\
-0.08 & -0.24 & 0.09 & site\_side-SCI\_GULL\_ISLE-W\\
\addlinespace
0.06 & -0.11 & 0.23 & site\_side-SCI\_HAZARDS-CEN\\
0.03 & -0.14 & 0.19 & site\_side-SCI\_HAZARDS-E\\
0.07 & -0.10 & 0.23 & site\_side-SCI\_HAZARDS-W\\
-0.02 & -0.21 & 0.18 & site\_side-SCI\_LITTLE\_SCORPION-E\\
0.28 & 0.09 & 0.47 & site\_side-SCI\_LITTLE\_SCORPION-W\\
\addlinespace
0.05 & -0.12 & 0.22 & site\_side-SCI\_PAINTED\_CAVE-CEN\\
0.16 & -0.01 & 0.33 & site\_side-SCI\_PAINTED\_CAVE-E\\
-0.13 & -0.30 & 0.05 & site\_side-SCI\_PAINTED\_CAVE-W\\
0.15 & -0.01 & 0.32 & site\_side-SCI\_PELICAN-CEN\\
0.17 & 0.00 & 0.34 & site\_side-SCI\_PELICAN-E\\
\addlinespace
-0.07 & -0.47 & 0.32 & site\_side-SCI\_PELICAN-FAR\_WEST\\
-0.12 & -0.29 & 0.04 & site\_side-SCI\_PELICAN-W\\
-0.29 & -0.49 & -0.09 & site\_side-SCI\_POTATO\_PASTURE-E\\
-0.34 & -0.54 & -0.15 & site\_side-SCI\_POTATO\_PASTURE-W\\
-0.24 & -0.44 & -0.05 & site\_side-SCI\_SAN\_PEDRO\_POINT-E\\
\addlinespace
0.25 & 0.06 & 0.43 & site\_side-SCI\_SAN\_PEDRO\_POINT-W\\
-0.06 & -0.25 & 0.13 & site\_side-SCI\_SCORPION\_ANCHORAGE-CEN\\
0.05 & -0.12 & 0.21 & site\_side-SCI\_SCORPION-E\\
-0.16 & -0.33 & 0.00 & site\_side-SCI\_SCORPION-W\\
-0.21 & -0.39 & -0.04 & site\_side-SCI\_VALLEY-CEN\\
\addlinespace
-0.95 & -1.14 & -0.76 & site\_side-SCI\_VALLEY-E\\
-0.58 & -0.76 & -0.40 & site\_side-SCI\_VALLEY-W\\
-0.58 & -0.75 & -0.41 & site\_side-SCI\_YELLOWBANKS-CEN\\
-0.65 & -0.97 & -0.34 & site\_side-SCI\_YELLOWBANKS-E\\
-0.43 & -0.59 & -0.26 & site\_side-SCI\_YELLOWBANKS-W\\
\addlinespace
-0.56 & -0.97 & -0.14 & site\_side-SMI\_BAY\_POINT-CEN\\
-1.52 & -1.71 & -1.32 & site\_side-SMI\_CROOK\_POINT-E\\
-1.36 & -1.55 & -1.17 & site\_side-SMI\_CROOK\_POINT-W\\
-1.67 & -1.86 & -1.48 & site\_side-SMI\_CUYLER-E\\
-0.97 & -1.14 & -0.79 & site\_side-SMI\_CUYLER-W\\
\addlinespace
-0.57 & -0.75 & -0.40 & site\_side-SMI\_HARRIS\_PT\_RESERVE-E\\
-1.43 & -1.62 & -1.25 & site\_side-SMI\_HARRIS\_PT\_RESERVE-W\\
-0.68 & -0.86 & -0.50 & site\_side-SMI\_TYLER\_BIGHT-E\\
-1.00 & -1.18 & -0.82 & site\_side-SMI\_TYLER\_BIGHT-W\\
-1.58 & -1.88 & -1.28 & site\_side-SRI\_BEACON\_REEF-E\\
\addlinespace
-1.77 & -2.12 & -1.43 & site\_side-SRI\_BEACON\_REEF-W\\
-1.08 & -1.37 & -0.80 & site\_side-SRI\_BEE\_ROCK-E\\
-1.16 & -1.46 & -0.85 & site\_side-SRI\_BEE\_ROCK-W\\
-1.21 & -1.67 & -0.74 & site\_side-SRI\_CARRINGTON-CEN\\
-1.29 & -1.76 & -0.82 & site\_side-SRI\_CARRINGTON-E\\
\addlinespace
-1.26 & -1.63 & -0.88 & site\_side-SRI\_CARRINGTON-W\\
-0.40 & -0.58 & -0.21 & site\_side-SRI\_CHICKASAW-E\\
-0.62 & -0.80 & -0.43 & site\_side-SRI\_CHICKASAW-W\\
-0.88 & -1.05 & -0.70 & site\_side-SRI\_CLUSTER\_POINT-N\\
-1.31 & -1.49 & -1.12 & site\_side-SRI\_CLUSTER\_POINT-S\\
\addlinespace
-0.74 & -1.09 & -0.39 & site\_side-SRI\_FORD\_POINT-CEN\\
-0.64 & -0.83 & -0.45 & site\_side-SRI\_JOHNSONS\_LEE\_NORTH-E\\
-0.92 & -1.11 & -0.72 & site\_side-SRI\_JOHNSONS\_LEE\_NORTH-W\\
-0.35 & -0.51 & -0.18 & site\_side-SRI\_JOHNSONS\_LEE\_SOUTH-E\\
-0.96 & -1.13 & -0.79 & site\_side-SRI\_JOHNSONS\_LEE\_SOUTH-W\\
\addlinespace
-0.82 & -1.06 & -0.58 & site\_side-SRI\_JOLLA\_VIEJA-E\\
-1.31 & -1.57 & -1.04 & site\_side-SRI\_JOLLA\_VIEJA-W\\
-1.25 & -1.55 & -0.95 & site\_side-SRI\_MONACOS-E\\
-0.76 & -1.06 & -0.46 & site\_side-SRI\_MONACOS-W\\
-0.83 & -1.09 & -0.56 & site\_side-SRI\_RODES\_REEF-E\\
\addlinespace
-0.92 & -1.17 & -0.68 & site\_side-SRI\_RODES\_REEF-W\\
-0.66 & -0.83 & -0.49 & site\_side-SRI\_SOUTH\_POINT-E\\
-0.58 & -0.75 & -0.41 & site\_side-SRI\_SOUTH\_POINT-W\\
-0.87 & -1.07 & -0.68 & site\_side-SRI\_TRANCION\_CANYON-E\\
-0.88 & -1.08 & -0.69 & site\_side-SRI\_TRANCION\_CANYON-W\\
\addlinespace
-1.63 & -1.68 & -1.57 & level-CNMD\\
-1.46 & -1.50 & -1.43 & level-MID\\
0.01 & -0.04 & 0.06 & factor\_month-11\\
-0.10 & -0.17 & -0.03 & factor\_month-12\\
-0.01 & -0.24 & 0.23 & factor\_month-7\\
\addlinespace
0.09 & 0.05 & 0.13 & factor\_month-8\\
0.12 & 0.09 & 0.16 & factor\_month-9\\
-0.22 & -0.27 & -0.16 & cumulative\_n\_obs\_2\\
-1.18 & -1.71 & -0.65 & year\_classcode-bfre-2000\\
-0.62 & -1.11 & -0.12 & year\_classcode-bfre-2001\\
\addlinespace
-1.48 & -1.95 & -1.01 & year\_classcode-bfre-2002\\
-2.25 & -2.56 & -1.95 & year\_classcode-bfre-2003\\
-1.86 & -2.11 & -1.62 & year\_classcode-bfre-2004\\
-2.03 & -2.26 & -1.80 & year\_classcode-bfre-2005\\
-2.26 & -2.51 & -2.02 & year\_classcode-bfre-2006\\
\addlinespace
-1.81 & -2.04 & -1.58 & year\_classcode-bfre-2007\\
-2.53 & -2.81 & -2.25 & year\_classcode-bfre-2008\\
-2.17 & -2.39 & -1.95 & year\_classcode-bfre-2009\\
-2.42 & -2.68 & -2.17 & year\_classcode-bfre-2010\\
-2.18 & -2.45 & -1.91 & year\_classcode-bfre-2011\\
\addlinespace
-2.30 & -2.56 & -2.03 & year\_classcode-bfre-2012\\
-2.17 & -2.43 & -1.92 & year\_classcode-bfre-2013\\
-1.87 & -2.13 & -1.61 & year\_classcode-bfre-2014\\
-2.52 & -2.81 & -2.23 & year\_classcode-bfre-2015\\
-2.67 & -3.07 & -2.26 & year\_classcode-bfre-2016\\
\addlinespace
-2.39 & -2.71 & -2.08 & year\_classcode-bfre-2017\\
1.07 & 0.48 & 1.66 & year\_classcode-cpri-2000\\
0.67 & 0.00 & 1.33 & year\_classcode-cpri-2001\\
0.64 & -0.02 & 1.29 & year\_classcode-cpri-2002\\
0.60 & 0.10 & 1.10 & year\_classcode-cpri-2003\\
\addlinespace
0.67 & 0.15 & 1.19 & year\_classcode-cpri-2004\\
0.57 & 0.14 & 0.99 & year\_classcode-cpri-2005\\
0.59 & 0.11 & 1.07 & year\_classcode-cpri-2006\\
0.83 & 0.35 & 1.32 & year\_classcode-cpri-2007\\
0.77 & 0.30 & 1.24 & year\_classcode-cpri-2008\\
\addlinespace
1.06 & 0.64 & 1.48 & year\_classcode-cpri-2009\\
1.01 & 0.55 & 1.48 & year\_classcode-cpri-2010\\
1.33 & 0.80 & 1.85 & year\_classcode-cpri-2011\\
1.48 & 0.94 & 2.02 & year\_classcode-cpri-2012\\
1.07 & 0.43 & 1.72 & year\_classcode-cpri-2013\\
\addlinespace
1.10 & 0.53 & 1.68 & year\_classcode-cpri-2014\\
-1.32 & -1.73 & -0.91 & year\_classcode-cpri-2015\\
0.12 & -0.20 & 0.44 & year\_classcode-cpri-2016\\
0.46 & 0.13 & 0.79 & year\_classcode-cpri-2017\\
1.23 & 0.88 & 1.58 & year\_classcode-cpun-2000\\
\addlinespace
1.02 & 0.68 & 1.37 & year\_classcode-cpun-2001\\
0.39 & 0.10 & 0.68 & year\_classcode-cpun-2002\\
0.14 & -0.10 & 0.38 & year\_classcode-cpun-2003\\
-0.30 & -0.54 & -0.07 & year\_classcode-cpun-2004\\
-0.25 & -0.46 & -0.04 & year\_classcode-cpun-2005\\
\addlinespace
-0.41 & -0.63 & -0.20 & year\_classcode-cpun-2006\\
-0.37 & -0.58 & -0.15 & year\_classcode-cpun-2007\\
-0.73 & -0.95 & -0.52 & year\_classcode-cpun-2008\\
0.16 & -0.05 & 0.36 & year\_classcode-cpun-2009\\
0.17 & -0.05 & 0.39 & year\_classcode-cpun-2010\\
\addlinespace
0.35 & 0.12 & 0.58 & year\_classcode-cpun-2011\\
0.16 & -0.07 & 0.39 & year\_classcode-cpun-2012\\
0.13 & -0.11 & 0.36 & year\_classcode-cpun-2013\\
-0.24 & -0.47 & -0.02 & year\_classcode-cpun-2014\\
-0.36 & -0.57 & -0.14 & year\_classcode-cpun-2015\\
\addlinespace
0.04 & -0.17 & 0.25 & year\_classcode-cpun-2016\\
-0.22 & -0.44 & 0.00 & year\_classcode-cpun-2017\\
0.66 & 0.27 & 1.05 & year\_classcode-ejac-2000\\
0.41 & 0.01 & 0.81 & year\_classcode-ejac-2001\\
0.48 & 0.13 & 0.82 & year\_classcode-ejac-2002\\
\addlinespace
-0.18 & -0.45 & 0.10 & year\_classcode-ejac-2003\\
-0.42 & -0.68 & -0.17 & year\_classcode-ejac-2004\\
-0.07 & -0.29 & 0.16 & year\_classcode-ejac-2005\\
0.31 & 0.08 & 0.54 & year\_classcode-ejac-2006\\
-0.09 & -0.32 & 0.13 & year\_classcode-ejac-2007\\
\addlinespace
-0.90 & -1.13 & -0.67 & year\_classcode-ejac-2008\\
-0.13 & -0.35 & 0.09 & year\_classcode-ejac-2009\\
-0.05 & -0.28 & 0.18 & year\_classcode-ejac-2010\\
-0.01 & -0.25 & 0.22 & year\_classcode-ejac-2011\\
-0.44 & -0.68 & -0.20 & year\_classcode-ejac-2012\\
\addlinespace
-0.28 & -0.54 & -0.03 & year\_classcode-ejac-2013\\
-0.23 & -0.48 & 0.01 & year\_classcode-ejac-2014\\
-0.53 & -0.79 & -0.28 & year\_classcode-ejac-2015\\
-0.50 & -0.76 & -0.24 & year\_classcode-ejac-2016\\
-0.84 & -1.09 & -0.58 & year\_classcode-ejac-2017\\
\addlinespace
0.11 & -0.47 & 0.68 & year\_classcode-elat-2000\\
0.29 & -0.27 & 0.85 & year\_classcode-elat-2001\\
0.23 & -0.28 & 0.74 & year\_classcode-elat-2002\\
0.28 & -0.03 & 0.59 & year\_classcode-elat-2003\\
-0.26 & -0.52 & 0.01 & year\_classcode-elat-2004\\
\addlinespace
-0.26 & -0.48 & -0.04 & year\_classcode-elat-2005\\
0.27 & 0.05 & 0.49 & year\_classcode-elat-2006\\
0.12 & -0.12 & 0.37 & year\_classcode-elat-2007\\
-0.73 & -0.96 & -0.49 & year\_classcode-elat-2008\\
0.12 & -0.09 & 0.33 & year\_classcode-elat-2009\\
\addlinespace
0.05 & -0.18 & 0.28 & year\_classcode-elat-2010\\
0.02 & -0.20 & 0.25 & year\_classcode-elat-2011\\
-0.06 & -0.29 & 0.16 & year\_classcode-elat-2012\\
-0.18 & -0.46 & 0.10 & year\_classcode-elat-2013\\
-0.20 & -0.44 & 0.03 & year\_classcode-elat-2014\\
\addlinespace
-0.07 & -0.38 & 0.24 & year\_classcode-elat-2015\\
-0.70 & -1.04 & -0.37 & year\_classcode-elat-2016\\
-0.82 & -1.18 & -0.47 & year\_classcode-elat-2017\\
1.46 & 1.06 & 1.85 & year\_classcode-gnig-2000\\
1.48 & 1.11 & 1.86 & year\_classcode-gnig-2001\\
\addlinespace
1.25 & 0.92 & 1.59 & year\_classcode-gnig-2002\\
1.33 & 1.06 & 1.60 & year\_classcode-gnig-2003\\
1.01 & 0.77 & 1.25 & year\_classcode-gnig-2004\\
1.20 & 0.98 & 1.42 & year\_classcode-gnig-2005\\
1.22 & 1.00 & 1.44 & year\_classcode-gnig-2006\\
\addlinespace
1.27 & 1.05 & 1.49 & year\_classcode-gnig-2007\\
1.02 & 0.80 & 1.24 & year\_classcode-gnig-2008\\
1.32 & 1.10 & 1.53 & year\_classcode-gnig-2009\\
1.46 & 1.21 & 1.70 & year\_classcode-gnig-2010\\
1.34 & 1.10 & 1.59 & year\_classcode-gnig-2011\\
\addlinespace
1.46 & 1.22 & 1.71 & year\_classcode-gnig-2012\\
1.21 & 0.94 & 1.48 & year\_classcode-gnig-2013\\
1.28 & 1.04 & 1.52 & year\_classcode-gnig-2014\\
1.11 & 0.88 & 1.34 & year\_classcode-gnig-2015\\
1.31 & 1.07 & 1.55 & year\_classcode-gnig-2016\\
\addlinespace
1.29 & 1.06 & 1.52 & year\_classcode-gnig-2017\\
-0.39 & -1.89 & 1.10 & year\_classcode-hcar-2000\\
-1.50 & -2.40 & -0.61 & year\_classcode-hcar-2001\\
-1.99 & -2.70 & -1.28 & year\_classcode-hcar-2002\\
-1.44 & -1.93 & -0.95 & year\_classcode-hcar-2003\\
\addlinespace
-1.38 & -1.98 & -0.78 & year\_classcode-hcar-2004\\
-1.47 & -1.80 & -1.15 & year\_classcode-hcar-2005\\
-1.98 & -2.31 & -1.65 & year\_classcode-hcar-2006\\
-1.59 & -1.88 & -1.29 & year\_classcode-hcar-2007\\
-2.68 & -2.98 & -2.37 & year\_classcode-hcar-2008\\
\addlinespace
-1.61 & -1.92 & -1.31 & year\_classcode-hcar-2009\\
-1.73 & -2.11 & -1.36 & year\_classcode-hcar-2010\\
-1.56 & -2.11 & -1.01 & year\_classcode-hcar-2011\\
-1.90 & -2.34 & -1.46 & year\_classcode-hcar-2012\\
-2.00 & -2.47 & -1.53 & year\_classcode-hcar-2013\\
\addlinespace
-1.87 & -2.58 & -1.17 & year\_classcode-hcar-2014\\
-0.87 & -1.79 & 0.05 & year\_classcode-hcar-2015\\
-1.37 & -2.23 & -0.51 & year\_classcode-hcar-2016\\
-1.40 & -2.19 & -0.62 & year\_classcode-hcar-2017\\
-2.80 & -3.71 & -1.89 & year\_classcode-hros-2000\\
\addlinespace
-3.55 & -4.28 & -2.82 & year\_classcode-hros-2001\\
-2.91 & -3.86 & -1.97 & year\_classcode-hros-2002\\
-2.59 & -3.53 & -1.65 & year\_classcode-hros-2003\\
-4.13 & -4.46 & -3.80 & year\_classcode-hros-2004\\
-3.71 & -4.10 & -3.32 & year\_classcode-hros-2005\\
\addlinespace
-4.09 & -4.49 & -3.69 & year\_classcode-hros-2006\\
-3.89 & -4.23 & -3.56 & year\_classcode-hros-2007\\
-4.28 & -4.76 & -3.79 & year\_classcode-hros-2008\\
-3.66 & -3.97 & -3.36 & year\_classcode-hros-2009\\
-1.29 & -2.20 & -0.39 & year\_classcode-hros-2010\\
\addlinespace
-3.54 & -4.02 & -3.06 & year\_classcode-hros-2011\\
-3.59 & -3.99 & -3.18 & year\_classcode-hros-2012\\
-2.25 & -3.11 & -1.39 & year\_classcode-hros-2013\\
-3.54 & -4.02 & -3.06 & year\_classcode-hros-2014\\
-3.31 & -3.81 & -2.81 & year\_classcode-hros-2015\\
\addlinespace
-2.57 & -3.11 & -2.04 & year\_classcode-hros-2016\\
-3.94 & -4.32 & -3.57 & year\_classcode-hros-2017\\
1.73 & 1.38 & 2.07 & year\_classcode-hrub-2000\\
1.78 & 1.47 & 2.10 & year\_classcode-hrub-2001\\
1.47 & 1.18 & 1.77 & year\_classcode-hrub-2002\\
\addlinespace
1.37 & 1.10 & 1.63 & year\_classcode-hrub-2003\\
0.89 & 0.66 & 1.13 & year\_classcode-hrub-2004\\
1.03 & 0.82 & 1.25 & year\_classcode-hrub-2005\\
0.95 & 0.73 & 1.16 & year\_classcode-hrub-2006\\
1.10 & 0.88 & 1.31 & year\_classcode-hrub-2007\\
\addlinespace
0.89 & 0.67 & 1.10 & year\_classcode-hrub-2008\\
1.18 & 0.97 & 1.39 & year\_classcode-hrub-2009\\
1.07 & 0.83 & 1.31 & year\_classcode-hrub-2010\\
1.00 & 0.76 & 1.23 & year\_classcode-hrub-2011\\
1.05 & 0.82 & 1.29 & year\_classcode-hrub-2012\\
\addlinespace
1.02 & 0.78 & 1.26 & year\_classcode-hrub-2013\\
0.34 & 0.12 & 0.57 & year\_classcode-hrub-2014\\
0.49 & 0.27 & 0.71 & year\_classcode-hrub-2015\\
0.36 & 0.14 & 0.58 & year\_classcode-hrub-2016\\
0.70 & 0.48 & 0.91 & year\_classcode-hrub-2017\\
\addlinespace
-0.10 & -0.45 & 0.25 & year\_classcode-hsem-2000\\
0.04 & -0.34 & 0.41 & year\_classcode-hsem-2001\\
-0.29 & -0.68 & 0.11 & year\_classcode-hsem-2002\\
-0.49 & -0.90 & -0.08 & year\_classcode-hsem-2003\\
-0.31 & -0.68 & 0.06 & year\_classcode-hsem-2004\\
\addlinespace
-1.33 & -1.59 & -1.07 & year\_classcode-hsem-2005\\
-0.85 & -1.13 & -0.57 & year\_classcode-hsem-2006\\
-1.16 & -1.39 & -0.93 & year\_classcode-hsem-2007\\
-1.03 & -1.27 & -0.80 & year\_classcode-hsem-2008\\
-0.56 & -0.78 & -0.33 & year\_classcode-hsem-2009\\
\addlinespace
-0.40 & -0.68 & -0.12 & year\_classcode-hsem-2010\\
-0.26 & -0.55 & 0.02 & year\_classcode-hsem-2011\\
-0.13 & -0.52 & 0.27 & year\_classcode-hsem-2012\\
-0.75 & -1.16 & -0.34 & year\_classcode-hsem-2013\\
-1.88 & -2.17 & -1.59 & year\_classcode-hsem-2014\\
\addlinespace
-0.56 & -0.77 & -0.34 & year\_classcode-hsem-2015\\
-0.19 & -0.40 & 0.02 & year\_classcode-hsem-2016\\
-0.13 & -0.35 & 0.08 & year\_classcode-hsem-2017\\
0.87 & 0.43 & 1.32 & year\_classcode-mcal-2000\\
0.99 & 0.58 & 1.39 & year\_classcode-mcal-2001\\
\addlinespace
0.27 & -0.07 & 0.62 & year\_classcode-mcal-2002\\
0.47 & 0.18 & 0.77 & year\_classcode-mcal-2003\\
-0.05 & -0.32 & 0.22 & year\_classcode-mcal-2004\\
0.12 & -0.11 & 0.34 & year\_classcode-mcal-2005\\
-0.15 & -0.36 & 0.06 & year\_classcode-mcal-2006\\
\addlinespace
0.30 & 0.07 & 0.53 & year\_classcode-mcal-2007\\
-0.15 & -0.39 & 0.09 & year\_classcode-mcal-2008\\
0.43 & 0.19 & 0.68 & year\_classcode-mcal-2009\\
0.67 & 0.36 & 0.98 & year\_classcode-mcal-2010\\
0.44 & 0.11 & 0.76 & year\_classcode-mcal-2011\\
\addlinespace
0.71 & 0.36 & 1.07 & year\_classcode-mcal-2012\\
0.49 & 0.13 & 0.84 & year\_classcode-mcal-2013\\
0.11 & -0.14 & 0.36 & year\_classcode-mcal-2014\\
-0.07 & -0.29 & 0.14 & year\_classcode-mcal-2015\\
0.32 & 0.10 & 0.54 & year\_classcode-mcal-2016\\
\addlinespace
0.05 & -0.17 & 0.28 & year\_classcode-mcal-2017\\
0.23 & -0.13 & 0.59 & year\_classcode-ocal-2000\\
0.16 & -0.19 & 0.51 & year\_classcode-ocal-2001\\
-0.36 & -0.69 & -0.03 & year\_classcode-ocal-2002\\
-0.96 & -1.20 & -0.71 & year\_classcode-ocal-2003\\
\addlinespace
-0.93 & -1.15 & -0.70 & year\_classcode-ocal-2004\\
-0.81 & -1.02 & -0.61 & year\_classcode-ocal-2005\\
-0.45 & -0.66 & -0.25 & year\_classcode-ocal-2006\\
-0.50 & -0.71 & -0.30 & year\_classcode-ocal-2007\\
-0.94 & -1.14 & -0.73 & year\_classcode-ocal-2008\\
\addlinespace
-0.19 & -0.38 & 0.01 & year\_classcode-ocal-2009\\
-0.19 & -0.40 & 0.01 & year\_classcode-ocal-2010\\
-0.42 & -0.64 & -0.20 & year\_classcode-ocal-2011\\
-0.62 & -0.83 & -0.40 & year\_classcode-ocal-2012\\
-0.46 & -0.69 & -0.22 & year\_classcode-ocal-2013\\
\addlinespace
-0.65 & -0.87 & -0.44 & year\_classcode-ocal-2014\\
-0.71 & -0.92 & -0.50 & year\_classcode-ocal-2015\\
-0.63 & -0.84 & -0.41 & year\_classcode-ocal-2016\\
-0.57 & -0.79 & -0.35 & year\_classcode-ocal-2017\\
-1.12 & -1.60 & -0.64 & year\_classcode-opic-2000\\
\addlinespace
-1.72 & -2.17 & -1.28 & year\_classcode-opic-2001\\
0.23 & -0.84 & 1.30 & year\_classcode-opic-2002\\
-1.93 & -2.30 & -1.56 & year\_classcode-opic-2003\\
-1.68 & -1.91 & -1.46 & year\_classcode-opic-2004\\
-1.91 & -2.12 & -1.70 & year\_classcode-opic-2005\\
\addlinespace
-1.70 & -1.92 & -1.48 & year\_classcode-opic-2006\\
-2.09 & -2.30 & -1.88 & year\_classcode-opic-2007\\
-2.19 & -2.40 & -1.99 & year\_classcode-opic-2008\\
-2.07 & -2.27 & -1.87 & year\_classcode-opic-2009\\
-1.94 & -2.15 & -1.74 & year\_classcode-opic-2010\\
\addlinespace
-1.84 & -2.05 & -1.64 & year\_classcode-opic-2011\\
-1.80 & -2.01 & -1.60 & year\_classcode-opic-2012\\
-2.06 & -2.28 & -1.85 & year\_classcode-opic-2013\\
-2.35 & -2.56 & -2.13 & year\_classcode-opic-2014\\
-2.49 & -2.75 & -2.23 & year\_classcode-opic-2015\\
\addlinespace
-2.52 & -2.76 & -2.27 & year\_classcode-opic-2016\\
-2.39 & -2.64 & -2.14 & year\_classcode-opic-2017\\
1.39 & 1.07 & 1.71 & year\_classcode-pcla-2000\\
1.45 & 1.14 & 1.75 & year\_classcode-pcla-2001\\
1.26 & 0.98 & 1.53 & year\_classcode-pcla-2002\\
\addlinespace
0.86 & 0.61 & 1.11 & year\_classcode-pcla-2003\\
0.56 & 0.34 & 0.79 & year\_classcode-pcla-2004\\
0.03 & -0.17 & 0.24 & year\_classcode-pcla-2005\\
0.26 & 0.05 & 0.47 & year\_classcode-pcla-2006\\
0.46 & 0.26 & 0.67 & year\_classcode-pcla-2007\\
\addlinespace
0.35 & 0.14 & 0.55 & year\_classcode-pcla-2008\\
0.98 & 0.78 & 1.18 & year\_classcode-pcla-2009\\
0.83 & 0.61 & 1.06 & year\_classcode-pcla-2010\\
1.01 & 0.79 & 1.23 & year\_classcode-pcla-2011\\
1.02 & 0.80 & 1.24 & year\_classcode-pcla-2012\\
\addlinespace
0.86 & 0.63 & 1.08 & year\_classcode-pcla-2013\\
0.54 & 0.33 & 0.76 & year\_classcode-pcla-2014\\
0.49 & 0.28 & 0.70 & year\_classcode-pcla-2015\\
0.65 & 0.44 & 0.86 & year\_classcode-pcla-2016\\
0.78 & 0.58 & 0.99 & year\_classcode-pcla-2017\\
\addlinespace
-0.76 & -2.00 & 0.48 & year\_classcode-pfur-2000\\
-0.92 & -2.05 & 0.22 & year\_classcode-pfur-2001\\
-0.54 & -1.89 & 0.82 & year\_classcode-pfur-2002\\
-1.25 & -2.15 & -0.35 & year\_classcode-pfur-2003\\
-1.84 & -2.61 & -1.06 & year\_classcode-pfur-2004\\
\addlinespace
-0.67 & -1.37 & 0.03 & year\_classcode-pfur-2005\\
-1.00 & -1.74 & -0.26 & year\_classcode-pfur-2006\\
-0.67 & -1.74 & 0.39 & year\_classcode-pfur-2007\\
-1.52 & -2.54 & -0.51 & year\_classcode-pfur-2008\\
-1.45 & -2.10 & -0.79 & year\_classcode-pfur-2009\\
\addlinespace
-0.92 & -1.70 & -0.15 & year\_classcode-pfur-2010\\
-0.47 & -1.42 & 0.48 & year\_classcode-pfur-2011\\
-0.64 & -1.79 & 0.51 & year\_classcode-pfur-2012\\
0.37 & -1.00 & 1.75 & year\_classcode-pfur-2013\\
-0.75 & -1.83 & 0.33 & year\_classcode-pfur-2014\\
\addlinespace
0.51 & -0.83 & 1.86 & year\_classcode-pfur-2015\\
0.13 & -1.13 & 1.39 & year\_classcode-pfur-2016\\
-0.11 & -1.36 & 1.14 & year\_classcode-pfur-2017\\
0.95 & 0.13 & 1.77 & year\_classcode-rtox-2000\\
1.10 & 0.51 & 1.69 & year\_classcode-rtox-2001\\
\addlinespace
0.81 & 0.34 & 1.29 & year\_classcode-rtox-2002\\
0.79 & 0.36 & 1.22 & year\_classcode-rtox-2003\\
0.68 & 0.27 & 1.08 & year\_classcode-rtox-2004\\
0.94 & 0.61 & 1.26 & year\_classcode-rtox-2005\\
0.87 & 0.52 & 1.23 & year\_classcode-rtox-2006\\
\addlinespace
1.26 & 0.90 & 1.62 & year\_classcode-rtox-2007\\
1.40 & 0.96 & 1.83 & year\_classcode-rtox-2008\\
1.31 & 1.01 & 1.61 & year\_classcode-rtox-2009\\
1.12 & 0.78 & 1.47 & year\_classcode-rtox-2010\\
0.79 & 0.48 & 1.11 & year\_classcode-rtox-2011\\
\addlinespace
0.93 & 0.56 & 1.29 & year\_classcode-rtox-2012\\
0.90 & 0.49 & 1.31 & year\_classcode-rtox-2013\\
1.47 & 1.01 & 1.92 & year\_classcode-rtox-2014\\
0.51 & 0.07 & 0.95 & year\_classcode-rtox-2015\\
0.54 & 0.11 & 0.98 & year\_classcode-rtox-2016\\
\addlinespace
-0.18 & -0.78 & 0.41 & year\_classcode-rtox-2017\\
0.27 & -0.15 & 0.68 & year\_classcode-rvac-2000\\
0.29 & -0.11 & 0.70 & year\_classcode-rvac-2001\\
0.05 & -0.31 & 0.41 & year\_classcode-rvac-2002\\
-0.27 & -0.57 & 0.02 & year\_classcode-rvac-2003\\
\addlinespace
-0.39 & -0.68 & -0.10 & year\_classcode-rvac-2004\\
-0.12 & -0.36 & 0.12 & year\_classcode-rvac-2005\\
0.28 & 0.04 & 0.51 & year\_classcode-rvac-2006\\
-0.08 & -0.32 & 0.17 & year\_classcode-rvac-2007\\
-0.49 & -0.77 & -0.22 & year\_classcode-rvac-2008\\
\addlinespace
-0.20 & -0.44 & 0.04 & year\_classcode-rvac-2009\\
-0.13 & -0.38 & 0.11 & year\_classcode-rvac-2010\\
0.00 & -0.25 & 0.25 & year\_classcode-rvac-2011\\
-0.45 & -0.72 & -0.18 & year\_classcode-rvac-2012\\
0.16 & -0.13 & 0.45 & year\_classcode-rvac-2013\\
\addlinespace
-0.13 & -0.41 & 0.15 & year\_classcode-rvac-2014\\
-0.36 & -0.68 & -0.05 & year\_classcode-rvac-2015\\
-0.14 & -0.44 & 0.17 & year\_classcode-rvac-2016\\
-0.67 & -1.00 & -0.34 & year\_classcode-rvac-2017\\
0.78 & 0.24 & 1.31 & year\_classcode-satr-2000\\
\addlinespace
0.87 & 0.41 & 1.32 & year\_classcode-satr-2001\\
0.69 & 0.29 & 1.10 & year\_classcode-satr-2002\\
0.35 & 0.08 & 0.61 & year\_classcode-satr-2003\\
0.02 & -0.21 & 0.26 & year\_classcode-satr-2004\\
0.00 & -0.21 & 0.20 & year\_classcode-satr-2005\\
\addlinespace
0.39 & 0.18 & 0.59 & year\_classcode-satr-2006\\
0.25 & 0.04 & 0.45 & year\_classcode-satr-2007\\
-0.10 & -0.31 & 0.10 & year\_classcode-satr-2008\\
0.50 & 0.31 & 0.69 & year\_classcode-satr-2009\\
0.30 & 0.09 & 0.51 & year\_classcode-satr-2010\\
\addlinespace
0.47 & 0.27 & 0.67 & year\_classcode-satr-2011\\
0.26 & 0.05 & 0.47 & year\_classcode-satr-2012\\
0.56 & 0.31 & 0.80 & year\_classcode-satr-2013\\
0.30 & 0.05 & 0.54 & year\_classcode-satr-2014\\
0.53 & 0.25 & 0.80 & year\_classcode-satr-2015\\
\addlinespace
0.14 & -0.13 & 0.41 & year\_classcode-satr-2016\\
-0.20 & -0.47 & 0.07 & year\_classcode-satr-2017\\
0.27 & -0.74 & 1.28 & year\_classcode-saur-2000\\
-0.05 & -0.84 & 0.75 & year\_classcode-saur-2001\\
0.38 & -0.61 & 1.38 & year\_classcode-saur-2002\\
\addlinespace
-0.14 & -0.95 & 0.68 & year\_classcode-saur-2003\\
-0.22 & -1.04 & 0.60 & year\_classcode-saur-2004\\
0.06 & -0.75 & 0.86 & year\_classcode-saur-2005\\
-0.18 & -0.87 & 0.50 & year\_classcode-saur-2006\\
-0.15 & -0.78 & 0.48 & year\_classcode-saur-2007\\
\addlinespace
-0.27 & -0.96 & 0.43 & year\_classcode-saur-2008\\
0.23 & -0.43 & 0.89 & year\_classcode-saur-2009\\
0.48 & -0.39 & 1.36 & year\_classcode-saur-2010\\
-0.54 & -1.18 & 0.11 & year\_classcode-saur-2011\\
-0.48 & -1.12 & 0.16 & year\_classcode-saur-2012\\
\addlinespace
-0.07 & -0.74 & 0.59 & year\_classcode-saur-2013\\
0.07 & -0.63 & 0.77 & year\_classcode-saur-2014\\
-0.06 & -0.77 & 0.65 & year\_classcode-saur-2015\\
-0.68 & -1.51 & 0.16 & year\_classcode-saur-2016\\
-1.62 & -2.36 & -0.88 & year\_classcode-saur-2017\\
\addlinespace
0.07 & -0.99 & 1.13 & year\_classcode-scau-2000\\
0.39 & -0.70 & 1.48 & year\_classcode-scau-2001\\
0.10 & -0.87 & 1.07 & year\_classcode-scau-2002\\
-0.56 & -1.19 & 0.06 & year\_classcode-scau-2003\\
-0.38 & -0.95 & 0.20 & year\_classcode-scau-2004\\
\addlinespace
-0.16 & -0.71 & 0.39 & year\_classcode-scau-2005\\
-0.08 & -0.60 & 0.44 & year\_classcode-scau-2006\\
-0.05 & -0.56 & 0.45 & year\_classcode-scau-2007\\
-0.06 & -0.59 & 0.47 & year\_classcode-scau-2008\\
0.07 & -0.41 & 0.55 & year\_classcode-scau-2009\\
\addlinespace
-1.21 & -1.66 & -0.77 & year\_classcode-scau-2010\\
-1.63 & -2.03 & -1.24 & year\_classcode-scau-2011\\
-0.20 & -0.67 & 0.28 & year\_classcode-scau-2012\\
-0.99 & -1.54 & -0.45 & year\_classcode-scau-2013\\
-0.47 & -1.00 & 0.06 & year\_classcode-scau-2014\\
\addlinespace
-0.52 & -1.20 & 0.16 & year\_classcode-scau-2015\\
-0.05 & -0.60 & 0.50 & year\_classcode-scau-2016\\
-1.11 & -1.76 & -0.46 & year\_classcode-scau-2017\\
0.19 & -0.41 & 0.78 & year\_classcode-schr-2000\\
0.22 & -0.38 & 0.82 & year\_classcode-schr-2001\\
\addlinespace
0.23 & -0.34 & 0.81 & year\_classcode-schr-2002\\
0.19 & -0.29 & 0.68 & year\_classcode-schr-2003\\
-0.08 & -0.44 & 0.28 & year\_classcode-schr-2004\\
-0.03 & -0.35 & 0.30 & year\_classcode-schr-2005\\
0.07 & -0.27 & 0.41 & year\_classcode-schr-2006\\
\addlinespace
-0.12 & -0.44 & 0.20 & year\_classcode-schr-2007\\
-0.19 & -0.53 & 0.14 & year\_classcode-schr-2008\\
0.01 & -0.26 & 0.27 & year\_classcode-schr-2009\\
-0.43 & -0.77 & -0.08 & year\_classcode-schr-2010\\
-0.63 & -0.98 & -0.27 & year\_classcode-schr-2011\\
\addlinespace
-0.14 & -0.45 & 0.18 & year\_classcode-schr-2012\\
-0.12 & -0.48 & 0.24 & year\_classcode-schr-2013\\
-0.01 & -0.33 & 0.30 & year\_classcode-schr-2014\\
-0.13 & -0.49 & 0.23 & year\_classcode-schr-2015\\
-0.03 & -0.39 & 0.32 & year\_classcode-schr-2016\\
\addlinespace
-0.51 & -0.96 & -0.06 & year\_classcode-schr-2017\\
-0.03 & -0.42 & 0.36 & year\_classcode-smys-2000\\
0.06 & -0.35 & 0.47 & year\_classcode-smys-2001\\
0.10 & -0.28 & 0.49 & year\_classcode-smys-2002\\
-0.10 & -0.36 & 0.16 & year\_classcode-smys-2003\\
\addlinespace
-0.26 & -0.53 & 0.01 & year\_classcode-smys-2004\\
0.00 & -0.21 & 0.21 & year\_classcode-smys-2005\\
0.32 & 0.10 & 0.54 & year\_classcode-smys-2006\\
0.39 & 0.14 & 0.64 & year\_classcode-smys-2007\\
0.19 & -0.06 & 0.44 & year\_classcode-smys-2008\\
\addlinespace
0.77 & 0.51 & 1.04 & year\_classcode-smys-2009\\
0.47 & 0.19 & 0.76 & year\_classcode-smys-2010\\
0.21 & -0.01 & 0.43 & year\_classcode-smys-2011\\
0.26 & 0.01 & 0.51 & year\_classcode-smys-2012\\
0.07 & -0.17 & 0.31 & year\_classcode-smys-2013\\
\addlinespace
-0.17 & -0.38 & 0.05 & year\_classcode-smys-2014\\
-0.24 & -0.49 & 0.00 & year\_classcode-smys-2015\\
0.24 & -0.04 & 0.52 & year\_classcode-smys-2016\\
-0.45 & -0.77 & -0.13 & year\_classcode-smys-2017\\
1.02 & 0.61 & 1.43 & year\_classcode-spul-2000\\
\addlinespace
1.82 & 1.46 & 2.18 & year\_classcode-spul-2001\\
1.47 & 1.11 & 1.84 & year\_classcode-spul-2002\\
1.25 & 0.97 & 1.53 & year\_classcode-spul-2003\\
0.71 & 0.44 & 0.97 & year\_classcode-spul-2004\\
0.49 & 0.27 & 0.71 & year\_classcode-spul-2005\\
\addlinespace
0.84 & 0.62 & 1.07 & year\_classcode-spul-2006\\
1.17 & 0.94 & 1.39 & year\_classcode-spul-2007\\
0.90 & 0.67 & 1.13 & year\_classcode-spul-2008\\
0.70 & 0.49 & 0.91 & year\_classcode-spul-2009\\
0.84 & 0.61 & 1.07 & year\_classcode-spul-2010\\
\addlinespace
1.10 & 0.87 & 1.33 & year\_classcode-spul-2011\\
1.23 & 1.00 & 1.47 & year\_classcode-spul-2012\\
1.56 & 1.31 & 1.80 & year\_classcode-spul-2013\\
1.21 & 0.96 & 1.46 & year\_classcode-spul-2014\\
0.13 & -0.10 & 0.36 & year\_classcode-spul-2015\\
\addlinespace
0.39 & 0.17 & 0.61 & year\_classcode-spul-2016\\
0.78 & 0.56 & 1.00 & year\_classcode-spul-2017\\
-0.27 & -0.70 & 0.17 & year\_classcode-bfre-2000\\
-0.41 & -0.82 & 0.00 & year\_classcode-bfre-2001\\
-0.62 & -0.99 & -0.26 & year\_classcode-bfre-2002\\
\addlinespace
-0.05 & -0.29 & 0.18 & year\_classcode-bfre-2003\\
0.87 & 0.68 & 1.07 & year\_classcode-bfre-2004\\
0.64 & 0.48 & 0.80 & year\_classcode-bfre-2005\\
0.21 & 0.03 & 0.39 & year\_classcode-bfre-2006\\
0.62 & 0.47 & 0.78 & year\_classcode-bfre-2007\\
\addlinespace
-0.51 & -0.71 & -0.32 & year\_classcode-bfre-2008\\
0.76 & 0.61 & 0.91 & year\_classcode-bfre-2009\\
0.23 & 0.04 & 0.42 & year\_classcode-bfre-2010\\
-0.23 & -0.43 & -0.03 & year\_classcode-bfre-2011\\
-0.11 & -0.30 & 0.09 & year\_classcode-bfre-2012\\
\addlinespace
0.76 & 0.56 & 0.97 & year\_classcode-bfre-2013\\
0.13 & -0.07 & 0.32 & year\_classcode-bfre-2014\\
-0.41 & -0.63 & -0.20 & year\_classcode-bfre-2015\\
-1.32 & -1.60 & -1.04 & year\_classcode-bfre-2016\\
-0.58 & -0.81 & -0.35 & year\_classcode-bfre-2017\\
\addlinespace
-1.00 & -1.56 & -0.44 & year\_classcode-cpri-2000\\
-1.41 & -1.99 & -0.83 & year\_classcode-cpri-2001\\
-1.77 & -2.33 & -1.22 & year\_classcode-cpri-2002\\
-1.72 & -2.12 & -1.31 & year\_classcode-cpri-2003\\
-2.16 & -2.57 & -1.74 & year\_classcode-cpri-2004\\
\addlinespace
-2.00 & -2.32 & -1.67 & year\_classcode-cpri-2005\\
-2.16 & -2.53 & -1.80 & year\_classcode-cpri-2006\\
-2.33 & -2.70 & -1.96 & year\_classcode-cpri-2007\\
-2.28 & -2.64 & -1.91 & year\_classcode-cpri-2008\\
-2.06 & -2.38 & -1.74 & year\_classcode-cpri-2009\\
\addlinespace
-2.05 & -2.42 & -1.69 & year\_classcode-cpri-2010\\
-2.45 & -2.86 & -2.04 & year\_classcode-cpri-2011\\
-2.52 & -2.94 & -2.10 & year\_classcode-cpri-2012\\
-2.63 & -3.14 & -2.12 & year\_classcode-cpri-2013\\
-2.61 & -3.06 & -2.16 & year\_classcode-cpri-2014\\
\addlinespace
-1.60 & -1.90 & -1.29 & year\_classcode-cpri-2015\\
-0.89 & -1.14 & -0.64 & year\_classcode-cpri-2016\\
-0.93 & -1.19 & -0.68 & year\_classcode-cpri-2017\\
2.14 & 1.70 & 2.58 & year\_classcode-cpun-2000\\
1.51 & 1.15 & 1.88 & year\_classcode-cpun-2001\\
\addlinespace
1.70 & 1.41 & 1.99 & year\_classcode-cpun-2002\\
1.55 & 1.35 & 1.75 & year\_classcode-cpun-2003\\
1.07 & 0.89 & 1.26 & year\_classcode-cpun-2004\\
1.09 & 0.95 & 1.24 & year\_classcode-cpun-2005\\
1.21 & 1.06 & 1.36 & year\_classcode-cpun-2006\\
\addlinespace
0.89 & 0.74 & 1.03 & year\_classcode-cpun-2007\\
0.79 & 0.64 & 0.94 & year\_classcode-cpun-2008\\
1.44 & 1.30 & 1.58 & year\_classcode-cpun-2009\\
1.35 & 1.19 & 1.52 & year\_classcode-cpun-2010\\
0.87 & 0.70 & 1.04 & year\_classcode-cpun-2011\\
\addlinespace
0.77 & 0.60 & 0.94 & year\_classcode-cpun-2012\\
1.11 & 0.93 & 1.29 & year\_classcode-cpun-2013\\
1.07 & 0.90 & 1.25 & year\_classcode-cpun-2014\\
1.73 & 1.56 & 1.91 & year\_classcode-cpun-2015\\
2.26 & 2.08 & 2.44 & year\_classcode-cpun-2016\\
\addlinespace
1.82 & 1.65 & 2.00 & year\_classcode-cpun-2017\\
0.83 & 0.43 & 1.22 & year\_classcode-ejac-2000\\
0.36 & -0.01 & 0.72 & year\_classcode-ejac-2001\\
0.50 & 0.20 & 0.79 & year\_classcode-ejac-2002\\
0.70 & 0.49 & 0.91 & year\_classcode-ejac-2003\\
\addlinespace
0.80 & 0.62 & 0.99 & year\_classcode-ejac-2004\\
0.78 & 0.62 & 0.93 & year\_classcode-ejac-2005\\
0.69 & 0.53 & 0.85 & year\_classcode-ejac-2006\\
0.47 & 0.31 & 0.62 & year\_classcode-ejac-2007\\
0.68 & 0.53 & 0.83 & year\_classcode-ejac-2008\\
\addlinespace
0.79 & 0.64 & 0.93 & year\_classcode-ejac-2009\\
0.96 & 0.79 & 1.13 & year\_classcode-ejac-2010\\
0.75 & 0.58 & 0.93 & year\_classcode-ejac-2011\\
0.65 & 0.48 & 0.82 & year\_classcode-ejac-2012\\
0.63 & 0.44 & 0.82 & year\_classcode-ejac-2013\\
\addlinespace
0.63 & 0.45 & 0.81 & year\_classcode-ejac-2014\\
0.36 & 0.18 & 0.55 & year\_classcode-ejac-2015\\
0.34 & 0.16 & 0.52 & year\_classcode-ejac-2016\\
0.69 & 0.51 & 0.87 & year\_classcode-ejac-2017\\
-0.18 & -0.70 & 0.33 & year\_classcode-elat-2000\\
\addlinespace
-0.23 & -0.72 & 0.26 & year\_classcode-elat-2001\\
-0.30 & -0.75 & 0.15 & year\_classcode-elat-2002\\
-0.11 & -0.41 & 0.18 & year\_classcode-elat-2003\\
0.16 & -0.10 & 0.43 & year\_classcode-elat-2004\\
0.22 & -0.01 & 0.45 & year\_classcode-elat-2005\\
\addlinespace
0.24 & 0.00 & 0.47 & year\_classcode-elat-2006\\
-0.49 & -0.73 & -0.24 & year\_classcode-elat-2007\\
-0.09 & -0.32 & 0.14 & year\_classcode-elat-2008\\
0.24 & 0.02 & 0.47 & year\_classcode-elat-2009\\
0.46 & 0.21 & 0.70 & year\_classcode-elat-2010\\
\addlinespace
0.48 & 0.24 & 0.73 & year\_classcode-elat-2011\\
0.47 & 0.23 & 0.71 & year\_classcode-elat-2012\\
-0.02 & -0.29 & 0.24 & year\_classcode-elat-2013\\
0.47 & 0.22 & 0.72 & year\_classcode-elat-2014\\
-0.40 & -0.69 & -0.11 & year\_classcode-elat-2015\\
\addlinespace
-0.57 & -0.85 & -0.28 & year\_classcode-elat-2016\\
-0.37 & -0.66 & -0.08 & year\_classcode-elat-2017\\
0.11 & -0.18 & 0.39 & year\_classcode-gnig-2000\\
0.08 & -0.19 & 0.35 & year\_classcode-gnig-2001\\
-0.05 & -0.30 & 0.20 & year\_classcode-gnig-2002\\
\addlinespace
-0.04 & -0.24 & 0.17 & year\_classcode-gnig-2003\\
0.27 & 0.07 & 0.46 & year\_classcode-gnig-2004\\
0.12 & -0.04 & 0.27 & year\_classcode-gnig-2005\\
0.22 & 0.06 & 0.38 & year\_classcode-gnig-2006\\
0.06 & -0.09 & 0.22 & year\_classcode-gnig-2007\\
\addlinespace
-0.04 & -0.20 & 0.11 & year\_classcode-gnig-2008\\
0.10 & -0.05 & 0.26 & year\_classcode-gnig-2009\\
-0.12 & -0.30 & 0.06 & year\_classcode-gnig-2010\\
-0.19 & -0.38 & -0.01 & year\_classcode-gnig-2011\\
-0.22 & -0.40 & -0.03 & year\_classcode-gnig-2012\\
\addlinespace
-0.44 & -0.66 & -0.22 & year\_classcode-gnig-2013\\
-0.15 & -0.34 & 0.03 & year\_classcode-gnig-2014\\
0.08 & -0.09 & 0.26 & year\_classcode-gnig-2015\\
-0.06 & -0.24 & 0.12 & year\_classcode-gnig-2016\\
0.10 & -0.08 & 0.27 & year\_classcode-gnig-2017\\
\addlinespace
-2.28 & -3.76 & -0.79 & year\_classcode-hcar-2000\\
-0.92 & -1.75 & -0.10 & year\_classcode-hcar-2001\\
-0.70 & -1.34 & -0.06 & year\_classcode-hcar-2002\\
-1.24 & -1.67 & -0.82 & year\_classcode-hcar-2003\\
-2.11 & -2.60 & -1.63 & year\_classcode-hcar-2004\\
\addlinespace
-1.08 & -1.37 & -0.79 & year\_classcode-hcar-2005\\
-0.94 & -1.23 & -0.65 & year\_classcode-hcar-2006\\
-0.79 & -1.06 & -0.52 & year\_classcode-hcar-2007\\
-0.86 & -1.13 & -0.58 & year\_classcode-hcar-2008\\
-1.03 & -1.31 & -0.75 & year\_classcode-hcar-2009\\
\addlinespace
-1.17 & -1.50 & -0.84 & year\_classcode-hcar-2010\\
-2.20 & -2.64 & -1.76 & year\_classcode-hcar-2011\\
-1.66 & -2.03 & -1.29 & year\_classcode-hcar-2012\\
-1.41 & -1.80 & -1.02 & year\_classcode-hcar-2013\\
-2.47 & -2.99 & -1.94 & year\_classcode-hcar-2014\\
\addlinespace
-2.95 & -3.66 & -2.24 & year\_classcode-hcar-2015\\
-2.85 & -3.47 & -2.24 & year\_classcode-hcar-2016\\
-2.41 & -3.01 & -1.81 & year\_classcode-hcar-2017\\
-1.65 & -2.36 & -0.94 & year\_classcode-hros-2000\\
-1.13 & -1.67 & -0.59 & year\_classcode-hros-2001\\
\addlinespace
-2.30 & -3.00 & -1.60 & year\_classcode-hros-2002\\
-2.97 & -3.64 & -2.30 & year\_classcode-hros-2003\\
-0.21 & -0.44 & 0.02 & year\_classcode-hros-2004\\
-1.31 & -1.56 & -1.05 & year\_classcode-hros-2005\\
-1.27 & -1.53 & -1.01 & year\_classcode-hros-2006\\
\addlinespace
-0.91 & -1.13 & -0.70 & year\_classcode-hros-2007\\
-1.89 & -2.20 & -1.59 & year\_classcode-hros-2008\\
-0.68 & -0.87 & -0.48 & year\_classcode-hros-2009\\
-3.58 & -4.27 & -2.89 & year\_classcode-hros-2010\\
-1.77 & -2.08 & -1.45 & year\_classcode-hros-2011\\
\addlinespace
-1.30 & -1.57 & -1.03 & year\_classcode-hros-2012\\
-3.07 & -3.71 & -2.44 & year\_classcode-hros-2013\\
-1.74 & -2.07 & -1.42 & year\_classcode-hros-2014\\
-1.86 & -2.20 & -1.52 & year\_classcode-hros-2015\\
-2.06 & -2.44 & -1.69 & year\_classcode-hros-2016\\
\addlinespace
-0.92 & -1.17 & -0.66 & year\_classcode-hros-2017\\
0.23 & -0.09 & 0.56 & year\_classcode-hrub-2000\\
0.31 & 0.00 & 0.62 & year\_classcode-hrub-2001\\
0.15 & -0.11 & 0.42 & year\_classcode-hrub-2002\\
-0.16 & -0.39 & 0.06 & year\_classcode-hrub-2003\\
\addlinespace
0.09 & -0.11 & 0.29 & year\_classcode-hrub-2004\\
0.08 & -0.08 & 0.25 & year\_classcode-hrub-2005\\
0.18 & 0.01 & 0.35 & year\_classcode-hrub-2006\\
0.05 & -0.12 & 0.21 & year\_classcode-hrub-2007\\
0.07 & -0.09 & 0.23 & year\_classcode-hrub-2008\\
\addlinespace
-0.02 & -0.18 & 0.14 & year\_classcode-hrub-2009\\
-0.37 & -0.56 & -0.17 & year\_classcode-hrub-2010\\
-0.20 & -0.39 & -0.01 & year\_classcode-hrub-2011\\
-0.27 & -0.46 & -0.08 & year\_classcode-hrub-2012\\
-0.10 & -0.30 & 0.09 & year\_classcode-hrub-2013\\
\addlinespace
-0.06 & -0.24 & 0.13 & year\_classcode-hrub-2014\\
0.04 & -0.13 & 0.22 & year\_classcode-hrub-2015\\
0.31 & 0.13 & 0.50 & year\_classcode-hrub-2016\\
0.42 & 0.24 & 0.60 & year\_classcode-hrub-2017\\
0.31 & -0.11 & 0.72 & year\_classcode-hsem-2000\\
\addlinespace
-0.20 & -0.61 & 0.21 & year\_classcode-hsem-2001\\
-0.80 & -1.19 & -0.40 & year\_classcode-hsem-2002\\
-1.52 & -1.90 & -1.14 & year\_classcode-hsem-2003\\
-1.58 & -1.92 & -1.24 & year\_classcode-hsem-2004\\
-0.82 & -1.04 & -0.59 & year\_classcode-hsem-2005\\
\addlinespace
-1.03 & -1.28 & -0.78 & year\_classcode-hsem-2006\\
-0.10 & -0.28 & 0.09 & year\_classcode-hsem-2007\\
-0.32 & -0.52 & -0.13 & year\_classcode-hsem-2008\\
-0.29 & -0.48 & -0.11 & year\_classcode-hsem-2009\\
-1.03 & -1.29 & -0.78 & year\_classcode-hsem-2010\\
\addlinespace
-1.13 & -1.39 & -0.88 & year\_classcode-hsem-2011\\
-2.10 & -2.46 & -1.74 & year\_classcode-hsem-2012\\
-1.83 & -2.19 & -1.46 & year\_classcode-hsem-2013\\
-0.95 & -1.20 & -0.70 & year\_classcode-hsem-2014\\
0.54 & 0.36 & 0.73 & year\_classcode-hsem-2015\\
\addlinespace
0.75 & 0.56 & 0.93 & year\_classcode-hsem-2016\\
0.80 & 0.62 & 0.98 & year\_classcode-hsem-2017\\
-0.45 & -0.92 & 0.02 & year\_classcode-mcal-2000\\
-0.37 & -0.80 & 0.05 & year\_classcode-mcal-2001\\
-0.53 & -0.89 & -0.17 & year\_classcode-mcal-2002\\
\addlinespace
-0.74 & -1.03 & -0.46 & year\_classcode-mcal-2003\\
-0.82 & -1.08 & -0.57 & year\_classcode-mcal-2004\\
-0.40 & -0.59 & -0.21 & year\_classcode-mcal-2005\\
0.11 & -0.06 & 0.28 & year\_classcode-mcal-2006\\
-0.54 & -0.73 & -0.35 & year\_classcode-mcal-2007\\
\addlinespace
-0.86 & -1.07 & -0.65 & year\_classcode-mcal-2008\\
-0.96 & -1.17 & -0.75 & year\_classcode-mcal-2009\\
-1.42 & -1.70 & -1.14 & year\_classcode-mcal-2010\\
-1.69 & -1.99 & -1.39 & year\_classcode-mcal-2011\\
-1.89 & -2.21 & -1.57 & year\_classcode-mcal-2012\\
\addlinespace
-1.70 & -2.04 & -1.37 & year\_classcode-mcal-2013\\
-0.76 & -0.99 & -0.54 & year\_classcode-mcal-2014\\
0.24 & 0.06 & 0.43 & year\_classcode-mcal-2015\\
0.24 & 0.06 & 0.43 & year\_classcode-mcal-2016\\
-0.05 & -0.25 & 0.15 & year\_classcode-mcal-2017\\
\addlinespace
1.02 & 0.63 & 1.42 & year\_classcode-ocal-2000\\
0.82 & 0.47 & 1.18 & year\_classcode-ocal-2001\\
0.61 & 0.31 & 0.90 & year\_classcode-ocal-2002\\
1.27 & 1.07 & 1.47 & year\_classcode-ocal-2003\\
1.54 & 1.36 & 1.73 & year\_classcode-ocal-2004\\
\addlinespace
1.69 & 1.54 & 1.83 & year\_classcode-ocal-2005\\
1.57 & 1.42 & 1.72 & year\_classcode-ocal-2006\\
1.61 & 1.47 & 1.75 & year\_classcode-ocal-2007\\
1.67 & 1.53 & 1.81 & year\_classcode-ocal-2008\\
2.25 & 2.10 & 2.39 & year\_classcode-ocal-2009\\
\addlinespace
1.81 & 1.64 & 1.97 & year\_classcode-ocal-2010\\
1.10 & 0.93 & 1.27 & year\_classcode-ocal-2011\\
1.21 & 1.04 & 1.38 & year\_classcode-ocal-2012\\
0.83 & 0.64 & 1.01 & year\_classcode-ocal-2013\\
1.44 & 1.26 & 1.61 & year\_classcode-ocal-2014\\
\addlinespace
1.86 & 1.68 & 2.04 & year\_classcode-ocal-2015\\
1.48 & 1.31 & 1.65 & year\_classcode-ocal-2016\\
1.35 & 1.18 & 1.53 & year\_classcode-ocal-2017\\
-0.51 & -0.99 & -0.04 & year\_classcode-opic-2000\\
-0.44 & -0.87 & -0.02 & year\_classcode-opic-2001\\
\addlinespace
-2.51 & -3.29 & -1.73 & year\_classcode-opic-2002\\
-1.15 & -1.47 & -0.83 & year\_classcode-opic-2003\\
0.49 & 0.30 & 0.68 & year\_classcode-opic-2004\\
0.50 & 0.35 & 0.66 & year\_classcode-opic-2005\\
0.02 & -0.16 & 0.19 & year\_classcode-opic-2006\\
\addlinespace
0.17 & 0.01 & 0.33 & year\_classcode-opic-2007\\
0.54 & 0.39 & 0.70 & year\_classcode-opic-2008\\
0.64 & 0.50 & 0.79 & year\_classcode-opic-2009\\
1.07 & 0.91 & 1.24 & year\_classcode-opic-2010\\
1.04 & 0.87 & 1.22 & year\_classcode-opic-2011\\
\addlinespace
1.12 & 0.95 & 1.28 & year\_classcode-opic-2012\\
0.74 & 0.55 & 0.93 & year\_classcode-opic-2013\\
0.39 & 0.21 & 0.58 & year\_classcode-opic-2014\\
-0.70 & -0.93 & -0.48 & year\_classcode-opic-2015\\
-0.25 & -0.45 & -0.04 & year\_classcode-opic-2016\\
\addlinespace
-0.39 & -0.60 & -0.17 & year\_classcode-opic-2017\\
1.99 & 1.57 & 2.41 & year\_classcode-pcla-2000\\
1.86 & 1.49 & 2.23 & year\_classcode-pcla-2001\\
1.55 & 1.27 & 1.83 & year\_classcode-pcla-2002\\
0.97 & 0.76 & 1.18 & year\_classcode-pcla-2003\\
\addlinespace
1.21 & 1.02 & 1.41 & year\_classcode-pcla-2004\\
1.03 & 0.88 & 1.18 & year\_classcode-pcla-2005\\
1.26 & 1.10 & 1.41 & year\_classcode-pcla-2006\\
1.05 & 0.90 & 1.19 & year\_classcode-pcla-2007\\
1.16 & 1.01 & 1.31 & year\_classcode-pcla-2008\\
\addlinespace
1.34 & 1.19 & 1.48 & year\_classcode-pcla-2009\\
0.93 & 0.76 & 1.11 & year\_classcode-pcla-2010\\
0.95 & 0.77 & 1.13 & year\_classcode-pcla-2011\\
0.94 & 0.77 & 1.12 & year\_classcode-pcla-2012\\
0.90 & 0.71 & 1.08 & year\_classcode-pcla-2013\\
\addlinespace
1.15 & 0.98 & 1.33 & year\_classcode-pcla-2014\\
1.58 & 1.40 & 1.75 & year\_classcode-pcla-2015\\
1.59 & 1.42 & 1.77 & year\_classcode-pcla-2016\\
2.06 & 1.88 & 2.24 & year\_classcode-pcla-2017\\
-2.86 & -4.04 & -1.68 & year\_classcode-pfur-2000\\
\addlinespace
-2.72 & -3.75 & -1.69 & year\_classcode-pfur-2001\\
-3.77 & -5.05 & -2.48 & year\_classcode-pfur-2002\\
-3.01 & -3.72 & -2.30 & year\_classcode-pfur-2003\\
-2.63 & -3.14 & -2.12 & year\_classcode-pfur-2004\\
-2.99 & -3.52 & -2.47 & year\_classcode-pfur-2005\\
\addlinespace
-2.98 & -3.52 & -2.44 & year\_classcode-pfur-2006\\
-3.98 & -4.78 & -3.18 & year\_classcode-pfur-2007\\
-3.64 & -4.33 & -2.95 & year\_classcode-pfur-2008\\
-2.64 & -3.07 & -2.21 & year\_classcode-pfur-2009\\
-3.06 & -3.64 & -2.48 & year\_classcode-pfur-2010\\
\addlinespace
-3.90 & -4.69 & -3.11 & year\_classcode-pfur-2011\\
-4.19 & -5.10 & -3.27 & year\_classcode-pfur-2012\\
-5.09 & -6.51 & -3.67 & year\_classcode-pfur-2013\\
-4.08 & -4.96 & -3.20 & year\_classcode-pfur-2014\\
-5.23 & -6.60 & -3.87 & year\_classcode-pfur-2015\\
\addlinespace
-4.70 & -5.86 & -3.53 & year\_classcode-pfur-2016\\
-4.48 & -5.59 & -3.38 & year\_classcode-pfur-2017\\
-1.76 & -2.48 & -1.05 & year\_classcode-rtox-2000\\
-1.28 & -1.83 & -0.73 & year\_classcode-rtox-2001\\
-1.07 & -1.51 & -0.64 & year\_classcode-rtox-2002\\
\addlinespace
-1.57 & -1.94 & -1.19 & year\_classcode-rtox-2003\\
-1.71 & -2.05 & -1.36 & year\_classcode-rtox-2004\\
-1.53 & -1.79 & -1.26 & year\_classcode-rtox-2005\\
-1.66 & -1.95 & -1.37 & year\_classcode-rtox-2006\\
-1.92 & -2.22 & -1.62 & year\_classcode-rtox-2007\\
\addlinespace
-2.39 & -2.75 & -2.03 & year\_classcode-rtox-2008\\
-1.37 & -1.60 & -1.13 & year\_classcode-rtox-2009\\
-1.49 & -1.77 & -1.20 & year\_classcode-rtox-2010\\
-1.25 & -1.50 & -0.99 & year\_classcode-rtox-2011\\
-1.73 & -2.03 & -1.42 & year\_classcode-rtox-2012\\
\addlinespace
-1.75 & -2.09 & -1.41 & year\_classcode-rtox-2013\\
-2.28 & -2.66 & -1.90 & year\_classcode-rtox-2014\\
-2.18 & -2.55 & -1.82 & year\_classcode-rtox-2015\\
-2.08 & -2.44 & -1.71 & year\_classcode-rtox-2016\\
-2.58 & -3.02 & -2.13 & year\_classcode-rtox-2017\\
\addlinespace
-0.14 & -0.52 & 0.25 & year\_classcode-rvac-2000\\
-0.23 & -0.59 & 0.14 & year\_classcode-rvac-2001\\
-0.15 & -0.46 & 0.16 & year\_classcode-rvac-2002\\
0.09 & -0.13 & 0.32 & year\_classcode-rvac-2003\\
-0.13 & -0.34 & 0.08 & year\_classcode-rvac-2004\\
\addlinespace
0.03 & -0.13 & 0.20 & year\_classcode-rvac-2005\\
0.06 & -0.11 & 0.23 & year\_classcode-rvac-2006\\
-0.20 & -0.37 & -0.04 & year\_classcode-rvac-2007\\
-0.41 & -0.58 & -0.23 & year\_classcode-rvac-2008\\
-0.08 & -0.24 & 0.08 & year\_classcode-rvac-2009\\
\addlinespace
0.23 & 0.05 & 0.41 & year\_classcode-rvac-2010\\
0.03 & -0.15 & 0.21 & year\_classcode-rvac-2011\\
-0.02 & -0.21 & 0.16 & year\_classcode-rvac-2012\\
-0.51 & -0.74 & -0.29 & year\_classcode-rvac-2013\\
-0.41 & -0.62 & -0.21 & year\_classcode-rvac-2014\\
\addlinespace
-0.73 & -0.96 & -0.51 & year\_classcode-rvac-2015\\
-0.77 & -1.00 & -0.53 & year\_classcode-rvac-2016\\
-0.53 & -0.75 & -0.31 & year\_classcode-rvac-2017\\
0.87 & 0.33 & 1.42 & year\_classcode-satr-2000\\
1.19 & 0.73 & 1.66 & year\_classcode-satr-2001\\
\addlinespace
0.96 & 0.56 & 1.36 & year\_classcode-satr-2002\\
0.73 & 0.45 & 1.01 & year\_classcode-satr-2003\\
0.77 & 0.50 & 1.03 & year\_classcode-satr-2004\\
0.53 & 0.30 & 0.75 & year\_classcode-satr-2005\\
0.76 & 0.53 & 0.99 & year\_classcode-satr-2006\\
\addlinespace
0.38 & 0.15 & 0.60 & year\_classcode-satr-2007\\
0.34 & 0.12 & 0.57 & year\_classcode-satr-2008\\
0.88 & 0.66 & 1.10 & year\_classcode-satr-2009\\
1.11 & 0.87 & 1.35 & year\_classcode-satr-2010\\
1.35 & 1.11 & 1.59 & year\_classcode-satr-2011\\
\addlinespace
0.83 & 0.59 & 1.07 & year\_classcode-satr-2012\\
0.65 & 0.39 & 0.90 & year\_classcode-satr-2013\\
0.40 & 0.14 & 0.66 & year\_classcode-satr-2014\\
0.26 & -0.02 & 0.53 & year\_classcode-satr-2015\\
0.02 & -0.25 & 0.29 & year\_classcode-satr-2016\\
\addlinespace
0.33 & 0.06 & 0.61 & year\_classcode-satr-2017\\
-3.52 & -4.77 & -2.27 & year\_classcode-saur-2000\\
-2.64 & -3.56 & -1.72 & year\_classcode-saur-2001\\
-3.88 & -5.05 & -2.71 & year\_classcode-saur-2002\\
-3.46 & -4.30 & -2.63 & year\_classcode-saur-2003\\
\addlinespace
-3.69 & -4.50 & -2.88 & year\_classcode-saur-2004\\
-3.88 & -4.70 & -3.06 & year\_classcode-saur-2005\\
-3.20 & -3.85 & -2.55 & year\_classcode-saur-2006\\
-3.09 & -3.69 & -2.49 & year\_classcode-saur-2007\\
-3.42 & -4.09 & -2.75 & year\_classcode-saur-2008\\
\addlinespace
-3.35 & -3.99 & -2.72 & year\_classcode-saur-2009\\
-4.12 & -5.02 & -3.23 & year\_classcode-saur-2010\\
-2.98 & -3.53 & -2.43 & year\_classcode-saur-2011\\
-3.00 & -3.56 & -2.44 & year\_classcode-saur-2012\\
-3.04 & -3.71 & -2.38 & year\_classcode-saur-2013\\
\addlinespace
-3.52 & -4.21 & -2.83 & year\_classcode-saur-2014\\
-3.63 & -4.35 & -2.91 & year\_classcode-saur-2015\\
-3.47 & -4.18 & -2.76 & year\_classcode-saur-2016\\
-2.60 & -3.11 & -2.08 & year\_classcode-saur-2017\\
-1.69 & -2.82 & -0.55 & year\_classcode-scau-2000\\
\addlinespace
-2.12 & -3.30 & -0.95 & year\_classcode-scau-2001\\
-1.90 & -2.86 & -0.94 & year\_classcode-scau-2002\\
-1.44 & -1.89 & -0.99 & year\_classcode-scau-2003\\
-1.65 & -2.07 & -1.23 & year\_classcode-scau-2004\\
-2.12 & -2.51 & -1.73 & year\_classcode-scau-2005\\
\addlinespace
-1.79 & -2.16 & -1.43 & year\_classcode-scau-2006\\
-1.92 & -2.28 & -1.57 & year\_classcode-scau-2007\\
-2.06 & -2.44 & -1.69 & year\_classcode-scau-2008\\
-1.82 & -2.16 & -1.48 & year\_classcode-scau-2009\\
-0.94 & -1.25 & -0.63 & year\_classcode-scau-2010\\
\addlinespace
-0.53 & -0.82 & -0.25 & year\_classcode-scau-2011\\
-1.37 & -1.72 & -1.03 & year\_classcode-scau-2012\\
-1.32 & -1.69 & -0.95 & year\_classcode-scau-2013\\
-1.43 & -1.81 & -1.06 & year\_classcode-scau-2014\\
-2.03 & -2.52 & -1.53 & year\_classcode-scau-2015\\
\addlinespace
-1.66 & -2.07 & -1.26 & year\_classcode-scau-2016\\
-1.56 & -2.01 & -1.12 & year\_classcode-scau-2017\\
-1.72 & -2.70 & -0.74 & year\_classcode-schr-2000\\
-1.80 & -2.74 & -0.86 & year\_classcode-schr-2001\\
-1.80 & -2.64 & -0.95 & year\_classcode-schr-2002\\
\addlinespace
-2.23 & -2.83 & -1.62 & year\_classcode-schr-2003\\
-1.62 & -2.03 & -1.21 & year\_classcode-schr-2004\\
-1.87 & -2.22 & -1.51 & year\_classcode-schr-2005\\
-1.92 & -2.29 & -1.54 & year\_classcode-schr-2006\\
-1.73 & -2.07 & -1.39 & year\_classcode-schr-2007\\
\addlinespace
-1.85 & -2.21 & -1.50 & year\_classcode-schr-2008\\
-1.23 & -1.53 & -0.94 & year\_classcode-schr-2009\\
-1.34 & -1.69 & -1.00 & year\_classcode-schr-2010\\
-1.20 & -1.53 & -0.87 & year\_classcode-schr-2011\\
-1.40 & -1.74 & -1.06 & year\_classcode-schr-2012\\
\addlinespace
-1.61 & -2.01 & -1.21 & year\_classcode-schr-2013\\
-1.24 & -1.60 & -0.89 & year\_classcode-schr-2014\\
-1.43 & -1.83 & -1.02 & year\_classcode-schr-2015\\
-1.72 & -2.13 & -1.32 & year\_classcode-schr-2016\\
-1.53 & -1.96 & -1.09 & year\_classcode-schr-2017\\
\addlinespace
1.19 & 0.69 & 1.70 & year\_classcode-smys-2000\\
0.68 & 0.21 & 1.16 & year\_classcode-smys-2001\\
0.47 & 0.05 & 0.89 & year\_classcode-smys-2002\\
0.45 & 0.16 & 0.74 & year\_classcode-smys-2003\\
-0.20 & -0.49 & 0.08 & year\_classcode-smys-2004\\
\addlinespace
0.27 & 0.04 & 0.50 & year\_classcode-smys-2005\\
0.29 & 0.05 & 0.52 & year\_classcode-smys-2006\\
-0.51 & -0.76 & -0.26 & year\_classcode-smys-2007\\
-0.56 & -0.81 & -0.30 & year\_classcode-smys-2008\\
-0.68 & -0.93 & -0.42 & year\_classcode-smys-2009\\
\addlinespace
-0.55 & -0.83 & -0.27 & year\_classcode-smys-2010\\
0.49 & 0.24 & 0.74 & year\_classcode-smys-2011\\
-0.14 & -0.40 & 0.11 & year\_classcode-smys-2012\\
0.22 & -0.04 & 0.48 & year\_classcode-smys-2013\\
0.81 & 0.56 & 1.06 & year\_classcode-smys-2014\\
\addlinespace
0.32 & 0.05 & 0.58 & year\_classcode-smys-2015\\
-0.33 & -0.61 & -0.05 & year\_classcode-smys-2016\\
-0.51 & -0.83 & -0.20 & year\_classcode-smys-2017\\
0.66 & 0.27 & 1.05 & year\_classcode-spul-2000\\
0.93 & 0.57 & 1.28 & year\_classcode-spul-2001\\
\addlinespace
0.22 & -0.09 & 0.53 & year\_classcode-spul-2002\\
0.39 & 0.17 & 0.61 & year\_classcode-spul-2003\\
0.15 & -0.05 & 0.35 & year\_classcode-spul-2004\\
0.61 & 0.46 & 0.77 & year\_classcode-spul-2005\\
0.57 & 0.41 & 0.73 & year\_classcode-spul-2006\\
\addlinespace
0.47 & 0.32 & 0.62 & year\_classcode-spul-2007\\
0.29 & 0.13 & 0.44 & year\_classcode-spul-2008\\
0.96 & 0.82 & 1.11 & year\_classcode-spul-2009\\
0.71 & 0.53 & 0.88 & year\_classcode-spul-2010\\
0.69 & 0.52 & 0.87 & year\_classcode-spul-2011\\
\addlinespace
0.39 & 0.21 & 0.56 & year\_classcode-spul-2012\\
0.57 & 0.38 & 0.76 & year\_classcode-spul-2013\\
0.19 & 0.00 & 0.37 & year\_classcode-spul-2014\\
0.71 & 0.53 & 0.89 & year\_classcode-spul-2015\\
1.41 & 1.24 & 1.58 & year\_classcode-spul-2016\\
\addlinespace
1.37 & 1.20 & 1.55 & year\_classcode-spul-2017\\
0.00 & 0.00 & 0.00 & region\_cluster-1-ANA\\
0.00 & 0.00 & 0.00 & region\_cluster-1-SCI\\
0.00 & 0.00 & 0.00 & region\_cluster-1-SMI\\
0.00 & 0.00 & 0.00 & region\_cluster-1-SRI\\
\addlinespace
-0.79 & -1.04 & -0.54 & region\_cluster-3-ANA\\
-0.46 & -0.67 & -0.24 & region\_cluster-3-SCI\\
0.16 & -0.06 & 0.38 & region\_cluster-3-SMI\\
0.04 & -0.18 & 0.26 & region\_cluster-3-SRI\\
0.00 & 0.00 & 0.00 & region\_cluster-1-ANA\\
\addlinespace
0.00 & 0.00 & 0.00 & region\_cluster-1-SCI\\
0.00 & 0.00 & 0.00 & region\_cluster-1-SMI\\
0.00 & 0.00 & 0.00 & region\_cluster-1-SRI\\
-2.36 & -2.57 & -2.15 & region\_cluster-3-ANA\\
-1.03 & -1.22 & -0.84 & region\_cluster-3-SCI\\
\addlinespace
1.63 & 1.43 & 1.83 & region\_cluster-3-SMI\\
0.93 & 0.74 & 1.13 & region\_cluster-3-SRI\\
-0.19 & -0.68 & 0.31 & mpa\_effect\\
-0.08 & -0.55 & 0.40 & mpa\_effect\\
0.17 & -0.29 & 0.63 & mpa\_effect\\
\addlinespace
0.18 & -0.25 & 0.61 & mpa\_effect\\
-0.15 & -0.57 & 0.27 & mpa\_effect\\
-0.18 & -0.59 & 0.23 & mpa\_effect\\
0.11 & -0.30 & 0.53 & mpa\_effect\\
-0.13 & -0.54 & 0.29 & mpa\_effect\\
\addlinespace
0.10 & -0.32 & 0.53 & mpa\_effect\\
-0.04 & -0.45 & 0.37 & mpa\_effect\\
0.00 & -0.42 & 0.42 & mpa\_effect\\
0.29 & -0.13 & 0.71 & mpa\_effect\\
0.25 & -0.16 & 0.67 & mpa\_effect\\
\addlinespace
0.24 & -0.19 & 0.66 & mpa\_effect\\
0.38 & -0.05 & 0.81 & mpa\_effect\\
-0.19 & -0.63 & 0.24 & mpa\_effect\\
-0.15 & -0.57 & 0.28 & mpa\_effect\\
-0.39 & -0.82 & 0.05 & mpa\_effect\\*
\end{longtable}

Figures S\ref{fig:fe-plot}:S\ref{fig:targ-plot} present estimated effects for covariates included in the model, along with the raw estimated mean trends of the targeted and non-targeted species (while the difference between these trends is presented in our main results).

\newpage

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/fe-plot-1.pdf}
\caption{\label{fig:fe-plot}Estimated coefficients for non-spatial fixed effects in observation model (seeing) and observed model (seen)}
\end{figure}

\pagebreak

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/site-plot-1.pdf}
\caption{\label{fig:site-plot}Estimated coefficients for spatial random effects in observation model (seeing) and observed model (seen)}
\end{figure}

\pagebreak

\pagebreak

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/targ-plot-1.pdf}
\caption{\label{fig:targ-plot}Trends in standardized mean abundance of targeted and non-targeted species}
\end{figure}

\pagebreak

\hypertarget{regression-diagnostics}{%
\subsubsection{Regression Diagnostics}\label{regression-diagnostics}}

We include visual diagnostics of our estimation model. All coefficients passed convergence criteria for TMB.

Looking first at the predictions of the model for the positive observations in the data (i.e.~using the full model to predict biomass densities, and then comparing those predictions to cases where some positive biomass densities were observed), the model diagnostics show no clear problems. The \(R^2\) of the model is 0.43. Residuals do not exhibit trends, though some grouping the residuals is evident. The quantile-quantile plot suggest that on the assumption of log-normal errors on the observed densities is reasonable, though the model appears to have some slight problems estimating the highest observed densities (Fig.S\ref{fig:obs-v-pred-plot}).

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/obs-v-pred-plot-1.pdf}
\caption{\label{fig:obs-v-pred-plot}High level diagnostics for observed compontent of Delta-GLM: Observed vs predicted log densities (A), predicted log density vs residuals (B), and a normal qq-plot of the residuals (C)}
\end{figure}

In order to evaluate the ability of the model to estimate positive observations, we can compare the the binned predicted probability of a positive observation to the proportion of observations in that bin that recorded positive observations. If our model is working well, we would expect a group of fisheries that our model estimates on average should have a 50\% probability of a positive observation, then we should expect about 50\% of those observations to have positive observations. This is indeed what we see from the model (Fig.S\ref{fig:obs-prob-plot}).

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/obs-prob-plot-1.pdf}
\caption{\label{fig:obs-prob-plot}Binned mean predicted probability of detection provided by the first stage of the hurdle model vs observed proportion of positive detections}
\end{figure}

We can also examine the receiver-operator-curve (ROC) to assess the performance of the observation component of the model. The area under the curve (AUC) value for the model is 0.84 (on a scale of 0.5 to 1), indicating the model is an overall good predictor of whether or not a given observation of biomass densities will be positive or not.

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/roc-plot-1.pdf}
\caption{\label{fig:roc-plot}Receiver operating characteristic curve of predictions of positive biomass densities}
\end{figure}

\hypertarget{standardized-abundance-indices}{%
\subsubsection{Standardized Abundance Indices}\label{standardized-abundance-indices}}

Overall most species showed consistent trends in biomass densities across the different islands at which they have been observed (Fig.S\ref{fig:reg-pop}).

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/reg-pop-1.pdf}
\caption{\label{fig:reg-pop}Mean density by island by year for each fish species included in the analysis}
\end{figure}

The standardized indices of abundance generally did not very substantially from the raw mean densities by species over time. However, for some species, such as blue rockfish, the standardized abundance index suggests much higher biomass densities in the pre-MPA period than those reported in the raw data. We suspect this is largely a function of changes in sampling sites over time, that the standardization is better able to account for (Fig.S\ref{fig:raw-v-stand-plot}).

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/raw-v-stand-plot-1.pdf}
\caption{\label{fig:raw-v-stand-plot}Raw (points) and standardized (lines) indices of abundance for each of the fishes included in the analysis}
\end{figure}

We include a variety of environmental, observation, and temporal indicators in our model. Inclusion of highly co-linear variables in a model can inflate standard errors and obscure ``true'' effects. To account for this we calculated the Pearson's correlation coefficients for all of the continuous data included in our model to ensure that none of the included variables had correlation coefficients greater than 0.7, a general rule of thumb for co-inclusion of variables. We did not find problematic levels of correlation among any of our included continuous variables.

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/numeric-cor-1.pdf}
\caption{\label{fig:numeric-cor}Pearson correlation coefficients of continuos data included in the regression model}
\end{figure}

\hypertarget{alternative-estimation-models}{%
\subsection{Alternative estimation models}\label{alternative-estimation-models}}

Our proposed identification strategy attempts to control for non-MPA (and not directly modeled) related changes in abundances through the trend in the non-targeted species. However, a simpler alternative would be to simply compare densities before-and-after MPA implementation, while explicitly controlling for non-MPA related factors that we believe may have some effect on densities (a ``selection on observables'' strategy). To that end, we fit a mixed-effects regression that models log densities of targeted species only (positive observations only, for the sake of simplicity) as a function of temperature deviations, kelp cover, observer experience, random effects for species and region, and fixed effects for each year in the data (omitting the year 2000). The hypothesis here is that any non-MPA related factors that affect densities are accounted for in the observed variables included in the model.

Using this model, densities of targeted species appear to have been declining steadily since 2000, and appear to have plateaued off since the implementation of MPAs in 2003. Without an identification strategy such as the one employed in this study then, all we could conclude is that densities appear to be lower post-MPA, and have not increased substantially over time (Fig.S\ref{fig:obs-plot}).

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/obs-plot-1.pdf}
\caption{\label{fig:obs-plot}Selection on observables identification strategy. Plotted estimates are fixed effects of year on log-density (relative to the year 2000), controlling for observer experience, temperature deviations, and kelp cover, with random effects for species and region}
\end{figure}

The estimation model used for our main results is complicated. We feel this complexity is justified in order to best capture the uncertainty inherent in the challenging task of conducting underwater visual surveys, as well as the spatio-temporal nature of the underlying data. However, we also ran several simpler models in order to examine the sensitivity of our results to the model structure selected here.

Much of the complexity of our model comes from the integration of a standardized abundance index for each of the species, which are then compiled into a standardized index of abundance index for targeted species as a whole.

In one simpler approach, we aggregated the PISCO data into mean biomass densities by species and year across all the Channel Islands. Since all included species have at least some positive observations in each year, this removes the complication of dealing with the probability of detection problem in the raw data, while of course assuming that the net effects of all of the other factors included in our main model are on average zero (e.g.~observer skill, visibility, etc.). A difference-in-difference model fit to these simplified data show qualitatively similar (though much more uncertain) results to our full model. In particular, the model shows the same gradual increase in targeted densities relative to non-targeted until 2013, followed by an attenuation of this trend. Importantly, though, this simplified model fails to correct for some pre-MPA differences in the two groups, leading to negative ``effects'' of the MPAs being estimated before the MPAs themselves went in place. While these sort of anticipatory effects are certainly possible (McDermott et al. 2019), in this case we would suggest the likelier explanation---given that these anticipatory effects disappear in the full model---is that our full model provides important controls for pre-MPA characteristics (Fig.S\ref{fig:sortasimple-did})

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/sortasimple-did-1.pdf}
\caption{\label{fig:sortasimple-did}Results of simplified difference-in-difference regression. The model estimates the difference in the mean trend of densities of targeted and non-targeted species over time, controlling for the mean densities of each individual species group. Data are initially aggregated to the level of species-by-year.}
\end{figure}

As an even simpler approach, we aggregated the data to the level of targeted and non-targeted species, and estimated the divergence in their trends over time.

\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/unnamed-chunk-7-1.pdf}

We can also explore the effects of our species filtering by running a simplified regression but now including all the species in the database, no matter of how infrequently they are observed.

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/sortasimple-noilter-did-1.pdf}
\caption{\label{fig:sortasimple-noilter-did}Results of simplified difference-in-difference regression including all observed finfish. The model estimates the difference in the mean trend of densities of targeted and non-targeted species over time, controlling for the mean densities of each individual species group. Data are initially aggregated to the level of species-by-year.}
\end{figure}

\hypertarget{synthetic-controls}{%
\subsubsection{Synthetic controls}\label{synthetic-controls}}

Synthetic controls are an alternative method for attempting to estimate the causal effect of a policy intervention (Abadie, Diamond, and Hainmueller 2010). A difference-in-difference approach assumes that some observable group serves as an adequate control for the state of the treated group in an untreated world. In our default case, we assume that the mean standardized index of non-targeted species are our control for the targeted species. Alternatively, synthetic controls use timeseries of treated and non-treated groups before and after treatment to construct a new ``control'' group built by weighting the pre-treatment timeseries of un-treated observations (together with covariates) such that the synthetic control group matches the trends in the treated group pre-treatment.

We chose to present difference-in-difference as our main result since it better allows us to capture the uncertainty in the data generating process through our hierarchical model. However, we felt that it was worth exploring whether synthetic controls provided substantially different results than our default model.

For the first synthetic control, we pulled our standardized mean index of abundance for targeted species as a whole from our difference-in-difference model as our treated group. We then pull the standardized indices of abundance for each of the non-targeted groups from the difference-in-difference to use as the candidate untreated components for the synthetic controls. A complete synthetic control analysis would require more extensive validation of the methods, but we use this approach simply to explore whether we observe substantially divergent results in the synthetic control versus the difference-in-difference model.

We centered and scaled the candidate abundance indices to facilitate model convergence given the very few number of pre-treatment years available. The results of a synthetic control model are presented as the difference between the observed treatment outcome and the synthetic control (the difference in this case being in units of standard deviations). The model was not able to construct an adequate synthetic control at this level, as shown by the differences between the treated group and the synthetic control pre-treatment. However, we would note that the post-treatment results do show similarities with our main results, namely a lack of a clear divergence between the treatment and the control, and an upwards trend up through the earl 2010s followed by a decline (Fig.S\ref{fig:total-synth}).

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/total-synth-1.pdf}
\caption{\label{fig:total-synth}Difference in centered and scaled standardized targeted abundance and synthetic standardized targeted abundance}
\end{figure}

As an extension, we repeated this process, but now treating each targeted species individually as the treated group, and the non-targeted species as the non-targeted. This is intended to explore whether we see clearer signals for individual species than we do for the targeted class as a whole.

Overall we see similarly unclear results as the aggregate targeted synthetic control (and our main results). The synthetic control was better constructed for some individual species, but not clearly for any one, and most species showed some evidence of the upward-then-downward trend seen throughout our results (Fig.S\ref{fig:synth-classcode-fits}).

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/synth-classcode-fits-1.pdf}
\caption{\label{fig:synth-classcode-fits}Synthetic control gaps for each targeted species}
\end{figure}

\hypertarget{testing-model-assumptions}{%
\subsection{Testing Model Assumptions}\label{testing-model-assumptions}}

\hypertarget{simulation-testing}{%
\subsubsection{Simulation testing}\label{simulation-testing}}

We state that a difference-in-difference model using targeted and non-targeted species is capable (conditional on assumptions) of estimating the causal effect of MPAs. We simulated MPA outcomes to test this claim. We first test our estimation strategy under idealized circumstances, where recruitment is deterministic and PISCO divers all have constant and perfect observer skills. We simulate five species that vary only in their maximum size and length at maturity. For each of these species, we set one version that is targeted by fishing and one that is not. We set a constant fishing mortality rate for each simulated targeted species, and then ran two matched simulations, one with MPAs and one without. We then have our simulated divers sample data from each of these scenarios, and then pass the sampled biomass densities to a simplified version of our difference-in-difference model (omitting the probability of detection step). We can then compare the difference-in-difference estimates of the MPA effect to the true simulated effect. The difference in difference model is able to capture the simulated MPA effect under these circumstances (Fig.S\ref{fig:simple-did-test})

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/simple-did-test-1.pdf}
\caption{\label{fig:simple-did-test}Simulated mean (red dashed line) and individual species (solid lines) MPA effects over time, along with difference-in-difference estimated MPA effects (mean with 95\% confidence intervals)}
\end{figure}

We then simulated a more complex example. We use the actual targeted and non-targeted species from our model. We assign species predominately seen in the western Channel Islands as ``cold water'' and those in the eastern Channel Islands as ``warm water''. We allow for stochasticity in recruitment. We use El Niño data as a simulated environmental recruitment driver, where we assume that El Niño events produce negative recruitment shocks for cold water species and \emph{vice versa} for warm water species. We simulate three different divers each with different base skill levels, visual selectivities, and an evolving skill rate (such that observers get better over time). We hold fishing mortality rates constant for each species, although that fishing mortality affects each species differently because of intrinsic biological differences in maturity-at-age and steepness. We then test the ability of the difference-in-difference model to isolate the mean MPA effect across all of these targeted species, which our results show it is capable of doing (Fig.S\ref{fig:complex-did-test}).

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/complex-did-test-1.pdf}
\caption{\label{fig:complex-did-test}Simulated mean (red dashed line) and individual species (solid lines) MPA effects over time, along with difference-in-difference estimated MPA effects (mean with 95\% confidence intervals)}
\end{figure}

\hypertarget{sensitivity-to-missing-observations}{%
\subsubsection{Sensitivity to ``missing'' observations}\label{sensitivity-to-missing-observations}}

Appropriately addressing the problem of ``missing'' observations is a critical challenge in any field observation study. If no observations of a given fish species were recorded on a given transect, should the density of that species on that transect be marked as zero, and influence the estimate of the overall mean density accordingly? The obvious answer seems to be yes, but what if that species simply does not live in the environment covered by a particular transect, or was not present during the particular time of the diver's observation? For our base runs, we assign a value of zero density on a given transect for any fish species that has been observed at least once at a given site at any time in our data but was not observed on that particular transect. If that species was never observed at that site, we do not include a zero for that species. Our rationale for this is that given the shifting nature of the sampled sites, and the intensity of sampling at those sites, we do not want to skew density trends by changes in the amount of suitable habitat for a given species sampled. However, this is clearly a strong assumption. For example, perhaps the decreasing trend in mean densities from 2000 to 2004 is due to increased number of sites (and therefore zeros) included in the data. To assess the potential importance of this choice, we can compare the mean densities of targeted and non-targeted species over time with the added zeros to the mean densities using only positive observations (i.e.~not including any zeros in the data, (Fig.S\ref{fig:nozero-raw-trend}). The trends in the raw densities, and most importantly the mean trends of targeted and non-targeted fishes, are nearly identical whether or not zeros are added, providing strong evidence that our choice of how to incorporate missing observations into the data are not strongly influencing our overall results.

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/nozero-raw-trend-1.pdf}
\caption{\label{fig:nozero-raw-trend}Centered and scaled mean annual density, excluding zeros, of included fishes (points) and smoothed means of targeted and non-targeted groups (line) over time}
\end{figure}

\hypertarget{testing-sutva-with-convergent-cross-mapping}{%
\subsubsection{Testing SUTVA with Convergent Cross Mapping}\label{testing-sutva-with-convergent-cross-mapping}}

The difference-in-difference model also assumes that the targeted and non-targeted fishes do not directly or indirectly affect each other.This assumption is clearly violated on some level: all the fishes in this analysis are part of the same ecosystem and therefore interact to some degree. For example, if the protection of targeted predatory fishes results in increased mortality of non-targeted fishes, the model would attribute that as an increased regional effect (greater divergence between the abundance of targeted and non-targeted species). Given the time scale of analysis (15 years of protection), we do not feel that massive trophic cascades are likely to have developed yet, given both the pace and complexity of trophic cascade development (Babcock et al. 2010; Pershing et al. 2015). A complete assessment of evidence for trophic cascades in the Channel Islands is beyond the scope of this study, but to address this question somewhat we utilized convergent cross mapping \emph{sensu} Sugihara et al. (2012) to test for a significant causal signal between different broad trophic groups in the data, implemented in the \texttt{rEDM} package in R.

Convergent cross mapping is a nonlinear forecasting method that uses observed time series data to test for significant causal links between variables. Following methods laid out in Clark et al. (2015) and Sugihara et al. (2012), we pool the abundance of each broad trophic group by region (Fig.S\ref{fig:trophic-plot}. This uses the data from the islands as ``replicates'', requiring the assumption that the islands are all part of the same dynamic system, but allowing us to take advantage of the extra information provided by each island to further resolve the reconstructed manifolds. Using these aggregations, we then test whether the variables can be properly embedded, i.e., if they have predictable manifold dynamics. We do this through a simplex forecasting test, using an individual timeseries' own lags to build a manifold. For each timeseries, the ``best embedding dimension'' is an approximation of the dimensionality of the dynamic system, in other words, the number of dimensions that define and predict the evolving states of the timeseries. This analysis shows that only the carnivore, piscivore, and planktivores groups show evidence of significant predictability (that is, that past dynamics of these species groups can predict future dynamics, Fig.S\ref{fig:embed-plot}).

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/trophic-plot-1.pdf}
\caption{\label{fig:trophic-plot}Centered and scaled densities by broad trophic group and island over time}
\end{figure}

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/embed-plot-1.pdf}
\caption{\label{fig:embed-plot}Predictive skill as a function of embedding dimensions}
\end{figure}

Focusing on just these three groups then (removing herbivores), we can test for causal relationships between groups using convergent cross mapping and the logic of Takens' theorem of dynamic systems. Generalizations of Takens' theorem indicate that if two variables (in our case, species or physical variables) are part of the same dynamic system, their individual dynamics should reflect their relative causal influence (Sugihara et al. 2012). In other words, if one variable is causally forced by another, that forcing should leave a signature on the first time series. Convergent cross mapping (CCM) tests for causation by using the attractor/manifold built from the time series of one variable to predict another (hence the ``cross-mapping''). In simple terms, the \emph{causal effect of A on B is determined by how well B cross-maps A}.

There are two criteria for CCM to establish causality: First, and most obviously, predictive cross-map skill using all available data should be significantly greater than zero. Second, that predictability should be convergent. Convergence means that cross-mapped estimates improve with library length (the number of state-space vectors used to build the attractor), because the attractor is more fully resolved and therefore estimation error should decline. Convergence is key to distinguishing causation from simple or spurious correlation. If two variables are spuriously correlated and not causally linked, CCM should fail to satisfy this second criterion. Based on these criteria, there is little evidence of significant dynamic interactions between trophic groups (Fig.S\ref{fig:cross-map-cp}-\ref{fig:cross-map-pis}). Cross-mapping produced positive predictive skill, but was non-convergent for all cross-mappings with the exception of carnivores cross-mapping planktivores (providing some evidence that planktivore dynamics may be a driver of carnivore dynamics). This analysis provides evidence that trophic cascades are unlikely to be a significant driver of our results. It is important to note though that this analysis does not mean that trophic cascades could not emerge in this system, rather that we do not detect them with these data at this time.

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/cross-map-cp-1.pdf}
\caption{\label{fig:cross-map-cp}Cross mapping of effect of piscivores on carnivores (A) and carnivores on piscivores (B) in the PISCO data from 2000 to 2017. Shaded region show 95\% confidence interval}
\end{figure}

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/cross-map-pl-1.pdf}
\caption{\label{fig:cross-map-pl}Cross mapping of effect of planktivores on carnivores (A) and carnivores on planktivores (B) in the PISCO data from 2000 to 2017. Shaded region show 95\% confidence interval}
\end{figure}

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/cross-map-pis-1.pdf}
\caption{\label{fig:cross-map-pis}Cross mapping of effect of piscivores on carnivores (A) and carnivores on planktivores (B) in the PISCO data from 2000 to 2017. Shaded region show 95\% confidence interval}
\end{figure}

\hypertarget{repeat-analysis-with-kelp-forest-monitoring-data}{%
\subsection{Repeat Analysis with Kelp Forest Monitoring Data}\label{repeat-analysis-with-kelp-forest-monitoring-data}}

As a robustness check to our main results, we repeated our analysis utilizing data provided by the \href{https://science.nature.nps.gov/im/units/medn/monitor/kelpforest.cfm}{Kelp Forest Monitoring Program (KFM)} conducted in the Channel Islands. Despite having similar but different methods and survey locations, we find almost identical estimated trends in divergences between targeted and non-targeted species using the KFM data (Fig.S\ref{fig:kfm-did}).

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/kfm-did-1.pdf}
\caption{\label{fig:kfm-did}Estimated divergence in biomass densities of targeted and non-targeted fishes in the Channel Islands (i.e.~integrated across inside and outside of MPAs) using the KFM data . MPAs are implemented in 2003 (red dashed line). Estimates are from a regression on log(abundance index), and so estimated effects roughly correspond to percentage changes}
\end{figure}

\hypertarget{mpa-only-analysis}{%
\subsection{MPA Only Analysis}\label{mpa-only-analysis}}

Given trends in mean densities observed in the raw data, the ``regional conservation effect'' estimated by our model---defined as the divergence in trends between the targeted and non-targeted species across the Channel Islands region---is not surprising. By jumping through countless statistical hoops we reach a similar conclusion that we would just by looking at the divergences in the mean trends. The integration of data from inside and outside of MPAs is a possible explanation for this lack of a clear regional effect. If spillover is limited or has simply not developed yet, especially relative to the effect of fishing outside of MPAs, then it is possible that there is a clear positive effect inside the MPAs, a clear negative effect outside, and when we look across both types of sites we get an unclear average of the two.

To address this, we can first repeat some exploratory data analysis of trends in densities inside and outside the MPAs for targeted and non-targeted species. Caselle et al. (2015) provides a thorough look at this question of differences inside and outside of MPAs. We update that analysis here to account for our specific questions of trend divergence, potential differences in filtering methods, to include data up through 2017, and to utilize our estimation method on just the inside-MPA data. For all exploratory analyses, we consider the same top 23 consistently observed species. Looking first at simple trends in total mean biomass density across these species inside and outside of MPAs, we find evidence that biomass densities inside the MPAs is increasing faster (and is higher inside) than outside (Fig.S\ref{fig:bio-mpa}).

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/bio-mpa-1.pdf}
\caption{\label{fig:bio-mpa}Annual mean aggregate biomass density (summed across all fishes) inside and outside of eventual MPA locations over time. Red dashed line indicates MPA implementation year}
\end{figure}

Our proposed identification strategy here though is not that total biomass density should be different inside and outside, but that the non-targeted species should serve as the control to the targeted. If we believe that the MPA effects are greater inside the MPA, then we would expect to see stronger divergences in biomass densities between these two targeted and non-targeted fishes inside the MPAs than outside

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/div-mpa-1.pdf}
\caption{\label{fig:div-mpa}Trends in total biomass density inside and outside of eventual MPAs for targeted and non-targeted fishes. Red dashed line indicates MPA implementation year}
\end{figure}

Here we see a different picture. While there is some visual evidence that the targeted species were diverging from the non-targeted faster inside the MPAs than outside, both inside and outside we see that the trend in total biomass density of targeted species is trending downward, relative to the trend in the non-targeted species in recent years. This analysis is of total biomass density. However, our model estimates the mean difference in targeted and non-targeted species. Both have their advantages, but we chose the mean to reflect a hypothesis that the MPAs would provide positive benefits across all targeted species. The total biomass density could be strongly affected by a sharp increase or decrease in one or two species, even if the mean trend is different. Examining the mean trends though, we see the same results (Fig.S\ref{fig:mean-div-mpa}).

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/mean-div-mpa-1.pdf}
\caption{\label{fig:mean-div-mpa}Trends in mean total biomass density inside and outside of eventual MPAs for targeted and non-targeted fishes. Red dashed line indicates MPA implementation year}
\end{figure}

Lastly, we can examine both the mean and individual trend to check clear species-by-species outliers in the overall biomass density trends. This analysis shows noise, but overall the targeted and non-targeted species seem to be following similar trends within their respective groups

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/cs-trends-1.pdf}
\caption{\label{fig:cs-trends}Centered and scaled biomass density trends for each fish grouped by targeted and non targeted (pale lines) and fitted LOESS smoother (with 95\% confidence intervals around mean) and mean by targeted and non-targeted groups, inside and outside od MPAs. Red dashed line indicates MPA implementation year}
\end{figure}

These visual assessments suggest that similar to our results looking both inside and outside of MPAs, we would expect that our estimation model fitted only on data from inside eventual MPAs would reach similar conclusions as our results fitted to data from both inside and outside MPAs. To test this, we re-ran our analysis, but only using data from sites that are eventually placed inside MPAs. Our results reflect the same trends as displayed in the raw data and the statistical region-wide analysis, providing robust statistical support to the conclusions we would reach from visually examining the raw data (Fig.S \ref{fig:mpa-did}).

\begin{figure}
\centering
\includegraphics{ovando-regional-effects-of-mpas-si_files/figure-latex/mpa-did-1.pdf}
\caption{\label{fig:mpa-did}Estimated divergence in biomass densities of targeted and non-targeted fishes inside eventual Channel Islands MPAs. MPAs are implement in 2003 (red dashed line). Estimates are from a regression on log(abundance index), and so estimated effects roughly correspond to percentage changes}
\end{figure}

\hypertarget{references}{%
\subsection*{References}\label{references}}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-abadie2010}{}%
Abadie, Alberto, Alexis Diamond, and Jens Hainmueller. 2010. ``Synthetic Control Methods for Comparative Case Studies: Estimating the Effect of California's Tobacco Control Program.'' \emph{Journal of the American Statistical Association} 105 (490): 493--505. \url{https://doi.org/10.1198/jasa.2009.ap08746}.

\leavevmode\hypertarget{ref-babcock2010}{}%
Babcock, R. C., N. T. Shears, A. C. Alcala, N. S. Barrett, G. J. Edgar, K. D. Lafferty, T. R. McClanahan, and G. R. Russ. 2010. ``Decadal Trends in Marine Reserves Reveal Differential Rates of Change in Direct and Indirect Effects.'' \emph{Proceedings of the National Academy of Sciences} 107 (43): 18256--61. \url{https://doi.org/10.1073/pnas.0908012107}.

\leavevmode\hypertarget{ref-caselle2015}{}%
Caselle, Jennifer E., Andrew Rassweiler, Scott L. Hamilton, and Robert R. Warner. 2015. ``Recovery Trajectories of Kelp Forest Animals Are Rapid yet Spatially Variable Across a Network of Temperate Marine Protected Areas.'' \emph{Scientific Reports} 5: 14102.

\leavevmode\hypertarget{ref-clark2015}{}%
Clark, Adam Thomas, Hao Ye, Forest Isbell, Ethan R. Deyle, Jane Cowles, G. David Tilman, and George Sugihara. 2015. ``Spatial Convergent Cross Mapping to Detect Causal Relationships from Short Time Series.'' \emph{Ecology} 96 (5): 1174--81. \url{https://doi.org/10.1890/14-1479.1}.

\leavevmode\hypertarget{ref-leeper2018}{}%
Leeper, Thomas J. 2018. ``Tabulizer: Bindings for Tabula PDF Table Extractor Library.''

\leavevmode\hypertarget{ref-lter2017}{}%
LTER, Santa Barbara Coastal, Tom W Bell, Kyle C Cavanaugh, and David A Siegel. 2017. ``SBC LTER: Time Series of Quarterly NetCDF Files of Kelp Biomass in the Canopy from Landsat 5, 7 and 8, 1984 - 2016 (Ongoing).'' Environmental Data Initiative. \url{https://doi.org/10.6073/pasta/817d2c24ebd78621869e17d94ba0df0c}.

\leavevmode\hypertarget{ref-mcdermott2019}{}%
McDermott, Grant R., Kyle C. Meng, Gavin G. McDonald, and Christopher J. Costello. 2019. ``The Blue Paradox: Preemptive Overfishing in Marine Reserves.'' \emph{Proceedings of the National Academy of Sciences} 116 (12): 5319--25. \url{https://doi.org/10.1073/pnas.1802862115}.

\leavevmode\hypertarget{ref-pershing2015a}{}%
Pershing, Andrew J., Michael A. Alexander, Christina M. Hernandez, Lisa A. Kerr, Arnault Le Bris, Katherine E. Mills, Janet A. Nye, et al. 2015. ``Slow Adaptation in the Face of Rapid Warming Leads to Collapse of the Gulf of Maine Cod Fishery.'' \emph{Science} 350 (6262): 809--12. \url{https://doi.org/10.1126/science.aac9819}.

\leavevmode\hypertarget{ref-punt2000}{}%
Punt, André E., Terence I. Walker, Bruce L. Taylor, and Fred Pribac. 2000. ``Standardization of Catch and Effort Data in a Spatially-Structured Shark Fishery.'' \emph{Fisheries Research} 45 (2): 129--45. \url{https://doi.org/10.1016/S0165-7836(99)00106-X}.

\leavevmode\hypertarget{ref-sugihara2012}{}%
Sugihara, George, Robert May, Hao Ye, Chih-hao Hsieh, Ethan Deyle, Michael Fogarty, and Stephan Munch. 2012. ``Detecting Causality in Complex Ecosystems.'' \emph{Science} 338 (6106): 496--500. \url{https://doi.org/10.1126/science.1227079}.

\leavevmode\hypertarget{ref-thorson2017c}{}%
Thorson, James T., Stephan B. Munch, Jason M. Cope, and Jin Gao. 2017. ``Predicting Life History Parameters for All Fishes Worldwide.'' \emph{Ecological Applications}, n/a--n/a. \url{https://doi.org/10.1002/eap.1606}.

\end{document}

### 2019-10-09 documents/ovando-regional-effects-of-mpas.Rmd
Initial commit of cleaned up regional effects repo

---
title: The Regional Effects of Marine Protected Areas
author:
  - name: Daniel Ovando
    affiliation: a,1
    email: danovan@uw.edu
  - name: Jennifer E. Caselle
    affiliation: b
  - name: Christopher Costello
    affiliation: b   
  - name: Olivier Deschenes
    affiliation: b  
  - name: Steven D. Gaines
    affiliation: b
  - name: Ray Hilborn
    affiliation: a
  - name: Owen Liu
    affiliation: b        
address:
  - code: a
    address: University of Washington, School of Aquatic and Fishery Sciences
  - code: b
    address: University of California, Santa Barbara
corresponding_author:
  - code: 1
    text: "To whom correspondence should be addressed. E-mail: danovan@uw.edu"
lead_author_surname: Ovando
author_contributions: |
  D.O., S.G., and R.H. developed structure of simulation model. D.O., O.D., C.C., and J.C. developed estimation strategy. J.C. provided support on collecting and interpreting data. All simulations, statistics, and sensitivity analyses performed by D.O. and O.L. . All authors contributed to the manuscript. 
conflict_of_interest: |
  R.H's research program receives funding from environmental NGOs, foundations, fishing industry, governments and international agencies. All of these can be interpreted as a conflict of interest when evaluating fisheries policy. C.C., S.G., and O.D's research program includes funding from environmental NGOs and foundations with an interest in ocean conservation and food security
abstract: |
  Marine Protected Areas (MPAs) cover between 3-7% of the world's oceans, up from less than 1% in the year 2000. The Convention on Biological Diversity calls for 10% of coastal waters to be protected inside MPAs by 2020, while the International Union for Conservation of Nature calls for 30% protection by 2030. It is often clear that MPAs produce conservation benefits inside their borders, but many MPAs are also justified on the grounds that they also benefit the broader region outside their borders. Despite the expanding role of MPAs in marine resource management, we lack a clear understanding of what the regional effects of MPAs might be, and how we might detect them in scientific studies. We used a bio-economic simulation model to demonstrate how the regional effects of MPAs are affected by a suite of economic and environmental drivers and show how commonly used methods to measure inside-MPA effects are likely to fail to capture the regional impacts of protection. We present an empirical strategy for estimating regional MPA effects on biomass density using 16 years of data from inside and outside of MPAs within the Channel Islands, California, USA. We find no statistically clear effect of MPAs on aggregate mean biomass densities of targeted finfish throughout the study region. To make meaningful progress on MPA design and evaluation, the next generation of MPA science must address the challenge of understanding and estimating the regional effects of MPAs across the diversity of species in the ecosystem.
significance: |
  Healthy marine ecosystems are critical to the well-being of the planet. Marine protected areas, parts of the oceans protected from human activities such as fishing, are increasingly being used in an effort to conserve and manage these ecosystems. Our study pairs theory with empirical methods to examine the regional effects of MPAs on conservation outcomes. Using simulations, we show that the effects of marine protected areas can be highly variable and dependent on human drivers. Building off of MPA theory, we find no statistically clear effect of a network of MPAs, which our simulation analysis suggests is entirely reasonable. Our results lay a foundation for future research on the design and monitoring of marine protected areas. 
acknowledgements: |
  This study would not be possible without the work provided by PISCO divers over the years. DO thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments and technical support. 
keywords:
  - Marine Protected Areas
  - Conservation
  - Bio-economic modeling
  - Program Evaluation
## must be one of: pnasresearcharticle (usual two-column layout), pnasmathematics (one column layout), or pnasinvited (invited submissions only)
pnas_type: pnasresearcharticle
bibliography: My Library.bib
csl: pnas.csl
lineno: true
header-includes:
   - \usepackage{eso-pic,graphicx,transparent}
output:
  rticles::pnas_article: default
  bookdown::word_document2: default
  bookdown::pdf_document2: default
  bookdown::html_document2: default
params:
  run_name: ["v5.1"]
  sim_years: [50]
  burn_years: [20]
  num_patches: [25]
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dev = "CairoPDF", dpi = 600,
                      cache = FALSE)
# knitr::opts_knit$set(root.dir = "/Users/danovando/PhD/zissou") # not entirely reprodcible hack to deal with knit child in dissertation.. not idealk

```

<!-- why is it going to be hard ot detect regional effects of MPAs few if any empirical studies at net regional effects. does this mean that i doesn't exist, or impossible tot est, why? three fundamental stumbling blocks -->

```{r, include=FALSE}
library(sf)
library(ggmap)
library(viridis)
library(hrbrthemes)
library(scales)
library(patchwork)
library(rpart)
library(extrafont)
library(caret)
library(ggsci)
library(rEDM)
library(tidyverse)
extrafont::loadfonts()
functions <- list.files(here::here("functions"))

walk(functions, ~ here::here("functions", .x) %>% source()) # load local functions
save_plots <- TRUE

# sim_years <- 50
# 
# burn_years <- 20
# 
# num_patches <- 25

# run_name <- "v5.0"

run_name <- params$run_name

sim_years <- params$sim_years

burn_years <- params$burn_years

num_patches <- params$num_patches

run_dir <- here::here("results", run_name)

experiment_dir <- here::here("results",run_name, "experiments")

load(file = here::here("results",run_name, "rawish_zissou_data.Rdata"))

load(file = here::here("results",run_name, "model_runs.Rdata"))

processed_grid <-  read_rds(file.path(run_dir,"filtered_processed_grid.rds"))

load(file = file.path(run_dir ,"sim_grid.Rdata"))

cdfw_catches <-
  read_csv(file = here::here("data", 'cdfw-catches.csv')) %>%
  group_by(sci_name, year) %>%
  summarise(catch = sum(pounds_caught, na.rm = T)) %>%
  left_join(
    life_history_data %>% select(taxa, classcode, commonname) %>% mutate(taxa = tolower(taxa)),
    by = c("sci_name" = "taxa")
  ) %>%
  filter(!is.na(classcode))

# cdfw_catches %>%
#   group_by(commonname, classcode, sci_name) %>%
#   summarise(catch =sum(catch, na.rm = TRUE)) %>%
#   mutate(in_analysis = classcode %in% unique(cip_data$classcode)) %>%
#   View()

density_ratios <- readRDS(file.path(run_dir,"density_ratios.rds"))

outcomes <- readRDS(file.path(run_dir,"outcomes.rds"))

outcomes <- outcomes %>% 
  filter(!is.na(mpa_effect))

density_ratios <- density_ratios %>% 
  filter(!is.na(biased_density_ratio))

   eqo <- outcomes %>%
     filter(year == max(year))

facet_labels <- c(
  mpa_size = "Range in MPA",
  depletion = "Depletion"
)

channel_islands <- readRDS(here::here("data","channel_islands_map.RDS"))

ca_mpas <- sf::st_read(here::here("data","MPA_CA_Existing_160301")) %>%
  rmapshaper::ms_simplify() %>%
  sf::st_transform(crs = 4326)

zissou_theme <-
  theme_ipsum(
    base_size = 12,
    axis_title_size = 12,
    strip_text_size = 12,
    base_family = "Fira Sans"
  )

testplot <- ggplot()

panel_height = unit(1,"npc") - sum(ggplotGrob(testplot)[["heights"]][-3]) - unit(4,"line")

gc <- guide_colorbar(frame.colour = "black",
                     ticks.colour = "white",
                     barheight = panel_height)

  hgc <- guide_colorbar(frame.colour = "black",
                       ticks.colour = "black",
                       barwidth = 10)

    wide_hgc <- guide_colorbar(frame.colour = "black",
                       ticks.colour = "black",
                       barwidth = 25)

theme_set(zissou_theme)

models_worked <- model_runs$tmb_fit %>% map("error") %>% map_lgl(is_null)

model_runs <- model_runs %>%
  filter(models_worked) %>%
  mutate(tmb_fit = map(tmb_fit,"result")) %>%
  mutate(processed_fits = map(tmb_fit, process_fits)) %>%
  mutate(did_plot = map(processed_fits, "did_plot"))

base_run <- model_runs %>%
filter(var_names == "pisco_a", mpa_only == FALSE, center_scale == TRUE)

mpa_run <- model_runs %>%
  filter(data_source == "pisco",
         var_names == "pisco_a",
         mpa_only == TRUE)

kfm_run <- model_runs %>%
  filter(data_source == "kfm")

fitted_data <- base_run$data[[1]]

zissou_fit <- base_run$tmb_fit[[1]]

report <- base_run$tmb_fit[[1]]

site_data <- read_csv(here::here("data",'Final_Site_Table_UCSB.csv')) %>%
  magrittr::set_colnames(., tolower(colnames(.))) %>%
  select(
    site,
    side,
    mpagroup,
    mpa_status,
    reserve,
    region,
    year_mpa,
    mpaareanm2,
    lat_wgs84,
    lon_wgs84
  ) %>%
  rename(lat_wgs84 = lon_wgs84,
         lon_wgs84 = lat_wgs84) %>%
  unique() %>%
  mutate(eventual_mpa = (year_mpa > 0))

seen_non_nested_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seen_non_nested_betas")) %>%
  rename(group = variable) %>%
  mutate(variable  = zissou_fit$zissou_data$x_seen_non_nested %>% colnames())

seeing_non_nested_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seeing_non_nested_betas")) %>%
  rename(group = variable) %>%
  mutate(variable  = zissou_fit$zissou_data$x_seen_non_nested %>% colnames())

seen_year_species_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seen_year_species_betas")) %>%
  rename(group = variable) %>%
  mutate(variable = zissou_fit$zissou_data$x_seen_year_species %>% colnames())

seeing_year_species_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seeing_year_species_betas")) %>%
  # filter(variable == "seeing_year_species_betas") %>%
  rename(group = variable) %>%
  mutate(variable = zissou_fit$zissou_data$x_seen_year_species %>% colnames())

seen_region_cluster_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seen_region_cluster_betas")) %>%
  # filter(variable == "seen_region_cluster_betas") %>%
  rename(group = variable) %>%
  mutate(variable = zissou_fit$zissou_data$x_seen_region_cluster %>% colnames())

seeing_region_cluster_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seeing_region_cluster_betas")) %>%
  # filter(variable == "seeing_region_cluster_betas") %>%
  rename(group = variable) %>%
  mutate(variable = zissou_fit$zissou_data$x_seen_region_cluster %>% colnames())

did_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "mpa_effect")) %>%
  # filter(variable == "mpa_effect") %>%
  mutate(group = variable) %>%
  mutate(year = zissou_fit$did_data$year %>% unique())

 betas <- bind_rows(
   seen_non_nested_betas,
   seeing_non_nested_betas,
   seen_year_species_betas,
   seeing_year_species_betas,
   seen_region_cluster_betas,
   seeing_region_cluster_betas,
   did_betas %>% select(-year)
 ) %>%
   as_data_frame()

 non_nested_beta_plot <- betas %>%
   filter(str_detect(group, "non_nested")) %>%
   ggplot() +
   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
   geom_pointrange(aes(x = variable,
                       y = estimate,
                       ymin = lower,
                       ymax = upper)) +
   facet_wrap(~group) +
   coord_flip() +
   theme(axis.text.y = element_text(size = 10))

 year_species_effects_plots <- betas %>%
   filter(str_detect(group, "year_species_betas")) %>%
   mutate(year = str_replace_all(variable,"\\D","") %>% as.numeric()) %>%
   mutate(classcode = str_split(variable,'-', simplify = T)[,2]) %>%
   ggplot() +
   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
   geom_ribbon(aes(x = year, ymin = lower, ymax = upper, fill = classcode), alpha = 0.25) +
   geom_line(aes(x = year, y = estimate, color = classcode)) +
   facet_wrap(~group)

  region_cluster_plots <- betas %>%
   filter(str_detect(group, "region_cluster_betas")) %>%
   mutate(cluster = str_replace_all(variable,"\\D","")) %>%
   mutate(region = str_split(variable,'-', simplify = T)[,3]) %>%
   ggplot() +
    geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
    geom_pointrange(aes(x = region,
                        y = estimate,
                        ymin = lower,
                        ymax = upper, color = cluster)) +
    facet_wrap(~group)

top_species <- base_run$data[[1]]$classcode %>% unique()

life_summaries <- life_history_data %>% 
  select(classcode, tm, vbgf.k, vbgf.linf) %>% 
  rename(age_mature = tm, vbk = vbgf.k, linf = vbgf.linf) %>% 
  unique() %>% 
  filter(classcode %in% top_species) %>% 
  select(-classcode) %>% 
  gather(variable,value) %>% 
  group_by(variable) %>% 
  summarise(min_val = min(value, na.rm = TRUE),
         max_val = max(value, na.rm = TRUE))

 sim_life_history <- sim_grid %>%
    mutate(
      age_mature = map_dbl(fish, "age_mature"),
      vbk = map_dbl(fish, "vbk"),
      linf = map_dbl(fish, "linf")
    ) %>%
    select(scientific_name, age_mature, vbk, linf) %>%
    unique() %>% 
   filter(age_mature > life_summaries$min_val[life_summaries$variable == "age_mature"],
          age_mature < life_summaries$max_val[life_summaries$variable == "age_mature"],
          vbk > life_summaries$min_val[life_summaries$variable == "vbk"],
          vbk < life_summaries$max_val[life_summaries$variable == "vbk"],
          linf > life_summaries$min_val[life_summaries$variable == "linf"],
          linf < life_summaries$max_val[life_summaries$variable == "linf"])

  sim_plot <- outcomes %>%
    mutate(rsize = plyr::round_any(mpa_size, 0.25),
           reff = plyr::round_any(mpa_effect,0.2)) %>%
    filter(rsize == 0.25, depletion < 0.8, depletion > 0.2,
           scientific_name %in% sim_life_history$scientific_name
           ) %>%
    mutate(year = years_protected + 2001) %>%
    filter(year %in% did_betas$year, years_protected >1) %>% 
    group_by(year, reff) %>% 
    count() %>% 
    group_by(year) %>% 
    mutate(sc = n / sum(n))

  did_plot <- did_betas %>%
    ggplot() +
    geom_tile(data = sim_plot, aes(year, reff, fill = sc), alpha = 0.75) +
    geom_vline(aes(xintercept = 2003), color = 'red', linetype = 2, size = 2) +
    geom_hline(aes(yintercept = 0)) +
    geom_pointrange(aes(
      year,
      y = exp(estimate) - 1,
      ymin = exp(lower) - 1,
      ymax = exp(upper) - 1
    ),
    size = 1.5
    ) +
    labs(x = "Year", y = "Targeted Trend Divergence") +
    scale_fill_viridis(guide = wide_hgc,
                       name = "% of Annual Sims",
                       labels = scales::percent_format(accuracy = 1),
                       option = "magma") + 
    theme(legend.position = "top",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()
        )

cip_data <- pisco_data %>%
  left_join(site_data, by = c("site","side")) %>%
  filter(is.na(eventual_mpa) == F) %>%
  filter(region %in% c("ANA", "SCI","SRI",'SMI'),
         classcode %in% top_species)

# process outcomes

```

\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: danovan@uw.edu}

<!-- \AddToShipoutPictureFG{ -->
<!--   \AtPageCenter{% or \AtTextCenter -->
<!--     \makebox[0pt]{\rotatebox[origin=c]{45}{% -->
<!--       \scalebox{5}{\texttransparent{0.3}{DRAFT}}% -->
<!--     }} -->
<!--   } -->
<!-- } -->

No-take Marine Protected Areas (MPAs), spatial regions of the ocean in which fishing is prohibited, have a long history in the management of marine resources. Traditional cultures in Oceania utilized - often temporary - MPAs as "fish banks" for times of need [@johannes1978]. Modern MPAs were first established primarily as marine analogs to the terrestrial protection of iconic landscapes like Yellowstone or Kruger National Parks [@iucn1976;@watson2014b]. Over time our goals and expectations for MPAs have evolved; while all MPAs are expected to deliver conservation benefits within their borders, many modern MPAs are also established to  bolster fish populations throughout the region in which they are located (what we term “regional-scale effects”) [@gaines2010].

Indeed, many recent agreements to expand MPAs (most proximately, the Convention on Biological Diversity’s Strategic Plan for Biodiversity, which calls for 10% of coastal waters to be protected inside MPAs by 2020 and the International Union for Conservation of Nature call for 30% by 2030) rely on the belief that well-designed MPAs will achieve benefits both within, and outside, their borders. Despite these assumptions, our collective scientific understanding of the regional-scale conservation and fishery impacts of current and future MPAs is surprisingly limited.

What is the scientific evidence of the conservation effects of MPAs? Numerous studies provide evidence that well-enforced and appropriately sized MPAs can produce conservation benefits within their borders [@lester2009;@halpern2003; @edgar2014; @gerber2005]. As these conservation benefits accrue inside MPAs, theory holds that MPAs can affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort. Several studies have documented empirical evidence for the existence of adult or larval spillover affecting both abundance and fisheries [e.g. @goni2010; @halpern2009;@kay2012; @stobart2009; @mcclanahan2000; @russ1996; @thompson2017;@baetscher2019;@pelc2009], as well as alteration of fishing effort in reaction to [@costa2013;@mason2012;@murawski2005] and in anticipation of [@mcdermott2019] MPA placement. The potentially more important question, however, is not whether spillover occurs (it must to some degree in any realistic scenario), but what the net effects of spillover are and whether those effects are empirically detectable. From a fishery perspective, are spillover benefits sufficient to offset losses in fishing grounds and changes in responses of displaced fishers caused by an MPA? From a conservation perspective, how much does the buildup of fish inside an MPA increase biomass outside the protected area? Overall, what are the regional effects of MPAs?

As stakeholders around the world increasingly seek to use MPAs in marine resource management portfolios, and the performance of existing MPAs is evaluated, it is critical that we develop a better understanding of the magnitude and drivers of regional-scale MPA effects. To address this gap, this study examines two critical questions: 1) What do we expect the regional-scale conservation effects of MPAs to be and 2) When (and how) can we expect to empirically detect these effects? We answer these questions using a simulation analysis to frame the theoretical regional conservation impacts of MPAs. We then present an empirical test of the evidence for regional-level conservation effects of MPAs resulting from a network of closures put in place in the Channel Islands, California, USA, in 2003.

What Are the Regional Effects of MPAs? {.unnumbered}
----------------------

<!-- **key thing: a lot more ways for small things to happen than big things ** -->

<!-- xx need to add in discussion of other outcomes not covered here, e.g. resiliency, tourism, etc. xx -->

Much of the literature on the effects of Marine Protected Areas focuses on assessing the conservation effects within the borders of protected areas [e.g. @edgar2014]. While these within-MPA effects are vitally important for protecting rare species, biodiversity, critical habitats, and often tourism, they paint an incomplete picture of the overall population effects of MPAs. The organisms within the borders of protected areas are generally part of a broader population, the biological stock, connected through adult movement and larval dispersal. If the goal of conservationists or natural resource managers is to increase the total abundance or productivity of a resource, a broader question we should ask of MPAs is not just are there more fish inside their borders, but also how have the reserves affected abundances throughout the region in which they are located? This logic, that MPAs will have conservation benefits for most species beyond their borders, is implicit in all multilateral calls for MPA expansion.

We define the regional conservation effects of MPAs as the change in total biomass of fish (summing inside and outside of MPAs) relative to the total biomass of fish that would have occurred without the MPAs (acknowledging that other outcomes such as increased biodiversity or resiliency are also important to conservation but are beyond the scope of this analysis). Note that this definition of regional effects is in line with the recommendations of @ferraro2018, but differs from effects measured by for example differences in outcomes before and after MPAs.

We used a simulation model to examine the regional conservation effects of MPAs under a wide array of conditions that might reasonably be expected to affect these outcomes. The model explores combinations of MPA designs (e.g., sizes, number, and placement of MPAs), life histories (e.g., growth and mortality rates, adult movement, larval dispersal, recruitment variability and autocorrelation, and timing of density dependence), and fishing fleet dynamics (e.g., degree of fishing pressure, size selectivity of the fleet, reaction of fishing pressure to MPA placement) (See Table.S1 for a complete description of simulation variables). We use this model to explore how different drivers interact to affect the conservation and fishery outcomes of MPAs.

<!-- Numerous works have used simulation modeling to explore the effects of MPAs on outcomes such as fisheries yields and stock abundance (see @fulton2014 and references within),but these prior studies have either focused on the effect of a select set of drivers (such as dispersal rates) on outcomes, or have been highly localized assessments that incorporate multiple drivers but are tuned to one specific region. Our model captures critical bio-economic features, while still being flexible and fast enough to be generalized to a broad array of scenarios. We use this model to explore how different drivers interact to affect the conservation and fishery outcomes of MPAs. -->

### Theoretical Conservation Effects {#conservation .unnumbered}

As a thought experiment, imagine a region that has driven its fish populations to near extinction and then places 100\% of its waters inside a no-take MPA.  In that setting, we would expect the regional-scale conservation benefits to be massive. On the other hand, imagine implementing an MPA in an area where a sedentary species has been only lightly fished.  In that setting, it is possible to create a small net conservation loss if the concentration of fishing pressure outside the reserve has a greater effect than the biomass buildup inside the reserve. Within these broad bounds, numerous factors can act to affect the regional effects of MPAs. These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008], the strength and timing of density dependence in the population (e.g. pre- or post-settlement), how overfished the population would be without the MPA, and how fishing activity and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @gerber2005; @hilborn2004a; @gaines2010; @botsford2003]. In addition, even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through habitat and network effects [@gaines2010; @costello2010].

```{r}
short_term <- outcomes %>%
  group_by(experiment) %>% 
  filter(years_protected == max(years_protected), years_protected >= 0) %>% 
  ungroup() %>% 
  filter(!is.na(mpa_effect))

median_effect <- median(short_term$mpa_effect[short_term$year == max(short_term$year)], na.rm = TRUE)

median_pop_effect <- median(short_term$pop_effect[short_term$year == max(short_term$year)], na.rm = TRUE)

# median_fish_effect <- median(fishery_short_term$mpa_effect[short_term$year == max(short_term$year)], na.rm = TRUE)
# 
# median_fish_effect <- median(fishery_short_term$fishery_effect[short_term$year == max(short_term$year)], na.rm = TRUE)

pos_runs <- round(mean(short_term$pop_effect[short_term$year == max(short_term$year)] > 0),2)

# pos_fish_runs <- round(mean(fishery_short_term$msy_effect[fishery_short_term$year == max(fishery_short_term$year)] > 0, na.rm = TRUE),2)

pos_fish_effect <- round(mean(short_term$f[short_term$year == max(short_term$year)] > 0),2)

small_effect <- round( 100* median(short_term$mpa_effect[short_term$depletion <= 0.4]))

min_small_effect <- round( 100* min(short_term$mpa_effect[ short_term$depletion <= 0.4]))

max_small_effect <- round( 100* max(short_term$mpa_effect[ short_term$depletion <= 0.4]))

medium_effect <- round(100*median(short_term$mpa_effect[short_term$depletion > 0.4 & short_term$depletion <= 0.6]))

min_medium_effect <- round(100*min(short_term$mpa_effect[short_term$depletion > 0.4 & short_term$depletion <= 0.6]))

max_medium_effect <- round(100*max(short_term$mpa_effect[short_term$depletion > 0.4 & short_term$depletion <= 0.6]))

big_effect <- round(100*median(short_term$mpa_effect[short_term$depletion > 0.6]))

min_big_effect <- round(100*min(short_term$mpa_effect[short_term$depletion > 0.6]))

max_big_effect <- round(100*max(short_term$mpa_effect[short_term$depletion > 0.6]))

```

<!-- Across all simulations the median simulated equilibrium regional conservation effect, expressed as the percentage change in total biomass with the MPAs and without the MPAs, was `r round(100*median(eqo$mpa_effect))`%, with a minimum of less than `r round(100*min(eqo$mpa_effect))`% and a maximum of over `r round(100*max(eqo$mpa_effect))`%. The median time required for the simulated MPAs to cause a 10% increase in regional biomass was approximately 10 years. -->

We tested the effects of different combinations of these theoretical drivers of MPA conservation effects using over 10,000 simulated MPA scenarios. A striking outcome of this simulation analysis is how the level of fishing pressure can cause dramatic and predictable shifts in conservation outcomes. We measure fishing pressure by the degree of simulated depletion (percent of unfished biomass gone, such that a depletion of 0% means the population is at 100% of unfished biomass, 100% means 0% of unfished biomass remains) in the scenario without MPAs. When MPAs were placed in relatively unexploited fisheries (depletions less than 40%), the median regional conservation effect was barely noticeable (`r small_effect`%). For ecosystems fished near MSY levels (depletion of 40%-60%), the median conservation effect was `r medium_effect`%, while for overfished ecosystems (without-MPA depletions above 60%) the median regional conservation effect was over  `r big_effect`% (Fig.\ref{conservation-effects}-A). Across all of these groups though, effects as low as declines of -50% and as high as an increase of more than 100% were possible. It should be noted though that these runs represent the range of scenarios evaluated in our simulation, and that certain types of scenarios may be much more likely than others in the real world. To put these results into context, the FAO estimates that 7% of the worlds fisheries with status estimates fall into the relatively unexploited category, 60% fall into the fully fished category, and 33% fall into the heavily fished category [@fao2018], though works that include a broader range of fisheries estimate that 50% or more of stocks to fall into the heavily fished category [@costello2012;@rosenberg2017].

These simple results support the intuitive conclusion that bigger MPAs on overfished populations produce large positive conservation outcomes. However, when combined with other factors such as fleet dynamics, life history, and MPA size, the conservation effects of MPAs can vary widely in magnitude and timing, a finding corroborated by empirical evidence [@starr2015;@caselle2015;@nickols2019;@villasenor-derbez2019] (Fig.\ref{conservation-effects}-B).  MPAs protecting small proportions of a stock can produce small positive effects, large positive effects, or even small negative effects. As the area protected inside MPAs increases, positive effects become more likely, but even MPAs protecting most of a population's footprint can produce conservation effects as low as 0% and as high as 200% and above depending on the context in which they are placed. The degree of depletion (i.e. fishing pressure) has a clearer signal, with small conservation benefits of MPAs when depletion is low, and larger effects when depletion is high, but again even for severely depleted populations the conservation effects of MPAs can vary widely (although nearly always with population benefits).

While positive conservation outcomes were much more likely to occur across our simulations (`r percent(pos_runs, accuracy = 1)` of simulations), the `r percent(1 - pos_runs, accuracy = 1)` of runs that produced net conservation losses are of particular interest. These runs occurred almost exclusively when the simulated fleet followed a "constant-catch" fishing strategy. Under the constant-catch fleet model, fishing communities (or fishery policies) seek to catch the same amount regardless of the presence of an MPA. While a constant-catch greater than MSY is not possible over the long-term under the assumptions of our model, over the short-term a constant-catch scenario is not implausible. Subsistence fisheries may use a constant-catch style policy over the short-term, as they seek to ensure that their food needs are met. More industrial fisheries may have pre-arranged agreements with buyers to deliver set amounts of fish. Constant-catch dynamics might also occur in fisheries with constraining quotas that are not updated after the implementation of MPAs. These patterns suggest that coupling MPA implementation with appropriate management reforms that reduce the effects of constant-catch strategies can move the predicted conservation effects of MPAs to being nearly uniformly positive. The possibility of constant-catch strategies to result in net MPA conservation losses highlights the critical importance of understanding the broader management context in which MPAs are placed.

```{r conservation-effect, include = FALSE}

pop_time_plot <- outcomes %>%
  filter(years_protected > 0) %>%
  ggplot(aes(years_protected, mpa_effect)) +
  geom_bin2d() +
  scale_fill_viridis(guide = gc)

ylabs <- c(expression(""<= "-50%"), paste0(seq(-25,75, by = 25),"%"),expression("">= "100%") )

ybreaks = seq(-.50,1, by = .25)

pop_depletion_plot <- outcomes %>%
  filter(year == max(year)) %>%
  ggplot() +
  geom_bin2d(aes(depletion, mpa_effect), binwidth = c(.1, .1),
             show.legend = FALSE) +
  scale_fill_viridis(
  option = "A",
  # trans = "log10",
  guide = guide_colorbar(frame.colour = "black",
  frame.linewidth = 1,),
  name = "Median Effect"
  ) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1), name = "Depletion") +
  scale_y_continuous(labels = ylabs, breaks = ybreaks, name = "")

pop_depletion_plot <- outcomes %>%
    filter(year == max(year)) %>%
    mutate(
      rdep = cut(
        depletion,
        breaks = seq(0, 1, by = .2),
        labels = seq(0, .8, by = .2),
        include.lowest = TRUE
      )
      ,
      reff = cut(
        mpa_effect,
        breaks = seq(-.6, 1, by = .2),
        labels = seq(-.6, .8, by = .2),
        include.lowest = TRUE
      )
    ) %>%
    group_by(rdep, reff) %>%
    count() %>%
    group_by(rdep) %>%
    mutate(pn = n / sum(n)) %>%
    ungroup() %>%
    mutate(rdep = as.numeric(as.character(rdep)),
           reff = as.numeric(as.character(reff))) %>%
    ggplot(aes(rdep, reff, fill = pn)) +
    geom_raster(hjust = 1, vjust = 1, show.legend = FALSE) +
    scale_fill_viridis(
      option = "A",
      # trans = "log10",
      guide = hgc,
      name = "% of Sims",
      labels = scales::percent_format(accuracy = 1)
    ) +
    scale_x_continuous(labels = scales::percent_format(accuracy = 1),
                       name = "Depletion",
                       expand = expand_scale(add = c(0,0))) +
    scale_y_continuous(labels = ylabs,
                       breaks = ybreaks,
                       name = "")

  sim_plot <- outcomes %>%
    mutate(rsize = plyr::round_any(mpa_size, 0.25),
           reff = plyr::round_any(mpa_effect,0.2)) %>%
    filter(rsize == 0.25, depletion < 0.8, depletion > 0.2,
           scientific_name %in% sim_life_history$scientific_name
           ) %>%
    mutate(year = years_protected + 2001) %>%
    filter(year %in% did_betas$year, years_protected >1) %>% 
    group_by(year, reff) %>% 
    count() %>% 
    group_by(year) %>% 
    mutate(sc = n / sum(n))

  pop_size_plot <- outcomes %>%
    filter(year == max(year)) %>%
    mutate(
      rsize = cut(
        mpa_size,
        breaks = seq(0, 1, by = .2),
        labels = seq(0, .8, by = .2)
      )
      ,
      reff = cut(
        mpa_effect,
        breaks = seq(-.6, 1, by = .2),
        labels = seq(-.6, .8, by = .2)
      )
    ) %>%
    group_by(rsize, reff) %>%
    count() %>%
    group_by(rsize) %>%
    mutate(pn = n / sum(n)) %>%
    ungroup() %>%
    mutate(rsize = as.numeric(as.character(rsize)),
           reff = as.numeric(as.character(reff))) %>%
    ggplot(aes(rsize, reff, fill = pn)) +
    geom_raster(hjust = 1, vjust = 1) +
    scale_fill_viridis(
      option = "A",
      # trans = "log10",
      guide = hgc,
      name = "% of Sims",
      labels = scales::percent_format(accuracy = 1)
    ) +
    scale_x_continuous(labels = scales::percent_format(accuracy = 1),
                       name = "Range in MPA",
                       expand = expand_scale(add = c(0,0))) +
    scale_y_continuous(labels = ylabs,
                       breaks = ybreaks,
                       name = "Conservation Effect") +
    theme(legend.position = "top") +
    labs(title = 'B)')

 pop_facet_effect_plot <- outcomes %>%
  filter(year == max(year)) %>%
   select(mpa_effect, mpa_size, depletion) %>%
   gather(measure, value, -mpa_effect) %>%
   ggplot() +
   geom_bin2d(aes(value, mpa_effect), binwidth = c(.1, .1), show.legend = TRUE) +
  scale_fill_viridis(
  option = "A",
  # trans = "log10",
  guide = hgc,
  name = "# of Sims"
  ) +
   facet_wrap(~measure, labeller = labeller(measure = facet_labels), strip.position = "bottom") +
   scale_x_percent(name = "") +
   scale_y_percent(name = "Conservation Effect") + 
   theme(legend.position = "top")

  pop_depletion_and_size_plot <- outcomes %>%
    filter(years_protected >= 0) %>%
    group_by(experiment) %>%
    mutate(depletion = plyr::round_any(depletion[years_protected == 0], .1),
           mpa_size = plyr::round_any(mpa_size, .1)) %>%
    filter(year == max(year)) %>%
    group_by(depletion, mpa_size) %>%
    summarise(median_mpa_effect = median(mpa_effect, na.rm = TRUE)) %>%
    ggplot(aes(depletion, mpa_size, fill = median_mpa_effect)) +
    geom_tile() +
    # geom_contour(aes(z = median_mpa_effect, color = ..level..), show.legend = FALSE) +
    scale_fill_viridis(labels = scales::percent_format(accuracy = 1),
                       guide = hgc,
                       name = "Median Effect") +
  labs(x = "Depletion", y = "Range in MPA", title = "A)") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                       expand = c(0,0)) +
    scale_x_continuous(labels = scales::percent_format(accuracy = 1), expand = c(0,0), limits = c(0,1)) + 
    theme(legend.position = "top")

  # pop_depletion_and_size_plot <-  directlabels::direct.label(pop_depletion_and_size_plot, list("angled.boxes", box.alpha = 0.5))

  pop_combo_plot <-
    (pop_depletion_and_size_plot) + ((pop_size_plot) / pop_depletion_plot)  + plot_layout(widths = c(1.5, 1)) & theme(
    plot.margin = unit(c(0, 1, 0, 1), units = "lines"),
    axis.text.x = element_text(size = 8),
    legend.box.margin = unit(c(0, 0, 0, 0), units = "lines"),
    axis.text.y = element_text(size = 10)
    )

  ggsave(pop_combo_plot,filename = "figs/conservation-effect.png",
         width = 8,
         height = 5)

  pdf(
    file =  "figs/conservation-effect.pdf",
    width = 8,
    height = 5,
    useDingbats = FALSE
  )
  print(pop_combo_plot)
  dev.off()

```

\begin{figure*}%[tbhp]
  \centering
  \includegraphics[width=.9\linewidth]{figs/conservation-effect.png}
  \caption{Median (A) and range (B) of equilibrium regional MPA conservation effects (percent change in total biomass with MPAs relative to without MPAs) across a range of depletion and MPA sizes (and incorporating the full range of scenarios included in our study). "Range in MPA" is the percent of patches covered by an MPA, "Depletion" is the depletion that would have occurred in equilibrium without the MPA. Colors in panel B represent the percentage of simulations that fell into an effect size bin at each x-axis bin, such that the sum of all percentages for a given x-axis bin is 1.}
  \label{conservation-effects}
\end{figure*}

## How Can We Detect Regional Effects?

How can we empirically measure the regional effects of MPAs?  Intuition and bio-economic theory tells us that the regional conservation effects of MPAs are likely to be highly context dependent across locations and across species within locations, so careful thought must be devoted to an empirical design to detect their effects. The perfect experiment would involve two parallel worlds that were identical, except for the implementation of an MPA.  In world “A”, no MPA would be implemented, and in the facsimile world “B”, the MPA would be implemented. Both worlds would be tracked before and after MPA implementation, and the conservation and fishery outcomes would be compared after treatment. Instead of two parallel worlds, a similar experimental design would involve random placement of MPAs.  Unfortunately, neither of these experimental designs has been implemented, because parallel worlds do not exist and MPAs are never placed at random. Careful causal inference requires some other means of controlling for biases introduced by factors such as the MPA siting process and biological spillover and concentration of fishing effort  [@ferraro2018;@larsen2019].

<!-- From the fisheries perspective, catch rates before and after [e.g. @roberts2001a] or near and far from MPAs [e.g. @russ2004a] are sometimes presented as evidence for the fisheries effects of MPAs. However neither of these approaches can be counted on to provide reliable estimate of the net fishery effects of MPAs, due to failures to account for potential broader trends in catch rates independent of MPAs and the potentially confounding effects of effort redistribution. @smith2006a presents an econometrically robust method for estimating the fishery effects of MPAs, though this approach requires high-resolution data on comparable fishing effort both near and far from MPAs, data that is very rarely available. It is beyond the scope of this paper to delve deeper into detecting the net fishery effects of MPAs, and so we now focus on the question of estimating their regional conservation effects. -->

Conservation effects are typically estimated by comparing organism densities or biomass inside MPAs to densities or biomass in selected control sites outside MPAs, which we will refer to as response ratios. @lester2009 and @halpern2002 present meta-analyses of hundreds of such studies. These results often find massively higher densities and biomasses inside MPAs than outside [median biomass increases of  over 400%, @lester2009]. While response ratios are intended as a measure of the effectiveness of reserves within their borders, these result stand in contrast to the smaller region wide effect sizes predicted by our simulation model. This raises the question: How reliable are density or biomass ratios as estimators of regional conservation effects?

Control sites used in density or biomass ratios are often selected based on abiotic or ecological traits such as habitat characteristics [@ferraro2018]. However, selection of control sites is further complicated by the very spillover that MPAs are often intended to create. Export of adults or larvae from the MPA to the "control" site affects their status as controls, as does displacement of fishing effort from MPAs to control sites. In theory, control sites far enough away to negate both biological spillover and concentration by the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs may be challenging. While these concerns have been stated previously [e.g. @halpern2004], the MPA evaluation literature has by and large been unable to adequately address them, often due to the very real challenges of acquiring the necessary data [@ferraro2018].

Failure to account for these spillover effects can result in a biased estimate of the true effect of a policy such as MPA placement [@larsen2019]. In an ideal setting, the control site used in a response ratio would be a perfect counterfactual for what would have occurred without the MPA. We used our simulation model to approximate this scenario, calculating the response ratio as the ratio of densities inside the MPAs relative to the overall density from the paired simulation without MPAs multiplied by the area of the population in an MPA (to account for the fact that we would expect the effectiveness of a response ratio as a proxy for regional conservation effects to scale with the size of the MPA relative to the range of the species, Fig.\ref{density-ratio}-A). In this idealized case, the response ratio is often a reasonable estimator for the regional conservation effect.

<!-- However, even under this ideal example the response ratio is not a perfect estimator due to the fact that this ratio contains no information on the effect of the MPAs outside their borders.  -->

What happens when the “control” sites are in fact affected by the application of the MPA through biological spillover and fishing concentration? In order to address this possibility, we calculated a new response ratio as the mean density inside MPAs relative to the mean density outside the MPAs (both weighted by distance from MPA borders and scaled by MPA size). We only include simulations in which habitat and larval dispersal rates are identical inside and outside of MPAs, to approximate a scenario in which treatment and control sites have been paired by ecological characteristics. Under these circumstances the response ratio is a somewhat biased and very inaccurate estimator of the true regional effect of MPAs. While on average high response ratios correspond with high true regional effects, the variance is high, so the response ratio alone cannot be used as a reliable proxy for the true regional effect, without accounting for the sources of biases discussed here. For higher movement rates, response ratios were frequently near 0\% (potentially leading stakeholders to conclude that the MPA had been ineffective), when in fact in many cases the true effect of the MPA was highly positive. For highly sedentary species, extremely high response ratios could create the appearance of massive conservation gains when in fact the net effect on the regional population has been near zero or even negative (Fig.\ref{density-ratio}-B).

We note that spatial before-after-control-impact (BACI) studies present a potential improvement over response ratios, controlling for differences in mean densities in the treated and control regions as well as MPA independent trends, but are much rarer due to the need for extensive pre-MPA monitoring. However, spatial BACI still requires that the MPAs do not affect the control sites, and as such has many of the same limitations as response ratios as an estimator of regional conservation effects.

To the extent that we can select control sites that are both sufficiently similar to the region treated with MPAs and are unaffected by biological or economic spillover, response ratios may be reliable estimators of regional conservation effects. In reality these conditions are unlikely to hold but for a few select cases. This become especially difficult when networks of MPAs are implemented, such as has occurred in California. In this case, control sites are generally located within a few kilometers from MPA borders, making biological and economic spillover highly likely. In these circumstances, without robust statistical controls, response ratios may provide an unreliable estimate of the regional conservation effect of marine protected areas.

```{r dr-vs-mpa-effect-plot, include = FALSE}

dr_plot <- density_ratios %>%
  filter(sprinkler == FALSE, mpa_habfactor == 1) %>%
  select(experiment,
         years_protected,
         mpa_effect,
         adult_movement,
         mpa_size,
         contains("density_ratio")) %>%
  mutate(adult_movement = adult_movement / max(adult_movement)) %>%
  gather(scenario, density_ratio, contains("density_ratio")) %>%
  group_by(experiment, scenario) %>%
  filter(years_protected == max(years_protected)) %>%
  mutate(density_ratio = density_ratio * mpa_size) %>% 
  ggplot(aes(y = pmin(4, mpa_effect), x = pmin(4, density_ratio - 1))) +
    geom_hline(aes(yintercept = 0)) + 
  geom_vline(aes(xintercept = 0)) + 
  geom_abline(aes(slope = 1, intercept = 0), linetype = 2, size = 1.5, color = "black") +
  geom_point(aes(color = adult_movement, size = mpa_size), alpha = 0.5, show.legend = TRUE)  +
    geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ rev(scenario), labeller = labeller(`rev(scenario)` = c(biased_density_ratio = "A) Inside vs. Experiment",true_density_ratio = "B) Inside vs. Outside"))) +
  labs(y = "True % Increase in Biomass", x = "% Response Ratio") +
  # scale_x_percent(labels = c(percent(seq(-1,1, by = 1)),">200%")) +
  # scale_y_percent(labels = c(percent(seq(-1,1, by = 1)),">200%")) +
  scale_color_viridis(guide = wide_hgc, labels = scales::percent_format(accuracy = 1), name = "Fish Movement",
                      option = "cividis") + 
  scale_x_continuous(labels = percent) + 
    scale_y_continuous(labels = percent) + 
  theme(legend.position = "top") + 
  guides(size = FALSE)

pdf(file =  "figs/density-ratio.pdf",width = 8, height = 5, useDingbats = TRUE)
print(dr_plot)
dev.off()

```

\begin{figure*}%[tbhp]
  \centering
  \includegraphics[width=.9\linewidth]{figs/density-ratio.pdf}
  \caption{Simulated MPA-size weighted response ratios (x-axis) plotted against true regional change in biomass caused by MPAs (y-axis). Each point represents a simulation, and color represents the movement rate of fish in the simulation (expressed as a percentage of the patches a fish is likely to move in one time step). The black line shows the one:one fit line, the blue line is a linear fit of the relationship between observed response ratios and true regional change in biomass. Inside vs. Experiment simulates a true control, while Inside Vs. Outside uses a control that may be affected by the treatment.}
  \label{density-ratio}
\end{figure*}

```{r}
fish_v_fishing <- processed_grid %>%
  select(-density_ratio) %>%
  unnest() %>%
  rename(fishery_effect = mpa_effect1) %>%
  group_by(experiment) %>%
  mutate(year = 1:length(year)) %>%
  ungroup() %>%
  mutate(years_protected = year - year_mpa + 1) %>%
  left_join(outcomes %>% select(experiment, year, depletion),
            by = c("experiment", "year"))

fish_v_fishing_plot <- fish_v_fishing %>%
  filter(years_protected == max(years_protected)) %>%
  ggplot(aes(pmin(2, mpa_effect), pmin(2, fishery_effect))) +
  geom_abline(aes(slope = 1, intercept = 0)) +
  geom_smooth(method = "lm") +
  geom_hex()

```

<!-- Our strategy for identifying the effect of the MPAs in the Channel Islands is to use biomass densities of non-targeted species as our control for broader factors affecting biomass densities of fishes throughout the Islands besides the effect of the MPAs. Simply examining densities of tarted species throughout the islands before and after MPAs might confuse changes in regional densities that would have occurred anyway without the MPAs with MPA effects (since at the regional scale we do not have a spatial control for the effects of the MPAs). We assume that over the 13 years of MPA protection covered in this study non-targeted species were relatively unaffected by the MPAs (i.e. indirect effects of MPAs on species interactions are negligible), and can can serve as an indicator for the overall changes in fish abundance throughout the Channel Islands that would have occurred without the MPAs. Raw densities calculated by PISCO provide visual support for this hypothesis, in that both the targeted and non-targeted species groups share common trends (Fig.\ref{pop-trends}, see Materials and Methods for statistical support for this assumption). -->

## MPA Effects in the Channel Islands {#estimating .unnumbered}

```{r ci-map, fig.cap = "Map of study region and sampling locations. Shaded polygons indicate location of MPAs. Points represent sampling locations, and color indicates the number of observations recorded at a given point", include = FALSE}

ci_map <- pisco_data %>%
  left_join(site_data, by = c("site","side")) %>%
  filter(is.na(eventual_mpa) == F) %>%
  filter(region %in% c("ANA", "SCI","SRI",'SMI'),
         classcode %in% top_species) %>%
  group_by(lat_wgs84, lon_wgs84, site, region, eventual_mpa) %>%
  count()

ci_map <-  ci_map %>%
  dplyr::mutate(geometry = purrr::map2(lon_wgs84, lat_wgs84, ~ sf::st_point(x = c(.x, .y), dim = 'XY'))) %>%
  ungroup() %>%
  mutate(geometry = sf::st_sfc(geometry, crs = 4326)) %>%
  sf::st_sf()

bbox <- sf::st_bbox(ci_map)

pisco_ci_map_plot <-  ggmap(channel_islands, alpha = 0.5) +
  geom_sf(data = ca_mpas, inherit.aes = FALSE, alpha = 0.5) +
  geom_sf(data = ci_map,inherit.aes = FALSE,
  aes(fill = eventual_mpa, color = eventual_mpa),
  shape = 21, 
  alpha = 0.5,
  show.legend = FALSE,
  size = 1.5) +
  coord_sf(xlim = c(bbox['xmin'] - .3, bbox['xmax'] + .06),
  ylim = c(bbox['ymin'] - .1, bbox['ymax'] + .1)) +
  # scale_color_gradient(low = "white", high = "orangered",
  # guide = guide_colorbar(frame.colour = "black",frame.linewidth = 1,
  # barheight = unit(13, "lines")),
  # name = "Observations")  +
  labs(x = "", y = "") +
      ggspatial::annotation_scale(location = "br", width_hint = 0.4, text_col = "white") + 
  ggsci::scale_color_npg() + 
    ggsci::scale_color_npg()

# pisco_ci_map_plot <-  ggplot() +
#   geom_sf(data = tests2, fill = c("#FFD39B")) +
#   geom_sf(data = ca_mpas, inherit.aes = FALSE, alpha = 0.5) +
#   geom_sf(data = ci_map,inherit.aes = FALSE,
#   aes(color = eventual_mpa),
#   alpha = 0.75,
#   show.legend = FALSE,
#   size = 2) +
#   coord_sf(xlim = c(bbox['xmin'] - .3, bbox['xmax'] + .06),
#   ylim = c(bbox['ymin'] - .1, bbox['ymax'] + .1)) +
#   # scale_color_gradient(low = "white", high = "orangered",
#   # guide = guide_colorbar(frame.colour = "black",frame.linewidth = 1,
#   # barheight = unit(13, "lines")),
#   # name = "Observations")  +
#   labs(x = "", y = "") +
#       ggspatial::annotation_scale(location = "br", width_hint = 0.4, text_col = "white") + 
#   theme(panel.background = element_rect(fill = "lightblue"))
# 
pdf(file =  "figs/channel-islands.pdf",width = 8, height = 5, useDingbats = TRUE)
print(pisco_ci_map_plot)
dev.off()

```

\begin{figure}%[tbhp]
  \centering
  \includegraphics[width=1\linewidth]{figs/channel-islands.pdf}
  \caption{Map of PISCO sampling locations within the the Channel Islands National Marine Sanctuary, California, USA. Grey boxes indicated MPAs, points sampling locations, with color representing whether the site is inside an MPA (blue) or not (red)}
  \label{channel-islands}
\end{figure}

```{r}
ci_effects <- short_term %>%
  filter(years_protected == max(years_protected),
         mpa_size  <= 0.25, mpa_size >= 0.15)

med_ci_effect <- median(ci_effects$mpa_effect)

# rethinking::HPDI(ci_effects$mpa_effect, prob = 0.5)

quart_ci_effect <- quantile(ci_effects$mpa_effect)

```

Given these challenges, what is a viable alternative for estimating the conservation effects of MPAs? In a manner similar to @caselle2015, we use a difference-in-difference estimator comparing densities in finfish species targeted by fishing effort (i.e., those potentially affected by an MPA) from those not targeted before and after MPA implementation [see @larsen2019 for a succinct explanation of difference-in-difference estimators in ecological settings]. Targeted species in the Northern Channel Islands, California, include commercially important finfish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish. Each of these targeted species was the subject of prior bio-economic modeling related to the effects of MPAs in southern California [@costello2010d;@rassweiler2014]. But, we must note that the analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*).  Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species). The key assumptions of this approach are that a) within the time-frame of the model there are no significant interaction effects between the targeted and non-targeted species (which in fact we do not detect, see SI section 1.6.3), and b) that in the absence of the MPAs both the targeted and non-targeted groups of species would have exhibited similar trends in densities (the parallel-trends assumption). The advantage to this approach is that given the time-frame of the model (15 years), we believe that spillover effects of MPA placement on the non-targeted species are likely to be much less severe than the effects of biological spillover and fishing concentration that may bias the performance of estimators such as a response ratios or a spatial BACI design.

We applied this difference-in-difference strategy to a large dataset in California. We used empirical kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands with the ultimate goal of testing the regional effects of MPAs in a real world context. A network of MPAs covering approximately 20% of the islands’ waters was put in place in 2003 as part of the California Marine Life Protection Act (MLPA) (see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of the MLPA). PISCO conducts visual SCUBA surveys at a large number of rocky-reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\ref{channel-islands}-\ref{pop-trends}).

Although the Islands do not encompass the full population of any of these species, they encompass a large enough area to assess the regional MPA conservation effects we might see in the Channel Islands after 13 years of protection (the time span covered by our model).  The median regional conservation effect produced by our simulation model after 13 years of protection by MPAs covering between 15%-20% of a population's range was `r percent(med_ci_effect, accuracy = 1)`. This projected effect has an an interquartile range of `r percent(quart_ci_effect[2], accuracy = 1)` and `r percent(quart_ci_effect[4], accuracy = 1)`, although values as high as a 200% increase and as low as a 50% decrease were also produced. Our expectation then is that all else being equal we are likely looking for a modest but positive effect size.

```{r raw-trend, fig.cap="Centered and scaled mean annual biomass density of included species (faded lines) and smoothed means of targeted and non-targeted groups, and mean (darker lines) and 95% confidence interval of the mean (ribbon) over time", include = FALSE}
raw_did_plot <- base_run$data[[1]] %>%
  select(year, targeted,classcode, log_density, any_seen) %>%
  mutate(density = exp(log_density) * any_seen) %>%
  group_by(classcode,targeted, year) %>%
  summarise(md = mean(density)) %>%
  group_by(classcode,targeted) %>%
  mutate(smd = (md - mean(md)) / sd(md)) %>%
  ggplot(aes(year, smd, color = targeted == 1, fill = targeted == 1)) +
  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red", size = 1.5) +
  geom_line(aes(group = interaction(targeted, classcode)), alpha = 0.5) +
  geom_smooth(size = 2) +
  labs(y = "Centered and Scaled Mean Density", x = "Year") +
    scale_color_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
      scale_fill_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
    labs(y = "Centered and Scaled Mean Density", x = "Year") + 
  theme(legend.position = "top",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()
        )

raw_did_plot <- raw_did_plot +
      scale_color_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "") +
        scale_fill_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "")

pdf(file =  "figs/pop-trends.pdf",width = 8, height = 5, useDingbats = TRUE)
print(raw_did_plot)
dev.off()

```

\begin{figure}%[tbhp]
  \centering
  \includegraphics[width=1\linewidth]{figs/pop-trends.pdf}
  \caption{Centered and scaled mean annual biomass densities of included targeted (red) and non-targeted (blue) fish species (faded lines) and smoothed means with 95\% confidence intervals over time (darker lines and ribbons).}
  \label{pop-trends}
\end{figure}

The raw trends in the biomass densities appear to show parallel trends in the densities of targeted and non-targeted fishes pre-MPA, with the targeted group diverging first positively and then negatively over time (Fig.\ref{pop-trends}). However, these raw trends could be caused by any number of confounding variables besides the effects of the MPAs themselves, from changes in observer skill and sampling locations, to shifts in environmental conditions. Using our difference-in-difference model in an effort to control for these factors, we found no significant difference in the biomass densities of targeted and non-targeted fishes pre-MPA (before 2003), which provides support for the parallel trends assumption critical to the difference-in-difference method. Following the implementation of the MPAs in 2003, we see evidence of an increasing trend in densities of targeted relative to non-targeted fish, reaching a peak value in 2014. Following 2014, densities of targeted species appear to have declined relative to what we would expect based on the non-targeted group (Fig.\ref{did-plot}).

```{r did-plot, include = FALSE}

did_plot <- did_plot +
  scale_y_continuous(name = "~Divergence from Non-Targeted", labels = scales::percent_format(accuracy = 1))

ggsave(did_plot, filename = "figs/did_plot.png",
       width = 8, height = 5)

# pdf(file =  "figs/did_plot.pdf",width = 8, height = 5, useDingbats = TRUE)
# print(did_plot)
# dev.off()

```

\begin{figure}%[tbhp]
  \centering
  \includegraphics[width=1\linewidth]{figs/did_plot.png}
  \caption{Estimated divergence in biomass densities of targeted and non-targeted fishes throughout the Channel Islands (integrated across inside and outside of MPAs). MPAs are implemented in 2003 (red dashed line). Estimates are from a regression on log(biomass density), so estimated effects roughly correspond to percentage changes. Colored cells in the background indicate the frequency of MPA conservation effects predicted by our simulation modeling for MPAs similar in design and species composition to the Channel Islands.}
  \label{did-plot}
\end{figure}

Discussion {#discussion .unnumbered}
------

<!-- Studies such as @dinerstein2019 call for 30% of the oceans to be places in protected areas. What does simulation exercise suggest that the effects of such a closure might be on overall population biomass densities? XX. Suppose we assumed that all fisheries are both open-access and overfished (a clearly worst case scenario as many fisheries are well managed). In that case, we see that... -->

<!-- <!-- zeros may be an artifact of "error in variables", which can bias estimates towards zero [@larsen2019] -->

MPAs are an important part of the marine resource management toolbox. Under ideal circumstances they can protect individual species and ecosystem linkages, while supporting local economies through tourism and fishing opportunities. One rationale for the expansion of MPAs is that they will deliver net conservation benefits both inside *and outside* their borders. To date, this assumption is insufficiently tested, and this is the focus of our paper. Our results show that regional conservation benefits of MPAs are highly context dependent and in many circumstances, are likely to be so small that they are nearly impossible to detect empirically. Indeed, this is exactly what we found in our empirical case study from the Channel Islands, California, USA.

<!-- Our simulation results present a clear outcome from the perspective of fishery effects: Under the assumptions of our model, MPAs should not be expected to provide improvements in fisheries catches unless the stock is experiencing substantial overfishing (Fig.\ref{catch-msy-effects}-B). This general result has been discussed before [@hilborn2004], but our results provide the most comprehensive simulation testing of this concept to date. MPAs placed in overfished fisheries frequently produced catch gains of 0-25%, though larger and smaller effects were still possible. MPA size also presents an interesting tradeoff from a fisheries perspective. MPAs that cover less than 25% of the population range, which is true for nearly all MPAs currently on the planet, very rarely produced substantial fishery losses, but also never produced substantial fishery gains. Simulated MPAs  that covered between 25% & 50% of species ranges produced the potential for higher fishery gains, but also opened up risk of greater fishery loses. For MPAs covering more than 50% of a species range, fishery losses quickly become much more likely than fishery gains. -->

<!-- We only measure one aspect of fishery outcomes here (catches). MPAs can affect many other outcomes of importance to fisheries, include profits (e.g. if institutions do not create derby fishing and MPAs improve catch per unit effort along the borders of reserves), resiliency, and research. Our model does not account for potentially important factors such as species interactions Overall, our fishery-specific results suggest a relatively simple rule of thumb: unless there is clear evidence of overfishing, MPAs should not be expected to produce improvements in fishery catches without clear explanation for the mechanism for these increases (e.g. improvements in habitat). In assessing the potential fishery benefits of MPAs, stakeholders must make explicit the mechanisms by which these benefits are expected to be produced (for example through increased catchability of fish near MPA borders), and what evidence exists for the existence of those mechanisms within the local context. -->

We find regional conservation gains from MPAs in `r percent(pos_runs, accuracy = 1)` of our simulations (Fig.\ref{conservation-effects}). MPAs covering less than 25% of the population range were unlikely to produce regional conservation gains above 10% unless the stock would have been severely overfished without the MPA. Given that few marine species on the planet have more than 25% of their entire range protected in MPAs, this suggests that large regional gains are unlikely to be common unless the overall extent of MPAs grows globally. Large MPAs protecting highly depleted stocks almost always produced large regional conservation gains, as should be expected. The median conservation effect of `r  round(100*median(eqo$mpa_effect))`\% stands in contrast to the large within-MPA effects on biomass densities reported by @halpern2003a and @lester2009. This difference is due in part to the fact that within-MPA effects are likely to be much larger than regional-MPA effects, but may also reflect the many challenges of translating response ratios into regional conservation effects. Our simulation analysis shows that without proper controls (in particular for the effects of fleet redistribution), response ratios well over 100% are entirely plausible even when the actual conservation effect is much lower (and *vice versa*).

<!-- While our results provide important insight on the expected fishery and conservation effects of MPAs, they also highlight the critical importance of explicitly modeling the links between the biological effects of MPAs and the ways that humans respond to them. One of our clearest findings was that MPAs can cause net conservation losses when implemented in a "constant-catch" fishery, as fishers must dramatically increase their fishing effort in the waters still available to them to maintain their desired landings. The focus of much of the MPA design and modeling literature has historically been on biological and ecological such as growth rates and connectivity [citation], though studies such as @hastings1999 and @sanchirico2006b incorporate assumptions about outside-MPA fleet dynamics as well. Our results make clear that economic and ecological drivers must be included in any attempt assess the effects of MPAs MPAs: in particular, we need to explicitly understand the the behavior of fishing fleets if we hope to estimate the conservation or fishery effects of MPAs [see @costa2013 for an example of this process]. Large-scale empirical evidence confirms that predicting the effects of MPAs depends on understanding the human context in which they find themselves [@cinner2018].  -->

<!-- The question of what is a "meaningful" conservation gain is of course context dependent and depends on the priorities and preferences of stakeholders. However, under any set of preferences and objectives managers and stakeholders would of course like to know whether policy interventions such as MPAs are having their intended effects. While the importance of say a 0-10% regional conservation gain will vary, there can be little doubt that an effect size within this range will be very difficult to detect in a noisy natural system (xx citations on noise and effect size xx), and our empirical assessment of MPA conservation effects in the Channel Islands demonstrates this.  -->

<!-- . That we were unable to detect a significant divergence in the biomass densities of targeted species throughout the Channel Islands post MPA implementation, relative to the densities that the trends of the non-targeted species would suggest we should observe, should not come as a great surprise then.  -->

We were not able to detect a statistically significant difference in the densities of targeted and non-targeted finfish species over the 13 years of MPA protection in the Channel Islands covered by our analysis. The data (and model assumptions) provide support for both negative and positive MPA effects, although the size of these estimated effects is well within the range that our simulation analysis suggests are plausible. At face value, this lack of a clear result may seem surprising given the size and carefully studied nature of this MPA network.

However, our simulation analysis suggests that this result should perhaps be expected. The Channel Islands MPAs cover approximately 20% of the waters in the Channel Islands, and while formal stock assessments are not available for many of the targeted species in our analysis, what evidence we have suggests that, as a group, these fish are on average not heavily overfished. Some species, such as California sheephead and blue rockfish were likely below target levels during the period [@alonzo2004;@dick2017], but projections based upon the overall average response across all species will likely suggest modest benefits even if a subset of species could experience much larger population gains. Harkening back to our simulations, we expect the average percentage difference in densities of targeted species with and without MPAs to be modest (Fig.\ref{did-plot}). Effects of this size are likely to be challenging to detect empirically given the large natural variation of marine ecosystems (especially temperate reefs) and the observation error inherent in infrequent annual survey programs such as those provided by PISCO.

```{r}
tmdata <- life_history_data %>% 
  filter(targeted == 1, classcode %in% top_species) %>% 
  select(tm, everything())

median_tm = round(median(tmdata$tm))
```

In addition, our simulations examine percentage increases in biomass densities. As an extreme example, an increase in biomass densities from 2kg/m^2^ to 4kg/m^2^ would be reported as a 100% increase. While this chance produces a large percentage increase, this is a small change in absolute biomass densities relative to the variance in the observation process itself (see Fig.S1 for a companion to Fig.\ref{conservation-effects} scaled by absolute population size). Beyond these challenges, the median estimated age at sexual maturity for the targeted species included in this study is `r median_tm` years, meaning that the span of this analysis represents less than three generations of MPA protection for half of the the measured species. Ongoing monitoring may yet reveal clearer effects. Analysis of more rapidly growing and maturing species, e.g. spiny lobster, may also reveal clearer signals.

While the MPA effect estimated by our model was not significant in any one year, the positive trend from 2005 to 2014 is evident, as is the sharp decline from 2015-2017. To what should we attribute this apparent shift? We cannot reject the parallel trends assumption between the targeted and non-targeted species in the years before the MPAs (Fig.\ref{did-plot}). But, the Channel Islands region (and the entire West coast of the USA) experienced a dramatic ‘marine heatwave’ beginning in 2014 and persisting through 2016, resulting in part in extremely elevated water temperatures throughout the region  [@gentemann2017]. Biogeographic differences in the distributions of targeted and non-targeted species may confound the observed effects of MPAs. Many of the non-targeted species have warm thermal affinities and have increased in numbers since the heatwave [@freedman2019]. However, the targeted group is made up mostly of fishes with cold-water affinities, such as members of the genus Sebastes. As such, we hypothesize that the recent evidence for a decline in densities of targeted species is due to environmental conditions that disproportionately affect the targeted group (and not for example due to concentrated fishing pressure outside the reserves). This hypothesis is supported by the sharp declines in densities of targeted species seen inside the MPAs themselves (see Fig.S27-S29), suggesting a driver other than fishing may be at play. The Channel Islands case study demonstrates the challenge of detecting small effect sizes in nature and highlights the potential for small effect sizes to be overwhelmed by additional environmental shocks to the system.

<!-- Our simulation model accounts for many factors that theory suggest should be critical drivers of MPA effects, namely adult and larval movement rates, nature of density dependence, and fleet dynamics. However, currently our model does not contain many other potentially important factors, such as species interactions. If stakeholders expect an effect that our model says is unlikely or impossible given local context, they should provide a clear hypothesis and evidence as to what factors not included in this model would cause their expected effect. For example, stakeholders might assert that larger conservation effects than predicted by our simulation model might occur in a location due to protection of critical spawning aggregation sites [@roberts2002;@erisman2015], but relevant evidence supporting this belief must be presented. It is also important to remember that our conservation results refer to the regional effects of MPAs both inside and outside of their borders. The buildup of biomass within the border of well enforced reserves is well documented [@lester2009], we show though that the regional effects may be much smaller and harder to detect than these within-MPA outcomes. -->

What do our results imply about the future of MPA science? Our simulation model is by no means exhaustive (for example it ignores features such as species interactions, habitat effects, and climate feedback), but captures many of the core factors theorized to affect MPA performance. Our results show that while attributes such as MPA size and fishing pressure are important factors in determining the effects of protection, local fleet dynamics and the movement rates of adult fish can dramatically affect the outcomes of protected areas as well  (Fig.\ref{conservation-effects}). Far from being a simple tool with clear outcomes, the effects of MPAs can be highly context dependent, requiring - we would argue - a bio-economic model of at minimum the complexity presented here to help communities design and set expectations of MPAs at the tactical level. While users may not be able to parameterize every aspect of a model such as that presented here, working with stakeholders to visualize the implications of, for example, different fleet responses to MPA implementation is a critical step in MPA design. Readers can use our simulation model to explore the effects of MPAs through an interactive web application available [here](https://danovando.shinyapps.io/simmpa/).

Once simulation models have been used to help design an MPA, how can users evaluate whether it is achieving their objectives? Response ratios are commonly used as evidence for conservation outcomes of MPAs [@lester2009], but as suggested in @ferraro2018 and @larsen2019 and demonstrated here, without careful attention to the design of control sites (accounting for example for the displacement of effort by MPAs), response ratios may be highly unreliable estimators of regional MPA effects. As @larsen2019 suggests, there are many potential alternatives for estimating the effects of MPAs that better account for the challenges of causal inference. We applied one such approach here (a difference-in-difference estimator), and yet were still unable to reach robust conclusions as to the effect of MPAs on the density of targeted species in the Channel Islands, due to the likely small size of the true effect relative to the strength and variability of environmental drivers.

While this does not mean that all MPAs will face similar challenges in estimating their effects, our results in the rigorously studied Channel Islands system make clear that in many instances empirically detecting a clear regional effect of MPAs may not be possible. How then should stakeholders go about adaptively monitoring and managing MPAs? Simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can perhaps be tuned to focus on the species groups that have the highest chance of a detectable effect size [@nickols2019]. Expanding data collection to include robust monitoring of fleet dynamics may help statistical approaches to isolate the true effect of MPAs on conservation outcomes by allowing for improved control of the effects of fishing dynamics, and allow managers to take into account potential negative interactions between MPAs and fleet dynamics such as those that may occur under constant-catch dynamics. Whenever possible monitoring programs should be implemented prior to MPA implementation to provide a pre-treatment benchmark. Non-equilibrium analyses (Fig.\ref{did-plot}) also help set expectations for effect sizes over time. Beyond that, educating communities about the challenges of estimating the effects of MPAs can help set expectations, so that a lack of a clear effect is not necessarily viewed as a failure of the program, but rather considered in the context of reasonable expectations. While this paper has focused on the conservation outcomes of MPAs, future work must also address the challenge of predicting and estimating the fishery impacts of protected areas.

<!-- Our results provide a unique survey of the likely regional effects of MPAs to fisheries and conservation, and places our empirical assessment of the regional effects of the Channel Islands MPAs in context. The Channel Islands are an intensely studied system, and the challenge of identifying a clear regional effect of the network of MPAs placed there in 2003 may seem surprising. However, our results show that in fact a smaller effect size, from the perspective of regional conservation gains, is to be expected in this system, and therefore the true effect will be very challenging to separate from environmental noise. -->

<!-- Modeling effort such as this can then help manager and stakeholders to set realistic expectations of regional effects for any given MPA network design, and importantly, help inform the design of monitoring programs. For example, monitoring may be targeted at the species and fleets that modeling suggests may provide the clearest indication of MPA-mediated effects. For cases where bio-economic modeling suggests small potential for MPA driven regional density effects, monitoring efforts can be targeted around detecting potential negative effects should they arise, i.e. evidence that the model is wrong, rather than exerting massive amounts of effort on what theory and modeling suggest may be a small and potentially undetectable effect size. -->

<!-- Policy making is inherently an exercise in utilizing best available theory, experience, and modeling to make decisions that are often extremely difficult to empirically test. Causal inference in coupled human and natural systems is incredibly challenging, and our results highlight the particular difficulties in estimating the causal effects of MPAs on regional conservation and fishery outcomes. Bio-economic simulation models such as ours can support the policy evaluation process by helping stakeholders and communities set realistic expectations for what MPAs might achieve, help managers understand the scale and magnitude of MPA effects, and help scientists design cost-effective monitoring plans that stand the best chance possible of detecting these effects.  -->

As the number and size of global MPA networks increase, it is critical that we both set appropriate expectations for their outcomes, and plan how we will monitor the performance of these protected areas over time. While the history of MPA science has made important strides in helping us understand the dynamics of protected areas, the future of MPA science must directly tackle the challenge of evaluating the performance of these MPAs, and adapting their design as needed to best achieve objectives. Commonly employed metrics such as response ratios may be applicable in some circumstances, but can have severe shortcomings as metrics of regional conservation effects. Dependence on unreliable estimators of MPA effects may lead to stakeholders incorrectly attributing negative environmental shocks as MPA failures, or interpreting data arising from scorched earth fishing outside MPAs as a conservation success. Bio-economic modeling can help frame community expectations, reducing the potential for a reduction in support if unrealistic conservation or fishery gains are not realized. Statistical approaches that explicitly address complications such as the spatial spillover effects of MPAs (such as the difference-in-difference approach used here) may give users an improved understanding of the performance of their MPAs. Effective use of simulation and statistics to clearly communicate what we should expect and what we can detect from MPAs is critical in ensuring that MPAs play effective roles in fisheries management and marine conservation.

<!-- Despite our best efforts we are unlikely to ever truly "know" the effect of our efforts to mitigate climate change, but must instead rely on comparisons to best available modeling outcomes to understand how effective our policies have been. Similarly, given the complexities of marine systems much of our decisions on MPA design will have to be based on effective modeling. While we are hardly the first to point out that bio-economic models are a critical tool for MPA design, our results help indicate a minimum floor of model complexity to provide candid assessments of regional MPA effects.  -->

<!-- While factors such as MPA size and degree of depletion are especially strong drivers, for all but the most extreme cases of each a wide array of effects, from negative to highly positive,  are possible based on the complex interaction of factors such as fleet dynamics, movement rates, and recruitment timing. Confronting these interactions by considering the likely parameter space for a given region is a critical step then in understanding what likely regional effects of MPAs are. While models such as ours do require large numbers of parameters that may be challenging to obtain, our results show that working with communities to confront these uncertainties is preferable to sweeping them under the rug in favor of simpler models that are easier to parameter but miss details that our results show can have dramatic effect on expected outcomes.  -->

<!-- We focus mostly on regional conservation gains in this paper. However, fisheries spillover is often another important factor to consider (i.e. are fisheries better off with the MPAs than they would have been without them). The fishery benefits of MPAs are just as (and likely more) intensely debated than the regional conservation outcomes [@hilborn2004; @roberts2001;@sala2018b;@hilborn2018]. We only address fisheries affects briefly in this study, but our results highlight important tradeoffs and synergies between conservation and fishery spillover effects of MPAs. The good news from a fisheries perspective is also fairly obvious: Both the regional conservation and fishery benefits are expected to be greatest when the fished population is in an extremely depleted state pre-MPA, even over a 15 year time horizon, even for larger MPAs (though further work is needed to compare MPAs to alternative fisheries management strategies in these cases). For cases where a valuable and formerly abundant species is overfished, a large MPA may then provide large conservation and fishery gains for that species, while potentially having smaller impacts on other less depleted species. Our simulation results also do identify though a large parameter space where MPAs create tradeoffs between moderate conservation gains and moderate fishing losses. This type of projection analysis can help managers consider where in this space they may be. The most critical point with regards to conservation and fishery effects from our simulation analysis is that the conservation or fishery effects of MPAs cannot be reliably estimated without some knowledge and consideration of the dynamics of the fishing fleet outside the MPAs: over the short-term open access vs constant-catch dynamics can make the difference between a substantial conservation and fisheries win to a more depleted stock with more expensive fishing.  -->

<!-- This process provides a unique survey of the likely regional effects of MPAs to fisheries and conservation, and places our empirical assessment of the regional effects of the Channel Islands MPAs in context. The Channel Islands are an intensely studied system, and the challenge of identifying a clear regional effect of the network of MPAs placed there in 2003 may seem surprising. However, our results show that in fact a smaller effect size, from the perspective of regional conservation gains, is to be expected in this system, and therefore the true effect will be very challenging to separate from environmental noise. The solution then though is not to give up on detecting effects, but rather to shift focus from identifying a specific effect size and instead use simulation analysis to appropriately set expectations for conservation and fishing stakeholders, and design monitoring programs around the species and situations that serve as effective indicators of the ability of an MPA network to achieve its objectives.  -->

## Materials and Methods

We present here critical characteristics of our simulation model and regression approach. Further details can be found in the Supplementary Information. All analysis were conducted in R [@rcoreteam2019]. Our main difference-in-difference model was fit using Template Model Builder in R [@kristensen2016]. A completely reproducible version of our results can be accessed and run at XX.

### Simulation Model

Our bio-economic model simulates the effect of MPAs on a spatially explicit age-structured representation of a single species. Readers can explore the functionality of the model using an online tool available [here](https://danovando.shinyapps.io/simmpa/). See Table.S1 for a complete description of simulation states. The model consists of 25 patches with wrapped edges (picture the waters around a circular island). For any one simulation we randomly pull a species and its associated life history (growth, mortality, maturity) from the `FishLife` [@thorson2017c] package in R. We pair these data with randomly selected values between 0.6 and 0.95 for Beverton-Holt steepness [as parameterized in @mace1994], as well as larval and adult dispersal rates (where at the lowest values adults and larvae stay within the patch they were created, and at the highest values they are capable of moving throughout the system in each time step). We also randomly assign whether adults have density dependent movement (meaning that adult biomass preferentially moves towards patches with lower adult biomass as opposed to random dispersal), as well as one of three potential types of density dependence [building off of @babcock2011]:

  1. Local density dependence: Density dependence occurs independently in each patch, and recruits then disperse to nearby patches

  2. Global density dependence: Density dependence is a function of the sum of spawning biomass across all patches, and recruits are then distributed according to habitat quality

  3. Post-dispersal density dependence: Larvae are distributed throughout the system, and then density dependence occurs based on the density of adult biomass at the destination patch

We allow for three potential siting strategies for MPAs. In the first, MPAs are randomly placed. In the second, we assume that MPAs are placed in preferentially better habit (unfished recruitment is four times greater inside MPA locations). In the third, we allow for scenarios in which MPAs are placed in hotspots of larval dispersal. In this scenario, patches in which MPAs will be placed have larval dispersal rates four times greater than patches that do not become MPAs.

Each simulation is assigned a random set of fleet dynamics from one of three categories: constant-catch, constant harvest rate, or open-access. Under constant-catch, the fleet attempts to catch a randomly selected multiplier of maximum sustainable yield (MSY) each year (note that values of MSY greater than 1 will crash the population eventually, but allows for overfishing dynamics to be observed when shorter time periods are considered). Under a constant harvest rate the fleet captures a randomly selected fraction between 0.01\% and 98\% of the population in each time step. Under open access, fishing effort is allowed to expand and contract in response to the profitability of the fishery, where

\begin{equation}
  E_{t} = E_{t-1}\times\theta{PPUE_{t-1}}
  \label{eqn:oa}
\end{equation}

and

\begin{equation}
  PPUE{t} = \frac{\sum_{i=1}^{I}pCatch_{i,t} - cE_{i,t}^\beta}{\sum_{i=1}^I E_{i,t}}
  \label{eqn:profits}
\end{equation}

To shift the equilibrium of the open access model, we fix price *p* at 1 and $\beta$ at 2, and adjust the cost parameter to achieve the selected fishing mortality rate at bionomic equilibrium. These combinations of fleet models allow each simulation to achieve randomly selected levels of fishing pressure through different processes.

<!-- While at equilibrium in the absence of an MPA each of these fleet models can achieve the same level of depletion, each of these fleet models interacts with MPAs in a different manner.  -->

Along with the fleet dynamics model, each simulation is assigned a random fleet dispersal scenario: uniform dispersal (where the total effort of the fleet is divided evenly among all open patches), catch dispersal (where the total effort of the fleet is divided according to the catchable biomass in each available patch), and profit dispersal (where the total effort of the fleet is divided according to the profit per unit effort in each available patch).

Lastly each simulation is assigned an MPA scenario, defined by the number and size of MPAs, the placement of those MPAs, and the year that the MPAs are put in place. Many MPA models assume equilibrium conditions prior to the MPA, and then measure equilibrium outcomes. While these are important scenarios to understand, they do not reflect the reality of many MPAs, which are often placed in non-equilibrium conditions, and evaluated over the early years of their existence. Each simulation starts the population off at unfished equilibrium and then beings to apply the fleet model. The MPAs are then placed during the randomly selected start year, allowing some runs to explore how the early dynamics of the MPA play out when the fishery and population they are placed on is not already at equilibrium. Fishing effort in displaced by MPAs can either concentrate outside or leave the fishery.

Each simulation is run to equilibrium with and without the selected MPA strategy (holding all else constant). We then measure the difference in biomass and fishery catches in each time step in the scenario with and without the MPAs to calculate the conservation and fishery effects of the MPAs over time. We filter out any scenario that results in a collapse of the population (biomass less than 10% of unfished biomass) prior to the implementation of MPAs, and any scenario in which the open-access dynamics devolve into chaos.

### Difference in Difference Regression

We use a difference-in-difference regression style regression to attempt to estimate the causal effect of MPAs on regional targeted fish biomass. Empirically, this amounts to estimating the pre-post MPA difference in the biomass densities of targeted species net the same difference for non-targeted species in the Channel Islands.

The simplified form of this model is

\begin{equation}
  log(d_{i}) = \beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + e_{i}
\label{eq:did}
\end{equation}

where $d_i$ is the biomass density at observation *i*, *T* indicates whether the observation *i* is for a targeted ($T = 1$) or non-targeted ($T = 0$) species, and *MPA* marks whether observation *i* is in a pre MPA ($MPA = 0$) or post MPA ($MPA = 1$) state. Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment (*MPA*) on the treated (targeted species).

<!-- The purpose of this regression approach is to control for unobserved time-invariant variables that are correlated with the treatment (MPA) and affect the outcome  (biomass densities). This approach depends on the assumption that the the untreated group (in this case non-targeted species) serve as a proxy for the trend in post-treatment densities in the absence of the MPAs (where the intercept terms for the targeted group allows for different mean densities). -->

We briefly assess two of the most critical assumptions of this model here: that the treated and non-treated groups have parallel trends, and that the effect of the treatment on the treated does not tangentially affect the untreated (i.e. that the implementation of MPAs does not indirectly affect the non-targeted group somehow). While the parallel trends assumption cannot be formally proven, we can examine its validity using the data from the years before the MPAs were put in place in 2003 (since after that we no longer expect the trends in the observed data to be parallel). We do not detect any significant differences in the trends of the biomass densities of the targeted and non-targeted species in the years before the MPAs, meaning that we do not have evidence to reject the assumption that densities of targeted and non-targeted species were following similar trajectories before the MPAs (though the model does not estimate a precise zero effect either).

With regards to the second assumption, all of the species in this empirical analysis exist within an ecosystem, and as such affect each other through mechanisms such as predation, competition, and habitat modification. Conceivably then, protection of carnivorous targeted species inside MPAs could drive down the density of non-targeted prey species, serving in that case to positively bias our estimate of the effect of MPAs on the targeted species. While we know dynamics such as this have to exist on some level, we find it unlikely that these effects have had enough time to manifest in the 13 years of post-MPA data used in our analysis [@babcock2010; @pershing2015a].

We used convergent cross mapping (CCM), in the manner of @clark2015, to test for the possibility of the trophic cascades biasing our results. Generalizations of Takens' theorem indicate that if two variables (in our case, species or physical variables) are part of the same dynamic system, their individual dynamics should reflect their relative causal influence. In other words, if one variable is causally forced by another, that forcing should leave a signature on the first time series. Convergent cross mapping (CCM) tests for causation by using the attractor/manifold built from the time series of one variable to predict another (hence the "cross-mapping"). In simple terms, the causal effect of A on B is determined by how well B cross-maps A. CCM then allows us to test for causal relationships in the timeseries of densities of targeted and non-targeted species. Our results found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (see SI for additional information in CCM testing). However, the longer MPAs are in place, the greater the possibility that substantial species interactions that can affect use of non-targeted species as a control may arise.

While Equation.\ref{eq:did} presents the general form of our model, in practice the estimation model is much more involved. At the rawest level, the data are counts of finfish in 2cm length bins along a 30m x 2m transect at various sites and depths. These length bins are converted to biomass, and then biomass densities, by converting length to weights using available allometric data and dividing by the transect area. Our goal is to estimate the effect of the MPAs on these biomass densities of fish throughout the Channel Islands. We fit this model using a hierarchical mixed-effect framework using Template Model Builder [TMB, @kristensen2016] in R. The model consists of three levels, the first (starting from the "bottom") being transect-level densities of fish species observed by PISCO, which are standardized into a standardized biomass abundance index, accounting for both probability of detection and expected density as a result of changes in both abundance and covariates such as visibility and observer experience [see @maunder2004]. For the second stage, we break the abundance indices into targeted and non-targeted species (per the classifications in the PISCO data), and estimate the mean trend of each group (targeted and non-targeted) over time. In the third step, we estimate the difference in the de-meaned trend between the targeted and non-targeted fishes (controlling for factors such as water temperature, kelp cover, and commercial fishery catches), that under the assumptions of the model reflects the causal effect of the MPAs on the outcome of interest (in this case regional biomass density of targeted fishes). All three of these steps are integrated into the same estimation model, in order to propagate uncertainty through the model correctly. Simpler versions of the estimation model, as well as a synthetic control identification strategy, were assessed as well and showed results consistent with the model in our main results (see SI).

<!-- \showmatmethods -->
\showacknow

&nbsp;
&nbsp;
&nbsp;

### 2019-10-09 documents/ovando-regional-effects-of-mpas.tex
Initial commit of cleaned up regional effects repo

--- 2019-10-09 documents/ovando-regional-effects-of-mpas.Rmd
@@
----
-title: The Regional Effects of Marine Protected Areas
-author:
-  - name: Daniel Ovando
-    affiliation: a,1
-    email: danovan@uw.edu
-  - name: Jennifer E. Caselle
-    affiliation: b
-  - name: Christopher Costello
-    affiliation: b   
-  - name: Olivier Deschenes
-    affiliation: b  
-  - name: Steven D. Gaines
-    affiliation: b
-  - name: Ray Hilborn
-    affiliation: a
-  - name: Owen Liu
-    affiliation: b        
-address:
-  - code: a
-    address: University of Washington, School of Aquatic and Fishery Sciences
-  - code: b
-    address: University of California, Santa Barbara
-corresponding_author:
-  - code: 1
-    text: "To whom correspondence should be addressed. E-mail: danovan@uw.edu"
-lead_author_surname: Ovando
-author_contributions: |
-  D.O., S.G., and R.H.
-developed structure of simulation model.
-D.O., O.D., C.C., and J.C.
-developed estimation strategy.
-J.C.
-provided support on collecting and interpreting data.
-All simulations, statistics, and sensitivity analyses performed by D.O.
-and O.L.
-.
-All authors contributed to the manuscript.
-conflict_of_interest: |
-  R.H's research program receives funding from environmental NGOs, foundations, fishing industry, governments and international agencies.
-All of these can be interpreted as a conflict of interest when evaluating fisheries policy.
-C.C., S.G., and O.D's research program includes funding from environmental NGOs and foundations with an interest in ocean conservation and food security
-abstract: |
-  Marine Protected Areas (MPAs) cover between 3-7% of the world's oceans, up from less than 1% in the year 2000.
-The Convention on Biological Diversity calls for 10% of coastal waters to be protected inside MPAs by 2020, while the International Union for Conservation of Nature calls for 30% protection by 2030.
-It is often clear that MPAs produce conservation benefits inside their borders, but many MPAs are also justified on the grounds that they also benefit the broader region outside their borders.
-Despite the expanding role of MPAs in marine resource management, we lack a clear understanding of what the regional effects of MPAs might be, and how we might detect them in scientific studies.
-We used a bio-economic simulation model to demonstrate how the regional effects of MPAs are affected by a suite of economic and environmental drivers and show how commonly used methods to measure inside-MPA effects are likely to fail to capture the regional impacts of protection.
-We present an empirical strategy for estimating regional MPA effects on biomass density using 16 years of data from inside and outside of MPAs within the Channel Islands, California, USA.
-We find no statistically clear effect of MPAs on aggregate mean biomass densities of targeted finfish throughout the study region.
-To make meaningful progress on MPA design and evaluation, the next generation of MPA science must address the challenge of understanding and estimating the regional effects of MPAs across the diversity of species in the ecosystem.
-significance: |
-  Healthy marine ecosystems are critical to the well-being of the planet.
-Marine protected areas, parts of the oceans protected from human activities such as fishing, are increasingly being used in an effort to conserve and manage these ecosystems.
-Our study pairs theory with empirical methods to examine the regional effects of MPAs on conservation outcomes.
-Using simulations, we show that the effects of marine protected areas can be highly variable and dependent on human drivers.
-Building off of MPA theory, we find no statistically clear effect of a network of MPAs, which our simulation analysis suggests is entirely reasonable.
-Our results lay a foundation for future research on the design and monitoring of marine protected areas.
-acknowledgements: |
-  This study would not be possible without the work provided by PISCO divers over the years. DO thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments and technical support. 
-keywords:
-  - Marine Protected Areas
-  - Conservation
-  - Bio-economic modeling
-  - Program Evaluation
-## must be one of: pnasresearcharticle (usual two-column layout), pnasmathematics (one column layout), or pnasinvited (invited submissions only)
-pnas_type: pnasresearcharticle
-bibliography: My Library.bib
-csl: pnas.csl
-lineno: true
-header-includes:
-   - \usepackage{eso-pic,graphicx,transparent}
-output:
-  rticles::pnas_article: default
-  bookdown::word_document2: default
-  bookdown::pdf_document2: default
-  bookdown::html_document2: default
-params:
-  run_name: ["v5.1"]
-  sim_years: [50]
-  burn_years: [20]
-  num_patches: [25]
----
-
-```{r, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dev = "CairoPDF", dpi = 600,
-                      cache = FALSE)
-# knitr::opts_knit$set(root.dir = "/Users/danovando/PhD/zissou") # not entirely reprodcible hack to deal with knit child in dissertation.. not idealk
-
-```
-
-<!-- why is it going to be hard ot detect regional effects of MPAs few if any empirical studies at net regional effects.
-does this mean that i doesn't exist, or impossible tot est, why?
-three fundamental stumbling blocks -->
-
-```{r, include=FALSE}
-library(sf)
-library(ggmap)
-library(viridis)
-library(hrbrthemes)
-library(scales)
-library(patchwork)
-library(rpart)
-library(extrafont)
-library(caret)
-library(ggsci)
-library(rEDM)
-library(tidyverse)
-extrafont::loadfonts()
-functions <- list.files(here::here("functions"))
-
-walk(functions, ~ here::here("functions", .x) %>% source()) # load local functions
-save_plots <- TRUE
-
-# sim_years <- 50
-# 
-# burn_years <- 20
-# 
-# num_patches <- 25
-
-# run_name <- "v5.0"
-
-run_name <- params$run_name
-
-sim_years <- params$sim_years
-
-burn_years <- params$burn_years
-
-num_patches <- params$num_patches
-
-run_dir <- here::here("results", run_name)
-
-experiment_dir <- here::here("results",run_name, "experiments")
-
-load(file = here::here("results",run_name, "rawish_zissou_data.Rdata"))
-
-load(file = here::here("results",run_name, "model_runs.Rdata"))
-
-processed_grid <-  read_rds(file.path(run_dir,"filtered_processed_grid.rds"))
-
-load(file = file.path(run_dir ,"sim_grid.Rdata"))
-
-cdfw_catches <-
-  read_csv(file = here::here("data", 'cdfw-catches.csv')) %>%
-  group_by(sci_name, year) %>%
-  summarise(catch = sum(pounds_caught, na.rm = T)) %>%
-  left_join(
-    life_history_data %>% select(taxa, classcode, commonname) %>% mutate(taxa = tolower(taxa)),
-    by = c("sci_name" = "taxa")
-  ) %>%
-  filter(!is.na(classcode))
-
-# cdfw_catches %>%
-#   group_by(commonname, classcode, sci_name) %>%
-#   summarise(catch =sum(catch, na.rm = TRUE)) %>%
-#   mutate(in_analysis = classcode %in% unique(cip_data$classcode)) %>%
-#   View()
-
-density_ratios <- readRDS(file.path(run_dir,"density_ratios.rds"))
-
-outcomes <- readRDS(file.path(run_dir,"outcomes.rds"))
-
-outcomes <- outcomes %>% 
-  filter(!is.na(mpa_effect))
-
-density_ratios <- density_ratios %>% 
-  filter(!is.na(biased_density_ratio))
-
-   eqo <- outcomes %>%
-     filter(year == max(year))
-
-facet_labels <- c(
-  mpa_size = "Range in MPA",
-  depletion = "Depletion"
-)
-
-channel_islands <- readRDS(here::here("data","channel_islands_map.RDS"))
-
-ca_mpas <- sf::st_read(here::here("data","MPA_CA_Existing_160301")) %>%
-  rmapshaper::ms_simplify() %>%
-  sf::st_transform(crs = 4326)
-
-zissou_theme <-
-  theme_ipsum(
-    base_size = 12,
-    axis_title_size = 12,
-    strip_text_size = 12,
-    base_family = "Fira Sans"
-  )
-
-testplot <- ggplot()
-
-panel_height = unit(1,"npc") - sum(ggplotGrob(testplot)[["heights"]][-3]) - unit(4,"line")
-
-gc <- guide_colorbar(frame.colour = "black",
-                     ticks.colour = "white",
-                     barheight = panel_height)
-
-  hgc <- guide_colorbar(frame.colour = "black",
-                       ticks.colour = "black",
-                       barwidth = 10)
-
-    wide_hgc <- guide_colorbar(frame.colour = "black",
-                       ticks.colour = "black",
-                       barwidth = 25)
-
-theme_set(zissou_theme)
-
-models_worked <- model_runs$tmb_fit %>% map("error") %>% map_lgl(is_null)
-
-model_runs <- model_runs %>%
-  filter(models_worked) %>%
-  mutate(tmb_fit = map(tmb_fit,"result")) %>%
-  mutate(processed_fits = map(tmb_fit, process_fits)) %>%
-  mutate(did_plot = map(processed_fits, "did_plot"))
-
-base_run <- model_runs %>%
-filter(var_names == "pisco_a", mpa_only == FALSE, center_scale == TRUE)
-
-mpa_run <- model_runs %>%
-  filter(data_source == "pisco",
-         var_names == "pisco_a",
-         mpa_only == TRUE)
-
-kfm_run <- model_runs %>%
-  filter(data_source == "kfm")
-
-fitted_data <- base_run$data[[1]]
-
-zissou_fit <- base_run$tmb_fit[[1]]
-
-report <- base_run$tmb_fit[[1]]
-
-site_data <- read_csv(here::here("data",'Final_Site_Table_UCSB.csv')) %>%
-  magrittr::set_colnames(., tolower(colnames(.))) %>%
-  select(
-    site,
-    side,
-    mpagroup,
-    mpa_status,
-    reserve,
-    region,
-    year_mpa,
-    mpaareanm2,
-    lat_wgs84,
-    lon_wgs84
-  ) %>%
-  rename(lat_wgs84 = lon_wgs84,
-         lon_wgs84 = lat_wgs84) %>%
-  unique() %>%
-  mutate(eventual_mpa = (year_mpa > 0))
-
-seen_non_nested_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "seen_non_nested_betas")) %>%
-  rename(group = variable) %>%
-  mutate(variable  = zissou_fit$zissou_data$x_seen_non_nested %>% colnames())
-
-seeing_non_nested_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "seeing_non_nested_betas")) %>%
-  rename(group = variable) %>%
-  mutate(variable  = zissou_fit$zissou_data$x_seen_non_nested %>% colnames())
-
-seen_year_species_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "seen_year_species_betas")) %>%
-  rename(group = variable) %>%
-  mutate(variable = zissou_fit$zissou_data$x_seen_year_species %>% colnames())
-
-seeing_year_species_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "seeing_year_species_betas")) %>%
-  # filter(variable == "seeing_year_species_betas") %>%
-  rename(group = variable) %>%
-  mutate(variable = zissou_fit$zissou_data$x_seen_year_species %>% colnames())
-
-seen_region_cluster_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "seen_region_cluster_betas")) %>%
-  # filter(variable == "seen_region_cluster_betas") %>%
-  rename(group = variable) %>%
-  mutate(variable = zissou_fit$zissou_data$x_seen_region_cluster %>% colnames())
-
-seeing_region_cluster_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "seeing_region_cluster_betas")) %>%
-  # filter(variable == "seeing_region_cluster_betas") %>%
-  rename(group = variable) %>%
-  mutate(variable = zissou_fit$zissou_data$x_seen_region_cluster %>% colnames())
-
-did_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "mpa_effect")) %>%
-  # filter(variable == "mpa_effect") %>%
-  mutate(group = variable) %>%
-  mutate(year = zissou_fit$did_data$year %>% unique())
-
- betas <- bind_rows(
-   seen_non_nested_betas,
-   seeing_non_nested_betas,
-   seen_year_species_betas,
-   seeing_year_species_betas,
-   seen_region_cluster_betas,
-   seeing_region_cluster_betas,
-   did_betas %>% select(-year)
- ) %>%
-   as_data_frame()
-
- non_nested_beta_plot <- betas %>%
-   filter(str_detect(group, "non_nested")) %>%
-   ggplot() +
-   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
-   geom_pointrange(aes(x = variable,
-                       y = estimate,
-                       ymin = lower,
-                       ymax = upper)) +
-   facet_wrap(~group) +
-   coord_flip() +
-   theme(axis.text.y = element_text(size = 10))
-
- year_species_effects_plots <- betas %>%
-   filter(str_detect(group, "year_species_betas")) %>%
-   mutate(year = str_replace_all(variable,"\\D","") %>% as.numeric()) %>%
-   mutate(classcode = str_split(variable,'-', simplify = T)[,2]) %>%
-   ggplot() +
-   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
-   geom_ribbon(aes(x = year, ymin = lower, ymax = upper, fill = classcode), alpha = 0.25) +
-   geom_line(aes(x = year, y = estimate, color = classcode)) +
-   facet_wrap(~group)
-
-  region_cluster_plots <- betas %>%
-   filter(str_detect(group, "region_cluster_betas")) %>%
-   mutate(cluster = str_replace_all(variable,"\\D","")) %>%
-   mutate(region = str_split(variable,'-', simplify = T)[,3]) %>%
-   ggplot() +
-    geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
-    geom_pointrange(aes(x = region,
-                        y = estimate,
-                        ymin = lower,
-                        ymax = upper, color = cluster)) +
-    facet_wrap(~group)
-
-top_species <- base_run$data[[1]]$classcode %>% unique()
-
-life_summaries <- life_history_data %>% 
-  select(classcode, tm, vbgf.k, vbgf.linf) %>% 
-  rename(age_mature = tm, vbk = vbgf.k, linf = vbgf.linf) %>% 
-  unique() %>% 
-  filter(classcode %in% top_species) %>% 
-  select(-classcode) %>% 
-  gather(variable,value) %>% 
-  group_by(variable) %>% 
-  summarise(min_val = min(value, na.rm = TRUE),
-         max_val = max(value, na.rm = TRUE))
-
- sim_life_history <- sim_grid %>%
-    mutate(
-      age_mature = map_dbl(fish, "age_mature"),
-      vbk = map_dbl(fish, "vbk"),
-      linf = map_dbl(fish, "linf")
-    ) %>%
-    select(scientific_name, age_mature, vbk, linf) %>%
-    unique() %>% 
-   filter(age_mature > life_summaries$min_val[life_summaries$variable == "age_mature"],
-          age_mature < life_summaries$max_val[life_summaries$variable == "age_mature"],
-          vbk > life_summaries$min_val[life_summaries$variable == "vbk"],
-          vbk < life_summaries$max_val[life_summaries$variable == "vbk"],
-          linf > life_summaries$min_val[life_summaries$variable == "linf"],
-          linf < life_summaries$max_val[life_summaries$variable == "linf"])
-
-  sim_plot <- outcomes %>%
-    mutate(rsize = plyr::round_any(mpa_size, 0.25),
-           reff = plyr::round_any(mpa_effect,0.2)) %>%
-    filter(rsize == 0.25, depletion < 0.8, depletion > 0.2,
-           scientific_name %in% sim_life_history$scientific_name
-           ) %>%
-    mutate(year = years_protected + 2001) %>%
-    filter(year %in% did_betas$year, years_protected >1) %>% 
-    group_by(year, reff) %>% 
-    count() %>% 
-    group_by(year) %>% 
-    mutate(sc = n / sum(n))
-
-  did_plot <- did_betas %>%
-    ggplot() +
-    geom_tile(data = sim_plot, aes(year, reff, fill = sc), alpha = 0.75) +
-    geom_vline(aes(xintercept = 2003), color = 'red', linetype = 2, size = 2) +
-    geom_hline(aes(yintercept = 0)) +
-    geom_pointrange(aes(
-      year,
-      y = exp(estimate) - 1,
-      ymin = exp(lower) - 1,
-      ymax = exp(upper) - 1
-    ),
-    size = 1.5
-    ) +
-    labs(x = "Year", y = "Targeted Trend Divergence") +
-    scale_fill_viridis(guide = wide_hgc,
-                       name = "% of Annual Sims",
-                       labels = scales::percent_format(accuracy = 1),
-                       option = "magma") + 
-    theme(legend.position = "top",
-        panel.grid.major.x = element_blank(),
-        panel.grid.minor.x = element_blank()
-        )
-
-cip_data <- pisco_data %>%
-  left_join(site_data, by = c("site","side")) %>%
-  filter(is.na(eventual_mpa) == F) %>%
-  filter(region %in% c("ANA", "SCI","SRI",'SMI'),
-         classcode %in% top_species)
-
-# process outcomes
-
-```
+\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
+
+%% Some pieces required from the pandoc template
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+% Use the lineno option to display guide line numbers if required.
+% Note that the use of elements such as single-column equations
+% may affect the guide line number alignment.
+
+\usepackage[T1]{fontenc}
+\usepackage[utf8]{inputenc}
+
+\usepackage{eso-pic,graphicx,transparent}
+
+\templatetype{pnasresearcharticle}  % Choose template
+
+\title{The Regional Effects of Marine Protected Areas}
+
+\author[a,1]{Daniel Ovando}
+\author[b]{Jennifer E. Caselle}
+\author[b]{Christopher Costello}
+\author[b]{Olivier Deschenes}
+\author[b]{Steven D. Gaines}
+\author[a]{Ray Hilborn}
+\author[b]{Owen Liu}
+
+  \affil[a]{University of Washington, School of Aquatic and Fishery Science}
+  \affil[b]{University of California, Santa Barbara}
+
+% Please give the surname of the lead author for the running footer
+\leadauthor{Ovando}
+
+% Please add here a significance statement to explain the relevance of your work
+\significancestatement{Healthy marine ecosystems are critical to the well-being of the planet.
+Marine protected areas, parts of the oceans protected from human
+activities such as fishing, are increasingly being used in an effort to
+conserve and manage these ecosystems. Our study pairs theory with
+empirical methods to examine the regional effects of MPAs on
+conservation and fishery outcomes. Using simulations, we show that the
+effects of marine protected areas can be highly variable and dependent
+on human drivers. Building off of MPA theory, we find no statistically
+clear effect of a network of MPAs, which our simulation analysis
+suggests is entirely reasonable. Our results lay a foundation for future
+research on the design and monitoring of marine protected areas.}
+
+\authorcontributions{D.O., S.G., and R.H. developed structure of simulation model. D.O.,
+O.D., C.C., and J.C. developed estimation strategy. J.C. provided
+support on collecting and interpreting data. All simulations,
+statistics, and sensitivity analyses performed by D.O. and O.L. . All
+authors contributed to the manuscript.}
+
+\authordeclaration{R.H's research program receives funding from environmental NGOs,
+foundations, fishing industry, governments and international agencies.
+All of these can be interpreted as a conflict of interest when
+evaluating fisheries policy. C.C., S.G., and O.D's research program
+includes funding from environmental NGOs and foundations with an
+interest in ocean conservation and food security}
+
+\correspondingauthor{\textsuperscript{} }
+
+% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
+ \keywords{  Marine Protected Areas |  Conservation |  Bio-economic modeling |  Program Evaluation |  Channel Islands National Marine Sanctuary  }
+
+\begin{abstract}
+Marine Protected Areas (MPAs) cover between 3-7\% of the world's oceans,
+up from less than 1\% in the year 2000. The Convention on Biological
+Diversity calls for 10\% of coastal waters to be protected inside MPAs
+by 2020, while the International Union for Conservation of Nature calls
+for 30\% protection by 2030. It is often clear that MPAs produce
+conservation benefits inside their borders, but many MPAs are also
+justified on the grounds that they also benefit the broader region
+outside their borders. Despite the expanding role of MPAs in marine
+resource management, we lack a clear understanding of what the regional
+effects of MPAs might be, and how we might detect them in scientific
+studies. We used a bio-economic simulation model to demonstrate how the
+regional effects of MPAs are affected by a suite of economic and
+environmental drivers and show how commonly used methods to measure
+inside-MPA effects are likely to fail to capture the regional impacts of
+protection. We present an empirical strategy for estimating regional MPA
+effects on biomass density using 16 years of data from inside and
+outside of MPAs within the Channel Islands, California, USA. We find no
+statistically clear effect of MPAs on aggregate mean biomass densities
+of targeted finfish throughout the study region. To make meaningful
+progress on MPA design and evaluation, the next generation of MPA
+science must address the challenge of understanding and estimating the
+regional effects of MPAs across the diversity of species in the
+ecosystem.
+\end{abstract}
+
+\dates{This manuscript was compiled on \today}
+\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
+
+\begin{document}
+
+% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
+% You should only change this length when you've finalised the article contents.
+\verticaladjustment{-2pt}
+
+\maketitle
+\thispagestyle{firststyle}
+\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
+
+% If your first paragraph (i.e.
+with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation.
+If this is the case, add \parshape=0 to the end of the list environment.
+
+\acknow{This study would not be possible without the work provided by PISCO
+divers over the years. DO thanks Cody Szuwalski, Julia Lawson, and André
+Punt for helpful comments and technical support.}
 
@@
 
-<!-- \AddToShipoutPictureFG{ -->
-<!--   \AtPageCenter{% or \AtTextCenter -->
-<!--     \makebox[0pt]{\rotatebox[origin=c]{45}{% -->
-<!--       \scalebox{5}{\texttransparent{0.3}{DRAFT}}% -->
-<!--     }} -->
-<!--   } -->
-<!-- } -->
-
-No-take Marine Protected Areas (MPAs), spatial regions of the ocean in which fishing is prohibited, have a long history in the management of marine resources.
-Traditional cultures in Oceania utilized - often temporary - MPAs as "fish banks" for times of need [@johannes1978].
-Modern MPAs were first established primarily as marine analogs to the terrestrial protection of iconic landscapes like Yellowstone or Kruger National Parks [@iucn1976;@watson2014b].
-Over time our goals and expectations for MPAs have evolved; while all MPAs are expected to deliver conservation benefits within their borders, many modern MPAs are also established to  bolster fish populations throughout the region in which they are located (what we term “regional-scale effects”) [@gaines2010].
-
-Indeed, many recent agreements to expand MPAs (most proximately, the Convention on Biological Diversity’s Strategic Plan for Biodiversity, which calls for 10% of coastal waters to be protected inside MPAs by 2020 and the International Union for Conservation of Nature call for 30% by 2030) rely on the belief that well-designed MPAs will achieve benefits both within, and outside, their borders.
-Despite these assumptions, our collective scientific understanding of the regional-scale conservation and fishery impacts of current and future MPAs is surprisingly limited.
+No-take Marine Protected Areas (MPAs), spatial regions of the ocean in
+which fishing is prohibited, have a long history in the management of
+marine resources. Traditional cultures in Oceania utilized - often
+temporary - MPAs as ``fish banks'' for times of need (1). Modern MPAs
+were first established primarily as marine analogs to the terrestrial
+protection of iconic landscapes like Yellowstone or Kruger National
+Parks (2, 3). Over time our goals and expectations for MPAs have
+evolved; while all MPAs are expected to deliver conservation benefits
+within their borders, many modern MPAs are also established to bolster
+fish populations throughout the region in which they are located (what
+we term ``regional-scale effects'') (4).
+
+Indeed, many recent agreements to expand MPAs (most proximately, the
+Convention on Biological Diversity's Strategic Plan for Biodiversity,
+which calls for 10\% of coastal waters to be protected inside MPAs by
+2020 and the International Union for Conservation of Nature call for
+30\% by 2030) rely on the belief that well-designed MPAs will achieve
+benefits both within, and outside, their borders. Despite these
+assumptions, our collective scientific understanding of the
+regional-scale conservation and fishery impacts of current and future
+MPAs is surprisingly limited.
 
 What is the scientific evidence of the conservation effects of MPAs?
-Numerous studies provide evidence that well-enforced and appropriately sized MPAs can produce conservation benefits within their borders [@lester2009;@halpern2003; @edgar2014; @gerber2005].
-As these conservation benefits accrue inside MPAs, theory holds that MPAs can affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
-Several studies have documented empirical evidence for the existence of adult or larval spillover affecting both abundance and fisheries [e.g.
-@goni2010; @halpern2009;@kay2012; @stobart2009; @mcclanahan2000; @russ1996; @thompson2017;@baetscher2019;@pelc2009], as well as alteration of fishing effort in reaction to [@costa2013;@mason2012;@murawski2005] and in anticipation of [@mcdermott2019] MPA placement.
-The potentially more important question, however, is not whether spillover occurs (it must to some degree in any realistic scenario), but what the net effects of spillover are and whether those effects are empirically detectable.
-From a fishery perspective, are spillover benefits sufficient to offset losses in fishing grounds and changes in responses of displaced fishers caused by an MPA?
-From a conservation perspective, how much does the buildup of fish inside an MPA increase biomass outside the protected area?
-Overall, what are the regional effects of MPAs?
-
-As stakeholders around the world increasingly seek to use MPAs in marine resource management portfolios, and the performance of existing MPAs is evaluated, it is critical that we develop a better understanding of the magnitude and drivers of regional-scale MPA effects.
-To address this gap, this study examines two critical questions: 1) What do we expect the regional-scale conservation effects of MPAs to be and 2) When (and how) can we expect to empirically detect these effects?
-We answer these questions using a simulation analysis to frame the theoretical regional conservation impacts of MPAs.
-We then present an empirical test of the evidence for regional-level conservation effects of MPAs resulting from a network of closures put in place in the Channel Islands, California, USA, in 2003.
-
-What Are the Regional Effects of MPAs? {.unnumbered}
-----------------------
-
-<!-- **key thing: a lot more ways for small things to happen than big things ** -->
-
-<!-- xx need to add in discussion of other outcomes not covered here, e.g. resiliency, tourism, etc. xx -->
-
-Much of the literature on the effects of Marine Protected Areas focuses on assessing the conservation effects within the borders of protected areas [e.g.
-@edgar2014].
-While these within-MPA effects are vitally important for protecting rare species, biodiversity, critical habitats, and often tourism, they paint an incomplete picture of the overall population effects of MPAs.
-The organisms within the borders of protected areas are generally part of a broader population, the biological stock, connected through adult movement and larval dispersal.
-If the goal of conservationists or natural resource managers is to increase the total abundance or productivity of a resource, a broader question we should ask of MPAs is not just are there more fish inside their borders, but also how have the reserves affected abundances throughout the region in which they are located?
-This logic, that MPAs will have conservation benefits for most species beyond their borders, is implicit in all multilateral calls for MPA expansion.
-
-We define the regional conservation effects of MPAs as the change in total biomass of fish (summing inside and outside of MPAs) relative to the total biomass of fish that would have occurred without the MPAs (acknowledging that other outcomes such as increased biodiversity or resiliency are also important to conservation but are beyond the scope of this analysis).
-Note that this definition of regional effects is in line with the recommendations of @ferraro2018, but differs from effects measured by for example differences in outcomes before and after MPAs.
-
-We used a simulation model to examine the regional conservation effects of MPAs under a wide array of conditions that might reasonably be expected to affect these outcomes.
-The model explores combinations of MPA designs (e.g., sizes, number, and placement of MPAs), life histories (e.g., growth and mortality rates, adult movement, larval dispersal, recruitment variability and autocorrelation, and timing of density dependence), and fishing fleet dynamics (e.g., degree of fishing pressure, size selectivity of the fleet, reaction of fishing pressure to MPA placement) (See Table.S1 for a complete description of simulation variables).
-We use this model to explore how different drivers interact to affect the conservation and fishery outcomes of MPAs.
-
-<!-- Numerous works have used simulation modeling to explore the effects of MPAs on outcomes such as fisheries yields and stock abundance (see @fulton2014 and references within),but these prior studies have either focused on the effect of a select set of drivers (such as dispersal rates) on outcomes, or have been highly localized assessments that incorporate multiple drivers but are tuned to one specific region.
-Our model captures critical bio-economic features, while still being flexible and fast enough to be generalized to a broad array of scenarios.
-We use this model to explore how different drivers interact to affect the conservation and fishery outcomes of MPAs.
--->
-
-### Theoretical Conservation Effects {#conservation .unnumbered}
-
-As a thought experiment, imagine a region that has driven its fish populations to near extinction and then places 100\% of its waters inside a no-take MPA.
-In that setting, we would expect the regional-scale conservation benefits to be massive.
-On the other hand, imagine implementing an MPA in an area where a sedentary species has been only lightly fished.
-In that setting, it is possible to create a small net conservation loss if the concentration of fishing pressure outside the reserve has a greater effect than the biomass buildup inside the reserve.
-Within these broad bounds, numerous factors can act to affect the regional effects of MPAs.
-These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008], the strength and timing of density dependence in the population (e.g.
-pre- or post-settlement), how overfished the population would be without the MPA, and how fishing activity and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @gerber2005; @hilborn2004a; @gaines2010; @botsford2003].
-In addition, even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through habitat and network effects [@gaines2010; @costello2010].
-
-```{r}
-short_term <- outcomes %>%
-  group_by(experiment) %>% 
-  filter(years_protected == max(years_protected), years_protected >= 0) %>% 
-  ungroup() %>% 
-  filter(!is.na(mpa_effect))
-
-median_effect <- median(short_term$mpa_effect[short_term$year == max(short_term$year)], na.rm = TRUE)
-
-median_pop_effect <- median(short_term$pop_effect[short_term$year == max(short_term$year)], na.rm = TRUE)
-
-# median_fish_effect <- median(fishery_short_term$mpa_effect[short_term$year == max(short_term$year)], na.rm = TRUE)
-# 
-# median_fish_effect <- median(fishery_short_term$fishery_effect[short_term$year == max(short_term$year)], na.rm = TRUE)
-
-pos_runs <- round(mean(short_term$pop_effect[short_term$year == max(short_term$year)] > 0),2)
-
-# pos_fish_runs <- round(mean(fishery_short_term$msy_effect[fishery_short_term$year == max(fishery_short_term$year)] > 0, na.rm = TRUE),2)
-
-pos_fish_effect <- round(mean(short_term$f[short_term$year == max(short_term$year)] > 0),2)
-
-small_effect <- round( 100* median(short_term$mpa_effect[short_term$depletion <= 0.4]))
-
-min_small_effect <- round( 100* min(short_term$mpa_effect[ short_term$depletion <= 0.4]))
-
-max_small_effect <- round( 100* max(short_term$mpa_effect[ short_term$depletion <= 0.4]))
-
-medium_effect <- round(100*median(short_term$mpa_effect[short_term$depletion > 0.4 & short_term$depletion <= 0.6]))
-
-min_medium_effect <- round(100*min(short_term$mpa_effect[short_term$depletion > 0.4 & short_term$depletion <= 0.6]))
-
-max_medium_effect <- round(100*max(short_term$mpa_effect[short_term$depletion > 0.4 & short_term$depletion <= 0.6]))
-
-big_effect <- round(100*median(short_term$mpa_effect[short_term$depletion > 0.6]))
-
-min_big_effect <- round(100*min(short_term$mpa_effect[short_term$depletion > 0.6]))
-
-max_big_effect <- round(100*max(short_term$mpa_effect[short_term$depletion > 0.6]))
-
-```
-
-<!-- Across all simulations the median simulated equilibrium regional conservation effect, expressed as the percentage change in total biomass with the MPAs and without the MPAs, was `r round(100*median(eqo$mpa_effect))`%, with a minimum of less than `r round(100*min(eqo$mpa_effect))`% and a maximum of over `r round(100*max(eqo$mpa_effect))`%.
-The median time required for the simulated MPAs to cause a 10% increase in regional biomass was approximately 10 years.
--->
-
-We tested the effects of different combinations of these theoretical drivers of MPA conservation effects using over 10,000 simulated MPA scenarios.
-A striking outcome of this simulation analysis is how the level of fishing pressure can cause dramatic and predictable shifts in conservation outcomes.
-We measure fishing pressure by the degree of simulated depletion (percent of unfished biomass gone, such that a depletion of 0% means the population is at 100% of unfished biomass, 100% means 0% of unfished biomass remains) in the scenario without MPAs.
-When MPAs were placed in relatively unexploited fisheries (depletions less than 40%), the median regional conservation effect was barely noticeable (`r small_effect`%).
-For ecosystems fished near MSY levels (depletion of 40%-60%), the median conservation effect was `r medium_effect`%, while for overfished ecosystems (without-MPA depletions above 60%) the median regional conservation effect was over  `r big_effect`% (Fig.\ref{conservation-effects}-A).
-Across all of these groups though, effects as low as declines of -50% and as high as an increase of more than 100% were possible.
-It should be noted though that these runs represent the range of scenarios evaluated in our simulation, and that certain types of scenarios may be much more likely than others in the real world.
-To put these results into context, the FAO estimates that 7% of the worlds fisheries with status estimates fall into the relatively unexploited category, 60% fall into the fully fished category, and 33% fall into the heavily fished category [@fao2018], though works that include a broader range of fisheries estimate that 50% or more of stocks to fall into the heavily fished category [@costello2012;@rosenberg2017].
-
-These simple results support the intuitive conclusion that bigger MPAs on overfished populations produce large positive conservation outcomes.
-However, when combined with other factors such as fleet dynamics, life history, and MPA size, the conservation effects of MPAs can vary widely in magnitude and timing, a finding corroborated by empirical evidence [@starr2015;@caselle2015;@nickols2019;@villasenor-derbez2019] (Fig.\ref{conservation-effects}-B).
-MPAs protecting small proportions of a stock can produce small positive effects, large positive effects, or even small negative effects.
-As the area protected inside MPAs increases, positive effects become more likely, but even MPAs protecting most of a population's footprint can produce conservation effects as low as 0% and as high as 200% and above depending on the context in which they are placed.
-The degree of depletion (i.e.
-fishing pressure) has a clearer signal, with small conservation benefits of MPAs when depletion is low, and larger effects when depletion is high, but again even for severely depleted populations the conservation effects of MPAs can vary widely (although nearly always with population benefits).
-
-While positive conservation outcomes were much more likely to occur across our simulations (`r percent(pos_runs, accuracy = 1)` of simulations), the `r percent(1 - pos_runs, accuracy = 1)` of runs that produced net conservation losses are of particular interest.
-These runs occurred almost exclusively when the simulated fleet followed a "constant-catch" fishing strategy.
-Under the constant-catch fleet model, fishing communities (or fishery policies) seek to catch the same amount regardless of the presence of an MPA.
-While a constant-catch greater than MSY is not possible over the long-term under the assumptions of our model, over the short-term a constant-catch scenario is not implausible.
-Subsistence fisheries may use a constant-catch style policy over the short-term, as they seek to ensure that their food needs are met.
-More industrial fisheries may have pre-arranged agreements with buyers to deliver set amounts of fish.
-Constant-catch dynamics might also occur in fisheries with constraining quotas that are not updated after the implementation of MPAs.
-These patterns suggest that coupling MPA implementation with appropriate management reforms that reduce the effects of constant-catch strategies can move the predicted conservation effects of MPAs to being nearly uniformly positive.
-The possibility of constant-catch strategies to result in net MPA conservation losses highlights the critical importance of understanding the broader management context in which MPAs are placed.
-
-```{r conservation-effect, include = FALSE}
-
-pop_time_plot <- outcomes %>%
-  filter(years_protected > 0) %>%
-  ggplot(aes(years_protected, mpa_effect)) +
-  geom_bin2d() +
-  scale_fill_viridis(guide = gc)
-
-ylabs <- c(expression(""<= "-50%"), paste0(seq(-25,75, by = 25),"%"),expression("">= "100%") )
-
-ybreaks = seq(-.50,1, by = .25)
-
-pop_depletion_plot <- outcomes %>%
-  filter(year == max(year)) %>%
-  ggplot() +
-  geom_bin2d(aes(depletion, mpa_effect), binwidth = c(.1, .1),
-             show.legend = FALSE) +
-  scale_fill_viridis(
-  option = "A",
-  # trans = "log10",
-  guide = guide_colorbar(frame.colour = "black",
-  frame.linewidth = 1,),
-  name = "Median Effect"
-  ) +
-  scale_x_continuous(labels = scales::percent_format(accuracy = 1), name = "Depletion") +
-  scale_y_continuous(labels = ylabs, breaks = ybreaks, name = "")
-
-pop_depletion_plot <- outcomes %>%
-    filter(year == max(year)) %>%
-    mutate(
-      rdep = cut(
-        depletion,
-        breaks = seq(0, 1, by = .2),
-        labels = seq(0, .8, by = .2),
-        include.lowest = TRUE
-      )
-      ,
-      reff = cut(
-        mpa_effect,
-        breaks = seq(-.6, 1, by = .2),
-        labels = seq(-.6, .8, by = .2),
-        include.lowest = TRUE
-      )
-    ) %>%
-    group_by(rdep, reff) %>%
-    count() %>%
-    group_by(rdep) %>%
-    mutate(pn = n / sum(n)) %>%
-    ungroup() %>%
-    mutate(rdep = as.numeric(as.character(rdep)),
-           reff = as.numeric(as.character(reff))) %>%
-    ggplot(aes(rdep, reff, fill = pn)) +
-    geom_raster(hjust = 1, vjust = 1, show.legend = FALSE) +
-    scale_fill_viridis(
-      option = "A",
-      # trans = "log10",
-      guide = hgc,
-      name = "% of Sims",
-      labels = scales::percent_format(accuracy = 1)
-    ) +
-    scale_x_continuous(labels = scales::percent_format(accuracy = 1),
-                       name = "Depletion",
-                       expand = expand_scale(add = c(0,0))) +
-    scale_y_continuous(labels = ylabs,
-                       breaks = ybreaks,
-                       name = "")
-
-  sim_plot <- outcomes %>%
-    mutate(rsize = plyr::round_any(mpa_size, 0.25),
-           reff = plyr::round_any(mpa_effect,0.2)) %>%
-    filter(rsize == 0.25, depletion < 0.8, depletion > 0.2,
-           scientific_name %in% sim_life_history$scientific_name
-           ) %>%
-    mutate(year = years_protected + 2001) %>%
-    filter(year %in% did_betas$year, years_protected >1) %>% 
-    group_by(year, reff) %>% 
-    count() %>% 
-    group_by(year) %>% 
-    mutate(sc = n / sum(n))
-
-  pop_size_plot <- outcomes %>%
-    filter(year == max(year)) %>%
-    mutate(
-      rsize = cut(
-        mpa_size,
-        breaks = seq(0, 1, by = .2),
-        labels = seq(0, .8, by = .2)
-      )
-      ,
-      reff = cut(
-        mpa_effect,
-        breaks = seq(-.6, 1, by = .2),
-        labels = seq(-.6, .8, by = .2)
-      )
-    ) %>%
-    group_by(rsize, reff) %>%
-    count() %>%
-    group_by(rsize) %>%
-    mutate(pn = n / sum(n)) %>%
-    ungroup() %>%
-    mutate(rsize = as.numeric(as.character(rsize)),
-           reff = as.numeric(as.character(reff))) %>%
-    ggplot(aes(rsize, reff, fill = pn)) +
-    geom_raster(hjust = 1, vjust = 1) +
-    scale_fill_viridis(
-      option = "A",
-      # trans = "log10",
-      guide = hgc,
-      name = "% of Sims",
-      labels = scales::percent_format(accuracy = 1)
-    ) +
-    scale_x_continuous(labels = scales::percent_format(accuracy = 1),
-                       name = "Range in MPA",
-                       expand = expand_scale(add = c(0,0))) +
-    scale_y_continuous(labels = ylabs,
-                       breaks = ybreaks,
-                       name = "Conservation Effect") +
-    theme(legend.position = "top") +
-    labs(title = 'B)')
-
- pop_facet_effect_plot <- outcomes %>%
-  filter(year == max(year)) %>%
-   select(mpa_effect, mpa_size, depletion) %>%
-   gather(measure, value, -mpa_effect) %>%
-   ggplot() +
-   geom_bin2d(aes(value, mpa_effect), binwidth = c(.1, .1), show.legend = TRUE) +
-  scale_fill_viridis(
-  option = "A",
-  # trans = "log10",
-  guide = hgc,
-  name = "# of Sims"
-  ) +
-   facet_wrap(~measure, labeller = labeller(measure = facet_labels), strip.position = "bottom") +
-   scale_x_percent(name = "") +
-   scale_y_percent(name = "Conservation Effect") + 
-   theme(legend.position = "top")
-
-  pop_depletion_and_size_plot <- outcomes %>%
-    filter(years_protected >= 0) %>%
-    group_by(experiment) %>%
-    mutate(depletion = plyr::round_any(depletion[years_protected == 0], .1),
-           mpa_size = plyr::round_any(mpa_size, .1)) %>%
-    filter(year == max(year)) %>%
-    group_by(depletion, mpa_size) %>%
-    summarise(median_mpa_effect = median(mpa_effect, na.rm = TRUE)) %>%
-    ggplot(aes(depletion, mpa_size, fill = median_mpa_effect)) +
-    geom_tile() +
-    # geom_contour(aes(z = median_mpa_effect, color = ..level..), show.legend = FALSE) +
-    scale_fill_viridis(labels = scales::percent_format(accuracy = 1),
-                       guide = hgc,
-                       name = "Median Effect") +
-  labs(x = "Depletion", y = "Range in MPA", title = "A)") +
-    scale_y_continuous(labels = scales::percent_format(accuracy = 1),
-                       expand = c(0,0)) +
-    scale_x_continuous(labels = scales::percent_format(accuracy = 1), expand = c(0,0), limits = c(0,1)) + 
-    theme(legend.position = "top")
-
-  # pop_depletion_and_size_plot <-  directlabels::direct.label(pop_depletion_and_size_plot, list("angled.boxes", box.alpha = 0.5))
-
-  pop_combo_plot <-
-    (pop_depletion_and_size_plot) + ((pop_size_plot) / pop_depletion_plot)  + plot_layout(widths = c(1.5, 1)) & theme(
-    plot.margin = unit(c(0, 1, 0, 1), units = "lines"),
-    axis.text.x = element_text(size = 8),
-    legend.box.margin = unit(c(0, 0, 0, 0), units = "lines"),
-    axis.text.y = element_text(size = 10)
-    )
-
-  ggsave(pop_combo_plot,filename = "figs/conservation-effect.png",
-         width = 8,
-         height = 5)
-
-  pdf(
-    file =  "figs/conservation-effect.pdf",
-    width = 8,
-    height = 5,
-    useDingbats = FALSE
-  )
-  print(pop_combo_plot)
-  dev.off()
-
-```
+Numerous studies provide evidence that well-enforced and appropriately
+sized MPAs can produce conservation benefits within their borders
+(5--8). As these conservation benefits accrue inside MPAs, theory holds
+that MPAs can affect the waters beyond their borders through the
+spillover of adult and larval fish from the protected to the fished
+areas, as well as through displacement of fishing effort. Several
+studies have documented empirical evidence for the existence of adult or
+larval spillover affecting both abundance and fisheries (9--17), as well
+as alteration of fishing effort in reaction to (18--20) and in
+anticipation of (21) MPA placement. The potentially more important
+question, however, is not whether spillover occurs (it must to some
+degree in any realistic scenario), but what the net effects of spillover
+are and whether those effects are empirically detectable. From a fishery
+perspective, are spillover benefits sufficient to offset losses in
+fishing grounds and changes in responses of displaced fishers caused by
+an MPA? From a conservation perspective, how much does the buildup of
+fish inside an MPA increase biomass outside the protected area? Overall,
+what are the regional effects of MPAs?
+
+As stakeholders around the world increasingly seek to use MPAs in marine
+resource management portfolios, and the performance of existing MPAs is
+evaluated, it is critical that we develop a better understanding of the
+magnitude and drivers of regional-scale MPA effects. To address this
+gap, this study examines two critical questions: 1) What do we expect
+the regional-scale conservation effects of MPAs to be and 2) When (and
+how) can we expect to empirically detect these effects? We answer these
+questions using a simulation analysis to frame the theoretical regional
+conservation impacts of MPAs. We then present an empirical test of the
+evidence for regional-level conservation effects of MPAs resulting from
+a network of closures put in place in the Channel Islands, California,
+USA, in 2003.
+
+\hypertarget{what-are-the-regional-effects-of-mpas}{%
+\subsection*{What Are the Regional Effects of
+MPAs?}\label{what-are-the-regional-effects-of-mpas}}
+\addcontentsline{toc}{subsection}{What Are the Regional Effects of
+MPAs?}
+
+Much of the literature on the effects of Marine Protected Areas focuses
+on assessing the conservation effects within the borders of protected
+areas (7). While these within-MPA effects are vitally important for
+protecting rare species, biodiversity, critical habitats, and often
+tourism, they paint an incomplete picture of the overall population
+effects of MPAs. The organisms within the borders of protected areas are
+generally part of a broader population, the biological stock, connected
+through adult movement and larval dispersal. If the goal of
+conservationists or natural resource managers is to increase the total
+abundance or productivity of a resource, a broader question we should
+ask of MPAs is not just are there more fish inside their borders, but
+also how have the reserves affected abundances throughout the region in
+which they are located? This logic, that MPAs will have conservation
+benefits for most species beyond their borders, is implicit in all
+multilateral calls for MPA expansion.
+
+We define the regional conservation effects of MPAs as the change in
+total biomass of fish (summing inside and outside of MPAs) relative to
+the total biomass of fish that would have occurred without the MPAs
+(acknowledging that other outcomes such as increased biodiversity or
+resiliency are also important to conservation but are beyond the scope
+of this analysis). Note that this definition of regional effects is in
+line with the recommendations of (22), but differs from effects measured
+by for example differences in outcomes before and after MPAs.
+
+We used a simulation model to examine the regional conservation effects
+of MPAs under a wide array of conditions that might reasonably be
+expected to affect these outcomes. The model explores combinations of
+MPA designs (e.g., sizes, number, and placement of MPAs), life histories
+(e.g., growth and mortality rates, adult movement, larval dispersal,
+recruitment variability and autocorrelation, and timing of density
+dependence), and fishing fleet dynamics (e.g., degree of fishing
+pressure, size selectivity of the fleet, reaction of fishing pressure to
+MPA placement) (See Table.S1 for a complete description of simulation
+variables). We use this model to explore how different drivers interact
+to affect the conservation and fishery outcomes of MPAs.
+
+\hypertarget{conservation}{%
+\subsubsection*{Theoretical Conservation Effects}\label{conservation}}
+\addcontentsline{toc}{subsubsection}{Theoretical Conservation Effects}
+
+As a thought experiment, imagine a region that has driven its fish
+populations to near extinction and then places 100\% of its waters
+inside a no-take MPA. In that setting, we would expect the
+regional-scale conservation benefits to be massive. On the other hand,
+imagine implementing an MPA in an area where a sedentary species has
+been only lightly fished. In that setting, it is possible to create a
+small net conservation loss if the concentration of fishing pressure
+outside the reserve has a greater effect than the biomass buildup inside
+the reserve. Within these broad bounds, numerous factors can act to
+affect the regional effects of MPAs. These include the scale of adult
+and larval dispersal relative to the size of the MPAs ({\textbf{???}},
+23--25), the strength and timing of density dependence in the population
+(e.g.~pre- or post-settlement), how overfished the population would be
+without the MPA, and how fishing activity and management responds to the
+implementation of the MPAs (4, 8, 26--32). In addition, even for the
+same total area of MPAs, the location and spacing of the MPAs can have a
+profound influence on their cumulative impact through habitat and
+network effects (4, 33).
+
+We tested the effects of different combinations of these theoretical
+drivers of MPA conservation effects using over 10,000 simulated MPA
+scenarios. A striking outcome of this simulation analysis is how the
+level of fishing pressure can cause dramatic and predictable shifts in
+conservation outcomes. We measure fishing pressure by the degree of
+simulated depletion (percent of unfished biomass gone, such that a
+depletion of 0\% means the population is at 100\% of unfished biomass,
+100\% means 0\% of unfished biomass remains) in the scenario without
+MPAs. When MPAs were placed in relatively unexploited fisheries
+(depletions less than 40\%), the median regional conservation effect was
+barely noticeable (4\%). For ecosystems fished near MSY levels
+(depletion of 40\%-60\%), the median conservation effect was 28\%, while
+for overfished ecosystems (without-MPA depletions above 60\%) the median
+regional conservation effect was over 100\%
+(Fig.\ref{conservation-effects}-A). Across all of these groups though,
+effects as low as declines of -50\% and as high as an increase of more
+than 100\% were possible. It should be noted though that these runs
+represent the range of scenarios evaluated in our simulation, and that
+certain types of scenarios may be much more likely than others in the
+real world. To put these results into context, the FAO estimates that
+7\% of the worlds fisheries with status estimates fall into the
+relatively unexploited category, 60\% fall into the fully fished
+category, and 33\% fall into the heavily fished category (34), though
+works that include a broader range of fisheries estimate that 50\% or
+more of stocks to fall into the heavily fished category (35, 36).
+
+These simple results support the intuitive conclusion that bigger MPAs
+on overfished populations produce large positive conservation outcomes.
+However, when combined with other factors such as fleet dynamics, life
+history, and MPA size, the conservation effects of MPAs can vary widely
+in magnitude and timing, a finding corroborated by empirical evidence
+(37--40) (Fig.\ref{conservation-effects}-B). MPAs protecting small
+proportions of a stock can produce small positive effects, large
+positive effects, or even small negative effects. As the area protected
+inside MPAs increases, positive effects become more likely, but even
+MPAs protecting most of a population's footprint can produce
+conservation effects as low as 0\% and as high as 200\% and above
+depending on the context in which they are placed. The degree of
+depletion (i.e.~fishing pressure) has a clearer signal, with small
+conservation benefits of MPAs when depletion is low, and larger effects
+when depletion is high, but again even for severely depleted populations
+the conservation effects of MPAs can vary widely (although nearly always
+with population benefits).
+
+While positive conservation outcomes were much more likely to occur
+across our simulations (94\% of simulations), the 6\% of runs that
+produced net conservation losses are of particular interest. These runs
+occurred almost exclusively when the simulated fleet followed a
+``constant-catch'' fishing strategy. Under the constant-catch fleet
+model, fishing communities (or fishery policies) seek to catch the same
+amount regardless of the presence of an MPA. While a constant-catch
+greater than MSY is not possible over the long-term under the
+assumptions of our model, over the short-term a constant-catch scenario
+is not implausible. Subsistence fisheries may use a constant-catch style
+policy over the short-term, as they seek to ensure that their food needs
+are met. More industrial fisheries may have pre-arranged agreements with
+buyers to deliver set amounts of fish. Constant-catch dynamics might
+also occur in fisheries with constraining quotas that are not updated
+after the implementation of MPAs. These patterns suggest that coupling
+MPA implementation with appropriate management reforms that reduce the
+effects of constant-catch strategies can move the predicted conservation
+effects of MPAs to being nearly uniformly positive. The possibility of
+constant-catch strategies to result in net MPA conservation losses
+highlights the critical importance of understanding the broader
+management context in which MPAs are placed.
 
@@
 
-## How Can We Detect Regional Effects?
-
-How can we empirically measure the regional effects of MPAs?
-Intuition and bio-economic theory tells us that the regional conservation effects of MPAs are likely to be highly context dependent across locations and across species within locations, so careful thought must be devoted to an empirical design to detect their effects.
-The perfect experiment would involve two parallel worlds that were identical, except for the implementation of an MPA.
-In world “A”, no MPA would be implemented, and in the facsimile world “B”, the MPA would be implemented.
-Both worlds would be tracked before and after MPA implementation, and the conservation and fishery outcomes would be compared after treatment.
-Instead of two parallel worlds, a similar experimental design would involve random placement of MPAs.
-Unfortunately, neither of these experimental designs has been implemented, because parallel worlds do not exist and MPAs are never placed at random.
-Careful causal inference requires some other means of controlling for biases introduced by factors such as the MPA siting process and biological spillover and concentration of fishing effort  [@ferraro2018;@larsen2019].
-
-<!-- From the fisheries perspective, catch rates before and after [e.g.
-@roberts2001a] or near and far from MPAs [e.g.
-@russ2004a] are sometimes presented as evidence for the fisheries effects of MPAs.
-However neither of these approaches can be counted on to provide reliable estimate of the net fishery effects of MPAs, due to failures to account for potential broader trends in catch rates independent of MPAs and the potentially confounding effects of effort redistribution.
-@smith2006a presents an econometrically robust method for estimating the fishery effects of MPAs, though this approach requires high-resolution data on comparable fishing effort both near and far from MPAs, data that is very rarely available.
-It is beyond the scope of this paper to delve deeper into detecting the net fishery effects of MPAs, and so we now focus on the question of estimating their regional conservation effects.
--->
-
-Conservation effects are typically estimated by comparing organism densities or biomass inside MPAs to densities or biomass in selected control sites outside MPAs, which we will refer to as response ratios.
-@lester2009 and @halpern2002 present meta-analyses of hundreds of such studies.
-These results often find massively higher densities and biomasses inside MPAs than outside [median biomass increases of  over 400%, @lester2009].
-While response ratios are intended as a measure of the effectiveness of reserves within their borders, these result stand in contrast to the smaller region wide effect sizes predicted by our simulation model.
-This raises the question: How reliable are density or biomass ratios as estimators of regional conservation effects?
-
-Control sites used in density or biomass ratios are often selected based on abiotic or ecological traits such as habitat characteristics [@ferraro2018].
-However, selection of control sites is further complicated by the very spillover that MPAs are often intended to create.
-Export of adults or larvae from the MPA to the "control" site affects their status as controls, as does displacement of fishing effort from MPAs to control sites.
-In theory, control sites far enough away to negate both biological spillover and concentration by the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs may be challenging.
-While these concerns have been stated previously [e.g.
-@halpern2004], the MPA evaluation literature has by and large been unable to adequately address them, often due to the very real challenges of acquiring the necessary data [@ferraro2018].
-
-Failure to account for these spillover effects can result in a biased estimate of the true effect of a policy such as MPA placement [@larsen2019].
-In an ideal setting, the control site used in a response ratio would be a perfect counterfactual for what would have occurred without the MPA.
-We used our simulation model to approximate this scenario, calculating the response ratio as the ratio of densities inside the MPAs relative to the overall density from the paired simulation without MPAs multiplied by the area of the population in an MPA (to account for the fact that we would expect the effectiveness of a response ratio as a proxy for regional conservation effects to scale with the size of the MPA relative to the range of the species, Fig.\ref{density-ratio}-A).
-In this idealized case, the response ratio is often a reasonable estimator for the regional conservation effect.
-
-<!-- However, even under this ideal example the response ratio is not a perfect estimator due to the fact that this ratio contains no information on the effect of the MPAs outside their borders.  -->
-
-What happens when the “control” sites are in fact affected by the application of the MPA through biological spillover and fishing concentration?
-In order to address this possibility, we calculated a new response ratio as the mean density inside MPAs relative to the mean density outside the MPAs (both weighted by distance from MPA borders and scaled by MPA size).
-We only include simulations in which habitat and larval dispersal rates are identical inside and outside of MPAs, to approximate a scenario in which treatment and control sites have been paired by ecological characteristics.
-Under these circumstances the response ratio is a somewhat biased and very inaccurate estimator of the true regional effect of MPAs.
-While on average high response ratios correspond with high true regional effects, the variance is high, so the response ratio alone cannot be used as a reliable proxy for the true regional effect, without accounting for the sources of biases discussed here.
-For higher movement rates, response ratios were frequently near 0\% (potentially leading stakeholders to conclude that the MPA had been ineffective), when in fact in many cases the true effect of the MPA was highly positive.
-For highly sedentary species, extremely high response ratios could create the appearance of massive conservation gains when in fact the net effect on the regional population has been near zero or even negative (Fig.\ref{density-ratio}-B).
-
-We note that spatial before-after-control-impact (BACI) studies present a potential improvement over response ratios, controlling for differences in mean densities in the treated and control regions as well as MPA independent trends, but are much rarer due to the need for extensive pre-MPA monitoring.
-However, spatial BACI still requires that the MPAs do not affect the control sites, and as such has many of the same limitations as response ratios as an estimator of regional conservation effects.
-
-To the extent that we can select control sites that are both sufficiently similar to the region treated with MPAs and are unaffected by biological or economic spillover, response ratios may be reliable estimators of regional conservation effects.
-In reality these conditions are unlikely to hold but for a few select cases.
-This become especially difficult when networks of MPAs are implemented, such as has occurred in California.
-In this case, control sites are generally located within a few kilometers from MPA borders, making biological and economic spillover highly likely.
-In these circumstances, without robust statistical controls, response ratios may provide an unreliable estimate of the regional conservation effect of marine protected areas.
-
-```{r dr-vs-mpa-effect-plot, include = FALSE}
-
-dr_plot <- density_ratios %>%
-  filter(sprinkler == FALSE, mpa_habfactor == 1) %>%
-  select(experiment,
-         years_protected,
-         mpa_effect,
-         adult_movement,
-         mpa_size,
-         contains("density_ratio")) %>%
-  mutate(adult_movement = adult_movement / max(adult_movement)) %>%
-  gather(scenario, density_ratio, contains("density_ratio")) %>%
-  group_by(experiment, scenario) %>%
-  filter(years_protected == max(years_protected)) %>%
-  mutate(density_ratio = density_ratio * mpa_size) %>% 
-  ggplot(aes(y = pmin(4, mpa_effect), x = pmin(4, density_ratio - 1))) +
-    geom_hline(aes(yintercept = 0)) + 
-  geom_vline(aes(xintercept = 0)) + 
-  geom_abline(aes(slope = 1, intercept = 0), linetype = 2, size = 1.5, color = "black") +
-  geom_point(aes(color = adult_movement, size = mpa_size), alpha = 0.5, show.legend = TRUE)  +
-    geom_smooth(method = "lm", se = FALSE) +
-  facet_wrap(~ rev(scenario), labeller = labeller(`rev(scenario)` = c(biased_density_ratio = "A) Inside vs. Experiment",true_density_ratio = "B) Inside vs. Outside"))) +
-  labs(y = "True % Increase in Biomass", x = "% Response Ratio") +
-  # scale_x_percent(labels = c(percent(seq(-1,1, by = 1)),">200%")) +
-  # scale_y_percent(labels = c(percent(seq(-1,1, by = 1)),">200%")) +
-  scale_color_viridis(guide = wide_hgc, labels = scales::percent_format(accuracy = 1), name = "Fish Movement",
-                      option = "cividis") + 
-  scale_x_continuous(labels = percent) + 
-    scale_y_continuous(labels = percent) + 
-  theme(legend.position = "top") + 
-  guides(size = FALSE)
-
-pdf(file =  "figs/density-ratio.pdf",width = 8, height = 5, useDingbats = TRUE)
-print(dr_plot)
-dev.off()
-
-```
+\hypertarget{how-can-we-detect-regional-effects}{%
+\subsection{How Can We Detect Regional
+Effects?}\label{how-can-we-detect-regional-effects}}
+
+How can we empirically measure the regional effects of MPAs? Intuition
+and bio-economic theory tells us that the regional conservation effects
+of MPAs are likely to be highly context dependent across locations and
+across species within locations, so careful thought must be devoted to
+an empirical design to detect their effects. The perfect experiment
+would involve two parallel worlds that were identical, except for the
+implementation of an MPA. In world ``A'', no MPA would be implemented,
+and in the facsimile world ``B'', the MPA would be implemented. Both
+worlds would be tracked before and after MPA implementation, and the
+conservation and fishery outcomes would be compared after treatment.
+Instead of two parallel worlds, a similar experimental design would
+involve random placement of MPAs. Unfortunately, neither of these
+experimental designs has been implemented, because parallel worlds do
+not exist and MPAs are never placed at random. Careful causal inference
+requires some other means of controlling for biases introduced by
+factors such as the MPA siting process and biological spillover and
+concentration of fishing effort (22, 41).
+
+Conservation effects are typically estimated by comparing organism
+densities or biomass inside MPAs to densities or biomass in selected
+control sites outside MPAs, which we will refer to as response ratios.
+(5) and (42) present meta-analyses of hundreds of such studies. These
+results often find massively higher densities and biomasses inside MPAs
+than outside (5). While response ratios are intended as a measure of the
+effectiveness of reserves within their borders, these result stand in
+contrast to the smaller region wide effect sizes predicted by our
+simulation model. This raises the question: How reliable are density or
+biomass ratios as estimators of regional conservation effects?
+
+Control sites used in density or biomass ratios are often selected based
+on abiotic or ecological traits such as habitat characteristics (22).
+However, selection of control sites is further complicated by the very
+spillover that MPAs are often intended to create. Export of adults or
+larvae from the MPA to the ``control'' site affects their status as
+controls, as does displacement of fishing effort from MPAs to control
+sites. In theory, control sites far enough away to negate both
+biological spillover and concentration by the fishing fleet could be
+selected, but finding suitably far sites that are also appropriate
+proxies for the ecological and economic context of the MPAs may be
+challenging. While these concerns have been stated previously (43), the
+MPA evaluation literature has by and large been unable to adequately
+address them, often due to the very real challenges of acquiring the
+necessary data (22).
+
+Failure to account for these spillover effects can result in a biased
+estimate of the true effect of a policy such as MPA placement (41). In
+an ideal setting, the control site used in a response ratio would be a
+perfect counterfactual for what would have occurred without the MPA. We
+used our simulation model to approximate this scenario, calculating the
+response ratio as the ratio of densities inside the MPAs relative to the
+overall density from the paired simulation without MPAs multiplied by
+the area of the population in an MPA (to account for the fact that we
+would expect the effectiveness of a response ratio as a proxy for
+regional conservation effects to scale with the size of the MPA relative
+to the range of the species, Fig.\ref{density-ratio}-A). In this
+idealized case, the response ratio is often a reasonable estimator for
+the regional conservation effect.
+
+What happens when the ``control'' sites are in fact affected by the
+application of the MPA through biological spillover and fishing
+concentration? In order to address this possibility, we calculated a new
+response ratio as the mean density inside MPAs relative to the mean
+density outside the MPAs (both weighted by distance from MPA borders and
+scaled by MPA size). We only include simulations in which habitat and
+larval dispersal rates are identical inside and outside of MPAs, to
+approximate a scenario in which treatment and control sites have been
+paired by ecological characteristics. Under these circumstances the
+response ratio is a somewhat biased and very inaccurate estimator of the
+true regional effect of MPAs. While on average high response ratios
+correspond with high true regional effects, the variance is high, so the
+response ratio alone cannot be used as a reliable proxy for the true
+regional effect, without accounting for the sources of biases discussed
+here. For higher movement rates, response ratios were frequently near
+0\% (potentially leading stakeholders to conclude that the MPA had been
+ineffective), when in fact in many cases the true effect of the MPA was
+highly positive. For highly sedentary species, extremely high response
+ratios could create the appearance of massive conservation gains when in
+fact the net effect on the regional population has been near zero or
+even negative (Fig.\ref{density-ratio}-B).
+
+We note that spatial before-after-control-impact (BACI) studies present
+a potential improvement over response ratios, controlling for
+differences in mean densities in the treated and control regions as well
+as MPA independent trends, but are much rarer due to the need for
+extensive pre-MPA monitoring. However, spatial BACI still requires that
+the MPAs do not affect the control sites, and as such has many of the
+same limitations as response ratios as an estimator of regional
+conservation effects.
+
+To the extent that we can select control sites that are both
+sufficiently similar to the region treated with MPAs and are unaffected
+by biological or economic spillover, response ratios may be reliable
+estimators of regional conservation effects. In reality these conditions
+are unlikely to hold but for a few select cases. This become especially
+difficult when networks of MPAs are implemented, such as has occurred in
+California. In this case, control sites are generally located within a
+few kilometers from MPA borders, making biological and economic
+spillover highly likely. In these circumstances, without robust
+statistical controls, response ratios may provide an unreliable estimate
+of the regional conservation effect of marine protected areas.
 
@@
 
-```{r}
-fish_v_fishing <- processed_grid %>%
-  select(-density_ratio) %>%
-  unnest() %>%
-  rename(fishery_effect = mpa_effect1) %>%
-  group_by(experiment) %>%
-  mutate(year = 1:length(year)) %>%
-  ungroup() %>%
-  mutate(years_protected = year - year_mpa + 1) %>%
-  left_join(outcomes %>% select(experiment, year, depletion),
-            by = c("experiment", "year"))
-
-fish_v_fishing_plot <- fish_v_fishing %>%
-  filter(years_protected == max(years_protected)) %>%
-  ggplot(aes(pmin(2, mpa_effect), pmin(2, fishery_effect))) +
-  geom_abline(aes(slope = 1, intercept = 0)) +
-  geom_smooth(method = "lm") +
-  geom_hex()
-
-```
-
-<!-- Our strategy for identifying the effect of the MPAs in the Channel Islands is to use biomass densities of non-targeted species as our control for broader factors affecting biomass densities of fishes throughout the Islands besides the effect of the MPAs.
-Simply examining densities of tarted species throughout the islands before and after MPAs might confuse changes in regional densities that would have occurred anyway without the MPAs with MPA effects (since at the regional scale we do not have a spatial control for the effects of the MPAs).
-We assume that over the 13 years of MPA protection covered in this study non-targeted species were relatively unaffected by the MPAs (i.e.
-indirect effects of MPAs on species interactions are negligible), and can can serve as an indicator for the overall changes in fish abundance throughout the Channel Islands that would have occurred without the MPAs.
-Raw densities calculated by PISCO provide visual support for this hypothesis, in that both the targeted and non-targeted species groups share common trends (Fig.\ref{pop-trends}, see Materials and Methods for statistical support for this assumption).
--->
-
-## MPA Effects in the Channel Islands {#estimating .unnumbered}
-
-```{r ci-map, fig.cap = "Map of study region and sampling locations.
-Shaded polygons indicate location of MPAs.
-Points represent sampling locations, and color indicates the number of observations recorded at a given point", include = FALSE}
-
-ci_map <- pisco_data %>%
-  left_join(site_data, by = c("site","side")) %>%
-  filter(is.na(eventual_mpa) == F) %>%
-  filter(region %in% c("ANA", "SCI","SRI",'SMI'),
-         classcode %in% top_species) %>%
-  group_by(lat_wgs84, lon_wgs84, site, region, eventual_mpa) %>%
-  count()
-
-ci_map <-  ci_map %>%
-  dplyr::mutate(geometry = purrr::map2(lon_wgs84, lat_wgs84, ~ sf::st_point(x = c(.x, .y), dim = 'XY'))) %>%
-  ungroup() %>%
-  mutate(geometry = sf::st_sfc(geometry, crs = 4326)) %>%
-  sf::st_sf()
-
-bbox <- sf::st_bbox(ci_map)
-
-pisco_ci_map_plot <-  ggmap(channel_islands, alpha = 0.5) +
-  geom_sf(data = ca_mpas, inherit.aes = FALSE, alpha = 0.5) +
-  geom_sf(data = ci_map,inherit.aes = FALSE,
-  aes(fill = eventual_mpa, color = eventual_mpa),
-  shape = 21, 
-  alpha = 0.5,
-  show.legend = FALSE,
-  size = 1.5) +
-  coord_sf(xlim = c(bbox['xmin'] - .3, bbox['xmax'] + .06),
-  ylim = c(bbox['ymin'] - .1, bbox['ymax'] + .1)) +
-  # scale_color_gradient(low = "white", high = "orangered",
-  # guide = guide_colorbar(frame.colour = "black",frame.linewidth = 1,
-  # barheight = unit(13, "lines")),
-  # name = "Observations")  +
-  labs(x = "", y = "") +
-      ggspatial::annotation_scale(location = "br", width_hint = 0.4, text_col = "white") + 
-  ggsci::scale_color_npg() + 
-    ggsci::scale_color_npg()
-
-# pisco_ci_map_plot <-  ggplot() +
-#   geom_sf(data = tests2, fill = c("#FFD39B")) +
-#   geom_sf(data = ca_mpas, inherit.aes = FALSE, alpha = 0.5) +
-#   geom_sf(data = ci_map,inherit.aes = FALSE,
-#   aes(color = eventual_mpa),
-#   alpha = 0.75,
-#   show.legend = FALSE,
-#   size = 2) +
-#   coord_sf(xlim = c(bbox['xmin'] - .3, bbox['xmax'] + .06),
-#   ylim = c(bbox['ymin'] - .1, bbox['ymax'] + .1)) +
-#   # scale_color_gradient(low = "white", high = "orangered",
-#   # guide = guide_colorbar(frame.colour = "black",frame.linewidth = 1,
-#   # barheight = unit(13, "lines")),
-#   # name = "Observations")  +
-#   labs(x = "", y = "") +
-#       ggspatial::annotation_scale(location = "br", width_hint = 0.4, text_col = "white") + 
-#   theme(panel.background = element_rect(fill = "lightblue"))
-# 
-pdf(file =  "figs/channel-islands.pdf",width = 8, height = 5, useDingbats = TRUE)
-print(pisco_ci_map_plot)
-dev.off()
-
-```
+\hypertarget{estimating}{%
+\subsection*{MPA Effects in the Channel Islands}\label{estimating}}
+\addcontentsline{toc}{subsection}{MPA Effects in the Channel Islands}
 
@@
 
-```{r}
-ci_effects <- short_term %>%
-  filter(years_protected == max(years_protected),
-         mpa_size  <= 0.25, mpa_size >= 0.15)
-
-med_ci_effect <- median(ci_effects$mpa_effect)
-
-# rethinking::HPDI(ci_effects$mpa_effect, prob = 0.5)
-
-quart_ci_effect <- quantile(ci_effects$mpa_effect)
-
-```
-
-Given these challenges, what is a viable alternative for estimating the conservation effects of MPAs?
-In a manner similar to @caselle2015, we use a difference-in-difference estimator comparing densities in finfish species targeted by fishing effort (i.e., those potentially affected by an MPA) from those not targeted before and after MPA implementation [see @larsen2019 for a succinct explanation of difference-in-difference estimators in ecological settings].
-Targeted species in the Northern Channel Islands, California, include commercially important finfish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
-Each of these targeted species was the subject of prior bio-economic modeling related to the effects of MPAs in southern California [@costello2010d;@rassweiler2014].
-But, we must note that the analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*).
-Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species).
-The key assumptions of this approach are that a) within the time-frame of the model there are no significant interaction effects between the targeted and non-targeted species (which in fact we do not detect, see SI section 1.6.3), and b) that in the absence of the MPAs both the targeted and non-targeted groups of species would have exhibited similar trends in densities (the parallel-trends assumption).
-The advantage to this approach is that given the time-frame of the model (15 years), we believe that spillover effects of MPA placement on the non-targeted species are likely to be much less severe than the effects of biological spillover and fishing concentration that may bias the performance of estimators such as a response ratios or a spatial BACI design.
-
-We applied this difference-in-difference strategy to a large dataset in California.
-We used empirical kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands with the ultimate goal of testing the regional effects of MPAs in a real world context.
-A network of MPAs covering approximately 20% of the islands’ waters was put in place in 2003 as part of the California Marine Life Protection Act (MLPA) (see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of the MLPA).
-PISCO conducts visual SCUBA surveys at a large number of rocky-reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\ref{channel-islands}-\ref{pop-trends}).
-
-Although the Islands do not encompass the full population of any of these species, they encompass a large enough area to assess the regional MPA conservation effects we might see in the Channel Islands after 13 years of protection (the time span covered by our model).
-The median regional conservation effect produced by our simulation model after 13 years of protection by MPAs covering between 15%-20% of a population's range was `r percent(med_ci_effect, accuracy = 1)`.
-This projected effect has an an interquartile range of `r percent(quart_ci_effect[2], accuracy = 1)` and `r percent(quart_ci_effect[4], accuracy = 1)`, although values as high as a 200% increase and as low as a 50% decrease were also produced.
-Our expectation then is that all else being equal we are likely looking for a modest but positive effect size.
-
-```{r raw-trend, fig.cap="Centered and scaled mean annual biomass density of included species (faded lines) and smoothed means of targeted and non-targeted groups, and mean (darker lines) and 95% confidence interval of the mean (ribbon) over time", include = FALSE}
-raw_did_plot <- base_run$data[[1]] %>%
-  select(year, targeted,classcode, log_density, any_seen) %>%
-  mutate(density = exp(log_density) * any_seen) %>%
-  group_by(classcode,targeted, year) %>%
-  summarise(md = mean(density)) %>%
-  group_by(classcode,targeted) %>%
-  mutate(smd = (md - mean(md)) / sd(md)) %>%
-  ggplot(aes(year, smd, color = targeted == 1, fill = targeted == 1)) +
-  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red", size = 1.5) +
-  geom_line(aes(group = interaction(targeted, classcode)), alpha = 0.5) +
-  geom_smooth(size = 2) +
-  labs(y = "Centered and Scaled Mean Density", x = "Year") +
-    scale_color_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
-      scale_fill_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
-    labs(y = "Centered and Scaled Mean Density", x = "Year") + 
-  theme(legend.position = "top",
-        panel.grid.major.x = element_blank(),
-        panel.grid.minor.x = element_blank()
-        )
-
-raw_did_plot <- raw_did_plot +
-      scale_color_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "") +
-        scale_fill_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "")
-
-pdf(file =  "figs/pop-trends.pdf",width = 8, height = 5, useDingbats = TRUE)
-print(raw_did_plot)
-dev.off()
-
-```
+Given these challenges, what is a viable alternative for estimating the
+conservation effects of MPAs? In a manner similar to (38), we use a
+difference-in-difference estimator comparing densities in finfish
+species targeted by fishing effort (i.e., those potentially affected by
+an MPA) from those not targeted before and after MPA implementation
+(41). Targeted species in the Northern Channel Islands, California,
+include commercially important finfish such as California sheephead
+(\emph{Semicossyphus pulcher}), and copper (\emph{Sebastes caurinus})
+and blue (\emph{Sebastes mystinus}) rockfish. Each of these targeted
+species was the subject of prior bio-economic modeling related to the
+effects of MPAs in southern California (44, 45). But, we must note that
+the analysis omits species from important invertebrate fisheries
+including red urchin (\emph{Mesocentrotus franciscanus}) and spiny
+lobster (\emph{Panulirus interruptus}). Non-targeted species include
+garibaldi (\emph{Hypsypops rubicundus}), halfmoons (\emph{Medialuna
+californiensis}), and blacksmith (\emph{Chromis punctipinnis}) (See
+Table.S2 for a complete list of species). The key assumptions of this
+approach are that a) within the time-frame of the model there are no
+significant interaction effects between the targeted and non-targeted
+species (which in fact we do not detect, see SI section 1.6.3), and b)
+that in the absence of the MPAs both the targeted and non-targeted
+groups of species would have exhibited similar trends in densities (the
+parallel-trends assumption). The advantage to this approach is that
+given the time-frame of the model (15 years), we believe that spillover
+effects of MPA placement on the non-targeted species are likely to be
+much less severe than the effects of biological spillover and fishing
+concentration that may bias the performance of estimators such as a
+response ratios or a spatial BACI design.
+
+We applied this difference-in-difference strategy to a large dataset in
+California. We used empirical kelp forest survey data from the
+Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO)
+monitoring in the Northern Channel Islands with the ultimate goal of
+testing the regional effects of MPAs in a real world context. A network
+of MPAs covering approximately 20\% of the islands' waters was put in
+place in 2003 as part of the California Marine Life Protection Act
+(MLPA) (see (46), (47), (48), and (49) for information on the creation
+of the MLPA). PISCO conducts visual SCUBA surveys at a large number of
+rocky-reef and kelp forest sites inside and outside of MPAs throughout
+the Channel Islands, producing estimates of densities of fishes that are
+both targeted and non-targeted by fishing
+(Fig.\ref{channel-islands}-\ref{pop-trends}).
+
+Although the Islands do not encompass the full population of any of
+these species, they encompass a large enough area to assess the regional
+MPA conservation effects we might see in the Channel Islands after 13
+years of protection (the time span covered by our model). The median
+regional conservation effect produced by our simulation model after 13
+years of protection by MPAs covering between 15\%-20\% of a population's
+range was 25\%. This projected effect has an an interquartile range of
+8\% and 100\%, although values as high as a 200\% increase and as low as
+a 50\% decrease were also produced. Our expectation then is that all
+else being equal we are likely looking for a modest but positive effect
+size.
 
@@
 
-The raw trends in the biomass densities appear to show parallel trends in the densities of targeted and non-targeted fishes pre-MPA, with the targeted group diverging first positively and then negatively over time (Fig.\ref{pop-trends}).
-However, these raw trends could be caused by any number of confounding variables besides the effects of the MPAs themselves, from changes in observer skill and sampling locations, to shifts in environmental conditions.
-Using our difference-in-difference model in an effort to control for these factors, we found no significant difference in the biomass densities of targeted and non-targeted fishes pre-MPA (before 2003), which provides support for the parallel trends assumption critical to the difference-in-difference method.
-Following the implementation of the MPAs in 2003, we see evidence of an increasing trend in densities of targeted relative to non-targeted fish, reaching a peak value in 2014.
-Following 2014, densities of targeted species appear to have declined relative to what we would expect based on the non-targeted group (Fig.\ref{did-plot}).
-
-```{r did-plot, include = FALSE}
-
-did_plot <- did_plot +
-  scale_y_continuous(name = "~Divergence from Non-Targeted", labels = scales::percent_format(accuracy = 1))
-
-ggsave(did_plot, filename = "figs/did_plot.png",
-       width = 8, height = 5)
-
-# pdf(file =  "figs/did_plot.pdf",width = 8, height = 5, useDingbats = TRUE)
-# print(did_plot)
-# dev.off()
-
-```
+The raw trends in the biomass densities appear to show parallel trends
+in the densities of targeted and non-targeted fishes pre-MPA, with the
+targeted group diverging first positively and then negatively over time
+(Fig.\ref{pop-trends}). However, these raw trends could be caused by any
+number of confounding variables besides the effects of the MPAs
+themselves, from changes in observer skill and sampling locations, to
+shifts in environmental conditions. Using our difference-in-difference
+model in an effort to control for these factors, we found no significant
+difference in the biomass densities of targeted and non-targeted fishes
+pre-MPA (before 2003), which provides support for the parallel trends
+assumption critical to the difference-in-difference method. Following
+the implementation of the MPAs in 2003, we see evidence of an increasing
+trend in densities of targeted relative to non-targeted fish, reaching a
+peak value in 2014. Following 2014, densities of targeted species appear
+to have declined relative to what we would expect based on the
+non-targeted group (Fig.\ref{did-plot}).
 
@@
 
-Discussion {#discussion .unnumbered}
-------
-
-<!-- Studies such as @dinerstein2019 call for 30% of the oceans to be places in protected areas.
-What does simulation exercise suggest that the effects of such a closure might be on overall population biomass densities?
-XX.
-Suppose we assumed that all fisheries are both open-access and overfished (a clearly worst case scenario as many fisheries are well managed).
-In that case, we see that...
--->
-
-<!-- <!-- zeros may be an artifact of "error in variables", which can bias estimates towards zero [@larsen2019] -->
+\hypertarget{discussion}{%
+\subsection*{Discussion}\label{discussion}}
+\addcontentsline{toc}{subsection}{Discussion}
 
 MPAs are an important part of the marine resource management toolbox.
-Under ideal circumstances they can protect individual species and ecosystem linkages, while supporting local economies through tourism and fishing opportunities.
-One rationale for the expansion of MPAs is that they will deliver net conservation benefits both inside *and outside* their borders.
-To date, this assumption is insufficiently tested, and this is the focus of our paper.
-Our results show that regional conservation benefits of MPAs are highly context dependent and in many circumstances, are likely to be so small that they are nearly impossible to detect empirically.
-Indeed, this is exactly what we found in our empirical case study from the Channel Islands, California, USA.
-
-<!-- Our simulation results present a clear outcome from the perspective of fishery effects: Under the assumptions of our model, MPAs should not be expected to provide improvements in fisheries catches unless the stock is experiencing substantial overfishing (Fig.\ref{catch-msy-effects}-B).
-This general result has been discussed before [@hilborn2004], but our results provide the most comprehensive simulation testing of this concept to date.
-MPAs placed in overfished fisheries frequently produced catch gains of 0-25%, though larger and smaller effects were still possible.
-MPA size also presents an interesting tradeoff from a fisheries perspective.
-MPAs that cover less than 25% of the population range, which is true for nearly all MPAs currently on the planet, very rarely produced substantial fishery losses, but also never produced substantial fishery gains.
-Simulated MPAs  that covered between 25% & 50% of species ranges produced the potential for higher fishery gains, but also opened up risk of greater fishery loses.
-For MPAs covering more than 50% of a species range, fishery losses quickly become much more likely than fishery gains.
--->
-
-<!-- We only measure one aspect of fishery outcomes here (catches).
-MPAs can affect many other outcomes of importance to fisheries, include profits (e.g.
-if institutions do not create derby fishing and MPAs improve catch per unit effort along the borders of reserves), resiliency, and research.
-Our model does not account for potentially important factors such as species interactions Overall, our fishery-specific results suggest a relatively simple rule of thumb: unless there is clear evidence of overfishing, MPAs should not be expected to produce improvements in fishery catches without clear explanation for the mechanism for these increases (e.g.
-improvements in habitat).
-In assessing the potential fishery benefits of MPAs, stakeholders must make explicit the mechanisms by which these benefits are expected to be produced (for example through increased catchability of fish near MPA borders), and what evidence exists for the existence of those mechanisms within the local context.
--->
-
-We find regional conservation gains from MPAs in `r percent(pos_runs, accuracy = 1)` of our simulations (Fig.\ref{conservation-effects}).
-MPAs covering less than 25% of the population range were unlikely to produce regional conservation gains above 10% unless the stock would have been severely overfished without the MPA.
-Given that few marine species on the planet have more than 25% of their entire range protected in MPAs, this suggests that large regional gains are unlikely to be common unless the overall extent of MPAs grows globally.
-Large MPAs protecting highly depleted stocks almost always produced large regional conservation gains, as should be expected.
-The median conservation effect of `r  round(100*median(eqo$mpa_effect))`\% stands in contrast to the large within-MPA effects on biomass densities reported by @halpern2003a and @lester2009.
-This difference is due in part to the fact that within-MPA effects are likely to be much larger than regional-MPA effects, but may also reflect the many challenges of translating response ratios into regional conservation effects.
-Our simulation analysis shows that without proper controls (in particular for the effects of fleet redistribution), response ratios well over 100% are entirely plausible even when the actual conservation effect is much lower (and *vice versa*).
-
-<!-- While our results provide important insight on the expected fishery and conservation effects of MPAs, they also highlight the critical importance of explicitly modeling the links between the biological effects of MPAs and the ways that humans respond to them.
-One of our clearest findings was that MPAs can cause net conservation losses when implemented in a "constant-catch" fishery, as fishers must dramatically increase their fishing effort in the waters still available to them to maintain their desired landings.
-The focus of much of the MPA design and modeling literature has historically been on biological and ecological such as growth rates and connectivity [citation], though studies such as @hastings1999 and @sanchirico2006b incorporate assumptions about outside-MPA fleet dynamics as well.
-Our results make clear that economic and ecological drivers must be included in any attempt assess the effects of MPAs MPAs: in particular, we need to explicitly understand the the behavior of fishing fleets if we hope to estimate the conservation or fishery effects of MPAs [see @costa2013 for an example of this process].
-Large-scale empirical evidence confirms that predicting the effects of MPAs depends on understanding the human context in which they find themselves [@cinner2018].
--->
-
-<!-- The question of what is a "meaningful" conservation gain is of course context dependent and depends on the priorities and preferences of stakeholders.
-However, under any set of preferences and objectives managers and stakeholders would of course like to know whether policy interventions such as MPAs are having their intended effects.
-While the importance of say a 0-10% regional conservation gain will vary, there can be little doubt that an effect size within this range will be very difficult to detect in a noisy natural system (xx citations on noise and effect size xx), and our empirical assessment of MPA conservation effects in the Channel Islands demonstrates this.
--->
-
-<!-- .
-That we were unable to detect a significant divergence in the biomass densities of targeted species throughout the Channel Islands post MPA implementation, relative to the densities that the trends of the non-targeted species would suggest we should observe, should not come as a great surprise then.
--->
-
-We were not able to detect a statistically significant difference in the densities of targeted and non-targeted finfish species over the 13 years of MPA protection in the Channel Islands covered by our analysis.
-The data (and model assumptions) provide support for both negative and positive MPA effects, although the size of these estimated effects is well within the range that our simulation analysis suggests are plausible.
-At face value, this lack of a clear result may seem surprising given the size and carefully studied nature of this MPA network.
-
-However, our simulation analysis suggests that this result should perhaps be expected.
-The Channel Islands MPAs cover approximately 20% of the waters in the Channel Islands, and while formal stock assessments are not available for many of the targeted species in our analysis, what evidence we have suggests that, as a group, these fish are on average not heavily overfished.
-Some species, such as California sheephead and blue rockfish were likely below target levels during the period [@alonzo2004;@dick2017], but projections based upon the overall average response across all species will likely suggest modest benefits even if a subset of species could experience much larger population gains.
-Harkening back to our simulations, we expect the average percentage difference in densities of targeted species with and without MPAs to be modest (Fig.\ref{did-plot}).
-Effects of this size are likely to be challenging to detect empirically given the large natural variation of marine ecosystems (especially temperate reefs) and the observation error inherent in infrequent annual survey programs such as those provided by PISCO.
-
-```{r}
-tmdata <- life_history_data %>% 
-  filter(targeted == 1, classcode %in% top_species) %>% 
-  select(tm, everything())
-
-median_tm = round(median(tmdata$tm))
-```
-
-In addition, our simulations examine percentage increases in biomass densities.
-As an extreme example, an increase in biomass densities from 2kg/m^2^ to 4kg/m^2^ would be reported as a 100% increase.
-While this chance produces a large percentage increase, this is a small change in absolute biomass densities relative to the variance in the observation process itself (see Fig.S1 for a companion to Fig.\ref{conservation-effects} scaled by absolute population size).
-Beyond these challenges, the median estimated age at sexual maturity for the targeted species included in this study is `r median_tm` years, meaning that the span of this analysis represents less than three generations of MPA protection for half of the the measured species.
-Ongoing monitoring may yet reveal clearer effects.
-Analysis of more rapidly growing and maturing species, e.g.
-spiny lobster, may also reveal clearer signals.
-
-While the MPA effect estimated by our model was not significant in any one year, the positive trend from 2005 to 2014 is evident, as is the sharp decline from 2015-2017.
-To what should we attribute this apparent shift?
-We cannot reject the parallel trends assumption between the targeted and non-targeted species in the years before the MPAs (Fig.\ref{did-plot}).
-But, the Channel Islands region (and the entire West coast of the USA) experienced a dramatic ‘marine heatwave’ beginning in 2014 and persisting through 2016, resulting in part in extremely elevated water temperatures throughout the region  [@gentemann2017].
-Biogeographic differences in the distributions of targeted and non-targeted species may confound the observed effects of MPAs.
-Many of the non-targeted species have warm thermal affinities and have increased in numbers since the heatwave [@freedman2019].
-However, the targeted group is made up mostly of fishes with cold-water affinities, such as members of the genus Sebastes.
-As such, we hypothesize that the recent evidence for a decline in densities of targeted species is due to environmental conditions that disproportionately affect the targeted group (and not for example due to concentrated fishing pressure outside the reserves).
-This hypothesis is supported by the sharp declines in densities of targeted species seen inside the MPAs themselves (see Fig.S27-S29), suggesting a driver other than fishing may be at play.
-The Channel Islands case study demonstrates the challenge of detecting small effect sizes in nature and highlights the potential for small effect sizes to be overwhelmed by additional environmental shocks to the system.
-
-<!-- Our simulation model accounts for many factors that theory suggest should be critical drivers of MPA effects, namely adult and larval movement rates, nature of density dependence, and fleet dynamics.
-However, currently our model does not contain many other potentially important factors, such as species interactions.
-If stakeholders expect an effect that our model says is unlikely or impossible given local context, they should provide a clear hypothesis and evidence as to what factors not included in this model would cause their expected effect.
-For example, stakeholders might assert that larger conservation effects than predicted by our simulation model might occur in a location due to protection of critical spawning aggregation sites [@roberts2002;@erisman2015], but relevant evidence supporting this belief must be presented.
-It is also important to remember that our conservation results refer to the regional effects of MPAs both inside and outside of their borders.
-The buildup of biomass within the border of well enforced reserves is well documented [@lester2009], we show though that the regional effects may be much smaller and harder to detect than these within-MPA outcomes.
--->
-
-What do our results imply about the future of MPA science?
-Our simulation model is by no means exhaustive (for example it ignores features such as species interactions, habitat effects, and climate feedback), but captures many of the core factors theorized to affect MPA performance.
-Our results show that while attributes such as MPA size and fishing pressure are important factors in determining the effects of protection, local fleet dynamics and the movement rates of adult fish can dramatically affect the outcomes of protected areas as well  (Fig.\ref{conservation-effects}).
-Far from being a simple tool with clear outcomes, the effects of MPAs can be highly context dependent, requiring - we would argue - a bio-economic model of at minimum the complexity presented here to help communities design and set expectations of MPAs at the tactical level.
-While users may not be able to parameterize every aspect of a model such as that presented here, working with stakeholders to visualize the implications of, for example, different fleet responses to MPA implementation is a critical step in MPA design.
-Readers can use our simulation model to explore the effects of MPAs through an interactive web application available [here](https://danovando.shinyapps.io/simmpa/).
-
-Once simulation models have been used to help design an MPA, how can users evaluate whether it is achieving their objectives?
-Response ratios are commonly used as evidence for conservation outcomes of MPAs [@lester2009], but as suggested in @ferraro2018 and @larsen2019 and demonstrated here, without careful attention to the design of control sites (accounting for example for the displacement of effort by MPAs), response ratios may be highly unreliable estimators of regional MPA effects.
-As @larsen2019 suggests, there are many potential alternatives for estimating the effects of MPAs that better account for the challenges of causal inference.
-We applied one such approach here (a difference-in-difference estimator), and yet were still unable to reach robust conclusions as to the effect of MPAs on the density of targeted species in the Channel Islands, due to the likely small size of the true effect relative to the strength and variability of environmental drivers.
-
-While this does not mean that all MPAs will face similar challenges in estimating their effects, our results in the rigorously studied Channel Islands system make clear that in many instances empirically detecting a clear regional effect of MPAs may not be possible.
-How then should stakeholders go about adaptively monitoring and managing MPAs?
-Simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can perhaps be tuned to focus on the species groups that have the highest chance of a detectable effect size [@nickols2019].
-Expanding data collection to include robust monitoring of fleet dynamics may help statistical approaches to isolate the true effect of MPAs on conservation outcomes by allowing for improved control of the effects of fishing dynamics, and allow managers to take into account potential negative interactions between MPAs and fleet dynamics such as those that may occur under constant-catch dynamics.
-Whenever possible monitoring programs should be implemented prior to MPA implementation to provide a pre-treatment benchmark.
-Non-equilibrium analyses (Fig.\ref{did-plot}) also help set expectations for effect sizes over time.
-Beyond that, educating communities about the challenges of estimating the effects of MPAs can help set expectations, so that a lack of a clear effect is not necessarily viewed as a failure of the program, but rather considered in the context of reasonable expectations.
-While this paper has focused on the conservation outcomes of MPAs, future work must also address the challenge of predicting and estimating the fishery impacts of protected areas.
-
-<!-- Our results provide a unique survey of the likely regional effects of MPAs to fisheries and conservation, and places our empirical assessment of the regional effects of the Channel Islands MPAs in context.
-The Channel Islands are an intensely studied system, and the challenge of identifying a clear regional effect of the network of MPAs placed there in 2003 may seem surprising.
-However, our results show that in fact a smaller effect size, from the perspective of regional conservation gains, is to be expected in this system, and therefore the true effect will be very challenging to separate from environmental noise.
--->
-
-<!-- Modeling effort such as this can then help manager and stakeholders to set realistic expectations of regional effects for any given MPA network design, and importantly, help inform the design of monitoring programs.
-For example, monitoring may be targeted at the species and fleets that modeling suggests may provide the clearest indication of MPA-mediated effects.
-For cases where bio-economic modeling suggests small potential for MPA driven regional density effects, monitoring efforts can be targeted around detecting potential negative effects should they arise, i.e.
-evidence that the model is wrong, rather than exerting massive amounts of effort on what theory and modeling suggest may be a small and potentially undetectable effect size.
--->
-
-<!-- Policy making is inherently an exercise in utilizing best available theory, experience, and modeling to make decisions that are often extremely difficult to empirically test.
-Causal inference in coupled human and natural systems is incredibly challenging, and our results highlight the particular difficulties in estimating the causal effects of MPAs on regional conservation and fishery outcomes.
-Bio-economic simulation models such as ours can support the policy evaluation process by helping stakeholders and communities set realistic expectations for what MPAs might achieve, help managers understand the scale and magnitude of MPA effects, and help scientists design cost-effective monitoring plans that stand the best chance possible of detecting these effects.
--->
-
-As the number and size of global MPA networks increase, it is critical that we both set appropriate expectations for their outcomes, and plan how we will monitor the performance of these protected areas over time.
-While the history of MPA science has made important strides in helping us understand the dynamics of protected areas, the future of MPA science must directly tackle the challenge of evaluating the performance of these MPAs, and adapting their design as needed to best achieve objectives.
-Commonly employed metrics such as response ratios may be applicable in some circumstances, but can have severe shortcomings as metrics of regional conservation effects.
-Dependence on unreliable estimators of MPA effects may lead to stakeholders incorrectly attributing negative environmental shocks as MPA failures, or interpreting data arising from scorched earth fishing outside MPAs as a conservation success.
-Bio-economic modeling can help frame community expectations, reducing the potential for a reduction in support if unrealistic conservation or fishery gains are not realized.
-Statistical approaches that explicitly address complications such as the spatial spillover effects of MPAs (such as the difference-in-difference approach used here) may give users an improved understanding of the performance of their MPAs.
-Effective use of simulation and statistics to clearly communicate what we should expect and what we can detect from MPAs is critical in ensuring that MPAs play effective roles in fisheries management and marine conservation.
-
-<!-- Despite our best efforts we are unlikely to ever truly "know" the effect of our efforts to mitigate climate change, but must instead rely on comparisons to best available modeling outcomes to understand how effective our policies have been.
-Similarly, given the complexities of marine systems much of our decisions on MPA design will have to be based on effective modeling.
-While we are hardly the first to point out that bio-economic models are a critical tool for MPA design, our results help indicate a minimum floor of model complexity to provide candid assessments of regional MPA effects.
--->
-
-<!-- While factors such as MPA size and degree of depletion are especially strong drivers, for all but the most extreme cases of each a wide array of effects, from negative to highly positive,  are possible based on the complex interaction of factors such as fleet dynamics, movement rates, and recruitment timing.
-Confronting these interactions by considering the likely parameter space for a given region is a critical step then in understanding what likely regional effects of MPAs are.
-While models such as ours do require large numbers of parameters that may be challenging to obtain, our results show that working with communities to confront these uncertainties is preferable to sweeping them under the rug in favor of simpler models that are easier to parameter but miss details that our results show can have dramatic effect on expected outcomes.
--->
-
-<!-- We focus mostly on regional conservation gains in this paper.
-However, fisheries spillover is often another important factor to consider (i.e.
-are fisheries better off with the MPAs than they would have been without them).
-The fishery benefits of MPAs are just as (and likely more) intensely debated than the regional conservation outcomes [@hilborn2004; @roberts2001;@sala2018b;@hilborn2018].
-We only address fisheries affects briefly in this study, but our results highlight important tradeoffs and synergies between conservation and fishery spillover effects of MPAs.
-The good news from a fisheries perspective is also fairly obvious: Both the regional conservation and fishery benefits are expected to be greatest when the fished population is in an extremely depleted state pre-MPA, even over a 15 year time horizon, even for larger MPAs (though further work is needed to compare MPAs to alternative fisheries management strategies in these cases).
-For cases where a valuable and formerly abundant species is overfished, a large MPA may then provide large conservation and fishery gains for that species, while potentially having smaller impacts on other less depleted species.
-Our simulation results also do identify though a large parameter space where MPAs create tradeoffs between moderate conservation gains and moderate fishing losses.
-This type of projection analysis can help managers consider where in this space they may be.
-The most critical point with regards to conservation and fishery effects from our simulation analysis is that the conservation or fishery effects of MPAs cannot be reliably estimated without some knowledge and consideration of the dynamics of the fishing fleet outside the MPAs: over the short-term open access vs constant-catch dynamics can make the difference between a substantial conservation and fisheries win to a more depleted stock with more expensive fishing.
--->
-
-<!-- This process provides a unique survey of the likely regional effects of MPAs to fisheries and conservation, and places our empirical assessment of the regional effects of the Channel Islands MPAs in context.
-The Channel Islands are an intensely studied system, and the challenge of identifying a clear regional effect of the network of MPAs placed there in 2003 may seem surprising.
-However, our results show that in fact a smaller effect size, from the perspective of regional conservation gains, is to be expected in this system, and therefore the true effect will be very challenging to separate from environmental noise.
-The solution then though is not to give up on detecting effects, but rather to shift focus from identifying a specific effect size and instead use simulation analysis to appropriately set expectations for conservation and fishing stakeholders, and design monitoring programs around the species and situations that serve as effective indicators of the ability of an MPA network to achieve its objectives.
--->
-
-## Materials and Methods
-
-We present here critical characteristics of our simulation model and regression approach.
-Further details can be found in the Supplementary Information.
-All analysis were conducted in R [@rcoreteam2019].
-Our main difference-in-difference model was fit using Template Model Builder in R [@kristensen2016].
-A completely reproducible version of our results can be accessed and run at XX.
-
-### Simulation Model
-
-Our bio-economic model simulates the effect of MPAs on a spatially explicit age-structured representation of a single species.
-Readers can explore the functionality of the model using an online tool available [here](https://danovando.shinyapps.io/simmpa/).
-See Table.S1 for a complete description of simulation states.
-The model consists of 25 patches with wrapped edges (picture the waters around a circular island).
-For any one simulation we randomly pull a species and its associated life history (growth, mortality, maturity) from the `FishLife` [@thorson2017c] package in R.
-We pair these data with randomly selected values between 0.6 and 0.95 for Beverton-Holt steepness [as parameterized in @mace1994], as well as larval and adult dispersal rates (where at the lowest values adults and larvae stay within the patch they were created, and at the highest values they are capable of moving throughout the system in each time step).
-We also randomly assign whether adults have density dependent movement (meaning that adult biomass preferentially moves towards patches with lower adult biomass as opposed to random dispersal), as well as one of three potential types of density dependence [building off of @babcock2011]:
-
-  1. Local density dependence: Density dependence occurs independently in each patch, and recruits then disperse to nearby patches
-
-  2. Global density dependence: Density dependence is a function of the sum of spawning biomass across all patches, and recruits are then distributed according to habitat quality
-
-  3. Post-dispersal density dependence: Larvae are distributed throughout the system, and then density dependence occurs based on the density of adult biomass at the destination patch
-
-We allow for three potential siting strategies for MPAs.
-In the first, MPAs are randomly placed.
-In the second, we assume that MPAs are placed in preferentially better habit (unfished recruitment is four times greater inside MPA locations).
-In the third, we allow for scenarios in which MPAs are placed in hotspots of larval dispersal.
-In this scenario, patches in which MPAs will be placed have larval dispersal rates four times greater than patches that do not become MPAs.
-
-Each simulation is assigned a random set of fleet dynamics from one of three categories: constant-catch, constant harvest rate, or open-access.
-Under constant-catch, the fleet attempts to catch a randomly selected multiplier of maximum sustainable yield (MSY) each year (note that values of MSY greater than 1 will crash the population eventually, but allows for overfishing dynamics to be observed when shorter time periods are considered).
-Under a constant harvest rate the fleet captures a randomly selected fraction between 0.01\% and 98\% of the population in each time step.
-Under open access, fishing effort is allowed to expand and contract in response to the profitability of the fishery, where
+Under ideal circumstances they can protect individual species and
+ecosystem linkages, while supporting local economies through tourism and
+fishing opportunities. One rationale for the expansion of MPAs is that
+they will deliver net conservation benefits both inside \emph{and
+outside} their borders. To date, this assumption is insufficiently
+tested, and this is the focus of our paper. Our results show that
+regional conservation benefits of MPAs are highly context dependent and
+in many circumstances, are likely to be so small that they are nearly
+impossible to detect empirically. Indeed, this is exactly what we found
+in our empirical case study from the Channel Islands, California, USA.
+
+We find regional conservation gains from MPAs in 94\% of our simulations
+(Fig.\ref{conservation-effects}). MPAs covering less than 25\% of the
+population range were unlikely to produce regional conservation gains
+above 10\% unless the stock would have been severely overfished without
+the MPA. Given that few marine species on the planet have more than 25\%
+of their entire range protected in MPAs, this suggests that large
+regional gains are unlikely to be common unless the overall extent of
+MPAs grows globally. Large MPAs protecting highly depleted stocks almost
+always produced large regional conservation gains, as should be
+expected. The median conservation effect of 36\% stands in contrast to
+the large within-MPA effects on biomass densities reported by (50) and
+(5). This difference is due in part to the fact that within-MPA effects
+are likely to be much larger than regional-MPA effects, but may also
+reflect the many challenges of translating response ratios into regional
+conservation effects. Our simulation analysis shows that without proper
+controls (in particular for the effects of fleet redistribution),
+response ratios well over 100\% are entirely plausible even when the
+actual conservation effect is much lower (and \emph{vice versa}).
+
+We were not able to detect a statistically significant difference in the
+densities of targeted and non-targeted finfish species over the 13 years
+of MPA protection in the Channel Islands covered by our analysis. The
+data (and model assumptions) provide support for both negative and
+positive MPA effects, although the size of these estimated effects is
+well within the range that our simulation analysis suggests are
+plausible. At face value, this lack of a clear result may seem
+surprising given the size and carefully studied nature of this MPA
+network.
+
+However, our simulation analysis suggests that this result should
+perhaps be expected. The Channel Islands MPAs cover approximately 20\%
+of the waters in the Channel Islands, and while formal stock assessments
+are not available for many of the targeted species in our analysis, what
+evidence we have suggests that, as a group, these fish are on average
+not heavily overfished. Some species, such as California sheephead and
+blue rockfish were likely below target levels during the period (51,
+52), but projections based upon the overall average response across all
+species will likely suggest modest benefits even if a subset of species
+could experience much larger population gains. Harkening back to our
+simulations, we expect the average percentage difference in densities of
+targeted species with and without MPAs to be modest
+(Fig.\ref{did-plot}). Effects of this size are likely to be challenging
+to detect empirically given the large natural variation of marine
+ecosystems (especially temperate reefs) and the observation error
+inherent in infrequent annual survey programs such as those provided by
+PISCO.
+
+In addition, our simulations examine percentage increases in biomass
+densities. As an extreme example, an increase in biomass densities from
+2kg/m\textsuperscript{2} to 4kg/m\textsuperscript{2} would be reported
+as a 100\% increase. While this chance produces a large percentage
+increase, this is a small change in absolute biomass densities relative
+to the variance in the observation process itself (see Fig.S1 for a
+companion to Fig.\ref{conservation-effects} scaled by absolute
+population size). Beyond these challenges, the median estimated age at
+sexual maturity for the targeted species included in this study is 5
+years, meaning that the span of this analysis represents less than three
+generations of MPA protection for half of the the measured species.
+Ongoing monitoring may yet reveal clearer effects. Analysis of more
+rapidly growing and maturing species, e.g.~spiny lobster, may also
+reveal clearer signals.
+
+While the MPA effect estimated by our model was not significant in any
+one year, the positive trend from 2005 to 2014 is evident, as is the
+sharp decline from 2015-2017. To what should we attribute this apparent
+shift? We cannot reject the parallel trends assumption between the
+targeted and non-targeted species in the years before the MPAs
+(Fig.\ref{did-plot}). But, the Channel Islands region (and the entire
+West coast of the USA) experienced a dramatic `marine heatwave'
+beginning in 2014 and persisting through 2016, resulting in part in
+extremely elevated water temperatures throughout the region (53).
+Biogeographic differences in the distributions of targeted and
+non-targeted species may confound the observed effects of MPAs. Many of
+the non-targeted species have warm thermal affinities and have increased
+in numbers since the heatwave (54). However, the targeted group is made
+up mostly of fishes with cold-water affinities, such as members of the
+genus Sebastes. As such, we hypothesize that the recent evidence for a
+decline in densities of targeted species is due to environmental
+conditions that disproportionately affect the targeted group (and not
+for example due to concentrated fishing pressure outside the reserves).
+This hypothesis is supported by the sharp declines in densities of
+targeted species seen inside the MPAs themselves (see Fig.S27-S29),
+suggesting a driver other than fishing may be at play. The Channel
+Islands case study demonstrates the challenge of detecting small effect
+sizes in nature and highlights the potential for small effect sizes to
+be overwhelmed by additional environmental shocks to the system.
+
+What do our results imply about the future of MPA science? Our
+simulation model is by no means exhaustive (for example it ignores
+features such as species interactions, habitat effects, and climate
+feedback), but captures many of the core factors theorized to affect MPA
+performance. Our results show that while attributes such as MPA size and
+fishing pressure are important factors in determining the effects of
+protection, local fleet dynamics and the movement rates of adult fish
+can dramatically affect the outcomes of protected areas as well
+(Fig.\ref{conservation-effects}). Far from being a simple tool with
+clear outcomes, the effects of MPAs can be highly context dependent,
+requiring - we would argue - a bio-economic model of at minimum the
+complexity presented here to help communities design and set
+expectations of MPAs at the tactical level. While users may not be able
+to parameterize every aspect of a model such as that presented here,
+working with stakeholders to visualize the implications of, for example,
+different fleet responses to MPA implementation is a critical step in
+MPA design. Readers can use our simulation model to explore the effects
+of MPAs through an interactive web application available
+\href{https://danovando.shinyapps.io/simmpa/}{here}.
+
+Once simulation models have been used to help design an MPA, how can
+users evaluate whether it is achieving their objectives? Response ratios
+are commonly used as evidence for conservation outcomes of MPAs (5), but
+as suggested in (22) and (41) and demonstrated here, without careful
+attention to the design of control sites (accounting for example for the
+displacement of effort by MPAs), response ratios may be highly
+unreliable estimators of regional MPA effects. As (41) suggests, there
+are many potential alternatives for estimating the effects of MPAs that
+better account for the challenges of causal inference. We applied one
+such approach here (a difference-in-difference estimator), and yet were
+still unable to reach robust conclusions as to the effect of MPAs on the
+density of targeted species in the Channel Islands, due to the likely
+small size of the true effect relative to the strength and variability
+of environmental drivers.
+
+While this does not mean that all MPAs will face similar challenges in
+estimating their effects, our results in the rigorously studied Channel
+Islands system make clear that in many instances empirically detecting a
+clear regional effect of MPAs may not be possible. How then should
+stakeholders go about adaptively monitoring and managing MPAs?
+Simulation modeling can help inform the range of effect sizes that may
+be expected, and monitoring programs can perhaps be tuned to focus on
+the species groups that have the highest chance of a detectable effect
+size (39). Expanding data collection to include robust monitoring of
+fleet dynamics may help statistical approaches to isolate the true
+effect of MPAs on conservation outcomes by allowing for improved control
+of the effects of fishing dynamics, and allow managers to take into
+account potential negative interactions between MPAs and fleet dynamics
+such as those that may occur under constant-catch dynamics. Whenever
+possible monitoring programs should be implemented prior to MPA
+implementation to provide a pre-treatment benchmark. Non-equilibrium
+analyses (Fig.\ref{did-plot}) also help set expectations for effect
+sizes over time. Beyond that, educating communities about the challenges
+of estimating the effects of MPAs can help set expectations, so that a
+lack of a clear effect is not necessarily viewed as a failure of the
+program, but rather considered in the context of reasonable
+expectations. While this paper has focused on the conservation outcomes
+of MPAs, future work must also address the challenge of predicting and
+estimating the fishery impacts of protected areas.
+
+As the number and size of global MPA networks increase, it is critical
+that we both set appropriate expectations for their outcomes, and plan
+how we will monitor the performance of these protected areas over time.
+While the history of MPA science has made important strides in helping
+us understand the dynamics of protected areas, the future of MPA science
+must directly tackle the challenge of evaluating the performance of
+these MPAs, and adapting their design as needed to best achieve
+objectives. Commonly employed metrics such as response ratios may be
+applicable in some circumstances, but can have severe shortcomings as
+metrics of regional conservation effects. Dependence on unreliable
+estimators of MPA effects may lead to stakeholders incorrectly
+attributing negative environmental shocks as MPA failures, or
+interpreting data arising from scorched earth fishing outside MPAs as a
+conservation success. Bio-economic modeling can help frame community
+expectations, reducing the potential for a reduction in support if
+unrealistic conservation or fishery gains are not realized. Statistical
+approaches that explicitly address complications such as the spatial
+spillover effects of MPAs (such as the difference-in-difference approach
+used here) may give users an improved understanding of the performance
+of their MPAs. Effective use of simulation and statistics to clearly
+communicate what we should expect and what we can detect from MPAs is
+critical in ensuring that MPAs play effective roles in fisheries
+management and marine conservation.
+
+\hypertarget{materials-and-methods}{%
+\subsection{Materials and Methods}\label{materials-and-methods}}
+
+We present here critical characteristics of our simulation model and
+regression approach. Further details can be found in the Supplementary
+Information. All analysis were conducted in R (55). Our main
+difference-in-difference model was fit using Template Model Builder in R
+(56). A completely reproducible version of our results can be accessed
+and run at XX.
+
+\hypertarget{simulation-model}{%
+\subsubsection{Simulation Model}\label{simulation-model}}
+
+Our bio-economic model simulates the effect of MPAs on a spatially
+explicit age-structured representation of a single species. Readers can
+explore the functionality of the model using an online tool available
+\href{https://danovando.shinyapps.io/simmpa/}{here}. See Table.S1 for a
+complete description of simulation states. The model consists of 25
+patches with wrapped edges (picture the waters around a circular
+island). For any one simulation we randomly pull a species and its
+associated life history (growth, mortality, maturity) from the
+\texttt{FishLife} (57) package in R. We pair these data with randomly
+selected values between 0.6 and 0.95 for Beverton-Holt steepness (58),
+as well as larval and adult dispersal rates (where at the lowest values
+adults and larvae stay within the patch they were created, and at the
+highest values they are capable of moving throughout the system in each
+time step). We also randomly assign whether adults have density
+dependent movement (meaning that adult biomass preferentially moves
+towards patches with lower adult biomass as opposed to random
+dispersal), as well as one of three potential types of density
+dependence (59):
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Local density dependence: Density dependence occurs independently in
+  each patch, and recruits then disperse to nearby patches
+\item
+  Global density dependence: Density dependence is a function of the sum
+  of spawning biomass across all patches, and recruits are then
+  distributed according to habitat quality
+\item
+  Post-dispersal density dependence: Larvae are distributed throughout
+  the system, and then density dependence occurs based on the density of
+  adult biomass at the destination patch
+\end{enumerate}
+
+We allow for three potential siting strategies for MPAs. In the first,
+MPAs are randomly placed. In the second, we assume that MPAs are placed
+in preferentially better habit (unfished recruitment is four times
+greater inside MPA locations). In the third, we allow for scenarios in
+which MPAs are placed in hotspots of larval dispersal. In this scenario,
+patches in which MPAs will be placed have larval dispersal rates four
+times greater than patches that do not become MPAs.
+
+Each simulation is assigned a random set of fleet dynamics from one of
+three categories: constant-catch, constant harvest rate, or open-access.
+Under constant-catch, the fleet attempts to catch a randomly selected
+multiplier of maximum sustainable yield (MSY) each year (note that
+values of MSY greater than 1 will crash the population eventually, but
+allows for overfishing dynamics to be observed when shorter time periods
+are considered). Under a constant harvest rate the fleet captures a
+randomly selected fraction between 0.01\% and 98\% of the population in
+each time step. Under open access, fishing effort is allowed to expand
+and contract in response to the profitability of the fishery, where
 
@@
 
-To shift the equilibrium of the open access model, we fix price *p* at 1 and $\beta$ at 2, and adjust the cost parameter to achieve the selected fishing mortality rate at bionomic equilibrium.
-These combinations of fleet models allow each simulation to achieve randomly selected levels of fishing pressure through different processes.
-
-<!-- While at equilibrium in the absence of an MPA each of these fleet models can achieve the same level of depletion, each of these fleet models interacts with MPAs in a different manner.  -->
-
-Along with the fleet dynamics model, each simulation is assigned a random fleet dispersal scenario: uniform dispersal (where the total effort of the fleet is divided evenly among all open patches), catch dispersal (where the total effort of the fleet is divided according to the catchable biomass in each available patch), and profit dispersal (where the total effort of the fleet is divided according to the profit per unit effort in each available patch).
-
-Lastly each simulation is assigned an MPA scenario, defined by the number and size of MPAs, the placement of those MPAs, and the year that the MPAs are put in place.
-Many MPA models assume equilibrium conditions prior to the MPA, and then measure equilibrium outcomes.
-While these are important scenarios to understand, they do not reflect the reality of many MPAs, which are often placed in non-equilibrium conditions, and evaluated over the early years of their existence.
-Each simulation starts the population off at unfished equilibrium and then beings to apply the fleet model.
-The MPAs are then placed during the randomly selected start year, allowing some runs to explore how the early dynamics of the MPA play out when the fishery and population they are placed on is not already at equilibrium.
-Fishing effort in displaced by MPAs can either concentrate outside or leave the fishery.
-
-Each simulation is run to equilibrium with and without the selected MPA strategy (holding all else constant).
-We then measure the difference in biomass and fishery catches in each time step in the scenario with and without the MPAs to calculate the conservation and fishery effects of the MPAs over time.
-We filter out any scenario that results in a collapse of the population (biomass less than 10% of unfished biomass) prior to the implementation of MPAs, and any scenario in which the open-access dynamics devolve into chaos.
-
-### Difference in Difference Regression
-
-We use a difference-in-difference regression style regression to attempt to estimate the causal effect of MPAs on regional targeted fish biomass.
-Empirically, this amounts to estimating the pre-post MPA difference in the biomass densities of targeted species net the same difference for non-targeted species in the Channel Islands.
+To shift the equilibrium of the open access model, we fix price \emph{p}
+at 1 and \(\beta\) at 2, and adjust the cost parameter to achieve the
+selected fishing mortality rate at bionomic equilibrium. These
+combinations of fleet models allow each simulation to achieve randomly
+selected levels of fishing pressure through different processes.
+
+Along with the fleet dynamics model, each simulation is assigned a
+random fleet dispersal scenario: uniform dispersal (where the total
+effort of the fleet is divided evenly among all open patches), catch
+dispersal (where the total effort of the fleet is divided according to
+the catchable biomass in each available patch), and profit dispersal
+(where the total effort of the fleet is divided according to the profit
+per unit effort in each available patch).
+
+Lastly each simulation is assigned an MPA scenario, defined by the
+number and size of MPAs, the placement of those MPAs, and the year that
+the MPAs are put in place. Many MPA models assume equilibrium conditions
+prior to the MPA, and then measure equilibrium outcomes. While these are
+important scenarios to understand, they do not reflect the reality of
+many MPAs, which are often placed in non-equilibrium conditions, and
+evaluated over the early years of their existence. Each simulation
+starts the population off at unfished equilibrium and then beings to
+apply the fleet model. The MPAs are then placed during the randomly
+selected start year, allowing some runs to explore how the early
+dynamics of the MPA play out when the fishery and population they are
+placed on is not already at equilibrium. Fishing effort in displaced by
+MPAs can either concentrate outside or leave the fishery.
+
+Each simulation is run to equilibrium with and without the selected MPA
+strategy (holding all else constant). We then measure the difference in
+biomass and fishery catches in each time step in the scenario with and
+without the MPAs to calculate the conservation and fishery effects of
+the MPAs over time. We filter out any scenario that results in a
+collapse of the population (biomass less than 10\% of unfished biomass)
+prior to the implementation of MPAs, and any scenario in which the
+open-access dynamics devolve into chaos.
+
+\hypertarget{difference-in-difference-regression}{%
+\subsubsection{Difference in Difference
+Regression}\label{difference-in-difference-regression}}
+
+We use a difference-in-difference regression style regression to attempt
+to estimate the causal effect of MPAs on regional targeted fish biomass.
+Empirically, this amounts to estimating the pre-post MPA difference in
+the biomass densities of targeted species net the same difference for
+non-targeted species in the Channel Islands.
 
@@
 
-where $d_i$ is the biomass density at observation *i*, *T* indicates whether the observation *i* is for a targeted ($T = 1$) or non-targeted ($T = 0$) species, and *MPA* marks whether observation *i* is in a pre MPA ($MPA = 0$) or post MPA ($MPA = 1$) state.
-Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment (*MPA*) on the treated (targeted species).
-
-<!-- The purpose of this regression approach is to control for unobserved time-invariant variables that are correlated with the treatment (MPA) and affect the outcome  (biomass densities).
-This approach depends on the assumption that the the untreated group (in this case non-targeted species) serve as a proxy for the trend in post-treatment densities in the absence of the MPAs (where the intercept terms for the targeted group allows for different mean densities).
--->
-
-We briefly assess two of the most critical assumptions of this model here: that the treated and non-treated groups have parallel trends, and that the effect of the treatment on the treated does not tangentially affect the untreated (i.e.
-that the implementation of MPAs does not indirectly affect the non-targeted group somehow).
-While the parallel trends assumption cannot be formally proven, we can examine its validity using the data from the years before the MPAs were put in place in 2003 (since after that we no longer expect the trends in the observed data to be parallel).
-We do not detect any significant differences in the trends of the biomass densities of the targeted and non-targeted species in the years before the MPAs, meaning that we do not have evidence to reject the assumption that densities of targeted and non-targeted species were following similar trajectories before the MPAs (though the model does not estimate a precise zero effect either).
-
-With regards to the second assumption, all of the species in this empirical analysis exist within an ecosystem, and as such affect each other through mechanisms such as predation, competition, and habitat modification.
-Conceivably then, protection of carnivorous targeted species inside MPAs could drive down the density of non-targeted prey species, serving in that case to positively bias our estimate of the effect of MPAs on the targeted species.
-While we know dynamics such as this have to exist on some level, we find it unlikely that these effects have had enough time to manifest in the 13 years of post-MPA data used in our analysis [@babcock2010; @pershing2015a].
-
-We used convergent cross mapping (CCM), in the manner of @clark2015, to test for the possibility of the trophic cascades biasing our results.
-Generalizations of Takens' theorem indicate that if two variables (in our case, species or physical variables) are part of the same dynamic system, their individual dynamics should reflect their relative causal influence.
-In other words, if one variable is causally forced by another, that forcing should leave a signature on the first time series.
-Convergent cross mapping (CCM) tests for causation by using the attractor/manifold built from the time series of one variable to predict another (hence the "cross-mapping").
-In simple terms, the causal effect of A on B is determined by how well B cross-maps A.
-CCM then allows us to test for causal relationships in the timeseries of densities of targeted and non-targeted species.
-Our results found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (see SI for additional information in CCM testing).
-However, the longer MPAs are in place, the greater the possibility that substantial species interactions that can affect use of non-targeted species as a control may arise.
-
-While Equation.\ref{eq:did} presents the general form of our model, in practice the estimation model is much more involved.
-At the rawest level, the data are counts of finfish in 2cm length bins along a 30m x 2m transect at various sites and depths.
-These length bins are converted to biomass, and then biomass densities, by converting length to weights using available allometric data and dividing by the transect area.
-Our goal is to estimate the effect of the MPAs on these biomass densities of fish throughout the Channel Islands.
-We fit this model using a hierarchical mixed-effect framework using Template Model Builder [TMB, @kristensen2016] in R.
-The model consists of three levels, the first (starting from the "bottom") being transect-level densities of fish species observed by PISCO, which are standardized into a standardized biomass abundance index, accounting for both probability of detection and expected density as a result of changes in both abundance and covariates such as visibility and observer experience [see @maunder2004].
-For the second stage, we break the abundance indices into targeted and non-targeted species (per the classifications in the PISCO data), and estimate the mean trend of each group (targeted and non-targeted) over time.
-In the third step, we estimate the difference in the de-meaned trend between the targeted and non-targeted fishes (controlling for factors such as water temperature, kelp cover, and commercial fishery catches), that under the assumptions of the model reflects the causal effect of the MPAs on the outcome of interest (in this case regional biomass density of targeted fishes).
-All three of these steps are integrated into the same estimation model, in order to propagate uncertainty through the model correctly.
-Simpler versions of the estimation model, as well as a synthetic control identification strategy, were assessed as well and showed results consistent with the model in our main results (see SI).
-
-<!-- \showmatmethods -->
+where \(d_i\) is the biomass density at observation \emph{i}, \emph{T}
+indicates whether the observation \emph{i} is for a targeted (\(T = 1\))
+or non-targeted (\(T = 0\)) species, and \emph{MPA} marks whether
+observation \emph{i} is in a pre MPA (\(MPA = 0\)) or post MPA
+(\(MPA = 1\)) state. Under the assumptions of this model, \(\beta_3\) is
+the causal effect of the treatment (\emph{MPA}) on the treated (targeted
+species).
+
+We briefly assess two of the most critical assumptions of this model
+here: that the treated and non-treated groups have parallel trends, and
+that the effect of the treatment on the treated does not tangentially
+affect the untreated (i.e.~that the implementation of MPAs does not
+indirectly affect the non-targeted group somehow). While the parallel
+trends assumption cannot be formally proven, we can examine its validity
+using the data from the years before the MPAs were put in place in 2003
+(since after that we no longer expect the trends in the observed data to
+be parallel). We do not detect any significant differences in the trends
+of the biomass densities of the targeted and non-targeted species in the
+years before the MPAs, meaning that we do not have evidence to reject
+the assumption that densities of targeted and non-targeted species were
+following similar trajectories before the MPAs (though the model does
+not estimate a precise zero effect either).
+
+With regards to the second assumption, all of the species in this
+empirical analysis exist within an ecosystem, and as such affect each
+other through mechanisms such as predation, competition, and habitat
+modification. Conceivably then, protection of carnivorous targeted
+species inside MPAs could drive down the density of non-targeted prey
+species, serving in that case to positively bias our estimate of the
+effect of MPAs on the targeted species. While we know dynamics such as
+this have to exist on some level, we find it unlikely that these effects
+have had enough time to manifest in the 13 years of post-MPA data used
+in our analysis (60, 61).
+
+We used convergent cross mapping (CCM), in the manner of (62), to test
+for the possibility of the trophic cascades biasing our results.
+Generalizations of Takens' theorem indicate that if two variables (in
+our case, species or physical variables) are part of the same dynamic
+system, their individual dynamics should reflect their relative causal
+influence. In other words, if one variable is causally forced by
+another, that forcing should leave a signature on the first time series.
+Convergent cross mapping (CCM) tests for causation by using the
+attractor/manifold built from the time series of one variable to predict
+another (hence the ``cross-mapping''). In simple terms, the causal
+effect of A on B is determined by how well B cross-maps A. CCM then
+allows us to test for causal relationships in the timeseries of
+densities of targeted and non-targeted species. Our results found no
+significant cross-mappings between targeted and non-targeted species,
+indicating that while clearly there are interactions between these
+groups on some level, the effects within the timespan of the data are
+not pronounced enough to be of concern to our results (see SI for
+additional information in CCM testing). However, the longer MPAs are in
+place, the greater the possibility that substantial species interactions
+that can affect use of non-targeted species as a control may arise.
+
+While Equation.\ref{eq:did} presents the general form of our model, in
+practice the estimation model is much more involved. At the rawest
+level, the data are counts of finfish in 2cm length bins along a 30m x
+2m transect at various sites and depths. These length bins are converted
+to biomass, and then biomass densities, by converting length to weights
+using available allometric data and dividing by the transect area. Our
+goal is to estimate the effect of the MPAs on these biomass densities of
+fish throughout the Channel Islands. We fit this model using a
+hierarchical mixed-effect framework using Template Model Builder (56) in
+R. The model consists of three levels, the first (starting from the
+``bottom'') being transect-level densities of fish species observed by
+PISCO, which are standardized into a standardized biomass abundance
+index, accounting for both probability of detection and expected density
+as a result of changes in both abundance and covariates such as
+visibility and observer experience (63). For the second stage, we break
+the abundance indices into targeted and non-targeted species (per the
+classifications in the PISCO data), and estimate the mean trend of each
+group (targeted and non-targeted) over time. In the third step, we
+estimate the difference in the de-meaned trend between the targeted and
+non-targeted fishes (controlling for factors such as water temperature,
+kelp cover, and commercial fishery catches), that under the assumptions
+of the model reflects the causal effect of the MPAs on the outcome of
+interest (in this case regional biomass density of targeted fishes). All
+three of these steps are integrated into the same estimation model, in
+order to propagate uncertainty through the model correctly. Simpler
+versions of the estimation model, as well as a synthetic control
+identification strategy, were assessed as well and showed results
+consistent with the model in our main results (see SI).
+
 \showacknow
 
-&nbsp;
-&nbsp;
-&nbsp;
+~ ~ ~
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-johannes1978}{}%
+1. Johannes RE (1978) Traditional Marine Conservation Methods in Oceania
+and Their Demise. \emph{Annual Review of Ecology and Systematics}
+9(1):349--364.
+
+\leavevmode\hypertarget{ref-iucn1976}{}%
+2. IUCN (1976) \emph{IUCN yearbook, 1975-76 : Annual report of the
+International Union for Conservation of Nature and Natural Resources for
+1975 and for January-May 1976}.
+
+\leavevmode\hypertarget{ref-watson2014b}{}%
+3. Watson JEM, Dudley N, Segan DB, Hockings M (2014) The performance and
+potential of protected areas. \emph{Nature} 515(7525):67--73.
+
+\leavevmode\hypertarget{ref-gaines2010}{}%
+4. Gaines SD, White C, Carr MH, Palumbi SR (2010) Designing Marine
+Reserve Networks for Both Conservation and Fisheries Management.
+\emph{Proceedings of the National Academy of Sciences}
+107(43):18286--18293.
+
+\leavevmode\hypertarget{ref-lester2009}{}%
+5. Lester S, et al. (2009) Biological effects within no-take marine
+reserves: A global synthesis. \emph{Marine Ecology Progress Series}
+384:33--46.
+
+\leavevmode\hypertarget{ref-halpern2003}{}%
+6. Halpern BS, Warner RR (2003) Review Paper. Matching marine reserve
+design to reserve objectives. \emph{Proceedings of the Royal Society of
+London Series B: Biological Sciences} 270(1527):1871--1878.
+
+\leavevmode\hypertarget{ref-edgar2014}{}%
+7. Edgar GJ, et al. (2014) Global conservation outcomes depend on marine
+protected areas with five key features. \emph{Nature} 506(7487):216.
+
+\leavevmode\hypertarget{ref-gerber2005}{}%
+8. Gerber LR, Heppell SS, Ballantyne F, Sala E (2005) The role of
+dispersal and demography in determining the efficacy of marine reserves.
+\emph{Canadian Journal of Fisheries and Aquatic Sciences}
+62(4):863--871.
+
+\leavevmode\hypertarget{ref-goni2010}{}%
+9. Goni R, Hilborn R, D\textbackslash{}'ıaz D, Mallol S, Adlerstein S
+(2010) Net contribution of spillover from a marine reserve to fishery
+catches. \emph{Mar Ecol Prog Ser} 400:233--243.
+
+\leavevmode\hypertarget{ref-halpern2009}{}%
+10. Halpern BS, Lester SE, Kellner JB (2009) Spillover from marine
+reserves and the replenishment of fished stocks. \emph{Environmental
+Conservation} 36(04):268--276.
+
+\leavevmode\hypertarget{ref-kay2012}{}%
+11. Kay MC, Lenihan HS, Kotchen MJ, Miller CJ (2012) Effects of marine
+reserves on California spiny lobster are robust and modified by
+fine-scale habitat features and distance from reserve borders.
+\emph{Marine Ecology Progress Series} 451:137--150.
+
+\leavevmode\hypertarget{ref-stobart2009}{}%
+12. Stobart B, et al. (2009) Long-term and spillover effects of a marine
+protected area on an exploited fish community. \emph{Marine Ecology
+Progress Series} 384:47--60.
+
+\leavevmode\hypertarget{ref-mcclanahan2000}{}%
+13. McClanahan TR, Mangi S (2000) Spillover Of Exploitable Fishes From A
+Marine Park And Its Effect On The Adjacent Fishery. \emph{Ecological
+Applications} 10(6):1792--1805.
+
+\leavevmode\hypertarget{ref-russ1996}{}%
+14. Russ G, Alcala A (1996) Do marine reserves export adult fish
+biomass? Evidence from Apo Island, central Philippines. \emph{Marine
+Ecology Progress Series} 132:1--9.
+
+\leavevmode\hypertarget{ref-thompson2017}{}%
+15. Thompson AR, Chen DC, Guo LW, Hyde JR, Watson W (2017) Larval
+abundances of rockfishes that were historically targeted by fishing
+increased over 16 years in association with a large marine protected
+area. \emph{Royal Society Open Science} 4(9):170639.
+
+\leavevmode\hypertarget{ref-baetscher2019}{}%
+16. Baetscher DS, et al. (2019) Dispersal of a nearshore marine fish
+connects marine reserves and adjacent fished areas along an open coast.
+\emph{Molecular Ecology} 28(7):1611--1623.
+
+\leavevmode\hypertarget{ref-pelc2009}{}%
+17. Pelc R, Baskett M, Tanci T, Gaines S, Warner R (2009) Quantifying
+larval export from South African marine reserves. \emph{Marine Ecology
+Progress Series} 394:65--78.
+
+\leavevmode\hypertarget{ref-costa2013}{}%
+18. Costa BH e, et al. (2013) Fishers' Behaviour in Response to the
+Implementation of a Marine Protected Area. \emph{PLOS ONE} 8(6):e65057.
+
+\leavevmode\hypertarget{ref-mason2012}{}%
+19. Mason J, Kosaka R, Mamula A, Speir C (2012) Effort changes around a
+marine reserve: The case of the California Rockfish Conservation Area.
+\emph{Marine Policy} 36(5):1054--1063.
+
+\leavevmode\hypertarget{ref-murawski2005}{}%
+20. Murawski SA, Wigley SE, Fogarty MJ, Rago PJ, Mountain DG (2005)
+Effort distribution and catch patterns adjacent to temperate MPAs.
+\emph{ICES Journal of Marine Science} 62(6):1150--1167.
+
+\leavevmode\hypertarget{ref-mcdermott2019}{}%
+21. McDermott GR, Meng KC, McDonald GG, Costello CJ (2019) The blue
+paradox: Preemptive overfishing in marine reserves. \emph{Proceedings of
+the National Academy of Sciences} 116(12):5319--5325.
+
+\leavevmode\hypertarget{ref-ferraro2018}{}%
+22. Ferraro PJ, Sanchirico JN, Smith MD (2018) Causal inference in
+coupled human and natural systems. \emph{Proceedings of the National
+Academy of Sciences}:201805563.
+
+\leavevmode\hypertarget{ref-gaines2003}{}%
+23. Gaines SD, Gaylord B, Largier JL (2003) Avoiding Current Oversights
+in Marine Reserve Design. \emph{Ecological Applications} 13(sp1):32--46.
+
+\leavevmode\hypertarget{ref-botsford2008}{}%
+24. Botsford LW, et al. (2008) Connectivity, sustainability, and yield:
+Bridging the gap between conventional fisheries management and marine
+protected areas. \emph{Reviews in Fish Biology and Fisheries}
+19(1):69--95.
+
+\leavevmode\hypertarget{ref-difranco2018}{}%
+25. Di Franco A, et al. (2018) Linking home ranges to protected area
+size: The case study of the Mediterranean Sea. \emph{Biological
+Conservation} 221:175--181.
+
+\leavevmode\hypertarget{ref-hilborn2004}{}%
+26. Hilborn R, Punt AE, Orensanz J (2004) Beyond band-aids in fisheries
+management: Fixing world fisheries. \emph{Bulletin of Marine Science}
+74(3):493--507.
+
+\leavevmode\hypertarget{ref-hilborn1992}{}%
+27. Hilborn R (1992) Can fisheries agencies learn from experience?
+\emph{Fisheries} 17(4):6--14.
+
+\leavevmode\hypertarget{ref-walters2004}{}%
+28. Walters CJ, Martell SJ (2004) \emph{Fisheries ecology and
+management} (Princeton University Press).
+
+\leavevmode\hypertarget{ref-hastings2003}{}%
+29. Hastings A, Botsford LW (2003) Comparing Designs Of Marine Reserves
+For Fisheries And For Biodiversity. \emph{Ecological Applications}
+13(sp1):65--70.
+
+\leavevmode\hypertarget{ref-gerber2003}{}%
+30. Gerber LR, et al. (2003) Population Models For Marine Reserve
+Design: A Retrospective And Prospective Synthesis. \emph{Ecological
+Applications} 13(sp1):47--64.
+
+\leavevmode\hypertarget{ref-hilborn2004a}{}%
+31. Hilborn R, et al. (2004) When can marine reserves improve fisheries
+management? \emph{Ocean \& Coastal Management} 47(34):197--205.
+
+\leavevmode\hypertarget{ref-botsford2003}{}%
+32. Botsford LW, Micheli F, Hastings A (2003) Principles for the Design
+of Marine Reserves. \emph{Ecological Applications} 13(sp1):25--31.
+
+\leavevmode\hypertarget{ref-costello2010}{}%
+33. Costello C, Lynham J, Lester SE, Gaines SD (2010) Economic
+incentives and global fisheries sustainability. \emph{Resource} 2.
+
+\leavevmode\hypertarget{ref-fao2018}{}%
+34. FAO (2018) \emph{The State of World Fisheries and Aquaculture 2018:
+Meeting the sustainable development goals} (FAO, Rome, Italy).
+
+\leavevmode\hypertarget{ref-costello2012}{}%
+35. Costello C, Gaines S, Gerber LR (2012) Conservation science: A
+market approach to saving the whales. \emph{Nature} 481(7380):139--140.
+
+\leavevmode\hypertarget{ref-rosenberg2017}{}%
+36. Rosenberg AA, et al. (2017) Applying a New Ensemble Approach to
+Estimating Stock Status of Marine Fisheries around the World.
+\emph{Conservation Letters}:n/a--n/a.
+
+\leavevmode\hypertarget{ref-starr2015}{}%
+37. Starr RM, et al. (2015) Variation in Responses of Fishes across
+Multiple Reserves within a Network of Marine Protected Areas in
+Temperate Waters. \emph{PLoS ONE} 10(3):e0118502.
+
+\leavevmode\hypertarget{ref-caselle2015}{}%
+38. Caselle JE, Rassweiler A, Hamilton SL, Warner RR (2015) Recovery
+trajectories of kelp forest animals are rapid yet spatially variable
+across a network of temperate marine protected areas. \emph{Scientific
+reports} 5:14102.
+
+\leavevmode\hypertarget{ref-nickols2019}{}%
+39. Nickols KJ, et al. (2019) Setting ecological expectations for
+adaptive management of marine protected areas. \emph{Journal of Applied
+Ecology} 0(0).
+doi:\href{https://doi.org/10.1111/1365-2664.13463}{10.1111/1365-2664.13463}.
+
+\leavevmode\hypertarget{ref-villasenor-derbez2019}{}%
+40. Villaseñor-Derbez JC, et al. (2019) An interdisciplinary evaluation
+of community-based TURF-reserves. \emph{PLOS ONE} 14(8):e0221660.
+
+\leavevmode\hypertarget{ref-larsen2019}{}%
+41. Larsen AE, Meng K, Kendall BE (2019) Causal analysis in
+controlImpact ecological studies with observational data. \emph{Methods
+in Ecology and Evolution} 0(0).
+doi:\href{https://doi.org/10.1111/2041-210X.13190}{10.1111/2041-210X.13190}.
+
+\leavevmode\hypertarget{ref-halpern2002}{}%
+42. Halpern BS, Warner RR (2002) Marine reserves have rapid and lasting
+effects. \emph{Ecology Letters} 5(3):361--366.
+
+\leavevmode\hypertarget{ref-halpern2004}{}%
+43. Halpern BS, Gaines SD, Warner RR (2004) Confounding effects of the
+export of production and the displacement of fishing effort from marine
+reserves. \emph{Ecological Applications} 14(4):1248--1256.
+
+\leavevmode\hypertarget{ref-costello2010d}{}%
+44. Costello C, et al. (2010) The value of spatial information in MPA
+network design. \emph{Proceedings of the National Academy of Sciences}
+107(43):18294--18299.
+
+\leavevmode\hypertarget{ref-rassweiler2014}{}%
+45. Rassweiler A, Costello C, Hilborn R, Siegel DA (2014) Integrating
+scientific guidance into marine spatial planning. \emph{Proceedings of
+the Royal Society B: Biological Sciences} 281(1781):20132252.
+
+\leavevmode\hypertarget{ref-osmond2010}{}%
+46. Osmond M, Airame S, Caldwell M, Day J (2010) ``Lessons for marine
+conservation planning: A comparison of three marine protected area
+planning processes''. \emph{Ocean \& Coastal Management} 53(2):41--51.
+
+\leavevmode\hypertarget{ref-kirlin2013}{}%
+47. Kirlin J, et al. (2013) California's Marine Life Protection Act
+Initiative: Supporting implementation of legislation establishing a
+statewide network of marine protected areas. \emph{Ocean \& Coastal
+Management} 74:3--13.
+
+\leavevmode\hypertarget{ref-botsford2014}{}%
+48. Botsford LW, White JW, Carr MH, Caselle JE (2014) Chapter Six -
+Marine Protected Area Networks in California, USA. \emph{Advances in
+Marine Biology}, Marine Managed Areas and Fisheries., eds Johnson ML,
+Sandell J (Academic Press), pp 205--251.
+
+\leavevmode\hypertarget{ref-hilborn2012}{}%
+49. Hilborn R (2012) The role of science in MPA establishment in
+California: A personal perspective. \emph{Environmental Conservation}
+39(03):195--198.
+
+\leavevmode\hypertarget{ref-halpern2003a}{}%
+50. Halpern BS (2003) The impact of marine reserves: Do reserves work
+and does reserve size matter? \emph{Ecological Applications}
+13(sp1):117--137.
+
+\leavevmode\hypertarget{ref-alonzo2004}{}%
+51. Alonzo S, Key M, Ish T, MacCall AD (2004) \emph{Status of the
+California Sheephead Stock for 2004}.
+
+\leavevmode\hypertarget{ref-dick2017}{}%
+52. Dick EJ, et al. (2017) \emph{The Combined Status of Blue and Deacon
+Rockfishes in U.S. Waters off California and Oregon in 2017} (Pacific
+Fishery Management Council, Portland, OR).
+
+\leavevmode\hypertarget{ref-gentemann2017}{}%
+53. Gentemann CL, Fewings MR, García-Reyes M (2017) Satellite sea
+surface temperatures along the West Coast of the United States during
+the 20142016 northeast Pacific marine heat wave. \emph{Geophysical
+Research Letters} 44(1):312--319.
+
+\leavevmode\hypertarget{ref-freedman2019}{}%
+54. Freedman R (2019) Understanding the Efficacy of Spatial Management
+on Emerging Threats. PhD thesis (University of California, Santa
+Barbara).
+
+\leavevmode\hypertarget{ref-rcoreteam2019}{}%
+55. R Core Team (2019) R: A Language and Environment for Statistical
+Computing.
+
+\leavevmode\hypertarget{ref-kristensen2016}{}%
+56. Kristensen K, Nielsen A, Berg CW, Skaug H, Bell BM (2016)
+\textbf{TMB} : Automatic Differentiation and Laplace Approximation.
+\emph{Journal of Statistical Software} 70(5).
+doi:\href{https://doi.org/10.18637/jss.v070.i05}{10.18637/jss.v070.i05}.
+
+\leavevmode\hypertarget{ref-thorson2017c}{}%
+57. Thorson JT, Munch SB, Cope JM, Gao J (2017) Predicting life history
+parameters for all fishes worldwide. \emph{Ecological
+Applications}:n/a--n/a.
+
+\leavevmode\hypertarget{ref-mace1994}{}%
+58. Mace PM (1994) Relationships between Common Biological Reference
+Points Used as Thresholds and Targets of Fisheries Management
+Strategies. \emph{Canadian Journal of Fisheries and Aquatic Sciences}
+51(1):110--122.
+
+\leavevmode\hypertarget{ref-babcock2011}{}%
+59. Babcock EA, MacCall AD (2011) How useful is the ratio of fish
+density outside versus inside no-take marine reserves as a metric for
+fishery management control rules? \emph{Canadian journal of fisheries
+and aquatic sciences} 68(2):343--359.
+
+\leavevmode\hypertarget{ref-babcock2010}{}%
+60. Babcock RC, et al. (2010) Decadal trends in marine reserves reveal
+differential rates of change in direct and indirect effects.
+\emph{Proceedings of the National Academy of Sciences}
+107(43):18256--18261.
+
+\leavevmode\hypertarget{ref-pershing2015a}{}%
+61. Pershing AJ, et al. (2015) Slow adaptation in the face of rapid
+warming leads to collapse of the Gulf of Maine cod fishery.
+\emph{Science} 350(6262):809--812.
+
+\leavevmode\hypertarget{ref-clark2015}{}%
+62. Clark AT, et al. (2015) Spatial convergent cross mapping to detect
+causal relationships from short time series. \emph{Ecology}
+96(5):1174--1181.
+
+\leavevmode\hypertarget{ref-maunder2004}{}%
+63. Maunder MN, Langley AD (2004) Integrating the standardization of
+catch-per-unit-of-effort into stock assessment models: Testing a
+population dynamics model and using multiple data types. \emph{Fisheries
+Research} 70(23):389--395.
+
+% Bibliography
+% \bibliography{pnas-sample}
+
+\end{document}

### 2019-10-09 functions/add-missing-years.R

@@ -1,0 +1,54 @@
+add_mising_years <- function(data, candidate_years){
+
+
+  missing_some <- data %>%
+    group_by(classcode) %>%
+    summarise(missing_some = n_distinct(year) < length(candidate_years)) %>%
+    filter(missing_some)
+
+  fillfoo <- function(data, years_missing){
+
+    blank <- data[1,]
+
+    blank <- map_df(years_missing, ~ mutate(blank,factor_year = as.character(.x)), blank = blank)
+
+    blank$year <- as.numeric(blank$factor_year)
+
+    data <- data %>%
+      bind_rows(blank) %>%
+      arrange(numeric_classcode, year)
+
+    data$density_g_m2 <- exp(data$log_density)
+
+    data$density_g_m2[data$year %in% years_missing] <-  NA
+
+    density_trend <- data %>%
+      group_by(year) %>%
+      summarise(mean_density = mean(density_g_m2)) %>%
+      ungroup() %>%
+      mutate(interped_density = zoo::na.approx(mean_density, rule = 2)) %>%
+      select(-mean_density)
+
+    data <- data %>%
+      left_join(density_trend, by = "year") %>%
+    mutate(log_density = ifelse(is.na(log_density), log(interped_density),log_density )) %>%
+      select(-interped_density,-density_g_m2)
+
+    return(data)
+}
+
+  needs_filling <- data %>%
+    filter(classcode %in% missing_some$classcode) %>%
+    nest(-classcode) %>%
+    mutate(years_missing = map(data,~candidate_years[!candidate_years %in% unique(.x$year)], candidate_years =candidate_years )) %>%
+    mutate(filled_data = map2(data, years_missing, fillfoo)) %>%
+    select(classcode, filled_data) %>%
+    unnest()
+
+  out <- data %>%
+    filter(!classcode %in% missing_some$classcode) %>%
+    bind_rows(needs_filling) %>%
+    arrange(classcode, year)
+
+  return(out)
+}

### 2019-10-09 functions/calculate-baci.R

@@ -1,0 +1,44 @@
+#' calculate baci
+#'
+#' calculates before-after-control-impact metric
+#'
+#' @param dat the simulation jdata
+#' @param distance a matrix of distance to/from MPA border
+#' @param type one of 'biased' or "unbiased" where biased is the world with mpa, unbiased without
+#' @param before_window the number of years to average over for the 'before' treatment
+#'
+#' @return a dataframe of BACI estimates over time
+#' @export
+#'
+calculate_baci <- function(dat, distance_grid = distance_grid,type = "biased", before_window = 1){
+
+  # results$mpa_experiment[[1]]$raw_outcomes -> dat
+
+  ex <- ifelse(type == "biased","with-mpa","no-mpa")
+
+  mpa_year <- min(dat$year[dat$mpa == TRUE])
+
+  sum_dat <- dat %>%
+    filter(experiment == ex) %>%
+    left_join(distance_grid, by = c("patch" = "from")) %>%
+    group_by(year, patch, eventual_mpa) %>%
+    summarise(biomass = sum(biomass),
+              min_fished_dist = unique(min_fished_dist),
+              min_dist = unique(min_dist)) %>%
+    group_by(year) %>%
+    summarise(treated_biomass = weighted.mean(biomass[eventual_mpa == TRUE], min_fished_dist[eventual_mpa == TRUE]),
+              control_biomass = weighted.mean(biomass[eventual_mpa == FALSE], min_dist[eventual_mpa == FALSE]))
+
+  before_treatment <-
+    mean(sum_dat$treated_biomass[sum_dat$year < mpa_year &
+                           sum_dat$year >= mpa_year - before_window])
+
+  before_control <-
+    mean(sum_dat$control_biomass[sum_dat$year < mpa_year &
+                                   sum_dat$year >= mpa_year - before_window])
+
+  baci <- sum_dat %>%
+    group_by(year) %>%
+    summarise(baci = (treated_biomass - before_treatment) - (control_biomass - before_control))
+
+}

### 2019-10-09 functions/center_scale.R

@@ -1,0 +1,10 @@
+center_scale <- function(x, xname = '', omit_names = 'none') {
+  if (is.numeric(x) & !all(unique(x) %in% c(1, 0)) &
+      !xname %in% omit_names) {
+    x <- (x - mean(x, na.rm = T)) / (2 * sd(x, na.rm = T))
+
+  }
+
+  return(x)
+
+}

### 2019-10-09 functions/check-ahnold.R

@@ -1,0 +1,80 @@
+check_ahnold <- function(length_to_density_data, abundance_indices, did_models,min_years = 10, min_year = 2000){
+
+  reg_data <- abundance_indices %>%
+    filter(data_source == 'length_to_density', population_filtering == 'all',
+           population_structure == 'one-pop') %>%
+    select(data, classcode) %>%
+    unnest()
+
+  did_data <- did_models %>%
+    filter(data_source == 'length_to_density', population_filtering == 'all',
+           population_structure == 'one-pop',
+           timing == 'years', abundance_source == 'glm_abundance_index',
+           complexity == 'kitchen_sink') %>%
+    select(data) %>%
+    unnest()
+
+  check_species_counts <- length_to_density_data %>%
+    group_by(year, month, day, site, side, zone, level, transect) %>%
+    summarise(num_species = n_distinct(classcode)) %>%
+    group_by(site) %>%
+    summarise(species_counts_per_event = n_distinct(num_species))
+
+  if (all(check_species_counts$species_counts_per_event == 1)){
+    print('pass: adding zeros worked correctly - all events have same # species')
+
+  }else{
+    print('fail: different number of species per event - check add_missing_fish')
+
+  }
+
+  check_species_coverage <- reg_data %>%
+    group_by(classcode) %>%
+    summarise(nyears = n_distinct(year),
+              min_year = min(year),
+              max_year = max(year))
+
+  if (all(check_species_coverage$nyears >= min_years)){
+    print('pass: all species have minimum number of observed years')
+
+  } else{
+
+    print('fail: some species have too few years - check filterfoo')
+  }
+
+  if (all(check_species_coverage$min_year <= min_year)){
+
+    print(glue::glue('pass: all species start in {min_year}'))
+
+  } else{
+    print('fail: some species do not start in min_year - check filterfoo')
+  }
+
+  reg_data %>%
+    group_by(targeted) %>%
+    summarise(number_of_species = n_distinct(classcode)) %>%
+    knitr::kable() %>%
+    print()
+
+  catch_check <- did_data %>%
+    group_by(targeted) %>%
+    summarise(total_reported_catch = sum(catch, na.rm =T))
+
+  print(knitr::kable(catch_check))
+
+  if (catch_check$total_reported_catch[catch_check$targeted == 0] > 0){
+
+    print('Some "non-targeted" species have reported CDFW catches')
+
+    did_data %>%
[…]

### 2019-10-09 functions/compare-trend-foo.R

@@ -1,0 +1,88 @@
+compare_trend_foo <- function(common_name, data, run_dir) {
+  data <- data %>%
+    select(
+      population_structure,
+      data_source,
+      population_filtering,
+      abundance_index,
+      raw_abundance_trend
+    ) %>%
+    mutate(
+      simple_abundance_index = map2(abundance_index, population_structure, simplify_trend),
+      simple_raw_index = map2(raw_abundance_trend, population_structure, simplify_trend)
+    ) %>%
+    select(-abundance_index, -raw_abundance_trend) %>%
+    gather(abundance_type, data, contains('_index')) %>%
+    unnest()
+
+  length_to_density_pop_trends_plot <- data %>%
+    filter(data_source == 'length_to_density') %>%
+    group_by(population_level, abundance_type, population_structure,population_filtering) %>%
+    mutate(abundance_index = abundance_index / max(abundance_index)) %>%
+    ggplot(aes(
+      year,
+      abundance_index,
+      color = population_level,
+      linetype = abundance_type
+    )) +
+    geom_line() +
+    facet_grid(population_structure ~ population_filtering) +
+    labs(title = common_name)
+
+  raw_density_pop_trends_plot <- data %>%
+    filter(data_source == 'supplied_density') %>%
+    group_by(population_level, abundance_type, population_structure,population_filtering) %>%
+    mutate(abundance_index = abundance_index / max(abundance_index)) %>%
+    ggplot(aes(
+      year,
+      abundance_index,
+      color = population_level,
+      linetype = abundance_type
+    )) +
+    geom_line() +
+    facet_grid(population_structure ~ population_filtering) +
+    labs(title = common_name)
+
+  kfm_density_pop_trends_plot <- data %>%
+    filter(data_source == 'kfm_density') %>%
+    group_by(population_level, abundance_type, population_structure,population_filtering) %>%
+    mutate(abundance_index = abundance_index / max(abundance_index)) %>%
+    ggplot(aes(
+      year,
+      abundance_index,
+      color = population_level,
+      linetype = abundance_type
+    )) +
+    geom_line() +
+    facet_grid(population_structure ~ population_filtering) +
+    labs(title = common_name)
+
+  ggsave(
+    file = paste0(
+      run_dir,
+      '/',
+      common_name,
+      '-length_to_density_pop_trends.pdf'
+    ),
[…]

### 2019-10-09 functions/create-abundance-index.R

@@ -1,0 +1,149 @@
+create_abundance_index <-
+  function(seen_model,
+           seeing_model,
+           seeing_aug,
+           seen_aug,
+           pop_structure) {
+
+
+    # find seen omitted terms
+    # omitted_levels <-
+    #   data_frame(factor_terms = names(seen_model$xlevels),
+    #              omitted_term = NA)
+    #
+    # seen_coefs <- broom::tidy(seen_model)
+    # for (i in seq_along(1:nrow(omitted_levels))) {
+    #   possible_levels <- seen_model$xlevels[[i]]
+    #
+    #   estimated_levels <- seen_coefs %>%
+    #     filter(str_detect(term, omitted_levels$factor_terms[i])) %>% {
+    #       .$term
+    #     }
+    #
+    #   estimated_levels <-
+    #     str_replace(estimated_levels, omitted_levels$factor_terms[i], '')
+    #
+    #   omitted_level <-
+    #     possible_levels[!possible_levels %in% estimated_levels]
+    #
+    #   omitted_levels$omitted_term[i] <- omitted_level
+    #
+    # }
+    #
+    # omitted_levels <- omitted_levels %>%
+    #   mutate(omitted_name = map2_chr(factor_terms, omitted_term, ~ paste(.x, .y, sep = '_')))
+    #
+    # intercept_term <-
+    #   paste(omitted_levels$omitted_name, collapse = '-') # a whole ton of work to double check what the intercept is
+
+    # extract and convert intercept and year terms the old fashioned way
+    # abundance_index <- seen_coefs %>%
+    #   filter(term == "(Intercept)" | str_detect(term, 'factor_year')) %>%
+    #   mutate(term = str_replace(term, 'factor_year',''),
+    #          abundance_index = exp(estimate + std.error ^ 2 / 2)) %>%
+    #   select(term, abundance_index) %>%
+    #   filter(term != '(Intercept)') %>%
+    #   ungroup() #%>%
+    #mutate(abundance_index = center_scale(abundance_index))
+
+    # in order to not drop a year use model to predict abundance over time for the most frequent factor levels held constant
+
+    # seen_aug <-   seen_model %>%
+    #   broom::augment()
+    if (pop_structure == 'one-pop') {
+      seen_series <-
+        create_reference_case(seen_aug = seen_aug, seen_model = seen_model) %>%
+        mutate(smearing_term = mean(exp(seen_aug$.resid)))
+    }
+    if (pop_structure == 'regional-pops')
[…]

### 2019-10-09 functions/create-samples.R

@@ -1,0 +1,154 @@
+create_samples <- function(fishes,
+                           divers,
+                           f_v_m = 1.5,
+                           fleet_q = 0.001,
+                           num_patches = 2,
+                           year_mpa = 50,
+                           sim_years = 100,
+                           burn_years = 50,
+                           mpa_size = 0.5,
+                           samples = 2,
+                           time_step = 1,
+                           enviro_strength = 1,
+                           rec_driver = 'stochastic',
+                           sigma_r = 0,
+                           cores = 1) {
+  simple_fish <- fishes %>%
+    mutate(fish = pmap(
+      list(
+        linf = loo,
+        vbk = k,
+        m = m,
+        length_50_mature = lm
+      ),
+      create_fish,
+      time_step = time_step,
+      sigma_r = sigma_r
+    )) %>%
+    mutate(constant_f = map_dbl(fish, ~ .x$m *  f_v_m, f_v_m = f_v_m)) %>%
+    mutate(constant_effort = num_patches / 2 * ((constant_f / fleet_q * targeted) / time_step)) %>%
+    mutate(fleet = map2(
+      constant_effort,
+      fish,
+      ~ create_fleet(
+        initial_effort = .x,
+        fish = .y,
+        q = fleet_q
+      ),
+      fleet_q = fleet_q
+    ))
+
+  simple_fish <- simple_fish %>%
+    mutate(
+      mpa_experiment = pmap(
+        list(
+          fish = fish,
+          fleet = fleet,
+          enviro = enviro
+        ),
+        run_mpa_experiment,
+        year_mpa = year_mpa,
+        sim_years = sim_years,
+        num_patches = num_patches,
+        burn_years =  burn_years,
+        mpa_size = mpa_size,
+        enviro_strength = enviro_strength,
+        rec_driver = rec_driver
+      )
+    )
+
+  simple_fish <- simple_fish %>%
+    mutate(net_outcomes = map(mpa_experiment, 'outcomes')) %>%
+    mutate(raw_outcomes = map(mpa_experiment, 'raw_outcomes')) %>%
+    mutate(mpa_plot = map(net_outcomes, ~ .x %>%
+                            ungroup() %>%
+                            ggplot(
+                              aes(year, ssb, color = experiment, linetype = experiment)
+                            ) +
[…]

### 2019-10-09 functions/diagnostic-plots.R

@@ -1,0 +1,114 @@
+diagnostic_plots <- function(model){
+
+  # loo_model <- rstanarm::kfold(model, K = 10)
+  safe_bayes <- safely(rstanarm::bayes_R2)
+
+  a <- safe_bayes(model)
+
+  if (is.null(a$error)){
+
+  loo_model <- rstanarm::loo(model)
+
+  r2 <- rstanarm::bayes_R2(model)
+
+  r2_hist_plot <- data_frame(r2 = r2) %>%
+    ggplot(aes(r2)) +
+    geom_histogram(color = 'black', fill = 'grey')
+
+  poster_predictive_plot <- rstanarm::pp_check(model)
+
+  } else{
+
+    loo_model <- NA
+
+    r2 <- NA
+
+    r2_hist_plot <- ggplot()
+
+    poster_predictive_plot <- ggplot()
+
+  }
+
+
+  augmod <- broom::augment(model)
+
+  normal_qq_plot <- augmod %>%
+    ggplot(aes(sample = .resid)) +
+    stat_qq() +
+    stat_qq_line(color = 'red')
+
+  hist_resid_plot <- augmod %>%
+    ggplot(aes(.resid)) +
+    geom_histogram() +
+    geom_vline(aes(xintercept = 0), color = 'blue', linetype = 2) +
+    geom_vline(aes(xintercept = mean(.resid)), color = 'red', linetype = 3)
+
+
+  resid_v_fitted_plot <- augmod %>%
+    ggplot(aes(.fitted, .resid)) +
+    geom_point() +
+    geom_hline(aes(yintercept = 0), color = 'red')
+
+  obs_v_predicted_plot <- augmod %>%
+    ggplot(aes(log_abundance_index, .fitted)) +
+    geom_point() +
+    geom_abline(aes(slope = 1, intercept = 0), color = 'red') +
+    geom_smooth(method = 'lm') +
+    labs(title = glue::glue('Mean Bayesian R2 is {round(mean(r2),2)}'))
+
+
+  comp_plot <-
+    (obs_v_predicted_plot + labs(title = 'A')) + (r2_hist_plot+labs(title = 'B')) +
+    plot_layout(ncol = 1,
+                nrow = 2,
+                heights = c(3, 1))
+
+  if (class(model)[[1]] == 'lmerMod'){
+
+    ind_vars <- model@frame %>% colnames()
+
+  }else{
+
+  ind_vars <- (model$terms %>% as.character())[3]
+
[…]

### 2019-10-09 functions/estimate-net-did.R

@@ -1,0 +1,45 @@
+estimate_net_did <- function(seen_model,
+                             seeing_model){
+
+
+seen_model <- abundance_models$seen_model[[1]]
+
+seeing_model <- abundance_models$seen_model[[1]]
+
+
+a <- abundance_models %>%
+  filter(data_source == 'length_to_density',
+         population_structure == 'one-pop',
+         population_filtering == 'all'
+         ) %>%
+  select(classcode,data) %>%
+  unnest()
+
+seen_data <- a %>%
+  filter(any_seen== T) %>%
+  mutate(targeted = (targeted == 'Targeted') %>% as.numeric) %>%
+  left_join(enso, by = 'year') %>%
+  left_join(pdo, by = 'year') %>%
+  left_join(ci_catches, by = c('classcode', 'year')) %>%
+  mutate(catch = ifelse(is.na(catch), 0, catch),
+         post_mpa = year >=2003)
+
+
+
+
+test <- lm(log_density ~  targeted + post_mpa +targeted:factor_year + region + mean_vis + factor_month +
+             trunc_observer + cumulative_n_obs + level + surge + mean_temp +
+             mean_enso + mean_pdo + lag1_enso + lag2_enso + lag1_pdo + lag2_pdo, data = seen_data)
+
+dids <- test %>%
+  broom::tidy() %>%
+  filter(str_detect(term,'targeted:')) %>%
+  mutate(year = str_replace_all(term,'\\D','') %>% as.numeric())
+
+
+dids %>%
+  ggplot() +
+  geom_pointrange(aes(x = year, y = estimate, ymin =  estimate - 1.96*std.error,
+                      ymax = estimate + 1.96*std.error))
+
+}

### 2019-10-09 functions/fit-ahnold.R

@@ -1,0 +1,270 @@
+fit_ahnold <- function(data,
+                       non_nested_variables =  c(
+                         'targeted',
+                         'factor_year',
+                         'level',
+                         'factor_month',
+                         'cumulative_n_obs',
+                         'temp_deviation',
+                         'surge',
+                         'mean_canopy',
+                         'mean_depth'
+                       ),
+                       script_name,
+                       seed = 42,
+                       run_dir,
+                       use_tmb = T,
+                       fixed_did = T,
+                       include_intercept = T,
+                       center_and_scale = T,
+                       fixed_regions = F) {
+
+
+  numeric_species_key <-
+    data_frame(classcode = unique(data$classcode)) %>%
+    arrange(classcode) %>%
+    mutate(numeric_classcode = 1:nrow(.))
+
+
+  seen_has_important <- data %>%
+    filter(any_seen == T) %>%
+    select(non_nested_variables) %>%
+    mutate(index = 1:nrow(.)) %>%
+    na.omit()
+
+  seeing_has_important <- data %>%
+    select(non_nested_variables) %>%
+    mutate(index = 1:nrow(.)) %>%
+    na.omit()
+
+  seen_data <- data %>%
+    filter(any_seen == T) %>%
+    left_join(numeric_species_key, by = "classcode") %>%
+    slice(seen_has_important$index)
+
+  log_density <- seen_data$log_density
+
+  seen_data <- seen_data %>%
+    select(-log_density)
+
+  seeing_data <- data %>%
+    left_join(numeric_species_key, by = "classcode") %>%
+    slice(seeing_has_important$index)
+
+  any_seen <- seeing_data$any_seen
+
+  seeing_data <- seeing_data %>%
+    select(-any_seen)
+
+
+  # prepare data for c++ ----------------------------------------------------
+
+
+browser()
+  seen_cdata <-
+    make_c_worthy(seen_data,
+                  non_nested_vars = non_nested_variables,
+                  fixed_regions = fixed_regions,
+                  include_intercept = include_intercept,
+                  fixed_did = fixed_did,
[…]

### 2019-10-09 functions/fit-depletion.R

@@ -1,0 +1,26 @@
+
+fit_depletion <- function(target_depletion,scientific_name, target_catch, linf,sim_years,burn_years,
+                   num_patches,
+                   larval_movement,
+                   adult_movement,
+                   density_dependence_form,
+                   steepness)
+{
+  nlminb(
+    c(log(1000*target_catch), log(200)),
+    tune_f,
+    target_depletion = target_depletion,
+    target_catch = target_catch,
+    scientific_name = scientific_name,
+    linf = NA,
+    alpha = 0.5,
+    sim_years = sim_years,
+    burn_years = burn_years,
+    num_patches = num_patches,
+    adult_movement = adult_movement,
+    larval_movement = larval_movement,
+    steepness = steepness,
+    density_dependence_form = density_dependence_form
+  )$par
+
+}

### 2019-10-09 functions/fit-did.R

@@ -1,0 +1,99 @@
+fit_did <- function(did_data, timing, complexity, dirty_dishes,chains = 4, cores = 4) {
+  did_data <- did_data %>%
+    nest(-population_structure,
+         -data_source,
+         -population_filtering,
+         -abundance_source)
+
+
+  # test correlations inspect abundance trends ---------------------------------------
+  # did_data <- did_data %>%
+  #   mutate(correlation_tests = map(data, test_parallel_trends)) %>%
+  #   mutate(
+  #     pre_correlation = map_dbl(correlation_tests, ~ .x$overall_correlation_test$estimate),
+  #     pre_correlation_signif = map_dbl(correlation_tests, ~ .x$overall_correlation_test$p.value)
+  #   )
+
+
+  # fit DiD estimator on abundance indicies ---------------------------------
+
+
+  if (timing == 'years'){
+
+    did_term <- 'targeted:factor_year'
+
+  }
+  if (timing == 'generations'){
+
+    did_term <- paste(c('generations_protected',
+                        'targeted:generations_protected'), collapse = '+')
+
+  }
+  if (timing == 'recruits'){
+
+    did_term <- paste(c('recruits_protected',
+                        'targeted:recruits_protected'), collapse = '+')
+
+  }
+
+  if (complexity == 'bare_bones'){
+
+    did_reg <-
+      paste0('log_abundance_index ~', paste(
+        c(
+          'targeted',
+          'factor_year',
+          did_term
+        ),
+        collapse = '+'
+      ))
+
+  }
+  if (complexity == 'kitchen_sink'){
+
+    did_reg <-
+      paste0('log_abundance_index ~', paste(
+        c(
+          'targeted',
+          'factor_year',
+          did_term,
+          dirty_dishes),
+        collapse = '+'
+      ))
+  }
+  if (str_detect(did_reg,'\\|')){ # if there are random effects
+
+    # fitfoo <- rstanarm::stan_glmer
+
+    did_models <- did_data %>%
+      mutate(did_reg = did_reg) %>%
+      mutate(did_model = map2(
+        data,
+        did_reg,
+        ~ rstanarm::stan_glmer(
+          .y,
+          data = .x,
+          chains = chains,
+          cores = cores
+        )
+      ))
+  } else {
+
+    # fitfoo <- rstanarm::stan_glm
+
+    did_models <- did_data %>%
+      mutate(did_reg = did_reg) %>%
+      mutate(did_model = map2(
[…]

### 2019-10-09 functions/fit-fish.R

@@ -1,0 +1,104 @@
+fit_fish <-
+  function(data,
+           pop_structure,
+           pop_filter,
+           ind_covars,
+           dep_var,
+           family,
+           fit_seen = T,
+           consistent_sites,
+           model_type = 'glm') {
+    if (pop_filter == 'mpa-only') {
+      data <- data %>%
+        filter(eventual_mpa == T)
+
+    }
+    if (pop_filter == 'consistent-sites') {
+      data <- data %>%
+        filter(site %in% consistent_sites$site)
+    }
+
+    if (fit_seen == T) {
+      data <- data %>%
+        filter(any_seen == fit_seen & is.na(any_seen) == F)
+    } else{
+      data <- data %>%
+        filter(is.na(any_seen) == F)
+    }
+
+
+
+    if (pop_structure == 'one-pop') {
+      pop_term <- 'factor_year'
+
+      if (model_type == 'glmer') {
+        pop_term <- '(1|factor_year)'
+      }
+
+    }
+    if (pop_structure == 'regional-pops') {
+
+      num_levels <- length(unique(data$region))
+
+      pop_term <- 'factor_year:region'
+
+      if (num_levels > 1) {
+        pop_term <- 'factor_year:region'
+      } else {
+        pop_term <- 'factor_year'
+      }
+
+
+      if (model_type == 'glmer') {
+
+        if (num_levels > 1) {
+          pop_term <- '(factor_year|region)'
+        } else {
+          pop_term <- '(1|factor_year)'
+        }
+      }
+
+    }
+
+    if (pop_structure == 'mpa-pops') {
+
+      num_levels <- length(unique(data$eventual_mpa))
+
+      pop_term <- 'factor_year:eventual_mpa'
+
+      if (num_levels > 1) {
+        pop_term <- '(factor_year:eventual_mpa)'
+      } else {
+        pop_term <- 'factor_year'
+      }
+
+      if (model_type == 'glmer') {
+
+        if (num_levels > 1) {
+          pop_term <- '(factor_year|eventual_mpa)'
+        } else {
+          pop_term <- '(1|factor_year)'
+        }
+      }
+    }
+    ind_vars <- paste(c(pop_term, ind_covars), collapse = '+')
+    reg_fmla <- paste(dep_var, ind_vars, sep = '~') %>% as.formula()
+
+    if (model_type == 'gam') {
+
+    }
+    if (model_type == 'glm') {
[…]

### 2019-10-09 functions/fit-sim-mpa.R

@@ -1,0 +1,270 @@
+fit_sim_mpa <- function(data,
+                       non_nested_variables =  c(
+                         'targeted',
+                         'factor_year',
+                         'level',
+                         'factor_month',
+                         'cumulative_n_obs',
+                         'temp_deviation',
+                         'surge',
+                         'mean_canopy',
+                         'mean_depth'
+                       ),
+                       script_name,
+                       seed = 42,
+                       run_dir,
+                       use_tmb = T,
+                       fixed_did = T,
+                       include_intercept = T,
+                       center_and_scale = T,
+                       fixed_regions = F) {
+
+
+  numeric_species_key <-
+    data_frame(classcode = unique(data$classcode)) %>%
+    arrange(classcode) %>%
+    mutate(numeric_classcode = 1:nrow(.))
+
+
+  seen_has_important <- data %>%
+    filter(any_seen == T) %>%
+    select(non_nested_variables) %>%
+    mutate(index = 1:nrow(.)) %>%
+    na.omit()
+#
+#   seeing_has_important <- data %>%
+#     select(non_nested_variables) %>%
+#     mutate(index = 1:nrow(.)) %>%
+#     na.omit()
+
+  seen_data <- data %>%
+    filter(any_seen == T) %>%
+    left_join(numeric_species_key, by = "classcode") %>%
+    slice(seen_has_important$index)
+
+  log_density <- seen_data$log_density
+
+  seen_data <- seen_data %>%
+    select(-log_density)
+
+  # seeing_data <- data %>%
+  #   left_join(numeric_species_key, by = "classcode") %>%
+  #   slice(seeing_has_important$index)
+
+  # any_seen <- seeing_data$any_seen
+
+  # seeing_data <- seeing_data %>%
+  #   select(-any_seen)
+
+
+  # prepare data for c++ ----------------------------------------------------
+
+
+
+  seen_cdata <-
+    make_sim_c_worthy(seen_data,
+                  non_nested_vars = non_nested_variables,
+                  fixed_regions = fixed_regions,
+                  include_intercept = include_intercept,
+                  fixed_did = fixed_did,
[…]

### 2019-10-09 functions/fit-zissou.R

@@ -1,0 +1,370 @@
+fit_zissou <- function(data,
+                       non_nested_variables =  c(
+                         'targeted',
+                         'factor_year',
+                         'level',
+                         'factor_month',
+                         'cumulative_n_obs',
+                         'temp_deviation',
+                         'surge',
+                         'mean_canopy',
+                         'mean_depth'
+                       ),
+                       non_nested_did_variables = c(
+                         "temp", "kelp", "lag_catch"
+                       ),
+                       script_name,
+                       seed = 42,
+                       run_dir,
+                       use_tmb = T,
+                       fixed_did = T,
+                       include_intercept = T,
+                       center_scale = T,
+                       fixed_regions = F,
+                       mpa_only = F,
+                       bin_years = FALSE) {
+
+  # data <- data %>%
+  #   filter(classcode != "opic")
+
+  if (bin_years == TRUE){
+
+    data$year <- plyr::round_any(data$year, 3)
+
+    data$factor_year <- factor(data$year)
+
+  }
+
+
+  if (mpa_only == T){
+
+    data <- data %>%
+      filter(eventual_mpa == T)
+  }
+
+
+  if (center_scale == T){
+
+    data_recipe <- recipes::recipe(log_density ~ ., data = data) %>%
+      recipes::step_center(all_numeric(), -all_outcomes(),-year,-targeted,-geographic_cluster) %>%
+      recipes::step_scale(all_numeric(), -all_outcomes(),-year,-targeted,-geographic_cluster)
+
+    prepped_data <- recipes::prep(data_recipe, data, retain = TRUE)
+
+    data <- recipes::juice(prepped_data)
+
+  }
+
+  numeric_species_key <-
+    data_frame(classcode = unique(data$classcode)) %>%
+    arrange(classcode) %>%
+    mutate(numeric_classcode = 1:nrow(.))
+
+
+  seen_has_important <- data %>%
+    filter(any_seen == T) %>%
+    select(non_nested_variables) %>%
+    mutate(index = 1:nrow(.)) %>%
+    na.omit()
+
+  seeing_has_important <- data %>%
+    select(non_nested_variables) %>%
+    mutate(index = 1:nrow(.)) %>%
[…]

### 2019-10-09 functions/get-fish-life.R

@@ -1,0 +1,18 @@
+get_fish_life <- function(genus, species) {
+  Predict = Plot_taxa(
+    Search_species(Genus = genus, Species = species)$match_taxonomy,
+    mfrow = c(2, 2),
+    partial_match = T,
+    verbose = F
+  )
+  out <- Predict[[1]]$Mean_pred %>%
+    as.matrix() %>%
+    t() %>%
+    as.data.frame()
+
+  out[colnames(out) != 'Temperature'] <-
+    exp(out[colnames(out) != 'Temperature'])
+
+  return(out)
+
+}

### 2019-10-09 functions/length-to-weight.R

@@ -1,0 +1,100 @@
+#' length_to_weight
+#' Esimate weight from one row of observations in pisco data, where one
+#' observation is a defined by an integer count of fish of one species
+#' of a given length or within a range of lengths defined by a min, max
+#' and mean length
+#'
+#' @param mean_length the mean length of observations
+#' @param min_length min observed length
+#' @param max_length max observed length
+#' @param count total count for that observation
+#' @param weight_a a coefficient for weight
+#' @param weight_b b coefficient for weight
+#' @param length_units units of length, e.g. cm, mm
+#' @param weight_units output units for weight, e.g. kg
+#' @param length_for_weight_units units from length to weight relationship
+#' @param length_type_for_weight type of length, e.g. SL, TL
+#' @param tl_sl_a params for conversion to/from SL to TL
+#' @param tl_sl_b params for conversion to/from SL to TL
+#' @param tl_sl_type params for conversion to/from SL to TL
+#' @param tl_sl_formula params for conversion to/from SL to TL
+#' @param length_prob type of distirbution to use to generate multiple length comps
+#'
+#' @return a real number of total estiamted weight for that observation
+#' @export
+#'
+length_to_weight <-
+  function(mean_length,
+           min_length,
+           max_length,
+           count,
+           weight_a,
+           weight_b,
+           length_units = 'cm',
+           weight_units = 'g',
+           length_for_weight_units = 'mm',
+           length_type_for_weight,
+           tl_sl_a,
+           tl_sl_b,
+           tl_sl_type,
+           tl_sl_formula,
+           length_prob = "multinomial") {
+    #
+    # generate_lengths <- function(count,mean_length, min_length, max_length){
+
+    if (is.na(count) | count == 0){
+
+      outweight <-  0
+    }  else {
+
+      if (is.na(min_length) |
+          is.na(max_length)) {
+        #generate distribution of lengths
+
+        lengths <-  rep(mean_length, count)
+
+      } else{
+
+        if(length_prob == "unif"){
+
+        lengths <- runif(count, min = min_length, max = max_length)
+        } else {
+
+          probs <- rep(1, round(max_length) - round(min_length) + 1)
+
+        lengths <- as.numeric(rmultinom(1, count, prob = probs))
+
+        lengths <- data_frame(length = round(min_length):round(max_length), counts = lengths) %>%
[…]

### 2019-10-09 functions/make-c-worthy.R

@@ -1,0 +1,108 @@
+make_c_worthy <- function(subdata,
+                          non_nested_vars,
+                          center_and_scale = T,
+                          fixed_did = T,
+                          fixed_regions = T,
+                          include_intercept = T,
+                          numeric_species_key){
+
+
+  x_non_nested <- subdata %>%
+    select(non_nested_vars) %>%{
+      if (include_intercept == T){
+        mutate(.,intercept = 1)
+      } else{
+        .
+      }
+    } %>%
+    spread_factors(drop_one = include_intercept) %>%{
+      if (center_and_scale == T){
+        purrrlyr::dmap(.,center_scale)
+
+      } else
+      {
+        .
+      }
+    }
+
+  if (any(str_detect(non_nested_vars,"cumulative_n_obs"))){
+    x_non_nested <-  x_non_nested %>%
+      mutate(cumulative_n_obs_2 = cumulative_n_obs ^ 2) #%>%
+  }
+
+ # x_enso <- subdata %>%
+ #    select(classcode, enso) %>%
+ #    mutate(index = 1:nrow(.)) %>%
+ #    spread(classcode, enso, fill = 0) %>%
+ #    arrange(index) %>%
+ #    select(-index) %>%
+ #    set_names(paste0("enso_",colnames(.)))
+
+ # x_pdo <- subdata %>%
+ #   select(classcode, pdo) %>%
+ #   mutate(index = 1:nrow(.)) %>%
+ #   spread(classcode, pdo, fill = 0) %>%
+ #   arrange(index) %>%
+ #   select(-index) %>%
+ #   set_names(paste0("pdo_",colnames(.)))
+
+ # x_non_nested <- x_non_nested %>%
+   # bind_cols(x_enso,
+   #           x_pdo)
+
+ x_did <- subdata %>%
+   select(targeted, factor_year) %>%
+   mutate(index = 1:nrow(.)) %>%
+   spread(factor_year, targeted, fill = 0) %>%
+   arrange(index) %>% {
+     if (fixed_did == T){
+       select(.,-(1:2))
+     } else{
+       select(., -1)
+     }
+   }
+
+
+ x_year_species <- subdata %>%
+   mutate(year_classcode = paste(classcode, year, sep = "-")) %>%
+   select(year_classcode) %>%
+   spread_factors(drop_one = F)
+
+ species_index <- subdata$numeric_classcode
+
+ year_species_index <-
+   data_frame(classcode = colnames(x_year_species)) %>%
[…]

### 2019-10-09 functions/make-sim-c-worthy.R

@@ -1,0 +1,106 @@
+make_sim_c_worthy <- function(subdata,
+                          non_nested_vars,
+                          center_and_scale = T,
+                          fixed_did = T,
+                          fixed_regions = T,
+                          include_intercept = T,
+                          numeric_species_key){
+
+
+  x_non_nested <- subdata %>%
+    select(non_nested_vars) %>%{
+      if (include_intercept == T){
+        mutate(.,intercept = 1)
+      } else{
+        .
+      }
+    } %>%
+    spread_factors(drop_one = T) %>%{
+      if (center_and_scale == T){
+        purrrlyr::dmap(.,center_scale)
+
+      } else
+      {
+        .
+      }
+    } %>%
+    # mutate(cumulative_n_obs_2 = cumulative_n_obs ^ 2) %>%
+    mutate(targeted = subdata$targeted)
+
+
+ # x_enso <- subdata %>%
+ #    select(classcode, enso) %>%
+ #    mutate(index = 1:nrow(.)) %>%
+ #    spread(classcode, enso, fill = 0) %>%
+ #    arrange(index) %>%
+ #    select(-index) %>%
+ #    set_names(paste0("enso_",colnames(.)))
+
+ # x_pdo <- subdata %>%
+ #   select(classcode, pdo) %>%
+ #   mutate(index = 1:nrow(.)) %>%
+ #   spread(classcode, pdo, fill = 0) %>%
+ #   arrange(index) %>%
+ #   select(-index) %>%
+ #   set_names(paste0("pdo_",colnames(.)))
+
+ # x_non_nested <- x_non_nested %>%
+ #   bind_cols(x_enso,
+ #             x_pdo)
+
+ x_did <- subdata %>%
+   select(targeted, factor_year) %>%
+   mutate(index = 1:nrow(.)) %>%
+   spread(factor_year, targeted, fill = 0) %>%
+   arrange(index) %>% {
+     if (fixed_did == T){
+       select(.,-(1:2))
+     } else{
+       select(., -1)
+     }
+   }
+
+
+ x_year_species <- subdata %>%
+   mutate(year_classcode = paste(classcode, year, sep = "-")) %>%
+   select(year_classcode) %>%
+   spread_factors(drop_one = F)
+
+ species_index <- subdata$numeric_classcode
+
+ year_species_index <-
+   data_frame(classcode = colnames(x_year_species)) %>%
+   mutate(classcode = str_extract_all(classcode, "(?<=-).*(?=-)", "")) %>%
[…]

### 2019-10-09 functions/mlpa_delta_likelihood.R

@@ -1,0 +1,76 @@
+mlpa_delta_likelihood <- function(parm, Data,reg_model = 'tobit')
+{
+  ### Separate out parameters for observed density  ----
+  beta <- parm[Data$pos_den_beta]
+
+  parm[Data$pos_den_sigma] <- interval(parm[Data$pos_den_sigma], 1e-100, Inf)
+
+  #   parm[Data$pos_den_sigma] <- exp(parm[Data$pos_den_sigma])
+
+  sigma_year <- parm[Data$pos_sigma_year]
+
+  sigma_bi_year <- parm[Data$pos_sigma_bi_year]
+
+  sigma_region <- parm[Data$pos_sigma_region]
+
+  sigma_density <- parm[Data$pos_sigma_density]
+
+  ### Log-Priors for observed ----
+
+  year_priors <- sum(dnorm(parm[Data$pos_den_time_terms],0,sigma_year, log = T))
+
+  bi_year_priors <- sum(dnorm(parm[Data$pos_bi_time_terms],0,sigma_bi_year, log = T))
+
+
+  sigma_year_prior <- dgamma(sigma_year, 2,0.5, log = T)
+
+  sigma_bi_year_prior <- dgamma(sigma_bi_year, 2,0.5, log = T)
+
+  region_priors <- sum(dnorm(parm[Data$pos_den_region_terms],
+                             0,sigma_region, log = T))
+
+  sigma_region_prior <- dgamma(sigma_region, 2,.5, log = T)
+
+  sigma_density_prior <- dgamma(sigma_density, 2,.5, log = T)
+
+
+  ### Hurdle Log-Likelihood ----
+
+  bi_beta <- parm[Data$beta_to_use_binom]
+
+  bi_dat <- Data$bi_reg_mat
+
+  bi_hat <- pmin(10,bi_dat %*% bi_beta)
+# browser()
+#   bi_hat <- pmin(10,rowSums(bi_dat * bi_beta))
+#
+
+  prob_hat <- exp(bi_hat)/(1 + exp(bi_hat))
+
+  bi_loglike <-  sum(dbinom(Data$binom_dep_var,Data$sites_checked,pmax(1e-15,prob_hat), log = T)) # actual values of 1 return -inf if not fulfilled
+
+#   bi_loglike <-  sum(dbinom(Data$binom_dep_var,1,pmax(1e-15,prob_hat), log = T)) # actual values of 1 return -inf if not fulfilled
+
[…]

### 2019-10-09 functions/mlpa_likelihood.R

@@ -1,0 +1,57 @@
+mlpa_likelihood <- function(parm, Data,reg_model = 'tobit')
+{
+  ### Parameters
+  #         a <- proc.time()
+  beta <- parm[Data$pos.beta]
+  sigma <- interval(parm[Data$pos.sigma], 1e-100, Inf)
+
+  parm[Data$pos.sigma] <- sigma
+  ### Log-Priors
+
+  year_priors <- sum(dnorm(parm[Data$pos.time_terms],0,parm[Data$parm.names == 'sigma_year'], log = T))
+
+  sigma_year_prior <- dnorm(parm[Data$parm.names == 'sigma_year'], 1,.2, log = T)
+
+#   sigma_year_prior <- dhalfcauchy(parm[Data$parm.names == 'sigma_year'],25,log = T) #think through this
+
+  region_priors <- sum(dnorm(parm[Data$parm.names %in% Data$site_vars ],
+                             0,parm[Data$parm.names == 'sigma_region'], log = T))
+
+  sigma_region_prior <- dnorm(parm[Data$parm.names == 'sigma_region'], .1,.2, log = T)
+
+  sigma_density_prior <- dnorm(parm[Data$parm.names == 'sigma_density'], .1,.2, log = T)
+
+
+  ### Log-Likelihood
+  mu <- Data$reg_dat %*% beta
+
+  y_star <- Data$dep_var
+
+  #     if (reg_model == 'tobit')
+  #     {
+
+  min_val <- min(Data$dep_var)
+
+  censored <- y_star <= min_val
+
+
+#   thing <- NULL
+#   for (i in 1:sum(censored))
+#   {
+#     thing[i] <- rtnorm(1,mean = mu[censored][i],sd = parm[Data$parm.names == 'sigma_density'], upper = min_val)
+#   }
+#   browser()
+
+  y_star[censored] <- rtnorm(sum(censored),mean = mu[censored],sd = parm[Data$parm.names == 'sigma_density'], upper = min_val)
+  # }
+
+  LL <- sum(dnorm(y_star, mu, parm[Data$parm.names == 'sigma_density'], log=TRUE))
+  ### Log-Posterior
+  LP <- LL + year_priors + sigma_year_prior +  region_priors + sigma_region_prior + sigma_density_prior
+
+  Modelout <- list(LP=LP, Dev=-2*LL, Monitor=LP,
+                   yhat=rnorm(length(mu), mu, sigma), parm=parm)
+
+  #         show(proc.time() - a)
+  return(Modelout)
+}

### 2019-10-09 functions/mlpa_mcmc.R

@@ -1,0 +1,123 @@
+#' Run MCMC on mosquitoes
+#'
+#' \code{Mozzy_MCMC} runs an MCMC on
+#' mosquito count data
+#' @param par_init initival vector of parameter guesses
+#' @param varnames names of the parameters
+#' @param dat are the raw stream-count-distance data
+#' @param covar is the variance-covariance matrix of pars
+#' @param n_sim number of simulations including burn in
+#' @param n_burn number of simulations to burn in
+#' @param n_thin the number of iterations to thin by
+#' @param vcov_augment factor to multiply vcov by in the burn in period
+#' @param prog_bar T or F to output progress bar
+#' @param seed if you want to set a specific seed for this run
+#' @param targ_accept_rate the target acceptance rate to tune the model
+#' @param jumpyness parameter to modify acceptance ratio
+
+mlpa_mcmc <- function(par_init,parm.names,dat,vcov,n_sim=1000,n_burn=0,n_thin=1,
+                     vcov_augment = 1,prog_bar = T,jumpyness = 1, seed = NA,targ_accept_rate = 0.2)
+{
+
+  if (is.na(seed) == F)
+  {
+    set.seed(seed)
+  }
+
+  # Set things up ----
+
+  par_curr <- par_init
+
+  if (class(vcov) == 'data.frame'){ vcov <- as.matrix(vcov) }
+
+  likelihoods <- mlpa_delta_likelihood(parm = par_curr, Data = dat)
+
+  ll_curr <- likelihoods$LP
+
+  deviance_curr <- likelihoods$Dev
+
+  outvars <- c(tolower(parm.names),'ll','deviance')
+
+  par_out <- matrix(NA, nrow = (n_sim - n_burn), ncol = length(outvars))
+
+  i_count <- 0
+
+  if(prog_bar == T)
+  {
+    pb <- txtProgressBar(min = 1, max = n_sim, style = 3)
+  }
+
+  #   a <- proc.time()
+
+
+  sel_count <- 0
+
+  orig_vcov <- vcov
+
+  vcov_scalar <- vcov_augment
+
+  track_rate <- rep(NA,n_sim)
+
+  ### run MCMC----
+  for (i in 1 : n_sim)
+  {
+
+    vcov <- orig_vcov * vcov_scalar
+
+    par_next <- rmvnorm(n = 1, mean= par_curr)
+
+    colnames(par_next) <- parm.names
+
[…]

### 2019-10-09 functions/mpa_jags_model.R

@@ -1,0 +1,18 @@
+# mlpa_delta_likelihood <- function(parm, Data,reg_model = 'tobit')
+# {
+
+'model{
+  for (i in 1:n){
+
+    den_hat[n] <- den_dat[n,1:d] %*% den_beta
+
+    dep_var ~ dnorm(den_hat[n],sigma_density)^(any_seen)
+
+    binom_hat[n] <- min(10, binom_dat[n,1:b] %*% binom_beta)
+
+    prob_hat[n] <- exp(binom_hat[n])/(1 + exp(binom_hat[n]))
+
+    binom_dep_var ~ dbinom(sites_checked,max(1e-15,prob_hat))
+
+  }
+}'

### 2019-10-09 functions/mpa_presentation_figs.R

@@ -1,0 +1,303 @@
+mpa_presentation_figs <- function(run, out,font_size = 18, font_family = 'Helvetica',
+                                  base_theme = theme_grey(), fig_width = 10, fig_height = 8)
+{
+
+
+  plot_theme <- base_theme +
+    theme(text = element_text(size = font_size, family = font_family),
+          axis.title.y = element_text(angle = 0, hjust = 0))
+
+  runfolder <- run
+
+  runpath <- paste('results/',runfolder,'/', sep = '')
+
+  out_folder <- paste(runpath,out,'/',sep = '')
+
+
+  if (dir.exists(out_folder) == F)
+  {
+    dir.create(out_folder, recursive = T)
+  }
+
+  load(paste(runpath,'reg_data.Rdata', sep = ''))
+
+  load(paste(runpath,'MCMC results.Rdata', sep = ''))
+
+  load(paste(runpath,'Processed MCMC results.Rdata', sep = ''))
+
+  .simpleCap <- function(x) {
+    s <- strsplit(x, " ")[[1]]
+    paste(toupper(substring(s, 1, 1)), substring(s, 2),
+          sep = "", collapse = " ")
+  }
+
+  reg_data$species <- sapply(reg_data$species, .simpleCap, USE.NAMES = F)
+
+  species_summary <- reg_data %>%
+    group_by(species) %>%
+    summarize(observations = length(species),fished = unique(fished == 1)) %>%
+    mutate(obs_order = order(observations),
+           species = factor(species, levels = species[obs_order]))
+
+  species_summary$fished[species_summary$fished == TRUE] <- 'Yes'
+
+  species_summary$fished[species_summary$fished == FALSE] <- 'No'
+
+
+  density_plot <- ggplot(reg_data, aes(log_density)) +
+    geom_histogram(fill = 'lightseagreen', color = 'black') +
+    ylab('Log Density (tons/hectare') +
+    xlab('# of Observations') +
+    plot_theme
+
+  ggsave(filename = paste(out_folder,'Density Summary Plot.pdf', sep = ''), density_plot, width = fig_width,
+         height = fig_height)
+
+  species_plot <- ggplot(species_summary,aes(species,observations, fill = fished)) +
+    geom_bar(stat = 'identity', color = 'black') +
+    coord_flip() +
+    ylab('Observations') +
[…]

### 2019-10-09 functions/plot-did.R

@@ -1,0 +1,60 @@
+plot_did <- function(model){
+
+
+ check <-  model$stan_summary %>%
+    as.data.frame() %>%
+    mutate(term = rownames(.)) %>%
+    filter(str_detect(term, 'targeted'))
+
+ if (any(str_detect(check$term, 'generations'))){
+
+   did_terms <- model$stan_summary %>%
+     as.data.frame() %>%
+     mutate(term = rownames(.)) %>%
+     filter(str_detect(term, 'targeted')) %>%
+     mutate(year = map_dbl(term, ~str_extract_all(.x,'(?<=(protected)).*') %>% as.numeric())) %>%
+     filter(!is.na(year))
+
+ } else{
+
+   did_terms <- model$stan_summary %>%
+     as.data.frame() %>%
+     mutate(term = rownames(.)) %>%
+     filter(str_detect(term, 'targeted')) %>%
+     mutate(year = map_dbl(term, ~str_replace_all(.x,'\\D','') %>% as.numeric())) %>%
+     filter(!is.na(year))
+
+ }
+
+
+  did_terms %>%
+    ggplot() +
+    geom_hline(aes(yintercept = 0), color = 'red') +
+    # geom_vline(aes(xintercept = 2003), color = 'blue') +
+    geom_pointrange(aes(
+      x = year,
+      y = mean,
+      ymin = `2.5%`,
+      ymax = `97.5%`
+    ))
+
+  # unfished_terms <- model$stan_summary %>%
+  #   as.data.frame() %>%
+  #   mutate(term = rownames(.)) %>%
+  #   filter(str_detect(term, 'factor_year') & !str_detect(term,'targeted')) %>%
+  #   mutate(year = map_dbl(term, ~str_replace_all(.x,'\\D','') %>% as.numeric())) %>%
+  #   filter(!is.na(year))
+  #
+  # unfished_terms %>%
+  #   ggplot() +
+  #   geom_hline(aes(yintercept = 0), color = 'red') +
+  #   geom_pointrange(aes(
+  #     x = year,
+  #     y = mean,
+  #     ymin = `2.5%`,
+  #     ymax = `97.5%`
+  #   ))
+
+
+
+}

### 2019-10-09 functions/prep_demon.R

@@ -1,0 +1,48 @@
+#' Prepare objects required for Laplaces Demon
+#'
+#' \code{prep_demon} takes data and options and
+#' returns a list with components required by Laplaces Demon
+#' @param demondat data to be used
+#' @param pos_vars possible variables to include in the model
+#' @param  scale_numerics scale T or F to center and scale variables
+
+
+prep_demon <- function(demondat, pos_vars,scale_numerics = F,constant = T)
+{
+  ind_vars <- demondat[,pos_vars] #pull out things you need
+
+  var_types <- sapply(ind_vars,class)
+
+  factors <- pos_vars[which(var_types == 'character' | var_types == 'factor') ]
+
+  binaries <- ind_vars %>%
+    gather('var','value',convert = T) %>%
+    subset(!var %in% factors) %>%
+    mutate(value = as.numeric(value)) %>%
+    group_by(var) %>%
+    summarise(is_binary = sum(value == 0 | value == 1) == length(value)) %>%
+    subset(is_binary == T)
+
+  numerics <- pos_vars[which(var_types == 'numeric')]
+
+  leave_alone <- numerics[grepl('fished',numerics) | grepl('year',numerics) | grepl('temp',numerics)| grepl('lag',numerics)]
+
+  numerics <- numerics[!numerics %in% leave_alone]
+
+  if (scale_numerics == T)
+  {
+    for (j in 1:length(numerics)){
+      ind_vars[,numerics[j]] <- CenterScale(as.matrix(ind_vars[,numerics[j]]))
+    }
+  }
+  for (f in 1:length(factors))
+  {
+    ind_vars <- spread_factor(ind_vars,var = factors[f])
+  }
+
+  if (constant == T){
+    ind_vars$constant <- 1
+  }
+
+  return(ind_vars)
+}

### 2019-10-09 functions/process-fits.R

@@ -1,0 +1,124 @@
+process_fits <- function(zissou_fit) {
+  seen_non_nested_betas <- zissou_fit$zissou_estimates %>%
+    filter(stringr::str_detect(variable, "seen_non_nested_betas")) %>%
+    rename(group = variable) %>%
+    mutate(variable  = zissou_fit$zissou_data$x_seen_non_nested %>% colnames())
+
+  seeing_non_nested_betas <- zissou_fit$zissou_estimates %>%
+    filter(stringr::str_detect(variable, "seeing_non_nested_betas")) %>%
+    rename(group = variable) %>%
+    mutate(variable  = zissou_fit$zissou_data$x_seen_non_nested %>% colnames())
+
+  seen_year_species_betas <- zissou_fit$zissou_estimates %>%
+    filter(stringr::str_detect(variable, "seen_year_species_betas")) %>%
+    rename(group = variable) %>%
+    mutate(variable = zissou_fit$zissou_data$x_seen_year_species %>% colnames())
+
+  seeing_year_species_betas <- zissou_fit$zissou_estimates %>%
+    filter(stringr::str_detect(variable, "seeing_year_species_betas")) %>%
+    # filter(variable == "seeing_year_species_betas") %>%
+    rename(group = variable) %>%
+    mutate(variable = zissou_fit$zissou_data$x_seen_year_species %>% colnames())
+
+  seen_region_cluster_betas <- zissou_fit$zissou_estimates %>%
+    filter(stringr::str_detect(variable, "seen_region_cluster_betas")) %>%
+    # filter(variable == "seen_region_cluster_betas") %>%
+    rename(group = variable) %>%
+    mutate(variable = zissou_fit$zissou_data$x_seen_region_cluster %>% colnames())
+
+  seeing_region_cluster_betas <- zissou_fit$zissou_estimates %>%
+    filter(stringr::str_detect(variable, "seeing_region_cluster_betas")) %>%
+    # filter(variable == "seeing_region_cluster_betas") %>%
+    rename(group = variable) %>%
+    mutate(variable = zissou_fit$zissou_data$x_seen_region_cluster %>% colnames())
+
+  did_betas <- zissou_fit$zissou_estimates %>%
+    filter(stringr::str_detect(variable, "mpa_effect")) %>%
+    # filter(variable == "mpa_effect") %>%
+    mutate(group = variable) %>%
[…]

### 2019-10-09 functions/process_demon.R

@@ -1,0 +1,429 @@
+#' analyze demon object
+#'
+#' \code{process_demon} loads saved data
+#' from run_demon and produces diagnostics
+#' fits etc
+#' @param runfolder the folder where results are stored
+#' @param fontsize numeric fontsize to use
+#' @param post_sample_size the desired final number of samples
+#' from the MCMC after burning and thinning
+#' @param burn proportion 0 to 1 of iterations to burn
+
+process_demon <- function(runfolder,fontsize = 14,post_sample_size = 1000, burn = 0.6)
+{
+
+  runpath <- paste('results/',runfolder,'/', sep = '')
+
+  load(paste(runpath,'reg_data.Rdata', sep = ''))
+
+  load(paste(runpath,'MCMC results.Rdata', sep = ''))
+
+  plot_theme <- theme_classic() + theme(text = element_text(size = fontsize, family = 'Helvetica'))
+
+  post <- reg_results$post
+
+  #   post <- bayes_reg$demon_fit$Posterior1
+
+  its <- dim(post)[1]
+  thinned_post <- thin_mcmc(post[(burn * its):its, ], thin_every = (its*(1-burn))/post_sample_size)
+  ggmcmc(ggs(mcmc(thinned_post)), file = paste(runpath,'ggMCMC Diagnostics.pdf', sep = ''))
+
+  predictions <- apply_demon(demonpost = thinned_post,
+                             dat = reg_results$Data, raw_data = reg_data)
+
+
+  # Make three/4 plots:
+  # map/variogram of residuals by site/region
+  # residuals by time
+  # overlayed distributions of residuals by species group
+
+  # Diagnostic plots ----
+
+  resids_by_time_plot <- predictions$post_dat %>%
+    subset(mean_density >0) %>%
+    group_by(year) %>%
+    mutate(mean_mean_resid = mean(mean_resid)) %>%
+    ggplot(aes(factor(year),mean_resid)) +
+    geom_boxplot() +
+    geom_point(aes(factor(year),mean_mean_resid),
+               shape = 17) +
+    geom_hline(aes(yintercept = 0), color = 'red') +
+    xlab('Year') +
+    ylab('Residuals') +
+    plot_theme
+
+  resids_by_site_plot <- predictions$post_dat %>%
+    subset(mean_density >0) %>%
+    group_by(site) %>%
[…]

### 2019-10-09 functions/refit_lw.R

@@ -1,0 +1,32 @@
+refit_lw <- function(linf, a, b, linf_factor) {
+   old_lengths <- seq(1, linf, length.out = 100)
+
+  pesos <- a * old_lengths ^ b
+
+  new_lengths <- old_lengths * linf_factor
+
+  fitfoo <- function(ws, sizes, peso) {
+    new_weight <- ws[1] * sizes ^ ws[2]
+
+    ss <- sum((new_weight - peso) ^ 2)
+
+    return(ss)
+
+  }
+
+  news <-
+    nlminb(
+      start = c(a, b),
+      fitfoo,
+      sizes = new_lengths,
+      peso = pesos,
+      lower = c(0, 2),
+      upper = c(1e3, 4)
+    )
+
+  out <- list(a = news$par[1], b = news$par[2])
+
+  return(out)
+
+
+}

### 2019-10-09 functions/run-mpa-experiment.R

@@ -1,0 +1,84 @@
+run_mpa_experiment <- function(fish,
+                     fleet,
+                     year_mpa,
+                     mpa_size,
+                     sim_years,
+                     num_patches,
+                     burn_years,
+                     sprinkler = FALSE,
+                     mpa_habfactor = 1,
+                     enviro = NA,
+                     enviro_strength = NA,
+                     rec_driver = 'stochastic',
+                     random_mpa = FALSE,
+                     min_size = 1,
+                     simseed = 42) {
+
+
+  set.seed(simseed)
+  wi_mpa <-
+    sim_fishery(
+      fish = fish,
+      fleet = fleet,
+      manager = create_manager(year_mpa = year_mpa, mpa_size = mpa_size),
+      sim_years = sim_years,
+      num_patches = num_patches,
+      burn_years = burn_years,
+      enviro = enviro,
+      enviro_strength = enviro_strength,
+      rec_driver = rec_driver,
+      sprinkler = sprinkler,
+      mpa_habfactor = mpa_habfactor,
+      tune_costs = FALSE,
+      est_msy = FALSE,
+      random_mpa = random_mpa,
+      min_size = min_size
+    ) %>%
+    mutate(experiment = 'with-mpa')
+
+  mpa_locations <- unique(wi_mpa$patch[wi_mpa$mpa == TRUE])
+
+  set.seed(simseed)
+
+  no_mpa <-
+    sim_fishery(
+      fish = fish,
+      fleet = fleet,
+      manager = create_manager(year_mpa = year_mpa, mpa_size = 0,
+                               mpa_locations = mpa_locations),
+      sim_years = sim_years,
+      num_patches = num_patches,
+      burn_years = burn_years,
+      enviro = enviro,
+      enviro_strength = enviro_strength,
+      rec_driver = rec_driver,
+      sprinkler = sprinkler,
+      mpa_habfactor = mpa_habfactor,
+      tune_costs = FALSE,
+      est_msy = FALSE,
+      random_mpa = random_mpa,
+      min_size = min_size
+    ) %>%
+    mutate(experiment = 'no-mpa')
+
+
+  outcomes <- no_mpa %>%
+    bind_rows(wi_mpa) %>%
+    group_by(year, experiment) %>%
+    summarise(
+      numbers = sum(numbers),
+      biomass = sum(biomass),
+      ssb = sum(ssb),
[…]

### 2019-10-09 functions/run_delta_demon.R

@@ -1,0 +1,340 @@
+#' Function to run MLPA demon using a
+#' delta/hurdle method
+#'
+#' \code{run_delta_demon} uses Laplaces Demon to
+#' fit a bayesian hierarchechal model to the MLPA
+#' data
+#' @param dat dataframe of regression variables
+#' @param dep_var the name of the dependent variable
+#' in the regression
+#' @param pos_vars vector of variables to be potentially included
+#' in the regression
+#' @param delta_vars vector of variables to be included in the
+#' binomial part of the regression
+#' @param iterations the number of runs for the MCMC
+#' @param status the fraction of iterations at which to display status
+#' @param thin the thinning rate for the MCMC
+#' @param burn the amount of the chain to burn off
+#' @param scale_numerics T or F to center and scale numeric covariates
+#' @param runpath the folder to store results
+#' @param acceptance_rate the target acceptance rate
+#' @param method Summon Demon for standard MCMC, Summon Reversible Demon for
+#' reversible MCMC
+#'
+
+run_delta_demon <- function(dat,dep_var,pos_vars,delta_vars,iterations = 1000,status = .05,thin = 1,burn = .5,
+                            scale_numerics = F,
+                            runpath, acceptance_rate  = 0.234, method = 'Summon Demon',
+                            num_chains = 2) {
+
+
+#   dat = reg_data #[1:6000,]
+#   method = 'Jagged Demon'
+#   dep_var = dep_var
+#   pos_vars = pos_vars
+#   delta_vars = delta_vars
+#   runpath = runpath
+#   scale_numerics = scale_numerics
+#   iterations = its
+#   status = .01
+#   acceptance_rate = 0.43
+#   thin = its/1e4
+
+
+  # Convert data to regression format----
+  observed_dat <- prep_demon(dat,pos_vars = pos_vars, scale_numerics = scale_numerics)
+
+  binom_dat <- prep_demon(dat,pos_vars = delta_vars, scale_numerics = scale_numerics)
+
+  vars_to_use_binom <- colnames(binom_dat)
+
+  colnames(binom_dat) <- paste('bi',colnames(binom_dat), sep = '.')
+
[…]

### 2019-10-09 functions/run_demon.R

@@ -1,0 +1,98 @@
+#' Function to run MLPA demon
+#'
+#' \code{run_mlpa_demon} uses Laplaces Demon to
+#' fit a bayesian hierarchechal model to the MLPA
+#' data
+#'
+
+run_mlpa_demon <- function(dat,dep_var,pos_vars,iterations = 1000,status = .05,thin = 1,burn = .5,
+                           scale_numerics = F,
+                           runpath, acceptance_rate  = 0.234, method = 'Summon Demon') {
+
+  observed_dep_var <- dat[,dep_var] > min(dat[,dep_var])
+
+  reg_dat <- prep_demon(dat,pos_vars = pos_vars, scale_numerics = scale_numerics)
+
+  binom_dat <- prep_demon(dat,pos_vars = pos_vars, scale_numerics = scale_numerics)
+
+  reg_frame <- data.frame(dat[,dep_var],reg_dat)
+
+  fmla <- as.formula(paste("log_density ~ ", paste(colnames(reg_dat), collapse= "+"),'- 1'))
+
+  tobit_reg <- tobit(fmla,
+                     data = reg_frame,
+                     left = min(dat[,dep_var], na.rm = T))
+
+  vcov <- tobit_reg$var
+
+  vcov <- vcov[rownames(vcov) != 'Log(scale)', colnames(vcov) != 'Log(scale)']
+
+  J <- dim(reg_dat)[2]
+
+  names <- colnames(reg_dat)
+
+  sigmas <- c('sigma_density','sigma_year','sigma_region')
+
+  #   mus <- c('mu_year')
+
+  time_vars <- names[grepl('_year.2',names)]
+
+  site_vars <- names[grepl('region.',names)]
+
+  species_vars <- names(grepl('trophic.',names))
+
+  mon.names <- "LP"
+
+  parm.names <- c(colnames(reg_dat),sigmas)
+
+  full_vcov <- diag(x = .1, nrow = length(parm.names), ncol = length(parm.names))
+
+  rownames(full_vcov) <- parm.names
+
+  colnames(full_vcov) <- parm.names
+
+  full_vcov[rownames(full_vcov) %in% colnames(reg_dat), colnames(full_vcov) %in% colnames(reg_dat) ] <- vcov
+
+  parm <- rep(0,length(parm.names))
+
+  pos.beta <- which(parm.names %in% colnames(reg_dat))
+
+  pos.sigma <- which(parm.names %in% sigmas)
+
[…]

### 2019-10-09 functions/scrape_enso.R

@@ -1,0 +1,27 @@
+#' \code{scrape_enso} extracts El Niño data from
+# [here]('https://www.esrl.noaa.gov/psd/enso/mei/table.html)
+#'
+#' @return a saved csv of ENSO idices
+#' @export
+#' @import readr
+#'
+#' @examples
+#' scrape_enso(outdir = '../data/')
+scrape_enso <- function(outdir = '../data/'){
+
+
+    # enso <- readr::read_lines('https://www.esrl.noaa.gov/psd/gcos_wgsp/Timeseries/Data/nino34.long.anom.data')
+
+    enso <- readr::read_table('https://www.esrl.noaa.gov/psd/gcos_wgsp/Timeseries/Data/nino34.long.anom.data',
+                              col_names = F, skip = 1,n_max = 2017-1870 + 1) %>%
+      set_names(c('year',1:12)) %>%
+      gather(month, enso,-year) %>%
+      mutate(month =  as.numeric(month)) %>%
+      mutate(date = lubridate::ymd(paste(year,month,1, sep = '-'))) %>%
+      mutate(enso = ifelse(enso == -99.99, NA, enso))
+
+
+readr::write_csv(enso, path = paste0(outdir,'enso.csv'))
+
+
+}

### 2019-10-09 functions/scrape_pdo.R

@@ -1,0 +1,20 @@
+#' \code{scrape_pdo} extracts pacific decadal oscilation data from
+# [here]('https://www.esrl.noaa.gov/psd/gcos_wgsp/Timeseries/Data/pdo.long.data')
+#'
+#' @return a saved csv of PDO idices
+#' @export
+#'
+#' @examples
+#' scrape_pdo(outdir = '../data/')
+scrape_pdo <- function(outdir = '../data/'){
+
+  pdo <- read_table("https://www.esrl.noaa.gov/psd/gcos_wgsp/Timeseries/Data/pdo.long.data",
+                    na = c("-99.99", "99.99",'-99'), skip = 1, n_max = lubridate::year(Sys.time()) - 1900,
+                    col_names = c("year", 1:12)) %>%
+    gather(month, pdo, -year) %>%
+    mutate(month = as.double(month),
+           date = lubridate::ymd(paste(year,month,'01', sep = '-')))
+  readr::write_csv(pdo, path = paste0(outdir,'pdo.csv'))
+
+
+}

### 2019-10-09 functions/select-dishes.R

@@ -1,0 +1,117 @@
+select_dishes <- function(did_data, run_dir) {
+
+  did_data <- did_data %>%
+    filter(population_filtering == 'all',
+           data_source == 'length_to_density',
+           population_structure == 'one-pop',
+           abundance_source == 'glm_abundance_index')
+
+time_enviro_correlations <- did_data %>%
+  select(year, mean_enso, mean_pdo, mean_annual_kelp, mean_annual_temp) %>%
+  cor()
+
+pdf(glue::glue("{run_dir}/did_covariate_correlations.pdf"))
+corrplot::corrplot(time_enviro_correlations)
+dev.off()
+
+
+# dishes <- 'loo + (mean_pdo + lag1_pdo + lag2_pdo + lag3_pdo + lag4_pdo + mean_annual_kelp + temp_deviation |classcode)'
+
+dishes <- c('loo + (mean_enso + mean_annual_kelp + temp_deviation +mean_pdo + lag1_pdo + lag2_pdo + lag3_pdo + lag4_pdo - 1 |classcode)',
+            '(mean_enso + mean_annual_kelp + temp_deviation +mean_pdo + lag1_pdo + lag2_pdo + lag3_pdo + lag4_pdo - 1 |classcode)',
+            'loo + (mean_enso + mean_annual_kelp + temp_deviation - 1|classcode)',
+            'loo + (mean_enso + lag1_enso + lag2_enso + lag3_enso + lag4_enso + mean_annual_kelp + temp_deviation - 1 |classcode)',
+            'loo + (mean_pdo + lag1_pdo + lag2_pdo + lag3_pdo + lag4_pdo + mean_annual_kelp + temp_deviation - 1 |classcode)',
+            'loo + mean_enso + lag1_enso + lag2_enso + lag3_enso + lag4_enso + mean_annual_kelp + temp_deviation',
+            'loo + mean_pdo + lag1_pdo + lag2_pdo + lag3_pdo + lag4_pdo + mean_annual_kelp + temp_deviation',
+            'loo + mean_enso + mean_annual_kelp + temp_deviation' ,
+            'loo + (mean_enso + mean_annual_kelp + temp_deviation - 1|geographic_cluster)')
+
+did_models <-
+  cross_df(
+    list(
+      did_data = list(did_data),
+      timing = c('years'),
+      complexity = c('kitchen_sink'),
+      dirty_dishes = dishes
+    )
+  )
+
+
[…]

### 2019-10-09 functions/sim-scuba-steve.R

@@ -1,0 +1,42 @@
+sim_scuba_steve <- function(pop,diver, fish, patches, effort = 1, cv = 0){
+
+pop_at_age <- pop %>%
+  group_by(age) %>%
+  summarise(number = sum(numbers))
+
+pop_at_age <- pop_at_age %>%
+  mutate(mean_length = fish$linf * (1 - exp(-fish$vbk * (age - fish$t0)))) %>%
+  mutate(selectivity = (1 / (1 + exp(-log(
+  19
+) * ((mean_length - diver$sel_size_50) / (diver$sel_size_delta)
+)))))
+
+pop_at_age <- pop_at_age %>%
+  mutate(potential_seen = number * (1 - exp(-diver$q * effort * selectivity)))
+
+pop_at_age <- pop_at_age %>%
+  mutate(numbers_seen = rmultinom(1, sum(potential_seen), potential_seen) %>% as.numeric())
+
+
+age_samples <- pop_at_age
+
+
+length_samples <- spasm::sample_lengths(age_samples, sample_col = quo(numbers_seen),sample_type = 'other',
+                                        cv = cv,
+                                        k = fish$vbk,
+                                        linf = fish$linf,
+                                        t0 = fish$t0,
+                                        percent_sampled = 1,
+                                        linf_buffer = 1.25
+                                        ) %>%
+  mutate(weight = numbers * (fish$weight_a * length_bin ^fish$weight_b))
+
+
+
+
+return(list(age_samples = age_samples,
+            length_samples = length_samples))
+
+
+
+}

### 2019-10-09 functions/spread-factors.R

@@ -1,0 +1,23 @@
+spread_factors <- function(data, drop_one = T) {
+  factors <-
+    colnames(data)[map_lgl(data, ~ class(.x) %in% c('character', 'factor'))]
+  for (i in seq_along(factors)) {
+    var <- factors[i]
+
+    data <- data %>%
+      ungroup() %>%
+      mutate(dummy = 1, index = 1:nrow(.)) %>%
+      spread_(var, 'dummy', fill = 0, sep = 'dummy') %>%
+      arrange(index) %>% {
+        if (drop_one == T) {
+          dplyr::select(., -index, -contains('dummy')[1]) #drop the first dummy
+        } else{
+          dplyr::select(., -index)
+        }
+      } %>%
+      set_names(str_replace(colnames(.), 'dummy', '-'))
+
+  }
+
+  return(data)
+}

### 2019-10-09 functions/study-mpa.R

@@ -1,0 +1,98 @@
+#' study mpas
+#'
+#' study mpas calculates density ratios, BACI experiments, etc. from simulated MPAs
+#'
+#' @param results results from MPA experiments generated as part of zissou
+#'
+#' @return a list object with MPA studies appended
+#' @export
+#'
+study_mpa <- function(results){
+
+
+calc_mpa_effect <- function(outcomes) {
+  mpa_effect <- outcomes %>%
+    group_by(year) %>%
+    mutate(mpa_size = max(percent_mpa)) %>%
+    ungroup() %>%
+    select(year, experiment, biomass, mpa_size) %>%
+    spread(experiment, biomass) %>%
+    mutate(mpa_effect = `with-mpa` / `no-mpa` - 1) # %>%
+  # select(year, mpa_size, mpa_effect)
+}
+
+
+calc_abs_mpa_effect <- function(outcomes) {
+  mpa_effect <- outcomes %>%
+    group_by(year) %>%
+    mutate(mpa_size = max(percent_mpa)) %>%
+    ungroup() %>%
+    select(year, experiment, biomass, mpa_size) %>%
+    spread(experiment, biomass) %>%
+    mutate(mpa_effect = `with-mpa` - `no-mpa`) # %>%
+  # select(year, mpa_size, mpa_effect)
+}
+
+
+calc_mpa_fishery_effect <- function(outcomes) {
+  mpa_effect <- outcomes %>%
+    group_by(year) %>%
+    mutate(mpa_size = max(percent_mpa)) %>%
+    ungroup() %>%
+    select(year, experiment, catch, mpa_size) %>%
+    spread(experiment, catch) %>%
+    mutate(mpa_effect = `with-mpa` / `no-mpa` - 1)
+}
+
+
+patch_weight <-  results$mpa_experiment[[1]]$raw_outcomes %>%
+  filter(year == max(year), experiment == "with-mpa") %>%
+  group_by(patch) %>%
+  summarise(mpa = unique(mpa))
+
+# weight fished mpa scenario densities by distance from nearest MPA
+distance_grid <-
+  expand.grid(from = 1:num_patches, to = 1:num_patches) %>%
+  as.data.frame() %>%
+  mutate(distance = purrr::map2_dbl(from, to, ~ min(
+    c(abs(.x - .y),
+      .x + num_patches - .y,
+      num_patches - .x + .y)
+  ))) %>%
[…]

### 2019-10-09 functions/test-performance.R

@@ -1,0 +1,181 @@
+test_performance <-
+  function(fishes,
+           year_mpa,
+           min_year = 75,
+           max_year = 100,
+           time_step = 1) {
+    simple_data <- fishes %>%
+      select(loo, k, lm, m, targeted, classcode, commonname, pisco_samples) %>%
+      unnest() %>%
+      filter(year > min_year, year < max_year) %>%
+      select(-pop,-sampled_lengths,-diver_stats) %>%
+      mutate(year = year * time_step) %>%
+      mutate(subyear = year - floor(year),
+             year = floor(year)) %>%
+      mutate(
+        factor_year = as.factor(year),
+        log_density = log(density),
+        logical_targeted = targeted > 0,
+        any_seen = density > 0,
+        region = patches,
+        post_mpa = year >= year_mpa
+      )
+
+    years_protected <- unique(simple_data$year) - year_mpa
+
+    bins <-
+      c(seq(min(years_protected),-1, by = 5), seq(0, max(years_protected), by = 5))
+
+    year_block <- data_frame(year = unique(simple_data$year)) %>%
+      mutate(years_protected = year - year_mpa) %>%
+      mutate(protected_block = cut(
+        years_protected,
+        breaks = bins,
+        include.lowest = T
+      )) %>%
+      select(-years_protected)
+
+    simple_data <- simple_data %>%
+      left_join(year_block, by = "year")
+
+    true_effect <- fishes %>%
+      select(classcode, targeted, mpa_effect) %>%
+      unnest() %>%
+      # mutate(mpa_effect = log(`with-mpa`) - log(`no-mpa`)) %>%
+      filter(year > min_year, targeted == 1) %>%
+      mutate(post_mpa = year > year_mpa) %>%
+      mutate(year = year * time_step) %>%
+      mutate(subyear = year - floor(year),
+             year = floor(year))
+
+
+    bare_bones_model <-
+      lm(log_density ~ targeted + protected_block + targeted:protected_block,
+         data = simple_data)
+    if (n_distinct(simple_data$diver) > 1) {
+      # mixed_effect_model <-
+      #   lme4::lmer(
+      #     log_density ~ (1 + enso|classcode) + diver + targeted*protected_block ,
+      #     data = simple_data,
+      #     verbose = 1
+      #   )
+
+      mixed_effect_model <-
+        lme4::lmer(
[…]

### 2019-10-09 functions/thin_mcmc.R

@@ -1,0 +1,7 @@
+thin_mcmc <- function(chains, thin_every =  1)
+{
+
+  thinned_chains <- chains[(1:dim(chains)[1]) %% thin_every == 0,] #only select chains divisible by the thinning interval
+
+  return(thinned_chains)
+}

### 2019-10-09 functions/tune-f.R

@@ -1,0 +1,79 @@
+
+tune_f <-
+  function(params,
+           target_depletion,
+           linf,
+           scientific_name,
+           adult_movement,
+           larval_movement,
+           steepness,
+           density_dependence_form,
+           target_catch = 100,
+           alpha = 0.5,
+           sim_years = sim_years,
+           burn_years = burn_years,
+           num_patches = num_patches) {
+
+    scientific_name <- gsub("(^)([[:alpha:]])", "\\1\\U\\2", scientific_name, perl=TRUE)
+
+    fish <- create_fish(
+      scientific_name = scientific_name,
+      linf = linf,
+      m = NA,
+      r0 = exp(params[1]),
+      adult_movement = adult_movement,
+      larval_movement = larval_movement,
+      density_dependence_form = density_dependence_form,
+      steepness = steepness,
+      query_fishlife = T
+    )
+
+    manager <- create_manager(year_mpa = 999)
+    unfished <-
+      sim_fishery(
+        fish = fish,
+        fleet = create_fleet(initial_effort = 0, fish = fish, fleet_model = "constant-effort"),
+        manager = manager,
+        num_patches = num_patches,
+        sim_years = sim_years,
+        burn_year = burn_years
+      )
+
+    ssb0 <- unfished %>%
+      filter(year == max(year)) %>%
+      summarise(ssb = sum(ssb)) %>% {
+        .$ssb
+      }
+
+    fished <-
+      sim_fishery(
+        fish = fish,
+        fleet = create_fleet(initial_effort = exp(params[2]), fish = fish, fleet_model = "constant-effort"),
+        manager = manager,
+        sim_years = sim_years,
+        burn_year = burn_years,
+        num_patches = num_patches
+      )
+
+    ssb <- fished %>%
+      filter(year == max(year)) %>%
+      summarise(ssb = sum(ssb)) %>% {
+        .$ssb
+      }
+
+    fish_caught <- fished %>%
+      filter(year == max(year)) %>%
+      summarise(caught = sum(biomass_caught)) %>% {
+        .$caught
+      }
+
+
+    depletion <- ssb / ssb0
+
+    ss <-
+      alpha * (log(depletion) - log(target_depletion)) ^ 2 + (1 - alpha) * (log(fish_caught) - log(target_catch)) ^
+      2
+
+    print(ss)
+    return(ss)
+  }

### 2019-10-09 functions/tune-fishery.R

@@ -1,0 +1,376 @@
+#' assign tuned variables
+#'
+#' tunes a control variable to achieve a target depletion level
+#'
+#' @param control_variable
+#' @param target_depletion
+#' @param fish
+#' @param fleet
+#' @param sim_years
+#' @param burn_years
+#' @param num_patches
+#'
+#' @return ss for depletion
+#' @export
+#'
+#' @examples
+tune_fishery <- function( f_v_m,
+                               fish,
+                               fleet,
+                               sim_years = 50,
+                               burn_years = 1,
+                               num_patches = 10,
+                          sprinkler = sprinkler,
+                          mpa_habfactor = mpa_habfactor){
+  
+
+  if (fleet$fleet_model == "constant-catch"){
+
+    tol <- 100
+
+    lower <- 0
+
+    upper <- 400
+
+    golden <- (sqrt(5) -1)/2
+
+    best <- 1000
+
+    delta_best <- 100
+
+    counter <-  0
+
+    while(delta_best > tol | counter < 20) {
+
+      counter <- counter + 1
+
+      constant <- (1 - golden) * (upper - lower)
+
+      x1 <- lower + constant
+
+      x2 <- upper - constant
+
+      yield_1 <- estimate_msy(x1, fish = fish, fleet = fleet)
+
+      yield_2 <- estimate_msy(x2, fish = fish, fleet = fleet)
+
+      delta_best <-  (best -  min(yield_1,yield_2))^2
+
+      best <- min(yield_1,yield_2)
+
+      if (yield_1 < yield_2){
+
+        lower <- lower
+
+        upper <- x2
+      } else{
+
+        lower <- x1
+
+        upper <- upper
+
+      }
+
+    } # close golden while
+
+    msy_fit <- nlminb(mean(c(lower, upper)), estimate_msy, fish = fish, fleet = fleet, lower = 0,
+                      num_patches = 1)
+
+    fleet$e_msy <- msy_fit$par
+
+    fish$msy <- -msy_fit$objective
+
+    fish$b_msy <- estimate_msy(fleet$e_msy, fish = fish, fleet = fleet, use = "other")
+
+    msy <- fish$msy
+
+    fleet$target_catch <- msy * f_v_m
+
+
+  } else if (fleet$fleet_model == "constant-effort"){
+
[…]

### 2019-10-09 functions/vasterize-pisco-data.R

@@ -1,0 +1,324 @@
+#' Apply VAST to survey data
+#'
+#' \code{vasterize_pisco_data} takes survey data and applies the VAST model
+#' to provide standardized spatio-temporal abundance indicies
+#'
+#' @param raw_data the survey data in question
+#' @param region the general region, see VAST documentation
+#' @param run_dir the directory results will be stored in
+#' @param version the vast version to use
+#' @param method grid or mesh method
+#' @param strata
+#' @param n_x
+#' @param randomseed
+#' @param nstart
+#' @param iter.max
+#' @param omega1
+#' @param epsilon1
+#' @param omega2
+#' @param epsilon2
+#' @param beta1
+#' @param beta2
+#' @param vessel
+#' @param vessel_year
+#' @param obs_model a vector with the statistical model for the "delta" model
+#' @param sd_site_density
+#' @param sd_site_logdensity
+#' @param calculate_range
+#' @param calculate_evenness
+#' @param calculate_effective_area
+#' @param calculate_cov_se
+#' @param calculate_synchrony
+#' @param calculate_coherence
+#'
+#' @return a list of cleaned up outputs from VAST
+#' @export
+#'
+vasterize_pisco_data <- function(raw_data,
+                                 region = 'California_current',
+                                 run_dir = '../results/',
+                                 version = "VAST_v2_4_0",
+                                 method = 'Mesh',
+                                 strata = "All_areas",
+                                 n_x = 1000,
+                                 randomseed = 1,
+                                 nstart = 100,
+                                 iter.max = 1e3,
+                                 omega1 = 1,
+                                 epsilon1 = 1,
+                                 omega2 = 1,
+                                 epsilon2 = 1,
+                                 beta1 = 0,
+                                 beta2 = 0,
+                                 vessel = 0,
+                                 vessel_year = 0,
+                                 obs_model = c(2, 0),
+                                 sd_site_density = 0,
+                                 sd_site_logdensity = 0,
[…]

### 2019-10-09 make-regional-mpas.R

@@ -1,0 +1,3457 @@
+# Run Complete Zissou Analysis --------------------------------------------
+
+
+# libraries ---------------------------------------------------------------
+set.seed(42)
+library(scales)
+library(viridis)
+library(ggmap)
+library(forcats)
+library(stringr)
+library(lubridate)
+library(lme4)
+library(TMB)
+library(FishLife)
+library(patchwork)
+library(rstan)
+library(rstanarm)
+library(extrafont)
+library(hrbrthemes)
+library(here)
+library(doParallel)
+library(spasm)
+library(furrr)
+library(ggsci)
+library(recipes)
+library(sf)
+library(ggmap)
+library(rEDM)
+library(bayesplot)
+library(tidyverse)
+extrafont::loadfonts()
+
+
+
+
+functions <- list.files(here::here("functions"))
+
+walk(functions, ~ here::here("functions", .x) %>% source()) # load local functions
+
+
+# options -----------------------------------------------------------------
+
+run_name <- 'v5.0'
+
+run_description <- "now with recruitment deviates and autocorrleation - testing everything works post rebase with master after that screwup before mering back into master"
+
+# the following analysis run the complete contents of "regional-effects-of-mpas". Each section depends on the out
+# outcomes of the prior section, but will load relevant saved files.
+# So, once you've run simulate_mpas, you can set it to FALSE and validata_mpas will work
+
+
+run_did <- TRUE # run difference in difference on data from the CINMS
+
+simulate_mpas <- FALSE # simulate MPA outcomes
+
+validate_mpas <- FALSE
+
+process_results <- FALSE
+
+knit_paper <- TRUE
+
+sim_years <- 50
+
+num_patches <- 50
+
+n_cores <- 4
+
+# prepare run -------------------------------------------------------------
+
+run_dir <- here::here("results", run_name)
+
+if (dir.exists(run_dir) == F) {
+  dir.create(run_dir, recursive = T)
+}
+
+data_dir <- "data"
+
+write(run_description,
+      file = paste(run_dir, 'RUN_DESCRIPTION.txt', sep = '/'))
+
[…]

### 2019-10-09 src/fit_simmpa.cpp

@@ -1,0 +1,203 @@
+
+// TMB attempt at fitting ahnold
+#include <TMB.hpp>
+template<class Type>
+Type objective_function<Type>::operator() ()
+{
+  /////////load data/////////
+
+  // seen data
+
+  DATA_MATRIX(x_seen_non_nested); // non nested part of data
+
+  DATA_MATRIX(x_seen_did); // difference in difference terms
+
+  DATA_MATRIX(x_seen_year_species); // year random effects
+
+  DATA_MATRIX(x_seen_region_cluster);
+
+  DATA_IVECTOR(year_species_index); // index the same rows as x_seen showing what species is in that row
+
+  DATA_IVECTOR(seen_species_index); // index the same rows as x_seen showing what species is in that row
+
+  DATA_IVECTOR(region_cluster_index); // index the same rows as x_seen showing what species is in that row
+
+  DATA_VECTOR(log_density); // observed log densities
+
+  // seeing data
+
+  // DATA_MATRIX(x_seeing_non_nested); // non nested part of data
+
+  // DATA_MATRIX(x_seeing_did); // difference in difference terms
+
+  // DATA_MATRIX(x_seeing_year_species); // year random effects
+
+  // DATA_MATRIX(x_seeing_region_cluster); // REGION CLUSTERS
+
+  // DATA_VECTOR(any_seen); // observed log densities
+
+
+
+  // did data
+
+  DATA_MATRIX(standard_non_nested); // standardized non nested part of data
+
+  DATA_MATRIX(standard_did_with_mpa); // standardized did estimators with mpa
+
+  DATA_MATRIX(standard_did_without_mpa); // standardized did estimators without mpa
+
+  DATA_MATRIX(standard_year_species); // standardized did estimators without mpa
+
+  DATA_MATRIX(standard_region_cluster); // STANDARD REGION CLUSTERS
+
+  /////////define parameters/////////
+
+  PARAMETER_VECTOR(seen_non_nested_betas); // NON DID BETAS
+
+  PARAMETER_VECTOR(seen_did_betas);  // DIFFERENCE IN DIFFERENCE BETAS
+
+  PARAMETER_VECTOR(seen_density_species_sigma);
+
+  PARAMETER_VECTOR(seen_year_species_betas);
+
+  PARAMETER_VECTOR(seen_year_species_sigmas);
+
+  PARAMETER_VECTOR(seen_region_cluster_betas);
+
[…]

### 2019-10-09 src/fit_zissou.cpp

@@ -1,0 +1,259 @@
+
+// TMB attempt at fitting ahnold
+#include <TMB.hpp>
+template<class Type>
+Type objective_function<Type>::operator() ()
+{
+  /////////load data/////////
+
+  // seen data
+
+  DATA_MATRIX(x_seen_non_nested); // non nested part of data
+
+  // DATA_MATRIX(x_seen_did); // difference in difference terms
+
+  DATA_MATRIX(x_seen_year_species); // year random effects
+
+  DATA_MATRIX(x_seen_region_cluster);
+
+  DATA_IVECTOR(year_species_index); // index the same rows as x_seen showing what species is in that row
+
+  DATA_IVECTOR(seen_species_index); // index the same rows as x_seen showing what species is in that row
+
+  DATA_IVECTOR(region_cluster_index); // index the same rows as x_seen showing what species is in that row
+
+  DATA_VECTOR(log_density); // observed log densities
+
+  // seeing data
+
+  DATA_MATRIX(x_seeing_non_nested); // non nested part of data
+
+  DATA_MATRIX(x_seeing_year_species); // year random effects
+
+  DATA_MATRIX(x_seeing_region_cluster); // REGION CLUSTERS
+
+  DATA_VECTOR(any_seen); // observed log densities
+
+
+
+  // standardized matrices data
+
+  DATA_MATRIX(standard_non_nested); // standardized non nested part of data
+
+  DATA_MATRIX(standard_year_species); // standardized did estimators without mpa
+
+  DATA_MATRIX(standard_region_cluster); // STANDARD REGION CLUSTERS
+
+  // did data
+
+  DATA_MATRIX(non_nested_did_data);
+
+  DATA_MATRIX(targeted_year_did_data);
+
+  DATA_MATRIX(nontargeted_year_did_data);
+
+  DATA_MATRIX(species_did_data);
+
+
+
+  /////////define parameters/////////
+
+  PARAMETER_VECTOR(seen_non_nested_betas); // NON DID BETAS
+
+  // PARAMETER_VECTOR(seen_did_betas);  // DIFFERENCE IN DIFFERENCE BETAS
+
+  PARAMETER_VECTOR(seen_density_species_sigma);
+
+  PARAMETER_VECTOR(seen_year_species_betas);
+
+  PARAMETER_VECTOR(seen_year_species_sigmas);
+
+  PARAMETER_VECTOR(seen_region_cluster_betas);
+
[…]

### 2019-10-10 documents/ovando-regional-effects-of-mpas-si.Rmd
initial submission

@@
     - \usepackage[left]{lineno}
+    - \usepackage{xcolor}
     - \linenumbers
@@
 
-The computing environment needed to reproduce the results presented here is stored in a Docker container available [here]()XX.
-Interested users can load the container on their system to reproduce the complete analysis and manuscript.
-Analysis were performed in `r R.version.string`.
+All code needed to reproduce our main results and manuscript can be found at https://github.com/DanOvando/regional-effects-of-mpas.
+A fully reproducible environment for running this analysis and compiling the manuscript will be made available through Code Ocean at https://codeocean.com/capsule/5233105/tree.
+All analysis were performed in `r R.version.string`.
 Package versions are shown in Table.S\@ref(tab:pkg-info).
@@
 The operating model is a spatial single-species age-structured bio-economic model.
-The operating model itself is organized as an R package, which can be found and installed [here](https://github.com/DanOvando/spasm).
-Users can explore the functionality of the operating mode through an interactive web application [here](https://danovando.shinyapps.io/simmpa/).
+The operating model itself is organized as an R package, which can be found and installed at https://github.com/DanOvando/spasm.
+Users can explore the functionality of the operating mode through an interactive web application at https://danovando.shinyapps.io/simmpa/.
 
@@
 
+At the rawest level, the data are counts of finfish in 2cm length bins along a 30m x 2m transect at various sites and depths.
+These length bins are converted to biomass, and then biomass densities, by converting length to weights using available allometric data and dividing by the transect area.
+Our goal is to estimate the effect of the MPAs on these biomass densities of fish throughout the Channel Islands.
+We fit this model using a hierarchical mixed-effect framework using Template Model Builder [TMB, @kristensen2016] in R.
+The model consists of three levels, the first (starting from the "bottom") being transect-level densities of fish species observed by PISCO, which are standardized into a standardized biomass abundance index, accounting for both probability of detection and expected density as a result of changes in both abundance and covariates such as visibility and observer experience [see @maunder2004].
+For the second stage, we break the abundance indices into targeted and non-targeted species (per the classifications in the PISCO data), and estimate the mean trend of each group (targeted and non-targeted) over time.
+In the third stage, we estimate the difference in the de-meaned trend between the targeted and non-targeted fishes (controlling for factors such as water temperature, kelp cover, and commercial fishery catches), that under the assumptions of the model reflects the causal effect of the MPAs on the outcome of interest (in this case regional biomass density of targeted fishes).
+All three of these steps are integrated into the same estimation model, in order to propagate uncertainty through the model correctly.
+Simpler versions of the estimation model, as well as a synthetic control identification strategy, were assessed as well and showed results consistent with the model in our main results.
+
 ### Data

### 2019-10-10 documents/ovando-regional-effects-of-mpas-si.tex
initial submission

@@
 \usepackage[left]{lineno}
+\usepackage{xcolor}
 \linenumbers
@@
 
-The computing environment needed to reproduce the results presented here is stored in a Docker container available \href{}{here}XX.
-Interested users can load the container on their system to reproduce the complete analysis and manuscript.
-Analysis were performed in R version 3.6.1 (2019-07-05).
+All code needed to reproduce our main results and manuscript can be found at \url{https://github.com/DanOvando/regional-effects-of-mpas}.
+A fully reproducible environment for running this analysis and compiling the manuscript will be made available through Code Ocean at \url{https://codeocean.com/capsule/5233105/tree}.
+All analysis were performed in R version 3.6.1 (2019-07-05).
 Package versions are shown in Table.S\ref{tab:pkg-info}.
@@
 The operating model is a spatial single-species age-structured bio-economic model.
-The operating model itself is organized as an R package, which can be found and installed \href{https://github.com/DanOvando/spasm}{here}.
-Users can explore the functionality of the operating mode through an interactive web application \href{https://danovando.shinyapps.io/simmpa/}{here}.
+The operating model itself is organized as an R package, which can be found and installed at \url{https://github.com/DanOvando/spasm}.
+Users can explore the functionality of the operating mode through an interactive web application at \url{https://danovando.shinyapps.io/simmpa/}.
 

### 2019-10-10 documents/ovando-regional-effects-of-mpas.Rmd
initial submission

@@
 All of these can be interpreted as a conflict of interest when evaluating fisheries policy.
-C.C., S.G., and O.D's research program includes funding from environmental NGOs and foundations with an interest in ocean conservation and food security
+C.C., S.G., and O.D's research program includes funding from environmental NGOs and foundations with an interest in ocean conservation and food security.
+Funding for D.O.
+was partly provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship.
 abstract: |
@@
 Our study pairs theory with empirical methods to examine the regional effects of MPAs on conservation outcomes.
-Using simulations, we show that the effects of marine protected areas can be highly variable and dependent on human drivers.
-Building off of MPA theory, we find no statistically clear effect of a network of MPAs, which our simulation analysis suggests is entirely reasonable.
+Using simulations, we show that the regional effects of marine protected areas can be highly variable and dependent on human drivers.
+We find no statistically clear effect of a network of MPAs after 13 years of protection, which our simulation analysis suggests is actually to be expected.
 Our results lay a foundation for future research on the design and monitoring of marine protected areas.
 acknowledgements: |
-  This study would not be possible without the work provided by PISCO divers over the years. DO thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments and technical support. 
+  Funding for the simulation model was provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship.
+This study would not be possible without the work provided by PISCO divers over the years.
+DO thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments and technical support.
 keywords:
@@
 header-includes:
-   - \usepackage{eso-pic,graphicx,transparent}
+   - \usepackage{eso-pic,graphicx,transparent,xcolor}
 output:
@@
 params:
-  run_name: ["v5.1"]
+  run_name: ["v5.0"]
   sim_years: [50]
@@
   num_patches: [25]
+linkcolor: blue
 ---
@@
 ```
-
-<!-- why is it going to be hard ot detect regional effects of MPAs few if any empirical studies at net regional effects.
-does this mean that i doesn't exist, or impossible tot est, why?
-three fundamental stumbling blocks -->
 
@@
 
-<!-- \AddToShipoutPictureFG{ -->
-<!--   \AtPageCenter{% or \AtTextCenter -->
-<!--     \makebox[0pt]{\rotatebox[origin=c]{45}{% -->
-<!--       \scalebox{5}{\texttransparent{0.3}{DRAFT}}% -->
-<!--     }} -->
-<!--   } -->
-<!-- } -->
-
 No-take Marine Protected Areas (MPAs), spatial regions of the ocean in which fishing is prohibited, have a long history in the management of marine resources.
@@
 Modern MPAs were first established primarily as marine analogs to the terrestrial protection of iconic landscapes like Yellowstone or Kruger National Parks [@iucn1976;@watson2014b].
-Over time our goals and expectations for MPAs have evolved; while all MPAs are expected to deliver conservation benefits within their borders, many modern MPAs are also established to  bolster fish populations throughout the region in which they are located (what we term “regional-scale effects”) [@gaines2010].
+Over time our goals and expectations for MPAs have evolved; while all MPAs are expected to deliver conservation benefits within their borders, many modern MPAs are also established to  bolster fish populations throughout the region in which they are located [@gaines2010].
 
@@
 
-<!-- **key thing: a lot more ways for small things to happen than big things ** -->
-
-<!-- xx need to add in discussion of other outcomes not covered here, e.g. resiliency, tourism, etc. xx -->
-
-Much of the literature on the effects of Marine Protected Areas focuses on assessing the conservation effects within the borders of protected areas [e.g.
+Much of the literature on the effects of MPAs focuses on assessing the conservation effects within the borders of protected areas [e.g.
 @edgar2014].
 While these within-MPA effects are vitally important for protecting rare species, biodiversity, critical habitats, and often tourism, they paint an incomplete picture of the overall population effects of MPAs.
-The organisms within the borders of protected areas are generally part of a broader population, the biological stock, connected through adult movement and larval dispersal.
+The organisms within the borders of protected areas are generally part of a broader biological stock, connected through adult movement and larval dispersal.
 If the goal of conservationists or natural resource managers is to increase the total abundance or productivity of a resource, a broader question we should ask of MPAs is not just are there more fish inside their borders, but also how have the reserves affected abundances throughout the region in which they are located?
@@
 We define the regional conservation effects of MPAs as the change in total biomass of fish (summing inside and outside of MPAs) relative to the total biomass of fish that would have occurred without the MPAs (acknowledging that other outcomes such as increased biodiversity or resiliency are also important to conservation but are beyond the scope of this analysis).
-Note that this definition of regional effects is in line with the recommendations of @ferraro2018, but differs from effects measured by for example differences in outcomes before and after MPAs.
 
 We used a simulation model to examine the regional conservation effects of MPAs under a wide array of conditions that might reasonably be expected to affect these outcomes.
-The model explores combinations of MPA designs (e.g., sizes, number, and placement of MPAs), life histories (e.g., growth and mortality rates, adult movement, larval dispersal, recruitment variability and autocorrelation, and timing of density dependence), and fishing fleet dynamics (e.g., degree of fishing pressure, size selectivity of the fleet, reaction of fishing pressure to MPA placement) (See Table.S1 for a complete description of simulation variables).
-We use this model to explore how different drivers interact to affect the conservation and fishery outcomes of MPAs.
-
-<!-- Numerous works have used simulation modeling to explore the effects of MPAs on outcomes such as fisheries yields and stock abundance (see @fulton2014 and references within),but these prior studies have either focused on the effect of a select set of drivers (such as dispersal rates) on outcomes, or have been highly localized assessments that incorporate multiple drivers but are tuned to one specific region.
-Our model captures critical bio-economic features, while still being flexible and fast enough to be generalized to a broad array of scenarios.
-We use this model to explore how different drivers interact to affect the conservation and fishery outcomes of MPAs.
--->
+The model explores combinations of MPA designs, life histories, and fishing fleet dynamics (See Table.S1 for a complete description of simulation variables).
+We use this model to explore how different drivers interact to affect the conservation outcomes of MPAs.
 
@@
 
-<!-- Across all simulations the median simulated equilibrium regional conservation effect, expressed as the percentage change in total biomass with the MPAs and without the MPAs, was `r round(100*median(eqo$mpa_effect))`%, with a minimum of less than `r round(100*min(eqo$mpa_effect))`% and a maximum of over `r round(100*max(eqo$mpa_effect))`%.
-The median time required for the simulated MPAs to cause a 10% increase in regional biomass was approximately 10 years.
--->
-
 We tested the effects of different combinations of these theoretical drivers of MPA conservation effects using over 10,000 simulated MPA scenarios.
@@
 However, when combined with other factors such as fleet dynamics, life history, and MPA size, the conservation effects of MPAs can vary widely in magnitude and timing, a finding corroborated by empirical evidence [@starr2015;@caselle2015;@nickols2019;@villasenor-derbez2019] (Fig.\ref{conservation-effects}-B).
-MPAs protecting small proportions of a stock can produce small positive effects, large positive effects, or even small negative effects.
-As the area protected inside MPAs increases, positive effects become more likely, but even MPAs protecting most of a population's footprint can produce conservation effects as low as 0% and as high as 200% and above depending on the context in which they are placed.
+As the area protected inside MPAs increases, positive effects become more likely, but even MPAs protecting most of a population's footprint can produce conservation effects as low as 0% and as high as 200%.
 The degree of depletion (i.e.
@@
 Instead of two parallel worlds, a similar experimental design would involve random placement of MPAs.
+
 Unfortunately, neither of these experimental designs has been implemented, because parallel worlds do not exist and MPAs are never placed at random.
 Careful causal inference requires some other means of controlling for biases introduced by factors such as the MPA siting process and biological spillover and concentration of fishing effort  [@ferraro2018;@larsen2019].
-
-<!-- From the fisheries perspective, catch rates before and after [e.g.
-@roberts2001a] or near and far from MPAs [e.g.
-@russ2004a] are sometimes presented as evidence for the fisheries effects of MPAs.
-However neither of these approaches can be counted on to provide reliable estimate of the net fishery effects of MPAs, due to failures to account for potential broader trends in catch rates independent of MPAs and the potentially confounding effects of effort redistribution.
-@smith2006a presents an econometrically robust method for estimating the fishery effects of MPAs, though this approach requires high-resolution data on comparable fishing effort both near and far from MPAs, data that is very rarely available.
-It is beyond the scope of this paper to delve deeper into detecting the net fishery effects of MPAs, and so we now focus on the question of estimating their regional conservation effects.
--->
 
@@
 
-<!-- However, even under this ideal example the response ratio is not a perfect estimator due to the fact that this ratio contains no information on the effect of the MPAs outside their borders.  -->
-
 What happens when the “control” sites are in fact affected by the application of the MPA through biological spillover and fishing concentration?
@@
 Under these circumstances the response ratio is a somewhat biased and very inaccurate estimator of the true regional effect of MPAs.
+
 While on average high response ratios correspond with high true regional effects, the variance is high, so the response ratio alone cannot be used as a reliable proxy for the true regional effect, without accounting for the sources of biases discussed here.
@@
 
-To the extent that we can select control sites that are both sufficiently similar to the region treated with MPAs and are unaffected by biological or economic spillover, response ratios may be reliable estimators of regional conservation effects.
+<!-- To the extent that we can select control sites that are both sufficiently similar to the region treated with MPAs and are unaffected by biological or economic spillover, response ratios may be reliable estimators of regional conservation effects.
 In reality these conditions are unlikely to hold but for a few select cases.
-This become especially difficult when networks of MPAs are implemented, such as has occurred in California.
-In this case, control sites are generally located within a few kilometers from MPA borders, making biological and economic spillover highly likely.
+This becomes especially difficult when networks of MPAs are implemented, such as has occurred in California.
+In this case, control sites are generally located within a few kilometers from MPA borders, making biological spillover and fishing concentration highly likely.
 In these circumstances, without robust statistical controls, response ratios may provide an unreliable estimate of the regional conservation effect of marine protected areas.
+-->
 
 ```{r dr-vs-mpa-effect-plot, include = FALSE}
+
+ylabs <- c(expression(""<= "-100%"), paste0(seq(0,300, by = 100),"%"),expression("">= "400%") )
 
@@
                       option = "cividis") + 
-  scale_x_continuous(labels = percent) + 
-    scale_y_continuous(labels = percent) + 
+  scale_x_continuous(labels = ylabs) + 
+    scale_y_continuous(labels = ylabs) + 
   theme(legend.position = "top") + 
@@
 
-pdf(file =  "figs/density-ratio.pdf",width = 8, height = 5, useDingbats = TRUE)
-print(dr_plot)
-dev.off()
+  ggsave(dr_plot,filename = "figs/density-ratio.png",
+         width = 8,
+         height = 5)
+
+# pdf(file =  "figs/density-ratio.png",width = 8, height = 5, useDingbats = TRUE)
+# print(dr_plot)
+# dev.off()
 
@@
   \centering
-  \includegraphics[width=.9\linewidth]{figs/density-ratio.pdf}
+  \includegraphics[width=.9\linewidth]{figs/density-ratio.png}
   \caption{Simulated MPA-size weighted response ratios (x-axis) plotted against true regional change in biomass caused by MPAs (y-axis).
@@
 ```
-
-<!-- Our strategy for identifying the effect of the MPAs in the Channel Islands is to use biomass densities of non-targeted species as our control for broader factors affecting biomass densities of fishes throughout the Islands besides the effect of the MPAs.
-Simply examining densities of tarted species throughout the islands before and after MPAs might confuse changes in regional densities that would have occurred anyway without the MPAs with MPA effects (since at the regional scale we do not have a spatial control for the effects of the MPAs).
-We assume that over the 13 years of MPA protection covered in this study non-targeted species were relatively unaffected by the MPAs (i.e.
-indirect effects of MPAs on species interactions are negligible), and can can serve as an indicator for the overall changes in fish abundance throughout the Channel Islands that would have occurred without the MPAs.
-Raw densities calculated by PISCO provide visual support for this hypothesis, in that both the targeted and non-targeted species groups share common trends (Fig.\ref{pop-trends}, see Materials and Methods for statistical support for this assumption).
--->
 
@@
 Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species).
-The key assumptions of this approach are that a) within the time-frame of the model there are no significant interaction effects between the targeted and non-targeted species (which in fact we do not detect, see SI section 1.6.3), and b) that in the absence of the MPAs both the targeted and non-targeted groups of species would have exhibited similar trends in densities (the parallel-trends assumption).
-The advantage to this approach is that given the time-frame of the model (15 years), we believe that spillover effects of MPA placement on the non-targeted species are likely to be much less severe than the effects of biological spillover and fishing concentration that may bias the performance of estimators such as a response ratios or a spatial BACI design.
+The key assumptions of this approach are that a) within the time-frame of the model there are no significant interaction effects between the targeted and non-targeted species (which in fact we do not detect, see SI), and b) that in the absence of the MPAs both the targeted and non-targeted groups of species would have exhibited similar trends in densities.
+The advantage to this approach is that given the time-frame of the model (16 years), we believe that spillover effects of MPA placement on the non-targeted species are likely to be much less severe than the effects of biological spillover and fishing concentration that may bias the performance of estimators such as a response ratios or a spatial BACI design.
 
@@
   \includegraphics[width=1\linewidth]{figs/pop-trends.pdf}
-  \caption{Centered and scaled mean annual biomass densities of included targeted (red) and non-targeted (blue) fish species (faded lines) and smoothed means with 95\% confidence intervals over time (darker lines and ribbons).}
+  \caption{Centered and scaled mean annual biomass densities of included targeted (red) and non-targeted (blue) fish species (faded lines) and smoothed means with 95\% confidence intervals over time (darker lines and ribbons).
+Vertical red dashed line shows year of MPA implementation.}
   \label{pop-trends}
@@
 However, these raw trends could be caused by any number of confounding variables besides the effects of the MPAs themselves, from changes in observer skill and sampling locations, to shifts in environmental conditions.
-Using our difference-in-difference model in an effort to control for these factors, we found no significant difference in the biomass densities of targeted and non-targeted fishes pre-MPA (before 2003), which provides support for the parallel trends assumption critical to the difference-in-difference method.
+Using our difference-in-difference model in an effort to control for these factors, we found no significant difference in the biomass densities of targeted and non-targeted fishes pre-MPA (before 2003).
 Following the implementation of the MPAs in 2003, we see evidence of an increasing trend in densities of targeted relative to non-targeted fish, reaching a peak value in 2014.
-Following 2014, densities of targeted species appear to have declined relative to what we would expect based on the non-targeted group (Fig.\ref{did-plot}).
+After 2014, densities of targeted species appear to have declined relative to what we would expect based on the non-targeted group (Fig.\ref{did-plot}).
 
@@
 MPAs are implemented in 2003 (red dashed line).
-Estimates are from a regression on log(biomass density), so estimated effects roughly correspond to percentage changes.
-Colored cells in the background indicate the frequency of MPA conservation effects predicted by our simulation modeling for MPAs similar in design and species composition to the Channel Islands.}
+Colored cells in the background indicate the frequency of MPA conservation effects predicted by our simulation modeling for MPAs similar in design and species composition to the Channel Islands in each year.}
   \label{did-plot}
@@
 ------
-
-<!-- Studies such as @dinerstein2019 call for 30% of the oceans to be places in protected areas.
-What does simulation exercise suggest that the effects of such a closure might be on overall population biomass densities?
-XX.
-Suppose we assumed that all fisheries are both open-access and overfished (a clearly worst case scenario as many fisheries are well managed).
-In that case, we see that...
--->
-
-<!-- <!-- zeros may be an artifact of "error in variables", which can bias estimates towards zero [@larsen2019] -->
 
@@
 
-<!-- Our simulation results present a clear outcome from the perspective of fishery effects: Under the assumptions of our model, MPAs should not be expected to provide improvements in fisheries catches unless the stock is experiencing substantial overfishing (Fig.\ref{catch-msy-effects}-B).
-This general result has been discussed before [@hilborn2004], but our results provide the most comprehensive simulation testing of this concept to date.
-MPAs placed in overfished fisheries frequently produced catch gains of 0-25%, though larger and smaller effects were still possible.
-MPA size also presents an interesting tradeoff from a fisheries perspective.
-MPAs that cover less than 25% of the population range, which is true for nearly all MPAs currently on the planet, very rarely produced substantial fishery losses, but also never produced substantial fishery gains.
-Simulated MPAs  that covered between 25% & 50% of species ranges produced the potential for higher fishery gains, but also opened up risk of greater fishery loses.
-For MPAs covering more than 50% of a species range, fishery losses quickly become much more likely than fishery gains.
--->
-
-<!-- We only measure one aspect of fishery outcomes here (catches).
-MPAs can affect many other outcomes of importance to fisheries, include profits (e.g.
-if institutions do not create derby fishing and MPAs improve catch per unit effort along the borders of reserves), resiliency, and research.
-Our model does not account for potentially important factors such as species interactions Overall, our fishery-specific results suggest a relatively simple rule of thumb: unless there is clear evidence of overfishing, MPAs should not be expected to produce improvements in fishery catches without clear explanation for the mechanism for these increases (e.g.
-improvements in habitat).
-In assessing the potential fishery benefits of MPAs, stakeholders must make explicit the mechanisms by which these benefits are expected to be produced (for example through increased catchability of fish near MPA borders), and what evidence exists for the existence of those mechanisms within the local context.
--->
-
 We find regional conservation gains from MPAs in `r percent(pos_runs, accuracy = 1)` of our simulations (Fig.\ref{conservation-effects}).
@@
 Large MPAs protecting highly depleted stocks almost always produced large regional conservation gains, as should be expected.
+
 The median conservation effect of `r round(100*median(eqo$mpa_effect))`\% stands in contrast to the large within-MPA effects on biomass densities reported by @halpern2003a and @lester2009.
@@
 
-<!-- While our results provide important insight on the expected fishery and conservation effects of MPAs, they also highlight the critical importance of explicitly modeling the links between the biological effects of MPAs and the ways that humans respond to them.
-One of our clearest findings was that MPAs can cause net conservation losses when implemented in a "constant-catch" fishery, as fishers must dramatically increase their fishing effort in the waters still available to them to maintain their desired landings.
-The focus of much of the MPA design and modeling literature has historically been on biological and ecological such as growth rates and connectivity [citation], though studies such as @hastings1999 and @sanchirico2006b incorporate assumptions about outside-MPA fleet dynamics as well.
-Our results make clear that economic and ecological drivers must be included in any attempt assess the effects of MPAs MPAs: in particular, we need to explicitly understand the the behavior of fishing fleets if we hope to estimate the conservation or fishery effects of MPAs [see @costa2013 for an example of this process].
-Large-scale empirical evidence confirms that predicting the effects of MPAs depends on understanding the human context in which they find themselves [@cinner2018].
--->
-
-<!-- The question of what is a "meaningful" conservation gain is of course context dependent and depends on the priorities and preferences of stakeholders.
-However, under any set of preferences and objectives managers and stakeholders would of course like to know whether policy interventions such as MPAs are having their intended effects.
-While the importance of say a 0-10% regional conservation gain will vary, there can be little doubt that an effect size within this range will be very difficult to detect in a noisy natural system (xx citations on noise and effect size xx), and our empirical assessment of MPA conservation effects in the Channel Islands demonstrates this.
--->
-
-<!-- .
-That we were unable to detect a significant divergence in the biomass densities of targeted species throughout the Channel Islands post MPA implementation, relative to the densities that the trends of the non-targeted species would suggest we should observe, should not come as a great surprise then.
--->
-
 We were not able to detect a statistically significant difference in the densities of targeted and non-targeted finfish species over the 13 years of MPA protection in the Channel Islands covered by our analysis.
@@
 
-However, our simulation analysis suggests that this result should perhaps be expected.
+Our simulation analysis suggests that this result should perhaps be expected.
 The Channel Islands MPAs cover approximately 20% of the waters in the Channel Islands, and while formal stock assessments are not available for many of the targeted species in our analysis, what evidence we have suggests that, as a group, these fish are on average not heavily overfished.
@@
 Many of the non-targeted species have warm thermal affinities and have increased in numbers since the heatwave [@freedman2019].
-However, the targeted group is made up mostly of fishes with cold-water affinities, such as members of the genus Sebastes.
+The targeted group is made up mostly of fishes with cold-water affinities, such as members of the genus *Sebastes*.
 As such, we hypothesize that the recent evidence for a decline in densities of targeted species is due to environmental conditions that disproportionately affect the targeted group (and not for example due to concentrated fishing pressure outside the reserves).
 This hypothesis is supported by the sharp declines in densities of targeted species seen inside the MPAs themselves (see Fig.S27-S29), suggesting a driver other than fishing may be at play.
-The Channel Islands case study demonstrates the challenge of detecting small effect sizes in nature and highlights the potential for small effect sizes to be overwhelmed by additional environmental shocks to the system.
-
-<!-- Our simulation model accounts for many factors that theory suggest should be critical drivers of MPA effects, namely adult and larval movement rates, nature of density dependence, and fleet dynamics.
-However, currently our model does not contain many other potentially important factors, such as species interactions.
-If stakeholders expect an effect that our model says is unlikely or impossible given local context, they should provide a clear hypothesis and evidence as to what factors not included in this model would cause their expected effect.
-For example, stakeholders might assert that larger conservation effects than predicted by our simulation model might occur in a location due to protection of critical spawning aggregation sites [@roberts2002;@erisman2015], but relevant evidence supporting this belief must be presented.
-It is also important to remember that our conservation results refer to the regional effects of MPAs both inside and outside of their borders.
-The buildup of biomass within the border of well enforced reserves is well documented [@lester2009], we show though that the regional effects may be much smaller and harder to detect than these within-MPA outcomes.
--->
 
@@
 
-<!-- Our results provide a unique survey of the likely regional effects of MPAs to fisheries and conservation, and places our empirical assessment of the regional effects of the Channel Islands MPAs in context.
-The Channel Islands are an intensely studied system, and the challenge of identifying a clear regional effect of the network of MPAs placed there in 2003 may seem surprising.
-However, our results show that in fact a smaller effect size, from the perspective of regional conservation gains, is to be expected in this system, and therefore the true effect will be very challenging to separate from environmental noise.
--->
-
-<!-- Modeling effort such as this can then help manager and stakeholders to set realistic expectations of regional effects for any given MPA network design, and importantly, help inform the design of monitoring programs.
-For example, monitoring may be targeted at the species and fleets that modeling suggests may provide the clearest indication of MPA-mediated effects.
-For cases where bio-economic modeling suggests small potential for MPA driven regional density effects, monitoring efforts can be targeted around detecting potential negative effects should they arise, i.e.
-evidence that the model is wrong, rather than exerting massive amounts of effort on what theory and modeling suggest may be a small and potentially undetectable effect size.
--->
-
-<!-- Policy making is inherently an exercise in utilizing best available theory, experience, and modeling to make decisions that are often extremely difficult to empirically test.
-Causal inference in coupled human and natural systems is incredibly challenging, and our results highlight the particular difficulties in estimating the causal effects of MPAs on regional conservation and fishery outcomes.
-Bio-economic simulation models such as ours can support the policy evaluation process by helping stakeholders and communities set realistic expectations for what MPAs might achieve, help managers understand the scale and magnitude of MPA effects, and help scientists design cost-effective monitoring plans that stand the best chance possible of detecting these effects.
--->
-
 As the number and size of global MPA networks increase, it is critical that we both set appropriate expectations for their outcomes, and plan how we will monitor the performance of these protected areas over time.
@@
 Commonly employed metrics such as response ratios may be applicable in some circumstances, but can have severe shortcomings as metrics of regional conservation effects.
-Dependence on unreliable estimators of MPA effects may lead to stakeholders incorrectly attributing negative environmental shocks as MPA failures, or interpreting data arising from scorched earth fishing outside MPAs as a conservation success.
+Dependence on unreliable estimators of MPA effects may lead to stakeholders incorrectly attributing negative environmental shocks as MPA failures, or interpreting data arising from scorched-earth fishing outside MPAs as a conservation success.
 Bio-economic modeling can help frame community expectations, reducing the potential for a reduction in support if unrealistic conservation or fishery gains are not realized.
@@
 Effective use of simulation and statistics to clearly communicate what we should expect and what we can detect from MPAs is critical in ensuring that MPAs play effective roles in fisheries management and marine conservation.
-
-<!-- Despite our best efforts we are unlikely to ever truly "know" the effect of our efforts to mitigate climate change, but must instead rely on comparisons to best available modeling outcomes to understand how effective our policies have been.
-Similarly, given the complexities of marine systems much of our decisions on MPA design will have to be based on effective modeling.
-While we are hardly the first to point out that bio-economic models are a critical tool for MPA design, our results help indicate a minimum floor of model complexity to provide candid assessments of regional MPA effects.
--->
-
-<!-- While factors such as MPA size and degree of depletion are especially strong drivers, for all but the most extreme cases of each a wide array of effects, from negative to highly positive,  are possible based on the complex interaction of factors such as fleet dynamics, movement rates, and recruitment timing.
-Confronting these interactions by considering the likely parameter space for a given region is a critical step then in understanding what likely regional effects of MPAs are.
-While models such as ours do require large numbers of parameters that may be challenging to obtain, our results show that working with communities to confront these uncertainties is preferable to sweeping them under the rug in favor of simpler models that are easier to parameter but miss details that our results show can have dramatic effect on expected outcomes.
--->
-
-<!-- We focus mostly on regional conservation gains in this paper.
-However, fisheries spillover is often another important factor to consider (i.e.
-are fisheries better off with the MPAs than they would have been without them).
-The fishery benefits of MPAs are just as (and likely more) intensely debated than the regional conservation outcomes [@hilborn2004; @roberts2001;@sala2018b;@hilborn2018].
-We only address fisheries affects briefly in this study, but our results highlight important tradeoffs and synergies between conservation and fishery spillover effects of MPAs.
-The good news from a fisheries perspective is also fairly obvious: Both the regional conservation and fishery benefits are expected to be greatest when the fished population is in an extremely depleted state pre-MPA, even over a 15 year time horizon, even for larger MPAs (though further work is needed to compare MPAs to alternative fisheries management strategies in these cases).
-For cases where a valuable and formerly abundant species is overfished, a large MPA may then provide large conservation and fishery gains for that species, while potentially having smaller impacts on other less depleted species.
-Our simulation results also do identify though a large parameter space where MPAs create tradeoffs between moderate conservation gains and moderate fishing losses.
-This type of projection analysis can help managers consider where in this space they may be.
-The most critical point with regards to conservation and fishery effects from our simulation analysis is that the conservation or fishery effects of MPAs cannot be reliably estimated without some knowledge and consideration of the dynamics of the fishing fleet outside the MPAs: over the short-term open access vs constant-catch dynamics can make the difference between a substantial conservation and fisheries win to a more depleted stock with more expensive fishing.
--->
-
-<!-- This process provides a unique survey of the likely regional effects of MPAs to fisheries and conservation, and places our empirical assessment of the regional effects of the Channel Islands MPAs in context.
-The Channel Islands are an intensely studied system, and the challenge of identifying a clear regional effect of the network of MPAs placed there in 2003 may seem surprising.
-However, our results show that in fact a smaller effect size, from the perspective of regional conservation gains, is to be expected in this system, and therefore the true effect will be very challenging to separate from environmental noise.
-The solution then though is not to give up on detecting effects, but rather to shift focus from identifying a specific effect size and instead use simulation analysis to appropriately set expectations for conservation and fishing stakeholders, and design monitoring programs around the species and situations that serve as effective indicators of the ability of an MPA network to achieve its objectives.
--->
 
@@
 Our main difference-in-difference model was fit using Template Model Builder in R [@kristensen2016].
-A completely reproducible version of our results can be accessed and run at XX.
+Code needed to replicate results can be found [here](https://github.com/DanOvando/regional-effects-of-mpas).
 
@@
 The model consists of 25 patches with wrapped edges (picture the waters around a circular island).
-For any one simulation we randomly pull a species and its associated life history (growth, mortality, maturity) from the `FishLife` [@thorson2017c] package in R.
-We pair these data with randomly selected values between 0.6 and 0.95 for Beverton-Holt steepness [as parameterized in @mace1994], as well as larval and adult dispersal rates (where at the lowest values adults and larvae stay within the patch they were created, and at the highest values they are capable of moving throughout the system in each time step).
-We also randomly assign whether adults have density dependent movement (meaning that adult biomass preferentially moves towards patches with lower adult biomass as opposed to random dispersal), as well as one of three potential types of density dependence [building off of @babcock2011]:
+For any one simulation we randomly pull a species and its associated life history from the `FishLife` [@thorson2017c] package in R.
+We pair these data with randomly selected values between 0.6 and 0.95 for Beverton-Holt steepness [as parameterized in @mace1994], as well as larval and adult dispersal rates.
+We also randomly assign whether adults have density dependent movement, as well as one of three potential types of density dependence [building off of @babcock2011]:
 
@@
 
-Each simulation is assigned a random set of fleet dynamics from one of three categories: constant-catch, constant harvest rate, or open-access.
-Under constant-catch, the fleet attempts to catch a randomly selected multiplier of maximum sustainable yield (MSY) each year (note that values of MSY greater than 1 will crash the population eventually, but allows for overfishing dynamics to be observed when shorter time periods are considered).
-Under a constant harvest rate the fleet captures a randomly selected fraction between 0.01\% and 98\% of the population in each time step.
-Under open access, fishing effort is allowed to expand and contract in response to the profitability of the fishery, where
-
-\begin{equation}
-  E_{t} = E_{t-1}\times\theta{PPUE_{t-1}}
-  \label{eqn:oa}
-\end{equation}
-
-and
-
-\begin{equation}
-  PPUE{t} = \frac{\sum_{i=1}^{I}pCatch_{i,t} - cE_{i,t}^\beta}{\sum_{i=1}^I E_{i,t}}
-  \label{eqn:profits}
-\end{equation}
-
-To shift the equilibrium of the open access model, we fix price *p* at 1 and $\beta$ at 2, and adjust the cost parameter to achieve the selected fishing mortality rate at bionomic equilibrium.
-These combinations of fleet models allow each simulation to achieve randomly selected levels of fishing pressure through different processes.
-
-<!-- While at equilibrium in the absence of an MPA each of these fleet models can achieve the same level of depletion, each of these fleet models interacts with MPAs in a different manner.  -->
-
 Along with the fleet dynamics model, each simulation is assigned a random fleet dispersal scenario: uniform dispersal (where the total effort of the fleet is divided evenly among all open patches), catch dispersal (where the total effort of the fleet is divided according to the catchable biomass in each available patch), and profit dispersal (where the total effort of the fleet is divided according to the profit per unit effort in each available patch).
@@
 Lastly each simulation is assigned an MPA scenario, defined by the number and size of MPAs, the placement of those MPAs, and the year that the MPAs are put in place.
-Many MPA models assume equilibrium conditions prior to the MPA, and then measure equilibrium outcomes.
-While these are important scenarios to understand, they do not reflect the reality of many MPAs, which are often placed in non-equilibrium conditions, and evaluated over the early years of their existence.
 Each simulation starts the population off at unfished equilibrium and then beings to apply the fleet model.
@@
 Fishing effort in displaced by MPAs can either concentrate outside or leave the fishery.
-
 Each simulation is run to equilibrium with and without the selected MPA strategy (holding all else constant).
 We then measure the difference in biomass and fishery catches in each time step in the scenario with and without the MPAs to calculate the conservation and fishery effects of the MPAs over time.
-We filter out any scenario that results in a collapse of the population (biomass less than 10% of unfished biomass) prior to the implementation of MPAs, and any scenario in which the open-access dynamics devolve into chaos.
 
@@
 Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment (*MPA*) on the treated (targeted species).
-
-<!-- The purpose of this regression approach is to control for unobserved time-invariant variables that are correlated with the treatment (MPA) and affect the outcome  (biomass densities).
-This approach depends on the assumption that the the untreated group (in this case non-targeted species) serve as a proxy for the trend in post-treatment densities in the absence of the MPAs (where the intercept terms for the targeted group allows for different mean densities).
--->
-
-We briefly assess two of the most critical assumptions of this model here: that the treated and non-treated groups have parallel trends, and that the effect of the treatment on the treated does not tangentially affect the untreated (i.e.
-that the implementation of MPAs does not indirectly affect the non-targeted group somehow).
-While the parallel trends assumption cannot be formally proven, we can examine its validity using the data from the years before the MPAs were put in place in 2003 (since after that we no longer expect the trends in the observed data to be parallel).
-We do not detect any significant differences in the trends of the biomass densities of the targeted and non-targeted species in the years before the MPAs, meaning that we do not have evidence to reject the assumption that densities of targeted and non-targeted species were following similar trajectories before the MPAs (though the model does not estimate a precise zero effect either).
+The full form of the estimation model is much more involved, integrating uncertainty in a hierarchical model from the individual observation level up to the mean trends of targeted and non-targeted species.
+See SI for details of estimation model.
+
+We briefly assess two of the most critical assumptions of this model here: that the treated and non-treated groups have parallel trends, and that the effect of the treatment on the treated does not tangentially affect the untreated.
+While the parallel trends assumption cannot be formally proven, we can examine its validity using the data from the years before the MPAs were put in place in 2003.
+We do not detect any significant differences in the trends of the biomass densities of the targeted and non-targeted species in the years before the MPAs.
 
 With regards to the second assumption, all of the species in this empirical analysis exist within an ecosystem, and as such affect each other through mechanisms such as predation, competition, and habitat modification.
-Conceivably then, protection of carnivorous targeted species inside MPAs could drive down the density of non-targeted prey species, serving in that case to positively bias our estimate of the effect of MPAs on the targeted species.
-While we know dynamics such as this have to exist on some level, we find it unlikely that these effects have had enough time to manifest in the 13 years of post-MPA data used in our analysis [@babcock2010; @pershing2015a].
+We find it unlikely that these effects have had enough time to manifest in a meaningful way in the 13 years of post-MPA data used in our analysis [@babcock2010; @pershing2015a].
 
 We used convergent cross mapping (CCM), in the manner of @clark2015, to test for the possibility of the trophic cascades biasing our results.
-Generalizations of Takens' theorem indicate that if two variables (in our case, species or physical variables) are part of the same dynamic system, their individual dynamics should reflect their relative causal influence.
-In other words, if one variable is causally forced by another, that forcing should leave a signature on the first time series.
+Generalizations of Takens' theorem indicate that if two variables are part of the same dynamic system, their individual dynamics should reflect their relative causal influence.
 Convergent cross mapping (CCM) tests for causation by using the attractor/manifold built from the time series of one variable to predict another (hence the "cross-mapping").
-In simple terms, the causal effect of A on B is determined by how well B cross-maps A.
 CCM then allows us to test for causal relationships in the timeseries of densities of targeted and non-targeted species.
@@
 
-While Equation.\ref{eq:did} presents the general form of our model, in practice the estimation model is much more involved.
-At the rawest level, the data are counts of finfish in 2cm length bins along a 30m x 2m transect at various sites and depths.
-These length bins are converted to biomass, and then biomass densities, by converting length to weights using available allometric data and dividing by the transect area.
-Our goal is to estimate the effect of the MPAs on these biomass densities of fish throughout the Channel Islands.
-We fit this model using a hierarchical mixed-effect framework using Template Model Builder [TMB, @kristensen2016] in R.
-The model consists of three levels, the first (starting from the "bottom") being transect-level densities of fish species observed by PISCO, which are standardized into a standardized biomass abundance index, accounting for both probability of detection and expected density as a result of changes in both abundance and covariates such as visibility and observer experience [see @maunder2004].
-For the second stage, we break the abundance indices into targeted and non-targeted species (per the classifications in the PISCO data), and estimate the mean trend of each group (targeted and non-targeted) over time.
-In the third step, we estimate the difference in the de-meaned trend between the targeted and non-targeted fishes (controlling for factors such as water temperature, kelp cover, and commercial fishery catches), that under the assumptions of the model reflects the causal effect of the MPAs on the outcome of interest (in this case regional biomass density of targeted fishes).
-All three of these steps are integrated into the same estimation model, in order to propagate uncertainty through the model correctly.
-Simpler versions of the estimation model, as well as a synthetic control identification strategy, were assessed as well and showed results consistent with the model in our main results (see SI).
-
 <!-- \showmatmethods -->

### 2019-10-10 documents/ovando-regional-effects-of-mpas.tex
initial submission

@@
 
-\usepackage{eso-pic,graphicx,transparent}
+\usepackage{eso-pic,graphicx,transparent,xcolor}
 
@@
 
-  \affil[a]{University of Washington, School of Aquatic and Fishery Science}
+  \affil[a]{University of Washington, School of Aquatic and Fishery Sciences}
   \affil[b]{University of California, Santa Barbara}
@@
 empirical methods to examine the regional effects of MPAs on
-conservation and fishery outcomes. Using simulations, we show that the
+conservation outcomes. Using simulations, we show that the regional
 effects of marine protected areas can be highly variable and dependent
-on human drivers. Building off of MPA theory, we find no statistically
-clear effect of a network of MPAs, which our simulation analysis
-suggests is entirely reasonable. Our results lay a foundation for future
-research on the design and monitoring of marine protected areas.}
+on human drivers. We find no statistically clear effect of a network of
+MPAs after 13 years of protection, which our simulation analysis
+suggests is actually to be expected. Our results lay a foundation for
+future research on the design and monitoring of marine protected areas.}
 
@@
 includes funding from environmental NGOs and foundations with an
-interest in ocean conservation and food security}
+interest in ocean conservation and food security. Funding for D.O. was
+partly provided by the NMFS-Sea Grant Population and Ecosystem Dynamics
+Fellowship.}
 
@@
 % Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
- \keywords{  Marine Protected Areas |  Conservation |  Bio-economic modeling |  Program Evaluation |  Channel Islands National Marine Sanctuary  }
+ \keywords{  Marine Protected Areas |  Conservation |  Bio-economic modeling |  Program Evaluation  }
 
@@
 
-\acknow{This study would not be possible without the work provided by PISCO
-divers over the years. DO thanks Cody Szuwalski, Julia Lawson, and André
-Punt for helpful comments and technical support.}
+\acknow{Funding for the simulation model was provided by the NMFS-Sea Grant
+Population and Ecosystem Dynamics Fellowship. This study would not be
+possible without the work provided by PISCO divers over the years. DO
+thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments
+and technical support.}
 
@@
 within their borders, many modern MPAs are also established to bolster
-fish populations throughout the region in which they are located (what
-we term ``regional-scale effects'') (4).
+fish populations throughout the region in which they are located (4).
 
@@
 
-Much of the literature on the effects of Marine Protected Areas focuses
-on assessing the conservation effects within the borders of protected
-areas (7). While these within-MPA effects are vitally important for
-protecting rare species, biodiversity, critical habitats, and often
-tourism, they paint an incomplete picture of the overall population
-effects of MPAs. The organisms within the borders of protected areas are
-generally part of a broader population, the biological stock, connected
-through adult movement and larval dispersal. If the goal of
-conservationists or natural resource managers is to increase the total
-abundance or productivity of a resource, a broader question we should
-ask of MPAs is not just are there more fish inside their borders, but
-also how have the reserves affected abundances throughout the region in
-which they are located? This logic, that MPAs will have conservation
-benefits for most species beyond their borders, is implicit in all
-multilateral calls for MPA expansion.
+Much of the literature on the effects of MPAs focuses on assessing the
+conservation effects within the borders of protected areas (7). While
+these within-MPA effects are vitally important for protecting rare
+species, biodiversity, critical habitats, and often tourism, they paint
+an incomplete picture of the overall population effects of MPAs. The
+organisms within the borders of protected areas are generally part of a
+broader biological stock, connected through adult movement and larval
+dispersal. If the goal of conservationists or natural resource managers
+is to increase the total abundance or productivity of a resource, a
+broader question we should ask of MPAs is not just are there more fish
+inside their borders, but also how have the reserves affected abundances
+throughout the region in which they are located? This logic, that MPAs
+will have conservation benefits for most species beyond their borders,
+is implicit in all multilateral calls for MPA expansion.
 
@@
 resiliency are also important to conservation but are beyond the scope
-of this analysis). Note that this definition of regional effects is in
-line with the recommendations of (22), but differs from effects measured
-by for example differences in outcomes before and after MPAs.
+of this analysis).
 
@@
 expected to affect these outcomes. The model explores combinations of
-MPA designs (e.g., sizes, number, and placement of MPAs), life histories
-(e.g., growth and mortality rates, adult movement, larval dispersal,
-recruitment variability and autocorrelation, and timing of density
-dependence), and fishing fleet dynamics (e.g., degree of fishing
-pressure, size selectivity of the fleet, reaction of fishing pressure to
-MPA placement) (See Table.S1 for a complete description of simulation
-variables). We use this model to explore how different drivers interact
-to affect the conservation and fishery outcomes of MPAs.
+MPA designs, life histories, and fishing fleet dynamics (See Table.S1
+for a complete description of simulation variables). We use this model
+to explore how different drivers interact to affect the conservation
+outcomes of MPAs.
 
@@
 affect the regional effects of MPAs. These include the scale of adult
-and larval dispersal relative to the size of the MPAs ({\textbf{???}},
-23--25), the strength and timing of density dependence in the population
-(e.g.~pre- or post-settlement), how overfished the population would be
-without the MPA, and how fishing activity and management responds to the
+and larval dispersal relative to the size of the MPAs (22--25), the
+strength and timing of density dependence in the population (e.g.~pre-
+or post-settlement), how overfished the population would be without the
+MPA, and how fishing activity and management responds to the
 implementation of the MPAs (4, 8, 26--32). In addition, even for the
@@
 in magnitude and timing, a finding corroborated by empirical evidence
-(37--40) (Fig.\ref{conservation-effects}-B). MPAs protecting small
-proportions of a stock can produce small positive effects, large
-positive effects, or even small negative effects. As the area protected
+(37--40) (Fig.\ref{conservation-effects}-B). As the area protected
 inside MPAs increases, positive effects become more likely, but even
 MPAs protecting most of a population's footprint can produce
-conservation effects as low as 0\% and as high as 200\% and above
-depending on the context in which they are placed. The degree of
+conservation effects as low as 0\% and as high as 200\%. The degree of
 depletion (i.e.~fishing pressure) has a clearer signal, with small
@@
 Instead of two parallel worlds, a similar experimental design would
-involve random placement of MPAs. Unfortunately, neither of these
-experimental designs has been implemented, because parallel worlds do
-not exist and MPAs are never placed at random. Careful causal inference
-requires some other means of controlling for biases introduced by
-factors such as the MPA siting process and biological spillover and
-concentration of fishing effort (22, 41).
+involve random placement of MPAs.
+
+Unfortunately, neither of these experimental designs has been
+implemented, because parallel worlds do not exist and MPAs are never
+placed at random. Careful causal inference requires some other means of
+controlling for biases introduced by factors such as the MPA siting
+process and biological spillover and concentration of fishing effort
+(41, 42).
 
@@
 control sites outside MPAs, which we will refer to as response ratios.
-(5) and (42) present meta-analyses of hundreds of such studies. These
+(5) and (43) present meta-analyses of hundreds of such studies. These
 results often find massively higher densities and biomasses inside MPAs
@@
 Control sites used in density or biomass ratios are often selected based
-on abiotic or ecological traits such as habitat characteristics (22).
+on abiotic or ecological traits such as habitat characteristics (41).
 However, selection of control sites is further complicated by the very
@@
 proxies for the ecological and economic context of the MPAs may be
-challenging. While these concerns have been stated previously (43), the
+challenging. While these concerns have been stated previously (44), the
 MPA evaluation literature has by and large been unable to adequately
 address them, often due to the very real challenges of acquiring the
-necessary data (22).
+necessary data (41).
 
 Failure to account for these spillover effects can result in a biased
-estimate of the true effect of a policy such as MPA placement (41). In
+estimate of the true effect of a policy such as MPA placement (42). In
 an ideal setting, the control site used in a response ratio would be a
@@
 response ratio is a somewhat biased and very inaccurate estimator of the
-true regional effect of MPAs. While on average high response ratios
-correspond with high true regional effects, the variance is high, so the
-response ratio alone cannot be used as a reliable proxy for the true
-regional effect, without accounting for the sources of biases discussed
-here. For higher movement rates, response ratios were frequently near
-0\% (potentially leading stakeholders to conclude that the MPA had been
-ineffective), when in fact in many cases the true effect of the MPA was
-highly positive. For highly sedentary species, extremely high response
-ratios could create the appearance of massive conservation gains when in
-fact the net effect on the regional population has been near zero or
-even negative (Fig.\ref{density-ratio}-B).
+true regional effect of MPAs.
+
+While on average high response ratios correspond with high true regional
+effects, the variance is high, so the response ratio alone cannot be
+used as a reliable proxy for the true regional effect, without
+accounting for the sources of biases discussed here. For higher movement
+rates, response ratios were frequently near 0\% (potentially leading
+stakeholders to conclude that the MPA had been ineffective), when in
+fact in many cases the true effect of the MPA was highly positive. For
+highly sedentary species, extremely high response ratios could create
+the appearance of massive conservation gains when in fact the net effect
+on the regional population has been near zero or even negative
+(Fig.\ref{density-ratio}-B).
 
@@
 
-To the extent that we can select control sites that are both
-sufficiently similar to the region treated with MPAs and are unaffected
-by biological or economic spillover, response ratios may be reliable
-estimators of regional conservation effects. In reality these conditions
-are unlikely to hold but for a few select cases. This become especially
-difficult when networks of MPAs are implemented, such as has occurred in
-California. In this case, control sites are generally located within a
-few kilometers from MPA borders, making biological and economic
-spillover highly likely. In these circumstances, without robust
-statistical controls, response ratios may provide an unreliable estimate
-of the regional conservation effect of marine protected areas.
-
 \begin{figure*}%[tbhp]
   \centering
-  \includegraphics[width=.9\linewidth]{figs/density-ratio.pdf}
+  \includegraphics[width=.9\linewidth]{figs/density-ratio.png}
   \caption{Simulated MPA-size weighted response ratios (x-axis) plotted against true regional change in biomass caused by MPAs (y-axis).
@@
 an MPA) from those not targeted before and after MPA implementation
-(41). Targeted species in the Northern Channel Islands, California,
+(42). Targeted species in the Northern Channel Islands, California,
 include commercially important finfish such as California sheephead
@@
 species was the subject of prior bio-economic modeling related to the
-effects of MPAs in southern California (44, 45). But, we must note that
+effects of MPAs in southern California (45, 46). But, we must note that
 the analysis omits species from important invertebrate fisheries
@@
 significant interaction effects between the targeted and non-targeted
-species (which in fact we do not detect, see SI section 1.6.3), and b)
-that in the absence of the MPAs both the targeted and non-targeted
-groups of species would have exhibited similar trends in densities (the
-parallel-trends assumption). The advantage to this approach is that
-given the time-frame of the model (15 years), we believe that spillover
-effects of MPA placement on the non-targeted species are likely to be
-much less severe than the effects of biological spillover and fishing
-concentration that may bias the performance of estimators such as a
-response ratios or a spatial BACI design.
+species (which in fact we do not detect, see SI), and b) that in the
+absence of the MPAs both the targeted and non-targeted groups of species
+would have exhibited similar trends in densities. The advantage to this
+approach is that given the time-frame of the model (16 years), we
+believe that spillover effects of MPA placement on the non-targeted
+species are likely to be much less severe than the effects of biological
+spillover and fishing concentration that may bias the performance of
+estimators such as a response ratios or a spatial BACI design.
 
@@
 place in 2003 as part of the California Marine Life Protection Act
-(MLPA) (see (46), (47), (48), and (49) for information on the creation
+(MLPA) (see (47), (48), (49), and (50) for information on the creation
 of the MLPA). PISCO conducts visual SCUBA surveys at a large number of
@@
 difference in the biomass densities of targeted and non-targeted fishes
-pre-MPA (before 2003), which provides support for the parallel trends
-assumption critical to the difference-in-difference method. Following
-the implementation of the MPAs in 2003, we see evidence of an increasing
-trend in densities of targeted relative to non-targeted fish, reaching a
-peak value in 2014. Following 2014, densities of targeted species appear
-to have declined relative to what we would expect based on the
-non-targeted group (Fig.\ref{did-plot}).
+pre-MPA (before 2003). Following the implementation of the MPAs in 2003,
+we see evidence of an increasing trend in densities of targeted relative
+to non-targeted fish, reaching a peak value in 2014. After 2014,
+densities of targeted species appear to have declined relative to what
+we would expect based on the non-targeted group (Fig.\ref{did-plot}).
 
@@
 always produced large regional conservation gains, as should be
-expected. The median conservation effect of 36\% stands in contrast to
-the large within-MPA effects on biomass densities reported by (50) and
-(5). This difference is due in part to the fact that within-MPA effects
-are likely to be much larger than regional-MPA effects, but may also
-reflect the many challenges of translating response ratios into regional
+expected.
+
+The median conservation effect of 36\% stands in contrast to the large
+within-MPA effects on biomass densities reported by (51) and (5). This
+difference is due in part to the fact that within-MPA effects are likely
+to be much larger than regional-MPA effects, but may also reflect the
+many challenges of translating response ratios into regional
 conservation effects. Our simulation analysis shows that without proper
@@
 
-However, our simulation analysis suggests that this result should
-perhaps be expected. The Channel Islands MPAs cover approximately 20\%
-of the waters in the Channel Islands, and while formal stock assessments
-are not available for many of the targeted species in our analysis, what
+Our simulation analysis suggests that this result should perhaps be
+expected. The Channel Islands MPAs cover approximately 20\% of the
+waters in the Channel Islands, and while formal stock assessments are
+not available for many of the targeted species in our analysis, what
 evidence we have suggests that, as a group, these fish are on average
 not heavily overfished. Some species, such as California sheephead and
-blue rockfish were likely below target levels during the period (51,
-52), but projections based upon the overall average response across all
+blue rockfish were likely below target levels during the period (52,
+53), but projections based upon the overall average response across all
 species will likely suggest modest benefits even if a subset of species
@@
 beginning in 2014 and persisting through 2016, resulting in part in
-extremely elevated water temperatures throughout the region (53).
+extremely elevated water temperatures throughout the region (54).
 Biogeographic differences in the distributions of targeted and
@@
 the non-targeted species have warm thermal affinities and have increased
-in numbers since the heatwave (54). However, the targeted group is made
-up mostly of fishes with cold-water affinities, such as members of the
-genus Sebastes. As such, we hypothesize that the recent evidence for a
+in numbers since the heatwave (55). The targeted group is made up mostly
+of fishes with cold-water affinities, such as members of the genus
+\emph{Sebastes}. As such, we hypothesize that the recent evidence for a
 decline in densities of targeted species is due to environmental
@@
 targeted species seen inside the MPAs themselves (see Fig.S27-S29),
-suggesting a driver other than fishing may be at play. The Channel
-Islands case study demonstrates the challenge of detecting small effect
-sizes in nature and highlights the potential for small effect sizes to
-be overwhelmed by additional environmental shocks to the system.
+suggesting a driver other than fishing may be at play.
 
@@
 are commonly used as evidence for conservation outcomes of MPAs (5), but
-as suggested in (22) and (41) and demonstrated here, without careful
+as suggested in (41) and (42) and demonstrated here, without careful
 attention to the design of control sites (accounting for example for the
 displacement of effort by MPAs), response ratios may be highly
-unreliable estimators of regional MPA effects. As (41) suggests, there
+unreliable estimators of regional MPA effects. As (42) suggests, there
 are many potential alternatives for estimating the effects of MPAs that
@@
 attributing negative environmental shocks as MPA failures, or
-interpreting data arising from scorched earth fishing outside MPAs as a
+interpreting data arising from scorched-earth fishing outside MPAs as a
 conservation success. Bio-economic modeling can help frame community
@@
 regression approach. Further details can be found in the Supplementary
-Information. All analysis were conducted in R (55). Our main
+Information. All analysis were conducted in R (56). Our main
 difference-in-difference model was fit using Template Model Builder in R
-(56). A completely reproducible version of our results can be accessed
-and run at XX.
+(57). Code needed to replicate results can be found
+\href{https://github.com/DanOvando/regional-effects-of-mpas}{here}.
 
@@
 island). For any one simulation we randomly pull a species and its
-associated life history (growth, mortality, maturity) from the
-\texttt{FishLife} (57) package in R. We pair these data with randomly
-selected values between 0.6 and 0.95 for Beverton-Holt steepness (58),
-as well as larval and adult dispersal rates (where at the lowest values
-adults and larvae stay within the patch they were created, and at the
-highest values they are capable of moving throughout the system in each
-time step). We also randomly assign whether adults have density
-dependent movement (meaning that adult biomass preferentially moves
-towards patches with lower adult biomass as opposed to random
-dispersal), as well as one of three potential types of density
-dependence (59):
+associated life history from the \texttt{FishLife} (58) package in R. We
+pair these data with randomly selected values between 0.6 and 0.95 for
+Beverton-Holt steepness (59), as well as larval and adult dispersal
+rates. We also randomly assign whether adults have density dependent
+movement, as well as one of three potential types of density dependence
+(60):
 
@@
 
-Each simulation is assigned a random set of fleet dynamics from one of
-three categories: constant-catch, constant harvest rate, or open-access.
-Under constant-catch, the fleet attempts to catch a randomly selected
-multiplier of maximum sustainable yield (MSY) each year (note that
-values of MSY greater than 1 will crash the population eventually, but
-allows for overfishing dynamics to be observed when shorter time periods
-are considered). Under a constant harvest rate the fleet captures a
-randomly selected fraction between 0.01\% and 98\% of the population in
-each time step. Under open access, fishing effort is allowed to expand
-and contract in response to the profitability of the fishery, where
-
-\begin{equation}
-  E_{t} = E_{t-1}\times\theta{PPUE_{t-1}}
-  \label{eqn:oa}
-\end{equation}
-
-and
-
-\begin{equation}
-  PPUE{t} = \frac{\sum_{i=1}^{I}pCatch_{i,t} - cE_{i,t}^\beta}{\sum_{i=1}^I E_{i,t}}
-  \label{eqn:profits}
-\end{equation}
-
-To shift the equilibrium of the open access model, we fix price \emph{p}
-at 1 and \(\beta\) at 2, and adjust the cost parameter to achieve the
-selected fishing mortality rate at bionomic equilibrium. These
-combinations of fleet models allow each simulation to achieve randomly
-selected levels of fishing pressure through different processes.
-
 Along with the fleet dynamics model, each simulation is assigned a
@@
 number and size of MPAs, the placement of those MPAs, and the year that
-the MPAs are put in place. Many MPA models assume equilibrium conditions
-prior to the MPA, and then measure equilibrium outcomes. While these are
-important scenarios to understand, they do not reflect the reality of
-many MPAs, which are often placed in non-equilibrium conditions, and
-evaluated over the early years of their existence. Each simulation
-starts the population off at unfished equilibrium and then beings to
-apply the fleet model. The MPAs are then placed during the randomly
-selected start year, allowing some runs to explore how the early
-dynamics of the MPA play out when the fishery and population they are
-placed on is not already at equilibrium. Fishing effort in displaced by
-MPAs can either concentrate outside or leave the fishery.
-
-Each simulation is run to equilibrium with and without the selected MPA
-strategy (holding all else constant). We then measure the difference in
-biomass and fishery catches in each time step in the scenario with and
-without the MPAs to calculate the conservation and fishery effects of
-the MPAs over time. We filter out any scenario that results in a
-collapse of the population (biomass less than 10\% of unfished biomass)
-prior to the implementation of MPAs, and any scenario in which the
-open-access dynamics devolve into chaos.
+the MPAs are put in place. Each simulation starts the population off at
+unfished equilibrium and then beings to apply the fleet model. The MPAs
+are then placed during the randomly selected start year, allowing some
+runs to explore how the early dynamics of the MPA play out when the
+fishery and population they are placed on is not already at equilibrium.
+Fishing effort in displaced by MPAs can either concentrate outside or
+leave the fishery. Each simulation is run to equilibrium with and
+without the selected MPA strategy (holding all else constant). We then
+measure the difference in biomass and fishery catches in each time step
+in the scenario with and without the MPAs to calculate the conservation
+and fishery effects of the MPAs over time.
 
@@
 the causal effect of the treatment (\emph{MPA}) on the treated (targeted
-species).
+species). The full form of the estimation model is much more involved,
+integrating uncertainty in a hierarchical model from the individual
+observation level up to the mean trends of targeted and non-targeted
+species. See SI for details of estimation model.
 
@@
 that the effect of the treatment on the treated does not tangentially
-affect the untreated (i.e.~that the implementation of MPAs does not
-indirectly affect the non-targeted group somehow). While the parallel
-trends assumption cannot be formally proven, we can examine its validity
-using the data from the years before the MPAs were put in place in 2003
-(since after that we no longer expect the trends in the observed data to
-be parallel). We do not detect any significant differences in the trends
-of the biomass densities of the targeted and non-targeted species in the
-years before the MPAs, meaning that we do not have evidence to reject
-the assumption that densities of targeted and non-targeted species were
-following similar trajectories before the MPAs (though the model does
-not estimate a precise zero effect either).
+affect the untreated. While the parallel trends assumption cannot be
+formally proven, we can examine its validity using the data from the
+years before the MPAs were put in place in 2003. We do not detect any
+significant differences in the trends of the biomass densities of the
+targeted and non-targeted species in the years before the MPAs.
 
@@
 other through mechanisms such as predation, competition, and habitat
-modification. Conceivably then, protection of carnivorous targeted
-species inside MPAs could drive down the density of non-targeted prey
-species, serving in that case to positively bias our estimate of the
-effect of MPAs on the targeted species. While we know dynamics such as
-this have to exist on some level, we find it unlikely that these effects
-have had enough time to manifest in the 13 years of post-MPA data used
-in our analysis (60, 61).
-
-We used convergent cross mapping (CCM), in the manner of (62), to test
+modification. We find it unlikely that these effects have had enough
+time to manifest in a meaningful way in the 13 years of post-MPA data
+used in our analysis (61, 62).
+
+We used convergent cross mapping (CCM), in the manner of (63), to test
 for the possibility of the trophic cascades biasing our results.
-Generalizations of Takens' theorem indicate that if two variables (in
-our case, species or physical variables) are part of the same dynamic
-system, their individual dynamics should reflect their relative causal
-influence. In other words, if one variable is causally forced by
-another, that forcing should leave a signature on the first time series.
-Convergent cross mapping (CCM) tests for causation by using the
-attractor/manifold built from the time series of one variable to predict
-another (hence the ``cross-mapping''). In simple terms, the causal
-effect of A on B is determined by how well B cross-maps A. CCM then
-allows us to test for causal relationships in the timeseries of
+Generalizations of Takens' theorem indicate that if two variables are
+part of the same dynamic system, their individual dynamics should
+reflect their relative causal influence. Convergent cross mapping (CCM)
+tests for causation by using the attractor/manifold built from the time
+series of one variable to predict another (hence the ``cross-mapping'').
+CCM then allows us to test for causal relationships in the timeseries of
 densities of targeted and non-targeted species. Our results found no
@@
 
-While Equation.\ref{eq:did} presents the general form of our model, in
-practice the estimation model is much more involved. At the rawest
-level, the data are counts of finfish in 2cm length bins along a 30m x
-2m transect at various sites and depths. These length bins are converted
-to biomass, and then biomass densities, by converting length to weights
-using available allometric data and dividing by the transect area. Our
-goal is to estimate the effect of the MPAs on these biomass densities of
-fish throughout the Channel Islands. We fit this model using a
-hierarchical mixed-effect framework using Template Model Builder (56) in
-R. The model consists of three levels, the first (starting from the
-``bottom'') being transect-level densities of fish species observed by
-PISCO, which are standardized into a standardized biomass abundance
-index, accounting for both probability of detection and expected density
-as a result of changes in both abundance and covariates such as
-visibility and observer experience (63). For the second stage, we break
-the abundance indices into targeted and non-targeted species (per the
-classifications in the PISCO data), and estimate the mean trend of each
-group (targeted and non-targeted) over time. In the third step, we
-estimate the difference in the de-meaned trend between the targeted and
-non-targeted fishes (controlling for factors such as water temperature,
-kelp cover, and commercial fishery catches), that under the assumptions
-of the model reflects the causal effect of the MPAs on the outcome of
-interest (in this case regional biomass density of targeted fishes). All
-three of these steps are integrated into the same estimation model, in
-order to propagate uncertainty through the model correctly. Simpler
-versions of the estimation model, as well as a synthetic control
-identification strategy, were assessed as well and showed results
-consistent with the model in our main results (see SI).
-
 \showacknow
@@
 
-\leavevmode\hypertarget{ref-ferraro2018}{}%
-22. Ferraro PJ, Sanchirico JN, Smith MD (2018) Causal inference in
-coupled human and natural systems. \emph{Proceedings of the National
-Academy of Sciences}:201805563.
-
 \leavevmode\hypertarget{ref-gaines2003}{}%
-23. Gaines SD, Gaylord B, Largier JL (2003) Avoiding Current Oversights
+22. Gaines SD, Gaylord B, Largier JL (2003) Avoiding Current Oversights
 in Marine Reserve Design. \emph{Ecological Applications} 13(sp1):32--46.
@@
 \leavevmode\hypertarget{ref-botsford2008}{}%
-24. Botsford LW, et al. (2008) Connectivity, sustainability, and yield:
+23. Botsford LW, et al. (2008) Connectivity, sustainability, and yield:
 Bridging the gap between conventional fisheries management and marine
@@
 \leavevmode\hypertarget{ref-difranco2018}{}%
-25. Di Franco A, et al. (2018) Linking home ranges to protected area
+24. Di Franco A, et al. (2018) Linking home ranges to protected area
 size: The case study of the Mediterranean Sea. \emph{Biological
 Conservation} 221:175--181.
+
+\leavevmode\hypertarget{ref-mcgilliard2008}{}%
+25. McGilliard CR, Hilborn R (2008) Modeling no-take marine reserves in
+regulated fisheries: Assessing the role of larval dispersal.
+\emph{Canadian Journal of Fisheries and Aquatic Sciences}
+65(11):2509--2523.
 
@@
 
+\leavevmode\hypertarget{ref-ferraro2018}{}%
+41. Ferraro PJ, Sanchirico JN, Smith MD (2018) Causal inference in
+coupled human and natural systems. \emph{Proceedings of the National
+Academy of Sciences}:201805563.
+
 \leavevmode\hypertarget{ref-larsen2019}{}%
-41. Larsen AE, Meng K, Kendall BE (2019) Causal analysis in
+42. Larsen AE, Meng K, Kendall BE (2019) Causal analysis in
 controlImpact ecological studies with observational data. \emph{Methods
@@
 \leavevmode\hypertarget{ref-halpern2002}{}%
-42. Halpern BS, Warner RR (2002) Marine reserves have rapid and lasting
+43. Halpern BS, Warner RR (2002) Marine reserves have rapid and lasting
 effects. \emph{Ecology Letters} 5(3):361--366.
@@
 \leavevmode\hypertarget{ref-halpern2004}{}%
-43. Halpern BS, Gaines SD, Warner RR (2004) Confounding effects of the
+44. Halpern BS, Gaines SD, Warner RR (2004) Confounding effects of the
 export of production and the displacement of fishing effort from marine
@@
 \leavevmode\hypertarget{ref-costello2010d}{}%
-44. Costello C, et al. (2010) The value of spatial information in MPA
+45. Costello C, et al. (2010) The value of spatial information in MPA
 network design. \emph{Proceedings of the National Academy of Sciences}
@@
 \leavevmode\hypertarget{ref-rassweiler2014}{}%
-45. Rassweiler A, Costello C, Hilborn R, Siegel DA (2014) Integrating
+46. Rassweiler A, Costello C, Hilborn R, Siegel DA (2014) Integrating
 scientific guidance into marine spatial planning. \emph{Proceedings of
@@
 \leavevmode\hypertarget{ref-osmond2010}{}%
-46. Osmond M, Airame S, Caldwell M, Day J (2010) ``Lessons for marine
+47. Osmond M, Airame S, Caldwell M, Day J (2010) ``Lessons for marine
 conservation planning: A comparison of three marine protected area
@@
 \leavevmode\hypertarget{ref-kirlin2013}{}%
-47. Kirlin J, et al. (2013) California's Marine Life Protection Act
+48. Kirlin J, et al. (2013) California's Marine Life Protection Act
 Initiative: Supporting implementation of legislation establishing a
@@
 \leavevmode\hypertarget{ref-botsford2014}{}%
-48. Botsford LW, White JW, Carr MH, Caselle JE (2014) Chapter Six -
+49. Botsford LW, White JW, Carr MH, Caselle JE (2014) Chapter Six -
 Marine Protected Area Networks in California, USA. \emph{Advances in
@@
 \leavevmode\hypertarget{ref-hilborn2012}{}%
-49. Hilborn R (2012) The role of science in MPA establishment in
+50. Hilborn R (2012) The role of science in MPA establishment in
 California: A personal perspective. \emph{Environmental Conservation}
@@
 \leavevmode\hypertarget{ref-halpern2003a}{}%
-50. Halpern BS (2003) The impact of marine reserves: Do reserves work
+51. Halpern BS (2003) The impact of marine reserves: Do reserves work
 and does reserve size matter? \emph{Ecological Applications}
@@
 \leavevmode\hypertarget{ref-alonzo2004}{}%
-51. Alonzo S, Key M, Ish T, MacCall AD (2004) \emph{Status of the
+52. Alonzo S, Key M, Ish T, MacCall AD (2004) \emph{Status of the
 California Sheephead Stock for 2004}.
@@
 \leavevmode\hypertarget{ref-dick2017}{}%
-52. Dick EJ, et al. (2017) \emph{The Combined Status of Blue and Deacon
+53. Dick EJ, et al. (2017) \emph{The Combined Status of Blue and Deacon
 Rockfishes in U.S. Waters off California and Oregon in 2017} (Pacific
@@
 \leavevmode\hypertarget{ref-gentemann2017}{}%
-53. Gentemann CL, Fewings MR, García-Reyes M (2017) Satellite sea
+54. Gentemann CL, Fewings MR, García-Reyes M (2017) Satellite sea
 surface temperatures along the West Coast of the United States during
@@
 \leavevmode\hypertarget{ref-freedman2019}{}%
-54. Freedman R (2019) Understanding the Efficacy of Spatial Management
+55. Freedman R (2019) Understanding the Efficacy of Spatial Management
 on Emerging Threats. PhD thesis (University of California, Santa
@@
 \leavevmode\hypertarget{ref-rcoreteam2019}{}%
-55. R Core Team (2019) R: A Language and Environment for Statistical
+56. R Core Team (2019) R: A Language and Environment for Statistical
 Computing.
@@
 \leavevmode\hypertarget{ref-kristensen2016}{}%
-56. Kristensen K, Nielsen A, Berg CW, Skaug H, Bell BM (2016)
+57. Kristensen K, Nielsen A, Berg CW, Skaug H, Bell BM (2016)
 \textbf{TMB} : Automatic Differentiation and Laplace Approximation.
@@
 \leavevmode\hypertarget{ref-thorson2017c}{}%
-57. Thorson JT, Munch SB, Cope JM, Gao J (2017) Predicting life history
+58. Thorson JT, Munch SB, Cope JM, Gao J (2017) Predicting life history
 parameters for all fishes worldwide. \emph{Ecological
@@
 \leavevmode\hypertarget{ref-mace1994}{}%
-58. Mace PM (1994) Relationships between Common Biological Reference
+59. Mace PM (1994) Relationships between Common Biological Reference
 Points Used as Thresholds and Targets of Fisheries Management
@@
 \leavevmode\hypertarget{ref-babcock2011}{}%
-59. Babcock EA, MacCall AD (2011) How useful is the ratio of fish
+60. Babcock EA, MacCall AD (2011) How useful is the ratio of fish
 density outside versus inside no-take marine reserves as a metric for
@@
 \leavevmode\hypertarget{ref-babcock2010}{}%
-60. Babcock RC, et al. (2010) Decadal trends in marine reserves reveal
+61. Babcock RC, et al. (2010) Decadal trends in marine reserves reveal
 differential rates of change in direct and indirect effects.
@@
 \leavevmode\hypertarget{ref-pershing2015a}{}%
-61. Pershing AJ, et al. (2015) Slow adaptation in the face of rapid
+62. Pershing AJ, et al. (2015) Slow adaptation in the face of rapid
 warming leads to collapse of the Gulf of Maine cod fishery.
@@
 \leavevmode\hypertarget{ref-clark2015}{}%
-62. Clark AT, et al. (2015) Spatial convergent cross mapping to detect
+63. Clark AT, et al. (2015) Spatial convergent cross mapping to detect
 causal relationships from short time series. \emph{Ecology}
@@
 
-\leavevmode\hypertarget{ref-maunder2004}{}%
-63. Maunder MN, Langley AD (2004) Integrating the standardization of
-catch-per-unit-of-effort into stock assessment models: Testing a
-population dynamics model and using multiple data types. \emph{Fisheries
-Research} 70(23):389--395.
-
 % Bibliography

### 2020-01-16 documents/ovando-regional-effects-of-mpas-si.Rmd
passing to squid

@@
 base_run <- model_runs %>%
-filter(var_names == "pisco_a", mpa_only == FALSE, center_scale == TRUE)
+filter(var_names == "pisco_a", data_to_use == "all", center_scale == TRUE)
 
@@
          var_names == "pisco_a",
-         mpa_only == TRUE)
+         data_to_use == "mpa_only")
 

### 2020-01-16 documents/ovando-regional-effects-of-mpas.Rmd
passing to squid

@@
 \correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: danovan@uw.edu}
+
+<!-- \AddToShipoutPictureFG{ -->
+<!--   \AtPageCenter{% or \AtTextCenter -->
+<!--     \makebox[0pt]{\rotatebox[origin=c]{45}{% -->
+<!--       \scalebox{5}{\texttransparent{0.3}{DRAFT}}% -->
+<!--     }} -->
+<!--   } -->
+<!-- } -->
 

### 2020-01-16 functions/fit-zissou.R

@@ -21,7 +21,7 @@ fit_zissou <- function(data,
                        include_intercept = T,
                        center_scale = T,
                        fixed_regions = F,
-                       mpa_only = F,
+                       data_to_use = 'all',
                        bin_years = FALSE) {
 
   # data <- data %>%
@@ -36,12 +36,16 @@ fit_zissou <- function(data,
   }
 
 
-  if (mpa_only == T){
+  if (data_to_use == "mpa_only"){
 
     data <- data %>%
-      filter(eventual_mpa == T)
+      filter(eventual_mpa == TRUE)
   }
 
+  if (data_to_use == "fished_only"){
+    data <- data %>%
+      filter(eventual_mpa == FALSE)
+  }
 
   if (center_scale == T){
 

### 2020-01-16 make-regional-mpas.R

@@ -119,8 +119,6 @@ channel_islands_only <- T # only include channel islands, leave T
 
 min_year <- 1999 # included years must be greater than this
 
-mpa_only <- FALSE
-
 occurance_ranking_cutoff <- 0.5
 
 small_num <-  0 # no idea
@@ -1121,17 +1119,16 @@ model_runs <- cross_df(
   list(
     data_source = abundance_data$data_source,
     var_names = var_options$var_names,
-    mpa_only = c(TRUE,FALSE),
+    data_to_use = c("all","mpa_only", "fished_only"),
     center_scale = c(TRUE)
   )
 ) %>%
   left_join(var_options, by = "var_names") %>%
   left_join(abundance_data, by = "data_source") %>%
   filter(!(data_source == "kfm" & str_detect(var_names,"pisco"))) %>%
-  filter(!(data_source == "kfm" & mpa_only == TRUE))
+  filter(!(data_source == "kfm" & (data_to_use %in% c("mpa_only","fished_only"))))
 
 
-# model_runs <- model_runs %>% filter(data_source == "kfm")
 if (run_tmb == T){
 
   if (file.exists("fit-progress.txt")){
@@ -1155,7 +1152,7 @@ if (run_tmb == T){
 
     fits <- sfz(data = model_runs$data[[i]],
                 non_nested_variables = model_runs$non_nested_variables[[i]],
-                mpa_only = model_runs$mpa_only[[i]],
+                data_to_use = model_runs$data_to_use[[i]],
                 center_scale = model_runs$center_scale[[i]],
                 run_dir = run_dir,
                 script_name = script_name,
@@ -1980,12 +1977,12 @@ if (process_results == TRUE){
     mutate(did_plot = map(processed_fits, "did_plot"))
 
   base_run <- model_runs %>%
-    filter(var_names == "pisco_a", mpa_only == FALSE, center_scale == TRUE)
+    filter(var_names == "pisco_a", data_to_use == "all", center_scale == TRUE)
 
   mpa_run <- model_runs %>%
     filter(data_source == "pisco",
            var_names == "pisco_a",
-           mpa_only == TRUE)
+           data_to_use == "mpa_only")
 
   kfm_run <- model_runs %>%
     filter(data_source == "kfm")

### 2020-02-20 documents/regional-mpa-notebook.Rmd
passing to squid post empirical density ratio estimation

---
title: "regional-mpa-notebook"
author: "Dan"
date: "2/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

### 2020-02-20 functions/process-zissou-fit.R

@@ -1,0 +1,87 @@
+process_zissou_fit <- function(fit){
+
+seen_non_nested_betas <- fit$zissou_estimates %>%
+  filter(stringr::str_detect(variable, "seen_non_nested_betas")) %>%
+  rename(group = variable) %>%
+  mutate(variable  = fit$zissou_data$x_seen_non_nested %>% colnames())
+
+seeing_non_nested_betas <- fit$zissou_estimates %>%
+  filter(stringr::str_detect(variable, "seeing_non_nested_betas")) %>%
+  rename(group = variable) %>%
+  mutate(variable  = fit$zissou_data$x_seen_non_nested %>% colnames())
+
+seen_year_species_betas <- fit$zissou_estimates %>%
+  filter(stringr::str_detect(variable, "seen_year_species_betas")) %>%
+  rename(group = variable) %>%
+  mutate(variable = fit$zissou_data$x_seen_year_species %>% colnames())
+
+seeing_year_species_betas <- fit$zissou_estimates %>%
+  filter(stringr::str_detect(variable, "seeing_year_species_betas")) %>%
+  # filter(variable == "seeing_year_species_betas") %>%
+  rename(group = variable) %>%
+  mutate(variable = fit$zissou_data$x_seen_year_species %>% colnames())
+
+seen_region_cluster_betas <- fit$zissou_estimates %>%
+  filter(stringr::str_detect(variable, "seen_region_cluster_betas")) %>%
+  # filter(variable == "seen_region_cluster_betas") %>%
+  rename(group = variable) %>%
+  mutate(variable = fit$zissou_data$x_seen_region_cluster %>% colnames())
+
+seeing_region_cluster_betas <- fit$zissou_estimates %>%
+  filter(stringr::str_detect(variable, "seeing_region_cluster_betas")) %>%
+  # filter(variable == "seeing_region_cluster_betas") %>%
+  rename(group = variable) %>%
+  mutate(variable = fit$zissou_data$x_seen_region_cluster %>% colnames())
+
+did_betas <- fit$zissou_estimates %>%
+  filter(stringr::str_detect(variable, "mpa_effect")) %>%
+  # filter(variable == "mpa_effect") %>%
+  mutate(group = variable) %>%
+  mutate(year = fit$did_data$year %>% unique())
+
+
+
+non_targeted_abundance_betas <- fit$zissou_estimates %>%
[…]

### 2020-02-20 make-regional-mpas.R

@@ -49,13 +49,13 @@ run_description <- "PNAS R and R run with MPA and fishery-only runs "
 # So, once you've run simulate_mpas, you can set it to FALSE and validata_mpas will work
 
 
-run_did <- TRUE # run difference in difference on data from the CINMS
+run_did <- FALSE # run difference in difference on data from the CINMS
 
 simulate_mpas <- FALSE # simulate MPA outcomes
 
 validate_mpas <- FALSE
 
-process_results <- FALSE
+process_results <- TRUE
 
 knit_paper <- FALSE
 
@@ -773,8 +773,8 @@ pisco_data <- pisco_data %>%
            as.factor() %>% as.numeric()) %>%
   group_by(sampling_event,
            classcode) %>%
+  mutate(total_biomass_g = sum(total_biomass_g)) %>% 
   mutate(
-    total_biomass_g = sum(total_biomass_g),
     density_g_m2 = total_biomass_g / 60,
     total_count = sum(count),
     mean_length = mean(fish_tl, na.rm = T),
@@ -1993,6 +1993,12 @@ if (process_results == TRUE){
            var_names == "pisco_a",
            data_to_use == "mpa_only")
 
+  
+  fished_run <- model_runs %>%
+    filter(data_source == "pisco",
+           var_names == "pisco_a",
+           data_to_use == "fished_only")
+  
   kfm_run <- model_runs %>%
     filter(data_source == "kfm")
 
@@ -2168,6 +2174,246 @@ if (process_results == TRUE){
 
   processed_grid$adult_movement <- (2 * processed_grid$adult_movement) / num_patches
 
+  # calculate response ratios
+  
+  years <- unique(fitted_data$year)
+  
+  years <- zissou_fit$did_data$year %>% unique()
+  
+  
+  outside_abundance <-
+    tibble(
+      abundance_hat =   fished_run$tmb_fit[[1]]$zissou_report$abundance_hat,
+      classcode = rep(unique(fitted_data$classcode), each = length(years))
+    )  %>% 
+    left_join(life_history_data %>% select(classcode, targeted), by = "classcode") %>% 
+    mutate(mpa = FALSE) %>% 
+    group_by(classcode) %>%
+    mutate(year = years) %>%
+    ungroup()
+  
+  inside_abundance <-
+    tibble(
+      abundance_hat =   mpa_run$tmb_fit[[1]]$zissou_report$abundance_hat,
[…]

### 2020-02-20 test-vast.R

@@ -1,0 +1,205 @@
+library(tidyverse)
+library(TMB)
+#library(VAST)
+library(sdmTMB)
+
+
+# assume loading in data already
+
+pisco_abundance_data <-
+  abundance_data$data[abundance_data$data_source == "pisco"][[1]]
+
+
+
+empirical_biomass_densities  <- pisco_abundance_data %>%
+  filter(classcode %in% unique(fitted_data$classcode)) %>%
+  group_by(year, site_side, region, zone, eventual_mpa, classcode, targeted) %>%
+  summarise(total_classcode_density = sum(exp(log_density)),
+            mean_experience  = mean(cumulative_n_obs),
+            mean_vis = mean(mean_vis),
+            mean_depth = mean(mean_depth),
+            mean_canopy = mean(mean_canopy)) %>%
+  group_by(year, site_side, classcode) %>%
+  summarise(
+    total_biomass_density = sum(total_classcode_density),
+    mean_biomass_density = mean(total_classcode_density),
+    mean_experience  = mean(mean_experience),
+    mean_vis = mean(mean_vis),
+    mean_depth = mean(mean_depth),
+    mean_canopy = mean(mean_canopy)
+  ) %>%
+  ungroup()
+
+empirical_biomass_densities %>% ggplot(aes(mean_biomass_density)) + geom_histogram() +
+  facet_wrap( ~ classcode, scales = "free_x") + 
+  theme_minimal()
+
+
+classcodes <- unique(empirical_biomass_densities$classcode)
+
+
+standardizer <- function(temp_classscode, pknots = 0.5){
+
+temp <- empirical_biomass_densities %>%
+  filter(classcode == temp_classscode) %>%
+  # select(year, site_side, mean_biomass_density) %>%
+  left_join(
+    site_data %>% select(site, side, lon_wgs84, lat_wgs84) %>% unite(col = site_side, site:side, sep = '-'),
+    by = c("site_side")
+  ) %>%
+  rename(X = lon_wgs84,
+         Y = lat_wgs84) %>% 
+  mutate(mean_experience = scale(mean_experience)) %>% 
+  mutate(mean_experience_2 = mean_experience^2,
+         mean_vis = scale(mean_vis))
+
+temp %>%
+  ggplot(aes(X, Y, color = mean_biomass_density)) +
+  geom_point() +
[…]

### 2020-03-11 documents/cover-letter.Rmd
back to squid

---
title: "regional-mpa-cover-letter"
author: "Dan"
date: "2/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##

Dear Dr. Ovando,

I regret to inform you that the PNAS Editorial Board has rejected your manuscript [MS# 2019-17437]. The expert who served as editor obtained 2 reviews, which are included below. After careful consideration, the editor decided that we cannot accept your manuscript.

However, because the reviewers think the work is of interest and the editor concurs, we are willing to consider a resubmission that constructively addresses all of the concerns raised in the critiques. The paper would have to satisfy both the reviewers and the editor, and new criticisms could arise upon re-evaluation. We cannot guarantee success and will be unable to consider further resubmissions.

Please submit the resubmission as you would a new paper. Your resubmission will be given a new tracking number. In your cover letter and the Comments to Editorial Staff, please specify that the paper is a resubmission and include your original tracking number. Additionally, please provide a "resubmission/revision" cover letter, including a point-by-point response to the editor and reviewers' comments. If you have any questions, please contact our office.

Thank you for submitting your work to PNAS.

Sincerely yours,
May R. Berenbaum
Editor-in-Chief

Editor's Remarks to Author:

This is a fascinating manuscript because it examines an iconic MPA, for which there may be more detailed biological and physical data than any other MPA.

The simulation model that is applied to the case study is also rich in detail.

One referee, who is a world-renowned fisheries scientist with quantitative chops and deep knowledge, challenges the novelty and significance of the paper, holding it to a "high PNAS-bar", while also recognizing that as a case study it needs to be published. I hold this referee in high regard, and upon a close reading of the manuscript myself, I had to agree that when I was leading a science group at NOAA fifteen years ago, which is when MPA's were a hot new topic, it was widely appreciated that the outcome of any MPA would be dramatically impacted by where displaced vessels go and movement rates of adult fish.

In order for this to be published in PNAS, you need to be very clear on the manuscript's novel contribution, and also place it better in the salient literature. How relevant are references 33 and 35 compared to the references provided by referee #2?

The one difference I have with referee #2 is that I am not sure to what extent the richness or many, many different species, of before and after data you have for the Channel Islands MPA has ever been matched in other MPA study? In other words, there is some chance the empirical data may be what are most novel - although those data do not come to the forefront in this paper.

It could be that this warrants two papers rather than one. I would have liked to seen a paper focusing on the results of the difference in differences analyses. For maximum impact these would be presented in a naïve form because that is illustrative to practitioners-as it is "what they expect", and in the more rigorous form after accounting for the absence of any true control site and modeling the fact that an MPA affects the so-called, but non-existent, control. The naïve results are worthy of attention. If the empirical data were elevated, then the Supplementary materials need to include more information on the sampling, spatial and temporal detail, etc. As it is, I have no idea how much confidence to put in the fish-by-fish biomass data.

Then the great merit of the model is it should tell the manager what they should expect, and what they need to do if they want to make sure any newly instituted MPA is performing as hoped for. There are several marine conservation NGO's who are operating with totally naïve and misguided expectations about what will be evident after an MPA is established. Written up well, this research could mitigate some dangerous delusions. To do this the model would be used to discuss:

What would you expect to see even if an MPA was working well?

What would you expect to see if it an MPA was ineffective and doomed to be ineffective?

What does all this mean for the expectations we put on MPA's?

All of the above questions are touched on in this manuscript -- but not direct enough for a manager or NGO to take note of them. Lines 642 to 651 are noteworthy, but are kind of lost in their placement.

Reviewer Comments:
Reviewer #1:

Suitable Quality?: Yes
Sufficient General Interest?: Yes
Conclusions Justified?: Yes
Clearly Written?: Yes
Procedures Described?: Yes
Supplemental Material Warranted?: Yes

Comments on Significance Statement:

I think it could use some clarification/specificity. Since fish biomass was the only 'conservation outcome' analyzed I think that should be stated in the significance statement as well as abstract. Further I think an additional word in the case study sentence would be helpful to demonstrate that was the empirical part of the study, additional words in caps: "We find no statistically clear effect of ONE EXISTING network of MPAs in FISH BIOMASS after 13 years of protection, which our simulation analysis suggests is actually to be expected."

Comments:

- I think that greater emphasis needs to be placed up front, abstract, statement of significance, introduction..perhaps even title, that this model and paper only addresses fish biomass...not other 'conservation outcomes'. You do define this but not until page 2, line 148-154. I recommend moving this up.

- I was surprised that there was no discussion about enforcement being a potential confounding factor. Currently the paper assumes that the Channel Islands MPAs are fully enforced, without poaching. This may or may not be the case and should be stated as such. You do address well other system wide impacts like water temperature (lines 557-8, 691+). Perhaps enforcement could be added to the discussion around line 711-13.

- Your discussion of the model also emphasizes that stock status prior to MPA designation is very key to outcomes; in the SI most of the species included for the Channel Island review have unknown status---which influences your discussion, and may better be described.

- There is extra punctuation in the footnote around line 117-118.

Reviewer #2:

Suitable Quality?: No
Sufficient General Interest?: No
Conclusions Justified?: Yes
Clearly Written?: Yes
Procedures Described?: Yes
Supplemental Material Warranted?: Yes

Comments:

Measuring the performance of marine protected areas is an important area of research. The authors undertake a rigorous analysis of one set of reserves in Southern California. The case study is interesting and robustly done, but it is not clear what general lessons from the modeling and empirical application are novel. Overall, the authors need to do a better job explaining why this paper makes a contribution worthy of PNAS.

Some text from the paper: "The potentially more important question, however, is not whether spillover occurs (it must to some degree in any realistic scenario), but what the net effects of spillover are and whether those effects are empirically detectable. From a fishery perspective, are spillover benefits sufficient to offset losses in fishing grounds and changes in responses of displaced fishers caused by an MPA? From a conservation perspective, how much does the buildup of fish inside an MPA increase biomass outside the protected area? Overall, what are the regional effects of MPAs?"
...
"To address this gap, this study examines two critical questions: 1) What do we expect the regional-scale conservation effects of MPAs to be and 2) When (and how) can we expect to empirically detect these effects?"

Yes, these are important questions but they are not new to ecology, economics, and fishery science. You are not the first to ask them and point out their importance; see articles at bottom for some examples. What are you doing to answer them that is better than all of the other papers that come before you? And why is your case study an exemplary place to answer the questions? The conclusions on line 709-728 are well-known in the MPA literature.

The idea that structural modeling / simulation model has an important role to play to understand empirical signals from data is not novel in general and specifically with respect to marine reserves, see included citations. While you do a lot of perturbations of the single species model, there is really nothing new in the modeling and results. I would also be more cautious regarding your conclusion about your model predicts what should be expected in the data analysis. It could be simply luck that there is agreement between the two.

What about using the simulation model to demonstrate the validity of the technique used to identify the MPA effect used in the paper, which was pioneered in publication in 38? Or to understand whether the implication of the statement on line 539, that is, species ranges exceed the Island and your sampling domain, is important for measuring the impacts of MPAs outside of their boundaries? In sum, in its current form, I would move all of the simulation model to the SI (lines 131-252) and mention in the discussion section that these results are consistent with the predictions of a stylized single species model.

The discussion on how to detect changes outside the MPA (lines 254-347) follows very closely the discussion in Ferraro et al (#41 in your list). It seems like you can remove most of this text and point the reader to that paper.

To me the potential contribution of the paper starts on line 350, which is way too late in a paper of this size.

Is part of assumption b) (line 372) that the dispersal characteristics of the fished and non-fished species need to be similar? I could envision a scenario where the two sets of species have different dispersal characteristics, non-fished have smaller dispersal ranges than fished species, for example. How does this impact your identification strategy? If the trends are similar regardless of the mechanistic process by which it occurs, does it matter for the empirics? More discussion on both assumptions (a) and (b) and what they imply about the biological mechanisms seems warranted.

The use of visual data surveys has, of course, problems that introduce errors in dependent and independent variables that could be correlated with features of the fished and non-fished species, e.g., size distribution. How can you be sure that this is not driving your results? Could you use the simulation model to investigate this?

The answer to your question on line 709 that appears on lines 714-719 has been known in the literature on MPAs for over 20 years; see, e.g., papers by Martin Smith below. Given that your question is what "your results imply about the future of marine science", my answer is not very much. Again, you did a rigorous empirical case study that should be published. I am however finding it difficult to see a path to publication in PNAS.

Examples of relevant literature, you should expand your net to find other relevant literature to better distinguish your contribution:

White, J. Wilson, et al. "Decision analysis for designing marine protected areas for multiple species with uncertain fishery status." Ecological Applications 20.6 (2010): 1523-1541.

White, J. W., Botsford, L. W., Baskett, M. L., Barnett, L. A., Barr, R. J. and Hastings, A. (2011), Linking models with monitoring data for assessing performance of no‐take marine reserves. Frontiers in Ecology and the Environment, 9: 390-399. doi:10.1890/100138

Burgess, S. C., Nickols, K. J., Griesemer, C. D., Barnett, L. A., Dedrick, A. G., Satterthwaite, E. V., Yamane, L. , Morgan, S. G., White, J. W. and Botsford, L. W. (2014), Beyond connectivity: how empirical methods can quantify population persistence to improve marine protected‐area design. Ecological Applications, 24: 257-270. doi:10.1890/13-0710.1

Smith, M. D., & Wilen, J. E. (2003). Economic impacts of marine reserves: the importance of spatial behavior. Journal of Environmental Economics and Management, 46(2), 183-206. doi:https://doi.org/10.1016/S0095-0696(03)00024-X

Smith, M. D., Zhang, J., & Coleman, F. C. (2006). Effectiveness of marine reserves for large-scale fisheries management. Canadian Journal of Fisheries and Aquatic Sciences, 63(1), 153-164. doi:10.1139/f05-205

Kaplan, K. A., L. Yamane, L. W. Botsford, M. L. Baskett, A. Hastings, S. Worden, and J. W. White. 2019. Setting expected timelines of fished population recovery for the adaptive management of a marine protected area network. Ecological Applications 29(6):e01949. 10.1002/eap.1949

### 2020-03-11 documents/ovando-regional-effects-of-mpas-v2.Rmd
back to squid

--- 2020-01-16 documents/ovando-regional-effects-of-mpas.Rmd
@@
 It is often clear that MPAs produce conservation benefits inside their borders, but many MPAs are also justified on the grounds that they also benefit the broader region outside their borders.
-Despite the expanding role of MPAs in marine resource management, we lack a clear understanding of what the regional effects of MPAs might be, and how we might detect them in scientific studies.
-We used a bio-economic simulation model to demonstrate how the regional effects of MPAs are affected by a suite of economic and environmental drivers and show how commonly used methods to measure inside-MPA effects are likely to fail to capture the regional impacts of protection.
-We present an empirical strategy for estimating regional MPA effects on biomass density using 16 years of data from inside and outside of MPAs within the Channel Islands, California, USA.
-We find no statistically clear effect of MPAs on aggregate mean biomass densities of targeted finfish throughout the study region.
-To make meaningful progress on MPA design and evaluation, the next generation of MPA science must address the challenge of understanding and estimating the regional effects of MPAs across the diversity of species in the ecosystem.
+The conservation effects of MPAs are most commonly evaluated using response ratios---measures of biomass densities inside their borders, relative to biomass densities in reference sites outside their borders.
+Studies of this nature have provided broad evidence that MPAs produce conservation benefits within their borders.
+While these gains can be vitally important, marine populations are rarely contained entirely within the borders of MPAs; therefore a critical question is not only do MPAs produce conservation gains inside their borders, but how do they affect the broader region in which they are located?
+The Channel Islands National Marine Sanctuary provides a clear example of this challenge.
+A network of MPAs covering roughly 20% of the Islands’ waters was put in place in 2003, with a goal of providing regional conservation and fishery benefits.
+Response ratios from the region indicate that the Channel Island MPAs have increased biomass densities inside the MPAs.
+However, we are unable to find a clear effect of these same MPAs at the regional scale across multiple species.Building off of existing theory, we use a bio-economic simulation model to explain this discrepancy, and demonstrate under what conditions we likely can and cannot clearly estimate the regional conservation effects of MPAs.
+We show that MPA networks covering 25% or less of a region are likely to produce regional increases in fish biomass on the order of 10%-20%, a meaningful effect but also one that can easily be overwhelmed by environmental shocks.
+Our results provide a novel assessment of the regional effects of a large and iconic Marine Protected Area network, and provide guidance for managers charged with monitoring and adapting MPAs.
 significance: |
@@
 params:
-  run_name: ["v5.0"]
+  run_name: ["v6.0"]
   sim_years: [50]
@@
 
-load(file = here::here("results",run_name, "model_runs.Rdata"))
+model_runs <- read_rds( file.path(run_dir,"did_fits.rds"))
+
+# load(file = here::here("results",run_name, "model_runs.Rdata"))
 
@@
 
-models_worked <- model_runs$tmb_fit %>% map("error") %>% map_lgl(is_null)
-
-model_runs <- model_runs %>%
-  filter(models_worked) %>%
-  mutate(tmb_fit = map(tmb_fit,"result")) %>%
-  mutate(processed_fits = map(tmb_fit, process_fits)) %>%
-  mutate(did_plot = map(processed_fits, "did_plot"))
+# models_worked <- model_runs$tmb_fit %>% map("error") %>% map_lgl(is_null)
+
+# model_runs <- model_runs %>%
+#   filter(models_worked) %>%
+#   mutate(tmb_fit = map(tmb_fit,"result")) %>%
+#   mutate(processed_fits = map(tmb_fit, process_fits)) %>%
+#   mutate(did_plot = map(processed_fits, "did_plot"))
 
@@
 
+fishsed_run <- model_runs %>%
+  filter(data_source == "pisco",
+         var_names == "pisco_a",
+         data_to_use == "fished_only")
+
 kfm_run <- model_runs %>%
@@
 
-fitted_data <- base_run$data[[1]]
-
-zissou_fit <- base_run$tmb_fit[[1]]
-
-report <- base_run$tmb_fit[[1]]
+fitted_data <- abundance_data$data[abundance_data$data_source == "pisco"][[1]]
 
@@
 
-seen_non_nested_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "seen_non_nested_betas")) %>%
-  rename(group = variable) %>%
-  mutate(variable  = zissou_fit$zissou_data$x_seen_non_nested %>% colnames())
-
-seeing_non_nested_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "seeing_non_nested_betas")) %>%
-  rename(group = variable) %>%
-  mutate(variable  = zissou_fit$zissou_data$x_seen_non_nested %>% colnames())
-
-seen_year_species_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "seen_year_species_betas")) %>%
-  rename(group = variable) %>%
-  mutate(variable = zissou_fit$zissou_data$x_seen_year_species %>% colnames())
-
-seeing_year_species_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "seeing_year_species_betas")) %>%
-  # filter(variable == "seeing_year_species_betas") %>%
-  rename(group = variable) %>%
-  mutate(variable = zissou_fit$zissou_data$x_seen_year_species %>% colnames())
-
-seen_region_cluster_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "seen_region_cluster_betas")) %>%
-  # filter(variable == "seen_region_cluster_betas") %>%
-  rename(group = variable) %>%
-  mutate(variable = zissou_fit$zissou_data$x_seen_region_cluster %>% colnames())
-
-seeing_region_cluster_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "seeing_region_cluster_betas")) %>%
-  # filter(variable == "seeing_region_cluster_betas") %>%
-  rename(group = variable) %>%
-  mutate(variable = zissou_fit$zissou_data$x_seen_region_cluster %>% colnames())
-
-did_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "mpa_effect")) %>%
-  # filter(variable == "mpa_effect") %>%
-  mutate(group = variable) %>%
-  mutate(year = zissou_fit$did_data$year %>% unique())
-
- betas <- bind_rows(
-   seen_non_nested_betas,
-   seeing_non_nested_betas,
-   seen_year_species_betas,
-   seeing_year_species_betas,
-   seen_region_cluster_betas,
-   seeing_region_cluster_betas,
-   did_betas %>% select(-year)
- ) %>%
-   as_data_frame()
-
- non_nested_beta_plot <- betas %>%
-   filter(str_detect(group, "non_nested")) %>%
-   ggplot() +
-   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
-   geom_pointrange(aes(x = variable,
-                       y = estimate,
-                       ymin = lower,
-                       ymax = upper)) +
-   facet_wrap(~group) +
-   coord_flip() +
-   theme(axis.text.y = element_text(size = 10))
-
- year_species_effects_plots <- betas %>%
-   filter(str_detect(group, "year_species_betas")) %>%
-   mutate(year = str_replace_all(variable,"\\D","") %>% as.numeric()) %>%
-   mutate(classcode = str_split(variable,'-', simplify = T)[,2]) %>%
-   ggplot() +
-   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
-   geom_ribbon(aes(x = year, ymin = lower, ymax = upper, fill = classcode), alpha = 0.25) +
-   geom_line(aes(x = year, y = estimate, color = classcode)) +
-   facet_wrap(~group)
-
-  region_cluster_plots <- betas %>%
-   filter(str_detect(group, "region_cluster_betas")) %>%
-   mutate(cluster = str_replace_all(variable,"\\D","")) %>%
-   mutate(region = str_split(variable,'-', simplify = T)[,3]) %>%
-   ggplot() +
-    geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
-    geom_pointrange(aes(x = region,
-                        y = estimate,
-                        ymin = lower,
-                        ymax = upper, color = cluster)) +
-    facet_wrap(~group)
-
-top_species <- base_run$data[[1]]$classcode %>% unique()
+top_species <- fitted_data$classcode %>% unique()
 
@@
 
-  sim_plot <- outcomes %>%
-    mutate(rsize = plyr::round_any(mpa_size, 0.25),
-           reff = plyr::round_any(mpa_effect,0.2)) %>%
-    filter(rsize == 0.25, depletion < 0.8, depletion > 0.2,
-           scientific_name %in% sim_life_history$scientific_name
-           ) %>%
-    mutate(year = years_protected + 2001) %>%
-    filter(year %in% did_betas$year, years_protected >1) %>% 
-    group_by(year, reff) %>% 
-    count() %>% 
-    group_by(year) %>% 
-    mutate(sc = n / sum(n))
-
-  did_plot <- did_betas %>%
-    ggplot() +
-    geom_tile(data = sim_plot, aes(year, reff, fill = sc), alpha = 0.75) +
-    geom_vline(aes(xintercept = 2003), color = 'red', linetype = 2, size = 2) +
-    geom_hline(aes(yintercept = 0)) +
-    geom_pointrange(aes(
-      year,
-      y = exp(estimate) - 1,
-      ymin = exp(lower) - 1,
-      ymax = exp(upper) - 1
-    ),
-    size = 1.5
-    ) +
-    labs(x = "Year", y = "Targeted Trend Divergence") +
-    scale_fill_viridis(guide = wide_hgc,
-                       name = "% of Annual Sims",
-                       labels = scales::percent_format(accuracy = 1),
-                       option = "magma") + 
-    theme(legend.position = "top",
-        panel.grid.major.x = element_blank(),
-        panel.grid.minor.x = element_blank()
-        )
-
-cip_data <- pisco_data %>%
-  left_join(site_data, by = c("site","side")) %>%
-  filter(is.na(eventual_mpa) == F) %>%
-  filter(region %in% c("ANA", "SCI","SRI",'SMI'),
-         classcode %in% top_species)
+#   sim_plot <- outcomes %>%
+#     mutate(rsize = plyr::round_any(mpa_size, 0.25),
+#            reff = plyr::round_any(mpa_effect,0.2)) %>%
+#     filter(rsize == 0.25, depletion < 0.8, depletion > 0.2,
+#            scientific_name %in% sim_life_history$scientific_name
+#            ) %>%
+#     mutate(year = years_protected + 2001) %>%
+#     filter(year %in% did_betas$year, years_protected >1) %>% 
+#     group_by(year, reff) %>% 
+#     count() %>% 
+#     group_by(year) %>% 
+#     mutate(sc = n / sum(n))
+# 
+#   did_plot <- did_betas %>%
+#     ggplot() +
+#     geom_tile(data = sim_plot, aes(year, reff, fill = sc), alpha = 0.75) +
+#     geom_vline(aes(xintercept = 2003), color = 'red', linetype = 2, size = 2) +
+#     geom_hline(aes(yintercept = 0)) +
+#     geom_pointrange(aes(
+#       year,
+#       y = exp(estimate) - 1,
+#       ymin = exp(lower) - 1,
+#       ymax = exp(upper) - 1
+#     ),
+#     size = 1.5
+#     ) +
+#     labs(x = "Year", y = "Targeted Trend Divergence") +
+#     scale_fill_viridis(guide = wide_hgc,
+#                        name = "% of Annual Sims",
+#                        labels = scales::percent_format(accuracy = 1),
+#                        option = "magma") + 
+#     theme(legend.position = "top",
+#         panel.grid.major.x = element_blank(),
+#         panel.grid.minor.x = element_blank()
+#         )
+#   
+# 
+# 
+# 
+# cip_data <- pisco_data %>%
+#   left_join(site_data, by = c("site","side")) %>%
+#   filter(is.na(eventual_mpa) == F) %>%
+#   filter(region %in% c("ANA", "SCI","SRI",'SMI'),
+#          classcode %in% top_species)
 
@@
 
-Indeed, many recent agreements to expand MPAs (most proximately, the Convention on Biological Diversity’s Strategic Plan for Biodiversity, which calls for 10% of coastal waters to be protected inside MPAs by 2020 and the International Union for Conservation of Nature call for 30% by 2030) rely on the belief that well-designed MPAs will achieve benefits both within, and outside, their borders.
+Indeed, many recent agreements to expand MPAs (most proximately, the Convention on Biological Diversity’s Strategic Plan for Biodiversity, which calls for 10% of coastal waters to be protected inside MPAs by 2020 and the International Union for Conservation of Nature call for 30% by 2030) are based on the expectaion that well-designed MPAs will achieve benefits both within, and outside, their borders.
 Despite these assumptions, our collective scientific understanding of the regional-scale conservation and fishery impacts of current and future MPAs is surprisingly limited.
 
-What is the scientific evidence of the conservation effects of MPAs?
 Numerous studies provide evidence that well-enforced and appropriately sized MPAs can produce conservation benefits within their borders [@lester2009;@halpern2003; @edgar2014; @gerber2005].
@@
 As stakeholders around the world increasingly seek to use MPAs in marine resource management portfolios, and the performance of existing MPAs is evaluated, it is critical that we develop a better understanding of the magnitude and drivers of regional-scale MPA effects.
-To address this gap, this study examines two critical questions: 1) What do we expect the regional-scale conservation effects of MPAs to be and 2) When (and how) can we expect to empirically detect these effects?
-We answer these questions using a simulation analysis to frame the theoretical regional conservation impacts of MPAs.
-We then present an empirical test of the evidence for regional-level conservation effects of MPAs resulting from a network of closures put in place in the Channel Islands, California, USA, in 2003.
+The Channel Islands National Marine Sanctuary, California, USA provides an ideal case study to address this need.
+A network of protected areas covering approximately 20% of the Islands' waters qas put in place in 2003 (see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of these MPAs).
+We use data from the first 15 years of protection to provide what is to our knowledge the first large-scale empirical assessment of the regional effect of a large MPA network on a wide array of fin-fish species.
+In contrast to clear differences in biomass densities observed inside and outside of well protected MPAs [@lester2009], we are unable to detect a clear regional effect from the Channel Islands MPAs.
+We build off of existing MPA theory to consider why this might be, and provide guidance for scientists and managers as to when and how we might expect to estimate the regional conservation effects of MPAs.
 
@@
 
-Much of the literature on the effects of MPAs focuses on assessing the conservation effects within the borders of protected areas [e.g.
+The empirical MPA literature focuses on assessing the conservation effects within the borders of protected areas [e.g.
 @edgar2014].
@@
 
-We used a simulation model to examine the regional conservation effects of MPAs under a wide array of conditions that might reasonably be expected to affect these outcomes.
-The model explores combinations of MPA designs, life histories, and fishing fleet dynamics (See Table.S1 for a complete description of simulation variables).
-We use this model to explore how different drivers interact to affect the conservation outcomes of MPAs.
-
-### Theoretical Conservation Effects {#conservation .unnumbered}
-
-As a thought experiment, imagine a region that has driven its fish populations to near extinction and then places 100\% of its waters inside a no-take MPA.
+<!-- As a thought experiment, imagine a region that has driven its fish populations to near extinction and then places 100\% of its waters inside a no-take MPA.
 In that setting, we would expect the regional-scale conservation benefits to be massive.
@@
 In that setting, it is possible to create a small net conservation loss if the concentration of fishing pressure outside the reserve has a greater effect than the biomass buildup inside the reserve.
-Within these broad bounds, numerous factors can act to affect the regional effects of MPAs.
-These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008], the strength and timing of density dependence in the population (e.g.
-pre- or post-settlement), how overfished the population would be without the MPA, and how fishing activity and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @gerber2005; @hilborn2004a; @gaines2010; @botsford2003].
+Within these broad bounds,  -->
+
+Numerous factors can affect the regional effects of MPAs.
+These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008], larval dispersal patternss and the strength and timing of density dependence in the population (e.g.
+pre- or post-settlement, @burgess), how overfished the population would be without the MPA, and how fishing activity and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @gerber2005; @hilborn2004a; @gaines2010; @botsford2003;@smith2003;@smith2006].
 In addition, even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through habitat and network effects [@gaines2010; @costello2010].
@@
 ```
+
+Given this existing body of theory, what should we expect *a priori* the effect size of the Channel Islands MPAs to be?
+The species are xx.
+We utilize a spatial bio-econonomic model to simulate the expected effects of an MPA network stylized to resemble the Channel Islands (See SI for details of simulation framework).
+For this group, MPA theory suggests that regional effects on the order of XX% are reasonable for such a system.
+
+Although the Islands do not encompass the full population of any of these species, they encompass a large enough area to assess the regional MPA conservation effects we might see in the Channel Islands after 13 years of protection (the time span covered by our model).
+The median regional conservation effect produced by our simulation model after 13 years of protection by MPAs covering between 15%-20% of a population's range was `r percent(med_ci_effect, accuracy = 1)`.
+This projected effect has an an interquartile range of `r percent(quart_ci_effect[2], accuracy = 1)` and `r percent(quart_ci_effect[4], accuracy = 1)`, although values as high as a 200% increase and as low as a 50% decrease were also produced.
+Our expectation then is that all else being equal we are likely looking for a modest but positive effect size.
+
+** reference HARKing, of course DiD and theory matching together doens't prove anything, but it's an important step to take **
+
+## How Can We Detect Regional Effects?
+
+What evidence do we have  for the effects of the Channel Island MPAs?
+@caslle2015 examined changes in biomass densities of species targeted by fishing activity inside and outside of the MPAs over time (response ratios), and compared these changes to comparable trends in non-targeted species.
+They found a statistically significant increase in the response ratios of targeted speices over time, and *evidenceXX* that this increase is smaller in the non-targeted species.
+Updating the results of @caselle2015 through 2018 shows a continuation of this trend (Fig.\@ref(fig:XX)).
+
+Why do these response ratio results not serve as sufficient evidence for the regional effects of the Channel Island MPAs?
+Control sites used in calculating response ratios are often selected based on abiotic or ecological traits such as habitat characteristics [@ferraro2018].
+However, selection of control sites is further complicated by the very spillover that MPAs are often intended to create.
+Export of adults or larvae from the MPA to the "control" site affects their status as controls, as does displacement of fishing effort from MPAs to control sites.
+In theory, control sites far enough away to negate both biological spillover and concentration by the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs may be challenging.
+While these concerns have been stated previously [e.g.
+@halpern2004], the MPA evaluation literature has by and large been unable to adequately address them, often due to the very real challenges of acquiring the necessary data [@ferraro2018].
+
+In the case of the Channel Islands MPAs, control sites are often located within XX Km of the MPAs, suggesting that they are susceptible to biological spillover and concentration of fishing effort resulting from the MPAs.
+Given these complications, we can certainly attribute the response ratios initially reported in @caselle2015 as evidence that the MPAs are indeed providing effective protection of targeted fish biomass inside their borders.
+However, it is unclear how well these response ratio results serve as in indicator of overall regional effects of the MPAs.
+
+To illustrate this problem, we used our stylized bio-economic model to simulate response ratio trajectories for the species and MPA coverage present in the Channel Islands.
+We then compared the range of plausible response ratios generated by our model to the true regional effect of the MPAs in each simulation.
+When response ratios were calculated from true control sites completely unaffected by the MPAs, they are a highly reliable measure of regional conservation effects.
+However, once we allow for the possibility of MPAs affecting the control sites through ecological and economic mechanisms, response ratios were generally a poor predictor of regional conservation effects (Fig.\@ref(fig:)).
+This was particularly true in the early years of MPA implementation; while biomass increases caused by MPAs are constrained by the growth rates of fish individuals and populations, reductions in biomass caused by concentrated fishing outside of reserves can occur quite quickly.
+As a result, response ratios can indeed serve as a measure of MPAs ability to provide effective within-border protection, but particular in the first XX years they are a highly imprecise measure of the net effect of the MPAs at a regional scale.
+
+<!-- If response ratios are not a broadly applicable solution to the challenge of estimating regional conservation effects, what then can we do?
+How then can we empirically measure the regional effects of MPAs?
+Intuition and bio-economic theory tells us that the regional conservation effects of MPAs are likely to be highly context dependent across locations and across species within locations, so careful thought must be devoted to an empirical design to detect their effects.
+The perfect experiment would involve two parallel worlds that were identical, except for the implementation of an MPA.
+In world “A”, no MPA would be implemented, and in the facsimile world “B”, the MPA would be implemented.
+Both worlds would be tracked before and after MPA implementation, and the conservation and fishery outcomes would be compared after treatment.
+Instead of two parallel worlds, a similar experimental design would involve random placement of MPAs.
+-->
+
+## MPA Effects in the Channel Islands {#estimating .unnumbered}
+
+Given these challenges, what is a viable alternative for estimating the conservation effects of MPAs?
+Building off of @caselle2015, we use a difference-in-difference estimator comparing densities in finfish species targeted by fishing effort (i.e., those potentially affected by an MPA) from those not targeted before and after MPA implementation [see @larsen2019 for a succinct explanation of difference-in-difference estimators in ecological settings].
+Targeted species in the Northern Channel Islands, California, include commercially important finfish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
+Each of these targeted species was the subject of prior bio-economic modeling related to the effects of MPAs in southern California [@costello2010d;@rassweiler2014].
+But, the analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*).
+Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species).
+
+```{r raw-trend, fig.cap="Centered and scaled mean annual biomass density of included species (faded lines) and smoothed means of targeted and non-targeted groups, and mean (darker lines) and 95% confidence interval of the mean (ribbon) over time", include = FALSE}
+raw_did_plot <- base_run$data[[1]] %>%
+  select(year, targeted,classcode, log_density, any_seen) %>%
+  mutate(density = exp(log_density) * any_seen) %>%
+  group_by(classcode,targeted, year) %>%
+  summarise(md = mean(density)) %>%
+  group_by(classcode,targeted) %>%
+  mutate(smd = (md - mean(md)) / sd(md)) %>%
+  ggplot(aes(year, smd, color = targeted == 1, fill = targeted == 1)) +
+  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red", size = 1.5) +
+  geom_line(aes(group = interaction(targeted, classcode)), alpha = 0.5) +
+  geom_smooth(size = 2) +
+  labs(y = "Centered and Scaled Mean Density", x = "Year") +
+    scale_color_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
+      scale_fill_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
+    labs(y = "Centered and Scaled Mean Density", x = "Year") + 
+  theme(legend.position = "top",
+        panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank()
+        )
+
+raw_did_plot <- raw_did_plot +
+      scale_color_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "") +
+        scale_fill_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "")
+
+pdf(file =  "figs/pop-trends.pdf",width = 8, height = 5, useDingbats = TRUE)
+print(raw_did_plot)
+dev.off()
+
+```
+
+\begin{figure}%[tbhp]
+  \centering
+  \includegraphics[width=1\linewidth]{figs/pop-trends.pdf}
+  \caption{Centered and scaled mean annual biomass densities of included targeted (red) and non-targeted (blue) fish species (faded lines) and smoothed means with 95\% confidence intervals over time (darker lines and ribbons).
+Vertical red dashed line shows year of MPA implementation.}
+  \label{pop-trends}
+\end{figure}
+
+Using this model, we estimate...
+XX.
+As we might expect from bio-economic modeling, we detect no significant difference between targeted and non-targeted biomass densities for the first XX years of protection.
+However, over time the model assigns greater and greater probability to MPA-mediated increases in targeted fin-fish biomass densities, up until XX, at which point th trend reverses itself.
+We detect no significant MPA effect between 2015 and 2018 (Fig.\@ref(fig:)).
+
+<!-- The raw trends in the biomass densities appear to show parallel trends in the densities of targeted and non-targeted fishes pre-MPA, with the targeted group diverging first positively and then negatively over time (Fig.\ref{pop-trends}).
+However, these raw trends could be caused by any number of confounding variables besides the effects of the MPAs themselves, from changes in observer skill and sampling locations, to shifts in environmental conditions.
+Using our difference-in-difference model in an effort to control for these factors, we found no significant difference in the biomass densities of targeted and non-targeted fishes pre-MPA (before 2003).
+Following the implementation of the MPAs in 2003, we see evidence of an increasing trend in densities of targeted relative to non-targeted fish, reaching a peak value in 2014.
+After 2014, densities of targeted species appear to have declined relative to what we would expect based on the non-targeted group (Fig.\ref{did-plot}).
+-->
+
+```{r did-plot, include = FALSE}
+
+did_plot <- did_plot +
+  scale_y_continuous(name = "~Divergence from Non-Targeted", labels = scales::percent_format(accuracy = 1))
+
+ggsave(did_plot, filename = "figs/did_plot.png",
+       width = 8, height = 5)
+
+# pdf(file =  "figs/did_plot.pdf",width = 8, height = 5, useDingbats = TRUE)
+# print(did_plot)
+# dev.off()
+
+```
+
+\begin{figure}%[tbhp]
+  \centering
+  \includegraphics[width=1\linewidth]{figs/did_plot.png}
+  \caption{Estimated divergence in biomass densities of targeted and non-targeted fishes throughout the Channel Islands (integrated across inside and outside of MPAs).
+MPAs are implemented in 2003 (red dashed line).
+Colored cells in the background indicate the frequency of MPA conservation effects predicted by our simulation modeling for MPAs similar in design and species composition to the Channel Islands in each year.}
+  \label{did-plot}
+\end{figure}
+
+<!-- To the extent that we can select control sites that are both sufficiently similar to the region treated with MPAs and are unaffected by biological or economic spillover, response ratios may be reliable estimators of regional conservation effects.
+In reality these conditions are unlikely to hold but for a few select cases.
+This becomes especially difficult when networks of MPAs are implemented, such as has occurred in California.
+In this case, control sites are generally located within a few kilometers from MPA borders, making biological spillover and fishing concentration highly likely.
+In these circumstances, without robust statistical controls, response ratios may provide an unreliable estimate of the regional conservation effect of marine protected areas.
+-->
+
+```{r dr-vs-mpa-effect-plot, include = FALSE}
+
+ylabs <- c(expression(""<= "-100%"), paste0(seq(0,300, by = 100),"%"),expression("">= "400%") )
+
+dr_plot <- density_ratios %>%
+  filter(sprinkler == FALSE, mpa_habfactor == 1) %>%
+  select(experiment,
+         years_protected,
+         mpa_effect,
+         adult_movement,
+         mpa_size,
+         contains("density_ratio")) %>%
+  mutate(adult_movement = adult_movement / max(adult_movement)) %>%
+  gather(scenario, density_ratio, contains("density_ratio")) %>%
+  group_by(experiment, scenario) %>%
+  filter(years_protected == max(years_protected)) %>%
+  mutate(density_ratio = density_ratio * mpa_size) %>% 
+  ggplot(aes(y = pmin(4, mpa_effect), x = pmin(4, density_ratio - 1))) +
+    geom_hline(aes(yintercept = 0)) + 
+  geom_vline(aes(xintercept = 0)) + 
+  geom_abline(aes(slope = 1, intercept = 0), linetype = 2, size = 1.5, color = "black") +
+  geom_point(aes(color = adult_movement, size = mpa_size), alpha = 0.5, show.legend = TRUE)  +
+    geom_smooth(method = "lm", se = FALSE) +
+  facet_wrap(~ rev(scenario), labeller = labeller(`rev(scenario)` = c(biased_density_ratio = "A) Inside vs. Experiment",true_density_ratio = "B) Inside vs. Outside"))) +
+  labs(y = "True % Increase in Biomass", x = "% Response Ratio") +
+  # scale_x_percent(labels = c(percent(seq(-1,1, by = 1)),">200%")) +
+  # scale_y_percent(labels = c(percent(seq(-1,1, by = 1)),">200%")) +
+  scale_color_viridis(guide = wide_hgc, labels = scales::percent_format(accuracy = 1), name = "Fish Movement",
+                      option = "cividis") + 
+  scale_x_continuous(labels = ylabs) + 
+    scale_y_continuous(labels = ylabs) + 
+  theme(legend.position = "top") + 
+  guides(size = FALSE)
+
+  ggsave(dr_plot,filename = "figs/density-ratio.png",
+         width = 8,
+         height = 5)
+
+# pdf(file =  "figs/density-ratio.png",width = 8, height = 5, useDingbats = TRUE)
+# print(dr_plot)
+# dev.off()
+
+```
+
+\begin{figure*}%[tbhp]
+  \centering
+  \includegraphics[width=.9\linewidth]{figs/density-ratio.png}
+  \caption{Simulated MPA-size weighted response ratios (x-axis) plotted against true regional change in biomass caused by MPAs (y-axis).
+Each point represents a simulation, and color represents the movement rate of fish in the simulation (expressed as a percentage of the patches a fish is likely to move in one time step).
+The black line shows the one:one fit line, the blue line is a linear fit of the relationship between observed response ratios and true regional change in biomass.
+Inside vs.
+Experiment simulates a true control, while Inside Vs.
+Outside uses a control that may be affected by the treatment.}
+  \label{density-ratio}
+\end{figure*}
+
+```{r}
+fish_v_fishing <- processed_grid %>%
+  select(-density_ratio) %>%
+  unnest() %>%
+  rename(fishery_effect = mpa_effect1) %>%
+  group_by(experiment) %>%
+  mutate(year = 1:length(year)) %>%
+  ungroup() %>%
+  mutate(years_protected = year - year_mpa + 1) %>%
+  left_join(outcomes %>% select(experiment, year, depletion),
+            by = c("experiment", "year"))
+
+fish_v_fishing_plot <- fish_v_fishing %>%
+  filter(years_protected == max(years_protected)) %>%
+  ggplot(aes(pmin(2, mpa_effect), pmin(2, fishery_effect))) +
+  geom_abline(aes(slope = 1, intercept = 0)) +
+  geom_smooth(method = "lm") +
+  geom_hex()
+
+```
+
+```{r ci-map, fig.cap = "Map of study region and sampling locations.
+Shaded polygons indicate location of MPAs.
+Points represent sampling locations, and color indicates the number of observations recorded at a given point", include = FALSE}
+
+ci_map <- pisco_data %>%
+  left_join(site_data, by = c("site","side")) %>%
+  filter(is.na(eventual_mpa) == F) %>%
+  filter(region %in% c("ANA", "SCI","SRI",'SMI'),
+         classcode %in% top_species) %>%
+  group_by(lat_wgs84, lon_wgs84, site, region, eventual_mpa) %>%
+  count()
+
+ci_map <-  ci_map %>%
+  dplyr::mutate(geometry = purrr::map2(lon_wgs84, lat_wgs84, ~ sf::st_point(x = c(.x, .y), dim = 'XY'))) %>%
+  ungroup() %>%
+  mutate(geometry = sf::st_sfc(geometry, crs = 4326)) %>%
+  sf::st_sf()
+
+bbox <- sf::st_bbox(ci_map)
+
+pisco_ci_map_plot <-  ggmap(channel_islands, alpha = 0.5) +
+  geom_sf(data = ca_mpas, inherit.aes = FALSE, alpha = 0.5) +
+  geom_sf(data = ci_map,inherit.aes = FALSE,
+  aes(fill = eventual_mpa, color = eventual_mpa),
+  shape = 21, 
+  alpha = 0.5,
+  show.legend = FALSE,
+  size = 1.5) +
+  coord_sf(xlim = c(bbox['xmin'] - .3, bbox['xmax'] + .06),
+  ylim = c(bbox['ymin'] - .1, bbox['ymax'] + .1)) +
+  # scale_color_gradient(low = "white", high = "orangered",
+  # guide = guide_colorbar(frame.colour = "black",frame.linewidth = 1,
+  # barheight = unit(13, "lines")),
+  # name = "Observations")  +
+  labs(x = "", y = "") +
+      ggspatial::annotation_scale(location = "br", width_hint = 0.4, text_col = "white") + 
+  ggsci::scale_color_npg() + 
+    ggsci::scale_color_npg()
+
+# pisco_ci_map_plot <-  ggplot() +
+#   geom_sf(data = tests2, fill = c("#FFD39B")) +
+#   geom_sf(data = ca_mpas, inherit.aes = FALSE, alpha = 0.5) +
+#   geom_sf(data = ci_map,inherit.aes = FALSE,
+#   aes(color = eventual_mpa),
+#   alpha = 0.75,
+#   show.legend = FALSE,
+#   size = 2) +
+#   coord_sf(xlim = c(bbox['xmin'] - .3, bbox['xmax'] + .06),
+#   ylim = c(bbox['ymin'] - .1, bbox['ymax'] + .1)) +
+#   # scale_color_gradient(low = "white", high = "orangered",
+#   # guide = guide_colorbar(frame.colour = "black",frame.linewidth = 1,
+#   # barheight = unit(13, "lines")),
+#   # name = "Observations")  +
+#   labs(x = "", y = "") +
+#       ggspatial::annotation_scale(location = "br", width_hint = 0.4, text_col = "white") + 
+#   theme(panel.background = element_rect(fill = "lightblue"))
+# 
+pdf(file =  "figs/channel-islands.pdf",width = 8, height = 5, useDingbats = TRUE)
+print(pisco_ci_map_plot)
+dev.off()
+
+```
+
+\begin{figure}%[tbhp]
+  \centering
+  \includegraphics[width=1\linewidth]{figs/channel-islands.pdf}
+  \caption{Map of PISCO sampling locations within the the Channel Islands National Marine Sanctuary, California, USA.
+Grey boxes indicated MPAs, points sampling locations, with color representing whether the site is inside an MPA (blue) or not (red)}
+  \label{channel-islands}
+\end{figure}
+
+```{r}
+ci_effects <- short_term %>%
+  filter(years_protected == max(years_protected),
+         mpa_size  <= 0.25, mpa_size >= 0.15)
+
+med_ci_effect <- median(ci_effects$mpa_effect)
+
+# rethinking::HPDI(ci_effects$mpa_effect, prob = 0.5)
+
+quart_ci_effect <- quantile(ci_effects$mpa_effect)
+
+```
+
+Discussion {#discussion .unnumbered}
+------
+
+1. Why are the results from the channel islands imporant and what do they tell us
+
+2. Put the channel islands results in context -
+
+### What Should We Expect from MPAs?
 
@@
 
-## How Can We Detect Regional Effects?
-
-How can we empirically measure the regional effects of MPAs?
-Intuition and bio-economic theory tells us that the regional conservation effects of MPAs are likely to be highly context dependent across locations and across species within locations, so careful thought must be devoted to an empirical design to detect their effects.
-The perfect experiment would involve two parallel worlds that were identical, except for the implementation of an MPA.
-In world “A”, no MPA would be implemented, and in the facsimile world “B”, the MPA would be implemented.
-Both worlds would be tracked before and after MPA implementation, and the conservation and fishery outcomes would be compared after treatment.
-Instead of two parallel worlds, a similar experimental design would involve random placement of MPAs.
+### How can we use response ratios
 
@@
 Careful causal inference requires some other means of controlling for biases introduced by factors such as the MPA siting process and biological spillover and concentration of fishing effort  [@ferraro2018;@larsen2019].
-
 Conservation effects are typically estimated by comparing organism densities or biomass inside MPAs to densities or biomass in selected control sites outside MPAs, which we will refer to as response ratios.
@@
 
-Control sites used in density or biomass ratios are often selected based on abiotic or ecological traits such as habitat characteristics [@ferraro2018].
-However, selection of control sites is further complicated by the very spillover that MPAs are often intended to create.
-Export of adults or larvae from the MPA to the "control" site affects their status as controls, as does displacement of fishing effort from MPAs to control sites.
-In theory, control sites far enough away to negate both biological spillover and concentration by the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs may be challenging.
-While these concerns have been stated previously [e.g.
-@halpern2004], the MPA evaluation literature has by and large been unable to adequately address them, often due to the very real challenges of acquiring the necessary data [@ferraro2018].
-
 Failure to account for these spillover effects can result in a biased estimate of the true effect of a policy such as MPA placement [@larsen2019].
@@
 However, spatial BACI still requires that the MPAs do not affect the control sites, and as such has many of the same limitations as response ratios as an estimator of regional conservation effects.
-
-<!-- To the extent that we can select control sites that are both sufficiently similar to the region treated with MPAs and are unaffected by biological or economic spillover, response ratios may be reliable estimators of regional conservation effects.
-In reality these conditions are unlikely to hold but for a few select cases.
-This becomes especially difficult when networks of MPAs are implemented, such as has occurred in California.
-In this case, control sites are generally located within a few kilometers from MPA borders, making biological spillover and fishing concentration highly likely.
-In these circumstances, without robust statistical controls, response ratios may provide an unreliable estimate of the regional conservation effect of marine protected areas.
--->
-
-```{r dr-vs-mpa-effect-plot, include = FALSE}
-
-ylabs <- c(expression(""<= "-100%"), paste0(seq(0,300, by = 100),"%"),expression("">= "400%") )
-
-dr_plot <- density_ratios %>%
-  filter(sprinkler == FALSE, mpa_habfactor == 1) %>%
-  select(experiment,
-         years_protected,
-         mpa_effect,
-         adult_movement,
-         mpa_size,
-         contains("density_ratio")) %>%
-  mutate(adult_movement = adult_movement / max(adult_movement)) %>%
-  gather(scenario, density_ratio, contains("density_ratio")) %>%
-  group_by(experiment, scenario) %>%
-  filter(years_protected == max(years_protected)) %>%
-  mutate(density_ratio = density_ratio * mpa_size) %>% 
-  ggplot(aes(y = pmin(4, mpa_effect), x = pmin(4, density_ratio - 1))) +
-    geom_hline(aes(yintercept = 0)) + 
-  geom_vline(aes(xintercept = 0)) + 
-  geom_abline(aes(slope = 1, intercept = 0), linetype = 2, size = 1.5, color = "black") +
-  geom_point(aes(color = adult_movement, size = mpa_size), alpha = 0.5, show.legend = TRUE)  +
-    geom_smooth(method = "lm", se = FALSE) +
-  facet_wrap(~ rev(scenario), labeller = labeller(`rev(scenario)` = c(biased_density_ratio = "A) Inside vs. Experiment",true_density_ratio = "B) Inside vs. Outside"))) +
-  labs(y = "True % Increase in Biomass", x = "% Response Ratio") +
-  # scale_x_percent(labels = c(percent(seq(-1,1, by = 1)),">200%")) +
-  # scale_y_percent(labels = c(percent(seq(-1,1, by = 1)),">200%")) +
-  scale_color_viridis(guide = wide_hgc, labels = scales::percent_format(accuracy = 1), name = "Fish Movement",
-                      option = "cividis") + 
-  scale_x_continuous(labels = ylabs) + 
-    scale_y_continuous(labels = ylabs) + 
-  theme(legend.position = "top") + 
-  guides(size = FALSE)
-
-  ggsave(dr_plot,filename = "figs/density-ratio.png",
-         width = 8,
-         height = 5)
-
-# pdf(file =  "figs/density-ratio.png",width = 8, height = 5, useDingbats = TRUE)
-# print(dr_plot)
-# dev.off()
-
-```
-
-\begin{figure*}%[tbhp]
-  \centering
-  \includegraphics[width=.9\linewidth]{figs/density-ratio.png}
-  \caption{Simulated MPA-size weighted response ratios (x-axis) plotted against true regional change in biomass caused by MPAs (y-axis).
-Each point represents a simulation, and color represents the movement rate of fish in the simulation (expressed as a percentage of the patches a fish is likely to move in one time step).
-The black line shows the one:one fit line, the blue line is a linear fit of the relationship between observed response ratios and true regional change in biomass.
-Inside vs.
-Experiment simulates a true control, while Inside Vs.
-Outside uses a control that may be affected by the treatment.}
-  \label{density-ratio}
-\end{figure*}
-
-```{r}
-fish_v_fishing <- processed_grid %>%
-  select(-density_ratio) %>%
-  unnest() %>%
-  rename(fishery_effect = mpa_effect1) %>%
-  group_by(experiment) %>%
-  mutate(year = 1:length(year)) %>%
-  ungroup() %>%
-  mutate(years_protected = year - year_mpa + 1) %>%
-  left_join(outcomes %>% select(experiment, year, depletion),
-            by = c("experiment", "year"))
-
-fish_v_fishing_plot <- fish_v_fishing %>%
-  filter(years_protected == max(years_protected)) %>%
-  ggplot(aes(pmin(2, mpa_effect), pmin(2, fishery_effect))) +
-  geom_abline(aes(slope = 1, intercept = 0)) +
-  geom_smooth(method = "lm") +
-  geom_hex()
-
-```
-
-## MPA Effects in the Channel Islands {#estimating .unnumbered}
-
-```{r ci-map, fig.cap = "Map of study region and sampling locations.
-Shaded polygons indicate location of MPAs.
-Points represent sampling locations, and color indicates the number of observations recorded at a given point", include = FALSE}
-
-ci_map <- pisco_data %>%
-  left_join(site_data, by = c("site","side")) %>%
-  filter(is.na(eventual_mpa) == F) %>%
-  filter(region %in% c("ANA", "SCI","SRI",'SMI'),
-         classcode %in% top_species) %>%
-  group_by(lat_wgs84, lon_wgs84, site, region, eventual_mpa) %>%
-  count()
-
-ci_map <-  ci_map %>%
-  dplyr::mutate(geometry = purrr::map2(lon_wgs84, lat_wgs84, ~ sf::st_point(x = c(.x, .y), dim = 'XY'))) %>%
-  ungroup() %>%
-  mutate(geometry = sf::st_sfc(geometry, crs = 4326)) %>%
-  sf::st_sf()
-
-bbox <- sf::st_bbox(ci_map)
-
-pisco_ci_map_plot <-  ggmap(channel_islands, alpha = 0.5) +
-  geom_sf(data = ca_mpas, inherit.aes = FALSE, alpha = 0.5) +
-  geom_sf(data = ci_map,inherit.aes = FALSE,
-  aes(fill = eventual_mpa, color = eventual_mpa),
-  shape = 21, 
-  alpha = 0.5,
-  show.legend = FALSE,
-  size = 1.5) +
-  coord_sf(xlim = c(bbox['xmin'] - .3, bbox['xmax'] + .06),
-  ylim = c(bbox['ymin'] - .1, bbox['ymax'] + .1)) +
-  # scale_color_gradient(low = "white", high = "orangered",
-  # guide = guide_colorbar(frame.colour = "black",frame.linewidth = 1,
-  # barheight = unit(13, "lines")),
-  # name = "Observations")  +
-  labs(x = "", y = "") +
-      ggspatial::annotation_scale(location = "br", width_hint = 0.4, text_col = "white") + 
-  ggsci::scale_color_npg() + 
-    ggsci::scale_color_npg()
-
-# pisco_ci_map_plot <-  ggplot() +
-#   geom_sf(data = tests2, fill = c("#FFD39B")) +
-#   geom_sf(data = ca_mpas, inherit.aes = FALSE, alpha = 0.5) +
-#   geom_sf(data = ci_map,inherit.aes = FALSE,
-#   aes(color = eventual_mpa),
-#   alpha = 0.75,
-#   show.legend = FALSE,
-#   size = 2) +
-#   coord_sf(xlim = c(bbox['xmin'] - .3, bbox['xmax'] + .06),
-#   ylim = c(bbox['ymin'] - .1, bbox['ymax'] + .1)) +
-#   # scale_color_gradient(low = "white", high = "orangered",
-#   # guide = guide_colorbar(frame.colour = "black",frame.linewidth = 1,
-#   # barheight = unit(13, "lines")),
-#   # name = "Observations")  +
-#   labs(x = "", y = "") +
-#       ggspatial::annotation_scale(location = "br", width_hint = 0.4, text_col = "white") + 
-#   theme(panel.background = element_rect(fill = "lightblue"))
-# 
-pdf(file =  "figs/channel-islands.pdf",width = 8, height = 5, useDingbats = TRUE)
-print(pisco_ci_map_plot)
-dev.off()
-
-```
-
-\begin{figure}%[tbhp]
-  \centering
-  \includegraphics[width=1\linewidth]{figs/channel-islands.pdf}
-  \caption{Map of PISCO sampling locations within the the Channel Islands National Marine Sanctuary, California, USA.
-Grey boxes indicated MPAs, points sampling locations, with color representing whether the site is inside an MPA (blue) or not (red)}
-  \label{channel-islands}
-\end{figure}
-
-```{r}
-ci_effects <- short_term %>%
-  filter(years_protected == max(years_protected),
-         mpa_size  <= 0.25, mpa_size >= 0.15)
-
-med_ci_effect <- median(ci_effects$mpa_effect)
-
-# rethinking::HPDI(ci_effects$mpa_effect, prob = 0.5)
-
-quart_ci_effect <- quantile(ci_effects$mpa_effect)
-
-```
-
-Given these challenges, what is a viable alternative for estimating the conservation effects of MPAs?
-In a manner similar to @caselle2015, we use a difference-in-difference estimator comparing densities in finfish species targeted by fishing effort (i.e., those potentially affected by an MPA) from those not targeted before and after MPA implementation [see @larsen2019 for a succinct explanation of difference-in-difference estimators in ecological settings].
-Targeted species in the Northern Channel Islands, California, include commercially important finfish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
-Each of these targeted species was the subject of prior bio-economic modeling related to the effects of MPAs in southern California [@costello2010d;@rassweiler2014].
-But, we must note that the analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*).
-Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species).
-The key assumptions of this approach are that a) within the time-frame of the model there are no significant interaction effects between the targeted and non-targeted species (which in fact we do not detect, see SI), and b) that in the absence of the MPAs both the targeted and non-targeted groups of species would have exhibited similar trends in densities.
-The advantage to this approach is that given the time-frame of the model (16 years), we believe that spillover effects of MPA placement on the non-targeted species are likely to be much less severe than the effects of biological spillover and fishing concentration that may bias the performance of estimators such as a response ratios or a spatial BACI design.
-
-We applied this difference-in-difference strategy to a large dataset in California.
-We used empirical kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands with the ultimate goal of testing the regional effects of MPAs in a real world context.
-A network of MPAs covering approximately 20% of the islands’ waters was put in place in 2003 as part of the California Marine Life Protection Act (MLPA) (see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of the MLPA).
-PISCO conducts visual SCUBA surveys at a large number of rocky-reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\ref{channel-islands}-\ref{pop-trends}).
-
-Although the Islands do not encompass the full population of any of these species, they encompass a large enough area to assess the regional MPA conservation effects we might see in the Channel Islands after 13 years of protection (the time span covered by our model).
-The median regional conservation effect produced by our simulation model after 13 years of protection by MPAs covering between 15%-20% of a population's range was `r percent(med_ci_effect, accuracy = 1)`.
-This projected effect has an an interquartile range of `r percent(quart_ci_effect[2], accuracy = 1)` and `r percent(quart_ci_effect[4], accuracy = 1)`, although values as high as a 200% increase and as low as a 50% decrease were also produced.
-Our expectation then is that all else being equal we are likely looking for a modest but positive effect size.
-
-```{r raw-trend, fig.cap="Centered and scaled mean annual biomass density of included species (faded lines) and smoothed means of targeted and non-targeted groups, and mean (darker lines) and 95% confidence interval of the mean (ribbon) over time", include = FALSE}
-raw_did_plot <- base_run$data[[1]] %>%
-  select(year, targeted,classcode, log_density, any_seen) %>%
-  mutate(density = exp(log_density) * any_seen) %>%
-  group_by(classcode,targeted, year) %>%
-  summarise(md = mean(density)) %>%
-  group_by(classcode,targeted) %>%
-  mutate(smd = (md - mean(md)) / sd(md)) %>%
-  ggplot(aes(year, smd, color = targeted == 1, fill = targeted == 1)) +
-  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red", size = 1.5) +
-  geom_line(aes(group = interaction(targeted, classcode)), alpha = 0.5) +
-  geom_smooth(size = 2) +
-  labs(y = "Centered and Scaled Mean Density", x = "Year") +
-    scale_color_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
-      scale_fill_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
-    labs(y = "Centered and Scaled Mean Density", x = "Year") + 
-  theme(legend.position = "top",
-        panel.grid.major.x = element_blank(),
-        panel.grid.minor.x = element_blank()
-        )
-
-raw_did_plot <- raw_did_plot +
-      scale_color_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "") +
-        scale_fill_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "")
-
-pdf(file =  "figs/pop-trends.pdf",width = 8, height = 5, useDingbats = TRUE)
-print(raw_did_plot)
-dev.off()
-
-```
-
-\begin{figure}%[tbhp]
-  \centering
-  \includegraphics[width=1\linewidth]{figs/pop-trends.pdf}
-  \caption{Centered and scaled mean annual biomass densities of included targeted (red) and non-targeted (blue) fish species (faded lines) and smoothed means with 95\% confidence intervals over time (darker lines and ribbons).
-Vertical red dashed line shows year of MPA implementation.}
-  \label{pop-trends}
-\end{figure}
-
-The raw trends in the biomass densities appear to show parallel trends in the densities of targeted and non-targeted fishes pre-MPA, with the targeted group diverging first positively and then negatively over time (Fig.\ref{pop-trends}).
-However, these raw trends could be caused by any number of confounding variables besides the effects of the MPAs themselves, from changes in observer skill and sampling locations, to shifts in environmental conditions.
-Using our difference-in-difference model in an effort to control for these factors, we found no significant difference in the biomass densities of targeted and non-targeted fishes pre-MPA (before 2003).
-Following the implementation of the MPAs in 2003, we see evidence of an increasing trend in densities of targeted relative to non-targeted fish, reaching a peak value in 2014.
-After 2014, densities of targeted species appear to have declined relative to what we would expect based on the non-targeted group (Fig.\ref{did-plot}).
-
-```{r did-plot, include = FALSE}
-
-did_plot <- did_plot +
-  scale_y_continuous(name = "~Divergence from Non-Targeted", labels = scales::percent_format(accuracy = 1))
-
-ggsave(did_plot, filename = "figs/did_plot.png",
-       width = 8, height = 5)
-
-# pdf(file =  "figs/did_plot.pdf",width = 8, height = 5, useDingbats = TRUE)
-# print(did_plot)
-# dev.off()
-
-```
-
-\begin{figure}%[tbhp]
-  \centering
-  \includegraphics[width=1\linewidth]{figs/did_plot.png}
-  \caption{Estimated divergence in biomass densities of targeted and non-targeted fishes throughout the Channel Islands (integrated across inside and outside of MPAs).
-MPAs are implemented in 2003 (red dashed line).
-Colored cells in the background indicate the frequency of MPA conservation effects predicted by our simulation modeling for MPAs similar in design and species composition to the Channel Islands in each year.}
-  \label{did-plot}
-\end{figure}
-
-Discussion {#discussion .unnumbered}
-------
 
@@
 
+We applied this difference-in-difference strategy to a large dataset in California.
+We used empirical kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands with the ultimate goal of testing the regional effects of MPAs in a real world context.
+A network of MPAs covering approximately 20% of the islands’ waters was put in place in 2003 as part of the California Marine Life Protection Act (MLPA) (see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of the MLPA).
+PISCO conducts visual SCUBA surveys at a large number of rocky-reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\ref{channel-islands}-\ref{pop-trends}).
+
+The key assumptions of this approach are that a) within the time-frame of the model there are no significant interaction effects between the targeted and non-targeted species (which in fact we do not detect, see SI), and b) that in the absence of the MPAs both the targeted and non-targeted groups of species would have exhibited similar trends in densities.
+The advantage to this approach is that given the time-frame of the model (18 years), we believe that spillover effects of MPA placement on the non-targeted species are likely to be much less severe than the effects of biological spillover and fishing concentration that may bias the performance of estimators such as a response ratios or a spatial BACI design.
+
 We use a difference-in-difference regression style regression to attempt to estimate the causal effect of MPAs on regional targeted fish biomass.

### 2020-03-11 documents/regional-mpa-notebook.Rmd
back to squid

@@
 knitr::opts_chunk$set(echo = TRUE)
+```
+
+important_lh <- life_history_data %>% 
+  filter(classcode %in% unique(pisco_abundance_data$classcode)) %>% 
+  mutate(m_v_k = m / k, lm_v_loo = lm / loo ) %>% 
+  select(m_v_k, lm_v_loo, temperature, classcode, commonname, targeted)
+
+lh_pca <- prcomp(important_lh %>% select(-commonname, -classcode, -targeted))
+
+# ggbiplot::ggbiplot(lh_pca, labels = important_lh$commonname, groups = important_lh$targeted == 1)
+
+# unlo("plyr")
+
+lg_pca_scores <- lh_pca$x %>% 
+  as_tibble() %>% 
+  mutate(commonname = important_lh$commonname, classcode = important_lh$classcode) %>% 
+  pivot_longer(contains("PC"), names_to = "pc", values_to = "pcscore") %>% 
+  group_by(pc) %>% 
+  mutate(pcscore = scale(pcscore)) %>% 
+  filter(pc == "PC1") %>% 
+  filter(pcscore <1, pcscore > -1)
+
+simple_pca_did_data <- pisco_abundance_data %>%
+  filter(classcode %in% unique(lg_pca_scores$classcode),
+         site_side %in% unique(consistent_sites$site_side)) %>% 
+  group_by(year, site,side, region, zone, transect,eventual_mpa, classcode, targeted) %>%
+  summarise(total_classcode_density = sum(exp(log_density))) %>%  # sum density across all levels of a transect
+  group_by(year, site,side, region,eventual_mpa, classcode, targeted) %>%
+  summarise(md = mean(total_classcode_density)) %>% # calculate mean density per year site, side, species, averaging over zone, transect
+  group_by(year, site,side,region, eventual_mpa, targeted) %>% 
+  summarise(total_biomass_density = (sum(md) / 1e6) * 10000, # calculate total and mean biomass densities across all species per year site side
+            mean_biomass_density = (mean(md) / 1e6) * 10000) %>% 
+  ungroup() %>% 
+  mutate(fyear = factor(year)) %>% 
+  mutate(fyear = relevel(fyear, "2003")) %>% 
+  mutate(site_side = paste(site, side, sep = "-")) %>% 
+  mutate(year_bins = cut(year, year_bins))
+
+pca_did_reg <- stan_glmer(log(total_biomass_density + 1e-6) ~ targeted*year_bins + (1|site_side), data = simple_pca_did_data)
+
+pca_did_reg_gamma <- stan_glmer((total_biomass_density + 1e-6) ~ targeted*year_bins + (1|site_side), data = simple_pca_did_data,
+                                family = Gamma(link = "log"))
+
+# loo_compare(loo(simple_did_reg), loo(simple_did_reg_gamma))
+
+pca_did_results <- tidybayes::tidy_draws(pca_did_reg) %>% 
+  select(contains("."), contains("targeted:")) %>% 
+  pivot_longer(contains("targeted"), names_to = "year", values_to = "did", names_prefix = "targeted:year_bins") %>% 
+  mutate(did =  exp(did) - 1) %>% 
+  group_by(year) %>% 
+  mutate(prank = percent_rank(did)) %>% 
+  ungroup() %>% 
+  filter(prank >= 0.025, prank <= 0.975)
+
+agg_pca_did_plot <-  pca_did_results %>%
+  ggplot(aes(year, did)) +
+  geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
+  tidybayes::stat_halfeye(alpha = 0.7) + 
+  scale_y_continuous(labels = percent, name = "Estimated MPA Effect")
+```{r setup, include=FALSE}

### 2020-03-11 functions/estimate_did.R

@@ -1,0 +1,309 @@
+estimate_did <-
+  function(data,
+           site_data,
+           life_history_data,
+           bin_width = 3,
+           min_classcode_years = 15,
+           consistent_sites_only = TRUE,
+           cdfw_catches,
+           data_to_use = "all",
+           data_source = "pisco") {
+    
+    # data <- pisco_abundance_data
+    # 
+    # data <- kfm_abundance_data
+    if (data_to_use == "mpa_only"){
+      
+      data <- data %>% 
+        filter(eventual_mpa == TRUE)
+      
+    } else if (data_to_use == "fished_only"){
+      
+      data <- data %>% 
+        filter(eventual_mpa == FALSE)
+    }
+    
+    annual_catches <- cdfw_catches %>%
+      filter(classcode %in% unique(data$classcode)) %>%
+      group_by(year) %>%
+      summarise(var_catch = sum(catch)) %>%
+      mutate(var_lag_catch = lag(var_catch, 1))
+    
+    annual_catches$var_lag_catch[1] <-   annual_catches$var_lag_catch[2] #assume 1999 catches = 2000 catches, not crazy given AC in totals
+    
+    consistent_did_sites <- data %>%
+      select(site_side, year) %>%
+      unique() %>%
+      group_by(site_side) %>%
+      mutate(has_all = all((2003:2012) %in% year)) %>% # results change dramatically if you filter until present time...
+      ungroup() %>%
+      filter(has_all)
+    
+    consistent_classcodes <- data %>%
+      group_by(classcode) %>%
+      filter(any_seen) %>%
+      summarise(n_years = n_distinct(year)) %>%
+      ungroup() %>%
+      filter(n_years >= min_classcode_years)
+    
+    year_bins <-
+      c(1999, seq(2003, max(pisco_abundance_data$year) + 1, by = bin_width))
+    
+    did_data <- data %>% {
+      if (consistent_sites_only == TRUE) {
+        filter(.,
+               site_side %in% unique(consistent_did_sites$site_side))
+      } else{
+        .
+      }
+    } %>%
+    filter(classcode %in% consistent_classcodes$classcode) 
+    
+    if (data_source == "pisco"){
+    
+    did_data <- did_data %>% 
+      group_by(year,
+               site_side,
+               region,
+               zone,
+               transect,
+               eventual_mpa,
+               classcode,
+               targeted) %>%
+      summarise(
[…]

### 2020-03-11 make-regional-mpas.R

@@ -29,6 +29,7 @@ library(rEDM)
 library(bayesplot)
 library(tidyverse)
 extrafont::loadfonts()
+rstan_options(auto_write = TRUE)
 
 
 
@@ -40,7 +41,7 @@ walk(functions, ~ here::here("functions", .x) %>% source()) # load local functio
 
 # options -----------------------------------------------------------------
 
-run_name <- 'v6.0'
+run_name <- 'v5.1'
 
 run_description <- "PNAS R and R run with MPA and fishery-only runs "
 
@@ -49,13 +50,17 @@ run_description <- "PNAS R and R run with MPA and fishery-only runs "
 # So, once you've run simulate_mpas, you can set it to FALSE and validata_mpas will work
 
 
-run_did <- FALSE # run difference in difference on data from the CINMS
+run_did <- TRUE # run difference in difference on data from the CINMS
+
+run_tmb <- FALSE
 
 simulate_mpas <- FALSE # simulate MPA outcomes
 
 validate_mpas <- FALSE
 
 process_results <- TRUE
+
+get_cdfw_catches <-  FALSE
 
 knit_paper <- FALSE
 
@@ -101,9 +106,6 @@ if (!dir.exists(fig_dir)){
 # run MLPA difference-in-difference ---------------------------------------
 if (run_did == TRUE){
 
-rstan_options(auto_write = TRUE)
-
-run_tmb <- FALSE
 
 run_length_to_density <-  FALSE
 
@@ -261,8 +263,133 @@ life_history_data <- life_history_data %>%
   mutate(targeted = ifelse(is.na(caselle_targeted), targeted, caselle_targeted))
 
 
-cdfw_catches <-
-  read_csv(file = here::here("data", 'cdfw-catches.csv')) %>%
+if (get_cdfw_catches){
+  
+  file_names <- list.files(here('data','cdfw-data'))
+  
+  
+  process_cdfw <- function(file_name) {
+    dat <-
+      tabulizer::extract_tables(here("data","cdfw-data", file_name))
+    
+    year <- str_extract(file_name, pattern = '(?<=landings).*(?=_)')
+    
+    year <- as.numeric(paste0('20', year))
+    
+    collapse_pages <- function(x) {
+      missing_all <-
+        map_lgl(x %>% as_tibble, ~ all(str_count(.x) == 0)) == F
+      
+      x <- x[, missing_all]
+      
[…]

### 2020-03-12 documents/ovando-regional-effects-of-mpas-v2.Rmd
passing to abalone

@@
 
+  load(file.path(run_dir, 'abundance_data.Rdata'))
+
 model_runs <- read_rds( file.path(run_dir,"did_fits.rds"))
@@
 load(file = file.path(run_dir ,"sim_grid.Rdata"))
+
+load(file = file.path(run_dir,"plots.RData"))
 
@@
 )
-
-channel_islands <- readRDS(here::here("data","channel_islands_map.RDS"))
-
-ca_mpas <- sf::st_read(here::here("data","MPA_CA_Existing_160301")) %>%
-  rmapshaper::ms_simplify() %>%
-  sf::st_transform(crs = 4326)
 
@@
 A network of protected areas covering approximately 20% of the Islands' waters qas put in place in 2003 (see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of these MPAs).
-We use data from the first 15 years of protection to provide what is to our knowledge the first large-scale empirical assessment of the regional effect of a large MPA network on a wide array of fin-fish species.
+We use data from the first 14 years of protection to provide what is to our knowledge the first large-scale empirical assessment of the regional effect of a large MPA network on a wide array of fin-fish species.
 In contrast to clear differences in biomass densities observed inside and outside of well protected MPAs [@lester2009], we are unable to detect a clear regional effect from the Channel Islands MPAs.
@@
 These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008], larval dispersal patternss and the strength and timing of density dependence in the population (e.g.
-pre- or post-settlement, @burgess), how overfished the population would be without the MPA, and how fishing activity and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @gerber2005; @hilborn2004a; @gaines2010; @botsford2003;@smith2003;@smith2006].
+pre- or post-settlement, @burgess), how overfished the population would be without the MPA, and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @gerber2005; @hilborn2004a; @gaines2010; @botsford2003;@smith2003;@smith2006].
 In addition, even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through habitat and network effects [@gaines2010; @costello2010].
@@
 
-Although the Islands do not encompass the full population of any of these species, they encompass a large enough area to assess the regional MPA conservation effects we might see in the Channel Islands after 13 years of protection (the time span covered by our model).
+Although the Islands do not encompass the full population of any of these species, they encompass a large enough area to assess the regional MPA conservation effects we might see in the Channel Islands after 14 years of protection (the time span covered by our model).
 The median regional conservation effect produced by our simulation model after 13 years of protection by MPAs covering between 15%-20% of a population's range was `r percent(med_ci_effect, accuracy = 1)`.
@@
 What evidence do we have  for the effects of the Channel Island MPAs?
-@caslle2015 examined changes in biomass densities of species targeted by fishing activity inside and outside of the MPAs over time (response ratios), and compared these changes to comparable trends in non-targeted species.
-They found a statistically significant increase in the response ratios of targeted speices over time, and *evidenceXX* that this increase is smaller in the non-targeted species.
+@caslle2015 examined changes in biomass densities of species targeted by fishing activity inside and outside of the MPAs over time (generally termed "response ratios"), and compared these changes to the trends in non-targeted species.
+They found a statistically significant increase in the response ratios of targeted speices over time, and evidence that this increase is smaller in the non-targeted species.
 Updating the results of @caselle2015 through 2018 shows a continuation of this trend (Fig.\@ref(fig:XX)).
@@
 
-<!-- If response ratios are not a broadly applicable solution to the challenge of estimating regional conservation effects, what then can we do?
-How then can we empirically measure the regional effects of MPAs?
-Intuition and bio-economic theory tells us that the regional conservation effects of MPAs are likely to be highly context dependent across locations and across species within locations, so careful thought must be devoted to an empirical design to detect their effects.
-The perfect experiment would involve two parallel worlds that were identical, except for the implementation of an MPA.
-In world “A”, no MPA would be implemented, and in the facsimile world “B”, the MPA would be implemented.
-Both worlds would be tracked before and after MPA implementation, and the conservation and fishery outcomes would be compared after treatment.
-Instead of two parallel worlds, a similar experimental design would involve random placement of MPAs.
--->
-
-## MPA Effects in the Channel Islands {#estimating .unnumbered}
-
-Given these challenges, what is a viable alternative for estimating the conservation effects of MPAs?
-Building off of @caselle2015, we use a difference-in-difference estimator comparing densities in finfish species targeted by fishing effort (i.e., those potentially affected by an MPA) from those not targeted before and after MPA implementation [see @larsen2019 for a succinct explanation of difference-in-difference estimators in ecological settings].
-Targeted species in the Northern Channel Islands, California, include commercially important finfish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
-Each of these targeted species was the subject of prior bio-economic modeling related to the effects of MPAs in southern California [@costello2010d;@rassweiler2014].
-But, the analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*).
-Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species).
-
-```{r raw-trend, fig.cap="Centered and scaled mean annual biomass density of included species (faded lines) and smoothed means of targeted and non-targeted groups, and mean (darker lines) and 95% confidence interval of the mean (ribbon) over time", include = FALSE}
-raw_did_plot <- base_run$data[[1]] %>%
-  select(year, targeted,classcode, log_density, any_seen) %>%
-  mutate(density = exp(log_density) * any_seen) %>%
-  group_by(classcode,targeted, year) %>%
-  summarise(md = mean(density)) %>%
-  group_by(classcode,targeted) %>%
-  mutate(smd = (md - mean(md)) / sd(md)) %>%
-  ggplot(aes(year, smd, color = targeted == 1, fill = targeted == 1)) +
-  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red", size = 1.5) +
-  geom_line(aes(group = interaction(targeted, classcode)), alpha = 0.5) +
-  geom_smooth(size = 2) +
-  labs(y = "Centered and Scaled Mean Density", x = "Year") +
-    scale_color_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
-      scale_fill_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
-    labs(y = "Centered and Scaled Mean Density", x = "Year") + 
-  theme(legend.position = "top",
-        panel.grid.major.x = element_blank(),
-        panel.grid.minor.x = element_blank()
-        )
-
-raw_did_plot <- raw_did_plot +
-      scale_color_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "") +
-        scale_fill_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "")
-
-pdf(file =  "figs/pop-trends.pdf",width = 8, height = 5, useDingbats = TRUE)
-print(raw_did_plot)
-dev.off()
-
-```
-
-\begin{figure}%[tbhp]
-  \centering
-  \includegraphics[width=1\linewidth]{figs/pop-trends.pdf}
-  \caption{Centered and scaled mean annual biomass densities of included targeted (red) and non-targeted (blue) fish species (faded lines) and smoothed means with 95\% confidence intervals over time (darker lines and ribbons).
-Vertical red dashed line shows year of MPA implementation.}
-  \label{pop-trends}
-\end{figure}
-
-Using this model, we estimate...
-XX.
-As we might expect from bio-economic modeling, we detect no significant difference between targeted and non-targeted biomass densities for the first XX years of protection.
-However, over time the model assigns greater and greater probability to MPA-mediated increases in targeted fin-fish biomass densities, up until XX, at which point th trend reverses itself.
-We detect no significant MPA effect between 2015 and 2018 (Fig.\@ref(fig:)).
-
-<!-- The raw trends in the biomass densities appear to show parallel trends in the densities of targeted and non-targeted fishes pre-MPA, with the targeted group diverging first positively and then negatively over time (Fig.\ref{pop-trends}).
-However, these raw trends could be caused by any number of confounding variables besides the effects of the MPAs themselves, from changes in observer skill and sampling locations, to shifts in environmental conditions.
-Using our difference-in-difference model in an effort to control for these factors, we found no significant difference in the biomass densities of targeted and non-targeted fishes pre-MPA (before 2003).
-Following the implementation of the MPAs in 2003, we see evidence of an increasing trend in densities of targeted relative to non-targeted fish, reaching a peak value in 2014.
-After 2014, densities of targeted species appear to have declined relative to what we would expect based on the non-targeted group (Fig.\ref{did-plot}).
--->
-
-```{r did-plot, include = FALSE}
-
-did_plot <- did_plot +
-  scale_y_continuous(name = "~Divergence from Non-Targeted", labels = scales::percent_format(accuracy = 1))
-
-ggsave(did_plot, filename = "figs/did_plot.png",
-       width = 8, height = 5)
-
-# pdf(file =  "figs/did_plot.pdf",width = 8, height = 5, useDingbats = TRUE)
-# print(did_plot)
-# dev.off()
-
-```
-
-\begin{figure}%[tbhp]
-  \centering
-  \includegraphics[width=1\linewidth]{figs/did_plot.png}
-  \caption{Estimated divergence in biomass densities of targeted and non-targeted fishes throughout the Channel Islands (integrated across inside and outside of MPAs).
-MPAs are implemented in 2003 (red dashed line).
-Colored cells in the background indicate the frequency of MPA conservation effects predicted by our simulation modeling for MPAs similar in design and species composition to the Channel Islands in each year.}
-  \label{did-plot}
-\end{figure}
-
-<!-- To the extent that we can select control sites that are both sufficiently similar to the region treated with MPAs and are unaffected by biological or economic spillover, response ratios may be reliable estimators of regional conservation effects.
-In reality these conditions are unlikely to hold but for a few select cases.
-This becomes especially difficult when networks of MPAs are implemented, such as has occurred in California.
-In this case, control sites are generally located within a few kilometers from MPA borders, making biological spillover and fishing concentration highly likely.
-In these circumstances, without robust statistical controls, response ratios may provide an unreliable estimate of the regional conservation effect of marine protected areas.
--->
-
 ```{r dr-vs-mpa-effect-plot, include = FALSE}
@@
 ylabs <- c(expression(""<= "-100%"), paste0(seq(0,300, by = 100),"%"),expression("">= "400%") )
+
+density_ratios %>% 
+  ggplot(aes(years_protected, true_density_ratio)) + 
+  geom_point()
 
@@
 
+<!-- If response ratios are not a broadly applicable solution to the challenge of estimating regional conservation effects, what then can we do?
+How then can we empirically measure the regional effects of MPAs?
+Intuition and bio-economic theory tells us that the regional conservation effects of MPAs are likely to be highly context dependent across locations and across species within locations, so careful thought must be devoted to an empirical design to detect their effects.
+The perfect experiment would involve two parallel worlds that were identical, except for the implementation of an MPA.
+In world “A”, no MPA would be implemented, and in the facsimile world “B”, the MPA would be implemented.
+Both worlds would be tracked before and after MPA implementation, and the conservation and fishery outcomes would be compared after treatment.
+Instead of two parallel worlds, a similar experimental design would involve random placement of MPAs.
+-->
+
+## MPA Effects in the Channel Islands {#estimating .unnumbered}
+
+Given these challenges, what is a viable alternative for estimating the conservation effects of MPAs?
+Building off of @caselle2015, we use a difference-in-difference estimator comparing densities in finfish species targeted by fishing effort (i.e., those potentially affected by an MPA) from those not targeted before and after MPA implementation [see @larsen2019 for a succinct explanation of difference-in-difference estimators in ecological settings].
+Targeted species in the Northern Channel Islands, California, include commercially important finfish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
+Each of these targeted species was the subject of prior bio-economic modeling related to the effects of MPAs in southern California [@costello2010d;@rassweiler2014].
+But, the analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*).
+Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species).
+
+```{r raw-trend, fig.cap="Centered and scaled mean annual biomass density of included species (faded lines) and smoothed means of targeted and non-targeted groups, and mean (darker lines) and 95% confidence interval of the mean (ribbon) over time", include = FALSE}
+raw_did_plot <- base_run$data[[1]] %>%
+  select(year, targeted,classcode, log_density, any_seen) %>%
+  mutate(density = exp(log_density) * any_seen) %>%
+  group_by(classcode,targeted, year) %>%
+  summarise(md = mean(density)) %>%
+  group_by(classcode,targeted) %>%
+  mutate(smd = (md - mean(md)) / sd(md)) %>%
+  ggplot(aes(year, smd, color = targeted == 1, fill = targeted == 1)) +
+  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red", size = 1.5) +
+  geom_line(aes(group = interaction(targeted, classcode)), alpha = 0.5) +
+  geom_smooth(size = 2) +
+  labs(y = "Centered and Scaled Mean Density", x = "Year") +
+    scale_color_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
+      scale_fill_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
+    labs(y = "Centered and Scaled Mean Density", x = "Year") + 
+  theme(legend.position = "top",
+        panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank()
+        )
+
+raw_did_plot <- raw_did_plot +
+      scale_color_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "") +
+        scale_fill_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "")
+
+pdf(file =  "figs/pop-trends.pdf",width = 8, height = 5, useDingbats = TRUE)
+print(raw_did_plot)
+dev.off()
+
+```
+
+\begin{figure}%[tbhp]
+  \centering
+  \includegraphics[width=1\linewidth]{figs/pop-trends.pdf}
+  \caption{Centered and scaled mean annual biomass densities of included targeted (red) and non-targeted (blue) fish species (faded lines) and smoothed means with 95\% confidence intervals over time (darker lines and ribbons).
+Vertical red dashed line shows year of MPA implementation.}
+  \label{pop-trends}
+\end{figure}
+
+Using this model, we estimate...
+XX.
+As we might expect from bio-economic modeling, we detect no significant difference between targeted and non-targeted biomass densities for the first XX years of protection.
+However, over time the model assigns greater and greater probability to MPA-mediated increases in targeted fin-fish biomass densities, up until XX, at which point th trend reverses itself.
+We detect no significant MPA effect between 2015 and 2018 (Fig.\@ref(fig:)).
+
+<!-- The raw trends in the biomass densities appear to show parallel trends in the densities of targeted and non-targeted fishes pre-MPA, with the targeted group diverging first positively and then negatively over time (Fig.\ref{pop-trends}).
+However, these raw trends could be caused by any number of confounding variables besides the effects of the MPAs themselves, from changes in observer skill and sampling locations, to shifts in environmental conditions.
+Using our difference-in-difference model in an effort to control for these factors, we found no significant difference in the biomass densities of targeted and non-targeted fishes pre-MPA (before 2003).
+Following the implementation of the MPAs in 2003, we see evidence of an increasing trend in densities of targeted relative to non-targeted fish, reaching a peak value in 2014.
+After 2014, densities of targeted species appear to have declined relative to what we would expect based on the non-targeted group (Fig.\ref{did-plot}).
+-->
+
+```{r did-plot, include = FALSE}
+
+did_plot <- did_plot +
+  scale_y_continuous(name = "~Divergence from Non-Targeted", labels = scales::percent_format(accuracy = 1))
+
+ggsave(did_plot, filename = "figs/did_plot.png",
+       width = 8, height = 5)
+
+# pdf(file =  "figs/did_plot.pdf",width = 8, height = 5, useDingbats = TRUE)
+# print(did_plot)
+# dev.off()
+
+```
+
+\begin{figure}%[tbhp]
+  \centering
+  \includegraphics[width=1\linewidth]{figs/did_plot.png}
+  \caption{Estimated divergence in biomass densities of targeted and non-targeted fishes throughout the Channel Islands (integrated across inside and outside of MPAs).
+MPAs are implemented in 2003 (red dashed line).
+Colored cells in the background indicate the frequency of MPA conservation effects predicted by our simulation modeling for MPAs similar in design and species composition to the Channel Islands in each year.}
+  \label{did-plot}
+\end{figure}
+
+<!-- To the extent that we can select control sites that are both sufficiently similar to the region treated with MPAs and are unaffected by biological or economic spillover, response ratios may be reliable estimators of regional conservation effects.
+In reality these conditions are unlikely to hold but for a few select cases.
+This becomes especially difficult when networks of MPAs are implemented, such as has occurred in California.
+In this case, control sites are generally located within a few kilometers from MPA borders, making biological spillover and fishing concentration highly likely.
+In these circumstances, without robust statistical controls, response ratios may provide an unreliable estimate of the regional conservation effect of marine protected areas.
+-->
+
 ```{r}
@@
 
-ci_map <- pisco_data %>%
+sample_sites <- pisco_data %>%
   left_join(site_data, by = c("site","side")) %>%
@@
 
-ci_map <-  ci_map %>%
-  dplyr::mutate(geometry = purrr::map2(lon_wgs84, lat_wgs84, ~ sf::st_point(x = c(.x, .y), dim = 'XY'))) %>%
-  ungroup() %>%
-  mutate(geometry = sf::st_sfc(geometry, crs = 4326)) %>%
-  sf::st_sf()
-
-bbox <- sf::st_bbox(ci_map)
-
-pisco_ci_map_plot <-  ggmap(channel_islands, alpha = 0.5) +
-  geom_sf(data = ca_mpas, inherit.aes = FALSE, alpha = 0.5) +
-  geom_sf(data = ci_map,inherit.aes = FALSE,
-  aes(fill = eventual_mpa, color = eventual_mpa),
-  shape = 21, 
-  alpha = 0.5,
-  show.legend = FALSE,
-  size = 1.5) +
-  coord_sf(xlim = c(bbox['xmin'] - .3, bbox['xmax'] + .06),
-  ylim = c(bbox['ymin'] - .1, bbox['ymax'] + .1)) +
+channel_islands <- sf::st_read(here::here("data","cinms_py2","cinms_py.shp"))
+
+rnaturalearth::ne_coastline()
+
+ci_bbox <- sf::st_bbox(channel_islands)
+
+sample_sites <-  sample_sites %>%
+  st_as_sf(coords = c("lon_wgs84", "lat_wgs84"),
+           crs = 4326)
+
+california <- rnaturalearth::ne_states(country = "united states of america", returnclass = "sf") %>% 
+  filter(name == "California") %>% 
+    sf::st_transform(crs = sf::st_crs(channel_islands))
+
+channel_islands_mpas <- sf::st_read(here::here("data","MPA_CA_Existing_160301"))%>%
+  sf::st_transform(crs = sf::st_crs(channel_islands))
+
+pisco_ci_map_plot <-  ggplot() + 
+  geom_sf(data = california, fill = "lightgrey") +
+  geom_sf(data = channel_islands_mpas, fill = "lightblue") +
+ geom_sf(data = sample_sites,shape = 21, fill = "darkgrey",
+                                      size = 2, alpha = .75) +
+
   # scale_color_gradient(low = "white", high = "orangered",
@@
   labs(x = "", y = "") +
-      ggspatial::annotation_scale(location = "br", width_hint = 0.4, text_col = "white") + 
-  ggsci::scale_color_npg() + 
+  scale_y_continuous(limits = c(33.8, 34.2)) +
+  scale_x_continuous(limits = c(-120.75,-119.25)) +
+      ggspatial::annotation_scale(location = "br") + 
+  #   coord_sf(xlim = c(bbox['xmin'] - .3, bbox['xmax'] + .06),
+  # ylim = c(bbox['ymin'] - .1, bbox['ymax'] + .1)) +
     ggsci::scale_color_npg()

### 2020-03-12 functions/estimate_did.R

@@ -306,7 +306,7 @@ estimate_did <-
     #                           .width = c(0.5, 0.95)) +
     #   scale_y_continuous(labels = percent, name = "Estimated MPA Effect") +
     #   scale_x_discrete(name = "Year Bin")
-    
+    # 
     out <- list(did_results = did_results,
                 did_reg = did_reg
                 )

### 2020-03-12 functions/run-mpa-experiment.R

@@ -75,9 +75,11 @@ run_mpa_experiment <- function(fish,
       effort = sum(effort)
     ) %>%
     ungroup()
+  
   raw_outcomes <- no_mpa %>%
-    bind_rows(wi_mpa)
-
+    bind_rows(wi_mpa) %>% 
+    mutate(eventual_mpa = patch %in% mpa_locations)
+  
   out <- list(outcomes = outcomes,
               raw_outcomes = raw_outcomes)
 

### 2020-03-12 make-regional-mpas.R

@@ -49,13 +49,13 @@ run_description <- "PNAS R and R with simplified DiD"
 # So, once you've run simulate_mpas, you can set it to FALSE and validata_mpas will work
 
 
-run_did <- TRUE # run difference in difference on data from the CINMS
+run_did <- FALSE # run difference in difference on data from the CINMS
 
 run_tmb <- FALSE
 
 simulate_mpas <- TRUE # simulate MPA outcomes
 
-validate_mpas <- TRUE
+validate_mpas <- FALSE
 
 process_results <- TRUE
 
@@ -67,7 +67,7 @@ sim_years <- 50
 
 num_patches <- 50
 
-n_cores <- 8
+n_cores <- 10
 
 # prepare run -------------------------------------------------------------
 
@@ -1353,13 +1353,27 @@ model_runs <- model_runs %>%
   
 did_fits <- model_runs %>% 
   select(-data)
+
+plot_did <- function(x){
+  
+  mpa_effect_plot <-  x$did_results %>%
+    ggplot(aes(year, did)) +
+    geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
+    tidybayes::stat_halfeye(alpha = 0.7,
+                            .width = c(0.5, 0.95)) +
+    scale_y_continuous(labels = percent, name = "Estimated MPA Effect") +
+    scale_x_discrete(name = "Year Bin")
+  
+}
+
+did_fits <- did_fits %>% 
+  mutate(did_plot = map(did_fit, plot_did))
   
 # print(object.size(did_fits), units = "Mb")
 
 
 write_rds(did_fits, path = file.path(run_dir,"did_fits.rds"))
 
-test <- read_rds( file.path(run_dir,"did_fits.rds"))
 }
 # simulate mpa outcomes ---------------------------------------------------
  if (simulate_mpas == TRUE){
@@ -1376,7 +1390,7 @@ test <- read_rds( file.path(run_dir,"did_fits.rds"))
 
    create_grid <- TRUE
 
-   samps <- 20
+   samps <- 15000
 
    grid_search <-  FALSE
 
@@ -2025,8 +2039,11 @@ if (process_results == TRUE){
   write_rds(outcomes, file.path(run_dir,"outcomes.rds"))
 
   density_ratios <- processed_grid %>%
-    select(-fishery_effect) %>%
-    unnest() %>%
+    select(-absolute_mpa_effect, fishery_effect) %>%
[…]

### 2020-03-13 documents/ovando-regional-effects-of-mpas-v2.Rmd
hopefully resolved major git blowup

@@
 params:
-  run_name: ["v6.0"]
+  run_name: ["temp"]
   sim_years: [50]
@@
 
+*as we move away from "set it and forget it" management towards adaptive management of marine ecosystems, it is critically important that we develop robust strategies for monitoring the effectiveness of our management intercentions*
+
 No-take Marine Protected Areas (MPAs), spatial regions of the ocean in which fishing is prohibited, have a long history in the management of marine resources.
@@
 Numerous factors can affect the regional effects of MPAs.
-These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008], larval dispersal patternss and the strength and timing of density dependence in the population (e.g.
-pre- or post-settlement, @burgess), how overfished the population would be without the MPA, and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @gerber2005; @hilborn2004a; @gaines2010; @botsford2003;@smith2003;@smith2006].
+These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], larval dispersal patternss and the strength and timing of density dependence in the population (e.g.
+pre- or post-settlement, @burgess), how overfished the population would be without the MPA, and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003;@smith2006;@walters2000].
 In addition, even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through habitat and network effects [@gaines2010; @costello2010].
@@
 Given this existing body of theory, what should we expect *a priori* the effect size of the Channel Islands MPAs to be?
-The species are xx.
-We utilize a spatial bio-econonomic model to simulate the expected effects of an MPA network stylized to resemble the Channel Islands (See SI for details of simulation framework).
-For this group, MPA theory suggests that regional effects on the order of XX% are reasonable for such a system.
-
-Although the Islands do not encompass the full population of any of these species, they encompass a large enough area to assess the regional MPA conservation effects we might see in the Channel Islands after 14 years of protection (the time span covered by our model).
-The median regional conservation effect produced by our simulation model after 13 years of protection by MPAs covering between 15%-20% of a population's range was `r percent(med_ci_effect, accuracy = 1)`.
+The targeted species in our database span a range of life histories, but are largely made up of perches and rockfish.
+The size of the MPA relative to the adult and larval dispersal of the species, as well as the level of fishing pressure pre and post MPA implemntation are two critical drivers of MPA effect sizes.
+The MPAs in this study cover 21% of the the waters of the Northern Channel Islands, a region spanning XX km.
+While detailed dispersal studies are not available for all of these species, what information we have suggest that while adults of some of these species are likely to exhibit some site fidelity, larvae are likely distributed beyond the Channel Islands.
+Therefore, if we assume that the population of the targeted species is at least equal to the extent of the Northern Channel Islands, we can assume that at most the MPAs cover roughly 20% of the targeted populations' ranges.
+Turning to the critical question of fishing pressure, formal stock assessments are largely lacking for these species.
+However, what evidence we have suggests mostly moderate fihing pressure, with some species such as California sheephead experiencing high levels of fishing mortality pre-MPA, and others such as blue rockfish only moderate XX.
+
+Considering the two variables of coverage and fishing pressure then, we might expect that the Channel Islands would produce a small to moderate effect size over time.
+However, theory also tells us that many other variables can interact to affect the outcomes of MPAs.
+To address this, we utilized a spatial bio-econonomic model to simulate the expected effects of an MPA network stylized to resemble the Channel Islands on the group of targeted species covered in our analysis (See SI for details of simulation framework).
+For this group, the median regional conservation effect produced by our simulation model after 14 years of protection by MPAs covering between XX%-20% of a population's range was `r percent(med_ci_effect, accuracy = 1)`.
 This projected effect has an an interquartile range of `r percent(quart_ci_effect[2], accuracy = 1)` and `r percent(quart_ci_effect[4], accuracy = 1)`, although values as high as a 200% increase and as low as a 50% decrease were also produced.
@@
 
-What evidence do we have  for the effects of the Channel Island MPAs?
+Having established a theoretical grounding for the likely effect of the Channel Island MPAs, what evidence do we have for the empirical effects of this protected area network?
 @caslle2015 examined changes in biomass densities of species targeted by fishing activity inside and outside of the MPAs over time (generally termed "response ratios"), and compared these changes to the trends in non-targeted species.
 They found a statistically significant increase in the response ratios of targeted speices over time, and evidence that this increase is smaller in the non-targeted species.
-Updating the results of @caselle2015 through 2018 shows a continuation of this trend (Fig.\@ref(fig:XX)).
+Updating the results of @caselle2015 through 2017 shows the same increasing trend in the resonse ratios of targeted species (Fig.\@ref(fig:XX)).
 
@@
 
-In the case of the Channel Islands MPAs, control sites are often located within XX Km of the MPAs, suggesting that they are susceptible to biological spillover and concentration of fishing effort resulting from the MPAs.
-Given these complications, we can certainly attribute the response ratios initially reported in @caselle2015 as evidence that the MPAs are indeed providing effective protection of targeted fish biomass inside their borders.
+In the case of the Channel Islands MPAs, control sites are often located within only a few kilometers of an MPA, suggesting that they are susceptible to biological spillover and concentration of fishing effort resulting from the MPAs.
+Given these complications, we can certainly interpret the response ratios reported in @caselle2015 and updated here as evidence that the MPAs are indeed providing effective protection of targeted fish biomass inside their borders.
 However, it is unclear how well these response ratio results serve as in indicator of overall regional effects of the MPAs.
 
-To illustrate this problem, we used our stylized bio-economic model to simulate response ratio trajectories for the species and MPA coverage present in the Channel Islands.
-We then compared the range of plausible response ratios generated by our model to the true regional effect of the MPAs in each simulation.
+To illustrate this problem, we used our bio-economic model to simulate response ratio trajectories for the species and MPA coverage present in the Channel Islands.
+We then compared the range of plausible response ratios generated by our model to the true regional effect of the MPAs in each simulation over time.
 When response ratios were calculated from true control sites completely unaffected by the MPAs, they are a highly reliable measure of regional conservation effects.
 However, once we allow for the possibility of MPAs affecting the control sites through ecological and economic mechanisms, response ratios were generally a poor predictor of regional conservation effects (Fig.\@ref(fig:)).
-This was particularly true in the early years of MPA implementation; while biomass increases caused by MPAs are constrained by the growth rates of fish individuals and populations, reductions in biomass caused by concentrated fishing outside of reserves can occur quite quickly.
+This was particularly true in the early years of MPA implementation; while biomass increases caused by MPAs are constrained by the growth rates of fish individuals and populations, reductions in biomass caused by concentrated fishing outside of reserves can occur quite quickly XX.
 As a result, response ratios can indeed serve as a measure of MPAs ability to provide effective within-border protection, but particular in the first XX years they are a highly imprecise measure of the net effect of the MPAs at a regional scale.
+
+xx new figure.
+time on x axis.
+posteriors distribution of desponse ratio on the y axis.
+For matching response ratios in each time bin -STORY: in a perfect world response ratios over time are are a reasonable proxy for regional MPA effects, but not in a realistic world .
+KISS.
+Ridgeplot of respoonse ratio, biased DR, and unbiased DR over time, where simulated DR are sampled in proportion to observed response ratio
 
@@
 
-## MPA Effects in the Channel Islands {#estimating .unnumbered}
-
-Given these challenges, what is a viable alternative for estimating the conservation effects of MPAs?
+<!-- ## MPA Effects in the Channel Islands {#estimating .unnumbered} -->
+
+Given these challenges, what is a viable alternative for estimating the regional conservation effects of MPAs?
 Building off of @caselle2015, we use a difference-in-difference estimator comparing densities in finfish species targeted by fishing effort (i.e., those potentially affected by an MPA) from those not targeted before and after MPA implementation [see @larsen2019 for a succinct explanation of difference-in-difference estimators in ecological settings].
@@
 
-```{r raw-trend, fig.cap="Centered and scaled mean annual biomass density of included species (faded lines) and smoothed means of targeted and non-targeted groups, and mean (darker lines) and 95% confidence interval of the mean (ribbon) over time", include = FALSE}
-raw_did_plot <- base_run$data[[1]] %>%
-  select(year, targeted,classcode, log_density, any_seen) %>%
-  mutate(density = exp(log_density) * any_seen) %>%
-  group_by(classcode,targeted, year) %>%
-  summarise(md = mean(density)) %>%
-  group_by(classcode,targeted) %>%
-  mutate(smd = (md - mean(md)) / sd(md)) %>%
-  ggplot(aes(year, smd, color = targeted == 1, fill = targeted == 1)) +
-  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red", size = 1.5) +
-  geom_line(aes(group = interaction(targeted, classcode)), alpha = 0.5) +
-  geom_smooth(size = 2) +
-  labs(y = "Centered and Scaled Mean Density", x = "Year") +
-    scale_color_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
-      scale_fill_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
-    labs(y = "Centered and Scaled Mean Density", x = "Year") + 
-  theme(legend.position = "top",
-        panel.grid.major.x = element_blank(),
-        panel.grid.minor.x = element_blank()
-        )
-
-raw_did_plot <- raw_did_plot +
-      scale_color_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "") +
-        scale_fill_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "")
-
-pdf(file =  "figs/pop-trends.pdf",width = 8, height = 5, useDingbats = TRUE)
-print(raw_did_plot)
-dev.off()
-
-```
-
-\begin{figure}%[tbhp]
-  \centering
-  \includegraphics[width=1\linewidth]{figs/pop-trends.pdf}
-  \caption{Centered and scaled mean annual biomass densities of included targeted (red) and non-targeted (blue) fish species (faded lines) and smoothed means with 95\% confidence intervals over time (darker lines and ribbons).
-Vertical red dashed line shows year of MPA implementation.}
-  \label{pop-trends}
-\end{figure}
-
-Using this model, we estimate...
-XX.
-As we might expect from bio-economic modeling, we detect no significant difference between targeted and non-targeted biomass densities for the first XX years of protection.
-However, over time the model assigns greater and greater probability to MPA-mediated increases in targeted fin-fish biomass densities, up until XX, at which point th trend reverses itself.
-We detect no significant MPA effect between 2015 and 2018 (Fig.\@ref(fig:)).
-
-<!-- The raw trends in the biomass densities appear to show parallel trends in the densities of targeted and non-targeted fishes pre-MPA, with the targeted group diverging first positively and then negatively over time (Fig.\ref{pop-trends}).
-However, these raw trends could be caused by any number of confounding variables besides the effects of the MPAs themselves, from changes in observer skill and sampling locations, to shifts in environmental conditions.
-Using our difference-in-difference model in an effort to control for these factors, we found no significant difference in the biomass densities of targeted and non-targeted fishes pre-MPA (before 2003).
-Following the implementation of the MPAs in 2003, we see evidence of an increasing trend in densities of targeted relative to non-targeted fish, reaching a peak value in 2014.
-After 2014, densities of targeted species appear to have declined relative to what we would expect based on the non-targeted group (Fig.\ref{did-plot}).
--->
-
-```{r did-plot, include = FALSE}
-
-did_plot <- did_plot +
-  scale_y_continuous(name = "~Divergence from Non-Targeted", labels = scales::percent_format(accuracy = 1))
-
-ggsave(did_plot, filename = "figs/did_plot.png",
-       width = 8, height = 5)
-
-# pdf(file =  "figs/did_plot.pdf",width = 8, height = 5, useDingbats = TRUE)
-# print(did_plot)
-# dev.off()
-
-```
-
-\begin{figure}%[tbhp]
-  \centering
-  \includegraphics[width=1\linewidth]{figs/did_plot.png}
-  \caption{Estimated divergence in biomass densities of targeted and non-targeted fishes throughout the Channel Islands (integrated across inside and outside of MPAs).
-MPAs are implemented in 2003 (red dashed line).
-Colored cells in the background indicate the frequency of MPA conservation effects predicted by our simulation modeling for MPAs similar in design and species composition to the Channel Islands in each year.}
-  \label{did-plot}
-\end{figure}
-
-<!-- To the extent that we can select control sites that are both sufficiently similar to the region treated with MPAs and are unaffected by biological or economic spillover, response ratios may be reliable estimators of regional conservation effects.
-In reality these conditions are unlikely to hold but for a few select cases.
-This becomes especially difficult when networks of MPAs are implemented, such as has occurred in California.
-In this case, control sites are generally located within a few kilometers from MPA borders, making biological spillover and fishing concentration highly likely.
-In these circumstances, without robust statistical controls, response ratios may provide an unreliable estimate of the regional conservation effect of marine protected areas.
--->
-
-```{r}
-fish_v_fishing <- processed_grid %>%
-  select(-density_ratio) %>%
-  unnest() %>%
-  rename(fishery_effect = mpa_effect1) %>%
-  group_by(experiment) %>%
-  mutate(year = 1:length(year)) %>%
-  ungroup() %>%
-  mutate(years_protected = year - year_mpa + 1) %>%
-  left_join(outcomes %>% select(experiment, year, depletion),
-            by = c("experiment", "year"))
-
-fish_v_fishing_plot <- fish_v_fishing %>%
-  filter(years_protected == max(years_protected)) %>%
-  ggplot(aes(pmin(2, mpa_effect), pmin(2, fishery_effect))) +
-  geom_abline(aes(slope = 1, intercept = 0)) +
-  geom_smooth(method = "lm") +
-  geom_hex()
-
-```
+The key assumption of this estimation strategy is that the non-targeted species serve as an effective control for the trajectory the biomass densities of targeted species would have taken in the absence of the MPAs (see SI for support for this assumption).
+This allows us to control for potential environmental shocks to the region, without the assumption that fished areas are unaffected by the MPA required to interpret response ratios as a measure of regional effects (though we do need to assume that non-targeted species are unaffected by the MPA, and that targeted and non-targeted species respond to enviornmental shocks in the same way, see SI).
+The result of this regression is an estimate of the effect of the MPAs on the average biomass densities of targeted species throughout the Channel Islands.
 
@@
 
+```{r raw-trend, fig.cap="Centered and scaled mean annual biomass density of included species (faded lines) and smoothed means of targeted and non-targeted groups, and mean (darker lines) and 95% confidence interval of the mean (ribbon) over time", include = FALSE}
+raw_did_plot <- base_run$data[[1]] %>%
+  select(year, targeted,classcode, log_density, any_seen) %>%
+  mutate(density = exp(log_density) * any_seen) %>%
+  group_by(classcode,targeted, year) %>%
+  summarise(md = mean(density)) %>%
+  group_by(classcode,targeted) %>%
+  mutate(smd = (md - mean(md)) / sd(md)) %>%
+  ggplot(aes(year, smd, color = targeted == 1, fill = targeted == 1)) +
+  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red", size = 1.5) +
+  geom_line(aes(group = interaction(targeted, classcode)), alpha = 0.5) +
+  geom_smooth(size = 2) +
+  labs(y = "Centered and Scaled Mean Density", x = "Year") +
+    scale_color_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
+      scale_fill_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
+    labs(y = "Centered and Scaled Mean Density", x = "Year") + 
+  theme(legend.position = "top",
+        panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank()
+        )
+
+raw_did_plot <- raw_did_plot +
+      scale_color_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "") +
+        scale_fill_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "")
+
+pdf(file =  "figs/pop-trends.pdf",width = 8, height = 5, useDingbats = TRUE)
+print(raw_did_plot)
+dev.off()
+
+```
+
+\begin{figure}%[tbhp]
+  \centering
+  \includegraphics[width=1\linewidth]{figs/pop-trends.pdf}
+  \caption{Centered and scaled mean annual biomass densities of included targeted (red) and non-targeted (blue) fish species (faded lines) and smoothed means with 95\% confidence intervals over time (darker lines and ribbons).
+Vertical red dashed line shows year of MPA implementation.}
+  \label{pop-trends}
+\end{figure}
+
+As we expected from our bio-economic modeling, we estimate a small but likely positive effect of the MPAs over the first three years of protection (median 30%, (95% credible interval XX to XX)).
+By the period 2009-2012, the median estimated effect size climbed to 70% (95% credible interval XX to XX), well in line with reasonable effect sizes predicted by theory.
+However, in the subsequent years the estimated effect size begins to dissapear, ending with a median effect of -10% for the years 2015-2018 (95% credible inverval XX to XX) (Fig.\@ref(fig:)).
+
+<!-- The raw trends in the biomass densities appear to show parallel trends in the densities of targeted and non-targeted fishes pre-MPA, with the targeted group diverging first positively and then negatively over time (Fig.\ref{pop-trends}).
+However, these raw trends could be caused by any number of confounding variables besides the effects of the MPAs themselves, from changes in observer skill and sampling locations, to shifts in environmental conditions.
+Using our difference-in-difference model in an effort to control for these factors, we found no significant difference in the biomass densities of targeted and non-targeted fishes pre-MPA (before 2003).
+Following the implementation of the MPAs in 2003, we see evidence of an increasing trend in densities of targeted relative to non-targeted fish, reaching a peak value in 2014.
+After 2014, densities of targeted species appear to have declined relative to what we would expect based on the non-targeted group (Fig.\ref{did-plot}).
+-->
+
+```{r did-plot, include = FALSE}
+
+base_run <- model_runs %>% 
+  filter(data_source == "pisco", var_names == "pisco_a", data_to_use == "all")
+
+did_plot <- base_run$did_plot[[1]] +
+  scale_y_continuous(name = "% Change in Biomass Density", labels = scales::percent_format(accuracy = 1))
+
+ggsave(did_plot, filename = "figs/did_plot.png",
+       width = 8, height = 5)
+
+# pdf(file =  "figs/did_plot.pdf",width = 8, height = 5, useDingbats = TRUE)
+# print(did_plot)
+# dev.off()
+
+```
+
+\begin{figure}%[tbhp]
+  \centering
+  \includegraphics[width=1\linewidth]{figs/did_plot.png}
+  \caption{Estimated divergence in biomass densities of targeted and non-targeted fishes throughout the Channel Islands (integrated across inside and outside of MPAs).
+MPAs are implemented in 2003 (red dashed line).
+Colored cells in the background indicate the frequency of MPA conservation effects predicted by our simulation modeling for MPAs similar in design and species composition to the Channel Islands in each year.}
+  \label{did-plot}
+\end{figure}
+
+<!-- To the extent that we can select control sites that are both sufficiently similar to the region treated with MPAs and are unaffected by biological or economic spillover, response ratios may be reliable estimators of regional conservation effects.
+In reality these conditions are unlikely to hold but for a few select cases.
+This becomes especially difficult when networks of MPAs are implemented, such as has occurred in California.
+In this case, control sites are generally located within a few kilometers from MPA borders, making biological spillover and fishing concentration highly likely.
+In these circumstances, without robust statistical controls, response ratios may provide an unreliable estimate of the regional conservation effect of marine protected areas.
+-->
+
+```{r}
+fish_v_fishing <- processed_grid %>%
+  select(-density_ratio) %>%
+  unnest() %>%
+  rename(fishery_effect = mpa_effect1) %>%
+  group_by(experiment) %>%
+  mutate(year = 1:length(year)) %>%
+  ungroup() %>%
+  mutate(years_protected = year - year_mpa + 1) %>%
+  left_join(outcomes %>% select(experiment, year, depletion),
+            by = c("experiment", "year"))
+
+fish_v_fishing_plot <- fish_v_fishing %>%
+  filter(years_protected == max(years_protected)) %>%
+  ggplot(aes(pmin(2, mpa_effect), pmin(2, fishery_effect))) +
+  geom_abline(aes(slope = 1, intercept = 0)) +
+  geom_smooth(method = "lm") +
+  geom_hex()
+
+```
+
 ```{r}
@@
 2. Put the channel islands results in context -
+
+The network of MPAs implemented in the Channel Islands in 2003 provide an ideal opportunity to examine the regional effects of proteced area networks.
+This analysis is to our knowledge the first empirical estimate of the regional effects of such a large MPA network on a broad assemblage of targeted finfish.
+Despite the well enforced nature of these MPAs (XX), and the presence of a rigorous scientific monitoring program, we are unable to estimate a clear effect of protection.
+While response ratios have targeted species have increased over time, providing evidence of effective *within MPA* protection, these response ratios are unlikely to be a realible indicator of the regional conservation effects.
+Using our difference-in-difference estimator, we estimate high probability for increasinly positivei MPA effects over the first nine years of protection, only to see these gains reversed from 2012 to 2017.
+For the most recent time period, we can detect no clear positive or negative effect of the Channel Island MPAs.
+
+How can we explain these results?
+The estimated effects from 2003 to 2012 are well in line with theoretical expectations given the species and network in question.
+How might we explain the most recent results though?
+One explanation may lie in fleet dynamics.
+Much of the theoretical literautre on MPAs assumes that reserves always produce net conservation gains: the only question is how much [@white2011].
+However, to our knowledge all of these models simulate fleet dynamics through assumptions about fishing mortality rates (e.g.
+concentration of fishing mortality @halpern2004).
+The assumption of these models is that fishers determine an amount off effort to exert, and distribute that effort outside the MPA in response to some function.
+
+An alternative an to our knowledge unexplored fleet model though is a "constant-catch" strategy.
+Under this model, fishers have a catch objective, and exert as much (or little) effort as needed to achieve that objective.
+While a constant-catch greater than MSY is not possible over the long-term under the assumptions of our model, over the short-term a constant-catch scenario is not implausible.
+Subsistence fisheries may use a constant-catch style policy over the short-term, as they seek to ensure that their food needs are met [XX].
+More industrial fisheries may have pre-arranged agreements with buyers to deliver set amounts of fish.
+Constant-catch dynamics might also occur in fisheries with constraining quotas that are not updated after the implementation of MPAs.
+Interestingly, when we simulated the effects of MPAs under different fleet models, the "constant-catch" scenario stood out as the only way for MPAs to actually produce a net conservation loss.
+While open-access fishing strategies can result in "scorched earth" scenarios where the only fish left are found inside the reserve, across all XX simulations the net effect of the reserves was still positive.
+Under a constant-catch scenario though, fishers have to fish much harder than before to get the same catch from a smaller part of the population, reducing the size structure of the population and producing a net conservation loss (Fig.XX).
+
+Can the the non-trivial probability that the Channel Islands MPAs caused a net conservation loss (taking the results at face value) in the most recent years be explained by the local fleets following a constant-catch strategy?
+While we do not have access to fine scale fishing data from the islands directly, reported catches for the species of interest in the Santa Barbara region in fact exhibit an overall downward trend in the years post reserve (see SI, XX).
+What then is another explanation?
+The parallel trends assumption of our model appears to be valid in the pre-MPA years, and we do not detect any clear evidence of interaction between the targeted and non-targeted species (see SI).
+Hoever, the Santa Barbara Channel experience a massive heatwave in XX.
+In addition, for many of the targeted species in our database the Channel Islands are at the warmer edge of their range.
+Conversely, many of the non-targeted species are sub-tropical.
+As a result, we hypothesize that the blob even of XX broke down the parllel trends assumption between the targeted and non-targeted species in our study, as the targeted species were more negatively affected by the warming temperatures.
+This is supported by the fact that we observe similar declines in targeted species both outside and inside the reserves.
+Assuming that the reserves are sufficiently well enforced and large to protect targeted species biomass, if the cause of recent declines was due to increases in fishing pressure we would expect to see substantial declines only in the fished areas (Fig.
+XX).
 

### 2020-03-13 make-regional-mpas.R

@@ -40,7 +40,7 @@ walk(functions, ~ here::here("functions", .x) %>% source()) # load local functio
 
 # options -----------------------------------------------------------------
 
-run_name <- 'v6.0'
+run_name <- 'temp'
 
 run_description <- "PNAS R and R with simplified DiD"
 
@@ -49,11 +49,11 @@ run_description <- "PNAS R and R with simplified DiD"
 # So, once you've run simulate_mpas, you can set it to FALSE and validata_mpas will work
 
 
-run_did <- FALSE # run difference in difference on data from the CINMS
+run_did <- TRUE # run difference in difference on data from the CINMS
 
 run_tmb <- FALSE
 
-simulate_mpas <- TRUE # simulate MPA outcomes
+simulate_mpas <- FALSE # simulate MPA outcomes
 
 validate_mpas <- FALSE
 
@@ -716,7 +716,7 @@ if (file.exists(here(data_dir,'pdo.csv'))) {
 # convert transect data to density estimates ------------------------------
 
 
-if (file.exists(file.path(run_dir,"pisco-data.Rdata")) == F |
+if (file.exists(here("data","pisco-data.Rdata")) == F |
     run_length_to_density == T) {
 
 
@@ -830,11 +830,11 @@ if (file.exists(file.path(run_dir,"pisco-data.Rdata")) == F |
     stop('multiple species per classcode')
   }
 
-  save(file = file.path(run_dir,'pisco-data.Rdata'),
+  save(file = here("data",'pisco-data.Rdata'),
        pisco_data)
 
 } else {
-  load(file.path(run_dir,'pisco-data.Rdata'))
+  load(here("data",'pisco-data.Rdata'))
 
 }
 
@@ -1321,8 +1321,6 @@ if (run_tmb == T){
 
 } else {
   
-  
-
   # load(file = file.path(run_dir, 'model_runs.Rdata'))
 }
 model_runs <- model_runs %>%
@@ -1390,7 +1388,7 @@ write_rds(did_fits, path = file.path(run_dir,"did_fits.rds"))
 
    create_grid <- TRUE
 
-   samps <- 15000
+   samps <- 20 
 
    grid_search <-  FALSE
 

### 2020-03-19 make-regional-mpas.R
figures rescued

@@ -29,7 +29,9 @@ library(rEDM)
 library(bayesplot)
 library(tabulizer)
 library(tidyverse)
+library(ggtext)
 extrafont::loadfonts()
+# extrafont::font_import()
 rstan_options(auto_write = TRUE)
 
 
@@ -88,9 +90,9 @@ plot_theme <- hrbrthemes::theme_ipsum(base_size = 14,
 
 theme_set(plot_theme)
 
-fig_name <- "presentations"
+fig_name <- "paper"
 
-fig_width <- 12
+fig_width <- 8
 
 fig_height <- fig_width / 1.333
 
@@ -2065,7 +2067,7 @@ if (process_results == TRUE){
     mutate(year = 1:length(year)) %>%
     ungroup() %>%
     mutate(years_protected = year - year_mpa + 1) %>%
-    mutate(mpa_effect = pmax(-.5,mpa_effect)) %>%
+    mutate(mpa_effect = pmax(-.5,pmin(mpa_effect,2.5))) %>%
     group_by(experiment) %>%
     mutate(b0 = `no-mpa`[year == min(year)]) %>%
     mutate(depletion = pmax(0,1 - `no-mpa`/b0)) %>%
@@ -2446,7 +2448,7 @@ if (process_results == TRUE){
   pop_depletion_plot <- outcomes %>%
     filter(years_protected == short_frame) %>%
     ggplot() +
-    geom_bin2d(aes(depletion, mpa_effect), binwidth = c(.05, .05),
+    geom_bin2d(aes(depletion, mpa_effect), binwidth = c(.05, .2),
                show.legend = FALSE) +
     scale_fill_viridis(
       option = "A",
@@ -2511,7 +2513,7 @@ if (process_results == TRUE){
   pop_size_plot <- outcomes %>%
     filter(years_protected == short_frame) %>%
     ggplot() +
-    geom_bin2d(aes(mpa_size, mpa_effect), binwidth = c(.05, .05), show.legend = TRUE) +
+    geom_bin2d(aes(mpa_size, mpa_effect), binwidth = c(.05, .2), show.legend = TRUE) +
     scale_fill_viridis(
       option = "A",
       trans = plot_trans,
@@ -2528,11 +2530,11 @@ if (process_results == TRUE){
     select(mpa_effect, mpa_size, depletion) %>%
     gather(measure, value, -mpa_effect) %>%
     ggplot() +
[…]

### 2020-03-24 Untitled/Untitled.Rmd
MAJOR restructure of paper, bayesian DiD up and running, added in "ability to estiamte" test, paper 80% restructured

This PNAS journal template is provided to help you write your work in the
correct journal format. Instructions for use are provided below.

Note: please start your introduction without including the word
"Introduction" as a section heading (except for math articles in the
Physical Sciences section); this heading is implied in the first
paragraphs.

Guide to using this template {.unnumbered}
========================================

Please note that whilst this template provides a preview of the typeset
manuscript for submission, to help in this preparation, it will not
necessarily be the final publication layout. For more detailed
information please see the [PNAS Information for
Authors](http://www.pnas.org/site/authors/format.xhtml).

Author Affiliations {#author-affiliations .unnumbered}
-------------------

Include department, institution, and complete address, with the
ZIP/postal code, for each author. Use lower case letters to match
authors with institutions, as shown in the example. Authors with an
ORCID ID may supply this information at submission.

Submitting Manuscripts {#submitting-manuscripts .unnumbered}
----------------------

All authors must submit their articles at
[PNAScentral](http://www.pnascentral.org/cgi-bin/main.plex). If you are
using Overleaf to write your article, you can use the "Submit to PNAS"
option in the top bar of the editor window.

Format {#format .unnumbered}
------

Many authors find it useful to organize their manuscripts with the
following order of sections; Title, Author Affiliation, Keywords,
Abstract, Significance Statement, Results, Discussion, Materials and
methods, Acknowledgments, and References. Other orders and headings are
permitted.

Manuscript Length {#manuscript-length .unnumbered}
-----------------

PNAS generally uses a two-column format averaging 67 characters,
including spaces, per line. The maximum length of a Direct Submission
research article is six pages and a PNAS PLUS research article is ten
pages including all text, spaces, and the number of characters displaced
by figures, tables, and equations. When submitting tables, figures,
and/or equations in addition to text, keep the text for your manuscript
under 39,000 characters (including spaces) for Direct Submissions and
72,000 characters (including spaces) for

<!-- Leave these lines as they are at the end of your .Rmd file to ensure placement of methods & acknowledgements sections before the references-->
\showmatmethods
\showacknow
\pnasbreak

### 2020-03-24 Untitled/Untitled.tex
MAJOR restructure of paper, bayesian DiD up and running, added in "ability to estiamte" test, paper 80% restructured

\documentclass[9pt,twocolumn,twoside,]{pnas-new}

%% Some pieces required from the pandoc template
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\templatetype{pnasresearcharticle}  % Choose template

\title{Template for preparing your research report submission to PNAS using
RMarkdown}

\author[a,1,2]{Alice Anonymous}
\author[a,b]{Bob Security}

  \affil[a]{Some Institute of Technology, Department, Street, City, State, Zip}
  \affil[b]{Another University Department, Street, City, State, Zip}

% Please give the surname of the lead author for the running footer
\leadauthor{Anonymous}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{Authors must submit a 120-word maximum statement about the significance
of their research paper written at a level understandable to an
undergraduate educated scientist outside their field of speciality. The
primary goal of the Significance Statement is to explain the relevance
of the work in broad context to a broad readership. The Significance
Statement appears in the paper itself and is required for all research
papers.}

\authorcontributions{Please provide details of author contributions here.}

\authordeclaration{Please declare any conflict of interest here.}

\equalauthors{\textsuperscript{} }

\correspondingauthor{\textsuperscript{} }

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
 \keywords{  one |  two |  optional |  optional |  optional  }

\begin{abstract}
Please provide an abstract of no more than 250 words in a single
paragraph. Abstracts should explain to the general reader the major
contributions of the article. References in the abstract must be cited
in full within the abstract itself and cited in the text.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.

\acknow{Please include your acknowledgments here, set in a single paragraph.
Please do not include any acknowledgments in the Supporting Information,
or anywhere else in the manuscript.}

\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: danovan@uw.edu}

This PNAS journal template is provided to help you write your work in
the correct journal format. Instructions for use are provided below.

Note: please start your introduction without including the word
``Introduction'' as a section heading (except for math articles in the
Physical Sciences section); this heading is implied in the first
paragraphs.

\hypertarget{guide-to-using-this-template}{%
\section*{Guide to using this
template}\label{guide-to-using-this-template}}
\addcontentsline{toc}{section}{Guide to using this template}

Please note that whilst this template provides a preview of the typeset
manuscript for submission, to help in this preparation, it will not
necessarily be the final publication layout. For more detailed
information please see the
\href{http://www.pnas.org/site/authors/format.xhtml}{PNAS Information
for Authors}.

\hypertarget{author-affiliations}{%
\subsection*{Author Affiliations}\label{author-affiliations}}
\addcontentsline{toc}{subsection}{Author Affiliations}

Include department, institution, and complete address, with the
ZIP/postal code, for each author. Use lower case letters to match
authors with institutions, as shown in the example. Authors with an
ORCID ID may supply this information at submission.

\hypertarget{submitting-manuscripts}{%
\subsection*{Submitting Manuscripts}\label{submitting-manuscripts}}
\addcontentsline{toc}{subsection}{Submitting Manuscripts}

All authors must submit their articles at
\href{http://www.pnascentral.org/cgi-bin/main.plex}{PNAScentral}. If you
are using Overleaf to write your article, you can use the ``Submit to
PNAS'' option in the top bar of the editor window.

\hypertarget{format}{%
\subsection*{Format}\label{format}}
\addcontentsline{toc}{subsection}{Format}

Many authors find it useful to organize their manuscripts with the
following order of sections; Title, Author Affiliation, Keywords,
Abstract, Significance Statement, Results, Discussion, Materials and
methods, Acknowledgments, and References. Other orders and headings are
permitted.

\hypertarget{manuscript-length}{%
\subsection*{Manuscript Length}\label{manuscript-length}}
\addcontentsline{toc}{subsection}{Manuscript Length}

PNAS generally uses a two-column format averaging 67 characters,
including spaces, per line. The maximum length of a Direct Submission
research article is six pages and a PNAS PLUS research article is ten
pages including all text, spaces, and the number of characters displaced
by figures, tables, and equations. When submitting tables, figures,
and/or equations in addition to text, keep the text for your manuscript
under 39,000 characters (including spaces) for Direct Submissions and
72,000 characters (including spaces) for

\showmatmethods
\showacknow
\pnasbreak

% Bibliography
% \bibliography{pnas-sample}

\end{document}

### 2020-03-24 documents/Untitled.Rmd
MAJOR restructure of paper, bayesian DiD up and running, added in "ability to estiamte" test, paper 80% restructured

--- 2020-03-11 documents/ovando-regional-effects-of-mpas-v2.Rmd
@@
 
-```{r, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dev = "CairoPDF", dpi = 600,
-                      cache = FALSE)
-# knitr::opts_knit$set(root.dir = "/Users/danovando/PhD/zissou") # not entirely reprodcible hack to deal with knit child in dissertation.. not idealk
-
-```
-
-```{r, include=FALSE}
-library(sf)
-library(ggmap)
-library(viridis)
-library(hrbrthemes)
-library(scales)
-library(patchwork)
-library(rpart)
-library(extrafont)
-library(caret)
-library(ggsci)
-library(rEDM)
-library(tidyverse)
-extrafont::loadfonts()
-functions <- list.files(here::here("functions"))
-
-walk(functions, ~ here::here("functions", .x) %>% source()) # load local functions
-save_plots <- TRUE
-
-# sim_years <- 50
-# 
-# burn_years <- 20
-# 
-# num_patches <- 25
-
-# run_name <- "v5.0"
-
-run_name <- params$run_name
-
-sim_years <- params$sim_years
-
-burn_years <- params$burn_years
-
-num_patches <- params$num_patches
-
-run_dir <- here::here("results", run_name)
-
-experiment_dir <- here::here("results",run_name, "experiments")
-
-load(file = here::here("results",run_name, "rawish_zissou_data.Rdata"))
-
-model_runs <- read_rds( file.path(run_dir,"did_fits.rds"))
-
-# load(file = here::here("results",run_name, "model_runs.Rdata"))
-
-processed_grid <-  read_rds(file.path(run_dir,"filtered_processed_grid.rds"))
-
-load(file = file.path(run_dir ,"sim_grid.Rdata"))
-
-cdfw_catches <-
-  read_csv(file = here::here("data", 'cdfw-catches.csv')) %>%
-  group_by(sci_name, year) %>%
-  summarise(catch = sum(pounds_caught, na.rm = T)) %>%
-  left_join(
-    life_history_data %>% select(taxa, classcode, commonname) %>% mutate(taxa = tolower(taxa)),
-    by = c("sci_name" = "taxa")
-  ) %>%
-  filter(!is.na(classcode))
-
-# cdfw_catches %>%
-#   group_by(commonname, classcode, sci_name) %>%
-#   summarise(catch =sum(catch, na.rm = TRUE)) %>%
-#   mutate(in_analysis = classcode %in% unique(cip_data$classcode)) %>%
-#   View()
-
-density_ratios <- readRDS(file.path(run_dir,"density_ratios.rds"))
-
-outcomes <- readRDS(file.path(run_dir,"outcomes.rds"))
-
-outcomes <- outcomes %>% 
-  filter(!is.na(mpa_effect))
-
-density_ratios <- density_ratios %>% 
-  filter(!is.na(biased_density_ratio))
-
-   eqo <- outcomes %>%
-     filter(year == max(year))
-
-facet_labels <- c(
-  mpa_size = "Range in MPA",
-  depletion = "Depletion"
-)
-
-channel_islands <- readRDS(here::here("data","channel_islands_map.RDS"))
-
-ca_mpas <- sf::st_read(here::here("data","MPA_CA_Existing_160301")) %>%
-  rmapshaper::ms_simplify() %>%
-  sf::st_transform(crs = 4326)
-
-zissou_theme <-
-  theme_ipsum(
-    base_size = 12,
-    axis_title_size = 12,
-    strip_text_size = 12,
-    base_family = "Fira Sans"
-  )
-
-testplot <- ggplot()
-
-panel_height = unit(1,"npc") - sum(ggplotGrob(testplot)[["heights"]][-3]) - unit(4,"line")
-
-gc <- guide_colorbar(frame.colour = "black",
-                     ticks.colour = "white",
-                     barheight = panel_height)
-
-  hgc <- guide_colorbar(frame.colour = "black",
-                       ticks.colour = "black",
-                       barwidth = 10)
-
-    wide_hgc <- guide_colorbar(frame.colour = "black",
-                       ticks.colour = "black",
-                       barwidth = 25)
-
-theme_set(zissou_theme)
-
-# models_worked <- model_runs$tmb_fit %>% map("error") %>% map_lgl(is_null)
-
-# model_runs <- model_runs %>%
-#   filter(models_worked) %>%
-#   mutate(tmb_fit = map(tmb_fit,"result")) %>%
-#   mutate(processed_fits = map(tmb_fit, process_fits)) %>%
-#   mutate(did_plot = map(processed_fits, "did_plot"))
-
-base_run <- model_runs %>%
-filter(var_names == "pisco_a", data_to_use == "all", center_scale == TRUE)
-
-mpa_run <- model_runs %>%
-  filter(data_source == "pisco",
-         var_names == "pisco_a",
-         data_to_use == "mpa_only")
-
-fishsed_run <- model_runs %>%
-  filter(data_source == "pisco",
-         var_names == "pisco_a",
-         data_to_use == "fished_only")
-
-kfm_run <- model_runs %>%
-  filter(data_source == "kfm")
-
-fitted_data <- abundance_data$data[abundance_data$data_source == "pisco"][[1]]
-
-site_data <- read_csv(here::here("data",'Final_Site_Table_UCSB.csv')) %>%
-  magrittr::set_colnames(., tolower(colnames(.))) %>%
-  select(
-    site,
-    side,
-    mpagroup,
-    mpa_status,
-    reserve,
-    region,
-    year_mpa,
-    mpaareanm2,
-    lat_wgs84,
-    lon_wgs84
-  ) %>%
-  rename(lat_wgs84 = lon_wgs84,
-         lon_wgs84 = lat_wgs84) %>%
-  unique() %>%
-  mutate(eventual_mpa = (year_mpa > 0))
-
-top_species <- fitted_data$classcode %>% unique()
-
-life_summaries <- life_history_data %>% 
-  select(classcode, tm, vbgf.k, vbgf.linf) %>% 
-  rename(age_mature = tm, vbk = vbgf.k, linf = vbgf.linf) %>% 
-  unique() %>% 
-  filter(classcode %in% top_species) %>% 
-  select(-classcode) %>% 
-  gather(variable,value) %>% 
-  group_by(variable) %>% 
-  summarise(min_val = min(value, na.rm = TRUE),
-         max_val = max(value, na.rm = TRUE))
-
- sim_life_history <- sim_grid %>%
-    mutate(
-      age_mature = map_dbl(fish, "age_mature"),
-      vbk = map_dbl(fish, "vbk"),
-      linf = map_dbl(fish, "linf")
-    ) %>%
-    select(scientific_name, age_mature, vbk, linf) %>%
-    unique() %>% 
-   filter(age_mature > life_summaries$min_val[life_summaries$variable == "age_mature"],
-          age_mature < life_summaries$max_val[life_summaries$variable == "age_mature"],
-          vbk > life_summaries$min_val[life_summaries$variable == "vbk"],
-          vbk < life_summaries$max_val[life_summaries$variable == "vbk"],
-          linf > life_summaries$min_val[life_summaries$variable == "linf"],
-          linf < life_summaries$max_val[life_summaries$variable == "linf"])
-
-#   sim_plot <- outcomes %>%
-#     mutate(rsize = plyr::round_any(mpa_size, 0.25),
-#            reff = plyr::round_any(mpa_effect,0.2)) %>%
-#     filter(rsize == 0.25, depletion < 0.8, depletion > 0.2,
-#            scientific_name %in% sim_life_history$scientific_name
-#            ) %>%
-#     mutate(year = years_protected + 2001) %>%
-#     filter(year %in% did_betas$year, years_protected >1) %>% 
-#     group_by(year, reff) %>% 
-#     count() %>% 
-#     group_by(year) %>% 
-#     mutate(sc = n / sum(n))
-# 
-#   did_plot <- did_betas %>%
-#     ggplot() +
-#     geom_tile(data = sim_plot, aes(year, reff, fill = sc), alpha = 0.75) +
-#     geom_vline(aes(xintercept = 2003), color = 'red', linetype = 2, size = 2) +
-#     geom_hline(aes(yintercept = 0)) +
-#     geom_pointrange(aes(
-#       year,
-#       y = exp(estimate) - 1,
-#       ymin = exp(lower) - 1,
-#       ymax = exp(upper) - 1
-#     ),
-#     size = 1.5
-#     ) +
-#     labs(x = "Year", y = "Targeted Trend Divergence") +
-#     scale_fill_viridis(guide = wide_hgc,
-#                        name = "% of Annual Sims",
-#                        labels = scales::percent_format(accuracy = 1),
-#                        option = "magma") + 
-#     theme(legend.position = "top",
-#         panel.grid.major.x = element_blank(),
-#         panel.grid.minor.x = element_blank()
-#         )
-#   
-# 
-# 
-# 
-# cip_data <- pisco_data %>%
-#   left_join(site_data, by = c("site","side")) %>%
-#   filter(is.na(eventual_mpa) == F) %>%
-#   filter(region %in% c("ANA", "SCI","SRI",'SMI'),
-#          classcode %in% top_species)
-
-# process outcomes
-
-```
-
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: danovan@uw.edu}
-
 <!-- \AddToShipoutPictureFG{ -->
@@
 <!-- } -->
-
-No-take Marine Protected Areas (MPAs), spatial regions of the ocean in which fishing is prohibited, have a long history in the management of marine resources.
-Traditional cultures in Oceania utilized - often temporary - MPAs as "fish banks" for times of need [@johannes1978].
-Modern MPAs were first established primarily as marine analogs to the terrestrial protection of iconic landscapes like Yellowstone or Kruger National Parks [@iucn1976;@watson2014b].
-Over time our goals and expectations for MPAs have evolved; while all MPAs are expected to deliver conservation benefits within their borders, many modern MPAs are also established to  bolster fish populations throughout the region in which they are located [@gaines2010].
-
-Indeed, many recent agreements to expand MPAs (most proximately, the Convention on Biological Diversity’s Strategic Plan for Biodiversity, which calls for 10% of coastal waters to be protected inside MPAs by 2020 and the International Union for Conservation of Nature call for 30% by 2030) are based on the expectaion that well-designed MPAs will achieve benefits both within, and outside, their borders.
-Despite these assumptions, our collective scientific understanding of the regional-scale conservation and fishery impacts of current and future MPAs is surprisingly limited.
-
-Numerous studies provide evidence that well-enforced and appropriately sized MPAs can produce conservation benefits within their borders [@lester2009;@halpern2003; @edgar2014; @gerber2005].
-As these conservation benefits accrue inside MPAs, theory holds that MPAs can affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
-Several studies have documented empirical evidence for the existence of adult or larval spillover affecting both abundance and fisheries [e.g.
-@goni2010; @halpern2009;@kay2012; @stobart2009; @mcclanahan2000; @russ1996; @thompson2017;@baetscher2019;@pelc2009], as well as alteration of fishing effort in reaction to [@costa2013;@mason2012;@murawski2005] and in anticipation of [@mcdermott2019] MPA placement.
-The potentially more important question, however, is not whether spillover occurs (it must to some degree in any realistic scenario), but what the net effects of spillover are and whether those effects are empirically detectable.
-From a fishery perspective, are spillover benefits sufficient to offset losses in fishing grounds and changes in responses of displaced fishers caused by an MPA?
-From a conservation perspective, how much does the buildup of fish inside an MPA increase biomass outside the protected area?
-Overall, what are the regional effects of MPAs?
-
-As stakeholders around the world increasingly seek to use MPAs in marine resource management portfolios, and the performance of existing MPAs is evaluated, it is critical that we develop a better understanding of the magnitude and drivers of regional-scale MPA effects.
-The Channel Islands National Marine Sanctuary, California, USA provides an ideal case study to address this need.
-A network of protected areas covering approximately 20% of the Islands' waters qas put in place in 2003 (see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of these MPAs).
-We use data from the first 15 years of protection to provide what is to our knowledge the first large-scale empirical assessment of the regional effect of a large MPA network on a wide array of fin-fish species.
-In contrast to clear differences in biomass densities observed inside and outside of well protected MPAs [@lester2009], we are unable to detect a clear regional effect from the Channel Islands MPAs.
-We build off of existing MPA theory to consider why this might be, and provide guidance for scientists and managers as to when and how we might expect to estimate the regional conservation effects of MPAs.
-
-What Are the Regional Effects of MPAs? {.unnumbered}
-----------------------
-
-The empirical MPA literature focuses on assessing the conservation effects within the borders of protected areas [e.g.
-@edgar2014].
-While these within-MPA effects are vitally important for protecting rare species, biodiversity, critical habitats, and often tourism, they paint an incomplete picture of the overall population effects of MPAs.
-The organisms within the borders of protected areas are generally part of a broader biological stock, connected through adult movement and larval dispersal.
-If the goal of conservationists or natural resource managers is to increase the total abundance or productivity of a resource, a broader question we should ask of MPAs is not just are there more fish inside their borders, but also how have the reserves affected abundances throughout the region in which they are located?
-This logic, that MPAs will have conservation benefits for most species beyond their borders, is implicit in all multilateral calls for MPA expansion.
-
-We define the regional conservation effects of MPAs as the change in total biomass of fish (summing inside and outside of MPAs) relative to the total biomass of fish that would have occurred without the MPAs (acknowledging that other outcomes such as increased biodiversity or resiliency are also important to conservation but are beyond the scope of this analysis).
-
-<!-- As a thought experiment, imagine a region that has driven its fish populations to near extinction and then places 100\% of its waters inside a no-take MPA.
-In that setting, we would expect the regional-scale conservation benefits to be massive.
-On the other hand, imagine implementing an MPA in an area where a sedentary species has been only lightly fished.
-In that setting, it is possible to create a small net conservation loss if the concentration of fishing pressure outside the reserve has a greater effect than the biomass buildup inside the reserve.
-Within these broad bounds,  -->
-
-Numerous factors can affect the regional effects of MPAs.
-These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008], larval dispersal patternss and the strength and timing of density dependence in the population (e.g.
-pre- or post-settlement, @burgess), how overfished the population would be without the MPA, and how fishing activity and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @gerber2005; @hilborn2004a; @gaines2010; @botsford2003;@smith2003;@smith2006].
-In addition, even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through habitat and network effects [@gaines2010; @costello2010].
-
-```{r}
-short_term <- outcomes %>%
-  group_by(experiment) %>% 
-  filter(years_protected == max(years_protected), years_protected >= 0) %>% 
-  ungroup() %>% 
-  filter(!is.na(mpa_effect))
-
-median_effect <- median(short_term$mpa_effect[short_term$year == max(short_term$year)], na.rm = TRUE)
-
-median_pop_effect <- median(short_term$pop_effect[short_term$year == max(short_term$year)], na.rm = TRUE)
-
-# median_fish_effect <- median(fishery_short_term$mpa_effect[short_term$year == max(short_term$year)], na.rm = TRUE)
-# 
-# median_fish_effect <- median(fishery_short_term$fishery_effect[short_term$year == max(short_term$year)], na.rm = TRUE)
-
-pos_runs <- round(mean(short_term$pop_effect[short_term$year == max(short_term$year)] > 0),2)
-
-# pos_fish_runs <- round(mean(fishery_short_term$msy_effect[fishery_short_term$year == max(fishery_short_term$year)] > 0, na.rm = TRUE),2)
-
-pos_fish_effect <- round(mean(short_term$f[short_term$year == max(short_term$year)] > 0),2)
-
-small_effect <- round( 100* median(short_term$mpa_effect[short_term$depletion <= 0.4]))
-
-min_small_effect <- round( 100* min(short_term$mpa_effect[ short_term$depletion <= 0.4]))
-
-max_small_effect <- round( 100* max(short_term$mpa_effect[ short_term$depletion <= 0.4]))
-
-medium_effect <- round(100*median(short_term$mpa_effect[short_term$depletion > 0.4 & short_term$depletion <= 0.6]))
-
-min_medium_effect <- round(100*min(short_term$mpa_effect[short_term$depletion > 0.4 & short_term$depletion <= 0.6]))
-
-max_medium_effect <- round(100*max(short_term$mpa_effect[short_term$depletion > 0.4 & short_term$depletion <= 0.6]))
-
-big_effect <- round(100*median(short_term$mpa_effect[short_term$depletion > 0.6]))
-
-min_big_effect <- round(100*min(short_term$mpa_effect[short_term$depletion > 0.6]))
-
-max_big_effect <- round(100*max(short_term$mpa_effect[short_term$depletion > 0.6]))
-
-```
-
-Given this existing body of theory, what should we expect *a priori* the effect size of the Channel Islands MPAs to be?
-The species are xx.
-We utilize a spatial bio-econonomic model to simulate the expected effects of an MPA network stylized to resemble the Channel Islands (See SI for details of simulation framework).
-For this group, MPA theory suggests that regional effects on the order of XX% are reasonable for such a system.
-
-Although the Islands do not encompass the full population of any of these species, they encompass a large enough area to assess the regional MPA conservation effects we might see in the Channel Islands after 13 years of protection (the time span covered by our model).
-The median regional conservation effect produced by our simulation model after 13 years of protection by MPAs covering between 15%-20% of a population's range was `r percent(med_ci_effect, accuracy = 1)`.
-This projected effect has an an interquartile range of `r percent(quart_ci_effect[2], accuracy = 1)` and `r percent(quart_ci_effect[4], accuracy = 1)`, although values as high as a 200% increase and as low as a 50% decrease were also produced.
-Our expectation then is that all else being equal we are likely looking for a modest but positive effect size.
-
-** reference HARKing, of course DiD and theory matching together doens't prove anything, but it's an important step to take **
-
-## How Can We Detect Regional Effects?
-
-What evidence do we have  for the effects of the Channel Island MPAs?
-@caslle2015 examined changes in biomass densities of species targeted by fishing activity inside and outside of the MPAs over time (response ratios), and compared these changes to comparable trends in non-targeted species.
-They found a statistically significant increase in the response ratios of targeted speices over time, and *evidenceXX* that this increase is smaller in the non-targeted species.
-Updating the results of @caselle2015 through 2018 shows a continuation of this trend (Fig.\@ref(fig:XX)).
-
-Why do these response ratio results not serve as sufficient evidence for the regional effects of the Channel Island MPAs?
-Control sites used in calculating response ratios are often selected based on abiotic or ecological traits such as habitat characteristics [@ferraro2018].
-However, selection of control sites is further complicated by the very spillover that MPAs are often intended to create.
-Export of adults or larvae from the MPA to the "control" site affects their status as controls, as does displacement of fishing effort from MPAs to control sites.
-In theory, control sites far enough away to negate both biological spillover and concentration by the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs may be challenging.
-While these concerns have been stated previously [e.g.
-@halpern2004], the MPA evaluation literature has by and large been unable to adequately address them, often due to the very real challenges of acquiring the necessary data [@ferraro2018].
-
-In the case of the Channel Islands MPAs, control sites are often located within XX Km of the MPAs, suggesting that they are susceptible to biological spillover and concentration of fishing effort resulting from the MPAs.
-Given these complications, we can certainly attribute the response ratios initially reported in @caselle2015 as evidence that the MPAs are indeed providing effective protection of targeted fish biomass inside their borders.
-However, it is unclear how well these response ratio results serve as in indicator of overall regional effects of the MPAs.
-
-To illustrate this problem, we used our stylized bio-economic model to simulate response ratio trajectories for the species and MPA coverage present in the Channel Islands.
-We then compared the range of plausible response ratios generated by our model to the true regional effect of the MPAs in each simulation.
-When response ratios were calculated from true control sites completely unaffected by the MPAs, they are a highly reliable measure of regional conservation effects.
-However, once we allow for the possibility of MPAs affecting the control sites through ecological and economic mechanisms, response ratios were generally a poor predictor of regional conservation effects (Fig.\@ref(fig:)).
-This was particularly true in the early years of MPA implementation; while biomass increases caused by MPAs are constrained by the growth rates of fish individuals and populations, reductions in biomass caused by concentrated fishing outside of reserves can occur quite quickly.
-As a result, response ratios can indeed serve as a measure of MPAs ability to provide effective within-border protection, but particular in the first XX years they are a highly imprecise measure of the net effect of the MPAs at a regional scale.
-
-<!-- If response ratios are not a broadly applicable solution to the challenge of estimating regional conservation effects, what then can we do?
-How then can we empirically measure the regional effects of MPAs?
-Intuition and bio-economic theory tells us that the regional conservation effects of MPAs are likely to be highly context dependent across locations and across species within locations, so careful thought must be devoted to an empirical design to detect their effects.
-The perfect experiment would involve two parallel worlds that were identical, except for the implementation of an MPA.
-In world “A”, no MPA would be implemented, and in the facsimile world “B”, the MPA would be implemented.
-Both worlds would be tracked before and after MPA implementation, and the conservation and fishery outcomes would be compared after treatment.
-Instead of two parallel worlds, a similar experimental design would involve random placement of MPAs.
--->
-
-## MPA Effects in the Channel Islands {#estimating .unnumbered}
-
-Given these challenges, what is a viable alternative for estimating the conservation effects of MPAs?
-Building off of @caselle2015, we use a difference-in-difference estimator comparing densities in finfish species targeted by fishing effort (i.e., those potentially affected by an MPA) from those not targeted before and after MPA implementation [see @larsen2019 for a succinct explanation of difference-in-difference estimators in ecological settings].
-Targeted species in the Northern Channel Islands, California, include commercially important finfish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
-Each of these targeted species was the subject of prior bio-economic modeling related to the effects of MPAs in southern California [@costello2010d;@rassweiler2014].
-But, the analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*).
-Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species).
-
-```{r raw-trend, fig.cap="Centered and scaled mean annual biomass density of included species (faded lines) and smoothed means of targeted and non-targeted groups, and mean (darker lines) and 95% confidence interval of the mean (ribbon) over time", include = FALSE}
-raw_did_plot <- base_run$data[[1]] %>%
-  select(year, targeted,classcode, log_density, any_seen) %>%
-  mutate(density = exp(log_density) * any_seen) %>%
-  group_by(classcode,targeted, year) %>%
-  summarise(md = mean(density)) %>%
-  group_by(classcode,targeted) %>%
-  mutate(smd = (md - mean(md)) / sd(md)) %>%
-  ggplot(aes(year, smd, color = targeted == 1, fill = targeted == 1)) +
-  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red", size = 1.5) +
-  geom_line(aes(group = interaction(targeted, classcode)), alpha = 0.5) +
-  geom_smooth(size = 2) +
-  labs(y = "Centered and Scaled Mean Density", x = "Year") +
-    scale_color_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
-      scale_fill_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
-    labs(y = "Centered and Scaled Mean Density", x = "Year") + 
-  theme(legend.position = "top",
-        panel.grid.major.x = element_blank(),
-        panel.grid.minor.x = element_blank()
-        )
-
-raw_did_plot <- raw_did_plot +
-      scale_color_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "") +
-        scale_fill_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "")
-
-pdf(file =  "figs/pop-trends.pdf",width = 8, height = 5, useDingbats = TRUE)
-print(raw_did_plot)
-dev.off()
-
-```
-
-\begin{figure}%[tbhp]
-  \centering
-  \includegraphics[width=1\linewidth]{figs/pop-trends.pdf}
-  \caption{Centered and scaled mean annual biomass densities of included targeted (red) and non-targeted (blue) fish species (faded lines) and smoothed means with 95\% confidence intervals over time (darker lines and ribbons).
-Vertical red dashed line shows year of MPA implementation.}
-  \label{pop-trends}
-\end{figure}
-
-Using this model, we estimate...
-XX.
-As we might expect from bio-economic modeling, we detect no significant difference between targeted and non-targeted biomass densities for the first XX years of protection.
-However, over time the model assigns greater and greater probability to MPA-mediated increases in targeted fin-fish biomass densities, up until XX, at which point th trend reverses itself.
-We detect no significant MPA effect between 2015 and 2018 (Fig.\@ref(fig:)).
-
-<!-- The raw trends in the biomass densities appear to show parallel trends in the densities of targeted and non-targeted fishes pre-MPA, with the targeted group diverging first positively and then negatively over time (Fig.\ref{pop-trends}).
-However, these raw trends could be caused by any number of confounding variables besides the effects of the MPAs themselves, from changes in observer skill and sampling locations, to shifts in environmental conditions.
-Using our difference-in-difference model in an effort to control for these factors, we found no significant difference in the biomass densities of targeted and non-targeted fishes pre-MPA (before 2003).
-Following the implementation of the MPAs in 2003, we see evidence of an increasing trend in densities of targeted relative to non-targeted fish, reaching a peak value in 2014.
-After 2014, densities of targeted species appear to have declined relative to what we would expect based on the non-targeted group (Fig.\ref{did-plot}).
--->
-
-```{r did-plot, include = FALSE}
-
-did_plot <- did_plot +
-  scale_y_continuous(name = "~Divergence from Non-Targeted", labels = scales::percent_format(accuracy = 1))
-
-ggsave(did_plot, filename = "figs/did_plot.png",
-       width = 8, height = 5)
-
-# pdf(file =  "figs/did_plot.pdf",width = 8, height = 5, useDingbats = TRUE)
-# print(did_plot)
-# dev.off()
-
-```
-
-\begin{figure}%[tbhp]
-  \centering
-  \includegraphics[width=1\linewidth]{figs/did_plot.png}
-  \caption{Estimated divergence in biomass densities of targeted and non-targeted fishes throughout the Channel Islands (integrated across inside and outside of MPAs).
-MPAs are implemented in 2003 (red dashed line).
-Colored cells in the background indicate the frequency of MPA conservation effects predicted by our simulation modeling for MPAs similar in design and species composition to the Channel Islands in each year.}
-  \label{did-plot}
-\end{figure}
-
-<!-- To the extent that we can select control sites that are both sufficiently similar to the region treated with MPAs and are unaffected by biological or economic spillover, response ratios may be reliable estimators of regional conservation effects.
-In reality these conditions are unlikely to hold but for a few select cases.
-This becomes especially difficult when networks of MPAs are implemented, such as has occurred in California.
-In this case, control sites are generally located within a few kilometers from MPA borders, making biological spillover and fishing concentration highly likely.
-In these circumstances, without robust statistical controls, response ratios may provide an unreliable estimate of the regional conservation effect of marine protected areas.
--->
-
-```{r dr-vs-mpa-effect-plot, include = FALSE}
-
-ylabs <- c(expression(""<= "-100%"), paste0(seq(0,300, by = 100),"%"),expression("">= "400%") )
-
-dr_plot <- density_ratios %>%
-  filter(sprinkler == FALSE, mpa_habfactor == 1) %>%
-  select(experiment,
-         years_protected,
-         mpa_effect,
-         adult_movement,
-         mpa_size,
-         contains("density_ratio")) %>%
-  mutate(adult_movement = adult_movement / max(adult_movement)) %>%
-  gather(scenario, density_ratio, contains("density_ratio")) %>%
-  group_by(experiment, scenario) %>%
-  filter(years_protected == max(years_protected)) %>%
-  mutate(density_ratio = density_ratio * mpa_size) %>% 
-  ggplot(aes(y = pmin(4, mpa_effect), x = pmin(4, density_ratio - 1))) +
-    geom_hline(aes(yintercept = 0)) + 
-  geom_vline(aes(xintercept = 0)) + 
-  geom_abline(aes(slope = 1, intercept = 0), linetype = 2, size = 1.5, color = "black") +
-  geom_point(aes(color = adult_movement, size = mpa_size), alpha = 0.5, show.legend = TRUE)  +
-    geom_smooth(method = "lm", se = FALSE) +
-  facet_wrap(~ rev(scenario), labeller = labeller(`rev(scenario)` = c(biased_density_ratio = "A) Inside vs. Experiment",true_density_ratio = "B) Inside vs. Outside"))) +
-  labs(y = "True % Increase in Biomass", x = "% Response Ratio") +
-  # scale_x_percent(labels = c(percent(seq(-1,1, by = 1)),">200%")) +
-  # scale_y_percent(labels = c(percent(seq(-1,1, by = 1)),">200%")) +
-  scale_color_viridis(guide = wide_hgc, labels = scales::percent_format(accuracy = 1), name = "Fish Movement",
-                      option = "cividis") + 
-  scale_x_continuous(labels = ylabs) + 
-    scale_y_continuous(labels = ylabs) + 
-  theme(legend.position = "top") + 
-  guides(size = FALSE)
-
-  ggsave(dr_plot,filename = "figs/density-ratio.png",
-         width = 8,
-         height = 5)
-
-# pdf(file =  "figs/density-ratio.png",width = 8, height = 5, useDingbats = TRUE)
-# print(dr_plot)
-# dev.off()
-
-```
-
-\begin{figure*}%[tbhp]
-  \centering
-  \includegraphics[width=.9\linewidth]{figs/density-ratio.png}
-  \caption{Simulated MPA-size weighted response ratios (x-axis) plotted against true regional change in biomass caused by MPAs (y-axis).
-Each point represents a simulation, and color represents the movement rate of fish in the simulation (expressed as a percentage of the patches a fish is likely to move in one time step).
-The black line shows the one:one fit line, the blue line is a linear fit of the relationship between observed response ratios and true regional change in biomass.
-Inside vs.
-Experiment simulates a true control, while Inside Vs.
-Outside uses a control that may be affected by the treatment.}
-  \label{density-ratio}
-\end{figure*}
-
-```{r}
-fish_v_fishing <- processed_grid %>%
-  select(-density_ratio) %>%
-  unnest() %>%
-  rename(fishery_effect = mpa_effect1) %>%
-  group_by(experiment) %>%
-  mutate(year = 1:length(year)) %>%
-  ungroup() %>%
-  mutate(years_protected = year - year_mpa + 1) %>%
-  left_join(outcomes %>% select(experiment, year, depletion),
-            by = c("experiment", "year"))
-
-fish_v_fishing_plot <- fish_v_fishing %>%
-  filter(years_protected == max(years_protected)) %>%
-  ggplot(aes(pmin(2, mpa_effect), pmin(2, fishery_effect))) +
-  geom_abline(aes(slope = 1, intercept = 0)) +
-  geom_smooth(method = "lm") +
-  geom_hex()
-
-```
-
-```{r ci-map, fig.cap = "Map of study region and sampling locations.
-Shaded polygons indicate location of MPAs.
-Points represent sampling locations, and color indicates the number of observations recorded at a given point", include = FALSE}
-
-ci_map <- pisco_data %>%
-  left_join(site_data, by = c("site","side")) %>%
-  filter(is.na(eventual_mpa) == F) %>%
-  filter(region %in% c("ANA", "SCI","SRI",'SMI'),
-         classcode %in% top_species) %>%
-  group_by(lat_wgs84, lon_wgs84, site, region, eventual_mpa) %>%
-  count()
-
-ci_map <-  ci_map %>%
-  dplyr::mutate(geometry = purrr::map2(lon_wgs84, lat_wgs84, ~ sf::st_point(x = c(.x, .y), dim = 'XY'))) %>%
-  ungroup() %>%
-  mutate(geometry = sf::st_sfc(geometry, crs = 4326)) %>%
-  sf::st_sf()
-
-bbox <- sf::st_bbox(ci_map)
-
-pisco_ci_map_plot <-  ggmap(channel_islands, alpha = 0.5) +
-  geom_sf(data = ca_mpas, inherit.aes = FALSE, alpha = 0.5) +
-  geom_sf(data = ci_map,inherit.aes = FALSE,
-  aes(fill = eventual_mpa, color = eventual_mpa),
-  shape = 21, 
-  alpha = 0.5,
-  show.legend = FALSE,
-  size = 1.5) +
-  coord_sf(xlim = c(bbox['xmin'] - .3, bbox['xmax'] + .06),
-  ylim = c(bbox['ymin'] - .1, bbox['ymax'] + .1)) +
-  # scale_color_gradient(low = "white", high = "orangered",
-  # guide = guide_colorbar(frame.colour = "black",frame.linewidth = 1,
-  # barheight = unit(13, "lines")),
-  # name = "Observations")  +
-  labs(x = "", y = "") +
-      ggspatial::annotation_scale(location = "br", width_hint = 0.4, text_col = "white") + 
-  ggsci::scale_color_npg() + 
-    ggsci::scale_color_npg()
-
-# pisco_ci_map_plot <-  ggplot() +
-#   geom_sf(data = tests2, fill = c("#FFD39B")) +
-#   geom_sf(data = ca_mpas, inherit.aes = FALSE, alpha = 0.5) +
-#   geom_sf(data = ci_map,inherit.aes = FALSE,
-#   aes(color = eventual_mpa),
-#   alpha = 0.75,
-#   show.legend = FALSE,
-#   size = 2) +
-#   coord_sf(xlim = c(bbox['xmin'] - .3, bbox['xmax'] + .06),
-#   ylim = c(bbox['ymin'] - .1, bbox['ymax'] + .1)) +
-#   # scale_color_gradient(low = "white", high = "orangered",
-#   # guide = guide_colorbar(frame.colour = "black",frame.linewidth = 1,
-#   # barheight = unit(13, "lines")),
-#   # name = "Observations")  +
-#   labs(x = "", y = "") +
-#       ggspatial::annotation_scale(location = "br", width_hint = 0.4, text_col = "white") + 
-#   theme(panel.background = element_rect(fill = "lightblue"))
-# 
-pdf(file =  "figs/channel-islands.pdf",width = 8, height = 5, useDingbats = TRUE)
-print(pisco_ci_map_plot)
-dev.off()
-
-```
-
-\begin{figure}%[tbhp]
-  \centering
-  \includegraphics[width=1\linewidth]{figs/channel-islands.pdf}
-  \caption{Map of PISCO sampling locations within the the Channel Islands National Marine Sanctuary, California, USA.
-Grey boxes indicated MPAs, points sampling locations, with color representing whether the site is inside an MPA (blue) or not (red)}
-  \label{channel-islands}
-\end{figure}
-
-```{r}
-ci_effects <- short_term %>%
-  filter(years_protected == max(years_protected),
-         mpa_size  <= 0.25, mpa_size >= 0.15)
-
-med_ci_effect <- median(ci_effects$mpa_effect)
-
-# rethinking::HPDI(ci_effects$mpa_effect, prob = 0.5)
-
-quart_ci_effect <- quantile(ci_effects$mpa_effect)
-
-```
-
-Discussion {#discussion .unnumbered}
-------
-
-1. Why are the results from the channel islands imporant and what do they tell us
-
-2. Put the channel islands results in context -
-
-### What Should We Expect from MPAs?
-
-We tested the effects of different combinations of these theoretical drivers of MPA conservation effects using over 10,000 simulated MPA scenarios.
-A striking outcome of this simulation analysis is how the level of fishing pressure can cause dramatic and predictable shifts in conservation outcomes.
-We measure fishing pressure by the degree of simulated depletion (percent of unfished biomass gone, such that a depletion of 0% means the population is at 100% of unfished biomass, 100% means 0% of unfished biomass remains) in the scenario without MPAs.
-When MPAs were placed in relatively unexploited fisheries (depletions less than 40%), the median regional conservation effect was barely noticeable (`r small_effect`%).
-For ecosystems fished near MSY levels (depletion of 40%-60%), the median conservation effect was `r medium_effect`%, while for overfished ecosystems (without-MPA depletions above 60%) the median regional conservation effect was over  `r big_effect`% (Fig.\ref{conservation-effects}-A).
-Across all of these groups though, effects as low as declines of -50% and as high as an increase of more than 100% were possible.
-It should be noted though that these runs represent the range of scenarios evaluated in our simulation, and that certain types of scenarios may be much more likely than others in the real world.
-To put these results into context, the FAO estimates that 7% of the worlds fisheries with status estimates fall into the relatively unexploited category, 60% fall into the fully fished category, and 33% fall into the heavily fished category [@fao2018], though works that include a broader range of fisheries estimate that 50% or more of stocks to fall into the heavily fished category [@costello2012;@rosenberg2017].
-
-These simple results support the intuitive conclusion that bigger MPAs on overfished populations produce large positive conservation outcomes.
-However, when combined with other factors such as fleet dynamics, life history, and MPA size, the conservation effects of MPAs can vary widely in magnitude and timing, a finding corroborated by empirical evidence [@starr2015;@caselle2015;@nickols2019;@villasenor-derbez2019] (Fig.\ref{conservation-effects}-B).
-As the area protected inside MPAs increases, positive effects become more likely, but even MPAs protecting most of a population's footprint can produce conservation effects as low as 0% and as high as 200%.
-The degree of depletion (i.e.
-fishing pressure) has a clearer signal, with small conservation benefits of MPAs when depletion is low, and larger effects when depletion is high, but again even for severely depleted populations the conservation effects of MPAs can vary widely (although nearly always with population benefits).
-
-While positive conservation outcomes were much more likely to occur across our simulations (`r percent(pos_runs, accuracy = 1)` of simulations), the `r percent(1 - pos_runs, accuracy = 1)` of runs that produced net conservation losses are of particular interest.
-These runs occurred almost exclusively when the simulated fleet followed a "constant-catch" fishing strategy.
-Under the constant-catch fleet model, fishing communities (or fishery policies) seek to catch the same amount regardless of the presence of an MPA.
-While a constant-catch greater than MSY is not possible over the long-term under the assumptions of our model, over the short-term a constant-catch scenario is not implausible.
-Subsistence fisheries may use a constant-catch style policy over the short-term, as they seek to ensure that their food needs are met.
-More industrial fisheries may have pre-arranged agreements with buyers to deliver set amounts of fish.
-Constant-catch dynamics might also occur in fisheries with constraining quotas that are not updated after the implementation of MPAs.
-These patterns suggest that coupling MPA implementation with appropriate management reforms that reduce the effects of constant-catch strategies can move the predicted conservation effects of MPAs to being nearly uniformly positive.
-The possibility of constant-catch strategies to result in net MPA conservation losses highlights the critical importance of understanding the broader management context in which MPAs are placed.
-
-```{r conservation-effect, include = FALSE}
-
-pop_time_plot <- outcomes %>%
-  filter(years_protected > 0) %>%
-  ggplot(aes(years_protected, mpa_effect)) +
-  geom_bin2d() +
-  scale_fill_viridis(guide = gc)
-
-ylabs <- c(expression(""<= "-50%"), paste0(seq(-25,75, by = 25),"%"),expression("">= "100%") )
-
-ybreaks = seq(-.50,1, by = .25)
-
-pop_depletion_plot <- outcomes %>%
-  filter(year == max(year)) %>%
-  ggplot() +
-  geom_bin2d(aes(depletion, mpa_effect), binwidth = c(.1, .1),
-             show.legend = FALSE) +
-  scale_fill_viridis(
-  option = "A",
-  # trans = "log10",
-  guide = guide_colorbar(frame.colour = "black",
-  frame.linewidth = 1,),
-  name = "Median Effect"
-  ) +
-  scale_x_continuous(labels = scales::percent_format(accuracy = 1), name = "Depletion") +
-  scale_y_continuous(labels = ylabs, breaks = ybreaks, name = "")
-
-pop_depletion_plot <- outcomes %>%
-    filter(year == max(year)) %>%
-    mutate(
-      rdep = cut(
-        depletion,
-        breaks = seq(0, 1, by = .2),
-        labels = seq(0, .8, by = .2),
-        include.lowest = TRUE
-      )
-      ,
-      reff = cut(
-        mpa_effect,
-        breaks = seq(-.6, 1, by = .2),
-        labels = seq(-.6, .8, by = .2),
-        include.lowest = TRUE
-      )
-    ) %>%
-    group_by(rdep, reff) %>%
-    count() %>%
-    group_by(rdep) %>%
-    mutate(pn = n / sum(n)) %>%
-    ungroup() %>%
-    mutate(rdep = as.numeric(as.character(rdep)),
-           reff = as.numeric(as.character(reff))) %>%
-    ggplot(aes(rdep, reff, fill = pn)) +
-    geom_raster(hjust = 1, vjust = 1, show.legend = FALSE) +
-    scale_fill_viridis(
-      option = "A",
-      # trans = "log10",
-      guide = hgc,
-      name = "% of Sims",
-      labels = scales::percent_format(accuracy = 1)
-    ) +
-    scale_x_continuous(labels = scales::percent_format(accuracy = 1),
-                       name = "Depletion",
-                       expand = expand_scale(add = c(0,0))) +
-    scale_y_continuous(labels = ylabs,
-                       breaks = ybreaks,
-                       name = "")
-
-  sim_plot <- outcomes %>%
-    mutate(rsize = plyr::round_any(mpa_size, 0.25),
-           reff = plyr::round_any(mpa_effect,0.2)) %>%
-    filter(rsize == 0.25, depletion < 0.8, depletion > 0.2,
-           scientific_name %in% sim_life_history$scientific_name
-           ) %>%
-    mutate(year = years_protected + 2001) %>%
-    filter(year %in% did_betas$year, years_protected >1) %>% 
-    group_by(year, reff) %>% 
-    count() %>% 
-    group_by(year) %>% 
-    mutate(sc = n / sum(n))
-
-  pop_size_plot <- outcomes %>%
-    filter(year == max(year)) %>%
-    mutate(
-      rsize = cut(
-        mpa_size,
-        breaks = seq(0, 1, by = .2),
-        labels = seq(0, .8, by = .2)
-      )
-      ,
-      reff = cut(
-        mpa_effect,
-        breaks = seq(-.6, 1, by = .2),
-        labels = seq(-.6, .8, by = .2)
-      )
-    ) %>%
-    group_by(rsize, reff) %>%
-    count() %>%
-    group_by(rsize) %>%
-    mutate(pn = n / sum(n)) %>%
-    ungroup() %>%
-    mutate(rsize = as.numeric(as.character(rsize)),
-           reff = as.numeric(as.character(reff))) %>%
-    ggplot(aes(rsize, reff, fill = pn)) +
-    geom_raster(hjust = 1, vjust = 1) +
-    scale_fill_viridis(
-      option = "A",
-      # trans = "log10",
-      guide = hgc,
-      name = "% of Sims",
-      labels = scales::percent_format(accuracy = 1)
-    ) +
-    scale_x_continuous(labels = scales::percent_format(accuracy = 1),
-                       name = "Range in MPA",
-                       expand = expand_scale(add = c(0,0))) +
-    scale_y_continuous(labels = ylabs,
-                       breaks = ybreaks,
-                       name = "Conservation Effect") +
-    theme(legend.position = "top") +
-    labs(title = 'B)')
-
- pop_facet_effect_plot <- outcomes %>%
-  filter(year == max(year)) %>%
-   select(mpa_effect, mpa_size, depletion) %>%
-   gather(measure, value, -mpa_effect) %>%
-   ggplot() +
-   geom_bin2d(aes(value, mpa_effect), binwidth = c(.1, .1), show.legend = TRUE) +
-  scale_fill_viridis(
-  option = "A",
-  # trans = "log10",
-  guide = hgc,
-  name = "# of Sims"
-  ) +
-   facet_wrap(~measure, labeller = labeller(measure = facet_labels), strip.position = "bottom") +
-   scale_x_percent(name = "") +
-   scale_y_percent(name = "Conservation Effect") + 
-   theme(legend.position = "top")
-
-  pop_depletion_and_size_plot <- outcomes %>%
-    filter(years_protected >= 0) %>%
-    group_by(experiment) %>%
-    mutate(depletion = plyr::round_any(depletion[years_protected == 0], .1),
-           mpa_size = plyr::round_any(mpa_size, .1)) %>%
-    filter(year == max(year)) %>%
-    group_by(depletion, mpa_size) %>%
-    summarise(median_mpa_effect = median(mpa_effect, na.rm = TRUE)) %>%
-    ggplot(aes(depletion, mpa_size, fill = median_mpa_effect)) +
-    geom_tile() +
-    # geom_contour(aes(z = median_mpa_effect, color = ..level..), show.legend = FALSE) +
-    scale_fill_viridis(labels = scales::percent_format(accuracy = 1),
-                       guide = hgc,
-                       name = "Median Effect") +
-  labs(x = "Depletion", y = "Range in MPA", title = "A)") +
-    scale_y_continuous(labels = scales::percent_format(accuracy = 1),
-                       expand = c(0,0)) +
-    scale_x_continuous(labels = scales::percent_format(accuracy = 1), expand = c(0,0), limits = c(0,1)) + 
-    theme(legend.position = "top")
-
-  # pop_depletion_and_size_plot <-  directlabels::direct.label(pop_depletion_and_size_plot, list("angled.boxes", box.alpha = 0.5))
-
-  pop_combo_plot <-
-    (pop_depletion_and_size_plot) + ((pop_size_plot) / pop_depletion_plot)  + plot_layout(widths = c(1.5, 1)) & theme(
-    plot.margin = unit(c(0, 1, 0, 1), units = "lines"),
-    axis.text.x = element_text(size = 8),
-    legend.box.margin = unit(c(0, 0, 0, 0), units = "lines"),
-    axis.text.y = element_text(size = 10)
-    )
-
-  ggsave(pop_combo_plot,filename = "figs/conservation-effect.png",
-         width = 8,
-         height = 5)
-
-  pdf(
-    file =  "figs/conservation-effect.pdf",
-    width = 8,
-    height = 5,
-    useDingbats = FALSE
-  )
-  print(pop_combo_plot)
-  dev.off()
-
-```
-
-\begin{figure*}%[tbhp]
-  \centering
-  \includegraphics[width=.9\linewidth]{figs/conservation-effect.png}
-  \caption{Median (A) and range (B) of equilibrium regional MPA conservation effects (percent change in total biomass with MPAs relative to without MPAs) across a range of depletion and MPA sizes (and incorporating the full range of scenarios included in our study).
-"Range in MPA" is the percent of patches covered by an MPA, "Depletion" is the depletion that would have occurred in equilibrium without the MPA.
-Colors in panel B represent the percentage of simulations that fell into an effect size bin at each x-axis bin, such that the sum of all percentages for a given x-axis bin is 1.}
-  \label{conservation-effects}
-\end{figure*}
-
-### How can we use response ratios
-
-Unfortunately, neither of these experimental designs has been implemented, because parallel worlds do not exist and MPAs are never placed at random.
-Careful causal inference requires some other means of controlling for biases introduced by factors such as the MPA siting process and biological spillover and concentration of fishing effort  [@ferraro2018;@larsen2019].
-Conservation effects are typically estimated by comparing organism densities or biomass inside MPAs to densities or biomass in selected control sites outside MPAs, which we will refer to as response ratios.
-@lester2009 and @halpern2002 present meta-analyses of hundreds of such studies.
-These results often find massively higher densities and biomasses inside MPAs than outside [median biomass increases of  over 400%, @lester2009].
-While response ratios are intended as a measure of the effectiveness of reserves within their borders, these result stand in contrast to the smaller region wide effect sizes predicted by our simulation model.
-This raises the question: How reliable are density or biomass ratios as estimators of regional conservation effects?
-
-Failure to account for these spillover effects can result in a biased estimate of the true effect of a policy such as MPA placement [@larsen2019].
-In an ideal setting, the control site used in a response ratio would be a perfect counterfactual for what would have occurred without the MPA.
-We used our simulation model to approximate this scenario, calculating the response ratio as the ratio of densities inside the MPAs relative to the overall density from the paired simulation without MPAs multiplied by the area of the population in an MPA (to account for the fact that we would expect the effectiveness of a response ratio as a proxy for regional conservation effects to scale with the size of the MPA relative to the range of the species, Fig.\ref{density-ratio}-A).
-In this idealized case, the response ratio is often a reasonable estimator for the regional conservation effect.
-
-What happens when the “control” sites are in fact affected by the application of the MPA through biological spillover and fishing concentration?
-In order to address this possibility, we calculated a new response ratio as the mean density inside MPAs relative to the mean density outside the MPAs (both weighted by distance from MPA borders and scaled by MPA size).
-We only include simulations in which habitat and larval dispersal rates are identical inside and outside of MPAs, to approximate a scenario in which treatment and control sites have been paired by ecological characteristics.
-Under these circumstances the response ratio is a somewhat biased and very inaccurate estimator of the true regional effect of MPAs.
-
-While on average high response ratios correspond with high true regional effects, the variance is high, so the response ratio alone cannot be used as a reliable proxy for the true regional effect, without accounting for the sources of biases discussed here.
-For higher movement rates, response ratios were frequently near 0\% (potentially leading stakeholders to conclude that the MPA had been ineffective), when in fact in many cases the true effect of the MPA was highly positive.
-For highly sedentary species, extremely high response ratios could create the appearance of massive conservation gains when in fact the net effect on the regional population has been near zero or even negative (Fig.\ref{density-ratio}-B).
-
-We note that spatial before-after-control-impact (BACI) studies present a potential improvement over response ratios, controlling for differences in mean densities in the treated and control regions as well as MPA independent trends, but are much rarer due to the need for extensive pre-MPA monitoring.
-However, spatial BACI still requires that the MPAs do not affect the control sites, and as such has many of the same limitations as response ratios as an estimator of regional conservation effects.
-
-MPAs are an important part of the marine resource management toolbox.
-Under ideal circumstances they can protect individual species and ecosystem linkages, while supporting local economies through tourism and fishing opportunities.
-One rationale for the expansion of MPAs is that they will deliver net conservation benefits both inside *and outside* their borders.
-To date, this assumption is insufficiently tested, and this is the focus of our paper.
-Our results show that regional conservation benefits of MPAs are highly context dependent and in many circumstances, are likely to be so small that they are nearly impossible to detect empirically.
-Indeed, this is exactly what we found in our empirical case study from the Channel Islands, California, USA.
-
-We find regional conservation gains from MPAs in `r percent(pos_runs, accuracy = 1)` of our simulations (Fig.\ref{conservation-effects}).
-MPAs covering less than 25% of the population range were unlikely to produce regional conservation gains above 10% unless the stock would have been severely overfished without the MPA.
-Given that few marine species on the planet have more than 25% of their entire range protected in MPAs, this suggests that large regional gains are unlikely to be common unless the overall extent of MPAs grows globally.
-Large MPAs protecting highly depleted stocks almost always produced large regional conservation gains, as should be expected.
-
-The median conservation effect of `r round(100*median(eqo$mpa_effect))`\% stands in contrast to the large within-MPA effects on biomass densities reported by @halpern2003a and @lester2009.
-This difference is due in part to the fact that within-MPA effects are likely to be much larger than regional-MPA effects, but may also reflect the many challenges of translating response ratios into regional conservation effects.
-Our simulation analysis shows that without proper controls (in particular for the effects of fleet redistribution), response ratios well over 100% are entirely plausible even when the actual conservation effect is much lower (and *vice versa*).
-
-We were not able to detect a statistically significant difference in the densities of targeted and non-targeted finfish species over the 13 years of MPA protection in the Channel Islands covered by our analysis.
-The data (and model assumptions) provide support for both negative and positive MPA effects, although the size of these estimated effects is well within the range that our simulation analysis suggests are plausible.
-At face value, this lack of a clear result may seem surprising given the size and carefully studied nature of this MPA network.
-
-Our simulation analysis suggests that this result should perhaps be expected.
-The Channel Islands MPAs cover approximately 20% of the waters in the Channel Islands, and while formal stock assessments are not available for many of the targeted species in our analysis, what evidence we have suggests that, as a group, these fish are on average not heavily overfished.
-Some species, such as California sheephead and blue rockfish were likely below target levels during the period [@alonzo2004;@dick2017], but projections based upon the overall average response across all species will likely suggest modest benefits even if a subset of species could experience much larger population gains.
-Harkening back to our simulations, we expect the average percentage difference in densities of targeted species with and without MPAs to be modest (Fig.\ref{did-plot}).
-Effects of this size are likely to be challenging to detect empirically given the large natural variation of marine ecosystems (especially temperate reefs) and the observation error inherent in infrequent annual survey programs such as those provided by PISCO.
-
-```{r}
-tmdata <- life_history_data %>% 
-  filter(targeted == 1, classcode %in% top_species) %>% 
-  select(tm, everything())
-
-median_tm = round(median(tmdata$tm))
-```
-
-In addition, our simulations examine percentage increases in biomass densities.
-As an extreme example, an increase in biomass densities from 2kg/m^2^ to 4kg/m^2^ would be reported as a 100% increase.
-While this chance produces a large percentage increase, this is a small change in absolute biomass densities relative to the variance in the observation process itself (see Fig.S1 for a companion to Fig.\ref{conservation-effects} scaled by absolute population size).
-Beyond these challenges, the median estimated age at sexual maturity for the targeted species included in this study is `r median_tm` years, meaning that the span of this analysis represents less than three generations of MPA protection for half of the the measured species.
-Ongoing monitoring may yet reveal clearer effects.
-Analysis of more rapidly growing and maturing species, e.g.
-spiny lobster, may also reveal clearer signals.
-
-While the MPA effect estimated by our model was not significant in any one year, the positive trend from 2005 to 2014 is evident, as is the sharp decline from 2015-2017.
-To what should we attribute this apparent shift?
-We cannot reject the parallel trends assumption between the targeted and non-targeted species in the years before the MPAs (Fig.\ref{did-plot}).
-But, the Channel Islands region (and the entire West coast of the USA) experienced a dramatic ‘marine heatwave’ beginning in 2014 and persisting through 2016, resulting in part in extremely elevated water temperatures throughout the region  [@gentemann2017].
-Biogeographic differences in the distributions of targeted and non-targeted species may confound the observed effects of MPAs.
-Many of the non-targeted species have warm thermal affinities and have increased in numbers since the heatwave [@freedman2019].
-The targeted group is made up mostly of fishes with cold-water affinities, such as members of the genus *Sebastes*.
-As such, we hypothesize that the recent evidence for a decline in densities of targeted species is due to environmental conditions that disproportionately affect the targeted group (and not for example due to concentrated fishing pressure outside the reserves).
-This hypothesis is supported by the sharp declines in densities of targeted species seen inside the MPAs themselves (see Fig.S27-S29), suggesting a driver other than fishing may be at play.
-
-What do our results imply about the future of MPA science?
-Our simulation model is by no means exhaustive (for example it ignores features such as species interactions, habitat effects, and climate feedback), but captures many of the core factors theorized to affect MPA performance.
-Our results show that while attributes such as MPA size and fishing pressure are important factors in determining the effects of protection, local fleet dynamics and the movement rates of adult fish can dramatically affect the outcomes of protected areas as well  (Fig.\ref{conservation-effects}).
-Far from being a simple tool with clear outcomes, the effects of MPAs can be highly context dependent, requiring - we would argue - a bio-economic model of at minimum the complexity presented here to help communities design and set expectations of MPAs at the tactical level.
-While users may not be able to parameterize every aspect of a model such as that presented here, working with stakeholders to visualize the implications of, for example, different fleet responses to MPA implementation is a critical step in MPA design.
-Readers can use our simulation model to explore the effects of MPAs through an interactive web application available [here](https://danovando.shinyapps.io/simmpa/).
-
-Once simulation models have been used to help design an MPA, how can users evaluate whether it is achieving their objectives?
-Response ratios are commonly used as evidence for conservation outcomes of MPAs [@lester2009], but as suggested in @ferraro2018 and @larsen2019 and demonstrated here, without careful attention to the design of control sites (accounting for example for the displacement of effort by MPAs), response ratios may be highly unreliable estimators of regional MPA effects.
-As @larsen2019 suggests, there are many potential alternatives for estimating the effects of MPAs that better account for the challenges of causal inference.
-We applied one such approach here (a difference-in-difference estimator), and yet were still unable to reach robust conclusions as to the effect of MPAs on the density of targeted species in the Channel Islands, due to the likely small size of the true effect relative to the strength and variability of environmental drivers.
-
-While this does not mean that all MPAs will face similar challenges in estimating their effects, our results in the rigorously studied Channel Islands system make clear that in many instances empirically detecting a clear regional effect of MPAs may not be possible.
-How then should stakeholders go about adaptively monitoring and managing MPAs?
-Simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can perhaps be tuned to focus on the species groups that have the highest chance of a detectable effect size [@nickols2019].
-Expanding data collection to include robust monitoring of fleet dynamics may help statistical approaches to isolate the true effect of MPAs on conservation outcomes by allowing for improved control of the effects of fishing dynamics, and allow managers to take into account potential negative interactions between MPAs and fleet dynamics such as those that may occur under constant-catch dynamics.
-Whenever possible monitoring programs should be implemented prior to MPA implementation to provide a pre-treatment benchmark.
-Non-equilibrium analyses (Fig.\ref{did-plot}) also help set expectations for effect sizes over time.
-Beyond that, educating communities about the challenges of estimating the effects of MPAs can help set expectations, so that a lack of a clear effect is not necessarily viewed as a failure of the program, but rather considered in the context of reasonable expectations.
-While this paper has focused on the conservation outcomes of MPAs, future work must also address the challenge of predicting and estimating the fishery impacts of protected areas.
-
-As the number and size of global MPA networks increase, it is critical that we both set appropriate expectations for their outcomes, and plan how we will monitor the performance of these protected areas over time.
-While the history of MPA science has made important strides in helping us understand the dynamics of protected areas, the future of MPA science must directly tackle the challenge of evaluating the performance of these MPAs, and adapting their design as needed to best achieve objectives.
-Commonly employed metrics such as response ratios may be applicable in some circumstances, but can have severe shortcomings as metrics of regional conservation effects.
-Dependence on unreliable estimators of MPA effects may lead to stakeholders incorrectly attributing negative environmental shocks as MPA failures, or interpreting data arising from scorched-earth fishing outside MPAs as a conservation success.
-Bio-economic modeling can help frame community expectations, reducing the potential for a reduction in support if unrealistic conservation or fishery gains are not realized.
-Statistical approaches that explicitly address complications such as the spatial spillover effects of MPAs (such as the difference-in-difference approach used here) may give users an improved understanding of the performance of their MPAs.
-Effective use of simulation and statistics to clearly communicate what we should expect and what we can detect from MPAs is critical in ensuring that MPAs play effective roles in fisheries management and marine conservation.
-
-## Materials and Methods
-
-We present here critical characteristics of our simulation model and regression approach.
-Further details can be found in the Supplementary Information.
-All analysis were conducted in R [@rcoreteam2019].
-Our main difference-in-difference model was fit using Template Model Builder in R [@kristensen2016].
-Code needed to replicate results can be found [here](https://github.com/DanOvando/regional-effects-of-mpas).
-
-### Simulation Model
-
-Our bio-economic model simulates the effect of MPAs on a spatially explicit age-structured representation of a single species.
-Readers can explore the functionality of the model using an online tool available [here](https://danovando.shinyapps.io/simmpa/).
-See Table.S1 for a complete description of simulation states.
-The model consists of 25 patches with wrapped edges (picture the waters around a circular island).
-For any one simulation we randomly pull a species and its associated life history from the `FishLife` [@thorson2017c] package in R.
-We pair these data with randomly selected values between 0.6 and 0.95 for Beverton-Holt steepness [as parameterized in @mace1994], as well as larval and adult dispersal rates.
-We also randomly assign whether adults have density dependent movement, as well as one of three potential types of density dependence [building off of @babcock2011]:
-
-  1. Local density dependence: Density dependence occurs independently in each patch, and recruits then disperse to nearby patches
-
-  2. Global density dependence: Density dependence is a function of the sum of spawning biomass across all patches, and recruits are then distributed according to habitat quality
-
-  3. Post-dispersal density dependence: Larvae are distributed throughout the system, and then density dependence occurs based on the density of adult biomass at the destination patch
-
-We allow for three potential siting strategies for MPAs.
-In the first, MPAs are randomly placed.
-In the second, we assume that MPAs are placed in preferentially better habit (unfished recruitment is four times greater inside MPA locations).
-In the third, we allow for scenarios in which MPAs are placed in hotspots of larval dispersal.
-In this scenario, patches in which MPAs will be placed have larval dispersal rates four times greater than patches that do not become MPAs.
-
-Along with the fleet dynamics model, each simulation is assigned a random fleet dispersal scenario: uniform dispersal (where the total effort of the fleet is divided evenly among all open patches), catch dispersal (where the total effort of the fleet is divided according to the catchable biomass in each available patch), and profit dispersal (where the total effort of the fleet is divided according to the profit per unit effort in each available patch).
-
-Lastly each simulation is assigned an MPA scenario, defined by the number and size of MPAs, the placement of those MPAs, and the year that the MPAs are put in place.
-Each simulation starts the population off at unfished equilibrium and then beings to apply the fleet model.
-The MPAs are then placed during the randomly selected start year, allowing some runs to explore how the early dynamics of the MPA play out when the fishery and population they are placed on is not already at equilibrium.
-Fishing effort in displaced by MPAs can either concentrate outside or leave the fishery.
-Each simulation is run to equilibrium with and without the selected MPA strategy (holding all else constant).
-We then measure the difference in biomass and fishery catches in each time step in the scenario with and without the MPAs to calculate the conservation and fishery effects of the MPAs over time.
-
-### Difference in Difference Regression
-
-We applied this difference-in-difference strategy to a large dataset in California.
-We used empirical kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands with the ultimate goal of testing the regional effects of MPAs in a real world context.
-A network of MPAs covering approximately 20% of the islands’ waters was put in place in 2003 as part of the California Marine Life Protection Act (MLPA) (see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of the MLPA).
-PISCO conducts visual SCUBA surveys at a large number of rocky-reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\ref{channel-islands}-\ref{pop-trends}).
-
-The key assumptions of this approach are that a) within the time-frame of the model there are no significant interaction effects between the targeted and non-targeted species (which in fact we do not detect, see SI), and b) that in the absence of the MPAs both the targeted and non-targeted groups of species would have exhibited similar trends in densities.
-The advantage to this approach is that given the time-frame of the model (18 years), we believe that spillover effects of MPA placement on the non-targeted species are likely to be much less severe than the effects of biological spillover and fishing concentration that may bias the performance of estimators such as a response ratios or a spatial BACI design.
-
-We use a difference-in-difference regression style regression to attempt to estimate the causal effect of MPAs on regional targeted fish biomass.
-Empirically, this amounts to estimating the pre-post MPA difference in the biomass densities of targeted species net the same difference for non-targeted species in the Channel Islands.
-
-The simplified form of this model is
-
-\begin{equation}
-  log(d_{i}) = \beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + e_{i}
-\label{eq:did}
-\end{equation}
-
-where $d_i$ is the biomass density at observation *i*, *T* indicates whether the observation *i* is for a targeted ($T = 1$) or non-targeted ($T = 0$) species, and *MPA* marks whether observation *i* is in a pre MPA ($MPA = 0$) or post MPA ($MPA = 1$) state.
-Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment (*MPA*) on the treated (targeted species).
-The full form of the estimation model is much more involved, integrating uncertainty in a hierarchical model from the individual observation level up to the mean trends of targeted and non-targeted species.
-See SI for details of estimation model.
-
-We briefly assess two of the most critical assumptions of this model here: that the treated and non-treated groups have parallel trends, and that the effect of the treatment on the treated does not tangentially affect the untreated.
-While the parallel trends assumption cannot be formally proven, we can examine its validity using the data from the years before the MPAs were put in place in 2003.
-We do not detect any significant differences in the trends of the biomass densities of the targeted and non-targeted species in the years before the MPAs.
-
-With regards to the second assumption, all of the species in this empirical analysis exist within an ecosystem, and as such affect each other through mechanisms such as predation, competition, and habitat modification.
-We find it unlikely that these effects have had enough time to manifest in a meaningful way in the 13 years of post-MPA data used in our analysis [@babcock2010; @pershing2015a].
-
-We used convergent cross mapping (CCM), in the manner of @clark2015, to test for the possibility of the trophic cascades biasing our results.
-Generalizations of Takens' theorem indicate that if two variables are part of the same dynamic system, their individual dynamics should reflect their relative causal influence.
-Convergent cross mapping (CCM) tests for causation by using the attractor/manifold built from the time series of one variable to predict another (hence the "cross-mapping").
-CCM then allows us to test for causal relationships in the timeseries of densities of targeted and non-targeted species.
-Our results found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (see SI for additional information in CCM testing).
-However, the longer MPAs are in place, the greater the possibility that substantial species interactions that can affect use of non-targeted species as a control may arise.
-
-<!-- \showmatmethods -->
-\showacknow
-
-&nbsp;
-&nbsp;
-&nbsp;

### 2020-03-24 documents/Untitled.tex
MAJOR restructure of paper, bayesian DiD up and running, added in "ability to estiamte" test, paper 80% restructured

--- 2019-10-10 documents/ovando-regional-effects-of-mpas.tex
@@
 justified on the grounds that they also benefit the broader region
-outside their borders. Despite the expanding role of MPAs in marine
-resource management, we lack a clear understanding of what the regional
-effects of MPAs might be, and how we might detect them in scientific
-studies. We used a bio-economic simulation model to demonstrate how the
-regional effects of MPAs are affected by a suite of economic and
-environmental drivers and show how commonly used methods to measure
-inside-MPA effects are likely to fail to capture the regional impacts of
-protection. We present an empirical strategy for estimating regional MPA
-effects on biomass density using 16 years of data from inside and
-outside of MPAs within the Channel Islands, California, USA. We find no
-statistically clear effect of MPAs on aggregate mean biomass densities
-of targeted finfish throughout the study region. To make meaningful
-progress on MPA design and evaluation, the next generation of MPA
-science must address the challenge of understanding and estimating the
-regional effects of MPAs across the diversity of species in the
-ecosystem.
+outside their borders. The conservation effects of MPAs are most
+commonly evaluated using response ratios---measures of biomass densities
+inside their borders, relative to biomass densities in reference sites
+outside their borders. Studies of this nature have provided broad
+evidence that MPAs produce conservation benefits within their borders.
+While these gains can be vitally important, marine populations are
+rarely contained entirely within the borders of MPAs; therefore a
+critical question is not only do MPAs produce conservation gains inside
+their borders, but how do they affect the broader region in which they
+are located? The Channel Islands National Marine Sanctuary provides a
+clear example of this challenge. A network of MPAs covering roughly 20\%
+of the Islands' waters was put in place in 2003, with a goal of
+providing regional conservation and fishery benefits. Response ratios
+from the region indicate that the Channel Island MPAs have increased
+biomass densities inside the MPAs. However, we are unable to find a
+clear effect of these same MPAs at the regional scale across multiple
+species.Building off of existing theory, we use a bio-economic
+simulation model to explain this discrepancy, and demonstrate under what
+conditions we likely can and cannot clearly estimate the regional
+conservation effects of MPAs. We show that MPA networks covering 25\% or
+less of a region are likely to produce regional increases in fish
+biomass on the order of 10\%-20\%, a meaningful effect but also one that
+can easily be overwhelmed by environmental shocks. Our results provide a
+novel assessment of the regional effects of a large and iconic Marine
+Protected Area network, and provide guidance for managers charged with
+monitoring and adapting MPAs.
 \end{abstract}
@@
 
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: danovan@uw.edu}
-
-No-take Marine Protected Areas (MPAs), spatial regions of the ocean in
-which fishing is prohibited, have a long history in the management of
-marine resources. Traditional cultures in Oceania utilized - often
-temporary - MPAs as ``fish banks'' for times of need (1). Modern MPAs
-were first established primarily as marine analogs to the terrestrial
-protection of iconic landscapes like Yellowstone or Kruger National
-Parks (2, 3). Over time our goals and expectations for MPAs have
-evolved; while all MPAs are expected to deliver conservation benefits
-within their borders, many modern MPAs are also established to bolster
-fish populations throughout the region in which they are located (4).
-
-Indeed, many recent agreements to expand MPAs (most proximately, the
-Convention on Biological Diversity's Strategic Plan for Biodiversity,
-which calls for 10\% of coastal waters to be protected inside MPAs by
-2020 and the International Union for Conservation of Nature call for
-30\% by 2030) rely on the belief that well-designed MPAs will achieve
-benefits both within, and outside, their borders. Despite these
-assumptions, our collective scientific understanding of the
-regional-scale conservation and fishery impacts of current and future
-MPAs is surprisingly limited.
-
-What is the scientific evidence of the conservation effects of MPAs?
-Numerous studies provide evidence that well-enforced and appropriately
-sized MPAs can produce conservation benefits within their borders
-(5--8). As these conservation benefits accrue inside MPAs, theory holds
-that MPAs can affect the waters beyond their borders through the
-spillover of adult and larval fish from the protected to the fished
-areas, as well as through displacement of fishing effort. Several
-studies have documented empirical evidence for the existence of adult or
-larval spillover affecting both abundance and fisheries (9--17), as well
-as alteration of fishing effort in reaction to (18--20) and in
-anticipation of (21) MPA placement. The potentially more important
-question, however, is not whether spillover occurs (it must to some
-degree in any realistic scenario), but what the net effects of spillover
-are and whether those effects are empirically detectable. From a fishery
-perspective, are spillover benefits sufficient to offset losses in
-fishing grounds and changes in responses of displaced fishers caused by
-an MPA? From a conservation perspective, how much does the buildup of
-fish inside an MPA increase biomass outside the protected area? Overall,
-what are the regional effects of MPAs?
-
-As stakeholders around the world increasingly seek to use MPAs in marine
-resource management portfolios, and the performance of existing MPAs is
-evaluated, it is critical that we develop a better understanding of the
-magnitude and drivers of regional-scale MPA effects. To address this
-gap, this study examines two critical questions: 1) What do we expect
-the regional-scale conservation effects of MPAs to be and 2) When (and
-how) can we expect to empirically detect these effects? We answer these
-questions using a simulation analysis to frame the theoretical regional
-conservation impacts of MPAs. We then present an empirical test of the
-evidence for regional-level conservation effects of MPAs resulting from
-a network of closures put in place in the Channel Islands, California,
-USA, in 2003.
-
-\hypertarget{what-are-the-regional-effects-of-mpas}{%
-\subsection*{What Are the Regional Effects of
-MPAs?}\label{what-are-the-regional-effects-of-mpas}}
-\addcontentsline{toc}{subsection}{What Are the Regional Effects of
-MPAs?}
-
-Much of the literature on the effects of MPAs focuses on assessing the
-conservation effects within the borders of protected areas (7). While
-these within-MPA effects are vitally important for protecting rare
-species, biodiversity, critical habitats, and often tourism, they paint
-an incomplete picture of the overall population effects of MPAs. The
-organisms within the borders of protected areas are generally part of a
-broader biological stock, connected through adult movement and larval
-dispersal. If the goal of conservationists or natural resource managers
-is to increase the total abundance or productivity of a resource, a
-broader question we should ask of MPAs is not just are there more fish
-inside their borders, but also how have the reserves affected abundances
-throughout the region in which they are located? This logic, that MPAs
-will have conservation benefits for most species beyond their borders,
-is implicit in all multilateral calls for MPA expansion.
-
-We define the regional conservation effects of MPAs as the change in
-total biomass of fish (summing inside and outside of MPAs) relative to
-the total biomass of fish that would have occurred without the MPAs
-(acknowledging that other outcomes such as increased biodiversity or
-resiliency are also important to conservation but are beyond the scope
-of this analysis).
-
-We used a simulation model to examine the regional conservation effects
-of MPAs under a wide array of conditions that might reasonably be
-expected to affect these outcomes. The model explores combinations of
-MPA designs, life histories, and fishing fleet dynamics (See Table.S1
-for a complete description of simulation variables). We use this model
-to explore how different drivers interact to affect the conservation
-outcomes of MPAs.
-
-\hypertarget{conservation}{%
-\subsubsection*{Theoretical Conservation Effects}\label{conservation}}
-\addcontentsline{toc}{subsubsection}{Theoretical Conservation Effects}
-
-As a thought experiment, imagine a region that has driven its fish
-populations to near extinction and then places 100\% of its waters
-inside a no-take MPA. In that setting, we would expect the
-regional-scale conservation benefits to be massive. On the other hand,
-imagine implementing an MPA in an area where a sedentary species has
-been only lightly fished. In that setting, it is possible to create a
-small net conservation loss if the concentration of fishing pressure
-outside the reserve has a greater effect than the biomass buildup inside
-the reserve. Within these broad bounds, numerous factors can act to
-affect the regional effects of MPAs. These include the scale of adult
-and larval dispersal relative to the size of the MPAs (22--25), the
-strength and timing of density dependence in the population (e.g.~pre-
-or post-settlement), how overfished the population would be without the
-MPA, and how fishing activity and management responds to the
-implementation of the MPAs (4, 8, 26--32). In addition, even for the
-same total area of MPAs, the location and spacing of the MPAs can have a
-profound influence on their cumulative impact through habitat and
-network effects (4, 33).
-
-We tested the effects of different combinations of these theoretical
-drivers of MPA conservation effects using over 10,000 simulated MPA
-scenarios. A striking outcome of this simulation analysis is how the
-level of fishing pressure can cause dramatic and predictable shifts in
-conservation outcomes. We measure fishing pressure by the degree of
-simulated depletion (percent of unfished biomass gone, such that a
-depletion of 0\% means the population is at 100\% of unfished biomass,
-100\% means 0\% of unfished biomass remains) in the scenario without
-MPAs. When MPAs were placed in relatively unexploited fisheries
-(depletions less than 40\%), the median regional conservation effect was
-barely noticeable (4\%). For ecosystems fished near MSY levels
-(depletion of 40\%-60\%), the median conservation effect was 28\%, while
-for overfished ecosystems (without-MPA depletions above 60\%) the median
-regional conservation effect was over 100\%
-(Fig.\ref{conservation-effects}-A). Across all of these groups though,
-effects as low as declines of -50\% and as high as an increase of more
-than 100\% were possible. It should be noted though that these runs
-represent the range of scenarios evaluated in our simulation, and that
-certain types of scenarios may be much more likely than others in the
-real world. To put these results into context, the FAO estimates that
-7\% of the worlds fisheries with status estimates fall into the
-relatively unexploited category, 60\% fall into the fully fished
-category, and 33\% fall into the heavily fished category (34), though
-works that include a broader range of fisheries estimate that 50\% or
-more of stocks to fall into the heavily fished category (35, 36).
-
-These simple results support the intuitive conclusion that bigger MPAs
-on overfished populations produce large positive conservation outcomes.
-However, when combined with other factors such as fleet dynamics, life
-history, and MPA size, the conservation effects of MPAs can vary widely
-in magnitude and timing, a finding corroborated by empirical evidence
-(37--40) (Fig.\ref{conservation-effects}-B). As the area protected
-inside MPAs increases, positive effects become more likely, but even
-MPAs protecting most of a population's footprint can produce
-conservation effects as low as 0\% and as high as 200\%. The degree of
-depletion (i.e.~fishing pressure) has a clearer signal, with small
-conservation benefits of MPAs when depletion is low, and larger effects
-when depletion is high, but again even for severely depleted populations
-the conservation effects of MPAs can vary widely (although nearly always
-with population benefits).
-
-While positive conservation outcomes were much more likely to occur
-across our simulations (94\% of simulations), the 6\% of runs that
-produced net conservation losses are of particular interest. These runs
-occurred almost exclusively when the simulated fleet followed a
-``constant-catch'' fishing strategy. Under the constant-catch fleet
-model, fishing communities (or fishery policies) seek to catch the same
-amount regardless of the presence of an MPA. While a constant-catch
-greater than MSY is not possible over the long-term under the
-assumptions of our model, over the short-term a constant-catch scenario
-is not implausible. Subsistence fisheries may use a constant-catch style
-policy over the short-term, as they seek to ensure that their food needs
-are met. More industrial fisheries may have pre-arranged agreements with
-buyers to deliver set amounts of fish. Constant-catch dynamics might
-also occur in fisheries with constraining quotas that are not updated
-after the implementation of MPAs. These patterns suggest that coupling
-MPA implementation with appropriate management reforms that reduce the
-effects of constant-catch strategies can move the predicted conservation
-effects of MPAs to being nearly uniformly positive. The possibility of
-constant-catch strategies to result in net MPA conservation losses
-highlights the critical importance of understanding the broader
-management context in which MPAs are placed.
-
-\begin{figure*}%[tbhp]
-  \centering
-  \includegraphics[width=.9\linewidth]{figs/conservation-effect.png}
-  \caption{Median (A) and range (B) of equilibrium regional MPA conservation effects (percent change in total biomass with MPAs relative to without MPAs) across a range of depletion and MPA sizes (and incorporating the full range of scenarios included in our study).
-"Range in MPA" is the percent of patches covered by an MPA, "Depletion" is the depletion that would have occurred in equilibrium without the MPA.
-Colors in panel B represent the percentage of simulations that fell into an effect size bin at each x-axis bin, such that the sum of all percentages for a given x-axis bin is 1.}
-  \label{conservation-effects}
-\end{figure*}
-
-\hypertarget{how-can-we-detect-regional-effects}{%
-\subsection{How Can We Detect Regional
-Effects?}\label{how-can-we-detect-regional-effects}}
-
-How can we empirically measure the regional effects of MPAs? Intuition
-and bio-economic theory tells us that the regional conservation effects
-of MPAs are likely to be highly context dependent across locations and
-across species within locations, so careful thought must be devoted to
-an empirical design to detect their effects. The perfect experiment
-would involve two parallel worlds that were identical, except for the
-implementation of an MPA. In world ``A'', no MPA would be implemented,
-and in the facsimile world ``B'', the MPA would be implemented. Both
-worlds would be tracked before and after MPA implementation, and the
-conservation and fishery outcomes would be compared after treatment.
-Instead of two parallel worlds, a similar experimental design would
-involve random placement of MPAs.
-
-Unfortunately, neither of these experimental designs has been
-implemented, because parallel worlds do not exist and MPAs are never
-placed at random. Careful causal inference requires some other means of
-controlling for biases introduced by factors such as the MPA siting
-process and biological spillover and concentration of fishing effort
-(41, 42).
-
-Conservation effects are typically estimated by comparing organism
-densities or biomass inside MPAs to densities or biomass in selected
-control sites outside MPAs, which we will refer to as response ratios.
-(5) and (43) present meta-analyses of hundreds of such studies. These
-results often find massively higher densities and biomasses inside MPAs
-than outside (5). While response ratios are intended as a measure of the
-effectiveness of reserves within their borders, these result stand in
-contrast to the smaller region wide effect sizes predicted by our
-simulation model. This raises the question: How reliable are density or
-biomass ratios as estimators of regional conservation effects?
-
-Control sites used in density or biomass ratios are often selected based
-on abiotic or ecological traits such as habitat characteristics (41).
-However, selection of control sites is further complicated by the very
-spillover that MPAs are often intended to create. Export of adults or
-larvae from the MPA to the ``control'' site affects their status as
-controls, as does displacement of fishing effort from MPAs to control
-sites. In theory, control sites far enough away to negate both
-biological spillover and concentration by the fishing fleet could be
-selected, but finding suitably far sites that are also appropriate
-proxies for the ecological and economic context of the MPAs may be
-challenging. While these concerns have been stated previously (44), the
-MPA evaluation literature has by and large been unable to adequately
-address them, often due to the very real challenges of acquiring the
-necessary data (41).
-
-Failure to account for these spillover effects can result in a biased
-estimate of the true effect of a policy such as MPA placement (42). In
-an ideal setting, the control site used in a response ratio would be a
-perfect counterfactual for what would have occurred without the MPA. We
-used our simulation model to approximate this scenario, calculating the
-response ratio as the ratio of densities inside the MPAs relative to the
-overall density from the paired simulation without MPAs multiplied by
-the area of the population in an MPA (to account for the fact that we
-would expect the effectiveness of a response ratio as a proxy for
-regional conservation effects to scale with the size of the MPA relative
-to the range of the species, Fig.\ref{density-ratio}-A). In this
-idealized case, the response ratio is often a reasonable estimator for
-the regional conservation effect.
-
-What happens when the ``control'' sites are in fact affected by the
-application of the MPA through biological spillover and fishing
-concentration? In order to address this possibility, we calculated a new
-response ratio as the mean density inside MPAs relative to the mean
-density outside the MPAs (both weighted by distance from MPA borders and
-scaled by MPA size). We only include simulations in which habitat and
-larval dispersal rates are identical inside and outside of MPAs, to
-approximate a scenario in which treatment and control sites have been
-paired by ecological characteristics. Under these circumstances the
-response ratio is a somewhat biased and very inaccurate estimator of the
-true regional effect of MPAs.
-
-While on average high response ratios correspond with high true regional
-effects, the variance is high, so the response ratio alone cannot be
-used as a reliable proxy for the true regional effect, without
-accounting for the sources of biases discussed here. For higher movement
-rates, response ratios were frequently near 0\% (potentially leading
-stakeholders to conclude that the MPA had been ineffective), when in
-fact in many cases the true effect of the MPA was highly positive. For
-highly sedentary species, extremely high response ratios could create
-the appearance of massive conservation gains when in fact the net effect
-on the regional population has been near zero or even negative
-(Fig.\ref{density-ratio}-B).
-
-We note that spatial before-after-control-impact (BACI) studies present
-a potential improvement over response ratios, controlling for
-differences in mean densities in the treated and control regions as well
-as MPA independent trends, but are much rarer due to the need for
-extensive pre-MPA monitoring. However, spatial BACI still requires that
-the MPAs do not affect the control sites, and as such has many of the
-same limitations as response ratios as an estimator of regional
-conservation effects.
-
-\begin{figure*}%[tbhp]
-  \centering
-  \includegraphics[width=.9\linewidth]{figs/density-ratio.png}
-  \caption{Simulated MPA-size weighted response ratios (x-axis) plotted against true regional change in biomass caused by MPAs (y-axis).
-Each point represents a simulation, and color represents the movement rate of fish in the simulation (expressed as a percentage of the patches a fish is likely to move in one time step).
-The black line shows the one:one fit line, the blue line is a linear fit of the relationship between observed response ratios and true regional change in biomass.
-Inside vs.
-Experiment simulates a true control, while Inside Vs.
-Outside uses a control that may be affected by the treatment.}
-  \label{density-ratio}
-\end{figure*}
-
-\hypertarget{estimating}{%
-\subsection*{MPA Effects in the Channel Islands}\label{estimating}}
-\addcontentsline{toc}{subsection}{MPA Effects in the Channel Islands}
-
-\begin{figure}%[tbhp]
-  \centering
-  \includegraphics[width=1\linewidth]{figs/channel-islands.pdf}
-  \caption{Map of PISCO sampling locations within the the Channel Islands National Marine Sanctuary, California, USA.
-Grey boxes indicated MPAs, points sampling locations, with color representing whether the site is inside an MPA (blue) or not (red)}
-  \label{channel-islands}
-\end{figure}
-
-Given these challenges, what is a viable alternative for estimating the
-conservation effects of MPAs? In a manner similar to (38), we use a
-difference-in-difference estimator comparing densities in finfish
-species targeted by fishing effort (i.e., those potentially affected by
-an MPA) from those not targeted before and after MPA implementation
-(42). Targeted species in the Northern Channel Islands, California,
-include commercially important finfish such as California sheephead
-(\emph{Semicossyphus pulcher}), and copper (\emph{Sebastes caurinus})
-and blue (\emph{Sebastes mystinus}) rockfish. Each of these targeted
-species was the subject of prior bio-economic modeling related to the
-effects of MPAs in southern California (45, 46). But, we must note that
-the analysis omits species from important invertebrate fisheries
-including red urchin (\emph{Mesocentrotus franciscanus}) and spiny
-lobster (\emph{Panulirus interruptus}). Non-targeted species include
-garibaldi (\emph{Hypsypops rubicundus}), halfmoons (\emph{Medialuna
-californiensis}), and blacksmith (\emph{Chromis punctipinnis}) (See
-Table.S2 for a complete list of species). The key assumptions of this
-approach are that a) within the time-frame of the model there are no
-significant interaction effects between the targeted and non-targeted
-species (which in fact we do not detect, see SI), and b) that in the
-absence of the MPAs both the targeted and non-targeted groups of species
-would have exhibited similar trends in densities. The advantage to this
-approach is that given the time-frame of the model (16 years), we
-believe that spillover effects of MPA placement on the non-targeted
-species are likely to be much less severe than the effects of biological
-spillover and fishing concentration that may bias the performance of
-estimators such as a response ratios or a spatial BACI design.
-
-We applied this difference-in-difference strategy to a large dataset in
-California. We used empirical kelp forest survey data from the
-Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO)
-monitoring in the Northern Channel Islands with the ultimate goal of
-testing the regional effects of MPAs in a real world context. A network
-of MPAs covering approximately 20\% of the islands' waters was put in
-place in 2003 as part of the California Marine Life Protection Act
-(MLPA) (see (47), (48), (49), and (50) for information on the creation
-of the MLPA). PISCO conducts visual SCUBA surveys at a large number of
-rocky-reef and kelp forest sites inside and outside of MPAs throughout
-the Channel Islands, producing estimates of densities of fishes that are
-both targeted and non-targeted by fishing
-(Fig.\ref{channel-islands}-\ref{pop-trends}).
-
-Although the Islands do not encompass the full population of any of
-these species, they encompass a large enough area to assess the regional
-MPA conservation effects we might see in the Channel Islands after 13
-years of protection (the time span covered by our model). The median
-regional conservation effect produced by our simulation model after 13
-years of protection by MPAs covering between 15\%-20\% of a population's
-range was 25\%. This projected effect has an an interquartile range of
-8\% and 100\%, although values as high as a 200\% increase and as low as
-a 50\% decrease were also produced. Our expectation then is that all
-else being equal we are likely looking for a modest but positive effect
-size.
-
-\begin{figure}%[tbhp]
-  \centering
-  \includegraphics[width=1\linewidth]{figs/pop-trends.pdf}
-  \caption{Centered and scaled mean annual biomass densities of included targeted (red) and non-targeted (blue) fish species (faded lines) and smoothed means with 95\% confidence intervals over time (darker lines and ribbons).
-Vertical red dashed line shows year of MPA implementation.}
-  \label{pop-trends}
-\end{figure}
-
-The raw trends in the biomass densities appear to show parallel trends
-in the densities of targeted and non-targeted fishes pre-MPA, with the
-targeted group diverging first positively and then negatively over time
-(Fig.\ref{pop-trends}). However, these raw trends could be caused by any
-number of confounding variables besides the effects of the MPAs
-themselves, from changes in observer skill and sampling locations, to
-shifts in environmental conditions. Using our difference-in-difference
-model in an effort to control for these factors, we found no significant
-difference in the biomass densities of targeted and non-targeted fishes
-pre-MPA (before 2003). Following the implementation of the MPAs in 2003,
-we see evidence of an increasing trend in densities of targeted relative
-to non-targeted fish, reaching a peak value in 2014. After 2014,
-densities of targeted species appear to have declined relative to what
-we would expect based on the non-targeted group (Fig.\ref{did-plot}).
-
-\begin{figure}%[tbhp]
-  \centering
-  \includegraphics[width=1\linewidth]{figs/did_plot.png}
-  \caption{Estimated divergence in biomass densities of targeted and non-targeted fishes throughout the Channel Islands (integrated across inside and outside of MPAs).
-MPAs are implemented in 2003 (red dashed line).
-Colored cells in the background indicate the frequency of MPA conservation effects predicted by our simulation modeling for MPAs similar in design and species composition to the Channel Islands in each year.}
-  \label{did-plot}
-\end{figure}
-
-\hypertarget{discussion}{%
-\subsection*{Discussion}\label{discussion}}
-\addcontentsline{toc}{subsection}{Discussion}
-
-MPAs are an important part of the marine resource management toolbox.
-Under ideal circumstances they can protect individual species and
-ecosystem linkages, while supporting local economies through tourism and
-fishing opportunities. One rationale for the expansion of MPAs is that
-they will deliver net conservation benefits both inside \emph{and
-outside} their borders. To date, this assumption is insufficiently
-tested, and this is the focus of our paper. Our results show that
-regional conservation benefits of MPAs are highly context dependent and
-in many circumstances, are likely to be so small that they are nearly
-impossible to detect empirically. Indeed, this is exactly what we found
-in our empirical case study from the Channel Islands, California, USA.
-
-We find regional conservation gains from MPAs in 94\% of our simulations
-(Fig.\ref{conservation-effects}). MPAs covering less than 25\% of the
-population range were unlikely to produce regional conservation gains
-above 10\% unless the stock would have been severely overfished without
-the MPA. Given that few marine species on the planet have more than 25\%
-of their entire range protected in MPAs, this suggests that large
-regional gains are unlikely to be common unless the overall extent of
-MPAs grows globally. Large MPAs protecting highly depleted stocks almost
-always produced large regional conservation gains, as should be
-expected.
-
-The median conservation effect of 36\% stands in contrast to the large
-within-MPA effects on biomass densities reported by (51) and (5). This
-difference is due in part to the fact that within-MPA effects are likely
-to be much larger than regional-MPA effects, but may also reflect the
-many challenges of translating response ratios into regional
-conservation effects. Our simulation analysis shows that without proper
-controls (in particular for the effects of fleet redistribution),
-response ratios well over 100\% are entirely plausible even when the
-actual conservation effect is much lower (and \emph{vice versa}).
-
-We were not able to detect a statistically significant difference in the
-densities of targeted and non-targeted finfish species over the 13 years
-of MPA protection in the Channel Islands covered by our analysis. The
-data (and model assumptions) provide support for both negative and
-positive MPA effects, although the size of these estimated effects is
-well within the range that our simulation analysis suggests are
-plausible. At face value, this lack of a clear result may seem
-surprising given the size and carefully studied nature of this MPA
-network.
-
-Our simulation analysis suggests that this result should perhaps be
-expected. The Channel Islands MPAs cover approximately 20\% of the
-waters in the Channel Islands, and while formal stock assessments are
-not available for many of the targeted species in our analysis, what
-evidence we have suggests that, as a group, these fish are on average
-not heavily overfished. Some species, such as California sheephead and
-blue rockfish were likely below target levels during the period (52,
-53), but projections based upon the overall average response across all
-species will likely suggest modest benefits even if a subset of species
-could experience much larger population gains. Harkening back to our
-simulations, we expect the average percentage difference in densities of
-targeted species with and without MPAs to be modest
-(Fig.\ref{did-plot}). Effects of this size are likely to be challenging
-to detect empirically given the large natural variation of marine
-ecosystems (especially temperate reefs) and the observation error
-inherent in infrequent annual survey programs such as those provided by
-PISCO.
-
-In addition, our simulations examine percentage increases in biomass
-densities. As an extreme example, an increase in biomass densities from
-2kg/m\textsuperscript{2} to 4kg/m\textsuperscript{2} would be reported
-as a 100\% increase. While this chance produces a large percentage
-increase, this is a small change in absolute biomass densities relative
-to the variance in the observation process itself (see Fig.S1 for a
-companion to Fig.\ref{conservation-effects} scaled by absolute
-population size). Beyond these challenges, the median estimated age at
-sexual maturity for the targeted species included in this study is 5
-years, meaning that the span of this analysis represents less than three
-generations of MPA protection for half of the the measured species.
-Ongoing monitoring may yet reveal clearer effects. Analysis of more
-rapidly growing and maturing species, e.g.~spiny lobster, may also
-reveal clearer signals.
-
-While the MPA effect estimated by our model was not significant in any
-one year, the positive trend from 2005 to 2014 is evident, as is the
-sharp decline from 2015-2017. To what should we attribute this apparent
-shift? We cannot reject the parallel trends assumption between the
-targeted and non-targeted species in the years before the MPAs
-(Fig.\ref{did-plot}). But, the Channel Islands region (and the entire
-West coast of the USA) experienced a dramatic `marine heatwave'
-beginning in 2014 and persisting through 2016, resulting in part in
-extremely elevated water temperatures throughout the region (54).
-Biogeographic differences in the distributions of targeted and
-non-targeted species may confound the observed effects of MPAs. Many of
-the non-targeted species have warm thermal affinities and have increased
-in numbers since the heatwave (55). The targeted group is made up mostly
-of fishes with cold-water affinities, such as members of the genus
-\emph{Sebastes}. As such, we hypothesize that the recent evidence for a
-decline in densities of targeted species is due to environmental
-conditions that disproportionately affect the targeted group (and not
-for example due to concentrated fishing pressure outside the reserves).
-This hypothesis is supported by the sharp declines in densities of
-targeted species seen inside the MPAs themselves (see Fig.S27-S29),
-suggesting a driver other than fishing may be at play.
-
-What do our results imply about the future of MPA science? Our
-simulation model is by no means exhaustive (for example it ignores
-features such as species interactions, habitat effects, and climate
-feedback), but captures many of the core factors theorized to affect MPA
-performance. Our results show that while attributes such as MPA size and
-fishing pressure are important factors in determining the effects of
-protection, local fleet dynamics and the movement rates of adult fish
-can dramatically affect the outcomes of protected areas as well
-(Fig.\ref{conservation-effects}). Far from being a simple tool with
-clear outcomes, the effects of MPAs can be highly context dependent,
-requiring - we would argue - a bio-economic model of at minimum the
-complexity presented here to help communities design and set
-expectations of MPAs at the tactical level. While users may not be able
-to parameterize every aspect of a model such as that presented here,
-working with stakeholders to visualize the implications of, for example,
-different fleet responses to MPA implementation is a critical step in
-MPA design. Readers can use our simulation model to explore the effects
-of MPAs through an interactive web application available
-\href{https://danovando.shinyapps.io/simmpa/}{here}.
-
-Once simulation models have been used to help design an MPA, how can
-users evaluate whether it is achieving their objectives? Response ratios
-are commonly used as evidence for conservation outcomes of MPAs (5), but
-as suggested in (41) and (42) and demonstrated here, without careful
-attention to the design of control sites (accounting for example for the
-displacement of effort by MPAs), response ratios may be highly
-unreliable estimators of regional MPA effects. As (42) suggests, there
-are many potential alternatives for estimating the effects of MPAs that
-better account for the challenges of causal inference. We applied one
-such approach here (a difference-in-difference estimator), and yet were
-still unable to reach robust conclusions as to the effect of MPAs on the
-density of targeted species in the Channel Islands, due to the likely
-small size of the true effect relative to the strength and variability
-of environmental drivers.
-
-While this does not mean that all MPAs will face similar challenges in
-estimating their effects, our results in the rigorously studied Channel
-Islands system make clear that in many instances empirically detecting a
-clear regional effect of MPAs may not be possible. How then should
-stakeholders go about adaptively monitoring and managing MPAs?
-Simulation modeling can help inform the range of effect sizes that may
-be expected, and monitoring programs can perhaps be tuned to focus on
-the species groups that have the highest chance of a detectable effect
-size (39). Expanding data collection to include robust monitoring of
-fleet dynamics may help statistical approaches to isolate the true
-effect of MPAs on conservation outcomes by allowing for improved control
-of the effects of fishing dynamics, and allow managers to take into
-account potential negative interactions between MPAs and fleet dynamics
-such as those that may occur under constant-catch dynamics. Whenever
-possible monitoring programs should be implemented prior to MPA
-implementation to provide a pre-treatment benchmark. Non-equilibrium
-analyses (Fig.\ref{did-plot}) also help set expectations for effect
-sizes over time. Beyond that, educating communities about the challenges
-of estimating the effects of MPAs can help set expectations, so that a
-lack of a clear effect is not necessarily viewed as a failure of the
-program, but rather considered in the context of reasonable
-expectations. While this paper has focused on the conservation outcomes
-of MPAs, future work must also address the challenge of predicting and
-estimating the fishery impacts of protected areas.
-
-As the number and size of global MPA networks increase, it is critical
-that we both set appropriate expectations for their outcomes, and plan
-how we will monitor the performance of these protected areas over time.
-While the history of MPA science has made important strides in helping
-us understand the dynamics of protected areas, the future of MPA science
-must directly tackle the challenge of evaluating the performance of
-these MPAs, and adapting their design as needed to best achieve
-objectives. Commonly employed metrics such as response ratios may be
-applicable in some circumstances, but can have severe shortcomings as
-metrics of regional conservation effects. Dependence on unreliable
-estimators of MPA effects may lead to stakeholders incorrectly
-attributing negative environmental shocks as MPA failures, or
-interpreting data arising from scorched-earth fishing outside MPAs as a
-conservation success. Bio-economic modeling can help frame community
-expectations, reducing the potential for a reduction in support if
-unrealistic conservation or fishery gains are not realized. Statistical
-approaches that explicitly address complications such as the spatial
-spillover effects of MPAs (such as the difference-in-difference approach
-used here) may give users an improved understanding of the performance
-of their MPAs. Effective use of simulation and statistics to clearly
-communicate what we should expect and what we can detect from MPAs is
-critical in ensuring that MPAs play effective roles in fisheries
-management and marine conservation.
-
-\hypertarget{materials-and-methods}{%
-\subsection{Materials and Methods}\label{materials-and-methods}}
-
-We present here critical characteristics of our simulation model and
-regression approach. Further details can be found in the Supplementary
-Information. All analysis were conducted in R (56). Our main
-difference-in-difference model was fit using Template Model Builder in R
-(57). Code needed to replicate results can be found
-\href{https://github.com/DanOvando/regional-effects-of-mpas}{here}.
-
-\hypertarget{simulation-model}{%
-\subsubsection{Simulation Model}\label{simulation-model}}
-
-Our bio-economic model simulates the effect of MPAs on a spatially
-explicit age-structured representation of a single species. Readers can
-explore the functionality of the model using an online tool available
-\href{https://danovando.shinyapps.io/simmpa/}{here}. See Table.S1 for a
-complete description of simulation states. The model consists of 25
-patches with wrapped edges (picture the waters around a circular
-island). For any one simulation we randomly pull a species and its
-associated life history from the \texttt{FishLife} (58) package in R. We
-pair these data with randomly selected values between 0.6 and 0.95 for
-Beverton-Holt steepness (59), as well as larval and adult dispersal
-rates. We also randomly assign whether adults have density dependent
-movement, as well as one of three potential types of density dependence
-(60):
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  Local density dependence: Density dependence occurs independently in
-  each patch, and recruits then disperse to nearby patches
-\item
-  Global density dependence: Density dependence is a function of the sum
-  of spawning biomass across all patches, and recruits are then
-  distributed according to habitat quality
-\item
-  Post-dispersal density dependence: Larvae are distributed throughout
-  the system, and then density dependence occurs based on the density of
-  adult biomass at the destination patch
-\end{enumerate}
-
-We allow for three potential siting strategies for MPAs. In the first,
-MPAs are randomly placed. In the second, we assume that MPAs are placed
-in preferentially better habit (unfished recruitment is four times
-greater inside MPA locations). In the third, we allow for scenarios in
-which MPAs are placed in hotspots of larval dispersal. In this scenario,
-patches in which MPAs will be placed have larval dispersal rates four
-times greater than patches that do not become MPAs.
-
-Along with the fleet dynamics model, each simulation is assigned a
-random fleet dispersal scenario: uniform dispersal (where the total
-effort of the fleet is divided evenly among all open patches), catch
-dispersal (where the total effort of the fleet is divided according to
-the catchable biomass in each available patch), and profit dispersal
-(where the total effort of the fleet is divided according to the profit
-per unit effort in each available patch).
-
-Lastly each simulation is assigned an MPA scenario, defined by the
-number and size of MPAs, the placement of those MPAs, and the year that
-the MPAs are put in place. Each simulation starts the population off at
-unfished equilibrium and then beings to apply the fleet model. The MPAs
-are then placed during the randomly selected start year, allowing some
-runs to explore how the early dynamics of the MPA play out when the
-fishery and population they are placed on is not already at equilibrium.
-Fishing effort in displaced by MPAs can either concentrate outside or
-leave the fishery. Each simulation is run to equilibrium with and
-without the selected MPA strategy (holding all else constant). We then
-measure the difference in biomass and fishery catches in each time step
-in the scenario with and without the MPAs to calculate the conservation
-and fishery effects of the MPAs over time.
-
-\hypertarget{difference-in-difference-regression}{%
-\subsubsection{Difference in Difference
-Regression}\label{difference-in-difference-regression}}
-
-We use a difference-in-difference regression style regression to attempt
-to estimate the causal effect of MPAs on regional targeted fish biomass.
-Empirically, this amounts to estimating the pre-post MPA difference in
-the biomass densities of targeted species net the same difference for
-non-targeted species in the Channel Islands.
-
-The simplified form of this model is
-
-\begin{equation}
-  log(d_{i}) = \beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + e_{i}
-\label{eq:did}
-\end{equation}
-
-where \(d_i\) is the biomass density at observation \emph{i}, \emph{T}
-indicates whether the observation \emph{i} is for a targeted (\(T = 1\))
-or non-targeted (\(T = 0\)) species, and \emph{MPA} marks whether
-observation \emph{i} is in a pre MPA (\(MPA = 0\)) or post MPA
-(\(MPA = 1\)) state. Under the assumptions of this model, \(\beta_3\) is
-the causal effect of the treatment (\emph{MPA}) on the treated (targeted
-species). The full form of the estimation model is much more involved,
-integrating uncertainty in a hierarchical model from the individual
-observation level up to the mean trends of targeted and non-targeted
-species. See SI for details of estimation model.
-
-We briefly assess two of the most critical assumptions of this model
-here: that the treated and non-treated groups have parallel trends, and
-that the effect of the treatment on the treated does not tangentially
-affect the untreated. While the parallel trends assumption cannot be
-formally proven, we can examine its validity using the data from the
-years before the MPAs were put in place in 2003. We do not detect any
-significant differences in the trends of the biomass densities of the
-targeted and non-targeted species in the years before the MPAs.
-
-With regards to the second assumption, all of the species in this
-empirical analysis exist within an ecosystem, and as such affect each
-other through mechanisms such as predation, competition, and habitat
-modification. We find it unlikely that these effects have had enough
-time to manifest in a meaningful way in the 13 years of post-MPA data
-used in our analysis (61, 62).
-
-We used convergent cross mapping (CCM), in the manner of (63), to test
-for the possibility of the trophic cascades biasing our results.
-Generalizations of Takens' theorem indicate that if two variables are
-part of the same dynamic system, their individual dynamics should
-reflect their relative causal influence. Convergent cross mapping (CCM)
-tests for causation by using the attractor/manifold built from the time
-series of one variable to predict another (hence the ``cross-mapping'').
-CCM then allows us to test for causal relationships in the timeseries of
-densities of targeted and non-targeted species. Our results found no
-significant cross-mappings between targeted and non-targeted species,
-indicating that while clearly there are interactions between these
-groups on some level, the effects within the timespan of the data are
-not pronounced enough to be of concern to our results (see SI for
-additional information in CCM testing). However, the longer MPAs are in
-place, the greater the possibility that substantial species interactions
-that can affect use of non-targeted species as a control may arise.
-
-\showacknow
-
-~ ~ ~
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-johannes1978}{}%
-1. Johannes RE (1978) Traditional Marine Conservation Methods in Oceania
-and Their Demise. \emph{Annual Review of Ecology and Systematics}
-9(1):349--364.
-
-\leavevmode\hypertarget{ref-iucn1976}{}%
-2. IUCN (1976) \emph{IUCN yearbook, 1975-76 : Annual report of the
-International Union for Conservation of Nature and Natural Resources for
-1975 and for January-May 1976}.
-
-\leavevmode\hypertarget{ref-watson2014b}{}%
-3. Watson JEM, Dudley N, Segan DB, Hockings M (2014) The performance and
-potential of protected areas. \emph{Nature} 515(7525):67--73.
-
-\leavevmode\hypertarget{ref-gaines2010}{}%
-4. Gaines SD, White C, Carr MH, Palumbi SR (2010) Designing Marine
-Reserve Networks for Both Conservation and Fisheries Management.
-\emph{Proceedings of the National Academy of Sciences}
-107(43):18286--18293.
-
-\leavevmode\hypertarget{ref-lester2009}{}%
-5. Lester S, et al. (2009) Biological effects within no-take marine
-reserves: A global synthesis. \emph{Marine Ecology Progress Series}
-384:33--46.
-
-\leavevmode\hypertarget{ref-halpern2003}{}%
-6. Halpern BS, Warner RR (2003) Review Paper. Matching marine reserve
-design to reserve objectives. \emph{Proceedings of the Royal Society of
-London Series B: Biological Sciences} 270(1527):1871--1878.
-
-\leavevmode\hypertarget{ref-edgar2014}{}%
-7. Edgar GJ, et al. (2014) Global conservation outcomes depend on marine
-protected areas with five key features. \emph{Nature} 506(7487):216.
-
-\leavevmode\hypertarget{ref-gerber2005}{}%
-8. Gerber LR, Heppell SS, Ballantyne F, Sala E (2005) The role of
-dispersal and demography in determining the efficacy of marine reserves.
-\emph{Canadian Journal of Fisheries and Aquatic Sciences}
-62(4):863--871.
-
-\leavevmode\hypertarget{ref-goni2010}{}%
-9. Goni R, Hilborn R, D\textbackslash{}'ıaz D, Mallol S, Adlerstein S
-(2010) Net contribution of spillover from a marine reserve to fishery
-catches. \emph{Mar Ecol Prog Ser} 400:233--243.
-
-\leavevmode\hypertarget{ref-halpern2009}{}%
-10. Halpern BS, Lester SE, Kellner JB (2009) Spillover from marine
-reserves and the replenishment of fished stocks. \emph{Environmental
-Conservation} 36(04):268--276.
-
-\leavevmode\hypertarget{ref-kay2012}{}%
-11. Kay MC, Lenihan HS, Kotchen MJ, Miller CJ (2012) Effects of marine
-reserves on California spiny lobster are robust and modified by
-fine-scale habitat features and distance from reserve borders.
-\emph{Marine Ecology Progress Series} 451:137--150.
-
-\leavevmode\hypertarget{ref-stobart2009}{}%
-12. Stobart B, et al. (2009) Long-term and spillover effects of a marine
-protected area on an exploited fish community. \emph{Marine Ecology
-Progress Series} 384:47--60.
-
-\leavevmode\hypertarget{ref-mcclanahan2000}{}%
-13. McClanahan TR, Mangi S (2000) Spillover Of Exploitable Fishes From A
-Marine Park And Its Effect On The Adjacent Fishery. \emph{Ecological
-Applications} 10(6):1792--1805.
-
-\leavevmode\hypertarget{ref-russ1996}{}%
-14. Russ G, Alcala A (1996) Do marine reserves export adult fish
-biomass? Evidence from Apo Island, central Philippines. \emph{Marine
-Ecology Progress Series} 132:1--9.
-
-\leavevmode\hypertarget{ref-thompson2017}{}%
-15. Thompson AR, Chen DC, Guo LW, Hyde JR, Watson W (2017) Larval
-abundances of rockfishes that were historically targeted by fishing
-increased over 16 years in association with a large marine protected
-area. \emph{Royal Society Open Science} 4(9):170639.
-
-\leavevmode\hypertarget{ref-baetscher2019}{}%
-16. Baetscher DS, et al. (2019) Dispersal of a nearshore marine fish
-connects marine reserves and adjacent fished areas along an open coast.
-\emph{Molecular Ecology} 28(7):1611--1623.
-
-\leavevmode\hypertarget{ref-pelc2009}{}%
-17. Pelc R, Baskett M, Tanci T, Gaines S, Warner R (2009) Quantifying
-larval export from South African marine reserves. \emph{Marine Ecology
-Progress Series} 394:65--78.
-
-\leavevmode\hypertarget{ref-costa2013}{}%
-18. Costa BH e, et al. (2013) Fishers' Behaviour in Response to the
-Implementation of a Marine Protected Area. \emph{PLOS ONE} 8(6):e65057.
-
-\leavevmode\hypertarget{ref-mason2012}{}%
-19. Mason J, Kosaka R, Mamula A, Speir C (2012) Effort changes around a
-marine reserve: The case of the California Rockfish Conservation Area.
-\emph{Marine Policy} 36(5):1054--1063.
-
-\leavevmode\hypertarget{ref-murawski2005}{}%
-20. Murawski SA, Wigley SE, Fogarty MJ, Rago PJ, Mountain DG (2005)
-Effort distribution and catch patterns adjacent to temperate MPAs.
-\emph{ICES Journal of Marine Science} 62(6):1150--1167.
-
-\leavevmode\hypertarget{ref-mcdermott2019}{}%
-21. McDermott GR, Meng KC, McDonald GG, Costello CJ (2019) The blue
-paradox: Preemptive overfishing in marine reserves. \emph{Proceedings of
-the National Academy of Sciences} 116(12):5319--5325.
-
-\leavevmode\hypertarget{ref-gaines2003}{}%
-22. Gaines SD, Gaylord B, Largier JL (2003) Avoiding Current Oversights
-in Marine Reserve Design. \emph{Ecological Applications} 13(sp1):32--46.
-
-\leavevmode\hypertarget{ref-botsford2008}{}%
-23. Botsford LW, et al. (2008) Connectivity, sustainability, and yield:
-Bridging the gap between conventional fisheries management and marine
-protected areas. \emph{Reviews in Fish Biology and Fisheries}
-19(1):69--95.
-
-\leavevmode\hypertarget{ref-difranco2018}{}%
-24. Di Franco A, et al. (2018) Linking home ranges to protected area
-size: The case study of the Mediterranean Sea. \emph{Biological
-Conservation} 221:175--181.
-
-\leavevmode\hypertarget{ref-mcgilliard2008}{}%
-25. McGilliard CR, Hilborn R (2008) Modeling no-take marine reserves in
-regulated fisheries: Assessing the role of larval dispersal.
-\emph{Canadian Journal of Fisheries and Aquatic Sciences}
-65(11):2509--2523.
-
-\leavevmode\hypertarget{ref-hilborn2004}{}%
-26. Hilborn R, Punt AE, Orensanz J (2004) Beyond band-aids in fisheries
-management: Fixing world fisheries. \emph{Bulletin of Marine Science}
-74(3):493--507.
-
-\leavevmode\hypertarget{ref-hilborn1992}{}%
-27. Hilborn R (1992) Can fisheries agencies learn from experience?
-\emph{Fisheries} 17(4):6--14.
-
-\leavevmode\hypertarget{ref-walters2004}{}%
-28. Walters CJ, Martell SJ (2004) \emph{Fisheries ecology and
-management} (Princeton University Press).
-
-\leavevmode\hypertarget{ref-hastings2003}{}%
-29. Hastings A, Botsford LW (2003) Comparing Designs Of Marine Reserves
-For Fisheries And For Biodiversity. \emph{Ecological Applications}
-13(sp1):65--70.
-
-\leavevmode\hypertarget{ref-gerber2003}{}%
-30. Gerber LR, et al. (2003) Population Models For Marine Reserve
-Design: A Retrospective And Prospective Synthesis. \emph{Ecological
-Applications} 13(sp1):47--64.
-
-\leavevmode\hypertarget{ref-hilborn2004a}{}%
-31. Hilborn R, et al. (2004) When can marine reserves improve fisheries
-management? \emph{Ocean \& Coastal Management} 47(34):197--205.
-
-\leavevmode\hypertarget{ref-botsford2003}{}%
-32. Botsford LW, Micheli F, Hastings A (2003) Principles for the Design
-of Marine Reserves. \emph{Ecological Applications} 13(sp1):25--31.
-
-\leavevmode\hypertarget{ref-costello2010}{}%
-33. Costello C, Lynham J, Lester SE, Gaines SD (2010) Economic
-incentives and global fisheries sustainability. \emph{Resource} 2.
-
-\leavevmode\hypertarget{ref-fao2018}{}%
-34. FAO (2018) \emph{The State of World Fisheries and Aquaculture 2018:
-Meeting the sustainable development goals} (FAO, Rome, Italy).
-
-\leavevmode\hypertarget{ref-costello2012}{}%
-35. Costello C, Gaines S, Gerber LR (2012) Conservation science: A
-market approach to saving the whales. \emph{Nature} 481(7380):139--140.
-
-\leavevmode\hypertarget{ref-rosenberg2017}{}%
-36. Rosenberg AA, et al. (2017) Applying a New Ensemble Approach to
-Estimating Stock Status of Marine Fisheries around the World.
-\emph{Conservation Letters}:n/a--n/a.
-
-\leavevmode\hypertarget{ref-starr2015}{}%
-37. Starr RM, et al. (2015) Variation in Responses of Fishes across
-Multiple Reserves within a Network of Marine Protected Areas in
-Temperate Waters. \emph{PLoS ONE} 10(3):e0118502.
-
-\leavevmode\hypertarget{ref-caselle2015}{}%
-38. Caselle JE, Rassweiler A, Hamilton SL, Warner RR (2015) Recovery
-trajectories of kelp forest animals are rapid yet spatially variable
-across a network of temperate marine protected areas. \emph{Scientific
-reports} 5:14102.
-
-\leavevmode\hypertarget{ref-nickols2019}{}%
-39. Nickols KJ, et al. (2019) Setting ecological expectations for
-adaptive management of marine protected areas. \emph{Journal of Applied
-Ecology} 0(0).
-doi:\href{https://doi.org/10.1111/1365-2664.13463}{10.1111/1365-2664.13463}.
-
-\leavevmode\hypertarget{ref-villasenor-derbez2019}{}%
-40. Villaseñor-Derbez JC, et al. (2019) An interdisciplinary evaluation
-of community-based TURF-reserves. \emph{PLOS ONE} 14(8):e0221660.
-
-\leavevmode\hypertarget{ref-ferraro2018}{}%
-41. Ferraro PJ, Sanchirico JN, Smith MD (2018) Causal inference in
-coupled human and natural systems. \emph{Proceedings of the National
-Academy of Sciences}:201805563.
-
-\leavevmode\hypertarget{ref-larsen2019}{}%
-42. Larsen AE, Meng K, Kendall BE (2019) Causal analysis in
-controlImpact ecological studies with observational data. \emph{Methods
-in Ecology and Evolution} 0(0).
-doi:\href{https://doi.org/10.1111/2041-210X.13190}{10.1111/2041-210X.13190}.
-
-\leavevmode\hypertarget{ref-halpern2002}{}%
-43. Halpern BS, Warner RR (2002) Marine reserves have rapid and lasting
-effects. \emph{Ecology Letters} 5(3):361--366.
-
-\leavevmode\hypertarget{ref-halpern2004}{}%
-44. Halpern BS, Gaines SD, Warner RR (2004) Confounding effects of the
-export of production and the displacement of fishing effort from marine
-reserves. \emph{Ecological Applications} 14(4):1248--1256.
-
-\leavevmode\hypertarget{ref-costello2010d}{}%
-45. Costello C, et al. (2010) The value of spatial information in MPA
-network design. \emph{Proceedings of the National Academy of Sciences}
-107(43):18294--18299.
-
-\leavevmode\hypertarget{ref-rassweiler2014}{}%
-46. Rassweiler A, Costello C, Hilborn R, Siegel DA (2014) Integrating
-scientific guidance into marine spatial planning. \emph{Proceedings of
-the Royal Society B: Biological Sciences} 281(1781):20132252.
-
-\leavevmode\hypertarget{ref-osmond2010}{}%
-47. Osmond M, Airame S, Caldwell M, Day J (2010) ``Lessons for marine
-conservation planning: A comparison of three marine protected area
-planning processes''. \emph{Ocean \& Coastal Management} 53(2):41--51.
-
-\leavevmode\hypertarget{ref-kirlin2013}{}%
-48. Kirlin J, et al. (2013) California's Marine Life Protection Act
-Initiative: Supporting implementation of legislation establishing a
-statewide network of marine protected areas. \emph{Ocean \& Coastal
-Management} 74:3--13.
-
-\leavevmode\hypertarget{ref-botsford2014}{}%
-49. Botsford LW, White JW, Carr MH, Caselle JE (2014) Chapter Six -
-Marine Protected Area Networks in California, USA. \emph{Advances in
-Marine Biology}, Marine Managed Areas and Fisheries., eds Johnson ML,
-Sandell J (Academic Press), pp 205--251.
-
-\leavevmode\hypertarget{ref-hilborn2012}{}%
-50. Hilborn R (2012) The role of science in MPA establishment in
-California: A personal perspective. \emph{Environmental Conservation}
-39(03):195--198.
-
-\leavevmode\hypertarget{ref-halpern2003a}{}%
-51. Halpern BS (2003) The impact of marine reserves: Do reserves work
-and does reserve size matter? \emph{Ecological Applications}
-13(sp1):117--137.
-
-\leavevmode\hypertarget{ref-alonzo2004}{}%
-52. Alonzo S, Key M, Ish T, MacCall AD (2004) \emph{Status of the
-California Sheephead Stock for 2004}.
-
-\leavevmode\hypertarget{ref-dick2017}{}%
-53. Dick EJ, et al. (2017) \emph{The Combined Status of Blue and Deacon
-Rockfishes in U.S. Waters off California and Oregon in 2017} (Pacific
-Fishery Management Council, Portland, OR).
-
-\leavevmode\hypertarget{ref-gentemann2017}{}%
-54. Gentemann CL, Fewings MR, García-Reyes M (2017) Satellite sea
-surface temperatures along the West Coast of the United States during
-the 20142016 northeast Pacific marine heat wave. \emph{Geophysical
-Research Letters} 44(1):312--319.
-
-\leavevmode\hypertarget{ref-freedman2019}{}%
-55. Freedman R (2019) Understanding the Efficacy of Spatial Management
-on Emerging Threats. PhD thesis (University of California, Santa
-Barbara).
-
-\leavevmode\hypertarget{ref-rcoreteam2019}{}%
-56. R Core Team (2019) R: A Language and Environment for Statistical
-Computing.
-
-\leavevmode\hypertarget{ref-kristensen2016}{}%
-57. Kristensen K, Nielsen A, Berg CW, Skaug H, Bell BM (2016)
-\textbf{TMB} : Automatic Differentiation and Laplace Approximation.
-\emph{Journal of Statistical Software} 70(5).
-doi:\href{https://doi.org/10.18637/jss.v070.i05}{10.18637/jss.v070.i05}.
-
-\leavevmode\hypertarget{ref-thorson2017c}{}%
-58. Thorson JT, Munch SB, Cope JM, Gao J (2017) Predicting life history
-parameters for all fishes worldwide. \emph{Ecological
-Applications}:n/a--n/a.
-
-\leavevmode\hypertarget{ref-mace1994}{}%
-59. Mace PM (1994) Relationships between Common Biological Reference
-Points Used as Thresholds and Targets of Fisheries Management
-Strategies. \emph{Canadian Journal of Fisheries and Aquatic Sciences}
-51(1):110--122.
-
-\leavevmode\hypertarget{ref-babcock2011}{}%
-60. Babcock EA, MacCall AD (2011) How useful is the ratio of fish
-density outside versus inside no-take marine reserves as a metric for
-fishery management control rules? \emph{Canadian journal of fisheries
-and aquatic sciences} 68(2):343--359.
-
-\leavevmode\hypertarget{ref-babcock2010}{}%
-61. Babcock RC, et al. (2010) Decadal trends in marine reserves reveal
-differential rates of change in direct and indirect effects.
-\emph{Proceedings of the National Academy of Sciences}
-107(43):18256--18261.
-
-\leavevmode\hypertarget{ref-pershing2015a}{}%
-62. Pershing AJ, et al. (2015) Slow adaptation in the face of rapid
-warming leads to collapse of the Gulf of Maine cod fishery.
-\emph{Science} 350(6262):809--812.
-
-\leavevmode\hypertarget{ref-clark2015}{}%
-63. Clark AT, et al. (2015) Spatial convergent cross mapping to detect
-causal relationships from short time series. \emph{Ecology}
-96(5):1174--1181.
-
 % Bibliography

### 2020-03-24 documents/ovando-regional-effects-of-mpas-v2.Rmd
figures rescued
MAJOR restructure of paper, bayesian DiD up and running, added in "ability to estiamte" test, paper 80% restructured

--- 2020-03-12 documents/ovando-regional-effects-of-mpas-v2.Rmd
@@
 output:
+  bookdown::pdf_document2: default
+  bookdown::word_document2: default
   rticles::pnas_article: default
-  bookdown::word_document2: default
-  bookdown::pdf_document2: default
   bookdown::html_document2: default
@@
 library(tidyverse)
+library(ggtext)
 extrafont::loadfonts()
@@
 load(file = file.path(run_dir,"plots.RData"))
+
+valgrid <- read_rds(file.path(run_dir, "valgrid.rds"))
 
@@
 
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: danovan@uw.edu}
+<!-- \correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: danovan@uw.edu} -->
 
@@
 <!-- } -->
+
+*as we move away from "set it and forget it" management towards adaptive management of marine ecosystems, it is critically important that we develop robust strategies for monitoring the effectiveness of our management intercentions*
 
@@
 Given this existing body of theory, what should we expect *a priori* the effect size of the Channel Islands MPAs to be?
-The species are xx.
-We utilize a spatial bio-econonomic model to simulate the expected effects of an MPA network stylized to resemble the Channel Islands (See SI for details of simulation framework).
-For this group, MPA theory suggests that regional effects on the order of XX% are reasonable for such a system.
-
-Although the Islands do not encompass the full population of any of these species, they encompass a large enough area to assess the regional MPA conservation effects we might see in the Channel Islands after 14 years of protection (the time span covered by our model).
-The median regional conservation effect produced by our simulation model after 13 years of protection by MPAs covering between 15%-20% of a population's range was `r percent(med_ci_effect, accuracy = 1)`.
-This projected effect has an an interquartile range of `r percent(quart_ci_effect[2], accuracy = 1)` and `r percent(quart_ci_effect[4], accuracy = 1)`, although values as high as a 200% increase and as low as a 50% decrease were also produced.
-Our expectation then is that all else being equal we are likely looking for a modest but positive effect size.
-
-** reference HARKing, of course DiD and theory matching together doens't prove anything, but it's an important step to take **
+While we know many parameters will affect the expected effects of the MPAs, theory suggests that life history, adult and larval dispersal relative to MPA size, and fishing pressure will be key drivers.
+The targeted species in our database span a range of life histories, but are largely made up of fishes in the genus XX (e.g.) and **Sebastes** (e.g.
+), with a median Von Bertalanffy growth coefficient of XX, and a median age at maturity of XX [values drawn from @thorson2017c].
+The MPAs in this study cover roughly 21% of the the waters of the Northern Channel Islands, a region spanning roughly 90km east to west.
+While detailed dispersal studies are not available for all of the species covered by our study, what information we have suggest that while adults of some of these species are likely to exhibit some site fidelity [@love2011], larvae are likely distributed beyond the northern Channel Islands [@rassweiler2012].
+Therefore, if we assume that the population of the targeted species is at least equal to the extent of the Northern Channel Islands, we can assume that at most the MPAs cover roughly 20% of the targeted populations' ranges.
+Turning to the critical question of fishing pressure, formal stock assessments are largely lacking for these species.
+However, what evidence we have suggests mostly moderate fishing pressure, with some species such as California sheephead (XX) and blue rockfish (xx) experiencing high levels of fishing mortality / biomass  levels beloe target levels during the early 2000s pre-MPA [@alonzo2004;@dick2017].
+
+Considering the combinatino of slow growth, 20% or lower MPA coverage, and moderate fishing pressure, we might expect that the northern Channel Islands MPAs would produce a small to moderate effect size over time.
+However, theory also tells us that many other variables can interact to affect the outcomes of MPAs.
+To address this, we utilized a spatial bio-econonomic model to simulate the expected effects of an MPA network stylized to resemble the Channel Islands on the group of targeted species covered in our analysis (See SI for details of simulation framework).
+
+**reference HARKing, of course DiD and theory matching together doens't prove anything, but it's an important step to take**
 
@@
 
-What evidence do we have  for the effects of the Channel Island MPAs?
-@caslle2015 examined changes in biomass densities of species targeted by fishing activity inside and outside of the MPAs over time (generally termed "response ratios"), and compared these changes to the trends in non-targeted species.
+Having established a theoretical grounding for the likely effect of the Channel Island MPAs, what empirical evidence do we have for the effects of this protected area network?
+@caslle2015 examined changes in biomass densities of species targeted by fishing activity inside and outside of the MPAs over time (a measurement generally termed "response ratios"), and compared these changes to the trends in non-targeted species.
 They found a statistically significant increase in the response ratios of targeted speices over time, and evidence that this increase is smaller in the non-targeted species.
-Updating the results of @caselle2015 through 2018 shows a continuation of this trend (Fig.\@ref(fig:XX)).
+Updating the results of @caselle2015 through 2017 shows the same increasing trend in the resonse ratios of targeted species (Fig.\@ref(fig:XX)).
 
@@
 While these concerns have been stated previously [e.g.
-@halpern2004], the MPA evaluation literature has by and large been unable to adequately address them, often due to the very real challenges of acquiring the necessary data [@ferraro2018].
-
-In the case of the Channel Islands MPAs, control sites are often located within XX Km of the MPAs, suggesting that they are susceptible to biological spillover and concentration of fishing effort resulting from the MPAs.
-Given these complications, we can certainly attribute the response ratios initially reported in @caselle2015 as evidence that the MPAs are indeed providing effective protection of targeted fish biomass inside their borders.
+@halpern2004], the MPA evaluation literature has by and large been unable to adequately address them, often due to the very real challenges of identifying and sampling adequate control sites that are truly independent of the MPAs in question [@ferraro2018].
+
+In the case of the Channel Islands MPAs, control sites are often located within only a few kilometers of an MPA, suggesting that they are susceptible to biological spillover and concentration of fishing effort resulting from the MPAs.
+Given these complications, we can certainly interpret the response ratios reported in @caselle2015 and updated here as evidence that the MPAs are indeed providing effective protection of targeted fish biomass inside their borders, particularly in light of the general increase in respones ratios over time.
 However, it is unclear how well these response ratio results serve as in indicator of overall regional effects of the MPAs.
 
-To illustrate this problem, we used our stylized bio-economic model to simulate response ratio trajectories for the species and MPA coverage present in the Channel Islands.
-We then compared the range of plausible response ratios generated by our model to the true regional effect of the MPAs in each simulation.
-When response ratios were calculated from true control sites completely unaffected by the MPAs, they are a highly reliable measure of regional conservation effects.
-However, once we allow for the possibility of MPAs affecting the control sites through ecological and economic mechanisms, response ratios were generally a poor predictor of regional conservation effects (Fig.\@ref(fig:)).
-This was particularly true in the early years of MPA implementation; while biomass increases caused by MPAs are constrained by the growth rates of fish individuals and populations, reductions in biomass caused by concentrated fishing outside of reserves can occur quite quickly.
-As a result, response ratios can indeed serve as a measure of MPAs ability to provide effective within-border protection, but particular in the first XX years they are a highly imprecise measure of the net effect of the MPAs at a regional scale.
-
-```{r dr-vs-mpa-effect-plot, include = FALSE}
-
-ylabs <- c(expression(""<= "-100%"), paste0(seq(0,300, by = 100),"%"),expression("">= "400%") )
-
-density_ratios %>% 
-  ggplot(aes(years_protected, true_density_ratio)) + 
-  geom_point()
-
-dr_plot <- density_ratios %>%
-  filter(sprinkler == FALSE, mpa_habfactor == 1) %>%
-  select(experiment,
-         years_protected,
-         mpa_effect,
-         adult_movement,
-         mpa_size,
-         contains("density_ratio")) %>%
-  mutate(adult_movement = adult_movement / max(adult_movement)) %>%
-  gather(scenario, density_ratio, contains("density_ratio")) %>%
-  group_by(experiment, scenario) %>%
-  filter(years_protected == max(years_protected)) %>%
-  mutate(density_ratio = density_ratio * mpa_size) %>% 
-  ggplot(aes(y = pmin(4, mpa_effect), x = pmin(4, density_ratio - 1))) +
-    geom_hline(aes(yintercept = 0)) + 
-  geom_vline(aes(xintercept = 0)) + 
-  geom_abline(aes(slope = 1, intercept = 0), linetype = 2, size = 1.5, color = "black") +
-  geom_point(aes(color = adult_movement, size = mpa_size), alpha = 0.5, show.legend = TRUE)  +
-    geom_smooth(method = "lm", se = FALSE) +
-  facet_wrap(~ rev(scenario), labeller = labeller(`rev(scenario)` = c(biased_density_ratio = "A) Inside vs. Experiment",true_density_ratio = "B) Inside vs. Outside"))) +
-  labs(y = "True % Increase in Biomass", x = "% Response Ratio") +
-  # scale_x_percent(labels = c(percent(seq(-1,1, by = 1)),">200%")) +
-  # scale_y_percent(labels = c(percent(seq(-1,1, by = 1)),">200%")) +
-  scale_color_viridis(guide = wide_hgc, labels = scales::percent_format(accuracy = 1), name = "Fish Movement",
-                      option = "cividis") + 
-  scale_x_continuous(labels = ylabs) + 
-    scale_y_continuous(labels = ylabs) + 
-  theme(legend.position = "top") + 
-  guides(size = FALSE)
-
-  ggsave(dr_plot,filename = "figs/density-ratio.png",
-         width = 8,
-         height = 5)
+To illustrate this problem, we used our bio-economic model to simulate response ratio trajectories for species and MPA coverage representative of the northern Channel Islands.
+These simulations cover the life-histories and MPA sizes seen in the Channel Islands, but vary in key unknowns such as the actual degree of fishing pressure, the timing of density dependence, and the fleet responses to the MPAs.
+For each year of protection, we paired our simlated response ratios to the estimated posterior probablity distributinos of the Channel Island response ratios.
+For example, if the mean response ratio in the year 2006 is 1, we found all simulations with three years of protection that had simulated response ratios near 1, and then pulled out the "true" regional MPA effect from each of those simulations.
+This provides us with a distribution of simulated regional MPA effects that could plausibly generate the types of response ratios actually observed in the Channel Islands.
+
+As reported in @caselle2015, we see evidence for an increasing trend in respones ratios of targeted species over time (Fig.\@ref(fig:rr-plot)).
+This provides strong evidence that the MPAs are providing protected for targeted species within their borders.
+However, our simulation results show thatresponse ratios trends we observe in the data could plausibly be produced by a wide range of regional MPA effects (Fig.\@ref(fig:rr-plot)).
+Response ratios well over 1 were associated with regional MPA effects generally elss than 25%, and many simulations produced large response ratios but regional MPA effects close to zero.
+This can occur if for example fishing pressure is only moderate, adult movement is low, larval dispersal is high, and displaced fishing effort concentrates around the border of the MPAs.
+
+```{r rr-plot, include = FALSE}
+
+response_ratio_plot
+
+ggsave(
+  response_ratio_plot,
+  filename = here::here("documents","figs","response_ratio_plot.png"),
+  width = 4,
+  height = 6
+)
 
@@
   \centering
-  \includegraphics[width=.9\linewidth]{figs/density-ratio.png}
-  \caption{Simulated MPA-size weighted response ratios (x-axis) plotted against true regional change in biomass caused by MPAs (y-axis).
-Each point represents a simulation, and color represents the movement rate of fish in the simulation (expressed as a percentage of the patches a fish is likely to move in one time step).
-The black line shows the one:one fit line, the blue line is a linear fit of the relationship between observed response ratios and true regional change in biomass.
-Inside vs.
-Experiment simulates a true control, while Inside Vs.
-Outside uses a control that may be affected by the treatment.}
-  \label{density-ratio}
+  \includegraphics[width=.9\linewidth]{figs/response_ratio_plot.png}
+  \caption{90% Posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis) in blue.
+Simulated MPA effects associated with simulated response ratios matched to empirical response ratios in red.
+For response ratios, a value of zero indicates that biomass densities of targeted and non-targeted species are identical inside and outside MPAs, a value of 1 indicates that biomass densities of targeted species are 100% greater inside MPAs than outside.
+For MPA effect, a value of zero indicates that mean biomass densities are identical in the with- and without- MPA scenarios.
+A value of 1 indicates that mean biomass densities are 100% greater in the scenario with MPAs than the scenario without MPAs.}
+  \label{rr-plot}
 \end{figure*}
 
-<!-- If response ratios are not a broadly applicable solution to the challenge of estimating regional conservation effects, what then can we do?
-How then can we empirically measure the regional effects of MPAs?
-Intuition and bio-economic theory tells us that the regional conservation effects of MPAs are likely to be highly context dependent across locations and across species within locations, so careful thought must be devoted to an empirical design to detect their effects.
-The perfect experiment would involve two parallel worlds that were identical, except for the implementation of an MPA.
-In world “A”, no MPA would be implemented, and in the facsimile world “B”, the MPA would be implemented.
-Both worlds would be tracked before and after MPA implementation, and the conservation and fishery outcomes would be compared after treatment.
-Instead of two parallel worlds, a similar experimental design would involve random placement of MPAs.
--->
-
-## MPA Effects in the Channel Islands {#estimating .unnumbered}
-
-Given these challenges, what is a viable alternative for estimating the conservation effects of MPAs?
-Building off of @caselle2015, we use a difference-in-difference estimator comparing densities in finfish species targeted by fishing effort (i.e., those potentially affected by an MPA) from those not targeted before and after MPA implementation [see @larsen2019 for a succinct explanation of difference-in-difference estimators in ecological settings].
+While we see empirical evidence of increasing response ratios of targeted species in the Channel Islands, theory indicates that this may not be a reliable indicator of the regional effect of the MPAs.
+An ideal "control" for estimating the regional effects of MPAs would a) control for regional environmental shocks unrelated to the MPAs b) be unaffected by treatment bias (e.g.
+preferential placement of MPAs in ideal habitat) c) be unaffected by the MPAs themselves.
+While spatial response-ratios can help control for environmental shocks and, in theory, treatment bias, at the scale of the Channel Islands it is unlikely that we can strongly say that outside-MPA sites are themselves unaffected by the MPAs we are trying to evaluate.
+
+Building off of @caselle2015, we propose an alternative identification strategy utilizing biomass densities of species that are not directly targeted by fishing as our control group (non-targeted), and biomass densities of species targeted by fishing as our treatment group.
 Targeted species in the Northern Channel Islands, California, include commercially important finfish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
@@
 Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species).
-
-```{r raw-trend, fig.cap="Centered and scaled mean annual biomass density of included species (faded lines) and smoothed means of targeted and non-targeted groups, and mean (darker lines) and 95% confidence interval of the mean (ribbon) over time", include = FALSE}
-raw_did_plot <- base_run$data[[1]] %>%
-  select(year, targeted,classcode, log_density, any_seen) %>%
-  mutate(density = exp(log_density) * any_seen) %>%
-  group_by(classcode,targeted, year) %>%
-  summarise(md = mean(density)) %>%
-  group_by(classcode,targeted) %>%
-  mutate(smd = (md - mean(md)) / sd(md)) %>%
-  ggplot(aes(year, smd, color = targeted == 1, fill = targeted == 1)) +
-  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red", size = 1.5) +
-  geom_line(aes(group = interaction(targeted, classcode)), alpha = 0.5) +
-  geom_smooth(size = 2) +
-  labs(y = "Centered and Scaled Mean Density", x = "Year") +
-    scale_color_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
-      scale_fill_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
-    labs(y = "Centered and Scaled Mean Density", x = "Year") + 
-  theme(legend.position = "top",
-        panel.grid.major.x = element_blank(),
-        panel.grid.minor.x = element_blank()
+We use a Bayesian difference-in-difference estimator to estimate any difference in mean total biomass densities of finfish species targeted by fishing effort (i.e., those potentially affected by an MPA) and those species not targeted by fishing before and after MPA implementation [see @larsen2019 for a succinct explanation of difference-in-difference estimators in ecological settings].
+The result of this regression is an estimate of the effect of the MPAs on the mean total biomass densities of targeted species throughout the Channel Islands.
+
+As conventional wisdom and theory would suggest, over the first three years of implementation (2003-2006), the effects of the MPAs is unclear, with support for a small negative effect to a substantially positive effect (with much higher probability of a small positive effect).
+Over the next six years we see signs of an increasingly positive MPA effect, peaking in 2009-2011 with a median estimate of MPA effect of 75% on mean total bioamss of targeted species.
+These empirical estimates are in line with the range of outcomes that our simulation model suggests are plausible given the rough characteristics of the northern Channel Island MPAs.
+However, in the subsequent years the trend reverses itself, and for the years 2015-2017 we once again see no clear effect of the MPAs Fig.\@ref(fig:mpa-effect)
+
+```{r mpa-effect, include=FALSE}
+
+mpa_effect_plot
+
+ggsave(
+  mpa_effect_plot,
+  filename = here::here("documents","figs","mpa_effect_plot.png"),
+  width = 6,
+  height = 4
 )
 
-raw_did_plot <- raw_did_plot +
-      scale_color_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "") +
-        scale_fill_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "")
-
-pdf(file =  "figs/pop-trends.pdf",width = 8, height = 5, useDingbats = TRUE)
-print(raw_did_plot)
-dev.off()
-
 ```
 
-\begin{figure}%[tbhp]
+\begin{figure*}%[tbhp]
   \centering
-  \includegraphics[width=1\linewidth]{figs/pop-trends.pdf}
-  \caption{Centered and scaled mean annual biomass densities of included targeted (red) and non-targeted (blue) fish species (faded lines) and smoothed means with 95\% confidence intervals over time (darker lines and ribbons).
-Vertical red dashed line shows year of MPA implementation.}
-  \label{pop-trends}
-\end{figure}
-
-Using this model, we estimate...
-XX.
-As we might expect from bio-economic modeling, we detect no significant difference between targeted and non-targeted biomass densities for the first XX years of protection.
-However, over time the model assigns greater and greater probability to MPA-mediated increases in targeted fin-fish biomass densities, up until XX, at which point th trend reverses itself.
-We detect no significant MPA effect between 2015 and 2018 (Fig.\@ref(fig:)).
-
-<!-- The raw trends in the biomass densities appear to show parallel trends in the densities of targeted and non-targeted fishes pre-MPA, with the targeted group diverging first positively and then negatively over time (Fig.\ref{pop-trends}).
-However, these raw trends could be caused by any number of confounding variables besides the effects of the MPAs themselves, from changes in observer skill and sampling locations, to shifts in environmental conditions.
-Using our difference-in-difference model in an effort to control for these factors, we found no significant difference in the biomass densities of targeted and non-targeted fishes pre-MPA (before 2003).
-Following the implementation of the MPAs in 2003, we see evidence of an increasing trend in densities of targeted relative to non-targeted fish, reaching a peak value in 2014.
-After 2014, densities of targeted species appear to have declined relative to what we would expect based on the non-targeted group (Fig.\ref{did-plot}).
--->
-
-```{r did-plot, include = FALSE}
-
-did_plot <- did_plot +
-  scale_y_continuous(name = "~Divergence from Non-Targeted", labels = scales::percent_format(accuracy = 1))
-
-ggsave(did_plot, filename = "figs/did_plot.png",
-       width = 8, height = 5)
-
-# pdf(file =  "figs/did_plot.pdf",width = 8, height = 5, useDingbats = TRUE)
-# print(did_plot)
-# dev.off()
+  \includegraphics[width=.9\linewidth]{`r here::here("documents","figs","mpa_effect_plot.png")`}
+  \caption{Results of difference-in-difference regression estimating the regional effect of the northern Channel Island MPAs on mean total biomass densities of targeted species (difference in mean total biomass density of targeted species over time relative to expected levels using non-targeted species as a control).
+Grey distibutions show 90%XX posterior probability distribution of estimated MPA effect; red point is median estimated effect, thicker red section 50% credible interval, thinner red line 90% credible interval).
+Blue distributions in background show range of MPA effects produced by simulation model tuned to reflect the dybnamics of the Northern Channel Island MPAs (black dashed line is median simulated value).
+Results are estimated in three blocks, including years greater than or equal to left-hand value and less than right-hand value.}
+  \label{mpa-effect}
+\end{figure*}
+
+Discussion {#discussion .unnumbered}
+------
+
+1. Why are the results from the channel islands imporant and what do they tell us
+
+2. Put the channel islands results in context -
+
+Ideally then we want a control for broad environmental shocks to the region that is independent of factors such as selection bias, biological spillover, and fishing concentration.
+
+1. multiple lines of evidence
+
+2. Theory before empirics
+
+3. Patience
+
+The network of MPAs implemented in the Northern Channel Islands in 2003 provides an ideal opportunity to examine the regional effects of proteced area networks.
+This analysis is to our knowledge the first empirical estimate of the regional effects of such a large MPA network on a broad assemblage of targeted finfish.
+Despite the well enforced nature of these MPAs [xx], and the presence of a rigorous scientific monitoring program, we are unable to estimate a clear effect of protection.
+While response ratios have targeted species have increased over time, providing evidence of effective *within MPA* protection, these response ratios are on their own an unrealible indicator of the regional conservation effects of the MPAs.
+Our difference-in-difference estimates a high probability of increasinly positive MPA effects over the first nine years of protection, only to see these estimated gains reversed from 2012 to 2017.
+For the most recent time period, we can detect no clear positive or negative effect of the Channel Island MPAs Fig.\@ref(fig:mpa-effect)
+
+How can we explain these results?
+The estimated effects from 2003 to 2014 are well in line with theoretical expectations given the species and network in question.
+How might we explain the most recent results though?
+One explanation may lie in fleet dynamics.
+Much of the theoretical literautre on MPAs assumes that all else being equal bigger reserves produce bigger conservation gains [@white2011].
+However, to our knowledge all of these models simulate fleet dynamics through assumptions about fishing mortality rates (e.g.
+concentration of fishing mortality @halpern2004).
+The assumption of these models is that fishers determine an amount off effort to exert, and distribute that effort outside the MPA in response to some function.
+
+An alternative and to our knowledge unexplored (in the context of MPAs) fleet model though is a "constant-catch" strategy.
+Under this model, fishers have a catch objective, and exert as much (or little) effort as needed to achieve that objective.
+While a constant-catch greater than MSY is not possible over the long-term under the assumptions of our model, over the short-term a constant-catch scenario is not implausible.
+Subsistence fisheries may use a constant-catch style policy over the short-term, as they seek to ensure that their food needs are met.
+More industrial fisheries may have pre-arranged agreements with buyers to deliver set amounts of fish.
+Constant-catch dynamics might also occur in fisheries with constraining quotas that are not updated after the implementation of MPAs.
+Interestingly, when we simulated the effects of MPAs under different fleet models, the "constant-catch" scenario stood out as the only way for MPAs to actually produce a net conservation loss (Fig.\@ref(fig:XX)).
+While open-access fishing strategies can result in "scorched earth" scenarios where the only fish left are found inside the reserve, across all `r n_distinct(outcomes$experiment)` simulations the net effect of the reserves was still positive.
+Under a constant-catch scenario though, fishers have to fish much harder than before to get the same catch from a smaller part of the population, reducing the size structure of the population and, under XX% of our constant-catch simulations producing a net conservation loss (Fig.XX).
+XX
+
+Can the the non-trivial probability that the Channel Islands MPAs caused a net conservation loss (taking the difference-in-difference results at face value) in the most recent years be explained by the local fleets following a constant-catch strategy?
+While we do not have access to fine scale fishing data from the islands directly, reported catches for the species of interest in the Santa Barbara region in fact exhibit an overall downward trend in the years post reserve (see SI, XX).
+We can most likely rule a negative MPA effect caused by a constant-catch fishing strategy then.
+What then is another explanation?
+The parallel trends assumption of our model appears to be valid initial years of protection, and we do not detect any clear evidence of interaction between the targeted and non-targeted species (e.g.
+depressed levels of non-targeted species as a result of predation from increased biomass densities of targeted species, see Fig.SXX).
+
+However, the Santa Barbara Channel experience a massive heatwave in XX.
+The Channle Islands represent the warmest edge of the distribution of many of the targeted species in our study.
+Conversely, many of the represented non-targeted species are sub-tropical.
+As a result, we hypothesize that the blob even of XX broke down the parllel trends assumption between the targeted and non-targeted species in our study, as the targeted species were more negatively affected by the warming temperatures.
+This is supported by the fact that we observe similar declines in targeted species both outside and inside the reserves (Fig.XX).
+Assuming that the reserves are sufficiently well enforced and large to protect targeted species biomass (an assumption supported by the response ratio results), if the cause of recent declines was due to increases in fishing pressure we would expect to see substantial declines only in the fished areas (Fig.
+XX).
+
+### When Can We Detect the Effects of MPAs?
+
+The Northern Channel Islands would appear to be an ideal system to study the effects of MPA networks on regional conservation outcomes.
+Simulation analysis built on existing MPA theory suggests that these reserves were likely to have a positive but modest effect (50% of simulated effects sizes less than 20% after 15 years of protection) on mean total biomass of targeted species.
+From 2003 (the year of MPA implementation) to 2011, our empirical results estimate a a highly uncertain but overall positive effect of the Channel Islands MPA, roughly in line with what MPA theory would suggest.
+However, these gains dissapear over the most recent years of available data, a phenomenon likely due to dramatic changes in the Channel Islands environment over the last decade that disproportinately affected the targeted species complex.
+Combined with the challenges we have demonstrated of using response ratios as a measure of regional impacts, we are left with an unclear picture of the overall effect of the Channel Island MPAs on biomass densities of targeted species after 14 years of protection.
+
+What does this result in a system as well-studied and well-enforced as the Channel Islands suggest for our ability to assess the  effects of other MPAs around the world?
+First, much of the empirical MPA literature has focused on the relatively clear signals provided by response ratios [@lester2009;].
+While we often see biomass densities of targeted species on the order of 400% higher inside MPAs than outside, as discussed in @ferraro2018 and expanded on here response ratios even as high as these are not a reliable indicator of the regional effects of MPAs.
+While response ratios can provide important evidence as to the effectivness of the MPAs in protecting fishes within their borders, if our interest is in estimating the effect of MPAs on the broader ecosystems in which they are located we must look for additional lines of evidence.
+
+The difference-in-difference strategy utilized here presents an alternative strategy, that while not without its own strict caveats presents some potential improvements over response-ratios as a means of estimating regional conservation effects.
+Given the natural variability of marine ecosystems, and the strong challenges of obtaining accurate samples from marine environments, how large of an effect would an MPA network have to allow a difference-in-difference strategy such as this to be a reliable measure of MPA effects?
+
+To provide guidance on this important question, we simulated data from a range of scenarios with increasing MPA effect size along with varying degrees of observation error and natrual recruitment variation.
+As an added measure, we include scenarios in which the sampled species go through recruitment regimes, which may be positive for both targeted and non-targeted species, or positive for non-targeted species and negative for targeted species.
+We then used a simlpe Bayesian difference-in-difference regression on these simulated data and estimated the percent error in the true MPA effect.
+
+XX .
+Mean absolote percent error dropped below XX% once the true MPA effect exceeded XX%.
+While this if of course no guarantee that any study of such an effect size will be successful (many model runs still ahd high error even at this effect size), our results help illustrate the effect size that may be required if managers which to have a given degree of confidence in their results.
+XX This is merely an illustrative exercise, omitting clearly critical factors such as detection probability and sampling strategy.
+However, since nearly any omission which one can think of would make an MPA effect harder to detect, not easier, these results serve as a useful floor for the likely uncertainty in estimating MPA effects.
+
+```{r}
+
+ggsave(
+  validation_plot,
+  filename = here::here("documents","figs","validation_plot.png"),
+  width = 6,
+  height = 4
+)
 
@@
 
-\begin{figure}%[tbhp]
+\begin{figure*}%[tbhp]
   \centering
-  \includegraphics[width=1\linewidth]{figs/did_plot.png}
-  \caption{Estimated divergence in biomass densities of targeted and non-targeted fishes throughout the Channel Islands (integrated across inside and outside of MPAs).
-MPAs are implemented in 2003 (red dashed line).
-Colored cells in the background indicate the frequency of MPA conservation effects predicted by our simulation modeling for MPAs similar in design and species composition to the Channel Islands in each year.}
-  \label{did-plot}
-\end{figure}
-
-<!-- To the extent that we can select control sites that are both sufficiently similar to the region treated with MPAs and are unaffected by biological or economic spillover, response ratios may be reliable estimators of regional conservation effects.
-In reality these conditions are unlikely to hold but for a few select cases.
-This becomes especially difficult when networks of MPAs are implemented, such as has occurred in California.
-In this case, control sites are generally located within a few kilometers from MPA borders, making biological spillover and fishing concentration highly likely.
-In these circumstances, without robust statistical controls, response ratios may provide an unreliable estimate of the regional conservation effect of marine protected areas.
--->
+  \includegraphics[width=.9\linewidth]{`r here::here("documents","figs","validation_plot.png")`}
+  \caption{Validation}
+  \label{val-plot}
+\end{figure*}
+
+This finding begs an important question then: when should we expect to see MPA effects big enough to stand a reasonable chance of detection?
+To address this, we simulated `r n_distinct(outcomes$experiment)` MPA scenarios across a wide range of life histories, network designs, and fishing dynamics (see SI for a full description of scenarios).
+Using these simulations, we can consider what we might expect the regional effects of an MPA network to be under a with range of plausible circumstances.
+Interested users can explore the use of this model for simulation at [danovando.shinyapps.io/simmpa](https://danovando.shinyapps.io/simmpa/).
+
+Suppose that we are willing to tolerate a MAPE of XX%.
+Two of the most critical drivers of MPA performance are the size of the MPA and the degree of fishing pressure.
+Looking across these two variables, our simulations suggest if the MPA network covers XX% or more of a species range and/or pre-MPA depletion is greater XX we might expect an effect size with a reasonable chance of detection.
+While recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, for commercial finfish many coastal MPA networks are likely to cover areas more in line with the Channel Islands (20%) (Fig.\@ref(fig:XX)).
+
+These results would seem to suggest some rather simple rules of thumb: put an MPA of sufficient size on a fishing population and one can expect large results.
+However, as discussed earlier, the MPA literature has highlighted a large number of variables beyond simply size and fishing pressure that can affect performance.
+To address this, we can
+
+plot error as a function of MPA effect.
 
 ```{r}
-fish_v_fishing <- processed_grid %>%
-  select(-density_ratio) %>%
-  unnest() %>%
-  rename(fishery_effect = mpa_effect1) %>%
-  group_by(experiment) %>%
-  mutate(year = 1:length(year)) %>%
-  ungroup() %>%
-  mutate(years_protected = year - year_mpa + 1) %>%
-  left_join(outcomes %>% select(experiment, year, depletion),
-            by = c("experiment", "year"))
-
-fish_v_fishing_plot <- fish_v_fishing %>%
-  filter(years_protected == max(years_protected)) %>%
-  ggplot(aes(pmin(2, mpa_effect), pmin(2, fishery_effect))) +
-  geom_abline(aes(slope = 1, intercept = 0)) +
-  geom_smooth(method = "lm") +
-  geom_hex()
+
+ggsave(
+  expected_mpa_effect_plot,
+  filename = here::here("documents","figs","expected_mpa_effect.png"),
+  width = 6,
+  height = 4
+)
 
 ```
+
+\begin{figure*}%[tbhp]
+  \centering
+  \includegraphics[width=.9\linewidth]{`r here::here("documents","figs","expected_mpa_effect.png")`}
+  \caption{Expectation}
+  \label{expect-plot}
+\end{figure*}
+
+we hav shown that even in a well studied network, MPA effects on the order 0f 0-25% will be very challenging to detect.
+Under what circumstances might we expect MPA effects large enough to facilitate detection even in the face of observation and process error?
+
+THe persistently high response ratios suggest that despite overall decreases in targeted biomass densities inside and outside MPAs, the MPAs may still be providing protection within their borders.
+However
+
+we hav shown that even in a well studied network, MPA effects on the order 0f 0-25% will be very challenging to detect.
+Under what circumstances might we expect MPA effects large enough to facilitate detection even in the face of observation and process error?
+
+We tested the effects of different combinations of these theoretical drivers of MPA conservation effects using over 10,000 simulated MPA scenarios.
+A striking outcome of this simulation analysis is how the level of fishing pressure can cause dramatic and predictable shifts in conservation outcomes.
+We measure fishing pressure by the degree of simulated depletion (percent of unfished biomass gone, such that a depletion of 0% means the population is at 100% of unfished biomass, 100% means 0% of unfished biomass remains) in the scenario without MPAs.
+When MPAs were placed in relatively unexploited fisheries (depletions less than 40%), the median regional conservation effect was barely noticeable (`r small_effect`%).
+For ecosystems fished near MSY levels (depletion of 40%-60%), the median conservation effect was `r medium_effect`%, while for overfished ecosystems (without-MPA depletions above 60%) the median regional conservation effect was over  `r big_effect`% (Fig.\ref{conservation-effects}-A).
+Across all of these groups though, effects as low as declines of -50% and as high as an increase of more than 100% were possible.
+It should be noted though that these runs represent the range of scenarios evaluated in our simulation, and that certain types of scenarios may be much more likely than others in the real world.
+To put these results into context, the FAO estimates that 7% of the worlds fisheries with status estimates fall into the relatively unexploited category, 60% fall into the fully fished category, and 33% fall into the heavily fished category [@fao2018], though works that include a broader range of fisheries estimate that 50% or more of stocks to fall into the heavily fished category [@costello2012;@rosenberg2017].
+
+These simple results support the intuitive conclusion that bigger MPAs on overfished populations produce large positive conservation outcomes.
+However, when combined with other factors such as fleet dynamics, life history, and MPA size, the conservation effects of MPAs can vary widely in magnitude and timing, a finding corroborated by empirical evidence [@starr2015;@caselle2015;@nickols2019;@villasenor-derbez2019] (Fig.\ref{conservation-effects}-B).
+As the area protected inside MPAs increases, positive effects become more likely, but even MPAs protecting most of a population's footprint can produce conservation effects as low as 0% and as high as 200%.
+The degree of depletion (i.e.
+fishing pressure) has a clearer signal, with small conservation benefits of MPAs when depletion is low, and larger effects when depletion is high, but again even for severely depleted populations the conservation effects of MPAs can vary widely (although nearly always with population benefits).
+
+While positive conservation outcomes were much more likely to occur across our simulations (`r percent(pos_runs, accuracy = 1)` of simulations), the `r percent(1 - pos_runs, accuracy = 1)` of runs that produced net conservation losses are of particular interest.
+These runs occurred almost exclusively when the simulated fleet followed a "constant-catch" fishing strategy.
+Under the constant-catch fleet model, fishing communities (or fishery policies) seek to catch the same amount regardless of the presence of an MPA.
+While a constant-catch greater than MSY is not possible over the long-term under the assumptions of our model, over the short-term a constant-catch scenario is not implausible.
+Subsistence fisheries may use a constant-catch style policy over the short-term, as they seek to ensure that their food needs are met.
+More industrial fisheries may have pre-arranged agreements with buyers to deliver set amounts of fish.
+Constant-catch dynamics might also occur in fisheries with constraining quotas that are not updated after the implementation of MPAs.
+These patterns suggest that coupling MPA implementation with appropriate management reforms that reduce the effects of constant-catch strategies can move the predicted conservation effects of MPAs to being nearly uniformly positive.
+The possibility of constant-catch strategies to result in net MPA conservation losses highlights the critical importance of understanding the broader management context in which MPAs are placed.
+
+### How can we use response ratios
+
+Unfortunately, neither of these experimental designs has been implemented, because parallel worlds do not exist and MPAs are never placed at random.
+Careful causal inference requires some other means of controlling for biases introduced by factors such as the MPA siting process and biological spillover and concentration of fishing effort  [@ferraro2018;@larsen2019].
+Conservation effects are typically estimated by comparing organism densities or biomass inside MPAs to densities or biomass in selected control sites outside MPAs, which we will refer to as response ratios.
+@lester2009 and @halpern2002 present meta-analyses of hundreds of such studies.
+These results often find massively higher densities and biomasses inside MPAs than outside [median biomass increases of  over 400%, @lester2009].
+While response ratios are intended as a measure of the effectiveness of reserves within their borders, these result stand in contrast to the smaller region wide effect sizes predicted by our simulation model.
+This raises the question: How reliable are density or biomass ratios as estimators of regional conservation effects?
+
+Failure to account for these spillover effects can result in a biased estimate of the true effect of a policy such as MPA placement [@larsen2019].
+In an ideal setting, the control site used in a response ratio would be a perfect counterfactual for what would have occurred without the MPA.
+We used our simulation model to approximate this scenario, calculating the response ratio as the ratio of densities inside the MPAs relative to the overall density from the paired simulation without MPAs multiplied by the area of the population in an MPA (to account for the fact that we would expect the effectiveness of a response ratio as a proxy for regional conservation effects to scale with the size of the MPA relative to the range of the species, Fig.\ref{density-ratio}-A).
+In this idealized case, the response ratio is often a reasonable estimator for the regional conservation effect.
+
+What happens when the “control” sites are in fact affected by the application of the MPA through biological spillover and fishing concentration?
+In order to address this possibility, we calculated a new response ratio as the mean density inside MPAs relative to the mean density outside the MPAs (both weighted by distance from MPA borders and scaled by MPA size).
+We only include simulations in which habitat and larval dispersal rates are identical inside and outside of MPAs, to approximate a scenario in which treatment and control sites have been paired by ecological characteristics.
+Under these circumstances the response ratio is a somewhat biased and very inaccurate estimator of the true regional effect of MPAs.
+
+While on average high response ratios correspond with high true regional effects, the variance is high, so the response ratio alone cannot be used as a reliable proxy for the true regional effect, without accounting for the sources of biases discussed here.
+For higher movement rates, response ratios were frequently near 0\% (potentially leading stakeholders to conclude that the MPA had been ineffective), when in fact in many cases the true effect of the MPA was highly positive.
+For highly sedentary species, extremely high response ratios could create the appearance of massive conservation gains when in fact the net effect on the regional population has been near zero or even negative (Fig.\ref{density-ratio}-B).
+
+We note that spatial before-after-control-impact (BACI) studies present a potential improvement over response ratios, controlling for differences in mean densities in the treated and control regions as well as MPA independent trends, but are much rarer due to the need for extensive pre-MPA monitoring.
+However, spatial BACI still requires that the MPAs do not affect the control sites, and as such has many of the same limitations as response ratios as an estimator of regional conservation effects.
+
+MPAs are an important part of the marine resource management toolbox.
+Under ideal circumstances they can protect individual species and ecosystem linkages, while supporting local economies through tourism and fishing opportunities.
+One rationale for the expansion of MPAs is that they will deliver net conservation benefits both inside *and outside* their borders.
+To date, this assumption is insufficiently tested, and this is the focus of our paper.
+Our results show that regional conservation benefits of MPAs are highly context dependent and in many circumstances, are likely to be so small that they are nearly impossible to detect empirically.
+Indeed, this is exactly what we found in our empirical case study from the Channel Islands, California, USA.
+
+We find regional conservation gains from MPAs in `r percent(pos_runs, accuracy = 1)` of our simulations (Fig.\ref{conservation-effects}).
+MPAs covering less than 25% of the population range were unlikely to produce regional conservation gains above 10% unless the stock would have been severely overfished without the MPA.
+Given that few marine species on the planet have more than 25% of their entire range protected in MPAs, this suggests that large regional gains are unlikely to be common unless the overall extent of MPAs grows globally.
+Large MPAs protecting highly depleted stocks almost always produced large regional conservation gains, as should be expected.
+
+The median conservation effect of `r round(100*median(eqo$mpa_effect))`\% stands in contrast to the large within-MPA effects on biomass densities reported by @halpern2003a and @lester2009.
+This difference is due in part to the fact that within-MPA effects are likely to be much larger than regional-MPA effects, but may also reflect the many challenges of translating response ratios into regional conservation effects.
+Our simulation analysis shows that without proper controls (in particular for the effects of fleet redistribution), response ratios well over 100% are entirely plausible even when the actual conservation effect is much lower (and *vice versa*).
+
+We were not able to detect a statistically significant difference in the densities of targeted and non-targeted finfish species over the 13 years of MPA protection in the Channel Islands covered by our analysis.
+The data (and model assumptions) provide support for both negative and positive MPA effects, although the size of these estimated effects is well within the range that our simulation analysis suggests are plausible.
+At face value, this lack of a clear result may seem surprising given the size and carefully studied nature of this MPA network.
+
+Our simulation analysis suggests that this result should perhaps be expected.
+The Channel Islands MPAs cover approximately 20% of the waters in the Channel Islands, and while formal stock assessments are not available for many of the targeted species in our analysis, what evidence we have suggests that, as a group, these fish are on average not heavily overfished.
+Some species, such as California sheephead and blue rockfish were likely below target levels during the period [@alonzo2004;@dick2017], but projections based upon the overall average response across all species will likely suggest modest benefits even if a subset of species could experience much larger population gains.
+Harkening back to our simulations, we expect the average percentage difference in densities of targeted species with and without MPAs to be modest (Fig.\ref{did-plot}).
+Effects of this size are likely to be challenging to detect empirically given the large natural variation of marine ecosystems (especially temperate reefs) and the observation error inherent in infrequent annual survey programs such as those provided by PISCO.
+
+```{r}
+tmdata <- life_history_data %>% 
+  filter(targeted == 1, classcode %in% top_species) %>% 
+  select(tm, everything())
+
+median_tm = round(median(tmdata$tm))
+```
+
+In addition, our simulations examine percentage increases in biomass densities.
+As an extreme example, an increase in biomass densities from 2kg/m^2^ to 4kg/m^2^ would be reported as a 100% increase.
+While this chance produces a large percentage increase, this is a small change in absolute biomass densities relative to the variance in the observation process itself (see Fig.S1 for a companion to Fig.\ref{conservation-effects} scaled by absolute population size).
+Beyond these challenges, the median estimated age at sexual maturity for the targeted species included in this study is `r median_tm` years, meaning that the span of this analysis represents less than three generations of MPA protection for half of the the measured species.
+Ongoing monitoring may yet reveal clearer effects.
+Analysis of more rapidly growing and maturing species, e.g.
+spiny lobster, may also reveal clearer signals.
+
+While the MPA effect estimated by our model was not significant in any one year, the positive trend from 2005 to 2014 is evident, as is the sharp decline from 2015-2017.
+To what should we attribute this apparent shift?
+We cannot reject the parallel trends assumption between the targeted and non-targeted species in the years before the MPAs (Fig.\ref{did-plot}).
+But, the Channel Islands region (and the entire West coast of the USA) experienced a dramatic ‘marine heatwave’ beginning in 2014 and persisting through 2016, resulting in part in extremely elevated water temperatures throughout the region  [@gentemann2017].
+Biogeographic differences in the distributions of targeted and non-targeted species may confound the observed effects of MPAs.
+Many of the non-targeted species have warm thermal affinities and have increased in numbers since the heatwave [@freedman2019].
+The targeted group is made up mostly of fishes with cold-water affinities, such as members of the genus *Sebastes*.
+As such, we hypothesize that the recent evidence for a decline in densities of targeted species is due to environmental conditions that disproportionately affect the targeted group (and not for example due to concentrated fishing pressure outside the reserves).
+This hypothesis is supported by the sharp declines in densities of targeted species seen inside the MPAs themselves (see Fig.S27-S29), suggesting a driver other than fishing may be at play.
+
+What do our results imply about the future of MPA science?
+Our simulation model is by no means exhaustive (for example it ignores features such as species interactions, habitat effects, and climate feedback), but captures many of the core factors theorized to affect MPA performance.
+Our results show that while attributes such as MPA size and fishing pressure are important factors in determining the effects of protection, local fleet dynamics and the movement rates of adult fish can dramatically affect the outcomes of protected areas as well  (Fig.\ref{conservation-effects}).
+Far from being a simple tool with clear outcomes, the effects of MPAs can be highly context dependent, requiring - we would argue - a bio-economic model of at minimum the complexity presented here to help communities design and set expectations of MPAs at the tactical level.
+While users may not be able to parameterize every aspect of a model such as that presented here, working with stakeholders to visualize the implications of, for example, different fleet responses to MPA implementation is a critical step in MPA design.
+Readers can use our simulation model to explore the effects of MPAs through an interactive web application available [here](https://danovando.shinyapps.io/simmpa/).
+
+Once simulation models have been used to help design an MPA, how can users evaluate whether it is achieving their objectives?
+Response ratios are commonly used as evidence for conservation outcomes of MPAs [@lester2009], but as suggested in @ferraro2018 and @larsen2019 and demonstrated here, without careful attention to the design of control sites (accounting for example for the displacement of effort by MPAs), response ratios may be highly unreliable estimators of regional MPA effects.
+As @larsen2019 suggests, there are many potential alternatives for estimating the effects of MPAs that better account for the challenges of causal inference.
+We applied one such approach here (a difference-in-difference estimator), and yet were still unable to reach robust conclusions as to the effect of MPAs on the density of targeted species in the Channel Islands, due to the likely small size of the true effect relative to the strength and variability of environmental drivers.
+
+While this does not mean that all MPAs will face similar challenges in estimating their effects, our results in the rigorously studied Channel Islands system make clear that in many instances empirically detecting a clear regional effect of MPAs may not be possible.
+How then should stakeholders go about adaptively monitoring and managing MPAs?
+Simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can perhaps be tuned to focus on the species groups that have the highest chance of a detectable effect size [@nickols2019].
+Expanding data collection to include robust monitoring of fleet dynamics may help statistical approaches to isolate the true effect of MPAs on conservation outcomes by allowing for improved control of the effects of fishing dynamics, and allow managers to take into account potential negative interactions between MPAs and fleet dynamics such as those that may occur under constant-catch dynamics.
+Whenever possible monitoring programs should be implemented prior to MPA implementation to provide a pre-treatment benchmark.
+Non-equilibrium analyses (Fig.\ref{did-plot}) also help set expectations for effect sizes over time.
+Beyond that, educating communities about the challenges of estimating the effects of MPAs can help set expectations, so that a lack of a clear effect is not necessarily viewed as a failure of the program, but rather considered in the context of reasonable expectations.
+While this paper has focused on the conservation outcomes of MPAs, future work must also address the challenge of predicting and estimating the fishery impacts of protected areas.
+
+As the number and size of global MPA networks increase, it is critical that we both set appropriate expectations for their outcomes, and plan how we will monitor the performance of these protected areas over time.
+While the history of MPA science has made important strides in helping us understand the dynamics of protected areas, the future of MPA science must directly tackle the challenge of evaluating the performance of these MPAs, and adapting their design as needed to best achieve objectives.
+Commonly employed metrics such as response ratios may be applicable in some circumstances, but can have severe shortcomings as metrics of regional conservation effects.
+Dependence on unreliable estimators of MPA effects may lead to stakeholders incorrectly attributing negative environmental shocks as MPA failures, or interpreting data arising from scorched-earth fishing outside MPAs as a conservation success.
+Bio-economic modeling can help frame community expectations, reducing the potential for a reduction in support if unrealistic conservation or fishery gains are not realized.
+Statistical approaches that explicitly address complications such as the spatial spillover effects of MPAs (such as the difference-in-difference approach used here) may give users an improved understanding of the performance of their MPAs.
+Effective use of simulation and statistics to clearly communicate what we should expect and what we can detect from MPAs is critical in ensuring that MPAs play effective roles in fisheries management and marine conservation.
+
+## Materials and Methods
+
+We present here critical characteristics of our simulation model and regression approach.
+Further details can be found in the Supplementary Information.
+All analysis were conducted in R [@rcoreteam2019].
+Our main difference-in-difference model was fit using Template Model Builder in R [@kristensen2016].
+Code needed to replicate results can be found [here](https://github.com/DanOvando/regional-effects-of-mpas).
+
+### Simulation Model
+
+Our bio-economic model simulates the effect of MPAs on a spatially explicit age-structured representation of a single species.
+Readers can explore the functionality of the model using an online tool available [here](https://danovando.shinyapps.io/simmpa/).
+See Table.S1 for a complete description of simulation states.
+The model consists of 25 patches with wrapped edges (picture the waters around a circular island).
+For any one simulation we randomly pull a species and its associated life history from the `FishLife` [@thorson2017c] package in R.
+We pair these data with randomly selected values between 0.6 and 0.95 for Beverton-Holt steepness [as parameterized in @mace1994], as well as larval and adult dispersal rates.
+We also randomly assign whether adults have density dependent movement, as well as one of three potential types of density dependence [building off of @babcock2011]:
+
+  1. Local density dependence: Density dependence occurs independently in each patch, and recruits then disperse to nearby patches
+
+  2. Global density dependence: Density dependence is a function of the sum of spawning biomass across all patches, and recruits are then distributed according to habitat quality
+
+  3. Post-dispersal density dependence: Larvae are distributed throughout the system, and then density dependence occurs based on the density of adult biomass at the destination patch
+
+We allow for three potential siting strategies for MPAs.
+In the first, MPAs are randomly placed.
+In the second, we assume that MPAs are placed in preferentially better habit (unfished recruitment is four times greater inside MPA locations).
+In the third, we allow for scenarios in which MPAs are placed in hotspots of larval dispersal.
+In this scenario, patches in which MPAs will be placed have larval dispersal rates four times greater than patches that do not become MPAs.
+
+Along with the fleet dynamics model, each simulation is assigned a random fleet dispersal scenario: uniform dispersal (where the total effort of the fleet is divided evenly among all open patches), catch dispersal (where the total effort of the fleet is divided according to the catchable biomass in each available patch), and profit dispersal (where the total effort of the fleet is divided according to the profit per unit effort in each available patch).
+
+Lastly each simulation is assigned an MPA scenario, defined by the number and size of MPAs, the placement of those MPAs, and the year that the MPAs are put in place.
+Each simulation starts the population off at unfished equilibrium and then beings to apply the fleet model.
+The MPAs are then placed during the randomly selected start year, allowing some runs to explore how the early dynamics of the MPA play out when the fishery and population they are placed on is not already at equilibrium.
+Fishing effort in displaced by MPAs can either concentrate outside or leave the fishery.
+Each simulation is run to equilibrium with and without the selected MPA strategy (holding all else constant).
+We then measure the difference in biomass and fishery catches in each time step in the scenario with and without the MPAs to calculate the conservation and fishery effects of the MPAs over time.
+
+### Difference in Difference Regression
 
@@
 channel_islands <- sf::st_read(here::here("data","cinms_py2","cinms_py.shp"))
-
-rnaturalearth::ne_coastline()
 
@@
 
-```{r}
-ci_effects <- short_term %>%
-  filter(years_protected == max(years_protected),
-         mpa_size  <= 0.25, mpa_size >= 0.15)
-
-med_ci_effect <- median(ci_effects$mpa_effect)
-
-# rethinking::HPDI(ci_effects$mpa_effect, prob = 0.5)
-
-quart_ci_effect <- quantile(ci_effects$mpa_effect)
+```{r raw-trend, fig.cap="Centered and scaled mean annual biomass density of included species (faded lines) and smoothed means of targeted and non-targeted groups, and mean (darker lines) and 95% confidence interval of the mean (ribbon) over time", include = FALSE}
+raw_did_plot <- fitted_data %>%
+  select(year, targeted,classcode, log_density, any_seen) %>%
+  mutate(density = exp(log_density) * any_seen) %>%
+  group_by(classcode,targeted, year) %>%
+  summarise(md = mean(density)) %>%
+  group_by(classcode,targeted) %>%
+  mutate(smd = (md - mean(md)) / sd(md)) %>%
+  ggplot(aes(year, smd, color = targeted == 1, fill = targeted == 1)) +
+  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red", size = 1.5) +
+  geom_line(aes(group = interaction(targeted, classcode)), alpha = 0.5) +
+  geom_smooth(size = 2) +
+  labs(y = "Centered and Scaled Mean Density", x = "Year") +
+    scale_color_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
+      scale_fill_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
+    labs(y = "Centered and Scaled Mean Density", x = "Year") + 
+  theme(legend.position = "top",
+        panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank()
+        )
+
+raw_did_plot <- raw_did_plot +
+      scale_color_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "") +
+        scale_fill_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "")
+
+pdf(file =  "figs/pop-trends.pdf",width = 8, height = 5, useDingbats = TRUE)
+print(raw_did_plot)
+dev.off()
 
@@
 
-Discussion {#discussion .unnumbered}
-------
-
-1. Why are the results from the channel islands imporant and what do they tell us
-
-2. Put the channel islands results in context -
-
-### What Should We Expect from MPAs?
-
-We tested the effects of different combinations of these theoretical drivers of MPA conservation effects using over 10,000 simulated MPA scenarios.
-A striking outcome of this simulation analysis is how the level of fishing pressure can cause dramatic and predictable shifts in conservation outcomes.
-We measure fishing pressure by the degree of simulated depletion (percent of unfished biomass gone, such that a depletion of 0% means the population is at 100% of unfished biomass, 100% means 0% of unfished biomass remains) in the scenario without MPAs.
-When MPAs were placed in relatively unexploited fisheries (depletions less than 40%), the median regional conservation effect was barely noticeable (`r small_effect`%).
-For ecosystems fished near MSY levels (depletion of 40%-60%), the median conservation effect was `r medium_effect`%, while for overfished ecosystems (without-MPA depletions above 60%) the median regional conservation effect was over  `r big_effect`% (Fig.\ref{conservation-effects}-A).
-Across all of these groups though, effects as low as declines of -50% and as high as an increase of more than 100% were possible.
-It should be noted though that these runs represent the range of scenarios evaluated in our simulation, and that certain types of scenarios may be much more likely than others in the real world.
-To put these results into context, the FAO estimates that 7% of the worlds fisheries with status estimates fall into the relatively unexploited category, 60% fall into the fully fished category, and 33% fall into the heavily fished category [@fao2018], though works that include a broader range of fisheries estimate that 50% or more of stocks to fall into the heavily fished category [@costello2012;@rosenberg2017].
-
-These simple results support the intuitive conclusion that bigger MPAs on overfished populations produce large positive conservation outcomes.
-However, when combined with other factors such as fleet dynamics, life history, and MPA size, the conservation effects of MPAs can vary widely in magnitude and timing, a finding corroborated by empirical evidence [@starr2015;@caselle2015;@nickols2019;@villasenor-derbez2019] (Fig.\ref{conservation-effects}-B).
-As the area protected inside MPAs increases, positive effects become more likely, but even MPAs protecting most of a population's footprint can produce conservation effects as low as 0% and as high as 200%.
-The degree of depletion (i.e.
-fishing pressure) has a clearer signal, with small conservation benefits of MPAs when depletion is low, and larger effects when depletion is high, but again even for severely depleted populations the conservation effects of MPAs can vary widely (although nearly always with population benefits).
-
-While positive conservation outcomes were much more likely to occur across our simulations (`r percent(pos_runs, accuracy = 1)` of simulations), the `r percent(1 - pos_runs, accuracy = 1)` of runs that produced net conservation losses are of particular interest.
-These runs occurred almost exclusively when the simulated fleet followed a "constant-catch" fishing strategy.
-Under the constant-catch fleet model, fishing communities (or fishery policies) seek to catch the same amount regardless of the presence of an MPA.
-While a constant-catch greater than MSY is not possible over the long-term under the assumptions of our model, over the short-term a constant-catch scenario is not implausible.
-Subsistence fisheries may use a constant-catch style policy over the short-term, as they seek to ensure that their food needs are met.
-More industrial fisheries may have pre-arranged agreements with buyers to deliver set amounts of fish.
-Constant-catch dynamics might also occur in fisheries with constraining quotas that are not updated after the implementation of MPAs.
-These patterns suggest that coupling MPA implementation with appropriate management reforms that reduce the effects of constant-catch strategies can move the predicted conservation effects of MPAs to being nearly uniformly positive.
-The possibility of constant-catch strategies to result in net MPA conservation losses highlights the critical importance of understanding the broader management context in which MPAs are placed.
-
-```{r conservation-effect, include = FALSE}
-
-pop_time_plot <- outcomes %>%
-  filter(years_protected > 0) %>%
-  ggplot(aes(years_protected, mpa_effect)) +
-  geom_bin2d() +
-  scale_fill_viridis(guide = gc)
-
-ylabs <- c(expression(""<= "-50%"), paste0(seq(-25,75, by = 25),"%"),expression("">= "100%") )
-
-ybreaks = seq(-.50,1, by = .25)
-
-pop_depletion_plot <- outcomes %>%
-  filter(year == max(year)) %>%
-  ggplot() +
-  geom_bin2d(aes(depletion, mpa_effect), binwidth = c(.1, .1),
-             show.legend = FALSE) +
-  scale_fill_viridis(
-  option = "A",
-  # trans = "log10",
-  guide = guide_colorbar(frame.colour = "black",
-  frame.linewidth = 1,),
-  name = "Median Effect"
-  ) +
-  scale_x_continuous(labels = scales::percent_format(accuracy = 1), name = "Depletion") +
-  scale_y_continuous(labels = ylabs, breaks = ybreaks, name = "")
-
-pop_depletion_plot <- outcomes %>%
-    filter(year == max(year)) %>%
-    mutate(
-      rdep = cut(
-        depletion,
-        breaks = seq(0, 1, by = .2),
-        labels = seq(0, .8, by = .2),
-        include.lowest = TRUE
-      )
-      ,
-      reff = cut(
-        mpa_effect,
-        breaks = seq(-.6, 1, by = .2),
-        labels = seq(-.6, .8, by = .2),
-        include.lowest = TRUE
-      )
-    ) %>%
-    group_by(rdep, reff) %>%
-    count() %>%
-    group_by(rdep) %>%
-    mutate(pn = n / sum(n)) %>%
-    ungroup() %>%
-    mutate(rdep = as.numeric(as.character(rdep)),
-           reff = as.numeric(as.character(reff))) %>%
-    ggplot(aes(rdep, reff, fill = pn)) +
-    geom_raster(hjust = 1, vjust = 1, show.legend = FALSE) +
-    scale_fill_viridis(
-      option = "A",
-      # trans = "log10",
-      guide = hgc,
-      name = "% of Sims",
-      labels = scales::percent_format(accuracy = 1)
-    ) +
-    scale_x_continuous(labels = scales::percent_format(accuracy = 1),
-                       name = "Depletion",
-                       expand = expand_scale(add = c(0,0))) +
-    scale_y_continuous(labels = ylabs,
-                       breaks = ybreaks,
-                       name = "")
-
-  sim_plot <- outcomes %>%
-    mutate(rsize = plyr::round_any(mpa_size, 0.25),
-           reff = plyr::round_any(mpa_effect,0.2)) %>%
-    filter(rsize == 0.25, depletion < 0.8, depletion > 0.2,
-           scientific_name %in% sim_life_history$scientific_name
-           ) %>%
-    mutate(year = years_protected + 2001) %>%
-    filter(year %in% did_betas$year, years_protected >1) %>% 
-    group_by(year, reff) %>% 
-    count() %>% 
-    group_by(year) %>% 
-    mutate(sc = n / sum(n))
-
-  pop_size_plot <- outcomes %>%
-    filter(year == max(year)) %>%
-    mutate(
-      rsize = cut(
-        mpa_size,
-        breaks = seq(0, 1, by = .2),
-        labels = seq(0, .8, by = .2)
-      )
-      ,
-      reff = cut(
-        mpa_effect,
-        breaks = seq(-.6, 1, by = .2),
-        labels = seq(-.6, .8, by = .2)
-      )
-    ) %>%
-    group_by(rsize, reff) %>%
-    count() %>%
-    group_by(rsize) %>%
-    mutate(pn = n / sum(n)) %>%
-    ungroup() %>%
-    mutate(rsize = as.numeric(as.character(rsize)),
-           reff = as.numeric(as.character(reff))) %>%
-    ggplot(aes(rsize, reff, fill = pn)) +
-    geom_raster(hjust = 1, vjust = 1) +
-    scale_fill_viridis(
-      option = "A",
-      # trans = "log10",
-      guide = hgc,
-      name = "% of Sims",
-      labels = scales::percent_format(accuracy = 1)
-    ) +
-    scale_x_continuous(labels = scales::percent_format(accuracy = 1),
-                       name = "Range in MPA",
-                       expand = expand_scale(add = c(0,0))) +
-    scale_y_continuous(labels = ylabs,
-                       breaks = ybreaks,
-                       name = "Conservation Effect") +
-    theme(legend.position = "top") +
-    labs(title = 'B)')
-
- pop_facet_effect_plot <- outcomes %>%
-  filter(year == max(year)) %>%
-   select(mpa_effect, mpa_size, depletion) %>%
-   gather(measure, value, -mpa_effect) %>%
-   ggplot() +
-   geom_bin2d(aes(value, mpa_effect), binwidth = c(.1, .1), show.legend = TRUE) +
-  scale_fill_viridis(
-  option = "A",
-  # trans = "log10",
-  guide = hgc,
-  name = "# of Sims"
-  ) +
-   facet_wrap(~measure, labeller = labeller(measure = facet_labels), strip.position = "bottom") +
-   scale_x_percent(name = "") +
-   scale_y_percent(name = "Conservation Effect") + 
-   theme(legend.position = "top")
-
-  pop_depletion_and_size_plot <- outcomes %>%
-    filter(years_protected >= 0) %>%
-    group_by(experiment) %>%
-    mutate(depletion = plyr::round_any(depletion[years_protected == 0], .1),
-           mpa_size = plyr::round_any(mpa_size, .1)) %>%
-    filter(year == max(year)) %>%
-    group_by(depletion, mpa_size) %>%
-    summarise(median_mpa_effect = median(mpa_effect, na.rm = TRUE)) %>%
-    ggplot(aes(depletion, mpa_size, fill = median_mpa_effect)) +
-    geom_tile() +
-    # geom_contour(aes(z = median_mpa_effect, color = ..level..), show.legend = FALSE) +
-    scale_fill_viridis(labels = scales::percent_format(accuracy = 1),
-                       guide = hgc,
-                       name = "Median Effect") +
-  labs(x = "Depletion", y = "Range in MPA", title = "A)") +
-    scale_y_continuous(labels = scales::percent_format(accuracy = 1),
-                       expand = c(0,0)) +
-    scale_x_continuous(labels = scales::percent_format(accuracy = 1), expand = c(0,0), limits = c(0,1)) + 
-    theme(legend.position = "top")
-
-  # pop_depletion_and_size_plot <-  directlabels::direct.label(pop_depletion_and_size_plot, list("angled.boxes", box.alpha = 0.5))
-
-  pop_combo_plot <-
-    (pop_depletion_and_size_plot) + ((pop_size_plot) / pop_depletion_plot)  + plot_layout(widths = c(1.5, 1)) & theme(
-    plot.margin = unit(c(0, 1, 0, 1), units = "lines"),
-    axis.text.x = element_text(size = 8),
-    legend.box.margin = unit(c(0, 0, 0, 0), units = "lines"),
-    axis.text.y = element_text(size = 10)
-    )
-
-  ggsave(pop_combo_plot,filename = "figs/conservation-effect.png",
-         width = 8,
-         height = 5)
-
-  pdf(
-    file =  "figs/conservation-effect.pdf",
-    width = 8,
-    height = 5,
-    useDingbats = FALSE
-  )
-  print(pop_combo_plot)
-  dev.off()
-
-```
-
-\begin{figure*}%[tbhp]
+\begin{figure}%[tbhp]
   \centering
-  \includegraphics[width=.9\linewidth]{figs/conservation-effect.png}
-  \caption{Median (A) and range (B) of equilibrium regional MPA conservation effects (percent change in total biomass with MPAs relative to without MPAs) across a range of depletion and MPA sizes (and incorporating the full range of scenarios included in our study).
-"Range in MPA" is the percent of patches covered by an MPA, "Depletion" is the depletion that would have occurred in equilibrium without the MPA.
-Colors in panel B represent the percentage of simulations that fell into an effect size bin at each x-axis bin, such that the sum of all percentages for a given x-axis bin is 1.}
-  \label{conservation-effects}
-\end{figure*}
-
-### How can we use response ratios
-
-Unfortunately, neither of these experimental designs has been implemented, because parallel worlds do not exist and MPAs are never placed at random.
-Careful causal inference requires some other means of controlling for biases introduced by factors such as the MPA siting process and biological spillover and concentration of fishing effort  [@ferraro2018;@larsen2019].
-Conservation effects are typically estimated by comparing organism densities or biomass inside MPAs to densities or biomass in selected control sites outside MPAs, which we will refer to as response ratios.
-@lester2009 and @halpern2002 present meta-analyses of hundreds of such studies.
-These results often find massively higher densities and biomasses inside MPAs than outside [median biomass increases of  over 400%, @lester2009].
-While response ratios are intended as a measure of the effectiveness of reserves within their borders, these result stand in contrast to the smaller region wide effect sizes predicted by our simulation model.
-This raises the question: How reliable are density or biomass ratios as estimators of regional conservation effects?
-
-Failure to account for these spillover effects can result in a biased estimate of the true effect of a policy such as MPA placement [@larsen2019].
-In an ideal setting, the control site used in a response ratio would be a perfect counterfactual for what would have occurred without the MPA.
-We used our simulation model to approximate this scenario, calculating the response ratio as the ratio of densities inside the MPAs relative to the overall density from the paired simulation without MPAs multiplied by the area of the population in an MPA (to account for the fact that we would expect the effectiveness of a response ratio as a proxy for regional conservation effects to scale with the size of the MPA relative to the range of the species, Fig.\ref{density-ratio}-A).
-In this idealized case, the response ratio is often a reasonable estimator for the regional conservation effect.
-
-What happens when the “control” sites are in fact affected by the application of the MPA through biological spillover and fishing concentration?
-In order to address this possibility, we calculated a new response ratio as the mean density inside MPAs relative to the mean density outside the MPAs (both weighted by distance from MPA borders and scaled by MPA size).
-We only include simulations in which habitat and larval dispersal rates are identical inside and outside of MPAs, to approximate a scenario in which treatment and control sites have been paired by ecological characteristics.
-Under these circumstances the response ratio is a somewhat biased and very inaccurate estimator of the true regional effect of MPAs.
-
-While on average high response ratios correspond with high true regional effects, the variance is high, so the response ratio alone cannot be used as a reliable proxy for the true regional effect, without accounting for the sources of biases discussed here.
-For higher movement rates, response ratios were frequently near 0\% (potentially leading stakeholders to conclude that the MPA had been ineffective), when in fact in many cases the true effect of the MPA was highly positive.
-For highly sedentary species, extremely high response ratios could create the appearance of massive conservation gains when in fact the net effect on the regional population has been near zero or even negative (Fig.\ref{density-ratio}-B).
-
-We note that spatial before-after-control-impact (BACI) studies present a potential improvement over response ratios, controlling for differences in mean densities in the treated and control regions as well as MPA independent trends, but are much rarer due to the need for extensive pre-MPA monitoring.
-However, spatial BACI still requires that the MPAs do not affect the control sites, and as such has many of the same limitations as response ratios as an estimator of regional conservation effects.
-
-MPAs are an important part of the marine resource management toolbox.
-Under ideal circumstances they can protect individual species and ecosystem linkages, while supporting local economies through tourism and fishing opportunities.
-One rationale for the expansion of MPAs is that they will deliver net conservation benefits both inside *and outside* their borders.
-To date, this assumption is insufficiently tested, and this is the focus of our paper.
-Our results show that regional conservation benefits of MPAs are highly context dependent and in many circumstances, are likely to be so small that they are nearly impossible to detect empirically.
-Indeed, this is exactly what we found in our empirical case study from the Channel Islands, California, USA.
-
-We find regional conservation gains from MPAs in `r percent(pos_runs, accuracy = 1)` of our simulations (Fig.\ref{conservation-effects}).
-MPAs covering less than 25% of the population range were unlikely to produce regional conservation gains above 10% unless the stock would have been severely overfished without the MPA.
-Given that few marine species on the planet have more than 25% of their entire range protected in MPAs, this suggests that large regional gains are unlikely to be common unless the overall extent of MPAs grows globally.
-Large MPAs protecting highly depleted stocks almost always produced large regional conservation gains, as should be expected.
-
-The median conservation effect of `r round(100*median(eqo$mpa_effect))`\% stands in contrast to the large within-MPA effects on biomass densities reported by @halpern2003a and @lester2009.
-This difference is due in part to the fact that within-MPA effects are likely to be much larger than regional-MPA effects, but may also reflect the many challenges of translating response ratios into regional conservation effects.
-Our simulation analysis shows that without proper controls (in particular for the effects of fleet redistribution), response ratios well over 100% are entirely plausible even when the actual conservation effect is much lower (and *vice versa*).
-
-We were not able to detect a statistically significant difference in the densities of targeted and non-targeted finfish species over the 13 years of MPA protection in the Channel Islands covered by our analysis.
-The data (and model assumptions) provide support for both negative and positive MPA effects, although the size of these estimated effects is well within the range that our simulation analysis suggests are plausible.
-At face value, this lack of a clear result may seem surprising given the size and carefully studied nature of this MPA network.
-
-Our simulation analysis suggests that this result should perhaps be expected.
-The Channel Islands MPAs cover approximately 20% of the waters in the Channel Islands, and while formal stock assessments are not available for many of the targeted species in our analysis, what evidence we have suggests that, as a group, these fish are on average not heavily overfished.
-Some species, such as California sheephead and blue rockfish were likely below target levels during the period [@alonzo2004;@dick2017], but projections based upon the overall average response across all species will likely suggest modest benefits even if a subset of species could experience much larger population gains.
-Harkening back to our simulations, we expect the average percentage difference in densities of targeted species with and without MPAs to be modest (Fig.\ref{did-plot}).
-Effects of this size are likely to be challenging to detect empirically given the large natural variation of marine ecosystems (especially temperate reefs) and the observation error inherent in infrequent annual survey programs such as those provided by PISCO.
-
-```{r}
-tmdata <- life_history_data %>% 
-  filter(targeted == 1, classcode %in% top_species) %>% 
-  select(tm, everything())
-
-median_tm = round(median(tmdata$tm))
-```
-
-In addition, our simulations examine percentage increases in biomass densities.
-As an extreme example, an increase in biomass densities from 2kg/m^2^ to 4kg/m^2^ would be reported as a 100% increase.
-While this chance produces a large percentage increase, this is a small change in absolute biomass densities relative to the variance in the observation process itself (see Fig.S1 for a companion to Fig.\ref{conservation-effects} scaled by absolute population size).
-Beyond these challenges, the median estimated age at sexual maturity for the targeted species included in this study is `r median_tm` years, meaning that the span of this analysis represents less than three generations of MPA protection for half of the the measured species.
-Ongoing monitoring may yet reveal clearer effects.
-Analysis of more rapidly growing and maturing species, e.g.
-spiny lobster, may also reveal clearer signals.
-
-While the MPA effect estimated by our model was not significant in any one year, the positive trend from 2005 to 2014 is evident, as is the sharp decline from 2015-2017.
-To what should we attribute this apparent shift?
-We cannot reject the parallel trends assumption between the targeted and non-targeted species in the years before the MPAs (Fig.\ref{did-plot}).
-But, the Channel Islands region (and the entire West coast of the USA) experienced a dramatic ‘marine heatwave’ beginning in 2014 and persisting through 2016, resulting in part in extremely elevated water temperatures throughout the region  [@gentemann2017].
-Biogeographic differences in the distributions of targeted and non-targeted species may confound the observed effects of MPAs.
-Many of the non-targeted species have warm thermal affinities and have increased in numbers since the heatwave [@freedman2019].
-The targeted group is made up mostly of fishes with cold-water affinities, such as members of the genus *Sebastes*.
-As such, we hypothesize that the recent evidence for a decline in densities of targeted species is due to environmental conditions that disproportionately affect the targeted group (and not for example due to concentrated fishing pressure outside the reserves).
-This hypothesis is supported by the sharp declines in densities of targeted species seen inside the MPAs themselves (see Fig.S27-S29), suggesting a driver other than fishing may be at play.
-
-What do our results imply about the future of MPA science?
-Our simulation model is by no means exhaustive (for example it ignores features such as species interactions, habitat effects, and climate feedback), but captures many of the core factors theorized to affect MPA performance.
-Our results show that while attributes such as MPA size and fishing pressure are important factors in determining the effects of protection, local fleet dynamics and the movement rates of adult fish can dramatically affect the outcomes of protected areas as well  (Fig.\ref{conservation-effects}).
-Far from being a simple tool with clear outcomes, the effects of MPAs can be highly context dependent, requiring - we would argue - a bio-economic model of at minimum the complexity presented here to help communities design and set expectations of MPAs at the tactical level.
-While users may not be able to parameterize every aspect of a model such as that presented here, working with stakeholders to visualize the implications of, for example, different fleet responses to MPA implementation is a critical step in MPA design.
-Readers can use our simulation model to explore the effects of MPAs through an interactive web application available [here](https://danovando.shinyapps.io/simmpa/).
-
-Once simulation models have been used to help design an MPA, how can users evaluate whether it is achieving their objectives?
-Response ratios are commonly used as evidence for conservation outcomes of MPAs [@lester2009], but as suggested in @ferraro2018 and @larsen2019 and demonstrated here, without careful attention to the design of control sites (accounting for example for the displacement of effort by MPAs), response ratios may be highly unreliable estimators of regional MPA effects.
-As @larsen2019 suggests, there are many potential alternatives for estimating the effects of MPAs that better account for the challenges of causal inference.
-We applied one such approach here (a difference-in-difference estimator), and yet were still unable to reach robust conclusions as to the effect of MPAs on the density of targeted species in the Channel Islands, due to the likely small size of the true effect relative to the strength and variability of environmental drivers.
-
-While this does not mean that all MPAs will face similar challenges in estimating their effects, our results in the rigorously studied Channel Islands system make clear that in many instances empirically detecting a clear regional effect of MPAs may not be possible.
-How then should stakeholders go about adaptively monitoring and managing MPAs?
-Simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can perhaps be tuned to focus on the species groups that have the highest chance of a detectable effect size [@nickols2019].
-Expanding data collection to include robust monitoring of fleet dynamics may help statistical approaches to isolate the true effect of MPAs on conservation outcomes by allowing for improved control of the effects of fishing dynamics, and allow managers to take into account potential negative interactions between MPAs and fleet dynamics such as those that may occur under constant-catch dynamics.
-Whenever possible monitoring programs should be implemented prior to MPA implementation to provide a pre-treatment benchmark.
-Non-equilibrium analyses (Fig.\ref{did-plot}) also help set expectations for effect sizes over time.
-Beyond that, educating communities about the challenges of estimating the effects of MPAs can help set expectations, so that a lack of a clear effect is not necessarily viewed as a failure of the program, but rather considered in the context of reasonable expectations.
-While this paper has focused on the conservation outcomes of MPAs, future work must also address the challenge of predicting and estimating the fishery impacts of protected areas.
-
-As the number and size of global MPA networks increase, it is critical that we both set appropriate expectations for their outcomes, and plan how we will monitor the performance of these protected areas over time.
-While the history of MPA science has made important strides in helping us understand the dynamics of protected areas, the future of MPA science must directly tackle the challenge of evaluating the performance of these MPAs, and adapting their design as needed to best achieve objectives.
-Commonly employed metrics such as response ratios may be applicable in some circumstances, but can have severe shortcomings as metrics of regional conservation effects.
-Dependence on unreliable estimators of MPA effects may lead to stakeholders incorrectly attributing negative environmental shocks as MPA failures, or interpreting data arising from scorched-earth fishing outside MPAs as a conservation success.
-Bio-economic modeling can help frame community expectations, reducing the potential for a reduction in support if unrealistic conservation or fishery gains are not realized.
-Statistical approaches that explicitly address complications such as the spatial spillover effects of MPAs (such as the difference-in-difference approach used here) may give users an improved understanding of the performance of their MPAs.
-Effective use of simulation and statistics to clearly communicate what we should expect and what we can detect from MPAs is critical in ensuring that MPAs play effective roles in fisheries management and marine conservation.
-
-## Materials and Methods
-
-We present here critical characteristics of our simulation model and regression approach.
-Further details can be found in the Supplementary Information.
-All analysis were conducted in R [@rcoreteam2019].
-Our main difference-in-difference model was fit using Template Model Builder in R [@kristensen2016].
-Code needed to replicate results can be found [here](https://github.com/DanOvando/regional-effects-of-mpas).
-
-### Simulation Model
-
-Our bio-economic model simulates the effect of MPAs on a spatially explicit age-structured representation of a single species.
-Readers can explore the functionality of the model using an online tool available [here](https://danovando.shinyapps.io/simmpa/).
-See Table.S1 for a complete description of simulation states.
-The model consists of 25 patches with wrapped edges (picture the waters around a circular island).
-For any one simulation we randomly pull a species and its associated life history from the `FishLife` [@thorson2017c] package in R.
-We pair these data with randomly selected values between 0.6 and 0.95 for Beverton-Holt steepness [as parameterized in @mace1994], as well as larval and adult dispersal rates.
-We also randomly assign whether adults have density dependent movement, as well as one of three potential types of density dependence [building off of @babcock2011]:
-
-  1. Local density dependence: Density dependence occurs independently in each patch, and recruits then disperse to nearby patches
-
-  2. Global density dependence: Density dependence is a function of the sum of spawning biomass across all patches, and recruits are then distributed according to habitat quality
-
-  3. Post-dispersal density dependence: Larvae are distributed throughout the system, and then density dependence occurs based on the density of adult biomass at the destination patch
-
-We allow for three potential siting strategies for MPAs.
-In the first, MPAs are randomly placed.
-In the second, we assume that MPAs are placed in preferentially better habit (unfished recruitment is four times greater inside MPA locations).
-In the third, we allow for scenarios in which MPAs are placed in hotspots of larval dispersal.
-In this scenario, patches in which MPAs will be placed have larval dispersal rates four times greater than patches that do not become MPAs.
-
-Along with the fleet dynamics model, each simulation is assigned a random fleet dispersal scenario: uniform dispersal (where the total effort of the fleet is divided evenly among all open patches), catch dispersal (where the total effort of the fleet is divided according to the catchable biomass in each available patch), and profit dispersal (where the total effort of the fleet is divided according to the profit per unit effort in each available patch).
-
-Lastly each simulation is assigned an MPA scenario, defined by the number and size of MPAs, the placement of those MPAs, and the year that the MPAs are put in place.
-Each simulation starts the population off at unfished equilibrium and then beings to apply the fleet model.
-The MPAs are then placed during the randomly selected start year, allowing some runs to explore how the early dynamics of the MPA play out when the fishery and population they are placed on is not already at equilibrium.
-Fishing effort in displaced by MPAs can either concentrate outside or leave the fishery.
-Each simulation is run to equilibrium with and without the selected MPA strategy (holding all else constant).
-We then measure the difference in biomass and fishery catches in each time step in the scenario with and without the MPAs to calculate the conservation and fishery effects of the MPAs over time.
-
-### Difference in Difference Regression
+  \includegraphics[width=1\linewidth]{figs/pop-trends.pdf}
+  \caption{Centered and scaled mean annual biomass densities of included targeted (red) and non-targeted (blue) fish species (faded lines) and smoothed means with 95\% confidence intervals over time (darker lines and ribbons).
+Vertical red dashed line shows year of MPA implementation.}
+  \label{pop-trends}
+\end{figure}
 
@@
 <!-- \showmatmethods -->
-\showacknow
+<!-- \showacknow -->
 

### 2020-03-24 functions/create-samples.R

@@ -23,7 +23,10 @@ create_samples <- function(fishes,
       ),
       create_fish,
       time_step = time_step,
-      sigma_r = sigma_r
+      sigma_r = sigma_r,
+      larval_movement = round(num_patches  *.25),
+      adult_movement = round(num_patches * .05),
+      steepness = 0.8
     )) %>%
     mutate(constant_f = map_dbl(fish, ~ .x$m *  f_v_m, f_v_m = f_v_m)) %>%
     mutate(constant_effort =  ((constant_f / fleet_q) / time_step) * num_patches * targeted) %>%
@@ -34,7 +37,7 @@ create_samples <- function(fishes,
         initial_effort = .x,
         fish = .y,
         q = fleet_q,
-        length_50_sel = .y$length_50_mature
+        length_50_sel = .y$length_50_mature * .9
       ),
       fleet_q = fleet_q
     ))
@@ -58,7 +61,6 @@ create_samples <- function(fishes,
       )
     )
   
-
   simple_fish <- simple_fish %>%
     mutate(net_outcomes = map(mpa_experiment, 'outcomes')) %>%
     mutate(raw_outcomes = map(mpa_experiment, 'raw_outcomes')) %>%

### 2020-03-24 functions/estimate_did.R

@@ -127,7 +127,8 @@ estimate_did <-
     }
     did_data <- purrrlyr::dmap_at(did_data, vars, ~ scale(.x)) %>%
      purrrlyr::dmap_at(vars, nafoo) %>%
-      filter(total_biomass_density > 0) %>% 
+      mutate(total_biomass_density = total_biomass_density + 1e-6,
+             mean_biomass_density = mean_biomass_density + 1e-6) %>% 
       group_by(site_side, targeted) %>%
       mutate(scaled_total_biomass_density = scale(log(total_biomass_density)))
     # filter(region != "SMI")
@@ -479,12 +480,12 @@ estimate_did <-
       ungroup() 
     
     # 
-    mpa_effect_plot <-  did_results %>%
-      ggplot(aes(year, did)) +
-      geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
-      tidybayes::stat_halfeye(alpha = 0.7,
-                              .width = c(0.5, 0.95)) +
-      scale_y_continuous(labels = percent, name = "Estimated MPA Effect") +
+    # mpa_effect_plot <-  did_results %>%
+    #   ggplot(aes(year, did)) +
+    #   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
+    #   tidybayes::stat_halfeye(alpha = 0.7,
+    #                           .width = c(0.5, 0.95)) +
+    #   scale_y_continuous(labels = percent, name = "Estimated MPA Effect") +
       scale_x_discrete(name = "Year Bin")
     # 
     out <- list(did_results = did_results,

### 2020-03-24 functions/test-performance.R

@@ -4,14 +4,16 @@ test_performance <-
            min_year = 75,
            max_year = 100,
            time_step = 1,
-           sigma_obs = 0) {
+           sigma_obs = 0,
+           samps_per_patch = 1) {
     
     
     simple_data <- fishes %>%
       select(loo, k, lm, m, targeted, classcode, commonname,raw_outcomes) %>%
       unnest(cols = raw_outcomes) %>%
       group_by(year, patch, experiment, targeted) %>% 
-      summarise(biomass_density = rlnorm(1,log(sum(biomass) + 1e-3), sigma_obs)) %>% 
+      summarise(biomass_density = list(rlnorm(samps_per_patch,log(sum(biomass) / 100 + 1e-3), sigma_obs))) %>% 
+      unnest(cols = biomass_density) %>% 
       ungroup() %>% 
       filter(year > min_year, year < max_year,
              experiment == "with-mpa") %>%
@@ -24,7 +26,7 @@ test_performance <-
         logical_targeted = targeted > 0,
         post_mpa = year >= year_mpa
       )
-
+    
     years_protected <- unique(simple_data$year) - year_mpa
 
     bins <-
@@ -57,58 +59,66 @@ test_performance <-
       mutate(post_mpa = year > year_mpa) %>%
       mutate(year = year * time_step) %>%
       mutate(subyear = year - floor(year),
-             year = floor(year))
+             year = floor(year)) %>% 
+      filter(targeted == 1)
 # 
-    did_model <- stan_glmer(biomass_density ~ targeted * protected_block + (1|patch),
-                            data = simple_data,
-                            chains = 4,
-                            cores = 4,
-                            family = Gamma(link = "log"))
+      env <- new.env(parent = parent.frame())
+      
+      env$simple_data <- simple_data
     
-    # did_model <- stan_glm(log(biomass_density) ~ targeted * protected_block,
+        # did_model <- with(env, {stan_glmer(biomass_density ~ targeted * factor_year + (1|patch),
+        #                     data = simple_data,
+        #                     iter = 5000,
+        #                     chains = 4,
+        #                     cores = 4,
+        #                     prior = normal(0,2.5, autoscale = TRUE),
+        #                     family = Gamma(link = "log"))}
+        # )
+
+        
+        did_model <- with(env, {stan_glm(biomass_density ~ targeted * factor_year,
+                                           data = simple_data,
+                                           iter = 3000,
+                                           chains = 4,
[…]

### 2020-03-24 make-regional-mpas.R

@@ -57,7 +57,7 @@ run_tmb <- FALSE
 
 simulate_mpas <- FALSE # simulate MPA outcomes
 
-simulate_channel_islands <- TRUE # simulate MPA outcomes
+simulate_channel_islands <- FALSE # simulate MPA outcomes
 
 validate_mpas <- FALSE
 
@@ -1797,13 +1797,13 @@ if (simulate_channel_islands == TRUE){
   
   num_patches <- 50
   
-  run_experiments <- TRUE
+  run_experiments <- FALSE
   
-  create_grid <- TRUE
+  create_grid <- FALSE
   
   save_experiment <- TRUE
   
-  samps <- 20
+  samps <- 2000
   
 
   
@@ -2101,7 +2101,6 @@ if (simulate_channel_islands == TRUE){
     sim_grid <- read_rds( file.path(run_dir, "ci_sim_grid.rds"))
     
   } # close run experiments
-  
   message("finished channel islands mpa experiments")
   # process outcomes --------------------------------------------------------
 
@@ -2161,21 +2160,6 @@ if (simulate_channel_islands == TRUE){
 # validate estimation strategy --------------------------------------------
 if (validate_mpas == TRUE){
 
-  simulate_samples <- TRUE
-
-  burn_years <- 1
-
-  sim_years <- 99
-
-  year_mpa <- 50
-
-  num_patches <-  2
-
-  mpa_size <- 0.5
-
-  n_samples <- 5
-
-  time_step <-  1
 
   # prepare data ------------------------------------------------------------
 
@@ -2185,77 +2169,28 @@ if (validate_mpas == TRUE){
 
   enso <- read_csv(here::here('data','enso.csv'))
 
-  pisco <- abundance_data$data[[1]]
+  validate_foo <- function(sigma_r,
+                           sigma_obs,
+                           env_strength,
+                           f_v_m,
+                           mpa_size,
+                           num_patches = 50,
+                           year_mpa = 50,
+                           sim_years = 69,
+                           burn_years = 1,
+                           time_step = 1) {
+    
   
-  enviro <- rep(c(rep(0,5), rep(1,5)), (sim_years + burn_years) / 10)
+  pisco <- abundance_data$data[[1]]
+    
+  enviro <- rep(c(rep(0,5), rep(env_strength,5)), (sim_years + burn_years) / 10)
 
[…]

### 2020-03-26 documents/ovando-regional-effects-of-mpas-v2.Rmd
passing to squid since sf on mac is broken beyond belief

@@
 output:
+  rticles::pnas_article: default
+  bookdown::word_document2: default
   bookdown::pdf_document2: default
-  bookdown::word_document2: default
-  rticles::pnas_article: default
   bookdown::html_document2: default
@@
 
-As stakeholders around the world increasingly seek to use MPAs in marine resource management portfolios, and the performance of existing MPAs is evaluated, it is critical that we develop a better understanding of the magnitude and drivers of regional-scale MPA effects.
-The Channel Islands National Marine Sanctuary, California, USA provides an ideal case study to address this need.
-A network of protected areas covering approximately 20% of the Islands' waters qas put in place in 2003 (see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of these MPAs).
-We use data from the first 14 years of protection to provide what is to our knowledge the first large-scale empirical assessment of the regional effect of a large MPA network on a wide array of fin-fish species.
-In contrast to clear differences in biomass densities observed inside and outside of well protected MPAs [@lester2009], we are unable to detect a clear regional effect from the Channel Islands MPAs.
+As stakeholders around the world increasingly seek to use MPAs in marine resource management portfolios, it is critical that we develop a better understanding of the magnitude and drivers of regional-scale MPA effects.
+This is particuarly true in the context of a changing climate, in which "set it and forget it" management strategies are increasinly untenable; we need to be able to monitor the performance of MPAs in order to adapt them as dictated by a shifting environment.
+The MPAs within the Northern Channel Islands National Marine Sanctuary, California, USA (which we will refer to as the Channel Islands from now on) provides an ideal case study to address this need.
+A network of protected areas covering approximately 20% of the Islands' waters was put in place in 2003 [see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of these MPAs].
+This MPA network has subsequently been used as a model case study in protected area design around the world XX.
+
+We use data from the first 14 years of protection to provide what is to our knowledge the first empirical assessment of the regional effect of a large MPA network on a wide array of finfish species.
+In contrast to clear differences in biomass densities observed inside and outside of well protected MPAs both globally [@lester2009] and in the Channel Islands [@caselle2014] we are unable to detect a clear regional effect from the Channel Islands MPAs.
 We build off of existing MPA theory to consider why this might be, and provide guidance for scientists and managers as to when and how we might expect to estimate the regional conservation effects of MPAs.
@@
 
-The empirical MPA literature focuses on assessing the conservation effects within the borders of protected areas [e.g.
+The empirical MPA literature has generally focused on assessing conservation effects within the borders of protected areas [e.g.
 @edgar2014].
@@
 The organisms within the borders of protected areas are generally part of a broader biological stock, connected through adult movement and larval dispersal.
-If the goal of conservationists or natural resource managers is to increase the total abundance or productivity of a resource, a broader question we should ask of MPAs is not just are there more fish inside their borders, but also how have the reserves affected abundances throughout the region in which they are located?
+If the goal of conservationists or natural resource managers is to increase the total abundance or productivity of a resource, a broader question we should ask of MPAs is not just are there more fish inside their borders, but also how have reserves affected populations throughout the region in which they are located?
 This logic, that MPAs will have conservation benefits for most species beyond their borders, is implicit in all multilateral calls for MPA expansion.
 
-We define the regional conservation effects of MPAs as the change in total biomass of fish (summing inside and outside of MPAs) relative to the total biomass of fish that would have occurred without the MPAs (acknowledging that other outcomes such as increased biodiversity or resiliency are also important to conservation but are beyond the scope of this analysis).
-
-<!-- As a thought experiment, imagine a region that has driven its fish populations to near extinction and then places 100\% of its waters inside a no-take MPA.
-In that setting, we would expect the regional-scale conservation benefits to be massive.
-On the other hand, imagine implementing an MPA in an area where a sedentary species has been only lightly fished.
-In that setting, it is possible to create a small net conservation loss if the concentration of fishing pressure outside the reserve has a greater effect than the biomass buildup inside the reserve.
-Within these broad bounds,  -->
-
+We define the regional conservation effects of MPAs as the change in total biomass densities of fish (averaging inside and outside of MPAs) relative to the total biomass densities of fish that would have occurred without the MPAs (acknowledging that other outcomes such as increased biodiversity or resiliency are also important to conservation but are beyond the scope of this analysis).
 Numerous factors can affect the regional effects of MPAs.
-These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], larval dispersal patternss and the strength and timing of density dependence in the population (e.g.
-pre- or post-settlement, @burgess), how overfished the population would be without the MPA, and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003;@smith2006;@walters2000].
+These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], larval dispersal patterns and the strength and timing of density dependence in the population (e.g.
+pre- or post-settlement, @burgess2014), how overfished the population would be without the MPA, and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003;@smith2006;@walters2000;@white2011].
 In addition, even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through habitat and network effects [@gaines2010; @costello2010].
-
-```{r}
-short_term <- outcomes %>%
-  group_by(experiment) %>% 
-  filter(years_protected == max(years_protected), years_protected >= 0) %>% 
-  ungroup() %>% 
-  filter(!is.na(mpa_effect))
-
-median_effect <- median(short_term$mpa_effect[short_term$year == max(short_term$year)], na.rm = TRUE)
-
-median_pop_effect <- median(short_term$pop_effect[short_term$year == max(short_term$year)], na.rm = TRUE)
-
-# median_fish_effect <- median(fishery_short_term$mpa_effect[short_term$year == max(short_term$year)], na.rm = TRUE)
-# 
-# median_fish_effect <- median(fishery_short_term$fishery_effect[short_term$year == max(short_term$year)], na.rm = TRUE)
-
-pos_runs <- round(mean(short_term$pop_effect[short_term$year == max(short_term$year)] > 0),2)
-
-# pos_fish_runs <- round(mean(fishery_short_term$msy_effect[fishery_short_term$year == max(fishery_short_term$year)] > 0, na.rm = TRUE),2)
-
-pos_fish_effect <- round(mean(short_term$f[short_term$year == max(short_term$year)] > 0),2)
-
-small_effect <- round( 100* median(short_term$mpa_effect[short_term$depletion <= 0.4]))
-
-min_small_effect <- round( 100* min(short_term$mpa_effect[ short_term$depletion <= 0.4]))
-
-max_small_effect <- round( 100* max(short_term$mpa_effect[ short_term$depletion <= 0.4]))
-
-medium_effect <- round(100*median(short_term$mpa_effect[short_term$depletion > 0.4 & short_term$depletion <= 0.6]))
-
-min_medium_effect <- round(100*min(short_term$mpa_effect[short_term$depletion > 0.4 & short_term$depletion <= 0.6]))
-
-max_medium_effect <- round(100*max(short_term$mpa_effect[short_term$depletion > 0.4 & short_term$depletion <= 0.6]))
-
-big_effect <- round(100*median(short_term$mpa_effect[short_term$depletion > 0.6]))
-
-min_big_effect <- round(100*min(short_term$mpa_effect[short_term$depletion > 0.6]))
-
-max_big_effect <- round(100*max(short_term$mpa_effect[short_term$depletion > 0.6]))
-
-```
 
@@
 While we know many parameters will affect the expected effects of the MPAs, theory suggests that life history, adult and larval dispersal relative to MPA size, and fishing pressure will be key drivers.
-The targeted species in our database span a range of life histories, but are largely made up of fishes in the genus XX (e.g.) and **Sebastes** (e.g.
-), with a median Von Bertalanffy growth coefficient of XX, and a median age at maturity of XX [values drawn from @thorson2017c].
-The MPAs in this study cover roughly 21% of the the waters of the Northern Channel Islands, a region spanning roughly 90km east to west.
-While detailed dispersal studies are not available for all of the species covered by our study, what information we have suggest that while adults of some of these species are likely to exhibit some site fidelity [@love2011], larvae are likely distributed beyond the northern Channel Islands [@rassweiler2012].
-Therefore, if we assume that the population of the targeted species is at least equal to the extent of the Northern Channel Islands, we can assume that at most the MPAs cover roughly 20% of the targeted populations' ranges.
+The targeted species in our database span a range of life histories, but are largely made up of fishes in the perch and rockfish complexes, with a mean Von Bertalanffy growth coefficient of `r round(mean(used_data$k),2)`, and a mean age at maturity of `r round(mean(used_data$tm),0)` years [values drawn from @thorson2017c].
+
+The MPAs in this study cover 21% of the the surface waters of the Northern Channel Islands, a region spanning roughly 90km east to west.
+While detailed dispersal studies are not available for all of the species covered by our study, what information we have suggest that while adults of some of these species are likely to exhibit site fidelity [@love2011], larvae are likely distributed beyond the Channel Islands [@rassweiler2012].
+Therefore, if we assume that the population of the targeted species is at least equal to the extent of the Channel Islands, we can assume that at most the MPAs cover roughly 20% of the targeted populations' ranges.
+
 Turning to the critical question of fishing pressure, formal stock assessments are largely lacking for these species.
-However, what evidence we have suggests mostly moderate fishing pressure, with some species such as California sheephead (XX) and blue rockfish (xx) experiencing high levels of fishing mortality / biomass  levels beloe target levels during the early 2000s pre-MPA [@alonzo2004;@dick2017].
+However, what evidence we have suggests mostly moderate fishing pressure, with some species such as California sheephead (*Semicossyphus pulcher*) and blue rockfish (*Sebastes mystinus*) experiencing high levels of fishing mortality / biomass  levels below target levels during the early 2000s [@alonzo2004;@dick2017].
 
-Considering the combinatino of slow growth, 20% or lower MPA coverage, and moderate fishing pressure, we might expect that the northern Channel Islands MPAs would produce a small to moderate effect size over time.
-However, theory also tells us that many other variables can interact to affect the outcomes of MPAs.
+We integrated these network design features, life history traits, and exploitation histories into a spatially explicit bio-economic simulation model, and used this model to generate expected outcomes for the Channel Islands MPAs.
+Our results suggest that while a wide range of outcomes are plausible, from 0% to upwards of 200%, across the majority of our simulations the effect size was 25% or less, with a median simulated effect size of roughly 20%; a potentially small value given the natural variability and sampling challenges of marine environments.
+
+<!-- However, theory also tells us that many other variables can interact to affect the outcomes of MPAs.
 To address this, we utilized a spatial bio-econonomic model to simulate the expected effects of an MPA network stylized to resemble the Channel Islands on the group of targeted species covered in our analysis (See SI for details of simulation framework).
+-->
 
-**reference HARKing, of course DiD and theory matching together doens't prove anything, but it's an important step to take**
+<!-- **reference HARKing, of course DiD and theory matching together doens't prove anything, but it's an important step to take** -->
 
@@
 
-Having established a theoretical grounding for the likely effect of the Channel Island MPAs, what empirical evidence do we have for the effects of this protected area network?
-@caslle2015 examined changes in biomass densities of species targeted by fishing activity inside and outside of the MPAs over time (a measurement generally termed "response ratios"), and compared these changes to the trends in non-targeted species.
-They found a statistically significant increase in the response ratios of targeted speices over time, and evidence that this increase is smaller in the non-targeted species.
-Updating the results of @caselle2015 through 2017 shows the same increasing trend in the resonse ratios of targeted species (Fig.\@ref(fig:XX)).
+Having established a theoretical grounding for the likely biomass density outcomes of the Channel Island MPAs, what empirical evidence do we have of the effects?
+@caslle2015 examined changes in biomass densities of species targeted by fishing activity inside and outside of the MPAs over time (a metric generally termed a "response ratio"), and compared these changes to the trends in non-targeted species.
+They found a statistically significant increase in the response ratios of targeted species over time, and evidence that this increase is smaller in the non-targeted species.
+Updating the results of @caselle2015 through 2017 with a Bayesian response-ratio regression shows a continuation of the increasing trend in the response ratios of targeted species (Fig.\@ref(fig:XX)).
 
@@
 While these concerns have been stated previously [e.g.
-@halpern2004], the MPA evaluation literature has by and large been unable to adequately address them, often due to the very real challenges of identifying and sampling adequate control sites that are truly independent of the MPAs in question [@ferraro2018].
+@halpern2004;@white2011], the MPA evaluation literature has by and large been unable to adequately address them, often due to the very real logistical challenges of identifying and sampling adequate control sites that are truly independent of the MPAs in question.
+As a result of these challenges, response ratios can be a highly imprecise and biased measure of the regional conservation effect of an MPA network [@ferraro2018].
 
-In the case of the Channel Islands MPAs, control sites are often located within only a few kilometers of an MPA, suggesting that they are susceptible to biological spillover and concentration of fishing effort resulting from the MPAs.
-Given these complications, we can certainly interpret the response ratios reported in @caselle2015 and updated here as evidence that the MPAs are indeed providing effective protection of targeted fish biomass inside their borders, particularly in light of the general increase in respones ratios over time.
+In the case of the Channel Islands MPAs, control sites are often located within a few kilometers of an MPA, suggesting that they are susceptible to biological spillover and concentration of fishing effort resulting from the MPAs.
+Given these complications, we can certainly interpret the response ratios reported in @caselle2015, and updated here, as evidence that the MPAs are providing effective protection of targeted fish biomass inside their borders, particularly in light of the general increase in response ratios over time.
 However, it is unclear how well these response ratio results serve as in indicator of overall regional effects of the MPAs.
 
-To illustrate this problem, we used our bio-economic model to simulate response ratio trajectories for species and MPA coverage representative of the northern Channel Islands.
-These simulations cover the life-histories and MPA sizes seen in the Channel Islands, but vary in key unknowns such as the actual degree of fishing pressure, the timing of density dependence, and the fleet responses to the MPAs.
-For each year of protection, we paired our simlated response ratios to the estimated posterior probablity distributinos of the Channel Island response ratios.
-For example, if the mean response ratio in the year 2006 is 1, we found all simulations with three years of protection that had simulated response ratios near 1, and then pulled out the "true" regional MPA effect from each of those simulations.
+To illustrate this problem, we used our bio-economic model to simulate response ratio trajectories for species and MPA coverage representative of the Channel Islands (see methods and SI for details of model structure).
+These simulations cover the life-histories and MPA sizes seen in the Channel Islands, but vary in key unknowns such as the actual degree of fishing pressure, the timing of density dependence, and the fishing fleets response to the MPAs.
+For each year of protection, we paired our simulated response ratios to the estimated posterior probability distributions of the Channel Island response ratios.
+For example, if the mean estimated response ratio in the year 2006 is one, we found all simulations with three years of protection that had simulated response ratios near one, and then pulled out the "true" simulated regional MPA effect from each of those simulations.
 This provides us with a distribution of simulated regional MPA effects that could plausibly generate the types of response ratios actually observed in the Channel Islands.
 
-As reported in @caselle2015, we see evidence for an increasing trend in respones ratios of targeted species over time (Fig.\@ref(fig:rr-plot)).
-This provides strong evidence that the MPAs are providing protected for targeted species within their borders.
-However, our simulation results show thatresponse ratios trends we observe in the data could plausibly be produced by a wide range of regional MPA effects (Fig.\@ref(fig:rr-plot)).
-Response ratios well over 1 were associated with regional MPA effects generally elss than 25%, and many simulations produced large response ratios but regional MPA effects close to zero.
-This can occur if for example fishing pressure is only moderate, adult movement is low, larval dispersal is high, and displaced fishing effort concentrates around the border of the MPAs.
-
+As reported in @caselle2015, we see evidence for an increasing trend in response ratios of targeted species over time (Fig.\@ref(fig:rr-plot)).
+This provides strong evidence that the MPAs are providing protection for targeted species within their borders.
+However, our simulation results show that the response ratio trends we observe in the data could plausibly be produced by a wide range of regional MPA effects (Fig.\@ref(fig:rr-plot)).
+Response ratios well over one were associated with regional MPA effects generally less than 25%, and many simulations produced large response ratios but regional MPA effects close to 0% This can occur if for example fishing pressure is only moderate, adult movement is low, larval dispersal is high, and displaced fishing effort concentrates around the border of the MPAs.
 ```{r rr-plot, include = FALSE}
@@
 
-While we see empirical evidence of increasing response ratios of targeted species in the Channel Islands, theory indicates that this may not be a reliable indicator of the regional effect of the MPAs.
-An ideal "control" for estimating the regional effects of MPAs would a) control for regional environmental shocks unrelated to the MPAs b) be unaffected by treatment bias (e.g.
-preferential placement of MPAs in ideal habitat) c) be unaffected by the MPAs themselves.
-While spatial response-ratios can help control for environmental shocks and, in theory, treatment bias, at the scale of the Channel Islands it is unlikely that we can strongly say that outside-MPA sites are themselves unaffected by the MPAs we are trying to evaluate.
+Given the potential unreliability of response ratios for the task, in an ideal world how would we empirically measure the regional effects of MPAs?
+The perfect experiment would involve two parallel worlds that were identical, except for the implementation of an MPA.
+In world “A”, no MPA would be implemented, and in the facsimile world “B”, the MPA would be implemented.
+Both worlds would be tracked before and after MPA implementation, and the biomass densities would be compared after treatment.
+Instead of two parallel worlds, a similar experimental design would involve random placement of MPAs.
+Unfortunately, neither of these experimental designs has been implemented, because parallel worlds do not exist and MPAs are to our knowledge never placed at random.
+Careful causal inference requires some other means of controlling for biases introduced by factors such as the unobserved environmental shocks, the MPA siting process, biological spillover, and concentration of fishing effort  [@ferraro2018;@larsen2019].
 
-Building off of @caselle2015, we propose an alternative identification strategy utilizing biomass densities of species that are not directly targeted by fishing as our control group (non-targeted), and biomass densities of species targeted by fishing as our treatment group.
-Targeted species in the Northern Channel Islands, California, include commercially important finfish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
+Building off of the concepts explored in @caselle2015, we used an identification strategy utilizing biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 0])` species that are not directly targeted by fishing as our control group (non-targeted), and biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 1])` species targeted by fishing as our treatment group.
+Targeted species in the Channel Islands, include commercially important fin-fish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
 Each of these targeted species was the subject of prior bio-economic modeling related to the effects of MPAs in southern California [@costello2010d;@rassweiler2014].
-But, the analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*).
-Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species).
-We use a Bayesian difference-in-difference estimator to estimate any difference in mean total biomass densities of finfish species targeted by fishing effort (i.e., those potentially affected by an MPA) and those species not targeted by fishing before and after MPA implementation [see @larsen2019 for a succinct explanation of difference-in-difference estimators in ecological settings].
-The result of this regression is an estimate of the effect of the MPAs on the mean total biomass densities of targeted species throughout the Channel Islands.
+It should be noted though that our analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*).
+Included non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species).
+We used a Bayesian difference-in-difference regression to estimate any difference in mean total biomass densities of fin-fish species targeted by fishing effort (i.e., those potentially affected by an MPA) and those species not targeted by fishing before and after MPA implementation [see @larsen2019 for a succinct explanation of difference-in-difference estimators in ecological settings].
+The result of this regression is, conditional on the assumptions of the model, an estimate of the effect of the MPAs on the mean total biomass densities of targeted species throughout the Channel Islands.
 
-As conventional wisdom and theory would suggest, over the first three years of implementation (2003-2006), the effects of the MPAs is unclear, with support for a small negative effect to a substantially positive effect (with much higher probability of a small positive effect).
-Over the next six years we see signs of an increasingly positive MPA effect, peaking in 2009-2011 with a median estimate of MPA effect of 75% on mean total bioamss of targeted species.
-These empirical estimates are in line with the range of outcomes that our simulation model suggests are plausible given the rough characteristics of the northern Channel Island MPAs.
-However, in the subsequent years the trend reverses itself, and for the years 2015-2017 we once again see no clear effect of the MPAs Fig.\@ref(fig:mpa-effect)
+```{r}
+did_model <- model_runs %>%
+  filter(data_source == "pisco",
+         var_names == "pisco_a",
+         data_to_use == "all") %>% {
+           .$did_fit[[1]]$did_reg
+         }
+did_coefs <- did_model %>% broom::tidy(intervals = TRUE)
+
+```
+
+As conventional wisdom and theory would suggest, over the first three years of implementation (2003-2006), the effects of the MPAs are unclear, with support for a small negative effect to a substantially positive effect, with much higher probability of a small positive effect (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`,
+90\% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`
+- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`).
+Over the next six years the model estimates greater probabilities of an increasingly positive MPA effect, peaking in 2009-2011 with a median estimate of MPA effect of a `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)` increase in mean total biomass density of targeted species (90% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)`
+- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)`.
+These empirical estimates are in line with (though in the upper range of) the outcomes that our simulation model suggests are plausible given the rough characteristics of the northern Channel Island MPAs.
+However, in the subsequent years the trend reverses itself, and for the years 2015-2017 we once again see no clear effect of the MPAs (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`,
+90\% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`
+- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`.) Fig.\@ref(fig:mpa-effect)
 
@@
   \caption{Results of difference-in-difference regression estimating the regional effect of the northern Channel Island MPAs on mean total biomass densities of targeted species (difference in mean total biomass density of targeted species over time relative to expected levels using non-targeted species as a control).
-Grey distibutions show 90%XX posterior probability distribution of estimated MPA effect; red point is median estimated effect, thicker red section 50% credible interval, thinner red line 90% credible interval).
-Blue distributions in background show range of MPA effects produced by simulation model tuned to reflect the dybnamics of the Northern Channel Island MPAs (black dashed line is median simulated value).
-Results are estimated in three blocks, including years greater than or equal to left-hand value and less than right-hand value.}
+Grey distributions show posterior probability distribution of estimated MPA effect; red point is median estimated effect, thicker red section 50% credible interval, thinner red line 90% credible interval).
+Blue distributions in background show range of MPA effects produced by simulation model tuned to reflect the dynamics of the Northern Channel Island MPAs (black dashed line is median simulated value).
+Results are estimated in blocks of three years, including years greater than or equal to left-hand value and less than right-hand value.}
   \label{mpa-effect}
@@
 
-1. Why are the results from the channel islands imporant and what do they tell us
+<!-- 1. Why are the results from the channel islands important and what do they tell us -->
 
-2. Put the channel islands results in context -
+<!-- 2. Put the channel islands results in context -  -->
 
-Ideally then we want a control for broad environmental shocks to the region that is independent of factors such as selection bias, biological spillover, and fishing concentration.
+<!-- Ideally then we want a control for broad environmental shocks to the region that is independent of factors such as selection bias, biological spillover, and fishing concentration.  -->
 
-1. multiple lines of evidence
+<!-- 1. multiple lines of evidence -->
 
-2. Theory before empirics
+<!-- 2. Theory before empirics -->
 
-3. Patience
+<!-- 3. Patience -->
 
-The network of MPAs implemented in the Northern Channel Islands in 2003 provides an ideal opportunity to examine the regional effects of proteced area networks.
-This analysis is to our knowledge the first empirical estimate of the regional effects of such a large MPA network on a broad assemblage of targeted finfish.
+The network of MPAs implemented in the Northern Channel Islands in 2003 provides an ideal opportunity to examine the regional effects of protected area networks.
+This analysis is to our knowledge the first empirical estimate of the regional effects of a large and iconic MPA network on a broad assemblage of targeted fin-fish.
 Despite the well enforced nature of these MPAs [xx], and the presence of a rigorous scientific monitoring program, we are unable to estimate a clear effect of protection.
-While response ratios have targeted species have increased over time, providing evidence of effective *within MPA* protection, these response ratios are on their own an unrealible indicator of the regional conservation effects of the MPAs.
-Our difference-in-difference estimates a high probability of increasinly positive MPA effects over the first nine years of protection, only to see these estimated gains reversed from 2012 to 2017.
-For the most recent time period, we can detect no clear positive or negative effect of the Channel Island MPAs Fig.\@ref(fig:mpa-effect)
+While response ratios of targeted species have increased over time, providing evidence of effective *within MPA* protection, these response ratios are on their own an unreliable indicator of the regional conservation effects of the MPAs.
+Our difference-in-difference regression estimates a high probability of increasingly positive MPA effects over the first nine years of protection, only to see these estimated gains reversed from 2012 to 2017.
+For the most recent time period, we can detect no clear positive or negative regional effect of the Channel Island MPAs on biomass densities of targeted fin-fish (Fig.\@ref(fig:mpa-effect)).
 
@@
 One explanation may lie in fleet dynamics.
-Much of the theoretical literautre on MPAs assumes that all else being equal bigger reserves produce bigger conservation gains [@white2011].
+Much of the theoretical literature on MPAs assumes that all else being equal bigger reserves produce bigger conservation gains [@white2011 and references therin].
 However, to our knowledge all of these models simulate fleet dynamics through assumptions about fishing mortality rates (e.g.
@@
 
-An alternative and to our knowledge unexplored (in the context of MPAs) fleet model though is a "constant-catch" strategy.
+```{r}
+positive_outcomes <- outcomes %>% 
+  filter(years_protected == 15) %>% 
+  group_by(fleet_model) %>% 
+  summarise(pp = mean(mpa_effect > 0))
+
+```
+
+An alternative and to our knowledge unexplored (in the context of MPA simulation) fleet model though is a "constant-catch" strategy.
 Under this model, fishers have a catch objective, and exert as much (or little) effort as needed to achieve that objective.
@@
 Constant-catch dynamics might also occur in fisheries with constraining quotas that are not updated after the implementation of MPAs.
-Interestingly, when we simulated the effects of MPAs under different fleet models, the "constant-catch" scenario stood out as the only way for MPAs to actually produce a net conservation loss (Fig.\@ref(fig:XX)).
-While open-access fishing strategies can result in "scorched earth" scenarios where the only fish left are found inside the reserve, across all `r n_distinct(outcomes$experiment)` simulations the net effect of the reserves was still positive.
-Under a constant-catch scenario though, fishers have to fish much harder than before to get the same catch from a smaller part of the population, reducing the size structure of the population and, under XX% of our constant-catch simulations producing a net conservation loss (Fig.XX).
-XX
+While open-access fishing strategies can result in "scorched earth" scenarios where the only fish left are found inside the reserve, across `r percent(positive_outcomes$pp[positive_outcomes$fleet_model == "open-access"], accuracy = 1)` simulations the net effect of the reserves was still positive.
+Under a constant-catch scenario though, fishers have to fish much harder than before to get the same catch from a smaller part of the population, reducing the size structure of the population and subsequently causing net conservation loss under `r percent(1 - positive_outcomes$pp[positive_outcomes$fleet_model == "constant-catch"], accuracy = 1)` of our constant-catch simulations.
+This is an important and often overlooked possibility, especially as MPAs are increasinly implemented in quota-managed fisheries.
 
-Can the the non-trivial probability that the Channel Islands MPAs caused a net conservation loss (taking the difference-in-difference results at face value) in the most recent years be explained by the local fleets following a constant-catch strategy?
-While we do not have access to fine scale fishing data from the islands directly, reported catches for the species of interest in the Santa Barbara region in fact exhibit an overall downward trend in the years post reserve (see SI, XX).
+While we do not have access to fine scale fishing data from the Northern Channel Islands alone, reported catches for the species of interest in the Santa Barbara region in fact exhibit an overall downward trend in the years post reserve (see SI).
 We can most likely rule a negative MPA effect caused by a constant-catch fishing strategy then.
-What then is another explanation?
-The parallel trends assumption of our model appears to be valid initial years of protection, and we do not detect any clear evidence of interaction between the targeted and non-targeted species (e.g.
+What is another possible explanation for the recent downward trend in the estimated MPA effects?
+We do not see evidence to reject the parallel trends assumption of our difference-in-difference model in initial years of the MPAs, and we do not detect any clear evidence of interaction between the targeted and non-targeted species (e.g.
 depressed levels of non-targeted species as a result of predation from increased biomass densities of targeted species, see Fig.SXX).
 
-However, the Santa Barbara Channel experience a massive heatwave in XX.
-The Channle Islands represent the warmest edge of the distribution of many of the targeted species in our study.
-Conversely, many of the represented non-targeted species are sub-tropical.
-As a result, we hypothesize that the blob even of XX broke down the parllel trends assumption between the targeted and non-targeted species in our study, as the targeted species were more negatively affected by the warming temperatures.
-This is supported by the fact that we observe similar declines in targeted species both outside and inside the reserves (Fig.XX).
+But, the Channel Islands region (and the entire West coast of the USA) experienced a dramatic ‘marine heatwave’ beginning in 2014 and persisting through 2016, resulting in part in extremely elevated water temperatures throughout the region  [@gentemann2017].
+Biogeographic differences in the distributions of targeted and non-targeted species may confound the observed effects of MPAs.
+Many of the non-targeted species have warm thermal affinities and have increased in numbers since the heatwave [@freedman2019].
+The targeted group is made up mostly of fishes with cold-water affinities, such as members of the genus *Sebastes*.
+As such, we hypothesize that the recent evidence for a decline in densities of targeted species is due to environmental conditions that disproportionately affect the targeted group (and not for example due to concentrated fishing pressure outside the reserves).
+
+This hypothesis is supported by the sharp declines in densities of targeted species seen inside the MPAs themselves (see Fig.XX).
 Assuming that the reserves are sufficiently well enforced and large to protect targeted species biomass (an assumption supported by the response ratio results), if the cause of recent declines was due to increases in fishing pressure we would expect to see substantial declines only in the fished areas (Fig.
@@
 
+```{r}
+raw_biomass_density_plot
+
+ggsave(
+  raw_biomass_density_plot,
+  filename = here::here("documents","figs","raw_biomass_density_plot.png"),
+  width = 6,
+  height = 6
+)
+
+```
+
+\begin{figure*}%[tbhp]
+  \centering
+  \includegraphics[width=.9\linewidth]{`r here::here("documents","figs","raw_biomass_density_plot.png")`}
+  \caption{Raw Biomass Trends}
+  \label{mpa-effect}
+\end{figure*}
+
 ### When Can We Detect the Effects of MPAs?
 
-The Northern Channel Islands would appear to be an ideal system to study the effects of MPA networks on regional conservation outcomes.
-Simulation analysis built on existing MPA theory suggests that these reserves were likely to have a positive but modest effect (50% of simulated effects sizes less than 20% after 15 years of protection) on mean total biomass of targeted species.
-From 2003 (the year of MPA implementation) to 2011, our empirical results estimate a a highly uncertain but overall positive effect of the Channel Islands MPA, roughly in line with what MPA theory would suggest.
-However, these gains dissapear over the most recent years of available data, a phenomenon likely due to dramatic changes in the Channel Islands environment over the last decade that disproportinately affected the targeted species complex.
-Combined with the challenges we have demonstrated of using response ratios as a measure of regional impacts, we are left with an unclear picture of the overall effect of the Channel Island MPAs on biomass densities of targeted species after 14 years of protection.
+Containing a carefully designed, well-enforced, and well-studied MPA network, The Channel Islands would seem at face value to be an ideal location to study the regional conservation effects of protected areas.
+The persistently high response ratios suggest that despite overall decreases in targeted biomass densities inside and outside MPAs, the MPAs may still be providing protection within their borders.
+But, as we have shown here these response ratios are not necessarily an indicator of regional effects.
+The difference-in-difference strategy utilized here presents an alternative identification strategy, that while not without its own strict caveats presents some potential improvements over response-ratios as a means of estimating regional conservation effects.
+While we estimate a highly uncertain but overall positive effect at first, we are unable to detect a robust signal from 2012-2017.
+After 14 years of MPA protection we are left without a clear picture of the effect of the Channel Island MPA network on biomass densities of targeted fin-fish species.
 
-What does this result in a system as well-studied and well-enforced as the Channel Islands suggest for our ability to assess the  effects of other MPAs around the world?
-First, much of the empirical MPA literature has focused on the relatively clear signals provided by response ratios [@lester2009;].
-While we often see biomass densities of targeted species on the order of 400% higher inside MPAs than outside, as discussed in @ferraro2018 and expanded on here response ratios even as high as these are not a reliable indicator of the regional effects of MPAs.
-While response ratios can provide important evidence as to the effectivness of the MPAs in protecting fishes within their borders, if our interest is in estimating the effect of MPAs on the broader ecosystems in which they are located we must look for additional lines of evidence.
-
-The difference-in-difference strategy utilized here presents an alternative strategy, that while not without its own strict caveats presents some potential improvements over response-ratios as a means of estimating regional conservation effects.
-Given the natural variability of marine ecosystems, and the strong challenges of obtaining accurate samples from marine environments, how large of an effect would an MPA network have to allow a difference-in-difference strategy such as this to be a reliable measure of MPA effects?
-
-To provide guidance on this important question, we simulated data from a range of scenarios with increasing MPA effect size along with varying degrees of observation error and natrual recruitment variation.
-As an added measure, we include scenarios in which the sampled species go through recruitment regimes, which may be positive for both targeted and non-targeted species, or positive for non-targeted species and negative for targeted species.
-We then used a simlpe Bayesian difference-in-difference regression on these simulated data and estimated the percent error in the true MPA effect.
-
-XX .
-Mean absolote percent error dropped below XX% once the true MPA effect exceeded XX%.
-While this if of course no guarantee that any study of such an effect size will be successful (many model runs still ahd high error even at this effect size), our results help illustrate the effect size that may be required if managers which to have a given degree of confidence in their results.
-XX This is merely an illustrative exercise, omitting clearly critical factors such as detection probability and sampling strategy.
-However, since nearly any omission which one can think of would make an MPA effect harder to detect, not easier, these results serve as a useful floor for the likely uncertainty in estimating MPA effects.
+Our simulation results suggest though that we should not be surprised by this result.
+The Channel Islands MPAs cover approximately 20% of the waters in the Channel Islands, and while formal stock assessments are not available for many of the targeted species in our analysis, what evidence we have suggests that, as a group, these fish are on average not heavily overfished.
+Some species, such as California sheephead and blue rockfish were likely below target levels during the period [@alonzo2004;@dick2017], but projections based upon the overall average response across all species will likely suggest modest benefits even if a subset of species could experience much larger population gains.
+Harkening back to our simulations, we expect the average percentage difference in densities of targeted species with and without MPAs to be modest (Fig.\ref{fig:did-plot}).
+Effects of this size are likely to be challenging to detect empirically given the large natural variation of marine ecosystems (especially temperate reefs) and the observation error inherent in infrequent annual survey programs such as those provided by PISCO.
 
 ```{r}
+tmdata <- life_history_data %>% 
+  filter(targeted == 1, classcode %in% top_species) %>% 
+  select(tm, everything())
 
+median_tm = round(median(tmdata$tm))
+```
+
+As an additional complication, our simulations and difference-in-difference model examine percentage changes in biomass densities.
+As an extreme example, an increase in biomass densities from .2kg/m^2^ to .4kg/m^2^ would be produce as a 100% increase.
+While a large percentage increase, this is a small change in absolute biomass densities relative to the variance in the observation process itself (see Fig.S1 for a companion to Fig.\ref{conservation-effects} scaled by absolute population size).
+Beyond these challenges, the median estimated age at sexual maturity for the targeted species included in this study is `r median_tm` years, meaning that the span of this analysis represents less than three generations of MPA protection for half of the the measured species.
+Ongoing monitoring may yet reveal clearer effects.
+Analysis of more rapidly growing and maturing species, e.g.
+spiny lobster, may also reveal clearer signals.
+
+Given the natural variability of marine ecosystems, and the strong challenges of obtaining accurate samples from oceanic environments, how large of an effect would an MPA network have to allow a difference-in-difference strategy such as this to be a reliable measure of MPA effects?
+To provide guidance on this important question, we used our bio-economic model to simulate data from a range of scenarios with increasing MPA effect size, along with increasing degrees of observation error and natural recruitment variation.
+As an added measure, we include scenarios in which the sampled species go through recruitment regimes, which may be positive for both targeted and non-targeted species, or positive for non-targeted species and negative for targeted species.
+We then used a simple Bayesian difference-in-difference regression on these simulated data, and estimated the percent error between the estimated and true MPA effect.
+
+While unbiased across simulations, the difference-in-difference model struggled severely when MPA effect sizes were less than 25% and the model was faced with observation and process errors (Fig.\@ref(fig:val-plot)).
+Even larger effect sizes commonly misestiamted the true MPA effect by 50% or more.
+Obtaining a mean absolute percent error (MAPE) of 25% or less across our simulated datasets required an MPA effect of at least 30%.
+This is merely an illustrative exercise, omitting clearly critical factors such as detection probability and sampling strategy.
+However, since nearly any omission which one can think of would make an MPA effect harder to detect, not easier, these results serve as a useful floor for the likely difficulty in estimating MPA effects.
+In the context of the Channel Islands, given the potential effect size produced by out simulation model these results suggest that we might expect to be unable to detect a clear effect.
+
+<!-- XX need to think about whether MAPE is the correct error here - maybe need something like the variance of the estimator? -->
+
+```{r, include = FALSE}
+
+validation_plot
 ggsave(
@@
 
-This finding begs an important question then: when should we expect to see MPA effects big enough to stand a reasonable chance of detection?
+This finding begs an important question: when should we expect to see MPA effects big enough to stand a reasonable chance of detection?
 To address this, we simulated `r n_distinct(outcomes$experiment)` MPA scenarios across a wide range of life histories, network designs, and fishing dynamics (see SI for a full description of scenarios).
-Using these simulations, we can consider what we might expect the regional effects of an MPA network to be under a with range of plausible circumstances.
-Interested users can explore the use of this model for simulation at [danovando.shinyapps.io/simmpa](https://danovando.shinyapps.io/simmpa/).
+Using these simulations, we can consider what we might expect the regional effects of an MPA network to be under a range of plausible circumstances.
 
-Suppose that we are willing to tolerate a MAPE of XX%.
+Suppose that we are willing to tolerate a MAPE of 25%.
+Our analysis suggests that we would need an MPA effect size of at least 30% to achieve this.
 Two of the most critical drivers of MPA performance are the size of the MPA and the degree of fishing pressure.
-Looking across these two variables, our simulations suggest if the MPA network covers XX% or more of a species range and/or pre-MPA depletion is greater XX we might expect an effect size with a reasonable chance of detection.
-While recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, for commercial finfish many coastal MPA networks are likely to cover areas more in line with the Channel Islands (20%) (Fig.\@ref(fig:XX)).
+Looking across these two variables, our simulations suggest if the MPA network covers 25% or more of a species range and/or pre-MPA depletion is greater 60% we might expect an effect size with our target MAPE.
+While recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, for near-shore commercial fin-fish many MPA networks are likely to cover areas more in line with the Channel Islands (20%) (Fig.\@ref(fig:XX)-A).
 
-These results would seem to suggest some rather simple rules of thumb: put an MPA of sufficient size on a fishing population and one can expect large results.
-However, as discussed earlier, the MPA literature has highlighted a large number of variables beyond simply size and fishing pressure that can affect performance.
-To address this, we can
+These initial simulation results would seem to suggest some rather simple rules of thumb: put an MPA of sufficient size on a at least somewhat overexploited population and one can expect large and potentially detectable results.
+However, as we discussed earlier, the MPA literature highlights a large number of variables beyond simply size and fishing pressure that can affect performance.
+To address this, we can examine the variability in expected MPA effects across each of the two major axes (pre-MPA depletion and MPA size)  (Fig.\@ref(fig:XX)-B).
+As both MPA size and pre-MPA depletion increase, the potential for a large MPA effect increases.
+However, for both variables even at extremely large or extremely small values a wide range of MPA effects were possible (though as we might expect the effect of pre-MPA depletion, in other words fishing pressure, was much clearer than MPA size alone).
 
-plot error as a function of MPA effect.
+To put these results into context, the FAO estimates that 7% of the worlds fisheries with status estimates fall into the relatively unexploited category (roughly depletion less than 50%), 60% fall into the fully fished category (roughly depletion 50%-70%), and 33% fall into the heavily fished category (roughly depletion greater than 70%) [@fao2018, assuming a Bmsy/B0 ratio of roughly 40%, per @thorson2012], though works that include a broader range of fisheries estimate that 50% or more of stocks to fall into the heavily fished category [@costello2012;@rosenberg2017].
+Taken together then, the regional effects of MPA networks covering 25% or less of a species range may be difficult to detect in many placed with already well-managed fisheries, while for that size we might expect clearer effects in less-managed locations (though that of course ignores the complication of compliance with MPA regulations).
+Within these broad guidelines though a wide range of outcomes are possible: as a starting place users can use the bio-economic MPA simulation model developed for this paper to explore potential outcomes for specific MPAs using an interactive web application available at [danovando.shinyapps.io/simmpa](https://danovando.shinyapps.io/simmpa/).
 
-```{r}
+```{r, include=FALSE}
+expected_mpa_effect_plot
 
@@
 
-we hav shown that even in a well studied network, MPA effects on the order 0f 0-25% will be very challenging to detect.
-Under what circumstances might we expect MPA effects large enough to facilitate detection even in the face of observation and process error?
-
-THe persistently high response ratios suggest that despite overall decreases in targeted biomass densities inside and outside MPAs, the MPAs may still be providing protection within their borders.
-However
-
-we hav shown that even in a well studied network, MPA effects on the order 0f 0-25% will be very challenging to detect.
-Under what circumstances might we expect MPA effects large enough to facilitate detection even in the face of observation and process error?
-
-We tested the effects of different combinations of these theoretical drivers of MPA conservation effects using over 10,000 simulated MPA scenarios.
-A striking outcome of this simulation analysis is how the level of fishing pressure can cause dramatic and predictable shifts in conservation outcomes.
-We measure fishing pressure by the degree of simulated depletion (percent of unfished biomass gone, such that a depletion of 0% means the population is at 100% of unfished biomass, 100% means 0% of unfished biomass remains) in the scenario without MPAs.
-When MPAs were placed in relatively unexploited fisheries (depletions less than 40%), the median regional conservation effect was barely noticeable (`r small_effect`%).
-For ecosystems fished near MSY levels (depletion of 40%-60%), the median conservation effect was `r medium_effect`%, while for overfished ecosystems (without-MPA depletions above 60%) the median regional conservation effect was over  `r big_effect`% (Fig.\ref{conservation-effects}-A).
-Across all of these groups though, effects as low as declines of -50% and as high as an increase of more than 100% were possible.
-It should be noted though that these runs represent the range of scenarios evaluated in our simulation, and that certain types of scenarios may be much more likely than others in the real world.
-To put these results into context, the FAO estimates that 7% of the worlds fisheries with status estimates fall into the relatively unexploited category, 60% fall into the fully fished category, and 33% fall into the heavily fished category [@fao2018], though works that include a broader range of fisheries estimate that 50% or more of stocks to fall into the heavily fished category [@costello2012;@rosenberg2017].
-
-These simple results support the intuitive conclusion that bigger MPAs on overfished populations produce large positive conservation outcomes.
-However, when combined with other factors such as fleet dynamics, life history, and MPA size, the conservation effects of MPAs can vary widely in magnitude and timing, a finding corroborated by empirical evidence [@starr2015;@caselle2015;@nickols2019;@villasenor-derbez2019] (Fig.\ref{conservation-effects}-B).
-As the area protected inside MPAs increases, positive effects become more likely, but even MPAs protecting most of a population's footprint can produce conservation effects as low as 0% and as high as 200%.
-The degree of depletion (i.e.
-fishing pressure) has a clearer signal, with small conservation benefits of MPAs when depletion is low, and larger effects when depletion is high, but again even for severely depleted populations the conservation effects of MPAs can vary widely (although nearly always with population benefits).
-
-While positive conservation outcomes were much more likely to occur across our simulations (`r percent(pos_runs, accuracy = 1)` of simulations), the `r percent(1 - pos_runs, accuracy = 1)` of runs that produced net conservation losses are of particular interest.
-These runs occurred almost exclusively when the simulated fleet followed a "constant-catch" fishing strategy.
-Under the constant-catch fleet model, fishing communities (or fishery policies) seek to catch the same amount regardless of the presence of an MPA.
-While a constant-catch greater than MSY is not possible over the long-term under the assumptions of our model, over the short-term a constant-catch scenario is not implausible.
-Subsistence fisheries may use a constant-catch style policy over the short-term, as they seek to ensure that their food needs are met.
-More industrial fisheries may have pre-arranged agreements with buyers to deliver set amounts of fish.
-Constant-catch dynamics might also occur in fisheries with constraining quotas that are not updated after the implementation of MPAs.
-These patterns suggest that coupling MPA implementation with appropriate management reforms that reduce the effects of constant-catch strategies can move the predicted conservation effects of MPAs to being nearly uniformly positive.
-The possibility of constant-catch strategies to result in net MPA conservation losses highlights the critical importance of understanding the broader management context in which MPAs are placed.
-
-### How can we use response ratios
-
-Unfortunately, neither of these experimental designs has been implemented, because parallel worlds do not exist and MPAs are never placed at random.
-Careful causal inference requires some other means of controlling for biases introduced by factors such as the MPA siting process and biological spillover and concentration of fishing effort  [@ferraro2018;@larsen2019].
-Conservation effects are typically estimated by comparing organism densities or biomass inside MPAs to densities or biomass in selected control sites outside MPAs, which we will refer to as response ratios.
-@lester2009 and @halpern2002 present meta-analyses of hundreds of such studies.
-These results often find massively higher densities and biomasses inside MPAs than outside [median biomass increases of  over 400%, @lester2009].
-While response ratios are intended as a measure of the effectiveness of reserves within their borders, these result stand in contrast to the smaller region wide effect sizes predicted by our simulation model.
-This raises the question: How reliable are density or biomass ratios as estimators of regional conservation effects?
-
-Failure to account for these spillover effects can result in a biased estimate of the true effect of a policy such as MPA placement [@larsen2019].
-In an ideal setting, the control site used in a response ratio would be a perfect counterfactual for what would have occurred without the MPA.
-We used our simulation model to approximate this scenario, calculating the response ratio as the ratio of densities inside the MPAs relative to the overall density from the paired simulation without MPAs multiplied by the area of the population in an MPA (to account for the fact that we would expect the effectiveness of a response ratio as a proxy for regional conservation effects to scale with the size of the MPA relative to the range of the species, Fig.\ref{density-ratio}-A).
-In this idealized case, the response ratio is often a reasonable estimator for the regional conservation effect.
-
-What happens when the “control” sites are in fact affected by the application of the MPA through biological spillover and fishing concentration?
-In order to address this possibility, we calculated a new response ratio as the mean density inside MPAs relative to the mean density outside the MPAs (both weighted by distance from MPA borders and scaled by MPA size).
-We only include simulations in which habitat and larval dispersal rates are identical inside and outside of MPAs, to approximate a scenario in which treatment and control sites have been paired by ecological characteristics.
-Under these circumstances the response ratio is a somewhat biased and very inaccurate estimator of the true regional effect of MPAs.
-
-While on average high response ratios correspond with high true regional effects, the variance is high, so the response ratio alone cannot be used as a reliable proxy for the true regional effect, without accounting for the sources of biases discussed here.
-For higher movement rates, response ratios were frequently near 0\% (potentially leading stakeholders to conclude that the MPA had been ineffective), when in fact in many cases the true effect of the MPA was highly positive.
-For highly sedentary species, extremely high response ratios could create the appearance of massive conservation gains when in fact the net effect on the regional population has been near zero or even negative (Fig.\ref{density-ratio}-B).
-
-We note that spatial before-after-control-impact (BACI) studies present a potential improvement over response ratios, controlling for differences in mean densities in the treated and control regions as well as MPA independent trends, but are much rarer due to the need for extensive pre-MPA monitoring.
-However, spatial BACI still requires that the MPAs do not affect the control sites, and as such has many of the same limitations as response ratios as an estimator of regional conservation effects.
+Conclusions {#conclusion .unnumbered}
+------
 
@@
 To date, this assumption is insufficiently tested, and this is the focus of our paper.
-Our results show that regional conservation benefits of MPAs are highly context dependent and in many circumstances, are likely to be so small that they are nearly impossible to detect empirically.
+Our results show that regional conservation benefits of MPAs are highly context dependent and in many circumstances, are likely to be small enough that they are nearly impossible to detect empirically.
 Indeed, this is exactly what we found in our empirical case study from the Channel Islands, California, USA.
 
-We find regional conservation gains from MPAs in `r percent(pos_runs, accuracy = 1)` of our simulations (Fig.\ref{conservation-effects}).
-MPAs covering less than 25% of the population range were unlikely to produce regional conservation gains above 10% unless the stock would have been severely overfished without the MPA.
-Given that few marine species on the planet have more than 25% of their entire range protected in MPAs, this suggests that large regional gains are unlikely to be common unless the overall extent of MPAs grows globally.
-Large MPAs protecting highly depleted stocks almost always produced large regional conservation gains, as should be expected.
-
-The median conservation effect of `r round(100*median(eqo$mpa_effect))`\% stands in contrast to the large within-MPA effects on biomass densities reported by @halpern2003a and @lester2009.
-This difference is due in part to the fact that within-MPA effects are likely to be much larger than regional-MPA effects, but may also reflect the many challenges of translating response ratios into regional conservation effects.
-Our simulation analysis shows that without proper controls (in particular for the effects of fleet redistribution), response ratios well over 100% are entirely plausible even when the actual conservation effect is much lower (and *vice versa*).
-
-We were not able to detect a statistically significant difference in the densities of targeted and non-targeted finfish species over the 13 years of MPA protection in the Channel Islands covered by our analysis.
+<!-- We were not able to detect a statistically significant difference in the biomass densities of targeted and non-targeted fin-fish species over the 14 years of MPA protection in the Channel Islands covered by our analysis.
 The data (and model assumptions) provide support for both negative and positive MPA effects, although the size of these estimated effects is well within the range that our simulation analysis suggests are plausible.
 At face value, this lack of a clear result may seem surprising given the size and carefully studied nature of this MPA network.
-
-Our simulation analysis suggests that this result should perhaps be expected.
-The Channel Islands MPAs cover approximately 20% of the waters in the Channel Islands, and while formal stock assessments are not available for many of the targeted species in our analysis, what evidence we have suggests that, as a group, these fish are on average not heavily overfished.
-Some species, such as California sheephead and blue rockfish were likely below target levels during the period [@alonzo2004;@dick2017], but projections based upon the overall average response across all species will likely suggest modest benefits even if a subset of species could experience much larger population gains.
-Harkening back to our simulations, we expect the average percentage difference in densities of targeted species with and without MPAs to be modest (Fig.\ref{did-plot}).
-Effects of this size are likely to be challenging to detect empirically given the large natural variation of marine ecosystems (especially temperate reefs) and the observation error inherent in infrequent annual survey programs such as those provided by PISCO.
-
-```{r}
-tmdata <- life_history_data %>% 
-  filter(targeted == 1, classcode %in% top_species) %>% 
-  select(tm, everything())
-
-median_tm = round(median(tmdata$tm))
-```
-
-In addition, our simulations examine percentage increases in biomass densities.
-As an extreme example, an increase in biomass densities from 2kg/m^2^ to 4kg/m^2^ would be reported as a 100% increase.
-While this chance produces a large percentage increase, this is a small change in absolute biomass densities relative to the variance in the observation process itself (see Fig.S1 for a companion to Fig.\ref{conservation-effects} scaled by absolute population size).
-Beyond these challenges, the median estimated age at sexual maturity for the targeted species included in this study is `r median_tm` years, meaning that the span of this analysis represents less than three generations of MPA protection for half of the the measured species.
-Ongoing monitoring may yet reveal clearer effects.
-Analysis of more rapidly growing and maturing species, e.g.
-spiny lobster, may also reveal clearer signals.
-
-While the MPA effect estimated by our model was not significant in any one year, the positive trend from 2005 to 2014 is evident, as is the sharp decline from 2015-2017.
-To what should we attribute this apparent shift?
-We cannot reject the parallel trends assumption between the targeted and non-targeted species in the years before the MPAs (Fig.\ref{did-plot}).
-But, the Channel Islands region (and the entire West coast of the USA) experienced a dramatic ‘marine heatwave’ beginning in 2014 and persisting through 2016, resulting in part in extremely elevated water temperatures throughout the region  [@gentemann2017].
-Biogeographic differences in the distributions of targeted and non-targeted species may confound the observed effects of MPAs.
-Many of the non-targeted species have warm thermal affinities and have increased in numbers since the heatwave [@freedman2019].
-The targeted group is made up mostly of fishes with cold-water affinities, such as members of the genus *Sebastes*.
-As such, we hypothesize that the recent evidence for a decline in densities of targeted species is due to environmental conditions that disproportionately affect the targeted group (and not for example due to concentrated fishing pressure outside the reserves).
-This hypothesis is supported by the sharp declines in densities of targeted species seen inside the MPAs themselves (see Fig.S27-S29), suggesting a driver other than fishing may be at play.
+-->
 
@@
 Our simulation model is by no means exhaustive (for example it ignores features such as species interactions, habitat effects, and climate feedback), but captures many of the core factors theorized to affect MPA performance.
-Our results show that while attributes such as MPA size and fishing pressure are important factors in determining the effects of protection, local fleet dynamics and the movement rates of adult fish can dramatically affect the outcomes of protected areas as well  (Fig.\ref{conservation-effects}).
+While attributes such as MPA size and fishing pressure are important factors in determining the effects of protection, local fleet dynamics and the movement rates of adult fish can dramatically affect the outcomes of protected areas as well  (Fig.\ref{conservation-effects}).
 Far from being a simple tool with clear outcomes, the effects of MPAs can be highly context dependent, requiring - we would argue - a bio-economic model of at minimum the complexity presented here to help communities design and set expectations of MPAs at the tactical level.
 While users may not be able to parameterize every aspect of a model such as that presented here, working with stakeholders to visualize the implications of, for example, different fleet responses to MPA implementation is a critical step in MPA design.
-Readers can use our simulation model to explore the effects of MPAs through an interactive web application available [here](https://danovando.shinyapps.io/simmpa/).
 
 Once simulation models have been used to help design an MPA, how can users evaluate whether it is achieving their objectives?
-Response ratios are commonly used as evidence for conservation outcomes of MPAs [@lester2009], but as suggested in @ferraro2018 and @larsen2019 and demonstrated here, without careful attention to the design of control sites (accounting for example for the displacement of effort by MPAs), response ratios may be highly unreliable estimators of regional MPA effects.
-As @larsen2019 suggests, there are many potential alternatives for estimating the effects of MPAs that better account for the challenges of causal inference.
-We applied one such approach here (a difference-in-difference estimator), and yet were still unable to reach robust conclusions as to the effect of MPAs on the density of targeted species in the Channel Islands, due to the likely small size of the true effect relative to the strength and variability of environmental drivers.
+Response ratios are commonly used as evidence for conservation outcomes of MPAs; @lester2009 and @halpern2002 present meta-analyses of hundreds of such studies.
+These results often find massively higher densities and biomass inside MPAs than outside [median biomass increases of  over 400%, @lester2009].
+But as suggested in @white2011, @ferraro2018, @larsen2019 and further demonstrated here, without careful attention to the design of control sites (accounting for example for the displacement of fishing effort by MPAs), response ratios may be highly unreliable estimators of regional MPA effects.
+When MPAs affect nearby control sites used in response ratios through biological spillover or concentration of fishing effort, it is entirely possible to for MPA to produce massive response ratios while simultaneously having minimal effects on the entire population partially protected by the MPAs.
+As @larsen2019 suggests, there are many potential alternatives for estimating the effects of MPAs that better account for the challenges of causal inference (though that may be more data-intensive).
+We applied one such approach here (a difference-in-difference estimator), and yet were still unable to reach robust conclusions as to the effect of MPAs on the biomass density of targeted finfish in the Channel Islands, due to the likely small size of the true effect relative to the strength and variability of environmental drivers.
 
-While this does not mean that all MPAs will face similar challenges in estimating their effects, our results in the rigorously studied Channel Islands system make clear that in many instances empirically detecting a clear regional effect of MPAs may not be possible.
+While this does not mean that all MPAs will face similar challenges in estimating their effects, our results in the rigorously studied and iconic Channel Islands Marine Protected Area network make clear that in many instances empirically detecting a clear regional effect of MPAs may not be possible.
 How then should stakeholders go about adaptively monitoring and managing MPAs?
 Simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can perhaps be tuned to focus on the species groups that have the highest chance of a detectable effect size [@nickols2019].
-Expanding data collection to include robust monitoring of fleet dynamics may help statistical approaches to isolate the true effect of MPAs on conservation outcomes by allowing for improved control of the effects of fishing dynamics, and allow managers to take into account potential negative interactions between MPAs and fleet dynamics such as those that may occur under constant-catch dynamics.
+Expanding data collection to include robust monitoring of spatio-temporal fleet dynamics may help assess the validity of control sites used in response ratios, support the direct inclusion of these fleet dynamics into statistical models, and allow managers to take into account potential negative interactions between MPAs and fleet dynamics such as those that may occur under constant-catch dynamics.
 Whenever possible monitoring programs should be implemented prior to MPA implementation to provide a pre-treatment benchmark.
-Non-equilibrium analyses (Fig.\ref{did-plot}) also help set expectations for effect sizes over time.
-Beyond that, educating communities about the challenges of estimating the effects of MPAs can help set expectations, so that a lack of a clear effect is not necessarily viewed as a failure of the program, but rather considered in the context of reasonable expectations.
+Non-equilibrium analyses also help set expectations for effect sizes over time [@nickols2019,Fig.\ref{did-plot}].
+Educating communities about the challenges of estimating the effects of MPAs can help set expectations, so that a lack of a clear effect is not necessarily viewed as a failure of the program, or large positive result based solely on response ratios as a clear sign of success and subsequent increased fishing opportunity.
+Rather, results and subsequent management actions (such as adaptation of MPA network design) must be considered in the context of reasonable expectations for the MPAs in question.
 While this paper has focused on the conservation outcomes of MPAs, future work must also address the challenge of predicting and estimating the fishery impacts of protected areas.
@@
 As the number and size of global MPA networks increase, it is critical that we both set appropriate expectations for their outcomes, and plan how we will monitor the performance of these protected areas over time.
-While the history of MPA science has made important strides in helping us understand the dynamics of protected areas, the future of MPA science must directly tackle the challenge of evaluating the performance of these MPAs, and adapting their design as needed to best achieve objectives.
+While the history of MPA science has made important strides in helping us understand the dynamics of protected areas, the future of MPA science must directly tackle the challenge of evaluating the performance of these MPAs at the regional scale, and adapting their design as needed to best achieve objectives.
+This is particularly true if communities are depending on MPAs as their fisheries management tool.
 Commonly employed metrics such as response ratios may be applicable in some circumstances, but can have severe shortcomings as metrics of regional conservation effects.
 Dependence on unreliable estimators of MPA effects may lead to stakeholders incorrectly attributing negative environmental shocks as MPA failures, or interpreting data arising from scorched-earth fishing outside MPAs as a conservation success.
-Bio-economic modeling can help frame community expectations, reducing the potential for a reduction in support if unrealistic conservation or fishery gains are not realized.
-Statistical approaches that explicitly address complications such as the spatial spillover effects of MPAs (such as the difference-in-difference approach used here) may give users an improved understanding of the performance of their MPAs.
-Effective use of simulation and statistics to clearly communicate what we should expect and what we can detect from MPAs is critical in ensuring that MPAs play effective roles in fisheries management and marine conservation.
+Bio-economic modeling can help frame community expectations, reducing the potential for a reduction in support if unrealistic conservation or fishery expectations are not realized.
+Statistical approaches that explicitly address complications such as the spatial spillover effects of MPAs (such as the difference-in-difference approach used here) may give users an improved understanding of the performance of their MPAs, but even they may struggle when expected effect sizes are small.
+Clearly communicating what we should expect, and what we can detect,m from MPAs is critical in ensuring that MPAs play effective roles in fisheries management and marine conservation.
 
@@
 All analysis were conducted in R [@rcoreteam2019].
-Our main difference-in-difference model was fit using Template Model Builder in R [@kristensen2016].
+Our main difference-in-difference model was fit using Stan [@carpenter2017] using the `rstanarm` package [@goodrich2020].
 Code needed to replicate results can be found [here](https://github.com/DanOvando/regional-effects-of-mpas).
@@
 
+Each simulation is randomly assigned a fleet model of the form
+
+  1. Open access: fishing effort changes in response to profit-per-unit-effort
+
+  2. Constant effort: total fishing effort is constant over time (unless altered by MPA displacement model)
+
+  3. Constant catch: the fleet exerts as much effort as need to achieve a target catch
+
+This fleet model is paired with a gear selectivity ranging from .1 to 1.5 of the length at 50% maturity for the species in question, and the fleet model is tuned to achieve a target fishing mortality relative to natural mortality ratio at equilibrium.
+
 Along with the fleet dynamics model, each simulation is assigned a random fleet dispersal scenario: uniform dispersal (where the total effort of the fleet is divided evenly among all open patches), catch dispersal (where the total effort of the fleet is divided according to the catchable biomass in each available patch), and profit dispersal (where the total effort of the fleet is divided according to the profit per unit effort in each available patch).
@@
 Each simulation is run to equilibrium with and without the selected MPA strategy (holding all else constant).
-We then measure the difference in biomass and fishery catches in each time step in the scenario with and without the MPAs to calculate the conservation and fishery effects of the MPAs over time.
+We then measure the difference in biomass densities each time step in the scenario with and without the MPAs to calculate the conservation and fishery effects of the MPAs over time.
 
@@
 
-```{r ci-map, fig.cap = "Map of study region and sampling locations.
-Shaded polygons indicate location of MPAs.
-Points represent sampling locations, and color indicates the number of observations recorded at a given point", include = FALSE}
+```{r ci-map, include = FALSE}
 
-sample_sites <- pisco_data %>%
-  left_join(site_data, by = c("site","side")) %>%
-  filter(is.na(eventual_mpa) == F) %>%
-  filter(region %in% c("ANA", "SCI","SRI",'SMI'),
-         classcode %in% top_species) %>%
-  group_by(lat_wgs84, lon_wgs84, site, region, eventual_mpa) %>%
-  count()
+sample_location_plot
 
-channel_islands <- sf::st_read(here::here("data","cinms_py2","cinms_py.shp"))
-
-ci_bbox <- sf::st_bbox(channel_islands)
-
-sample_sites <-  sample_sites %>%
-  st_as_sf(coords = c("lon_wgs84", "lat_wgs84"),
-           crs = 4326)
-
-california <- rnaturalearth::ne_states(country = "united states of america", returnclass = "sf") %>% 
-  filter(name == "California") %>% 
-    sf::st_transform(crs = sf::st_crs(channel_islands))
-
-channel_islands_mpas <- sf::st_read(here::here("data","MPA_CA_Existing_160301"))%>%
-  sf::st_transform(crs = sf::st_crs(channel_islands))
-
-pisco_ci_map_plot <-  ggplot() + 
-  geom_sf(data = california, fill = "lightgrey") +
-  geom_sf(data = channel_islands_mpas, fill = "lightblue") +
- geom_sf(data = sample_sites,shape = 21, fill = "darkgrey",
-                                      size = 2, alpha = .75) +
-
-  # scale_color_gradient(low = "white", high = "orangered",
-  # guide = guide_colorbar(frame.colour = "black",frame.linewidth = 1,
-  # barheight = unit(13, "lines")),
-  # name = "Observations")  +
-  labs(x = "", y = "") +
-  scale_y_continuous(limits = c(33.8, 34.2)) +
-  scale_x_continuous(limits = c(-120.75,-119.25)) +
-      ggspatial::annotation_scale(location = "br") + 
-  #   coord_sf(xlim = c(bbox['xmin'] - .3, bbox['xmax'] + .06),
-  # ylim = c(bbox['ymin'] - .1, bbox['ymax'] + .1)) +
-    ggsci::scale_color_npg()
-
-# pisco_ci_map_plot <-  ggplot() +
-#   geom_sf(data = tests2, fill = c("#FFD39B")) +
-#   geom_sf(data = ca_mpas, inherit.aes = FALSE, alpha = 0.5) +
-#   geom_sf(data = ci_map,inherit.aes = FALSE,
-#   aes(color = eventual_mpa),
-#   alpha = 0.75,
-#   show.legend = FALSE,
-#   size = 2) +
-#   coord_sf(xlim = c(bbox['xmin'] - .3, bbox['xmax'] + .06),
-#   ylim = c(bbox['ymin'] - .1, bbox['ymax'] + .1)) +
-#   # scale_color_gradient(low = "white", high = "orangered",
-#   # guide = guide_colorbar(frame.colour = "black",frame.linewidth = 1,
-#   # barheight = unit(13, "lines")),
-#   # name = "Observations")  +
-#   labs(x = "", y = "") +
-#       ggspatial::annotation_scale(location = "br", width_hint = 0.4, text_col = "white") + 
-#   theme(panel.background = element_rect(fill = "lightblue"))
-# 
-pdf(file =  "figs/channel-islands.pdf",width = 8, height = 5, useDingbats = TRUE)
-print(pisco_ci_map_plot)
-dev.off()
+ggsave(
+  sample_location_plot,
+  filename = here::here("documents","figs","sample_location_plot.png"),
+  width = 6,
+  height = 4
+)
 
@@
   \centering
-  \includegraphics[width=1\linewidth]{figs/channel-islands.pdf}
+  \includegraphics[width=1\linewidth]{`r here::here("documents","figs","sample_location_plot.png")`}
   \caption{Map of PISCO sampling locations within the the Channel Islands National Marine Sanctuary, California, USA.
@@
 
-```{r raw-trend, fig.cap="Centered and scaled mean annual biomass density of included species (faded lines) and smoothed means of targeted and non-targeted groups, and mean (darker lines) and 95% confidence interval of the mean (ribbon) over time", include = FALSE}
-raw_did_plot <- fitted_data %>%
-  select(year, targeted,classcode, log_density, any_seen) %>%
-  mutate(density = exp(log_density) * any_seen) %>%
-  group_by(classcode,targeted, year) %>%
-  summarise(md = mean(density)) %>%
-  group_by(classcode,targeted) %>%
-  mutate(smd = (md - mean(md)) / sd(md)) %>%
-  ggplot(aes(year, smd, color = targeted == 1, fill = targeted == 1)) +
-  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red", size = 1.5) +
-  geom_line(aes(group = interaction(targeted, classcode)), alpha = 0.5) +
-  geom_smooth(size = 2) +
-  labs(y = "Centered and Scaled Mean Density", x = "Year") +
-    scale_color_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
-      scale_fill_manual(values = c("steelblue", "tomato") ,name = "Targeted?") +
-    labs(y = "Centered and Scaled Mean Density", x = "Year") + 
-  theme(legend.position = "top",
-        panel.grid.major.x = element_blank(),
-        panel.grid.minor.x = element_blank()
-        )
-
-raw_did_plot <- raw_did_plot +
-      scale_color_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "") +
-        scale_fill_manual(values = c("steelblue", "tomato"), labels = c("Non-Targeted", "Targeted"), name = "")
-
-pdf(file =  "figs/pop-trends.pdf",width = 8, height = 5, useDingbats = TRUE)
-print(raw_did_plot)
-dev.off()
-
-```
-
-\begin{figure}%[tbhp]
-  \centering
-  \includegraphics[width=1\linewidth]{figs/pop-trends.pdf}
-  \caption{Centered and scaled mean annual biomass densities of included targeted (red) and non-targeted (blue) fish species (faded lines) and smoothed means with 95\% confidence intervals over time (darker lines and ribbons).
-Vertical red dashed line shows year of MPA implementation.}
-  \label{pop-trends}
-\end{figure}
-
-We applied this difference-in-difference strategy to a large dataset in California.
-We used empirical kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands with the ultimate goal of testing the regional effects of MPAs in a real world context.
+ The difference-in-difference model used empirical kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands.
 A network of MPAs covering approximately 20% of the islands’ waters was put in place in 2003 as part of the California Marine Life Protection Act (MLPA) (see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of the MLPA).
 PISCO conducts visual SCUBA surveys at a large number of rocky-reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\ref{channel-islands}-\ref{pop-trends}).
+.
+The details of the monitoring program are described in XX
 
-The key assumptions of this approach are that a) within the time-frame of the model there are no significant interaction effects between the targeted and non-targeted species (which in fact we do not detect, see SI), and b) that in the absence of the MPAs both the targeted and non-targeted groups of species would have exhibited similar trends in densities.
-The advantage to this approach is that given the time-frame of the model (18 years), we believe that spillover effects of MPA placement on the non-targeted species are likely to be much less severe than the effects of biological spillover and fishing concentration that may bias the performance of estimators such as a response ratios or a spatial BACI design.
+The key assumptions of the difference-in-difference model are that a) within the time-frame of the model there are no significant interaction effects between the targeted and non-targeted species (which in fact we do not detect, see SI), and b) that in the absence of the MPAs both the targeted and non-targeted groups of species would have exhibited similar trends in densities.
+The advantage to this approach is that given the time-frame of the model (14 years with MPAs), we believe that spillover effects of MPA placement on the non-targeted species are likely to be much less severe than the effects of biological spillover and fishing concentration that may bias the performance of estimators such as a response ratios.
 
-We use a difference-in-difference regression style regression to attempt to estimate the causal effect of MPAs on regional targeted fish biomass.
-Empirically, this amounts to estimating the pre-post MPA difference in the biomass densities of targeted species net the same difference for non-targeted species in the Channel Islands.
+The regression amounts to estimating the pre-post MPA difference in the biomass densities of targeted species net the same difference for non-targeted species in the Channel Islands.
 
 The simplified form of this model is
+XX adjust to GAMMA glm
 
 \begin{equation}
-  log(d_{i}) = \beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + e_{i}
+  d_{i} = Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i},shape, scale)
 \label{eq:did}
@@
 where $d_i$ is the biomass density at observation *i*, *T* indicates whether the observation *i* is for a targeted ($T = 1$) or non-targeted ($T = 0$) species, and *MPA* marks whether observation *i* is in a pre MPA ($MPA = 0$) or post MPA ($MPA = 1$) state.
+is a vector of coefficients for additional control variables in matrix *X* such as water visibility and observer experience.
 Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment (*MPA*) on the treated (targeted species).
-The full form of the estimation model is much more involved, integrating uncertainty in a hierarchical model from the individual observation level up to the mean trends of targeted and non-targeted species.
-See SI for details of estimation model.
+The shape and scale parameters of the Gamma distribution are estimated as well.
+See SI for further details of the estimation model.
 
@@
 We do not detect any significant differences in the trends of the biomass densities of the targeted and non-targeted species in the years before the MPAs.
+XX
 

### 2020-03-26 documents/ovando-regional-effects-of-mpas-v2.tex
passing to squid since sf on mac is broken beyond belief

--- 2019-10-10 documents/ovando-regional-effects-of-mpas.tex
@@
 
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: danovan@uw.edu}
+\emph{as we move away from ``set it and forget it'' management towards
+adaptive management of marine ecosystems, it is critically important
+that we develop robust strategies for monitoring the effectiveness of
+our management intercentions}
 
@@
 2020 and the International Union for Conservation of Nature call for
-30\% by 2030) rely on the belief that well-designed MPAs will achieve
-benefits both within, and outside, their borders. Despite these
+30\% by 2030) are based on the expectaion that well-designed MPAs will
+achieve benefits both within, and outside, their borders. Despite these
 assumptions, our collective scientific understanding of the
@@
 
-What is the scientific evidence of the conservation effects of MPAs?
 Numerous studies provide evidence that well-enforced and appropriately
@@
 As stakeholders around the world increasingly seek to use MPAs in marine
-resource management portfolios, and the performance of existing MPAs is
-evaluated, it is critical that we develop a better understanding of the
-magnitude and drivers of regional-scale MPA effects. To address this
-gap, this study examines two critical questions: 1) What do we expect
-the regional-scale conservation effects of MPAs to be and 2) When (and
-how) can we expect to empirically detect these effects? We answer these
-questions using a simulation analysis to frame the theoretical regional
-conservation impacts of MPAs. We then present an empirical test of the
-evidence for regional-level conservation effects of MPAs resulting from
-a network of closures put in place in the Channel Islands, California,
-USA, in 2003.
+resource management portfolios, it is critical that we develop a better
+understanding of the magnitude and drivers of regional-scale MPA
+effects. This is particuarly true in the context of a changing climate,
+in which ``set it and forget it'' management strategies are increasinly
+untenable; we need to be able to monitor the performance of MPAs in
+order to adapt them as dictated by a shifting environment. The Channel
+Islands National Marine Sanctuary, California, USA provides an ideal
+case study to address this need. A network of protected areas covering
+approximately 20\% of the Islands' waters was put in place in 2003 (22).
+This MPA network has subsequently been used as a model case study in
+protected area design around the world XX.
+
+We use data from the first 14 years of protection to provide what is to
+our knowledge the first empirical assessment of the regional effect of a
+large MPA network on a wide array of finfish species. In contrast to
+clear differences in biomass densities observed inside and outside of
+well protected MPAs both globally (5) and in the Channel Islands
+({\textbf{???}}) we are unable to detect a clear regional effect from
+the Channel Islands MPAs. We build off of existing MPA theory to
+consider why this might be, and provide guidance for scientists and
+managers as to when and how we might expect to estimate the regional
+conservation effects of MPAs.
 
@@
 
-Much of the literature on the effects of MPAs focuses on assessing the
+The empirical MPA literature has generally focused on assessing
 conservation effects within the borders of protected areas (7). While
@@
 broader question we should ask of MPAs is not just are there more fish
-inside their borders, but also how have the reserves affected abundances
+inside their borders, but also how have reserves affected populations
 throughout the region in which they are located? This logic, that MPAs
@@
 We define the regional conservation effects of MPAs as the change in
-total biomass of fish (summing inside and outside of MPAs) relative to
-the total biomass of fish that would have occurred without the MPAs
-(acknowledging that other outcomes such as increased biodiversity or
-resiliency are also important to conservation but are beyond the scope
-of this analysis).
-
-We used a simulation model to examine the regional conservation effects
-of MPAs under a wide array of conditions that might reasonably be
-expected to affect these outcomes. The model explores combinations of
-MPA designs, life histories, and fishing fleet dynamics (See Table.S1
-for a complete description of simulation variables). We use this model
-to explore how different drivers interact to affect the conservation
-outcomes of MPAs.
-
-\hypertarget{conservation}{%
-\subsubsection*{Theoretical Conservation Effects}\label{conservation}}
-\addcontentsline{toc}{subsubsection}{Theoretical Conservation Effects}
-
-As a thought experiment, imagine a region that has driven its fish
-populations to near extinction and then places 100\% of its waters
-inside a no-take MPA. In that setting, we would expect the
-regional-scale conservation benefits to be massive. On the other hand,
-imagine implementing an MPA in an area where a sedentary species has
-been only lightly fished. In that setting, it is possible to create a
-small net conservation loss if the concentration of fishing pressure
-outside the reserve has a greater effect than the biomass buildup inside
-the reserve. Within these broad bounds, numerous factors can act to
-affect the regional effects of MPAs. These include the scale of adult
-and larval dispersal relative to the size of the MPAs (22--25), the
-strength and timing of density dependence in the population (e.g.~pre-
-or post-settlement), how overfished the population would be without the
-MPA, and how fishing activity and management responds to the
-implementation of the MPAs (4, 8, 26--32). In addition, even for the
-same total area of MPAs, the location and spacing of the MPAs can have a
-profound influence on their cumulative impact through habitat and
-network effects (4, 33).
-
-We tested the effects of different combinations of these theoretical
-drivers of MPA conservation effects using over 10,000 simulated MPA
-scenarios. A striking outcome of this simulation analysis is how the
-level of fishing pressure can cause dramatic and predictable shifts in
-conservation outcomes. We measure fishing pressure by the degree of
-simulated depletion (percent of unfished biomass gone, such that a
-depletion of 0\% means the population is at 100\% of unfished biomass,
-100\% means 0\% of unfished biomass remains) in the scenario without
-MPAs. When MPAs were placed in relatively unexploited fisheries
-(depletions less than 40\%), the median regional conservation effect was
-barely noticeable (4\%). For ecosystems fished near MSY levels
-(depletion of 40\%-60\%), the median conservation effect was 28\%, while
-for overfished ecosystems (without-MPA depletions above 60\%) the median
-regional conservation effect was over 100\%
-(Fig.\ref{conservation-effects}-A). Across all of these groups though,
-effects as low as declines of -50\% and as high as an increase of more
-than 100\% were possible. It should be noted though that these runs
-represent the range of scenarios evaluated in our simulation, and that
-certain types of scenarios may be much more likely than others in the
-real world. To put these results into context, the FAO estimates that
-7\% of the worlds fisheries with status estimates fall into the
-relatively unexploited category, 60\% fall into the fully fished
-category, and 33\% fall into the heavily fished category (34), though
-works that include a broader range of fisheries estimate that 50\% or
-more of stocks to fall into the heavily fished category (35, 36).
-
-These simple results support the intuitive conclusion that bigger MPAs
-on overfished populations produce large positive conservation outcomes.
-However, when combined with other factors such as fleet dynamics, life
-history, and MPA size, the conservation effects of MPAs can vary widely
-in magnitude and timing, a finding corroborated by empirical evidence
-(37--40) (Fig.\ref{conservation-effects}-B). As the area protected
-inside MPAs increases, positive effects become more likely, but even
-MPAs protecting most of a population's footprint can produce
-conservation effects as low as 0\% and as high as 200\%. The degree of
-depletion (i.e.~fishing pressure) has a clearer signal, with small
-conservation benefits of MPAs when depletion is low, and larger effects
-when depletion is high, but again even for severely depleted populations
-the conservation effects of MPAs can vary widely (although nearly always
-with population benefits).
-
-While positive conservation outcomes were much more likely to occur
-across our simulations (94\% of simulations), the 6\% of runs that
-produced net conservation losses are of particular interest. These runs
-occurred almost exclusively when the simulated fleet followed a
-``constant-catch'' fishing strategy. Under the constant-catch fleet
-model, fishing communities (or fishery policies) seek to catch the same
-amount regardless of the presence of an MPA. While a constant-catch
-greater than MSY is not possible over the long-term under the
-assumptions of our model, over the short-term a constant-catch scenario
-is not implausible. Subsistence fisheries may use a constant-catch style
-policy over the short-term, as they seek to ensure that their food needs
-are met. More industrial fisheries may have pre-arranged agreements with
-buyers to deliver set amounts of fish. Constant-catch dynamics might
-also occur in fisheries with constraining quotas that are not updated
-after the implementation of MPAs. These patterns suggest that coupling
-MPA implementation with appropriate management reforms that reduce the
-effects of constant-catch strategies can move the predicted conservation
-effects of MPAs to being nearly uniformly positive. The possibility of
-constant-catch strategies to result in net MPA conservation losses
-highlights the critical importance of understanding the broader
-management context in which MPAs are placed.
-
-\begin{figure*}%[tbhp]
-  \centering
-  \includegraphics[width=.9\linewidth]{figs/conservation-effect.png}
-  \caption{Median (A) and range (B) of equilibrium regional MPA conservation effects (percent change in total biomass with MPAs relative to without MPAs) across a range of depletion and MPA sizes (and incorporating the full range of scenarios included in our study).
-"Range in MPA" is the percent of patches covered by an MPA, "Depletion" is the depletion that would have occurred in equilibrium without the MPA.
-Colors in panel B represent the percentage of simulations that fell into an effect size bin at each x-axis bin, such that the sum of all percentages for a given x-axis bin is 1.}
-  \label{conservation-effects}
-\end{figure*}
+total biomass densities of fish (averaging inside and outside of MPAs)
+relative to the total biomass densities of fish that would have occurred
+without the MPAs (acknowledging that other outcomes such as increased
+biodiversity or resiliency are also important to conservation but are
+beyond the scope of this analysis). Numerous factors can affect the
+regional effects of MPAs. These include the scale of adult and larval
+dispersal relative to the size of the MPAs (8, 26--29), larval dispersal
+patterns and the strength and timing of density dependence in the
+population (e.g.~pre- or post-settlement, \textbf{???}), how overfished
+the population would be without the MPA, and how fishing and management
+responds to the implementation of the MPAs ({\textbf{???}}, 4, 30--39).
+In addition, even for the same total area of MPAs, the location and
+spacing of the MPAs can have a profound influence on their cumulative
+impact through habitat and network effects (4, 40).
+
+Given this existing body of theory, what should we expect \emph{a
+priori} the effect size of the Channel Islands MPAs to be? While we know
+many parameters will affect the expected effects of the MPAs, theory
+suggests that life history, adult and larval dispersal relative to MPA
+size, and fishing pressure will be key drivers. The targeted species in
+our database span a range of life histories, but are largely made up of
+fishes in the genus XX (e.g.) and \textbf{Sebastes} (e.g.~), with a
+median Von Bertalanffy growth coefficient of XX, and a median age at
+maturity of XX (41). The MPAs in this study cover 21\% of the the
+surface waters of the Northern Channel Islands, a region spanning
+roughly 90km east to west. While detailed dispersal studies are not
+available for all of the species covered by our study, what information
+we have suggest that while adults of some of these species are likely to
+exhibit some site fidelity (42), larvae are likely distributed beyond
+the northern Channel Islands (43). Therefore, if we assume that the
+population of the targeted species is at least equal to the extent of
+the Northern Channel Islands, we can assume that at most the MPAs cover
+roughly 20\% of the targeted populations' ranges.
+
+Turning to the critical question of fishing pressure, formal stock
+assessments are largely lacking for these species. However, what
+evidence we have suggests mostly moderate fishing pressure, with some
+species such as California sheephead (\emph{Semicossyphus pulcher}) and
+blue rockfish (\emph{Sebastes mystinus}) experiencing high levels of
+fishing mortality / biomass levels below target levels during the early
+2000s (44, 45). We integrated these network design features, life
+history traits, and exploitation histories into a spatially explicit
+bio-economic simulation model, and used this model to generate expected
+outcomes for the Channel Islands MPAs. Our results suggest that while a
+wide range of outcomes are plausible, from 0\% to upwards of 200\%,
+across the majority of our simulations the effect size was 25\% or less,
+with a median simulated effect size of roughly 20\%; a potentially small
+value given the natural variability and sampling challenges of marine
+environments.
 
@@
 
-How can we empirically measure the regional effects of MPAs? Intuition
-and bio-economic theory tells us that the regional conservation effects
-of MPAs are likely to be highly context dependent across locations and
-across species within locations, so careful thought must be devoted to
-an empirical design to detect their effects. The perfect experiment
-would involve two parallel worlds that were identical, except for the
-implementation of an MPA. In world ``A'', no MPA would be implemented,
-and in the facsimile world ``B'', the MPA would be implemented. Both
-worlds would be tracked before and after MPA implementation, and the
-conservation and fishery outcomes would be compared after treatment.
-Instead of two parallel worlds, a similar experimental design would
-involve random placement of MPAs.
-
-Unfortunately, neither of these experimental designs has been
-implemented, because parallel worlds do not exist and MPAs are never
-placed at random. Careful causal inference requires some other means of
-controlling for biases introduced by factors such as the MPA siting
-process and biological spillover and concentration of fishing effort
-(41, 42).
-
-Conservation effects are typically estimated by comparing organism
-densities or biomass inside MPAs to densities or biomass in selected
-control sites outside MPAs, which we will refer to as response ratios.
-(5) and (43) present meta-analyses of hundreds of such studies. These
-results often find massively higher densities and biomasses inside MPAs
-than outside (5). While response ratios are intended as a measure of the
-effectiveness of reserves within their borders, these result stand in
-contrast to the smaller region wide effect sizes predicted by our
-simulation model. This raises the question: How reliable are density or
-biomass ratios as estimators of regional conservation effects?
-
-Control sites used in density or biomass ratios are often selected based
-on abiotic or ecological traits such as habitat characteristics (41).
-However, selection of control sites is further complicated by the very
-spillover that MPAs are often intended to create. Export of adults or
-larvae from the MPA to the ``control'' site affects their status as
-controls, as does displacement of fishing effort from MPAs to control
-sites. In theory, control sites far enough away to negate both
-biological spillover and concentration by the fishing fleet could be
-selected, but finding suitably far sites that are also appropriate
-proxies for the ecological and economic context of the MPAs may be
-challenging. While these concerns have been stated previously (44), the
-MPA evaluation literature has by and large been unable to adequately
-address them, often due to the very real challenges of acquiring the
-necessary data (41).
-
-Failure to account for these spillover effects can result in a biased
-estimate of the true effect of a policy such as MPA placement (42). In
-an ideal setting, the control site used in a response ratio would be a
-perfect counterfactual for what would have occurred without the MPA. We
-used our simulation model to approximate this scenario, calculating the
-response ratio as the ratio of densities inside the MPAs relative to the
-overall density from the paired simulation without MPAs multiplied by
-the area of the population in an MPA (to account for the fact that we
-would expect the effectiveness of a response ratio as a proxy for
-regional conservation effects to scale with the size of the MPA relative
-to the range of the species, Fig.\ref{density-ratio}-A). In this
-idealized case, the response ratio is often a reasonable estimator for
-the regional conservation effect.
-
-What happens when the ``control'' sites are in fact affected by the
-application of the MPA through biological spillover and fishing
-concentration? In order to address this possibility, we calculated a new
-response ratio as the mean density inside MPAs relative to the mean
-density outside the MPAs (both weighted by distance from MPA borders and
-scaled by MPA size). We only include simulations in which habitat and
-larval dispersal rates are identical inside and outside of MPAs, to
-approximate a scenario in which treatment and control sites have been
-paired by ecological characteristics. Under these circumstances the
-response ratio is a somewhat biased and very inaccurate estimator of the
-true regional effect of MPAs.
-
-While on average high response ratios correspond with high true regional
-effects, the variance is high, so the response ratio alone cannot be
-used as a reliable proxy for the true regional effect, without
-accounting for the sources of biases discussed here. For higher movement
-rates, response ratios were frequently near 0\% (potentially leading
-stakeholders to conclude that the MPA had been ineffective), when in
-fact in many cases the true effect of the MPA was highly positive. For
-highly sedentary species, extremely high response ratios could create
-the appearance of massive conservation gains when in fact the net effect
-on the regional population has been near zero or even negative
-(Fig.\ref{density-ratio}-B).
-
-We note that spatial before-after-control-impact (BACI) studies present
-a potential improvement over response ratios, controlling for
-differences in mean densities in the treated and control regions as well
-as MPA independent trends, but are much rarer due to the need for
-extensive pre-MPA monitoring. However, spatial BACI still requires that
-the MPAs do not affect the control sites, and as such has many of the
-same limitations as response ratios as an estimator of regional
-conservation effects.
-
-\begin{figure*}%[tbhp]
-  \centering
-  \includegraphics[width=.9\linewidth]{figs/density-ratio.png}
-  \caption{Simulated MPA-size weighted response ratios (x-axis) plotted against true regional change in biomass caused by MPAs (y-axis).
-Each point represents a simulation, and color represents the movement rate of fish in the simulation (expressed as a percentage of the patches a fish is likely to move in one time step).
-The black line shows the one:one fit line, the blue line is a linear fit of the relationship between observed response ratios and true regional change in biomass.
-Inside vs.
-Experiment simulates a true control, while Inside Vs.
-Outside uses a control that may be affected by the treatment.}
-  \label{density-ratio}
-\end{figure*}
-
-\hypertarget{estimating}{%
-\subsection*{MPA Effects in the Channel Islands}\label{estimating}}
-\addcontentsline{toc}{subsection}{MPA Effects in the Channel Islands}
-
-\begin{figure}%[tbhp]
-  \centering
-  \includegraphics[width=1\linewidth]{figs/channel-islands.pdf}
-  \caption{Map of PISCO sampling locations within the the Channel Islands National Marine Sanctuary, California, USA.
-Grey boxes indicated MPAs, points sampling locations, with color representing whether the site is inside an MPA (blue) or not (red)}
-  \label{channel-islands}
-\end{figure}
-
-Given these challenges, what is a viable alternative for estimating the
-conservation effects of MPAs? In a manner similar to (38), we use a
-difference-in-difference estimator comparing densities in finfish
+Having established a theoretical grounding for the likely biomass
+density outcomes of the Channel Island MPAs, what empirical evidence do
+we have of the effects? \textbf{???} examined changes in biomass
+densities of species targeted by fishing activity inside and outside of
+the MPAs over time (a metric generally termed a ``response ratio''), and
+compared these changes to the trends in non-targeted species. They found
+a statistically significant increase in the response ratios of targeted
+species over time, and evidence that this increase is smaller in the
+non-targeted species. Updating the results of (46) through 2017 with a
+Bayesian response-ratio regression shows a continuation of the
+increasing trend in the response ratios of targeted species
+(Fig.@ref(fig:XX)).
+
+Why do these response ratio results not serve as sufficient evidence for
+the regional effects of the Channel Island MPAs? Control sites used in
+calculating response ratios are often selected based on abiotic or
+ecological traits such as habitat characteristics (47). However,
+selection of control sites is further complicated by the very spillover
+that MPAs are often intended to create. Export of adults or larvae from
+the MPA to the ``control'' site affects their status as controls, as
+does displacement of fishing effort from MPAs to control sites. In
+theory, control sites far enough away to negate both biological
+spillover and concentration by the fishing fleet could be selected, but
+finding suitably far sites that are also appropriate proxies for the
+ecological and economic context of the MPAs may be challenging. While
+these concerns have been stated previously (48), the MPA evaluation
+literature has by and large been unable to adequately address them,
+often due to the very real logistical challenges of identifying and
+sampling adequate control sites that are truly independent of the MPAs
+in question. As a result of these challenges, response ratios can be a
+highly imprecise and biased measure of the regional conservation effect
+of an MPA network (47).
+
+In the case of the Channel Islands MPAs, control sites are often located
+within only a few kilometers of an MPA, suggesting that they are
+susceptible to biological spillover and concentration of fishing effort
+resulting from the MPAs. Given these complications, we can certainly
+interpret the response ratios reported in (46) and updated here as
+evidence that the MPAs are providing effective protection of targeted
+fish biomass inside their borders, particularly in light of the general
+increase in response ratios over time. However, it is unclear how well
+these response ratio results serve as in indicator of overall regional
+effects of the MPAs.
+
+To illustrate this problem, we used a bio-economic model to simulate
+response ratio trajectories for species and MPA coverage representative
+of the northern Channel Islands (see methods and SI for details of model
+structure). These simulations cover the life-histories and MPA sizes
+seen in the Channel Islands, but vary in key unknowns such as the actual
+degree of fishing pressure, the timing of density dependence, and the
+fleet responses to the MPAs. For each year of protection, we paired our
+simulated response ratios to the estimated posterior probability
+distributions of the Channel Island response ratios. For example, if the
+mean estimated response ratio in the year 2006 is one, we found all
+simulations with three years of protection that had simulated response
+ratios near one, and then pulled out the ``true'' simulated regional MPA
+effect from each of those simulations. This provides us with a
+distribution of simulated regional MPA effects that could plausibly
+generate the types of response ratios actually observed in the Channel
+Islands.
+
+As reported in (46), we see evidence for an increasing trend in response
+ratios of targeted species over time (Fig.@ref(fig:rr-plot)). This
+provides strong evidence that the MPAs are providing protection for
+targeted species within their borders. However, our simulation results
+show that the response ratio trends we observe in the data could
+plausibly be produced by a wide range of regional MPA effects
+(Fig.@ref(fig:rr-plot)). Response ratios well over one were associated
+with regional MPA effects generally less than 25\%, and many simulations
+produced large response ratios but regional MPA effects close to 0\%
+This can occur if for example fishing pressure is only moderate, adult
+movement is low, larval dispersal is high, and displaced fishing effort
+concentrates around the border of the MPAs.
+
+\textbackslash begin\{figure\emph{\}\%{[}tbhp{]} \centering
+\includegraphics[width=.9\linewidth]{figs/response_ratio_plot.png}
+\textbackslash caption\{90\% Posterior probability distributions of
+response ratios for targeted species (x-axis) over time (y-axis) in
+blue. Simulated MPA effects associated with simulated response ratios
+matched to empirical response ratios in red. For response ratios, a
+value of zero indicates that biomass densities of targeted and
+non-targeted species are identical inside and outside MPAs, a value of 1
+indicates that biomass densities of targeted species are 100\% greater
+inside MPAs than outside. For MPA effect, a value of zero indicates that
+mean biomass densities are identical in the with- and without- MPA
+scenarios. A value of 1 indicates that mean biomass densities are 100\%
+greater in the scenario with MPAs than the scenario without MPAs.\}
+\label{rr-plot} \textbackslash end\{figure}\}
+
+Given the potential unreliability of response ratios for the task, in an
+ideal world how would we empirically measure the regional effects of
+MPAs? The perfect experiment would involve two parallel worlds that were
+identical, except for the implementation of an MPA. In world ``A'', no
+MPA would be implemented, and in the facsimile world ``B'', the MPA
+would be implemented. Both worlds would be tracked before and after MPA
+implementation, and the biomass densities would be compared after
+treatment. Instead of two parallel worlds, a similar experimental design
+would involve random placement of MPAs. Unfortunately, neither of these
+experimental designs has been implemented, because parallel worlds do
+not exist and MPAs are to our knowledge never placed at random. Careful
+causal inference requires some other means of controlling for biases
+introduced by factors such as the unobserved environmental shocks, the
+MPA siting process, biological spillover, and concentration of fishing
+effort (47, 49).
+
+Building off of the concepts explored in (46), we propose an alternative
+identification strategy utilizing biomass densities of XX species that
+are not directly targeted by fishing as our control group
+(non-targeted), and biomass densities of XX species targeted by fishing
+as our treatment group. Targeted species in the Northern Channel
+Islands, California, include commercially important fin-fish such as
+California sheephead (\emph{Semicossyphus pulcher}), and copper
+(\emph{Sebastes caurinus}) and blue (\emph{Sebastes mystinus}) rockfish.
+Each of these targeted species was the subject of prior bio-economic
+modeling related to the effects of MPAs in southern California (50, 51).
+It should be noted though that our analysis omits species from important
+invertebrate fisheries including red urchin (\emph{Mesocentrotus
+franciscanus}) and spiny lobster (\emph{Panulirus interruptus}).
+Included non-targeted species include garibaldi (\emph{Hypsypops
+rubicundus}), halfmoons (\emph{Medialuna californiensis}), and
+blacksmith (\emph{Chromis punctipinnis}) (See Table.S2 for a complete
+list of species). We used a Bayesian difference-in-difference regression
+to estimate any difference in mean total biomass densities of fin-fish
 species targeted by fishing effort (i.e., those potentially affected by
-an MPA) from those not targeted before and after MPA implementation
-(42). Targeted species in the Northern Channel Islands, California,
-include commercially important finfish such as California sheephead
-(\emph{Semicossyphus pulcher}), and copper (\emph{Sebastes caurinus})
-and blue (\emph{Sebastes mystinus}) rockfish. Each of these targeted
-species was the subject of prior bio-economic modeling related to the
-effects of MPAs in southern California (45, 46). But, we must note that
-the analysis omits species from important invertebrate fisheries
-including red urchin (\emph{Mesocentrotus franciscanus}) and spiny
-lobster (\emph{Panulirus interruptus}). Non-targeted species include
-garibaldi (\emph{Hypsypops rubicundus}), halfmoons (\emph{Medialuna
-californiensis}), and blacksmith (\emph{Chromis punctipinnis}) (See
-Table.S2 for a complete list of species). The key assumptions of this
-approach are that a) within the time-frame of the model there are no
-significant interaction effects between the targeted and non-targeted
-species (which in fact we do not detect, see SI), and b) that in the
-absence of the MPAs both the targeted and non-targeted groups of species
-would have exhibited similar trends in densities. The advantage to this
-approach is that given the time-frame of the model (16 years), we
-believe that spillover effects of MPA placement on the non-targeted
-species are likely to be much less severe than the effects of biological
-spillover and fishing concentration that may bias the performance of
-estimators such as a response ratios or a spatial BACI design.
-
-We applied this difference-in-difference strategy to a large dataset in
-California. We used empirical kelp forest survey data from the
-Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO)
-monitoring in the Northern Channel Islands with the ultimate goal of
-testing the regional effects of MPAs in a real world context. A network
-of MPAs covering approximately 20\% of the islands' waters was put in
-place in 2003 as part of the California Marine Life Protection Act
-(MLPA) (see (47), (48), (49), and (50) for information on the creation
-of the MLPA). PISCO conducts visual SCUBA surveys at a large number of
-rocky-reef and kelp forest sites inside and outside of MPAs throughout
-the Channel Islands, producing estimates of densities of fishes that are
-both targeted and non-targeted by fishing
-(Fig.\ref{channel-islands}-\ref{pop-trends}).
-
-Although the Islands do not encompass the full population of any of
-these species, they encompass a large enough area to assess the regional
-MPA conservation effects we might see in the Channel Islands after 13
-years of protection (the time span covered by our model). The median
-regional conservation effect produced by our simulation model after 13
-years of protection by MPAs covering between 15\%-20\% of a population's
-range was 25\%. This projected effect has an an interquartile range of
-8\% and 100\%, although values as high as a 200\% increase and as low as
-a 50\% decrease were also produced. Our expectation then is that all
-else being equal we are likely looking for a modest but positive effect
-size.
-
-\begin{figure}%[tbhp]
-  \centering
-  \includegraphics[width=1\linewidth]{figs/pop-trends.pdf}
-  \caption{Centered and scaled mean annual biomass densities of included targeted (red) and non-targeted (blue) fish species (faded lines) and smoothed means with 95\% confidence intervals over time (darker lines and ribbons).
-Vertical red dashed line shows year of MPA implementation.}
-  \label{pop-trends}
-\end{figure}
-
-The raw trends in the biomass densities appear to show parallel trends
-in the densities of targeted and non-targeted fishes pre-MPA, with the
-targeted group diverging first positively and then negatively over time
-(Fig.\ref{pop-trends}). However, these raw trends could be caused by any
-number of confounding variables besides the effects of the MPAs
-themselves, from changes in observer skill and sampling locations, to
-shifts in environmental conditions. Using our difference-in-difference
-model in an effort to control for these factors, we found no significant
-difference in the biomass densities of targeted and non-targeted fishes
-pre-MPA (before 2003). Following the implementation of the MPAs in 2003,
-we see evidence of an increasing trend in densities of targeted relative
-to non-targeted fish, reaching a peak value in 2014. After 2014,
-densities of targeted species appear to have declined relative to what
-we would expect based on the non-targeted group (Fig.\ref{did-plot}).
-
-\begin{figure}%[tbhp]
-  \centering
-  \includegraphics[width=1\linewidth]{figs/did_plot.png}
-  \caption{Estimated divergence in biomass densities of targeted and non-targeted fishes throughout the Channel Islands (integrated across inside and outside of MPAs).
-MPAs are implemented in 2003 (red dashed line).
-Colored cells in the background indicate the frequency of MPA conservation effects predicted by our simulation modeling for MPAs similar in design and species composition to the Channel Islands in each year.}
-  \label{did-plot}
-\end{figure}
+an MPA) and those species not targeted by fishing before and after MPA
+implementation (49). The result of this regression is, conditional on
+the assumptions of the model, an estimate of the effect of the MPAs on
+the mean total biomass densities of targeted species throughout the
+Channel Islands.
+
+As conventional wisdom and theory would suggest, over the first three
+years of implementation (2003-2006), the effects of the MPAs are
+unclear, with support for a small negative effect to a substantially
+positive effect (with much higher probability of a small positive
+effect). Over the next six years the model estimates greater
+probabilities of an increasingly positive MPA effect, peaking in
+2009-2011 with a median estimate of MPA effect of a xx increase in mean
+total biomass of targeted species (90\% credible interval 37\% - 128\%.
+These empirical estimates are in line with (though in the upper range
+of) the outcomes that our simulation model suggests are plausible given
+the rough characteristics of the northern Channel Island MPAs. However,
+in the subsequent years the trend reverses itself, and for the years
+2015-2017 we once again see no clear effect of the MPAs
+Fig.@ref(fig:mpa-effect)
+
+\textbackslash begin\{figure\emph{\}\%{[}tbhp{]} \centering
+\includegraphics[width=.9\linewidth]{/Users/danovan/projects/regional-effects-of-mpas/documents/figs/mpa_effect_plot.png}
+\textbackslash caption\{Results of difference-in-difference regression
+estimating the regional effect of the northern Channel Island MPAs on
+mean total biomass densities of targeted species (difference in mean
+total biomass density of targeted species over time relative to expected
+levels using non-targeted species as a control). Grey distributions show
+90\%XX posterior probability distribution of estimated MPA effect; red
+point is median estimated effect, thicker red section 50\% credible
+interval, thinner red line 90\% credible interval). Blue distributions
+in background show range of MPA effects produced by simulation model
+tuned to reflect the dynamics of the Northern Channel Island MPAs (black
+dashed line is median simulated value). Results are estimated in three
+blocks, including years greater than or equal to left-hand value and
+less than right-hand value.\} \label{mpa-effect}
+\textbackslash end\{figure}\}
 
@@
 \addcontentsline{toc}{subsection}{Discussion}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Why are the results from the channel islands important and what do
+  they tell us
+\item
+  Put the channel islands results in context -
+\end{enumerate}
+
+Ideally then we want a control for broad environmental shocks to the
+region that is independent of factors such as selection bias, biological
+spillover, and fishing concentration.
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  multiple lines of evidence
+\item
+  Theory before empirics
+\item
+  Patience
+\end{enumerate}
+
+The network of MPAs implemented in the Northern Channel Islands in 2003
+provides an ideal opportunity to examine the regional effects of
+protected area networks. This analysis is to our knowledge the first
+empirical estimate of the regional effects of a large and iconic MPA
+network on a broad assemblage of targeted fin-fish. Despite the well
+enforced nature of these MPAs {[}xx{]}, and the presence of a rigorous
+scientific monitoring program, we are unable to estimate a clear effect
+of protection. While response ratios have targeted species have
+increased over time, providing evidence of effective \emph{within MPA}
+protection, these response ratios are on their own an unreliable
+indicator of the regional conservation effects of the MPAs. Our
+difference-in-difference regression estimates a high probability of
+increasingly positive MPA effects over the first nine years of
+protection, only to see these estimated gains reversed from 2012 to
+2017. For the most recent time period, we can detect no clear positive
+or negative regional effect of the Channel Island MPAs on biomass
+densities of targeted fin-fish (Fig.@ref(fig:mpa-effect)).
+
+How can we explain these results? The estimated effects from 2003 to
+2014 are well in line with theoretical expectations given the species
+and network in question. How might we explain the most recent results
+though? One explanation may lie in fleet dynamics. Much of the
+theoretical literature on MPAs assumes that all else being equal bigger
+reserves produce bigger conservation gains ({\textbf{???}}). However, to
+our knowledge all of these models simulate fleet dynamics through
+assumptions about fishing mortality rates (e.g.~concentration of fishing
+mortality (48)). The assumption of these models is that fishers
+determine an amount off effort to exert, and distribute that effort
+outside the MPA in response to some function.
+
+An alternative and to our knowledge unexplored (in the context of MPAs)
+fleet model though is a ``constant-catch'' strategy. Under this model,
+fishers have a catch objective, and exert as much (or little) effort as
+needed to achieve that objective. While a constant-catch greater than
+MSY is not possible over the long-term under the assumptions of our
+model, over the short-term a constant-catch scenario is not implausible.
+Subsistence fisheries may use a constant-catch style policy over the
+short-term, as they seek to ensure that their food needs are met. More
+industrial fisheries may have pre-arranged agreements with buyers to
+deliver set amounts of fish. Constant-catch dynamics might also occur in
+fisheries with constraining quotas that are not updated after the
+implementation of MPAs. Interestingly, when we simulated the effects of
+MPAs under different fleet models, the ``constant-catch'' scenario stood
+out as the only way for MPAs to actually produce a net conservation loss
+(Fig.@ref(fig:XX)). While open-access fishing strategies can result in
+``scorched earth'' scenarios where the only fish left are found inside
+the reserve, across 94\% simulations the net effect of the reserves was
+still positive. Under a constant-catch scenario though, fishers have to
+fish much harder than before to get the same catch from a smaller part
+of the population, reducing the size structure of the population and
+subsequently causing net conservation loss under 70\% of our
+constant-catch simulations.
+
+While we do not have access to fine scale fishing data from the Northern
+Channel Islands alone, reported catches for the species of interest in
+the Santa Barbara region in fact exhibit an overall downward trend in
+the years post reserve (see SI, XX). We can most likely rule a negative
+MPA effect caused by a constant-catch fishing strategy then. What then
+is another explanation for the recent downward trend in the estimated
+MPA effects? We do not see evidence to reject the parallel trends
+assumption of our difference-in-difference model in initial years of the
+MPAs, and we do not detect any clear evidence of interaction between the
+targeted and non-targeted species (e.g.~depressed levels of non-targeted
+species as a result of predation from increased biomass densities of
+targeted species, see Fig.SXX).
+
+However, the Santa Barbara Channel experience a massive heatwave in XX.
+The Channel Islands represent the warmest edge of the distribution of
+many of the targeted species in our study. Conversely, many of the
+represented non-targeted species are sub-tropical. As a result, we
+hypothesize that the blob even of XX broke down the parallel trends
+assumption between the targeted and non-targeted species in our study,
+as the targeted species were more negatively affected by the warming
+temperatures. This is supported by the fact that we observe similar
+declines in targeted species both outside and inside the reserves
+(Fig.XX). Assuming that the reserves are sufficiently well enforced and
+large to protect targeted species biomass (an assumption supported by
+the response ratio results), if the cause of recent declines was due to
+increases in fishing pressure we would expect to see substantial
+declines only in the fished areas (Fig. XX).
+
+\includegraphics{ovando-regional-effects-of-mpas-v2_files/figure-latex/unnamed-chunk-5-1.pdf}
+
+\begin{figure*}%[tbhp]
+  \centering
+  \includegraphics[width=.9\linewidth]{/Users/danovan/projects/regional-effects-of-mpas/documents/figs/raw_biomass_density_plot.png}
+  \caption{Raw Biomass Trends}
+  \label{mpa-effect}
+\end{figure*}
+
+\hypertarget{when-can-we-detect-the-effects-of-mpas}{%
+\subsubsection{When Can We Detect the Effects of
+MPAs?}\label{when-can-we-detect-the-effects-of-mpas}}
+
+Containing a carefully designed, well-enforced, and well-studied MPA
+network, The Channel Islands would seem at face value to be an ideal
+location to study the regional conservation effects of protected areas.
+The persistently high response ratios suggest that despite overall
+decreases in targeted biomass densities inside and outside MPAs, the
+MPAs may still be providing protection within their borders. But, as we
+have shown here these response ratios are not necessarily a sign of the
+magnitude of any regional effects. The difference-in-difference strategy
+utilized here presents an alternative strategy, that while not without
+its own strict caveats presents some potential improvements over
+response-ratios as a means of estimating regional conservation effects.
+While we estimate a highly uncertain but overall positive effect at
+first, we are unable to detect a robust signal from 2012-2017. After 14
+years of MPA protection we are left without a clear picture of the
+effect of the Channel Island MPA network on biomass densities of
+targeted fin-fish species.
+
+Our simulation results suggest though that we should not be surprised by
+this result. The Channel Islands MPAs cover approximately 20\% of the
+waters in the Channel Islands, and while formal stock assessments are
+not available for many of the targeted species in our analysis, what
+evidence we have suggests that, as a group, these fish are on average
+not heavily overfished. Some species, such as California sheephead and
+blue rockfish were likely below target levels during the period (44,
+45), but projections based upon the overall average response across all
+species will likely suggest modest benefits even if a subset of species
+could experience much larger population gains. Harkening back to our
+simulations, we expect the average percentage difference in densities of
+targeted species with and without MPAs to be modest
+(Fig.\ref{fig:did-plot}). Effects of this size are likely to be
+challenging to detect empirically given the large natural variation of
+marine ecosystems (especially temperate reefs) and the observation error
+inherent in infrequent annual survey programs such as those provided by
+PISCO.
+
+As an additional complication, our simulations difference-in-difference
+model examine percentage changes in biomass densities. As an extreme
+example, an increase in biomass densities from .2kg/m\textsuperscript{2}
+to .4kg/m\textsuperscript{2} would be produce as a 100\% increase. While
+a large percentage increase, this is a small change in absolute biomass
+densities relative to the variance in the observation process itself
+(see Fig.S1 for a companion to Fig.\ref{conservation-effects} scaled by
+absolute population size). Beyond these challenges, the median estimated
+age at sexual maturity for the targeted species included in this study
+is 6 years, meaning that the span of this analysis represents less than
+three generations of MPA protection for half of the the measured
+species. Ongoing monitoring may yet reveal clearer effects. Analysis of
+more rapidly growing and maturing species, e.g.~spiny lobster, may also
+reveal clearer signals.
+
+Given the natural variability of marine ecosystems, and the strong
+challenges of obtaining accurate samples from oceanic environments, how
+large of an effect would an MPA network have to allow a
+difference-in-difference strategy such as this to be a reliable measure
+of MPA effects? To provide guidance on this important question, we used
+our bio-economic model to simulate data from a range of scenarios with
+increasing MPA effect size, along with increasing degrees of observation
+error and natural recruitment variation. As an added measure, we include
+scenarios in which the sampled species go through recruitment regimes,
+which may be positive for both targeted and non-targeted species, or
+positive for non-targeted species and negative for targeted species. We
+then used a simple Bayesian difference-in-difference regression on these
+simulated data, and estimated the percent error between the estimated
+and true MPA effect.
+
+While unbiased across simulations, the difference-in-difference model
+struggled severely when MPA effect sizes were less than 25\% and the
+model was faced with observation and process errors
+(Fig.@ref(fig:val-plot)). Even larger effect sizes commonly misestiamted
+the true MPA effect by 50\% or more. Obtaining a mean absolute percent
+error (MAPE) of 25\% or less across our simulated datasets required an
+MPA effect of at least 30\%. This is merely an illustrative exercise,
+omitting clearly critical factors such as detection probability and
+sampling strategy. However, since nearly any omission which one can
+think of would make an MPA effect harder to detect, not easier, these
+results serve as a useful floor for the likely difficulty in estimating
+MPA effects. In the context of the Channel Islands, given the potential
+effect size produced by out simulation model these results suggest that
+we might expect to be unable to detect a clear effect. XX need to think
+about whether MAPE is the correct error here - maybe need something like
+the variance of the estimator?
+
+\begin{figure*}%[tbhp]
+  \centering
+  \includegraphics[width=.9\linewidth]{/Users/danovan/projects/regional-effects-of-mpas/documents/figs/validation_plot.png}
+  \caption{Validation}
+  \label{val-plot}
+\end{figure*}
+
+This finding begs an important question: when should we expect to see
+MPA effects big enough to stand a reasonable chance of detection? To
+address this, we simulated 9252 MPA scenarios across a wide range of
+life histories, network designs, and fishing dynamics (see SI for a full
+description of scenarios). Using these simulations, we can consider what
+we might expect the regional effects of an MPA network to be under a
+range of plausible circumstances.
+
+Suppose that we are willing to tolerate a MAPE of 25\%. Two of the most
+critical drivers of MPA performance are the size of the MPA and the
+degree of fishing pressure. Looking across these two variables, our
+simulations suggest if the MPA network covers 25\% or more of a species
+range and/or pre-MPA depletion is greater 60\% we might expect an effect
+size with a reasonable chance of detection. While recently some
+extremely large MPAs have been enacted that may indeed reach into the
+higher levels of MPA coverage, for near-shore commercial fin-fish many
+MPA networks are likely to cover areas more in line with the Channel
+Islands (20\%) (Fig.@ref(fig:XX)-A).
+
+These initial simulation results would seem to suggest some rather
+simple rules of thumb: put an MPA of sufficient size on a at least
+somewhat overexploited population and one can expect large and
+potentially detectable results. However, as we discussed earlier, the
+MPA literature highlights a large number of variables beyond simply size
+and fishing pressure that can affect performance. To address this, we
+can examine the variability in expected MPA effects across each of the
+two major axes (pre-MPA depletion and MPA size) (Fig.@ref(fig:XX)-B). As
+both MPA size and pre-MPA depletion increase, the potential for a large
+MPA effect increases. However, we see that for both variables, even at
+extremely large or extremely small values a wide range of MPA effects
+were possible, though as we might expect this effect of pre-MPA
+depletion, in other words fishing pressure were much clearer: it is hard
+for an MPA to have a massive effect when the stock is not heavily
+fished.
+
+To put these results into further context, the FAO estimates that 7\% of
+the worlds fisheries with status estimates fall into the relatively
+unexploited category (roughly depletion less than 50\%), 60\% fall into
+the fully fished category (roughly depletion 50\%-70\%), and 33\% fall
+into the heavily fished category (roughly depletion greater than 70\%)
+(52), though works that include a broader range of fisheries estimate
+that 50\% or more of stocks to fall into the heavily fished category
+(53, 54). Taken together then, the regional effects of MPA networks
+covering 25\% or less of a species range may be difficult to detect in
+many placed with already well-managed fisheries, while for that size we
+might expect clearer effects in less-managed locations (though that of
+course ignores the complication of compliance with MPA regulations).
+Within these broad guidelines though a wide range of outcomes are
+possible: as a starting place users can use the bio-economic MPA
+simulation model developed for this paper to explore potential outcomes
+for specific MPAs using an interactive web application available at
+\href{https://danovando.shinyapps.io/simmpa/}{danovando.shinyapps.io/simmpa}.
+
+\begin{figure*}%[tbhp]
+  \centering
+  \includegraphics[width=.9\linewidth]{/Users/danovan/projects/regional-effects-of-mpas/documents/figs/expected_mpa_effect.png}
+  \caption{Expectation}
+  \label{expect-plot}
+\end{figure*}
+
+\hypertarget{conclusion}{%
+\subsection*{Conclusions}\label{conclusion}}
+\addcontentsline{toc}{subsection}{Conclusions}
 
@@
 regional conservation benefits of MPAs are highly context dependent and
-in many circumstances, are likely to be so small that they are nearly
-impossible to detect empirically. Indeed, this is exactly what we found
-in our empirical case study from the Channel Islands, California, USA.
-
-We find regional conservation gains from MPAs in 94\% of our simulations
-(Fig.\ref{conservation-effects}). MPAs covering less than 25\% of the
-population range were unlikely to produce regional conservation gains
-above 10\% unless the stock would have been severely overfished without
-the MPA. Given that few marine species on the planet have more than 25\%
-of their entire range protected in MPAs, this suggests that large
-regional gains are unlikely to be common unless the overall extent of
-MPAs grows globally. Large MPAs protecting highly depleted stocks almost
-always produced large regional conservation gains, as should be
-expected.
-
-The median conservation effect of 36\% stands in contrast to the large
-within-MPA effects on biomass densities reported by (51) and (5). This
-difference is due in part to the fact that within-MPA effects are likely
-to be much larger than regional-MPA effects, but may also reflect the
-many challenges of translating response ratios into regional
-conservation effects. Our simulation analysis shows that without proper
-controls (in particular for the effects of fleet redistribution),
-response ratios well over 100\% are entirely plausible even when the
-actual conservation effect is much lower (and \emph{vice versa}).
-
-We were not able to detect a statistically significant difference in the
-densities of targeted and non-targeted finfish species over the 13 years
-of MPA protection in the Channel Islands covered by our analysis. The
-data (and model assumptions) provide support for both negative and
-positive MPA effects, although the size of these estimated effects is
-well within the range that our simulation analysis suggests are
-plausible. At face value, this lack of a clear result may seem
-surprising given the size and carefully studied nature of this MPA
-network.
-
-Our simulation analysis suggests that this result should perhaps be
-expected. The Channel Islands MPAs cover approximately 20\% of the
-waters in the Channel Islands, and while formal stock assessments are
-not available for many of the targeted species in our analysis, what
-evidence we have suggests that, as a group, these fish are on average
-not heavily overfished. Some species, such as California sheephead and
-blue rockfish were likely below target levels during the period (52,
-53), but projections based upon the overall average response across all
-species will likely suggest modest benefits even if a subset of species
-could experience much larger population gains. Harkening back to our
-simulations, we expect the average percentage difference in densities of
-targeted species with and without MPAs to be modest
-(Fig.\ref{did-plot}). Effects of this size are likely to be challenging
-to detect empirically given the large natural variation of marine
-ecosystems (especially temperate reefs) and the observation error
-inherent in infrequent annual survey programs such as those provided by
-PISCO.
-
-In addition, our simulations examine percentage increases in biomass
-densities. As an extreme example, an increase in biomass densities from
-2kg/m\textsuperscript{2} to 4kg/m\textsuperscript{2} would be reported
-as a 100\% increase. While this chance produces a large percentage
-increase, this is a small change in absolute biomass densities relative
-to the variance in the observation process itself (see Fig.S1 for a
-companion to Fig.\ref{conservation-effects} scaled by absolute
-population size). Beyond these challenges, the median estimated age at
-sexual maturity for the targeted species included in this study is 5
-years, meaning that the span of this analysis represents less than three
-generations of MPA protection for half of the the measured species.
-Ongoing monitoring may yet reveal clearer effects. Analysis of more
-rapidly growing and maturing species, e.g.~spiny lobster, may also
-reveal clearer signals.
-
-While the MPA effect estimated by our model was not significant in any
-one year, the positive trend from 2005 to 2014 is evident, as is the
-sharp decline from 2015-2017. To what should we attribute this apparent
-shift? We cannot reject the parallel trends assumption between the
-targeted and non-targeted species in the years before the MPAs
-(Fig.\ref{did-plot}). But, the Channel Islands region (and the entire
-West coast of the USA) experienced a dramatic `marine heatwave'
-beginning in 2014 and persisting through 2016, resulting in part in
-extremely elevated water temperatures throughout the region (54).
-Biogeographic differences in the distributions of targeted and
-non-targeted species may confound the observed effects of MPAs. Many of
-the non-targeted species have warm thermal affinities and have increased
-in numbers since the heatwave (55). The targeted group is made up mostly
-of fishes with cold-water affinities, such as members of the genus
-\emph{Sebastes}. As such, we hypothesize that the recent evidence for a
-decline in densities of targeted species is due to environmental
-conditions that disproportionately affect the targeted group (and not
-for example due to concentrated fishing pressure outside the reserves).
-This hypothesis is supported by the sharp declines in densities of
-targeted species seen inside the MPAs themselves (see Fig.S27-S29),
-suggesting a driver other than fishing may be at play.
+in many circumstances, are likely to be small enough that they are
+nearly impossible to detect empirically. Indeed, this is exactly what we
+found in our empirical case study from the Channel Islands, California,
+USA.
 
@@
 users evaluate whether it is achieving their objectives? Response ratios
-are commonly used as evidence for conservation outcomes of MPAs (5), but
-as suggested in (41) and (42) and demonstrated here, without careful
-attention to the design of control sites (accounting for example for the
-displacement of effort by MPAs), response ratios may be highly
-unreliable estimators of regional MPA effects. As (42) suggests, there
-are many potential alternatives for estimating the effects of MPAs that
-better account for the challenges of causal inference. We applied one
-such approach here (a difference-in-difference estimator), and yet were
-still unable to reach robust conclusions as to the effect of MPAs on the
-density of targeted species in the Channel Islands, due to the likely
-small size of the true effect relative to the strength and variability
-of environmental drivers.
+are commonly used as evidence for conservation outcomes of MPAs; (5) and
+(55) present meta-analyses of hundreds of such studies. These results
+often find massively higher densities and biomass inside MPAs than
+outside (5). But as suggested in (47) and (49) and further demonstrated
+here, without careful attention to the design of control sites
+(accounting for example for the displacement of fishing effort by MPAs),
+response ratios may be highly unreliable estimators of regional MPA
+effects. When MPAs affect nearby control sites used in response ratios
+through biological spillover or concentration of fishing effort, it is
+entirely possible to for MPA to produce massive response ratios while
+simultaneously having minimal effects on the actual populations
+partially protected by the MPAs. As (49) suggests, there are many
+potential alternatives for estimating the effects of MPAs that better
+account for the challenges of causal inference (though that may be more
+data-intensive). We applied one such approach here (a
+difference-in-difference estimator), and yet were still unable to reach
+robust conclusions as to the effect of MPAs on the density of targeted
+species in the Channel Islands, due to the likely small size of the true
+effect relative to the strength and variability of environmental
+drivers.
 
 While this does not mean that all MPAs will face similar challenges in
-estimating their effects, our results in the rigorously studied Channel
-Islands system make clear that in many instances empirically detecting a
-clear regional effect of MPAs may not be possible. How then should
-stakeholders go about adaptively monitoring and managing MPAs?
-Simulation modeling can help inform the range of effect sizes that may
-be expected, and monitoring programs can perhaps be tuned to focus on
-the species groups that have the highest chance of a detectable effect
-size (39). Expanding data collection to include robust monitoring of
-fleet dynamics may help statistical approaches to isolate the true
-effect of MPAs on conservation outcomes by allowing for improved control
-of the effects of fishing dynamics, and allow managers to take into
-account potential negative interactions between MPAs and fleet dynamics
-such as those that may occur under constant-catch dynamics. Whenever
-possible monitoring programs should be implemented prior to MPA
-implementation to provide a pre-treatment benchmark. Non-equilibrium
-analyses (Fig.\ref{did-plot}) also help set expectations for effect
-sizes over time. Beyond that, educating communities about the challenges
+estimating their effects, our results in the rigorously studied and
+iconic Channel Islands Marine Protected Area network make clear that in
+many instances empirically detecting a clear regional effect of MPAs may
+not be possible. How then should stakeholders go about adaptively
+monitoring and managing MPAs? Simulation modeling can help inform the
+range of effect sizes that may be expected, and monitoring programs can
+perhaps be tuned to focus on the species groups that have the highest
+chance of a detectable effect size (56). Expanding data collection to
+include robust monitoring of spatio-temporal fleet dynamics may help
+assess the validity of control sites used in response ratios, support
+the direct inclusion of these fleet dynamics into statistical models,
+and allow managers to take into account potential negative interactions
+between MPAs and fleet dynamics such as those that may occur under
+constant-catch dynamics. Whenever possible monitoring programs should be
+implemented prior to MPA implementation to provide a pre-treatment
+benchmark. Non-equilibrium analyses also help set expectations for
+effect sizes over time (56). Educating communities about the challenges
 of estimating the effects of MPAs can help set expectations, so that a
 lack of a clear effect is not necessarily viewed as a failure of the
-program, but rather considered in the context of reasonable
-expectations. While this paper has focused on the conservation outcomes
-of MPAs, future work must also address the challenge of predicting and
+program, or a cause for premature adaptation of the network, but rather
+considered in the context of reasonable expectations for the MPAs in
+question. While this paper has focused on the conservation outcomes of
+MPAs, future work must also address the challenge of predicting and
 estimating the fishery impacts of protected areas.
@@
 must directly tackle the challenge of evaluating the performance of
-these MPAs, and adapting their design as needed to best achieve
-objectives. Commonly employed metrics such as response ratios may be
-applicable in some circumstances, but can have severe shortcomings as
-metrics of regional conservation effects. Dependence on unreliable
-estimators of MPA effects may lead to stakeholders incorrectly
-attributing negative environmental shocks as MPA failures, or
-interpreting data arising from scorched-earth fishing outside MPAs as a
-conservation success. Bio-economic modeling can help frame community
+these MPAs at the regional scale, and adapting their design as needed to
+best achieve objectives. Commonly employed metrics such as response
+ratios may be applicable in some circumstances, but can have severe
+shortcomings as metrics of regional conservation effects. Dependence on
+unreliable estimators of MPA effects may lead to stakeholders
+incorrectly attributing negative environmental shocks as MPA failures,
+or interpreting data arising from scorched-earth fishing outside MPAs as
+a conservation success. Bio-economic modeling can help frame community
 expectations, reducing the potential for a reduction in support if
-unrealistic conservation or fishery gains are not realized. Statistical
-approaches that explicitly address complications such as the spatial
-spillover effects of MPAs (such as the difference-in-difference approach
-used here) may give users an improved understanding of the performance
-of their MPAs. Effective use of simulation and statistics to clearly
-communicate what we should expect and what we can detect from MPAs is
-critical in ensuring that MPAs play effective roles in fisheries
-management and marine conservation.
+unrealistic conservation or fishery expectations are not realized.
+Statistical approaches that explicitly address complications such as the
+spatial spillover effects of MPAs (such as the difference-in-difference
+approach used here) may give users an improved understanding of the
+performance of their MPAs, but even they may struggle when expected
+effect sizes are small. Clearly communicating what we should expect, and
+what we can detect,m from MPAs is critical in ensuring that MPAs play
+effective roles in fisheries management and marine conservation.
 
@@
 regression approach. Further details can be found in the Supplementary
-Information. All analysis were conducted in R (56). Our main
-difference-in-difference model was fit using Template Model Builder in R
-(57). Code needed to replicate results can be found
+Information. All analysis were conducted in R (57). Our main
+difference-in-difference model was fit using Stan (58) using the
+\texttt{rstanarm} package ({\textbf{???}}). Code needed to replicate
+results can be found
 \href{https://github.com/DanOvando/regional-effects-of-mpas}{here}.
@@
 island). For any one simulation we randomly pull a species and its
-associated life history from the \texttt{FishLife} (58) package in R. We
+associated life history from the \texttt{FishLife} (41) package in R. We
 pair these data with randomly selected values between 0.6 and 0.95 for
@@
 
+Each simulation is randomly assigned a fleet model of the form
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Open access: fishing effort changes in response to
+  profit-per-unit-effort
+\item
+  Constant effort: total fishing effort is constant over time (unless
+  altered by MPA displacement model)
+\item
+  Constant catch: the fleet exerts as much effort as need to achieve a
+  target catch
+\end{enumerate}
+
+This fleet model is paired with a gear selectivity ranging from .1 to
+1.5 of the length at 50\% maturity for the species in question, and the
+fleet model is tuned to achieve a target fishing mortality relative to
+natural mortality ratio at equilibrium.
+
 Along with the fleet dynamics model, each simulation is assigned a
@@
 without the selected MPA strategy (holding all else constant). We then
-measure the difference in biomass and fishery catches in each time step
-in the scenario with and without the MPAs to calculate the conservation
-and fishery effects of the MPAs over time.
+measure the difference in biomass densities each time step in the
+scenario with and without the MPAs to calculate the conservation and
+fishery effects of the MPAs over time.
 
@@
 
-We use a difference-in-difference regression style regression to attempt
-to estimate the causal effect of MPAs on regional targeted fish biomass.
-Empirically, this amounts to estimating the pre-post MPA difference in
-the biomass densities of targeted species net the same difference for
+\begin{figure}%[tbhp]
+  \centering
+  \includegraphics[width=1\linewidth]{/Users/danovan/projects/regional-effects-of-mpas/documents/figs/sample_location_plot.png}
+  \caption{Map of PISCO sampling locations within the the Channel Islands National Marine Sanctuary, California, USA.
+Grey boxes indicated MPAs, points sampling locations, with color representing whether the site is inside an MPA (blue) or not (red)}
+  \label{channel-islands}
+\end{figure}
+
+The difference-in-difference model used empirical kelp forest survey
+data from the Partnership for Interdisciplinary Studies of Coastal
+Oceans (PISCO) monitoring in the Northern Channel Islands. A network of
+MPAs covering approximately 20\% of the islands' waters was put in place
+in 2003 as part of the California Marine Life Protection Act (MLPA) (see
+(22), (23), (24), and (25) for information on the creation of the MLPA).
+PISCO conducts visual SCUBA surveys at a large number of rocky-reef and
+kelp forest sites inside and outside of MPAs throughout the Channel
+Islands, producing estimates of densities of fishes that are both
+targeted and non-targeted by fishing
+(Fig.\ref{channel-islands}-\ref{pop-trends}). . The details of the
+monitoring program are described in XX
+
+The key assumptions of the difference-in-difference model are that a)
+within the time-frame of the model there are no significant interaction
+effects between the targeted and non-targeted species (which in fact we
+do not detect, see SI), and b) that in the absence of the MPAs both the
+targeted and non-targeted groups of species would have exhibited similar
+trends in densities. The advantage to this approach is that given the
+time-frame of the model (14 years with MPAs), we believe that spillover
+effects of MPA placement on the non-targeted species are likely to be
+much less severe than the effects of biological spillover and fishing
+concentration that may bias the performance of estimators such as a
+response ratios.
+
+The regression amounts to estimating the pre-post MPA difference in the
+biomass densities of targeted species net the same difference for
 non-targeted species in the Channel Islands.
 
-The simplified form of this model is
+The simplified form of this model is XX adjust to GAMMA glm
 
 \begin{equation}
-  log(d_{i}) = \beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + e_{i}
+  d_{i} = Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i},shape, scale)
 \label{eq:did}
@@
 observation \emph{i} is in a pre MPA (\(MPA = 0\)) or post MPA
-(\(MPA = 1\)) state. Under the assumptions of this model, \(\beta_3\) is
-the causal effect of the treatment (\emph{MPA}) on the treated (targeted
-species). The full form of the estimation model is much more involved,
-integrating uncertainty in a hierarchical model from the individual
-observation level up to the mean trends of targeted and non-targeted
-species. See SI for details of estimation model.
+(\(MPA = 1\)) state. is a vector of coefficients for additional control
+variables in matrix \emph{X} such as water visibility and observer
+experience. Under the assumptions of this model, \(\beta_3\) is the
+causal effect of the treatment (\emph{MPA}) on the treated (targeted
+species). The shape and scale parameters of the Gamma distribution are
+estimated as well. See SI for further details of the estimation model.
 
@@
 significant differences in the trends of the biomass densities of the
-targeted and non-targeted species in the years before the MPAs.
+targeted and non-targeted species in the years before the MPAs. XX
 
@@
 
-\showacknow
-
 ~ ~ ~
@@
 \leavevmode\hypertarget{ref-lester2009}{}%
-5. Lester S, et al. (2009) Biological effects within no-take marine
+5. Lester SE, et al. (2009) Biological effects within no-take marine
 reserves: A global synthesis. \emph{Marine Ecology Progress Series}
@@
 \leavevmode\hypertarget{ref-goni2010}{}%
-9. Goni R, Hilborn R, D\textbackslash{}'ıaz D, Mallol S, Adlerstein S
+9. Goni R, Hilborn R, D\textbackslash'ıaz D, Mallol S, Adlerstein S
 (2010) Net contribution of spillover from a marine reserve to fishery
@@
 
+\leavevmode\hypertarget{ref-osmond2010}{}%
+22. Osmond M, Airame S, Caldwell M, Day J (2010) ``Lessons for marine
+conservation planning: A comparison of three marine protected area
+planning processes''. \emph{Ocean \& Coastal Management} 53(2):41--51.
+
+\leavevmode\hypertarget{ref-kirlin2013}{}%
+23. Kirlin J, et al. (2013) California's Marine Life Protection Act
+Initiative: Supporting implementation of legislation establishing a
+statewide network of marine protected areas. \emph{Ocean \& Coastal
+Management} 74:3--13.
+
+\leavevmode\hypertarget{ref-botsford2014}{}%
+24. Botsford LW, White JW, Carr MH, Caselle JE (2014) Chapter Six -
+Marine Protected Area Networks in California, USA. \emph{Advances in
+Marine Biology}, Marine Managed Areas and Fisheries., eds Johnson ML,
+Sandell J (Academic Press), pp 205--251.
+
+\leavevmode\hypertarget{ref-hilborn2012}{}%
+25. Hilborn R (2012) The role of science in MPA establishment in
+California: A personal perspective. \emph{Environmental Conservation}
+39(03):195--198.
+
 \leavevmode\hypertarget{ref-gaines2003}{}%
-22. Gaines SD, Gaylord B, Largier JL (2003) Avoiding Current Oversights
+26. Gaines SD, Gaylord B, Largier JL (2003) Avoiding Current Oversights
 in Marine Reserve Design. \emph{Ecological Applications} 13(sp1):32--46.
@@
 \leavevmode\hypertarget{ref-botsford2008}{}%
-23. Botsford LW, et al. (2008) Connectivity, sustainability, and yield:
+27. Botsford LW, et al. (2008) Connectivity, sustainability, and yield:
 Bridging the gap between conventional fisheries management and marine
@@
 \leavevmode\hypertarget{ref-difranco2018}{}%
-24. Di Franco A, et al. (2018) Linking home ranges to protected area
+28. Di Franco A, et al. (2018) Linking home ranges to protected area
 size: The case study of the Mediterranean Sea. \emph{Biological
@@
 \leavevmode\hypertarget{ref-mcgilliard2008}{}%
-25. McGilliard CR, Hilborn R (2008) Modeling no-take marine reserves in
+29. McGilliard CR, Hilborn R (2008) Modeling no-take marine reserves in
 regulated fisheries: Assessing the role of larval dispersal.
@@
 \leavevmode\hypertarget{ref-hilborn2004}{}%
-26. Hilborn R, Punt AE, Orensanz J (2004) Beyond band-aids in fisheries
+30. Hilborn R, Punt AE, Orensanz J (2004) Beyond band-aids in fisheries
 management: Fixing world fisheries. \emph{Bulletin of Marine Science}
@@
 \leavevmode\hypertarget{ref-hilborn1992}{}%
-27. Hilborn R (1992) Can fisheries agencies learn from experience?
+31. Hilborn R (1992) Can fisheries agencies learn from experience?
 \emph{Fisheries} 17(4):6--14.
@@
 \leavevmode\hypertarget{ref-walters2004}{}%
-28. Walters CJ, Martell SJ (2004) \emph{Fisheries ecology and
+32. Walters CJ, Martell SJ (2004) \emph{Fisheries ecology and
 management} (Princeton University Press).
@@
 \leavevmode\hypertarget{ref-hastings2003}{}%
-29. Hastings A, Botsford LW (2003) Comparing Designs Of Marine Reserves
+33. Hastings A, Botsford LW (2003) Comparing Designs Of Marine Reserves
 For Fisheries And For Biodiversity. \emph{Ecological Applications}
@@
 \leavevmode\hypertarget{ref-gerber2003}{}%
-30. Gerber LR, et al. (2003) Population Models For Marine Reserve
+34. Gerber LR, et al. (2003) Population Models For Marine Reserve
 Design: A Retrospective And Prospective Synthesis. \emph{Ecological
@@
 \leavevmode\hypertarget{ref-hilborn2004a}{}%
-31. Hilborn R, et al. (2004) When can marine reserves improve fisheries
+35. Hilborn R, et al. (2004) When can marine reserves improve fisheries
 management? \emph{Ocean \& Coastal Management} 47(34):197--205.
@@
 \leavevmode\hypertarget{ref-botsford2003}{}%
-32. Botsford LW, Micheli F, Hastings A (2003) Principles for the Design
+36. Botsford LW, Micheli F, Hastings A (2003) Principles for the Design
 of Marine Reserves. \emph{Ecological Applications} 13(sp1):25--31.
 
+\leavevmode\hypertarget{ref-smith2003}{}%
+37. Smith MD, Wilen JE (2003) Economic impacts of marine reserves: The
+importance of spatial behavior. \emph{Journal of Environmental Economics
+and Management} 46(2):183--206.
+
+\leavevmode\hypertarget{ref-smith2006}{}%
+38. Smith K (2006) Distributions, relative abundances and reproductive
+biology of the deep-water crabs Hypothalassia acerba and Chaceon bicolor
+in southwestern Australia.
+
+\leavevmode\hypertarget{ref-walters2000}{}%
+39. Walters C (2000) Impacts of dispersal, ecological interactions, and
+fishing effort dynamics on efficacy of marine protected areas: How large
+should protected areas be? \emph{Bulletin of marine science}
+66(3):745--757.
+
 \leavevmode\hypertarget{ref-costello2010}{}%
-33. Costello C, Lynham J, Lester SE, Gaines SD (2010) Economic
+40. Costello C, Lynham J, Lester SE, Gaines SD (2010) Economic
 incentives and global fisheries sustainability. \emph{Resource} 2.
 
-\leavevmode\hypertarget{ref-fao2018}{}%
-34. FAO (2018) \emph{The State of World Fisheries and Aquaculture 2018:
-Meeting the sustainable development goals} (FAO, Rome, Italy).
-
-\leavevmode\hypertarget{ref-costello2012}{}%
-35. Costello C, Gaines S, Gerber LR (2012) Conservation science: A
-market approach to saving the whales. \emph{Nature} 481(7380):139--140.
-
-\leavevmode\hypertarget{ref-rosenberg2017}{}%
-36. Rosenberg AA, et al. (2017) Applying a New Ensemble Approach to
-Estimating Stock Status of Marine Fisheries around the World.
-\emph{Conservation Letters}:n/a--n/a.
-
-\leavevmode\hypertarget{ref-starr2015}{}%
-37. Starr RM, et al. (2015) Variation in Responses of Fishes across
-Multiple Reserves within a Network of Marine Protected Areas in
-Temperate Waters. \emph{PLoS ONE} 10(3):e0118502.
+\leavevmode\hypertarget{ref-thorson2017c}{}%
+41. Thorson JT, Munch SB, Cope JM, Gao J (2017) Predicting life history
+parameters for all fishes worldwide. \emph{Ecological
+Applications}:n/a--n/a.
+
+\leavevmode\hypertarget{ref-love2011}{}%
+42. Love MS (2011) \emph{Certainly More Than You Want to Know About The
+Fishes of The Pacific Coast: A Postmodern Experience} (Really Big
+Press).
+
+\leavevmode\hypertarget{ref-rassweiler2012}{}%
+43. Rassweiler A, Costello C, Siegel DA (2012) Marine protected areas
+and the value of spatially optimized fishery management.
+\emph{Proceedings of the National Academy of Sciences}
+109(29):11884--11889.
+
+\leavevmode\hypertarget{ref-alonzo2004}{}%
+44. Alonzo S, Key M, Ish T, MacCall AD (2004) \emph{Status of the
+California Sheephead Stock for 2004}.
+
+\leavevmode\hypertarget{ref-dick2017}{}%
+45. Dick EJ, et al. (2017) \emph{The Combined Status of Blue and Deacon
+Rockfishes in U.S. Waters off California and Oregon in 2017} (Pacific
+Fishery Management Council, Portland, OR).
 
 \leavevmode\hypertarget{ref-caselle2015}{}%
-38. Caselle JE, Rassweiler A, Hamilton SL, Warner RR (2015) Recovery
+46. Caselle JE, Rassweiler A, Hamilton SL, Warner RR (2015) Recovery
 trajectories of kelp forest animals are rapid yet spatially variable
@@
 
+\leavevmode\hypertarget{ref-ferraro2018}{}%
+47. Ferraro PJ, Sanchirico JN, Smith MD (2018) Causal inference in
+coupled human and natural systems. \emph{Proceedings of the National
+Academy of Sciences}:201805563.
+
+\leavevmode\hypertarget{ref-halpern2004}{}%
+48. Halpern BS, Gaines SD, Warner RR (2004) Confounding effects of the
+export of production and the displacement of fishing effort from marine
+reserves. \emph{Ecological Applications} 14(4):1248--1256.
+
+\leavevmode\hypertarget{ref-larsen2019}{}%
+49. Larsen AE, Meng K, Kendall BE (2019) Causal analysis in
+controlImpact ecological studies with observational data. \emph{Methods
+in Ecology and Evolution} 0(0).
+doi:\href{https://doi.org/10.1111/2041-210X.13190}{10.1111/2041-210X.13190}.
+
+\leavevmode\hypertarget{ref-costello2010d}{}%
+50. Costello C, et al. (2010) The value of spatial information in MPA
+network design. \emph{Proceedings of the National Academy of Sciences}
+107(43):18294--18299.
+
+\leavevmode\hypertarget{ref-rassweiler2014}{}%
+51. Rassweiler A, Costello C, Hilborn R, Siegel DA (2014) Integrating
+scientific guidance into marine spatial planning. \emph{Proceedings of
+the Royal Society B: Biological Sciences} 281(1781):20132252.
+
+\leavevmode\hypertarget{ref-fao2018}{}%
+52. FAO (2018) \emph{The State of World Fisheries and Aquaculture 2018:
+Meeting the sustainable development goals} (FAO, Rome, Italy).
+
+\leavevmode\hypertarget{ref-costello2012}{}%
+53. Costello C, Gaines S, Gerber LR (2012) Conservation science: A
+market approach to saving the whales. \emph{Nature} 481(7380):139--140.
+
+\leavevmode\hypertarget{ref-rosenberg2017}{}%
+54. Rosenberg AA, et al. (2017) Applying a New Ensemble Approach to
+Estimating Stock Status of Marine Fisheries around the World.
+\emph{Conservation Letters}:n/a--n/a.
+
+\leavevmode\hypertarget{ref-halpern2002}{}%
+55. Halpern BS, Warner RR (2002) Marine reserves have rapid and lasting
+effects. \emph{Ecology Letters} 5(3):361--366.
+
 \leavevmode\hypertarget{ref-nickols2019}{}%
-39. Nickols KJ, et al. (2019) Setting ecological expectations for
+56. Nickols KJ, et al. (2019) Setting ecological expectations for
 adaptive management of marine protected areas. \emph{Journal of Applied
@@
 
-\leavevmode\hypertarget{ref-villasenor-derbez2019}{}%
-40. Villaseñor-Derbez JC, et al. (2019) An interdisciplinary evaluation
-of community-based TURF-reserves. \emph{PLOS ONE} 14(8):e0221660.
-
-\leavevmode\hypertarget{ref-ferraro2018}{}%
-41. Ferraro PJ, Sanchirico JN, Smith MD (2018) Causal inference in
-coupled human and natural systems. \emph{Proceedings of the National
-Academy of Sciences}:201805563.
-
-\leavevmode\hypertarget{ref-larsen2019}{}%
-42. Larsen AE, Meng K, Kendall BE (2019) Causal analysis in
-controlImpact ecological studies with observational data. \emph{Methods
-in Ecology and Evolution} 0(0).
-doi:\href{https://doi.org/10.1111/2041-210X.13190}{10.1111/2041-210X.13190}.
-
-\leavevmode\hypertarget{ref-halpern2002}{}%
-43. Halpern BS, Warner RR (2002) Marine reserves have rapid and lasting
-effects. \emph{Ecology Letters} 5(3):361--366.
-
-\leavevmode\hypertarget{ref-halpern2004}{}%
-44. Halpern BS, Gaines SD, Warner RR (2004) Confounding effects of the
-export of production and the displacement of fishing effort from marine
-reserves. \emph{Ecological Applications} 14(4):1248--1256.
-
-\leavevmode\hypertarget{ref-costello2010d}{}%
-45. Costello C, et al. (2010) The value of spatial information in MPA
-network design. \emph{Proceedings of the National Academy of Sciences}
-107(43):18294--18299.
-
-\leavevmode\hypertarget{ref-rassweiler2014}{}%
-46. Rassweiler A, Costello C, Hilborn R, Siegel DA (2014) Integrating
-scientific guidance into marine spatial planning. \emph{Proceedings of
-the Royal Society B: Biological Sciences} 281(1781):20132252.
-
-\leavevmode\hypertarget{ref-osmond2010}{}%
-47. Osmond M, Airame S, Caldwell M, Day J (2010) ``Lessons for marine
-conservation planning: A comparison of three marine protected area
-planning processes''. \emph{Ocean \& Coastal Management} 53(2):41--51.
-
-\leavevmode\hypertarget{ref-kirlin2013}{}%
-48. Kirlin J, et al. (2013) California's Marine Life Protection Act
-Initiative: Supporting implementation of legislation establishing a
-statewide network of marine protected areas. \emph{Ocean \& Coastal
-Management} 74:3--13.
-
-\leavevmode\hypertarget{ref-botsford2014}{}%
-49. Botsford LW, White JW, Carr MH, Caselle JE (2014) Chapter Six -
-Marine Protected Area Networks in California, USA. \emph{Advances in
-Marine Biology}, Marine Managed Areas and Fisheries., eds Johnson ML,
-Sandell J (Academic Press), pp 205--251.
-
-\leavevmode\hypertarget{ref-hilborn2012}{}%
-50. Hilborn R (2012) The role of science in MPA establishment in
-California: A personal perspective. \emph{Environmental Conservation}
-39(03):195--198.
-
-\leavevmode\hypertarget{ref-halpern2003a}{}%
-51. Halpern BS (2003) The impact of marine reserves: Do reserves work
-and does reserve size matter? \emph{Ecological Applications}
-13(sp1):117--137.
-
-\leavevmode\hypertarget{ref-alonzo2004}{}%
-52. Alonzo S, Key M, Ish T, MacCall AD (2004) \emph{Status of the
-California Sheephead Stock for 2004}.
-
-\leavevmode\hypertarget{ref-dick2017}{}%
-53. Dick EJ, et al. (2017) \emph{The Combined Status of Blue and Deacon
-Rockfishes in U.S. Waters off California and Oregon in 2017} (Pacific
-Fishery Management Council, Portland, OR).
-
-\leavevmode\hypertarget{ref-gentemann2017}{}%
-54. Gentemann CL, Fewings MR, García-Reyes M (2017) Satellite sea
-surface temperatures along the West Coast of the United States during
-the 20142016 northeast Pacific marine heat wave. \emph{Geophysical
-Research Letters} 44(1):312--319.
-
-\leavevmode\hypertarget{ref-freedman2019}{}%
-55. Freedman R (2019) Understanding the Efficacy of Spatial Management
-on Emerging Threats. PhD thesis (University of California, Santa
-Barbara).
-
 \leavevmode\hypertarget{ref-rcoreteam2019}{}%
-56. R Core Team (2019) R: A Language and Environment for Statistical
+57. R Core Team (2019) R: A Language and Environment for Statistical
 Computing.
 
-\leavevmode\hypertarget{ref-kristensen2016}{}%
-57. Kristensen K, Nielsen A, Berg CW, Skaug H, Bell BM (2016)
-\textbf{TMB} : Automatic Differentiation and Laplace Approximation.
-\emph{Journal of Statistical Software} 70(5).
-doi:\href{https://doi.org/10.18637/jss.v070.i05}{10.18637/jss.v070.i05}.
-
-\leavevmode\hypertarget{ref-thorson2017c}{}%
-58. Thorson JT, Munch SB, Cope JM, Gao J (2017) Predicting life history
-parameters for all fishes worldwide. \emph{Ecological
-Applications}:n/a--n/a.
+\leavevmode\hypertarget{ref-carpenter2017}{}%
+58. Carpenter B, et al. (2017) \emph{Stan} : A Probabilistic Programming
+Language. \emph{Journal of Statistical Software} 76(1).
+doi:\href{https://doi.org/10.18637/jss.v076.i01}{10.18637/jss.v076.i01}.
 

### 2020-03-26 functions/estimate_did.R

@@ -70,7 +70,7 @@ estimate_did <-
 #         geom_histogram() + 
 #         facet_wrap(~region)
       
-    did_data <- did_data %>% 
+   classcode_level_data <- did_data %>% 
       group_by(year,
                site_side,
                region,
@@ -98,7 +98,10 @@ estimate_did <-
         var_surge = mean(var_surge),
         var_kelp = mean(var_kelp)
       ) %>% # calculate mean density per year site, side, species, averaging over zone, transect
-      group_by(year, site_side, region, eventual_mpa, targeted) %>%
+      ungroup()
+   
+     did_data <- classcode_level_data %>% 
+     group_by(year, site_side, region, eventual_mpa, targeted) %>%
       summarise(
         total_biomass_density = (sum(md) / 1e6) * 10000,
         # calculate total and mean biomass densities across all species per year site side
@@ -332,7 +335,7 @@ estimate_did <-
     
     
     } else if (data_source == "kfm"){
-      did_data <- did_data %>% 
+      classcode_level_data <- did_data %>% 
         group_by(year,
                  site_side,
                  region,
@@ -350,6 +353,10 @@ estimate_did <-
           var_temp = mean(var_temp),
           var_kelp = mean(var_kelp)
         ) %>% # calculate mean density per year site, side, species, averaging over zone, transect
+        ungroup()
+        
+      
+      did_data <- classcode_level_data %>% 
         group_by(year, site_side, region, eventual_mpa, targeted) %>%
         summarise(
           total_biomass_density = (sum(md) / 1e6) * 10000,
@@ -489,7 +496,9 @@ estimate_did <-
       scale_x_discrete(name = "Year Bin")
     # 
     out <- list(did_results = did_results,
-                did_reg = did_reg
+                did_reg = did_reg,
+                did_data = did_data,
+                classcode_level_data = classcode_level_data
                 )
     
   }

### 2020-03-26 make-regional-mpas.R

@@ -30,6 +30,7 @@ library(bayesplot)
 library(tabulizer)
 library(tidyverse)
 library(ggtext)
+library(ggrepel)
 extrafont::loadfonts()
 # extrafont::font_import()
 rstan_options(auto_write = TRUE)
@@ -1347,7 +1348,6 @@ model_runs <- model_runs %>%
     )
   )
 
-
 # did_worked <- map(model_runs$did_fit, "error") %>% map_lgl(is.null)
 
 # model_runs <- model_runs %>% 
@@ -1373,9 +1373,10 @@ plot_did <- function(x){
   
 }
 
-did_fits <- did_fits %>% 
-  mutate(did_plot = map(did_fit, plot_did))
-  
+
+# did_fits <- did_fits %>% 
+#   mutate(did_plot = map(did_fit, plot_did))
+#   
 # print(object.size(did_fits), units = "Mb")
 
 
@@ -2963,27 +2964,11 @@ if (process_results == TRUE){
     group_by(experiment) %>% 
     summarise(n = sum(years_protected ==0))
   
-  wtf <- outcomes %>% 
-    filter(experiment == 9508)
   
-  pop_depletion_and_size_plot <- outcomes %>%
-    filter(years_protected >= 0) %>%
-    group_by(experiment) %>%
-    mutate(depletion = plyr::round_any(depletion[years_protected == 0], .05),
-           mpa_size = plyr::round_any(mpa_size, .05)) %>%
-    filter(years_protected == short_frame) %>%
-    group_by(depletion, mpa_size) %>%
-    summarise(median_mpa_effect = median(mpa_effect)) %>%
-    ggplot(aes(depletion, mpa_size, fill = median_mpa_effect)) +
-    geom_tile() +
-    # geom_contour(aes(z = median_mpa_effect)) +
-    scale_fill_viridis(labels = percent,
-                       guide = hgc,
-                       name = "Median Effect") +
-    labs(x = "Pre-MPA Depletion", y = "Range in MPA") +
-    scale_y_continuous(labels = percent) +
-    scale_x_continuous(labels = percent) +
-    theme(legend.position = "top")
+
+  
+    
+  
 
   eq_pop_depletion_and_size_plot <- outcomes %>%
     filter(years_protected >= 0) %>%
@@ -3024,23 +3009,38 @@ if (process_results == TRUE){
 
 
 
+  
+  
   pop_depletion_and_size_plot <- outcomes %>%
     filter(years_protected >= 0) %>%
     group_by(experiment) %>%
[…]

### 2020-03-26 documents/Untitled.Rmd
migrated paper to simpler bookdown template

@@
 title: The Regional Effects of Marine Protected Areas
-author:
-  - name: Daniel Ovando
-    affiliation: a,1
-    email: danovan@uw.edu
-  - name: Jennifer E. Caselle
-    affiliation: b
-  - name: Christopher Costello
-    affiliation: b   
-  - name: Olivier Deschenes
-    affiliation: b  
-  - name: Steven D. Gaines
-    affiliation: b
-  - name: Ray Hilborn
-    affiliation: a
-  - name: Owen Liu
-    affiliation: b        
-address:
-  - code: a
-    address: University of Washington, School of Aquatic and Fishery Sciences
-  - code: b
-    address: University of California, Santa Barbara
-corresponding_author:
-  - code: 1
-    text: "To whom correspondence should be addressed. E-mail: danovan@uw.edu"
-lead_author_surname: Ovando
-author_contributions: |
-  D.O., S.G., and R.H.
-developed structure of simulation model.
-D.O., O.D., C.C., and J.C.
-developed estimation strategy.
-J.C.
-provided support on collecting and interpreting data.
-All simulations, statistics, and sensitivity analyses performed by D.O.
-and O.L.
-.
-All authors contributed to the manuscript.
-conflict_of_interest: |
-  R.H's research program receives funding from environmental NGOs, foundations, fishing industry, governments and international agencies.
-All of these can be interpreted as a conflict of interest when evaluating fisheries policy.
-C.C., S.G., and O.D's research program includes funding from environmental NGOs and foundations with an interest in ocean conservation and food security.
-Funding for D.O.
-was partly provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship.
-abstract: |
-  Marine Protected Areas (MPAs) cover between 3-7% of the world's oceans, up from less than 1% in the year 2000.
-The Convention on Biological Diversity calls for 10% of coastal waters to be protected inside MPAs by 2020, while the International Union for Conservation of Nature calls for 30% protection by 2030.
-It is often clear that MPAs produce conservation benefits inside their borders, but many MPAs are also justified on the grounds that they also benefit the broader region outside their borders.
-The conservation effects of MPAs are most commonly evaluated using response ratios---measures of biomass densities inside their borders, relative to biomass densities in reference sites outside their borders.
-Studies of this nature have provided broad evidence that MPAs produce conservation benefits within their borders.
-While these gains can be vitally important, marine populations are rarely contained entirely within the borders of MPAs; therefore a critical question is not only do MPAs produce conservation gains inside their borders, but how do they affect the broader region in which they are located?
-The Channel Islands National Marine Sanctuary provides a clear example of this challenge.
-A network of MPAs covering roughly 20% of the Islands’ waters was put in place in 2003, with a goal of providing regional conservation and fishery benefits.
-Response ratios from the region indicate that the Channel Island MPAs have increased biomass densities inside the MPAs.
-However, we are unable to find a clear effect of these same MPAs at the regional scale across multiple species.Building off of existing theory, we use a bio-economic simulation model to explain this discrepancy, and demonstrate under what conditions we likely can and cannot clearly estimate the regional conservation effects of MPAs.
-We show that MPA networks covering 25% or less of a region are likely to produce regional increases in fish biomass on the order of 10%-20%, a meaningful effect but also one that can easily be overwhelmed by environmental shocks.
-Our results provide a novel assessment of the regional effects of a large and iconic Marine Protected Area network, and provide guidance for managers charged with monitoring and adapting MPAs.
-significance: |
-  Healthy marine ecosystems are critical to the well-being of the planet.
-Marine protected areas, parts of the oceans protected from human activities such as fishing, are increasingly being used in an effort to conserve and manage these ecosystems.
-Our study pairs theory with empirical methods to examine the regional effects of MPAs on conservation outcomes.
-Using simulations, we show that the regional effects of marine protected areas can be highly variable and dependent on human drivers.
-We find no statistically clear effect of a network of MPAs after 13 years of protection, which our simulation analysis suggests is actually to be expected.
-Our results lay a foundation for future research on the design and monitoring of marine protected areas.
-acknowledgements: |
-  Funding for the simulation model was provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship.
-This study would not be possible without the work provided by PISCO divers over the years.
-DO thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments and technical support.
-keywords:
-  - Marine Protected Areas
-  - Conservation
-  - Bio-economic modeling
-  - Program Evaluation
-## must be one of: pnasresearcharticle (usual two-column layout), pnasmathematics (one column layout), or pnasinvited (invited submissions only)
-pnas_type: pnasresearcharticle
+author: Daniel Ovando, Jennifer E. Caselle, Christopher Costello, Olivier Deschenes, Steven D. Gaines, Ray Hilborn, Owen Liu
 bibliography: My Library.bib
 csl: pnas.csl
-lineno: true
-header-includes:
-   - \usepackage{eso-pic,graphicx,transparent,xcolor}
+
 output:
-  rticles::pnas_article: default
+  bookdown::pdf_document2: default
   bookdown::word_document2: default
-  bookdown::pdf_document2: default
   bookdown::html_document2: default
@@
 
-<!-- \AddToShipoutPictureFG{ -->
-<!--   \AtPageCenter{% or \AtTextCenter -->
-<!--     \makebox[0pt]{\rotatebox[origin=c]{45}{% -->
-<!--       \scalebox{5}{\texttransparent{0.3}{DRAFT}}% -->
-<!--     }} -->
-<!--   } -->
-<!-- } -->
+$$d_{i} \sim Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + B\mathbf{X} + B,shape, scale)$$

### 2020-03-26 documents/Untitled.tex
migrated paper to simpler bookdown template

@@
-\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
-
-%% Some pieces required from the pandoc template
+\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
+%
+\documentclass[]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provides euro and other symbols
+\else % if luatex or xelatex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage[]{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\usepackage{xcolor}
+\usepackage{hyperref}
+\hypersetup{
+            pdftitle={The Regional Effects of Marine Protected Areas},
+            pdfauthor={Daniel Ovando, Jennifer E. Caselle, Christopher Costello, Olivier Deschenes, Steven D. Gaines, Ray Hilborn, Owen Liu},
+            colorlinks=true,
+            linkcolor=blue,
+            filecolor=Maroon,
+            citecolor=Blue,
+            urlcolor=Blue,
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage[margin=1in]{geometry}
+\usepackage{longtable,booktabs}
+% Fix footnotes in tables (requires footnote package)
+\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{longtable}}{}
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
 \providecommand{\tightlist}{%
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-% Use the lineno option to display guide line numbers if required.
-% Note that the use of elements such as single-column equations
-% may affect the guide line number alignment.
-
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-
-\usepackage{eso-pic,graphicx,transparent,xcolor}
-
-\templatetype{pnasresearcharticle}  % Choose template
+% set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 
 \title{The Regional Effects of Marine Protected Areas}
-
-\author[a,1]{Daniel Ovando}
-\author[b]{Jennifer E. Caselle}
-\author[b]{Christopher Costello}
-\author[b]{Olivier Deschenes}
-\author[b]{Steven D. Gaines}
-\author[a]{Ray Hilborn}
-\author[b]{Owen Liu}
-
-  \affil[a]{University of Washington, School of Aquatic and Fishery Sciences}
-  \affil[b]{University of California, Santa Barbara}
-
-% Please give the surname of the lead author for the running footer
-\leadauthor{Ovando}
-
-% Please add here a significance statement to explain the relevance of your work
-\significancestatement{Healthy marine ecosystems are critical to the well-being of the planet.
-Marine protected areas, parts of the oceans protected from human
-activities such as fishing, are increasingly being used in an effort to
-conserve and manage these ecosystems. Our study pairs theory with
-empirical methods to examine the regional effects of MPAs on
-conservation outcomes. Using simulations, we show that the regional
-effects of marine protected areas can be highly variable and dependent
-on human drivers. We find no statistically clear effect of a network of
-MPAs after 13 years of protection, which our simulation analysis
-suggests is actually to be expected. Our results lay a foundation for
-future research on the design and monitoring of marine protected areas.}
-
-\authorcontributions{D.O., S.G., and R.H. developed structure of simulation model. D.O.,
-O.D., C.C., and J.C. developed estimation strategy. J.C. provided
-support on collecting and interpreting data. All simulations,
-statistics, and sensitivity analyses performed by D.O. and O.L. . All
-authors contributed to the manuscript.}
-
-\authordeclaration{R.H's research program receives funding from environmental NGOs,
-foundations, fishing industry, governments and international agencies.
-All of these can be interpreted as a conflict of interest when
-evaluating fisheries policy. C.C., S.G., and O.D's research program
-includes funding from environmental NGOs and foundations with an
-interest in ocean conservation and food security. Funding for D.O. was
-partly provided by the NMFS-Sea Grant Population and Ecosystem Dynamics
-Fellowship.}
-
-\correspondingauthor{\textsuperscript{} }
-
-% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
- \keywords{  Marine Protected Areas |  Conservation |  Bio-economic modeling |  Program Evaluation  }
-
-\begin{abstract}
-Marine Protected Areas (MPAs) cover between 3-7\% of the world's oceans,
-up from less than 1\% in the year 2000. The Convention on Biological
-Diversity calls for 10\% of coastal waters to be protected inside MPAs
-by 2020, while the International Union for Conservation of Nature calls
-for 30\% protection by 2030. It is often clear that MPAs produce
-conservation benefits inside their borders, but many MPAs are also
-justified on the grounds that they also benefit the broader region
-outside their borders. The conservation effects of MPAs are most
-commonly evaluated using response ratios---measures of biomass densities
-inside their borders, relative to biomass densities in reference sites
-outside their borders. Studies of this nature have provided broad
-evidence that MPAs produce conservation benefits within their borders.
-While these gains can be vitally important, marine populations are
-rarely contained entirely within the borders of MPAs; therefore a
-critical question is not only do MPAs produce conservation gains inside
-their borders, but how do they affect the broader region in which they
-are located? The Channel Islands National Marine Sanctuary provides a
-clear example of this challenge. A network of MPAs covering roughly 20\%
-of the Islands' waters was put in place in 2003, with a goal of
-providing regional conservation and fishery benefits. Response ratios
-from the region indicate that the Channel Island MPAs have increased
-biomass densities inside the MPAs. However, we are unable to find a
-clear effect of these same MPAs at the regional scale across multiple
-species.Building off of existing theory, we use a bio-economic
-simulation model to explain this discrepancy, and demonstrate under what
-conditions we likely can and cannot clearly estimate the regional
-conservation effects of MPAs. We show that MPA networks covering 25\% or
-less of a region are likely to produce regional increases in fish
-biomass on the order of 10\%-20\%, a meaningful effect but also one that
-can easily be overwhelmed by environmental shocks. Our results provide a
-novel assessment of the regional effects of a large and iconic Marine
-Protected Area network, and provide guidance for managers charged with
-monitoring and adapting MPAs.
-\end{abstract}
-
-\dates{This manuscript was compiled on \today}
-\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
+\author{Daniel Ovando, Jennifer E. Caselle, Christopher Costello, Olivier Deschenes, Steven D. Gaines, Ray Hilborn, Owen Liu}
+\date{}
 
 \begin{document}
+\maketitle
 
-% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
-% You should only change this length when you've finalised the article contents.
-\verticaladjustment{-2pt}
-
-\maketitle
-\thispagestyle{firststyle}
-\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
-
-% If your first paragraph (i.e.
-with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation.
-If this is the case, add \parshape=0 to the end of the list environment.
-
-\acknow{Funding for the simulation model was provided by the NMFS-Sea Grant
-Population and Ecosystem Dynamics Fellowship. This study would not be
-possible without the work provided by PISCO divers over the years. DO
-thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments
-and technical support.}
-
-% Bibliography
-% \bibliography{pnas-sample}
+{
+\hypersetup{linkcolor=}
+\setcounter{tocdepth}{2}
+\tableofcontents
+}
+d\_\{i\} = Gamma(e\^{}\{\beta\emph{0 + \beta\emph{1T}\{i\} + \beta\emph{2MPA}\{i\} + \beta}\{3\}T\_iMPA\_i + \Beta\bold{X}\},shape, scale)
 

### 2020-03-26 documents/ovando-regional-effects-of-mpas-v2.Rmd
migrated paper to simpler bookdown template

@@
 title: The Regional Effects of Marine Protected Areas
-author:
-  - name: Daniel Ovando
-    affiliation: a,1
-    email: danovan@uw.edu
-  - name: Jennifer E. Caselle
-    affiliation: b
-  - name: Christopher Costello
-    affiliation: b   
-  - name: Olivier Deschenes
-    affiliation: b  
-  - name: Steven D. Gaines
-    affiliation: b
-  - name: Ray Hilborn
-    affiliation: a
-  - name: Owen Liu
-    affiliation: b        
-address:
-  - code: a
-    address: University of Washington, School of Aquatic and Fishery Sciences
-  - code: b
-    address: University of California, Santa Barbara
-corresponding_author:
-  - code: 1
-    text: "To whom correspondence should be addressed. E-mail: danovan@uw.edu"
-lead_author_surname: Ovando
-author_contributions: |
-  D.O., S.G., and R.H.
-developed structure of simulation model.
-D.O., O.D., C.C., and J.C.
-developed estimation strategy.
-J.C.
-provided support on collecting and interpreting data.
-All simulations, statistics, and sensitivity analyses performed by D.O.
-and O.L.
-.
-All authors contributed to the manuscript.
-conflict_of_interest: |
-  R.H's research program receives funding from environmental NGOs, foundations, fishing industry, governments and international agencies.
-All of these can be interpreted as a conflict of interest when evaluating fisheries policy.
-C.C., S.G., and O.D's research program includes funding from environmental NGOs and foundations with an interest in ocean conservation and food security.
-Funding for D.O.
-was partly provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship.
-abstract: |
-  Marine Protected Areas (MPAs) cover between 3-7% of the world's oceans, up from less than 1% in the year 2000.
-The Convention on Biological Diversity calls for 10% of coastal waters to be protected inside MPAs by 2020, while the International Union for Conservation of Nature calls for 30% protection by 2030.
-It is often clear that MPAs produce conservation benefits inside their borders, but many MPAs are also justified on the grounds that they also benefit the broader region outside their borders.
-The conservation effects of MPAs are most commonly evaluated using response ratios---measures of biomass densities inside their borders, relative to biomass densities in reference sites outside their borders.
-Studies of this nature have provided broad evidence that MPAs produce conservation benefits within their borders.
-While these gains can be vitally important, marine populations are rarely contained entirely within the borders of MPAs; therefore a critical question is not only do MPAs produce conservation gains inside their borders, but how do they affect the broader region in which they are located?
-The Channel Islands National Marine Sanctuary provides a clear example of this challenge.
-A network of MPAs covering roughly 20% of the Islands’ waters was put in place in 2003, with a goal of providing regional conservation and fishery benefits.
-Response ratios from the region indicate that the Channel Island MPAs have increased biomass densities inside the MPAs.
-However, we are unable to find a clear effect of these same MPAs at the regional scale across multiple species.Building off of existing theory, we use a bio-economic simulation model to explain this discrepancy, and demonstrate under what conditions we likely can and cannot clearly estimate the regional conservation effects of MPAs.
-We show that MPA networks covering 25% or less of a region are likely to produce regional increases in fish biomass on the order of 10%-20%, a meaningful effect but also one that can easily be overwhelmed by environmental shocks.
-Our results provide a novel assessment of the regional effects of a large and iconic Marine Protected Area network, and provide guidance for managers charged with monitoring and adapting MPAs.
-significance: |
-  Healthy marine ecosystems are critical to the well-being of the planet.
-Marine protected areas, parts of the oceans protected from human activities such as fishing, are increasingly being used in an effort to conserve and manage these ecosystems.
-Our study pairs theory with empirical methods to examine the regional effects of MPAs on conservation outcomes.
-Using simulations, we show that the regional effects of marine protected areas can be highly variable and dependent on human drivers.
-We find no statistically clear effect of a network of MPAs after 13 years of protection, which our simulation analysis suggests is actually to be expected.
-Our results lay a foundation for future research on the design and monitoring of marine protected areas.
-acknowledgements: |
-  Funding for the simulation model was provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship.
-This study would not be possible without the work provided by PISCO divers over the years.
-DO thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments and technical support.
-keywords:
-  - Marine Protected Areas
-  - Conservation
-  - Bio-economic modeling
-  - Program Evaluation
-## must be one of: pnasresearcharticle (usual two-column layout), pnasmathematics (one column layout), or pnasinvited (invited submissions only)
-pnas_type: pnasresearcharticle
+author: Daniel Ovando, Jennifer E. Caselle, Christopher Costello, Olivier Deschenes, Steven D. Gaines, Ray Hilborn, Owen Liu
 bibliography: My Library.bib
 csl: pnas.csl
-lineno: true
-header-includes:
-   - \usepackage{eso-pic,graphicx,transparent,xcolor}
 output:
-  rticles::pnas_article: default
   bookdown::word_document2: default
@@
 linkcolor: blue
+header-includes:
+  - \usepackage{setspace}\doublespacing
+  - \usepackage{lineno}\linenumbers
 ---
@@
 ```{r, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dev = "CairoPDF", dpi = 600,
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
                       cache = FALSE)
@@
 
+panel_width = unit(1,"npc") - sum(ggplotGrob(testplot)[["widths"]][-3]) - unit(4,"line")
+
 gc <- guide_colorbar(frame.colour = "black",
@@
 
-fitted_data <- abundance_data$data[abundance_data$data_source == "pisco"][[1]]
+used_data <- abundance_data$data[abundance_data$data_source == "pisco"][[1]]
 
@@
 
-top_species <- fitted_data$classcode %>% unique()
+top_species <- used_data$classcode %>% unique()
 
@@
 
-<!-- \correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: danovan@uw.edu} -->
-
 <!-- \AddToShipoutPictureFG{ -->
@@
 
-*as we move away from "set it and forget it" management towards adaptive management of marine ecosystems, it is critically important that we develop robust strategies for monitoring the effectiveness of our management intercentions*
+# Front Matter
+
+## author_contributions: 
+  D.O., S.G., and R.H.
+developed structure of simulation model.
+D.O., O.D., C.C., and J.C.
+developed estimation strategy.
+J.C.
+provided support on collecting and interpreting data.
+All simulations, statistics, and sensitivity analyses performed by D.O.
+and O.L.
+.
+All authors contributed to the manuscript.
+
+## conflict_of_interest:
+  R.H's research program receives funding from environmental NGOs, foundations, fishing industry, governments and international agencies.
+All of these can be interpreted as a conflict of interest when evaluating fisheries policy.
+C.C., S.G., and O.D's research program includes funding from environmental NGOs and foundations with an interest in ocean conservation and food security.
+Funding for D.O.
+was partly provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship.
+
+## abstract
+  Marine Protected Areas (MPAs) cover between 3-7% of the world's oceans, up from less than 1\% in the year 2000.
+The Convention on Biological Diversity calls for 10% of coastal waters to be protected inside MPAs by 2020, while the International Union for Conservation of Nature calls for 30\% protection by 2030.
+It is often clear that MPAs produce conservation benefits inside their borders, but many MPAs are also justified on the grounds that they also benefit the broader region outside their borders.
+The conservation effects of MPAs are most commonly evaluated using response ratios---measures of biomass densities inside their borders, relative to biomass densities in reference sites outside their borders.
+Studies of this nature have provided broad evidence that MPAs produce conservation benefits within their borders.
+While these gains can be vitally important, marine populations are rarely contained entirely within the borders of MPAs; therefore a critical question is not only do MPAs produce conservation gains inside their borders, but how do they affect the broader region in which they are located?
+The Channel Islands National Marine Sanctuary provides a clear example of this challenge.
+A network of MPAs covering roughly 20\% of the Islands’ waters was put in place in 2003, with a goal of providing regional conservation and fishery benefits.
+Response ratios from the region indicate that the Channel Island MPAs have increased biomass densities inside the MPAs.
+However, we are unable to find a clear effect of these same MPAs at the regional scale across multiple species.Building off of existing theory, we use a bio-economic simulation model to explain this discrepancy, and demonstrate under what conditions we likely can and cannot clearly estimate the regional conservation effects of MPAs.
+We show that MPA networks covering 25\% or less of a region are likely to produce regional increases in fish biomass on the order of 10\%-20\%, a meaningful effect but also one that can easily be overwhelmed by environmental shocks.
+Our results provide a novel assessment of the regional effects of a large and iconic Marine Protected Area network, and provide guidance for managers charged with monitoring and adapting MPAs.
+
+## significance
+  Healthy marine ecosystems are critical to the well-being of the planet.
+Marine protected areas, parts of the oceans protected from human activities such as fishing, are increasingly being used in an effort to conserve and manage these ecosystems.
+Our study pairs theory with empirical methods to examine the regional effects of MPAs on conservation outcomes.
+Using simulations, we show that the regional effects of marine protected areas can be highly variable and dependent on human drivers.
+We find no statistically clear effect of a network of MPAs after 13 years of protection, which our simulation analysis suggests is actually to be expected.
+Our results lay a foundation for future research on the design and monitoring of marine protected areas.
+
+## acknowledgements
+  Funding for the simulation model was provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship.
+This study would not be possible without the work provided by PISCO divers over the years.
+DO thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments and technical support.
+
+## keywords:
+  - Marine Protected Areas
+  - Conservation
+  - Bio-economic modeling
+  - Program Evaluation
+
+# Body
 
@@
 Traditional cultures in Oceania utilized - often temporary - MPAs as "fish banks" for times of need [@johannes1978].
-Modern MPAs were first established primarily as marine analogs to the terrestrial protection of iconic landscapes like Yellowstone or Kruger National Parks [@iucn1976;@watson2014b].
+Modern MPAs were first established primarily as marine analogs to the terrestrial protection of iconic landscapes like Yellowstone or Kruger National Parks [@iucn1976;@watson2014a].
 Over time our goals and expectations for MPAs have evolved; while all MPAs are expected to deliver conservation benefits within their borders, many modern MPAs are also established to bolster fish populations throughout the region in which they are located [@gaines2010].
 
-Indeed, many recent agreements to expand MPAs (most proximately, the Convention on Biological Diversity’s Strategic Plan for Biodiversity, which calls for 10% of coastal waters to be protected inside MPAs by 2020 and the International Union for Conservation of Nature call for 30% by 2030) are based on the expectaion that well-designed MPAs will achieve benefits both within, and outside, their borders.
+Many recent agreements to expand MPAs (most proximately, the Convention on Biological Diversity’s Strategic Plan for Biodiversity, which calls for 10% of coastal waters to be protected inside MPAs by 2020 and the International Union for Conservation of Nature call for 30% by 2030) are based on the expectaion that well-designed MPAs will achieve benefits both within, and outside, their borders.
 Despite these assumptions, our collective scientific understanding of the regional-scale conservation and fishery impacts of current and future MPAs is surprisingly limited.
@@
 We use data from the first 14 years of protection to provide what is to our knowledge the first empirical assessment of the regional effect of a large MPA network on a wide array of finfish species.
-In contrast to clear differences in biomass densities observed inside and outside of well protected MPAs both globally [@lester2009] and in the Channel Islands [@caselle2014] we are unable to detect a clear regional effect from the Channel Islands MPAs.
+In contrast to clear differences in biomass densities observed inside and outside of well protected MPAs both globally [@lester2009] and in the Channel Islands [@caselle2015] we are unable to detect a clear regional effect from the Channel Islands MPAs.
 We build off of existing MPA theory to consider why this might be, and provide guidance for scientists and managers as to when and how we might expect to estimate the regional conservation effects of MPAs.
 
-What Are the Regional Effects of MPAs? {.unnumbered}
-----------------------
+## Results
+
+### What Are the Regional Effects of MPAs? {.unnumbered}
 
@@
 
-We define the regional conservation effects of MPAs as the change in total biomass densities of fish (averaging inside and outside of MPAs) relative to the total biomass densities of fish that would have occurred without the MPAs (acknowledging that other outcomes such as increased biodiversity or resiliency are also important to conservation but are beyond the scope of this analysis).
+We define the regional conservation effects of MPAs as the change in total biomass densities of fish relative to the total biomass densities of fish that would have occurred without the MPAs (acknowledging that other outcomes such as increased biodiversity or resiliency are also important to conservation but are beyond the scope of this analysis).
 Numerous factors can affect the regional effects of MPAs.
 These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], larval dispersal patterns and the strength and timing of density dependence in the population (e.g.
-pre- or post-settlement, @burgess2014), how overfished the population would be without the MPA, and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003;@smith2006;@walters2000;@white2011].
-In addition, even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through habitat and network effects [@gaines2010; @costello2010].
+pre- or post-settlement, @burgess2014), how overfished the population would be without the MPA, and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006;@walters2000a;@white2011].
+In addition, even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through habitat and network effects [@gaines2010; @costello2010a].
 
@@
 We integrated these network design features, life history traits, and exploitation histories into a spatially explicit bio-economic simulation model, and used this model to generate expected outcomes for the Channel Islands MPAs.
-Our results suggest that while a wide range of outcomes are plausible, from 0% to upwards of 200%, across the majority of our simulations the effect size was 25% or less, with a median simulated effect size of roughly 20%; a potentially small value given the natural variability and sampling challenges of marine environments.
+Our results suggest that while a wide range of outcomes are plausible, from 0% to upwards of 200%, across the majority of our simulations the effect size was 25\% or less, with a median simulated effect size of roughly 20\%; a potentially small value given the natural variability and sampling challenges of marine environments.
 
@@
 
-## How Can We Detect Regional Effects?
+### Effects of the Channel Islands MPA Network
 
@@
 They found a statistically significant increase in the response ratios of targeted species over time, and evidence that this increase is smaller in the non-targeted species.
-Updating the results of @caselle2015 through 2017 with a Bayesian response-ratio regression shows a continuation of the increasing trend in the response ratios of targeted species (Fig.\@ref(fig:XX)).
+Updating the results of @caselle2015 through 2017 with a Bayesian response-ratio regression shows a continuation of the increasing trend in the response ratios of targeted species (Fig.\@ref(fig:rr-plot)).
 
@@
 
-As reported in @caselle2015, we see evidence for an increasing trend in response ratios of targeted species over time (Fig.\@ref(fig:rr-plot)).
+As reported in @caselle2015, we see an increasing trend in response ratios of targeted species over time (Fig.\@ref(fig:rr-plot)).
 This provides strong evidence that the MPAs are providing protection for targeted species within their borders.
@@
 Response ratios well over one were associated with regional MPA effects generally less than 25%, and many simulations produced large response ratios but regional MPA effects close to 0% This can occur if for example fishing pressure is only moderate, adult movement is low, larval dispersal is high, and displaced fishing effort concentrates around the border of the MPAs.
-```{r rr-plot, include = FALSE}
+
+```{r rr-plot, include = TRUE, fig.cap="90% Posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis) in blue.
+Simulated MPA effects associated with simulated response ratios matched to empirical response ratios in red.
+For response ratios, a value of zero indicates that biomass densities of targeted and non-targeted species are identical inside and outside MPAs, a value of 1 indicates that biomass densities of targeted species are 100% greater inside MPAs than outside.
+For MPA effect, a value of zero indicates that mean biomass densities are identical in the with- and without- MPA scenarios.
+A value of 1 indicates that mean biomass densities are 100% greater in the scenario with MPAs than the scenario without MPAs."}
 
@@
 
-\begin{figure*}%[tbhp]
-  \centering
-  \includegraphics[width=.9\linewidth]{figs/response_ratio_plot.png}
-  \caption{90% Posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis) in blue.
-Simulated MPA effects associated with simulated response ratios matched to empirical response ratios in red.
-For response ratios, a value of zero indicates that biomass densities of targeted and non-targeted species are identical inside and outside MPAs, a value of 1 indicates that biomass densities of targeted species are 100% greater inside MPAs than outside.
-For MPA effect, a value of zero indicates that mean biomass densities are identical in the with- and without- MPA scenarios.
-A value of 1 indicates that mean biomass densities are 100% greater in the scenario with MPAs than the scenario without MPAs.}
-  \label{rr-plot}
-\end{figure*}
-
 Given the potential unreliability of response ratios for the task, in an ideal world how would we empirically measure the regional effects of MPAs?
@@
 Targeted species in the Channel Islands, include commercially important fin-fish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
-Each of these targeted species was the subject of prior bio-economic modeling related to the effects of MPAs in southern California [@costello2010d;@rassweiler2014].
+Each of these targeted species was the subject of prior bio-economic modeling related to the effects of MPAs in southern California [@costello2010a;@rassweiler2014].
 It should be noted though that our analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*).
@@
 
-```{r mpa-effect, include=FALSE}
+```{r mpa-effect, include=TRUE, fig.cap = "Results of difference-in-difference regression estimating the regional effect of the northern Channel Island MPAs on mean total biomass densities of targeted species (difference in mean total biomass density of targeted species over time relative to expected levels using non-targeted species as a control).
+Grey distributions show posterior probability distribution of estimated MPA effect; red point is median estimated effect, thicker red section 50% credible interval, thinner red line 90% credible interval).
+Blue distributions in background show range of MPA effects produced by simulation model tuned to reflect the dynamics of the Northern Channel Island MPAs (black dashed line is median simulated value).
+Results are estimated in blocks of three years, including years greater than or equal to left-hand value and less than right-hand value."}
 
@@
 
-\begin{figure*}%[tbhp]
-  \centering
-  \includegraphics[width=.9\linewidth]{`r here::here("documents","figs","mpa_effect_plot.png")`}
-  \caption{Results of difference-in-difference regression estimating the regional effect of the northern Channel Island MPAs on mean total biomass densities of targeted species (difference in mean total biomass density of targeted species over time relative to expected levels using non-targeted species as a control).
-Grey distributions show posterior probability distribution of estimated MPA effect; red point is median estimated effect, thicker red section 50% credible interval, thinner red line 90% credible interval).
-Blue distributions in background show range of MPA effects produced by simulation model tuned to reflect the dynamics of the Northern Channel Island MPAs (black dashed line is median simulated value).
-Results are estimated in blocks of three years, including years greater than or equal to left-hand value and less than right-hand value.}
-  \label{mpa-effect}
-\end{figure*}
-
-Discussion {#discussion .unnumbered}
-------
+## Discussion {#discussion .unnumbered}
 
@@
 This analysis is to our knowledge the first empirical estimate of the regional effects of a large and iconic MPA network on a broad assemblage of targeted fin-fish.
-Despite the well enforced nature of these MPAs [xx], and the presence of a rigorous scientific monitoring program, we are unable to estimate a clear effect of protection.
+Despite the well enforced nature of these MPAs [XX is there any citation for this XX], and the presence of a rigorous scientific monitoring program, we are unable to estimate a clear effect of protection.
 While response ratios of targeted species have increased over time, providing evidence of effective *within MPA* protection, these response ratios are on their own an unreliable indicator of the regional conservation effects of the MPAs.
@@
 How can we explain these results?
-The estimated effects from 2003 to 2014 are well in line with theoretical expectations given the species and network in question.
-How might we explain the most recent results though?
 One explanation may lie in fleet dynamics.
 Much of the theoretical literature on MPAs assumes that all else being equal bigger reserves produce bigger conservation gains [@white2011 and references therin].
-However, to our knowledge all of these models simulate fleet dynamics through assumptions about fishing mortality rates (e.g.
+However, these models generally simulate fleet dynamics through fishing mortality rates (e.g.
 concentration of fishing mortality @halpern2004).
@@
 
-While we do not have access to fine scale fishing data from the Northern Channel Islands alone, reported catches for the species of interest in the Santa Barbara region in fact exhibit an overall downward trend in the years post reserve (see SI).
+While we do not have access to fine scale fishing data from the Northern Channel Islands alone, reported catches for the species of interest in the Santa Barbara region in fact exhibit an overall downward trend in the years post reserve (see Fig.SXX).
 We can most likely rule a negative MPA effect caused by a constant-catch fishing strategy then.
 What is another possible explanation for the recent downward trend in the estimated MPA effects?
-We do not see evidence to reject the parallel trends assumption of our difference-in-difference model in initial years of the MPAs, and we do not detect any clear evidence of interaction between the targeted and non-targeted species (e.g.
-depressed levels of non-targeted species as a result of predation from increased biomass densities of targeted species, see Fig.SXX).
-
-But, the Channel Islands region (and the entire West coast of the USA) experienced a dramatic ‘marine heatwave’ beginning in 2014 and persisting through 2016, resulting in part in extremely elevated water temperatures throughout the region  [@gentemann2017].
+The Channel Islands region (and the entire West coast of the USA) experienced a dramatic ‘marine heatwave’ beginning in 2014 and persisting through 2016, resulting in part in extremely elevated water temperatures throughout the region  [@gentemann2017].
 Biogeographic differences in the distributions of targeted and non-targeted species may confound the observed effects of MPAs.
@@
 As such, we hypothesize that the recent evidence for a decline in densities of targeted species is due to environmental conditions that disproportionately affect the targeted group (and not for example due to concentrated fishing pressure outside the reserves).
+This hypothesis is supported by the sharp declines in densities of targeted species seen inside the MPAs themselves.
+If the cause of recent declines was due to increases in fishing pressure we would expect to see substantial declines only in the fished areas (Fig.\@ref(fig:raw-density)).
 
-This hypothesis is supported by the sharp declines in densities of targeted species seen inside the MPAs themselves (see Fig.XX).
-Assuming that the reserves are sufficiently well enforced and large to protect targeted species biomass (an assumption supported by the response ratio results), if the cause of recent declines was due to increases in fishing pressure we would expect to see substantial declines only in the fished areas (Fig.
-XX).
-
-```{r}
+```{r raw-density, fig.cap = "Centered and scaled trends in biomass densities of targeted and non-targeted finfish included in our stury.
+Top panel shows trends across all sites, with smaller background lines showing trends for each indivdual species.
+Bottom two panels show aggregate biomass density trends outside and inside MPAs"}
 raw_biomass_density_plot
@@
 ```
-
-\begin{figure*}%[tbhp]
-  \centering
-  \includegraphics[width=.9\linewidth]{`r here::here("documents","figs","raw_biomass_density_plot.png")`}
-  \caption{Raw Biomass Trends}
-  \label{mpa-effect}
-\end{figure*}
 
@@
 While a large percentage increase, this is a small change in absolute biomass densities relative to the variance in the observation process itself (see Fig.S1 for a companion to Fig.\ref{conservation-effects} scaled by absolute population size).
-Beyond these challenges, the median estimated age at sexual maturity for the targeted species included in this study is `r median_tm` years, meaning that the span of this analysis represents less than three generations of MPA protection for half of the the measured species.
+In addition, the median estimated age at sexual maturity for the targeted species included in this study is `r median_tm` years, meaning that the span of this analysis represents less than three generations of MPA protection for half of the the measured species.
 Ongoing monitoring may yet reveal clearer effects.
@@
 As an added measure, we include scenarios in which the sampled species go through recruitment regimes, which may be positive for both targeted and non-targeted species, or positive for non-targeted species and negative for targeted species.
-We then used a simple Bayesian difference-in-difference regression on these simulated data, and estimated the percent error between the estimated and true MPA effect.
+We then used a simple Bayesian difference-in-difference regression on these simulated data, and estimated the percent error between the posterior probability distribution of the estimated MPA effect and true simulated MPA effect.
 
 While unbiased across simulations, the difference-in-difference model struggled severely when MPA effect sizes were less than 25% and the model was faced with observation and process errors (Fig.\@ref(fig:val-plot)).
-Even larger effect sizes commonly misestiamted the true MPA effect by 50% or more.
+Even larger effect sizes commonly misestimated the true MPA effect by 50% or more.
 Obtaining a mean absolute percent error (MAPE) of 25% or less across our simulated datasets required an MPA effect of at least 30%.
@@
 
-```{r, include = FALSE}
+```{r val-plot, include = TRUE, fig.cap = "Distribution of percent error in posterior estimate of MPA effect (y-axis) plotted against true simulated MPA effect (x-axis).
+Color shows concentration of simulations.
+Black line shows mean absolute percent error as a function of true simulated MPA effect."}
 
@@
 
-\begin{figure*}%[tbhp]
-  \centering
-  \includegraphics[width=.9\linewidth]{`r here::here("documents","figs","validation_plot.png")`}
-  \caption{Validation}
-  \label{val-plot}
-\end{figure*}
-
 This finding begs an important question: when should we expect to see MPA effects big enough to stand a reasonable chance of detection?
-To address this, we simulated `r n_distinct(outcomes$experiment)` MPA scenarios across a wide range of life histories, network designs, and fishing dynamics (see SI for a full description of scenarios).
+We simulated `r n_distinct(outcomes$experiment)` MPA scenarios across a wide range of life histories, network designs, and fishing dynamics (see SI for a full description of scenarios).
 Using these simulations, we can consider what we might expect the regional effects of an MPA network to be under a range of plausible circumstances.
@@
 Two of the most critical drivers of MPA performance are the size of the MPA and the degree of fishing pressure.
-Looking across these two variables, our simulations suggest if the MPA network covers 25% or more of a species range and/or pre-MPA depletion is greater 60% we might expect an effect size with our target MAPE.
-While recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, for near-shore commercial fin-fish many MPA networks are likely to cover areas more in line with the Channel Islands (20%) (Fig.\@ref(fig:XX)-A).
+Looking across these two variables, our simulations suggest if the MPA network covers 25% or more of a species range and/or pre-MPA depletion is greater than 60% we might expect an effect size with our target MAPE.
+While recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, for near-shore commercial fin-fish many MPA networks are likely to cover areas more in line with the Channel Islands (20%), and as such have effect sizes that may be difficult to detect (Fig.\@ref(fig:sim-effects)-A).
 
-These initial simulation results would seem to suggest some rather simple rules of thumb: put an MPA of sufficient size on a at least somewhat overexploited population and one can expect large and potentially detectable results.
-However, as we discussed earlier, the MPA literature highlights a large number of variables beyond simply size and fishing pressure that can affect performance.
-To address this, we can examine the variability in expected MPA effects across each of the two major axes (pre-MPA depletion and MPA size)  (Fig.\@ref(fig:XX)-B).
+These initial simulation results would seem to suggest some rather simple rules of thumb: put an MPA of sufficient size on an at least somewhat overexploited population and one can expect large and potentially detectable results.
+However, the MPA literature highlights a large number of variables beyond simply size and fishing pressure that can affect performance.
+To address this, we examined the variability in expected MPA effects across each of the two major axes (pre-MPA depletion and MPA size).
 As both MPA size and pre-MPA depletion increase, the potential for a large MPA effect increases.
-However, for both variables even at extremely large or extremely small values a wide range of MPA effects were possible (though as we might expect the effect of pre-MPA depletion, in other words fishing pressure, was much clearer than MPA size alone).
+However, for both variables even at extremely large or extremely small values a wide range of MPA effects were possible (though as we might expect the effect of pre-MPA depletion, in other words fishing pressure, was much clearer than MPA size alone) (Fig.\@ref(fig:sim-effects)-B).
 
-To put these results into context, the FAO estimates that 7% of the worlds fisheries with status estimates fall into the relatively unexploited category (roughly depletion less than 50%), 60% fall into the fully fished category (roughly depletion 50%-70%), and 33% fall into the heavily fished category (roughly depletion greater than 70%) [@fao2018, assuming a Bmsy/B0 ratio of roughly 40%, per @thorson2012], though works that include a broader range of fisheries estimate that 50% or more of stocks to fall into the heavily fished category [@costello2012;@rosenberg2017].
-Taken together then, the regional effects of MPA networks covering 25% or less of a species range may be difficult to detect in many placed with already well-managed fisheries, while for that size we might expect clearer effects in less-managed locations (though that of course ignores the complication of compliance with MPA regulations).
+To put these results into context, the FAO estimates that 7% of the worlds fisheries with status estimates fall into the relatively unexploited category (roughly depletion less than 50%), 60% fall into the fully fished category (roughly depletion 50%-70%), and 33% fall into the heavily fished category (roughly depletion greater than 70%) [@fao2018, assuming a Bmsy/B0 ratio of 40%, per @thorson2012], though works that include a broader range of fisheries estimate that 50% or more of stocks to fall into the heavily fished category [@costello2012;@rosenberg2018].
+The regional effects of MPA networks covering 25% or less of a species range may be difficult to detect in many places with already well-managed fisheries, while for that size we might expect clearer effects in less-managed locations (though that of course ignores the complication of compliance with MPA regulations).
 Within these broad guidelines though a wide range of outcomes are possible: as a starting place users can use the bio-economic MPA simulation model developed for this paper to explore potential outcomes for specific MPAs using an interactive web application available at [danovando.shinyapps.io/simmpa](https://danovando.shinyapps.io/simmpa/).
 
-```{r, include=FALSE}
+```{r sim-effects, include=TRUE, fig.cap = "Simulated MPA effect sizes as a function of percent of species' range inside MPA, and pre-MPA depletion.
+Pre-MPA depletion is a measure of fishing pressure, where 0 means that the population is unfished, and 1 means that the population is extinct in the time period immediately prior to MPA implementation.
+A) shows median MPA effects across range in MPA and pre-mpa depletin.
+Panel B shows distribution of simulations across range in MPA and pre-mpa depletion separately."}
 expected_mpa_effect_plot
@@
 
-\begin{figure*}%[tbhp]
-  \centering
-  \includegraphics[width=.9\linewidth]{`r here::here("documents","figs","expected_mpa_effect.png")`}
-  \caption{Expectation}
-  \label{expect-plot}
-\end{figure*}
-
-Conclusions {#conclusion .unnumbered}
-------
+## Conclusions {#conclusion .unnumbered}
 
@@
 Our results show that regional conservation benefits of MPAs are highly context dependent and in many circumstances, are likely to be small enough that they are nearly impossible to detect empirically.
-Indeed, this is exactly what we found in our empirical case study from the Channel Islands, California, USA.
+Indeed, this is exactly what we found in our empirical case study from the large and well-studied network of MPAs in the Channel Islands, California, USA.
 
@@
 While attributes such as MPA size and fishing pressure are important factors in determining the effects of protection, local fleet dynamics and the movement rates of adult fish can dramatically affect the outcomes of protected areas as well  (Fig.\ref{conservation-effects}).
-Far from being a simple tool with clear outcomes, the effects of MPAs can be highly context dependent, requiring - we would argue - a bio-economic model of at minimum the complexity presented here to help communities design and set expectations of MPAs at the tactical level.
+The effects of MPAs can be highly context dependent, requiring - we would argue - a bio-economic model of at minimum the complexity presented here to help communities design and set expectations of MPAs at the tactical level.
 While users may not be able to parameterize every aspect of a model such as that presented here, working with stakeholders to visualize the implications of, for example, different fleet responses to MPA implementation is a critical step in MPA design.
@@
 How then should stakeholders go about adaptively monitoring and managing MPAs?
-Simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can perhaps be tuned to focus on the species groups that have the highest chance of a detectable effect size [@nickols2019].
+Simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can perhaps be tuned to focus on the species groups that have the highest chance of a detectable effect size over the early years of the reserve [@nickols2019].
 Expanding data collection to include robust monitoring of spatio-temporal fleet dynamics may help assess the validity of control sites used in response ratios, support the direct inclusion of these fleet dynamics into statistical models, and allow managers to take into account potential negative interactions between MPAs and fleet dynamics such as those that may occur under constant-catch dynamics.
@@
 While the history of MPA science has made important strides in helping us understand the dynamics of protected areas, the future of MPA science must directly tackle the challenge of evaluating the performance of these MPAs at the regional scale, and adapting their design as needed to best achieve objectives.
-This is particularly true if communities are depending on MPAs as their fisheries management tool.
+This is particularly true if communities are depending on MPAs as their primary marine resource management tool.
 Commonly employed metrics such as response ratios may be applicable in some circumstances, but can have severe shortcomings as metrics of regional conservation effects.
@@
 Statistical approaches that explicitly address complications such as the spatial spillover effects of MPAs (such as the difference-in-difference approach used here) may give users an improved understanding of the performance of their MPAs, but even they may struggle when expected effect sizes are small.
-Clearly communicating what we should expect, and what we can detect,m from MPAs is critical in ensuring that MPAs play effective roles in fisheries management and marine conservation.
+Clearly communicating what we should expect, and what we can detect, from MPAs is critical in ensuring that MPAs play effective roles in fisheries management and marine conservation.
 
@@
 
-```{r ci-map, include = FALSE}
+```{r ci-map, include = TRUE, fig.cap="Map of study region; the Northern Channel Islands, California, USA. Colors show binned number of PISCO sampling events across the time period of our study."}
 
@@
 
-\begin{figure}%[tbhp]
-  \centering
-  \includegraphics[width=1\linewidth]{`r here::here("documents","figs","sample_location_plot.png")`}
-  \caption{Map of PISCO sampling locations within the the Channel Islands National Marine Sanctuary, California, USA.
-Grey boxes indicated MPAs, points sampling locations, with color representing whether the site is inside an MPA (blue) or not (red)}
-  \label{channel-islands}
-\end{figure}
-
  The difference-in-difference model used empirical kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands.
 A network of MPAs covering approximately 20% of the islands’ waters was put in place in 2003 as part of the California Marine Life Protection Act (MLPA) (see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of the MLPA).
-PISCO conducts visual SCUBA surveys at a large number of rocky-reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\ref{channel-islands}-\ref{pop-trends}).
-.
-The details of the monitoring program are described in XX
+PISCO conducts visual SCUBA surveys at a large number of rocky-reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\ref{fig:ci-map}).
+The details of the monitoring program are described in @caselle2015.
 
@@
 The simplified form of this model is
-XX adjust to GAMMA glm
 
 \begin{equation}
-  d_{i} = Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i},shape, scale)
+  d_{i} \sim Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + \mathbf{B^cX_i} + \mathbf{B^sS_i}},shape, scale)
 \label{eq:did}
@@
 where $d_i$ is the biomass density at observation *i*, *T* indicates whether the observation *i* is for a targeted ($T = 1$) or non-targeted ($T = 0$) species, and *MPA* marks whether observation *i* is in a pre MPA ($MPA = 0$) or post MPA ($MPA = 1$) state.
-is a vector of coefficients for additional control variables in matrix *X* such as water visibility and observer experience.
+**B^c^** is a vector of coefficients for additional control variables in matrix *X* such as water visibility and observer experience.
+**B^s^** is a vector of hierarchical coefficients for each sampling location **S** , clustered by island.
 Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment (*MPA*) on the treated (targeted species).
@@
 We do not detect any significant differences in the trends of the biomass densities of the targeted and non-targeted species in the years before the MPAs.
-XX
-
 With regards to the second assumption, all of the species in this empirical analysis exist within an ecosystem, and as such affect each other through mechanisms such as predation, competition, and habitat modification.
@@
 However, the longer MPAs are in place, the greater the possibility that substantial species interactions that can affect use of non-targeted species as a control may arise.
-
-<!-- \showmatmethods -->
-<!-- \showacknow -->
-
-&nbsp;
-&nbsp;
-&nbsp;

### 2020-03-27 documents/Untitled.Rmd
version sent to coauthors

@@
 csl: pnas.csl
-
+test: |
+  thing
 output:
@@
 linkcolor: blue
+toc: true
 ---
 
-$$d_{i} \sim Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + B\mathbf{X} + B,shape, scale)$$
+# hello
+
+$$d_{i} \sim Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + \mathbf{B_cX_i} + B_s\mathbf{S_i}},shape, scale)$$

### 2020-03-27 documents/cover-letter.Rmd
version sent to coauthors

@@
 title: "regional-mpa-cover-letter"
-author: "Dan"
 date: "2/23/2020"
-output: html_document
+output:
+  word_document: default
+  html_document: default
+bibliography: My Library.bib
 ---
@@
 
-##
+# Responses
+
+Insert flowery statements thanking reviewers and editors for all their insights.
+
+Core response: We have fundamentally restructured the paper to place much greater emphasis on the empirical results.
+This is to our knowledge the first empirical assessment of the regional effects of a large and iconic MPA network on a wide assemblage of commercially and economically important species.
+Our inability to estimate a clear effect in such a well-enforced and well-studied local is we feel a vitally important "null-result" both for scientists and MPA managers to be aware of.
+As such we feel that it warrants publication in PNAS.
+While our results are more focused on the empirical results, we explicitly link our results to the large body of MPA theoretical work, using it to help illustrate why we might expect such a null result and help managers understand when they might be able to detect an effect.
+etc.
+
+## Reviewer 2
+
+- *Yes, these are important questions but they are not new to ecology, economics, and fishery science.
+You are not the first to ask them and point out their importance; see articles at bottom for some examples.
+What are you doing to answer them that is better than all of the other papers that come before you?
+And why is your case study an exemplary place to answer the questions?
+The conclusions on line 709-728 are well-known in the MPA literature.*
+  - We have fundamentally restructured the paper based around the importance of the Channel Islands as a case study in measuring MPA effectiveness.
+We thank the reviewer for pointing out gaps in our literature, and we have incorporated the studies they recommend.
+While as the reviewer points out numerous other studies have considered the general theory of MPAs, we feel this is one of the first to pair robust empirical estimation of an until-now unexamined question with a thorough review and simulation of the existing theory.
+
+- *I would also be more cautious regarding your conclusion about your model predicts what should be expected in the data analysis. It could be simply luck that there is agreement between the two.*
+  - We have made the language clearer that this concordance could be simple coincidence.
+However, in light of recent concerns in the scientific literature around hypothesizing-after-results (HARKING), we feel that it is critically important to ground our theoretical results in *a priori* predictions.
+In addition, as we mention in the discussion, we encourage MPA scientists and stakeholders to undertake such simulation model to consider the plausibility of their results.
+As the reviewer says, empirical results that match theoretical predictions are not proof that the empirical methods are robust, but they do provide am additional and important line of evidence: empirical results that fail to match theoretical predictions need to be carefully considered.
+
+- *What about using the simulation model to demonstrate the validity of the technique used to identify the MPA effect used in the paper, which was pioneered in publication in 38?
+Or to understand whether the implication of the statement on line 539, that is, species ranges exceed the Island and your sampling domain, is important for measuring the impacts of MPAs outside of their boundaries?
+In sum, in its current form, I would move all of the simulation model to the SI (lines 131-252) and mention in the discussion section that these results are consistent with the predictions of a stylized single species model.*
+    - We have made clearer that our results build off the techniques used in @caselle2015, and made clearer the fundamental differences between what we use here and the methods utilized in that paper.
+XX I don't understand the "or to understand...." point at all, anyone else?
+XX.
+We have dramatically reduced the amount of the text dedicated to the simulation model.
+However, we have left some substantial portions within the paper in the discussion.
+While components of these points have been made before, we feel that having a clear reference explaining what we might expect the regional effects of MPAs to be is an important and useful contribution.
+
+- *The discussion on how to detect changes outside the MPA (lines 254-347) follows very closely the discussion in Ferraro et al (#41 in your list).
+It seems like you can remove most of this text and point the reader to that paper*
+  - We have greatly reduced the amount of text addressing response ratios, and restructured the response-ratio analysis to be linked explicitly with the empirical response ratios (in response to comments from the editor).
+We now use the empirical response ratios from the channel islands as motivation to consider what regional effects might plausibly produce those.
+
+- *To me the potential contribution of the paper starts on line 350, which is way too late in a paper of this size*
+  - We have restructured the paper so that this point arrives much earlier (line 81)
+
+- *Is part of assumption b) (line 372) that the dispersal characteristics of the fished and non-fished species need to be similar?
+I could envision a scenario where the two sets of species have different dispersal characteristics, non-fished have smaller dispersal ranges than fished species, for example.
+How does this impact your identification strategy?
+If the trends are similar regardless of the mechanistic process by which it occurs, does it matter for the empirics?
+More discussion on both assumptions (a) and (b) and what they imply about the biological mechanisms seems warranted.*
+  - No, that is not a necessary assumption of the model.
+All that is needed is a parallel trends assumption: even if we are only sampling a small portion of the non-targeted species range, so long as the portion that we sample has a parallel trend to the targeted group the model is valid.
+The mechanism of the trend does not matter, so long as it is consistent: as we observe in the paper, mechanisms may matter in that parallel trends are valid earlier but potentially not so later.
+XX I need to flesh this out a bit more in the methods XX
+
+- *The use of visual data surveys has, of course, problems that introduce errors in dependent and independent variables that could be correlated with features of the fished and non-fished species, e.g., size distribution.
+How can you be sure that this is not driving your results?
+Could you use the simulation model to investigate this?*
+  - Visual survey data are indeed challenging, but also among the most common forms of data for MPA analysis.
+So long as these errors are constant over time they should not bias our results, though clearly they have important implications for the estimated error in our model.
+We have run our model at a wide array of resolutions, from the updated aggregated form we use here to a complete hierarchical framework that attempts to control for errors in the observation process.
+The fundamental results remains the same: a general upwards trend, followed by a downward trend, with the clearest signal in the years around 2010.
+We would actually be more concerned about the robustness of our results if we had found a precise but small effect size: the finding instead of a uncertain results suggests that it is unlikely that we would get fundamentally different conclusions by including *more* error.
+We have expanded the simulation testing of the method a bit to help illustrate how violations of the model assumptions play out, and how they are related to the magnitude of the effect size
+
+- *(The answer to your question on line 709 that appears on lines 714-719 has been known in the literature on MPAs for over 20 years; see, e.g., papers by Martin Smith below.
+Given that your question is what "your results imply about the future of marine science", my answer is not very much.
+Again, you did a rigorous empirical case study that should be published.
+I am however finding it difficult to see a path to publication in PNAS.)*
+  - We hope that our restructuring around the novel empirical results resolves this concern.
+We would also politely point out that we asked what our results implied about the future of MPA science specifically, not marine science *writ large*
+
+- *Examples of relevant literature, you should expand your net to find other relevant literature to better distinguish your contribution*
+  - We thank the reviewer for pointing out these important omissions, in particular @white2011 and @nickols2019. We have included each of these references in their appropriate portions of the paper
+
+## Reviewer 1
+
+- *I think that greater emphasis needs to be placed up front, abstract, statement of significance, introduction..perhaps even title, that this model and paper only addresses fish biomass...not other 'conservation outcomes'.
+You do define this but not until page 2, line 148-154.
+I recommend moving this up.*
+  - Done
+
+- *I was surprised that there was no discussion about enforcement being a potential confounding factor.
+Currently the paper assumes that the Channel Islands MPAs are fully enforced, without poaching.
+This may or may not be the case and should be stated as such.
+You do address well other system wide impacts like water temperature (lines 557-8, 691+).
+Perhaps enforcement could be added to the discussion around line 711-13.*
+  - We now explicitly address enforcement XX can we get references on enforced nature of MPAs?
+
+- *Your discussion of the model also emphasizes that stock status prior to MPA designation is very key to outcomes; in the SI most of the species included for the Channel Island review have unknown status---which influences your discussion, and may better be described.*
+  - We make clearer that we iterate over unknowns such as this
+
+# Original Reviews
 
@@
 
-However, because the reviewers think the work is of interest and the editor concurs, we are willing to consider a resubmission that constructively addresses all of the concerns raised in the critiques.
+However, because the reviewers think the work is of interest and the editor concurs, we are willing to consider a re-submission that constructively addresses all of the concerns raised in the critiques.
 The paper would have to satisfy both the reviewers and the editor, and new criticisms could arise upon re-evaluation.
-We cannot guarantee success and will be unable to consider further resubmissions.
+We cannot guarantee success and will be unable to consider further re-submissions.
 
-Please submit the resubmission as you would a new paper.
-Your resubmission will be given a new tracking number.
-In your cover letter and the Comments to Editorial Staff, please specify that the paper is a resubmission and include your original tracking number.
-Additionally, please provide a "resubmission/revision" cover letter, including a point-by-point response to the editor and reviewers' comments.
+Please submit the re-submission as you would a new paper.
+Your re-submission will be given a new tracking number.
+In your cover letter and the Comments to Editorial Staff, please specify that the paper is a re-submission and include your original tracking number.
+Additionally, please provide a "re-submission/revision" cover letter, including a point-by-point response to the editor and reviewers' comments.
 If you have any questions, please contact our office.
@@
 10.1002/eap.1949
+
+# References

### 2020-03-27 documents/ovando-regional-effects-of-mpas-v2.Rmd
version sent to coauthors

@@
 ---
-title: The Regional Effects of Marine Protected Areas
-author: Daniel Ovando, Jennifer E. Caselle, Christopher Costello, Olivier Deschenes, Steven D. Gaines, Ray Hilborn, Owen Liu
+title: Beers Awarded for a New Title
+author: 
+  - Daniel Ovando
+  - Jennifer E. Caselle
+  - Christopher Costello
+  - Olivier Deschenes
+  - Steven D. Gaines
+  - Ray Hilborn
+  - Owen Liu
+abstract: |
+   Marine Protected Areas (MPAs) cover between 3-7% of the world's oceans, up from less than 1\% in the year 2000.
+The Convention on Biological Diversity calls for 10% of coastal waters to be protected inside MPAs by 2020, while the International Union for Conservation of Nature calls for 30\% protection by 2030.
+It is often clear that MPAs produce conservation benefits inside their borders, but many MPAs are also justified on the grounds that they also benefit the broader region outside their borders.
+The conservation effects of MPAs are most commonly evaluated using response ratios---measures of biomass densities inside their borders, relative to biomass densities in reference sites outside their borders.
+Studies of this nature have provided broad evidence that MPAs produce conservation benefits within their borders.
+While these gains can be vitally important, marine populations are rarely contained entirely within the borders of MPAs; therefore a critical question is not only do MPAs produce conservation gains inside their borders, but how do they affect the broader region in which they are located?
+The Channel Islands National Marine Sanctuary provides a clear example of this challenge.
+A network of MPAs covering roughly 20\% of the Islands’ waters was put in place in 2003, with a goal of providing regional conservation and fishery benefits.
+Response ratios from the region indicate that the Channel Island MPAs have increased biomass densities inside the MPAs.
+However, we are unable to find a clear effect of these same MPAs at the regional scale across multiple species.Building off of existing theory, we use a bio-economic simulation model to explain this discrepancy, and demonstrate under what conditions we likely can and cannot clearly estimate the regional conservation effects of MPAs.
+We show that MPA networks covering 25\% or less of a region are likely to produce regional increases in fish biomass densities on the order of 25\% or less, a meaningful effect but also one that can easily be overwhelmed by environmental shocks.
+Our results provide a novel assessment of the regional effects of a large and iconic Marine Protected Area network, and provide guidance for managers charged with monitoring and adapting MPAs.
 bibliography: My Library.bib
@@
 output:
+  bookdown::pdf_document2: default
   bookdown::word_document2: default
-  bookdown::pdf_document2: default
   bookdown::html_document2: default
@@
 linkcolor: blue
+toc: false
 header-includes:
@@
 knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
-                      cache = FALSE)
+                      cache = FALSE, fig.width = 3.42, fig.asp = 1.5)
 # knitr::opts_knit$set(root.dir = "/Users/danovando/PhD/zissou") # not entirely reprodcible hack to deal with knit child in dissertation.. not idealk
@@
 # run_name <- "v5.0"
+
+width_1col <- 3.42 # inches
+
+asp_1col <- 1.25
+
+width_2col <- 7 # inches
+
+asp_2col <-  .75
+
+dpi <- 900
+
+theme_1col <- hrbrthemes::theme_ipsum(base_size = 11,
+                                      axis_title_size = 12)
+
+theme_2col <- hrbrthemes::theme_ipsum(base_size = 12,
+                                        axis_title_size = 14)
 
@@
 
-zissou_theme <-
-  theme_ipsum(
-    base_size = 12,
-    axis_title_size = 12,
-    strip_text_size = 12,
-    base_family = "Fira Sans"
-  )
-
 testplot <- ggplot()
@@
 
-theme_set(zissou_theme)
+theme_set(theme_1col)
 
@@
 
-## abstract
-  Marine Protected Areas (MPAs) cover between 3-7% of the world's oceans, up from less than 1\% in the year 2000.
-The Convention on Biological Diversity calls for 10% of coastal waters to be protected inside MPAs by 2020, while the International Union for Conservation of Nature calls for 30\% protection by 2030.
-It is often clear that MPAs produce conservation benefits inside their borders, but many MPAs are also justified on the grounds that they also benefit the broader region outside their borders.
-The conservation effects of MPAs are most commonly evaluated using response ratios---measures of biomass densities inside their borders, relative to biomass densities in reference sites outside their borders.
-Studies of this nature have provided broad evidence that MPAs produce conservation benefits within their borders.
-While these gains can be vitally important, marine populations are rarely contained entirely within the borders of MPAs; therefore a critical question is not only do MPAs produce conservation gains inside their borders, but how do they affect the broader region in which they are located?
-The Channel Islands National Marine Sanctuary provides a clear example of this challenge.
-A network of MPAs covering roughly 20\% of the Islands’ waters was put in place in 2003, with a goal of providing regional conservation and fishery benefits.
-Response ratios from the region indicate that the Channel Island MPAs have increased biomass densities inside the MPAs.
-However, we are unable to find a clear effect of these same MPAs at the regional scale across multiple species.Building off of existing theory, we use a bio-economic simulation model to explain this discrepancy, and demonstrate under what conditions we likely can and cannot clearly estimate the regional conservation effects of MPAs.
-We show that MPA networks covering 25\% or less of a region are likely to produce regional increases in fish biomass on the order of 10\%-20\%, a meaningful effect but also one that can easily be overwhelmed by environmental shocks.
-Our results provide a novel assessment of the regional effects of a large and iconic Marine Protected Area network, and provide guidance for managers charged with monitoring and adapting MPAs.
-
 ## significance
@@
 
-Many recent agreements to expand MPAs (most proximately, the Convention on Biological Diversity’s Strategic Plan for Biodiversity, which calls for 10% of coastal waters to be protected inside MPAs by 2020 and the International Union for Conservation of Nature call for 30% by 2030) are based on the expectaion that well-designed MPAs will achieve benefits both within, and outside, their borders.
+Many recent agreements to expand MPAs (most proximately, the Convention on Biological Diversity’s Strategic Plan for Biodiversity, which calls for 10% of coastal waters to be protected inside MPAs by 2020 and the International Union for Conservation of Nature call for 30% by 2030) are based on the expectation that well-designed MPAs will achieve benefits both within, and outside, their borders.
 Despite these assumptions, our collective scientific understanding of the regional-scale conservation and fishery impacts of current and future MPAs is surprisingly limited.
@@
 As stakeholders around the world increasingly seek to use MPAs in marine resource management portfolios, it is critical that we develop a better understanding of the magnitude and drivers of regional-scale MPA effects.
-This is particuarly true in the context of a changing climate, in which "set it and forget it" management strategies are increasinly untenable; we need to be able to monitor the performance of MPAs in order to adapt them as dictated by a shifting environment.
+This is particularly true in the context of a changing climate, in which "set it and forget it" management strategies are increasingly untenable; we need to be able to monitor the performance of MPAs in order to adapt them as dictated by a shifting environment.
 The MPAs within the Northern Channel Islands National Marine Sanctuary, California, USA (which we will refer to as the Channel Islands from now on) provides an ideal case study to address this need.
@@
 Having established a theoretical grounding for the likely biomass density outcomes of the Channel Island MPAs, what empirical evidence do we have of the effects?
-@caslle2015 examined changes in biomass densities of species targeted by fishing activity inside and outside of the MPAs over time (a metric generally termed a "response ratio"), and compared these changes to the trends in non-targeted species.
+@caselle2015 examined changes in biomass densities of species targeted by fishing activity inside and outside of the MPAs over time (a metric generally termed a "response ratio"), and compared these changes to the trends in non-targeted species.
 They found a statistically significant increase in the response ratios of targeted species over time, and evidence that this increase is smaller in the non-targeted species.
@@
 
-```{r rr-plot, include = TRUE, fig.cap="90% Posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis) in blue.
+```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="90% Posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis) in blue.
 Simulated MPA effects associated with simulated response ratios matched to empirical response ratios in red.
@@
 
-response_ratio_plot
+temp_plot <- response_ratio_plot + 
+  theme_1col + 
+  theme(axis.title.x = element_textbox(size = 10))
 
 ggsave(
-  response_ratio_plot,
+  temp_plot,
   filename = here::here("documents","figs","response_ratio_plot.png"),
-  width = 4,
-  height = 6
+  width = width_1col,
+  height = width_1col * asp_1col,
+  dpi = dpi
 )
@@
 These empirical estimates are in line with (though in the upper range of) the outcomes that our simulation model suggests are plausible given the rough characteristics of the northern Channel Island MPAs.
+While this concordance between theory and empirics is by no means proof of the robustness of our results, it is an important line of evidence.
 However, in the subsequent years the trend reverses itself, and for the years 2015-2017 we once again see no clear effect of the MPAs (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`,
@@
 
-```{r mpa-effect, include=TRUE, fig.cap = "Results of difference-in-difference regression estimating the regional effect of the northern Channel Island MPAs on mean total biomass densities of targeted species (difference in mean total biomass density of targeted species over time relative to expected levels using non-targeted species as a control).
+```{r mpa-effect, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Results of difference-in-difference regression estimating the regional effect of the northern Channel Island MPAs on mean total biomass densities of targeted species (difference in mean total biomass density of targeted species over time relative to expected levels using non-targeted species as a control).
 Grey distributions show posterior probability distribution of estimated MPA effect; red point is median estimated effect, thicker red section 50% credible interval, thinner red line 90% credible interval).
@@
   filename = here::here("documents","figs","mpa_effect_plot.png"),
-  width = 6,
-  height = 4
+  width = width_2col,
+  height = width_2col * asp_2col,
+  dpi = dpi
 )
@@
 One explanation may lie in fleet dynamics.
-Much of the theoretical literature on MPAs assumes that all else being equal bigger reserves produce bigger conservation gains [@white2011 and references therin].
+Much of the theoretical literature on MPAs assumes that all else being equal bigger reserves produce bigger conservation gains [@white2011 and references therein].
 However, these models generally simulate fleet dynamics through fishing mortality rates (e.g.
@@
 Under a constant-catch scenario though, fishers have to fish much harder than before to get the same catch from a smaller part of the population, reducing the size structure of the population and subsequently causing net conservation loss under `r percent(1 - positive_outcomes$pp[positive_outcomes$fleet_model == "constant-catch"], accuracy = 1)` of our constant-catch simulations.
-This is an important and often overlooked possibility, especially as MPAs are increasinly implemented in quota-managed fisheries.
+This is an important and often overlooked possibility, especially as MPAs are increasingly implemented in quota-managed fisheries.
 
@@
 
-```{r raw-density, fig.cap = "Centered and scaled trends in biomass densities of targeted and non-targeted finfish included in our stury.
+```{r raw-density, fig.width = width_1col, fig.asp=asp_1col,fig.cap = "Centered and scaled trends in biomass densities of targeted and non-targeted finfish included in our stury.
 Top panel shows trends across all sites, with smaller background lines showing trends for each indivdual species.
 Bottom two panels show aggregate biomass density trends outside and inside MPAs"}
-raw_biomass_density_plot
+
+temp_plot <-
+  (total_trend_plot / (mpa_trend_plot)) &
+     theme_classic()  & theme(legend.position = "top", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), units = "lines"),
+                              panel.grid = element_blank()
+)
+
+temp_plot
 
 ggsave(
-  raw_biomass_density_plot,
+temp_plot,
 filename = here::here("documents","figs","raw_biomass_density_plot.png"),
-  width = 6,
-  height = 6
+  width = width_1col,
+  height = width_1col * asp_1col,
+    dpi = dpi
 )
@@
 Some species, such as California sheephead and blue rockfish were likely below target levels during the period [@alonzo2004;@dick2017], but projections based upon the overall average response across all species will likely suggest modest benefits even if a subset of species could experience much larger population gains.
-Harkening back to our simulations, we expect the average percentage difference in densities of targeted species with and without MPAs to be modest (Fig.\ref{fig:did-plot}).
+Harkening back to our simulations, we expect the average percentage difference in densities of targeted species with and without MPAs to be modest (Fig.\ref{fig:mpa-effect}).
 Effects of this size are likely to be challenging to detect empirically given the large natural variation of marine ecosystems (especially temperate reefs) and the observation error inherent in infrequent annual survey programs such as those provided by PISCO.
@@
 As an extreme example, an increase in biomass densities from .2kg/m^2^ to .4kg/m^2^ would be produce as a 100% increase.
-While a large percentage increase, this is a small change in absolute biomass densities relative to the variance in the observation process itself (see Fig.S1 for a companion to Fig.\ref{conservation-effects} scaled by absolute population size).
+While a large percentage increase, this is a small change in absolute biomass densities relative to the variance in the observation process itself (see Fig.S1 for a companion to Fig.\ref{fig:sim-effects} scaled by absolute population size).
 In addition, the median estimated age at sexual maturity for the targeted species included in this study is `r median_tm` years, meaning that the span of this analysis represents less than three generations of MPA protection for half of the the measured species.
@@
 
-```{r val-plot, include = TRUE, fig.cap = "Distribution of percent error in posterior estimate of MPA effect (y-axis) plotted against true simulated MPA effect (x-axis).
+```{r val-plot, include = TRUE,fig.width = width_1col, fig.asp=asp_1col, fig.cap = "Distribution of percent error in posterior estimate of MPA effect (y-axis) plotted against true simulated MPA effect (x-axis).
 Color shows concentration of simulations.
@@
 
-validation_plot
+temp_plot <- validation_plot + 
+  theme_1col + 
+   theme(legend.position = "top") + 
+    scale_fill_viridis(option = "C",
+                     name = "Sims(#)",
+                     guide = hgc <- guide_colorbar(frame.colour = "black",
+                                                   ticks.colour = "black",
+                                                   barwidth = unit(5,"lines")))
+
+temp_plot
 ggsave(
-  validation_plot,
+  temp_plot,
   filename = here::here("documents","figs","validation_plot.png"),
-  width = 6,
-  height = 4
+  width = width_1col,
+  height = width_1col * asp_1col,
+    dpi = dpi
+
 )
@@
 
-```{r sim-effects, include=TRUE, fig.cap = "Simulated MPA effect sizes as a function of percent of species' range inside MPA, and pre-MPA depletion.
+```{r sim-effects, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Simulated MPA effect sizes as a function of percent of species' range inside MPA, and pre-MPA depletion.
 Pre-MPA depletion is a measure of fishing pressure, where 0 means that the population is unfished, and 1 means that the population is extinct in the time period immediately prior to MPA implementation.
@@
 Panel B shows distribution of simulations across range in MPA and pre-mpa depletion separately."}
+
 expected_mpa_effect_plot
-
 ggsave(
@@
   filename = here::here("documents","figs","expected_mpa_effect.png"),
-  width = 6,
-  height = 4
+  width = width_2col,
+  height = width_2col * asp_2col,
+    dpi = dpi
+
 )
@@
 Our simulation model is by no means exhaustive (for example it ignores features such as species interactions, habitat effects, and climate feedback), but captures many of the core factors theorized to affect MPA performance.
-While attributes such as MPA size and fishing pressure are important factors in determining the effects of protection, local fleet dynamics and the movement rates of adult fish can dramatically affect the outcomes of protected areas as well  (Fig.\ref{conservation-effects}).
+While attributes such as MPA size and fishing pressure are important factors in determining the effects of protection, local fleet dynamics and the movement rates of adult fish can dramatically affect the outcomes of protected areas as well  (Fig.\ref{fig:sim-effects}).
 The effects of MPAs can be highly context dependent, requiring - we would argue - a bio-economic model of at minimum the complexity presented here to help communities design and set expectations of MPAs at the tactical level.
@@
 
-```{r ci-map, include = TRUE, fig.cap="Map of study region; the Northern Channel Islands, California, USA. Colors show binned number of PISCO sampling events across the time period of our study."}
+```{r ci-map, include = TRUE,fig.width=width_1col, fig.asp=1.5, fig.cap="Map of study region; the Northern Channel Islands, California, USA.
+Colors show binned number of PISCO sampling events across the time period of our study.", include=TRUE}
 
-sample_location_plot
+temp_plot <- sample_location_plot + 
+  theme(axis.text = element_blank())
 
+temp_plot
 ggsave(
-  sample_location_plot,
+ temp_plot,
   filename = here::here("documents","figs","sample_location_plot.png"),
-  width = 6,
-  height = 4
+  width = width_1col,
+  height = width_1col * asp_1col,
+    dpi = dpi
+
 )

### 2020-04-09 documents/ovando-regional-effects-of-mpas-v2.Rmd
sncing with abalone

@@
 ---
-title: Beers Awarded for a New Title
+title: Regional Effects of Marine Protected Area Networks Are Unclear
 author: 
@@
   - \usepackage{lineno}\linenumbers
+  - \usepackage{draftwatermark}
 ---
@@
                       cache = FALSE, fig.width = 3.42, fig.asp = 1.5)
-# knitr::opts_knit$set(root.dir = "/Users/danovando/PhD/zissou") # not entirely reprodcible hack to deal with knit child in dissertation.. not idealk
 
@@
 
+temp_plot
 # pdf(file =  "figs/density-ratio.png",width = 8, height = 5, useDingbats = TRUE)

### 2020-04-17 documents/ovando-regional-effects-of-mpas-v2.Rmd

@@
 output:
+  bookdown::word_document2: default
   bookdown::pdf_document2: default
-  bookdown::word_document2: default
   bookdown::html_document2: default

### 2020-04-18 documents/ovando-regional-effects-of-mpas-v2.Rmd
fixed some package version issues

@@
 abstract: |
-   Marine Protected Areas (MPAs) cover between 3-7% of the world's oceans, up from less than 1\% in the year 2000.
-The Convention on Biological Diversity calls for 10% of coastal waters to be protected inside MPAs by 2020, while the International Union for Conservation of Nature calls for 30\% protection by 2030.
-It is often clear that MPAs produce conservation benefits inside their borders, but many MPAs are also justified on the grounds that they also benefit the broader region outside their borders.
-The conservation effects of MPAs are most commonly evaluated using response ratios---measures of biomass densities inside their borders, relative to biomass densities in reference sites outside their borders.
+   Marine Protected Areas (MPAs) cover 3-7% of the world's oceans, up from less than 1\% in the year 2000.
+The Convention on Biological Diversity called for 10% of coastal waters to be protected inside MPAs by 2020, while the International Union for Conservation of Nature is now calling for 30\% protection by 2030.
+It is often clear that MPAs produce conservation benefits inside their borders, but many MPAs are also justified on the grounds that they may benefit the broader region outside their borders.
+The conservation effects of MPAs are most commonly evaluated using response ratios---measures of numerical or biomass densities inside their borders, relative to these metrics in reference sites outside their borders.
 Studies of this nature have provided broad evidence that MPAs produce conservation benefits within their borders.
@@
 The Channel Islands National Marine Sanctuary provides a clear example of this challenge.
-A network of MPAs covering roughly 20\% of the Islands’ waters was put in place in 2003, with a goal of providing regional conservation and fishery benefits.
+A network of MPAs covering 20\% of the Islands’ waters was put in place in 2003, with a goal of providing regional conservation and fishery benefits.
 Response ratios from the region indicate that the Channel Island MPAs have increased biomass densities inside the MPAs.
-However, we are unable to find a clear effect of these same MPAs at the regional scale across multiple species.Building off of existing theory, we use a bio-economic simulation model to explain this discrepancy, and demonstrate under what conditions we likely can and cannot clearly estimate the regional conservation effects of MPAs.
-We show that MPA networks covering 25\% or less of a region are likely to produce regional increases in fish biomass densities on the order of 25\% or less, a meaningful effect but also one that can easily be overwhelmed by environmental shocks.
+We are unable to find a clear effect of these same MPAs on biomass densities throughout the Channel Islands region.
+We use a bio-economic simulation model to explain this discrepancy, and demonstrate under what conditions we likely can and cannot clearly estimate the regional conservation effects of MPAs.
+MPA networks covering 25\% or less of a region are likely to produce regional increases in fish biomass densities on the order of 25\% or less, a meaningful effect but also one that can easily be overwhelmed by environmental shocks.
 Our results provide a novel assessment of the regional effects of a large and iconic Marine Protected Area network, and provide guidance for managers charged with monitoring and adapting MPAs.
@@
   Healthy marine ecosystems are critical to the well-being of the planet.
-Marine protected areas, parts of the oceans protected from human activities such as fishing, are increasingly being used in an effort to conserve and manage these ecosystems.
+Marine protected areas, parts of the oceans protected from human activities such as fishing, are increasingly being used to conserve and manage these ecosystems.
 Our study pairs theory with empirical methods to examine the regional effects of MPAs on conservation outcomes.
-Using simulations, we show that the regional effects of marine protected areas can be highly variable and dependent on human drivers.
-We find no statistically clear effect of a network of MPAs after 13 years of protection, which our simulation analysis suggests is actually to be expected.
+We show that the regional effects of marine protected areas can be highly variable and dependent on human drivers.
+We find no statistically clear effect of a network of MPAs after 14 years of protection, which our simulation analysis suggests is actually to be expected.
 Our results lay a foundation for future research on the design and monitoring of marine protected areas.
@@
   Funding for the simulation model was provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship.
+Empirical data collection was funded primarily by the David and Lucille Packard Foundation in support of the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) with additional funding from the California Ocean Protection Council and California SeaGrant.
 This study would not be possible without the work provided by PISCO divers over the years.
-DO thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments and technical support.
+DO thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments and technical support and special thanks to Kathyrn Davis Koehn and Avrey Parsons-Field.
 
@@
 This is particularly true in the context of a changing climate, in which "set it and forget it" management strategies are increasingly untenable; we need to be able to monitor the performance of MPAs in order to adapt them as dictated by a shifting environment.
-The MPAs within the Northern Channel Islands National Marine Sanctuary, California, USA (which we will refer to as the Channel Islands from now on) provides an ideal case study to address this need.
+The MPAs within the Channel Islands National Marine Sanctuary, California, USA (which we will refer to as the Channel Islands from now on) provides an ideal case study to address this need.
 A network of protected areas covering approximately 20% of the Islands' waters was put in place in 2003 [see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of these MPAs].
-This MPA network has subsequently been used as a model case study in protected area design around the world XX.
+This MPA network has subsequently been used as a model case study in protected area design around the world [@botsford2014 and references therein].
 
-We use data from the first 14 years of protection to provide what is to our knowledge the first empirical assessment of the regional effect of a large MPA network on a wide array of finfish species.
+We use data from the first 14 years of protection to provide what is to our knowledge the first empirical assessment of the regional effect (defined here as the effect of MPAs on the mean biomass density of targeted finfish both inside and outside of MPAs) of a large MPA network on a wide array of finfish species.
 In contrast to clear differences in biomass densities observed inside and outside of well protected MPAs both globally [@lester2009] and in the Channel Islands [@caselle2015] we are unable to detect a clear regional effect from the Channel Islands MPAs.
@@
 
-## Results
+## Results & Discussion
 
@@
 
-We define the regional conservation effects of MPAs as the change in total biomass densities of fish relative to the total biomass densities of fish that would have occurred without the MPAs (acknowledging that other outcomes such as increased biodiversity or resiliency are also important to conservation but are beyond the scope of this analysis).
+We define the regional conservation effects of MPAs as the change in mean total biomass densities of targeted finfish both inside and outside of MPAs, relative to the mean total biomass densities of targeted finfish inside and outside of MPAs that would have occurred without the MPAs (acknowledging that other outcomes such as increased biodiversity or resiliency are also important to conservation but are beyond the scope of this analysis).
+Consider a evenly distributed population that has 50% of its range protected by an MPA.
+Suppose that the MPA increase biomass densities inside the reserves by 20%, and by 0% outside the reserve.
+The regional effect of the MPA would be 10%.
+
 Numerous factors can affect the regional effects of MPAs.
@@
 
-## Discussion {#discussion .unnumbered}
-
 <!-- 1. Why are the results from the channel islands important and what do they tell us -->
@@
 While we do not have access to fine scale fishing data from the Northern Channel Islands alone, reported catches for the species of interest in the Santa Barbara region in fact exhibit an overall downward trend in the years post reserve (see Fig.SXX).
-We can most likely rule a negative MPA effect caused by a constant-catch fishing strategy then.
+We can most likely rule out a negative MPA effect caused by a constant-catch fishing strategy then.
 What is another possible explanation for the recent downward trend in the estimated MPA effects?
@@
 Biogeographic differences in the distributions of targeted and non-targeted species may confound the observed effects of MPAs.
-Many of the non-targeted species have warm thermal affinities and have increased in numbers since the heatwave [@freedman2019].
+Many of the non-targeted species in the Channel Islands have warm thermal affinities and have increased in numbers since the heatwave [@freedman2019].
 The targeted group is made up mostly of fishes with cold-water affinities, such as members of the genus *Sebastes*.
 As such, we hypothesize that the recent evidence for a decline in densities of targeted species is due to environmental conditions that disproportionately affect the targeted group (and not for example due to concentrated fishing pressure outside the reserves).
-This hypothesis is supported by the sharp declines in densities of targeted species seen inside the MPAs themselves.
+This hypothesis is supported by the declining trend in mean biomass densities of targeted species seen inside the MPAs themselves.
 If the cause of recent declines was due to increases in fishing pressure we would expect to see substantial declines only in the fished areas (Fig.\@ref(fig:raw-density)).
@@
 While we estimate a highly uncertain but overall positive effect at first, we are unable to detect a robust signal from 2012-2017.
-After 14 years of MPA protection we are left without a clear picture of the effect of the Channel Island MPA network on biomass densities of targeted fin-fish species.
+We believe that the dissapearance of increasing probabilities of positive MPA effects estimated by our model are likely driven by the marine heatwave experienced by the region beginning in 2014.
+After 14 years of MPA protection and following a large environmental perturbation we are left without a clear picture of the effect of the Channel Island MPA network on biomass densities of targeted fin-fish species.
 
-Our simulation results suggest though that we should not be surprised by this result.
-The Channel Islands MPAs cover approximately 20% of the waters in the Channel Islands, and while formal stock assessments are not available for many of the targeted species in our analysis, what evidence we have suggests that, as a group, these fish are on average not heavily overfished.
+Our simulation results suggest that we should not be surprised by this result.
+The Channel Islands MPAs cover 20% of the surface waters in the Channel Islands, and while formal stock assessments are not available for many of the targeted species in our analysis, what evidence we have suggests that, as a group, these fish are on average not heavily overfished.
 Some species, such as California sheephead and blue rockfish were likely below target levels during the period [@alonzo2004;@dick2017], but projections based upon the overall average response across all species will likely suggest modest benefits even if a subset of species could experience much larger population gains.
-Harkening back to our simulations, we expect the average percentage difference in densities of targeted species with and without MPAs to be modest (Fig.\ref{fig:mpa-effect}).
-Effects of this size are likely to be challenging to detect empirically given the large natural variation of marine ecosystems (especially temperate reefs) and the observation error inherent in infrequent annual survey programs such as those provided by PISCO.
+Our simulations suggest that the average percentage difference in densities of targeted species with and without MPAs to be modest (Fig.\ref{fig:mpa-effect}).
+Effects of this size are likely to be challenging to detect empirically given the large natural variation of marine ecosystems (especially temperate reefs) and the observation error inherent in infrequent (e.g.
+annual) common in marine systems.
 
@@
 
-As an additional complication, our simulations and difference-in-difference model examine percentage changes in biomass densities.
-As an extreme example, an increase in biomass densities from .2kg/m^2^ to .4kg/m^2^ would be produce as a 100% increase.
-While a large percentage increase, this is a small change in absolute biomass densities relative to the variance in the observation process itself (see Fig.S1 for a companion to Fig.\ref{fig:sim-effects} scaled by absolute population size).
+As an additional complication, our simulations and difference-in-difference model examine percentage changes in biomass densities (as oppose to changes in total population size).
+As an extreme example, an increase in biomass densities from 0.2kg/m^2^ to 0.4kg/m^2^ would be produce as a 100% increase.
+While a large percentage increase, this is a small change in biomass densities relative to the variance in the observation process itself (see Fig.S1 for a companion to Fig.\ref{fig:sim-effects} scaled by absolute population size).
 In addition, the median estimated age at sexual maturity for the targeted species included in this study is `r median_tm` years, meaning that the span of this analysis represents less than three generations of MPA protection for half of the the measured species.
@@
 
-Given the natural variability of marine ecosystems, and the strong challenges of obtaining accurate samples from oceanic environments, how large of an effect would an MPA network have to allow a difference-in-difference strategy such as this to be a reliable measure of MPA effects?
+Given the natural variability of marine ecosystems, and the large challenges of obtaining accurate samples from oceanic environments, how large of an effect would an MPA network have to have in order to allow a difference-in-difference strategy such as this to be a reliable measure of MPA effects?
 To provide guidance on this important question, we used our bio-economic model to simulate data from a range of scenarios with increasing MPA effect size, along with increasing degrees of observation error and natural recruitment variation.
 As an added measure, we include scenarios in which the sampled species go through recruitment regimes, which may be positive for both targeted and non-targeted species, or positive for non-targeted species and negative for targeted species.
-We then used a simple Bayesian difference-in-difference regression on these simulated data, and estimated the percent error between the posterior probability distribution of the estimated MPA effect and true simulated MPA effect.
+We then used a simple Bayesian difference-in-difference regression styled after the full model used here on these simulated data, and estimated the percent error between the posterior probability distribution of the estimated MPA effect and true simulated MPA effect.
 
@@
 However, since nearly any omission which one can think of would make an MPA effect harder to detect, not easier, these results serve as a useful floor for the likely difficulty in estimating MPA effects.
-In the context of the Channel Islands, given the potential effect size produced by out simulation model these results suggest that we might expect to be unable to detect a clear effect.
+In the context of the Channel Islands, given the potential effect size produced by our simulation model these results suggest that we might expect to be unable to detect a clear effect.
 
@@
 Color shows concentration of simulations.
-Black line shows mean absolute percent error as a function of true simulated MPA effect."}
+Black line shows mean absolute percent error (MAPE) as a function of true simulated MPA effect."}
 
@@
 Two of the most critical drivers of MPA performance are the size of the MPA and the degree of fishing pressure.
-Looking across these two variables, our simulations suggest if the MPA network covers 25% or more of a species range and/or pre-MPA depletion is greater than 60% we might expect an effect size with our target MAPE.
+Looking across these two variables, our simulations suggest if the MPA network covers 25% or more of a species range and pre-MPA depletion is greater than 60% we might expect an effect size with our target MAPE.
 While recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, for near-shore commercial fin-fish many MPA networks are likely to cover areas more in line with the Channel Islands (20%), and as such have effect sizes that may be difficult to detect (Fig.\@ref(fig:sim-effects)-A).
 
-These initial simulation results would seem to suggest some rather simple rules of thumb: put an MPA of sufficient size on an at least somewhat overexploited population and one can expect large and potentially detectable results.
-However, the MPA literature highlights a large number of variables beyond simply size and fishing pressure that can affect performance.
+These initial simulation results suggest some rather simple rules of thumb: put an MPA of sufficient size on an at least somewhat overexploited population and one can expect large and potentially detectable results.
+The MPA literature highlights a large number of variables beyond simply size and fishing pressure that can affect performance.
 To address this, we examined the variability in expected MPA effects across each of the two major axes (pre-MPA depletion and MPA size).
@@
 
-To put these results into context, the FAO estimates that 7% of the worlds fisheries with status estimates fall into the relatively unexploited category (roughly depletion less than 50%), 60% fall into the fully fished category (roughly depletion 50%-70%), and 33% fall into the heavily fished category (roughly depletion greater than 70%) [@fao2018, assuming a Bmsy/B0 ratio of 40%, per @thorson2012], though works that include a broader range of fisheries estimate that 50% or more of stocks to fall into the heavily fished category [@costello2012;@rosenberg2018].
+To put these results into context, the FAO estimates that 7% of the worlds fisheries with status estimates fall into the relatively unexploited category (roughly depletion less than 50%), 60% fall into the fully fished category (roughly depletion 50%-70%), and 33% fall into the heavily fished category (roughly depletion greater than 70%) [@fao2018, assuming a biomass at maximum sustainable yield to unfished biomass ratio of 40%, per @thorson2012], though works that include a broader range of fisheries estimate that 50% or more of stocks to fall into the heavily fished category [@costello2012;@rosenberg2018].
 The regional effects of MPA networks covering 25% or less of a species range may be difficult to detect in many places with already well-managed fisheries, while for that size we might expect clearer effects in less-managed locations (though that of course ignores the complication of compliance with MPA regulations).
-Within these broad guidelines though a wide range of outcomes are possible: as a starting place users can use the bio-economic MPA simulation model developed for this paper to explore potential outcomes for specific MPAs using an interactive web application available at [danovando.shinyapps.io/simmpa](https://danovando.shinyapps.io/simmpa/).
+Within these broad guidelines a wide range of outcomes are possible based on local fleet and fish dynamics: as a starting place users can use the bio-economic MPA simulation model developed for this paper to explore potential outcomes for specific MPAs using an interactive web application available at [danovando.shinyapps.io/simmpa](https://danovando.shinyapps.io/simmpa/).
 
@@
 Our results show that regional conservation benefits of MPAs are highly context dependent and in many circumstances, are likely to be small enough that they are nearly impossible to detect empirically.
-Indeed, this is exactly what we found in our empirical case study from the large and well-studied network of MPAs in the Channel Islands, California, USA.
+Indeed, this is exactly what we found in our empirical case study from the relatively large and well-studied network of MPAs in the Channel Islands, California, USA.
 
@@
 These results often find massively higher densities and biomass inside MPAs than outside [median biomass increases of  over 400%, @lester2009].
-But as suggested in @white2011, @ferraro2018, @larsen2019 and further demonstrated here, without careful attention to the design of control sites (accounting for example for the displacement of fishing effort by MPAs), response ratios may be highly unreliable estimators of regional MPA effects.
+But as suggested in @white2011, @ferraro2018, @larsen2019 and further demonstrated here, without careful attention to the design of control sites (e.g.
+accounting for the displacement of fishing effort by MPAs), response ratios may be highly unreliable estimators of regional MPA effects.
 When MPAs affect nearby control sites used in response ratios through biological spillover or concentration of fishing effort, it is entirely possible to for MPA to produce massive response ratios while simultaneously having minimal effects on the entire population partially protected by the MPAs.
@@
 Whenever possible monitoring programs should be implemented prior to MPA implementation to provide a pre-treatment benchmark.
+
 Non-equilibrium analyses also help set expectations for effect sizes over time [@nickols2019,Fig.\ref{did-plot}].
 Educating communities about the challenges of estimating the effects of MPAs can help set expectations, so that a lack of a clear effect is not necessarily viewed as a failure of the program, or large positive result based solely on response ratios as a clear sign of success and subsequent increased fishing opportunity.
-Rather, results and subsequent management actions (such as adaptation of MPA network design) must be considered in the context of reasonable expectations for the MPAs in question.
+Rather, results and subsequent management actions (such as adaptation of MPA networks) must be considered in the context of reasonable expectations given the size, age, and degree of enforcement of the MPAs in question, together with the ecological and economic dynamics of a given system.
 While this paper has focused on the conservation outcomes of MPAs, future work must also address the challenge of predicting and estimating the fishery impacts of protected areas.
@@
 Dependence on unreliable estimators of MPA effects may lead to stakeholders incorrectly attributing negative environmental shocks as MPA failures, or interpreting data arising from scorched-earth fishing outside MPAs as a conservation success.
+Both of these scenarios would hinder the ability of MPAs to serve as effective marine resource management tools.
 Bio-economic modeling can help frame community expectations, reducing the potential for a reduction in support if unrealistic conservation or fishery expectations are not realized.
@@
 Our main difference-in-difference model was fit using Stan [@carpenter2017] using the `rstanarm` package [@goodrich2020].
-Code needed to replicate results can be found [here](https://github.com/DanOvando/regional-effects-of-mpas).
+All materials needed to replicate results can be found [here](https://github.com/DanOvando/regional-effects-of-mpas).
 
@@
 See Table.S1 for a complete description of simulation states.
-The model consists of 25 patches with wrapped edges (picture the waters around a circular island).
+The model consists of 50 patches with wrapped edges (picture the waters around a circular island).
 For any one simulation we randomly pull a species and its associated life history from the `FishLife` [@thorson2017c] package in R.
 We pair these data with randomly selected values between 0.6 and 0.95 for Beverton-Holt steepness [as parameterized in @mace1994], as well as larval and adult dispersal rates.
-We also randomly assign whether adults have density dependent movement, as well as one of three potential types of density dependence [building off of @babcock2011]:
+We randomly assign whether adult fish preferentally move towards patches with lower relative densities, as well as one of three potential types of recruitment density dependence [building off of @babcock2011]:
 
@@
 
-The regression amounts to estimating the pre-post MPA difference in the biomass densities of targeted species net the same difference for non-targeted species in the Channel Islands.
+The regression amounts to estimating the pre-post MPA difference in the biomass densities of targeted species minus the same difference for non-targeted species in the Channel Islands.
 
@@
 See SI for further details of the estimation model.
+xx% of the samples in the raw data come from within MPAs, while the MPAs themselves cover roughly 20% of the surface waters of the channel islands.
+As such, we weight each observation such that within MPA data are assigned a total weight of 0.2 and outside-MPA data a total weight of 0.8.
 
@@
 With regards to the second assumption, all of the species in this empirical analysis exist within an ecosystem, and as such affect each other through mechanisms such as predation, competition, and habitat modification.
-We find it unlikely that these effects have had enough time to manifest in a meaningful way in the 13 years of post-MPA data used in our analysis [@babcock2010; @pershing2015a].
+We find it unlikely that these effects have had enough time to manifest in a meaningful way in the 14 years of post-MPA data used in our analysis [@babcock2010; @pershing2015a].
 

### 2020-04-18 functions/estimate_did.R

@@ -10,7 +10,7 @@ estimate_did <-
            data_source = "pisco",
            chains = 4,
            cores = 4,
-           iter = 2000) {
+           iter = 5000) {
     
     # data <- pisco_abundance_data
     # 
@@ -185,11 +185,28 @@ estimate_did <-
     #     prior = normal(0, 2)
     #   )
     # })
+    
+    weights <- did_data %>% 
+      mutate(mpa_area = ifelse(eventual_mpa, 0.2, 0.8)) %>% 
+      group_by(year, eventual_mpa) %>% 
+      summarise(n = n_distinct(site_side) * 2, # 2 because targeted and non-targeted
+                mpa_area = unique(mpa_area)) %>% 
+      mutate(weight = mpa_area / n)
+    
+    did_data <- did_data %>% 
+      left_join(weights, by = c("year", "eventual_mpa"))
+    
+    # test %>% 
+    #   group_by(year, eventual_mpa) %>% 
+    #   summarise(sum(weight))
+    # 
+    
     did_reg <- with(env, {
       stan_glmer(
         total_biomass_density ~ targeted * year_bins + (site_side - 1 |
                                                           region) + var_tex + var_surge + var_kelp + var_catch + var_temp +
-          regional_temp_dev,
+        regional_temp_dev,
+        weights = did_data$weight,
         data = did_data,
         cores = cores,
         chains = chains,
@@ -397,10 +414,27 @@ estimate_did <-
       env$chains <- chains
       
       env$iter <-  iter
+      #re-weight data in proportion to MPA area
+      weights <- did_data %>% 
+        mutate(mpa_area = ifelse(eventual_mpa, 0.2, 0.8)) %>% 
+        group_by(year, eventual_mpa) %>% 
+        summarise(n = n_distinct(site_side) * 2, # 2 because targeted and non-targeted
+                  mpa_area = unique(mpa_area)) %>% 
+        mutate(weight = mpa_area / n)
+      
+      did_data <- did_data %>% 
+        left_join(weights, by = c("year", "eventual_mpa"))
+      
+      # did_data %>%
+      #   group_by(year, eventual_mpa) %>%
+      #   summarise(sum(weight)) %>% 
+      #   View()
+
       
       did_reg <- with(env,{
         stan_glmer(
           total_biomass_density ~ targeted * year_bins + (1 |region) + var_kelp + var_catch,
+          weights = did_data$weight,
           data = did_data,
           cores = cores,
[…]

### 2020-04-18 make-regional-mpas.R

@@ -45,7 +45,7 @@ walk(functions, ~ here::here("functions", .x) %>% source()) # load local functio
 
 run_name <- 'v6.0'
 
-run_description <- "PNAS R and R with simplified DiD"
+run_description <- "PNAS R and R with simplified DiD and weighting"
 
 # the following analysis run the complete contents of "regional-effects-of-mpas". Each section depends on the out
 # outcomes of the prior section, but will load relevant saved files.
@@ -111,13 +111,13 @@ hgc <- guide_colorbar(frame.colour = "black",
                       barwidth = panel_width)
 
 # viridis is the default colour/fill scale for ordered factors
-(v <- ggplot(faithfuld) +
-    geom_tile(aes(waiting, eruptions, fill = density)))
-v + scale_fill_viridis_c()
-v + scale_fill_viridis_c(option = "plasma", guide = hgc) +
-
-  theme(legend.position = 
-          "top")
+# (v <- ggplot(faithfuld) +
+#     geom_tile(aes(waiting, eruptions, fill = density)))
+# v + scale_fill_viridis_c()
+# v + scale_fill_viridis_c(option = "plasma", guide = hgc) +
+# 
+#   theme(legend.position = 
+#           "top")
 
 fig_name <- "paper"
 
@@ -4279,9 +4279,9 @@ save(file = file.path(run_dir,"plots.RData"), list = plots)
 # knit paper --------------------------------------------------------------
 if (knit_paper == TRUE){
 
-  rmarkdown::render(here::here("documents","ovando-regional-effects-of-mpas.Rmd"), params = list(run_name = run_name))
+  rmarkdown::render(here::here("documents","ovando-regional-effects-of-mpas-v2.Rmd"), params = list(run_name = run_name))
 
-  rmarkdown::render(here::here("documents","ovando-regional-effects-of-mpas-si.Rmd"), params = list(run_name = run_name))
+  # rmarkdown::render(here::here("documents","ovando-regional-effects-of-mpas-si.Rmd"), params = list(run_name = run_name))
 
 }
 

### 2020-04-19 documents/ovando-regional-effects-of-mpas-v2.Rmd
confirmed knit with updated edits, turning to cover letter

@@
 While these gains can be vitally important, marine populations are rarely contained entirely within the borders of MPAs; therefore a critical question is not only do MPAs produce conservation gains inside their borders, but how do they affect the broader region in which they are located?
-The Channel Islands National Marine Sanctuary provides a clear example of this challenge.
+The Channel Islands National Marine Sanctuary provides a clear example of the importance and challenges of estimating regional MPA effects.
 A network of MPAs covering 20\% of the Islands’ waters was put in place in 2003, with a goal of providing regional conservation and fishery benefits.
@@
 
-#   sim_plot <- outcomes %>%
-#     mutate(rsize = plyr::round_any(mpa_size, 0.25),
-#            reff = plyr::round_any(mpa_effect,0.2)) %>%
-#     filter(rsize == 0.25, depletion < 0.8, depletion > 0.2,
-#            scientific_name %in% sim_life_history$scientific_name
-#            ) %>%
-#     mutate(year = years_protected + 2001) %>%
-#     filter(year %in% did_betas$year, years_protected >1) %>% 
-#     group_by(year, reff) %>% 
-#     count() %>% 
-#     group_by(year) %>% 
-#     mutate(sc = n / sum(n))
-# 
-#   did_plot <- did_betas %>%
-#     ggplot() +
-#     geom_tile(data = sim_plot, aes(year, reff, fill = sc), alpha = 0.75) +
-#     geom_vline(aes(xintercept = 2003), color = 'red', linetype = 2, size = 2) +
-#     geom_hline(aes(yintercept = 0)) +
-#     geom_pointrange(aes(
-#       year,
-#       y = exp(estimate) - 1,
-#       ymin = exp(lower) - 1,
-#       ymax = exp(upper) - 1
-#     ),
-#     size = 1.5
-#     ) +
-#     labs(x = "Year", y = "Targeted Trend Divergence") +
-#     scale_fill_viridis(guide = wide_hgc,
-#                        name = "% of Annual Sims",
-#                        labels = scales::percent_format(accuracy = 1),
-#                        option = "magma") + 
-#     theme(legend.position = "top",
-#         panel.grid.major.x = element_blank(),
-#         panel.grid.minor.x = element_blank()
-#         )
-#   
-# 
-# 
-# 
-# cip_data <- pisco_data %>%
-#   left_join(site_data, by = c("site","side")) %>%
-#   filter(is.na(eventual_mpa) == F) %>%
-#   filter(region %in% c("ANA", "SCI","SRI",'SMI'),
-#          classcode %in% top_species)
+ year_bins <-
+  seq(2003, max(pisco_data$year) + 1, by = 3)
 
-# process outcomes
-
+ sim_mpa_effects <- outcomes %>% 
+  filter(set == "ci",
+         final_depletion > 0,
+         f_v_m > .5) %>% 
+  mutate(year = years_protected + 2002) %>% 
+  filter(year <= max(pisco_data$year)) %>% 
+  mutate(binned_year = cut(year, year_bins, include.lowest = FALSE))  %>% 
+  filter(year >= 2003,
+         !is.na(binned_year)) %>% 
+  mutate(mpa_effect = pmin(mpa_effect,3))
 ```
@@
   Healthy marine ecosystems are critical to the well-being of the planet.
-Marine protected areas, parts of the oceans protected from human activities such as fishing, are increasingly being used to conserve and manage these ecosystems.
-Our study pairs theory with empirical methods to examine the regional effects of MPAs on conservation outcomes.
-We show that the regional effects of marine protected areas can be highly variable and dependent on human drivers.
-We find no statistically clear effect of a network of MPAs after 14 years of protection, which our simulation analysis suggests is actually to be expected.
-Our results lay a foundation for future research on the design and monitoring of marine protected areas.
+Marine protected areas (MPAs), parts of the oceans protected from human activities such as fishing, are increasingly being used to conserve and manage these ecosystems.
+Our study pairs theory with empirical methods to show that the regional conservation effects of marine protected areas can be highly variable and dependent on human drivers.
+We find no statistically clear effect of a large and well protected network of MPAs in the Channel Island, California, after 14 years of protection, which our simulation analysis suggests is actually to be expected.
+MPA practitioners must clearly communicate expectations from MPA networks, and plan monitoring and adaptations strategies that directly address the challenges of estimating regional conservation effects of protected areas.
 
@@
 Empirical data collection was funded primarily by the David and Lucille Packard Foundation in support of the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) with additional funding from the California Ocean Protection Council and California SeaGrant.
-This study would not be possible without the work provided by PISCO divers over the years.
-DO thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments and technical support and special thanks to Kathyrn Davis Koehn and Avrey Parsons-Field.
+This study would not be possible without the work provided by PISCO divers over the years, with special thanks to to Kathyrn Davis Koehn and Avrey Parsons-Field.
+DO thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments and technical support
 
@@
 
-Many recent agreements to expand MPAs (most proximately, the Convention on Biological Diversity’s Strategic Plan for Biodiversity, which calls for 10% of coastal waters to be protected inside MPAs by 2020 and the International Union for Conservation of Nature call for 30% by 2030) are based on the expectation that well-designed MPAs will achieve benefits both within, and outside, their borders.
+Recent agreements to expand MPAs (most proximately, the Convention on Biological Diversity’s Strategic Plan for Biodiversity, which calls for 10% of coastal waters to be protected inside MPAs by 2020 and the International Union for Conservation of Nature call for 30% by 2030) are based on the expectation that well-designed MPAs will achieve benefits both within, and outside, their borders.
 Despite these assumptions, our collective scientific understanding of the regional-scale conservation and fishery impacts of current and future MPAs is surprisingly limited.
@@
 
-As stakeholders around the world increasingly seek to use MPAs in marine resource management portfolios, it is critical that we develop a better understanding of the magnitude and drivers of regional-scale MPA effects.
+As stakeholders around the world seek to use MPAs in marine resource management portfolios, it is critical that we develop a better understanding of the magnitude and drivers of regional-scale MPA effects.
 This is particularly true in the context of a changing climate, in which "set it and forget it" management strategies are increasingly untenable; we need to be able to monitor the performance of MPAs in order to adapt them as dictated by a shifting environment.
-The MPAs within the Channel Islands National Marine Sanctuary, California, USA (which we will refer to as the Channel Islands from now on) provides an ideal case study to address this need.
+The MPAs within the Channel Islands National Marine Sanctuary, California, USA (which we will refer to as the Channel Islands from now on) provide an ideal case study to address this need.
 A network of protected areas covering approximately 20% of the Islands' waters was put in place in 2003 [see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of these MPAs].
@@
 Suppose that the MPA increase biomass densities inside the reserves by 20%, and by 0% outside the reserve.
-The regional effect of the MPA would be 10%.
+By our metric the regional conservation effect of the MPA would be 10%.
 
@@
 These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], larval dispersal patterns and the strength and timing of density dependence in the population (e.g.
-pre- or post-settlement, @burgess2014), how overfished the population would be without the MPA, and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006;@walters2000a;@white2011].
-In addition, even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through habitat and network effects [@gaines2010; @costello2010a].
+pre- or post-settlement, @burgess2014), the age and degree of enforcement [@edgar2014],  how overfished the population would be without the MPA, and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006;@walters2000a;@white2011].
+In addition, even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through network effects [@gaines2010; @costello2010a].
 
@@
 The MPAs in this study cover 21% of the the surface waters of the Northern Channel Islands, a region spanning roughly 90km east to west.
-While detailed dispersal studies are not available for all of the species covered by our study, what information we have suggest that while adults of some of these species are likely to exhibit site fidelity [@love2011], larvae are likely distributed beyond the Channel Islands [@rassweiler2012].
+Detailed dispersal studies are not available for all of the species covered by our study, but what information we have suggest that while adults of some of these species are likely to exhibit site fidelity [@love2011], larvae are likely distributed beyond the Channel Islands [@rassweiler2012].
 Therefore, if we assume that the population of the targeted species is at least equal to the extent of the Channel Islands, we can assume that at most the MPAs cover roughly 20% of the targeted populations' ranges.
@@
 
+```{r}
+median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected> 13])
+```
+
 We integrated these network design features, life history traits, and exploitation histories into a spatially explicit bio-economic simulation model, and used this model to generate expected outcomes for the Channel Islands MPAs.
-Our results suggest that while a wide range of outcomes are plausible, from 0% to upwards of 200%, across the majority of our simulations the effect size was 25\% or less, with a median simulated effect size of roughly 20\%; a potentially small value given the natural variability and sampling challenges of marine environments.
+Our results suggest that while a wide range of outcomes are plausible, from 0% to upwards of 200%, the a median simulated effect size for a stylized version of the Channel Islands after over ten years of protection was roughly `r percent(median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected > 10]))`; a potentially small value from the perspective of empirical detectability given the natural variability and sampling challenges of marine environments.
 
@@
 In the case of the Channel Islands MPAs, control sites are often located within a few kilometers of an MPA, suggesting that they are susceptible to biological spillover and concentration of fishing effort resulting from the MPAs.
-Given these complications, we can certainly interpret the response ratios reported in @caselle2015, and updated here, as evidence that the MPAs are providing effective protection of targeted fish biomass inside their borders, particularly in light of the general increase in response ratios over time.
+Given these complications, we can interpret the response ratios reported in @caselle2015, and updated here, as evidence that the MPAs are providing effective protection of targeted fish biomass inside their borders, particularly in light of the general increase in response ratios over time.
 However, it is unclear how well these response ratio results serve as in indicator of overall regional effects of the MPAs.
 
-To illustrate this problem, we used our bio-economic model to simulate response ratio trajectories for species and MPA coverage representative of the Channel Islands (see methods and SI for details of model structure).
+To illustrate this problem, we simulated response ratio trajectories for species and MPA coverage representative of the Channel Islands (see methods and SI for details of model structure).
 These simulations cover the life-histories and MPA sizes seen in the Channel Islands, but vary in key unknowns such as the actual degree of fishing pressure, the timing of density dependence, and the fishing fleets response to the MPAs.
@@
 
-As reported in @caselle2015, we see an increasing trend in response ratios of targeted species over time (Fig.\@ref(fig:rr-plot)).
-This provides strong evidence that the MPAs are providing protection for targeted species within their borders.
-However, our simulation results show that the response ratio trends we observe in the data could plausibly be produced by a wide range of regional MPA effects (Fig.\@ref(fig:rr-plot)).
+The response ratio trends we observe in the data could plausibly be produced by a wide range of regional MPA effects (Fig.\@ref(fig:rr-plot)).
 Response ratios well over one were associated with regional MPA effects generally less than 25%, and many simulations produced large response ratios but regional MPA effects close to 0% This can occur if for example fishing pressure is only moderate, adult movement is low, larval dispersal is high, and displaced fishing effort concentrates around the border of the MPAs.
+The response ratio results previously published for this region do not serve as sufficient evidence for the regional conservation effects of the Channel Islands MPA network.
 
@@
 Included non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species).
-We used a Bayesian difference-in-difference regression to estimate any difference in mean total biomass densities of fin-fish species targeted by fishing effort (i.e., those potentially affected by an MPA) and those species not targeted by fishing before and after MPA implementation [see @larsen2019 for a succinct explanation of difference-in-difference estimators in ecological settings].
+We used a Bayesian difference-in-difference regression to estimate any difference in mean total biomass densities of fin-fish species targeted by fishing effort (i.e., those potentially affected by an MPA) and those species not targeted by fishing before and after MPA implementation [see @larsen2019 for an explanation of difference-in-difference estimators in ecological settings].
 The result of this regression is, conditional on the assumptions of the model, an estimate of the effect of the MPAs on the mean total biomass densities of targeted species throughout the Channel Islands.
@@
 
-As conventional wisdom and theory would suggest, over the first three years of implementation (2003-2006), the effects of the MPAs are unclear, with support for a small negative effect to a substantially positive effect, with much higher probability of a small positive effect (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`,
+Consistent with MPA theory, over the first three years of implementation (2003-2006), the effects of the MPAs are unclear, with support for a small negative effect to a substantially positive effect, with much higher probability of a small positive effect (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`,
 90\% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`
@@
 
-```{r mpa-effect, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Results of difference-in-difference regression estimating the regional effect of the northern Channel Island MPAs on mean total biomass densities of targeted species (difference in mean total biomass density of targeted species over time relative to expected levels using non-targeted species as a control).
+```{r mpa-effect, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Results of difference-in-difference regression estimating the regional effect of the Channel Island MPAs on mean total biomass densities of targeted species (difference in mean total biomass density of targeted species over time relative to expected levels using non-targeted species as a control).
 Grey distributions show posterior probability distribution of estimated MPA effect; red point is median estimated effect, thicker red section 50% credible interval, thinner red line 90% credible interval).
-Blue distributions in background show range of MPA effects produced by simulation model tuned to reflect the dynamics of the Northern Channel Island MPAs (black dashed line is median simulated value).
+Blue distributions in background show range of MPA effects produced by simulation model tuned to reflect the dynamics of the Channel Island MPAs (black dashed line is median simulated value).
 Results are estimated in blocks of three years, including years greater than or equal to left-hand value and less than right-hand value."}
@@
 
-The network of MPAs implemented in the Northern Channel Islands in 2003 provides an ideal opportunity to examine the regional effects of protected area networks.
+<!-- The network of MPAs implemented in the Northern Channel Islands in 2003 provides an ideal opportunity to examine the regional effects of protected area networks.
 This analysis is to our knowledge the first empirical estimate of the regional effects of a large and iconic MPA network on a broad assemblage of targeted fin-fish.
-Despite the well enforced nature of these MPAs [XX is there any citation for this XX], and the presence of a rigorous scientific monitoring program, we are unable to estimate a clear effect of protection.
+Despite the well enforced nature of these MPAs, and the presence of a rigorous scientific monitoring program, we are unable to estimate a clear effect of protection.
 While response ratios of targeted species have increased over time, providing evidence of effective *within MPA* protection, these response ratios are on their own an unreliable indicator of the regional conservation effects of the MPAs.
@@
 For the most recent time period, we can detect no clear positive or negative regional effect of the Channel Island MPAs on biomass densities of targeted fin-fish (Fig.\@ref(fig:mpa-effect)).
+-->
 
-How can we explain these results?
+How can we explain the lack of a clear regional MPA effect after over ten years of protection?
 One explanation may lie in fleet dynamics.
@@
 
-While we do not have access to fine scale fishing data from the Northern Channel Islands alone, reported catches for the species of interest in the Santa Barbara region in fact exhibit an overall downward trend in the years post reserve (see Fig.SXX).
+While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region in fact exhibit an overall downward trend in the years post reserve (see Fig.SXX).
 We can most likely rule out a negative MPA effect caused by a constant-catch fishing strategy then.
@@
 The targeted group is made up mostly of fishes with cold-water affinities, such as members of the genus *Sebastes*.
-As such, we hypothesize that the recent evidence for a decline in densities of targeted species is due to environmental conditions that disproportionately affect the targeted group (and not for example due to concentrated fishing pressure outside the reserves).
+We hypothesize that the recent evidence for a decline in densities of targeted species is due to environmental conditions that disproportionately affect the targeted group (and not for example due to concentrated fishing pressure outside the reserves).
 This hypothesis is supported by the declining trend in mean biomass densities of targeted species seen inside the MPAs themselves.
@@
 We believe that the dissapearance of increasing probabilities of positive MPA effects estimated by our model are likely driven by the marine heatwave experienced by the region beginning in 2014.
-After 14 years of MPA protection and following a large environmental perturbation we are left without a clear picture of the effect of the Channel Island MPA network on biomass densities of targeted fin-fish species.
+After 14 years of MPA protection and following a large environmental perturbation we are left without a clear picture of the effect of the Channel Island MPA network on biomass densities of targeted fin-fish species from either response ratios or the difference-in-difference model.
 
@@
 As an additional complication, our simulations and difference-in-difference model examine percentage changes in biomass densities (as oppose to changes in total population size).
-As an extreme example, an increase in biomass densities from 0.2kg/m^2^ to 0.4kg/m^2^ would be produce as a 100% increase.
-While a large percentage increase, this is a small change in biomass densities relative to the variance in the observation process itself (see Fig.S1 for a companion to Fig.\ref{fig:sim-effects} scaled by absolute population size).
+An increase in biomass densities from 0.02kg/m^2^ to 0.04kg/m^2^ would be produce as a 100% increase.
+While a large percentage effect, this is a small change in biomass densities relative to the variance in the observation process itself (see Fig.S1 for a companion to Fig.\ref{fig:sim-effects} scaled by absolute population size).
 In addition, the median estimated age at sexual maturity for the targeted species included in this study is `r median_tm` years, meaning that the span of this analysis represents less than three generations of MPA protection for half of the the measured species.
@@
 While unbiased across simulations, the difference-in-difference model struggled severely when MPA effect sizes were less than 25% and the model was faced with observation and process errors (Fig.\@ref(fig:val-plot)).
-Even larger effect sizes commonly misestimated the true MPA effect by 50% or more.
-Obtaining a mean absolute percent error (MAPE) of 25% or less across our simulated datasets required an MPA effect of at least 30%.
-This is merely an illustrative exercise, omitting clearly critical factors such as detection probability and sampling strategy.
+Even models fir to data generated from large effect sizes commonly misestimated the true MPA effect by 50% or more.
+Obtaining a mean absolute percent error (MAPE) of 25% or less across our simulated datasets required a regional MPA effect of at least 30%.
+This is merely an illustrative exercise, omitting critical factors such as detection probability and sampling strategy.
 However, since nearly any omission which one can think of would make an MPA effect harder to detect, not easier, these results serve as a useful floor for the likely difficulty in estimating MPA effects.
-In the context of the Channel Islands, given the potential effect size produced by our simulation model these results suggest that we might expect to be unable to detect a clear effect.
+In the context of the Channel Islands, given the potential effect size produced by our simulation model these results suggest that we might expect to be unable to precisely estimate the true regional effect of the MPAs.
 
@@
 We simulated `r n_distinct(outcomes$experiment)` MPA scenarios across a wide range of life histories, network designs, and fishing dynamics (see SI for a full description of scenarios).
-Using these simulations, we can consider what we might expect the regional effects of an MPA network to be under a range of plausible circumstances.
-
 Suppose that we are willing to tolerate a MAPE of 25%.
@@
 Two of the most critical drivers of MPA performance are the size of the MPA and the degree of fishing pressure.
-Looking across these two variables, our simulations suggest if the MPA network covers 25% or more of a species range and pre-MPA depletion is greater than 60% we might expect an effect size with our target MAPE.
-While recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, for near-shore commercial fin-fish many MPA networks are likely to cover areas more in line with the Channel Islands (20%), and as such have effect sizes that may be difficult to detect (Fig.\@ref(fig:sim-effects)-A).
+Looking across these two variables, if the MPA network covers 25% or more of a species range and pre-MPA depletion is greater than 60% we might expect an effect size with our target MAPE.
+While recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, for near-shore commercial fin-fish many MPA networks are likely to cover areas more in line with the Channel Islands (20%), and as such have regional effect sizes that may be difficult to detect (Fig.\@ref(fig:sim-effects)-A).
 
@@
 Our results show that regional conservation benefits of MPAs are highly context dependent and in many circumstances, are likely to be small enough that they are nearly impossible to detect empirically.
-Indeed, this is exactly what we found in our empirical case study from the relatively large and well-studied network of MPAs in the Channel Islands, California, USA.
+This is exactly what we found in our empirical case study from the relatively large and well-studied network of MPAs in the Channel Islands, California, USA.
 
@@
 
-While this does not mean that all MPAs will face similar challenges in estimating their effects, our results in the rigorously studied and iconic Channel Islands Marine Protected Area network make clear that in many instances empirically detecting a clear regional effect of MPAs may not be possible.
+While this does not mean that all MPAs will face similar challenges in estimating their effects, our results in the relatively large, well-enforced, and rigorously studied Channel Islands Marine Protected Area network make clear that in many instances empirically detecting a clear regional effect of MPAs may not be possible.
 How then should stakeholders go about adaptively monitoring and managing MPAs?
@@
 
+Stock assessment models [as described in @hiborn1992b] that account for the population within the MPAs [field2006a] may be able to answer the management relevant question of whether fishing mortality rates and biomass levels are in-line with management objectives [e.g.
+@nickols2019].
+However, such an approach does not necessarily shed light on whether the MPAs themselves caused the estimated state of the population, and of course are highly data intensive, potentially restricting our ability to provide stock-assessment based inference of MPA outcomes for broad arrays of targeted species.
+
 Non-equilibrium analyses also help set expectations for effect sizes over time [@nickols2019,Fig.\ref{did-plot}].
-Educating communities about the challenges of estimating the effects of MPAs can help set expectations, so that a lack of a clear effect is not necessarily viewed as a failure of the program, or large positive result based solely on response ratios as a clear sign of success and subsequent increased fishing opportunity.
+Educating communities about the challenges of estimating the effects of MPAs can help set expectations, so that a lack of a clear effect is not necessarily viewed as a failure of the program, or large positive result based solely on response ratios as a clear sign of success and subsequent relaxation of other fishery management strategies.
 Rather, results and subsequent management actions (such as adaptation of MPA networks) must be considered in the context of reasonable expectations given the size, age, and degree of enforcement of the MPAs in question, together with the ecological and economic dynamics of a given system.
@@
 As the number and size of global MPA networks increase, it is critical that we both set appropriate expectations for their outcomes, and plan how we will monitor the performance of these protected areas over time.
-While the history of MPA science has made important strides in helping us understand the dynamics of protected areas, the future of MPA science must directly tackle the challenge of evaluating the performance of these MPAs at the regional scale, and adapting their design as needed to best achieve objectives.
+While the history of MPA science has made important strides in helping us understand the dynamics of protected areas, the future of MPA science must directly tackle the challenge of evaluating the performance of these MPAs at the regional scale, a task which has to date not been widely addressed.
 This is particularly true if communities are depending on MPAs as their primary marine resource management tool.
@@
 Dependence on unreliable estimators of MPA effects may lead to stakeholders incorrectly attributing negative environmental shocks as MPA failures, or interpreting data arising from scorched-earth fishing outside MPAs as a conservation success.
-Both of these scenarios would hinder the ability of MPAs to serve as effective marine resource management tools.
+Both of these scenarios would hinder the ability of MPAs to serve as effective marine resource management tools at scale.
 Bio-economic modeling can help frame community expectations, reducing the potential for a reduction in support if unrealistic conservation or fishery expectations are not realized.
@@
 Readers can explore the functionality of the model using an online tool available [here](https://danovando.shinyapps.io/simmpa/).
-See Table.S1 for a complete description of simulation states.
 The model consists of 50 patches with wrapped edges (picture the waters around a circular island).
@@
 We pair these data with randomly selected values between 0.6 and 0.95 for Beverton-Holt steepness [as parameterized in @mace1994], as well as larval and adult dispersal rates.
-We randomly assign whether adult fish preferentally move towards patches with lower relative densities, as well as one of three potential types of recruitment density dependence [building off of @babcock2011]:
+We randomly assign whether adult fish preferentially move towards patches with lower relative densities, as well as one of three potential types of recruitment density dependence [building off of @babcock2011]:
 
@@
 In the second, we assume that MPAs are placed in preferentially better habit (unfished recruitment is four times greater inside MPA locations).
-In the third, we allow for scenarios in which MPAs are placed in hotspots of larval dispersal.
-In this scenario, patches in which MPAs will be placed have larval dispersal rates four times greater than patches that do not become MPAs.
+In the third, we allow for scenarios in which MPAs are placed in sources of larval dispersal.
 
@@
 Along with the fleet dynamics model, each simulation is assigned a random fleet dispersal scenario: uniform dispersal (where the total effort of the fleet is divided evenly among all open patches), catch dispersal (where the total effort of the fleet is divided according to the catchable biomass in each available patch), and profit dispersal (where the total effort of the fleet is divided according to the profit per unit effort in each available patch).
+Fishing effort that occurred inside MPAs prior to closer can either leave the fishery, or be distributed to the patches outside the MPAs.
 
-Lastly each simulation is assigned an MPA scenario, defined by the number and size of MPAs, the placement of those MPAs, and the year that the MPAs are put in place.
+Each simulation is assigned an MPA scenario, defined by the number and size of MPAs, the placement of those MPAs, and the year that the MPAs are put in place.
 Each simulation starts the population off at unfished equilibrium and then beings to apply the fleet model.
 The MPAs are then placed during the randomly selected start year, allowing some runs to explore how the early dynamics of the MPA play out when the fishery and population they are placed on is not already at equilibrium.
-Fishing effort in displaced by MPAs can either concentrate outside or leave the fishery.
 Each simulation is run to equilibrium with and without the selected MPA strategy (holding all else constant).
-We then measure the difference in biomass densities each time step in the scenario with and without the MPAs to calculate the conservation and fishery effects of the MPAs over time.
+We then measure the difference in biomass densities each time step in the scenario with and without the MPAs to calculate the regional conservation effect of the MPAs over time.
 
@@
 The key assumptions of the difference-in-difference model are that a) within the time-frame of the model there are no significant interaction effects between the targeted and non-targeted species (which in fact we do not detect, see SI), and b) that in the absence of the MPAs both the targeted and non-targeted groups of species would have exhibited similar trends in densities.
-The advantage to this approach is that given the time-frame of the model (14 years with MPAs), we believe that spillover effects of MPA placement on the non-targeted species are likely to be much less severe than the effects of biological spillover and fishing concentration that may bias the performance of estimators such as a response ratios.
 
@@
 See SI for further details of the estimation model.
-xx% of the samples in the raw data come from within MPAs, while the MPAs themselves cover roughly 20% of the surface waters of the channel islands.
+`r percent(mean(used_data$eventual_mpa))` of the samples in the raw data come from within MPAs, while the MPAs themselves cover 20% of the surface waters of the channel islands.
 As such, we weight each observation such that within MPA data are assigned a total weight of 0.2 and outside-MPA data a total weight of 0.8.
@@
 While the parallel trends assumption cannot be formally proven, we can examine its validity using the data from the years before the MPAs were put in place in 2003.
-We do not detect any significant differences in the trends of the biomass densities of the targeted and non-targeted species in the years before the MPAs.
+We do not detect any significant differences in the trends of the biomass densities of the targeted and non-targeted species in the years before the MPAs (Fig.SXX).
 With regards to the second assumption, all of the species in this empirical analysis exist within an ecosystem, and as such affect each other through mechanisms such as predation, competition, and habitat modification.

### 2020-04-19 documents/reviewer-responses.Rmd
version sent to coauthors

---
title: "Reviewer Responses for 'Unclear Regional Effects of a Large Marine Protected Area Network' (previously 'The Regional Effects of MPAs')"
author: 
  - Daniel Ovando
  - Jennifer E. Caselle
  - Christopher Costello
  - Olivier Deschenes
  - Steven D. Gaines
  - Ray Hilborn
  - Owen Liu
output:
  word_document: default
  html_document: default
  pdf_document: default
bibliography: My Library.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reviewer Responses

We thank the two reviewers and the Editor for their constructive responses. We have fundamentally restructured the paper to place much greater emphasis on the empirical results from the Channel Islands MPA network. This is to our knowledge the first empirical assessment of the regional effects of a large and iconic MPA network on a wide assemblage of commercially and economically important species. Our inability to estimate a clear effect in such a well-enforced and well-studied local is we feel a vitally important "null-result" both for scientists and MPA managers to be aware of. As such we feel that it warrants publication in PNAS.

While our results are more focused on the empirical results, we explicitly link our results to the large body of MPA theoretical work, using it to help illustrate why we might expect such a null result and help managers understand when they might be able to detect an effect. As MPAs are used as a core tool of marine resource management, our results provide a critical update to our empirical expectations for MPAs, that to date have been based almost entirely on inside-outside MPA response ratios. We hope you find the updated version of our manuscript improved thanks to the inclusions of the careful points raised by the prior revision of this paper.

<!-- In addition to the comments from the last round, we have made a change in the estimation model that we feel should be highlighted. The previous draft of this manuscript utilized a mixed-effect model to estimate the regional effects of MPAs, fitting to data at the level of individual transect observations. This model provides a robust accounting of uncertainty at multiple levels, but it also cumbersome and   -->

## Reviewer 1

- *I think that greater emphasis needs to be placed up front, abstract, statement of significance, introduction..perhaps even title, that this model and paper only addresses fish biomass...not other 'conservation outcomes'. You do define this but not until page 2, line 148-154. I recommend moving this up.*
  - We apologize for not making this clearer in our earlier draft, and have included this key caveat in the abstract, statement of significance, and introduction (Line 98)

- *I was surprised that there was no discussion about enforcement being a potential confounding factor. Currently the paper assumes that the Channel Islands MPAs are fully enforced, without poaching. This may or may not be the case and should be stated as such. You do address well other system wide impacts like water temperature (lines 557-8, 691+). Perhaps enforcement could be added to the discussion around line 711-13.*
  - This is indeed a critical point that we have endeavored to correct. We have included enforcement under our summary of key enabling conditions for effective MPAs (line 124), as well as the discussion (line 382). While we do not have published data on strength of enforcement within the MPAs, anecdotal evidence suggests it is relatively high, as well as evidence from the mainland California MPAs [@zellmer2018] indicates it is high. In addition, the presence of a clear increase in response ratios within the Channel Islands provides evidence that the MPAs  are large enough and sufficiently well-enforced as to provide meaningful protection within their borders (Line XX).

- *Your discussion of the model also emphasizes that stock status prior to MPA designation is very key to outcomes; in the SI most of the species included for the Channel Island review have unknown status---which influences your discussion, and may better be described.*
  - We make clearer that we iterate over unknowns such as pre-MPA stock status in our description of the Channel Islands simulations (line XX), and the implications of pre-MPA stock status in the discussion (Line XX).

## Reviewer 2

- *Yes, these are important questions but they are not new to ecology, economics, and fishery science. You are not the first to ask them and point out their importance; see articles at bottom for some examples. What are you doing to answer them that is better than all of the other papers that come before you? And why is your case study an exemplary place to answer the questions? The conclusions on line 709-728 are well-known in the MPA literature.*
  - We have fundamentally restructured the paper based around the importance of the Channel Islands as a case study in measuring MPA effectiveness. We thank the reviewer for pointing out gaps in our literature, and we have incorporated the studies they recommend. The theoretical points we raise have indeed been pointed out in the literature before, and we feel our new manuscript makes that clearer. The Channel Islands MPAs are an iconic network that has served as a model for protected areas around the world. It is also incredibly well-studied and well-enforced. While previous studies have looked the effectiveness of the Channel Islands MPAs within their borders, this is the first study in the Islands (and globally of which we are aware) to attempt an estimate of the effect of the MPAs at the regional scale (both inside and outside of the MPAs themselves). That we are unable to estimate a clear regional effect in even this system is an important result that we feel warrants publication in PNAS. As MPAs are increasingly turned to tools for ocean conservation, our results provide an important update and caution to expectations based solely on inside-outside comparison studies.

- *I would also be more cautious regarding your conclusion about your model predicts what should be expected in the data analysis. It could be simply luck that there is agreement between the two.*
  - We have made the language clearer that this concordance could be simple coincidence, and is not in any way proof of the validity of our results (Line XX) However, in light of recent concerns in the scientific literature around hypothesizing-after-results, we feel that it is critically important to ground our theoretical results in *a priori* predictions. In addition, as we mention in the discussion, we encourage MPA scientists and stakeholders to undertake such simulation modeling to consider the plausibility of their results. As the reviewer says, empirical results that match theoretical predictions are not proof that the empirical methods are robust, but they do provide am additional and important line of evidence: empirical results that fail to match theoretical predictions need to be carefully considered.

- *What about using the simulation model to demonstrate the validity of the technique used to identify the MPA effect used in the paper, which was pioneered in publication in 38? Or to understand whether the implication of the statement on line 539, that is, species ranges exceed the Island and your sampling domain, is important for measuring the impacts of MPAs outside of their boundaries? In sum, in its current form, I would move all of the simulation model to the SI (lines 131-252) and mention in the discussion section that these results are consistent with the predictions of a stylized single species model.*
    - We have greatly expanded the simulation validation of our method, which was previously included in the supplementary material. Figure 4 of the paper now demonstrates the validity (and shortcomings) of the model under a variety of plausible conditions. We would note that this method is related to that described in @caselle2015 (previously reference #38), but is not the same. Our method explicitly estimates the net difference in the targeted and non-targeted biomass densities pre-and-post MPA as an estimate of the mean effect of the MPAs  both inside and outside their borders. Caselle et al. 2015 used trends in targeted and non-targeted species inside and outside of MPAs to illustrate the effects of the MPAs within their borders. We stress this difference in an effort to make clear that our results and methods are not simply an update to Caselle et al. 2015, but address a fundamentally different question. We hope we have made this clearer in the text, e.g. on line XX.

    With regards to the specific question about the possibility of the species range exceeding the sampling domain, this would be reflected as a reduction in the percent of the population protected by the MPAs: a 25km^2 MPA may provide near total protection for a sedentary species with brood-rearing, but be insignificant to a migratory tuna. To that end, our assessment of model performance as a function of MPA effect size addresses the question of how MPA size relative to species range affects our models performance. All else being equal the smaller a portion of a population's range is protected by MPAs, the smaller the expected effect size, and by extension the more challenge the estimation model will have.

  We apologize that we did not present our prior simulation work in a sufficiently compelling manner, and failed to differentiate its results from the prior literature. As suggested by the reviewers and the editor, we have restructured the paper to place much greater emphasis on the critical results from the Channel Islands. However, we feel that the simulation results do provide vital context to our results, and provide tangible recommendations as to under what circumstances we might be able to estimate the regional effects of MPAs. The reviewer correctly points out that the individual components of our model are not themselves novel, and that the importance of fleet dynamics and fish dispersal have been previously discussed. As such as have greatly reduced the attention these results receive, except as a vehicle for the explicit consideration of the feasibility of estimating regional MPA effects. In addition, while these broad points have been made in the literature before, we feel there is value in cohesively demonstrating the ramifications of these diverse pieces of MPA theory (that generally are looked at only one or a few at a time) in once place, both through our Figure 5 and the accompanying web application. We hope that these modifications make clear the contribution of the simulation modeling.

- *The discussion on how to detect changes outside the MPA (lines 254-347) follows very closely the discussion in Ferraro et al (#41 in your list). It seems like you can remove most of this text and point the reader to that paper*
  - We have greatly reduced the amount of text addressing response ratios, and restructured the response-ratio analysis to be linked explicitly with the empirical response ratios observed in the Channel Islands (in response to comments from the editor), in Figure 1. We feel this strategy makes much clearer how the results of this paper expand beyond the findings of Caselle et al. 2015, and why the results of Caselle et al. 2015 are evidence of within-MPA effects, but not of the question of regional effects addressed in this manuscript. Given that much of the MPA literature relies on response ratios as the empirical evidence for the effectiveness of MPAs, we feel that it is critically important to directly address the challenges of response ratios as measures of regional effects in this paper. We hope that the new manner, directly linking theoretical response ratios to the levels observed in the Channel Islands.

- *To me the potential contribution of the paper starts on line 350, which is way too late in a paper of this size*
  - We have restructured the paper so that it now focuses on the Channel Islands case study, and as suggested have moved this discussion up to the introduction (line 97).

- *Is part of assumption b) (line 372) that the dispersal characteristics of the fished and non-fished species need to be similar? I could envision a scenario where the two sets of species have different dispersal characteristics, non-fished have smaller dispersal ranges than fished species, for example. How does this impact your identification strategy? If the trends are similar regardless of the mechanistic process by which it occurs, does it matter for the empirics? More discussion on both assumptions (a) and (b) and what they imply about the biological mechanisms seems warranted.*
  - The estimation model does not depend on the dispersal characteristics of the fished and non-fished species per-say. The model is operating at the level of biomass densities. As such, even if a non-targeted species has a much bigger range than the targeted species (or even the sampling region), so long as the parallel trends assumption holds within the sampled portion of the non-targeted species range, the identification strategy may be valid. Similarly, the model is agnostic as to the root causes of the parallel trends, so long as they exist. However, if for example the parallel trends are dependent on a particular temperature regime which during the study is disrupted (as is likely the case in our data), then the model may be reliable for the early years but not later. We have made this point clearer in the paper (line XX), and made clear that the lack of a clear effect in the recent years is likely due to an emerging violation of the parallel trends assumption, and not to a failure of the MPAs per say. However, this is still an important result, since it suggests that neither the response ratios nor the difference-in-difference method give us a robust estimate of the regional effects of the Channel Island MPAs. As we stress in the paper, that even in a network such as the Channel Islands we are left without a clear empirical effect at the regional level has critical ramifications for the design and monitoring of MPAs globally.

- *The use of visual data surveys has, of course, problems that introduce errors in dependent and independent variables that could be correlated with features of the fished and non-fished species, e.g., size distribution. How can you be sure that this is not driving your results? Could you use the simulation model to investigate this?*
  - Visual survey data are indeed challenging, but also often the only form of data available for MPA analysis. Our base model operates at the level of total biomass density of targeted and non-targeted species, at the site level. This is the finest spatial scale at which the sampling regime would be likely to provide an estimate of the variable in question, since while any individual transect is unlikely to observe all of the targeted / non-targeted, but aggregating across transects provides a better chance of this. These totals are derived from average biomass densities of individual species across multiple transects within each site, which are then summed across species to achieve the total mean biomass density at the site level. The averaging process in acts to smooth out some of the sampling variation, and reflects a broader discussion about the right degree of smoothing in spatio-temporal data.

How might this affect our results? So long as sampling errors are consistent over time, they should not bias the mean or the trends of our results (e.g. if we systemically under/over estimate the biomass of some species). There is no doubt though that we are underestimating the true magnitude of the credible intervals though. For example, the model assumes constant length-to-weight relationships over time, whereas in reality there is clearly variation both among individuals and over time. Accounting for this should add more noise to our data, further complicating the ability of our model to achieve precise estimates. Complicated relationships such as the MPAs themselves affecting the length-to-weight relationship would indeed skew our results, but we feel that is beyond the scope of this or any other MPA study of which we are aware to address.

  We have run our model at a wide array of resolutions, from the updated aggregated form we use here to a complete hierarchical framework that attempts to control for errors in the individual observation process (Fig.SXX). The fundamental results remains the same: a general upwards trend, followed by a downward trend, with the clearest signal in the years around 2010. We would actually be more concerned about the robustness of our results if we had found a precise but small effect size: such an effect could easily disappear if we accounted for errors in the data (e.g. through allowing for variations in the length-weight relationship). Our relatively uncertain findings instead suggests that it is unlikely that we would get fundamentally different conclusions by including more error; we would simply be even less certain than the uncertain state the model estimates we are in. We have expanded the simulation testing of the method a bit to help illustrate how violations of the model assumptions play out, and how they are related to the magnitude of the effect size (Figure 5). The full suite of alternative resolution models can be found in Figs.SXX-XX (all of which show the same general result), and a a simulation test of our model incorporating factors such as differing diver skill can be found in Fig.SXX (which shows that the model is capable of performing under such circumstances).

- *(The answer to your question on line 709 that appears on lines 714-719 has been known in the literature on MPAs for over 20 years; see, e.g., papers by Martin Smith below. Given that your question is what "your results imply about the future of marine science", my answer is not very much. Again, you did a rigorous empirical case study that should be published. I am however finding it difficult to see a path to publication in PNAS.)*
  - We hope that our restructuring around the novel empirical results resolves this concern. While we agree that framing our results around the future of marine science would be a step too far, we would point out that our earlier and present draft asked what our results imply for the future of MPA science more specifically, which we hope is less of a stretch.

  Given the increasing role of MPAs in conservation strategies, the lack of a clear regional effect in a place like the Channel Islands should be known and taken into account by MPA practitioners around the globe. We agree that we framed our simulation results poorly in previous drafts, but feel that they do compliment the empirical results to help managers understand why the Channel Islands results may not be surprising, and when they might expect to be able to detect and effect, and when not. See section 2.1.2 ("When can we detect the effects of MPAs") and our updated conclusions, in particular the paragraph beginning on line XX.

- *Examples of relevant literature, you should expand your net to find other relevant literature to better distinguish your contribution*
  - We thank the reviewer for pointing out these important omissions. We have included each of these references and their key messages in their appropriate portions of the paper.

# References

<!-- ## Editor -->

<!-- # Original Reviews  -->

<!-- Dear Dr. Ovando, -->

<!-- I regret to inform you that the PNAS Editorial Board has rejected your manuscript [MS# 2019-17437]. The expert who served as editor obtained 2 reviews, which are included below. After careful consideration, the editor decided that we cannot accept your manuscript. -->

<!-- However, because the reviewers think the work is of interest and the editor concurs, we are willing to consider a re-submission that constructively addresses all of the concerns raised in the critiques. The paper would have to satisfy both the reviewers and the editor, and new criticisms could arise upon re-evaluation. We cannot guarantee success and will be unable to consider further re-submissions. -->

<!-- Please submit the re-submission as you would a new paper. Your re-submission will be given a new tracking number. In your cover letter and the Comments to Editorial Staff, please specify that the paper is a re-submission and include your original tracking number. Additionally, please provide a "re-submission/revision" cover letter, including a point-by-point response to the editor and reviewers' comments. If you have any questions, please contact our office. -->

<!-- Thank you for submitting your work to PNAS. -->

<!-- Sincerely yours, -->
<!-- May R. Berenbaum -->
<!-- Editor-in-Chief -->

<!-- Editor's Remarks to Author: -->

<!-- This is a fascinating manuscript because it examines an iconic MPA, for which there may be more detailed biological and physical data than any other MPA. -->

<!-- The simulation model that is applied to the case study is also rich in detail. -->

<!-- One referee, who is a world-renowned fisheries scientist with quantitative chops and deep knowledge, challenges the novelty and significance of the paper, holding it to a "high PNAS-bar", while also recognizing that as a case study it needs to be published. I hold this referee in high regard, and upon a close reading of the manuscript myself, I had to agree that when I was leading a science group at NOAA fifteen years ago, which is when MPA's were a hot new topic, it was widely appreciated that the outcome of any MPA would be dramatically impacted by where displaced vessels go and movement rates of adult fish. -->

<!-- In order for this to be published in PNAS, you need to be very clear on the manuscript's novel contribution, and also place it better in the salient literature. How relevant are references 33 and 35 compared to the references provided by referee #2? -->

<!-- The one difference I have with referee #2 is that I am not sure to what extent the richness or many, many different species, of before and after data you have for the Channel Islands MPA has ever been matched in other MPA study? In other words, there is some chance the empirical data may be what are most novel - although those data do not come to the forefront in this paper. -->

<!-- It could be that this warrants two papers rather than one. I would have liked to seen a paper focusing on the results of the difference in differences analyses. For maximum impact these would be presented in a naïve form because that is illustrative to practitioners-as it is "what they expect", and in the more rigorous form after accounting for the absence of any true control site and modeling the fact that an MPA affects the so-called, but non-existent, control. The naïve results are worthy of attention. If the empirical data were elevated, then the Supplementary materials need to include more information on the sampling, spatial and temporal detail, etc. As it is, I have no idea how much confidence to put in the fish-by-fish biomass data. -->

<!-- Then the great merit of the model is it should tell the manager what they should expect, and what they need to do if they want to make sure any newly instituted MPA is performing as hoped for. There are several marine conservation NGO's who are operating with totally naïve and misguided expectations about what will be evident after an MPA is established. Written up well, this research could mitigate some dangerous delusions. To do this the model would be used to discuss: -->

<!-- What would you expect to see even if an MPA was working well? -->

<!-- What would you expect to see if it an MPA was ineffective and doomed to be ineffective? -->

<!-- What does all this mean for the expectations we put on MPA's? -->

<!-- All of the above questions are touched on in this manuscript -- but not direct enough for a manager or NGO to take note of them. Lines 642 to 651 are noteworthy, but are kind of lost in their placement. -->

<!-- Reviewer Comments: -->
<!-- Reviewer #1: -->

<!-- Suitable Quality?: Yes -->
<!-- Sufficient General Interest?: Yes -->
<!-- Conclusions Justified?: Yes -->
<!-- Clearly Written?: Yes -->
<!-- Procedures Described?: Yes -->
<!-- Supplemental Material Warranted?: Yes -->

<!-- Comments on Significance Statement: -->

<!-- I think it could use some clarification/specificity. Since fish biomass was the only 'conservation outcome' analyzed I think that should be stated in the significance statement as well as abstract. Further I think an additional word in the case study sentence would be helpful to demonstrate that was the empirical part of the study, additional words in caps: "We find no statistically clear effect of ONE EXISTING network of MPAs in FISH BIOMASS after 13 years of protection, which our simulation analysis suggests is actually to be expected." -->

<!-- Comments: -->

<!-- - I think that greater emphasis needs to be placed up front, abstract, statement of significance, introduction..perhaps even title, that this model and paper only addresses fish biomass...not other 'conservation outcomes'. You do define this but not until page 2, line 148-154. I recommend moving this up. -->

<!-- - I was surprised that there was no discussion about enforcement being a potential confounding factor. Currently the paper assumes that the Channel Islands MPAs are fully enforced, without poaching. This may or may not be the case and should be stated as such. You do address well other system wide impacts like water temperature (lines 557-8, 691+). Perhaps enforcement could be added to the discussion around line 711-13. -->

<!-- - Your discussion of the model also emphasizes that stock status prior to MPA designation is very key to outcomes; in the SI most of the species included for the Channel Island review have unknown status---which influences your discussion, and may better be described. -->

<!-- - There is extra punctuation in the footnote around line 117-118. -->

<!-- Reviewer #2: -->

<!-- Suitable Quality?: No -->
<!-- Sufficient General Interest?: No -->
<!-- Conclusions Justified?: Yes -->
<!-- Clearly Written?: Yes -->
<!-- Procedures Described?: Yes -->
<!-- Supplemental Material Warranted?: Yes -->

<!-- Comments: -->

<!-- Measuring the performance of marine protected areas is an important area of research. The authors undertake a rigorous analysis of one set of reserves in Southern California. The case study is interesting and robustly done, but it is not clear what general lessons from the modeling and empirical application are novel. Overall, the authors need to do a better job explaining why this paper makes a contribution worthy of PNAS. -->

<!-- Some text from the paper: "The potentially more important question, however, is not whether spillover occurs (it must to some degree in any realistic scenario), but what the net effects of spillover are and whether those effects are empirically detectable. From a fishery perspective, are spillover benefits sufficient to offset losses in fishing grounds and changes in responses of displaced fishers caused by an MPA? From a conservation perspective, how much does the buildup of fish inside an MPA increase biomass outside the protected area? Overall, what are the regional effects of MPAs?" -->
<!-- ... -->
<!-- "To address this gap, this study examines two critical questions: 1) What do we expect the regional-scale conservation effects of MPAs to be and 2) When (and how) can we expect to empirically detect these effects?" -->

<!-- Yes, these are important questions but they are not new to ecology, economics, and fishery science. You are not the first to ask them and point out their importance; see articles at bottom for some examples. What are you doing to answer them that is better than all of the other papers that come before you? And why is your case study an exemplary place to answer the questions? The conclusions on line 709-728 are well-known in the MPA literature. -->

<!-- The idea that structural modeling / simulation model has an important role to play to understand empirical signals from data is not novel in general and specifically with respect to marine reserves, see included citations. While you do a lot of perturbations of the single species model, there is really nothing new in the modeling and results. I would also be more cautious regarding your conclusion about your model predicts what should be expected in the data analysis. It could be simply luck that there is agreement between the two. -->

<!-- What about using the simulation model to demonstrate the validity of the technique used to identify the MPA effect used in the paper, which was pioneered in publication in 38? Or to understand whether the implication of the statement on line 539, that is, species ranges exceed the Island and your sampling domain, is important for measuring the impacts of MPAs outside of their boundaries? In sum, in its current form, I would move all of the simulation model to the SI (lines 131-252) and mention in the discussion section that these results are consistent with the predictions of a stylized single species model. -->

<!-- The discussion on how to detect changes outside the MPA (lines 254-347) follows very closely the discussion in Ferraro et al (#41 in your list). It seems like you can remove most of this text and point the reader to that paper. -->

<!-- To me the potential contribution of the paper starts on line 350, which is way too late in a paper of this size. -->

<!-- Is part of assumption b) (line 372) that the dispersal characteristics of the fished and non-fished species need to be similar? I could envision a scenario where the two sets of species have different dispersal characteristics, non-fished have smaller dispersal ranges than fished species, for example. How does this impact your identification strategy? If the trends are similar regardless of the mechanistic process by which it occurs, does it matter for the empirics? More discussion on both assumptions (a) and (b) and what they imply about the biological mechanisms seems warranted. -->

<!-- The use of visual data surveys has, of course, problems that introduce errors in dependent and independent variables that could be correlated with features of the fished and non-fished species, e.g., size distribution. How can you be sure that this is not driving your results? Could you use the simulation model to investigate this? -->

<!-- The answer to your question on line 709 that appears on lines 714-719 has been known in the literature on MPAs for over 20 years; see, e.g., papers by Martin Smith below. Given that your question is what "your results imply about the future of marine science", my answer is not very much. Again, you did a rigorous empirical case study that should be published. I am however finding it difficult to see a path to publication in PNAS. -->

<!-- Examples of relevant literature, you should expand your net to find other relevant literature to better distinguish your contribution: -->

<!-- White, J. Wilson, et al. "Decision analysis for designing marine protected areas for multiple species with uncertain fishery status." Ecological Applications 20.6 (2010): 1523-1541. -->

<!-- *White, J. W., Botsford, L. W., Baskett, M. L., Barnett, L. A., Barr, R. J. and Hastings, A. (2011), Linking models with monitoring data for assessing performance of no‐take marine reserves. Frontiers in Ecology and the Environment, 9: 390-399. doi:10.1890/100138* -->

<!-- *Burgess, S. C., Nickols, K. J., Griesemer, C. D., Barnett, L. A., Dedrick, A. G., Satterthwaite, E. V., Yamane, L. , Morgan, S. G., White, J. W. and Botsford, L. W. (2014), Beyond connectivity: how empirical methods can quantify population persistence to improve marine protected‐area design. Ecological Applications, 24: 257-270. doi:10.1890/13-0710.1* -->

<!-- *Smith, M. D., & Wilen, J. E. (2003). Economic impacts of marine reserves: the importance of spatial behavior. Journal of Environmental Economics and Management, 46(2), 183-206. doi:https://doi.org/10.1016/S0095-0696(03)00024-X* -->

<!-- *Smith, M. D., Zhang, J., & Coleman, F. C. (2006). Effectiveness of marine reserves for large-scale fisheries management. Canadian Journal of Fisheries and Aquatic Sciences, 63(1), 153-164. doi:10.1139/f05-205* -->

<!-- *Kaplan, K. A., L. Yamane, L. W. Botsford, M. L. Baskett, A. Hastings, S. Worden, and J. W. White. 2019. Setting expected timelines of fished population recovery for the adaptive management of a marine protected area network. Ecological Applications 29(6):e01949. 10.1002/eap.1949* -->

<!-- # References -->

### 2020-04-19 documents/ovando-regional-effects-of-mpas-v2.Rmd
version sent to coauthors

@@
 ---
-title: Regional Effects of Marine Protected Area Networks Are Unclear
+title: Unclear Regional Effects of a Large Marine Protected Area Network
+date: "`r Sys.Date()`"
 author: 
@@
 Response ratios from the region indicate that the Channel Island MPAs have increased biomass densities inside the MPAs.
-We are unable to find a clear effect of these same MPAs on biomass densities throughout the Channel Islands region.
+We are unable to find a clear effect of these same MPAs on biomass densities of targeted finfish throughout the Channel Islands region.
 We use a bio-economic simulation model to explain this discrepancy, and demonstrate under what conditions we likely can and cannot clearly estimate the regional conservation effects of MPAs.
@@
 Our study pairs theory with empirical methods to show that the regional conservation effects of marine protected areas can be highly variable and dependent on human drivers.
-We find no statistically clear effect of a large and well protected network of MPAs in the Channel Island, California, after 14 years of protection, which our simulation analysis suggests is actually to be expected.
+We find no statistically clear effect of a large and well protected network of MPAs in the Channel Island, California, on targeted finfish after 14 years of protection, which our simulation analysis suggests is actually to be expected.
 MPA practitioners must clearly communicate expectations from MPA networks, and plan monitoring and adaptations strategies that directly address the challenges of estimating regional conservation effects of protected areas.
@@
 This study would not be possible without the work provided by PISCO divers over the years, with special thanks to to Kathyrn Davis Koehn and Avrey Parsons-Field.
-DO thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments and technical support
+We also thank Cody Szuwalski, Julia Lawson, André Punt, and two anonymous reviewers for helpful comments and technical support.
 
@@
 These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], larval dispersal patterns and the strength and timing of density dependence in the population (e.g.
-pre- or post-settlement, @burgess2014), the age and degree of enforcement [@edgar2014],  how overfished the population would be without the MPA, and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006;@walters2000a;@white2011].
+pre- or post-settlement, @burgess2014), the age and degree of enforcement [@edgar2014],  how overfished the population would be without the MPA, and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006;@walters2000a;@white2011,;@white2010c].
 In addition, even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through network effects [@gaines2010; @costello2010a].
@@
 
-The MPAs in this study cover 21% of the the surface waters of the Northern Channel Islands, a region spanning roughly 90km east to west.
+The MPAs in this study cover 20% of the the surface waters of the Channel Islands, a region spanning roughly 90km east to west.
 Detailed dispersal studies are not available for all of the species covered by our study, but what information we have suggest that while adults of some of these species are likely to exhibit site fidelity [@love2011], larvae are likely distributed beyond the Channel Islands [@rassweiler2012].
@@
 
-Turning to the critical question of fishing pressure, formal stock assessments are largely lacking for these species.
+Formal stock assessments are largely lacking for these species.
 However, what evidence we have suggests mostly moderate fishing pressure, with some species such as California sheephead (*Semicossyphus pulcher*) and blue rockfish (*Sebastes mystinus*) experiencing high levels of fishing mortality / biomass  levels below target levels during the early 2000s [@alonzo2004;@dick2017].
 
-```{r}
-median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected> 13])
-```
-
-We integrated these network design features, life history traits, and exploitation histories into a spatially explicit bio-economic simulation model, and used this model to generate expected outcomes for the Channel Islands MPAs.
-Our results suggest that while a wide range of outcomes are plausible, from 0% to upwards of 200%, the a median simulated effect size for a stylized version of the Channel Islands after over ten years of protection was roughly `r percent(median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected > 10]))`; a potentially small value from the perspective of empirical detectability given the natural variability and sampling challenges of marine environments.
+We integrated these network design features, life history traits, and exploitation histories into a spatially explicit bio-economic simulation model, and used this model to generate expected outcomes for the Channel Islands MPAs, varying key unknowns such as the degree of pre-MPA depletion, movement rates, and fleet dynamics.
+While a wide range of outcomes are plausible, from 0% to upwards of 200%, the a median simulated effect size for a stylized version of the Channel Islands after over ten years of protection was roughly `r percent(median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected > 10]))`.
 
@@
 Updating the results of @caselle2015 through 2017 with a Bayesian response-ratio regression shows a continuation of the increasing trend in the response ratios of targeted species (Fig.\@ref(fig:rr-plot)).
+This provides evidence that the Channel Islands MPAs are large enough and sufficiently well-enforced as to provide meaningful protection within their borders.
 
@@
 In theory, control sites far enough away to negate both biological spillover and concentration by the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs may be challenging.
-While these concerns have been stated previously [e.g.
-@halpern2004;@white2011], the MPA evaluation literature has by and large been unable to adequately address them, often due to the very real logistical challenges of identifying and sampling adequate control sites that are truly independent of the MPAs in question.
 As a result of these challenges, response ratios can be a highly imprecise and biased measure of the regional conservation effect of an MPA network [@ferraro2018].
+
+<!-- While these concerns have been stated previously [e.g.
+@halpern2004;@white2011], the MPA evaluation literature has by and large been unable to adequately address them, often due to the very real logistical challenges of identifying and sampling adequate control sites that are truly independent of the MPAs in question -->
 
@@
 - `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)`.
-These empirical estimates are in line with (though in the upper range of) the outcomes that our simulation model suggests are plausible given the rough characteristics of the northern Channel Island MPAs.
+These empirical estimates are in line with the outcomes that our simulation model suggests are plausible.
 While this concordance between theory and empirics is by no means proof of the robustness of our results, it is an important line of evidence.
@@
 The Channel Islands region (and the entire West coast of the USA) experienced a dramatic ‘marine heatwave’ beginning in 2014 and persisting through 2016, resulting in part in extremely elevated water temperatures throughout the region  [@gentemann2017].
+Difference-in-difference models require a "parallel trends" assumption; in this case that in the absence of the MPAs targeted and non-targeted species would have exhibited similar trends in biomass densities.
+While this assumption cannot be proven, we find support for it in the initial years of the MPAs (Fig.SXX).
+
 Biogeographic differences in the distributions of targeted and non-targeted species may confound the observed effects of MPAs.
@@
 If the cause of recent declines was due to increases in fishing pressure we would expect to see substantial declines only in the fished areas (Fig.\@ref(fig:raw-density)).
+The net result of this is that while we see evidence for the validity of the parallel trends assumption in the early years of our data, it is likely that the non-targeted species reacted in a different way to the marine heatwave than the targeted species did, causing a violation of the parallel trends assumption in the post 2014 data.
 
@@
 We believe that the dissapearance of increasing probabilities of positive MPA effects estimated by our model are likely driven by the marine heatwave experienced by the region beginning in 2014.
+This likely violation of the parallel trends assumption means that post 2014 estimates of MPA effects may be unreliable.
 After 14 years of MPA protection and following a large environmental perturbation we are left without a clear picture of the effect of the Channel Island MPA network on biomass densities of targeted fin-fish species from either response ratios or the difference-in-difference model.
@@
 
-As an additional complication, our simulations and difference-in-difference model examine percentage changes in biomass densities (as oppose to changes in total population size).
+<!-- As an additional complication, our simulations and difference-in-difference model examine percentage changes in biomass densities (as oppose to changes in total population size).
 An increase in biomass densities from 0.02kg/m^2^ to 0.04kg/m^2^ would be produce as a 100% increase.
@@
 spiny lobster, may also reveal clearer signals.
+-->
 
 Given the natural variability of marine ecosystems, and the large challenges of obtaining accurate samples from oceanic environments, how large of an effect would an MPA network have to have in order to allow a difference-in-difference strategy such as this to be a reliable measure of MPA effects?
-To provide guidance on this important question, we used our bio-economic model to simulate data from a range of scenarios with increasing MPA effect size, along with increasing degrees of observation error and natural recruitment variation.
+We simulated data from a range of scenarios with increasing MPA effect size, along with increasing degrees of observation error and natural recruitment variation.
 As an added measure, we include scenarios in which the sampled species go through recruitment regimes, which may be positive for both targeted and non-targeted species, or positive for non-targeted species and negative for targeted species.
@@
 While unbiased across simulations, the difference-in-difference model struggled severely when MPA effect sizes were less than 25% and the model was faced with observation and process errors (Fig.\@ref(fig:val-plot)).
-Even models fir to data generated from large effect sizes commonly misestimated the true MPA effect by 50% or more.
+Even models fit to data generated from large effect sizes commonly misestimated the true MPA effect by 50% or more.
 Obtaining a mean absolute percent error (MAPE) of 25% or less across our simulated datasets required a regional MPA effect of at least 30%.
@@
 
-This finding begs an important question: when should we expect to see MPA effects big enough to stand a reasonable chance of detection?
+When should we expect to see MPA effects big enough to stand a reasonable chance of detection?
 We simulated `r n_distinct(outcomes$experiment)` MPA scenarios across a wide range of life histories, network designs, and fishing dynamics (see SI for a full description of scenarios).
@@
 
-These initial simulation results suggest some rather simple rules of thumb: put an MPA of sufficient size on an at least somewhat overexploited population and one can expect large and potentially detectable results.
 The MPA literature highlights a large number of variables beyond simply size and fishing pressure that can affect performance.
-To address this, we examined the variability in expected MPA effects across each of the two major axes (pre-MPA depletion and MPA size).
+We examined the variability in expected MPA effects across each of the two major axes (pre-MPA depletion and MPA size).
 As both MPA size and pre-MPA depletion increase, the potential for a large MPA effect increases.
@@
 
-What do our results imply about the future of MPA science?
-Our simulation model is by no means exhaustive (for example it ignores features such as species interactions, habitat effects, and climate feedback), but captures many of the core factors theorized to affect MPA performance.
-While attributes such as MPA size and fishing pressure are important factors in determining the effects of protection, local fleet dynamics and the movement rates of adult fish can dramatically affect the outcomes of protected areas as well  (Fig.\ref{fig:sim-effects}).
-The effects of MPAs can be highly context dependent, requiring - we would argue - a bio-economic model of at minimum the complexity presented here to help communities design and set expectations of MPAs at the tactical level.
-While users may not be able to parameterize every aspect of a model such as that presented here, working with stakeholders to visualize the implications of, for example, different fleet responses to MPA implementation is a critical step in MPA design.
+What do these results imply for the evaluation and management of MPAs around the globe?
+A large theoretical literature has pointed out the importance of a wide range of economic and environmental factors in determining the effectiveness of MPAs [e.g.
+@hilborn2004; @smith2006; @white2010c].
+Our simulation results allow users to place this collective theory into context, clarifying under what states of the world we might expect to see particular effect sizes, and demonstrating what effect sizes we might reasonably expect to be detectable in the real world.
+In the case of the Channel Islands, simulation modeling suggests for the network in question, an effect size on the order of 30% or smaller is likely, which our analysis in turn demonstrates may be challenging to detect.
+This cautionary tale is likely to hold for many coastal MPA networks around the world.
 
@@
 We applied one such approach here (a difference-in-difference estimator), and yet were still unable to reach robust conclusions as to the effect of MPAs on the biomass density of targeted finfish in the Channel Islands, due to the likely small size of the true effect relative to the strength and variability of environmental drivers.
+If pre-MPA data are available, data currently used in response ratios can easily be re-purposed for an analysis such as this one; we encourage holders of such data worldwide to conduct similar exercises to examine to what extent empirical estimates of regional conservation effects match up with MPA theory.
 
-While this does not mean that all MPAs will face similar challenges in estimating their effects, our results in the relatively large, well-enforced, and rigorously studied Channel Islands Marine Protected Area network make clear that in many instances empirically detecting a clear regional effect of MPAs may not be possible.
+While this does not mean that all MPAs will face similar challenges in estimating their effects, our results in the relatively large, well-enforced, and rigorously studied Channel Islands Marine Protected Area network make clear that in many instances empirically detecting a clear regional effect of MPAs on total biomass densities of targeted finfish may not be possible.
 How then should stakeholders go about adaptively monitoring and managing MPAs?
-Simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can perhaps be tuned to focus on the species groups that have the highest chance of a detectable effect size over the early years of the reserve [@nickols2019].
+Simulation modeling incorporating the diverse drivers represented here can help inform the range of effect sizes that may be expected, and monitoring programs can perhaps be tuned to focus on the species groups that have the highest chance of a detectable effect size over the early years of the reserve [@nickols2019].
 Expanding data collection to include robust monitoring of spatio-temporal fleet dynamics may help assess the validity of control sites used in response ratios, support the direct inclusion of these fleet dynamics into statistical models, and allow managers to take into account potential negative interactions between MPAs and fleet dynamics such as those that may occur under constant-catch dynamics.
@@
 
-Stock assessment models [as described in @hiborn1992b] that account for the population within the MPAs [field2006a] may be able to answer the management relevant question of whether fishing mortality rates and biomass levels are in-line with management objectives [e.g.
+Stock assessment models [as described in @hiborn1992b] that account for the population within the MPAs [field2006a] may be able to answer the management relevant question of whether fishing mortality rates and biomass levels from systems containing MPAs are in-line with management objectives [e.g.
 @nickols2019].
@@
 
-Non-equilibrium analyses also help set expectations for effect sizes over time [@nickols2019,Fig.\ref{did-plot}].
+Non-equilibrium analyses also help set expectations for effect sizes over time [@nickols2019,@kaplan2019,Fig.\ref{did-plot}].
 Educating communities about the challenges of estimating the effects of MPAs can help set expectations, so that a lack of a clear effect is not necessarily viewed as a failure of the program, or large positive result based solely on response ratios as a clear sign of success and subsequent relaxation of other fishery management strategies.
@@
 We pair these data with randomly selected values between 0.6 and 0.95 for Beverton-Holt steepness [as parameterized in @mace1994], as well as larval and adult dispersal rates.
-We randomly assign whether adult fish preferentially move towards patches with lower relative densities, as well as one of three potential types of recruitment density dependence [building off of @babcock2011]:
+We randomly assign whether adult fish preferentially move towards patches with lower relative densities, as well as one of three potential types of recruitment density dependence [building off of @babcock2011].
+Simulations are also assigned random fleet dynamics (open access ,constant effort, constant catch), responses to MPA (leave the fishery or concentrate outside MPAs), a fleet dispersal model (uniform in response to catch-per-unit-effort, in response to profit-per-unit-effort),
 
-  1. Local density dependence: Density dependence occurs independently in each patch, and recruits then disperse to nearby patches
+<!--   1. Local density dependence: Density dependence occurs independently in each patch, and recruits then disperse to nearby patches -->
 
-  2. Global density dependence: Density dependence is a function of the sum of spawning biomass across all patches, and recruits are then distributed according to habitat quality
+<!--   2. Global density dependence: Density dependence is a function of the sum of spawning biomass across all patches, and recruits are then distributed according to habitat quality -->
 
-  3. Post-dispersal density dependence: Larvae are distributed throughout the system, and then density dependence occurs based on the density of adult biomass at the destination patch
+<!--   3. Post-dispersal density dependence: Larvae are distributed throughout the system, and then density dependence occurs based on the density of adult biomass at the destination patch -->
 
-We allow for three potential siting strategies for MPAs.
+<!-- We allow for three potential siting strategies for MPAs.
 In the first, MPAs are randomly placed.
@@
 In the third, we allow for scenarios in which MPAs are placed in sources of larval dispersal.
+-->
 
-Each simulation is randomly assigned a fleet model of the form
+<!-- Each simulation is randomly assigned a fleet model of the form  -->
 
-  1. Open access: fishing effort changes in response to profit-per-unit-effort
+<!--   1. Open access: fishing effort changes in response to profit-per-unit-effort -->
 
-  2. Constant effort: total fishing effort is constant over time (unless altered by MPA displacement model)
+<!--   2. Constant effort: total fishing effort is constant over time (unless altered by MPA displacement model) -->
 
-  3. Constant catch: the fleet exerts as much effort as need to achieve a target catch
+<!--   3. Constant catch: the fleet exerts as much effort as need to achieve a target catch -->
 
-This fleet model is paired with a gear selectivity ranging from .1 to 1.5 of the length at 50% maturity for the species in question, and the fleet model is tuned to achieve a target fishing mortality relative to natural mortality ratio at equilibrium.
+<!-- This fleet model is paired with a gear selectivity ranging from .1 to 1.5 of the length at 50% maturity for the species in question, and the fleet model is tuned to achieve a target fishing mortality relative to natural mortality ratio at equilibrium.
+-->
 
-Along with the fleet dynamics model, each simulation is assigned a random fleet dispersal scenario: uniform dispersal (where the total effort of the fleet is divided evenly among all open patches), catch dispersal (where the total effort of the fleet is divided according to the catchable biomass in each available patch), and profit dispersal (where the total effort of the fleet is divided according to the profit per unit effort in each available patch).
+<!-- Along with the fleet dynamics model, each simulation is assigned a random fleet dispersal scenario: uniform dispersal (where the total effort of the fleet is divided evenly among all open patches), catch dispersal (where the total effort of the fleet is divided according to the catchable biomass in each available patch), and profit dispersal (where the total effort of the fleet is divided according to the profit per unit effort in each available patch).
 Fishing effort that occurred inside MPAs prior to closer can either leave the fishery, or be distributed to the patches outside the MPAs.
+-->
 
 Each simulation is assigned an MPA scenario, defined by the number and size of MPAs, the placement of those MPAs, and the year that the MPAs are put in place.
-Each simulation starts the population off at unfished equilibrium and then beings to apply the fleet model.
+The population begins at unfished equilibrium and then fishing effort is applied in accordance to the fleet model.
 The MPAs are then placed during the randomly selected start year, allowing some runs to explore how the early dynamics of the MPA play out when the fishery and population they are placed on is not already at equilibrium.
@@
 The key assumptions of the difference-in-difference model are that a) within the time-frame of the model there are no significant interaction effects between the targeted and non-targeted species (which in fact we do not detect, see SI), and b) that in the absence of the MPAs both the targeted and non-targeted groups of species would have exhibited similar trends in densities.
+It is important to note that the model is agnostic as to the root causes of the parallel trends, so long as they hold.
 
@@
 We do not detect any significant differences in the trends of the biomass densities of the targeted and non-targeted species in the years before the MPAs (Fig.SXX).
+However, we believe that the parallel trends assumption is likely violated in the post 2014 data.
+
 With regards to the second assumption, all of the species in this empirical analysis exist within an ecosystem, and as such affect each other through mechanisms such as predation, competition, and habitat modification.
 We find it unlikely that these effects have had enough time to manifest in a meaningful way in the 14 years of post-MPA data used in our analysis [@babcock2010; @pershing2015a].
-
 We used convergent cross mapping (CCM), in the manner of @clark2015, to test for the possibility of the trophic cascades biasing our results.

### 2020-04-21 documents/reviewer-responses.Rmd
dual run on squid

@@
 While our results are more focused on the empirical results, we explicitly link our results to the large body of MPA theoretical work, using it to help illustrate why we might expect such a null result and help managers understand when they might be able to detect an effect.
-As MPAs are used as a core tool of marine resource management, our results provide a critical update to our empirical expectations for MPAs, that to date have been based almost entirely on inside-outside MPA response ratios.
+As MPAs are used as a core tool of marine resource management, our results provide a useful update to our empirical expectations for MPAs, that to date have been based almost entirely on inside-outside MPA response ratios.
 We hope you find the updated version of our manuscript improved thanks to the inclusions of the careful points raised by the prior revision of this paper.

### 2020-04-21 documents/ovando-regional-effects-of-mpas-si-v2.Rmd
dual run on squid

--- 2019-10-09 documents/ovando-regional-effects-of-mpas-si.Rmd
@@
     - \usepackage[left]{lineno}
+    - \usepackage{xcolor}
     - \linenumbers
 params:
-  run_name: ["v5.0"]
+  run_name: ["v6.0"]
   sim_years: [50]
@@
 ```{r, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE)
-
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
+                      cache = FALSE, fig.width = 3.42, fig.asp = 1.5)
+
+```
+
+```{r, include=FALSE}
 library(sf)
@@
 library(tidyverse)
-library(Synth)
-library(spasm)
+library(ggtext)
+library(rstan)
+library(rstanarm)
+library(bayesplot)
 extrafont::loadfonts()
+rstan_options(auto_write = TRUE)
+
 functions <- list.files(here::here("functions"))
@@
 # num_patches <- 25
-# 
+
 # run_name <- "v5.0"
 
+width_1col <- 3.42 # inches
+
+asp_1col <- 1.25
+
+width_2col <- 7 # inches
+
+asp_2col <-  .75
+
+dpi <- 900
+
+theme_1col <- hrbrthemes::theme_ipsum(base_size = 11,
+                                      axis_title_size = 12)
+
+theme_2col <- hrbrthemes::theme_ipsum(base_size = 12,
+                                        axis_title_size = 14)
+
 run_name <- params$run_name
@@
 
-load(file = here::here("results",run_name, "model_runs.Rdata"))
-
-load(file = here::here("results",run_name,"processed_grid.Rdata"))
-
-load(file.path(run_dir,"abundance_data.Rdata"))
+load(file.path(run_dir, 'abundance_data.Rdata'))
+
+model_runs <- read_rds( file.path(run_dir,"did_fits.rds"))
+
+# load(file = here::here("results",run_name, "model_runs.Rdata"))
+
+processed_grid <-  read_rds(file.path(run_dir,"filtered_processed_grid.rds"))
+
+load(file = file.path(run_dir ,"sim_grid.Rdata"))
+
+load(file = file.path(run_dir,"plots.RData"))
+
+valgrid <- read_rds(file.path(run_dir, "valgrid.rds"))
+
+cdfw_catches <-
+  read_csv(file = here::here("data", 'cdfw-catches.csv')) %>%
+  group_by(sci_name, year) %>%
+  summarise(catch = sum(pounds_caught, na.rm = T)) %>%
+  left_join(
+    life_history_data %>% select(taxa, classcode, commonname) %>% mutate(taxa = tolower(taxa)),
+    by = c("sci_name" = "taxa")
+  ) %>%
+  filter(!is.na(classcode))
+
+# cdfw_catches %>%
+#   group_by(commonname, classcode, sci_name) %>%
+#   summarise(catch =sum(catch, na.rm = TRUE)) %>%
+#   mutate(in_analysis = classcode %in% unique(cip_data$classcode)) %>%
+#   View()
 
@@
 outcomes <- readRDS(file.path(run_dir,"outcomes.rds"))
+
+outcomes <- outcomes %>% 
+  filter(!is.na(mpa_effect))
+
+density_ratios <- density_ratios %>% 
+  filter(!is.na(biased_density_ratio))
 
@@
   mpa_size = "Range in MPA",
-  depletion = "Without-MPA Depletion"
+  depletion = "Depletion"
 )
 
-channel_islands <- readRDS(here::here("data","channel_islands_map.RDS"))
-
-ca_mpas <- sf::st_read(here::here("data","MPA_CA_Existing_160301")) %>%
-  rmapshaper::ms_simplify() %>%
-  sf::st_transform(crs = 4326)
-
-zissou_theme <-
-  theme_ipsum(
-    base_size = 12,
-    axis_title_size = 12,
-    strip_text_size = 12,
-    base_family = "Helvetica"
-  )
-
-zissou_theme <- zissou_theme +
-  theme(   panel.grid.major.x = element_blank(),
-    panel.grid.minor.x = element_blank())
-
 testplot <- ggplot()
@@
 panel_height = unit(1,"npc") - sum(ggplotGrob(testplot)[["heights"]][-3]) - unit(4,"line")
+
+panel_width = unit(1,"npc") - sum(ggplotGrob(testplot)[["widths"]][-3]) - unit(4,"line")
 
@@
 
-theme_set(zissou_theme)
-
-models_worked <- model_runs$tmb_fit %>% map("error") %>% map_lgl(is_null)
-
-model_runs <- model_runs %>%
-  filter(models_worked) %>%
-  mutate(tmb_fit = map(tmb_fit,"result")) %>%
-  mutate(processed_fits = map(tmb_fit, process_fits)) %>%
-  mutate(did_plot = map(processed_fits, "did_plot"))
+theme_set(theme_1col)
+
+# models_worked <- model_runs$tmb_fit %>% map("error") %>% map_lgl(is_null)
+
+# model_runs <- model_runs %>%
+#   filter(models_worked) %>%
+#   mutate(tmb_fit = map(tmb_fit,"result")) %>%
+#   mutate(processed_fits = map(tmb_fit, process_fits)) %>%
+#   mutate(did_plot = map(processed_fits, "did_plot"))
 
 base_run <- model_runs %>%
-filter(var_names == "pisco_a", mpa_only == FALSE, center_scale == TRUE)
+filter(var_names == "pisco_a", data_to_use == "all", center_scale == TRUE)
 
@@
          var_names == "pisco_a",
-         mpa_only == TRUE)
+         data_to_use == "mpa_only")
+
+fishsed_run <- model_runs %>%
+  filter(data_source == "pisco",
+         var_names == "pisco_a",
+         data_to_use == "fished_only")
 
@@
 
-fitted_data <- base_run$data[[1]]
-
-zissou_fit <- base_run$tmb_fit[[1]]
-
-report <- base_run$tmb_fit[[1]]
+used_data <- abundance_data$data[abundance_data$data_source == "pisco"][[1]]
+
+did_data <- model_runs$did_fit[model_runs$data_source == "pisco" &
+                                 model_runs$var_names == "pisco_a" &
+                                 model_runs$data_to_use == "all"][[1]]$did_data
+
+did_fit <- model_runs$did_fit[model_runs$data_source == "pisco" &
+                                 model_runs$var_names == "pisco_a" &
+                                 model_runs$data_to_use == "all"][[1]]$did_reg
+
+did_results <- model_runs$did_fit[model_runs$data_source == "pisco" &
+                                 model_runs$var_names == "pisco_a" &
+                                 model_runs$data_to_use == "all"][[1]]$did_results
 
@@
 
-zissou_fit$zissou_estimates <- zissou_fit$zissou_estimates %>%
-  mutate(variable = str_replace_all(variable, "(\\.).*$",""))
-
-report$zissou_estimates <- report$zissou_estimates %>% 
-    mutate(variable = str_replace_all(variable, "(\\.).*$",""))
-
-seen_non_nested_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "seen_non_nested_betas")) %>%
-  rename(group = variable) %>%
-  mutate(variable  = zissou_fit$zissou_data$x_seen_non_nested %>% colnames())
-
-seeing_non_nested_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "seeing_non_nested_betas")) %>%
-  rename(group = variable) %>%
-  mutate(variable  = zissou_fit$zissou_data$x_seen_non_nested %>% colnames())
-
-seen_year_species_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "seen_year_species_betas")) %>%
-  rename(group = variable) %>%
-  mutate(variable = zissou_fit$zissou_data$x_seen_year_species %>% colnames())
-
-seeing_year_species_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "seeing_year_species_betas")) %>%
-  # filter(variable == "seeing_year_species_betas") %>%
-  rename(group = variable) %>%
-  mutate(variable = zissou_fit$zissou_data$x_seen_year_species %>% colnames())
-
-seen_region_cluster_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "seen_region_cluster_betas")) %>%
-  # filter(variable == "seen_region_cluster_betas") %>%
-  rename(group = variable) %>%
-  mutate(variable = zissou_fit$zissou_data$x_seen_region_cluster %>% colnames())
-
-seeing_region_cluster_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "seeing_region_cluster_betas")) %>%
-  # filter(variable == "seeing_region_cluster_betas") %>%
-  rename(group = variable) %>%
-  mutate(variable = zissou_fit$zissou_data$x_seen_region_cluster %>% colnames())
-
-did_betas <- zissou_fit$zissou_estimates %>%
-  filter(stringr::str_detect(variable, "mpa_effect")) %>%
-  # filter(variable == "mpa_effect") %>%
-  mutate(group = variable) %>%
-  mutate(year = zissou_fit$did_data$year %>% unique())
-
- betas <- bind_rows(
-   seen_non_nested_betas,
-   seeing_non_nested_betas,
-   seen_year_species_betas,
-   seeing_year_species_betas,
-   seen_region_cluster_betas,
-   seeing_region_cluster_betas,
-   did_betas %>% select(-year)
+top_species <- used_data$classcode %>% unique()
+
+life_summaries <- life_history_data %>% 
+  select(classcode, tm, vbgf.k, vbgf.linf) %>% 
+  rename(age_mature = tm, vbk = vbgf.k, linf = vbgf.linf) %>% 
+  unique() %>% 
+  filter(classcode %in% top_species) %>% 
+  select(-classcode) %>% 
+  gather(variable,value) %>% 
+  group_by(variable) %>% 
+  summarise(min_val = min(value, na.rm = TRUE),
+         max_val = max(value, na.rm = TRUE))
+
+ sim_life_history <- sim_grid %>%
+    mutate(
+      age_mature = map_dbl(fish, "age_mature"),
+      vbk = map_dbl(fish, "vbk"),
+      linf = map_dbl(fish, "linf")
     ) %>%
-   as_data_frame()
+    select(scientific_name, age_mature, vbk, linf) %>%
+    unique() %>% 
+   filter(age_mature > life_summaries$min_val[life_summaries$variable == "age_mature"],
+          age_mature < life_summaries$max_val[life_summaries$variable == "age_mature"],
+          vbk > life_summaries$min_val[life_summaries$variable == "vbk"],
+          vbk < life_summaries$max_val[life_summaries$variable == "vbk"],
+          linf > life_summaries$min_val[life_summaries$variable == "linf"],
+          linf < life_summaries$max_val[life_summaries$variable == "linf"])
+
+ year_bins <-
+  seq(2003, max(pisco_data$year) + 1, by = 3)
+
+ sim_mpa_effects <- outcomes %>% 
+  filter(set == "ci",
+         final_depletion > 0,
+         f_v_m > .5) %>% 
+  mutate(year = years_protected + 2002) %>% 
+  filter(year <= max(pisco_data$year)) %>% 
+  mutate(binned_year = cut(year, year_bins, include.lowest = FALSE))  %>% 
+  filter(year >= 2003,
+         !is.na(binned_year)) %>% 
+  mutate(mpa_effect = pmin(mpa_effect,3))
+```
+
+# Supporting Information (SI)
+
+# SI Text {#si-text .unnumbered}
+
+## Computing environment
+
+All code needed to reproduce our main results and manuscript can be found at https://github.com/DanOvando/regional-effects-of-mpas.
+A fully reproducible environment for running this analysis and compiling the manuscript will be made available through Code Ocean at https://codeocean.com/capsule/5233105/tree.
+All analysis were performed in `r R.version.string`.
+Package versions are shown in Table.S\@ref(tab:pkg-info).
+
+```{r pkg-info, results = "asis", eval=FALSE}
+table <- devtools::package_info() %>% 
+  select(package, loadedversion, date,source)
+
+knitr::kable(table,
+             digits = 2,
+             caption = "Package versions and sources used in this paper",
+             longtable = TRUE,
+             booktabs = TRUE,
+             format = "latex",
+             col.names = colnames(table)) %>% 
+  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
+
+```
+
+## PISCO Data
+
+All fish data used in the primary difference-in-difference model were collected by PISCO. The dive transect survey methods are described in @caselle2015, provided below for ease of reference
+
+"Fish assemblages were surveyed annually as part of a long-term monitoring program conducted by the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) using standard underwater visual belt survey methods (www.piscoweb.org9).
+We analyzed data from 47 PISCO sites at the northern Channel Islands that were sampled annually from at least 2003 to 2012.
+We also excluded 3 sites on Anacapa where a much older MPA had already been established in 1978.
+MPAs on each island were sampled annually during June-October and we surveyed multiple sites inside and outside of any individual MPA.
+Details of MPA characteristics such as size and coastline extent are given in Hamilton et al.9.
+At each site, we conducted 8 to 12 fish transects that measured 30×2×2m at multiple levels in the water column: benthic, midwater, and kelp canopy (when present).
+Transects are laid out in a stratified random design, with multiple nonpermanent transects located in fixed strata (i.e., outer, middle, and inner edges of the reef).
+At each level in the water column, one SCUBA diver per transect counted and estimated the sizes of all fish to the nearest centimeter (total length), excluding small cryptic fishes"
+- @caselle2015
+
+We take a number of steps to translate the raw transect data into the total biomass densities used in this study.
+For the default run, we only include species that were observed at least once for at least 15 of the 18 years of available data.
+We also exclude "young of the year" observations due the challenges of correctly identifying and measuring these individuals.
+We omitted data from 1999 due to changes in the sampling procedures that occurred after 1999.
+Per recommendations from PISCO staff we omit observations from the canopy level of the transects (leaving the middle, bottom, and middle canopy levels).
+
+PISCO data report positive observations of fish.
+In order to use these data in our model we need to add in zeros for any transect that could have observed a given species of fish but did not.
+We assume that a fish could have been observed on a given transect if that species has ever been observed at that site in any time period in the data (PISCO data are organized by sites, with multiple transect at different locations within the borders of a site).
+If a species has never been observed at a site we assume that it does not occur at that site.
+
+Once zeros have been introduced to the database, we convert positive observations of fish from numbers to biomass.
+Each observation in the raw database lists the species, the number of individuals seen, and the size of those individuals (either as one value or as a mininum and maximum size for the group seen).
+PISCO staff compiled allometric information used to convert lengths to expected weights.
+For each observation then, we convert the observed lengths to weights per these relationships (accounting for variations in length types such as standard vs.
+total length).
+When minimum and maximum ranges were reported, we drew a number of samples equal to the number of observed fish in that group from a uniform distribution spanning the minimum and maximum reported size in that group.
+We assume all length-to-weight conversions are constant and deterministic.
+
+For each species at each transect, we then calculate the biomass density of that transect as the sum of the observed biomass divided by the transect area.
+We then average the biomass densities for each species across all the transects at a given site, and lastly sum these mean species biomass densities to achieve the total mean species biomass at the site level.
+
+We include several additional sources of data in our regression analysis.
+Temperature readings are included from the PISCO data for each transect.
+We also include PISCO data on the estimated surge and visibility.
+We augmented these data with information on kelp cover over time from the Santa Barbara Channel Long Term Ecological Research Network [@lter2017].
+We used a k-nearest neighbors algorithm to fill in missing kelp observations, and matched the interpolated kelp data to the PISCO data at the resolution of year-month-site (Fig.\@ref(fig:kelp-plot)).
+We include a variable capturing the mean cumulative number of observations across all observers conducting transects, in an effort to control for evoloving observer skill.
+
+<!-- Temperature data were augmented with data from `FishLife` [@thorson2017c] to include the estimated preferred temperature for a given species, so that we can include deviations from the preferred temperature envelope as a predictor in the model.
+This allows different temperatures to have different effects on each species (and is less computationally intensive than estimating species-temperature slopes) (Fig.\@ref(fig:tempdev-plot)).
+-->
+
+We also included lagged catch totals in the Santa Barbara region for the commercially harvest species in the database, in an effort to control for changes in density caused by changes in fishing pressure.
+Catches were pulled from the CDFW website (https://www.wildlife.ca.gov/Fishing/Commercial/Landings), and extracted using the `tabulizer` package in R [@leeper2018] (Fig.\@ref(fig:catches-plot)).
+
+```{r catches-plot, fig.cap = "Total CDFW reported commercial catches in the Santa Barbara region"}
+
+used_data %>% 
+  filter(targeted == 1) %>% 
+  group_by(commonname, year) %>% 
+  summarise(catches = unique(lag_catch)) %>% 
+  group_by(commonname) %>% 
+  mutate(total_catches = sum(catches)) %>% 
+  ungroup() %>% 
+  filter(total_catches > 0) %>% 
+  ggplot(aes(year, catches)) + 
+  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
+  geom_line() + 
+  facet_wrap(~commonname, scales = "free_y") + 
+  theme_minimal() + 
+  labs(x = "Year", y = "Catch (lbs)")
+
+```
+
+```{r kelp-plot, fig.cap="Mean kelp biomass by island over time from SBC LTER"}
+
+used_data %>% 
+  group_by(region,year) %>% 
+  summarise(mean_kelp = mean(interp_kelp, na.rm = TRUE)) %>% 
+  ggplot(aes(year, (mean_kelp), color = region)) + 
+  geom_line() + 
+  labs(x = "Year", y = "Mean Kelp Biomass")
+
+```
+
+```{r targ-trend-plot, fig.cap="Centered and scaled mean biomass densities of all targeted finfish in analysis before filtering"}
+
+targeted_trend_plot <- pisco_data %>% 
+  filter(targeted == 1) %>% 
+  group_by(year, classcode, targeted, commonname) %>% 
+  summarise(mean_biomass_density = mean(density_g_m2, na.rm = TRUE)) %>% 
+  group_by(classcode) %>% 
+  mutate(mean_biomass_density = scale(mean_biomass_density)) %>% 
+  ggplot(aes(year,mean_biomass_density)) + 
+  geom_hline(aes(yintercept = 0), linetype = 2) + 
+  geom_vline(aes(xintercept = 2003), color = "red") +
+  geom_line() + 
+  facet_wrap(~commonname) +
+  theme_minimal() + 
+  theme(strip.text = element_text(size = 6),
+        axis.text.x = element_text(size = 8)) + 
+  labs(x = "Year", y = "Centered and Scaled Mean Biomass Density")
+
+targeted_trend_plot
+```
+
+```{r nontarg-trend-plot, fig.cap="Centered and scaled mean biomass densities of all non-targeted finfish in analysis before filtering"}
+
+nontargeted_trend_plot <- pisco_data %>% 
+  filter(targeted == 0) %>% 
+  group_by(year, classcode, targeted, commonname) %>% 
+  summarise(mean_biomass_density = mean(density_g_m2, na.rm = TRUE)) %>% 
+  group_by(classcode) %>% 
+  mutate(mean_biomass_density = scale(mean_biomass_density)) %>% 
+  ggplot(aes(year,mean_biomass_density)) + 
+  geom_hline(aes(yintercept = 0), linetype = 2) + 
+  geom_vline(aes(xintercept = 2003), color = "red") +
+  geom_line() + 
+  facet_wrap(~commonname) +
+  theme_minimal() + 
+  theme(strip.text = element_text(size = 6),
+        axis.text.x = element_text(size = 8)) + 
+  labs(x = "Year", y = "Centered and Scaled Mean Biomass Density")
+
+nontargeted_trend_plot
+```
+
+```{r classcode}
+
+species_data <- used_data %>% 
+  select(classcode, commonname, taxa, targeted) %>% 
+  unique() %>% 
+  arrange(desc(targeted)) %>% 
+  mutate(targeted = targeted == 1) %>% 
+  rename(`Common Name` = commonname, 
+         `Scientific Name` = taxa,
+         `Targeted?` = targeted)
+
+status <- read_csv(here::here("data","stock-status-summaries.csv")) %>% 
+  select(classcode, `Stock Status`)
+
+species_data <- species_data %>% 
+  left_join(status, by = "classcode")
+
+knitr::kable(species_data,
+             digits = 2,
+             caption = "Species included in estimation model",
+             longtable = TRUE,
+             booktabs = TRUE,
+             col.names = colnames(species_data)) %>%
+  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
+
+```
+
+## Difference-in-Difference Model
+
+The difference-in-difference (DiD) regression amounts to estimating the pre-post MPA difference in the biomass densities of targeted species minus the same difference for non-targeted species in the Channel Islands.
+
+The simplified form of this model is
+
+\begin{equation}
+  d_{i} \sim Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + \mathbf{B^cX_i} + \mathbf{B^sS_i}},shape, scale)
+\label{eq:did}
+\end{equation}
+
+To provide greater detail on this process, we first conduct the data pre-processing described earlier.
+From there, we aggregated data to the level of total biomass density of targeted and non-targeted species at each site each year.
+As such, our base model estiamtes the effect of the MPAs on total biomass density of targeted species, though we also explore the effect on alternative specifications such as mean biomass of targeted species.
+the model was then fit using the `rstanarm` package in R (5000 iterations, 2500 warmup, 4 chains, adapt_delta = 0.85).
+
+The intercept prior was was determined using the `rstanarm` `autoscale` function.
+For the non-intercept terms, we manually set a normal(0,2) prior.
+This implies that coefficeints such as the MPA effect have a prior that provides support for an effect size centered on zero with a range of roughly -4 to 4.
+Since the covariates are centered and scaled, and the dependent variable is on the log-scale, this is an extremely diffuse prior.
+
+The full table of covariates and their posterior means and 89% credible interval can be found in Table.S\@ref(tab:did-coefs).
+
+```{r did-coefs, results = "asis"}
+
+did_fit %>% 
+  broom::tidy(intervals = TRUE, prob = 0.89) %>% 
+  select(term, estimate, lower, upper) %>% 
+  rename(mean = estimate, lower_89th_percentile = lower, upper_89th_percentile= upper) %>% 
+  knitr::kable(
+      digits = 2,
+            caption = "Posterior means and 89% credible interval for key model coefficients",
+             longtable = TRUE,
+             booktabs = TRUE) %>%
+  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
+```
+
+```{r prior-plot, fig.cap="Prior and posterior distributions of the estimated MPA effect over time"}
+
+priors <- tibble(year = unique(did_results$year)) %>% 
+  mutate(prior = list(rnorm(500,0,2))) %>% 
+  unnest(cols = prior)
+
+did_results %>% 
+  filter(between(prank, 0.05, 0.95)) %>% 
+  ggplot() + 
+  geom_density(aes(did, fill = "Posterior", y = ..scaled..), color = "transparent", alpha = 0.75) + 
+  geom_density(data = priors, aes(prior, fill = "prior", y = ..scaled..), color = "transparent",
+               alpha = 0.5) +
+  facet_wrap(~year) + 
+  scale_fill_discrete(name = '') + 
+  scale_y_continuous(name = "Scaled Density")
+
+```
+
+We include a series of standard visual diagnostic plot for Bayesian models below, with brief descriptions of the key outcome of the diagnostic included in the figure captions.
+The model had 0 divergences or max tree depth saturations.
+
+```{r, echo = FALSE}
+
+rstan::check_hmc_diagnostics(did_fit$stanfit)
+
+```
+
+```{r rhat-plot, fig.cap="Histogram of potential scale reduction statistic Rhat. All values are below 1.05, indicating there is not evidence of chain convergence failure"}
+
+data.frame(rhat = bayesplot::rhat(did_fit)) %>% 
+             ggplot(aes(rhat)) + 
+             geom_histogram() + 
+             geom_vline(aes(xintercept = 1.05))
+
+```
+
+```{r energy-plot, fig.cap = "HMC energy diagnostic plots. Closely matching histograms shows there is no evidence of excessively fat tails"}
+
+mcmc_nuts_energy(nuts_params(did_fit))
+
+```
+
+```{r}
+y <- (did_fit$y)
+
+yrep <- (rstanarm::posterior_predict(did_fit, draws = 500))
+```
+
+```{r, fig.cap = "Posterior Predictive distribution of total biomass density (blue lines) and observed distribution of total biomass density (red line)"}
+
+ppc_dens_overlay(y = y,
+                 yrep = yrep) +
+  scale_x_continuous(name = "Log(Total Biomass Density)") +
+  scale_colour_discrete(labels = c("Observed", "Posterior Predictive"),
+                       name = '')
+
+```
+
+```{r mean_test_plot, fig.cap = "Posterior predictive mean (distribution) and empirical mean of total biomass density. Posterior predictive mean adequately recovers empirical mean."}
+
+ppc_stat(y,  yrep, stat = "mean") 
+```
+
+```{r mean_test_plot, fig.cap = "Posterior predictive sum (distribution) and empirical sum of total biomass density. Posterior predictive sum adequately recovers empirical sum."}
+
+ppc_stat(y,  yrep, stat = "sum") 
+```
+
+```{r sd_test_plot, fig.cap = "Posterior predictive standard deviation (distribution) and empirical standard deviation of total biomass density.
+Posterior predictive standard deviation adequately recovers empirical standard deviation"}
+ppc_stat(y,  yrep, stat = "sd")
+
+```
+
+```{r min_test_plot, fig.cap = "Posterior predictive minimum (distribution) and empirical minimum of log total biomass density. Model slightly overpredicts minimum value in the data"}
+
+ppc_stat(y,  yrep, stat = "min") 
+```
+
+```{r max_test_plot, fig.cap = "Posterior predictive maximum (distribution) and empirical maximum of total biomass density. Model underestimates true maximum of the data."}
+ppc_stat(y,  yrep, stat = "max")
+
+```
+
+```{r resid-plot,fig.cap = "Posterior predictive total biomass densities plotted against mean posterior predictive residuals" }
+ppc_error_scatter_avg_vs_x(y, yrep, colMeans(yrep)) + 
+  scale_x_continuous(name = "Predicted total biomass density") + 
+  scale_y_continuous(name = "Mean Posterior Predictive Residuals") + 
+  geom_hline(yintercept = 0)
+```
+
+```{r resid-hist-plot, fig.cap = "Histograms of residuals for targeted and non-targeted obesrvations."}
+
+temp <- did_data %>% 
+  mutate(targ = ifelse(targeted == 1, "Targeted","Non-Targeted"))
+ppc_error_hist_grouped(y, yrep = yrep[1:2,], group = temp$targ) + 
+  geom_vline(aes(xintercept = 0)) + 
+  scale_x_continuous(name = "Residuals")
+```
+
+### Additional Difference-in-difference Runs
+
+We include a variety of additional model runs designed to explore the sensitivity of our key results to various assumptions and data processing steps included in our base results.
+Description and key message are provided inside figure captions.
+
+Our base run uses data collected by PISCO.
+As a robustness check to our main results, we repeated our analysis utilizing data provided by the [Kelp Forest Monitoring Program (KFM)](https://science.nature.nps.gov/im/units/medn/monitor/kelpforest.cfm) conducted in the Channel Islands.
+
+```{r, fig.cap="Model fit to PISCO data but only using variables also available for KFM data", results="hide"}
+
+val_runs <- model_runs %>% 
+  mutate(did_results = map(did_fit,"did_results"))
+
+ggplot() +
+ tidybayes::stat_halfeye(
+    data = val_runs$did_results[[2]],
+    aes(year, did, color = "Empirical Estimate"),
+    .width = c(0.5, 0.95),
+    slab_alpha = 0.95  ) +
+  scale_y_continuous(name = "MPA Effect") +
+  scale_x_discrete(name = element_blank()) +
+  theme(legend.position = "top") +
+  scale_fill_brewer(name = "Simulation Quantiles") + 
+  scale_color_manual(values = "red", name = '') + 
+  geom_hline(yintercept = 0)
+
+```
+
+```{r, fig.cap="Model fit to KFM data. Results are more uncertain than using PISCO data, but follow the same trend estimated from the PISCO data."}
+
+ggplot() +
+ tidybayes::stat_halfeye(
+    data = val_runs$did_results[[3]],
+    aes(year, did, color = "Empirical Estimate"),
+    .width = c(0.5, 0.95),
+    slab_alpha = 0.95  ) +
+  scale_y_continuous(name = "MPA Effect") +
+  scale_x_discrete(name = element_blank()) +
+  theme(legend.position = "top") +
+  scale_fill_brewer(name = "Simulation Quantiles") + 
+  scale_color_manual(values = "red", name = '') + 
+    geom_hline(yintercept = 0)
+
+```
+
+```{r mpa-fit-plot, fig.cap="Estimated MPA effects using PISCO data from inside MPAs only.
+The base model uses data from both inside and outside MPAs.
+These results show that the same general trend holds when only using data from inside the MPAs, though the model estimates greater probability of higher positive effects with only MPA data, compared to the model fit using all the data.",results = "hide"}
+
+ggplot() +
+ tidybayes::stat_halfeye(
+    data = val_runs$did_results[[4]],
+    aes(year, did, color = "Empirical Estimate"),
+    .width = c(0.5, 0.95),
+    slab_alpha = 0.95  ) +
+  scale_y_continuous(name = "MPA Effect") +
+  scale_x_discrete(name = element_blank()) +
+  theme(legend.position = "top") +
+  scale_fill_brewer(name = "Simulation Quantiles") + 
+  scale_color_manual(values = "red", name = '') + 
+  geom_hline(yintercept = 0)
+
+```
+
+```{r, fig.cap="Estimated MPA effects using PISCO data from outside MPAs only.
+The base model uses data from both inside and outside MPAs.
+These results show that the same general trend holds when only using data from outside the MPAs, though the model estimates less probability of high positive effects witout MPA data, compared to the model fit using all the data.",results = "hide"}
+
+ggplot() +
+ tidybayes::stat_halfeye(
+    data = val_runs$did_results[[6]],
+    aes(year, did, color = "Empirical Estimate"),
+    .width = c(0.5, 0.95),
+    slab_alpha = 0.95  ) +
+  scale_y_continuous(name = "MPA Effect") +
+  scale_x_discrete(name = element_blank()) +
+  theme(legend.position = "top") +
+  scale_fill_brewer(name = "Simulation Quantiles") + 
+  scale_color_manual(values = "red", name = '') + 
+    geom_hline(yintercept = 0)
+
+```
+
+```{r no-filter-plot, fig.cap="The base model only includes consistently observed species of finfish.
+For this run, we include species that have been seen in at least two years (down from fifteen).
+The results remain essentially unchanged.",results = "hide"}
+no_filter <- model_runs %>%
+  slice(1) %>%
+  left_join(abundance_data, by = "data_source") %>%
+  mutate(
+    did_fit = pmap(
+      list(
+        data = data,
+        data_to_use = data_to_use,
+        data_source = data_source
+      ),
+      estimate_did,
+      min_classcode_years = 2,
+      iter = 2500,
+      site_data = site_data,
+      cdfw_catches = cdfw_catches,
+      life_history_data = life_history_data
+    )
+  )
+
+ggplot() +
+ tidybayes::stat_halfeye(
+    data = no_filter$did_fit[[1]]$did_results,
+    aes(year, did, color = "Empirical Estimate"),
+    .width = c(0.5, 0.95),
+    slab_alpha = 0.95  ) +
+  scale_y_continuous(name = "MPA Effect") +
+  scale_x_discrete(name = element_blank()) +
+  theme(legend.position = "top") +
+  scale_fill_brewer(name = "Simulation Quantiles") + 
+  scale_color_manual(values = "red", name = '') + 
+    geom_hline(yintercept = 0)
+
+```
+
+```{r no-bin-plot, fig.cap="For ease of interpretation and model convergence, our default model estimates MPA effects in three-year bins.
+For this run we estimate the effect annually, relative to the year 2000.
+Results show the same overall trend and magnitude reported in our base model.", results = "hide", eval = FALSE}
+no_bin <- model_runs %>%
+  slice(1) %>%
+  left_join(abundance_data, by = "data_source") %>%
+  mutate(
+    did_fit = pmap(
+      list(
+        data = data,
+        data_to_use = data_to_use,
+        data_source = data_source
+      ),
+      estimate_did,
+      iter = 2500,
+      first_year = 2001,
+      bin_width = 1,
+      site_data = site_data,
+      cdfw_catches = cdfw_catches,
+      life_history_data = life_history_data
+    )
+  )
+
+ggplot() +
+ tidybayes::stat_halfeye(
+    data = no_bin$did_fit[[1]]$did_results,
+    aes(year, did, color = "Empirical Estimate"),
+    .width = c(0.5, 0.95),
+    slab_alpha = 0.95  ) +
+  scale_y_continuous(name = "MPA Effect") +
+  scale_x_discrete(name = element_blank()) +
+  theme(legend.position = "top") +
+  scale_fill_brewer(name = "Simulation Quantiles") + 
+  scale_color_manual(values = "red", name = '') + 
+    scale_x_discrete(guide = guide_axis(n.dodge = 2))
+
+```
+
+```{r mean-fit, fig.cap="Our base run estimates the effect of MPAs on total biomass densities of targeted finfish.
+For this run, we instead estimate the effect on mean biomass densities of targeted species.
+Results show the same pattern as our base run.
+"}
+mean_fit <- model_runs %>%
+  slice(1) %>%
+  left_join(abundance_data, by = "data_source") %>%
+  mutate(
+    did_fit = pmap(
+      list(
+        data = data,
+        data_to_use = data_to_use,
+        data_source = data_source
+      ),
+      estimate_did,
+      fit_mean = TRUE,
+      iter = 2500,
+      site_data = site_data,
+      cdfw_catches = cdfw_catches,
+      life_history_data = life_history_data
+    )
+  )
+
+ggplot() +
+ tidybayes::stat_halfeye(
+    data = mean_fit$did_fit[[1]]$did_results,
+    aes(year, did, color = "Empirical Estimate"),
+    .width = c(0.5, 0.95),
+    slab_alpha = 0.95  ) +
+  scale_y_continuous(name = "MPA Effect") +
+  scale_x_discrete(name = element_blank()) +
+  theme(legend.position = "top") +
+  scale_fill_brewer(name = "Simulation Quantiles") + 
+  scale_color_manual(values = "red", name = '') + 
+    scale_x_discrete(guide = guide_axis(n.dodge = 2))
+
+```
+
+ exclude zeros
+
+```{r no-zero-fit, fig.cap="Our base run estimates the effect of MPAs on total biomass densities of targeted finfish.
+For this run, we instead estimate the effect on mean biomass densities of targeted species.
+Results show the same pattern as our base run.
+"}
+
+nozero_fit <- model_runs %>%
+  slice(1) %>%
+  left_join(abundance_data, by = "data_source")
+
+nozero_fit$data[[1]] <- nozero_fit$data[[1]] %>% filter(any_seen == TRUE)
+
+nozero_fit <-  nozero_fit%>%
+  mutate(
+    did_fit = pmap(
+      list(
+        data = data,
+        data_to_use = data_to_use,
+        data_source = data_source
+      ),
+      estimate_did,
+      iter = 2500,
+      site_data = site_data,
+      cdfw_catches = cdfw_catches,
+      life_history_data = life_history_data
+    )
+  )
+
+ggplot() +
+ tidybayes::stat_halfeye(
+    data = nozero_fit$did_fit[[1]]$did_results,
+    aes(year, did, color = "Empirical Estimate"),
+    .width = c(0.5, 0.95),
+    slab_alpha = 0.95  ) +
+  scale_y_continuous(name = "MPA Effect") +
+  scale_x_discrete(name = element_blank()) +
+  theme(legend.position = "top") +
+  scale_fill_brewer(name = "Simulation Quantiles") + 
+  scale_color_manual(values = "red", name = '') + 
+    scale_x_discrete(guide = guide_axis(n.dodge = 2))
+
+```
+ selection on observables
+
+```{r}
+
+```
+
+ ### Transect Level Difference-in-difference Model
+
+Our base model aggregated the data to the level of total biomass densities of targeted and non-targeted species.
+This serves to average out some of the sampling noise from the indivudla transect level, and we would argue is the appropriate level of estimation for many key policy questions.
+However, this approach does oversimplify the data collection process, treating the total biomass densities at the year/site level as estimated without error.
+
+As an added robustness test, we include version of the model in which we operate at the level of individual transect observations.
+The raw data are estimated length compositions by fish species along a survey transect at a site.
+Lengths are converted to biomass per allometric relationships supplied by PISCO and supplemented by the `FishLife` [@thorson2017c] package in R where needed.
+We performed some minimal data filtering to reduce noise in the data.
+We only include species that were observed at least twice in each year of the dataset (2000-2017) somewhere in the core Channel Islands (Anacapa, Santa Cruz, Santa Rosa, San Miguel).
+While some data are available from 1999, per consultation with PISCO we omit those data due to changes in survey protocols.
+We assign species to targeted and non-targeted groups per the PISCO classifications.
+This filtering process results in `r filter_summary$no` non-targeted species and `r filter_summary$yes` targeted species remaining in the analysis.
+
+The simplified explanation of the estimation is a hierarchical model in which we first standardize the observed biomass densities into an abundance index of each species over time.
+The abundance indices in each year are assumed to be log-normally distributed with means and standard deviations for the targeted and non-targeted groups, giving an estimate of the mean densities of targeted and non-targeted species over time.
+We then calculate the difference between mean density of targeted species and the mean density of non-targeted species in each year.
+
+\newpage
+```{r cartoon, fig.cap= "Cartoon illustration of the hierarchichal difference-in-difference estimator", fig.width=3}
+knitr::include_graphics("figs/estimation-cartoon.png")
+```
+
+\pagebreak
+
+The first stage of the regression is a log-normal delta model.
+The model estimates two regressions, the first is a binomial generalized linear model (GLM) with a logit link estimating the probability of observing a given fish species at a observation *i* (transect at time *t*).
+The probability that a given species was observed *o* at a given observation is distributed
+
+\begin{equation}
+o_{s,i} \sim binomial(\frac{1}{1 +e^{-\beta^{o}{X}}})
+  (\#eq:binom)
+\end{equation}
+
+ where $\beta^{o}$ are the estimated coefficients for the observation model and *X* is a matrix of covariates that include random effects for each year in the data (2000 to 2017).
+
+The expected density *d* of positive observations is modeled per a log-normal distribution
+
+ \begin{equation}
+  log(d_{s,i}) \sim normal(\beta^{d}X, \sigma_s)
+   (\#eq:logd)
+ \end{equation}
+
+where $\beta^{d}$ are the estimated coefficients for the expected density model and *X* is the same matrix of covariates as used in the observation portion of the model and $\sigma_s$ allows for each species *s* to have different standard deviations.
+
+Our covariate matrix *X* contains both fixed and random effects.
+Fixed effects include the depth level of the transect, the sampling site, the month of the observation, the estimated surge at the transect, visibility, the depth of the transect, and the experience (and experience squared) of the diver conducting the transect.
+We classify each species into one of two clusters based on the mean longitude the species was encountered at, breaking the species into two groups: those primarily found in the western end of the Channel Islands those found more in the eastern end.
+We then estimate random effects for each island for each cluster
+
+\begin{equation}
+\beta_{island,cluster} \sim normal(0,\sigma_{cluster})
+  (\#eq:island)
+\end{equation}
+
+This allows the mean effect of each island to differ for each cluster, e.g.
+allowing San Miguel, the easternmost island, to have a higher mean density for eastern species than for more western species (if the data suggest it).
+
+The second critical component of the covariate matrix *X* are random effects for each year for each species
+
+\begin{equation}
+  \beta_{year,species} \sim normal(0,\sigma_{species})
+  (\#eq:species)
+\end{equation}
+
+These $\beta_{year,species}$ represent our "standardized" estimate of observed abundance of each species in each time step, controlling for the included covariates.
+
+However, we still need to account for changes in the probability of detection over time.
+For that, we create a standard matrix of with rows equal to the number of years and columns corresponding to each of the columns in *X*, holding everything fixed at mean (or most frequently observed level for factors) levels for all variables in X except for the year and species interaction indices.
+Calling this standardized matrix $X^{standard}$, the probability of observing a given species in year *y* is then
+
+\begin{equation}
+p_{s,y} = (\frac{1}{1 +e^{-\beta^{o}{X^{standard}}}})
+  (\#eq:pdetect)
+\end{equation}
+
+In the same manner as described by @punt2000, The standardized index of abundance for species *s* in year *y* then is
+
+\begin{equation}
+I_{species,year} = p_{species,year}e^{\beta_{species,year}}
+  (\#eq:index)
+\end{equation}
+
+The next phase of the model requires us to estimate the mean abundance of targeted and non-targeted species over time.
+The concept here is that each $I_{species,year}$ can be modeled by a regression that contains random effects for each year for targeted and non-targeted fishes, the assumption then being that there is a mean density for targeted and non-target species, and $I_{species,year}$ represent deviations from that mean.
+
+\begin{equation}
+log(I_{species,year}) \sim normal(\beta^{effect}X^{effect}, \sigma_I)
+  (\#eq:indexprior)
+\end{equation}
+
+$X^{effect}$ contains both fixed and random effects. The fixed effects include an intercept and the temperature deviation for a given species in a year, where temperature deviation is
+
+\begin{equation}
+t_{s,y} = (t^{pref}_{s} -  \bar{t_{y}})^2
+  (\#eq:tempdev)
+\end{equation}
+
+where $t^{pref}$ is the preferred temperature for species *s* (drawn form `FishLife`, @thorson2017c), and $\bar{t_{y}}$ is the mean temperature encountered by that species in year *y*.
+We also include as variables in the model the mean kelp cover experienced by a given species in a given year, as well as the total fishery catches reported in the previous year for that species in the Santa Barbara region [drawn from the California Department of Fish and Wildlife database].
+We also include random intercepts for each species in $X^{effect}$.
+The most important random effects are year effects for targeted and non-targeted species
+
+\begin{equation}
+\beta_{year,targeted} \sim normal(0,\sigma_{targeted})
+  (\#eq:betatarg)
+\end{equation}
+
+$\beta_{year,targeted}$ is the mean log density of targeted species in year *y*, controlling for included covariates.
+Therefore, the final step in the model, the divergence in the standardized abundance trends of targeted and non-targeted species is
+
+\begin{equation}
+divergence_{year} =  \beta_{year,targeted = 1} - \beta_{year,targeted = 0}
+  (\#eq:mpaeffect)
+\end{equation}
+
+The model is fit in TMB to integrate the uncertainty across all levels of the model, with standard errors for each coefficient in the model estimated through the Laplace approximation.
+
+A complete table of estimated coefficients can be seen in Table.S\@ref(tab:betatable).
+
+<!-- Stock status estimates are not available for many of the targeted species in the database.
+The California Department of Fish and Wildlife compiles status estimates for many nearshore finfish (https://www.wildlife.ca.gov/Conservation/Marine/Groundfish/Nearshore-Finfish), including recent stock assessments.
+For species not listed here we searched for available information on recent stock status from relevant stock assessments.
+-->
+
+```{r betatable}
+
+table <- betas %>% select(-std_error,-group)
+
+knitr::kable(table,
+             digits = 2,
+             caption = "Complete table of estimated coefficients from hierarchical difference-in-difference model",
+             longtable = TRUE,
+             booktabs = TRUE,
+             col.names = colnames(table)) %>% 
+  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
+
+```
+
+Figures S\@ref(fig:fe-plot):S\@ref(fig:targ-plot) present estimated effects for covariates included in the model, along with the raw estimated mean trends of the targeted and non-targeted species (while the difference between these trends is presented in our main results).
+
+\newpage
+
+```{r fe-plot, fig.cap = "Estimated coefficients for non-spatial fixed effects in observation model (seeing) and observed model (seen)"}
 
  non_nested_beta_plot <- betas %>%
-   filter(str_detect(group, "non_nested")) %>%
+   filter(str_detect(group, "non_nested") & !str_detect(variable, "site_side")) %>%
+   ggplot() +
+   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
+   geom_pointrange(aes(x =variable,
+                       y = estimate,
+                       ymin = lower,
+                       ymax = upper)) +
+   facet_wrap(~group, labeller=labeller(group=c(
+     seeing_non_nested_betas="Observation Model\n(seeing)",
+     seen_non_nested_betas = "Observed Model\n(seen)"
+   ))) +
+   coord_flip() +
+   theme(axis.text.y = element_text(size = 10))
+
+non_nested_beta_plot
+
+```
+
+\pagebreak
+
+```{r site-plot, fig.cap = "Estimated coefficients for spatial random effects in observation model (seeing) and observed model (seen)", fig.height=10}
+
+ non_nested_site_beta_plot <- betas %>%
+   filter(str_detect(group, "non_nested") & str_detect(variable, "site_side")) %>%
+  mutate(variable = tolower(variable)) %>% 
+  mutate(variable = str_replace(variable, "site_side","")) %>% 
+  mutate(island = str_replace(str_split(variable, '_', simplify = TRUE)[,1],'-','')
+ ) %>% 
    ggplot() +
@@
                        ymax = upper)) +
-   facet_wrap(~group) +
+   facet_grid(island~group, scales = "free",labeller=labeller(group=c(
+     seeing_non_nested_betas="Observation Model\n(seeing)",
+     seen_non_nested_betas = "Observed Model\n(seen)"),
+     island=c(anacapa="ANA",sci="SCI",smi="SMI",sri="SRI"))) +
    coord_flip() +
-   theme(axis.text.y = element_text(size = 10))
-
- year_species_effects_plots <- betas %>%
-   filter(str_detect(group, "year_species_betas")) %>%
-   mutate(year = str_replace_all(variable,"\\D","") %>% as.numeric()) %>%
-   mutate(classcode = str_split(variable,'-', simplify = T)[,2]) %>%
-   ggplot() +
-   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
-   geom_ribbon(aes(x = year, ymin = lower, ymax = upper, fill = classcode), alpha = 0.25) +
-   geom_line(aes(x = year, y = estimate, color = classcode)) +
-   facet_wrap(~group)
+   theme(axis.text.y = element_text(size = 10)) + 
+  theme_minimal()
+
+non_nested_site_beta_plot + 
+  theme(axis.text.y = element_text(size = 6))
+
+```
+
+\pagebreak
+
+```{r region-plot, fig.cap="Random effects for region by cluster", eval = FALSE}
 
@@
                         ymax = upper, color = cluster)) +
-    facet_wrap(~group)
-
-  sim_plot <- outcomes %>%
-    mutate(rsize = plyr::round_any(mpa_size, 0.25),
-           reff = plyr::round_any(mpa_effect,0.1)) %>%
-    filter(rsize == 0.25) %>%
-    mutate(year = years_protected + 2001) %>%
-    filter(year %in% did_betas$year) %>% 
-    group_by(year, reff) %>% 
-    count() %>% 
-    group_by(year) %>% 
-    mutate(sc = n / sum(n))
-
-  did_plot <- did_betas %>%
+    facet_wrap(~group,labeller=labeller(group=c(
+     seeing_region_cluster_betas="Observation Model\n(seeing)",
+     seen_region_cluster_betas = "Observed Model\n(seen)")))
+
+region_cluster_plots
+
+```
+
+\pagebreak
+
+```{r targ-plot, fig.cap = "Trends in standardized mean abundance of targeted and non-targeted species"}
+
+targeted_trends <- report$zissou_estimates %>% 
+  filter(str_detect(variable, "targeted_did")) %>% 
+  group_by(variable) %>% 
+  mutate(year =zissou_fit$did_data$year %>% unique(),
+         labvar=ifelse(variable=="nontargeted_did_betas","Nontargeted","Targeted")) %>% 
+  ungroup() %>% 
   ggplot() + 
-    geom_tile(data = sim_plot, aes(year, reff, fill = sc), alpha = 0.75) +
-    geom_vline(aes(xintercept = 2003), color = 'red', linetype = 2, size = 2) +
-    geom_hline(aes(yintercept = 0)) +
-    geom_pointrange(aes(
-      year,
-      y = exp(estimate) - 1,
-      ymin = exp(lower) - 1,
-      ymax = exp(upper) - 1
-    ),
-    size = 1.5
+  geom_line(aes(year, estimate, color = labvar),alpha = 0.5) +
+  geom_pointrange(aes(year, estimate, ymin = lower, ymax = upper, color = labvar)) + 
+  scale_color_d3(name = "") +
+  theme(legend.position = c(0.8,0.9))
+
+targeted_trends
+
+```
+
+\pagebreak
+
+#### Transect Level Regression Diagnostics
+
+We include visual diagnostics of our estimation model. All coefficients passed convergence criteria for TMB.
+
+Looking first at the predictions of the model for the positive observations in the data (i.e.
+using the full model to predict biomass densities, and then comparing those predictions to cases where some positive biomass densities were observed), the model diagnostics show no clear problems.
+The $R^2$ of the model is 0.43.
+Residuals do not exhibit trends, though some grouping the residuals is evident.
+The quantile-quantile plot suggest that on the assumption of log-normal errors on the observed densities is reasonable, though the model appears to have some slight problems estimating the highest observed densities (Fig.S\@ref(fig:obs-v-pred-plot)).
+
+```{r}
+
+fit_report <- base_run$tmb_fit[[1]]$zissou_report
+
+observed_seen <- base_run$tmb_fit[[1]]$zissou_data$log_density
+
+observed_seeing <- base_run$tmb_fit[[1]]$zissou_data$any_seen
+
+seen_diagnostics <- data_frame(observed = observed_seen, predicted = as.numeric(fit_report$log_density_hat)) %>% 
+  mutate(residuals = predicted - observed)
+
+seeing_diagnostics <- data_frame(observed = observed_seeing, predicted = as.numeric(fit_report$prob_seeing))
+
+```
+
+```{r, eval = FALSE}
+
+a = base_run$tmb_fit[[1]]$zissou_data$x_seen_non_nested %>% 
+  as_data_frame() %>% 
+  select(contains("mean_")) %>% 
+  mutate(resid = seen_diagnostics$residuals) %>% 
+  gather(variable, value, -resid) %>% 
+  ggplot(aes(value, resid)) + 
+  geom_point() + 
+  facet_wrap(~variable)
+
+a
+
+```
+
+```{r obs-v-pred-plot, fig.cap="High level diagnostics for observed compontent of Delta-GLM: Observed vs predicted log densities (A), predicted log density vs residuals (B), and a normal qq-plot of the residuals (C)"}
+
+obs_v_pred_plot <- seen_diagnostics %>% 
+  ggplot(aes(observed, predicted)) + 
+  geom_point(alpha = 0.5) + 
+  geom_abline(aes(intercept = 0, slope = 1), linetype = 2, color = "red") + 
+  labs(caption = glue::glue("R^2 is {round(yardstick::rsq(data = seen_diagnostics, truth = observed, estimate = predicted)$.estimate,2)}"), title= "A") + 
+  theme(plot.margin = unit(rep(.1,4), units = "lines"))
+
+fit_v_resid_plot <- seen_diagnostics %>% 
+  ggplot(aes(predicted, residuals)) + 
+  geom_point() + 
+  geom_hline(aes(yintercept = 0), linetype = 2, color = "red") + 
+  labs(title = "B") +
+    theme(plot.margin = unit(rep(.1,4), units = "lines"))
+
+qq_plot <- ggplot(seen_diagnostics, aes(sample = residuals)) +
+  stat_qq() +
+  stat_qq_line() + 
+  labs(title = "C") + 
+    theme(plot.margin = unit(rep(.1,4), units = "lines"))
+
+obs_v_pred_plot/(fit_v_resid_plot + qq_plot)
+
+```
+
+In order to evaluate the ability of the model to estimate positive observations, we can compare the the binned predicted probability of a positive observation to the proportion of observations in that bin that recorded positive observations.
+If our model is working well, we would expect a group of fisheries that our model estimates on average should have a 50% probability of a positive observation, then we should expect about 50% of those observations to have positive observations.
+This is indeed what we see from the model (Fig.S\@ref(fig:obs-prob-plot)).
+
+```{r obs-prob-plot, fig.cap = "Binned mean predicted probability of detection provided by the first stage of the hurdle model vs observed proportion of positive detections"}
+
+seeing_diag_plot <- seeing_diagnostics %>% 
+  ungroup() %>% 
+  mutate(binned_prob = ntile(predicted, 20)) %>% 
+  arrange(binned_prob) %>% 
+  group_by(binned_prob) %>% 
+  summarise(mean_prob_seen = mean(predicted),
+            percent_observed = mean(observed))  %>% 
+  ggplot() +
+  geom_point(aes(mean_prob_seen, percent_observed)) + 
+  geom_abline(aes(intercept = 0, slope = 1), color = "red", linetype = 2) + 
+  labs(x = "Binned Predicted Probability of Detection", y = "Proportion of Positive Observations")
+
+seeing_diag_plot
+```
+
+We can also examine the receiver-operator-curve (ROC) to assess the performance of the observation component of the model.
+The area under the curve (AUC) value for the model is 0.84 (on a scale of 0.5 to 1), indicating the model is an overall good predictor of whether or not a given observation of biomass densities will be positive or not.
+
+```{r roc-plot, fig.cap = "Receiver operating characteristic curve of predictions of positive biomass densities"}
+library(plotROC)
+rocplot <- ggplot(seeing_diagnostics, aes(m = predicted, d = observed))+ geom_roc(n.cuts=20,labels=FALSE) + 
+  geom_abline(aes(slope = 1, intercept = 0))
+
+roc_plot <- rocplot + style_roc(theme = zissou_theme)
+
+roc_plot
+```
+
+#### Standardized Abundance Indices
+
+Overall most species showed consistent trends in biomass densities across the different islands at which they have been observed (Fig.S\@ref(fig:reg-pop)).
+
+```{r reg-pop, fig.cap = "Mean density by island by year for each fish species included in the analysis"}
+
+reg_pop_plot <- fitted_data %>% 
+  group_by(region, year, classcode) %>% 
+  summarise(mean_density = mean(exp(log_density))) %>% 
+  mutate(log_mean_density = log(mean_density)) %>% 
+  ungroup() %>% 
+  group_by(classcode, region) %>% 
+  mutate(cs_mean_density = (mean_density - mean(mean_density)) / sd(mean_density)) %>% 
+  ungroup() %>% 
+  ggplot(aes(year, cs_mean_density, color = region)) + 
+  scale_color_npg(name="Region")+
+  geom_line() + 
+  labs(y="Centered and Scaled Density",x="Year")+
+  facet_wrap(~classcode, scales = "free_y") + 
+  theme_minimal()+
+  theme(axis.text=element_text(size=6))
+
+reg_pop_plot
+
+```
+
+The standardized indices of abundance generally did not very substantially from the raw mean densities by species over time.
+However, for some species, such as blue rockfish, the standardized abundance index suggests much higher biomass densities in the pre-MPA period than those reported in the raw data.
+We suspect this is largely a function of changes in sampling sites over time, that the standardization is better able to account for (Fig.S\@ref(fig:raw-v-stand-plot)).
+
+```{r raw-v-stand-plot, fig.cap = "Raw (points) and standardized (lines) indices of abundance for each of the fishes included in the analysis"}
+
+life_history <- fitted_data %>%
+  select(classcode) %>%
+  unique() %>%
+  mutate(numeric_classcode = as.numeric(as.factor(classcode)))
+
+raw <- fitted_data %>%
+  group_by(year, classcode, targeted) %>%
+  summarise(md = mean(exp(log_density)),
+            mt = mean(temp_deviation)) %>%
+  group_by(classcode) %>%
+  mutate(smd = (md - mean(md))/(sd(md))) %>%
+  ungroup() %>%
+  left_join(life_history, by = "classcode")
+
+years <- unique(fitted_data$year)
+
+years <- zissou_fit$did_data$year %>% unique()
+
+abundance_trends <- data_frame(abundance_hat = fit_report$abundance_hat,
+                               classcode = rep(1:n_distinct(fitted_data$classcode), each = length(years))) %>%
+  mutate(log_abundance_hat = log(abundance_hat)) %>%
+  group_by(classcode) %>%
+  mutate(year = years) %>%
+  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
+  ungroup() %>%
+  rename(numeric_classcode = classcode) %>%
+  left_join(life_history, by = "numeric_classcode" )
+
+# 
+# abundance_trends %>%
+#   ggplot(aes(year, log(abundance_hat), color = factor(classcode))) +
+#   geom_line(show.legend = F) +
+#   geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
+#   facet_wrap(~classcode) +
+#   theme_minimal()
+
+raw_v_std_plot <- abundance_trends %>%
+  filter(year %in% years) %>% 
+  ggplot(aes(year, scaled_abundance_hat, color = factor(classcode))) +
+  geom_line(show.legend = F) +
+  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
+  geom_point(data = raw,aes(year, smd, color = factor(classcode)), show.legend = F)+
+  facet_wrap(~classcode) +
+  labs(y="Raw (dots) and Standardized (lines) Abundance",x="Year")+
+  theme_minimal()+
+  theme(axis.text=element_text(size=6))
+
+raw_v_std_plot
+
+```
+
+We include a variety of environmental, observation, and temporal indicators in our model.
+Inclusion of highly co-linear variables in a model can inflate standard errors and obscure "true" effects.
+To account for this we calculated the Pearson's correlation coefficients for all of the continuous data included in our model to ensure that none of the included variables had correlation coefficients greater than 0.7, a general rule of thumb for co-inclusion of variables.
+We did not find problematic levels of correlation among any of our included continuous variables.
+
+```{r numeric-cor, fig.cap="Pearson correlation coefficients of continuos data included in the regression model"}
+
+numeric_cor_plot <- base_run$data[[1]] %>%
+  select_if(is.numeric) %>%
+  select(-log_density, -targeted) %>%
+  na.omit() %>%
+  corrr::correlate() %>%
+  gather("colname", "correlation", -rowname) %>%
+  ggplot(aes(rowname, colname, fill = correlation)) +
+  geom_tile() +
+  geom_text(aes(rowname, colname, label = round(correlation, 2))) +
+  scale_fill_gradient2(
+    low = "tomato",
+    mid = "white",
+    high = "steelblue",
+    midpoint = 0,
+    limits = c(-1, 1)
   ) +
-    labs(x = "Year", y = "Targeted Trend Divergence") +
-    scale_fill_viridis(guide = hgc,
-                       name = "% of Annual Sims",
-                       labels = scales::percent_format(accuracy = 1),
-                       option = "plasma") + 
-    theme(legend.position = "top")
-
-top_species <- base_run$data[[1]]$classcode %>% unique()
-
-cip_data <- pisco_data %>%
-  left_join(site_data, by = c("site","side")) %>%
-  filter(is.na(eventual_mpa) == F) %>%
-  filter(region %in% c("ANA", "SCI","SRI",'SMI'),
-         classcode %in% top_species)
-
-# process outcomes
-
-filter_summary <- fitted_data %>% 
-  select(classcode, targeted) %>% 
-  unique() %>% 
-  group_by(targeted) %>% 
-  count() %>% 
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(
+      angle = 45,
+      vjust = 1,
+      hjust = 1,
+    size = 6),
+    axis.text.y = element_text(size = 6),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank()
+  )
+
+numeric_cor_plot
+```
+
+#### Sensitivity to "missing" observations
+
+Appropriately addressing the problem of "missing" observations is a critical challenge in any field observation study.
+If no observations of a given fish species were recorded on a given transect, should the density of that species on that transect be marked as zero, and influence the estimate of the overall mean density accordingly?
+The obvious answer seems to be yes, but what if that species simply does not live in the environment covered by a particular transect, or was not present during the particular time of the diver's observation?
+For our base runs, we assign a value of zero density on a given transect for any fish species that has been observed at least once at a given site at any time in our data but was not observed on that particular transect.
+If that species was never observed at that site, we do not include a zero for that species.
+Our rationale for this is that given the shifting nature of the sampled sites, and the intensity of sampling at those sites, we do not want to skew density trends by changes in the amount of suitable habitat for a given species sampled.
+However, this is clearly a strong assumption.
+For example, perhaps the decreasing trend in mean densities from 2000 to 2004 is due to increased number of sites (and therefore zeros) included in the data.
+To assess the potential importance of this choice, we can compare the mean densities of targeted and non-targeted species over time with the added zeros to the mean densities using only positive observations (i.e.
+not including any zeros in the data, (Fig.S\@ref(fig:nozero-raw-trend)).
+The trends in the raw densities, and most importantly the mean trends of targeted and non-targeted fishes, are nearly identical whether or not zeros are added, providing strong evidence that our choice of how to incorporate missing observations into the data are not strongly influencing our overall results.
+
+```{r nozero-raw-trend, fig.cap = "Centered and scaled mean annual density, excluding zeros, of included fishes (points) and smoothed means of targeted and non-targeted groups (line) over time" }
+
+nozero_trend_plot <- base_run$data[[1]] %>% 
+  filter(any_seen == T) %>% 
+  select(year, targeted,classcode, log_density, any_seen) %>% 
+  mutate(density = exp(log_density) * any_seen) %>% 
+  group_by(classcode,targeted, year) %>% 
+  summarise(md = mean(density)) %>% 
+  group_by(classcode,targeted) %>% 
+  mutate(smd = (md - mean(md)) / sd(md)) %>% 
+  ggplot(aes(year, smd, color = targeted == 1, fill = targeted == 1)) + 
+  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
+  geom_line(aes(group = interaction(targeted, classcode)), alpha = 0.25) +
+  geom_smooth() +
+  scale_color_d3(name = "Targeted?") + 
+  scale_fill_d3(name = "Targeted?") +
+  labs(y = "Centered and Scaled Mean Density", x = "Year")
+
+nozero_trend_plot
+```
+
+## Testing Model Assumptions
+
+### Simulation testing
+
+We state that a difference-in-difference model using targeted and non-targeted species is capable (conditional on assumptions) of estimating the causal effect of MPAs.
+We simulated MPA outcomes to test this claim.
+We first test our estimation strategy under idealized circumstances, where recruitment is deterministic and PISCO divers all have constant and perfect observer skills.
+We simulate five species that vary only in their maximum size and length at maturity.
+For each of these species, we set one version that is targeted by fishing and one that is not.
+We set a constant fishing mortality rate for each simulated targeted species, and then ran two matched simulations, one with MPAs and one without.
+We then have our simulated divers sample data from each of these scenarios, and then pass the sampled biomass densities to a simplified version of our difference-in-difference model (omitting the probability of detection step).
+We can then compare the difference-in-difference estimates of the MPA effect to the true simulated effect.
+The difference in difference model is able to capture the simulated MPA effect under these circumstances (Fig.S\@ref(fig:simple-did-test))
+
+```{r simple-did-test, fig.cap = "Simulated mean (red dashed line) and individual species (solid lines) MPA effects over time, along with difference-in-difference estimated MPA effects (mean with 95% confidence intervals)"}
+
+load(file = file.path(run_dir,'simulated_did.Rdata'))
+
+simple_performance$mixed_effect_did + 
+  labs(title = '', x = "Years with MPA protection",
+       y = "MPA Effect")
+```
+
+We then simulated a more complex example.
+We use the actual targeted and non-targeted species from our model.
+We assign species predominately seen in the western Channel Islands as "cold water" and those in the eastern Channel Islands as "warm water".
+We allow for stochasticity in recruitment.
+We use El Niño data as a simulated environmental recruitment driver, where we assume that El Niño events produce negative recruitment shocks for cold water species and *vice versa* for warm water species.
+We simulate three different divers each with different base skill levels, visual selectivities, and an evolving skill rate (such that observers get better over time).
+We hold fishing mortality rates constant for each species, although that fishing mortality affects each species differently because of intrinsic biological differences in maturity-at-age and steepness.
+We then test the ability of the difference-in-difference model to isolate the mean MPA effect across all of these targeted species, which our results show it is capable of doing (Fig.S\@ref(fig:complex-did-test)).
+
+```{r complex-did-test,  fig.cap = "Simulated mean (red dashed line) and individual species (solid lines) MPA effects over time, along with difference-in-difference estimated MPA effects (mean with 95% confidence intervals)"}
+
+pisco_performance$mixed_effect_did + 
+    labs(title = '', x = "Years with MPA protection",
+       y = "MPA Effect")
+
+```
+
+### Testing SUTVA with Convergent Cross Mapping
+
+The difference-in-difference model also assumes that the targeted and non-targeted fishes do not directly or indirectly affect each other.This assumption is clearly violated on some level: all the fishes in this analysis are part of the same ecosystem and therefore interact to some degree.
+For example, if the protection of targeted predatory fishes results in increased mortality of non-targeted fishes, the model would attribute that as an increased regional effect (greater divergence between the abundance of targeted and non-targeted species).
+Given the time scale of analysis (15 years of protection), we do not feel that massive trophic cascades are likely to have developed yet, given both the pace and complexity of trophic cascade development [@babcock2010; @pershing2015a].
+A complete assessment of evidence for trophic cascades in the Channel Islands is beyond the scope of this study, but to address this question somewhat we utilized convergent cross mapping *sensu* @sugihara2012 to test for a significant causal signal between different broad trophic groups in the data, implemented in the `rEDM` package in R.
+
+Convergent cross mapping is a nonlinear forecasting method that uses observed time series data to test for significant causal links between variables.
+Following methods laid out in @clark2015 and @sugihara2012, we pool the abundance of each broad trophic group by region (Fig.S\@ref(fig:trophic-plot).
+This uses the data from the islands as "replicates", requiring the assumption that the islands are all part of the same dynamic system, but allowing us to take advantage of the extra information provided by each island to further resolve the reconstructed manifolds.
+Using these aggregations, we then test whether the variables can be properly embedded, i.e., if they have predictable manifold dynamics.
+We do this through a simplex forecasting test, using an individual timeseries' own lags to build a manifold.
+For each timeseries, the "best embedding dimension" is an approximation of the dimensionality of the dynamic system, in other words, the number of dimensions that define and predict the evolving states of the timeseries.
+This analysis shows that only the carnivore, piscivore, and planktivores groups show evidence of significant predictability (that is, that past dynamics of these species groups can predict future dynamics, Fig.S\@ref(fig:embed-plot)).
+
+```{r trophic-plot, fig.cap="Centered and scaled densities by broad trophic group and island over time"}
+
+trophic_groups <- cip_data %>% 
+  select(classcode, broadtrophic) %>% 
+  unique()
+
+# dat <- abundance_trends %>%
+#   left_join(trophic_groups, by = "classcode") %>%
+#   group_by(broadtrophic, year) %>%
+#   summarise(density = mean(scaled_abundance_hat)) %>%
+#   group_by(broadtrophic) %>%
+#   mutate(density = scale(density)) %>%
+#   ungroup() %>% 
+#   spread(broadtrophic, density) %>% 
+#   arrange(year) %>% 
+#   mutate(lag4_carnivore = lag(carnivore,4),
+#          lag4_piscovore = lag(piscivore,4)) %>% 
+#   na.omit() %>% 
+#   ungroup() %>% 
+#   filter(year > 1999) 
+# dat %>%
+#   ggplot(aes(year, density, color = broadtrophic)) +
+#   geom_line()
+
+dat <- cip_data %>%
+  group_by(broadtrophic, region, site, side, transect, year, eventual_mpa) %>%
+  summarise(density = sum(density_g_m2)) %>%
+  group_by(broadtrophic, region, site, year) %>%
+  summarise(mean_density = mean(density, na.rm = T)) %>%
+  group_by(broadtrophic,region, year) %>%
+  summarise(regional_density = mean(mean_density, na.rm = T)) %>% 
   ungroup() %>% 
-  mutate(targeted = ifelse(targeted == 1,"yes","no")) %>% 
-  spread(targeted, n)
-
-```
-
-# Supporting Information (SI)
-
-# SI Text {#si-text .unnumbered}
-
-## Computing environment
-
-The computing environment needed to reproduce the results presented here is stored in a Docker container available [here]()XX.
-Interested users can load the container on their system to reproduce the complete analysis and manuscript.
-Analysis were performed in `r R.version.string`.
-Package versions are shown in Table.S\@ref(tab:pkg-info).
-
-```{r pkg-info, results = "asis"}
-table <- devtools::package_info() %>% 
-  select(package, loadedversion, date,source)
-
-knitr::kable(table,
-             digits = 2,
-             caption = "Package versions and sources used in this paper",
-             longtable = TRUE,
-             booktabs = TRUE,
-             format = "latex",
-             col.names = colnames(table)) %>% 
-  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
+  spread(broadtrophic, regional_density) %>% 
+  group_by(region) %>% 
+  arrange(year) %>% 
+  mutate(lag4_carnivore = lag(carnivore,4),
+         lag4_piscovore = lag(piscivore,4)) %>% 
+  na.omit() %>% 
+  ungroup() %>% 
+  filter(year > 1999)
+
+ dat <- dat %>%
+  select(-lag4_carnivore,-lag4_piscovore)%>%
+  group_by(region)%>%
+  mutate_at(vars(carnivore:planktivore),funs(norm=(.-mean(.))/sd(.)))->dat
+
+group_trends_plot <- dat %>%
+  select(region,year,contains("norm"))%>%
+  gather("group","density",-region,-year)%>%
+  ggplot(aes(year,density,col=group))+
+  geom_line()+
+  geom_point() +
+  scale_color_locuszoom(name="Trophic Group",labels=c("carnivore","herbivore","piscivore","planktivore"))+
+  labs(x="year",y="Centered and Scaled Density")+
+  facet_wrap(~region)
+
+group_trends_plot
+```
+
+```{r embed-plot, fig.cap = "Predictive skill as a function of embedding dimensions"}
+
+## split data by region to analyze timeseries from each island separately
+ana.dat <- dat %>% filter(region=="ANA")
+sci.dat <- dat %>% filter(region=="SCI")
+smi.dat <- dat %>% filter(region=="SMI")
+sri.dat <- dat %>% filter(region=="SRI")
+
+datnest <- dat %>% group_by(region) %>% nest()
+
+datnorm <- dat %>% select(region,year,contains('norm')) %>% ungroup()
+
+# have to record the segments corresponding to each "replicate" so simplex algorithm does not try to make predictions crossing time barriers
+segs <- datnorm %>% mutate(ind=row_number()) %>% group_by(region) %>% summarise(first=first(ind),last=last(ind)) %>%
+  select(-region)
+
+var_names <- c("carnivore_norm","herbivore_norm","piscivore_norm","planktivore_norm")
+
+regions.combined.simp.list <- map(var_names,function(x){
+  temp <- datnorm %>% ungroup() %>% select(matches(x)) %>% as.data.frame()
+  out <- simplex(as.numeric(temp[,1]),E=1:10,lib=as.matrix(segs),silent=T) %>%
+    mutate(trophic=str_replace(x,"_norm",""))
+  out
+})
+
+embed_plot <- bind_rows(regions.combined.simp.list) %>%
+  ggplot(aes(E,rho,color=trophic))+
+  geom_line(size=2)+
+  facet_wrap(~trophic,nrow=2,scales="free_y")+
+  geom_hline(yintercept = 0,color="black")+
+  labs(x="Embedding Dimension (E)",y=expression(paste("Predictive Skill, ",rho)))+
+  scale_x_continuous(breaks=seq(0,12,by=2))+
+  scale_color_locuszoom()+
+  guides(color=F)
+
+embed_plot
+
+```
+
+Focusing on just these three groups then (removing herbivores), we can test for causal relationships between groups using convergent cross mapping and the logic of Takens' theorem of dynamic systems.
+Generalizations of Takens' theorem indicate that if two variables (in our case, species or physical variables) are part of the same dynamic system, their individual dynamics should reflect their relative causal influence [@sugihara2012].
+In other words, if one variable is causally forced by another, that forcing should leave a signature on the first time series.
+Convergent cross mapping (CCM) tests for causation by using the attractor/manifold built from the time series of one variable to predict another (hence the "cross-mapping").
+In simple terms, the *causal effect of A on B is determined by how well B cross-maps A*.
+
+There are two criteria for CCM to establish causality:  First, and most obviously, predictive cross-map skill using all available data should be significantly greater than zero.
+Second, that predictability should be convergent.
+Convergence means that cross-mapped estimates improve with library length (the number of state-space vectors used to build the attractor), because the attractor is more fully resolved and therefore estimation error should decline.
+Convergence is key to distinguishing causation from simple or spurious correlation.
+If two variables are spuriously correlated and not causally linked, CCM should fail to satisfy this second criterion.
+Based on these criteria, there is little evidence of significant dynamic interactions between trophic groups (Fig.S\@ref(fig:cross-map-cp)-\@ref(fig:cross-map-pis)).
+Cross-mapping produced positive predictive skill, but was non-convergent for all cross-mappings with the exception of carnivores cross-mapping planktivores (providing some evidence that planktivore dynamics may be a driver of carnivore dynamics).
+This analysis provides evidence that trophic cascades are unlikely to be a significant driver of our results.
+It is important to note though that this analysis does not mean that trophic cascades could not emerge in this system, rather that we do not detect them with these data at this time.
+
+```{r cross-map-fxn}
+# plots a cross mapping between two variables in the data, given their variable names (grp), their optimal embedding dimensions (from simplex forecasting), and their preferred names for plotting
+xmap <- function(grp1,grp2,E1,E2,name1,name2){
+    # cross-map group 1 to group 2
+
+  ccm1 <- ccm(datnorm,lib=as.matrix(segs),pred=as.matrix(segs),E=E1,lib_column= grp1,target_column = grp2,lib_sizes = c(10,25,50,75),num_samples=100,replace=T,silent=T,RNGseed = 41389)
+
+  # test for convergence (is predictability at max library size significantly greater than at minimum library size?)
+
+  rhomeans1 <- ccm1 %>% ccm_means()
+
+  conv_test1 <- t.test(ccm1$rho[ccm1$lib_size==10],ccm1$rho[ccm1$lib_size==75])$p.value
+  if((rhomeans1$rho[rhomeans1$lib_size==75]-rhomeans1$rho[rhomeans1$lib_size==10])<0.05) conv_test1 <- 1
+  conv_test1 <- ifelse(conv_test1<0.05,"Convergent","Non-convergent")
+
+  xmap1 <- ccm1 %>%
+    group_by(lib_size)%>%
+    summarise(rhomean=mean(rho,na.rm=T),upper=quantile(rho, 0.975),lower=quantile(rho, 0.025))%>%
+    ungroup()%>%
+    ggplot(aes(lib_size,rhomean))+
+    geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.3,fill="red")+
+    geom_line(color="darkorchid3")+
+    geom_hline(aes(yintercept = 0), linetype = 2) +
+    labs(x="Library Size",y=expression(paste(rho, " (predictive skill)")),title=paste0(name1," xmap ",name2,"\n","(",conv_test1,")"))+
+    theme(plot.title = element_text(size=6))
+
+  # cross map group 2 to group 1
+  # inspect the output of simplex from the previous step and use the best embedding dimension (highest rho) for the carnivore time series
+
+  ccm2 <- ccm(datnorm,lib=as.matrix(segs),pred=as.matrix(segs),E=E2,lib_column= grp2,target_column = grp1,lib_sizes = c(10,25,50,75),num_samples=100,replace=T,silent=T,RNGseed = 41389)
+
+  rhomeans2 <- ccm2 %>% ccm_means()
+
+  conv_test2 <- t.test(ccm2$rho[ccm2$lib_size==10],ccm2$rho[ccm2$lib_size==75])$p.value
+  if((rhomeans2$rho[rhomeans2$lib_size==75]-rhomeans2$rho[rhomeans2$lib_size==10])<0.05) conv_test2 <- 1
+  conv_test2 <- ifelse(conv_test2<0.05,"Convergent","Non-convergent")
+
+  xmap2 <- ccm2 %>%
+    group_by(lib_size)%>%
+    summarise(rhomean=mean(rho,na.rm=T),upper=quantile(rho, 0.975),lower=quantile(rho, 0.025))%>%
+    ungroup()%>%
+    ggplot(aes(lib_size,rhomean))+
+    geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.3,fill="red")+
+    geom_line(color="darkorchid3")+
+    geom_hline(aes(yintercept = 0), linetype = 2) +
+    labs(x="Library Size",y=expression(paste(rho, " (predictive skill)")),title=paste0(name2," xmap ",name1,"\n","(",conv_test2,")"))+
+    theme(plot.title = element_text(size=6))
+
+  cross_map_plot <- xmap1 + xmap2
+
+cross_map_plot
+}
+```
+
+```{r cross-map-cp, fig.cap = "Cross mapping of effect of piscivores on carnivores (A) and carnivores on piscivores (B) in the PISCO data from 2000 to 2017.
+Shaded region show 95% confidence interval"}
+
+car_pis<-xmap(grp1="carnivore_norm",grp2="piscivore_norm",E1=4,E2=4,name1="Carnivores",name2="Piscivores")
+car_pl <- xmap(grp1="carnivore_norm",grp2="planktivore_norm",E1=4,E2=4,name1="Carnivores",name2="Planktivores")
+pl_pis <- xmap(grp1="planktivore_norm",grp2="piscivore_norm",E1=4,E2=4,name1="Planktivores",name2="Piscivores")
+
+car_pis
+```
+
+```{r cross-map-pl, fig.cap = "Cross mapping of effect of planktivores on carnivores (A) and carnivores on planktivores (B) in the PISCO data from 2000 to 2017.
+Shaded region show 95% confidence interval"}
+
+car_pl
+
+```
+
+```{r cross-map-pis, fig.cap = "Cross mapping of effect of piscivores on carnivores (A) and carnivores on planktivores (B) in the PISCO data from 2000 to 2017.
+Shaded region show 95% confidence interval"}
+
+ pl_pis 
+```
+
+```{r trophic-cor-plot, fig.cap="Estimated correlation coefficients between broad trophic groups at the island by year resolution", eval = FALSE}
+
+trophic_effects_plot <- cip_data %>%
+  group_by(broadtrophic, region, site, side, transect, year, eventual_mpa) %>%
+  summarise(density = sum(density_g_m2)) %>%
+  group_by(broadtrophic, region, site, year) %>%
+  summarise(mean_density = mean(density, na.rm = T)) %>%
+  group_by(broadtrophic,region, year) %>%
+  summarise(regional_density = mean(mean_density, na.rm = T)) %>% 
+  ungroup() %>% 
+  spread(broadtrophic, regional_density) %>% 
+  group_by(region) %>% 
+  arrange(year) %>% 
+  mutate(lag4_carnivore = lag(carnivore,4),
+         lag4_piscovore = lag(piscivore,4)) %>% 
+  na.omit() %>% 
+  ungroup() %>% 
+  filter(year > 1999) %>% 
+  select(-year,-region) %>% 
+  corrr::correlate() %>% 
+   gather("colname","correlation",-rowname) %>% 
+  ggplot(aes(rowname, colname,fill = correlation)) + 
+  geom_tile() + 
+  geom_text(aes(rowname, colname, label = round(correlation,2))) + 
+  scale_fill_gradient2(low = "tomato", mid = "white", high = "steelblue", midpoint = 0, limits = c(-1,1),
+                       guide = guide_colorbar(frame.colour = "black",barheight = 13)) + 
+  theme(axis.text.x = element_text(angle = 45, vjust = 1,
+                                   hjust = 1),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank())
+
+trophic_effects_plot
 
@@
    facet_wrap(~measure, labeller = labeller(measure = facet_labels), strip.position = "bottom") +
-   scale_x_percent(name = "") +
-   scale_y_percent(name = "Pop. Conservation Effect") + 
+   scale_x_percent(name = "", expand = c(0,0)) +
+   scale_y_percent(name = "Pop. Conservation Effect", expand = c(0,0)) + 
    theme(legend.position = "top")
@@
 <!-- #  -->
-<!-- # ``` -->
-
-<!-- At the extreme sizes of MPAs there is clearly a tension between conservation and fishery goals.
-However, most MPAs are relatively small (XX), and therefore likely to cover lower percentages of a population as simulated here.
-A key question then is how do tradeoffs   -->
-
-```{r mpa-tradeoffs, eval = FALSE}
-
-tradeoffs <- fishery_outcomes %>%
-  select(experiment, year, years_protected,mpa_size, fishery_effect) %>%
-  left_join(outcomes %>%
-  select(experiment,year, mpa_effect), by = c("experiment","year")
-)
-
-vplot <- tibble(fishery_effect = -1:1, mpa_effect = -1:1) %>%
-  mutate(slope = ifelse(fishery_effect < 0, -1, 1)) %>%
-  mutate(vline = slope * fishery_effect)
-
-tradeoff_kobe_plot = tradeoffs %>%
-  group_by(experiment) %>%
-  filter(years_protected == max(years_protected)) %>%
-  ungroup() %>%
-  mutate(mpa_bin = cut(mpa_size, breaks = seq(0, 1, by = 0.25))) %>%
-  ggplot(aes(fishery_effect, mpa_effect)) +
-  geom_hline(aes(yintercept = 0), linetype = 2) +
-  geom_vline(aes(xintercept = 0), linetype = 2) +
-  geom_hex(bins = 5,    aes(fill = ..density..),color = "black") +
-    geom_line(data = vplot,aes(fishery_effect, vline), color = "red",
-              linetype = 2, size = 2) +
-
-  # geom_abline(aes(slope = -1, intercept = 0), linetype = 2, color = "red", size = 2) +
-  #   geom_abline(aes(slope = 1, intercept = 0), linetype = 2, color = "black", size = 2) +
-  scale_fill_distiller(
-    # option = "plasma",
-    labels = scales::percent_format(accuracy = 1),
-    name = "% of Sims",
-    guide = hgc
-  ) +
-  # geom_point(size = 2, alpha = 0.2, aes(color = factor(mpa_bin))) +
-  # scale_color_viridis_d() +
-  facet_wrap( ~ mpa_bin) +
-  theme_minimal() +
-  theme(legend.position = "top") +
-    scale_x_continuous(labels = scales::percent_format(accuracy = 1),
-                     name = "% Change Fishery") +
-      scale_y_continuous(labels = scales::percent_format(accuracy = 1),
-                     name = "% Change Conservation") + 
-  labs(caption = "facets are percent of range in MPA. Red dashed line shows points where fisheries losses/gains are equal to conservation gains")
-
-# suppose that utility is the sum of percentage gain in conservation and fishery
-
-tradeoff_bins <- tradeoffs %>%
-  group_by(experiment) %>%
-  filter(years_protected == max(years_protected)) %>%
-  mutate(tradeoff = mpa_effect + fishery_effect) %>%
-  mutate(tradeoff_bin = plyr::round_any(tradeoff,.25),
-         mpa_bin = plyr::round_any(mpa_size,.1)) %>%
-  group_by(tradeoff_bin, mpa_bin) %>%
-  count() %>%
-  group_by(mpa_bin) %>%
-  mutate(n = n / sum(n)) %>%
-  ungroup()
-
-tradeoff_summary <- tradeoffs %>%
-  group_by(experiment) %>%
-  filter(years_protected == max(years_protected)) %>%
-  mutate(tradeoff = mpa_effect + fishery_effect) %>%
-  mutate(mpa_bin = plyr::round_any(mpa_size,.1)) %>%
-  group_by(mpa_bin) %>%
-  summarise(net_tradeoff = mean(tradeoff > 0))
-
-tradeoff_density_plot <- tradeoff_bins %>%
-  ggplot(aes(mpa_bin, tradeoff_bin, fill = n)) +
-  # geom_hline(aes(yintercept = 0), linetype = 2, size = 2) +
-  geom_tile(alpha = 0.9) +
-  scale_fill_viridis(
-    option = "plasma",
-    labels = scales::percent_format(accuracy = 1),
-    name = "% of Sims",
-    guide = hgc
-  ) +
-  scale_x_continuous(labels = scales::percent_format(accuracy = 1),
-                     name = "% of Range in MPA") +
-  scale_y_continuous(name = "Net Utility") +
-  theme(legend.position = "top")
-
-tradeoff_summary_plot <- tradeoff_summary %>%
-  ggplot(aes(mpa_bin, net_tradeoff)) +
-  geom_col() +
-  scale_x_continuous(labels = scales::percent_format(accuracy = 1),
-                     name = "% of Range in MPA") +
-  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
-                     name = "% Positive Utility")
-
-# tradeoff_density_plot + tradeoff_summary_plot & theme(
-#     plot.margin = unit(c(0, 1, 0, 1), units = "lines"),
-#     axis.text.x = element_text(size = 8),
-#     legend.box.margin = unit(c(0, 0, 0, 0), units = "lines"),
-#     axis.text.y = element_text(size = 10)
-#     )
-
-```
-
-<!-- why is it going to be hard ot detect regional effects of MPAs few if any empirical studies at net regional effects.
-does this mean that i doesn't exist, or impossible tot est, why?
-three fundamental stumbling blocks -->
-
-## Estimation Model
-
-The goal of the estimation model is to estimate the difference in the mean densities of targeted and non-targeted species pre-and-post MPA implementation.
-
-### Data
-
-All observation data used in the estimation model were collected by PISCO.
-PISCO staff also compiled allometric information used to convert lengths to expected weights, and hence biomass densities.
-We do not account for error in this translation step Table.S\@ref(tab:classcode) lists the species included in the model.
-In order to be included in the estimation model, a species must have been observed at least twice a year every year for at least 14 years.
-We also omitted all observations of "young of the year" fish due to inability to identify these observations to the nearest species level in many cases.
-We omitted data from 1999 due to changes in the sampling procedures that occurred after 1999.
-Per recommendations from PISCO staff we omit observations from the canopy level of the transects (leaving the middle, bottom, and middle canopy levels).
-
-PISCO data report positive observations of fish.
-In order to use these data in our model we need to add in zeros for any transect that could have observed a given species of fish but did not.
-We assume that a fish could have been observed on a given transect if that species has ever been observed at that site in any time period in the data (PISCO data are organized by sites, with multiple transect at different locations within the borders of a site).
-If a species has never been observed at a site we assume that it does not occur at that site.
-
-Interested readers can examine the mean biomass densities of all species in Fig.\@ref(fig:targ-trend-plot)-\@ref(fig:nontarg-trend-plot).
-
-We include several additional sources of data in our regression analysis.
-Temperature readings are included from the PISCO data for each transect.
-We also include PISCO data on the estimated surge and visibility.
-We augmented these data with information on kelp cover over time from the Santa Barbara Channel Long Term Ecological Research Network [@lter2017].
-We used a k-nearest neighbors algorithm to fill in missing kelp observations, and matched the interpolated kelp data to the PISCO data at the resolution of year-month-site (Fig.\@ref(fig:kelp-plot)).
-
-Temperature data were augmented with data from `FishLife` [@thorson2017c] to include the estimated preferred temperature for a given species, so that we can include deviations from the preferred temperature envelope as a predictor in the model.
-This allows different temperatures to have different effects on each species (and is less computationally intensive than estimating species-temperature slopes) (Fig.\@ref(fig:tempdev-plot)).
-
-We also included lagged catch totals in the Santa Barbara region for the commercially harvest species in the database, in an effort to control for changes in density caused by changes in fishing pressure.
-Catches were pulled from the CDFW website (https://www.wildlife.ca.gov/Fishing/Commercial/Landings), and extracted using the `tabulizer` package in R [@leeper2018] (Fig.\@ref(fig:catches-plot)).
-
-```{r tempdev-plot, fig.cap = "Mean deviations from preferred temperature by species and year"}
-
-abundance_data$data[[1]] %>% 
-  group_by(commonname, year) %>% 
-  summarise(mean_tempdev = mean(temp_deviation)) %>% 
-  ggplot(aes(year, mean_tempdev)) + 
-  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
-  geom_line() + 
-  facet_wrap(~commonname, scales = "free_y") + 
-  theme_minimal() + 
-  labs(x = "Year", y = "Deviation from mean preferred temperature")
-
-```
-
-```{r catches-plot, fig.cap = "Total CDFW reported commercial catches in the Santa Barbara region"}
-
-abundance_data$data[[1]] %>% 
-  filter(targeted == 1) %>% 
-  group_by(commonname, year) %>% 
-  summarise(catches = unique(lag_catch)) %>% 
-  group_by(commonname) %>% 
-  mutate(total_catches = sum(catches)) %>% 
-  ungroup() %>% 
-  filter(total_catches > 0) %>% 
-  ggplot(aes(year, catches)) + 
-  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
-  geom_line() + 
-  facet_wrap(~commonname, scales = "free_y") + 
-  theme_minimal() + 
-  labs(x = "Year", y = "Catch (lbs)")
-
-```
-
-```{r kelp-plot, fig.cap="Mean kelp biomass by island over time from SBC LTER"}
-
-abundance_data$data[[1]] %>% 
-  group_by(region,year) %>% 
-  summarise(mean_kelp = mean(interp_kelp, na.rm = TRUE)) %>% 
-  ggplot(aes(year, (mean_kelp), color = region)) + 
-  geom_line() + 
-  labs(x = "Year", y = "Mean Kelp Biomass")
-
-```
-
-```{r targ-trend-plot, fig.cap="Centered and scaled mean biomass densities of all targeted finfish in analysis before filtering"}
-
-targeted_trend_plot <- pisco_data %>% 
-  filter(targeted == 1) %>% 
-  group_by(year, classcode, targeted, commonname) %>% 
-  summarise(mean_biomass_density = mean(density_g_m2, na.rm = TRUE)) %>% 
-  group_by(classcode) %>% 
-  mutate(mean_biomass_density = scale(mean_biomass_density)) %>% 
-  ggplot(aes(year,mean_biomass_density)) + 
-  geom_hline(aes(yintercept = 0), linetype = 2) + 
-  geom_vline(aes(xintercept = 2003), color = "red") +
-  geom_line() + 
-  facet_wrap(~commonname) +
-  theme_minimal() + 
-  theme(strip.text = element_text(size = 6),
-        axis.text.x = element_text(size = 8)) + 
-  labs(x = "Year", y = "Centered and Scaled Mean Biomass Density")
-
-targeted_trend_plot
-```
-
-```{r nontarg-trend-plot, fig.cap="Centered and scaled mean biomass densities of all non-targeted finfish in analysis before filtering"}
-
-nontargeted_trend_plot <- pisco_data %>% 
-  filter(targeted == 0) %>% 
-  group_by(year, classcode, targeted, commonname) %>% 
-  summarise(mean_biomass_density = mean(density_g_m2, na.rm = TRUE)) %>% 
-  group_by(classcode) %>% 
-  mutate(mean_biomass_density = scale(mean_biomass_density)) %>% 
-  ggplot(aes(year,mean_biomass_density)) + 
-  geom_hline(aes(yintercept = 0), linetype = 2) + 
-  geom_vline(aes(xintercept = 2003), color = "red") +
-  geom_line() + 
-  facet_wrap(~commonname) +
-  theme_minimal() + 
-  theme(strip.text = element_text(size = 6),
-        axis.text.x = element_text(size = 8)) + 
-  labs(x = "Year", y = "Centered and Scaled Mean Biomass Density")
-
-nontargeted_trend_plot
-```
-
-```{r classcode}
-
-species_data <- cip_data %>% 
-  select(classcode, commonname, taxa, targeted) %>% 
-  unique() %>% 
-  arrange(desc(targeted)) %>% 
-  mutate(targeted = targeted == 1) %>% 
-  rename(`Common Name` = commonname, 
-         `Scientific Name` = taxa,
-         `Targeted?` = targeted)
-
-status <- read_csv(here::here("data","stock-status-summaries.csv")) %>% 
-  select(classcode, `Stock Status`)
-
-species_data <- species_data %>% 
-  left_join(status, by = "classcode")
-
-knitr::kable(species_data,
-             digits = 2,
-             caption = "Species included in estimation model",
-             longtable = TRUE,
-             booktabs = TRUE,
-             col.names = colnames(species_data)) %>%
-  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
-
-```
-
-### Model
-
-The regression analysis uses a mixed-effects hierarchical model.
-The raw data are estimated length compositions by fish species along a survey transect at a site.
-Lengths are converted to biomass per allometric relationships supplied by PISCO and supplemented by the `FishLife` [@thorson2017c] package in R where needed.
-We performed some minimal data filtering to reduce noise in the data.
-We only include species that were observed at least twice in each year of the dataset (2000-2017) somewhere in the core Channel Islands (Anacapa, Santa Cruz, Santa Rosa, San Miguel).
-While some data are available from 1999, per consultation with PISCO we omit those data due to changes in survey protocols.
-We assign species to targeted and non-targeted groups per the PISCO classifications.
-This filtering process results in `r filter_summary$no` non-targeted species and `r filter_summary$yes` targeted species remaining in the analysis.
-
-The simplified explanation of the estimation is a hierarchical model in which we first standardize the observed biomass densities into an abundance index of each species over time.
-The abundance indices in each year are assumed to be log-normally distributed with means and standard deviations for the targeted and non-targeted groups, giving an estimate of the mean densities of targeted and non-targeted species over time.
-We then calculate the difference between mean density of targeted species and the mean density of non-targeted species in each year.
-
-\newpage
-```{r cartoon, fig.cap= "Cartoon illustration of the hierarchichal difference-in-difference estimator", fig.width=3}
-knitr::include_graphics("figs/estimation-cartoon.png")
-```
-
-\pagebreak
-
-The first stage of the regression is a log-normal delta model.
-The model estimates two regressions, the first is a binomial generalized linear model (GLM) with a logit link estimating the probability of observing a given fish species at a observation *i* (transect at time *t*).
-The probability that a given species was observed *o* at a given observation is distributed
-
-\begin{equation}
-o_{s,i} \sim binomial(\frac{1}{1 +e^{-\beta^{o}{X}}})
-  (\#eq:binom)
-\end{equation}
-
- where $\beta^{o}$ are the estimated coefficients for the observation model and *X* is a matrix of covariates that include random effects for each year in the data (2000 to 2017).
-
-The expected density *d* of positive observations is modeled per a log-normal distribution
-
- \begin{equation}
-  log(d_{s,i}) \sim normal(\beta^{d}X, \sigma_s)
-   (\#eq:logd)
- \end{equation}
-
-where $\beta^{d}$ are the estimated coefficients for the expected density model and *X* is the same matrix of covariates as used in the observation portion of the model and $\sigma_s$ allows for each species *s* to have different standard deviations.
-
-Our covariate matrix *X* contains both fixed and random effects.
-Fixed effects include the depth level of the transect, the sampling site, the month of the observation, the estimated surge at the transect, visibility, the depth of the transect, and the experience (and experience squared) of the diver conducting the transect.
-We classify each species into one of two clusters based on the mean longitude the species was encountered at, breaking the species into two groups: those primarily found in the western end of the Channel Islands those found more in the eastern end.
-We then estimate random effects for each island for each cluster
-
-\begin{equation}
-\beta_{island,cluster} \sim normal(0,\sigma_{cluster})
-  (\#eq:island)
-\end{equation}
-
-This allows the mean effect of each island to differ for each cluster, e.g.
-allowing San Miguel, the easternmost island, to have a higher mean density for eastern species than for more western species (if the data suggest it).
-
-The second critical component of the covariate matrix *X* are random effects for each year for each species
-
-\begin{equation}
-  \beta_{year,species} \sim normal(0,\sigma_{species})
-  (\#eq:species)
-\end{equation}
-
-These $\beta_{year,species}$ represent our "standardized" estimate of observed abundance of each species in each time step, controlling for the included covariates.
-
-However, we still need to account for changes in the probability of detection over time.
-For that, we create a standard matrix of with rows equal to the number of years and columns corresponding to each of the columns in *X*, holding everything fixed at mean (or most frequently observed level for factors) levels for all variables in X except for the year and species interaction indices.
-Calling this standardized matrix $X^{standard}$, the probability of observing a given species in year *y* is then
-
-\begin{equation}
-p_{s,y} = (\frac{1}{1 +e^{-\beta^{o}{X^{standard}}}})
-  (\#eq:pdetect)
-\end{equation}
-
-In the same manner as described by @punt2000, The standardized index of abundance for species *s* in year *y* then is
-
-\begin{equation}
-I_{species,year} = p_{species,year}e^{\beta_{species,year}}
-  (\#eq:index)
-\end{equation}
-
-The next phase of the model requires us to estimate the mean abundance of targeted and non-targeted species over time.
-The concept here is that each $I_{species,year}$ can be modeled by a regression that contains random effects for each year for targeted and non-targeted fishes, the assumption then being that there is a mean density for targeted and non-target species, and $I_{species,year}$ represent deviations from that mean.
-
-\begin{equation}
-log(I_{species,year}) \sim normal(\beta^{effect}X^{effect}, \sigma_I)
-  (\#eq:indexprior)
-\end{equation}
-
-$X^{effect}$ contains both fixed and random effects. The fixed effects include an intercept and the temperature deviation for a given species in a year, where temperature deviation is
-
-\begin{equation}
-t_{s,y} = (t^{pref}_{s} -  \bar{t_{y}})^2
-  (\#eq:tempdev)
-\end{equation}
-
-where $t^{pref}$ is the preferred temperature for species *s* (drawn form `FishLife`, @thorson2017c), and $\bar{t_{y}}$ is the mean temperature encountered by that species in year *y*.
-We also include as variables in the model the mean kelp cover experienced by a given species in a given year, as well as the total fishery catches reported in the previous year for that species in the Santa Barbara region [drawn from the California Department of Fish and Wildlife database].
-We also include random intercepts for each species in $X^{effect}$.
-The most important random effects are year effects for targeted and non-targeted species
-
-\begin{equation}
-\beta_{year,targeted} \sim normal(0,\sigma_{targeted})
-  (\#eq:betatarg)
-\end{equation}
-
-$\beta_{year,targeted}$ is the mean log density of targeted species in year *y*, controlling for included covariates.
-Therefore, the final step in the model, the divergence in the standardized abundance trends of targeted and non-targeted species is
-
-\begin{equation}
-divergence_{year} =  \beta_{year,targeted = 1} - \beta_{year,targeted = 0}
-  (\#eq:mpaeffect)
-\end{equation}
-
-The model is fit in TMB to integrate the uncertainty across all levels of the model, with standard errors for each coefficient in the model estimated through the Laplace approximation.
-
-A complete table of estimated coefficients can be seen in Table.S\@ref(tab:betatable).
-
-<!-- Stock status estimates are not available for many of the targeted species in the database.
-The California Department of Fish and Wildlife compiles status estimates for many nearshore finfish (https://www.wildlife.ca.gov/Conservation/Marine/Groundfish/Nearshore-Finfish), including recent stock assessments.
-For species not listed here we searched for available information on recent stock status from relevant stock assessments.
--->
-
-```{r betatable}
-
-table <- betas %>% select(-std_error,-group)
-
-knitr::kable(table,
-             digits = 2,
-             caption = "Complete table of estimated coefficients from hierarchical difference-in-difference model",
-             longtable = TRUE,
-             booktabs = TRUE,
-             col.names = colnames(table)) %>% 
-  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
-
-```
-
-Figures S\@ref(fig:fe-plot):S\@ref(fig:targ-plot) present estimated effects for covariates included in the model, along with the raw estimated mean trends of the targeted and non-targeted species (while the difference between these trends is presented in our main results).
-
-\newpage
-
-```{r fe-plot, fig.cap = "Estimated coefficients for non-spatial fixed effects in observation model (seeing) and observed model (seen)"}
-
- non_nested_beta_plot <- betas %>%
-   filter(str_detect(group, "non_nested") & !str_detect(variable, "site_side")) %>%
-   ggplot() +
-   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
-   geom_pointrange(aes(x =variable,
-                       y = estimate,
-                       ymin = lower,
-                       ymax = upper)) +
-   facet_wrap(~group, labeller=labeller(group=c(
-     seeing_non_nested_betas="Observation Model\n(seeing)",
-     seen_non_nested_betas = "Observed Model\n(seen)"
-   ))) +
-   coord_flip() +
-   theme(axis.text.y = element_text(size = 10))
-
-non_nested_beta_plot
-
-```
-
-\pagebreak
-
-```{r site-plot, fig.cap = "Estimated coefficients for spatial random effects in observation model (seeing) and observed model (seen)", fig.height=10}
-
- non_nested_site_beta_plot <- betas %>%
-   filter(str_detect(group, "non_nested") & str_detect(variable, "site_side")) %>%
-  mutate(variable = tolower(variable)) %>% 
-  mutate(variable = str_replace(variable, "site_side","")) %>% 
-  mutate(island = str_replace(str_split(variable, '_', simplify = TRUE)[,1],'-','')
- ) %>% 
-   ggplot() +
-   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
-   geom_pointrange(aes(x = variable,
-                       y = estimate,
-                       ymin = lower,
-                       ymax = upper)) +
-   facet_grid(island~group, scales = "free",labeller=labeller(group=c(
-     seeing_non_nested_betas="Observation Model\n(seeing)",
-     seen_non_nested_betas = "Observed Model\n(seen)"),
-     island=c(anacapa="ANA",sci="SCI",smi="SMI",sri="SRI"))) +
-   coord_flip() +
-   theme(axis.text.y = element_text(size = 10)) + 
-  theme_minimal()
-
-non_nested_site_beta_plot + 
-  theme(axis.text.y = element_text(size = 6))
-
-```
-
-\pagebreak
-
-```{r region-plot, fig.cap="Random effects for region by cluster", eval = FALSE}
-
-  region_cluster_plots <- betas %>%
-   filter(str_detect(group, "region_cluster_betas")) %>%
-   mutate(cluster = str_replace_all(variable,"\\D","")) %>%
-   mutate(region = str_split(variable,'-', simplify = T)[,3]) %>%
-   ggplot() +
-    geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
-    geom_pointrange(aes(x = region,
-                        y = estimate,
-                        ymin = lower,
-                        ymax = upper, color = cluster)) +
-    facet_wrap(~group,labeller=labeller(group=c(
-     seeing_region_cluster_betas="Observation Model\n(seeing)",
-     seen_region_cluster_betas = "Observed Model\n(seen)")))
-
-region_cluster_plots
-
-```
-
-\pagebreak
-
-```{r targ-plot, fig.cap = "Trends in standardized mean abundance of targeted and non-targeted species"}
-
-targeted_trends <- report$zissou_estimates %>% 
-  filter(str_detect(variable, "targeted_did")) %>% 
-  group_by(variable) %>% 
-  mutate(year =zissou_fit$did_data$year %>% unique(),
-         labvar=ifelse(variable=="nontargeted_did_betas","Nontargeted","Targeted")) %>% 
-  ungroup() %>% 
-  ggplot() + 
-  geom_line(aes(year, estimate, color = labvar),alpha = 0.5) +
-  geom_pointrange(aes(year, estimate, ymin = lower, ymax = upper, color = labvar)) + 
-  scale_color_d3(name = "") +
-  theme(legend.position = c(0.8,0.9))
-
-targeted_trends
-
-```
-
-\pagebreak
-
-### Regression Diagnostics
-
-We include visual diagnostics of our estimation model. All coefficients passed convergence criteria for TMB.
-
-Looking first at the predictions of the model for the positive observations in the data (i.e.
-using the full model to predict biomass densities, and then comparing those predictions to cases where some positive biomass densities were observed), the model diagnostics show no clear problems.
-The $R^2$ of the model is 0.43.
-Residuals do not exhibit trends, though some grouping the residuals is evident.
-The quantile-quantile plot suggest that on the assumption of log-normal errors on the observed densities is reasonable, though the model appears to have some slight problems estimating the highest observed densities (Fig.S\@ref(fig:obs-v-pred-plot)).
-
-```{r}
-
-fit_report <- base_run$tmb_fit[[1]]$zissou_report
-
-observed_seen <- base_run$tmb_fit[[1]]$zissou_data$log_density
-
-observed_seeing <- base_run$tmb_fit[[1]]$zissou_data$any_seen
-
-seen_diagnostics <- data_frame(observed = observed_seen, predicted = as.numeric(fit_report$log_density_hat)) %>% 
-  mutate(residuals = predicted - observed)
-
-seeing_diagnostics <- data_frame(observed = observed_seeing, predicted = as.numeric(fit_report$prob_seeing))
-
-```
-
-```{r, eval = FALSE}
-
-a = base_run$tmb_fit[[1]]$zissou_data$x_seen_non_nested %>% 
-  as_data_frame() %>% 
-  select(contains("mean_")) %>% 
-  mutate(resid = seen_diagnostics$residuals) %>% 
-  gather(variable, value, -resid) %>% 
-  ggplot(aes(value, resid)) + 
-  geom_point() + 
-  facet_wrap(~variable)
-
-a
-
-```
-
-```{r obs-v-pred-plot, fig.cap="High level diagnostics for observed compontent of Delta-GLM: Observed vs predicted log densities (A), predicted log density vs residuals (B), and a normal qq-plot of the residuals (C)"}
-
-obs_v_pred_plot <- seen_diagnostics %>% 
-  ggplot(aes(observed, predicted)) + 
-  geom_point(alpha = 0.5) + 
-  geom_abline(aes(intercept = 0, slope = 1), linetype = 2, color = "red") + 
-  labs(caption = glue::glue("R^2 is {round(yardstick::rsq(data = seen_diagnostics, truth = observed, estimate = predicted)$.estimate,2)}"), title= "A") + 
-  theme(plot.margin = unit(rep(.1,4), units = "lines"))
-
-fit_v_resid_plot <- seen_diagnostics %>% 
-  ggplot(aes(predicted, residuals)) + 
-  geom_point() + 
-  geom_hline(aes(yintercept = 0), linetype = 2, color = "red") + 
-  labs(title = "B") +
-    theme(plot.margin = unit(rep(.1,4), units = "lines"))
-
-qq_plot <- ggplot(seen_diagnostics, aes(sample = residuals)) +
-  stat_qq() +
-  stat_qq_line() + 
-  labs(title = "C") + 
-    theme(plot.margin = unit(rep(.1,4), units = "lines"))
-
-obs_v_pred_plot/(fit_v_resid_plot + qq_plot)
-
-```
-
-In order to evaluate the ability of the model to estimate positive observations, we can compare the the binned predicted probability of a positive observation to the proportion of observations in that bin that recorded positive observations.
-If our model is working well, we would expect a group of fisheries that our model estimates on average should have a 50% probability of a positive observation, then we should expect about 50% of those observations to have positive observations.
-This is indeed what we see from the model (Fig.S\@ref(fig:obs-prob-plot)).
-
-```{r obs-prob-plot, fig.cap = "Binned mean predicted probability of detection provided by the first stage of the hurdle model vs observed proportion of positive detections"}
-
-seeing_diag_plot <- seeing_diagnostics %>% 
-  ungroup() %>% 
-  mutate(binned_prob = ntile(predicted, 20)) %>% 
-  arrange(binned_prob) %>% 
-  group_by(binned_prob) %>% 
-  summarise(mean_prob_seen = mean(predicted),
-            percent_observed = mean(observed))  %>% 
-  ggplot() +
-  geom_point(aes(mean_prob_seen, percent_observed)) + 
-  geom_abline(aes(intercept = 0, slope = 1), color = "red", linetype = 2) + 
-  labs(x = "Binned Predicted Probability of Detection", y = "Proportion of Positive Observations")
-
-seeing_diag_plot
-```
-
-We can also examine the receiver-operator-curve (ROC) to assess the performance of the observation component of the model.
-The area under the curve (AUC) value for the model is 0.84 (on a scale of 0.5 to 1), indicating the model is an overall good predictor of whether or not a given observation of biomass densities will be positive or not.
-
-```{r roc-plot, fig.cap = "Receiver operating characteristic curve of predictions of positive biomass densities"}
-library(plotROC)
-rocplot <- ggplot(seeing_diagnostics, aes(m = predicted, d = observed))+ geom_roc(n.cuts=20,labels=FALSE) + 
-  geom_abline(aes(slope = 1, intercept = 0))
-
-roc_plot <- rocplot + style_roc(theme = zissou_theme)
-
-roc_plot
-```
-
-### Standardized Abundance Indices
-
-Overall most species showed consistent trends in biomass densities across the different islands at which they have been observed (Fig.S\@ref(fig:reg-pop)).
-
-```{r reg-pop, fig.cap = "Mean density by island by year for each fish species included in the analysis"}
-
-reg_pop_plot <- fitted_data %>% 
-  group_by(region, year, classcode) %>% 
-  summarise(mean_density = mean(exp(log_density))) %>% 
-  mutate(log_mean_density = log(mean_density)) %>% 
-  ungroup() %>% 
-  group_by(classcode, region) %>% 
-  mutate(cs_mean_density = (mean_density - mean(mean_density)) / sd(mean_density)) %>% 
-  ungroup() %>% 
-  ggplot(aes(year, cs_mean_density, color = region)) + 
-  scale_color_npg(name="Region")+
-  geom_line() + 
-  labs(y="Centered and Scaled Density",x="Year")+
-  facet_wrap(~classcode, scales = "free_y") + 
-  theme_minimal()+
-  theme(axis.text=element_text(size=6))
-
-reg_pop_plot
-
-```
-
-The standardized indices of abundance generally did not very substantially from the raw mean densities by species over time.
-However, for some species, such as blue rockfish, the standardized abundance index suggests much higher biomass densities in the pre-MPA period than those reported in the raw data.
-We suspect this is largely a function of changes in sampling sites over time, that the standardization is better able to account for (Fig.S\@ref(fig:raw-v-stand-plot)).
-
-```{r raw-v-stand-plot, fig.cap = "Raw (points) and standardized (lines) indices of abundance for each of the fishes included in the analysis"}
-
-life_history <- fitted_data %>%
-  select(classcode) %>%
-  unique() %>%
-  mutate(numeric_classcode = as.numeric(as.factor(classcode)))
-
-raw <- fitted_data %>%
-  group_by(year, classcode, targeted) %>%
-  summarise(md = mean(exp(log_density)),
-            mt = mean(temp_deviation)) %>%
-  group_by(classcode) %>%
-  mutate(smd = (md - mean(md))/(sd(md))) %>%
-  ungroup() %>%
-  left_join(life_history, by = "classcode")
-
-years <- unique(fitted_data$year)
-
-years <- zissou_fit$did_data$year %>% unique()
-
-abundance_trends <- data_frame(abundance_hat = fit_report$abundance_hat,
-                               classcode = rep(1:n_distinct(fitted_data$classcode), each = length(years))) %>%
-  mutate(log_abundance_hat = log(abundance_hat)) %>%
-  group_by(classcode) %>%
-  mutate(year = years) %>%
-  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
-  ungroup() %>%
-  rename(numeric_classcode = classcode) %>%
-  left_join(life_history, by = "numeric_classcode" )
-
-# 
-# abundance_trends %>%
-#   ggplot(aes(year, log(abundance_hat), color = factor(classcode))) +
-#   geom_line(show.legend = F) +
-#   geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
-#   facet_wrap(~classcode) +
-#   theme_minimal()
-
-raw_v_std_plot <- abundance_trends %>%
-  filter(year %in% years) %>% 
-  ggplot(aes(year, scaled_abundance_hat, color = factor(classcode))) +
-  geom_line(show.legend = F) +
-  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
-  geom_point(data = raw,aes(year, smd, color = factor(classcode)), show.legend = F)+
-  facet_wrap(~classcode) +
-  labs(y="Raw (dots) and Standardized (lines) Abundance",x="Year")+
-  theme_minimal()+
-  theme(axis.text=element_text(size=6))
-
-raw_v_std_plot
-
-```
-
-We include a variety of environmental, observation, and temporal indicators in our model.
-Inclusion of highly co-linear variables in a model can inflate standard errors and obscure "true" effects.
-To account for this we calculated the Pearson's correlation coefficients for all of the continuous data included in our model to ensure that none of the included variables had correlation coefficients greater than 0.7, a general rule of thumb for co-inclusion of variables.
-We did not find problematic levels of correlation among any of our included continuous variables.
-
-```{r numeric-cor, fig.cap="Pearson correlation coefficients of continuos data included in the regression model"}
-
-numeric_cor_plot <- base_run$data[[1]] %>%
-  select_if(is.numeric) %>%
-  select(-log_density, -targeted) %>%
-  na.omit() %>%
-  corrr::correlate() %>%
-  gather("colname", "correlation", -rowname) %>%
-  ggplot(aes(rowname, colname, fill = correlation)) +
-  geom_tile() +
-  geom_text(aes(rowname, colname, label = round(correlation, 2))) +
-  scale_fill_gradient2(
-    low = "tomato",
-    mid = "white",
-    high = "steelblue",
-    midpoint = 0,
-    limits = c(-1, 1)
-  ) +
-  theme_minimal() +
-  theme(
-    axis.text.x = element_text(
-      angle = 45,
-      vjust = 1,
-      hjust = 1,
-    size = 6),
-    axis.text.y = element_text(size = 6),
-    axis.title.x = element_blank(),
-    axis.title.y = element_blank()
-  )
-
-numeric_cor_plot
-```
-
-## Alternative estimation models
-
-Our proposed identification strategy attempts to control for non-MPA (and not directly modeled) related changes in abundances through the trend in the non-targeted species.
-However, a simpler alternative would be to simply compare densities before-and-after MPA implementation, while explicitly controlling for non-MPA related factors that we believe may have some effect on densities (a "selection on observables" strategy).
-To that end, we fit a mixed-effects regression that models log densities of targeted species only (positive observations only, for the sake of simplicity) as a function of temperature deviations, kelp cover, observer experience, random effects for species and region, and fixed effects for each year in the data (omitting the year 2000).
-The hypothesis here is that any non-MPA related factors that affect densities are accounted for in the observed variables included in the model.
-
-Using this model, densities of targeted species appear to have been declining steadily since 2000, and appear to have plateaued off since the implementation of MPAs in 2003.
-Without an identification strategy such as the one employed in this study then, all we could conclude is that densities appear to be lower post-MPA, and have not increased substantially over time (Fig.S\@ref(fig:obs-plot)).
-
-```{r obs-plot, fig.cap = "Selection on observables identification strategy.
-Plotted estimates are fixed effects of year on log-density (relative to the year 2000), controlling for observer experience, temperature deviations, and kelp cover, with random effects for species and region"}
-
-test <- lme4::lmer(log_density~ factor_year + cumulative_n_obs  + temp_deviation + interp_kelp + (1|classcode) + (1|region), data = fitted_data %>% filter(targeted == 1, any_seen == T))
-
-sel_on_obs_plot <- broom::tidy(test) %>% 
-  filter(str_detect(term, "factor_year")) %>% 
-  mutate(year = str_replace_all(term,"\\D","") %>% as.numeric()) %>% 
-  ggplot() + 
-  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
-  geom_pointrange(aes(year, estimate, ymin = estimate - 1.96*std.error, ymax = estimate + 1.96*std.error)) + 
-  labs(y = "Estimated Difference", x = "Year")
-
-sel_on_obs_plot
-```
-
-```{r, include=F}
-raw <- fitted_data %>%
-  group_by(year, classcode, region, targeted) %>%
-  summarise(md = mean(exp(log_density)),
-            mt = mean(temp_deviation)) %>%
-  group_by(classcode) %>%
-  mutate(smd = (md - mean(md))/(sd(md))) %>%
-  ungroup() %>%
-  left_join(life_history, by = "classcode")
-
-raw_region <- fitted_data %>%
-  group_by(year, region,classcode, targeted) %>%
-  summarise(md = mean(exp(log_density)),
-            mt = mean(temp_deviation)) %>%
-  group_by(classcode) %>%
-  mutate(smd = (md - mean(md))/(sd(md))) %>%
-  ungroup() %>%
-  left_join(life_history, by = "classcode")
-
-raw_mpa_only <- fitted_data %>%
-  filter(eventual_mpa == T) %>% 
-  group_by(year, classcode, targeted) %>%
-  summarise(md = mean(exp(log_density)),
-            mt = mean(temp_deviation)) %>%
-  group_by(classcode) %>%
-  mutate(smd = (md - mean(md))/(sd(md))) %>%
-  ungroup() %>%
-  left_join(life_history, by = "classcode")
-
-raw_nompa_only <- fitted_data %>%
-  filter(eventual_mpa == F) %>% 
-  group_by(year, classcode, targeted) %>%
-  summarise(md = mean(exp(log_density)),
-            mt = mean(temp_deviation)) %>%
-  group_by(classcode) %>%
-  mutate(smd = (md - mean(md))/(sd(md))) %>%
-  ungroup() %>%
-  left_join(life_history, by = "classcode")
-
-# basic_model <- rstanarm::stan_glm(log(md + 1e-3) ~ factor(year) + targeted + factor(year):targeted, data = raw)
-
-```
-
-The estimation model used for our main results is complicated.
-We feel this complexity is justified in order to best capture the uncertainty inherent in the challenging task of conducting underwater visual surveys, as well as the spatio-temporal nature of the underlying data.
-However, we also ran several simpler models in order to examine the sensitivity of our results to the model structure selected here.
-
-Much of the complexity of our model comes from the integration of a standardized abundance index for each of the species, which are then compiled into a standardized index of abundance index for targeted species as a whole.
-
-In one simpler approach, we aggregated the PISCO data into mean biomass densities by species and year across all the Channel Islands.
-Since all included species have at least some positive observations in each year, this removes the complication of dealing with the probability of detection problem in the raw data, while of course assuming that the net effects of all of the other factors included in our main model are on average zero (e.g.
-observer skill, visibility, etc.).
-A difference-in-difference model fit to these simplified data show qualitatively similar (though much more uncertain) results to our full model.
-In particular, the model shows the same gradual increase in targeted densities relative to non-targeted until 2013, followed by an attenuation of this trend.
-Importantly, though, this simplified model fails to correct for some pre-MPA differences in the two groups, leading to negative "effects" of the MPAs being estimated before the MPAs themselves went in place.
-While these sort of anticipatory effects are certainly possible [@mcdermott2019], in this case we would suggest the likelier explanation---given that these anticipatory effects disappear in the full model---is that our full model provides important controls for pre-MPA characteristics (Fig.S\@ref(fig:sortasimple-did))
-
-```{r, results = "hide"}
-
-year_mpa <- 2002
-
-  data <- abundance_data$data[abundance_data$data_source == "pisco"][[1]] %>%
-    mutate(density = exp(log_density) * any_seen) %>%
-    mutate(targ = case_when(targeted == 1 ~ "Targeted",  TRUE ~ "Non-Targeted"))
-
-  sortasimple <- data %>%
-    group_by(year, classcode) %>%
-    summarise(mean_density = mean(density),
-              targ = unique(targ)) %>%
-    mutate(mpa = year > year_mpa) %>%
-    ungroup() %>%
-    mutate(fyear = factor(year)) %>%
-    mutate(fyear = relevel(fyear, ref = "2003"))
-
-  sortasimple_reg <-
-    rstanarm::stan_glmer(
-      log(mean_density + 1e-3) ~ targ * fyear + (1 |
-                                                   classcode),
-      data = sortasimple,
-      cores = 4
-    )
-
-  sortasimple_plot <- bayesplot::mcmc_areas(as.matrix(sortasimple_reg),
-                                            regex_pars = ":fyear") +
-    geom_vline(aes(xintercept = 0), linetype = 2, color = "red") +
-    scale_y_discrete(labels = c(2000:2001, 2003:2017)) +
-    coord_flip() +
-    scale_x_percent() + 
-    zissou_theme + 
-    labs(x="Estimated Difference",y="Year")+
-    theme(axis.text.x = element_text(size = 8))
-
-```
-```{r sortasimple-did, fig.cap = "Results of simplified difference-in-difference regression.
-The model estimates the difference in the mean trend of densities of targeted and non-targeted species over time, controlling for the mean densities of each individual species group.
-Data are initially aggregated to the level of species-by-year."}
-sortasimple_plot
-```
-
-```{r species-by-species-did, results = "hide"}
-
-year_mpa <- 2002
-
-  data <- abundance_data$data[abundance_data$data_source == "pisco"][[1]] %>%
-    mutate(density = exp(log_density) * any_seen) %>%
-    mutate(targ = case_when(targeted == 1 ~ "Targeted",  TRUE ~ "Non-Targeted"))
-
-  sortasimple <- data %>%
-    group_by(year, classcode) %>%
-    summarise(mean_density = mean(density),
-              targ = unique(targ)) %>%
-    mutate(mpa = year > year_mpa) %>%
-    ungroup() %>%
-    mutate(fyear = factor(year)) %>%
-    mutate(fyear = relevel(fyear, ref = "2003"))
-
-  sortasimple_reg <-
-    rstanarm::stan_glmer(
-      log(mean_density + 1e-3) ~ targ * fyear + (1 |
-                                                   classcode),
-      data = sortasimple,
-      cores = 4
-    )
-
-  sortasimple_plot <- bayesplot::mcmc_areas(as.matrix(sortasimple_reg),
-                                            regex_pars = ":fyear") +
-    geom_vline(aes(xintercept = 0), linetype = 2, color = "red") +
-    scale_y_discrete(labels = c(2000:2001, 2003:2017)) +
-    coord_flip() +
-    scale_x_percent() + 
-    zissou_theme + 
-    labs(x="Estimated Difference",y="Year")+
-    theme(axis.text.x = element_text(size = 8))
-
-```
-
-As an even simpler approach, we aggregated the data to the level of targeted and non-targeted species, and estimated the divergence in their trends over time.
-
-```{r, results = "hide"}
-
-  data <- abundance_data$data[abundance_data$data_source == "pisco"][[1]] %>%
-    mutate(density = exp(log_density) * any_seen) %>%
-    mutate(targ = case_when(targeted == 1 ~ "Targeted",  TRUE ~ "Non-Targeted"))
-
-  supersimple <- data %>%
-    group_by(year, targ) %>%
-    summarise(mean_density = mean(density)) %>% 
-    mutate(mpa = year > year_mpa) %>%
-    ungroup() %>%
-    mutate(fyear = factor(year)) %>%
-    mutate(fyear = relevel(fyear, ref = "2002"))
-
-  supersimple_reg <-
-    rstanarm::stan_glm(
-      log(mean_density) ~ targ + fyear + targ:fyear,
-      data = supersimple,
-      cores = 4,
-      iter = 2000,
-      adapt_delta = 0.95
-    )
-
-    # supersimple_reg <-
-    # rstanarm::stan_glm(
-    #   log(mean_density + 1e-3) ~ targ * fyear,
-    #   data = supersimple,
-    #   cores = 4,
-    #   iter = 10000
-    # )
-
-  supersimple_plot <- bayesplot::mcmc_areas(as.matrix(supersimple_reg),
-                                            regex_pars = ":fyear") +
-    geom_vline(aes(xintercept = 0), linetype = 2, color = "red") +
-    scale_y_discrete(labels = c(2000:2001, 2003:2017)) +
-    coord_flip() +
-    scale_x_percent() + 
-    zissou_theme + 
-    labs(y="Year",x="Estimated Difference")+
-    theme(axis.text.x = element_text(size = 8))
-
-```
-
-```{r}
-supersimple_plot
-```
-
-We can also explore the effects of our species filtering by running a simplified regression but now including all the species in the database, no matter of how infrequently they are observed.
-
-```{r, results= "hide"}
-year_mpa <- 2002
-
-  sortasimple <- pisco_data %>%
-    group_by(year, classcode) %>%
-    summarise(mean_density = mean(density_g_m2, na.rm = TRUE),
-              targ = unique(targeted)) %>%
-    mutate(mpa = year > year_mpa) %>%
-    ungroup() %>%
-    mutate(fyear = factor(year)) %>%
-    mutate(fyear = relevel(fyear, ref = "2003"))
-
-  sortasimple_reg <-
-    rstanarm::stan_glmer(
-      log(mean_density + 1e-3) ~ targ * fyear + (1 |
-                                                   classcode),
-      data = sortasimple,
-      cores = 4
-    )
-
-  all_species_sortasimple_plot <- bayesplot::mcmc_areas(as.matrix(sortasimple_reg),
-                                            regex_pars = ":fyear") +
-    geom_vline(aes(xintercept = 0), linetype = 2, color = "red") +
-    scale_y_discrete(labels = c(1999:2001, 2003:2017)) +
-    coord_flip() +
-    scale_x_percent() + 
-    zissou_theme + 
-    labs(x="Estimated Difference",y="Year")+
-    theme(axis.text.x = element_text(size = 8))
-```
-
-```{r sortasimple-noilter-did, fig.cap = "Results of simplified difference-in-difference regression including all observed finfish.
-The model estimates the difference in the mean trend of densities of targeted and non-targeted species over time, controlling for the mean densities of each individual species group.
-Data are initially aggregated to the level of species-by-year."}
-all_species_sortasimple_plot
-```
-
-### Synthetic controls
-
-Synthetic controls are an alternative method for attempting to estimate the causal effect of a policy intervention [@abadie2010].
-A difference-in-difference approach assumes that some observable group serves as an adequate control for the state of the treated group in an untreated world.
-In our default case, we assume that the mean standardized index of non-targeted species are our control for the targeted species.
-Alternatively, synthetic controls use timeseries of treated and non-treated groups before and after treatment to construct a new "control" group built by weighting the pre-treatment timeseries of un-treated observations (together with covariates) such that the synthetic control group matches the trends in the treated group pre-treatment.
-
-We chose to present difference-in-difference as our main result since it better allows us to capture the uncertainty in the data generating process through our hierarchical model.
-However, we felt that it was worth exploring whether synthetic controls provided substantially different results than our default model.
-
-For the first synthetic control, we pulled our standardized mean index of abundance for targeted species as a whole from our difference-in-difference model as our treated group.
-We then pull the standardized indices of abundance for each of the non-targeted groups from the difference-in-difference to use as the candidate untreated components for the synthetic controls.
-A complete synthetic control analysis would require more extensive validation of the methods, but we use this approach simply to explore whether we observe substantially divergent results in the synthetic control versus the difference-in-difference model.
-
-We centered and scaled the candidate abundance indices to facilitate model convergence given the very few number of pre-treatment years available.
-The results of a synthetic control model are presented as the difference between the observed treatment outcome and the synthetic control (the difference in this case being in units of standard deviations).
-The model was not able to construct an adequate synthetic control at this level, as shown by the differences between the treated group and the synthetic control pre-treatment.
-However, we would note that the post-treatment results do show similarities with our main results, namely a lack of a clear divergence between the treatment and the control, and an upwards trend up through the earl 2010s followed by a decline (Fig.S\@ref(fig:total-synth)).
-
-```{r, include = FALSE}
-
-nontargeted_index <- data_frame(abundance_hat = fit_report$abundance_hat,
-                               classcode = rep(1:n_distinct(fitted_data$classcode), each = length(years))) %>%
-  mutate(log_abundance_hat = log(abundance_hat)) %>%
-  group_by(classcode) %>%
-  mutate(year =years) %>%
-  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
-  ungroup() %>%
-  rename(numeric_classcode = classcode) %>%
-  left_join(life_history, by = "numeric_classcode" ) %>% 
-  left_join(unique(abundance_data$data[[1]] %>% select(classcode, targeted)), by = "classcode") %>% 
-  filter(targeted == 0) %>% 
-  select(log_abundance_hat, year, classcode, targeted)
-
-targeted_index <- zissou_fit$zissou_estimates %>%
-  filter(
-    stringr::str_detect(variable, "(targeted_did_betas)"),!str_detect(variable, "non")
-  ) %>%
-  mutate(
-    year = years,
-    classcode = "targeted",
-    targeted = 1
-  ) %>%
-  rename(log_abundance_hat = estimate) %>% 
-  select(colnames(nontargeted_index))
-
-  # filter(variable == "seeing_year_species_betas") %>%
-
-nontargeted_covariates <- pisco_data %>% 
-  select(classcode, contains("mean"),year, targeted) %>% 
-  filter(targeted == 0) %>% 
-  select(-mean_length) %>% 
-  gather(variable, value, -classcode,-year,-targeted) %>% 
-  group_by(classcode, variable, year) %>% 
-  summarise(mean_value = mean(value, na.rm = TRUE)) %>% 
-  spread(variable, mean_value)
-
-targeted_covariates <- pisco_data %>% 
-  select(classcode, contains("mean"),year, targeted) %>% 
-  filter(targeted == 1) %>% 
-  mutate(classcode = "targeted") %>% 
-  select(-mean_length) %>% 
-  gather(variable, value, -classcode,-year,-targeted) %>% 
-  group_by(classcode, variable, year) %>% 
-  summarise(mean_value = mean(value, na.rm = TRUE)) %>% 
-  spread(variable, mean_value)
-
-synth_covariates <- nontargeted_covariates %>% 
-  bind_rows(targeted_covariates)
-
-synth_data <- targeted_index %>% 
-  bind_rows(nontargeted_index) %>% 
-  left_join(synth_covariates, by = c("classcode","year")) %>% 
-    mutate(numeric_classcode = as.numeric(factor(classcode))) %>% 
-  group_by(classcode) %>% 
-  mutate(log_abundance_hat = scale(exp(log_abundance_hat))) %>% 
-  ungroup() %>% 
-  as.data.frame()
-
-# synth_data %>%
-#   group_by(classcode) %>%
-#   mutate(index = (log_abundance_hat)) %>%
-#   ggplot(aes(
-#     year,
-#     index,
-#     color = factor(targeted),
-#     group = interaction(targeted, classcode)
-#   )) +
-#   geom_line()
-
-prepped_synth_data <-
-  Synth::dataprep(
-    foo = synth_data,
-    predictors = colnames(synth_data)[str_detect(colnames(synth_data),"mean")],
-    predictors.op = "mean",
-    dependent = "log_abundance_hat",
-    unit.variable = "numeric_classcode",
-    unit.names.variable = "classcode",
-    time.variable = "year",
-    treatment.identifier = "targeted",
-    controls.identifier = unique(synth_data$classcode)[unique(synth_data$classcode) != "targeted"],
-    time.predictors.prior = years[years <= 2003],
-    time.optimize.ssr = years[years <= 2003],
-    time.plot = years
-  )
-
-synth_fit <- Synth::synth(prepped_synth_data)
-
-gaps <-
-  rownames_to_column(
-    prepped_synth_data$Y1plot - (prepped_synth_data$Y0plot %*% synth_fit$solution.w) %>%
-      as.data.frame()
-  ) %>%
-  rename(gap = w.weight,
-         year = rowname) %>%
-  mutate(year = as.numeric(year)) %>%
-  mutate(synthetic = as.numeric(prepped_synth_data$Y0plot %*% synth_fit$solution.w)) %>% 
-  mutate(treatment = gap + synthetic)
-
-# path.plot(synth.res = synth_fit,
-#           dataprep.res = prepped_synth_data
-# )
-
-total_gap_plot <- gaps %>% 
-  ggplot(aes(year, gap)) + 
-  geom_hline(aes(yintercept = 0), linetype = 2) +
-  geom_vline(aes(xintercept = 2003), color = "red")+
-  geom_line() + 
-    labs(x = "Year", y = "Standard deviations from synthetic control")
-
-total_synth_plot <- gaps %>% 
-  select(year, synthetic, treatment) %>% 
-  gather(source, value, -year) %>% 
-  group_by(source) %>% 
-  # mutate(value = scale(value)) %>% 
-  ggplot(aes(year, value, color = source, linetype = source)) + 
-  geom_vline(aes(xintercept = 2003), color = "red")+
-  geom_line() + 
-    labs(x = "Year", y = "Standard deviations from synthetic control")
-
-```
-
-```{r  total-synth, fig.cap = "Difference in centered and scaled standardized targeted abundance and synthetic standardized targeted abundance"}
-total_gap_plot
-```
-
-As an extension, we repeated this process, but now treating each targeted species individually as the treated group, and the non-targeted species as the non-targeted.
-This is intended to explore whether we see clearer signals for individual species than we do for the targeted class as a whole.
-
-Overall we see similarly unclear results as the aggregate targeted synthetic control (and our main results).
-The synthetic control was better constructed for some individual species, but not clearly for any one, and most species showed some evidence of the upward-then-downward trend seen throughout our results (Fig.S\@ref(fig:synth-classcode-fits)).
-
-```{r, include = FALSE}
-
-synth_foo <-
-  function(treated_classcode,
-           synth_data,
-           targeted_classcodes) {
-    # treated_classcode <- targeted_classcodes[1]
-
-    temp_synth_data <- synth_data %>%
-      filter(classcode == treated_classcode |
-               !(classcode %in% targeted_classcodes))
-
-    prepped_synth_data <-
-      Synth::dataprep(
-        foo = temp_synth_data,
-        predictors = colnames(temp_synth_data)[str_detect(colnames(temp_synth_data), "mean")],
-        predictors.op = "mean",
-        dependent = "log_abundance_hat",
-        unit.variable = "numeric_classcode",
-        unit.names.variable = "classcode",
-        time.variable = "year",
-        treatment.identifier = treated_classcode,
-        controls.identifier = unique(temp_synth_data$classcode)[unique(temp_synth_data$classcode) != treated_classcode],
-        time.predictors.prior = c(2000:2003),
-        time.optimize.ssr = c(2000:2003),
-        time.plot = years
-      )
-
-    synth_fit <- Synth::synth(prepped_synth_data)
-
-    gaps <-
-      rownames_to_column(
-        prepped_synth_data$Y1plot - (prepped_synth_data$Y0plot %*% synth_fit$solution.w) %>%
-          as.data.frame()
-      ) %>%
-      rename(gap = w.weight,
-             year = rowname) %>%
-      mutate(year = as.numeric(year)) %>%
-      mutate(synthetic = as.numeric(prepped_synth_data$Y0plot %*% synth_fit$solution.w)) %>%
-      mutate(treatment = gap + synthetic)
-  }
-
-nontargeted_index <- data_frame(abundance_hat = fit_report$abundance_hat,
-                               classcode = rep(1:n_distinct(fitted_data$classcode), each = length(years))) %>%
-  mutate(log_abundance_hat = log(abundance_hat)) %>%
-  group_by(classcode) %>%
-  mutate(year = 1999 + 1:length(abundance_hat)) %>%
-  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
-  ungroup() %>%
-  rename(numeric_classcode = classcode) %>%
-  left_join(life_history, by = "numeric_classcode" ) %>% 
-  left_join(unique(abundance_data$data[[1]] %>% select(classcode, targeted)), by = "classcode") %>% 
-  select(log_abundance_hat, year, classcode, targeted)
-
-synth_covariates <- pisco_data %>% 
-  select(classcode, contains("mean"),year, targeted) %>% 
-  select(-mean_length) %>% 
-  gather(variable, value, -classcode,-year,-targeted) %>% 
-  group_by(classcode, variable, year) %>% 
-  summarise(mean_value = mean(value, na.rm = TRUE)) %>% 
-  spread(variable, mean_value)
-
-synth_data <- nontargeted_index %>% 
-  left_join(synth_covariates, by = c("classcode","year")) %>% 
-    mutate(numeric_classcode = as.numeric(factor(classcode))) %>% 
-  group_by(classcode) %>% 
-  mutate(log_abundance_hat = scale(exp(log_abundance_hat))) %>% 
-  ungroup() %>% 
-  as.data.frame()
-
-targeted_classcodes <- unique(synth_data$classcode[synth_data$targeted == 1])
-
-synth_classcode_fits <-
-  tibble(targeted_classcodes = targeted_classcodes) %>%
-  mutate(
-    synthetic_fit = map(
-      targeted_classcodes,
-      synth_foo,
-      synth_data = synth_data,
-      targeted_classcodes = targeted_classcodes
-    )
-  )
-
-synth_classcode_fits <- synth_classcode_fits %>% 
-  unnest()
-
-synth_classcode_fits_plot <- 
-synth_classcode_fits %>% 
-  ggplot(aes(year, gap)) + 
-  geom_hline(aes(yintercept = 0), linetype = 2) +
-  geom_vline(aes(xintercept = 2003), color = "red")+
-  geom_line() + 
-  facet_wrap(~targeted_classcodes) + 
-  labs(x = "Year", y = "Standard deviations from synthetic control")
-
-```
-
-```{r synth-classcode-fits, fig.cap="Synthetic control gaps for each targeted species"}
-synth_classcode_fits_plot + 
-  theme_minimal()
-```
-
-## Testing Model Assumptions
-
-### Simulation testing
-
-We state that a difference-in-difference model using targeted and non-targeted species is capable (conditional on assumptions) of estimating the causal effect of MPAs.
-We simulated MPA outcomes to test this claim.
-We first test our estimation strategy under idealized circumstances, where recruitment is deterministic and PISCO divers all have constant and perfect observer skills.
-We simulate five species that vary only in their maximum size and length at maturity.
-For each of these species, we set one version that is targeted by fishing and one that is not.
-We set a constant fishing mortality rate for each simulated targeted species, and then ran two matched simulations, one with MPAs and one without.
-We then have our simulated divers sample data from each of these scenarios, and then pass the sampled biomass densities to a simplified version of our difference-in-difference model (omitting the probability of detection step).
-We can then compare the difference-in-difference estimates of the MPA effect to the true simulated effect.
-The difference in difference model is able to capture the simulated MPA effect under these circumstances (Fig.S\@ref(fig:simple-did-test))
-
-```{r simple-did-test, fig.cap = "Simulated mean (red dashed line) and individual species (solid lines) MPA effects over time, along with difference-in-difference estimated MPA effects (mean with 95% confidence intervals)"}
-
-load(file = file.path(run_dir,'simulated_did.Rdata'))
-
-simple_performance$mixed_effect_did + 
-  labs(title = '', x = "Years with MPA protection",
-       y = "MPA Effect")
-```
-
-We then simulated a more complex example.
-We use the actual targeted and non-targeted species from our model.
-We assign species predominately seen in the western Channel Islands as "cold water" and those in the eastern Channel Islands as "warm water".
-We allow for stochasticity in recruitment.
-We use El Niño data as a simulated environmental recruitment driver, where we assume that El Niño events produce negative recruitment shocks for cold water species and *vice versa* for warm water species.
-We simulate three different divers each with different base skill levels, visual selectivities, and an evolving skill rate (such that observers get better over time).
-We hold fishing mortality rates constant for each species, although that fishing mortality affects each species differently because of intrinsic biological differences in maturity-at-age and steepness.
-We then test the ability of the difference-in-difference model to isolate the mean MPA effect across all of these targeted species, which our results show it is capable of doing (Fig.S\@ref(fig:complex-did-test)).
-
-```{r complex-did-test,  fig.cap = "Simulated mean (red dashed line) and individual species (solid lines) MPA effects over time, along with difference-in-difference estimated MPA effects (mean with 95% confidence intervals)"}
-
-pisco_performance$mixed_effect_did + 
-    labs(title = '', x = "Years with MPA protection",
-       y = "MPA Effect")
-
-```
-
-### Sensitivity to "missing" observations
-
-Appropriately addressing the problem of "missing" observations is a critical challenge in any field observation study.
-If no observations of a given fish species were recorded on a given transect, should the density of that species on that transect be marked as zero, and influence the estimate of the overall mean density accordingly?
-The obvious answer seems to be yes, but what if that species simply does not live in the environment covered by a particular transect, or was not present during the particular time of the diver's observation?
-For our base runs, we assign a value of zero density on a given transect for any fish species that has been observed at least once at a given site at any time in our data but was not observed on that particular transect.
-If that species was never observed at that site, we do not include a zero for that species.
-Our rationale for this is that given the shifting nature of the sampled sites, and the intensity of sampling at those sites, we do not want to skew density trends by changes in the amount of suitable habitat for a given species sampled.
-However, this is clearly a strong assumption.
-For example, perhaps the decreasing trend in mean densities from 2000 to 2004 is due to increased number of sites (and therefore zeros) included in the data.
-To assess the potential importance of this choice, we can compare the mean densities of targeted and non-targeted species over time with the added zeros to the mean densities using only positive observations (i.e.
-not including any zeros in the data, (Fig.S\@ref(fig:nozero-raw-trend)).
-The trends in the raw densities, and most importantly the mean trends of targeted and non-targeted fishes, are nearly identical whether or not zeros are added, providing strong evidence that our choice of how to incorporate missing observations into the data are not strongly influencing our overall results.
-
-```{r nozero-raw-trend, fig.cap = "Centered and scaled mean annual density, excluding zeros, of included fishes (points) and smoothed means of targeted and non-targeted groups (line) over time" }
-
-nozero_trend_plot <- base_run$data[[1]] %>% 
-  filter(any_seen == T) %>% 
-  select(year, targeted,classcode, log_density, any_seen) %>% 
-  mutate(density = exp(log_density) * any_seen) %>% 
-  group_by(classcode,targeted, year) %>% 
-  summarise(md = mean(density)) %>% 
-  group_by(classcode,targeted) %>% 
-  mutate(smd = (md - mean(md)) / sd(md)) %>% 
-  ggplot(aes(year, smd, color = targeted == 1, fill = targeted == 1)) + 
-  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
-  geom_line(aes(group = interaction(targeted, classcode)), alpha = 0.25) +
-  geom_smooth() +
-  scale_color_d3(name = "Targeted?") + 
-  scale_fill_d3(name = "Targeted?") +
-  labs(y = "Centered and Scaled Mean Density", x = "Year")
-
-nozero_trend_plot
-```
-
-### Testing SUTVA with Convergent Cross Mapping
-
-The difference-in-difference model also assumes that the targeted and non-targeted fishes do not directly or indirectly affect each other.This assumption is clearly violated on some level: all the fishes in this analysis are part of the same ecosystem and therefore interact to some degree.
-For example, if the protection of targeted predatory fishes results in increased mortality of non-targeted fishes, the model would attribute that as an increased regional effect (greater divergence between the abundance of targeted and non-targeted species).
-Given the time scale of analysis (15 years of protection), we do not feel that massive trophic cascades are likely to have developed yet, given both the pace and complexity of trophic cascade development [@babcock2010; @pershing2015a].
-A complete assessment of evidence for trophic cascades in the Channel Islands is beyond the scope of this study, but to address this question somewhat we utilized convergent cross mapping *sensu* @sugihara2012 to test for a significant causal signal between different broad trophic groups in the data, implemented in the `rEDM` package in R.
-
-Convergent cross mapping is a nonlinear forecasting method that uses observed time series data to test for significant causal links between variables.
-Following methods laid out in @clark2015 and @sugihara2012, we pool the abundance of each broad trophic group by region (Fig.S\@ref(fig:trophic-plot).
-This uses the data from the islands as "replicates", requiring the assumption that the islands are all part of the same dynamic system, but allowing us to take advantage of the extra information provided by each island to further resolve the reconstructed manifolds.
-Using these aggregations, we then test whether the variables can be properly embedded, i.e., if they have predictable manifold dynamics.
-We do this through a simplex forecasting test, using an individual timeseries' own lags to build a manifold.
-For each timeseries, the "best embedding dimension" is an approximation of the dimensionality of the dynamic system, in other words, the number of dimensions that define and predict the evolving states of the timeseries.
-This analysis shows that only the carnivore, piscivore, and planktivores groups show evidence of significant predictability (that is, that past dynamics of these species groups can predict future dynamics, Fig.S\@ref(fig:embed-plot)).
-
-```{r trophic-plot, fig.cap="Centered and scaled densities by broad trophic group and island over time"}
-
-trophic_groups <- cip_data %>% 
-  select(classcode, broadtrophic) %>% 
-  unique()
-
-# dat <- abundance_trends %>%
-#   left_join(trophic_groups, by = "classcode") %>%
-#   group_by(broadtrophic, year) %>%
-#   summarise(density = mean(scaled_abundance_hat)) %>%
-#   group_by(broadtrophic) %>%
-#   mutate(density = scale(density)) %>%
-#   ungroup() %>% 
-#   spread(broadtrophic, density) %>% 
-#   arrange(year) %>% 
-#   mutate(lag4_carnivore = lag(carnivore,4),
-#          lag4_piscovore = lag(piscivore,4)) %>% 
-#   na.omit() %>% 
-#   ungroup() %>% 
-#   filter(year > 1999) 
-# dat %>%
-#   ggplot(aes(year, density, color = broadtrophic)) +
-#   geom_line()
-
-dat <- cip_data %>%
-  group_by(broadtrophic, region, site, side, transect, year, eventual_mpa) %>%
-  summarise(density = sum(density_g_m2)) %>%
-  group_by(broadtrophic, region, site, year) %>%
-  summarise(mean_density = mean(density, na.rm = T)) %>%
-  group_by(broadtrophic,region, year) %>%
-  summarise(regional_density = mean(mean_density, na.rm = T)) %>% 
-  ungroup() %>% 
-  spread(broadtrophic, regional_density) %>% 
-  group_by(region) %>% 
-  arrange(year) %>% 
-  mutate(lag4_carnivore = lag(carnivore,4),
-         lag4_piscovore = lag(piscivore,4)) %>% 
-  na.omit() %>% 
-  ungroup() %>% 
-  filter(year > 1999)
-
- dat <- dat %>%
-  select(-lag4_carnivore,-lag4_piscovore)%>%
-  group_by(region)%>%
-  mutate_at(vars(carnivore:planktivore),funs(norm=(.-mean(.))/sd(.)))->dat
-
-group_trends_plot <- dat %>%
-  select(region,year,contains("norm"))%>%
-  gather("group","density",-region,-year)%>%
-  ggplot(aes(year,density,col=group))+
-  geom_line()+
-  geom_point() +
-  scale_color_locuszoom(name="Trophic Group",labels=c("carnivore","herbivore","piscivore","planktivore"))+
-  labs(x="year",y="Centered and Scaled Density")+
-  facet_wrap(~region)
-
-group_trends_plot
-```
-
-```{r embed-plot, fig.cap = "Predictive skill as a function of embedding dimensions"}
-
-## split data by region to analyze timeseries from each island separately
-ana.dat <- dat %>% filter(region=="ANA")
-sci.dat <- dat %>% filter(region=="SCI")
-smi.dat <- dat %>% filter(region=="SMI")
-sri.dat <- dat %>% filter(region=="SRI")
-
-datnest <- dat %>% group_by(region) %>% nest()
-
-datnorm <- dat %>% select(region,year,contains('norm')) %>% ungroup()
-
-# have to record the segments corresponding to each "replicate" so simplex algorithm does not try to make predictions crossing time barriers
-segs <- datnorm %>% mutate(ind=row_number()) %>% group_by(region) %>% summarise(first=first(ind),last=last(ind)) %>%
-  select(-region)
-
-var_names <- c("carnivore_norm","herbivore_norm","piscivore_norm","planktivore_norm")
-
-regions.combined.simp.list <- map(var_names,function(x){
-  temp <- datnorm %>% ungroup() %>% select(matches(x)) %>% as.data.frame()
-  out <- simplex(as.numeric(temp[,1]),E=1:10,lib=as.matrix(segs),silent=T) %>%
-    mutate(trophic=str_replace(x,"_norm",""))
-  out
-})
-
-embed_plot <- bind_rows(regions.combined.simp.list) %>%
-  ggplot(aes(E,rho,color=trophic))+
-  geom_line(size=2)+
-  facet_wrap(~trophic,nrow=2,scales="free_y")+
-  geom_hline(yintercept = 0,color="black")+
-  labs(x="Embedding Dimension (E)",y=expression(paste("Predictive Skill, ",rho)))+
-  scale_x_continuous(breaks=seq(0,12,by=2))+
-  scale_color_locuszoom()+
-  guides(color=F)
-
-embed_plot
-
-```
-
-Focusing on just these three groups then (removing herbivores), we can test for causal relationships between groups using convergent cross mapping and the logic of Takens' theorem of dynamic systems.
-Generalizations of Takens' theorem indicate that if two variables (in our case, species or physical variables) are part of the same dynamic system, their individual dynamics should reflect their relative causal influence [@sugihara2012].
-In other words, if one variable is causally forced by another, that forcing should leave a signature on the first time series.
-Convergent cross mapping (CCM) tests for causation by using the attractor/manifold built from the time series of one variable to predict another (hence the "cross-mapping").
-In simple terms, the *causal effect of A on B is determined by how well B cross-maps A*.
-
-There are two criteria for CCM to establish causality:  First, and most obviously, predictive cross-map skill using all available data should be significantly greater than zero.
-Second, that predictability should be convergent.
-Convergence means that cross-mapped estimates improve with library length (the number of state-space vectors used to build the attractor), because the attractor is more fully resolved and therefore estimation error should decline.
-Convergence is key to distinguishing causation from simple or spurious correlation.
-If two variables are spuriously correlated and not causally linked, CCM should fail to satisfy this second criterion.
-Based on these criteria, there is little evidence of significant dynamic interactions between trophic groups (Fig.S\@ref(fig:cross-map-cp)-\@ref(fig:cross-map-pis)).
-Cross-mapping produced positive predictive skill, but was non-convergent for all cross-mappings with the exception of carnivores cross-mapping planktivores (providing some evidence that planktivore dynamics may be a driver of carnivore dynamics).
-This analysis provides evidence that trophic cascades are unlikely to be a significant driver of our results.
-It is important to note though that this analysis does not mean that trophic cascades could not emerge in this system, rather that we do not detect them with these data at this time.
-
-```{r cross-map-fxn}
-# plots a cross mapping between two variables in the data, given their variable names (grp), their optimal embedding dimensions (from simplex forecasting), and their preferred names for plotting
-xmap <- function(grp1,grp2,E1,E2,name1,name2){
-    # cross-map group 1 to group 2
-
-  ccm1 <- ccm(datnorm,lib=as.matrix(segs),pred=as.matrix(segs),E=E1,lib_column= grp1,target_column = grp2,lib_sizes = c(10,25,50,75),num_samples=100,replace=T,silent=T,RNGseed = 41389)
-
-  # test for convergence (is predictability at max library size significantly greater than at minimum library size?)
-
-  rhomeans1 <- ccm1 %>% ccm_means()
-
-  conv_test1 <- t.test(ccm1$rho[ccm1$lib_size==10],ccm1$rho[ccm1$lib_size==75])$p.value
-  if((rhomeans1$rho[rhomeans1$lib_size==75]-rhomeans1$rho[rhomeans1$lib_size==10])<0.05) conv_test1 <- 1
-  conv_test1 <- ifelse(conv_test1<0.05,"Convergent","Non-convergent")
-
-  xmap1 <- ccm1 %>%
-    group_by(lib_size)%>%
-    summarise(rhomean=mean(rho,na.rm=T),upper=quantile(rho, 0.975),lower=quantile(rho, 0.025))%>%
-    ungroup()%>%
-    ggplot(aes(lib_size,rhomean))+
-    geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.3,fill="red")+
-    geom_line(color="darkorchid3")+
-    geom_hline(aes(yintercept = 0), linetype = 2) +
-    labs(x="Library Size",y=expression(paste(rho, " (predictive skill)")),title=paste0(name1," xmap ",name2,"\n","(",conv_test1,")"))+
-    theme(plot.title = element_text(size=6))
-
-  # cross map group 2 to group 1
-  # inspect the output of simplex from the previous step and use the best embedding dimension (highest rho) for the carnivore time series
-
-  ccm2 <- ccm(datnorm,lib=as.matrix(segs),pred=as.matrix(segs),E=E2,lib_column= grp2,target_column = grp1,lib_sizes = c(10,25,50,75),num_samples=100,replace=T,silent=T,RNGseed = 41389)
-
-  rhomeans2 <- ccm2 %>% ccm_means()
-
-  conv_test2 <- t.test(ccm2$rho[ccm2$lib_size==10],ccm2$rho[ccm2$lib_size==75])$p.value
-  if((rhomeans2$rho[rhomeans2$lib_size==75]-rhomeans2$rho[rhomeans2$lib_size==10])<0.05) conv_test2 <- 1
-  conv_test2 <- ifelse(conv_test2<0.05,"Convergent","Non-convergent")
-
-  xmap2 <- ccm2 %>%
-    group_by(lib_size)%>%
-    summarise(rhomean=mean(rho,na.rm=T),upper=quantile(rho, 0.975),lower=quantile(rho, 0.025))%>%
-    ungroup()%>%
-    ggplot(aes(lib_size,rhomean))+
-    geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.3,fill="red")+
-    geom_line(color="darkorchid3")+
-    geom_hline(aes(yintercept = 0), linetype = 2) +
-    labs(x="Library Size",y=expression(paste(rho, " (predictive skill)")),title=paste0(name2," xmap ",name1,"\n","(",conv_test2,")"))+
-    theme(plot.title = element_text(size=6))
-
-  cross_map_plot <- xmap1 + xmap2
-
-cross_map_plot
-}
-```
-
-```{r cross-map-cp, fig.cap = "Cross mapping of effect of piscivores on carnivores (A) and carnivores on piscivores (B) in the PISCO data from 2000 to 2017.
-Shaded region show 95% confidence interval"}
-
-car_pis<-xmap(grp1="carnivore_norm",grp2="piscivore_norm",E1=4,E2=4,name1="Carnivores",name2="Piscivores")
-car_pl <- xmap(grp1="carnivore_norm",grp2="planktivore_norm",E1=4,E2=4,name1="Carnivores",name2="Planktivores")
-pl_pis <- xmap(grp1="planktivore_norm",grp2="piscivore_norm",E1=4,E2=4,name1="Planktivores",name2="Piscivores")
-
-car_pis
-```
-
-```{r cross-map-pl, fig.cap = "Cross mapping of effect of planktivores on carnivores (A) and carnivores on planktivores (B) in the PISCO data from 2000 to 2017.
-Shaded region show 95% confidence interval"}
-
-car_pl
-
-```
-
-```{r cross-map-pis, fig.cap = "Cross mapping of effect of piscivores on carnivores (A) and carnivores on planktivores (B) in the PISCO data from 2000 to 2017.
-Shaded region show 95% confidence interval"}
-
- pl_pis 
-```
-
-```{r trophic-cor-plot, fig.cap="Estimated correlation coefficients between broad trophic groups at the island by year resolution", eval = FALSE}
-
-trophic_effects_plot <- cip_data %>%
-  group_by(broadtrophic, region, site, side, transect, year, eventual_mpa) %>%
-  summarise(density = sum(density_g_m2)) %>%
-  group_by(broadtrophic, region, site, year) %>%
-  summarise(mean_density = mean(density, na.rm = T)) %>%
-  group_by(broadtrophic,region, year) %>%
-  summarise(regional_density = mean(mean_density, na.rm = T)) %>% 
-  ungroup() %>% 
-  spread(broadtrophic, regional_density) %>% 
-  group_by(region) %>% 
-  arrange(year) %>% 
-  mutate(lag4_carnivore = lag(carnivore,4),
-         lag4_piscovore = lag(piscivore,4)) %>% 
-  na.omit() %>% 
-  ungroup() %>% 
-  filter(year > 1999) %>% 
-  select(-year,-region) %>% 
-  corrr::correlate() %>% 
-   gather("colname","correlation",-rowname) %>% 
-  ggplot(aes(rowname, colname,fill = correlation)) + 
-  geom_tile() + 
-  geom_text(aes(rowname, colname, label = round(correlation,2))) + 
-  scale_fill_gradient2(low = "tomato", mid = "white", high = "steelblue", midpoint = 0, limits = c(-1,1),
-                       guide = guide_colorbar(frame.colour = "black",barheight = 13)) + 
-  theme(axis.text.x = element_text(angle = 45, vjust = 1,
-                                   hjust = 1),
-        axis.title.x = element_blank(),
-        axis.title.y = element_blank())
-
-trophic_effects_plot
-
-```
-
-## Repeat Analysis with Kelp Forest Monitoring Data
-
-As a robustness check to our main results, we repeated our analysis utilizing data provided by the [Kelp Forest Monitoring Program (KFM)](https://science.nature.nps.gov/im/units/medn/monitor/kelpforest.cfm) conducted in the Channel Islands.
-Despite having similar but different methods and survey locations, we find almost identical estimated trends in divergences between targeted and non-targeted species using the KFM data (Fig.S\@ref(fig:kfm-did)).
-
-```{r kfm-did,fig.cap = "Estimated divergence in biomass densities of targeted and non-targeted fishes in the Channel Islands (i.e.
-integrated across inside and outside of MPAs) using the KFM data .
-MPAs are implemented in 2003 (red dashed line).
-Estimates are from a regression on log(abundance index), and so estimated effects roughly correspond to percentage changes"}
-
-kfm_did_plot <- kfm_run$did_plot[[1]] + 
-  scale_y_continuous(name = "Divergence from Non-Targeted", labels = percent) + 
-  labs(caption = "")
-
-kfm_did_plot
-
-```
-
-## MPA Only Analysis
-
-Given trends in mean densities observed in the raw data, the "regional conservation effect" estimated by our model---defined as the divergence in trends between the targeted and non-targeted species across the Channel Islands region---is not surprising.
-By jumping through countless statistical hoops we reach a similar conclusion that we would just by looking at the divergences in the mean trends.
-The integration of data from inside and outside of MPAs is a possible explanation for this lack of a clear regional effect.
-If spillover is limited or has simply not developed yet, especially relative to the effect of fishing outside of MPAs, then it is possible that there is a clear positive effect inside the MPAs, a clear negative effect outside, and when we look across both types of sites we get an unclear average of the two.
-
-To address this, we can first repeat some exploratory data analysis of trends in densities inside and outside the MPAs for targeted and non-targeted species.
-@caselle2015 provides a thorough look at this question of differences inside and outside of MPAs.
-We update that analysis here to account for our specific questions of trend divergence, potential differences in filtering methods, to include data up through 2017, and to utilize our estimation method on just the inside-MPA data.
-For all exploratory analyses, we consider the same top `r length(top_species)` consistently observed species.
-Looking first at simple trends in total mean biomass density across these species inside and outside of MPAs, we find evidence that biomass densities inside the MPAs is increasing faster (and is higher inside) than outside (Fig.S\@ref(fig:bio-mpa)).
-
-```{r bio-mpa, fig.cap= "Annual mean aggregate biomass density (summed across all fishes) inside and outside of eventual MPA locations over time. Red dashed line indicates MPA implementation year"}
-
-abundance_index_plot <- cip_data %>% 
-  group_by(targeted, region,site,side, transect, year, eventual_mpa, classcode) %>% 
-  summarise(density = sum(density_g_m2)) %>% 
-  group_by(targeted,classcode, region,site, year, eventual_mpa) %>% 
-  summarise(mean_density = mean(density, na.rm = T)) %>% 
-  group_by(targeted, classcode, year, eventual_mpa) %>% 
-  summarise(mmd = mean(mean_density, na.rm = T)) %>% 
-  group_by(year, eventual_mpa) %>% 
-  summarise(abundance_index = sum(mmd, na.rm = T)) %>% 
-  ggplot(aes(year, abundance_index, color = eventual_mpa)) + 
-  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
-  geom_line(size = 2) + 
-  scale_color_manual(values = c("#009ACD", "#548B54"), name = "" , labels = c("Non-MPA",'MPA')) + 
-  scale_y_continuous(name = "Total Abundance Index")+
-  labs(x="Year")
-
-abundance_index_plot
-
-```
-
-Our proposed identification strategy here though is not that total biomass density should be different inside and outside, but that the non-targeted species should serve as the control to the targeted.
-If we believe that the MPA effects are greater inside the MPA, then we would expect to see stronger divergences in biomass densities between these two targeted and non-targeted fishes inside the MPAs than outside
-
-```{r div-mpa, fig.cap="Trends in total biomass density inside and outside of eventual MPAs for targeted and non-targeted fishes. Red dashed line indicates MPA implementation year"}
-
-facet_labels <- c(
-  `TRUE` = "MPA",
-  `FALSE` = "Non-MPA"
-)
-
-mpa_abundance_plot <- cip_data %>% 
-  mutate(targeted = targeted == 1) %>% 
-  group_by(targeted, region,site,side, transect, year, eventual_mpa, classcode) %>% 
-  summarise(density = sum(density_g_m2)) %>% 
-  group_by(targeted,classcode, region,site, year, eventual_mpa) %>% 
-  summarise(mean_density = mean(density, na.rm = T)) %>% 
-  group_by(targeted, classcode, year, eventual_mpa) %>% 
-  summarise(mmd = mean(mean_density, na.rm = T)) %>% 
-  group_by(year, targeted, eventual_mpa) %>% 
-  summarise(abundance_index = sum(mmd, na.rm = T)) %>% 
-  ungroup() %>% 
-  group_by(targeted, eventual_mpa) %>% 
-  mutate(abundance_index = (abundance_index - mean(abundance_index)) / sd(abundance_index)) %>% 
-  ungroup() %>% 
-  rename(`Eventual MPA?` = eventual_mpa) %>% 
-  ggplot(aes(year, abundance_index, color = targeted)) + 
-  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
-  geom_line(size = 2) + 
-  scale_color_manual(values = c("steelblue", "tomato") ,name = "",
-                     labels = c("Non-Targeted","Targeted")) + 
-  facet_wrap(~`Eventual MPA?`, labeller = labeller(`Eventual MPA?` = facet_labels)) + 
-    scale_y_continuous(name = "Total Abundance Index")+
-  labs(x='Year')
-
-mpa_abundance_plot
-
-```
-
-Here we see a different picture.
-While there is some visual evidence that the targeted species were diverging from the non-targeted faster inside the MPAs than outside, both inside and outside we see that the trend in total biomass density of targeted species is trending downward, relative to the trend in the non-targeted species in recent years.
-This analysis is of total biomass density.
-However, our model estimates the mean difference in targeted and non-targeted species.
-Both have their advantages, but we chose the mean to reflect a hypothesis that the MPAs would provide positive benefits across all targeted species.
-The total biomass density could be strongly affected by a sharp increase or decrease in one or two species, even if the mean trend is different.
-Examining the mean trends though, we see the same results (Fig.S\@ref(fig:mean-div-mpa)).
-
-```{r mean-div-mpa, fig.cap = "Trends in mean total biomass density inside and outside of eventual MPAs for targeted and non-targeted fishes. Red dashed line indicates MPA implementation year"}
-
-targeted_mpa_plot <- cip_data %>% 
-  mutate(targeted = targeted == 1) %>% 
-  group_by(targeted, region,site,side, transect, year, eventual_mpa, classcode) %>% 
-  summarise(density = sum(density_g_m2)) %>% 
-  group_by(targeted,classcode, region,site, year, eventual_mpa) %>% 
-  summarise(mean_density = mean(density, na.rm = T)) %>% 
-  group_by(targeted, classcode, year, eventual_mpa) %>% 
-  summarise(mmd = mean(mean_density, na.rm = T)) %>% 
-  group_by(year, targeted, eventual_mpa) %>% 
-  summarise(abundance_index = mean(mmd, na.rm = T)) %>% 
-  ungroup() %>% 
-    group_by(targeted, eventual_mpa) %>% 
-  mutate(abundance_index = (abundance_index - mean(abundance_index)) / sd(abundance_index)) %>% 
-  ungroup() %>% 
-  rename(`Eventual MPA?` = eventual_mpa) %>% 
-  ggplot(aes(year, abundance_index, color = targeted)) + 
-  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
-  geom_line(size = 2) + 
-  scale_color_manual(values = c("steelblue", "tomato") ,name = "Targeted?") + 
-  facet_wrap(~`Eventual MPA?`, labeller = label_both) + 
-    scale_y_continuous(name = "Mean Abundance Index")+
-  labs(x="Year")
-
-targeted_mpa_plot
-```
-
-Lastly, we can examine both the mean and individual trend to check clear species-by-species outliers in the overall biomass density trends.
-This analysis shows noise, but overall the targeted and non-targeted species seem to be following similar trends within their respective groups
-
-```{r cs-trends, fig.cap = "Centered and scaled biomass density trends for each fish grouped by targeted and non targeted (pale lines) and fitted LOESS smoother (with 95% confidence intervals around mean) and mean by targeted and non-targeted groups, inside and outside od MPAs.
-Red dashed line indicates MPA implementation year"}
-
-species_mpa_trend_plot <- cip_data %>% 
-  group_by(year, classcode, eventual_mpa, targeted) %>% 
-  summarise(mean_density = mean(density_g_m2, na.rm = T)) %>% 
-  ungroup() %>% 
-  mutate(targeted = targeted == 1) %>% 
-  group_by(classcode, eventual_mpa) %>% 
-  mutate(cs_density = (mean_density - mean(mean_density)) / sd(mean_density)) %>% 
-  ungroup() %>% 
-    rename(`Eventual MPA?` = eventual_mpa) %>% 
-  ggplot(aes(year, cs_density, color = targeted, fill = targeted)) + 
-    geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
-  geom_line(aes(group = interaction(classcode, targeted)),alpha = 0.5) + 
-  geom_smooth() +
-    scale_color_manual(values = c("steelblue", "tomato") ,name = "", labels = c("Non-Targeted","Targeted")) +
-      scale_fill_manual(values = c("steelblue", "tomato") ,name = "",labels = c("Non-Targeted","Targeted")) +
-
-    facet_wrap(~`Eventual MPA?`, labeller = labeller(`Eventual MPA?` = facet_labels)) + 
-    labs(y = "Centered and Scaled Mean Density", x = "Year")
-
-species_mpa_trend_plot
-
-```
-
-These visual assessments suggest that similar to our results looking both inside and outside of MPAs, we would expect that our estimation model fitted only on data from inside eventual MPAs would reach similar conclusions as our results fitted to data from both inside and outside MPAs.
-To test this, we re-ran our analysis, but only using data from sites that are eventually placed inside MPAs.
-Our results reflect the same trends as displayed in the raw data and the statistical region-wide analysis, providing robust statistical support to the conclusions we would reach from visually examining the raw data (Fig.S \@ref(fig:mpa-did)).
-
-```{r mpa-did,fig.cap = "Estimated divergence in biomass densities of targeted and non-targeted fishes inside eventual Channel Islands MPAs.
-MPAs are implement in 2003 (red dashed line).
-Estimates are from a regression on log(abundance index), and so estimated effects roughly correspond to percentage changes" }
-
-mpa_did_plot <- mpa_run$did_plot[[1]] + 
-       labs(x = "Year", y = "Estimate of Regional Effect",
-          caption = "")
-
-mpa_did_plot
-
-```
 

### 2020-04-21 functions/estimate_did.R

@@ -8,8 +8,10 @@ estimate_did <-
            cdfw_catches,
            data_to_use = "all",
            data_source = "pisco",
+           fit_mean = TRUE,
            chains = 4,
            cores = 4,
+           first_year = 2003,
            iter = 5000) {
     
     # data <- pisco_abundance_data
@@ -51,7 +53,7 @@ estimate_did <-
       filter(n_years >= min_classcode_years)
     
     year_bins <-
-      c(1999, seq(2003, max(data$year) + 1, by = bin_width))
+      c(2000, seq(first_year, max(data$year) + 1, by = bin_width))
     
     did_data <- data %>% {
       if (consistent_sites_only == TRUE) {
@@ -62,6 +64,7 @@ estimate_did <-
       }
     } %>%
     filter(classcode %in% consistent_classcodes$classcode) 
+    
     # rm("data","cdfw_catches","life_history_data")
     if (data_source == "pisco"){
     
@@ -120,8 +123,18 @@ estimate_did <-
       left_join(annual_catches, by = "year") %>% 
       group_by(region) %>% 
       mutate(regional_temp_dev = scale(var_temp)) %>% 
-      ungroup()
+      ungroup() %>% 
+       mutate(var_tex_2 = var_tex^2,
+              var_lag_catch_2 = var_lag_catch^2,
+              regional_temp_dev_2 = regional_temp_dev^2,
+              var_temp_2 = var_temp^2)
 
+     if (fit_mean){
+       
+       did_data$total_biomass_density <- did_data$mean_biomass_density
+       
+     }
+     
     vars <- which(str_detect(colnames(did_data), "var_"))
     
     nafoo <- function(x){
@@ -130,8 +143,9 @@ estimate_did <-
     }
     did_data <- purrrlyr::dmap_at(did_data, vars, ~ scale(.x)) %>%
      purrrlyr::dmap_at(vars, nafoo) %>%
-      mutate(total_biomass_density = total_biomass_density + 1e-6,
-             mean_biomass_density = mean_biomass_density + 1e-6) %>% 
+      filter(total_biomass_density > 0) %>% 
+      mutate(total_biomass_density = total_biomass_density,
+             mean_biomass_density = mean_biomass_density) %>% 
       group_by(site_side, targeted) %>%
       mutate(scaled_total_biomass_density = scale(log(total_biomass_density)))
     # filter(region != "SMI")
@@ -200,23 +214,40 @@ estimate_did <-
[…]

### 2020-04-21 make-regional-mpas.R

@@ -11,6 +11,7 @@ library(stringr)
 library(lubridate)
 library(lme4)
 library(TMB)
+library(TMBhelper)
 library(FishLife)
 library(patchwork)
 library(rstan)
@@ -54,9 +55,9 @@ run_description <- "PNAS R and R with simplified DiD and weighting"
 # So, once you've run simulate_mpas, you can set it to FALSE and validata_mpas will work
 
 
-run_did <- FALSE # run difference in difference on data from the CINMS
+run_did <- TRUE # run difference in difference on data from the CINMS
 
-run_tmb <- FALSE
+run_tmb <- TRUE
 
 simulate_mpas <- FALSE # simulate MPA outcomes
 
@@ -74,7 +75,7 @@ sim_years <- 50
 
 num_patches <- 50
 
-n_cores <- 8
+n_cores <- 1
 
 # prepare run -------------------------------------------------------------
 
@@ -1296,7 +1297,6 @@ model_runs <- cross_df(
   filter(!(data_source == "kfm" & str_detect(var_names,"pisco"))) %>%
   filter(!(data_source == "kfm" & (data_to_use %in% c("mpa_only","fished_only"))))
 
-
 if (run_tmb == T){
 
   if (file.exists("fit-progress.txt")){
@@ -1306,14 +1306,18 @@ if (run_tmb == T){
   # browser()
   # future::plan(future::multiprocess, workers = 4)
 
-  doParallel::registerDoParallel(cores = n_cores)
+  tmb_runs <- model_runs %>%     
+    filter(var_names == "pisco_a", data_to_use == "all", center_scale == TRUE)
+
+  
+  doParallel::registerDoParallel(cores = 1)
   #
   # model_runs <- model_runs %>%
   #   slice(3)
 
-  compile(here::here("src", paste0(script_name, ".cpp")), "-O0") # what is the -O0?
+  TMB::compile(here::here("src","fit_zissou.cpp"), "-O0") # what is the -O0?
   
-  dyn.load(dynlib(here::here("src", script_name)))
+  dyn.load(dynlib(here::here("src","fit_zissou")))
   
   fits <- foreach::foreach(i = 1:nrow(model_runs)) %dopar% {
 
@@ -1344,19 +1348,13 @@ if (run_tmb == T){
     write(glue::glue("error message:{fits$error}"), file = "fit-errors.txt",
           append = T)
 
-    out <- fits
 
[…]

### 2020-04-22 documents/ovando-regional-effects-of-mpas-si-v2.Rmd
confirm knit of corrected supplemental information and paper

@@
 knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
-                      cache = FALSE, fig.width = 3.42, fig.asp = 1.5)
+                      cache = FALSE, fig.width = 7, fig.asp = .75)
 
@@
 
+tmb_runs <- read_rds( file.path(run_dir,"tmb_model_fits.rds"))
+
 # load(file = here::here("results",run_name, "model_runs.Rdata"))
@@
 used_data <- abundance_data$data[abundance_data$data_source == "pisco"][[1]]
+
+filter_summary <- used_data %>% 
+  select(classcode, targeted) %>% 
+  unique() %>% 
+  group_by(targeted) %>% 
+  count() %>% 
+  ungroup() %>% 
+  mutate(targeted = ifelse(targeted == 1,"yes","no")) %>% 
+  spread(targeted, n)
 
@@
   mutate(mpa_effect = pmin(mpa_effect,3))
+
+ filter_summary <- used_data %>% 
+  select(classcode, targeted) %>% 
+  unique() %>% 
+  group_by(targeted) %>% 
+  count() %>% 
+  ungroup() %>% 
+  mutate(targeted = ifelse(targeted == 1,"yes","no")) %>% 
+  spread(targeted, n)
+
 ```
@@
 
-```{r classcode}
+```{r classcode, eval = FALSE}
 
@@
 
-```{r did-coefs, results = "asis"}
+```{r did-coefs, results = "asis",eval = FALSE}
 
@@
 
-```{r mean_test_plot, fig.cap = "Posterior predictive sum (distribution) and empirical sum of total biomass density. Posterior predictive sum adequately recovers empirical sum."}
+```{r sum_test_plot, fig.cap = "Posterior predictive sum (distribution) and empirical sum of total biomass density. Posterior predictive sum adequately recovers empirical sum."}
 
@@
 For this run we estimate the effect annually, relative to the year 2000.
-Results show the same overall trend and magnitude reported in our base model.", results = "hide", eval = FALSE}
+Results show the same overall trend and magnitude reported in our base model.", results = "hide", eval = TRUE}
 no_bin <- model_runs %>%
@@
       cdfw_catches = cdfw_catches,
-      life_history_data = life_history_data
+      life_history_data = life_history_data,
+      refresh = 1000
     )
@@
 
- exclude zeros
+Appropriately addressing the problem of "missing" observations is a critical challenge in any field observation study.
+If no observations of a given fish species were recorded on a given transect, should the density of that species on that transect be marked as zero, and influence the estimate of the overall mean density accordingly?
+The obvious answer seems to be yes, but what if that species simply does not live in the environment covered by a particular transect, or was not present during the particular time of the diver's observation?
+For our base runs, we assign a value of zero density on a given transect for any fish species that has been observed at least once at a given site at any time in our data but was not observed on that particular transect.
+If that species was never observed at that site, we do not include a zero for that species.
+Our rationale for this is that given the shifting nature of the sampled sites, and the intensity of sampling at those sites, we do not want to skew density trends by changes in the amount of suitable habitat for a given species sampled.
+However, this is clearly a strong assumption.
+For example, perhaps the decreasing trend in mean densities from 2000 to 2004 is due to increased number of sites (and therefore zeros) included in the data.
+To assess the potential importance of this choice, we can compare the mean densities of targeted and non-targeted species over time with the added zeros to the mean densities using only positive observations (i.e.
+not including any zeros in the data, (Fig.S\@ref(fig:no-zero-fit)).
+The trends in the raw densities, and most importantly the mean trends of targeted and non-targeted fishes, are nearly identical whether or not zeros are added, providing strong evidence that our choice of how to incorporate missing observations into the data are not strongly influencing our overall results.
 
@@
 
+tmb_runs <- tmb_runs %>%
+  mutate(tmb_fit = map(tmb_fit,"result")) %>%
+  mutate(processed_fits = map(tmb_fit, process_fits)) %>%
+  mutate(did_plot = map(processed_fits, "did_plot"))
+
+betas <- tmb_runs$processed_fits[[1]]$betas
+
 table <- betas %>% select(-std_error,-group)
@@
    filter(str_detect(group, "non_nested") & !str_detect(variable, "site_side")) %>%
+  mutate(group = str_remove_all(group,"(?=\\.).*" )) %>% 
    ggplot() +
@@
    filter(str_detect(group, "non_nested") & str_detect(variable, "site_side")) %>%
+    mutate(group = str_remove_all(group,"(?=\\.).*" )) %>% 
   mutate(variable = tolower(variable)) %>% 
@@
    filter(str_detect(group, "region_cluster_betas")) %>%
+    mutate(group = str_remove_all(group,"(?=\\.).*" )) %>% 
    mutate(cluster = str_replace_all(variable,"\\D","")) %>%
@@
 
+report <- tmb_runs$tmb_fit[[1]]
+
 targeted_trends <- report$zissou_estimates %>% 
   filter(str_detect(variable, "targeted_did")) %>% 
+  mutate(variable = str_remove_all(variable,"(?=\\.).*" )) %>% 
   group_by(variable) %>% 
-  mutate(year =zissou_fit$did_data$year %>% unique(),
+  mutate(year =report$did_data$year %>% unique(),
          labvar=ifelse(variable=="nontargeted_did_betas","Nontargeted","Targeted")) %>% 
@@
 
-fit_report <- base_run$tmb_fit[[1]]$zissou_report
-
-observed_seen <- base_run$tmb_fit[[1]]$zissou_data$log_density
-
-observed_seeing <- base_run$tmb_fit[[1]]$zissou_data$any_seen
+fit_report <- tmb_runs$tmb_fit[[1]]$zissou_report
+
+observed_seen <- tmb_runs$tmb_fit[[1]]$zissou_data$log_density
+
+observed_seeing <- tmb_runs$tmb_fit[[1]]$zissou_data$any_seen
 
@@
 
-a = base_run$tmb_fit[[1]]$zissou_data$x_seen_non_nested %>% 
+a = tmb_runs$tmb_fit[[1]]$zissou_data$x_seen_non_nested %>% 
   as_data_frame() %>% 
@@
 
-roc_plot <- rocplot + style_roc(theme = zissou_theme)
+roc_plot <- rocplot + style_roc()
 
@@
 ```{r reg-pop, fig.cap = "Mean density by island by year for each fish species included in the analysis"}
+
+fitted_data <- used_data
 
@@
 
-years <- unique(fitted_data$year)
-
-years <- zissou_fit$did_data$year %>% unique()
-
-abundance_trends <- data_frame(abundance_hat = fit_report$abundance_hat,
-                               classcode = rep(1:n_distinct(fitted_data$classcode), each = length(years))) %>%
+years <- report$did_data$year %>% unique()
+
+abundance_trends <-
+  data_frame(
+    abundance_hat = fit_report$abundance_hat,
+    classcode = rep(1:n_distinct(fitted_data$classcode), each = length(years))
+  ) %>%
   mutate(log_abundance_hat = log(abundance_hat)) %>%
@@
   rename(numeric_classcode = classcode) %>%
-  left_join(life_history, by = "numeric_classcode" )
+  left_join(life_history, by = "numeric_classcode")
 
@@
 
-numeric_cor_plot <- base_run$data[[1]] %>%
+numeric_cor_plot <- used_data %>%
   select_if(is.numeric) %>%
-  select(-log_density, -targeted) %>%
+  select(contains("mean_"), n_obs) %>%
   na.omit() %>%
@@
 
-#### Sensitivity to "missing" observations
-
-Appropriately addressing the problem of "missing" observations is a critical challenge in any field observation study.
-If no observations of a given fish species were recorded on a given transect, should the density of that species on that transect be marked as zero, and influence the estimate of the overall mean density accordingly?
-The obvious answer seems to be yes, but what if that species simply does not live in the environment covered by a particular transect, or was not present during the particular time of the diver's observation?
-For our base runs, we assign a value of zero density on a given transect for any fish species that has been observed at least once at a given site at any time in our data but was not observed on that particular transect.
-If that species was never observed at that site, we do not include a zero for that species.
-Our rationale for this is that given the shifting nature of the sampled sites, and the intensity of sampling at those sites, we do not want to skew density trends by changes in the amount of suitable habitat for a given species sampled.
-However, this is clearly a strong assumption.
-For example, perhaps the decreasing trend in mean densities from 2000 to 2004 is due to increased number of sites (and therefore zeros) included in the data.
-To assess the potential importance of this choice, we can compare the mean densities of targeted and non-targeted species over time with the added zeros to the mean densities using only positive observations (i.e.
-not including any zeros in the data, (Fig.S\@ref(fig:nozero-raw-trend)).
-The trends in the raw densities, and most importantly the mean trends of targeted and non-targeted fishes, are nearly identical whether or not zeros are added, providing strong evidence that our choice of how to incorporate missing observations into the data are not strongly influencing our overall results.
-
-```{r nozero-raw-trend, fig.cap = "Centered and scaled mean annual density, excluding zeros, of included fishes (points) and smoothed means of targeted and non-targeted groups (line) over time" }
-
-nozero_trend_plot <- base_run$data[[1]] %>% 
-  filter(any_seen == T) %>% 
-  select(year, targeted,classcode, log_density, any_seen) %>% 
-  mutate(density = exp(log_density) * any_seen) %>% 
-  group_by(classcode,targeted, year) %>% 
-  summarise(md = mean(density)) %>% 
-  group_by(classcode,targeted) %>% 
-  mutate(smd = (md - mean(md)) / sd(md)) %>% 
-  ggplot(aes(year, smd, color = targeted == 1, fill = targeted == 1)) + 
-  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
-  geom_line(aes(group = interaction(targeted, classcode)), alpha = 0.25) +
-  geom_smooth() +
-  scale_color_d3(name = "Targeted?") + 
-  scale_fill_d3(name = "Targeted?") +
-  labs(y = "Centered and Scaled Mean Density", x = "Year")
-
-nozero_trend_plot
-```
-
 ## Testing Model Assumptions
@@
 ```{r trophic-plot, fig.cap="Centered and scaled densities by broad trophic group and island over time"}
-
-trophic_groups <- cip_data %>% 
+cip_data <- used_data
+trophic_groups <- used_data %>% 
   select(classcode, broadtrophic) %>% 
@@
 |steepness (h) | ~uniform(0.6,0.95)|
-|Adult movement ($\sigma_{s=a}$) | ~uniform(0,0.5 * P)|
-|Larval movement ($\sigma_{s=l}$) | ~uniform(0,0.5 * P)|
-|Recruitment variation ($\sigma_{r}$) |  $\in\{0,0.05,.1\}$|
-|Recruitment autocorrelation ($ac_{r}$) |  $\in\{0,0.05,.1\}$|
+|Adult movement ($\sigma_{s=a}$) | ~uniform(0,0.25 * P)|
+|Larval movement ($\sigma_{s=l}$) | ~uniform(0,0.25 * P)|
+|Recruitment variation ($\sigma_{r}$) |  $\in\{0,0.05,.1,.2\}$|
+|Recruitment autocorrelation ($ac_{r}$) |  $\in\{0,0.05,.1,.2\}$|
 |DD adult movement (dmod) | $\in\{0.25,1\}$|

### 2020-04-22 functions/estimate_did.R

@@ -9,8 +9,9 @@ estimate_did <-
            data_to_use = "all",
            data_source = "pisco",
            fit_mean = TRUE,
-           chains = 4,
-           cores = 4,
+           chains = 1,
+           cores = 1,
+           refresh = 500,
            first_year = 2003,
            iter = 5000) {
     
@@ -188,6 +189,8 @@ estimate_did <-
     
     env$iter <-  iter
     
+    env$refresh <- refresh
+    
     # did_reg <- with(env, {
     #   stan_glmer(
     #     log(total_biomass_density + 1e-6) ~ targeted * year_bins + (site_side - 1 |
@@ -227,7 +230,8 @@ estimate_did <-
         prior_intercept = normal(autoscale = TRUE),
         prior = normal(0, 2, autoscale = TRUE),
         iter = iter,
-        family = Gamma(link = "log")
+        family = Gamma(link = "log"),
+        refresh = refresh
       )
     })
     
@@ -445,6 +449,9 @@ estimate_did <-
       env$chains <- chains
       
       env$iter <-  iter
+      
+      env$refresh <- refresh
+      
       #re-weight data in proportion to MPA area
       weights <- did_data %>% 
         mutate(mpa_area = ifelse(eventual_mpa, 0.2, 0.8)) %>% 
@@ -472,7 +479,8 @@ estimate_did <-
           prior_intercept = normal(0, 2, autoscale = TRUE),
           prior = normal(0, 2),
           family = Gamma(link = "log"),
-          iter = iter
+          iter = iter,
+          refresh = refresh
         )})
       
       

### 2020-04-22 documents/reviewer-responses.Rmd
pre changing to population language commit

@@
 ---
-title: "Reviewer Responses for 'Unclear Regional Effects of a Large Marine Protected Area Network' (previously 'The Regional Effects of MPAs')"
+title: "Reviewer Responses for 'Population Effects of Marine Protected Areas Are Likely To Be Unclear' (previously 'The Regional Effects of MPAs')"
 author: 
@@
 As MPAs are used as a core tool of marine resource management, our results provide a useful update to our empirical expectations for MPAs, that to date have been based almost entirely on inside-outside MPA response ratios.
+
+As pointed out by the reviewers, many of the more theoretical points we make about the importance of fleet dynamics and the challenges of response ratios have been made in the past.
+However, we feel there is substantial value in revisiting those discussions in the context of a concrete empirical example, to provide scientists and practitioners alike a cohesive reference to observe and understand the full ramifications of these past findings on our practical ability to monitor and manage marine protected area networks.
+
 We hope you find the updated version of our manuscript improved thanks to the inclusions of the careful points raised by the prior revision of this paper.

### 2020-04-22 documents/Untitled.Rmd
pre changing to population language commit

--- 2020-03-26 documents/Untitled.Rmd
@@
 ---
-title: The Regional Effects of Marine Protected Areas
-author: Daniel Ovando, Jennifer E. Caselle, Christopher Costello, Olivier Deschenes, Steven D. Gaines, Ray Hilborn, Owen Liu
+title: "Supporting Information: The Regional Effects of Marine Protected Areas"
+author:  Daniel Ovando,Jennifer E. Caselle,Christopher Costello,Olivier Deschenes,Steven D. Gaines,Ray Hilborn,Owen Liu
+output: 
+  bookdown::pdf_document2:
+    keep_tex: true
+  pdf_document: default
 bibliography: My Library.bib
-csl: pnas.csl
-
-output:
-  bookdown::pdf_document2: default
-  bookdown::word_document2: default
-  bookdown::html_document2: default
+header-includes:
+    - \usepackage{longtable}
+    - \usepackage{booktabs}
+    - \usepackage{eso-pic}
+    - \usepackage{graphicx}
+    - \usepackage[left]{lineno}
+    - \usepackage{xcolor}
 params:
@@
   num_patches: [25]
-linkcolor: blue
+always_allow_html: true
 ---
 
-$$d_{i} \sim Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + B\mathbf{X} + B,shape, scale)$$
+```{r, include=FALSE}
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
+                      cache = FALSE, fig.width = 7, fig.asp = .75)
+
+```
+
+```{r, include=FALSE}
+library(sf)
+library(ggmap)
+library(viridis)
+library(hrbrthemes)
+library(scales)
+library(patchwork)
+library(rpart)
+library(extrafont)
+library(caret)
+library(ggsci)
+library(rEDM)
+library(tidyverse)
+library(ggtext)
+library(rstan)
+library(rstanarm)
+library(bayesplot)
+extrafont::loadfonts()
+rstan_options(auto_write = TRUE)
+
+functions <- list.files(here::here("functions"))
+
+walk(functions, ~ here::here("functions", .x) %>% source()) # load local functions
+save_plots <- TRUE
+
+# sim_years <- 50
+# 
+# burn_years <- 20
+# 
+# num_patches <- 25
+
+# run_name <- "v5.0"
+
+width_1col <- 3.42 # inches
+
+asp_1col <- 1.25
+
+width_2col <- 7 # inches
+
+asp_2col <-  .75
+
+dpi <- 900
+
+theme_1col <- hrbrthemes::theme_ipsum(base_size = 11,
+                                      axis_title_size = 12)
+
+theme_2col <- hrbrthemes::theme_ipsum(base_size = 12,
+                                        axis_title_size = 14)
+
+run_name <- params$run_name
+
+sim_years <- params$sim_years
+
+burn_years <- params$burn_years
+
+num_patches <- params$num_patches
+
+run_dir <- here::here("results", run_name)
+
+experiment_dir <- here::here("results",run_name, "experiments")
+
+load(file = here::here("results",run_name, "rawish_zissou_data.Rdata"))
+
+load(file.path(run_dir, 'abundance_data.Rdata'))
+
+model_runs <- read_rds( file.path(run_dir,"did_fits.rds"))
+
+tmb_runs <- read_rds( file.path(run_dir,"tmb_model_fits.rds"))
+
+# load(file = here::here("results",run_name, "model_runs.Rdata"))
+
+processed_grid <-  read_rds(file.path(run_dir,"filtered_processed_grid.rds"))
+
+load(file = file.path(run_dir ,"sim_grid.Rdata"))
+
+load(file = file.path(run_dir,"plots.RData"))
+
+valgrid <- read_rds(file.path(run_dir, "valgrid.rds"))
+
+cdfw_catches <-
+  read_csv(file = here::here("data", 'cdfw-catches.csv')) %>%
+  group_by(sci_name, year) %>%
+  summarise(catch = sum(pounds_caught, na.rm = T)) %>%
+  left_join(
+    life_history_data %>% select(taxa, classcode, commonname) %>% mutate(taxa = tolower(taxa)),
+    by = c("sci_name" = "taxa")
+  ) %>%
+  filter(!is.na(classcode))
+
+# cdfw_catches %>%
+#   group_by(commonname, classcode, sci_name) %>%
+#   summarise(catch =sum(catch, na.rm = TRUE)) %>%
+#   mutate(in_analysis = classcode %in% unique(cip_data$classcode)) %>%
+#   View()
+
+density_ratios <- readRDS(file.path(run_dir,"density_ratios.rds"))
+
+outcomes <- readRDS(file.path(run_dir,"outcomes.rds"))
+
+outcomes <- outcomes %>% 
+  filter(!is.na(mpa_effect))
+
+density_ratios <- density_ratios %>% 
+  filter(!is.na(biased_density_ratio))
+
+   eqo <- outcomes %>%
+     filter(year == max(year))
+
+facet_labels <- c(
+  mpa_size = "Range in MPA",
+  depletion = "Depletion"
+)
+
+testplot <- ggplot()
+
+panel_height = unit(1,"npc") - sum(ggplotGrob(testplot)[["heights"]][-3]) - unit(4,"line")
+
+panel_width = unit(1,"npc") - sum(ggplotGrob(testplot)[["widths"]][-3]) - unit(4,"line")
+
+gc <- guide_colorbar(frame.colour = "black",
+                     ticks.colour = "white",
+                     barheight = panel_height)
+
+  hgc <- guide_colorbar(frame.colour = "black",
+                       ticks.colour = "black",
+                       barwidth = 10)
+
+    wide_hgc <- guide_colorbar(frame.colour = "black",
+                       ticks.colour = "black",
+                       barwidth = 25)
+
+theme_set(theme_1col)
+
+# models_worked <- model_runs$tmb_fit %>% map("error") %>% map_lgl(is_null)
+
+# model_runs <- model_runs %>%
+#   filter(models_worked) %>%
+#   mutate(tmb_fit = map(tmb_fit,"result")) %>%
+#   mutate(processed_fits = map(tmb_fit, process_fits)) %>%
+#   mutate(did_plot = map(processed_fits, "did_plot"))
+
+base_run <- model_runs %>%
+filter(var_names == "pisco_a", data_to_use == "all", center_scale == TRUE)
+
+mpa_run <- model_runs %>%
+  filter(data_source == "pisco",
+         var_names == "pisco_a",
+         data_to_use == "mpa_only")
+
+fishsed_run <- model_runs %>%
+  filter(data_source == "pisco",
+         var_names == "pisco_a",
+         data_to_use == "fished_only")
+
+kfm_run <- model_runs %>%
+  filter(data_source == "kfm")
+
+used_data <- abundance_data$data[abundance_data$data_source == "pisco"][[1]]
+
+filter_summary <- used_data %>% 
+  select(classcode, targeted) %>% 
+  unique() %>% 
+  group_by(targeted) %>% 
+  count() %>% 
+  ungroup() %>% 
+  mutate(targeted = ifelse(targeted == 1,"yes","no")) %>% 
+  spread(targeted, n)
+
+did_data <- model_runs$did_fit[model_runs$data_source == "pisco" &
+                                 model_runs$var_names == "pisco_a" &
+                                 model_runs$data_to_use == "all"][[1]]$did_data
+
+did_fit <- model_runs$did_fit[model_runs$data_source == "pisco" &
+                                 model_runs$var_names == "pisco_a" &
+                                 model_runs$data_to_use == "all"][[1]]$did_reg
+
+did_results <- model_runs$did_fit[model_runs$data_source == "pisco" &
+                                 model_runs$var_names == "pisco_a" &
+                                 model_runs$data_to_use == "all"][[1]]$did_results
+
+site_data <- read_csv(here::here("data",'Final_Site_Table_UCSB.csv')) %>%
+  magrittr::set_colnames(., tolower(colnames(.))) %>%
+  select(
+    site,
+    side,
+    mpagroup,
+    mpa_status,
+    reserve,
+    region,
+    year_mpa,
+    mpaareanm2,
+    lat_wgs84,
+    lon_wgs84
+  ) %>%
+  rename(lat_wgs84 = lon_wgs84,
+         lon_wgs84 = lat_wgs84) %>%
+  unique() %>%
+  mutate(eventual_mpa = (year_mpa > 0))
+
+top_species <- used_data$classcode %>% unique()
+
+life_summaries <- life_history_data %>% 
+  select(classcode, tm, vbgf.k, vbgf.linf) %>% 
+  rename(age_mature = tm, vbk = vbgf.k, linf = vbgf.linf) %>% 
+  unique() %>% 
+  filter(classcode %in% top_species) %>% 
+  select(-classcode) %>% 
+  gather(variable,value) %>% 
+  group_by(variable) %>% 
+  summarise(min_val = min(value, na.rm = TRUE),
+         max_val = max(value, na.rm = TRUE))
+
+ sim_life_history <- sim_grid %>%
+    mutate(
+      age_mature = map_dbl(fish, "age_mature"),
+      vbk = map_dbl(fish, "vbk"),
+      linf = map_dbl(fish, "linf")
+    ) %>%
+    select(scientific_name, age_mature, vbk, linf) %>%
+    unique() %>% 
+   filter(age_mature > life_summaries$min_val[life_summaries$variable == "age_mature"],
+          age_mature < life_summaries$max_val[life_summaries$variable == "age_mature"],
+          vbk > life_summaries$min_val[life_summaries$variable == "vbk"],
+          vbk < life_summaries$max_val[life_summaries$variable == "vbk"],
+          linf > life_summaries$min_val[life_summaries$variable == "linf"],
+          linf < life_summaries$max_val[life_summaries$variable == "linf"])
+
+ year_bins <-
+  seq(2003, max(pisco_data$year) + 1, by = 3)
+
+ sim_mpa_effects <- outcomes %>% 
+  filter(set == "ci",
+         final_depletion > 0,
+         f_v_m > .5) %>% 
+  mutate(year = years_protected + 2002) %>% 
+  filter(year <= max(pisco_data$year)) %>% 
+  mutate(binned_year = cut(year, year_bins, include.lowest = FALSE))  %>% 
+  filter(year >= 2003,
+         !is.na(binned_year)) %>% 
+  mutate(mpa_effect = pmin(mpa_effect,3))
+
+ filter_summary <- used_data %>% 
+  select(classcode, targeted) %>% 
+  unique() %>% 
+  group_by(targeted) %>% 
+  count() %>% 
+  ungroup() %>% 
+  mutate(targeted = ifelse(targeted == 1,"yes","no")) %>% 
+  spread(targeted, n)
+
+```
+
+# Supporting Information (SI)
+
+# SI Text {#si-text .unnumbered}
+
+## Computing environment
+
+All code needed to reproduce our main results and manuscript can be found at https://github.com/DanOvando/regional-effects-of-mpas.
+All analysis were performed in `r R.version.string`.
+Package versions are shown in Table.S\@ref(tab:pkg-info).
+
+```{r pkg-info, results = "asis", eval=TRUE}
+table <- devtools::package_info() %>% 
+  select(package, loadedversion, date,source)
+
+knitr::kable(table,
+             digits = 2,
+             caption = "Package versions and sources used in this paper",
+             longtable = TRUE,
+             booktabs = TRUE,
+             format = "latex",
+             col.names = colnames(table)) %>% 
+  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
+
+```
+
+## PISCO Data
+
+All fish data used in the primary difference-in-difference model were collected by PISCO. The dive transect survey methods are described in @caselle2015, provided below for ease of reference
+
+"Fish assemblages were surveyed annually as part of a long-term monitoring program conducted by the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) using standard underwater visual belt survey methods (www.piscoweb.org9).
+We analyzed data from 47 PISCO sites at the northern Channel Islands that were sampled annually from at least 2003 to 2012.
+We also excluded 3 sites on Anacapa where a much older MPA had already been established in 1978.
+MPAs on each island were sampled annually during June-October and we surveyed multiple sites inside and outside of any individual MPA.
+Details of MPA characteristics such as size and coastline extent are given in Hamilton et al.9.
+At each site, we conducted 8 to 12 fish transects that measured 30×2×2m at multiple levels in the water column: benthic, midwater, and kelp canopy (when present).
+Transects are laid out in a stratified random design, with multiple nonpermanent transects located in fixed strata (i.e., outer, middle, and inner edges of the reef).
+At each level in the water column, one SCUBA diver per transect counted and estimated the sizes of all fish to the nearest centimeter (total length), excluding small cryptic fishes"
+- @caselle2015
+
+We take a number of steps to translate the raw transect data into the total biomass densities used in this study.
+For the default run, we only include species that were observed at least once for at least 15 of the 18 years of available data.
+We also exclude "young of the year" observations due the challenges of correctly identifying and measuring these individuals.
+We omitted data from 1999 due to changes in the sampling procedures that occurred after 1999.
+Per recommendations from PISCO staff we omit observations from the canopy level of the transects (leaving the middle, bottom, and middle canopy levels).
+
+PISCO data report positive observations of fish.
+In order to use these data in our model we need to add in zeros for any transect that could have observed a given species of fish but did not.
+We assume that a fish could have been observed on a given transect if that species has ever been observed at that site in any time period in the data (PISCO data are organized by sites, with multiple transect at different locations within the borders of a site).
+If a species has never been observed at a site we assume that it does not occur at that site.
+
+Once zeros have been introduced to the database, we convert positive observations of fish from numbers to biomass.
+Each observation in the raw database lists the species, the number of individuals seen, and the size of those individuals (either as one value or as a mininum and maximum size for the group seen).
+PISCO staff compiled allometric information used to convert lengths to expected weights.
+For each observation then, we convert the observed lengths to weights per these relationships (accounting for variations in length types such as standard vs.
+total length).
+When minimum and maximum ranges were reported, we drew a number of samples equal to the number of observed fish in that group from a uniform distribution spanning the minimum and maximum reported size in that group.
+We assume all length-to-weight conversions are constant and deterministic.
+
+For each species at each transect, we then calculate the biomass density of that transect as the sum of the observed biomass divided by the transect area.
+We then average the biomass densities for each species across all the transects at a given site, and lastly sum these mean species biomass densities to achieve the total mean species biomass at the site level.
+
+We include several additional sources of data in our regression analysis.
+Temperature readings are included from the PISCO data for each transect.
+We also include PISCO data on the estimated surge and visibility.
+We augmented these data with information on kelp cover over time from the Santa Barbara Channel Long Term Ecological Research Network [@lter2017].
+We used a k-nearest neighbors algorithm to fill in missing kelp observations, and matched the interpolated kelp data to the PISCO data at the resolution of year-month-site (Fig.\@ref(fig:kelp-plot)).
+We include a variable capturing the mean cumulative number of observations across all observers conducting transects, in an effort to control for evoloving observer skill.
+
+<!-- Temperature data were augmented with data from `FishLife` [@thorson2017c] to include the estimated preferred temperature for a given species, so that we can include deviations from the preferred temperature envelope as a predictor in the model.
+This allows different temperatures to have different effects on each species (and is less computationally intensive than estimating species-temperature slopes) (Fig.\@ref(fig:tempdev-plot)).
+-->
+
+We also included lagged catch totals in the Santa Barbara region for the commercially harvest species in the database, in an effort to control for changes in density caused by changes in fishing pressure.
+Catches were pulled from the CDFW website (https://www.wildlife.ca.gov/Fishing/Commercial/Landings), and extracted using the `tabulizer` package in R [@leeper2018] (Fig.\@ref(fig:catches-plot)).
+
+```{r catches-plot, fig.cap = "Total CDFW reported commercial catches in the Santa Barbara region"}
+
+used_data %>% 
+  filter(targeted == 1) %>% 
+  group_by(commonname, year) %>% 
+  summarise(catches = unique(lag_catch)) %>% 
+  group_by(commonname) %>% 
+  mutate(total_catches = sum(catches)) %>% 
+  ungroup() %>% 
+  filter(total_catches > 0) %>% 
+  ggplot(aes(year, catches)) + 
+  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
+  geom_line() + 
+  facet_wrap(~commonname, scales = "free_y") + 
+  theme_minimal() + 
+  labs(x = "Year", y = "Catch (lbs)")
+
+```
+
+```{r kelp-plot, fig.cap="Mean kelp biomass by island over time from SBC LTER"}
+
+used_data %>% 
+  group_by(region,year) %>% 
+  summarise(mean_kelp = mean(interp_kelp, na.rm = TRUE)) %>% 
+  ggplot(aes(year, (mean_kelp), color = region)) + 
+  geom_line() + 
+  labs(x = "Year", y = "Mean Kelp Biomass")
+
+```
+
+```{r targ-trend-plot, fig.cap="Centered and scaled mean biomass densities of all targeted finfish in analysis before filtering"}
+
+targeted_trend_plot <- pisco_data %>% 
+  filter(targeted == 1) %>% 
+  group_by(year, classcode, targeted, commonname) %>% 
+  summarise(mean_biomass_density = mean(density_g_m2, na.rm = TRUE)) %>% 
+  group_by(classcode) %>% 
+  mutate(mean_biomass_density = scale(mean_biomass_density)) %>% 
+  ggplot(aes(year,mean_biomass_density)) + 
+  geom_hline(aes(yintercept = 0), linetype = 2) + 
+  geom_vline(aes(xintercept = 2003), color = "red") +
+  geom_line() + 
+  facet_wrap(~commonname) +
+  theme_minimal() + 
+  theme(strip.text = element_text(size = 6),
+        axis.text.x = element_text(size = 8)) + 
+  labs(x = "Year", y = "Centered and Scaled Mean Biomass Density")
+
+targeted_trend_plot
+```
+
+```{r nontarg-trend-plot, fig.cap="Centered and scaled mean biomass densities of all non-targeted finfish in analysis before filtering"}
+
+nontargeted_trend_plot <- pisco_data %>% 
+  filter(targeted == 0) %>% 
+  group_by(year, classcode, targeted, commonname) %>% 
+  summarise(mean_biomass_density = mean(density_g_m2, na.rm = TRUE)) %>% 
+  group_by(classcode) %>% 
+  mutate(mean_biomass_density = scale(mean_biomass_density)) %>% 
+  ggplot(aes(year,mean_biomass_density)) + 
+  geom_hline(aes(yintercept = 0), linetype = 2) + 
+  geom_vline(aes(xintercept = 2003), color = "red") +
+  geom_line() + 
+  facet_wrap(~commonname) +
+  theme_minimal() + 
+  theme(strip.text = element_text(size = 6),
+        axis.text.x = element_text(size = 8)) + 
+  labs(x = "Year", y = "Centered and Scaled Mean Biomass Density")
+
+nontargeted_trend_plot
+```
+
+```{r classcode, eval = TRUE, results = "asis"}
+
+species_data <- used_data %>% 
+  select(classcode, commonname, taxa, targeted) %>% 
+  unique() %>% 
+  arrange(desc(targeted)) %>% 
+  mutate(targeted = targeted == 1) %>% 
+  rename(`Common Name` = commonname, 
+         `Scientific Name` = taxa,
+         `Targeted?` = targeted)
+
+status <- read_csv(here::here("data","stock-status-summaries.csv")) %>% 
+  select(classcode, `Stock Status`)
+
+species_data <- species_data %>% 
+  left_join(status, by = "classcode")
+
+knitr::kable(
+  species_data,
+  digits = 2,
+  caption = "Species included in estimation model",
+  longtable = TRUE,
+  booktabs = TRUE,
+  col.names = colnames(species_data)
+) %>%
+  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
+
+```
+
+## Difference-in-Difference Model
+
+The difference-in-difference (DiD) regression amounts to estimating the pre-post MPA difference in the biomass densities of targeted species minus the same difference for non-targeted species in the Channel Islands.
+
+The simplified form of this model is
+
+\begin{equation}
+  d_{i} \sim Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + \mathbf{B^cX_i} + \mathbf{B^sS_i}},shape, scale)
+\label{eq:did}
+\end{equation}
+
+To provide greater detail on this process, we first conduct the data pre-processing described earlier.
+From there, we aggregated data to the level of total biomass density of targeted and non-targeted species at each site each year.
+As such, our base model estiamtes the effect of the MPAs on total biomass density of targeted species, though we also explore the effect on alternative specifications such as mean biomass of targeted species.
+the model was then fit using the `rstanarm` package in R (5000 iterations, 2500 warmup, 4 chains, adapt_delta = 0.85).
+
+The intercept prior was was determined using the `rstanarm` `autoscale` function.
+For the non-intercept terms, we manually set a normal(0,2) prior.
+This implies that coefficeints such as the MPA effect have a prior that provides support for an effect size centered on zero with a range of roughly -4 to 4.
+Since the covariates are centered and scaled, and the dependent variable is on the log-scale, this is an extremely diffuse prior.
+
+The full table of covariates and their posterior means and 89% credible interval can be found in Table.S\@ref(tab:dids).
+
+```{r dids, results = "asis",eval = TRUE}
+
+did_fit %>%
+  broom::tidy(intervals = TRUE, prob = 0.89) %>%
+  select(term, estimate, lower, upper) %>%
+  rename(
+    mean = estimate,
+    lower_89th_percentile = lower,
+    upper_89th_percentile = upper
+  ) %>%
+  knitr::kable(
+    digits = 2,
+    caption = "Posterior means and credible interval for key model coefficients",
+    longtable = TRUE,
+    booktabs = TRUE ) %>%
+  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
+
+```

### 2020-04-22 documents/ovando-regional-effects-of-mpas-si-v2.Rmd
pre changing to population language commit

@@
 ---
-title: "Supporting Information: The Regional Effects of Marine Protected Areas"
-author:  Daniel Ovando,Jennifer E. Caselle,Christopher Costello,Olivier Deschenes,Steven D. Gaines,Ray Hilborn,Owen Liu
+title: "Supporting Information: Population Effects of Marine Protected Areas Are Likely To Be Unclear"
+author: 
+  - Daniel Ovando
+  - Jennifer E. Caselle
+  - Christopher Costello
+  - Olivier Deschenes
+  - Steven D. Gaines
+  - Ray Hilborn
+  - Owen Liu
+date: "`r Sys.Date()`"
 output: 
@@
     - \usepackage{xcolor}
-    - \linenumbers
 params:
@@
 All code needed to reproduce our main results and manuscript can be found at https://github.com/DanOvando/regional-effects-of-mpas.
-A fully reproducible environment for running this analysis and compiling the manuscript will be made available through Code Ocean at https://codeocean.com/capsule/5233105/tree.
 All analysis were performed in `r R.version.string`.
@@
 
-```{r pkg-info, results = "asis", eval=FALSE}
+```{r pkg-info, results = "asis", eval=TRUE}
 table <- devtools::package_info() %>% 
@@
 
-PISCO data report positive observations of fish.
+PISCO data report positive observations of fish, in the manner of number of individuals of a species seen within a particular size group.
 In order to use these data in our model we need to add in zeros for any transect that could have observed a given species of fish but did not.
@@
 
-For each species at each transect, we then calculate the biomass density of that transect as the sum of the observed biomass divided by the transect area.
+For each species at each transect, we calculate the biomass density of that transect as the sum of the observed biomass divided by the transect area.
 We then average the biomass densities for each species across all the transects at a given site, and lastly sum these mean species biomass densities to achieve the total mean species biomass at the site level.
@@
 
+The full list of included coefficients and their estimated values can be seen in Table.\@ref(tab:did-coefs).
+
 To provide greater detail on this process, we first conduct the data pre-processing described earlier.
-From there, we aggregated data to the level of total biomass density of targeted and non-targeted species at each site each year.
-As such, our base model estiamtes the effect of the MPAs on total biomass density of targeted species, though we also explore the effect on alternative specifications such as mean biomass of targeted species.
+From there, we aggregated data to the level of total mean biomass density of targeted and non-targeted species at each site each year.
+As such, our base model estiamtes the effect of the MPAs on total mean biomass density of targeted species, though we also explore the effect on alternative specifications such as mean biomass of targeted species.
 the model was then fit using the `rstanarm` package in R (5000 iterations, 2500 warmup, 4 chains, adapt_delta = 0.85).
@@
 This implies that coefficeints such as the MPA effect have a prior that provides support for an effect size centered on zero with a range of roughly -4 to 4.
-Since the covariates are centered and scaled, and the dependent variable is on the log-scale, this is an extremely diffuse prior.
+Since the covariates are centered and scaled, and the dependent variable is on the log-scale, this is an extremely diffuse prior (e.g.
+a one standard deviation change in an independent variable changes the dependent variable by 4 on the log scale).
 
 The full table of covariates and their posterior means and 89% credible interval can be found in Table.S\@ref(tab:did-coefs).
-
-```{r did-coefs, results = "asis",eval = FALSE}
-
+"var" refers to variable, "_2" indicates a squared term.
+tex is the mean cumulative number of observations of the observers collecting the raw data, surge is the mean reported surge across transects, kelp is the mean kelp coverage at the site, temp is the water temperature, regional_temp_dev is the water temperature at a given site scaled by the mean water temperature at that site.
+Targeted is a boolean marker indicating whether the given observation is for targeted or non-targeted species.
+
+```{r did-coefs, results = "asis",eval = TRUE}
 did_fit %>%
@@
   select(term, estimate, lower, upper) %>%
-  rename(mean = estimate, lower_89th_percentile = lower, upper_89th_percentile= upper) %>% 
+  rename(
+    mean = estimate,
+    lower_89th_ci = lower,
+    upper_89th_ci = upper
+  ) %>%
   knitr::kable(
     digits = 2,
-            caption = "Posterior means and 89% credible interval for key model coefficients",
+    caption = "Posterior means and credible interval for key model coefficients",
     longtable = TRUE,
-             booktabs = TRUE) %>%
+    booktabs = TRUE ) %>%
   kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
+
 ```
@@
                  yrep = yrep) +
-  scale_x_continuous(name = "Log(Total Biomass Density)") +
+  scale_x_continuous(name = "Total Biomass Density") +
   scale_colour_discrete(labels = c("Observed", "Posterior Predictive"),
@@
 
-```{r mean_test_plot, fig.cap = "Posterior predictive mean (distribution) and empirical mean of total biomass density. Posterior predictive mean adequately recovers empirical mean."}
+```{r mean_test_plot, fig.cap = "Posterior predictive mean (distribution) and empirical mean of total biomass density. Posterior predictive mean adequately covers empirical mean."}
 
@@
 
-```{r sum_test_plot, fig.cap = "Posterior predictive sum (distribution) and empirical sum of total biomass density. Posterior predictive sum adequately recovers empirical sum."}
+```{r sum_test_plot, fig.cap = "Posterior predictive sum (distribution) and empirical sum of total biomass density. Posterior predictive sum adequately covers empirical sum."}
 
@@
 ```{r sd_test_plot, fig.cap = "Posterior predictive standard deviation (distribution) and empirical standard deviation of total biomass density.
-Posterior predictive standard deviation adequately recovers empirical standard deviation"}
+Posterior predictive standard deviation adequately covers empirical standard deviation"}
 ppc_stat(y,  yrep, stat = "sd")
@@
 
-```{r, fig.cap="Model fit to PISCO data but only using variables also available for KFM data", results="hide"}
+We include one run using only variables availale in both datasets, to explore sensitivity of our results to selected covariates (Fig.\@ref(fig:kfm-data)).
+
+```{r kfm-data, fig.cap="Model fit to PISCO data but only using variables also available for KFM data", results="hide"}
 
@@
 
-```{r, fig.cap="Model fit to KFM data. Results are more uncertain than using PISCO data, but follow the same trend estimated from the PISCO data."}
+We fit an alternative model using data from the KFM program to test whether a different dataset from the same region provides different results.
+KFM based results are qualitatively similer to those from PISCO data (Fig.\@ref(fig:kfm-fit))
+
+```{r kfm-fit, fig.cap="Model fit to KFM data. Results are more uncertain than using PISCO data, but follow the same trend estimated from the PISCO data."}
 
@@
 
+We ran a version of the model using data from only inside MPAs.
+This serves as a test of whether our results are being entirely driven by the outside-MPA data, and allows us to compare results from inside MPAs to those outside (Fig.\@ref(fig:mpa-fit-plot)).
+
 ```{r mpa-fit-plot, fig.cap="Estimated MPA effects using PISCO data from inside MPAs only.
@@
 
-```{r, fig.cap="Estimated MPA effects using PISCO data from outside MPAs only.
+We ran a version of the model using data from only outside MPAs.
+This serves as a test of whether our results are being entirely driven by the inside-MPA data, and allows us to compare results from inside MPAs to those outside (Fig.\@ref(fig:fished-fit-plot)).
+
+```{r fished-fit-plot, fig.cap="Estimated MPA effects using PISCO data from outside MPAs only.
 The base model uses data from both inside and outside MPAs.
@@
 ```
+
+We include a model run using nearly all observed species, to test whether our results are being influenced by omissino of rarely observed species (Fig.\@ref(fig:no-filter-plot)).
 
@@
 ```
+
+For ease of interpretation and model convergence, our default model estimates MPA effects in three-year bins.
+For this run we estimate the effect annually, relative to the year 2000.
+Results show the same overall trend and magnitude reported in our base model (Fig.\@ref(fig:no-bin-plot)).
 
@@
 
+Our base run estimates the effect of MPAs on total biomass densities of targeted finfish.
+For this run, we instead estimate the effect on mean biomass densities of targeted species.
+Results show the same pattern as our base run (Fig.\@ref(fig:mean-fit)).
+
 ```{r mean-fit, fig.cap="Our base run estimates the effect of MPAs on total biomass densities of targeted finfish.
 For this run, we instead estimate the effect on mean biomass densities of targeted species.
-Results show the same pattern as our base run.
-"}
+Results show the same pattern as our base run.", results="hide"}
 mean_fit <- model_runs %>%
@@
 For this run, we instead estimate the effect on mean biomass densities of targeted species.
-Results show the same pattern as our base run.
-"}
+Results show the same pattern as our base run.", results = "hide"}
 
@@
 ```
- selection on observables
-
-```{r}
-
-```
 
@@
 
-```{r betatable}
+```{r betatable, results="asis"}
 
@@
 As depletion increases, relatively small changes in total biomass (relative to the variance in the observation process) can translate into large percent changes in biomass.
-For example, moving from a density of .2kg/m^2^ to .4kg/m^2^ translates to a 100% percent increase, but only a .2kg/m^2^ absolute increase, a small value to detect with a real observation program.
+For example, moving from a density of 0.02kg/m^2^ to 0.04kg/m^2^ translates to a 100% percent increase, but only a .02kg/m^2^ absolute increase, a small value to detect with a real observation program.
 

### 2020-04-22 documents/ovando-regional-effects-of-mpas-v2.Rmd
pre changing to population language commit

@@
 ---
-title: Unclear Regional Effects of a Large Marine Protected Area Network
+title: Population Effects of Marine Protected Areas Are Likely To Be Unclear
 date: "`r Sys.Date()`"
@@
 We use a bio-economic simulation model to explain this discrepancy, and demonstrate under what conditions we likely can and cannot clearly estimate the regional conservation effects of MPAs.
-MPA networks covering 25\% or less of a region are likely to produce regional increases in fish biomass densities on the order of 25\% or less, a meaningful effect but also one that can easily be overwhelmed by environmental shocks.
-Our results provide a novel assessment of the regional effects of a large and iconic Marine Protected Area network, and provide guidance for managers charged with monitoring and adapting MPAs.
+We show that MPA effect sizes less than 30% are likely to be difficult to detect, and the size of many MPA networks suggests that effect sizes may often be smaller than 30%.
+Our results provide a novel assessment of the regional effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.
 bibliography: My Library.bib
@@
 
-# sim_years <- 50
+# sim_years <- 50x
 # 
@@
   Healthy marine ecosystems are critical to the well-being of the planet.
-Marine protected areas (MPAs), parts of the oceans protected from human activities such as fishing, are increasingly being used to conserve and manage these ecosystems.
+Marine protected areas (MPAs) are increasingly being used to conserve and manage these ecosystems.
 Our study pairs theory with empirical methods to show that the regional conservation effects of marine protected areas can be highly variable and dependent on human drivers.
@@
 
-Recent agreements to expand MPAs (most proximately, the Convention on Biological Diversity’s Strategic Plan for Biodiversity, which calls for 10% of coastal waters to be protected inside MPAs by 2020 and the International Union for Conservation of Nature call for 30% by 2030) are based on the expectation that well-designed MPAs will achieve benefits both within, and outside, their borders.
-Despite these assumptions, our collective scientific understanding of the regional-scale conservation and fishery impacts of current and future MPAs is surprisingly limited.
+Recent international agreements to expand MPAs (most proximately, the Convention on Biological Diversity’s Strategic Plan for Biodiversity, which calls for 10% of coastal waters to be protected inside MPAs by 2020 and the International Union for Conservation of Nature call for 30% by 2030) are based on the expectation that well-designed MPAs will achieve benefits both within, and outside, their borders.
+Indeed, Numerous studies provide evidence that well-enforced and appropriately sized MPAs can produce conservation benefits within their borders [@lester2009;@halpern2003; @edgar2014; @gerber2005].
+However, our collective scientific understanding of the regional-scale conservation and fishery impacts of current and future MPAs is surprisingly limited.
 
-Numerous studies provide evidence that well-enforced and appropriately sized MPAs can produce conservation benefits within their borders [@lester2009;@halpern2003; @edgar2014; @gerber2005].
-As these conservation benefits accrue inside MPAs, theory holds that MPAs can affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
+As conservation benefits accrue inside MPAs, theory holds that MPAs can affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
 Several studies have documented empirical evidence for the existence of adult or larval spillover affecting both abundance and fisheries [e.g.
 @goni2010; @halpern2009;@kay2012; @stobart2009; @mcclanahan2000; @russ1996; @thompson2017;@baetscher2019;@pelc2009], as well as alteration of fishing effort in reaction to [@costa2013;@mason2012;@murawski2005] and in anticipation of [@mcdermott2019] MPA placement.
-The potentially more important question, however, is not whether spillover occurs (it must to some degree in any realistic scenario), but what the net effects of spillover are and whether those effects are empirically detectable.
+However, empirically detecting the regional net effects of spillover on fish populations and fisheries is much more difficult.
 From a fishery perspective, are spillover benefits sufficient to offset losses in fishing grounds and changes in responses of displaced fishers caused by an MPA?
@@
 This is particularly true in the context of a changing climate, in which "set it and forget it" management strategies are increasingly untenable; we need to be able to monitor the performance of MPAs in order to adapt them as dictated by a shifting environment.
-The MPAs within the Channel Islands National Marine Sanctuary, California, USA (which we will refer to as the Channel Islands from now on) provide an ideal case study to address this need.
+The MPAs within the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) provide an ideal case study to address this need.
 A network of protected areas covering approximately 20% of the Islands' waters was put in place in 2003 [see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of these MPAs].
@@
 
-We use data from the first 14 years of protection to provide what is to our knowledge the first empirical assessment of the regional effect (defined here as the effect of MPAs on the mean biomass density of targeted finfish both inside and outside of MPAs) of a large MPA network on a wide array of finfish species.
+We use data from the first 14 years of protection to provide what is to our knowledge the first empirical assessment of the regional effect of a large MPA network on a wide array of finfish species.
+For the purpose of this study, we define a regional effect as the effect of MPAs on the mean biomass density of targeted finfish both inside and outside of MPAs.
 In contrast to clear differences in biomass densities observed inside and outside of well protected MPAs both globally [@lester2009] and in the Channel Islands [@caselle2015] we are unable to detect a clear regional effect from the Channel Islands MPAs.
@@
 The organisms within the borders of protected areas are generally part of a broader biological stock, connected through adult movement and larval dispersal.
-If the goal of conservationists or natural resource managers is to increase the total abundance or productivity of a resource, a broader question we should ask of MPAs is not just are there more fish inside their borders, but also how have reserves affected populations throughout the region in which they are located?
-This logic, that MPAs will have conservation benefits for most species beyond their borders, is implicit in all multilateral calls for MPA expansion.
+The assumption that MPAs will have conservation benefits beyond their borders for most species is implicit in all multilateral calls for MPA expansion, but these benefits have not been widely measured.
 
-We define the regional conservation effects of MPAs as the change in mean total biomass densities of targeted finfish both inside and outside of MPAs, relative to the mean total biomass densities of targeted finfish inside and outside of MPAs that would have occurred without the MPAs (acknowledging that other outcomes such as increased biodiversity or resiliency are also important to conservation but are beyond the scope of this analysis).
-Consider a evenly distributed population that has 50% of its range protected by an MPA.
+We define the regional conservation effects of MPAs as the change in total mean biomass densities of targeted finfish both inside and outside of MPAs, relative to the total mean biomass densities of targeted finfish inside and outside of MPAs that would have occurred without the MPAs.
+While other outcomes such as increased biodiversity or resiliency are also important to conservation, these are beyond the scope of this analysis).
+As an example of a regional MPA effect, consider a evenly distributed population that has 50% of its range protected by an MPA.
 Suppose that the MPA increase biomass densities inside the reserves by 20%, and by 0% outside the reserve.
@@
 
-The MPAs in this study cover 20% of the the surface waters of the Channel Islands, a region spanning roughly 90km east to west.
+The MPAs in this study cover 20% of the surface waters of the Channel Islands, a region spanning roughly 90km east to west.
 Detailed dispersal studies are not available for all of the species covered by our study, but what information we have suggest that while adults of some of these species are likely to exhibit site fidelity [@love2011], larvae are likely distributed beyond the Channel Islands [@rassweiler2012].
@@
 Formal stock assessments are largely lacking for these species.
-However, what evidence we have suggests mostly moderate fishing pressure, with some species such as California sheephead (*Semicossyphus pulcher*) and blue rockfish (*Sebastes mystinus*) experiencing high levels of fishing mortality / biomass  levels below target levels during the early 2000s [@alonzo2004;@dick2017].
+However, what evidence we have suggests mostly moderate fishing pressure, with some species such as California sheephead (*Semicossyphus pulcher*) and blue rockfish (*Sebastes mystinus*) experiencing high levels of fishing mortality, likely resulting in biomass levels below management targets during the early 2000s [@alonzo2004;@dick2017].
 
-We integrated these network design features, life history traits, and exploitation histories into a spatially explicit bio-economic simulation model, and used this model to generate expected outcomes for the Channel Islands MPAs, varying key unknowns such as the degree of pre-MPA depletion, movement rates, and fleet dynamics.
-While a wide range of outcomes are plausible, from 0% to upwards of 200%, the a median simulated effect size for a stylized version of the Channel Islands after over ten years of protection was roughly `r percent(median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected > 10]))`.
+We integrated these network design features, life history traits, and exploitation histories into a spatially explicit bio-economic simulation model, and used this model to generate expected outcomes for the Channel Islands MPAs, varying key unknowns such as the degree of pre-MPA depletion, species' movement rates, and fleet dynamics.
+While a wide range of MPA effects are plausible, from 0% to upwards of 200%, the median simulated effect size for a stylized version of the Channel Islands after over ten years of protection was roughly `r percent(median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected > 10]))` (Fig.\@ref(fig:mpa-effect)).
 
@@
 
-Why do these response ratio results not serve as sufficient evidence for the regional effects of the Channel Island MPAs?
+Unfortunately, these response ratios do not provide sufficient evidence for significant regional-scale effects of the Channel Islands MPAs.
 Control sites used in calculating response ratios are often selected based on abiotic or ecological traits such as habitat characteristics [@ferraro2018].
-However, selection of control sites is further complicated by the very spillover that MPAs are often intended to create.
+Selection of these sites is complicated, though, by the very spillover that MPAs are often intended to create.
 Export of adults or larvae from the MPA to the "control" site affects their status as controls, as does displacement of fishing effort from MPAs to control sites.
@@
 
-In the case of the Channel Islands MPAs, control sites are often located within a few kilometers of an MPA, suggesting that they are susceptible to biological spillover and concentration of fishing effort resulting from the MPAs.
+In the case of the Channel Islands MPAs, control sites are often located within a few kilometers of an MPA, suggesting that they are susceptible to both local biological spillover and by concentration of fishing effort excluded from the MPAs.
 Given these complications, we can interpret the response ratios reported in @caselle2015, and updated here, as evidence that the MPAs are providing effective protection of targeted fish biomass inside their borders, particularly in light of the general increase in response ratios over time.
-However, it is unclear how well these response ratio results serve as in indicator of overall regional effects of the MPAs.
+But, we cannot use these response ratio results serve as a definitive indicator of overall regional effects of the MPAs.
 
@@
 These simulations cover the life-histories and MPA sizes seen in the Channel Islands, but vary in key unknowns such as the actual degree of fishing pressure, the timing of density dependence, and the fishing fleets response to the MPAs.
-For each year of protection, we paired our simulated response ratios to the estimated posterior probability distributions of the Channel Island response ratios.
-For example, if the mean estimated response ratio in the year 2006 is one, we found all simulations with three years of protection that had simulated response ratios near one, and then pulled out the "true" simulated regional MPA effect from each of those simulations.
+For each year of protection, we paired our simulated response ratios to the estimated posterior probability distributions of the  empirical Channel Island response ratios.
+For example, if the mean estimated response ratio in the year 2006 (year 3 after implementation) is one, we found all simulations with three years of protection that had simulated response ratios near one, and then pulled out the "true" simulated regional MPA effect from each of those simulations.
 This provides us with a distribution of simulated regional MPA effects that could plausibly generate the types of response ratios actually observed in the Channel Islands.
@@
 The response ratio trends we observe in the data could plausibly be produced by a wide range of regional MPA effects (Fig.\@ref(fig:rr-plot)).
-Response ratios well over one were associated with regional MPA effects generally less than 25%, and many simulations produced large response ratios but regional MPA effects close to 0% This can occur if for example fishing pressure is only moderate, adult movement is low, larval dispersal is high, and displaced fishing effort concentrates around the border of the MPAs.
+Response ratios well over one were associated with regional MPA effects generally less than 25%, and, importantly, many simulations produced large response ratios but regional MPA effects close to 0% This can occur if for example fishing pressure is only moderate, adult movement is low, larval dispersal is high, and displaced fishing effort concentrates around the border of the MPAs.
 The response ratio results previously published for this region do not serve as sufficient evidence for the regional conservation effects of the Channel Islands MPA network.
@@
 ```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="90% Posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis) in blue.
-Simulated MPA effects associated with simulated response ratios matched to empirical response ratios in red.
+Simulated MPA effect on biomass densities matched to empirical response ratios in red.
 For response ratios, a value of zero indicates that biomass densities of targeted and non-targeted species are identical inside and outside MPAs, a value of 1 indicates that biomass densities of targeted species are 100% greater inside MPAs than outside.
@@
 It should be noted though that our analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*).
-Included non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species).
-We used a Bayesian difference-in-difference regression to estimate any difference in mean total biomass densities of fin-fish species targeted by fishing effort (i.e., those potentially affected by an MPA) and those species not targeted by fishing before and after MPA implementation [see @larsen2019 for an explanation of difference-in-difference estimators in ecological settings].
-The result of this regression is, conditional on the assumptions of the model, an estimate of the effect of the MPAs on the mean total biomass densities of targeted species throughout the Channel Islands.
+Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species).
+We used a Bayesian difference-in-difference regression to estimate any difference in total mean biomass densities of fin-fish species targeted by fishing effort (i.e., those potentially affected by an MPA) and those species not targeted by fishing before and after MPA implementation [see @larsen2019 for an explanation of difference-in-difference estimators in ecological settings].
+The result of this regression is, conditional on the assumptions of the model, an estimate of the effect of the MPAs on the total mean biomass densities of targeted species throughout the Channel Islands.
 
@@
 - `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`).
-Over the next six years the model estimates greater probabilities of an increasingly positive MPA effect, peaking in 2009-2011 with a median estimate of MPA effect of a `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)` increase in mean total biomass density of targeted species (90% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)`
+Over the next six years the model estimates greater probabilities of an increasingly positive MPA effect, peaking in 2009-2011 with a median estimate of MPA effect of a `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)` increase in total mean biomass density of targeted species (90% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)`
 - `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)`.
@@
 
-```{r mpa-effect, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Results of difference-in-difference regression estimating the regional effect of the Channel Island MPAs on mean total biomass densities of targeted species (difference in mean total biomass density of targeted species over time relative to expected levels using non-targeted species as a control).
+```{r mpa-effect, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Results of difference-in-difference regression estimating the regional effect of the Channel Island MPAs on total mean biomass densities of targeted species (difference in total mean biomass density of targeted species over time relative to expected levels using non-targeted species as a control).
 Grey distributions show posterior probability distribution of estimated MPA effect; red point is median estimated effect, thicker red section 50% credible interval, thinner red line 90% credible interval).
@@
 One explanation may lie in fleet dynamics.
-Much of the theoretical literature on MPAs assumes that all else being equal bigger reserves produce bigger conservation gains [@white2011 and references therein].
+Much of the theoretical literature on MPAs assumes that all else being equal larger reserves produce larger conservation gains [@white2011 and references therein].
 However, these models generally simulate fleet dynamics through fishing mortality rates (e.g.
 concentration of fishing mortality @halpern2004).
-The assumption of these models is that fishers determine an amount off effort to exert, and distribute that effort outside the MPA in response to some function.
+The implication of this is that total catches will scale with the size of the available population: if fishing effort moves outside the reserve, total catch will go down if the pool of fishable biomass available to the fleet is lower outside the reserve.
 
@@
 
-An alternative and to our knowledge unexplored (in the context of MPA simulation) fleet model though is a "constant-catch" strategy.
-Under this model, fishers have a catch objective, and exert as much (or little) effort as needed to achieve that objective.
-While a constant-catch greater than MSY is not possible over the long-term under the assumptions of our model, over the short-term a constant-catch scenario is not implausible.
+Alternatively, though, fishers can pursue a  “constant-catch” rather than an effort-based strategy.
+This is an important and, to our knowledge, unexplored fleet model in MPA simulation.
+Under the constant catch model, fishers have a catch objective, and exert as much (or little) effort as needed to achieve that objective.
+While a constant-catch greater than the maximum sustainable catch from the population is not possible over the long-term under the assumptions of our model, over the short-term a constant-catch scenario is possible.
 Subsistence fisheries may use a constant-catch style policy over the short-term, as they seek to ensure that their food needs are met.
-More industrial fisheries may have pre-arranged agreements with buyers to deliver set amounts of fish.
+Industrial fisheries may have pre-arranged agreements with buyers to deliver set amounts of fish.
 Constant-catch dynamics might also occur in fisheries with constraining quotas that are not updated after the implementation of MPAs.
@@
 Under a constant-catch scenario though, fishers have to fish much harder than before to get the same catch from a smaller part of the population, reducing the size structure of the population and subsequently causing net conservation loss under `r percent(1 - positive_outcomes$pp[positive_outcomes$fleet_model == "constant-catch"], accuracy = 1)` of our constant-catch simulations.
-This is an important and often overlooked possibility, especially as MPAs are increasingly implemented in quota-managed fisheries.
+This is an important and often overlooked possibility, especially as MPAs are increasingly implemented in quota-managed fisheries [@liu2018].
 
-While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region in fact exhibit an overall downward trend in the years post reserve (see Fig.SXX).
-We can most likely rule out a negative MPA effect caused by a constant-catch fishing strategy then.
-What is another possible explanation for the recent downward trend in the estimated MPA effects?
+While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (see Fig.SXX) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
+Instead, another possibility for the observed decline in Channel Islands MPA effects is environmental disturbance.
 The Channel Islands region (and the entire West coast of the USA) experienced a dramatic ‘marine heatwave’ beginning in 2014 and persisting through 2016, resulting in part in extremely elevated water temperatures throughout the region  [@gentemann2017].
-Difference-in-difference models require a "parallel trends" assumption; in this case that in the absence of the MPAs targeted and non-targeted species would have exhibited similar trends in biomass densities.
-While this assumption cannot be proven, we find support for it in the initial years of the MPAs (Fig.SXX).
-
-Biogeographic differences in the distributions of targeted and non-targeted species may confound the observed effects of MPAs.
 Many of the non-targeted species in the Channel Islands have warm thermal affinities and have increased in numbers since the heatwave [@freedman2019].
@@
 We hypothesize that the recent evidence for a decline in densities of targeted species is due to environmental conditions that disproportionately affect the targeted group (and not for example due to concentrated fishing pressure outside the reserves).
-This hypothesis is supported by the declining trend in mean biomass densities of targeted species seen inside the MPAs themselves.
-If the cause of recent declines was due to increases in fishing pressure we would expect to see substantial declines only in the fished areas (Fig.\@ref(fig:raw-density)).
-The net result of this is that while we see evidence for the validity of the parallel trends assumption in the early years of our data, it is likely that the non-targeted species reacted in a different way to the marine heatwave than the targeted species did, causing a violation of the parallel trends assumption in the post 2014 data.
+The practical result of this is that in the presence of this marine heatwave the non-targeted species may no longer serve as an effective control for the evolution of biomass densities of tartegeted finfish in the absence of the MPAs (a violation of the "parallel trends" assumption required for Difference-in-difference models).
 
-```{r raw-density, fig.width = width_1col, fig.asp=asp_1col,fig.cap = "Centered and scaled trends in biomass densities of targeted and non-targeted finfish included in our stury.
-Top panel shows trends across all sites, with smaller background lines showing trends for each indivdual species.
+This hypothesis is supported by the fact that we estimate a similar decline in the estimated MPA effect both inside and outside of the MPAs (Fig.SXX) (as well as observe the decline in the raw data, Fig.\@ref(fig:raw-density)).
+If the cause of recent declines was due to increases in fishing pressure we would expect to see substantial declines only in the fished areas.
+While we see evidence for the validity of the parallel trends assumption in the early years of our data (Fig.\@ref(fig:mpa-effect), Fig.SXX), it is likely that the non-targeted species reacted in a different way to the marine heatwave than the targeted species did, causing a violation of the parallel trends assumption in the post 2014 data.
+
+```{r raw-density, fig.width = width_1col, fig.asp=asp_1col,fig.cap = "Centered and scaled trends in biomass densities of targeted and non-targeted finfish included in our study Top panel shows trends across all sites, with smaller background lines showing trends for each individual species.
 Bottom two panels show aggregate biomass density trends outside and inside MPAs"}
@@
 temp_plot <-
-  (total_trend_plot / (mpa_trend_plot)) &
-     theme_classic()  & theme(legend.position = "top", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), units = "lines"),
+  (total_trend_plot / (mpa_trend_plot)) & theme(legend.position = "top", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), units = "lines"),
                               panel.grid = element_blank()
@@
 
-Containing a carefully designed, well-enforced, and well-studied MPA network, The Channel Islands would seem at face value to be an ideal location to study the regional conservation effects of protected areas.
+Containing a carefully designed, well-enforced, and well-studied MPA network, the Channel Islands would seem at face value to be an ideal location to study the regional conservation effects of protected areas.
 The persistently high response ratios suggest that despite overall decreases in targeted biomass densities inside and outside MPAs, the MPAs may still be providing protection within their borders.
-But, as we have shown here these response ratios are not necessarily an indicator of regional effects.
+But, as we have shown here these response ratios are not necessarily an indicator of broader regional effects.
 The difference-in-difference strategy utilized here presents an alternative identification strategy, that while not without its own strict caveats presents some potential improvements over response-ratios as a means of estimating regional conservation effects.
-While we estimate a highly uncertain but overall positive effect at first, we are unable to detect a robust signal from 2012-2017.
-We believe that the dissapearance of increasing probabilities of positive MPA effects estimated by our model are likely driven by the marine heatwave experienced by the region beginning in 2014.
+While we estimate an uncertain but overall positive effect of the MPA network in its first few years, we are unable to detect a robust signal from 2012-2017.
+We believe that the disappearance of increasing probabilities of positive MPA effects estimated by our model are likely driven by the marine heatwave experienced by the region beginning in 2014.
 This likely violation of the parallel trends assumption means that post 2014 estimates of MPA effects may be unreliable.
@@
 
-<!-- As an additional complication, our simulations and difference-in-difference model examine percentage changes in biomass densities (as oppose to changes in total population size).
+Our simulations and difference-in-difference model examine percentage changes in biomass densities (as oppose to changes in total population size).
 An increase in biomass densities from 0.02kg/m^2^ to 0.04kg/m^2^ would be produce as a 100% increase.
-While a large percentage effect, this is a small change in biomass densities relative to the variance in the observation process itself (see Fig.S1 for a companion to Fig.\ref{fig:sim-effects} scaled by absolute population size).
+While a large percentage effect, this is a small change in biomass densities relative to the variance in the observation process itself.
+While an MPA providing 25% protection might be expected to produce a 30% increase in biomass densities, expressed as a percentage of unfished biomass this gain is on the order of 15% (Fig.SXX).
 In addition, the median estimated age at sexual maturity for the targeted species included in this study is `r median_tm` years, meaning that the span of this analysis represents less than three generations of MPA protection for half of the the measured species.
@@
 spiny lobster, may also reveal clearer signals.
--->
 
@@
 
-```{r val-plot, include = TRUE,fig.width = width_1col, fig.asp=asp_1col, fig.cap = "Distribution of percent error in posterior estimate of MPA effect (y-axis) plotted against true simulated MPA effect (x-axis).
+```{r val-plot, include = TRUE,fig.width = width_1col, fig.asp=asp_1col, fig.cap = "Distribution of percent error in posterior estimates of MPA effect (y-axis) plotted against true simulated MPA effect (x-axis).
 Color shows concentration of simulations.
@@
 
-```{r sim-effects, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Simulated MPA effect sizes as a function of percent of species' range inside MPA, and pre-MPA depletion.
+```{r sim-effects, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Simulated MPA effect sizes as a function of percent of species' range inside MPA (x-axis), and pre-MPA depletion (y-axis).
 Pre-MPA depletion is a measure of fishing pressure, where 0 means that the population is unfished, and 1 means that the population is extinct in the time period immediately prior to MPA implementation.
-A) shows median MPA effects across range in MPA and pre-mpa depletin.
-Panel B shows distribution of simulations across range in MPA and pre-mpa depletion separately."}
+Panel A) shows median MPA effects across range in MPA and pre-MPA depletion Panel B) shows distribution of simulations across range of MPA size and pre-MPA depletion separately."}
 
@@
 
-What do these results imply for the evaluation and management of MPAs around the globe?
 A large theoretical literature has pointed out the importance of a wide range of economic and environmental factors in determining the effectiveness of MPAs [e.g.
 @hilborn2004; @smith2006; @white2010c].
-Our simulation results allow users to place this collective theory into context, clarifying under what states of the world we might expect to see particular effect sizes, and demonstrating what effect sizes we might reasonably expect to be detectable in the real world.
+Our simulation model allows users to place this collective theory into context, clarifying under what states of the world we might expect to see particular effect sizes, and demonstrating what effect sizes we might reasonably expect to be detectable in the real world.
 In the case of the Channel Islands, simulation modeling suggests for the network in question, an effect size on the order of 30% or smaller is likely, which our analysis in turn demonstrates may be challenging to detect.
@@
 As @larsen2019 suggests, there are many potential alternatives for estimating the effects of MPAs that better account for the challenges of causal inference (though that may be more data-intensive).
-We applied one such approach here (a difference-in-difference estimator), and yet were still unable to reach robust conclusions as to the effect of MPAs on the biomass density of targeted finfish in the Channel Islands, due to the likely small size of the true effect relative to the strength and variability of environmental drivers.
+We applied one such approach here (a difference-in-difference estimator), and yet were still unable to reach robust conclusions as to the effect of MPAs on the biomass density of targeted finfish in the Channel Islands, due to the likely small size of the true effect relative to the influence of environmental variability.
 If pre-MPA data are available, data currently used in response ratios can easily be re-purposed for an analysis such as this one; we encourage holders of such data worldwide to conduct similar exercises to examine to what extent empirical estimates of regional conservation effects match up with MPA theory.
@@
 
-Stock assessment models [as described in @hiborn1992b] that account for the population within the MPAs [field2006a] may be able to answer the management relevant question of whether fishing mortality rates and biomass levels from systems containing MPAs are in-line with management objectives [e.g.
+Stock assessment models [as described in @hilborn1992b] that account for the population within the MPAs [@field2006a] may be able to answer the management relevant question of whether fishing mortality rates and biomass levels from systems containing MPAs are in-line with management objectives [e.g.
 @nickols2019].
@@
 See SI for further details of the estimation model.
-`r percent(mean(used_data$eventual_mpa))` of the samples in the raw data come from within MPAs, while the MPAs themselves cover 20% of the surface waters of the channel islands.
+`r percent(mean(used_data$eventual_mpa))` of the samples in the raw data come from within MPAs, while the MPAs themselves cover 20% of the surface waters of the Channel Islands.
 As such, we weight each observation such that within MPA data are assigned a total weight of 0.2 and outside-MPA data a total weight of 0.8.

### 2020-04-24 documents/reviewer-responses.Rmd
got synthetic controls working again

@@
 ---
-title: "Reviewer Responses for 'Population Effects of Marine Protected Areas Are Likely To Be Unclear' (previously 'The Regional Effects of MPAs')"
+title: "Reviewer Responses for 'The Population Effects of Marine Protected Areas Are Likely To Be Unclear' (previously 'The Regional Effects of MPAs')"
 author: 
@@
 We thank the two reviewers and the Editor for their constructive responses.
-We have fundamentally restructured the paper to place much greater emphasis on the empirical results from the Channel Islands MPA network.
+We have fundamentally restructured the paper to place much greater emphasis on the results from the Channel Islands Marine Protected Area (MPA) network.
 This is to our knowledge the first empirical assessment of the regional effects of a large and iconic MPA network on a wide assemblage of commercially and economically important species.
-Our inability to estimate a clear effect in such a well-enforced and well-studied local is we feel a vitally important "null-result" both for scientists and MPA managers to be aware of.
-As such we feel that it warrants publication in PNAS.
+Our inability to estimate a clear population-level effect in such a well-enforced and well-studied location is we feel an important "null-result" both for scientists and MPA managers to be aware of.
+As such we hope that it warrants publication in PNAS.
 
@@
 However, we feel there is substantial value in revisiting those discussions in the context of a concrete empirical example, to provide scientists and practitioners alike a cohesive reference to observe and understand the full ramifications of these past findings on our practical ability to monitor and manage marine protected area networks.
+We have strived to make clearer though that the theoretical section is a reflection of broad body of theoretical work.
 

### 2020-04-24 documents/ovando-et-al-pop-effects-of-mpas-si.Rmd
got synthetic controls working again

--- 2020-04-22 documents/ovando-regional-effects-of-mpas-si-v2.Rmd
@@
 ---
-title: "Supporting Information: Population Effects of Marine Protected Areas Are Likely To Be Unclear"
+title: "Supporting Information"
+subtitle: "Assessing the Population-level Conservation Effects of Marine Protected Areas"
 author: 
@@
 knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
-                      cache = FALSE, fig.width = 7, fig.asp = .75)
+                      cache = FALSE, fig.width = 7, fig.asp = .75,
+                      dev = "cairo_pdf")
 
@@
 library(bayesplot)
+library(Synth)
+
 extrafont::loadfonts()
@@
 We then calculate the difference between mean density of targeted species and the mean density of non-targeted species in each year.
+
+It is important to note that this model asks a slightly different question than our base model.
+The base model estimates the effect on total mean biomass of targeted finfish at the site level.
+This model estimates the effect on the mean biomass density of targeted finfish across all sites.
 
@@
 
+### Synthetic controls
+
+Synthetic controls are an alternative method for attempting to estimate the causal effect of a policy intervention [@abadie2010].
+A difference-in-difference approach assumes that some observable group serves as an adequate control for the state of the treated group in an untreated world.
+In our default case, we assume that the mean standardized index of non-targeted species are our control for the targeted species.
+Alternatively, synthetic controls use timeseries of treated and non-treated groups before and after treatment to construct a new "control" group built by weighting the pre-treatment timeseries of un-treated observations (together with covariates) such that the synthetic control group matches the trends in the treated group pre-treatment.
+
+We chose to present difference-in-difference as our main result since it better allows us to capture the uncertainty in the data generating process through our hierarchical model.
+However, we felt that it was worth exploring whether synthetic controls provided substantially different results than our default model.
+
+For the first synthetic control, we pulled our standardized mean index of abundance for targeted species as a whole from our difference-in-difference model as our treated group.
+We then pull the standardized indices of abundance for each of the non-targeted groups from the difference-in-difference to use as the candidate untreated components for the synthetic controls.
+A complete synthetic control analysis would require more extensive validation of the methods, but we use this approach simply to explore whether we observe substantially divergent results in the synthetic control versus the difference-in-difference model.
+
+We centered and scaled the candidate abundance indices to facilitate model convergence given the very few number of pre-treatment years available.
+The results of a synthetic control model are presented as the difference between the observed treatment outcome and the synthetic control (the difference in this case being in units of standard deviations).
+The model was not able to construct an adequate synthetic control at this level, as shown by the differences between the treated group and the synthetic control pre-treatment.
+However, we would note that the post-treatment results do show similarities with our main results, namely a lack of a clear divergence between the treatment and the control, and an upwards trend up through the earl 2010s followed by a decline (Fig.S\@ref(fig:total-synth)).
+
+```{r, include = FALSE}
+
+years <- unique(used_data$year)
+
+life_history <- used_data %>%
+  select(classcode) %>%
+  unique() %>%
+  mutate(numeric_classcode = as.numeric(as.factor(classcode)))
+
+nontargeted_index <- data_frame(abundance_hat = fit_report$abundance_hat,
+                               classcode = rep(1:n_distinct(used_data$classcode), each = length(years))) %>%
+  mutate(log_abundance_hat = log(abundance_hat)) %>%
+  group_by(classcode) %>%
+  mutate(year =years) %>%
+  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
+  ungroup() %>%
+  rename(numeric_classcode = classcode) %>%
+  left_join(life_history, by = "numeric_classcode" ) %>% 
+  left_join(unique(abundance_data$data[[1]] %>% select(classcode, targeted)), by = "classcode") %>% 
+  filter(targeted == 0) %>% 
+  select(log_abundance_hat, year, classcode, targeted)
+
+targeted_index <- tmb_runs$tmb_fit[[1]]$zissou_estimates %>%
+  filter(
+    stringr::str_detect(variable, "(targeted_did_betas)"),!str_detect(variable, "non")
+  ) %>%
+  mutate(
+    year = years,
+    classcode = "targeted",
+    targeted = 1
+  ) %>%
+  rename(log_abundance_hat = estimate) %>% 
+  select(colnames(nontargeted_index))
+
+  # filter(variable == "seeing_year_species_betas") %>%
+
+nontargeted_covariates <- used_data %>% 
+  select(classcode, contains("mean"),year, targeted) %>% 
+  filter(targeted == 0) %>% 
+  select(-mean_length) %>% 
+  gather(variable, value, -classcode,-year,-targeted) %>% 
+  group_by(classcode, variable, year) %>% 
+  summarise(mean_value = mean(value, na.rm = TRUE)) %>% 
+  spread(variable, mean_value)
+
+targeted_covariates <- used_data %>% 
+  select(classcode, contains("mean"),year, targeted) %>% 
+  filter(targeted == 1) %>% 
+  mutate(classcode = "targeted") %>% 
+  select(-mean_length) %>% 
+  gather(variable, value, -classcode,-year,-targeted) %>% 
+  group_by(classcode, variable, year) %>% 
+  summarise(mean_value = mean(value, na.rm = TRUE)) %>% 
+  spread(variable, mean_value)
+
+synth_covariates <- nontargeted_covariates %>% 
+  bind_rows(targeted_covariates)
+
+synth_data <- targeted_index %>% 
+  bind_rows(nontargeted_index) %>% 
+  left_join(synth_covariates, by = c("classcode","year")) %>% 
+    mutate(numeric_classcode = as.numeric(factor(classcode))) %>% 
+  group_by(classcode) %>% 
+  mutate(log_abundance_hat = scale(exp(log_abundance_hat))) %>% 
+  ungroup() %>% 
+  as.data.frame() %>% 
+  select(-rolling_mean)
+
+# synth_data %>%
+#   group_by(classcode) %>%
+#   mutate(index = (log_abundance_hat)) %>%
+#   ggplot(aes(
+#     year,
+#     index,
+#     color = factor(targeted),
+#     group = interaction(targeted, classcode)
+#   )) +
+#   geom_line()
+
+prepped_synth_data <-
+  Synth::dataprep(
+    foo = synth_data,
+    predictors = colnames(synth_data)[str_detect(colnames(synth_data),"mean")],
+    predictors.op = "mean",
+    dependent = "log_abundance_hat",
+    unit.variable = "numeric_classcode",
+    unit.names.variable = "classcode",
+    time.variable = "year",
+    treatment.identifier = "targeted",
+    controls.identifier = unique(synth_data$classcode)[unique(synth_data$classcode) != "targeted"],
+    time.predictors.prior = years[years <= 2003],
+    time.optimize.ssr = years[years <= 2003],
+    time.plot = years
+  )
+
+synth_fit <- Synth::synth(prepped_synth_data)
+
+gaps <-
+  rownames_to_column(
+    prepped_synth_data$Y1plot - (prepped_synth_data$Y0plot %*% synth_fit$solution.w) %>%
+      as.data.frame()
+  ) %>%
+  rename(gap = w.weight,
+         year = rowname) %>%
+  mutate(year = as.numeric(year)) %>%
+  mutate(synthetic = as.numeric(prepped_synth_data$Y0plot %*% synth_fit$solution.w)) %>% 
+  mutate(treatment = gap + synthetic)
+
+# path.plot(synth.res = synth_fit,
+#           dataprep.res = prepped_synth_data
+# )
+
+total_gap_plot <- gaps %>% 
+  ggplot(aes(year, gap)) + 
+  geom_hline(aes(yintercept = 0), linetype = 2) +
+  geom_vline(aes(xintercept = 2003), color = "red")+
+  geom_line() + 
+    labs(x = "Year", y = "Standard deviations from synthetic control")
+
+total_synth_plot <- gaps %>% 
+  select(year, synthetic, treatment) %>% 
+  gather(source, value, -year) %>% 
+  group_by(source) %>% 
+  # mutate(value = scale(value)) %>% 
+  ggplot(aes(year, value, color = source, linetype = source)) + 
+  geom_vline(aes(xintercept = 2003), color = "red")+
+  geom_line() + 
+    labs(x = "Year", y = "Standard deviations from synthetic control")
+
+```
+
+```{r raw-synth, include = FALSE}
+
+years <- unique(used_data$year)
+
+life_history <- used_data %>%
+  select(classcode) %>%
+  unique() %>%
+  mutate(numeric_classcode = as.numeric(as.factor(classcode)))
+
+nontargeted_index <- data_frame(abundance_hat = fit_report$abundance_hat,
+                               classcode = rep(1:n_distinct(used_data$classcode), each = length(years))) %>%
+  mutate(log_abundance_hat = log(abundance_hat)) %>%
+  group_by(classcode) %>%
+  mutate(year =years) %>%
+  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
+  ungroup() %>%
+  rename(numeric_classcode = classcode) %>%
+  left_join(life_history, by = "numeric_classcode" ) %>% 
+  left_join(unique(abundance_data$data[[1]] %>% select(classcode, targeted)), by = "classcode") %>% 
+  filter(targeted == 0) %>% 
+  select(log_abundance_hat, year, classcode, targeted)
+
+targeted_index <- tmb_runs$tmb_fit[[1]]$zissou_estimates %>%
+  filter(
+    stringr::str_detect(variable, "(targeted_did_betas)"),!str_detect(variable, "non")
+  ) %>%
+  mutate(
+    year = years,
+    classcode = "targeted",
+    targeted = 1
+  ) %>%
+  rename(log_abundance_hat = estimate) %>% 
+  select(colnames(nontargeted_index))
+
+  # filter(variable == "seeing_year_species_betas") %>%
+
+nontargeted_covariates <- used_data %>% 
+  select(classcode, contains("mean"),year, targeted) %>% 
+  filter(targeted == 0) %>% 
+  select(-mean_length) %>% 
+  gather(variable, value, -classcode,-year,-targeted) %>% 
+  group_by(classcode, variable, year) %>% 
+  summarise(mean_value = mean(value, na.rm = TRUE)) %>% 
+  spread(variable, mean_value)
+
+targeted_covariates <- used_data %>% 
+  select(classcode, contains("mean"),year, targeted) %>% 
+  filter(targeted == 1) %>% 
+  mutate(classcode = "targeted") %>% 
+  select(-mean_length) %>% 
+  gather(variable, value, -classcode,-year,-targeted) %>% 
+  group_by(classcode, variable, year) %>% 
+  summarise(mean_value = mean(value, na.rm = TRUE)) %>% 
+  spread(variable, mean_value)
+
+synth_covariates <- nontargeted_covariates %>% 
+  bind_rows(targeted_covariates)
+
+synth_data <- targeted_index %>% 
+  bind_rows(nontargeted_index) %>% 
+  left_join(synth_covariates, by = c("classcode","year")) %>% 
+    mutate(numeric_classcode = as.numeric(factor(classcode))) %>% 
+  group_by(classcode) %>% 
+  mutate(log_abundance_hat = scale(exp(log_abundance_hat))) %>% 
+  ungroup() %>% 
+  as.data.frame() %>% 
+  select(-rolling_mean)
+
+# synth_data %>%
+#   group_by(classcode) %>%
+#   mutate(index = (log_abundance_hat)) %>%
+#   ggplot(aes(
+#     year,
+#     index,
+#     color = factor(targeted),
+#     group = interaction(targeted, classcode)
+#   )) +
+#   geom_line()
+
+prepped_synth_data <-
+  Synth::dataprep(
+    foo = synth_data,
+    predictors = colnames(synth_data)[str_detect(colnames(synth_data),"mean")],
+    predictors.op = "mean",
+    dependent = "log_abundance_hat",
+    unit.variable = "numeric_classcode",
+    unit.names.variable = "classcode",
+    time.variable = "year",
+    treatment.identifier = "targeted",
+    controls.identifier = unique(synth_data$classcode)[unique(synth_data$classcode) != "targeted"],
+    time.predictors.prior = years[years <= 2003],
+    time.optimize.ssr = years[years <= 2003],
+    time.plot = years
+  )
+
+synth_fit <- Synth::synth(prepped_synth_data)
+
+gaps <-
+  rownames_to_column(
+    prepped_synth_data$Y1plot - (prepped_synth_data$Y0plot %*% synth_fit$solution.w) %>%
+      as.data.frame()
+  ) %>%
+  rename(gap = w.weight,
+         year = rowname) %>%
+  mutate(year = as.numeric(year)) %>%
+  mutate(synthetic = as.numeric(prepped_synth_data$Y0plot %*% synth_fit$solution.w)) %>% 
+  mutate(treatment = gap + synthetic)
+
+# path.plot(synth.res = synth_fit,
+#           dataprep.res = prepped_synth_data
+# )
+
+total_gap_plot <- gaps %>% 
+  ggplot(aes(year, gap)) + 
+  geom_hline(aes(yintercept = 0), linetype = 2) +
+  geom_vline(aes(xintercept = 2003), color = "red")+
+  geom_line() + 
+    labs(x = "Year", y = "Standard deviations from synthetic control")
+
+total_synth_plot <- gaps %>% 
+  select(year, synthetic, treatment) %>% 
+  gather(source, value, -year) %>% 
+  group_by(source) %>% 
+  # mutate(value = scale(value)) %>% 
+  ggplot(aes(year, value, color = source, linetype = source)) + 
+  geom_vline(aes(xintercept = 2003), color = "red")+
+  geom_line() + 
+    labs(x = "Year", y = "Standard deviations from synthetic control")
+
+```
+
+```{r  total-synth, fig.cap = "Difference in centered and scaled standardized targeted abundance and synthetic standardized targeted abundance"}
+total_gap_plot
+```
+
+As an extension, we repeated this process, but now treating each targeted species individually as the treated group, and the non-targeted species as the non-targeted.
+This is intended to explore whether we see clearer signals for individual species than we do for the targeted class as a whole.
+
+Overall we see similarly unclear results as the aggregate targeted synthetic control (and our main results).
+The synthetic control was better constructed for some individual species, but not clearly for any one, and most species showed some evidence of the upward-then-downward trend seen throughout our results (Fig.S\@ref(fig:synth-classcode-fits)).
+
+```{r, include = FALSE}
+
+synth_foo <-
+  function(treated_classcode,
+           synth_data,
+           targeted_classcodes) {
+    # treated_classcode <- targeted_classcodes[1]
+
+    temp_synth_data <- synth_data %>%
+      filter(classcode == treated_classcode |
+               !(classcode %in% targeted_classcodes))
+
+    prepped_synth_data <-
+      Synth::dataprep(
+        foo = temp_synth_data,
+        predictors = colnames(temp_synth_data)[str_detect(colnames(temp_synth_data), "mean")],
+        predictors.op = "mean",
+        dependent = "log_abundance_hat",
+        unit.variable = "numeric_classcode",
+        unit.names.variable = "classcode",
+        time.variable = "year",
+        treatment.identifier = treated_classcode,
+        controls.identifier = unique(temp_synth_data$classcode)[unique(temp_synth_data$classcode) != treated_classcode],
+        time.predictors.prior = c(2000:2003),
+        time.optimize.ssr = c(2000:2003),
+        time.plot = years
+      )
+
+    synth_fit <- Synth::synth(prepped_synth_data)
+
+    gaps <-
+      rownames_to_column(
+        prepped_synth_data$Y1plot - (prepped_synth_data$Y0plot %*% synth_fit$solution.w) %>%
+          as.data.frame()
+      ) %>%
+      rename(gap = w.weight,
+             year = rowname) %>%
+      mutate(year = as.numeric(year)) %>%
+      mutate(synthetic = as.numeric(prepped_synth_data$Y0plot %*% synth_fit$solution.w)) %>%
+      mutate(treatment = gap + synthetic)
+  }
+
+nontargeted_index <- data_frame(abundance_hat = fit_report$abundance_hat,
+                               classcode = rep(1:n_distinct(fitted_data$classcode), each = length(years))) %>%
+  mutate(log_abundance_hat = log(abundance_hat)) %>%
+  group_by(classcode) %>%
+  mutate(year = 1999 + 1:length(abundance_hat)) %>%
+  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
+  ungroup() %>%
+  rename(numeric_classcode = classcode) %>%
+  left_join(life_history, by = "numeric_classcode" ) %>% 
+  left_join(unique(abundance_data$data[[1]] %>% select(classcode, targeted)), by = "classcode") %>% 
+  select(log_abundance_hat, year, classcode, targeted)
+
+synth_covariates <- pisco_data %>% 
+  select(classcode, contains("mean"),year, targeted) %>% 
+  select(-mean_length) %>% 
+  gather(variable, value, -classcode,-year,-targeted) %>% 
+  group_by(classcode, variable, year) %>% 
+  summarise(mean_value = mean(value, na.rm = TRUE)) %>% 
+  spread(variable, mean_value)
+
+synth_data <- nontargeted_index %>% 
+  left_join(synth_covariates, by = c("classcode","year")) %>% 
+    mutate(numeric_classcode = as.numeric(factor(classcode))) %>% 
+  group_by(classcode) %>% 
+  mutate(log_abundance_hat = scale(exp(log_abundance_hat))) %>% 
+  ungroup() %>% 
+  as.data.frame()
+
+targeted_classcodes <- unique(synth_data$classcode[synth_data$targeted == 1])
+
+synth_classcode_fits <-
+  tibble(targeted_classcodes = targeted_classcodes) %>%
+  mutate(
+    synthetic_fit = map(
+      targeted_classcodes,
+      synth_foo,
+      synth_data = synth_data,
+      targeted_classcodes = targeted_classcodes
+    )
+  )
+
+synth_classcode_fits <- synth_classcode_fits %>% 
+  unnest()
+
+synth_classcode_fits_plot <- 
+synth_classcode_fits %>% 
+  ggplot(aes(year, gap)) + 
+  geom_hline(aes(yintercept = 0), linetype = 2) +
+  geom_vline(aes(xintercept = 2003), color = "red")+
+  geom_line() + 
+  facet_wrap(~targeted_classcodes) + 
+  labs(x = "Year", y = "Standard deviations from synthetic control")
+
+```
+
+```{r synth-classcode-fits, fig.cap="Synthetic control gaps for each targeted species"}
+synth_classcode_fits_plot + 
+  theme_minimal()
+```
+
 ## Testing Model Assumptions
@@
 The difference-in-difference model also assumes that the targeted and non-targeted fishes do not directly or indirectly affect each other.This assumption is clearly violated on some level: all the fishes in this analysis are part of the same ecosystem and therefore interact to some degree.
-For example, if the protection of targeted predatory fishes results in increased mortality of non-targeted fishes, the model would attribute that as an increased regional effect (greater divergence between the abundance of targeted and non-targeted species).
+For example, if the protection of targeted predatory fishes results in increased mortality of non-targeted fishes, the model would attribute that as an increased population effect (greater divergence between the abundance of targeted and non-targeted species).
 Given the time scale of analysis (15 years of protection), we do not feel that massive trophic cascades are likely to have developed yet, given both the pace and complexity of trophic cascade development [@babcock2010; @pershing2015a].
@@
 
+The model consists of 50 patches with wrapped edges (picture the waters around a circular island).
+For any one simulation we randomly pull a species and its associated life history from the `FishLife` [@thorson2017c] package in R.
+We pair these data with randomly selected values between 0.6 and 0.95 for Beverton-Holt steepness [as parameterized in @mace1994], as well as larval and adult dispersal rates.
+We randomly assign whether adult fish preferentially move towards patches with lower relative densities, as well as one of three potential types of recruitment density dependence [building off of @babcock2011].
+Simulations are also assigned random fleet dynamics (open access ,constant effort, constant catch), responses to MPA (leave the fishery or concentrate outside MPAs), a fleet dispersal model (uniform in response to catch-per-unit-effort, in response to profit-per-unit-effort)
+
+<!--   1. Local density dependence: Density dependence occurs independently in each patch, and recruits then disperse to nearby patches -->
+
+<!--   2. Global density dependence: Density dependence is a function of the sum of spawning biomass across all patches, and recruits are then distributed according to habitat quality -->
+
+<!--   3. Post-dispersal density dependence: Larvae are distributed throughout the system, and then density dependence occurs based on the density of adult biomass at the destination patch -->
+
+<!-- We allow for three potential siting strategies for MPAs.
+In the first, MPAs are randomly placed.
+In the second, we assume that MPAs are placed in preferentially better habit (unfished recruitment is four times greater inside MPA locations).
+In the third, we allow for scenarios in which MPAs are placed in sources of larval dispersal.
+-->
+
+<!-- Each simulation is randomly assigned a fleet model of the form  -->
+
+<!--   1. Open access: fishing effort changes in response to profit-per-unit-effort -->
+
+<!--   2. Constant effort: total fishing effort is constant over time (unless altered by MPA displacement model) -->
+
+<!--   3. Constant catch: the fleet exerts as much effort as need to achieve a target catch -->
+
+<!-- This fleet model is paired with a gear selectivity ranging from .1 to 1.5 of the length at 50% maturity for the species in question, and the fleet model is tuned to achieve a target fishing mortality relative to natural mortality ratio at equilibrium.
+-->
+
+<!-- Along with the fleet dynamics model, each simulation is assigned a random fleet dispersal scenario: uniform dispersal (where the total effort of the fleet is divided evenly among all open patches), catch dispersal (where the total effort of the fleet is divided according to the catchable biomass in each available patch), and profit dispersal (where the total effort of the fleet is divided according to the profit per unit effort in each available patch).
+Fishing effort that occurred inside MPAs prior to closer can either leave the fishery, or be distributed to the patches outside the MPAs.
+-->
+
+Each simulation is assigned an MPA scenario, defined by the number and size of MPAs, the placement of those MPAs, and the year that the MPAs are put in place.
+The population begins at unfished equilibrium and then fishing effort is applied in accordance to the fleet model.
+The MPAs are then placed during the randomly selected start year, allowing some runs to explore how the early dynamics of the MPA play out when the fishery and population they are placed on is not already at equilibrium.
+Each simulation is run to equilibrium with and without the selected MPA strategy (holding all else constant).
+We then measure the difference in biomass densities each time step in the scenario with and without the MPAs to calculate the population effect of the MPAs over time.
+
 The operating model is a spatial single-species age-structured bio-economic model.
@@
 
-```{r pop-conservation-effect, include = TRUE, fig.cap="Median (A) and range (B) of equilibrium regional MPA conservation effects (change in total biomass with MPAs relative to without MPAs as a percentage of unfished biomass) across a range of depletion and MPA sizes (and incorporating the full range of scenarios included in our study).
+```{r pop-conservation-effect, include = TRUE, fig.cap="Median (A) and range (B) of equilibrium population-level MPA conservation effects (change in total biomass with MPAs relative to without MPAs as a percentage of unfished biomass) across a range of depletion and MPA sizes (and incorporating the full range of scenarios included in our study).
 'Range in MPA' is the percent of patches covered by an MPA, 'Depletion' is the depletion that would have occurred in equilibrium without the MPA"}

### 2020-04-24 documents/ovando-et-al-pop-effects-of-mpas.Rmd
got synthetic controls working again

--- 2020-04-22 documents/ovando-regional-effects-of-mpas-v2.Rmd
@@
 ---
-title: Population Effects of Marine Protected Areas Are Likely To Be Unclear
+title: Assessing the Population-level Conservation Effects of Marine Protected Areas
 date: "`r Sys.Date()`"
@@
 abstract: |
-   Marine Protected Areas (MPAs) cover 3-7% of the world's oceans, up from less than 1\% in the year 2000.
-The Convention on Biological Diversity called for 10% of coastal waters to be protected inside MPAs by 2020, while the International Union for Conservation of Nature is now calling for 30\% protection by 2030.
-It is often clear that MPAs produce conservation benefits inside their borders, but many MPAs are also justified on the grounds that they may benefit the broader region outside their borders.
-The conservation effects of MPAs are most commonly evaluated using response ratios---measures of numerical or biomass densities inside their borders, relative to these metrics in reference sites outside their borders.
-Studies of this nature have provided broad evidence that MPAs produce conservation benefits within their borders.
-While these gains can be vitally important, marine populations are rarely contained entirely within the borders of MPAs; therefore a critical question is not only do MPAs produce conservation gains inside their borders, but how do they affect the broader region in which they are located?
-The Channel Islands National Marine Sanctuary provides a clear example of the importance and challenges of estimating regional MPA effects.
-A network of MPAs covering 20\% of the Islands’ waters was put in place in 2003, with a goal of providing regional conservation and fishery benefits.
-Response ratios from the region indicate that the Channel Island MPAs have increased biomass densities inside the MPAs.
-We are unable to find a clear effect of these same MPAs on biomass densities of targeted finfish throughout the Channel Islands region.
-We use a bio-economic simulation model to explain this discrepancy, and demonstrate under what conditions we likely can and cannot clearly estimate the regional conservation effects of MPAs.
-We show that MPA effect sizes less than 30% are likely to be difficult to detect, and the size of many MPA networks suggests that effect sizes may often be smaller than 30%.
-Our results provide a novel assessment of the regional effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.
+   Marine Protected Areas (MPAs) cover 3-7% of the world’s ocean, up from less than 1% in the year 2000, and international commitments call for 10%, 30%, and even 50% coverage.
+The premise underlying MPA expansion is that they conserve biodiversity, habitats, and fished populations.
+While numerous studies show that MPAs produce conservation benefits *inside* their borders, many MPAs are also justified on the grounds that they confer conservation benefits to the broader population *beyond* their borders.
+This paper examines the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and shows that even in cases when the conservation benefits are large, they are inherently difficult to detect empirically.
+A network of MPAs was put in place in The Channel Islands National Marine Sanctuary in 2003, with a goal of providing regional conservation and fishery benefits.
+Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level.
+MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present), and the size of many MPA networks suggests that effect sizes may often be smaller than 30%.
+Our results provide a novel assessment of the population-wide effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.
 bibliography: My Library.bib
@@
 knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
-                      cache = FALSE, fig.width = 3.42, fig.asp = 1.5)
+                      cache = FALSE, fig.width = 3.42, fig.asp = 1.5, dev = "cairo_pdf")
 
@@
 C.C., S.G., and O.D's research program includes funding from environmental NGOs and foundations with an interest in ocean conservation and food security.
-Funding for D.O.
-was partly provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship.
-
 ## significance
@@
 Marine protected areas (MPAs) are increasingly being used to conserve and manage these ecosystems.
-Our study pairs theory with empirical methods to show that the regional conservation effects of marine protected areas can be highly variable and dependent on human drivers.
+Our study pairs theory with empirical methods to show that the population-level conservation effects of marine protected areas can be highly variable and dependent on human drivers.
 We find no statistically clear effect of a large and well protected network of MPAs in the Channel Island, California, on targeted finfish after 14 years of protection, which our simulation analysis suggests is actually to be expected.
-MPA practitioners must clearly communicate expectations from MPA networks, and plan monitoring and adaptations strategies that directly address the challenges of estimating regional conservation effects of protected areas.
+MPA practitioners must clearly communicate expectations from MPA networks, and plan monitoring and adaptations strategies that directly address the challenges of estimating population-level conservation effects of protected areas.
 
@@
 ## keywords:
-  - Marine Protected Areas
+  - Marine Protected Area Networks
   - Conservation
@@
 Recent international agreements to expand MPAs (most proximately, the Convention on Biological Diversity’s Strategic Plan for Biodiversity, which calls for 10% of coastal waters to be protected inside MPAs by 2020 and the International Union for Conservation of Nature call for 30% by 2030) are based on the expectation that well-designed MPAs will achieve benefits both within, and outside, their borders.
-Indeed, Numerous studies provide evidence that well-enforced and appropriately sized MPAs can produce conservation benefits within their borders [@lester2009;@halpern2003; @edgar2014; @gerber2005].
-However, our collective scientific understanding of the regional-scale conservation and fishery impacts of current and future MPAs is surprisingly limited.
+Numerous studies provide evidence that well-enforced and appropriately sized MPAs can produce conservation benefits within their borders [@lester2009;@halpern2003; @edgar2014; @gerber2005].
+However, our collective scientific understanding of the regional-scale impacts of current and future MPAs on fish populations and fisheries is surprisingly limited.
 
 As conservation benefits accrue inside MPAs, theory holds that MPAs can affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
+These dynamics link fish both inside and outside of MPAs as part of a shared biological stock; the population.
 Several studies have documented empirical evidence for the existence of adult or larval spillover affecting both abundance and fisheries [e.g.
 @goni2010; @halpern2009;@kay2012; @stobart2009; @mcclanahan2000; @russ1996; @thompson2017;@baetscher2019;@pelc2009], as well as alteration of fishing effort in reaction to [@costa2013;@mason2012;@murawski2005] and in anticipation of [@mcdermott2019] MPA placement.
-However, empirically detecting the regional net effects of spillover on fish populations and fisheries is much more difficult.
+However, empirically detecting the net effects of spillover on entire fish populations and fisheries is much more difficult.
 From a fishery perspective, are spillover benefits sufficient to offset losses in fishing grounds and changes in responses of displaced fishers caused by an MPA?
 From a conservation perspective, how much does the buildup of fish inside an MPA increase biomass outside the protected area?
-Overall, what are the regional effects of MPAs?
+Overall, what are the population-level effects of MPAs?
 
-As stakeholders around the world seek to use MPAs in marine resource management portfolios, it is critical that we develop a better understanding of the magnitude and drivers of regional-scale MPA effects.
+As stakeholders around the world seek to use MPAs in marine resource management portfolios, it is critical that we develop a better understanding of the magnitude and drivers of population-level MPA effects.
 This is particularly true in the context of a changing climate, in which "set it and forget it" management strategies are increasingly untenable; we need to be able to monitor the performance of MPAs in order to adapt them as dictated by a shifting environment.
@@
 
-We use data from the first 14 years of protection to provide what is to our knowledge the first empirical assessment of the regional effect of a large MPA network on a wide array of finfish species.
-For the purpose of this study, we define a regional effect as the effect of MPAs on the mean biomass density of targeted finfish both inside and outside of MPAs.
-In contrast to clear differences in biomass densities observed inside and outside of well protected MPAs both globally [@lester2009] and in the Channel Islands [@caselle2015] we are unable to detect a clear regional effect from the Channel Islands MPAs.
-We build off of existing MPA theory to consider why this might be, and provide guidance for scientists and managers as to when and how we might expect to estimate the regional conservation effects of MPAs.
+We use data from the first 14 years of protection to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of finfish species.
+In contrast to clear differences in biomass densities observed inside and outside of well protected MPAs both globally [@lester2009] and in the Channel Islands [@caselle2015] we are unable to detect a clear population effect from the Channel Islands MPAs.
+We build off of existing MPA theory to consider why this might be, and provide guidance for scientists and managers as to when and how we might expect to estimate the population-level conservation effects of MPAs.
 
@@
 
-### What Are the Regional Effects of MPAs? {.unnumbered}
+### What Are the Population-Level Effects of MPAs? {.unnumbered}
 
@@
 
-We define the regional conservation effects of MPAs as the change in total mean biomass densities of targeted finfish both inside and outside of MPAs, relative to the total mean biomass densities of targeted finfish inside and outside of MPAs that would have occurred without the MPAs.
-While other outcomes such as increased biodiversity or resiliency are also important to conservation, these are beyond the scope of this analysis).
-As an example of a regional MPA effect, consider a evenly distributed population that has 50% of its range protected by an MPA.
+We define the population-level conservation effects of MPAs as the change in total mean biomass densities of targeted finfish both inside and outside of MPAs, relative to the total mean biomass densities of targeted finfish inside and outside of MPAs that would have occurred without the MPAs.
+As an example of a population-level MPA effect, consider a evenly distributed population that has 50% of its range protected by an MPA.
 Suppose that the MPA increase biomass densities inside the reserves by 20%, and by 0% outside the reserve.
-By our metric the regional conservation effect of the MPA would be 10%.
+By our metric the population conservation effect of the MPA would be 10%.
 
-Numerous factors can affect the regional effects of MPAs.
+Numerous factors can affect the population effects of MPAs.
 These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], larval dispersal patterns and the strength and timing of density dependence in the population (e.g.
@@
 
-Unfortunately, these response ratios do not provide sufficient evidence for significant regional-scale effects of the Channel Islands MPAs.
+Unfortunately, these response ratios do not provide sufficient evidence for significant population-level effects of the Channel Islands MPAs.
 Control sites used in calculating response ratios are often selected based on abiotic or ecological traits such as habitat characteristics [@ferraro2018].
@@
 In theory, control sites far enough away to negate both biological spillover and concentration by the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs may be challenging.
-As a result of these challenges, response ratios can be a highly imprecise and biased measure of the regional conservation effect of an MPA network [@ferraro2018].
+As a result of these challenges, response ratios can be a highly imprecise and biased measure of the population-level conservation effect of an MPA network [@ferraro2018].
 
@@
 Given these complications, we can interpret the response ratios reported in @caselle2015, and updated here, as evidence that the MPAs are providing effective protection of targeted fish biomass inside their borders, particularly in light of the general increase in response ratios over time.
-But, we cannot use these response ratio results serve as a definitive indicator of overall regional effects of the MPAs.
+But, we cannot use these response ratio results serve as a definitive indicator of overall population effects of the MPAs.
 
@@
 For each year of protection, we paired our simulated response ratios to the estimated posterior probability distributions of the  empirical Channel Island response ratios.
-For example, if the mean estimated response ratio in the year 2006 (year 3 after implementation) is one, we found all simulations with three years of protection that had simulated response ratios near one, and then pulled out the "true" simulated regional MPA effect from each of those simulations.
-This provides us with a distribution of simulated regional MPA effects that could plausibly generate the types of response ratios actually observed in the Channel Islands.
+For example, if the mean estimated response ratio in the year 2006 (year three after implementation) is one, we found all simulations that had simulated response ratios near one after three years of protection, and then pulled out the "true" simulated population-level MPA effect from each of those simulations.
+This provides us with a distribution of simulated population-level MPA effects that could plausibly generate the types of response ratios actually observed in the Channel Islands.
 
-The response ratio trends we observe in the data could plausibly be produced by a wide range of regional MPA effects (Fig.\@ref(fig:rr-plot)).
-Response ratios well over one were associated with regional MPA effects generally less than 25%, and, importantly, many simulations produced large response ratios but regional MPA effects close to 0% This can occur if for example fishing pressure is only moderate, adult movement is low, larval dispersal is high, and displaced fishing effort concentrates around the border of the MPAs.
-The response ratio results previously published for this region do not serve as sufficient evidence for the regional conservation effects of the Channel Islands MPA network.
+The response ratio trends we observe in the data could plausibly be produced by a wide range of population-level MPA effects (Fig.\@ref(fig:rr-plot)).
+Response ratios well over one were associated with population-level MPA effects generally less than 25%, and, importantly, many simulations produced large response ratios but population-level MPA effects close to 0% This can occur if for example fishing pressure is only moderate, adult movement is low, larval dispersal is high, and displaced fishing effort concentrates around the border of the MPAs.
 
 ```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="90% Posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis) in blue.
-Simulated MPA effect on biomass densities matched to empirical response ratios in red.
-For response ratios, a value of zero indicates that biomass densities of targeted and non-targeted species are identical inside and outside MPAs, a value of 1 indicates that biomass densities of targeted species are 100% greater inside MPAs than outside.
-For MPA effect, a value of zero indicates that mean biomass densities are identical in the with- and without- MPA scenarios.
-A value of 1 indicates that mean biomass densities are 100% greater in the scenario with MPAs than the scenario without MPAs."}
+Simulated population-level (pop.) effect on biomass densities matched to empirical response ratios in red.
+For response ratios, a value of zero indicates that biomass densities of targeted species are identical inside and outside MPAs, a value of one that biomass densities of targeted species are 100% inside MPAs relative to outside.
+For MPA population effect, a value of zero indicates that biomass densities are identical in the with- and without- MPA scenarios.
+A value of 1 indicates that biomass densities are 100% greater in the scenario with MPAs than the scenario without MPAs."}
 
@@
 
-Given the potential unreliability of response ratios for the task, in an ideal world how would we empirically measure the regional effects of MPAs?
+Given the potential unreliability of response ratios for the task, how can we empirically measure the population-level effects of MPAs?
 The perfect experiment would involve two parallel worlds that were identical, except for the implementation of an MPA.
@@
 
-Building off of the concepts explored in @caselle2015, we used an identification strategy utilizing biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 0])` species that are not directly targeted by fishing as our control group (non-targeted), and biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 1])` species targeted by fishing as our treatment group.
+Building off of the concepts explored in @caselle2015, we used an identification strategy utilizing biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 0])` species that are not directly targeted by fishing before and after MPA implementation as our control group (non-targeted), and biomass densities before and after MPA implementation of `r n_distinct(used_data$classcode[used_data$targeted == 1])` species targeted by fishing as our treatment group.
 Targeted species in the Channel Islands, include commercially important fin-fish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
 Each of these targeted species was the subject of prior bio-economic modeling related to the effects of MPAs in southern California [@costello2010a;@rassweiler2014].
-It should be noted though that our analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*).
+Our analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*) (but our methods could similarly be applied to these species given data).
 Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species).
@@
 Over the next six years the model estimates greater probabilities of an increasingly positive MPA effect, peaking in 2009-2011 with a median estimate of MPA effect of a `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)` increase in total mean biomass density of targeted species (90% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)`
-- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)`.
+- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)` (Fig.\@ref(fig:mpa-effect)).
 These empirical estimates are in line with the outcomes that our simulation model suggests are plausible.
@@
 
-```{r mpa-effect, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Results of difference-in-difference regression estimating the regional effect of the Channel Island MPAs on total mean biomass densities of targeted species (difference in total mean biomass density of targeted species over time relative to expected levels using non-targeted species as a control).
+```{r mpa-effect, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Results of difference-in-difference regression estimating the population effect of the Channel Island MPAs on total mean biomass densities of targeted species (difference in total mean biomass density of targeted species over time relative to expected levels using non-targeted species as a control).
 Grey distributions show posterior probability distribution of estimated MPA effect; red point is median estimated effect, thicker red section 50% credible interval, thinner red line 90% credible interval).
@@
 
-How can we explain the lack of a clear regional MPA effect after over ten years of protection?
+How can we explain the lack of a clear population-level MPA effect after over ten years of protection?
 One explanation may lie in fleet dynamics.
@@
 We hypothesize that the recent evidence for a decline in densities of targeted species is due to environmental conditions that disproportionately affect the targeted group (and not for example due to concentrated fishing pressure outside the reserves).
-The practical result of this is that in the presence of this marine heatwave the non-targeted species may no longer serve as an effective control for the evolution of biomass densities of tartegeted finfish in the absence of the MPAs (a violation of the "parallel trends" assumption required for Difference-in-difference models).
+The practical result of this is that in the presence of this marine heatwave the non-targeted species may no longer serve as an effective control for the evolution of biomass densities of targeted finfish in the absence of the MPAs (a violation of the "parallel trends" assumption required for Difference-in-difference models).
 
@@
 
-Containing a carefully designed, well-enforced, and well-studied MPA network, the Channel Islands would seem at face value to be an ideal location to study the regional conservation effects of protected areas.
+Containing a carefully designed, well-enforced, and well-studied MPA network, the Channel Islands would seem at face value to be an ideal location to study the population-level effects of protected areas.
 The persistently high response ratios suggest that despite overall decreases in targeted biomass densities inside and outside MPAs, the MPAs may still be providing protection within their borders.
-But, as we have shown here these response ratios are not necessarily an indicator of broader regional effects.
-The difference-in-difference strategy utilized here presents an alternative identification strategy, that while not without its own strict caveats presents some potential improvements over response-ratios as a means of estimating regional conservation effects.
+But, as we have shown here these response ratios are not necessarily an indicator of broader population-level effects.
+The difference-in-difference strategy utilized here presents an alternative identification strategy to response ratios , that while not without its own strict caveats presents some potential improvements over response-ratios as a means of estimating population-level effects.
 While we estimate an uncertain but overall positive effect of the MPA network in its first few years, we are unable to detect a robust signal from 2012-2017.
@@
 Even models fit to data generated from large effect sizes commonly misestimated the true MPA effect by 50% or more.
-Obtaining a mean absolute percent error (MAPE) of 25% or less across our simulated datasets required a regional MPA effect of at least 30%.
+Obtaining a mean absolute percent error (MAPE) of 25% or less across our simulated datasets required a population-level MPA effect of at least 30%.
 This is merely an illustrative exercise, omitting critical factors such as detection probability and sampling strategy.
 However, since nearly any omission which one can think of would make an MPA effect harder to detect, not easier, these results serve as a useful floor for the likely difficulty in estimating MPA effects.
-In the context of the Channel Islands, given the potential effect size produced by our simulation model these results suggest that we might expect to be unable to precisely estimate the true regional effect of the MPAs.
+In the context of the Channel Islands, given the potential effect size produced by our simulation model these results suggest that we might expect to be unable to precisely estimate the true population-level effect of the MPAs.
 
@@
 
-```{r val-plot, include = TRUE,fig.width = width_1col, fig.asp=asp_1col, fig.cap = "Distribution of percent error in posterior estimates of MPA effect (y-axis) plotted against true simulated MPA effect (x-axis).
+```{r val-plot, include = TRUE,fig.width = width_1col, fig.asp=asp_1col, fig.cap = "Distribution of percent error in posterior estimates of population-level MPA effect (y-axis) plotted against true simulated MPA effect (x-axis).
 Color shows concentration of simulations.
@@
 Looking across these two variables, if the MPA network covers 25% or more of a species range and pre-MPA depletion is greater than 60% we might expect an effect size with our target MAPE.
-While recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, for near-shore commercial fin-fish many MPA networks are likely to cover areas more in line with the Channel Islands (20%), and as such have regional effect sizes that may be difficult to detect (Fig.\@ref(fig:sim-effects)-A).
+While recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, for near-shore commercial fin-fish many MPA networks are likely to cover areas more in line with the Channel Islands (20%) or smaller, and as such have population-level effect sizes that may be difficult to detect (Fig.\@ref(fig:sim-effects)-A).
 
@@
 To put these results into context, the FAO estimates that 7% of the worlds fisheries with status estimates fall into the relatively unexploited category (roughly depletion less than 50%), 60% fall into the fully fished category (roughly depletion 50%-70%), and 33% fall into the heavily fished category (roughly depletion greater than 70%) [@fao2018, assuming a biomass at maximum sustainable yield to unfished biomass ratio of 40%, per @thorson2012], though works that include a broader range of fisheries estimate that 50% or more of stocks to fall into the heavily fished category [@costello2012;@rosenberg2018].
-The regional effects of MPA networks covering 25% or less of a species range may be difficult to detect in many places with already well-managed fisheries, while for that size we might expect clearer effects in less-managed locations (though that of course ignores the complication of compliance with MPA regulations).
-Within these broad guidelines a wide range of outcomes are possible based on local fleet and fish dynamics: as a starting place users can use the bio-economic MPA simulation model developed for this paper to explore potential outcomes for specific MPAs using an interactive web application available at [danovando.shinyapps.io/simmpa](https://danovando.shinyapps.io/simmpa/).
+The population-level effects of MPA networks covering 25% or less of a species range may be difficult to detect in many places with already well-managed fisheries, while for that size we might expect clearer effects in less-managed locations (though that of course ignores the complication of compliance with MPA regulations).
+Within these broad guidelines a wide range of outcomes are possible based on local fleet and fish dynamics: as a starting place users can use the bio-economic MPA simulation model developed for this paper to explore potential outcomes for specific MPAs using an interactive web application available [here](https://danovando.shinyapps.io/simmpa/).
 
-```{r sim-effects, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Simulated MPA effect sizes as a function of percent of species' range inside MPA (x-axis), and pre-MPA depletion (y-axis).
+```{r sim-effects, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Simulated population-level (pop.) MPA effect sizes as a function of percent of species' range inside MPA (x-axis), and pre-MPA depletion (y-axis).
 Pre-MPA depletion is a measure of fishing pressure, where 0 means that the population is unfished, and 1 means that the population is extinct in the time period immediately prior to MPA implementation.
@@
 To date, this assumption is insufficiently tested, and this is the focus of our paper.
-Our results show that regional conservation benefits of MPAs are highly context dependent and in many circumstances, are likely to be small enough that they are nearly impossible to detect empirically.
+Our results show that population-level conservation benefits of MPAs are highly context dependent and in many circumstances, are likely to be small enough that they are nearly impossible to detect empirically.
 This is exactly what we found in our empirical case study from the relatively large and well-studied network of MPAs in the Channel Islands, California, USA.
@@
 But as suggested in @white2011, @ferraro2018, @larsen2019 and further demonstrated here, without careful attention to the design of control sites (e.g.
-accounting for the displacement of fishing effort by MPAs), response ratios may be highly unreliable estimators of regional MPA effects.
-When MPAs affect nearby control sites used in response ratios through biological spillover or concentration of fishing effort, it is entirely possible to for MPA to produce massive response ratios while simultaneously having minimal effects on the entire population partially protected by the MPAs.
+accounting for the displacement of fishing effort by MPAs), response ratios may be highly unreliable estimators of population-level MPA effects.
+When MPAs affect nearby control sites used in response ratios through biological spillover or concentration of fishing effort, it is entirely possible to for MPA to produce massive response ratios while simultaneously having minimal effects on the entire population partially protected by the MPAs (and *vice versa*.
 As @larsen2019 suggests, there are many potential alternatives for estimating the effects of MPAs that better account for the challenges of causal inference (though that may be more data-intensive).
 We applied one such approach here (a difference-in-difference estimator), and yet were still unable to reach robust conclusions as to the effect of MPAs on the biomass density of targeted finfish in the Channel Islands, due to the likely small size of the true effect relative to the influence of environmental variability.
-If pre-MPA data are available, data currently used in response ratios can easily be re-purposed for an analysis such as this one; we encourage holders of such data worldwide to conduct similar exercises to examine to what extent empirical estimates of regional conservation effects match up with MPA theory.
+If pre-MPA data are available, data currently used in response ratios can easily be re-purposed for an analysis such as this one; we encourage holders of such data worldwide to conduct similar exercises to examine to what extent other empirical estimates of population-level MPA effects conform to MPA theory.
 
-While this does not mean that all MPAs will face similar challenges in estimating their effects, our results in the relatively large, well-enforced, and rigorously studied Channel Islands Marine Protected Area network make clear that in many instances empirically detecting a clear regional effect of MPAs on total biomass densities of targeted finfish may not be possible.
+While this does not mean that all MPAs will face similar challenges in estimating their effects, our results in the relatively large, well-enforced, and rigorously studied (though also relatively lightly fished) Channel Islands Marine Protected Area network make clear that in many instances empirically detecting a clear effect of MPAs on total biomass densities of targeted finfish at the population level may not be possible.
 How then should stakeholders go about adaptively monitoring and managing MPAs?
@@
 This is particularly true if communities are depending on MPAs as their primary marine resource management tool.
-Commonly employed metrics such as response ratios may be applicable in some circumstances, but can have severe shortcomings as metrics of regional conservation effects.
+Commonly employed metrics such as response ratios may be applicable in some circumstances,for example, when an MPA is designed for food security in the face of a disaster or for protection of unique or threatened species inside its borders, but can have shortcomings as metrics of population-level effects.
 Dependence on unreliable estimators of MPA effects may lead to stakeholders incorrectly attributing negative environmental shocks as MPA failures, or interpreting data arising from scorched-earth fishing outside MPAs as a conservation success.
@@
 Readers can explore the functionality of the model using an online tool available [here](https://danovando.shinyapps.io/simmpa/).
-The model consists of 50 patches with wrapped edges (picture the waters around a circular island).
-For any one simulation we randomly pull a species and its associated life history from the `FishLife` [@thorson2017c] package in R.
-We pair these data with randomly selected values between 0.6 and 0.95 for Beverton-Holt steepness [as parameterized in @mace1994], as well as larval and adult dispersal rates.
-We randomly assign whether adult fish preferentially move towards patches with lower relative densities, as well as one of three potential types of recruitment density dependence [building off of @babcock2011].
-Simulations are also assigned random fleet dynamics (open access ,constant effort, constant catch), responses to MPA (leave the fishery or concentrate outside MPAs), a fleet dispersal model (uniform in response to catch-per-unit-effort, in response to profit-per-unit-effort),
-
-<!--   1. Local density dependence: Density dependence occurs independently in each patch, and recruits then disperse to nearby patches -->
-
-<!--   2. Global density dependence: Density dependence is a function of the sum of spawning biomass across all patches, and recruits are then distributed according to habitat quality -->
-
-<!--   3. Post-dispersal density dependence: Larvae are distributed throughout the system, and then density dependence occurs based on the density of adult biomass at the destination patch -->
-
-<!-- We allow for three potential siting strategies for MPAs.
-In the first, MPAs are randomly placed.
-In the second, we assume that MPAs are placed in preferentially better habit (unfished recruitment is four times greater inside MPA locations).
-In the third, we allow for scenarios in which MPAs are placed in sources of larval dispersal.
--->
-
-<!-- Each simulation is randomly assigned a fleet model of the form  -->
-
-<!--   1. Open access: fishing effort changes in response to profit-per-unit-effort -->
-
-<!--   2. Constant effort: total fishing effort is constant over time (unless altered by MPA displacement model) -->
-
-<!--   3. Constant catch: the fleet exerts as much effort as need to achieve a target catch -->
-
-<!-- This fleet model is paired with a gear selectivity ranging from .1 to 1.5 of the length at 50% maturity for the species in question, and the fleet model is tuned to achieve a target fishing mortality relative to natural mortality ratio at equilibrium.
--->
-
-<!-- Along with the fleet dynamics model, each simulation is assigned a random fleet dispersal scenario: uniform dispersal (where the total effort of the fleet is divided evenly among all open patches), catch dispersal (where the total effort of the fleet is divided according to the catchable biomass in each available patch), and profit dispersal (where the total effort of the fleet is divided according to the profit per unit effort in each available patch).
-Fishing effort that occurred inside MPAs prior to closer can either leave the fishery, or be distributed to the patches outside the MPAs.
--->
-
-Each simulation is assigned an MPA scenario, defined by the number and size of MPAs, the placement of those MPAs, and the year that the MPAs are put in place.
-The population begins at unfished equilibrium and then fishing effort is applied in accordance to the fleet model.
-The MPAs are then placed during the randomly selected start year, allowing some runs to explore how the early dynamics of the MPA play out when the fishery and population they are placed on is not already at equilibrium.
-Each simulation is run to equilibrium with and without the selected MPA strategy (holding all else constant).
-We then measure the difference in biomass densities each time step in the scenario with and without the MPAs to calculate the regional conservation effect of the MPAs over time.
+Full details of the simulation model are available in the Supplementary Information.
 
@@
 Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment (*MPA*) on the treated (targeted species).
+$\beta_0$ is the mean total biomass of non-targeted species pre-MPA, and $\beta_1T$ is the mean total biomass of targeted species pre-MPA.
+$\beta_2MPA$ is the effect of the post-MPA period on the non-targeted species.
 The shape and scale parameters of the Gamma distribution are estimated as well.

### 2020-04-24 make-pop-effects-of-mpas.R

@@ -1,0 +1,4284 @@
+# Run Complete Zissou Analysis --------------------------------------------
+
+
+# libraries ---------------------------------------------------------------
+set.seed(42)
+library(scales)
+library(viridis)
+library(ggmap)
+library(forcats)
+library(stringr)
+library(lubridate)
+library(lme4)
+library(TMB)
+library(TMBhelper)
+library(FishLife)
+library(patchwork)
+library(rstan)
+library(rstanarm)
+library(extrafont)
+library(hrbrthemes)
+library(here)
+library(doParallel)
+library(spasm)
+library(furrr)
+library(ggsci)
+library(recipes)
+library(sf)
+library(ggmap)
+library(rEDM)
+library(bayesplot)
+library(tabulizer)
+library(tidyverse)
+library(ggtext)
+library(ggrepel)
+library(rnaturalearth)
+library(rnaturalearthhires)
+extrafont::loadfonts()
+# extrafont::font_import()
+rstan_options(auto_write = TRUE)
+Sys.unsetenv("PKG_CXXFLAGS")
+
+
+functions <- list.files(here::here("functions"))
+
+walk(functions, ~ here::here("functions", .x) %>% source()) # load local functions
+
+
+# options -----------------------------------------------------------------
+
+run_name <- 'v6.0'
+
+run_description <- "PNAS R and R with simplified DiD and weighting"
+
+# the following analysis run the complete contents of "regional-effects-of-mpas". Each section depends on the out
+# outcomes of the prior section, but will load relevant saved files.
+# So, once you've run simulate_mpas, you can set it to FALSE and validata_mpas will work
+
+
+run_did <- FALSE # run difference in difference on data from the CINMS
+
+run_tmb <- FALSE
+
+simulate_mpas <- FALSE # simulate MPA outcomes
+
+simulate_channel_islands <- FALSE # simulate MPA outcomes
+
+validate_mpas <- FALSE
+
+process_results <- TRUE
+
+knit_paper <- FALSE
+
+get_cdfw_catches <-  FALSE
+
+sim_years <- 50
+
+num_patches <- 50
+
+n_cores <- 1
+
+# prepare run -------------------------------------------------------------
+
+run_dir <- here::here("results", run_name)
+
+if (dir.exists(run_dir) == F) {
[…]

### 2020-04-24 documents/ovando-et-al-pop-effects-of-mpas-si.Rmd
add TMB plot back in

@@
 
+```{r tmb-fit, fig.cap="Difference-in-difference results from mixed effect model fit to the individual transect observation level of the data."}
+tmb_runs$did_plot[[1]] + 
+scale_y_continuous(name = "Estimated MPA Effect")
+```
+
 A complete table of estimated coefficients can be seen in Table.S\@ref(tab:betatable).
@@
   mutate(did_plot = map(processed_fits, "did_plot"))
+
+fit_report <- tmb_runs$tmb_fit[[1]]$zissou_report
+
+observed_seen <- tmb_runs$tmb_fit[[1]]$zissou_data$log_density
+
+observed_seeing <- tmb_runs$tmb_fit[[1]]$zissou_data$any_seen
+
+seen_diagnostics <- data_frame(observed = observed_seen, predicted = as.numeric(fit_report$log_density_hat)) %>% 
+  mutate(residuals = predicted - observed)
+
+seeing_diagnostics <- data_frame(observed = observed_seeing, predicted = as.numeric(fit_report$prob_seeing))
 
@@
 The quantile-quantile plot suggest that on the assumption of log-normal errors on the observed densities is reasonable, though the model appears to have some slight problems estimating the highest observed densities (Fig.S\@ref(fig:obs-v-pred-plot)).
-
-```{r}
-
-fit_report <- tmb_runs$tmb_fit[[1]]$zissou_report
-
-observed_seen <- tmb_runs$tmb_fit[[1]]$zissou_data$log_density
-
-observed_seeing <- tmb_runs$tmb_fit[[1]]$zissou_data$any_seen
-
-seen_diagnostics <- data_frame(observed = observed_seen, predicted = as.numeric(fit_report$log_density_hat)) %>% 
-  mutate(residuals = predicted - observed)
-
-seeing_diagnostics <- data_frame(observed = observed_seeing, predicted = as.numeric(fit_report$prob_seeing))
-
-```
 
@@
 
-```{r raw-synth, include = FALSE}
-
-years <- unique(used_data$year)
-
-life_history <- used_data %>%
-  select(classcode) %>%
-  unique() %>%
-  mutate(numeric_classcode = as.numeric(as.factor(classcode)))
-
-nontargeted_index <- data_frame(abundance_hat = fit_report$abundance_hat,
-                               classcode = rep(1:n_distinct(used_data$classcode), each = length(years))) %>%
-  mutate(log_abundance_hat = log(abundance_hat)) %>%
-  group_by(classcode) %>%
-  mutate(year =years) %>%
-  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
+```{r  total-synth, fig.cap = "Difference in centered and scaled standardized targeted abundance and synthetic standardized targeted abundance"}
+total_gap_plot
+```
+
+```{r raw-synth, include = FALSE, eval = FALSE}
+
+  consistent_classcodes <- used_data %>%
+      group_by(classcode, region) %>%
+      filter(any_seen) %>%
+      summarise(n_years = n_distinct(year)) %>%
       ungroup() %>%
-  rename(numeric_classcode = classcode) %>%
-  left_join(life_history, by = "numeric_classcode" ) %>% 
-  left_join(unique(abundance_data$data[[1]] %>% select(classcode, targeted)), by = "classcode") %>% 
-  filter(targeted == 0) %>% 
-  select(log_abundance_hat, year, classcode, targeted)
-
-targeted_index <- tmb_runs$tmb_fit[[1]]$zissou_estimates %>%
-  filter(
-    stringr::str_detect(variable, "(targeted_did_betas)"),!str_detect(variable, "non")
+      filter(n_years == n_distinct(used_data$year))
+
+synth_data <- used_data %>% 
+  filter(classcode %in% consistent_classcodes$classcode,
+         region %in% consistent_classcodes$region) %>% 
+  mutate(classcode = ifelse(targeted == 1, "targeted", classcode),
+         region = ifelse(targeted == 1, "targeted", region)) %>% 
+      group_by(year,
+               site_side,
+               region,
+               zone,
+               transect,
+               eventual_mpa,
+               classcode,
+               targeted) %>%
+      summarise(
+        total_classcode_density = sum(exp(log_density)),
+        var_tex = sum(cumulative_n_obs),
+        var_vis = mean(mean_vis),
+        var_temp = mean(mean_temp),
+        var_depth = mean(mean_depth),
+        var_surge = mean(surge),
+        var_kelp = mean(interp_kelp)
+      ) %>%  # sum density across all levels of a transect
+      group_by(year, site_side, region, eventual_mpa, classcode, targeted) %>%
+      summarise(
+        md = mean(total_classcode_density),
+        var_tex = mean(var_tex),
+        var_vis = mean(var_vis),
+        var_temp = mean(var_temp),
+        var_depth = mean(var_depth),
+        var_surge = mean(var_surge),
+        var_kelp = mean(var_kelp)
+      ) %>% # calculate mean density per year site, side, species, averaging over zone, transect
+      ungroup()
+
+ synth_data <- synth_data %>% 
+     group_by(year,region,classcode,targeted) %>%
+      summarise(
+        total_biomass_density = (sum(md) / 1e6) * 10000,
+        # calculate total and mean biomass densities across all species per year site side
+        mean_biomass_density = log((mean(md) / 1e6) * 10000 + 1e-3),
+        var_tex = mean(var_tex),
+        var_vis = mean(var_vis),
+        var_temp = mean(var_temp),
+        var_kelp = mean(var_kelp)
       ) %>%
-  mutate(
-    year = years,
-    classcode = "targeted",
-    targeted = 1
-  ) %>%
-  rename(log_abundance_hat = estimate) %>% 
-  select(colnames(nontargeted_index))
-
-  # filter(variable == "seeing_year_species_betas") %>%
-
-nontargeted_covariates <- used_data %>% 
-  select(classcode, contains("mean"),year, targeted) %>% 
-  filter(targeted == 0) %>% 
-  select(-mean_length) %>% 
-  gather(variable, value, -classcode,-year,-targeted) %>% 
-  group_by(classcode, variable, year) %>% 
-  summarise(mean_value = mean(value, na.rm = TRUE)) %>% 
-  spread(variable, mean_value)
-
-targeted_covariates <- used_data %>% 
-  select(classcode, contains("mean"),year, targeted) %>% 
-  filter(targeted == 1) %>% 
-  mutate(classcode = "targeted") %>% 
-  select(-mean_length) %>% 
-  gather(variable, value, -classcode,-year,-targeted) %>% 
-  group_by(classcode, variable, year) %>% 
-  summarise(mean_value = mean(value, na.rm = TRUE)) %>% 
-  spread(variable, mean_value)
-
-synth_covariates <- nontargeted_covariates %>% 
-  bind_rows(targeted_covariates)
-
-synth_data <- targeted_index %>% 
-  bind_rows(nontargeted_index) %>% 
-  left_join(synth_covariates, by = c("classcode","year")) %>% 
+   ungroup() %>% 
+   mutate(classcode = ifelse(targeted == 1, classcode, paste(classcode, region, sep = '_'))) %>% 
       mutate(numeric_classcode = as.numeric(factor(classcode))) %>% 
-  group_by(classcode) %>% 
-  mutate(log_abundance_hat = scale(exp(log_abundance_hat))) %>% 
-  ungroup() %>% 
-  as.data.frame() %>% 
-  select(-rolling_mean)
-
-# synth_data %>%
-#   group_by(classcode) %>%
-#   mutate(index = (log_abundance_hat)) %>%
-#   ggplot(aes(
-#     year,
-#     index,
-#     color = factor(targeted),
-#     group = interaction(targeted, classcode)
-#   )) +
-#   geom_line()
+   as.data.frame()
 
@@
     foo = synth_data,
-    predictors = colnames(synth_data)[str_detect(colnames(synth_data),"mean")],
+    predictors = colnames(synth_data)[str_detect(colnames(synth_data),"var")],
     predictors.op = "mean",
-    dependent = "log_abundance_hat",
+    dependent = "mean_biomass_density",
     unit.variable = "numeric_classcode",
@@
     treatment.identifier = "targeted",
-    controls.identifier = unique(synth_data$classcode)[unique(synth_data$classcode) != "targeted"],
+    controls.identifier = unique(synth_data$classcode[synth_data$targeted == 0]),
     time.predictors.prior = years[years <= 2003],
@@
   ggplot(aes(year, gap)) + 
-  geom_hline(aes(yintercept = 0), linetype = 2) +
+  # geom_hline(aes(yintercept = 0), linetype = 2) +
   geom_vline(aes(xintercept = 2003), color = "red")+
@@
 
-```
-
-```{r  total-synth, fig.cap = "Difference in centered and scaled standardized targeted abundance and synthetic standardized targeted abundance"}
-total_gap_plot
 ```
@@
 nontargeted_index <- data_frame(abundance_hat = fit_report$abundance_hat,
-                               classcode = rep(1:n_distinct(fitted_data$classcode), each = length(years))) %>%
+                               classcode = rep(1:n_distinct(used_data$classcode), each = length(years))) %>%
   mutate(log_abundance_hat = log(abundance_hat)) %>%

### 2020-04-24 documents/ovando-et-al-pop-effects-of-mpas.Rmd
add TMB plot back in

@@
 As such, we weight each observation such that within MPA data are assigned a total weight of 0.2 and outside-MPA data a total weight of 0.8.
+See SI for sensitivity analyses of various model structures and synthetic controls results.
 
@@
 We used convergent cross mapping (CCM), in the manner of @clark2015, to test for the possibility of the trophic cascades biasing our results.
-Generalizations of Takens' theorem indicate that if two variables are part of the same dynamic system, their individual dynamics should reflect their relative causal influence.
-Convergent cross mapping (CCM) tests for causation by using the attractor/manifold built from the time series of one variable to predict another (hence the "cross-mapping").
-CCM then allows us to test for causal relationships in the timeseries of densities of targeted and non-targeted species.
-Our results found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (see SI for additional information in CCM testing).
+Our results found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (Fig.SXX).
 However, the longer MPAs are in place, the greater the possibility that substantial species interactions that can affect use of non-targeted species as a control may arise.

### 2020-04-26 documents/ovando-et-al-pop-effects-of-mpas.Rmd
last round of edits completed

@@
 While numerous studies show that MPAs produce conservation benefits *inside* their borders, many MPAs are also justified on the grounds that they confer conservation benefits to the broader population *beyond* their borders.
-This paper examines the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and shows that even in cases when the conservation benefits are large, they are inherently difficult to detect empirically.
+This paper examines the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and shows that even in cases where the conservation benefits are large, they are inherently difficult to detect empirically.
 A network of MPAs was put in place in The Channel Islands National Marine Sanctuary in 2003, with a goal of providing regional conservation and fishery benefits.
@@
 
-## author_contributions: 
+## Author Contributions: 
   D.O., S.G., and R.H.
@@
 
-## conflict_of_interest:
+## Conflicts of Interest:
   R.H's research program receives funding from environmental NGOs, foundations, fishing industry, governments and international agencies.
@@
 C.C., S.G., and O.D's research program includes funding from environmental NGOs and foundations with an interest in ocean conservation and food security.
-## significance
+## Significance Statement
   Healthy marine ecosystems are critical to the well-being of the planet.
 Marine protected areas (MPAs) are increasingly being used to conserve and manage these ecosystems.
-Our study pairs theory with empirical methods to show that the population-level conservation effects of marine protected areas can be highly variable and dependent on human drivers.
-We find no statistically clear effect of a large and well protected network of MPAs in the Channel Island, California, on targeted finfish after 14 years of protection, which our simulation analysis suggests is actually to be expected.
+We pair theory with empirical methods to show that the population-level conservation effects of marine protected areas can be highly variable and challenging to detect.
+We find no statistically clear effect of a large and well protected network of MPAs in the Channel Island, California, on targeted finfish after 14 years of protection, which our simulation analysis suggests is in fact to be expected.
 MPA practitioners must clearly communicate expectations from MPA networks, and plan monitoring and adaptations strategies that directly address the challenges of estimating population-level conservation effects of protected areas.
@@
 
-We define the population-level conservation effects of MPAs as the change in total mean biomass densities of targeted finfish both inside and outside of MPAs, relative to the total mean biomass densities of targeted finfish inside and outside of MPAs that would have occurred without the MPAs.
+We define the population-level conservation effects of MPAs as the change in mean total biomass densities of targeted finfish both inside and outside of MPAs, relative to the mean total biomass densities of targeted finfish inside and outside of MPAs that would have occurred without the MPAs.
 As an example of a population-level MPA effect, consider a evenly distributed population that has 50% of its range protected by an MPA.
@@
 Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species).
-We used a Bayesian difference-in-difference regression to estimate any difference in total mean biomass densities of fin-fish species targeted by fishing effort (i.e., those potentially affected by an MPA) and those species not targeted by fishing before and after MPA implementation [see @larsen2019 for an explanation of difference-in-difference estimators in ecological settings].
-The result of this regression is, conditional on the assumptions of the model, an estimate of the effect of the MPAs on the total mean biomass densities of targeted species throughout the Channel Islands.
+We used a Bayesian difference-in-difference regression to estimate any difference in mean total biomass densities of fin-fish species targeted by fishing effort (i.e., those potentially affected by an MPA) and those species not targeted by fishing before and after MPA implementation [see @larsen2019 for an explanation of difference-in-difference estimators in ecological settings].
+The identification strategy control for unobserved environmental shocks to the system independent of the MPAs and similarly experienced by the targeted and non-targeted species.
+Conditional on the assumptions of the model, this regression produces an estimate of the effect of the MPAs on the mean total biomass densities of targeted species throughout the Channel Islands.
 
@@
          }
-did_coefs <- did_model %>% broom::tidy(intervals = TRUE)
+did_coefs <- did_model %>% broom::tidy(intervals = TRUE,conf.level = 0.9)
 
@@
 - `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`).
-Over the next six years the model estimates greater probabilities of an increasingly positive MPA effect, peaking in 2009-2011 with a median estimate of MPA effect of a `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)` increase in total mean biomass density of targeted species (90% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)`
+Over the next six years the model estimates greater probabilities of an increasingly positive MPA effect, peaking in 2009-2011 with a median estimate of MPA effect of a `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)` increase in mean total biomass density of targeted species (90% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)`
 - `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)` (Fig.\@ref(fig:mpa-effect)).
@@
 
-```{r mpa-effect, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Results of difference-in-difference regression estimating the population effect of the Channel Island MPAs on total mean biomass densities of targeted species (difference in total mean biomass density of targeted species over time relative to expected levels using non-targeted species as a control).
+```{r mpa-effect, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Results of difference-in-difference regression estimating the population effect of the Channel Island MPAs on mean total biomass densities of targeted species (difference in mean total biomass density of targeted species over time relative to expected levels using non-targeted species as a control).
 Grey distributions show posterior probability distribution of estimated MPA effect; red point is median estimated effect, thicker red section 50% credible interval, thinner red line 90% credible interval).
@@
 
-Our simulation results suggest that we should not be surprised by this result.
+  Simulation analysis suggests that we should not be surprised by this result.
 The Channel Islands MPAs cover 20% of the surface waters in the Channel Islands, and while formal stock assessments are not available for many of the targeted species in our analysis, what evidence we have suggests that, as a group, these fish are on average not heavily overfished.
 Some species, such as California sheephead and blue rockfish were likely below target levels during the period [@alonzo2004;@dick2017], but projections based upon the overall average response across all species will likely suggest modest benefits even if a subset of species could experience much larger population gains.
-Our simulations suggest that the average percentage difference in densities of targeted species with and without MPAs to be modest (Fig.\ref{fig:mpa-effect}).
-Effects of this size are likely to be challenging to detect empirically given the large natural variation of marine ecosystems (especially temperate reefs) and the observation error inherent in infrequent (e.g.
+Under such circumstances, the average percentage difference in densities of targeted species with and without MPAs is likely to be modest (Fig.\ref{fig:mpa-effect}).
+Effects of this size will often be challenging to detect empirically given the large natural variation of marine ecosystems (especially temperate reefs) and the observation error inherent in infrequent (e.g.
 annual) common in marine systems.
@@
 
-Our simulations and difference-in-difference model examine percentage changes in biomass densities (as oppose to changes in total population size).
-An increase in biomass densities from 0.02kg/m^2^ to 0.04kg/m^2^ would be produce as a 100% increase.
+Our simulations and difference-in-difference model examine percentage changes in biomass densities (as opposed to changes in total population size).
+An increase in biomass densities from 0.02kg/m^2^ to 0.04kg/m^2^ would be measured as a 100% increase.
 While a large percentage effect, this is a small change in biomass densities relative to the variance in the observation process itself.
-While an MPA providing 25% protection might be expected to produce a 30% increase in biomass densities, expressed as a percentage of unfished biomass this gain is on the order of 15% (Fig.SXX).
-In addition, the median estimated age at sexual maturity for the targeted species included in this study is `r median_tm` years, meaning that the span of this analysis represents less than three generations of MPA protection for half of the the measured species.
-Ongoing monitoring may yet reveal clearer effects.
+While an MPA providing 25% protection might be expected to produce a 30% increase in biomass densities, expressed as a percentage of unfished biomass, this gain is on the order of 15% (Fig.SXX).
+In addition, the median estimated age at sexual maturity for the targeted species included in this study is `r median_tm` years, meaning that the span of this analysis represents less than three generations of MPA protection for half of the measured species.
+Ongoing monitoring may yet reveal clearer effects, especially if the time series extends beyond the periods of environmental shocks.
 Analysis of more rapidly growing and maturing species, e.g.
@@
 Our analysis suggests that we would need an MPA effect size of at least 30% to achieve this.
-Two of the most critical drivers of MPA performance are the size of the MPA and the degree of fishing pressure.
+Two of the most critical drivers of MPA performance are the size of the MPA network and the degree of fishing pressure.
 Looking across these two variables, if the MPA network covers 25% or more of a species range and pre-MPA depletion is greater than 60% we might expect an effect size with our target MAPE.
-While recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, for near-shore commercial fin-fish many MPA networks are likely to cover areas more in line with the Channel Islands (20%) or smaller, and as such have population-level effect sizes that may be difficult to detect (Fig.\@ref(fig:sim-effects)-A).
+Although recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, most MPA networks for near-shore commercial finfish are likely to cover areas more in line with the Channel Islands (20%) or smaller, and as such have population-level effect sizes that may be difficult to detect unless they were experiencing extreme overfishing prior to MPA establishment (Fig.\@ref(fig:sim-effects)-A).
 
@@
 
-To put these results into context, the FAO estimates that 7% of the worlds fisheries with status estimates fall into the relatively unexploited category (roughly depletion less than 50%), 60% fall into the fully fished category (roughly depletion 50%-70%), and 33% fall into the heavily fished category (roughly depletion greater than 70%) [@fao2018, assuming a biomass at maximum sustainable yield to unfished biomass ratio of 40%, per @thorson2012], though works that include a broader range of fisheries estimate that 50% or more of stocks to fall into the heavily fished category [@costello2012;@rosenberg2018].
-The population-level effects of MPA networks covering 25% or less of a species range may be difficult to detect in many places with already well-managed fisheries, while for that size we might expect clearer effects in less-managed locations (though that of course ignores the complication of compliance with MPA regulations).
+To put these results into context, the FAO estimates that 7% of the worlds fisheries with status estimates fall into the relatively unexploited category (roughly depletion less than 50%), 60% fall into the fully fished category (roughly depletion 50%-70%), and 33% fall into the heavily fished category (roughly depletion greater than 70%) [@fao2018, assuming a biomass at maximum sustainable yield to unfished biomass ratio of 40%, per @thorson2012], although studies that include a broader range of fisheries estimate that 50% or more of stocks fall into the heavily fished category [@costello2012;@rosenberg2018].
+The population-level effects of MPA networks covering 25% or less of a species range may be difficult to detect in many places with already well-managed fisheries, while for that MPA network size we might expect clearer effects in less-managed locations (assuming compliance with MPA regulations).
 Within these broad guidelines a wide range of outcomes are possible based on local fleet and fish dynamics: as a starting place users can use the bio-economic MPA simulation model developed for this paper to explore potential outcomes for specific MPAs using an interactive web application available [here](https://danovando.shinyapps.io/simmpa/).
@@
 Our simulation model allows users to place this collective theory into context, clarifying under what states of the world we might expect to see particular effect sizes, and demonstrating what effect sizes we might reasonably expect to be detectable in the real world.
-In the case of the Channel Islands, simulation modeling suggests for the network in question, an effect size on the order of 30% or smaller is likely, which our analysis in turn demonstrates may be challenging to detect.
+In the case of the Channel Islands, simulation modeling suggests for the network in question, an effect size on the order of 30% or smaller is likely, which our analysis in turn demonstrates may be challenging to detect even with a comprehensive monitoring program.
 This cautionary tale is likely to hold for many coastal MPA networks around the world.
@@
 This is particularly true if communities are depending on MPAs as their primary marine resource management tool.
-Commonly employed metrics such as response ratios may be applicable in some circumstances,for example, when an MPA is designed for food security in the face of a disaster or for protection of unique or threatened species inside its borders, but can have shortcomings as metrics of population-level effects.
+Commonly employed metrics such as response ratios may be applicable in some circumstances,for example, when an MPA is designed for food security in the face of a disaster or for protection of unique or threatened species inside its borders.
+But, response ratios can also have shortcomings as metrics of population-level effects.
 Dependence on unreliable estimators of MPA effects may lead to stakeholders incorrectly attributing negative environmental shocks as MPA failures, or interpreting data arising from scorched-earth fishing outside MPAs as a conservation success.
@@
 Statistical approaches that explicitly address complications such as the spatial spillover effects of MPAs (such as the difference-in-difference approach used here) may give users an improved understanding of the performance of their MPAs, but even they may struggle when expected effect sizes are small.
-Clearly communicating what we should expect, and what we can detect, from MPAs is critical in ensuring that MPAs play effective roles in fisheries management and marine conservation.
+Clearly communicating what we should expect, and what we can detect, from MPAs is critical to ensuring that MPAs play effective roles in fisheries management and marine conservation.
 
@@
 Our main difference-in-difference model was fit using Stan [@carpenter2017] using the `rstanarm` package [@goodrich2020].
-All materials needed to replicate results can be found [here](https://github.com/DanOvando/regional-effects-of-mpas).
+All materials needed to reproduce results can be found [here](https://github.com/DanOvando/regional-effects-of-mpas).
 

### 2020-04-26 make-pop-effects-of-mpas.R

@@ -56,7 +56,7 @@ run_description <- "PNAS R and R with simplified DiD and weighting"
 # So, once you've run simulate_mpas, you can set it to FALSE and validata_mpas will work
 
 
-run_did <- TRUE # run difference in difference on data from the CINMS
+run_did <- FALSE # run difference in difference on data from the CINMS
 
 run_tmb <- FALSE
 
@@ -4093,6 +4093,8 @@ response_ratio_plot <-   targ_rr %>%
     fill = source),
   alpha = 0.8,
   stat = "binline",
+  color = "black",
+  size = .1,
   show.legend = FALSE,
   bins = 15) +
   scale_x_continuous(name = title, limits = c(NA, 2)) +
@@ -4188,7 +4190,7 @@ mpa_effect_plot <- ggplot() +
   tidybayes::stat_halfeye(
     data = did_results,
     aes(year, did, color = "Empirical Estimate"),
-    .width = c(0.5, 0.95),
+    .width = c(0.5, 0.9),
     slab_alpha = 0.95  ) +
   scale_y_continuous(labels =ylabs, breaks = ybreaks, name = "Population-Level MPA Effect") +
   scale_x_discrete(name = element_blank()) +

### 2020-04-26 test-vast.R

@@ -1,205 +1,0 @@
-library(tidyverse)
-library(TMB)
-#library(VAST)
-library(sdmTMB)
-
-
-# assume loading in data already
-
-pisco_abundance_data <-
-  abundance_data$data[abundance_data$data_source == "pisco"][[1]]
-
-
-
-empirical_biomass_densities  <- pisco_abundance_data %>%
-  filter(classcode %in% unique(fitted_data$classcode)) %>%
-  group_by(year, site_side, region, zone, eventual_mpa, classcode, targeted) %>%
-  summarise(total_classcode_density = sum(exp(log_density)),
-            mean_experience  = mean(cumulative_n_obs),
-            mean_vis = mean(mean_vis),
-            mean_depth = mean(mean_depth),
-            mean_canopy = mean(mean_canopy)) %>%
-  group_by(year, site_side, classcode) %>%
-  summarise(
-    total_biomass_density = sum(total_classcode_density),
-    mean_biomass_density = mean(total_classcode_density),
-    mean_experience  = mean(mean_experience),
-    mean_vis = mean(mean_vis),
-    mean_depth = mean(mean_depth),
-    mean_canopy = mean(mean_canopy)
-  ) %>%
-  ungroup()
-
-empirical_biomass_densities %>% ggplot(aes(mean_biomass_density)) + geom_histogram() +
-  facet_wrap( ~ classcode, scales = "free_x") + 
-  theme_minimal()
-
-
-classcodes <- unique(empirical_biomass_densities$classcode)
-
-
-standardizer <- function(temp_classscode, pknots = 0.5){
-
-temp <- empirical_biomass_densities %>%
-  filter(classcode == temp_classscode) %>%
-  # select(year, site_side, mean_biomass_density) %>%
-  left_join(
-    site_data %>% select(site, side, lon_wgs84, lat_wgs84) %>% unite(col = site_side, site:side, sep = '-'),
-    by = c("site_side")
-  ) %>%
-  rename(X = lon_wgs84,
-         Y = lat_wgs84) %>% 
-  mutate(mean_experience = scale(mean_experience)) %>% 
-  mutate(mean_experience_2 = mean_experience^2,
-         mean_vis = scale(mean_vis))
-
-temp %>%
-  ggplot(aes(X, Y, color = mean_biomass_density)) +
-  geom_point() +
[…]

### 2020-04-26 documents/reviewer-responses.Rmd
resubmited to PNAS

@@
 ---
-title: "Reviewer Responses for 'The Population Effects of Marine Protected Areas Are Likely To Be Unclear' (previously 'The Regional Effects of MPAs')"
+title: Reviewer Responses for "Assessing the Population-level Conservation Effects of Marine Protected Areas" (previously "The Regional Effects of MPAs")
 author: 
@@
 output:
+  bookdown::pdf_document2: default
+  html_document: default
   word_document: default
-  html_document: default
-  pdf_document: default
 bibliography: My Library.bib
@@
 
+This is a re-submission of MS# 2019-17437 (formerly titled "The Regional Effects of Marine Protected Areas").
 We thank the two reviewers and the Editor for their constructive responses.
 We have fundamentally restructured the paper to place much greater emphasis on the results from the Channel Islands Marine Protected Area (MPA) network.
-This is to our knowledge the first empirical assessment of the regional effects of a large and iconic MPA network on a wide assemblage of commercially and economically important species.
+This is to our knowledge the first empirical assessment of the population-level effects of a large and iconic MPA network on a wide assemblage of commercially and economically important species.
 Our inability to estimate a clear population-level effect in such a well-enforced and well-studied location is we feel an important "null-result" both for scientists and MPA managers to be aware of.
@@
 
-While our results are more focused on the empirical results, we explicitly link our results to the large body of MPA theoretical work, using it to help illustrate why we might expect such a null result and help managers understand when they might be able to detect an effect.
-As MPAs are used as a core tool of marine resource management, our results provide a useful update to our empirical expectations for MPAs, that to date have been based almost entirely on inside-outside MPA response ratios.
+While our results are more focused on the empirical results, we explicitly link our results to the large body of MPA theoretical work, using it to help illustrate why we might expect such a null-result and help managers understand when they might be able to detect the population-level effects of MPA networks.
+As MPAs are increasingly used as a core tool of marine resource management, our results provide a useful update to our empirical expectations for MPAs, that to date have been based almost entirely on inside-outside MPA response ratios.
 
@@
 However, we feel there is substantial value in revisiting those discussions in the context of a concrete empirical example, to provide scientists and practitioners alike a cohesive reference to observe and understand the full ramifications of these past findings on our practical ability to monitor and manage marine protected area networks.
-We have strived to make clearer though that the theoretical section is a reflection of broad body of theoretical work.
+We have tried to make clearer that the theoretical section is a reflection of broad body of previous work.
 
-We hope you find the updated version of our manuscript improved thanks to the inclusions of the careful points raised by the prior revision of this paper.
+We would like to call out two particular changes to the paper.
+We felt that term "regional effects" was needlessly vague and causing confusion.
+We have therefore re-framed our results around "population-level" effects, a more easily defined and we hope understandable metric.
+In particular, we feel that this framing places these results in clearer contrast to prior works such as Caselle et al.
+2015 that examined inside-outside MPA effects, as opposed to population-level effects.
+
+We have also replaced our empirical results with a slightly simplified model fit to aggregated data (as opposed to the raw transect-level fit reported before).
+The updated results are we feel much clearer to understand and equally empirically valid, and have the added benefit of allowing for a broader range of sensitivity analyses (all included in the Supplemental Information).
+They are also much easier for other users to reproduce.
+The transect-level results reported in the prior draft are still available in the Supplemental Information.
+
+We hope you find the updated version of our manuscript improved thanks to the inclusions of the careful points raised by the last round of reviews of this paper.
 
@@
 I recommend moving this up.*
-  - We apologize for not making this clearer in our earlier draft, and have included this key caveat in the abstract, statement of significance, and introduction (Line 98)
+  - We apologize for not making this clearer in our earlier draft, and have included this key caveat in the abstract, statement of significance, and introduction (Line 77)
 
@@
   - This is indeed a critical point that we have endeavored to correct.
-We have included enforcement under our summary of key enabling conditions for effective MPAs (line 124), as well as the discussion (line 382).
+We have included enforcement under our summary of key enabling conditions for effective MPAs (line 109), as well as the discussion (line 406).
 While we do not have published data on strength of enforcement within the MPAs, anecdotal evidence suggests it is relatively high, as well as evidence from the mainland California MPAs [@zellmer2018] indicates it is high.
-In addition, the presence of a clear increase in response ratios within the Channel Islands provides evidence that the MPAs  are large enough and sufficiently well-enforced as to provide meaningful protection within their borders (Line XX).
+In addition, the presence of a clear increase in response ratios within the Channel Islands provides evidence that the MPAs  are large enough and sufficiently well-enforced as to provide meaningful protection within their borders (Figure 1).
 
 - *Your discussion of the model also emphasizes that stock status prior to MPA designation is very key to outcomes; in the SI most of the species included for the Channel Island review have unknown status---which influences your discussion, and may better be described.*
-  - We make clearer that we iterate over unknowns such as pre-MPA stock status in our description of the Channel Islands simulations (line XX), and the implications of pre-MPA stock status in the discussion (Line XX).
+  - We make clearer that we iterate over unknowns such as pre-MPA stock status in our description of the Channel Islands simulations (line 132), and the implications of pre-MPA stock status in the discussion (Line 383, Figure 5).
 
@@
 The theoretical points we raise have indeed been pointed out in the literature before, and we feel our new manuscript makes that clearer.
-The Channel Islands MPAs are an iconic network that has served as a model for protected areas around the world.
-It is also incredibly well-studied and well-enforced.
-While previous studies have looked the effectiveness of the Channel Islands MPAs within their borders, this is the first study in the Islands (and globally of which we are aware) to attempt an estimate of the effect of the MPAs at the regional scale (both inside and outside of the MPAs themselves).
-That we are unable to estimate a clear regional effect in even this system is an important result that we feel warrants publication in PNAS.
+Specifically, we now use our simulation analysis to demonstrate when we might be able to detect population-level MPA effects, rather than focusing on the expected effect sizes as a function of traits that have been previously raised in the literature for
+  - The Channel Islands MPAs are an iconic network that has served as a model for protected areas around the world.
+It is also extremely well-studied and well-enforced.
+While previous studies have looked the effectiveness of the Channel Islands MPAs within their borders, this is the first study in the Islands (and globally of which we are aware) to attempt an estimate of the effect of the MPAs at the population scale (both inside and outside of the MPAs themselves).
+That we are unable to estimate a clear population-level effect in even this system is an important result that we feel warrants publication in PNAS.
 As MPAs are increasingly turned to tools for ocean conservation, our results provide an important update and caution to expectations based solely on inside-outside comparison studies.
@@
 - *I would also be more cautious regarding your conclusion about your model predicts what should be expected in the data analysis. It could be simply luck that there is agreement between the two.*
-  - We have made the language clearer that this concordance could be simple coincidence, and is not in any way proof of the validity of our results (Line XX) However, in light of recent concerns in the scientific literature around hypothesizing-after-results, we feel that it is critically important to ground our theoretical results in *a priori* predictions.
+  - We have made the language clearer that this concordance could be simple coincidence, and is not in any way proof of the validity of our results (Line 209) However, we feel that it is critically important to ground our theoretical results in theoretical predictions.
 In addition, as we mention in the discussion, we encourage MPA scientists and stakeholders to undertake such simulation modeling to consider the plausibility of their results.
@@
 We would note that this method is related to that described in @caselle2015 (previously reference #38), but is not the same.
-Our method explicitly estimates the net difference in the targeted and non-targeted biomass densities pre-and-post MPA as an estimate of the mean effect of the MPAs  both inside and outside their borders.
+Our method explicitly estimates the net difference in the targeted and non-targeted biomass densities pre-and-post MPA as an estimate of the mean population-level effect of the MPAs both inside and outside their borders.
 Caselle et al.
@@
 We hope we have made this clearer in the text, e.g.
-on line XX.
+on Line 89, Line 103, and Line 153.
 
-    With regards to the specific question about the possibility of the species range exceeding the sampling domain, this would be reflected as a reduction in the percent of the population protected by the MPAs: a 25km^2 MPA may provide near total protection for a sedentary species with brood-rearing, but be insignificant to a migratory tuna.
-To that end, our assessment of model performance as a function of MPA effect size addresses the question of how MPA size relative to species range affects our models performance.
+  - With regards to the specific question about the possibility of the species range exceeding the sampling domain, this would be reflected as a reduction in the percent of the population protected by the MPAs: a 25km^2 MPA may provide near total protection for a sedentary species with brood-rearing, but be insignificant to a migratory tuna.
+To that end, our assessment of model performance as a function of MPA effect size (Fig.4, Fig.5) addresses the question of how MPA size relative to species range affects our models performance.
 All else being equal the smaller a portion of a population's range is protected by MPAs, the smaller the expected effect size, and by extension the more challenge the estimation model will have.
 
-  We apologize that we did not present our prior simulation work in a sufficiently compelling manner, and failed to differentiate its results from the prior literature.
+  - We apologize that we did not present our prior simulation work in a sufficiently compelling manner, and failed to differentiate our results from the prior literature.
 As suggested by the reviewers and the editor, we have restructured the paper to place much greater emphasis on the critical results from the Channel Islands.
-However, we feel that the simulation results do provide vital context to our results, and provide tangible recommendations as to under what circumstances we might be able to estimate the regional effects of MPAs.
+However, we feel that the simulation results do provide vital context to our results, and provide tangible recommendations as to under what circumstances we might be able to estimate population-level effects of MPAs.
 The reviewer correctly points out that the individual components of our model are not themselves novel, and that the importance of fleet dynamics and fish dispersal have been previously discussed.
-As such as have greatly reduced the attention these results receive, except as a vehicle for the explicit consideration of the feasibility of estimating regional MPA effects.
+As such as have greatly reduced the attention these results receive, except as a vehicle for the explicit consideration of the feasibility of estimating population-level MPA effects.
 In addition, while these broad points have been made in the literature before, we feel there is value in cohesively demonstrating the ramifications of these diverse pieces of MPA theory (that generally are looked at only one or a few at a time) in once place, both through our Figure 5 and the accompanying web application.
@@
 It seems like you can remove most of this text and point the reader to that paper*
-  - We have greatly reduced the amount of text addressing response ratios, and restructured the response-ratio analysis to be linked explicitly with the empirical response ratios observed in the Channel Islands (in response to comments from the editor), in Figure 1.
+  - We have greatly reduced the amount of text addressing response ratios, and restructured the response-ratio analysis to be linked explicitly with the empirical response ratios observed in the Channel Islands (in response to comments from the Editor), in Figure 1.
 We feel this strategy makes much clearer how the results of this paper expand beyond the findings of Caselle et al.
 2015, and why the results of Caselle et al.
-2015 are evidence of within-MPA effects, but not of the question of regional effects addressed in this manuscript.
-Given that much of the MPA literature relies on response ratios as the empirical evidence for the effectiveness of MPAs, we feel that it is critically important to directly address the challenges of response ratios as measures of regional effects in this paper.
-We hope that the new manner, directly linking theoretical response ratios to the levels observed in the Channel Islands.
+2015 are evidence of within-MPA effects, but not of the question of population-level effects addressed in this manuscript.
+Given that much of the MPA literature relies on response ratios as the empirical evidence for the effectiveness of MPAs, we feel that it is critically important to directly address the challenges of response ratios as measures of population-level effects in this paper with an empirical example.
 
 - *To me the potential contribution of the paper starts on line 350, which is way too late in a paper of this size*
-  - We have restructured the paper so that it now focuses on the Channel Islands case study, and as suggested have moved this discussion up to the introduction (line 97).
+  - We have restructured the paper so that it now focuses on the Channel Islands case study, and as suggested have moved this discussion up to the introduction (Abstract, Significance Statement, Line 77).
 
@@
 The model is operating at the level of biomass densities.
-As such, even if a non-targeted species has a much bigger range than the targeted species (or even the sampling region), so long as the parallel trends assumption holds within the sampled portion of the non-targeted species range, the identification strategy may be valid.
+As such, even if non-targeted species have a much bigger range than the targeted species (or even the sampling region), so long as the parallel trends assumption holds within the sampled portion of the non-targeted species range, the identification strategy may be valid.
 Similarly, the model is agnostic as to the root causes of the parallel trends, so long as they exist.
-However, if for example the parallel trends are dependent on a particular temperature regime which during the study is disrupted (as is likely the case in our data), then the model may be reliable for the early years but not later.
-We have made this point clearer in the paper (line XX), and made clear that the lack of a clear effect in the recent years is likely due to an emerging violation of the parallel trends assumption, and not to a failure of the MPAs per say.
-However, this is still an important result, since it suggests that neither the response ratios nor the difference-in-difference method give us a robust estimate of the regional effects of the Channel Island MPAs.
-As we stress in the paper, that even in a network such as the Channel Islands we are left without a clear empirical effect at the regional level has critical ramifications for the design and monitoring of MPAs globally.
+However, if for example the parallel trends are dependent on a particular temperature regime which during the study is disrupted (as is likely the case in the later years of our data), then the model may be reliable for the early years but not later.
+We have made this point clearer in the paper (Line 276, Line 450), and made clear that the lack of a clear effect in the recent years is likely due to an emerging violation of the parallel trends assumption, and not to a failure of the MPAs per say.
+However, this is still an important result, since it suggests that neither the response ratios nor the difference-in-difference method give us a robust estimate of the population-level effects of the Channel Island MPAs.
+As we stress in the paper, that even in a network such as the Channel Islands we are left without a clear empirical effect at the population level has critical ramifications for the design and monitoring of MPAs globally.
 
@@
 These totals are derived from average biomass densities of individual species across multiple transects within each site, which are then summed across species to achieve the total mean biomass density at the site level.
-The averaging process in acts to smooth out some of the sampling variation, and reflects a broader discussion about the right degree of smoothing in spatio-temporal data.
+The averaging process acts to smooth out some of the sampling variation, and reflects a broader discussion about the right degree of smoothing in spatio-temporal data.
 
-How might this affect our results?
+  - How might this affect our results?
 So long as sampling errors are consistent over time, they should not bias the mean or the trends of our results (e.g.
 if we systemically under/over estimate the biomass of some species).
-There is no doubt though that we are underestimating the true magnitude of the credible intervals though.
+There is no doubt that we are underestimating the true magnitude of the credible intervals though.
 For example, the model assumes constant length-to-weight relationships over time, whereas in reality there is clearly variation both among individuals and over time.
 Accounting for this should add more noise to our data, further complicating the ability of our model to achieve precise estimates.
-Complicated relationships such as the MPAs themselves affecting the length-to-weight relationship would indeed skew our results, but we feel that is beyond the scope of this or any other MPA study of which we are aware to address.
+Complicated relationships such as the MPAs themselves affecting the length-to-weight relationship would indeed skew our results, but we feel that is beyond the scope of this or any other empirical MPA study of which we are aware to address.
 
-  We have run our model at a wide array of resolutions, from the updated aggregated form we use here to a complete hierarchical framework that attempts to control for errors in the individual observation process (Fig.SXX).
+  - We have run our model at a wide array of resolutions, from the updated aggregated form we use here to a complete hierarchical framework that attempts to control for errors in the individual observation process (Fig.4, Fig.S39).
 The fundamental results remains the same: a general upwards trend, followed by a downward trend, with the clearest signal in the years around 2010.
@@
 Our relatively uncertain findings instead suggests that it is unlikely that we would get fundamentally different conclusions by including more error; we would simply be even less certain than the uncertain state the model estimates we are in.
-We have expanded the simulation testing of the method a bit to help illustrate how violations of the model assumptions play out, and how they are related to the magnitude of the effect size (Figure 5).
-The full suite of alternative resolution models can be found in Figs.SXX-XX (all of which show the same general result), and a a simulation test of our model incorporating factors such as differing diver skill can be found in Fig.SXX (which shows that the model is capable of performing under such circumstances).
+We have expanded the simulation testing of the method a bit to help illustrate how violations of the model assumptions play out, and how they are related to the magnitude of the effect size (Fig.
+4).
+The full suite of alternative resolution models can be found in Figs.S16-26 (all of which show the same general result), and a a simulation test of our model incorporating factors such as differing diver skill can be found in Fig.S39 (which shows that the model is capable of performing under such circumstances).
 
@@
   - We hope that our restructuring around the novel empirical results resolves this concern.
-While we agree that framing our results around the future of marine science would be a step too far, we would point out that our earlier and present draft asked what our results imply for the future of MPA science more specifically, which we hope is less of a stretch.
-
-  Given the increasing role of MPAs in conservation strategies, the lack of a clear regional effect in a place like the Channel Islands should be known and taken into account by MPA practitioners around the globe.
+While we agree that framing our results around the future of marine science would be a step too far, our goal is to ask what our results imply for the future of MPA science more specifically, which we hope is less of a stretch.
+  - Given the increasing role of MPAs in conservation strategies, the lack of a clear population-level effect in a place like the Channel Islands should be known and taken into account by MPA practitioners around the globe.
 We agree that we framed our simulation results poorly in previous drafts, but feel that they do compliment the empirical results to help managers understand why the Channel Islands results may not be surprising, and when they might expect to be able to detect and effect, and when not.
-See section 2.1.2 ("When can we detect the effects of MPAs") and our updated conclusions, in particular the paragraph beginning on line XX.
+See section 2.1.2 ("When can we detect the effects of MPAs") and our updated conclusions, in particular the paragraph beginning on line 357.
+Works such as the papers by Martin Smith suggested by the reviewer are indeed critical to our understanding of MPAs, and we have taken steps to place our results in the context of such work.
+In particular, we advocate for more works such as Smith, Zhang, & Coleman 2006 for estimating the fishery effects of MPA networks (which differs from the population-level effects estimated here)
 
 - *Examples of relevant literature, you should expand your net to find other relevant literature to better distinguish your contribution*
+
   - We thank the reviewer for pointing out these important omissions. We have included each of these references and their key messages in their appropriate portions of the paper.

### 2020-04-26 documents/ovando-et-al-pop-effects-of-mpas-si.Rmd
resubmited to PNAS

@@
   - Owen Liu
-date: "`r Sys.Date()`"
 output: 
@@
 
-All code needed to reproduce our main results and manuscript can be found at https://github.com/DanOvando/regional-effects-of-mpas.
+All code needed to reproduce our main results and manuscript can be found at https://github.com/DanOvando/population-effects-of-mpas.
 All analysis were performed in `r R.version.string`.
@@
 
-"Fish assemblages were surveyed annually as part of a long-term monitoring program conducted by the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) using standard underwater visual belt survey methods (www.piscoweb.org9).
+"Fish assemblages were surveyed annually as part of a long-term monitoring program conducted by the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) using standard underwater visual belt survey methods (www.piscoweb.org).
 We analyzed data from 47 PISCO sites at the northern Channel Islands that were sampled annually from at least 2003 to 2012.
@@
 MPAs on each island were sampled annually during June-October and we surveyed multiple sites inside and outside of any individual MPA.
-Details of MPA characteristics such as size and coastline extent are given in Hamilton et al.9.
+Details of MPA characteristics such as size and coastline extent are given in Hamilton et al.
 At each site, we conducted 8 to 12 fish transects that measured 30×2×2m at multiple levels in the water column: benthic, midwater, and kelp canopy (when present).
@@
 Once zeros have been introduced to the database, we convert positive observations of fish from numbers to biomass.
-Each observation in the raw database lists the species, the number of individuals seen, and the size of those individuals (either as one value or as a mininum and maximum size for the group seen).
+Each observation in the raw database lists the species, the number of individuals seen, and the size of those individuals (either as one value or as a minimum and maximum size for the group seen).
 PISCO staff compiled allometric information used to convert lengths to expected weights.
@@
 We used a k-nearest neighbors algorithm to fill in missing kelp observations, and matched the interpolated kelp data to the PISCO data at the resolution of year-month-site (Fig.\@ref(fig:kelp-plot)).
-We include a variable capturing the mean cumulative number of observations across all observers conducting transects, in an effort to control for evoloving observer skill.
+We include a variable capturing the mean cumulative number of observations across all observers conducting transects, in an effort to control for evolving observer skill.
 
@@
 
-used_data %>% 
-  filter(targeted == 1) %>% 
-  group_by(commonname, year) %>% 
-  summarise(catches = unique(lag_catch)) %>% 
-  group_by(commonname) %>% 
-  mutate(total_catches = sum(catches)) %>% 
+cdfw_catches %>% 
+  group_by(classcode) %>% 
+  mutate(nc = sum(catch > 0)) %>% 
   ungroup() %>% 
-  filter(total_catches > 0) %>% 
-  ggplot(aes(year, catches)) + 
+  filter(classcode %in% unique(used_data$classcode), nc > 13) %>% 
+  ggplot(aes(year, catch)) + 
     geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
@@
 From there, we aggregated data to the level of total mean biomass density of targeted and non-targeted species at each site each year.
-As such, our base model estiamtes the effect of the MPAs on total mean biomass density of targeted species, though we also explore the effect on alternative specifications such as mean biomass of targeted species.
+As such, our base model estimates the effect of the MPAs on total mean biomass density of targeted species, though we also explore the effect on alternative specifications such as mean biomass of targeted species.
 the model was then fit using the `rstanarm` package in R (5000 iterations, 2500 warmup, 4 chains, adapt_delta = 0.85).
@@
 For the non-intercept terms, we manually set a normal(0,2) prior.
-This implies that coefficeints such as the MPA effect have a prior that provides support for an effect size centered on zero with a range of roughly -4 to 4.
+This implies that coefficients such as the MPA effect have a prior that provides support for an effect size centered on zero with a range of roughly -4 to 4.
 Since the covariates are centered and scaled, and the dependent variable is on the log-scale, this is an extremely diffuse prior (e.g.
 a one standard deviation change in an independent variable changes the dependent variable by 4 on the log scale).
+
+We weight each observation such that within MPA data are assigned a total weight of 0.2 and outside-MPA data a total weight of 0.8, in order to represent the proportion of the available habitat inside MPAs and outside.
+This makes the assumption that all areas throughout that Channel Islands are identical in terms of habitat / carrying capacity for the species in question.
+If detailed data on habitat are available, this method could be modified to weight each area by the available habitat in that area.
+See Fig.\@ref(fig:no-weight-plot) for a companion analysis without weighting (which implicitly assumes that the MPAs contain more of the available habitat than the outside-MPA areas, since the MPAs make up a disproportionate amount of the samples relative to their physcial area).
 
@@
 tex is the mean cumulative number of observations of the observers collecting the raw data, surge is the mean reported surge across transects, kelp is the mean kelp coverage at the site, temp is the water temperature, regional_temp_dev is the water temperature at a given site scaled by the mean water temperature at that site.
-Targeted is a boolean marker indicating whether the given observation is for targeted or non-targeted species.
+Targeted is a Boolean marker indicating whether the given observation is for targeted or non-targeted species.
 
@@
 
-We include one run using only variables availale in both datasets, to explore sensitivity of our results to selected covariates (Fig.\@ref(fig:kfm-data)).
+We include one run using only variables available in both datasets, to explore sensitivity of our results to selected covariates (Fig.\@ref(fig:kfm-data)).
 
@@
 We fit an alternative model using data from the KFM program to test whether a different dataset from the same region provides different results.
-KFM based results are qualitatively similer to those from PISCO data (Fig.\@ref(fig:kfm-fit))
+KFM based results are qualitatively similar to those from PISCO data (Fig.\@ref(fig:kfm-fit))
 
@@
 
-We include a model run using nearly all observed species, to test whether our results are being influenced by omissino of rarely observed species (Fig.\@ref(fig:no-filter-plot)).
+We include a model run using nearly all observed species, to test whether our results are being influenced by omission of rarely observed species (Fig.\@ref(fig:no-filter-plot)).
 
@@
     data = no_filter$did_fit[[1]]$did_results,
+    aes(year, did, color = "Empirical Estimate"),
+    .width = c(0.5, 0.95),
+    slab_alpha = 0.95  ) +
+  scale_y_continuous(name = "MPA Effect") +
+  scale_x_discrete(name = element_blank()) +
+  theme(legend.position = "top") +
+  scale_fill_brewer(name = "Simulation Quantiles") + 
+  scale_color_manual(values = "red", name = '') + 
+    geom_hline(yintercept = 0)
+
+```
+
+```{r no-weight-plot, fig.cap="Version of the model fit without reweighting samples in proportion to area",results = "hide"}
+no_weight <- model_runs %>%
+  slice(1) %>%
+  left_join(abundance_data, by = "data_source") %>%
+  mutate(
+    did_fit = pmap(
+      list(
+        data = data,
+        data_to_use = data_to_use,
+        data_source = data_source
+      ),
+      estimate_did,
+      iter = 2500,
+      site_data = site_data,
+      cdfw_catches = cdfw_catches,
+      life_history_data = life_history_data,
+      weight_samples = FALSE
+    )
+  )
+
+ggplot() +
+ tidybayes::stat_halfeye(
+    data = no_weight$did_fit[[1]]$did_results,
     aes(year, did, color = "Empirical Estimate"),
@@
 Our base model aggregated the data to the level of total biomass densities of targeted and non-targeted species.
-This serves to average out some of the sampling noise from the indivudla transect level, and we would argue is the appropriate level of estimation for many key policy questions.
+This serves to average out some of the sampling noise from the individual transect level, and we would argue is the appropriate level of estimation for many key policy questions.
 However, this approach does oversimplify the data collection process, treating the total biomass densities at the year/site level as estimated without error.
@@
 This model estimates the effect on the mean biomass density of targeted finfish across all sites.
+While this model attempts to account for finer-scale observation error than our aggregated model, it also introduces a large amount of sampling variation that cannot be explained by the model.
 
@@
 ```{r tmb-fit, fig.cap="Difference-in-difference results from mixed effect model fit to the individual transect observation level of the data."}
+tmb_runs <- tmb_runs %>%
+  mutate(tmb_fit = map(tmb_fit,"result")) %>%
+  mutate(processed_fits = map(tmb_fit, process_fits)) %>%
+  mutate(did_plot = map(processed_fits, "did_plot"))
+
 tmb_runs$did_plot[[1]] + 
@@
 ```{r betatable, results="asis"}
-
-tmb_runs <- tmb_runs %>%
-  mutate(tmb_fit = map(tmb_fit,"result")) %>%
-  mutate(processed_fits = map(tmb_fit, process_fits)) %>%
-  mutate(did_plot = map(processed_fits, "did_plot"))
 

### 2020-04-26 documents/ovando-et-al-pop-effects-of-mpas.Rmd
resubmited to PNAS

@@
 ---
-title: Assessing the Population-level Conservation Effects of Marine Protected Areas
-date: "`r Sys.Date()`"
+title: |
+  | Assessing the Population-level Conservation Effects of
+  | Marine Protected Areas
 author: 
@@
 MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present), and the size of many MPA networks suggests that effect sizes may often be smaller than 30%.
-Our results provide a novel assessment of the population-wide effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.
+Our results provide a novel assessment of the population-level effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.
 bibliography: My Library.bib
@@
   - \usepackage{lineno}\linenumbers
-  - \usepackage{draftwatermark}
 ---
@@
 
-## acknowledgements
+## Acknowledgements
   Funding for the simulation model was provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship.
@@
 
-## keywords:
+## Keywords:
   - Marine Protected Area Networks
@@
 These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], larval dispersal patterns and the strength and timing of density dependence in the population (e.g.
-pre- or post-settlement, @burgess2014), the age and degree of enforcement [@edgar2014],  how overfished the population would be without the MPA, and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006;@walters2000a;@white2011,;@white2010c].
+pre- or post-settlement, @burgess2014), the age and degree of enforcement [@edgar2014],  how overfished the population would be without the MPA, and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006a;@walters2000a;@white2011,;@white2010c].
 In addition, even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through network effects [@gaines2010; @costello2010a].
@@
 While we know many parameters will affect the expected effects of the MPAs, theory suggests that life history, adult and larval dispersal relative to MPA size, and fishing pressure will be key drivers.
-The targeted species in our database span a range of life histories, but are largely made up of fishes in the perch and rockfish complexes, with a mean Von Bertalanffy growth coefficient of `r round(mean(used_data$k),2)`, and a mean age at maturity of `r round(mean(used_data$tm),0)` years [values drawn from @thorson2017c].
+The targeted species in our database span a range of life histories, but are largely made up of fishes in the perch and rockfish complexes, with a mean Von Bertalanffy growth coefficient of `r round(mean(used_data$k),2)`, and a median age at maturity of `r round(median(used_data$tm),0)` years [values drawn from @thorson2017c].
 
@@
 
-While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (see Fig.SXX) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
-Instead, another possibility for the observed decline in Channel Islands MPA effects is environmental disturbance.
+While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (see Fig.S1) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
+However, some of the most common targeted species in the Channel Islands, such as blue rockfish and copper rockfish, have seen dramatic increases in reported catches in the Santa Barbara region since the year 2010 (although other such as California sheephead have seen declines relative to pre-MPA catch level, Fig.S1).
+To the extent that these increases in catches are due to external market forces, the dissapearance of clear positive population-level MPA effects estimated by our model could be due to exogenous increases in fishing pressure overwhelming any population-level increases caused by the MPA network.
+We lack a control though that would allow us to estimate the MPA effects relative to a counter-factual in which the same market forces increase catches but no MPAs are in place.
+
+Another possibility for the observed decline in Channel Islands MPA effects is environmental disturbance.
 The Channel Islands region (and the entire West coast of the USA) experienced a dramatic ‘marine heatwave’ beginning in 2014 and persisting through 2016, resulting in part in extremely elevated water temperatures throughout the region  [@gentemann2017].
@@
 
-This hypothesis is supported by the fact that we estimate a similar decline in the estimated MPA effect both inside and outside of the MPAs (Fig.SXX) (as well as observe the decline in the raw data, Fig.\@ref(fig:raw-density)).
+This hypothesis is supported by the fact that we estimate a similar decline in the estimated MPA effect both inside and outside of the MPAs (Fig.S18) (as well as observe the decline in the raw data, Fig.\@ref(fig:raw-density)).
 If the cause of recent declines was due to increases in fishing pressure we would expect to see substantial declines only in the fished areas.
-While we see evidence for the validity of the parallel trends assumption in the early years of our data (Fig.\@ref(fig:mpa-effect), Fig.SXX), it is likely that the non-targeted species reacted in a different way to the marine heatwave than the targeted species did, causing a violation of the parallel trends assumption in the post 2014 data.
+While we see evidence for the validity of the parallel trends assumption in the early years of our data (Fig.\@ref(fig:mpa-effect), Fig.S26), it is likely that the non-targeted species reacted in a different way to the marine heatwave than the targeted species did, causing a violation of the parallel trends assumption in the post 2014 data.
 
@@
 While we estimate an uncertain but overall positive effect of the MPA network in its first few years, we are unable to detect a robust signal from 2012-2017.
-We believe that the disappearance of increasing probabilities of positive MPA effects estimated by our model are likely driven by the marine heatwave experienced by the region beginning in 2014.
+We believe that the disappearance of increasing probabilities of positive MPA effects estimated by our model are likely driven by the marine heatwave experienced by the region beginning in 2014 (and potentially by exogenous increases in fishing mortality for key target species).
 This likely violation of the parallel trends assumption means that post 2014 estimates of MPA effects may be unreliable.
@@
 While a large percentage effect, this is a small change in biomass densities relative to the variance in the observation process itself.
-While an MPA providing 25% protection might be expected to produce a 30% increase in biomass densities, expressed as a percentage of unfished biomass, this gain is on the order of 15% (Fig.SXX).
+While an MPA providing 25% protection might be expected to produce a 30% increase in biomass densities, expressed as a percentage of unfished biomass, this gain is on the order of 15% (Fig.S45).
 In addition, the median estimated age at sexual maturity for the targeted species included in this study is `r median_tm` years, meaning that the span of this analysis represents less than three generations of MPA protection for half of the measured species.
@@
 When should we expect to see MPA effects big enough to stand a reasonable chance of detection?
-We simulated `r n_distinct(outcomes$experiment)` MPA scenarios across a wide range of life histories, network designs, and fishing dynamics (see SI for a full description of scenarios).
+We simulated `r n_distinct(outcomes$experiment)` MPA scenarios across a wide range of life histories, network designs, and fishing dynamics (see SI).
 Suppose that we are willing to tolerate a MAPE of 25%.
-Our analysis suggests that we would need an MPA effect size of at least 30% to achieve this.
-Two of the most critical drivers of MPA performance are the size of the MPA network and the degree of fishing pressure.
+Our analysis suggests that we would need a population-level MPA effect size of at least 30% to achieve this.
+Two of the most critical drivers of MPA effect size are the size of the MPA network and the degree of fishing pressure.
 Looking across these two variables, if the MPA network covers 25% or more of a species range and pre-MPA depletion is greater than 60% we might expect an effect size with our target MAPE.
@@
 A large theoretical literature has pointed out the importance of a wide range of economic and environmental factors in determining the effectiveness of MPAs [e.g.
-@hilborn2004; @smith2006; @white2010c].
+@hilborn2004; @smith2006a; @white2010c].
 Our simulation model allows users to place this collective theory into context, clarifying under what states of the world we might expect to see particular effect sizes, and demonstrating what effect sizes we might reasonably expect to be detectable in the real world.
@@
 Rather, results and subsequent management actions (such as adaptation of MPA networks) must be considered in the context of reasonable expectations given the size, age, and degree of enforcement of the MPAs in question, together with the ecological and economic dynamics of a given system.
-While this paper has focused on the conservation outcomes of MPAs, future work must also address the challenge of predicting and estimating the fishery impacts of protected areas.
+While this paper has focused on the conservation outcomes of MPAs, future work must also address the challenge of predicting and estimating the fishery impacts of protected areas [in the manner of @smith2006a].
 
@@
 While the parallel trends assumption cannot be formally proven, we can examine its validity using the data from the years before the MPAs were put in place in 2003.
-We do not detect any significant differences in the trends of the biomass densities of the targeted and non-targeted species in the years before the MPAs (Fig.SXX).
+We do not detect any significant differences in the trends of the biomass densities of the targeted and non-targeted species in the years before the MPAs (Fig.S26).
 However, we believe that the parallel trends assumption is likely violated in the post 2014 data.
@@
 We used convergent cross mapping (CCM), in the manner of @clark2015, to test for the possibility of the trophic cascades biasing our results.
-Our results found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (Fig.SXX).
+Our results found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (Fig.S42).
 However, the longer MPAs are in place, the greater the possibility that substantial species interactions that can affect use of non-targeted species as a control may arise.

### 2020-04-26 README.Rmd
resubmissted version to PNAS (#1)
* version re-submited to PNAS

@@
 
->> Marine Protected Areas (MPAs) cover between 3-7% of the world's oceans, up from less than 1% in the year 2000.
-The Convention on Biological Diversity calls for 10% of coastal waters to be protected inside MPAs by 2020, while the International Union for Conservation of Nature calls for 30% protection by 2030.
-It is often clear that MPAs produce conservation benefits inside their borders, but many MPAs are also justified on the grounds that they also benefit the broader region outside their borders.
-Despite the expanding role of MPAs in marine resource management, we lack a clear understanding of what the regional effects of MPAs might be, and how we might detect them in scientific studies.
-We used a bio-economic simulation model to demonstrate how the regional effects of MPAs are affected by a suite of economic and environmental drivers and show how commonly used methods to measure inside-MPA effects are likely to fail to capture the regional impacts of protection.
-We present an empirical strategy for estimating regional MPA effects on biomass density using 16 years of data from inside and outside of MPAs within the Channel Islands, California, USA.
-We find no statistically clear effect of MPAs on aggregate mean biomass densities of targeted finfish throughout the study region.
-To make meaningful progress on MPA design and evaluation, the next generation of MPA science must address the challenge of understanding and estimating the regional effects of MPAs across the diversity of species in the ecosystem.
+>>  Marine Protected Areas (MPAs) cover 3-7% of the world’s ocean, up from less than 1% in the year 2000, and international commitments call for 10%, 30%, and even 50% coverage.
+The premise underlying MPA expansion is that they conserve biodiversity, habitats, and fished populations.
+While numerous studies show that MPAs produce conservation benefits *inside* their borders, many MPAs are also justified on the grounds that they confer conservation benefits to the broader population *beyond* their borders.
+This paper examines the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and shows that even in cases when the conservation benefits are large, they are inherently difficult to detect empirically.
+A network of MPAs was put in place in The Channel Islands National Marine Sanctuary in 2003, with a goal of providing regional conservation and fishery benefits.
+Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level.
+MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present), and the size of many MPA networks suggests that effect sizes may often be smaller than 30%.
+Our results provide a novel assessment of the population-wide effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.

### 2020-04-26 documents/cover-letter.Rmd
resubmissted version to PNAS (#1)
* version re-submited to PNAS

--- 2020-03-11 documents/cover-letter.Rmd
@@
 
-##
+# Responses
+
+We thank the two reviewers and the Editor for their constructive responses.
+We have fundamentally restructured the paper to place much greater emphasis on the empirical results from the Channel Islands MPA network.
+This is to our knowledge the first empirical assessment of the regional effects of a large and iconic MPA network on a wide assemblage of commercially and economically important species.
+Our inability to estimate a clear effect in such a well-enforced and well-studied local is we feel a vitally important "null-result" both for scientists and MPA managers to be aware of.
+As such we feel that it warrants publication in PNAS.
+While our results are more focused on the empirical results, we explicitly link our results to the large body of MPA theoretical work, using it to help illustrate why we might expect such a null result and help managers understand when they might be able to detect an effect.
+As MPAs are used as a core tool of marine resource management, our results provide a critical update to our empirical expectations for MPAs, that to date have been based almost entirely on inside-outside MPA response ratios.
+We hope you find the updated version of our manuscript improved thanks to the inclusions of the careful points raised by the prior revision of this paper.
+
+## Reviewer 1
+
+- *I think that greater emphasis needs to be placed up front, abstract, statement of significance, introduction..perhaps even title, that this model and paper only addresses fish biomass...not other 'conservation outcomes'.
+You do define this but not until page 2, line 148-154.
+I recommend moving this up.*
+  - We apologize for not making this clearer in our earlier draft, and have included this key caveat in the abstract, statement of significance, and introduction (Line 98)
+
+- *I was surprised that there was no discussion about enforcement being a potential confounding factor.
+Currently the paper assumes that the Channel Islands MPAs are fully enforced, without poaching.
+This may or may not be the case and should be stated as such.
+You do address well other system wide impacts like water temperature (lines 557-8, 691+).
+Perhaps enforcement could be added to the discussion around line 711-13.*
+  - This is indeed a critical point that we have endeavored to correct.
+We have included enforcement under our summary of key enabling conditions for effective MPAs (line 124), as well as the discussion (line 382).
+While we do not have published data on strength of enforcement within the MPAs, anecdotal evidence suggests it is relatively high, as well as evidence from the mainland California MPAs (Zellmer et al.
+2018) indicates it is high.
+In addition, the presence of a clear increase in response ratios within the Channel Islands provides evidence that the MPAs  are large enough and sufficiently well-enforced as to provide meaningful protection within their borders (Line XX).
+
+- *Your discussion of the model also emphasizes that stock status prior to MPA designation is very key to outcomes; in the SI most of the species included for the Channel Island review have unknown status---which influences your discussion, and may better be described.*
+  - We make clearer that we iterate over unknowns such as pre-MPA stock status in our description of the Channel Islands simulations (line XX), and the implications of pre-MPA stock status in the discussion (Line XX).
+
+## Reviewer 2
+
+- *Yes, these are important questions but they are not new to ecology, economics, and fishery science.
+You are not the first to ask them and point out their importance; see articles at bottom for some examples.
+What are you doing to answer them that is better than all of the other papers that come before you?
+And why is your case study an exemplary place to answer the questions?
+The conclusions on line 709-728 are well-known in the MPA literature.*
+  - We have fundamentally restructured the paper based around the importance of the Channel Islands as a case study in measuring MPA effectiveness.
+We thank the reviewer for pointing out gaps in our literature, and we have incorporated the studies they recommend.
+The theoretical points we raise have indeed been pointed out in the literature before, and we feel our new manuscript makes that clearer.
+The Channel Islands MPAs are an iconic network that has served as a model for protected areas around the world.
+It is also incredibly well-studied and well-enforced.
+While previous studies have looked the effectiveness of the Channel Islands MPAs within their borders, this is the first study in the Islands (and globally of which we are aware) to attempt an estimate of the effect of the MPAs at the regional scale (both inside and outside of the MPAs themselves).
+That we are unable to estimate a clear regional effect in even this system is an important result that we feel warrants publication in PNAS.
+As MPAs are increasingly turned to tools for ocean conservation, our results provide an important update and caution to expectations based solely on inside-outside comparison studies.
+
+- *I would also be more cautious regarding your conclusion about your model predicts what should be expected in the data analysis. It could be simply luck that there is agreement between the two.*
+  - We have made the language clearer that this concordance could be simple coincidence, and is not in any way proof of the validity of our results (Line XX) However, in light of recent concerns in the scientific literature around hypothesizing-after-results, we feel that it is critically important to ground our theoretical results in *a priori* predictions.
+In addition, as we mention in the discussion, we encourage MPA scientists and stakeholders to undertake such simulation modeling to consider the plausibility of their results.
+As the reviewer says, empirical results that match theoretical predictions are not proof that the empirical methods are robust, but they do provide am additional and important line of evidence: empirical results that fail to match theoretical predictions need to be carefully considered.
+
+- *What about using the simulation model to demonstrate the validity of the technique used to identify the MPA effect used in the paper, which was pioneered in publication in 38?
+Or to understand whether the implication of the statement on line 539, that is, species ranges exceed the Island and your sampling domain, is important for measuring the impacts of MPAs outside of their boundaries?
+In sum, in its current form, I would move all of the simulation model to the SI (lines 131-252) and mention in the discussion section that these results are consistent with the predictions of a stylized single species model.*
+    - We have greatly expanded the simulation validation of our method, which was previously included in the supplementary material.
+Figure 4 of the paper now demonstrates the validity (and shortcomings) of the model under a variety of plausible conditions.
+We would note that this method is related to that described in Caselle et al.
+2015 (previously reference #38), but is not the same.
+Our method explicitly estimates the net difference in the targeted and non-targeted biomass densities pre-and-post MPA as an estimate of the mean effect of the MPAs  both inside and outside their borders.
+Caselle et al.
+2015 used trends in targeted and non-targeted species inside and outside of MPAs to illustrate the effects of the MPAs within their borders.
+We stress this difference in an effort to make clear that our results and methods are not simply an update to Caselle et al.
+2015, but address a fundamentally different question.
+We hope we have made this clearer in the text, e.g.
+on line XX.
+
+    With regards to the specific question about the possibility of the species range exceeding the sampling domain, this would be reflected as a reduction in the percent of the population protected by the MPAs: a 25km^2 MPA may provide near total protection for a sedentary species with brood-rearing, but be insignificant to a migratory tuna.
+To that end, our assessment of model performance as a function of MPA effect size addresses the question of hwo MPA size relative to species range affects our models performance.
+All else being equal the smaller a portion of a population's range is protected by MPAs, the smaller the expected effect size, and by extension the more challenge the estimation model will have.
+
+  We apologize that we did not present our prior simulation work in a sufficiently compelling manner, and failed to differentiate its results from the prior literature.
+As suggested by the reviewers and the editor, we have restructured the paper to place much greater emphasis on the critical results from the Channel Islands.
+However, we feel that the simulation results do provide vital context to our results, and provide tangible recommendations as to under what circumstances we might be able to estimate the regional effects of MPAs.
+The reviewer correctly points out that the individual components of our model are not themselves novel, and that the importance of fleet dynamics and fish dispersal have been previously discussed.
+As such as have greatly reduced the attention these results receive, except as a vehicle for the explicit consideration of the feasibility of estiamting reginoal MPA effects.
+In addition, while these broad points have been made in the literature before, we feel there is value in cohesively demonstrating the ramifications of these diverse pieces of MPA theory (that generally are looked at only one or a few at a time) in once place, both through our Figure 5 and the accompanying web application.
+We hope that these modifications make clear the contribution of the simulation modeling.
+
+- *The discussion on how to detect changes outside the MPA (lines 254-347) follows very closely the discussion in Ferraro et al (#41 in your list).
+It seems like you can remove most of this text and point the reader to that paper*
+  - We have greatly reduced the amount of text addressing response ratios, and restructured the response-ratio analysis to be linked explicitly with the empirical response ratios observed in the Channel Islands (in response to comments from the editor), in Figure 1.
+We feel this strategy makes much clearer how the results of this paper expand beyond the findings of Caselle et al.
+2015, and why the results of Caselle et al.
+2015 are evidence of within-MPA effects, but not of the question of regional effects addressed in this manuscript.
+Given that much of the MPA literature relies on response ratios as the empirical evidence for the effectiveness of MPAs, we feel that it is critically important to directly address the challenges of response ratios as measures of regional effects in this paper.
+We hope that the new manner, directly linking theoretical response ratios to the levels observed in the Channel Islands.
+
+- *To me the potential contribution of the paper starts on line 350, which is way too late in a paper of this size*
+  - We have restructured the paper so that it now focuses on the Channel Islands case study, and as suggested have moved this discussion up to the introduction (line 97).
+
+- *Is part of assumption b) (line 372) that the dispersal characteristics of the fished and non-fished species need to be similar?
+I could envision a scenario where the two sets of species have different dispersal characteristics, non-fished have smaller dispersal ranges than fished species, for example.
+How does this impact your identification strategy?
+If the trends are similar regardless of the mechanistic process by which it occurs, does it matter for the empirics?
+More discussion on both assumptions (a) and (b) and what they imply about the biological mechanisms seems warranted.*
+  - The estimation model does not depend on the dispersal characteristics of the fished and non-fished species per-say.
+The model is operating at the level of biomass densities.
+As such, even if a non-targeted species has a much bigger range than the targeted species (or even the sampling region), so long as the parallel trends assumption holds within the sampled portion of the non-targeted species range, the identification strategy may be valid.
+Similarly, the model is agnostic as to the root causes of the parallel trends, so long as they exist.
+However, if for example the parallel trends are dependent on a particular temperature regime which during the study is disrupted (as is likely the case in our data), then the model may be reliable for the early years but not later.
+We have made this point clearer in the paper (line XX), and made clear that the lack of a clear effect in the recent years is likely due to an emerging violation of the parallel trends assumption, and not to a failure of the MPAs per say.
+However, this is still an important result, since it suggests that neither the response ratios nor the difference-in-difference method give us a robust estimate of the regional effects of the Channel Island MPAs.
+As we stress in the paper, that even in a network such as the Channel Islands we ar left without a clear empirical effect at the regional level has critical ramificaitons for the design and monitoring of MPAs globally.
+
+- *The use of visual data surveys has, of course, problems that introduce errors in dependent and independent variables that could be correlated with features of the fished and non-fished species, e.g., size distribution.
+How can you be sure that this is not driving your results?
+Could you use the simulation model to investigate this?*
+  - Visual survey data are indeed challenging, but also often the only form of data available for MPA analysis.
+Our base model operates at the level of total biomass density of targeted and non-targeted species, at the site level.
+This is the finest spatial scale at which the sampling regime would be likely to provide an estimate of the variable in question, since while any individual transect is unlikely to observe all of the targeted / non-targeted, but aggregating across transects provides a better chance of this.
+These totals are derived from average biomass densities of indivudal species across multiple transects within each site, which are then summed across species to achieve the total mean biomass density at the site level.
+The averaging process in acts to smooth out some of the sampling variation, and reflects a broader discussion about the right degree of smoothing in spatio-temporal data.
+
+How might this affect our results?
+So long as sampling errors are consistent over time, they should not bias the mean or the trends of our results (e.g.
+if we systemically under/over estimate the biomass of some species).
+There is no doubt though that we are underestimating the true magnitude of the credible intervals though.
+For example, the model assumes constant length-to-weight relationships over time, whereas in reality there is clearly variation both among individuals and over time.
+Accounting for this should add more noise to our data, further complicating the ability of our model to achieve precise estimates.
+Complicated relationshops such as the MPAs themselves affecting the length-to-weight relationship would indeed skew our results, but we feel that is beyond the scope of this or any other MPA study of which we are aware to address.
+
+  We have run our model at a wide array of resolutions, from the updated aggregated form we use here to a complete hierarchical framework that attempts to control for errors in the individual observation process (Fig.SXX).
+The fundamental results remains the same: a general upwards trend, followed by a downward trend, with the clearest signal in the years around 2010.
+We would actually be more concerned about the robustness of our results if we had found a precise but small effect size: such an effect could easily dissapear if we accounted for errors in the data (e.g.
+through allowing for variations in the length-weight relationship).
+Our relatively uncertain findings instead suggests that it is unlikely that we would get fundamentally different conclusions by including more error; we would simply be even less certain than the uncertain state the model estiamtes we are in.
+We have expanded the simulation testing of the method a bit to help illustrate how violations of the model assumptions play out, and how they are related to the magnitude of the effect size (Figure 5).
+The full suite of alternative resolution models can be found in Figs.SXX-XX (all of which show the same general result), and a a simulation test of our model incorporating factors such as differing diver skill can be found in Fig.SXX (which shows that the model is capable of performign under such circumstances).
+
+- *(The answer to your question on line 709 that appears on lines 714-719 has been known in the literature on MPAs for over 20 years; see, e.g., papers by Martin Smith below.
+Given that your question is what "your results imply about the future of marine science", my answer is not very much.
+Again, you did a rigorous empirical case study that should be published.
+I am however finding it difficult to see a path to publication in PNAS.)*
+  - We hope that our restructuring around the novel empirical results resolves this concern.
+While we agree that framing our results around the future of marine science would be a step too far, we would point out that our earlier and present draft asked what our results imply for the future of MPA science more specifically, which we hope is less of a stretch.
+
+  Given the increasing role of MPAs in conservation strategies, the lack of a clear regional effect in a place like the Channel Islands should be known and taken into account by MPA practitioners around the globe.
+We agree that we framed our simulation results poorly in previous drafts, but feel that they do compliment the empirical results to help managers understand why the Channel Islands results may not be surprising, and when they might expect to be able to detect and effect, and when not.
+See section 2.1.2 ("When can we detect the effects of MPAs") and our updated conclusions, in particular the paragraph beginning on line XX.
+
+- *Examples of relevant literature, you should expand your net to find other relevant literature to better distinguish your contribution*
+  - We thank the reviewer for pointing out these important omissions, in particular @white2011, @smith2006, @nickols2019.
+We have included each of these references in their appropriate portions of the paper
+
+## Editor
+
+# Original Reviews
 
@@
 
-White, J.
+*White, J.
 W., Botsford, L.
@@
 Frontiers in Ecology and the Environment, 9: 390-399.
-doi:10.1890/100138
+doi:10.1890/100138*
 
-Burgess, S.
+*Burgess, S.
 C., Nickols, K.
@@
 Ecological Applications, 24: 257-270.
-doi:10.1890/13-0710.1
+doi:10.1890/13-0710.1*
 
-Smith, M.
+*Smith, M.
 D., & Wilen, J.
@@
 Journal of Environmental Economics and Management, 46(2), 183-206.
-doi:https://doi.org/10.1016/S0095-0696(03)00024-X
+doi:https://doi.org/10.1016/S0095-0696(03)00024-X*
 
-Smith, M.
+*Smith, M.
 D., Zhang, J., & Coleman, F.
@@
 Canadian Journal of Fisheries and Aquatic Sciences, 63(1), 153-164.
-doi:10.1139/f05-205
+doi:10.1139/f05-205*
 
-Kaplan, K.
+*Kaplan, K.
 A., L.
@@
 Ecological Applications 29(6):e01949.
-10.1002/eap.1949
+10.1002/eap.1949*
+
+# References

### 2020-04-26 documents/ovando-regional-effects-of-mpas-nopnas.Rmd
resubmissted version to PNAS (#1)
* version re-submited to PNAS

--- 2020-01-16 documents/ovando-regional-effects-of-mpas.Rmd
@@
 title: The Regional Effects of Marine Protected Areas
-author:
-  - name: Daniel Ovando
-    affiliation: a,1
-    email: danovan@uw.edu
-  - name: Jennifer E. Caselle
-    affiliation: b
-  - name: Christopher Costello
-    affiliation: b   
-  - name: Olivier Deschenes
-    affiliation: b  
-  - name: Steven D. Gaines
-    affiliation: b
-  - name: Ray Hilborn
-    affiliation: a
-  - name: Owen Liu
-    affiliation: b        
-address:
-  - code: a
-    address: University of Washington, School of Aquatic and Fishery Sciences
-  - code: b
-    address: University of California, Santa Barbara
-corresponding_author:
-  - code: 1
-    text: "To whom correspondence should be addressed. E-mail: danovan@uw.edu"
-lead_author_surname: Ovando
-author_contributions: |
-  D.O., S.G., and R.H.
-developed structure of simulation model.
-D.O., O.D., C.C., and J.C.
-developed estimation strategy.
-J.C.
-provided support on collecting and interpreting data.
-All simulations, statistics, and sensitivity analyses performed by D.O.
-and O.L.
-.
-All authors contributed to the manuscript.
-conflict_of_interest: |
-  R.H's research program receives funding from environmental NGOs, foundations, fishing industry, governments and international agencies.
-All of these can be interpreted as a conflict of interest when evaluating fisheries policy.
-C.C., S.G., and O.D's research program includes funding from environmental NGOs and foundations with an interest in ocean conservation and food security.
-Funding for D.O.
-was partly provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship.
-abstract: |
-  Marine Protected Areas (MPAs) cover between 3-7% of the world's oceans, up from less than 1% in the year 2000.
-The Convention on Biological Diversity calls for 10% of coastal waters to be protected inside MPAs by 2020, while the International Union for Conservation of Nature calls for 30% protection by 2030.
-It is often clear that MPAs produce conservation benefits inside their borders, but many MPAs are also justified on the grounds that they also benefit the broader region outside their borders.
-Despite the expanding role of MPAs in marine resource management, we lack a clear understanding of what the regional effects of MPAs might be, and how we might detect them in scientific studies.
-We used a bio-economic simulation model to demonstrate how the regional effects of MPAs are affected by a suite of economic and environmental drivers and show how commonly used methods to measure inside-MPA effects are likely to fail to capture the regional impacts of protection.
-We present an empirical strategy for estimating regional MPA effects on biomass density using 16 years of data from inside and outside of MPAs within the Channel Islands, California, USA.
-We find no statistically clear effect of MPAs on aggregate mean biomass densities of targeted finfish throughout the study region.
-To make meaningful progress on MPA design and evaluation, the next generation of MPA science must address the challenge of understanding and estimating the regional effects of MPAs across the diversity of species in the ecosystem.
-significance: |
-  Healthy marine ecosystems are critical to the well-being of the planet.
-Marine protected areas, parts of the oceans protected from human activities such as fishing, are increasingly being used in an effort to conserve and manage these ecosystems.
-Our study pairs theory with empirical methods to examine the regional effects of MPAs on conservation outcomes.
-Using simulations, we show that the regional effects of marine protected areas can be highly variable and dependent on human drivers.
-We find no statistically clear effect of a network of MPAs after 13 years of protection, which our simulation analysis suggests is actually to be expected.
-Our results lay a foundation for future research on the design and monitoring of marine protected areas.
-acknowledgements: |
-  Funding for the simulation model was provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship.
-This study would not be possible without the work provided by PISCO divers over the years.
-DO thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments and technical support.
-keywords:
-  - Marine Protected Areas
-  - Conservation
-  - Bio-economic modeling
-  - Program Evaluation
-## must be one of: pnasresearcharticle (usual two-column layout), pnasmathematics (one column layout), or pnasinvited (invited submissions only)
-pnas_type: pnasresearcharticle
+author: Daniel Ovando, Jennifer E. Caselle, Christopher Costello, Olivier Deschenes, Steven D. Gaines, Ray Hilborn, and Owen Liu    
+date: "Document compiled on `r Sys.Date()`"
 bibliography: My Library.bib
-csl: pnas.csl
-lineno: true
 header-includes:
-   - \usepackage{eso-pic,graphicx,transparent,xcolor}
+   - \usepackage{eso-pic,graphicx,transparent,lineno}
+   - \linenumbers
 output:
-  rticles::pnas_article: default
+  bookdown::pdf_document2: default
   bookdown::word_document2: default
-  bookdown::pdf_document2: default
   bookdown::html_document2: default
@@
   num_patches: [25]
-linkcolor: blue
 ---
@@
 
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: danovan@uw.edu}
-
-<!-- \AddToShipoutPictureFG{ -->
-<!--   \AtPageCenter{% or \AtTextCenter -->
-<!--     \makebox[0pt]{\rotatebox[origin=c]{45}{% -->
-<!--       \scalebox{5}{\texttransparent{0.3}{DRAFT}}% -->
-<!--     }} -->
-<!--   } -->
-<!-- } -->
+\AddToShipoutPictureFG{
+  \AtPageCenter{% or \AtTextCenter
+    \makebox[0pt]{\rotatebox[origin=c]{45}{%
+      \scalebox{5}{\texttransparent{0.3}{DRAFT}}%
+    }}
+  }
+}
+
+# Abstract
+
+Marine Protected Areas (MPAs) cover between 3-7\% of the world's oceans, up from less than 1% in the year 2000.
+The Convention on Biological Diversity calls for 10% of coastal waters to be protected inside MPAs by 2020, while the International Union for Conservation of Nature calls for 30\% protection by 2030.
+It is often clear that MPAs produce conservation benefits inside their borders, but many MPAs are also justified on the grounds that they also benefit the broader region outside their borders.
+Despite the expanding role of MPAs in marine resource management, we lack a clear understanding of what the regional effects of MPAs might be, and how we might detect them in scientific studies.
+We used a bio-economic simulation model to demonstrate how the regional effects of MPAs are affected by a suite of economic and environmental drivers and show how commonly used methods to measure inside-MPA effects are likely to fail to capture the regional impacts of protection.
+We present an empirical strategy for estimating regional MPA effects on biomass density using 16 years of data from inside and outside of MPAs within the Channel Islands, California, USA.
+We find no statistically clear effect of MPAs on aggregate mean biomass densities of targeted finfish throughout the study region.
+To make meaningful progress on MPA design and evaluation, the next generation of MPA science must address the challenge of understanding and estimating the regional effects of MPAs across the diversity of species in the ecosystem.
+
+# Significance
+
+Healthy marine ecosystems are critical to the well-being of the planet.
+Marine protected areas, parts of the oceans protected from human activities such as fishing, are increasingly being used in an effort to conserve and manage these ecosystems.
+Our study pairs theory with empirical methods to examine the regional effects of MPAs on conservation outcomes.
+Using simulations, we show that the regional effects of marine protected areas can be highly variable and dependent on human drivers.
+We find no statistically clear effect of a network of MPAs after 13 years of protection, which our simulation analysis suggests is actually to be expected.
+Our results lay a foundation for future research on the design and monitoring of marine protected areas.
+
+# Main Body
 
@@
 
-Indeed, many recent agreements to expand MPAs (most proximately, the Convention on Biological Diversity’s Strategic Plan for Biodiversity, which calls for 10% of coastal waters to be protected inside MPAs by 2020 and the International Union for Conservation of Nature call for 30% by 2030) rely on the belief that well-designed MPAs will achieve benefits both within, and outside, their borders.
+Indeed, many recent agreements to expand MPAs (most proximately, the Convention on Biological Diversity’s Strategic Plan for Biodiversity, which calls for 10\% of coastal waters to be protected inside MPAs by 2020 and the International Union for Conservation of Nature call for 30\% by 2030) rely on the belief that well-designed MPAs will achieve benefits both within, and outside, their borders.
 Despite these assumptions, our collective scientific understanding of the regional-scale conservation and fishery impacts of current and future MPAs is surprisingly limited.
@@
 
-<!-- \showmatmethods -->
-\showacknow
-
-&nbsp;
-&nbsp;
-&nbsp;
+# Acknowledgements
+
+Funding for the simulation model was provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship.
+This study would not be possible without the work provided by PISCO divers over the years.
+DO thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments and technical support.

### 2020-04-26 documents/ovando-regional-effects-of-mpas-si.tex
resubmissted version to PNAS (#1)
* version re-submited to PNAS

@@
 \usepackage{longtable,booktabs}
-\usepackage{graphicx,grffile}
+\usepackage{graphicx}
+% grffile has become a legacy package: https://ctan.org/pkg/grffile
+\IfFileExists{grffile.sty}{%
+\usepackage{grffile}
+}{}
 \makeatletter
@@
 
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+
 \usepackage{longtable}
@@
 backports & backports & 1.1.5 & 2019-10-02 & CRAN (R 3.6.1)\\
+base64enc & base64enc & 0.1-3 & 2015-07-28 & CRAN (R 3.6.0)\\
+bayesplot & bayesplot & 1.7.0 & 2019-05-23 & CRAN (R 3.6.0)\\
 bitops & bitops & 1.0-6 & 2013-08-17 & CRAN (R 3.6.0)\\
-bookdown & bookdown & 0.13 & 2019-08-21 & CRAN (R 3.6.0)\\
+\addlinespace
+bookdown & bookdown & 0.15 & 2019-11-12 & CRAN (R 3.6.0)\\
+boot & boot & 1.3-23 & 2019-07-05 & CRAN (R 3.6.0)\\
 broom & broom & 0.5.2 & 2019-04-07 & CRAN (R 3.6.0)\\
-\addlinespace
 callr & callr & 3.3.2 & 2019-09-22 & CRAN (R 3.6.0)\\
 caret & caret & 6.0-84 & 2019-04-27 & CRAN (R 3.6.0)\\
+\addlinespace
 cellranger & cellranger & 1.1.0 & 2016-07-27 & CRAN (R 3.6.0)\\
 class & class & 7.3-15 & 2019-01-01 & CRAN (R 3.6.1)\\
-classInt & classInt & 0.4-1 & 2019-08-06 & CRAN (R 3.6.0)\\
-\addlinespace
+classInt & classInt & 0.4-2 & 2019-10-17 & CRAN (R 3.6.0)\\
 cli & cli & 1.1.0 & 2019-03-19 & CRAN (R 3.6.0)\\
 codetools & codetools & 0.2-16 & 2018-12-24 & CRAN (R 3.6.1)\\
+\addlinespace
 colorspace & colorspace & 1.4-1 & 2019-03-18 & CRAN (R 3.6.0)\\
+colourpicker & colourpicker & 1.0 & 2017-09-27 & CRAN (R 3.6.0)\\
 crayon & crayon & 1.3.4 & 2017-09-16 & CRAN (R 3.6.0)\\
-crul & crul & 0.8.4 & 2019-08-02 & CRAN (R 3.6.0)\\
+crosstalk & crosstalk & 1.0.0 & 2016-12-21 & CRAN (R 3.6.0)\\
+crul & crul & 0.9.0 & 2019-11-06 & CRAN (R 3.6.0)\\
 \addlinespace
 curl & curl & 4.2 & 2019-09-24 & CRAN (R 3.6.0)\\
-data.table & data.table & 1.12.4 & 2019-10-03 & CRAN (R 3.6.1)\\
+data.table & data.table & 1.12.6 & 2019-10-18 & CRAN (R 3.6.0)\\
 DBI & DBI & 1.0.0 & 2018-05-02 & CRAN (R 3.6.0)\\
 desc & desc & 1.2.0 & 2018-05-01 & CRAN (R 3.6.0)\\
-devtools & devtools & 2.2.0 & 2019-09-07 & CRAN (R 3.6.0)\\
-\addlinespace
-digest & digest & 0.6.21 & 2019-09-20 & CRAN (R 3.6.0)\\
+devtools & devtools & 2.2.1 & 2019-09-24 & CRAN (R 3.6.0)\\
+\addlinespace
+digest & digest & 0.6.23 & 2019-11-23 & CRAN (R 3.6.0)\\
+doParallel & doParallel & 1.0.15 & 2019-08-02 & CRAN (R 3.6.0)\\
 dplyr & dplyr & 0.8.3 & 2019-07-04 & CRAN (R 3.6.0)\\
-DT & DT & 0.9 & 2019-09-17 & CRAN (R 3.6.1)\\
+DT & DT & 0.10 & 2019-11-12 & CRAN (R 3.6.0)\\
+dygraphs & dygraphs & 1.1.1.6 & 2018-07-11 & CRAN (R 3.6.0)\\
+\addlinespace
 e1071 & e1071 & 1.7-2 & 2019-06-05 & CRAN (R 3.6.0)\\
 ellipsis & ellipsis & 0.3.0 & 2019-09-20 & CRAN (R 3.6.0)\\
-\addlinespace
 evaluate & evaluate & 0.14 & 2019-05-28 & CRAN (R 3.6.0)\\
@@
 extrafontdb & extrafontdb & 1.0 & 2012-06-11 & CRAN (R 3.6.0)\\
+\addlinespace
+farver & farver & 2.0.1 & 2019-11-13 & CRAN (R 3.6.0)\\
+fastmap & fastmap & 1.0.1 & 2019-10-08 & CRAN (R 3.6.0)\\
+FishLife & FishLife & 2.0.0 & 2019-11-17 & Github (james-thorson/FishLife@3a6bdca)\\
 forcats & forcats & 0.4.0 & 2019-02-17 & CRAN (R 3.6.0)\\
@@
 fs & fs & 1.3.1 & 2019-05-06 & CRAN (R 3.6.0)\\
-gdtools & gdtools & 0.2.0 & 2019-09-03 & CRAN (R 3.6.0)\\
+furrr & furrr & 0.1.0 & 2018-05-16 & CRAN (R 3.6.0)\\
+future & future & 1.15.0 & 2019-11-08 & CRAN (R 3.6.0)\\
+gdtools & gdtools & 0.2.1 & 2019-10-14 & CRAN (R 3.6.0)\\
+\addlinespace
 generics & generics & 0.0.2 & 2018-11-29 & CRAN (R 3.6.0)\\
 geojson & geojson & 0.3.2 & 2019-01-31 & CRAN (R 3.6.0)\\
-\addlinespace
-geojsonio & geojsonio & 0.7.0 & 2019-04-25 & CRAN (R 3.6.0)\\
+geojsonio & geojsonio & 0.8.0 & 2019-10-29 & CRAN (R 3.6.0)\\
 geojsonlint & geojsonlint & 0.3.0 & 2019-02-08 & CRAN (R 3.6.0)\\
+gganimate & gganimate & 1.0.4 & 2019-11-18 & CRAN (R 3.6.1)\\
+\addlinespace
 ggmap & ggmap & 3.0.0.901 & 2019-10-05 & Github (dkahle/ggmap@37a8672)\\
 ggplot2 & ggplot2 & 3.2.1 & 2019-08-10 & CRAN (R 3.6.0)\\
+ggridges & ggridges & 0.5.1 & 2018-09-27 & CRAN (R 3.6.0)\\
 ggsci & ggsci & 2.9 & 2018-05-14 & CRAN (R 3.6.0)\\
-\addlinespace
 ggspatial & ggspatial & 1.0.3 & 2018-12-14 & CRAN (R 3.6.0)\\
+\addlinespace
+gh & gh & 1.0.1 & 2017-07-16 & CRAN (R 3.6.0)\\
+gifski & gifski & 0.8.6 & 2018-09-28 & CRAN (R 3.6.0)\\
+globals & globals & 0.12.4 & 2018-10-11 & CRAN (R 3.6.0)\\
 glue & glue & 1.3.1 & 2019-03-12 & CRAN (R 3.6.0)\\
 gower & gower & 0.2.1 & 2019-05-14 & CRAN (R 3.6.0)\\
+\addlinespace
 gridExtra & gridExtra & 2.3 & 2017-09-09 & CRAN (R 3.6.0)\\
 gtable & gtable & 0.3.0 & 2019-03-25 & CRAN (R 3.6.0)\\
-\addlinespace
-haven & haven & 2.1.1 & 2019-07-04 & CRAN (R 3.6.0)\\
+gtools & gtools & 3.8.1 & 2018-06-26 & CRAN (R 3.6.0)\\
+haven & haven & 2.2.0 & 2019-11-08 & CRAN (R 3.6.0)\\
 here & here & 0.1 & 2017-05-28 & CRAN (R 3.6.0)\\
-hms & hms & 0.5.1 & 2019-08-23 & CRAN (R 3.6.0)\\
+\addlinespace
+hms & hms & 0.5.2 & 2019-10-30 & CRAN (R 3.6.0)\\
 hrbrthemes & hrbrthemes & 0.6.0 & 2019-01-21 & CRAN (R 3.6.0)\\
 htmltools & htmltools & 0.4.0 & 2019-10-04 & CRAN (R 3.6.1)\\
-\addlinespace
-htmlwidgets & htmlwidgets & 1.3 & 2018-09-30 & CRAN (R 3.6.0)\\
+htmlwidgets & htmlwidgets & 1.5.1 & 2019-10-08 & CRAN (R 3.6.0)\\
 httpcode & httpcode & 0.2.0 & 2016-11-14 & CRAN (R 3.6.0)\\
+\addlinespace
+httpuv & httpuv & 1.5.2 & 2019-09-11 & CRAN (R 3.6.0)\\
 httr & httr & 1.4.1 & 2019-08-05 & CRAN (R 3.6.1)\\
+igraph & igraph & 1.2.4.1 & 2019-04-22 & CRAN (R 3.6.0)\\
+inline & inline & 0.3.15 & 2018-05-18 & CRAN (R 3.6.0)\\
 ipred & ipred & 0.9-9 & 2019-04-28 & CRAN (R 3.6.0)\\
+\addlinespace
 iterators & iterators & 1.0.12 & 2019-07-26 & CRAN (R 3.6.0)\\
-\addlinespace
-jpeg & jpeg & 0.1-8 & 2014-01-23 & CRAN (R 3.6.0)\\
+jpeg & jpeg & 0.1-8.1 & 2019-10-24 & CRAN (R 3.6.0)\\
 jqr & jqr & 1.1.0 & 2018-10-22 & CRAN (R 3.6.0)\\
@@
 jsonvalidate & jsonvalidate & 1.1.0 & 2019-06-25 & CRAN (R 3.6.0)\\
-kernlab & kernlab & 0.9-27 & 2018-08-10 & CRAN (R 3.6.0)\\
-\addlinespace
-KernSmooth & KernSmooth & 2.23-15 & 2015-06-29 & CRAN (R 3.6.1)\\
-knitr & knitr & 1.25 & 2019-09-18 & CRAN (R 3.6.1)\\
+\addlinespace
+kernlab & kernlab & 0.9-29 & 2019-11-12 & CRAN (R 3.6.0)\\
+KernSmooth & KernSmooth & 2.23-16 & 2019-10-15 & CRAN (R 3.6.0)\\
+knitr & knitr & 1.26 & 2019-11-12 & CRAN (R 3.6.0)\\
 labeling & labeling & 0.3 & 2014-08-23 & CRAN (R 3.6.0)\\
+later & later & 1.0.0 & 2019-10-04 & CRAN (R 3.6.1)\\
+\addlinespace
 lattice & lattice & 0.20-38 & 2018-11-04 & CRAN (R 3.6.1)\\
 lava & lava & 1.6.6 & 2019-08-01 & CRAN (R 3.6.0)\\
-\addlinespace
 lazyeval & lazyeval & 0.2.2 & 2019-03-15 & CRAN (R 3.6.0)\\
 lifecycle & lifecycle & 0.1.0 & 2019-08-01 & CRAN (R 3.6.0)\\
+listenv & listenv & 0.7.0 & 2018-01-21 & CRAN (R 3.6.0)\\
+\addlinespace
+lme4 & lme4 & 1.1-21 & 2019-03-05 & CRAN (R 3.6.0)\\
+loo & loo & 2.1.0 & 2019-03-13 & CRAN (R 3.6.0)\\
 lubridate & lubridate & 1.7.4 & 2018-04-11 & CRAN (R 3.6.0)\\
 magrittr & magrittr & 1.5 & 2014-11-22 & CRAN (R 3.6.0)\\
-maptools & maptools & 0.9-5 & 2019-02-18 & CRAN (R 3.6.0)\\
-\addlinespace
+maptools & maptools & 0.9-8 & 2019-10-05 & CRAN (R 3.6.0)\\
+\addlinespace
+markdown & markdown & 1.1 & 2019-08-07 & CRAN (R 3.6.0)\\
 MASS & MASS & 7.3-51.4 & 2019-03-31 & CRAN (R 3.6.1)\\
 Matrix & Matrix & 1.2-17 & 2019-03-22 & CRAN (R 3.6.1)\\
+matrixStats & matrixStats & 0.55.0 & 2019-09-07 & CRAN (R 3.6.0)\\
 memoise & memoise & 1.1.0 & 2017-04-21 & CRAN (R 3.6.0)\\
+\addlinespace
+mime & mime & 0.7 & 2019-06-11 & CRAN (R 3.6.0)\\
+miniUI & miniUI & 0.1.1.1 & 2018-05-18 & CRAN (R 3.6.0)\\
+minqa & minqa & 1.2.4 & 2014-10-09 & CRAN (R 3.6.0)\\
 ModelMetrics & ModelMetrics & 1.2.2 & 2018-11-03 & CRAN (R 3.6.0)\\
@@
 munsell & munsell & 0.5.0 & 2018-06-12 & CRAN (R 3.6.0)\\
-nlme & nlme & 3.1-141 & 2019-08-01 & CRAN (R 3.6.0)\\
+ncdf4 & ncdf4 & 1.17 & 2019-10-23 & CRAN (R 3.6.0)\\
+nlme & nlme & 3.1-142 & 2019-11-07 & CRAN (R 3.6.0)\\
+nloptr & nloptr & 1.2.1 & 2018-10-03 & CRAN (R 3.6.0)\\
 nnet & nnet & 7.3-12 & 2016-02-02 & CRAN (R 3.6.1)\\
+\addlinespace
 numDeriv & numDeriv & 2016.8-1.1 & 2019-06-06 & CRAN (R 3.6.0)\\
 optimx & optimx & 2018-7.10 & 2018-09-30 & CRAN (R 3.6.0)\\
-\addlinespace
 packrat & packrat & 0.5.0 & 2018-11-14 & CRAN (R 3.6.0)\\
-patchwork & patchwork & 0.0.1 & 2019-10-03 & Github (thomasp85/patchwork@36b4918)\\
+patchwork & patchwork & 1.0.0 & 2019-12-01 & CRAN (R 3.6.1)\\
 pillar & pillar & 1.4.2 & 2019-06-29 & CRAN (R 3.6.0)\\
-pkgbuild & pkgbuild & 1.0.5 & 2019-08-26 & CRAN (R 3.6.0)\\
+\addlinespace
+pkgbuild & pkgbuild & 1.0.6 & 2019-10-09 & CRAN (R 3.6.0)\\
 pkgconfig & pkgconfig & 2.0.3 & 2019-09-22 & CRAN (R 3.6.0)\\
-\addlinespace
 pkgload & pkgload & 1.0.2 & 2018-10-29 & CRAN (R 3.6.0)\\
+plotROC & plotROC & 2.2.1 & 2018-06-23 & CRAN (R 3.6.0)\\
 plyr & plyr & 1.8.4 & 2016-06-08 & CRAN (R 3.6.0)\\
+\addlinespace
 png & png & 0.1-7 & 2013-12-03 & CRAN (R 3.6.0)\\
@@
 processx & processx & 3.4.1 & 2019-07-18 & CRAN (R 3.6.0)\\
-\addlinespace
-prodlim & prodlim & 2018.04.18 & 2018-04-18 & CRAN (R 3.6.0)\\
+prodlim & prodlim & 2019.11.13 & 2019-11-17 & CRAN (R 3.6.1)\\
+progress & progress & 1.2.2 & 2019-05-16 & CRAN (R 3.6.0)\\
+\addlinespace
+promises & promises & 1.1.0 & 2019-10-04 & CRAN (R 3.6.1)\\
 ps & ps & 1.3.0 & 2018-12-21 & CRAN (R 3.6.0)\\
-purrr & purrr & 0.3.2 & 2019-03-15 & CRAN (R 3.6.0)\\
-R6 & R6 & 2.4.0 & 2019-02-14 & CRAN (R 3.6.0)\\
-Rcpp & Rcpp & 1.0.2 & 2019-07-25 & CRAN (R 3.6.0)\\
-\addlinespace
+purrr & purrr & 0.3.3 & 2019-10-18 & CRAN (R 3.6.0)\\
+R6 & R6 & 2.4.1 & 2019-11-12 & CRAN (R 3.6.0)\\
+RANN & RANN & 2.6.1 & 2019-01-08 & CRAN (R 3.6.0)\\
+\addlinespace
+Rcpp & Rcpp & 1.0.3 & 2019-11-08 & CRAN (R 3.6.0)\\
+RcppRoll & RcppRoll & 0.3.0 & 2018-06-05 & CRAN (R 3.6.0)\\
 readr & readr & 1.3.1 & 2018-12-21 & CRAN (R 3.6.0)\\
@@
 recipes & recipes & 0.1.7 & 2019-09-15 & CRAN (R 3.6.0)\\
+\addlinespace
 rEDM & rEDM & 0.7.4 & 2019-08-19 & Github (ha0ye/rEDM@88554a4)\\
 remotes & remotes & 2.1.0 & 2019-06-24 & CRAN (R 3.6.0)\\
-\addlinespace
 reshape2 & reshape2 & 1.4.3 & 2017-12-11 & CRAN (R 3.6.0)\\
-rgdal & rgdal & 1.4-4 & 2019-05-29 & CRAN (R 3.6.0)\\
-rgeos & rgeos & 0.5-1 & 2019-08-05 & CRAN (R 3.6.1)\\
+rfishbase & rfishbase & 3.0.4 & 2019-06-27 & CRAN (R 3.6.0)\\
+rgdal & rgdal & 1.4-7 & 2019-10-28 & CRAN (R 3.6.0)\\
+\addlinespace
+rgeos & rgeos & 0.5-2 & 2019-10-03 & CRAN (R 3.6.0)\\
 RgoogleMaps & RgoogleMaps & 1.4.4 & 2019-08-20 & CRAN (R 3.6.0)\\
 rjson & rjson & 0.2.20 & 2018-06-08 & CRAN (R 3.6.0)\\
-\addlinespace
-rlang & rlang & 0.4.0 & 2019-06-25 & CRAN (R 3.6.0)\\
+rlang & rlang & 0.4.2 & 2019-11-23 & CRAN (R 3.6.0)\\
 rmapshaper & rmapshaper & 0.4.1 & 2018-10-16 & CRAN (R 3.6.0)\\
-rmarkdown & rmarkdown & 1.16 & 2019-10-01 & CRAN (R 3.6.1)\\
+\addlinespace
+rmarkdown & rmarkdown & 1.17 & 2019-11-13 & CRAN (R 3.6.0)\\
 rpart & rpart & 4.1-15 & 2019-04-12 & CRAN (R 3.6.1)\\
 rprojroot & rprojroot & 1.3-2 & 2018-01-03 & CRAN (R 3.6.0)\\
-\addlinespace
+rsconnect & rsconnect & 0.8.15 & 2019-07-22 & CRAN (R 3.6.0)\\
+rstan & rstan & 2.19.2 & 2019-07-09 & CRAN (R 3.6.0)\\
+\addlinespace
+rstanarm & rstanarm & 2.19.2 & 2019-10-03 & CRAN (R 3.6.1)\\
+rstantools & rstantools & 2.0.0 & 2019-09-15 & CRAN (R 3.6.0)\\
 rstudioapi & rstudioapi & 0.10 & 2019-03-19 & CRAN (R 3.6.0)\\
-rticles & rticles & 0.10 & 2019-08-21 & CRAN (R 3.6.0)\\
+rticles & rticles & 0.12 & 2019-11-12 & CRAN (R 3.6.0)\\
 Rttf2pt1 & Rttf2pt1 & 1.3.7 & 2018-06-29 & CRAN (R 3.6.0)\\
-rvest & rvest & 0.3.4 & 2019-05-15 & CRAN (R 3.6.0)\\
-scales & scales & 1.0.0 & 2018-08-09 & CRAN (R 3.6.0)\\
-\addlinespace
+\addlinespace
+rvest & rvest & 0.3.5 & 2019-11-08 & CRAN (R 3.6.0)\\
+scales & scales & 1.1.0 & 2019-11-18 & CRAN (R 3.6.0)\\
 sessioninfo & sessioninfo & 1.1.1 & 2018-11-05 & CRAN (R 3.6.0)\\
 sf & sf & 0.8-0 & 2019-09-17 & CRAN (R 3.6.1)\\
-sp & sp & 1.3-1 & 2018-06-05 & CRAN (R 3.6.0)\\
-spasm & spasm & 0.1.0.9000 & 2019-10-09 & local\\
+shiny & shiny & 1.4.0 & 2019-10-10 & CRAN (R 3.6.0)\\
+\addlinespace
+shinyjs & shinyjs & 1.0 & 2018-01-08 & CRAN (R 3.6.0)\\
+shinystan & shinystan & 2.5.0 & 2018-05-01 & CRAN (R 3.6.0)\\
+shinythemes & shinythemes & 1.1.2 & 2018-11-06 & CRAN (R 3.6.0)\\
+sp & sp & 1.3-2 & 2019-11-07 & CRAN (R 3.6.0)\\
+spasm & spasm & 0.1.0.9000 & 2019-10-16 & local\\
+\addlinespace
+StanHeaders & StanHeaders & 2.19.0 & 2019-09-07 & CRAN (R 3.6.0)\\
 stringi & stringi & 1.4.3 & 2019-03-12 & CRAN (R 3.6.0)\\
-\addlinespace
 stringr & stringr & 1.4.0 & 2019-02-10 & CRAN (R 3.6.0)\\
-survival & survival & 2.44-1.1 & 2019-04-01 & CRAN (R 3.6.1)\\
+survival & survival & 3.1-7 & 2019-11-09 & CRAN (R 3.6.0)\\
 Synth & Synth & 1.1-5 & 2014-01-27 & CRAN (R 3.6.0)\\
+\addlinespace
 systemfonts & systemfonts & 0.1.1 & 2019-07-01 & CRAN (R 3.6.0)\\
-testthat & testthat & 2.2.1 & 2019-07-25 & CRAN (R 3.6.0)\\
-\addlinespace
+testthat & testthat & 2.3.0 & 2019-11-05 & CRAN (R 3.6.0)\\
+threejs & threejs & 0.3.1 & 2017-08-13 & CRAN (R 3.6.0)\\
 tibble & tibble & 2.1.3 & 2019-06-06 & CRAN (R 3.6.0)\\
 tidyr & tidyr & 1.0.0 & 2019-09-11 & CRAN (R 3.6.0)\\
+\addlinespace
 tidyselect & tidyselect & 0.2.5 & 2018-10-11 & CRAN (R 3.6.0)\\
@@
 timeDate & timeDate & 3043.102 & 2018-02-21 & CRAN (R 3.6.0)\\
-\addlinespace
-tinytex & tinytex & 0.16 & 2019-09-17 & CRAN (R 3.6.1)\\
-units & units & 0.6-4 & 2019-08-22 & CRAN (R 3.6.0)\\
+tinytex & tinytex & 0.17 & 2019-10-30 & CRAN (R 3.6.0)\\
+TMB & TMB & 1.7.15 & 2018-11-09 & CRAN (R 3.6.0)\\
+\addlinespace
+tweenr & tweenr & 1.0.1 & 2018-12-14 & CRAN (R 3.6.0)\\
+units & units & 0.6-5 & 2019-10-08 & CRAN (R 3.6.0)\\
 usethis & usethis & 1.5.1 & 2019-07-04 & CRAN (R 3.6.0)\\
@@
 withr & withr & 2.1.2 & 2018-03-15 & CRAN (R 3.6.0)\\
-xfun & xfun & 0.10 & 2019-10-01 & CRAN (R 3.6.1)\\
+xfun & xfun & 0.11 & 2019-11-12 & CRAN (R 3.6.0)\\
 xml2 & xml2 & 1.2.2 & 2019-08-09 & CRAN (R 3.6.0)\\
 \addlinespace
+xtable & xtable & 1.8-4 & 2019-04-21 & CRAN (R 3.6.0)\\
+xts & xts & 0.11-2 & 2018-11-05 & CRAN (R 3.6.0)\\
 yaml & yaml & 2.2.0 & 2018-07-25 & CRAN (R 3.6.0)\\
-zeallot & zeallot & 0.1.0 & 2018-01-28 & CRAN (R 3.6.0)\\*
+zeallot & zeallot & 0.1.0 & 2018-01-28 & CRAN (R 3.6.0)\\
+zoo & zoo & 1.8-6 & 2019-05-28 & CRAN (R 3.6.0)\\*
 \end{longtable}
@@
 \end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
-\textasciitilde{}uniform(0.6,0.95)\strut
+\textasciitilde uniform(0.6,0.95)\strut
 \end{minipage}\tabularnewline
@@
 \end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
-\textasciitilde{}uniform(0,0.5 * P)\strut
+\textasciitilde uniform(0,0.5 * P)\strut
 \end{minipage}\tabularnewline
@@
 \end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
-\textasciitilde{}uniform(0.01,1)\strut
+\textasciitilde uniform(0.01,1)\strut
 \end{minipage}\tabularnewline
@@
 \end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
-\textasciitilde{}uniform(0.01,4)\strut
+\textasciitilde uniform(0.01,4)\strut
 \end{minipage}\tabularnewline
@@
 \end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
-\textasciitilde{}uniform(0.1,1.25)\strut
+\textasciitilde uniform(0.1,1.25)\strut
 \end{minipage}\tabularnewline
@@
 \end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
-\textasciitilde{}round(uniform(5,0.66 * T))\strut
+\textasciitilde round(uniform(5,0.66 * T))\strut
 \end{minipage}\tabularnewline
@@
 \end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\centering
-\textasciitilde{}uniform(0.01,0.75)\strut
+\textasciitilde uniform(0.01,0.75)\strut
 \end{minipage}\tabularnewline
@@
 The goal of the estimation model is to estimate the difference in the mean densities of targeted and non-targeted species pre-and-post MPA implementation.
+
+At the rawest level, the data are counts of finfish in 2cm length bins along a 30m x 2m transect at various sites and depths.
+These length bins are converted to biomass, and then biomass densities, by converting length to weights using available allometric data and dividing by the transect area.
+Our goal is to estimate the effect of the MPAs on these biomass densities of fish throughout the Channel Islands.
+We fit this model using a hierarchical mixed-effect framework using Template Model Builder (TMB, Kristensen et al.
+2016) in R.
+The model consists of three levels, the first (starting from the ``bottom'') being transect-level densities of fish species observed by PISCO, which are standardized into a standardized biomass abundance index, accounting for both probability of detection and expected density as a result of changes in both abundance and covariates such as visibility and observer experience (see Maunder and Langley 2004).
+For the second stage, we break the abundance indices into targeted and non-targeted species (per the classifications in the PISCO data), and estimate the mean trend of each group (targeted and non-targeted) over time.
+In the third stage, we estimate the difference in the de-meaned trend between the targeted and non-targeted fishes (controlling for factors such as water temperature, kelp cover, and commercial fishery catches), that under the assumptions of the model reflects the causal effect of the MPAs on the outcome of interest (in this case regional biomass density of targeted fishes).
+All three of these steps are integrated into the same estimation model, in order to propagate uncertainty through the model correctly.
+Simpler versions of the estimation model, as well as a synthetic control identification strategy, were assessed as well and showed results consistent with the model in our main results.
 
@@
 \newpage
-
-\begin{figure}
-\centering
-\includegraphics{figs/estimation-cartoon.png}
-\caption{\label{fig:cartoon}Cartoon illustration of the hierarchichal difference-in-difference estimator}
+\begin{figure}
+\includegraphics[width=6in]{figs/estimation-cartoon} \caption{Cartoon illustration of the hierarchichal difference-in-difference estimator}\label{fig:cartoon}
 \end{figure}
@@
 
+\leavevmode\hypertarget{ref-kristensen2016}{}%
+Kristensen, Kasper, Anders Nielsen, Casper W.
+Berg, Hans Skaug, and Bradley M.
+Bell.
+2016.
+``\textbf{TMB} : Automatic Differentiation and Laplace Approximation.'' \emph{Journal of Statistical Software} 70 (5).
+\url{https://doi.org/10.18637/jss.v070.i05}.
+
 \leavevmode\hypertarget{ref-leeper2018}{}%
@@
 \url{https://doi.org/10.6073/pasta/817d2c24ebd78621869e17d94ba0df0c}.
+
+\leavevmode\hypertarget{ref-maunder2004}{}%
+Maunder, Mark N., and Adam D.
+Langley.
+2004.
+``Integrating the Standardization of Catch-Per-Unit-of-Effort into Stock Assessment Models: Testing a Population Dynamics Model and Using Multiple Data Types.'' \emph{Fisheries Research}, Models in Fisheries Research: GLMs, GAMS and GLMMs, 70 (23): 389--95.
+\url{https://doi.org/10.1016/j.fishres.2004.08.015}.
 

### 2020-04-26 functions/create-samples.R

@@ -1,5 +1,4 @@
 create_samples <- function(fishes,
-                           divers,
                            f_v_m = 1.5,
                            fleet_q = 0.001,
                            num_patches = 2,
@@ -12,7 +11,8 @@ create_samples <- function(fishes,
                            enviro_strength = 1,
                            rec_driver = 'stochastic',
                            sigma_r = 0,
-                           cores = 1) {
+                           cores = 1,
+                           cv = .1) {
   simple_fish <- fishes %>%
     mutate(fish = pmap(
       list(
@@ -23,21 +23,25 @@ create_samples <- function(fishes,
       ),
       create_fish,
       time_step = time_step,
-      sigma_r = sigma_r
+      sigma_r = sigma_r,
+      larval_movement = round(num_patches  *.25),
+      adult_movement = round(num_patches * .05),
+      steepness = 0.8
     )) %>%
     mutate(constant_f = map_dbl(fish, ~ .x$m *  f_v_m, f_v_m = f_v_m)) %>%
-    mutate(constant_effort = num_patches / 2 * ((constant_f / fleet_q * targeted) / time_step)) %>%
+    mutate(constant_effort =  ((constant_f / fleet_q) / time_step) * num_patches * targeted) %>%
     mutate(fleet = map2(
       constant_effort,
       fish,
       ~ create_fleet(
         initial_effort = .x,
         fish = .y,
-        q = fleet_q
+        q = fleet_q,
+        length_50_sel = .y$length_50_mature * .9
       ),
       fleet_q = fleet_q
     ))
-
+  
   simple_fish <- simple_fish %>%
     mutate(
       mpa_experiment = pmap(
@@ -56,7 +60,7 @@ create_samples <- function(fishes,
         rec_driver = rec_driver
       )
     )
-
+  
   simple_fish <- simple_fish %>%
     mutate(net_outcomes = map(mpa_experiment, 'outcomes')) %>%
     mutate(raw_outcomes = map(mpa_experiment, 'raw_outcomes')) %>%
@@ -82,72 +86,6 @@ create_samples <- function(fishes,
 
   simple_fish <- simple_fish %>%
     mutate(mpa_effect = map(raw_outcomes, calc_mpa_effect))
-
-  # cool
-  # ok now you need to think about how to set up your sampling regime.
-  # The idea now is that you can send out a bunch of observers to monitor the population each year...
-
[…]

### 2020-04-26 functions/create_diver_samples.R

@@ -1,0 +1,157 @@
+create_diver_samples <- function(fishes,
+                           divers,
+                           f_v_m = 1.5,
+                           fleet_q = 0.001,
+                           num_patches = 2,
+                           year_mpa = 50,
+                           sim_years = 100,
+                           burn_years = 50,
+                           mpa_size = 0.5,
+                           samples = 2,
+                           time_step = 1,
+                           enviro_strength = 1,
+                           rec_driver = 'stochastic',
+                           sigma_r = 0,
+                           cores = 1,
+                           cv = .1) {
+  simple_fish <- fishes %>%
+    mutate(fish = pmap(
+      list(
+        linf = loo,
+        vbk = k,
+        m = m,
+        length_50_mature = lm
+      ),
+      create_fish,
+      time_step = time_step,
+      sigma_r = sigma_r
+    )) %>%
+    mutate(constant_f = map_dbl(fish, ~ .x$m *  f_v_m, f_v_m = f_v_m)) %>%
+    mutate(constant_effort = num_patches / 2 * ((constant_f / fleet_q * targeted) / time_step)) %>%
+    mutate(fleet = map2(
+      constant_effort,
+      fish,
+      ~ create_fleet(
+        initial_effort = .x,
+        fish = .y,
+        q = fleet_q
+      ),
+      fleet_q = fleet_q
+    ))
+
+  simple_fish <- simple_fish %>%
+    mutate(
+      mpa_experiment = pmap(
+        list(
+          fish = fish,
+          fleet = fleet,
+          enviro = enviro
+        ),
+        run_mpa_experiment,
+        year_mpa = year_mpa,
+        sim_years = sim_years,
+        num_patches = num_patches,
+        burn_years =  burn_years,
+        mpa_size = mpa_size,
+        enviro_strength = enviro_strength,
+        rec_driver = rec_driver
+      )
+    )
+
+  simple_fish <- simple_fish %>%
+    mutate(net_outcomes = map(mpa_experiment, 'outcomes')) %>%
+    mutate(raw_outcomes = map(mpa_experiment, 'raw_outcomes')) %>%
+    mutate(mpa_plot = map(net_outcomes, ~ .x %>%
+                            ungroup() %>%
+                            ggplot(
+                              aes(year, ssb, color = experiment, linetype = experiment)
[…]

### 2020-04-26 functions/estimate_did.R

@@ -1,0 +1,590 @@
+estimate_did <-
+  function(data,
+           site_data,
+           life_history_data,
+           bin_width = 3,
+           min_classcode_years = 15,
+           consistent_sites_only = TRUE,
+           cdfw_catches,
+           data_to_use = "all",
+           data_source = "pisco",
+           fit_mean = TRUE,
+           chains = 1,
+           cores = 1,
+           refresh = 500,
+           first_year = 2003,
+           iter = 5000,
+           weight_samples = TRUE) {
+    # data <- pisco_abundance_data
+    #
+    # data <- kfm_abundance_data
+    
+    if (data_to_use == "mpa_only") {
+      data <- data %>%
+        filter(eventual_mpa == TRUE)
+      
+    } else if (data_to_use == "fished_only") {
+      data <- data %>%
+        filter(eventual_mpa == FALSE)
+    }
+    
+    annual_catches <- cdfw_catches %>%
+      filter(classcode %in% unique(data$classcode)) %>%
+      group_by(year) %>%
+      summarise(var_catch = sum(catch)) %>%
+      mutate(var_lag_catch = lag(var_catch, 1))
+    
+    annual_catches$var_lag_catch[1] <-
+      annual_catches$var_lag_catch[2] #assume 1999 catches = 2000 catches, not crazy given AC in totals
+    
+    consistent_did_sites <- data %>%
+      select(site_side, year) %>%
+      unique() %>%
+      group_by(site_side) %>%
+      mutate(has_all = all((2003:2003) %in% year)) %>% # results change dramatically if you filter until present time...
+      ungroup() %>%
+      filter(has_all)
+    
+    consistent_classcodes <- data %>%
+      group_by(classcode) %>%
+      filter(any_seen) %>%
+      summarise(n_years = n_distinct(year)) %>%
+      ungroup() %>%
+      filter(n_years >= min_classcode_years)
+    
+    year_bins <-
+      c(2000, seq(first_year, max(data$year) + 1, by = bin_width))
+    
+    did_data <- data %>% {
+      if (consistent_sites_only == TRUE) {
+        filter(.,
+               site_side %in% unique(consistent_did_sites$site_side))
+      } else{
+        .
+      }
+    } %>%
+      filter(classcode %in% consistent_classcodes$classcode)
+    
+    # rm("data","cdfw_catches","life_history_data")
+    if (data_source == "pisco") {
+      # did_data %>%
+      #         ggplot(aes(regional_temp_dev)) +
[…]

### 2020-04-26 functions/fit-ahnold.R

@@ -1,270 +1,0 @@
-fit_ahnold <- function(data,
-                       non_nested_variables =  c(
-                         'targeted',
-                         'factor_year',
-                         'level',
-                         'factor_month',
-                         'cumulative_n_obs',
-                         'temp_deviation',
-                         'surge',
-                         'mean_canopy',
-                         'mean_depth'
-                       ),
-                       script_name,
-                       seed = 42,
-                       run_dir,
-                       use_tmb = T,
-                       fixed_did = T,
-                       include_intercept = T,
-                       center_and_scale = T,
-                       fixed_regions = F) {
-
-
-  numeric_species_key <-
-    data_frame(classcode = unique(data$classcode)) %>%
-    arrange(classcode) %>%
-    mutate(numeric_classcode = 1:nrow(.))
-
-
-  seen_has_important <- data %>%
-    filter(any_seen == T) %>%
-    select(non_nested_variables) %>%
-    mutate(index = 1:nrow(.)) %>%
-    na.omit()
-
-  seeing_has_important <- data %>%
-    select(non_nested_variables) %>%
-    mutate(index = 1:nrow(.)) %>%
-    na.omit()
-
-  seen_data <- data %>%
-    filter(any_seen == T) %>%
-    left_join(numeric_species_key, by = "classcode") %>%
-    slice(seen_has_important$index)
-
-  log_density <- seen_data$log_density
-
-  seen_data <- seen_data %>%
-    select(-log_density)
-
-  seeing_data <- data %>%
-    left_join(numeric_species_key, by = "classcode") %>%
-    slice(seeing_has_important$index)
-
-  any_seen <- seeing_data$any_seen
-
-  seeing_data <- seeing_data %>%
-    select(-any_seen)
-
-
-  # prepare data for c++ ----------------------------------------------------
-
-
-browser()
-  seen_cdata <-
-    make_c_worthy(seen_data,
-                  non_nested_vars = non_nested_variables,
-                  fixed_regions = fixed_regions,
-                  include_intercept = include_intercept,
-                  fixed_did = fixed_did,
[…]

### 2020-04-26 functions/fit-zissou.R

@@ -21,7 +21,7 @@ fit_zissou <- function(data,
                        include_intercept = T,
                        center_scale = T,
                        fixed_regions = F,
-                       mpa_only = F,
+                       data_to_use = 'all',
                        bin_years = FALSE) {
 
   # data <- data %>%
@@ -36,12 +36,16 @@ fit_zissou <- function(data,
   }
 
 
-  if (mpa_only == T){
+  if (data_to_use == "mpa_only"){
 
     data <- data %>%
-      filter(eventual_mpa == T)
+      filter(eventual_mpa == TRUE)
   }
 
+  if (data_to_use == "fished_only"){
+    data <- data %>%
+      filter(eventual_mpa == FALSE)
+  }
 
   if (center_scale == T){
 
@@ -274,9 +278,9 @@ fit_zissou <- function(data,
 
   if (use_tmb == T) {
 
-    compile(here::here("src", paste0(script_name, ".cpp")), "-O0") # what is the -O0?
-
-    dyn.load(dynlib(here::here("src", script_name)))
+    # compile(here::here("src", paste0(script_name, ".cpp")), "-O0") # what is the -O0?
+    # 
+    # dyn.load(dynlib(here::here("src", script_name)))
 
     if (fixed_regions == T){
 

### 2020-04-26 functions/run-mpa-experiment.R

@@ -35,7 +35,7 @@ run_mpa_experiment <- function(fish,
       min_size = min_size
     ) %>%
     mutate(experiment = 'with-mpa')
-
+  
   mpa_locations <- unique(wi_mpa$patch[wi_mpa$mpa == TRUE])
 
   set.seed(simseed)
@@ -75,9 +75,11 @@ run_mpa_experiment <- function(fish,
       effort = sum(effort)
     ) %>%
     ungroup()
+  
   raw_outcomes <- no_mpa %>%
-    bind_rows(wi_mpa)
-
+    bind_rows(wi_mpa) %>% 
+    mutate(eventual_mpa = patch %in% mpa_locations)
+  
   out <- list(outcomes = outcomes,
               raw_outcomes = raw_outcomes)
 

### 2020-04-26 functions/study-mpa.R

@@ -77,7 +77,10 @@ density_ratio <- results$mpa_experiment[[1]]$raw_outcomes %>%
                                           experiment == "with-mpa"],
                                 w = min_fished_dist[eventual_mpa == TRUE &
                                                       experiment == "with-mpa"]),
-    nompa_density = mean(biomass[experiment == "no-mpa"]),
+    nompa_density =weighted.mean(biomass[eventual_mpa == TRUE &
+                                           experiment == "no-mpa"],
+                                 w = min_fished_dist[eventual_mpa == TRUE &
+                                                       experiment == "no-mpa"]),
     mpa_period = any(mpa[experiment == "with-mpa"] == TRUE)
   ) %>%
   ungroup() %>%

### 2020-04-26 functions/test-performance.R

@@ -3,24 +3,30 @@ test_performance <-
            year_mpa,
            min_year = 75,
            max_year = 100,
-           time_step = 1) {
+           time_step = 1,
+           sigma_obs = 0,
+           samps_per_patch = 1) {
+    
+    
     simple_data <- fishes %>%
-      select(loo, k, lm, m, targeted, classcode, commonname, pisco_samples) %>%
-      unnest() %>%
-      filter(year > min_year, year < max_year) %>%
-      select(-pop,-sampled_lengths,-diver_stats) %>%
+      select(loo, k, lm, m, targeted, classcode, commonname,raw_outcomes) %>%
+      unnest(cols = raw_outcomes) %>%
+      group_by(year, patch, experiment, targeted) %>% 
+      summarise(biomass_density = list(rlnorm(samps_per_patch,log(sum(biomass) / 100 + 1e-3), sigma_obs))) %>% 
+      unnest(cols = biomass_density) %>% 
+      ungroup() %>% 
+      filter(year > min_year, year < max_year,
+             experiment == "with-mpa") %>%
       mutate(year = year * time_step) %>%
       mutate(subyear = year - floor(year),
              year = floor(year)) %>%
       mutate(
         factor_year = as.factor(year),
-        log_density = log(density),
+        log_density = log(biomass_density),
         logical_targeted = targeted > 0,
-        any_seen = density > 0,
-        region = patches,
         post_mpa = year >= year_mpa
       )
-
+    
     years_protected <- unique(simple_data$year) - year_mpa
 
     bins <-
@@ -39,143 +45,84 @@ test_performance <-
       left_join(year_block, by = "year")
 
     true_effect <- fishes %>%
-      select(classcode, targeted, mpa_effect) %>%
-      unnest() %>%
-      # mutate(mpa_effect = log(`with-mpa`) - log(`no-mpa`)) %>%
-      filter(year > min_year, targeted == 1) %>%
+      select(targeted, classcode, commonname, raw_outcomes) %>%
+      unnest(cols = raw_outcomes) %>%
+      group_by(year, patch, experiment, targeted) %>%
+      summarise(b = sum(biomass)) %>%
+      group_by(year, experiment, targeted) %>%
+      summarise(mbd = mean(b)) %>%
+      ungroup() %>%
+      pivot_wider(names_from = experiment,
+                  values_from = mbd) %>%
[…]

### 2020-04-26 make-pop-effects-of-mpas.R

@@ -1,0 +1,4286 @@
+# Run Complete Zissou Analysis --------------------------------------------
+
+
+# libraries ---------------------------------------------------------------
+set.seed(42)
+library(scales)
+library(viridis)
+library(ggmap)
+library(forcats)
+library(stringr)
+library(lubridate)
+library(lme4)
+library(TMB)
+library(TMBhelper)
+library(FishLife)
+library(patchwork)
+library(rstan)
+library(rstanarm)
+library(extrafont)
+library(hrbrthemes)
+library(here)
+library(doParallel)
+library(spasm)
+library(furrr)
+library(ggsci)
+library(recipes)
+library(sf)
+library(ggmap)
+library(rEDM)
+library(bayesplot)
+library(tabulizer)
+library(tidyverse)
+library(ggtext)
+library(ggrepel)
+library(rnaturalearth)
+library(rnaturalearthhires)
+extrafont::loadfonts()
+# extrafont::font_import()
+rstan_options(auto_write = TRUE)
+Sys.unsetenv("PKG_CXXFLAGS")
+
+
+functions <- list.files(here::here("functions"))
+
+walk(functions, ~ here::here("functions", .x) %>% source()) # load local functions
+
+
+# options -----------------------------------------------------------------
+
+run_name <- 'v6.0'
+
+run_description <- "PNAS R and R with simplified DiD and weighting"
+
+# the following analysis run the complete contents of "regional-effects-of-mpas". Each section depends on the out
+# outcomes of the prior section, but will load relevant saved files.
+# So, once you've run simulate_mpas, you can set it to FALSE and validata_mpas will work
+
+
+run_did <- FALSE # run difference in difference on data from the CINMS
+
+run_tmb <- FALSE
+
+simulate_mpas <- FALSE # simulate MPA outcomes
+
+simulate_channel_islands <- FALSE # simulate MPA outcomes
+
+validate_mpas <- FALSE
+
+process_results <- TRUE
+
+knit_paper <- FALSE
+
+get_cdfw_catches <-  FALSE
+
+sim_years <- 50
+
+num_patches <- 50
+
+n_cores <- 1
+
+# prepare run -------------------------------------------------------------
+
+run_dir <- here::here("results", run_name)
+
+if (dir.exists(run_dir) == F) {
[…]

### 2020-04-26 make-regional-mpas.R

@@ -1,3457 +1,0 @@
-# Run Complete Zissou Analysis --------------------------------------------
-
-
-# libraries ---------------------------------------------------------------
-set.seed(42)
-library(scales)
-library(viridis)
-library(ggmap)
-library(forcats)
-library(stringr)
-library(lubridate)
-library(lme4)
-library(TMB)
-library(FishLife)
-library(patchwork)
-library(rstan)
-library(rstanarm)
-library(extrafont)
-library(hrbrthemes)
-library(here)
-library(doParallel)
-library(spasm)
-library(furrr)
-library(ggsci)
-library(recipes)
-library(sf)
-library(ggmap)
-library(rEDM)
-library(bayesplot)
-library(tidyverse)
-extrafont::loadfonts()
-
-
-
-
-functions <- list.files(here::here("functions"))
-
-walk(functions, ~ here::here("functions", .x) %>% source()) # load local functions
-
-
-# options -----------------------------------------------------------------
-
-run_name <- 'v5.0'
-
-run_description <- "now with recruitment deviates and autocorrleation - testing everything works post rebase with master after that screwup before mering back into master"
-
-# the following analysis run the complete contents of "regional-effects-of-mpas". Each section depends on the out
-# outcomes of the prior section, but will load relevant saved files.
-# So, once you've run simulate_mpas, you can set it to FALSE and validata_mpas will work
-
-
-run_did <- TRUE # run difference in difference on data from the CINMS
-
-simulate_mpas <- FALSE # simulate MPA outcomes
-
-validate_mpas <- FALSE
-
-process_results <- FALSE
-
-knit_paper <- TRUE
-
-sim_years <- 50
-
-num_patches <- 50
-
-n_cores <- 4
-
-# prepare run -------------------------------------------------------------
-
-run_dir <- here::here("results", run_name)
-
-if (dir.exists(run_dir) == F) {
-  dir.create(run_dir, recursive = T)
-}
-
-data_dir <- "data"
-
-write(run_description,
-      file = paste(run_dir, 'RUN_DESCRIPTION.txt', sep = '/'))
-
[…]

### 2020-04-26 README.Rmd
cleaning up repository

@@
 
-# The Regional Effects of Marine Protected Areas
+# Assessing the Population-level Conservation Effects of Marine Protected Areas
 
-This repo contains code and data needed to reproduce "The Regional Effects of Marine Protected Areas" by Ovando et al. In Prep.
+This repo contains code and data needed to reproduce "Assessing the Population-level Conservation Effects of Marine Protected Areas" by Ovando et al. In Prep.
+
+The script `make-pop-effects-of-mpas.R` will reproduce all results and manuscripts for this project.
 
@@
 MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present), and the size of many MPA networks suggests that effect sizes may often be smaller than 30%.
-Our results provide a novel assessment of the population-wide effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.
+Our results provide a novel assessment of the population-level effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.

### 2020-04-26 documents/ovando-et-al-pop-effects-of-mpas.Rmd
make sure documents/figs folder is present

@@
   mutate(mpa_effect = pmin(mpa_effect,3))
+
+ if (!dir.exists(here("documents", "figs"))){
+
+   dir.create(here("documents", "figs"))
+
+ }
 ```

### 2020-04-28 documents/ovando-et-al-pop-effects-of-mpas-si.Rmd
a bit of cleanup

@@
 ---
-title: "Supporting Information"
+title: "Supporting Information - Appendix"
 subtitle: "Assessing the Population-level Conservation Effects of Marine Protected Areas"
@@
 \newpage
-```{r cartoon, fig.cap= "Cartoon illustration of the hierarchichal difference-in-difference estimator", fig.width=3}
+```{r cartoon, fig.cap= "Cartoon illustration of the hierarchichal difference-in-difference estimator", fig.width=8}
 knitr::include_graphics("figs/estimation-cartoon.png")

### 2020-04-28 documents/ovando-et-al-pop-effects-of-mpas.Rmd
a bit of cleanup

@@
 output:
+  bookdown::word_document2:
+    reference_docx: template.docx
   bookdown::pdf_document2: default
-  bookdown::word_document2: default
   bookdown::html_document2: default
@@
 knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
-                      cache = FALSE, fig.width = 3.42, fig.asp = 1.5, dev = "cairo_pdf")
+                      cache = FALSE, fig.width = 3.42, fig.asp = 1.5, dev = "png")
 
@@
 library(ggtext)
+library(here)
 extrafont::loadfonts()
@@
 
+tmdata <- used_data %>% 
+  filter(targeted == 1) %>% 
+  select(tm, everything())
+
+median_tm = round(median(tmdata$tm))
+
 site_data <- read_csv(here::here("data",'Final_Site_Table_UCSB.csv')) %>%
@@
 
-# Body
+# Main Text
 
@@
 While we know many parameters will affect the expected effects of the MPAs, theory suggests that life history, adult and larval dispersal relative to MPA size, and fishing pressure will be key drivers.
-The targeted species in our database span a range of life histories, but are largely made up of fishes in the perch and rockfish complexes, with a mean Von Bertalanffy growth coefficient of `r round(mean(used_data$k),2)`, and a median age at maturity of `r round(median(used_data$tm),0)` years [values drawn from @thorson2017c].
+The targeted species in our database span a range of life histories, but are largely made up of fishes in the perch and rockfish complexes, with a mean Von Bertalanffy growth coefficient of `r round(mean(used_data$k[used_data$targeted == 1]),2)`, and a median age at maturity of `r round(median(used_data$tm[used_data$targeted == 1]),0)` years [values drawn from @thorson2017c].
 
@@
 
-To illustrate this problem, we simulated response ratio trajectories for species and MPA coverage representative of the Channel Islands (see methods and SI for details of model structure).
+To illustrate this problem, we simulated response ratio trajectories for species and MPA coverage representative of the Channel Islands (see methods and Supporting Information (SI) Appendix for details of model structure).
 These simulations cover the life-histories and MPA sizes seen in the Channel Islands, but vary in key unknowns such as the actual degree of fishing pressure, the timing of density dependence, and the fishing fleets response to the MPAs.
@@
 Response ratios well over one were associated with population-level MPA effects generally less than 25%, and, importantly, many simulations produced large response ratios but population-level MPA effects close to 0% This can occur if for example fishing pressure is only moderate, adult movement is low, larval dispersal is high, and displaced fishing effort concentrates around the border of the MPAs.
-
-```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="90% Posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis) in blue.
-Simulated population-level (pop.) effect on biomass densities matched to empirical response ratios in red.
-For response ratios, a value of zero indicates that biomass densities of targeted species are identical inside and outside MPAs, a value of one that biomass densities of targeted species are 100% inside MPAs relative to outside.
-For MPA population effect, a value of zero indicates that biomass densities are identical in the with- and without- MPA scenarios.
-A value of 1 indicates that biomass densities are 100% greater in the scenario with MPAs than the scenario without MPAs."}
-
-temp_plot <- response_ratio_plot + 
-  theme_1col + 
-  theme(axis.title.x = element_textbox(size = 10))
-
-ggsave(
-  temp_plot,
-  filename = here::here("documents","figs","response_ratio_plot.png"),
-  width = width_1col,
-  height = width_1col * asp_1col,
-  dpi = dpi
-)
-
-temp_plot
-# pdf(file =  "figs/density-ratio.png",width = 8, height = 5, useDingbats = TRUE)
-# print(dr_plot)
-# dev.off()
-
-```
 
@@
 - `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`.) Fig.\@ref(fig:mpa-effect)
-
-```{r mpa-effect, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Results of difference-in-difference regression estimating the population effect of the Channel Island MPAs on mean total biomass densities of targeted species (difference in mean total biomass density of targeted species over time relative to expected levels using non-targeted species as a control).
-Grey distributions show posterior probability distribution of estimated MPA effect; red point is median estimated effect, thicker red section 50% credible interval, thinner red line 90% credible interval).
-Blue distributions in background show range of MPA effects produced by simulation model tuned to reflect the dynamics of the Channel Island MPAs (black dashed line is median simulated value).
-Results are estimated in blocks of three years, including years greater than or equal to left-hand value and less than right-hand value."}
-
-mpa_effect_plot
-
-ggsave(
-  mpa_effect_plot,
-  filename = here::here("documents","figs","mpa_effect_plot.png"),
-  width = width_2col,
-  height = width_2col * asp_2col,
-  dpi = dpi
-)
-
-```
 
@@
 
-While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (see Fig.S1) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
-However, some of the most common targeted species in the Channel Islands, such as blue rockfish and copper rockfish, have seen dramatic increases in reported catches in the Santa Barbara region since the year 2010 (although other such as California sheephead have seen declines relative to pre-MPA catch level, Fig.S1).
+While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (see SI Appendix Fig.S1) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
+However, some of the most common targeted species in the Channel Islands, such as blue rockfish and copper rockfish, have seen dramatic increases in reported catches in the Santa Barbara region since the year 2010 (although other such as California sheephead have seen declines relative to pre-MPA catch level).
 To the extent that these increases in catches are due to external market forces, the dissapearance of clear positive population-level MPA effects estimated by our model could be due to exogenous increases in fishing pressure overwhelming any population-level increases caused by the MPA network.
@@
 
-This hypothesis is supported by the fact that we estimate a similar decline in the estimated MPA effect both inside and outside of the MPAs (Fig.S18) (as well as observe the decline in the raw data, Fig.\@ref(fig:raw-density)).
+This hypothesis is supported by the fact that we estimate a similar decline in the estimated MPA effect both inside and outside of the MPAs (SI Appendix Fig.S18) (as well as observe the decline in the raw data, Fig.\@ref(fig:raw-density)).
 If the cause of recent declines was due to increases in fishing pressure we would expect to see substantial declines only in the fished areas.
-While we see evidence for the validity of the parallel trends assumption in the early years of our data (Fig.\@ref(fig:mpa-effect), Fig.S26), it is likely that the non-targeted species reacted in a different way to the marine heatwave than the targeted species did, causing a violation of the parallel trends assumption in the post 2014 data.
-
-```{r raw-density, fig.width = width_1col, fig.asp=asp_1col,fig.cap = "Centered and scaled trends in biomass densities of targeted and non-targeted finfish included in our study Top panel shows trends across all sites, with smaller background lines showing trends for each individual species.
-Bottom two panels show aggregate biomass density trends outside and inside MPAs"}
-
-temp_plot <-
-  (total_trend_plot / (mpa_trend_plot)) & theme(legend.position = "top", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), units = "lines"),
-                              panel.grid = element_blank()
-)
-
-temp_plot
-
-ggsave(
-temp_plot,
-filename = here::here("documents","figs","raw_biomass_density_plot.png"),
-  width = width_1col,
-  height = width_1col * asp_1col,
-    dpi = dpi
-)
-
-```
+While we see evidence for the validity of the parallel trends assumption in the early years of our data (Fig.\@ref(fig:mpa-effect), SI Appendix Fig.S26), it is likely that the non-targeted species reacted in a different way to the marine heatwave than the targeted species did, causing a violation of the parallel trends assumption in the post 2014 data.
 
@@
 
-```{r}
-tmdata <- life_history_data %>% 
-  filter(targeted == 1, classcode %in% top_species) %>% 
-  select(tm, everything())
-
-median_tm = round(median(tmdata$tm))
-```
-
 Our simulations and difference-in-difference model examine percentage changes in biomass densities (as opposed to changes in total population size).
@@
 While a large percentage effect, this is a small change in biomass densities relative to the variance in the observation process itself.
-While an MPA providing 25% protection might be expected to produce a 30% increase in biomass densities, expressed as a percentage of unfished biomass, this gain is on the order of 15% (Fig.S45).
+While an MPA providing 25% protection might be expected to produce a 30% increase in biomass densities, expressed as a percentage of unfished biomass, this gain is on the order of 15% (SI Appendix Fig.S45).
 In addition, the median estimated age at sexual maturity for the targeted species included in this study is `r median_tm` years, meaning that the span of this analysis represents less than three generations of MPA protection for half of the measured species.
@@
 
-```{r val-plot, include = TRUE,fig.width = width_1col, fig.asp=asp_1col, fig.cap = "Distribution of percent error in posterior estimates of population-level MPA effect (y-axis) plotted against true simulated MPA effect (x-axis).
-Color shows concentration of simulations.
-Black line shows mean absolute percent error (MAPE) as a function of true simulated MPA effect."}
-
-temp_plot <- validation_plot + 
-  theme_1col + 
-   theme(legend.position = "top") + 
-    scale_fill_viridis(option = "C",
-                     name = "Sims(#)",
-                     guide = hgc <- guide_colorbar(frame.colour = "black",
-                                                   ticks.colour = "black",
-                                                   barwidth = unit(5,"lines")))
-
-temp_plot
-ggsave(
-  temp_plot,
-  filename = here::here("documents","figs","validation_plot.png"),
-  width = width_1col,
-  height = width_1col * asp_1col,
-    dpi = dpi
-
-)
-
-```
-
 When should we expect to see MPA effects big enough to stand a reasonable chance of detection?
-We simulated `r n_distinct(outcomes$experiment)` MPA scenarios across a wide range of life histories, network designs, and fishing dynamics (see SI).
+We simulated `r n_distinct(outcomes$experiment)` MPA scenarios across a wide range of life histories, network designs, and fishing dynamics.
 Suppose that we are willing to tolerate a MAPE of 25%.
@@
 The population-level effects of MPA networks covering 25% or less of a species range may be difficult to detect in many places with already well-managed fisheries, while for that MPA network size we might expect clearer effects in less-managed locations (assuming compliance with MPA regulations).
-Within these broad guidelines a wide range of outcomes are possible based on local fleet and fish dynamics: as a starting place users can use the bio-economic MPA simulation model developed for this paper to explore potential outcomes for specific MPAs using an interactive web application available [here](https://danovando.shinyapps.io/simmpa/).
-
-```{r sim-effects, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Simulated population-level (pop.) MPA effect sizes as a function of percent of species' range inside MPA (x-axis), and pre-MPA depletion (y-axis).
-Pre-MPA depletion is a measure of fishing pressure, where 0 means that the population is unfished, and 1 means that the population is extinct in the time period immediately prior to MPA implementation.
-Panel A) shows median MPA effects across range in MPA and pre-MPA depletion Panel B) shows distribution of simulations across range of MPA size and pre-MPA depletion separately."}
-
-expected_mpa_effect_plot
-ggsave(
-  expected_mpa_effect_plot,
-  filename = here::here("documents","figs","expected_mpa_effect.png"),
-  width = width_2col,
-  height = width_2col * asp_2col,
-    dpi = dpi
-
-)
-
-```
+Within these broad guidelines a wide range of outcomes are possible based on local fleet and fish dynamics: users can use the bio-economic MPA simulation model developed for this paper to explore potential outcomes for specific MPAs using an interactive web application available [danovando.shinyapps.io/simmpa/](https://danovando.shinyapps.io/simmpa/).
 
@@
 Non-equilibrium analyses also help set expectations for effect sizes over time [@nickols2019,@kaplan2019,Fig.\ref{did-plot}].
-Educating communities about the challenges of estimating the effects of MPAs can help set expectations, so that a lack of a clear effect is not necessarily viewed as a failure of the program, or large positive result based solely on response ratios as a clear sign of success and subsequent relaxation of other fishery management strategies.
+Educating communities about the challenges of estimating the effects of MPAs can help ensure that a lack of a clear effect is not necessarily viewed as a failure of the program, or large positive result based solely on response ratios as a clear sign of success and subsequent relaxation of other fishery management strategies.
 Rather, results and subsequent management actions (such as adaptation of MPA networks) must be considered in the context of reasonable expectations given the size, age, and degree of enforcement of the MPAs in question, together with the ecological and economic dynamics of a given system.
@@
 Our main difference-in-difference model was fit using Stan [@carpenter2017] using the `rstanarm` package [@goodrich2020].
-All materials needed to reproduce results can be found [here](https://github.com/DanOvando/regional-effects-of-mpas).
+All data and code needed to fully reproduce our results are publicly available at [github.com/DanOvando/population-effects-of-mpas](https://github.com/DanOvando/population-effects-of-mpas).
 
@@
 Our bio-economic model simulates the effect of MPAs on a spatially explicit age-structured representation of a single species.
-Readers can explore the functionality of the model using an online tool available [here](https://danovando.shinyapps.io/simmpa/).
+Readers can explore the functionality of the model using an online tool available at [danovando.shinyapps.io/simmpa/](https://danovando.shinyapps.io/simmpa/).
 Full details of the simulation model are available in the Supplementary Information.
@@
 ### Difference in Difference Regression
-
-```{r ci-map, include = TRUE,fig.width=width_1col, fig.asp=1.5, fig.cap="Map of study region; the Northern Channel Islands, California, USA.
-Colors show binned number of PISCO sampling events across the time period of our study.", include=TRUE}
-
-temp_plot <- sample_location_plot + 
-  theme(axis.text = element_blank())
-
-temp_plot
-ggsave(
- temp_plot,
-  filename = here::here("documents","figs","sample_location_plot.png"),
-  width = width_1col,
-  height = width_1col * asp_1col,
-    dpi = dpi
-
-)
-
-```
 
@@
 
-The key assumptions of the difference-in-difference model are that a) within the time-frame of the model there are no significant interaction effects between the targeted and non-targeted species (which in fact we do not detect, see SI), and b) that in the absence of the MPAs both the targeted and non-targeted groups of species would have exhibited similar trends in densities.
+The key assumptions of the difference-in-difference model are that a) within the time-frame of the model there are no significant interaction effects between the targeted and non-targeted species (which in fact we do not detect, see SI Appendix), and b) that in the absence of the MPAs both the targeted and non-targeted groups of species would have exhibited similar trends in densities.
 It is important to note that the model is agnostic as to the root causes of the parallel trends, so long as they hold.
@@
 The shape and scale parameters of the Gamma distribution are estimated as well.
-See SI for further details of the estimation model.
 `r percent(mean(used_data$eventual_mpa))` of the samples in the raw data come from within MPAs, while the MPAs themselves cover 20% of the surface waters of the Channel Islands.
 As such, we weight each observation such that within MPA data are assigned a total weight of 0.2 and outside-MPA data a total weight of 0.8.
-See SI for sensitivity analyses of various model structures and synthetic controls results.
+See SI Appendix for sensitivity analyses of various model structures and synthetic controls results.
 
@@
 While the parallel trends assumption cannot be formally proven, we can examine its validity using the data from the years before the MPAs were put in place in 2003.
-We do not detect any significant differences in the trends of the biomass densities of the targeted and non-targeted species in the years before the MPAs (Fig.S26).
+We do not detect any significant differences in the trends of the biomass densities of the targeted and non-targeted species in the years before the MPAs (SI Appendix Fig.S26).
 However, we believe that the parallel trends assumption is likely violated in the post 2014 data.
@@
 We used convergent cross mapping (CCM), in the manner of @clark2015, to test for the possibility of the trophic cascades biasing our results.
-Our results found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (Fig.S42).
+Our results found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (SI Appendix Fig.S42).
 However, the longer MPAs are in place, the greater the possibility that substantial species interactions that can affect use of non-targeted species as a control may arise.
+
+# Figures
+
+```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="90% Posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis) in blue.
+Simulated population-level (pop.) effect on biomass densities matched to empirical response ratios in red.
+For response ratios, a value of zero indicates that biomass densities of targeted species are identical inside and outside MPAs, a value of one that biomass densities of targeted species are 100% inside MPAs relative to outside.
+For MPA population effect, a value of zero indicates that biomass densities are identical in the with- and without- MPA scenarios.
+A value of 1 indicates that biomass densities are 100% greater in the scenario with MPAs than the scenario without MPAs."}
+
+temp_plot <- response_ratio_plot + 
+  theme_1col + 
+  theme(axis.title.x = element_textbox(size = 10))
+
+ggsave(
+  temp_plot,
+  filename = here::here("documents","figs","response_ratio_plot.pdf"),
+  device = cairo_pdf,
+  width = width_1col,
+  height = width_1col * asp_1col,
+  units = "in",
+)
+
+temp_plot
+# pdf(file =  "figs/density-ratio.png",width = 8, height = 5, useDingbats = TRUE)
+# print(dr_plot)
+# dev.off()
+
+```
+
+```{r mpa-effect, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Results of difference-in-difference regression estimating the population effect of the Channel Island MPAs on mean total biomass densities of targeted species (difference in mean total biomass density of targeted species over time relative to expected levels using non-targeted species as a control).
+Grey distributions show posterior probability distribution of estimated MPA effect; red point is median estimated effect, thicker red section 50% credible interval, thinner red line 90% credible interval).
+Blue distributions in background show range of MPA effects produced by simulation model tuned to reflect the dynamics of the Channel Island MPAs (black dashed line is median simulated value).
+Results are estimated in blocks of three years, including years greater than or equal to left-hand value and less than right-hand value."}
+
+mpa_effect_plot
+
+ggsave(
+  mpa_effect_plot,
+  filename = here::here("documents","figs","mpa_effect_plot.pdf"),
+  width = width_2col,
+  height = width_2col * asp_2col,
+  device = cairo_pdf
+)
+
+```
+
+```{r raw-density, fig.width = width_1col, fig.asp=asp_1col,fig.cap = "Centered and scaled trends in biomass densities of targeted and non-targeted finfish included in our study Top panel shows trends across all sites, with smaller background lines showing trends for each individual species.
+Bottom two panels show aggregate biomass density trends outside and inside MPAs"}
+
+temp_plot <-
+  (total_trend_plot / (mpa_trend_plot)) & theme(legend.position = "top", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), units = "lines"),
+                                                panel.grid = element_blank()
+  )
+
+temp_plot
+
+ggsave(
+  temp_plot,
+  filename = here::here("documents","figs","raw_biomass_density_plot.pdf"),
+  width = width_1col,
+  height = width_1col * asp_1col,
+  device = cairo_pdf
+)
+
+```
+
+```{r val-plot, include = TRUE,fig.width = width_1col, fig.asp=asp_1col, fig.cap = "Distribution of percent error in posterior estimates of population-level MPA effect (y-axis) plotted against true simulated MPA effect (x-axis).
+Color shows concentration of simulations.
+Black line shows mean absolute percent error (MAPE) as a function of true simulated MPA effect."}
+
+temp_plot <- validation_plot + 
+  theme_1col + 
+  theme(legend.position = "top") + 
+  scale_fill_viridis(option = "C",
+                     name = "Sims(#)",
+                     guide = hgc <- guide_colorbar(frame.colour = "black",
+                                                   ticks.colour = "black",
+                                                   barwidth = unit(5,"lines")))
+
+temp_plot
+ggsave(
+  temp_plot,
+  filename = here::here("documents","figs","validation_plot.pdf"),
+  width = width_1col,
+  height = width_1col * asp_1col,
+  device = cairo_pdf
+)
+
+```
+
+```{r sim-effects, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Simulated population-level (pop.) MPA effect sizes as a function of percent of species' range inside MPA (x-axis), and pre-MPA depletion (y-axis).
+Pre-MPA depletion is a measure of fishing pressure, where 0 means that the population is unfished, and 1 means that the population is extinct in the time period immediately prior to MPA implementation.
+Panel A) shows median MPA effects across range in MPA and pre-MPA depletion Panel B) shows distribution of simulations across range of MPA size and pre-MPA depletion separately."}
+
+expected_mpa_effect_plot
+ggsave(
+  expected_mpa_effect_plot,
+  filename = here::here("documents","figs","expected_mpa_effect.pdf"),
+  width = width_2col,
+  height = width_2col * asp_2col,
+  device = cairo_pdf
+
+)
+
+```
+
+```{r ci-map, include = TRUE,fig.width=width_1col, fig.asp=1.5, fig.cap="Map of study region; the Northern Channel Islands, California, USA.
+Colors show binned number of PISCO sampling events across the time period of our study.", include=TRUE}
+
+temp_plot <- sample_location_plot + 
+  theme(axis.text = element_blank())
+
+temp_plot
+ggsave(
+  temp_plot,
+  filename = here::here("documents","figs","sample_location_plot.pdf"),
+  width = width_1col,
+  height = width_1col * asp_1col,
+  device = cairo_pdf
+)
+
+```
+
+# References

### 2020-04-28 make-pop-effects-of-mpas.R

@@ -56,7 +56,7 @@ run_description <- "PNAS R and R with simplified DiD and weighting"
 # So, once you've run simulate_mpas, you can set it to FALSE and validata_mpas will work
 
 
-run_did <- FALSE # run difference in difference on data from the CINMS
+run_did <- TRUE # run difference in difference on data from the CINMS
 
 run_tmb <- FALSE
 
@@ -753,7 +753,7 @@ if (file.exists(here(data_dir,'pdo.csv'))) {
 # convert transect data to density estimates ------------------------------
 
 
-if (file.exists(here("data","pisco-data.Rdata")) == F |
+if (file.exists(here("data","processed-pisco-data.rds")) == F |
     run_length_to_density == T) {
 
 
@@ -867,17 +867,13 @@ if (file.exists(here("data","pisco-data.Rdata")) == F |
     stop('multiple species per classcode')
   }
 
-  save(file = here("data",'pisco-data.Rdata'),
-       pisco_data)
-
+  write_rds(pisco_data,path = file.path(run_dir,"processed-pisco-data.rds"))
+  
 } else {
-  load(here("data",'pisco-data.Rdata'))
+  
+  pisco_data <- read_rds(file.path(run_dir,"processed-pisco-data.rds"))
 
 }
-
-wtf <- pisco_data %>% 
-  filter(classcode == "gnig")
-
 
 # sum biomass across all observed sizes ---------------------------------
 
@@ -1377,7 +1373,7 @@ model_runs <- model_runs %>%
 #   mutate(did_fit = map(did_fit, "result"))
 
   browser()
-did_fits <- model_runs %>% 
+did_fits <- model_runs %>%
   select(-data)
 
 plot_did <- function(x){

### 2020-05-08 documents/ovando-et-al-pop-effects-of-mpas.Rmd
making R v4.0.0 compliant

@@
                                                    ticks.colour = "black",
-                                                   barwidth = unit(5,"lines")))
+                                                   barwidth = ggplot2::unit(5,"lines")))
 

### 2020-05-08 make-pop-effects-of-mpas.R

@@ -33,7 +33,7 @@ library(tidyverse)
 library(ggtext)
 library(ggrepel)
 library(rnaturalearth)
-library(rnaturalearthhires)
+# library(rnaturalearthhires)
 extrafont::loadfonts()
 # extrafont::font_import()
 rstan_options(auto_write = TRUE)
@@ -56,7 +56,7 @@ run_description <- "PNAS R and R with simplified DiD and weighting"
 # So, once you've run simulate_mpas, you can set it to FALSE and validata_mpas will work
 
 
-run_did <- TRUE # run difference in difference on data from the CINMS
+run_did <- FALSE # run difference in difference on data from the CINMS
 
 run_tmb <- FALSE
 
@@ -100,9 +100,9 @@ theme_set(plot_theme)
 
 testplot <- ggplot()
 
-panel_height = unit(1,"npc") - sum(ggplotGrob(testplot)[["heights"]][-3]) - unit(4,"line")
+panel_height = grid::unit(1,"npc") - sum(ggplotGrob(testplot)[["heights"]][-3]) - unit(4,"line")
 
-panel_width = unit(1,"npc") - sum(ggplotGrob(testplot)[["widths"]][-3]) - unit(5,"line")
+panel_width = grid::unit(1,"npc") - sum(ggplotGrob(testplot)[["widths"]][-3]) - unit(5,"line")
 
 
 gc <- guide_colorbar(frame.colour = "black",
@@ -3010,7 +3010,7 @@ if (process_results == TRUE){
     scale_fill_binned(type = "viridis", labels = label_percent(accuracy = 1),
                       name = "Pop. Effect",
                       breaks = seq(0,2, by = .25),
-                      guide = guide_bins(keywidth = unit(1.25, "lines"),
+                      guide = guide_bins(keywidth = ggplot2::unit(1.25, "lines"),
                                          reverse = FALSE,
                                          axis.colour = "white",
                                          axis.linewidth = 1)
@@ -3042,7 +3042,7 @@ if (process_results == TRUE){
       labels =  label_percent(accuracy = 1),
       guide = guide_colorbar(frame.colour = "black",
                              ticks.colour = "black",
-                             barwidth = unit(8, "lines")),
+                             barwidth = grid::unit(8, "lines")),
       name = "Sims(%)"
     ) +
     scale_x_continuous(labels = percent, name = "Range in MPA", expand = expansion(0,0)) +
@@ -3734,7 +3734,7 @@ if (process_results == TRUE){
[…]

### 2020-06-24 documents/ovando-et-al-pop-effects-of-mpas.Rmd
starting conservation biology submission

@@
 bibliography: My Library.bib
-csl: pnas.csl
+csl: conservation-biology.csl
 output:
@@
 
+ did_fit <- model_runs$did_fit[model_runs$data_source == "pisco" &
+                                 model_runs$var_names == "pisco_a" &
+                                 model_runs$data_to_use == "all"][[1]]$did_reg
+
  if (!dir.exists(here("documents", "figs"))){
@@
 
+## Introduction
+
 No-take Marine Protected Areas (MPAs), spatial regions of the ocean in which fishing is prohibited, have a long history in the management of marine resources.
@@
 
-We use data from the first 14 years of protection to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of finfish species.
-In contrast to clear differences in biomass densities observed inside and outside of well protected MPAs both globally [@lester2009] and in the Channel Islands [@caselle2015] we are unable to detect a clear population effect from the Channel Islands MPAs.
-We build off of existing MPA theory to consider why this might be, and provide guidance for scientists and managers as to when and how we might expect to estimate the population-level conservation effects of MPAs.
-
-## Results & Discussion
-
-### What Are the Population-Level Effects of MPAs? {.unnumbered}
-
 The empirical MPA literature has generally focused on assessing conservation effects within the borders of protected areas [e.g.
@@
 
-Given this existing body of theory, what should we expect *a priori* the effect size of the Channel Islands MPAs to be?
+We use data from the first 14 years of protection to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of finfish species.
+In contrast to clear differences in biomass densities observed inside and outside of well protected MPAs both globally [@lester2009] and in the Channel Islands [@caselle2015] we are unable to detect a clear population effect from the Channel Islands MPAs.
+We build off of existing MPA theory to consider why this might be, and provide guidance for scientists and managers as to when and how we might expect to estimate the population-level conservation effects of MPAs.
+
+## Methods
+
+Our methods consist of a spatially-explicit bio-economic simulation model and a Bayesian difference-in-difference regression.
+We use the simulation model to provide provide theoretical grounding and context for our empirical results.
+The difference-in-difference regression provides our estimate of the population-level conservation effect of the Channel Islands MPAs.
+All analysis were conducted in R [@rcoreteam2019].
+Our main difference-in-difference model was fit using Stan [@carpenter2017] using the `rstanarm` package [@goodrich2020].
+All data and code needed to fully reproduce our results are publicly available at [github.com/DanOvando/population-effects-of-mpas](https://github.com/DanOvando/population-effects-of-mpas).
+
+### Simulation Model
+
+Our bio-economic model simulates the effect of MPAs on a spatially explicit age-structured representation of a single species.Readers can explore the functionality of the model using an online tool available at [danovando.shinyapps.io/simmpa/](https://danovando.shinyapps.io/simmpa/).
+Detailed methods can be found in the online appendix.
+
+The simulation model consists of 50 patches with wrapped edges.
+For any one simulation we randomly pull a species and its associated life history from the `FishLife` [@thorson2017c] package.
+We pair these data with randomly selected values governing the characteristics of the simulation (Table.1).
+These include attributes of recruitment (variation, autocorrelation, and timing), movement (magnitude of larval and adult dispersal and density dependent movement), fleet dynamics (degree of fishing pressure, amount and spatial allocation of fishing effort), and MPA design (size and placement, habitat characteristics, larval source).
+
+The population begins at unfished equilibrium and then fishing effort is applied in accordance to the fleet model.
+The MPAs are then placed, and each simulation is run to equilibrium with and without the selected MPA strategy (holding all else constant).
+We then measure the difference in biomass densities in each time step in the scenario with and without the MPAs to calculate the population effect of the MPAs over time.
+
+We created one set of simulations specifically designed to reflect the dynamics of the Channel Islands.
+We only include species of the same genus as those targeted by fishing in the Channel Islands.
+The MPAs in this study cover 20% of the surface waters of the Channel Islands, a region spanning roughly 90km east to west.
+Detailed dispersal studies are not available for all of the species covered by our study, but what information we have suggest that while adults of some of these species are likely to exhibit site fidelity [@love2011], larvae are likely distributed beyond the Channel Islands [@rassweiler2012].
+Therefore, if we assume that the population of the targeted species is at least equal to the extent of the Channel Islands, we can assume that at most the MPAs cover roughly 20% of the targeted populations' ranges.
+As such we cap the MPA size in the Channel Islands specific simulation batch to 20%.
+
+### Difference in Difference Regression
+
+ The difference-in-difference model used empirical kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands.
+A network of MPAs covering approximately 20% of the islands’ waters was put in place in 2003 as part of the California Marine Life Protection Act (MLPA) (see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of the MLPA).
+PISCO conducts visual SCUBA surveys at a large number of rocky-reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\ref{fig:ci-map}).
+The details of the monitoring program are described in @caselle2015.
+
+Building off of the concepts explored in @caselle2015, we used an identification strategy utilizing biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 0])` species that are not directly targeted by fishing before and after MPA implementation as our control group (non-targeted), and biomass densities before and after MPA implementation of `r n_distinct(used_data$classcode[used_data$targeted == 1])` species targeted by fishing as our treatment group.
+Targeted species in the Channel Islands, include commercially important fin-fish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
+Each of these targeted species was the subject of prior bio-economic modeling related to the effects of MPAs in southern California [@costello2010a;@rassweiler2014].
+Our analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*) (but our methods could similarly be applied to these species given data).
+Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species).
+We used a Bayesian difference-in-difference regression to estimate any difference in mean total biomass densities of fin-fish species targeted by fishing effort (i.e., those potentially affected by an MPA) and those species not targeted by fishing before and after MPA implementation [see @larsen2019 for an explanation of difference-in-difference estimators in ecological settings].
+The identification strategy control for unobserved environmental shocks to the system independent of the MPAs and similarly experienced by the targeted and non-targeted species.
+Conditional on the assumptions of the model, this regression produces an estimate of the effect of the MPAs on the mean total biomass densities of targeted species throughout the Channel Islands.
+
+The regression amounts to estimating the pre-post MPA difference in the mean total biomass density of targeted species minus the same difference for non-targeted species in the Channel Islands.
+
+The simplified form of this model is
+
+\begin{equation}
+  d_{i} \sim Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + \mathbf{B^cX_i} + \mathbf{B^sS_i}},shape, scale)
+\label{eq:did}
+\end{equation}
+
+where $d_i$ is the biomass density at observation *i*, *T* indicates whether the observation *i* is for a targeted ($T = 1$) or non-targeted ($T = 0$) species, and *MPA* marks whether observation *i* is in a pre MPA ($MPA = 0$) or post MPA ($MPA = 1$) state.
+**B^c^** is a vector of coefficients for additional control variables in matrix *X* such as water visibility and observer experience.
+**B^s^** is a vector of hierarchical coefficients for each sampling location **S** , clustered by island.
+Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment (*MPA*) on the treated (targeted species).
+$\beta_0$ is the mean total biomass of non-targeted species pre-MPA, and $\beta_1T$ is the mean total biomass of targeted species pre-MPA.
+$\beta_2MPA$ is the effect of the post-MPA period on the non-targeted species.
+The shape and scale parameters of the Gamma distribution are estimated as well.
+
+`r percent(mean(used_data$eventual_mpa))` of the samples in the raw data come from within MPAs, while the MPAs themselves cover 20% of the surface waters of the Channel Islands.
+As such, we weight each observation such that within MPA data are assigned a total weight of 0.2 and outside-MPA data a total weight of 0.8.
+The intercept prior was was determined using the `rstanarm` `autoscale` function.
+For the non-intercept terms, we manually set a prior as a normal distribution with a mean of 0 and a standard deviation of 2.
+This implies that coefficients such as the MPA effect have a prior that provides support for an effect size centered on zero with a range of roughly -4 to 4.
+Since the covariates are centered and scaled, and the dependent variable is on the log-scale, this is an extremely diffuse prior (e.g.
+a one standard deviation change in an independent variable changes the dependent variable by 4 on the log scale).
+The full table of covariates and their posterior means and 89% credible interval can be found in Table.\@ref(tab:did-coefs).
+Diagnostic plots can be found in the online appendix, along with sensitivity analyses of various model structures estimation strategies.
+
+## Results
+
+### What Are the Population-Level Effects of MPAs? {.unnumbered}
+
+What should we expect *a priori* the population-level effect size of the Channel Islands MPAs to be?
 While we know many parameters will affect the expected effects of the MPAs, theory suggests that life history, adult and larval dispersal relative to MPA size, and fishing pressure will be key drivers.
@@
 
+*already said this*
 The MPAs in this study cover 20% of the surface waters of the Channel Islands, a region spanning roughly 90km east to west.
@@
 
+*reframe in light of new methods*
 We integrated these network design features, life history traits, and exploitation histories into a spatially explicit bio-economic simulation model, and used this model to generate expected outcomes for the Channel Islands MPAs, varying key unknowns such as the degree of pre-MPA depletion, species' movement rates, and fleet dynamics.
@@
 Having established a theoretical grounding for the likely biomass density outcomes of the Channel Island MPAs, what empirical evidence do we have of the effects?
-@caselle2015 examined changes in biomass densities of species targeted by fishing activity inside and outside of the MPAs over time (a metric generally termed a "response ratio"), and compared these changes to the trends in non-targeted species.
+@caselle2015 examined changes in response ratios of species targeted by fishing activity inside and outside of the MPAs over time and compared these changes to the trends in non-targeted species.
 They found a statistically significant increase in the response ratios of targeted species over time, and evidence that this increase is smaller in the non-targeted species.
-Updating the results of @caselle2015 through 2017 with a Bayesian response-ratio regression shows a continuation of the increasing trend in the response ratios of targeted species (Fig.\@ref(fig:rr-plot)).
+Updating the results of @caselle2015 through 2017 shows a continuation of the increasing trend in the response ratios of targeted species (Fig.\@ref(fig:rr-plot)).
 This provides evidence that the Channel Islands MPAs are large enough and sufficiently well-enforced as to provide meaningful protection within their borders.
 
-Unfortunately, these response ratios do not provide sufficient evidence for significant population-level effects of the Channel Islands MPAs.
+These response ratios alone do not provide sufficient evidence for significant population-level effects of the Channel Islands MPAs.
 Control sites used in calculating response ratios are often selected based on abiotic or ecological traits such as habitat characteristics [@ferraro2018].
@@
 
-To illustrate this problem, we simulated response ratio trajectories for species and MPA coverage representative of the Channel Islands (see methods and Supporting Information (SI) Appendix for details of model structure).
-These simulations cover the life-histories and MPA sizes seen in the Channel Islands, but vary in key unknowns such as the actual degree of fishing pressure, the timing of density dependence, and the fishing fleets response to the MPAs.
-For each year of protection, we paired our simulated response ratios to the estimated posterior probability distributions of the  empirical Channel Island response ratios.
+To illustrate this problem.
+For each year of protection, we paired our simulated Channel Islands response ratios to the estimated posterior probability distributions of the empirical Channel Island response ratios.
 For example, if the mean estimated response ratio in the year 2006 (year three after implementation) is one, we found all simulations that had simulated response ratios near one after three years of protection, and then pulled out the "true" simulated population-level MPA effect from each of those simulations.
@@
 Careful causal inference requires some other means of controlling for biases introduced by factors such as the unobserved environmental shocks, the MPA siting process, biological spillover, and concentration of fishing effort  [@ferraro2018;@larsen2019].
-
-Building off of the concepts explored in @caselle2015, we used an identification strategy utilizing biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 0])` species that are not directly targeted by fishing before and after MPA implementation as our control group (non-targeted), and biomass densities before and after MPA implementation of `r n_distinct(used_data$classcode[used_data$targeted == 1])` species targeted by fishing as our treatment group.
-Targeted species in the Channel Islands, include commercially important fin-fish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
-Each of these targeted species was the subject of prior bio-economic modeling related to the effects of MPAs in southern California [@costello2010a;@rassweiler2014].
-Our analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*) (but our methods could similarly be applied to these species given data).
-Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species).
-We used a Bayesian difference-in-difference regression to estimate any difference in mean total biomass densities of fin-fish species targeted by fishing effort (i.e., those potentially affected by an MPA) and those species not targeted by fishing before and after MPA implementation [see @larsen2019 for an explanation of difference-in-difference estimators in ecological settings].
-The identification strategy control for unobserved environmental shocks to the system independent of the MPAs and similarly experienced by the targeted and non-targeted species.
-Conditional on the assumptions of the model, this regression produces an estimate of the effect of the MPAs on the mean total biomass densities of targeted species throughout the Channel Islands.
+**bring this in earlier**
 
@@
 -->
+
+## Discussion {#discussion .unnumbered}
 
@@
 
-## Conclusions {#conclusion .unnumbered}
+<!-- The key assumptions of the difference-in-difference model are that a) within the time-frame of the model there are no significant interaction effects between the targeted and non-targeted species (which in fact we do not detect, see SI Appendix), and b) that in the absence of the MPAs both the targeted and non-targeted groups of species would have exhibited similar trends in densities.
+-->
+
+<!-- We briefly assess two of the most critical assumptions of this model here: that the treated and non-treated groups have parallel trends, and that the effect of the treatment on the treated does not tangentially affect the untreated.
+While the parallel trends assumption cannot be formally proven, we can examine its validity using the data from the years before the MPAs were put in place in 2003.
+We do not detect any significant differences in the trends of the biomass densities of the targeted and non-targeted species in the years before the MPAs (SI Appendix Fig.S26).
+However, we believe that the parallel trends assumption is likely violated in the post 2014 data.
+-->
+
+<!-- With regards to the second assumption, all of the species in this empirical analysis exist within an ecosystem, and as such affect each other through mechanisms such as predation, competition, and habitat modification.
+We find it unlikely that these effects have had enough time to manifest in a meaningful way in the 14 years of post-MPA data used in our analysis [@babcock2010; @pershing2015a].
+We used convergent cross mapping (CCM), in the manner of @clark2015, to test for the possibility of the trophic cascades biasing our results.
+Our results found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (SI Appendix Fig.S42).
+However, the longer MPAs are in place, the greater the possibility that substantial species interactions that can affect use of non-targeted species as a control may arise.
+-->
 
@@
 
-## Materials and Methods
+# Supporting Information
 
-We present here critical characteristics of our simulation model and regression approach.
-Further details can be found in the Supplementary Information.
-All analysis were conducted in R [@rcoreteam2019].
-Our main difference-in-difference model was fit using Stan [@carpenter2017] using the `rstanarm` package [@goodrich2020].
-All data and code needed to fully reproduce our results are publicly available at [github.com/DanOvando/population-effects-of-mpas](https://github.com/DanOvando/population-effects-of-mpas).
+A detailed description of the simulation model structure, as well as sensitivity analyses of our estimation model, are available in the online appendix.
 
-### Simulation Model
+# Tables
 
-Our bio-economic model simulates the effect of MPAs on a spatially explicit age-structured representation of a single species.
-Readers can explore the functionality of the model using an online tool available at [danovando.shinyapps.io/simmpa/](https://danovando.shinyapps.io/simmpa/).
-Full details of the simulation model are available in the Supplementary Information.
+I need to clean these up
 
-### Difference in Difference Regression
+Table.2 - Range of simulated variables
 
- The difference-in-difference model used empirical kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands.
-A network of MPAs covering approximately 20% of the islands’ waters was put in place in 2003 as part of the California Marine Life Protection Act (MLPA) (see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of the MLPA).
-PISCO conducts visual SCUBA surveys at a large number of rocky-reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\ref{fig:ci-map}).
-The details of the monitoring program are described in @caselle2015.
+| Variable | Distribution|
+|---------:|:-----------:|
+| Scientific Name | Drawn from all possible species in `FishLife` (@thorson2017c)|
+|steepness (h) | ~uniform(0.6,0.95)|
+|Adult movement ($\sigma_{s=a}$) | ~uniform(0,0.25 * P)|
+|Larval movement ($\sigma_{s=l}$) | ~uniform(0,0.25 * P)|
+|Recruitment variation ($\sigma_{r}$) |  $\in\{0,0.05,.1,.2\}$|
+|Recruitment autocorrelation ($ac_{r}$) |  $\in\{0,0.05,.1,.2\}$|
+|DD adult movement (dmod) | $\in\{0.25,1\}$|
+|Density-dependence timing | $\in\{local, global,post-dispersal\}$|
+|\% Patches in MPA | ~uniform(0.01,1)|
+|Initial fishing relative to natural mortality | ~uniform(0.01,4)|
+|Selectivity as a multiple of maturity length| ~uniform(0.1,1.25)|
+|Fleet model| $\in\{open-access, constant-effort,constant-catch\}$|
+|Spatial effort model| $\in\{uniform, biomass,profits\}$|
+|Years into simulation to start MPA| ~round(uniform(5,0.66 * T))|
+|MPA is sprinkler?| $\in\{TRUE,FALSE\}$|
+|Randomly place MPA?| $\in\{TRUE,FALSE\}$|
+|Fleet reaction to MPA| $\in\{concentrate, abandon-ship\}$|
+|Patchiness | ~uniform(0.01,0.75)|
+|MPA habitat factor| $\in\{1, 4\}$|
 
-The key assumptions of the difference-in-difference model are that a) within the time-frame of the model there are no significant interaction effects between the targeted and non-targeted species (which in fact we do not detect, see SI Appendix), and b) that in the absence of the MPAs both the targeted and non-targeted groups of species would have exhibited similar trends in densities.
-It is important to note that the model is agnostic as to the root causes of the parallel trends, so long as they hold.
+```{r did-coefs, results = "asis",eval = TRUE}
+did_fit %>%
+  broom::tidy(intervals = TRUE, prob = 0.89) %>%
+  select(term, estimate, lower, upper) %>%
+  rename(
+    mean = estimate,
+    lower_89th_ci = lower,
+    upper_89th_ci = upper
+  ) %>%
+  knitr::kable(
+    digits = 2,
+    caption = "Posterior means and credible interval for key model coefficients",
+    longtable = TRUE,
+    booktabs = TRUE ) %>%
+  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
 
-The regression amounts to estimating the pre-post MPA difference in the biomass densities of targeted species minus the same difference for non-targeted species in the Channel Islands.
-
-The simplified form of this model is
-
-\begin{equation}
-  d_{i} \sim Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + \mathbf{B^cX_i} + \mathbf{B^sS_i}},shape, scale)
-\label{eq:did}
-\end{equation}
-
-where $d_i$ is the biomass density at observation *i*, *T* indicates whether the observation *i* is for a targeted ($T = 1$) or non-targeted ($T = 0$) species, and *MPA* marks whether observation *i* is in a pre MPA ($MPA = 0$) or post MPA ($MPA = 1$) state.
-**B^c^** is a vector of coefficients for additional control variables in matrix *X* such as water visibility and observer experience.
-**B^s^** is a vector of hierarchical coefficients for each sampling location **S** , clustered by island.
-Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment (*MPA*) on the treated (targeted species).
-$\beta_0$ is the mean total biomass of non-targeted species pre-MPA, and $\beta_1T$ is the mean total biomass of targeted species pre-MPA.
-$\beta_2MPA$ is the effect of the post-MPA period on the non-targeted species.
-The shape and scale parameters of the Gamma distribution are estimated as well.
-`r percent(mean(used_data$eventual_mpa))` of the samples in the raw data come from within MPAs, while the MPAs themselves cover 20% of the surface waters of the Channel Islands.
-As such, we weight each observation such that within MPA data are assigned a total weight of 0.2 and outside-MPA data a total weight of 0.8.
-See SI Appendix for sensitivity analyses of various model structures and synthetic controls results.
-
-We briefly assess two of the most critical assumptions of this model here: that the treated and non-treated groups have parallel trends, and that the effect of the treatment on the treated does not tangentially affect the untreated.
-While the parallel trends assumption cannot be formally proven, we can examine its validity using the data from the years before the MPAs were put in place in 2003.
-We do not detect any significant differences in the trends of the biomass densities of the targeted and non-targeted species in the years before the MPAs (SI Appendix Fig.S26).
-However, we believe that the parallel trends assumption is likely violated in the post 2014 data.
-
-With regards to the second assumption, all of the species in this empirical analysis exist within an ecosystem, and as such affect each other through mechanisms such as predation, competition, and habitat modification.
-We find it unlikely that these effects have had enough time to manifest in a meaningful way in the 14 years of post-MPA data used in our analysis [@babcock2010; @pershing2015a].
-We used convergent cross mapping (CCM), in the manner of @clark2015, to test for the possibility of the trophic cascades biasing our results.
-Our results found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (SI Appendix Fig.S42).
-However, the longer MPAs are in place, the greater the possibility that substantial species interactions that can affect use of non-targeted species as a control may arise.
+```
 

### 2020-06-24 make-pop-effects-of-mpas.R

@@ -62,7 +62,7 @@ run_tmb <- FALSE
 
 simulate_mpas <- FALSE # simulate MPA outcomes
 
-simulate_channel_islands <- FALSE # simulate MPA outcomes
+simulate_channel_islands <- TRUE # simulate MPA outcomes
 
 validate_mpas <- FALSE
 
@@ -76,7 +76,7 @@ sim_years <- 50
 
 num_patches <- 50
 
-n_cores <- 1
+n_cores <- 6
 
 # prepare run -------------------------------------------------------------
 
@@ -1813,9 +1813,9 @@ if (simulate_channel_islands == TRUE){
   
   num_patches <- 50
   
-  run_experiments <- FALSE
+  run_experiments <- TRUE
   
-  create_grid <- FALSE
+  create_grid <- TRUE
   
   save_experiment <- TRUE
   
@@ -1883,7 +1883,7 @@ if (simulate_channel_islands == TRUE){
             larval_movement = sample(0:round(.25 * num_patches), samps, replace = T),
             density_movement_modifier = sample(c(0.25, 1), samps, replace = T),
             density_dependence_form = sample(1:3, samps, replace = T),
-            mpa_size = runif(samps, min = .05, max = .4),
+            mpa_size = runif(samps, min = .05, max = .2),
             f_v_m = runif(samps, min = 0.5, max = 1.5),
             fleet_model = sample(
               c("open-access", "constant-effort"),

### 2020-06-24 documents/ovando-et-al-pop-effects-of-mpas.Rmd
new structure complete, starting the cutting process. currently at 9260 words counting abstract and lit. cited, which count for conservation letters

@@
   filter(set == "ci",
-         final_depletion > 0,
-         f_v_m > .5) %>% 
+         final_depletion > 0) %>% 
   mutate(year = years_protected + 2002) %>% 
@@
 Recent international agreements to expand MPAs (most proximately, the Convention on Biological Diversity’s Strategic Plan for Biodiversity, which calls for 10% of coastal waters to be protected inside MPAs by 2020 and the International Union for Conservation of Nature call for 30% by 2030) are based on the expectation that well-designed MPAs will achieve benefits both within, and outside, their borders.
-Numerous studies provide evidence that well-enforced and appropriately sized MPAs can produce conservation benefits within their borders [@lester2009;@halpern2003; @edgar2014; @gerber2005].
-However, our collective scientific understanding of the regional-scale impacts of current and future MPAs on fish populations and fisheries is surprisingly limited.
-
-As conservation benefits accrue inside MPAs, theory holds that MPAs can affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
+The empirical MPA literature has generally focused on assessing conservation effects within the borders of protected areas [@lester2009;@halpern2003; @edgar2014; @gerber2005].As conservation benefits accrue inside MPAs, MPAs can affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
 These dynamics link fish both inside and outside of MPAs as part of a shared biological stock; the population.
@@
 @goni2010; @halpern2009;@kay2012; @stobart2009; @mcclanahan2000; @russ1996; @thompson2017;@baetscher2019;@pelc2009], as well as alteration of fishing effort in reaction to [@costa2013;@mason2012;@murawski2005] and in anticipation of [@mcdermott2019] MPA placement.
-However, empirically detecting the net effects of spillover on entire fish populations and fisheries is much more difficult.
+The assumption that MPAs will have conservation benefits beyond their borders for most species is implicit in all multilateral calls for MPA expansion, but our collective scientific understanding of the regional-scale impacts of current and future MPAs on fish populations and fisheries is surprisingly limited.
+
+Numerous factors can affect the population effects of MPAs.
+These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], the strength and timing of density dependence in the population (e.g.
+pre- or post-settlement, @burgess2014), the age and degree of enforcement [@edgar2014],  how overfished the population would be without the MPA, and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006a;@walters2000a;@white2011,;@white2010c].
+In addition, even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through network effects [@gaines2010; @costello2010a].
+
+Empirical evidence of the conservation outcomes of MPAs is made up almost entirely by versions of "response ratios", commonly measured as biomass densities of species inside MPAs relative to biomass densities of species in areas outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015].
+These studies have by and large found clear evidence that well enforced and sufficiently sized MPAs are associated with high response ratios.
+Why does this large body of empirical work not serve as sufficient evidence for the population-level effects of MPAs?
+
+Control sites used in calculating response ratios are often selected based on abiotic or ecological traits such as habitat characteristics [@ferraro2018].
+Selection of these sites is complicated, though, by the very spillover that MPAs are often intended to create.
+Export of adults or larvae from the MPA to the "control" site affects their status as controls, as does displacement of fishing effort from MPAs to control sites.
+In theory, control sites far enough away to negate both biological spillover and concentration by the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs may be challenging.
+As a result of these challenges, response ratios can be a highly imprecise and biased measure of the population-level conservation effect of an MPA network [@ferraro2018].
+
+Given the potential unreliability of response ratios for the task, how can we empirically measure the population-level effects of MPAs?
+The perfect experiment would involve two parallel worlds that were identical, except for the implementation of an MPA.
+In world “A”, no MPA would be implemented, and in the facsimile world “B”, the MPA would be implemented.
+Both worlds would be tracked before and after MPA implementation, and the biomass densities would be compared after treatment.
+Instead of two parallel worlds, a similar experimental design would involve random placement of MPAs.
+Unfortunately, neither of these experimental designs has been implemented, because parallel worlds do not exist and MPAs are to our knowledge never placed at random.
+Careful causal inference requires some other means of controlling for biases introduced by factors such as the unobserved environmental shocks, the MPA siting process, biological spillover, and concentration of fishing effort  [@ferraro2018;@larsen2019].
+
+As stakeholders around the world seek to use MPAs in marine resource management portfolios, it is critical that we develop a better understanding of the magnitude and drivers of population-level MPA effects.
+This is particularly true in the context of a changing climate, in which "set it and forget it" management strategies are increasingly untenable; we need to be able to monitor the performance of MPAs in order to adapt them as dictated by a shifting environment.
 From a fishery perspective, are spillover benefits sufficient to offset losses in fishing grounds and changes in responses of displaced fishers caused by an MPA?
@@
 
-As stakeholders around the world seek to use MPAs in marine resource management portfolios, it is critical that we develop a better understanding of the magnitude and drivers of population-level MPA effects.
-This is particularly true in the context of a changing climate, in which "set it and forget it" management strategies are increasingly untenable; we need to be able to monitor the performance of MPAs in order to adapt them as dictated by a shifting environment.
 The MPAs within the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) provide an ideal case study to address this need.
@@
 This MPA network has subsequently been used as a model case study in protected area design around the world [@botsford2014 and references therein].
-
-The empirical MPA literature has generally focused on assessing conservation effects within the borders of protected areas [e.g.
-@edgar2014].
-While these within-MPA effects are vitally important for protecting rare species, biodiversity, critical habitats, and often tourism, they paint an incomplete picture of the overall population effects of MPAs.
-The organisms within the borders of protected areas are generally part of a broader biological stock, connected through adult movement and larval dispersal.
-The assumption that MPAs will have conservation benefits beyond their borders for most species is implicit in all multilateral calls for MPA expansion, but these benefits have not been widely measured.
-
-We define the population-level conservation effects of MPAs as the change in mean total biomass densities of targeted finfish both inside and outside of MPAs, relative to the mean total biomass densities of targeted finfish inside and outside of MPAs that would have occurred without the MPAs.
-As an example of a population-level MPA effect, consider a evenly distributed population that has 50% of its range protected by an MPA.
-Suppose that the MPA increase biomass densities inside the reserves by 20%, and by 0% outside the reserve.
-By our metric the population conservation effect of the MPA would be 10%.
-
-Numerous factors can affect the population effects of MPAs.
-These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], larval dispersal patterns and the strength and timing of density dependence in the population (e.g.
-pre- or post-settlement, @burgess2014), the age and degree of enforcement [@edgar2014],  how overfished the population would be without the MPA, and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006a;@walters2000a;@white2011,;@white2010c].
-In addition, even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through network effects [@gaines2010; @costello2010a].
-
 We use data from the first 14 years of protection to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of finfish species.
@@
 
-Our methods consist of a spatially-explicit bio-economic simulation model and a Bayesian difference-in-difference regression.
+Our primary methods consist of a spatially-explicit bio-economic simulation model and a Bayesian difference-in-difference regression.
 We use the simulation model to provide provide theoretical grounding and context for our empirical results.
@@
 All analysis were conducted in R [@rcoreteam2019].
-Our main difference-in-difference model was fit using Stan [@carpenter2017] using the `rstanarm` package [@goodrich2020].
+Our difference-in-difference model was fit using Stan [@carpenter2017] using the `rstanarm` package [@goodrich2020].
 All data and code needed to fully reproduce our results are publicly available at [github.com/DanOvando/population-effects-of-mpas](https://github.com/DanOvando/population-effects-of-mpas).
@@
 
-Our bio-economic model simulates the effect of MPAs on a spatially explicit age-structured representation of a single species.Readers can explore the functionality of the model using an online tool available at [danovando.shinyapps.io/simmpa/](https://danovando.shinyapps.io/simmpa/).
+Our bio-economic model simulates the effect of MPAs on a spatially explicit age-structured representation of a fish population.
+Readers can explore the functionality of the model using an online tool available at [danovando.shinyapps.io/simmpa/](https://danovando.shinyapps.io/simmpa/).
 Detailed methods can be found in the online appendix.
@@
 These include attributes of recruitment (variation, autocorrelation, and timing), movement (magnitude of larval and adult dispersal and density dependent movement), fleet dynamics (degree of fishing pressure, amount and spatial allocation of fishing effort), and MPA design (size and placement, habitat characteristics, larval source).
-
 The population begins at unfished equilibrium and then fishing effort is applied in accordance to the fleet model.
@@
 We then measure the difference in biomass densities in each time step in the scenario with and without the MPAs to calculate the population effect of the MPAs over time.
+The results provide us a library of simulated MPA effects under a range of plausible biological and economic states of the world.
 
-We created one set of simulations specifically designed to reflect the dynamics of the Channel Islands.
+We also created one set of simulations specifically designed to reflect the dynamics of the Channel Islands.
 We only include species of the same genus as those targeted by fishing in the Channel Islands.
@@
  The difference-in-difference model used empirical kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands.
-A network of MPAs covering approximately 20% of the islands’ waters was put in place in 2003 as part of the California Marine Life Protection Act (MLPA) (see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of the MLPA).
 PISCO conducts visual SCUBA surveys at a large number of rocky-reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\ref{fig:ci-map}).
@@
 
+We define the population-level conservation effects of MPAs as the change in mean total biomass densities of targeted finfish both inside and outside of MPAs, relative to the mean total biomass densities of targeted finfish inside and outside of MPAs that would have occurred without the MPAs.
+As an example of a population-level MPA effect, consider a evenly distributed population that has 50% of its range protected by an MPA.
+Suppose that the MPA increase biomass densities inside the reserves by 20%, and by 0% outside the reserve.
+By our metric the population conservation effect of the MPA would be 10%, but a response ratio value of infinity.
 The regression amounts to estimating the pre-post MPA difference in the mean total biomass density of targeted species minus the same difference for non-targeted species in the Channel Islands.
@@
 
-### What Are the Population-Level Effects of MPAs? {.unnumbered}
+### Effects of the Channel Islands MPA Network
 
@@
 The targeted species in our database span a range of life histories, but are largely made up of fishes in the perch and rockfish complexes, with a mean Von Bertalanffy growth coefficient of `r round(mean(used_data$k[used_data$targeted == 1]),2)`, and a median age at maturity of `r round(median(used_data$tm[used_data$targeted == 1]),0)` years [values drawn from @thorson2017c].
-
-*already said this*
-The MPAs in this study cover 20% of the surface waters of the Channel Islands, a region spanning roughly 90km east to west.
-Detailed dispersal studies are not available for all of the species covered by our study, but what information we have suggest that while adults of some of these species are likely to exhibit site fidelity [@love2011], larvae are likely distributed beyond the Channel Islands [@rassweiler2012].
-Therefore, if we assume that the population of the targeted species is at least equal to the extent of the Channel Islands, we can assume that at most the MPAs cover roughly 20% of the targeted populations' ranges.
-
 Formal stock assessments are largely lacking for these species.
 However, what evidence we have suggests mostly moderate fishing pressure, with some species such as California sheephead (*Semicossyphus pulcher*) and blue rockfish (*Sebastes mystinus*) experiencing high levels of fishing mortality, likely resulting in biomass levels below management targets during the early 2000s [@alonzo2004;@dick2017].
-
-*reframe in light of new methods*
-We integrated these network design features, life history traits, and exploitation histories into a spatially explicit bio-economic simulation model, and used this model to generate expected outcomes for the Channel Islands MPAs, varying key unknowns such as the degree of pre-MPA depletion, species' movement rates, and fleet dynamics.
+We integrated these network design features, life history traits, and exploitation histories into our simulation model, varying key unknowns such as the degree of pre-MPA depletion, species' movement rates, and fleet dynamics.
 While a wide range of MPA effects are plausible, from 0% to upwards of 200%, the median simulated effect size for a stylized version of the Channel Islands after over ten years of protection was roughly `r percent(median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected > 10]))` (Fig.\@ref(fig:mpa-effect)).
-
-<!-- However, theory also tells us that many other variables can interact to affect the outcomes of MPAs.
-To address this, we utilized a spatial bio-econonomic model to simulate the expected effects of an MPA network stylized to resemble the Channel Islands on the group of targeted species covered in our analysis (See SI for details of simulation framework).
--->
-
-<!-- **reference HARKing, of course DiD and theory matching together doens't prove anything, but it's an important step to take** -->
-
-### Effects of the Channel Islands MPA Network
 
@@
 
-These response ratios alone do not provide sufficient evidence for significant population-level effects of the Channel Islands MPAs.
-Control sites used in calculating response ratios are often selected based on abiotic or ecological traits such as habitat characteristics [@ferraro2018].
-Selection of these sites is complicated, though, by the very spillover that MPAs are often intended to create.
-Export of adults or larvae from the MPA to the "control" site affects their status as controls, as does displacement of fishing effort from MPAs to control sites.
-In theory, control sites far enough away to negate both biological spillover and concentration by the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs may be challenging.
-As a result of these challenges, response ratios can be a highly imprecise and biased measure of the population-level conservation effect of an MPA network [@ferraro2018].
-
-<!-- While these concerns have been stated previously [e.g.
-@halpern2004;@white2011], the MPA evaluation literature has by and large been unable to adequately address them, often due to the very real logistical challenges of identifying and sampling adequate control sites that are truly independent of the MPAs in question -->
-
 In the case of the Channel Islands MPAs, control sites are often located within a few kilometers of an MPA, suggesting that they are susceptible to both local biological spillover and by concentration of fishing effort excluded from the MPAs.
@@
 
-To illustrate this problem.
-For each year of protection, we paired our simulated Channel Islands response ratios to the estimated posterior probability distributions of the empirical Channel Island response ratios.
+To illustrate this problem, we paired our simulated Channel Islands response ratios to the estimated posterior probability distributions of the empirical Channel Island response ratios.
 For example, if the mean estimated response ratio in the year 2006 (year three after implementation) is one, we found all simulations that had simulated response ratios near one after three years of protection, and then pulled out the "true" simulated population-level MPA effect from each of those simulations.
@@
 The response ratio trends we observe in the data could plausibly be produced by a wide range of population-level MPA effects (Fig.\@ref(fig:rr-plot)).
-Response ratios well over one were associated with population-level MPA effects generally less than 25%, and, importantly, many simulations produced large response ratios but population-level MPA effects close to 0% This can occur if for example fishing pressure is only moderate, adult movement is low, larval dispersal is high, and displaced fishing effort concentrates around the border of the MPAs.
-
-Given the potential unreliability of response ratios for the task, how can we empirically measure the population-level effects of MPAs?
-The perfect experiment would involve two parallel worlds that were identical, except for the implementation of an MPA.
-In world “A”, no MPA would be implemented, and in the facsimile world “B”, the MPA would be implemented.
-Both worlds would be tracked before and after MPA implementation, and the biomass densities would be compared after treatment.
-Instead of two parallel worlds, a similar experimental design would involve random placement of MPAs.
-Unfortunately, neither of these experimental designs has been implemented, because parallel worlds do not exist and MPAs are to our knowledge never placed at random.
-Careful causal inference requires some other means of controlling for biases introduced by factors such as the unobserved environmental shocks, the MPA siting process, biological spillover, and concentration of fishing effort  [@ferraro2018;@larsen2019].
-**bring this in earlier**
+Response ratios well over one were associated with population-level MPA effects generally less than 25%, and, importantly, many simulations produced large response ratios but population-level MPA effects close to 0%.
+This can occur if for example fishing pressure is only moderate, adult movement is low, larval dispersal is high, and displaced fishing effort concentrates around the border of the MPAs.
 
@@
 
+If response ratios are not clear evidence of the population level effect of the Channel Islands MPAs, what does our difference-in-difference strategy find?
 Consistent with MPA theory, over the first three years of implementation (2003-2006), the effects of the MPAs are unclear, with support for a small negative effect to a substantially positive effect, with much higher probability of a small positive effect (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`,
@@
 90\% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`
-- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`.) Fig.\@ref(fig:mpa-effect)
-
-<!-- 1. Why are the results from the channel islands important and what do they tell us -->
-
-<!-- 2. Put the channel islands results in context -  -->
-
-<!-- Ideally then we want a control for broad environmental shocks to the region that is independent of factors such as selection bias, biological spillover, and fishing concentration.  -->
-
-<!-- 1. multiple lines of evidence -->
-
-<!-- 2. Theory before empirics -->
-
-<!-- 3. Patience -->
-
-<!-- The network of MPAs implemented in the Northern Channel Islands in 2003 provides an ideal opportunity to examine the regional effects of protected area networks.
-This analysis is to our knowledge the first empirical estimate of the regional effects of a large and iconic MPA network on a broad assemblage of targeted fin-fish.
-Despite the well enforced nature of these MPAs, and the presence of a rigorous scientific monitoring program, we are unable to estimate a clear effect of protection.
-While response ratios of targeted species have increased over time, providing evidence of effective *within MPA* protection, these response ratios are on their own an unreliable indicator of the regional conservation effects of the MPAs.
-Our difference-in-difference regression estimates a high probability of increasingly positive MPA effects over the first nine years of protection, only to see these estimated gains reversed from 2012 to 2017.
-For the most recent time period, we can detect no clear positive or negative regional effect of the Channel Island MPAs on biomass densities of targeted fin-fish (Fig.\@ref(fig:mpa-effect)).
--->
+- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`.) Fig.\@ref(fig:mpa-effect).
 
@@
 
-How can we explain the lack of a clear population-level MPA effect after over ten years of protection?
+How can we explain the lack of a clear population-level MPA effect in the Channel Islands after over ten years of protection?
 One explanation may lie in fleet dynamics.
@@
 Constant-catch dynamics might also occur in fisheries with constraining quotas that are not updated after the implementation of MPAs.
-While open-access fishing strategies can result in "scorched earth" scenarios where the only fish left are found inside the reserve, across `r percent(positive_outcomes$pp[positive_outcomes$fleet_model == "open-access"], accuracy = 1)` simulations the net effect of the reserves was still positive.
-Under a constant-catch scenario though, fishers have to fish much harder than before to get the same catch from a smaller part of the population, reducing the size structure of the population and subsequently causing net conservation loss under `r percent(1 - positive_outcomes$pp[positive_outcomes$fleet_model == "constant-catch"], accuracy = 1)` of our constant-catch simulations.
+Under a constant-catch though, fishers have to fish much harder than before to get the same catch from a smaller part of the population, reducing the size structure of the population and subsequently causing net conservation loss under `r percent(1 - positive_outcomes$pp[positive_outcomes$fleet_model == "constant-catch"], accuracy = 1)` of our constant-catch simulations.
 This is an important and often overlooked possibility, especially as MPAs are increasingly implemented in quota-managed fisheries [@liu2018].
@@
 
-### When Can We Detect the Effects of MPAs?
-
 Containing a carefully designed, well-enforced, and well-studied MPA network, the Channel Islands would seem at face value to be an ideal location to study the population-level effects of protected areas.
@@
 We believe that the disappearance of increasing probabilities of positive MPA effects estimated by our model are likely driven by the marine heatwave experienced by the region beginning in 2014 (and potentially by exogenous increases in fishing mortality for key target species).
-This likely violation of the parallel trends assumption means that post 2014 estimates of MPA effects may be unreliable.
+This likely violation of the parallel trends assumption means that our post 2014 estimates of MPA effects may be unreliable.
 After 14 years of MPA protection and following a large environmental perturbation we are left without a clear picture of the effect of the Channel Island MPA network on biomass densities of targeted fin-fish species from either response ratios or the difference-in-difference model.
@@
 
+Could trophic cascades exlpain the lack of a clear population-level MPA effect in the Channel Islands?
+All of the species in this empirical analysis exist within an ecosystem, and as such affect each other through mechanisms such as predation, competition, and habitat modification.
+We find it unlikely that these effects have had enough time to manifest in a meaningful way in the 14 years of post-MPA data used in our analysis [@babcock2010; @pershing2015a].
+We used convergent cross mapping (CCM), in the manner of @clark2015, to test for the possibility of the trophic cascades biasing our results.
+Our results found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (SI Appendix Fig.S42).
+However, the longer MPAs are in place, the greater the possibility that substantial species interactions that can affect use of non-targeted species as a control may arise.
+
 Given the natural variability of marine ecosystems, and the large challenges of obtaining accurate samples from oceanic environments, how large of an effect would an MPA network have to have in order to allow a difference-in-difference strategy such as this to be a reliable measure of MPA effects?
@@
 In the context of the Channel Islands, given the potential effect size produced by our simulation model these results suggest that we might expect to be unable to precisely estimate the true population-level effect of the MPAs.
-
-<!-- XX need to think about whether MAPE is the correct error here - maybe need something like the variance of the estimator? -->
 
@@
 The population-level effects of MPA networks covering 25% or less of a species range may be difficult to detect in many places with already well-managed fisheries, while for that MPA network size we might expect clearer effects in less-managed locations (assuming compliance with MPA regulations).
-Within these broad guidelines a wide range of outcomes are possible based on local fleet and fish dynamics: users can use the bio-economic MPA simulation model developed for this paper to explore potential outcomes for specific MPAs using an interactive web application available [danovando.shinyapps.io/simmpa/](https://danovando.shinyapps.io/simmpa/).
-
-<!-- The key assumptions of the difference-in-difference model are that a) within the time-frame of the model there are no significant interaction effects between the targeted and non-targeted species (which in fact we do not detect, see SI Appendix), and b) that in the absence of the MPAs both the targeted and non-targeted groups of species would have exhibited similar trends in densities.
--->
-
-<!-- We briefly assess two of the most critical assumptions of this model here: that the treated and non-treated groups have parallel trends, and that the effect of the treatment on the treated does not tangentially affect the untreated.
-While the parallel trends assumption cannot be formally proven, we can examine its validity using the data from the years before the MPAs were put in place in 2003.
-We do not detect any significant differences in the trends of the biomass densities of the targeted and non-targeted species in the years before the MPAs (SI Appendix Fig.S26).
-However, we believe that the parallel trends assumption is likely violated in the post 2014 data.
--->
-
-<!-- With regards to the second assumption, all of the species in this empirical analysis exist within an ecosystem, and as such affect each other through mechanisms such as predation, competition, and habitat modification.
-We find it unlikely that these effects have had enough time to manifest in a meaningful way in the 14 years of post-MPA data used in our analysis [@babcock2010; @pershing2015a].
-We used convergent cross mapping (CCM), in the manner of @clark2015, to test for the possibility of the trophic cascades biasing our results.
-Our results found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (SI Appendix Fig.S42).
-However, the longer MPAs are in place, the greater the possibility that substantial species interactions that can affect use of non-targeted species as a control may arise.
--->
+Within these broad guidelines a wide range of outcomes are possible based on local fleet and fish dynamics.
 
@@
 This is exactly what we found in our empirical case study from the relatively large and well-studied network of MPAs in the Channel Islands, California, USA.
-
-<!-- We were not able to detect a statistically significant difference in the biomass densities of targeted and non-targeted fin-fish species over the 14 years of MPA protection in the Channel Islands covered by our analysis.
-The data (and model assumptions) provide support for both negative and positive MPA effects, although the size of these estimated effects is well within the range that our simulation analysis suggests are plausible.
-At face value, this lack of a clear result may seem surprising given the size and carefully studied nature of this MPA network.
--->
 
@@
 accounting for the displacement of fishing effort by MPAs), response ratios may be highly unreliable estimators of population-level MPA effects.
-When MPAs affect nearby control sites used in response ratios through biological spillover or concentration of fishing effort, it is entirely possible to for MPA to produce massive response ratios while simultaneously having minimal effects on the entire population partially protected by the MPAs (and *vice versa*.
+When MPAs affect nearby control sites used in response ratios through biological spillover or concentration of fishing effort, it is entirely possible to for MPA to produce massive response ratios while simultaneously having minimal effects on the entire population partially protected by the MPAs (and *vice versa*).
+
 As @larsen2019 suggests, there are many potential alternatives for estimating the effects of MPAs that better account for the challenges of causal inference (though that may be more data-intensive).
-We applied one such approach here (a difference-in-difference estimator), and yet were still unable to reach robust conclusions as to the effect of MPAs on the biomass density of targeted finfish in the Channel Islands, due to the likely small size of the true effect relative to the influence of environmental variability.
+We applied one such approach here (a difference-in-difference estimator, see online appendix for alternative strategies such as synthetic controls [@abadie2010] that show similar results), and yet were still unable to reach robust conclusions as to the effect of MPAs on the total biomass density of targeted finfish in the Channel Islands, due to the likely small size of the true effect relative to the influence of environmental variability.
 If pre-MPA data are available, data currently used in response ratios can easily be re-purposed for an analysis such as this one; we encourage holders of such data worldwide to conduct similar exercises to examine to what extent other empirical estimates of population-level MPA effects conform to MPA theory.
@@
 
+```{r ci-map, include = TRUE,fig.width=width_1col, fig.asp=1.5, fig.cap="Map of study region; the Northern Channel Islands, California, USA.
+Colors show binned number of PISCO sampling events across the time period of our study.", include=TRUE}
+
+temp_plot <- sample_location_plot + 
+  theme(axis.text = element_blank())
+
+temp_plot
+ggsave(
+  temp_plot,
+  filename = here::here("documents","figs","sample_location_plot.pdf"),
+  width = width_1col,
+  height = width_1col * asp_1col,
+  device = cairo_pdf
+)
+
+```
+
+```{r raw-density, fig.width = width_1col, fig.asp=asp_1col,fig.cap = "Centered and scaled trends in biomass densities of targeted and non-targeted finfish included in our study Top panel shows trends across all sites, with smaller background lines showing trends for each individual species.
+Bottom two panels show aggregate biomass density trends outside and inside MPAs"}
+
+temp_plot <-
+  (total_trend_plot / (mpa_trend_plot)) & theme(legend.position = "top", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), units = "lines"),
+                                                panel.grid = element_blank()
+  )
+
+temp_plot
+
+ggsave(
+  temp_plot,
+  filename = here::here("documents","figs","raw_biomass_density_plot.pdf"),
+  width = width_1col,
+  height = width_1col * asp_1col,
+  device = cairo_pdf
+)
+
+```
+
 ```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="90% Posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis) in blue.
@@
   height = width_2col * asp_2col,
-  device = cairo_pdf
-)
-
-```
-
-```{r raw-density, fig.width = width_1col, fig.asp=asp_1col,fig.cap = "Centered and scaled trends in biomass densities of targeted and non-targeted finfish included in our study Top panel shows trends across all sites, with smaller background lines showing trends for each individual species.
-Bottom two panels show aggregate biomass density trends outside and inside MPAs"}
-
-temp_plot <-
-  (total_trend_plot / (mpa_trend_plot)) & theme(legend.position = "top", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), units = "lines"),
-                                                panel.grid = element_blank()
-  )
-
-temp_plot
-
-ggsave(
-  temp_plot,
-  filename = here::here("documents","figs","raw_biomass_density_plot.pdf"),
-  width = width_1col,
-  height = width_1col * asp_1col,
   device = cairo_pdf
@@
 
-```{r ci-map, include = TRUE,fig.width=width_1col, fig.asp=1.5, fig.cap="Map of study region; the Northern Channel Islands, California, USA.
-Colors show binned number of PISCO sampling events across the time period of our study.", include=TRUE}
-
-temp_plot <- sample_location_plot + 
-  theme(axis.text = element_blank())
-
-temp_plot
-ggsave(
-  temp_plot,
-  filename = here::here("documents","figs","sample_location_plot.pdf"),
-  width = width_1col,
-  height = width_1col * asp_1col,
-  device = cairo_pdf
-)
-
-```
-
 # References

### 2020-06-24 make-pop-effects-of-mpas.R

@@ -62,7 +62,7 @@ run_tmb <- FALSE
 
 simulate_mpas <- FALSE # simulate MPA outcomes
 
-simulate_channel_islands <- TRUE # simulate MPA outcomes
+simulate_channel_islands <- FALSE # simulate MPA outcomes
 
 validate_mpas <- FALSE
 
@@ -2369,6 +2369,8 @@ if (process_results == TRUE){
   
   processed_grid$adult_movement <- (2 * processed_grid$adult_movement) / num_patches
   
+
+
   
 #
 #   zissou_theme <-
@@ -2404,7 +2406,6 @@ if (process_results == TRUE){
 
   processed_grid <- processed_grid %>%
     filter(!experiment %in% unique(bad_sims$experiment))
-
   
   write_rds(processed_grid, file.path(run_dir,"filtered_processed_grid.rds"))
 
@@ -2466,7 +2467,7 @@ if (process_results == TRUE){
     mutate(pop_effect = pmin(1,(`with-mpa` - `no-mpa`) / b0))
 
   write_rds(outcomes, file.path(run_dir,"outcomes.rds"))
-
+  
   density_ratios <- processed_grid %>%
     select(-absolute_mpa_effect, fishery_effect) %>% 
     mutate(mpa_effect = map(mpa_effect,~.x %>% select(-year,-mpa_size))) %>% 
@@ -4196,8 +4197,6 @@ mpa_effect_plot <- ggplot() +
 
 mpa_effect_plot
 
-
-# try it as a band?
 
 ## simulation results
 

### 2020-06-24 documents/ovando-et-al-pop-effects-of-mpas.Rmd
done all I can do

@@
 Over time our goals and expectations for MPAs have evolved; while all MPAs are expected to deliver conservation benefits within their borders, many modern MPAs are also established to bolster fish populations throughout the region in which they are located [@gaines2010].
-
 Recent international agreements to expand MPAs (most proximately, the Convention on Biological Diversity’s Strategic Plan for Biodiversity, which calls for 10% of coastal waters to be protected inside MPAs by 2020 and the International Union for Conservation of Nature call for 30% by 2030) are based on the expectation that well-designed MPAs will achieve benefits both within, and outside, their borders.
@@
 @goni2010; @halpern2009;@kay2012; @stobart2009; @mcclanahan2000; @russ1996; @thompson2017;@baetscher2019;@pelc2009], as well as alteration of fishing effort in reaction to [@costa2013;@mason2012;@murawski2005] and in anticipation of [@mcdermott2019] MPA placement.
-The assumption that MPAs will have conservation benefits beyond their borders for most species is implicit in all multilateral calls for MPA expansion, but our collective scientific understanding of the regional-scale impacts of current and future MPAs on fish populations and fisheries is surprisingly limited.
 
@@
 Control sites used in calculating response ratios are often selected based on abiotic or ecological traits such as habitat characteristics [@ferraro2018].
-Selection of these sites is complicated, though, by the very spillover that MPAs are often intended to create.
+Selection of these sites is complicated by the very spillover that MPAs are often intended to create.
 Export of adults or larvae from the MPA to the "control" site affects their status as controls, as does displacement of fishing effort from MPAs to control sites.
@@
 As a result of these challenges, response ratios can be a highly imprecise and biased measure of the population-level conservation effect of an MPA network [@ferraro2018].
+This means that our collective scientific understanding of the regional-scale impacts of current and future MPAs on fish populations and fisheries is surprisingly limited.
 
-Given the potential unreliability of response ratios for the task, how can we empirically measure the population-level effects of MPAs?
+<!-- Given the potential unreliability of response ratios for the task, how can we empirically measure the population-level effects of MPAs?
 The perfect experiment would involve two parallel worlds that were identical, except for the implementation of an MPA.
@@
 Careful causal inference requires some other means of controlling for biases introduced by factors such as the unobserved environmental shocks, the MPA siting process, biological spillover, and concentration of fishing effort  [@ferraro2018;@larsen2019].
+-->
 
-As stakeholders around the world seek to use MPAs in marine resource management portfolios, it is critical that we develop a better understanding of the magnitude and drivers of population-level MPA effects.
+<!-- As stakeholders around the world seek to use MPAs in marine resource management portfolios, it is critical that we develop a better understanding of the magnitude and drivers of population-level MPA effects.
 This is particularly true in the context of a changing climate, in which "set it and forget it" management strategies are increasingly untenable; we need to be able to monitor the performance of MPAs in order to adapt them as dictated by a shifting environment.
@@
 Overall, what are the population-level effects of MPAs?
+-->
 
@@
 Our analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*) (but our methods could similarly be applied to these species given data).
-Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species).
+Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*).
 We used a Bayesian difference-in-difference regression to estimate any difference in mean total biomass densities of fin-fish species targeted by fishing effort (i.e., those potentially affected by an MPA) and those species not targeted by fishing before and after MPA implementation [see @larsen2019 for an explanation of difference-in-difference estimators in ecological settings].
-The identification strategy control for unobserved environmental shocks to the system independent of the MPAs and similarly experienced by the targeted and non-targeted species.
+This identification strategy controls for unobserved environmental shocks to the system independent of the MPAs and similarly experienced by the targeted and non-targeted species.
 Conditional on the assumptions of the model, this regression produces an estimate of the effect of the MPAs on the mean total biomass densities of targeted species throughout the Channel Islands.
@@
 Suppose that the MPA increase biomass densities inside the reserves by 20%, and by 0% outside the reserve.
-By our metric the population conservation effect of the MPA would be 10%, but a response ratio value of infinity.
+By our metric the population conservation effect of the MPA would be 10%.
 The regression amounts to estimating the pre-post MPA difference in the mean total biomass density of targeted species minus the same difference for non-targeted species in the Channel Islands.
@@
 The shape and scale parameters of the Gamma distribution are estimated as well.
+Regression details and diagnostic plots are avilable in the online appendix.
+The full table of covariates and their posterior means and 89% credible interval can be found in Table.\@ref(tab:did-coefs).
 
-`r percent(mean(used_data$eventual_mpa))` of the samples in the raw data come from within MPAs, while the MPAs themselves cover 20% of the surface waters of the Channel Islands.
+<!-- r percent(mean(used_data$eventual_mpa))` of the samples in the raw data come from within MPAs, while the MPAs themselves cover 20% of the surface waters of the Channel Islands.
 As such, we weight each observation such that within MPA data are assigned a total weight of 0.2 and outside-MPA data a total weight of 0.8.
@@
 Since the covariates are centered and scaled, and the dependent variable is on the log-scale, this is an extremely diffuse prior (e.g.
-a one standard deviation change in an independent variable changes the dependent variable by 4 on the log scale).
-The full table of covariates and their posterior means and 89% credible interval can be found in Table.\@ref(tab:did-coefs).
-Diagnostic plots can be found in the online appendix, along with sensitivity analyses of various model structures estimation strategies.
+a one standard deviation change in an independent variable changes the dependent variable by 4 on the log scale) -->
 
@@
 We integrated these network design features, life history traits, and exploitation histories into our simulation model, varying key unknowns such as the degree of pre-MPA depletion, species' movement rates, and fleet dynamics.
-While a wide range of MPA effects are plausible, from 0% to upwards of 200%, the median simulated effect size for a stylized version of the Channel Islands after over ten years of protection was roughly `r percent(median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected > 10]))` (Fig.\@ref(fig:mpa-effect)).
+While a wide range of MPA effects are plausible, from 0% to upwards of 150%, the median simulated effect size for a stylized version of the Channel Islands after over ten years of protection was roughly `r percent(median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected > 10]))` (Fig.\@ref(fig:mpa-effect)).
 
@@
 
-How can we explain the lack of a clear population-level MPA effect in the Channel Islands after over ten years of protection?
-One explanation may lie in fleet dynamics.
+Containing a carefully designed, well-enforced, and well-studied MPA network, the Channel Islands would seem at face value to be an ideal location to study the population-level effects of protected areas.
+The persistently high response ratios suggest that despite overall decreases in targeted biomass densities inside and outside MPAs relative to pre-MPA measurements, the MPAs may still be providing protection within their borders.
+But, as we have shown here these response ratios are not necessarily an indicator of broader population-level effects.
+The difference-in-difference strategy utilized here presents an alternative identification strategy to response ratios, that while not without its own strict caveats presents some potential improvements over response-ratios as a means of estimating population-level effects.
+While we estimate an uncertain but overall positive effect of the MPA network in its first few years, we are unable to detect a robust signal from 2012-2017.
+After 14 years of MPA protection and following a large environmental perturbation we are left without a clear picture of the effect of the Channel Island MPA network on biomass densities of targeted fin-fish species from either response ratios or the difference-in-difference model.
+
+How can we explain this result?
 Much of the theoretical literature on MPAs assumes that all else being equal larger reserves produce larger conservation gains [@white2011 and references therein].
@@
 
-Alternatively, though, fishers can pursue a  “constant-catch” rather than an effort-based strategy.
-This is an important and, to our knowledge, unexplored fleet model in MPA simulation.
+Alternatively, fishers can pursue a  “constant-catch” rather than an effort-based strategy.
 Under the constant catch model, fishers have a catch objective, and exert as much (or little) effort as needed to achieve that objective.
@@
 Subsistence fisheries may use a constant-catch style policy over the short-term, as they seek to ensure that their food needs are met.
-Industrial fisheries may have pre-arranged agreements with buyers to deliver set amounts of fish.
 Constant-catch dynamics might also occur in fisheries with constraining quotas that are not updated after the implementation of MPAs.
@@
 
-This hypothesis is supported by the fact that we estimate a similar decline in the estimated MPA effect both inside and outside of the MPAs (SI Appendix Fig.S18) (as well as observe the decline in the raw data, Fig.\@ref(fig:raw-density)).
+<!-- This hypothesis is supported by the fact that we estimate a similar decline in the estimated MPA effect both inside and outside of the MPAs (SI Appendix Fig.S18) (as well as observe the decline in the raw data, Fig.\@ref(fig:raw-density)).
 If the cause of recent declines was due to increases in fishing pressure we would expect to see substantial declines only in the fished areas.
 While we see evidence for the validity of the parallel trends assumption in the early years of our data (Fig.\@ref(fig:mpa-effect), SI Appendix Fig.S26), it is likely that the non-targeted species reacted in a different way to the marine heatwave than the targeted species did, causing a violation of the parallel trends assumption in the post 2014 data.
+-->
 
-Containing a carefully designed, well-enforced, and well-studied MPA network, the Channel Islands would seem at face value to be an ideal location to study the population-level effects of protected areas.
-The persistently high response ratios suggest that despite overall decreases in targeted biomass densities inside and outside MPAs, the MPAs may still be providing protection within their borders.
-But, as we have shown here these response ratios are not necessarily an indicator of broader population-level effects.
-The difference-in-difference strategy utilized here presents an alternative identification strategy to response ratios , that while not without its own strict caveats presents some potential improvements over response-ratios as a means of estimating population-level effects.
-While we estimate an uncertain but overall positive effect of the MPA network in its first few years, we are unable to detect a robust signal from 2012-2017.
-We believe that the disappearance of increasing probabilities of positive MPA effects estimated by our model are likely driven by the marine heatwave experienced by the region beginning in 2014 (and potentially by exogenous increases in fishing mortality for key target species).
-This likely violation of the parallel trends assumption means that our post 2014 estimates of MPA effects may be unreliable.
-After 14 years of MPA protection and following a large environmental perturbation we are left without a clear picture of the effect of the Channel Island MPA network on biomass densities of targeted fin-fish species from either response ratios or the difference-in-difference model.
-
-  Simulation analysis suggests that we should not be surprised by this result.
+<!-- Simulation analysis suggests that we should not be surprised by this result.
 The Channel Islands MPAs cover 20% of the surface waters in the Channel Islands, and while formal stock assessments are not available for many of the targeted species in our analysis, what evidence we have suggests that, as a group, these fish are on average not heavily overfished.
@@
 annual) common in marine systems.
+-->
 
-Our simulations and difference-in-difference model examine percentage changes in biomass densities (as opposed to changes in total population size).
+<!-- Our simulations and difference-in-difference model examine percentage changes in biomass densities (as opposed to changes in total population size).
 An increase in biomass densities from 0.02kg/m^2^ to 0.04kg/m^2^ would be measured as a 100% increase.
@@
 spiny lobster, may also reveal clearer signals.
+-->
 
-Could trophic cascades exlpain the lack of a clear population-level MPA effect in the Channel Islands?
+Could trophic cascades explain the lack of a clear population-level MPA effect in the Channel Islands?
 All of the species in this empirical analysis exist within an ecosystem, and as such affect each other through mechanisms such as predation, competition, and habitat modification.
@@
 We then used a simple Bayesian difference-in-difference regression styled after the full model used here on these simulated data, and estimated the percent error between the posterior probability distribution of the estimated MPA effect and true simulated MPA effect.
-
 While unbiased across simulations, the difference-in-difference model struggled severely when MPA effect sizes were less than 25% and the model was faced with observation and process errors (Fig.\@ref(fig:val-plot)).
@@
 
-The MPA literature highlights a large number of variables beyond simply size and fishing pressure that can affect performance.
+<!-- The MPA literature highlights a large number of variables beyond simply size and fishing pressure that can affect performance.
 We examined the variability in expected MPA effects across each of the two major axes (pre-MPA depletion and MPA size).
@@
 However, for both variables even at extremely large or extremely small values a wide range of MPA effects were possible (though as we might expect the effect of pre-MPA depletion, in other words fishing pressure, was much clearer than MPA size alone) (Fig.\@ref(fig:sim-effects)-B).
+-->
 
@@
 
-MPAs are an important part of the marine resource management toolbox.
+<!-- MPAs are an important part of the marine resource management toolbox.
 Under ideal circumstances they can protect individual species and ecosystem linkages, while supporting local economies through tourism and fishing opportunities.
@@
 This is exactly what we found in our empirical case study from the relatively large and well-studied network of MPAs in the Channel Islands, California, USA.
+-->
 
-A large theoretical literature has pointed out the importance of a wide range of economic and environmental factors in determining the effectiveness of MPAs [e.g.
+<!-- A large theoretical literature has pointed out the importance of a wide range of economic and environmental factors in determining the effectiveness of MPAs [e.g.
 @hilborn2004; @smith2006a; @white2010c].
@@
 In the case of the Channel Islands, simulation modeling suggests for the network in question, an effect size on the order of 30% or smaller is likely, which our analysis in turn demonstrates may be challenging to detect even with a comprehensive monitoring program.
-This cautionary tale is likely to hold for many coastal MPA networks around the world.
+-->
 
-Once simulation models have been used to help design an MPA, how can users evaluate whether it is achieving their objectives?
+<!-- Once simulation models have been used to help design an MPA, how can users evaluate whether it is achieving their objectives?  -->
+How then should stakeholders think about anticipating and estimating the population-level effects of MPAs?
 Response ratios are commonly used as evidence for conservation outcomes of MPAs; @lester2009 and @halpern2002 present meta-analyses of hundreds of such studies.
 These results often find massively higher densities and biomass inside MPAs than outside [median biomass increases of  over 400%, @lester2009].
-But as suggested in @white2011, @ferraro2018, @larsen2019 and further demonstrated here, without careful attention to the design of control sites (e.g.
-accounting for the displacement of fishing effort by MPAs), response ratios may be highly unreliable estimators of population-level MPA effects.
-When MPAs affect nearby control sites used in response ratios through biological spillover or concentration of fishing effort, it is entirely possible to for MPA to produce massive response ratios while simultaneously having minimal effects on the entire population partially protected by the MPAs (and *vice versa*).
+But as suggested in @white2011, @ferraro2018, @larsen2019 and further demonstrated here, When MPAs affect nearby control sites used in response ratios through biological spillover or concentration of fishing effort, it is entirely possible to for MPA to produce massive response ratios while simultaneously having minimal effects on the entire population partially protected by the MPAs (and *vice versa*).
 
@@
 We applied one such approach here (a difference-in-difference estimator, see online appendix for alternative strategies such as synthetic controls [@abadie2010] that show similar results), and yet were still unable to reach robust conclusions as to the effect of MPAs on the total biomass density of targeted finfish in the Channel Islands, due to the likely small size of the true effect relative to the influence of environmental variability.
-If pre-MPA data are available, data currently used in response ratios can easily be re-purposed for an analysis such as this one; we encourage holders of such data worldwide to conduct similar exercises to examine to what extent other empirical estimates of population-level MPA effects conform to MPA theory.
 

### 2020-07-08 documents/ovando-et-al-pop-effects-of-mpas.Rmd
Owen's edits on main manuscript

@@
 We pair theory with empirical methods to show that the population-level conservation effects of marine protected areas can be highly variable and challenging to detect.
-We find no statistically clear effect of a large and well protected network of MPAs in the Channel Island, California, on targeted finfish after 14 years of protection, which our simulation analysis suggests is in fact to be expected.
+We find no statistically clear effect of a large and well protected network of MPAs in the Channel Island, California, on targeted fin-fish after 14 years of protection, which our simulation analysis suggests is in fact to be expected.
 MPA practitioners must clearly communicate expectations from MPA networks, and plan monitoring and adaptations strategies that directly address the challenges of estimating population-level conservation effects of protected areas.
@@
 Traditional cultures in Oceania utilized - often temporary - MPAs as "fish banks" for times of need [@johannes1978].
-Modern MPAs were first established primarily as marine analogs to the terrestrial protection of iconic landscapes like Yellowstone or Kruger National Parks [@iucn1976;@watson2014a].
-Over time our goals and expectations for MPAs have evolved; while all MPAs are expected to deliver conservation benefits within their borders, many modern MPAs are also established to bolster fish populations throughout the region in which they are located [@gaines2010].
-Recent international agreements to expand MPAs (most proximately, the Convention on Biological Diversity’s Strategic Plan for Biodiversity, which calls for 10% of coastal waters to be protected inside MPAs by 2020 and the International Union for Conservation of Nature call for 30% by 2030) are based on the expectation that well-designed MPAs will achieve benefits both within, and outside, their borders.
-The empirical MPA literature has generally focused on assessing conservation effects within the borders of protected areas [@lester2009;@halpern2003; @edgar2014; @gerber2005].As conservation benefits accrue inside MPAs, MPAs can affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
-These dynamics link fish both inside and outside of MPAs as part of a shared biological stock; the population.
-Several studies have documented empirical evidence for the existence of adult or larval spillover affecting both abundance and fisheries [e.g.
-@goni2010; @halpern2009;@kay2012; @stobart2009; @mcclanahan2000; @russ1996; @thompson2017;@baetscher2019;@pelc2009], as well as alteration of fishing effort in reaction to [@costa2013;@mason2012;@murawski2005] and in anticipation of [@mcdermott2019] MPA placement.
+Modern MPAs were first established primarily as marine analogs to the terrestrial protection of iconic landscapes [@iucn1976;@watson2014a].
+Over time our goals and expectations for MPAs have evolved and become more ambitious.
+Recent international agreements to expand MPAs are based on the assumption that well-designed MPAs will achieve benefits both within, and outside, their borders[@gaines2010].
 
-Numerous factors can affect the population effects of MPAs.
-These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], the strength and timing of density dependence in the population (e.g.
-pre- or post-settlement, @burgess2014), the age and degree of enforcement [@edgar2014],  how overfished the population would be without the MPA, and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006a;@walters2000a;@white2011,;@white2010c].
+The empirical MPA literature has generally focused on assessing conservation effects within MPA borders [@lester2009;@halpern2003; @edgar2014; @gerber2005].As conservation benefits accrue inside MPAs, MPAs can affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
+These dynamics link fish stocks inside MPAs to those outside.
+Several studies have documented empirical evidence for the existence of adult or larval fish spillover affecting both fish abundance and fisheries productivity [e.g.
+@goni2010; @halpern2009;@kay2012; @stobart2009; @mcclanahan2000; @russ1996; @thompson2017;@baetscher2019;@pelc2009].
+MPAs can also cause alterations in fishing effort [@costa2013;@mason2012;@murawski2005;@mcdermott2019].
+
+Numerous factors influence how MPAs affect region-wide fish populations.
+These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], the strength and timing of density dependence [@burgess2014], the degree of enforcement [@edgar2014], how overfished the population would be without the MPA, and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006a;@walters2000a;@white2011,;@white2010c].
 In addition, even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through network effects [@gaines2010; @costello2010a].
 
-Empirical evidence of the conservation outcomes of MPAs is made up almost entirely by versions of "response ratios", commonly measured as biomass densities of species inside MPAs relative to biomass densities of species in areas outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015].
+Empirical evidence of the conservation outcomes of MPAs is made up almost entirely by versions of "response ratios", commonly measured as the ratio of biomass densities of species inside relative to outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015].
 These studies have by and large found clear evidence that well enforced and sufficiently sized MPAs are associated with high response ratios.
@@
 
-Control sites used in calculating response ratios are often selected based on abiotic or ecological traits such as habitat characteristics [@ferraro2018].
-Selection of these sites is complicated by the very spillover that MPAs are often intended to create.
-Export of adults or larvae from the MPA to the "control" site affects their status as controls, as does displacement of fishing effort from MPAs to control sites.
-In theory, control sites far enough away to negate both biological spillover and concentration by the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs may be challenging.
-As a result of these challenges, response ratios can be a highly imprecise and biased measure of the population-level conservation effect of an MPA network [@ferraro2018].
-This means that our collective scientific understanding of the regional-scale impacts of current and future MPAs on fish populations and fisheries is surprisingly limited.
+Control sites used in calculating response ratios are often selected based on habitat characteristics [@ferraro2018].
+However, export of adults or larvae from the MPA to these "control" site affects their status as controls, as does displacement of fishing effort from MPAs to control sites.
+In theory, control sites far enough away to negate both biological spillover and concentration by the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs is challenging.
+As a result, response ratios can be a highly imprecise and biased measure of the population-level conservation effect of an MPA network [@ferraro2018].
+This means that our collective scientific understanding of the regional-scale impacts of MPAs is surprisingly limited.
 
@@
 
-The MPAs within the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) provide an ideal case study to address this need.
+The MPAs within the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) provide an opportunity to improve our understanding of population-level MPA effects.
 A network of protected areas covering approximately 20% of the Islands' waters was put in place in 2003 [see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of these MPAs].
-This MPA network has subsequently been used as a model case study in protected area design around the world [@botsford2014 and references therein].
-We use data from the first 14 years of protection to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of finfish species.
-In contrast to clear differences in biomass densities observed inside and outside of well protected MPAs both globally [@lester2009] and in the Channel Islands [@caselle2015] we are unable to detect a clear population effect from the Channel Islands MPAs.
+The network has been used as a model case study in protected area design around the world [@botsford2014 and references therein].
+We use data from the first 14 years of protection to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of fin-fish species.
+In contrast to clear differences in biomass densities observed inside and outside of well-protected MPAs both globally [@lester2009] and in the Channel Islands [@caselle2015] we are unable to detect a clear population effect from the Channel Islands MPAs.
 We build off of existing MPA theory to consider why this might be, and provide guidance for scientists and managers as to when and how we might expect to estimate the population-level conservation effects of MPAs.
@@
 The simulation model consists of 50 patches with wrapped edges.
-For any one simulation we randomly pull a species and its associated life history from the `FishLife` [@thorson2017c] package.
+For any one simulation we randomly pull a species and its associated life history traits from the `FishLife` [@thorson2017c] package.
 We pair these data with randomly selected values governing the characteristics of the simulation (Table.1).
-These include attributes of recruitment (variation, autocorrelation, and timing), movement (magnitude of larval and adult dispersal and density dependent movement), fleet dynamics (degree of fishing pressure, amount and spatial allocation of fishing effort), and MPA design (size and placement, habitat characteristics, larval source).
-The population begins at unfished equilibrium and then fishing effort is applied in accordance to the fleet model.
-The MPAs are then placed, and each simulation is run to equilibrium with and without the selected MPA strategy (holding all else constant).
-We then measure the difference in biomass densities in each time step in the scenario with and without the MPAs to calculate the population effect of the MPAs over time.
-The results provide us a library of simulated MPA effects under a range of plausible biological and economic states of the world.
+These include parameters governing recruitment (variation, autocorrelation, and timing), movement (magnitude of larval and adult dispersal and density dependent movement), fleet dynamics (degree of fishing pressure, distribution of fishing effort), and MPA design (size and placement, habitat characteristics, larval source).
+The population begins at unfished equilibrium.
+After fishing effort is applied, each simulation is run to equilibrium with and without the selected MPA design (holding all else constant).
+In each time step, we measure the difference in biomass densities between the scenario with and without the MPAs to calculate the population effect of the MPAs over time.
+Simulation results provide a library of plausible MPA effects for a range of biological and economic assumptions.
 
-We also created one set of simulations specifically designed to reflect the dynamics of the Channel Islands.
+One set of simulations is specifically designed to reflect the dynamics of the Channel Islands.
 We only include species of the same genus as those targeted by fishing in the Channel Islands.
-The MPAs in this study cover 20% of the surface waters of the Channel Islands, a region spanning roughly 90km east to west.
+The MPAs in this study cover 20% of the surface waters of the Channel Islands.
 Detailed dispersal studies are not available for all of the species covered by our study, but what information we have suggest that while adults of some of these species are likely to exhibit site fidelity [@love2011], larvae are likely distributed beyond the Channel Islands [@rassweiler2012].
 Therefore, if we assume that the population of the targeted species is at least equal to the extent of the Channel Islands, we can assume that at most the MPAs cover roughly 20% of the targeted populations' ranges.
-As such we cap the MPA size in the Channel Islands specific simulation batch to 20%.
+As such we cap the MPA size in the Channel Islands simulation to 20%.
 
@@
 The difference-in-difference model used empirical kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands.
-PISCO conducts visual SCUBA surveys at a large number of rocky-reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\ref{fig:ci-map}).
+PISCO conducts visual SCUBA surveys at a large number of rocky reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\ref{fig:ci-map}).
 The details of the monitoring program are described in @caselle2015.
 
-Building off of the concepts explored in @caselle2015, we used an identification strategy utilizing biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 0])` species that are not directly targeted by fishing before and after MPA implementation as our control group (non-targeted), and biomass densities before and after MPA implementation of `r n_distinct(used_data$classcode[used_data$targeted == 1])` species targeted by fishing as our treatment group.
-Targeted species in the Channel Islands, include commercially important fin-fish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
+Building off of the concepts explored in @caselle2015, we used an identification strategy utilizing biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 0])` species that are not directly targeted by fishing as our control group (non-targeted), and biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 1])` species targeted by fishing as our treatment group.
+Targeted species in the Channel Islands include fin-fish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
 Each of these targeted species was the subject of prior bio-economic modeling related to the effects of MPAs in southern California [@costello2010a;@rassweiler2014].
-Our analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*) (but our methods could similarly be applied to these species given data).
+Our analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*), although our methods could similarly be applied to these species given data.
 Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*).
 We used a Bayesian difference-in-difference regression to estimate any difference in mean total biomass densities of fin-fish species targeted by fishing effort (i.e., those potentially affected by an MPA) and those species not targeted by fishing before and after MPA implementation [see @larsen2019 for an explanation of difference-in-difference estimators in ecological settings].
-This identification strategy controls for unobserved environmental shocks to the system independent of the MPAs and similarly experienced by the targeted and non-targeted species.
+This identification strategy controls for unobserved environmental shocks to the system that are independent of the MPAs.
 Conditional on the assumptions of the model, this regression produces an estimate of the effect of the MPAs on the mean total biomass densities of targeted species throughout the Channel Islands.
 
-We define the population-level conservation effects of MPAs as the change in mean total biomass densities of targeted finfish both inside and outside of MPAs, relative to the mean total biomass densities of targeted finfish inside and outside of MPAs that would have occurred without the MPAs.
-As an example of a population-level MPA effect, consider a evenly distributed population that has 50% of its range protected by an MPA.
-Suppose that the MPA increase biomass densities inside the reserves by 20%, and by 0% outside the reserve.
+We define the population-level conservation effects of MPAs as the change in mean total biomass densities of targeted fin-fish both inside and outside of MPAs, relative to the mean total biomass densities of targeted fin-fish inside and outside of MPAs that would have occurred without the MPAs.
+For example, consider an evenly distributed population that has 50% of its range protected by an MPA.
+Suppose that the MPA increases biomass densities inside its boundaries by 20%, and by 0% outside the reserve.
 By our metric the population conservation effect of the MPA would be 10%.
@@
 
-What should we expect *a priori* the population-level effect size of the Channel Islands MPAs to be?
-While we know many parameters will affect the expected effects of the MPAs, theory suggests that life history, adult and larval dispersal relative to MPA size, and fishing pressure will be key drivers.
+Our Channel Islands-specific simulation provides a set of outcomes that can feasibly bound our expectations for the empirical effects of the MPA network.
 The targeted species in our database span a range of life histories, but are largely made up of fishes in the perch and rockfish complexes, with a mean Von Bertalanffy growth coefficient of `r round(mean(used_data$k[used_data$targeted == 1]),2)`, and a median age at maturity of `r round(median(used_data$tm[used_data$targeted == 1]),0)` years [values drawn from @thorson2017c].
-Formal stock assessments are largely lacking for these species.
-However, what evidence we have suggests mostly moderate fishing pressure, with some species such as California sheephead (*Semicossyphus pulcher*) and blue rockfish (*Sebastes mystinus*) experiencing high levels of fishing mortality, likely resulting in biomass levels below management targets during the early 2000s [@alonzo2004;@dick2017].
+Although stock assessments are largely lacking for these species, what evidence we have suggests mostly moderate fishing pressure, with high fishing mortality for some species such as California sheephead (*Semicossyphus pulcher*) and blue rockfish (*Sebastes mystinus*).
+For many species, this likely means that biomass levels during the early 2000s were below target management levels[@alonzo2004;@dick2017].
 We integrated these network design features, life history traits, and exploitation histories into our simulation model, varying key unknowns such as the degree of pre-MPA depletion, species' movement rates, and fleet dynamics.
-While a wide range of MPA effects are plausible, from 0% to upwards of 150%, the median simulated effect size for a stylized version of the Channel Islands after over ten years of protection was roughly `r percent(median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected > 10]))` (Fig.\@ref(fig:mpa-effect)).
+While a wide range of MPA effects are plausible, from 0% to upwards of 150%, the median simulated effect size for this stylized version of the Channel Islands after over ten years of protection was roughly `r percent(median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected > 10]))` (Fig.\@ref(fig:mpa-effect)).
 
-Having established a theoretical grounding for the likely biomass density outcomes of the Channel Island MPAs, what empirical evidence do we have of the effects?
-@caselle2015 examined changes in response ratios of species targeted by fishing activity inside and outside of the MPAs over time and compared these changes to the trends in non-targeted species.
-They found a statistically significant increase in the response ratios of targeted species over time, and evidence that this increase is smaller in the non-targeted species.
+We compared this range of plausible outcomees to the empirical data.
+Previous research [@caselle2015] found a statistically significant increase in the response ratios of targeted species over time, and evidence that this increase is smaller in the non-targeted species.
 Updating the results of @caselle2015 through 2017 shows a continuation of the increasing trend in the response ratios of targeted species (Fig.\@ref(fig:rr-plot)).
@@
 
+These response ratios cannot, however, be used as a definitive indicator of overall population effects of the MPAs.
 In the case of the Channel Islands MPAs, control sites are often located within a few kilometers of an MPA, suggesting that they are susceptible to both local biological spillover and by concentration of fishing effort excluded from the MPAs.
-Given these complications, we can interpret the response ratios reported in @caselle2015, and updated here, as evidence that the MPAs are providing effective protection of targeted fish biomass inside their borders, particularly in light of the general increase in response ratios over time.
-But, we cannot use these response ratio results serve as a definitive indicator of overall population effects of the MPAs.
+Given these complications, the population-wide effects of the Channel Islands MPAs are difficult to determine with response ratios.
 
 To illustrate this problem, we paired our simulated Channel Islands response ratios to the estimated posterior probability distributions of the empirical Channel Island response ratios.
-For example, if the mean estimated response ratio in the year 2006 (year three after implementation) is one, we found all simulations that had simulated response ratios near one after three years of protection, and then pulled out the "true" simulated population-level MPA effect from each of those simulations.
+For example, if the mean estimated response ratio in the year 2006 (year three after implementation) is one, we found all simulations that had response ratios near one after three years of protection, and then pulled out the "true" simulated population-level MPA effect from each of those simulations.
 This provides us with a distribution of simulated population-level MPA effects that could plausibly generate the types of response ratios actually observed in the Channel Islands.
@@
 The response ratio trends we observe in the data could plausibly be produced by a wide range of population-level MPA effects (Fig.\@ref(fig:rr-plot)).
-Response ratios well over one were associated with population-level MPA effects generally less than 25%, and, importantly, many simulations produced large response ratios but population-level MPA effects close to 0%.
+Response ratios well over one were associated with "true" population-level MPA effects generally less than 25%, and, importantly, many simulations produced large response ratios but population-level MPA effects close to 0%.
 This can occur if for example fishing pressure is only moderate, adult movement is low, larval dispersal is high, and displaced fishing effort concentrates around the border of the MPAs.
@@
 
-If response ratios are not clear evidence of the population level effect of the Channel Islands MPAs, what does our difference-in-difference strategy find?
+Our empirical difference-in-difference regression provides a new perspective on population-wide MPA effects.
 Consistent with MPA theory, over the first three years of implementation (2003-2006), the effects of the MPAs are unclear, with support for a small negative effect to a substantially positive effect, with much higher probability of a small positive effect (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`,
@@
 These empirical estimates are in line with the outcomes that our simulation model suggests are plausible.
-While this concordance between theory and empirics is by no means proof of the robustness of our results, it is an important line of evidence.
-However, in the subsequent years the trend reverses itself, and for the years 2015-2017 we once again see no clear effect of the MPAs (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`,
+While this concordance between theory and empirics is not proof of the robustness of our results, it is an important line of evidence.
+However, in the subsequent years the trend reverses itself, and in 2015-2017 we once again see no clear effect of the MPAs (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`,
 90\% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`
@@
 
-Containing a carefully designed, well-enforced, and well-studied MPA network, the Channel Islands would seem at face value to be an ideal location to study the population-level effects of protected areas.
+Containing a carefully designed, well-enforced, and well-studied MPA network, the Channel Islands seems to be an ideal location to study the population-level effects of protected areas.
 The persistently high response ratios suggest that despite overall decreases in targeted biomass densities inside and outside MPAs relative to pre-MPA measurements, the MPAs may still be providing protection within their borders.
 But, as we have shown here these response ratios are not necessarily an indicator of broader population-level effects.
-The difference-in-difference strategy utilized here presents an alternative identification strategy to response ratios, that while not without its own strict caveats presents some potential improvements over response-ratios as a means of estimating population-level effects.
-While we estimate an uncertain but overall positive effect of the MPA network in its first few years, we are unable to detect a robust signal from 2012-2017.
+The difference-in-difference strategy presents an alternative identification strategy.
+It is not without its own strict caveats, but provides some potential improvements over response-ratios as a means of estimating population-level effects.
+While we estimate an uncertain but overall positive effect of the MPA network in its first few years, we are unable to detect a robust positive signal from 2012-2017.
 After 14 years of MPA protection and following a large environmental perturbation we are left without a clear picture of the effect of the Channel Island MPA network on biomass densities of targeted fin-fish species from either response ratios or the difference-in-difference model.
@@
 How can we explain this result?
-Much of the theoretical literature on MPAs assumes that all else being equal larger reserves produce larger conservation gains [@white2011 and references therein].
+Much of the theoretical literature on MPAs assumes that larger reserves produce larger conservation gains [@white2011 and references therein].
 However, these models generally simulate fleet dynamics through fishing mortality rates (e.g.
@@
 
-Alternatively, fishers can pursue a  “constant-catch” rather than an effort-based strategy.
+Alternatively, though, fishers can pursue a  “constant-catch” rather than an effort-based strategy.
 Under the constant catch model, fishers have a catch objective, and exert as much (or little) effort as needed to achieve that objective.
-While a constant-catch greater than the maximum sustainable catch from the population is not possible over the long-term under the assumptions of our model, over the short-term a constant-catch scenario is possible.
+While a constant-catch greater than the maximum sustainable catch from the population is not possible over the long-term, over the short-term a constant-catch scenario is possible.
 Subsistence fisheries may use a constant-catch style policy over the short-term, as they seek to ensure that their food needs are met.
 Constant-catch dynamics might also occur in fisheries with constraining quotas that are not updated after the implementation of MPAs.
-Under a constant-catch though, fishers have to fish much harder than before to get the same catch from a smaller part of the population, reducing the size structure of the population and subsequently causing net conservation loss under `r percent(1 - positive_outcomes$pp[positive_outcomes$fleet_model == "constant-catch"], accuracy = 1)` of our constant-catch simulations.
+Fishers pursuing constant catch under declining fish population size have to continually fish harder to achieve the same catch from a smaller part of the population, reducing the size structure of the population and subsequently causing net conservation loss under `r percent(1 - positive_outcomes$pp[positive_outcomes$fleet_model == "constant-catch"], accuracy = 1)` of our constant-catch simulations.
 This is an important and often overlooked possibility, especially as MPAs are increasingly implemented in quota-managed fisheries [@liu2018].
@@
 However, some of the most common targeted species in the Channel Islands, such as blue rockfish and copper rockfish, have seen dramatic increases in reported catches in the Santa Barbara region since the year 2010 (although other such as California sheephead have seen declines relative to pre-MPA catch level).
-To the extent that these increases in catches are due to external market forces, the dissapearance of clear positive population-level MPA effects estimated by our model could be due to exogenous increases in fishing pressure overwhelming any population-level increases caused by the MPA network.
-We lack a control though that would allow us to estimate the MPA effects relative to a counter-factual in which the same market forces increase catches but no MPAs are in place.
+To the extent that these increases in catches are due to external market forces, the disappearance of clear positive population-level MPA effects estimated by our model could be due to exogenous increases in fishing pressure overwhelming any population-level increases caused by the MPA network.
+Unfortunately, we lack a control that would allow us to estimate the MPA effects relative to a counterfactual in which the same market forces increase catches but no MPAs are in place.
 
@@
 The targeted group is made up mostly of fishes with cold-water affinities, such as members of the genus *Sebastes*.
-We hypothesize that the recent evidence for a decline in densities of targeted species is due to environmental conditions that disproportionately affect the targeted group (and not for example due to concentrated fishing pressure outside the reserves).
-The practical result of this is that in the presence of this marine heatwave the non-targeted species may no longer serve as an effective control for the evolution of biomass densities of targeted finfish in the absence of the MPAs (a violation of the "parallel trends" assumption required for Difference-in-difference models).
+We hypothesize that the recent evidence for a decline in densities of targeted species is due to environmental conditions that disproportionately affected the targeted group.
+The practical result of this is that in the presence of this marine heatwave the non-targeted species may no longer serve as an effective control for the evolution of biomass densities of targeted fin-fish in the absence of the MPAs (a violation of the "parallel trends" assumption required for difference-in-difference models).
 
@@
 
-Could trophic cascades explain the lack of a clear population-level MPA effect in the Channel Islands?
-All of the species in this empirical analysis exist within an ecosystem, and as such affect each other through mechanisms such as predation, competition, and habitat modification.
-We find it unlikely that these effects have had enough time to manifest in a meaningful way in the 14 years of post-MPA data used in our analysis [@babcock2010; @pershing2015a].
-We used convergent cross mapping (CCM), in the manner of @clark2015, to test for the possibility of the trophic cascades biasing our results.
+Trophic cascades are another possible explanation for our equivocal results.
+All of the species in this empirical analysis may affect each other through mechanisms such as predation, competition, and habitat modification.
+We used convergent cross mapping (CCM), in the manner of @clark2015, to test for significant dynamic interactions between species and therfore the possibility of the trophic cascades biasing our results.
 Our results found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (SI Appendix Fig.S42).
@@
 
-Given the natural variability of marine ecosystems, and the large challenges of obtaining accurate samples from oceanic environments, how large of an effect would an MPA network have to have in order to allow a difference-in-difference strategy such as this to be a reliable measure of MPA effects?
-We simulated data from a range of scenarios with increasing MPA effect size, along with increasing degrees of observation error and natural recruitment variation.
-As an added measure, we include scenarios in which the sampled species go through recruitment regimes, which may be positive for both targeted and non-targeted species, or positive for non-targeted species and negative for targeted species.
-We then used a simple Bayesian difference-in-difference regression styled after the full model used here on these simulated data, and estimated the percent error between the posterior probability distribution of the estimated MPA effect and true simulated MPA effect.
-While unbiased across simulations, the difference-in-difference model struggled severely when MPA effect sizes were less than 25% and the model was faced with observation and process errors (Fig.\@ref(fig:val-plot)).
-Even models fit to data generated from large effect sizes commonly misestimated the true MPA effect by 50% or more.
+Our finding of small MPA effects in recent years is therefore likely a result of both changing fishing dynamics and environmental disturbance.
+Given the natural variability of marine ecosystems, how large of an effect would an MPA network have to have in order to allow a difference-in-difference strategy such as this to be a reliable measure of MPA effects?
+Our library of simulation results allows us to explore how MPA effects, and our ability to measure them, might vary with observation error and natural fish recruitment variation (including the possibility of recruitment regimes).
+Using a simple Bayesian difference-in-difference regression on these simulated data, we estimated the percent error between the posterior probability distribution of the estimated MPA effect from the regression and the true simulated MPA effect.
+
+While unbiased across simulations, the difference-in-difference model struggled severely when MPA effect sizes were less than 25% and the model was faced with both observation and process errors (Fig.\@ref(fig:val-plot)).
+Even models fit to data generated from large effect sizes commonly mis-estimated the true MPA effect by 50% or more.
 Obtaining a mean absolute percent error (MAPE) of 25% or less across our simulated datasets required a population-level MPA effect of at least 30%.
@@
 However, since nearly any omission which one can think of would make an MPA effect harder to detect, not easier, these results serve as a useful floor for the likely difficulty in estimating MPA effects.
-In the context of the Channel Islands, given the potential effect size produced by our simulation model these results suggest that we might expect to be unable to precisely estimate the true population-level effect of the MPAs.
+In the context of the Channel Islands, this finding unfortunately suggests that we should not be surprised at our difficulty in precisely estimating the population-level effect of the MPAs.
 
-When should we expect to see MPA effects big enough to stand a reasonable chance of detection?
-We simulated `r n_distinct(outcomes$experiment)` MPA scenarios across a wide range of life histories, network designs, and fishing dynamics.
-Suppose that we are willing to tolerate a MAPE of 25%.
-Our analysis suggests that we would need a population-level MPA effect size of at least 30% to achieve this.
 Two of the most critical drivers of MPA effect size are the size of the MPA network and the degree of fishing pressure.
-Looking across these two variables, if the MPA network covers 25% or more of a species range and pre-MPA depletion is greater than 60% we might expect an effect size with our target MAPE.
-Although recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, most MPA networks for near-shore commercial finfish are likely to cover areas more in line with the Channel Islands (20%) or smaller, and as such have population-level effect sizes that may be difficult to detect unless they were experiencing extreme overfishing prior to MPA establishment (Fig.\@ref(fig:sim-effects)-A).
+Looking across these two variables, the MPA network must be large (25% or more of a species range) and the target species overfished (pre-MPA depletion greater than about 60%) to achieve an effect size with an MAPE of 25%.
+Although recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, most MPA networks for near-shore commercial fin-fish are likely to cover areas more in line with the Channel Islands (20%) or smaller.
+As such, many MPA networks are expected to have population-level effect sizes that are difficult to detect unless target species were experiencing extreme overfishing prior to MPA establishment (Fig.\@ref(fig:sim-effects)-A).
 
@@
 -->
-
-To put these results into context, the FAO estimates that 7% of the worlds fisheries with status estimates fall into the relatively unexploited category (roughly depletion less than 50%), 60% fall into the fully fished category (roughly depletion 50%-70%), and 33% fall into the heavily fished category (roughly depletion greater than 70%) [@fao2018, assuming a biomass at maximum sustainable yield to unfished biomass ratio of 40%, per @thorson2012], although studies that include a broader range of fisheries estimate that 50% or more of stocks fall into the heavily fished category [@costello2012;@rosenberg2018].
-The population-level effects of MPA networks covering 25% or less of a species range may be difficult to detect in many places with already well-managed fisheries, while for that MPA network size we might expect clearer effects in less-managed locations (assuming compliance with MPA regulations).
-Within these broad guidelines a wide range of outcomes are possible based on local fleet and fish dynamics.
 
@@
 <!-- Once simulation models have been used to help design an MPA, how can users evaluate whether it is achieving their objectives?  -->
-How then should stakeholders think about anticipating and estimating the population-level effects of MPAs?
+How then should stakeholders reconsider the population-level effects of MPAs?
 Response ratios are commonly used as evidence for conservation outcomes of MPAs; @lester2009 and @halpern2002 present meta-analyses of hundreds of such studies.
 These results often find massively higher densities and biomass inside MPAs than outside [median biomass increases of  over 400%, @lester2009].
-But as suggested in @white2011, @ferraro2018, @larsen2019 and further demonstrated here, When MPAs affect nearby control sites used in response ratios through biological spillover or concentration of fishing effort, it is entirely possible to for MPA to produce massive response ratios while simultaneously having minimal effects on the entire population partially protected by the MPAs (and *vice versa*).
+But as suggested in @white2011, @ferraro2018, @larsen2019 and further demonstrated here, when MPAs affect nearby control sites used in response ratios through biological spillover or concentration of fishing effort, it is entirely possible to for MPA to produce massive response ratios while simultaneously having minimal effects on the entire population partially protected by the MPAs (and *vice versa*).
 
 As @larsen2019 suggests, there are many potential alternatives for estimating the effects of MPAs that better account for the challenges of causal inference (though that may be more data-intensive).
-We applied one such approach here (a difference-in-difference estimator, see online appendix for alternative strategies such as synthetic controls [@abadie2010] that show similar results), and yet were still unable to reach robust conclusions as to the effect of MPAs on the total biomass density of targeted finfish in the Channel Islands, due to the likely small size of the true effect relative to the influence of environmental variability.
+We applied one such approach here (a difference-in-difference estimator, see online appendix for alternative strategies such as synthetic controls [@abadie2010] that show similar results), and yet were still unable to reach robust conclusions as to the effect of MPAs on the total biomass density of targeted fin-fish in the Channel Islands, due to the likely small size of the true effect relative to the influence of environmental variability.
 
-While this does not mean that all MPAs will face similar challenges in estimating their effects, our results in the relatively large, well-enforced, and rigorously studied (though also relatively lightly fished) Channel Islands Marine Protected Area network make clear that in many instances empirically detecting a clear effect of MPAs on total biomass densities of targeted finfish at the population level may not be possible.
-How then should stakeholders go about adaptively monitoring and managing MPAs?
-Simulation modeling incorporating the diverse drivers represented here can help inform the range of effect sizes that may be expected, and monitoring programs can perhaps be tuned to focus on the species groups that have the highest chance of a detectable effect size over the early years of the reserve [@nickols2019].
+While this does not mean that all MPAs will face similar challenges in estimating their effects, our results in the relatively large, well-enforced, and rigorously studied (though also relatively lightly fished) Channel Islands Marine Protected Area network make clear that in many instances empirically detecting a clear effect of MPAs on total biomass densities of targeted fin-fish at the population level may not be possible.
+In the absence of clear empirical findings, simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can be tuned to focus on the species groups that have the highest chance of a detectable effect size over the early years of the reserve [@nickols2019].
 Expanding data collection to include robust monitoring of spatio-temporal fleet dynamics may help assess the validity of control sites used in response ratios, support the direct inclusion of these fleet dynamics into statistical models, and allow managers to take into account potential negative interactions between MPAs and fleet dynamics such as those that may occur under constant-catch dynamics.
@@
 Educating communities about the challenges of estimating the effects of MPAs can help ensure that a lack of a clear effect is not necessarily viewed as a failure of the program, or large positive result based solely on response ratios as a clear sign of success and subsequent relaxation of other fishery management strategies.
-Rather, results and subsequent management actions (such as adaptation of MPA networks) must be considered in the context of reasonable expectations given the size, age, and degree of enforcement of the MPAs in question, together with the ecological and economic dynamics of a given system.
+Rather, results and subsequent management actions must be considered in the context of reasonable expectations given the size, age, and degree of enforcement of the MPAs in question, together with the ecological and economic dynamics of a given system.
 While this paper has focused on the conservation outcomes of MPAs, future work must also address the challenge of predicting and estimating the fishery impacts of protected areas [in the manner of @smith2006a].
@@
 As the number and size of global MPA networks increase, it is critical that we both set appropriate expectations for their outcomes, and plan how we will monitor the performance of these protected areas over time.
-While the history of MPA science has made important strides in helping us understand the dynamics of protected areas, the future of MPA science must directly tackle the challenge of evaluating the performance of these MPAs at the regional scale, a task which has to date not been widely addressed.
+While the history of MPA science has made important strides in helping us understand the dynamics of protected areas, the future of MPA science must directly tackle the challenge of evaluating the performance of these MPAs at the regional scale.
 This is particularly true if communities are depending on MPAs as their primary marine resource management tool.
-Commonly employed metrics such as response ratios may be applicable in some circumstances,for example, when an MPA is designed for food security in the face of a disaster or for protection of unique or threatened species inside its borders.
-But, response ratios can also have shortcomings as metrics of population-level effects.
+Commonly employed metrics such as response ratios may be applicable in some circumstances, but are vulnerable to inaccuracy or misuse as metrics of population-level effects.
 Dependence on unreliable estimators of MPA effects may lead to stakeholders incorrectly attributing negative environmental shocks as MPA failures, or interpreting data arising from scorched-earth fishing outside MPAs as a conservation success.
@@
 Bio-economic modeling can help frame community expectations, reducing the potential for a reduction in support if unrealistic conservation or fishery expectations are not realized.
-Statistical approaches that explicitly address complications such as the spatial spillover effects of MPAs (such as the difference-in-difference approach used here) may give users an improved understanding of the performance of their MPAs, but even they may struggle when expected effect sizes are small.
+Statistical approaches like difference-in-difference regression that explicitly address complications such as the spatial spillover effects of MPAs (such as the difference-in-difference approach used here) may give users an improved understanding of the performance of their MPAs, but even they may struggle when expected effect sizes are small.
 Clearly communicating what we should expect, and what we can detect, from MPAs is critical to ensuring that MPAs play effective roles in fisheries management and marine conservation.
@@
 
-```{r raw-density, fig.width = width_1col, fig.asp=asp_1col,fig.cap = "Centered and scaled trends in biomass densities of targeted and non-targeted finfish included in our study Top panel shows trends across all sites, with smaller background lines showing trends for each individual species.
+```{r raw-density, fig.width = width_1col, fig.asp=asp_1col,fig.cap = "Centered and scaled trends in biomass densities of targeted and non-targeted fin-fish included in our study Top panel shows trends across all sites, with smaller background lines showing trends for each individual species.
 Bottom two panels show aggregate biomass density trends outside and inside MPAs"}

### 2020-07-09 documents/ovando-et-al-pop-effects-of-mpas.Rmd
Cons bio owenedits (#3)
* Owen's edits on main manuscript

* a few edits to owen's edits and adding in renv, merging back into conservation biology branch

Co-authored-by: Owen.Liu <1571646680@mil>

@@
 While numerous studies show that MPAs produce conservation benefits *inside* their borders, many MPAs are also justified on the grounds that they confer conservation benefits to the broader population *beyond* their borders.
-This paper examines the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and shows that even in cases where the conservation benefits are large, they are inherently difficult to detect empirically.
+This paper examines the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and shows that even in cases where the population benefits are large, they are inherently difficult to detect empirically.
 A network of MPAs was put in place in The Channel Islands National Marine Sanctuary in 2003, with a goal of providing regional conservation and fishery benefits.
@@
   - \usepackage{lineno}\linenumbers
+always_allow_html: true
 ---
@@
 
+* fix that mentino of inside and outside stocks
+
+* make clearer the idea that size of parallel trends violation exceeds effect size of MPA
+
 # Front Matter
@@
 The empirical MPA literature has generally focused on assessing conservation effects within MPA borders [@lester2009;@halpern2003; @edgar2014; @gerber2005].As conservation benefits accrue inside MPAs, MPAs can affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
-These dynamics link fish stocks inside MPAs to those outside.
+These dynamics link the state of fishes inside MPAs to those outside.
 Several studies have documented empirical evidence for the existence of adult or larval fish spillover affecting both fish abundance and fisheries productivity [e.g.
@@
 Numerous factors influence how MPAs affect region-wide fish populations.
-These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], the strength and timing of density dependence [@burgess2014], the degree of enforcement [@edgar2014], how overfished the population would be without the MPA, and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006a;@walters2000a;@white2011,;@white2010c].
+These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], the strength and timing of density dependence [@burgess2014], the degree of enforcement [@edgar2014], how overfished the population would be without the MPA, the time span under evaluation [@ovando2016;@moffitt2013], and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006a;@walters2000a;@white2011,;@white2010c].
 In addition, even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through network effects [@gaines2010; @costello2010a].
 
-Empirical evidence of the conservation outcomes of MPAs is made up almost entirely by versions of "response ratios", commonly measured as the ratio of biomass densities of species inside relative to outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015].
+The empirical evidence of the conservation outcomes of MPAs is made up almost entirely by versions of "response ratios", commonly measured as the ratio of biomass densities of species inside relative to outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015].
 These studies have by and large found clear evidence that well enforced and sufficiently sized MPAs are associated with high response ratios.
@@
 The shape and scale parameters of the Gamma distribution are estimated as well.
-Regression details and diagnostic plots are avilable in the online appendix.
+Regression details and diagnostic plots are available in the online appendix.
 The full table of covariates and their posterior means and 89% credible interval can be found in Table.\@ref(tab:did-coefs).
@@
 Our Channel Islands-specific simulation provides a set of outcomes that can feasibly bound our expectations for the empirical effects of the MPA network.
-The targeted species in our database span a range of life histories, but are largely made up of fishes in the perch and rockfish complexes, with a mean Von Bertalanffy growth coefficient of `r round(mean(used_data$k[used_data$targeted == 1]),2)`, and a median age at maturity of `r round(median(used_data$tm[used_data$targeted == 1]),0)` years [values drawn from @thorson2017c].
+The targeted species in our database span a range of life histories, but are largely made up of fishes in the perch and rockfish complexes, with a mean Von Bertalanffy growth coefficient of `r round(mean(used_data$k[used_data$targeted == 1]),2)`, and a median age at maturity of `r round(median(used_data$tm[used_data$targeted == 1]),0)` years.
 Although stock assessments are largely lacking for these species, what evidence we have suggests mostly moderate fishing pressure, with high fishing mortality for some species such as California sheephead (*Semicossyphus pulcher*) and blue rockfish (*Sebastes mystinus*).
-For many species, this likely means that biomass levels during the early 2000s were below target management levels[@alonzo2004;@dick2017].
+For many species, this likely means that biomass levels during the early 2000s were below target management levels [@alonzo2004;@dick2017].
 We integrated these network design features, life history traits, and exploitation histories into our simulation model, varying key unknowns such as the degree of pre-MPA depletion, species' movement rates, and fleet dynamics.
@@
 
-We compared this range of plausible outcomees to the empirical data.
+We compared this range of plausible outcomes to the empirical data.
 Previous research [@caselle2015] found a statistically significant increase in the response ratios of targeted species over time, and evidence that this increase is smaller in the non-targeted species.
@@
 While we estimate an uncertain but overall positive effect of the MPA network in its first few years, we are unable to detect a robust positive signal from 2012-2017.
-After 14 years of MPA protection and following a large environmental perturbation we are left without a clear picture of the effect of the Channel Island MPA network on biomass densities of targeted fin-fish species from either response ratios or the difference-in-difference model.
+After 14 years of MPA protection we are left without a clear picture of the effect of the Channel Island MPA network on biomass densities of targeted fin-fish species from either response ratios or the difference-in-difference model.
 
@@
 We hypothesize that the recent evidence for a decline in densities of targeted species is due to environmental conditions that disproportionately affected the targeted group.
-The practical result of this is that in the presence of this marine heatwave the non-targeted species may no longer serve as an effective control for the evolution of biomass densities of targeted fin-fish in the absence of the MPAs (a violation of the "parallel trends" assumption required for difference-in-difference models).
+The practical result of this is that in the presence of this marine heatwave the non-targeted species may no longer serve as an effective control for the evolution of biomass densities of targeted fin-fish in the absence of the MPAs, given the magnitude of the environmental shock relative to the size of the population-level MPA effect.
 
@@
 All of the species in this empirical analysis may affect each other through mechanisms such as predation, competition, and habitat modification.
-We used convergent cross mapping (CCM), in the manner of @clark2015, to test for significant dynamic interactions between species and therfore the possibility of the trophic cascades biasing our results.
+We used convergent cross mapping (CCM), in the manner of @clark2015, to test for significant dynamic interactions between species and therefore the possibility of the trophic cascades biasing our results.
 Our results found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (SI Appendix Fig.S42).
@@
 These results often find massively higher densities and biomass inside MPAs than outside [median biomass increases of  over 400%, @lester2009].
-But as suggested in @white2011, @ferraro2018, @larsen2019 and further demonstrated here, when MPAs affect nearby control sites used in response ratios through biological spillover or concentration of fishing effort, it is entirely possible to for MPA to produce massive response ratios while simultaneously having minimal effects on the entire population partially protected by the MPAs (and *vice versa*).
+But as suggested in @white2011, @ferraro2018, @larsen2019, @moffitt2013 and further demonstrated here, when MPAs affect nearby control sites used in response ratios through biological spillover or concentration of fishing effort, it is entirely possible to for MPA to produce massive response ratios while simultaneously having minimal effects on the entire population partially protected by the MPAs (and *vice versa*).
 
@@
 While this does not mean that all MPAs will face similar challenges in estimating their effects, our results in the relatively large, well-enforced, and rigorously studied (though also relatively lightly fished) Channel Islands Marine Protected Area network make clear that in many instances empirically detecting a clear effect of MPAs on total biomass densities of targeted fin-fish at the population level may not be possible.
-In the absence of clear empirical findings, simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can be tuned to focus on the species groups that have the highest chance of a detectable effect size over the early years of the reserve [@nickols2019].
+In the absence of clear empirical findings, simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can be tuned to focus on the species groups that have the highest chance of a detectable effect size over the early years of the reserve.
 Expanding data collection to include robust monitoring of spatio-temporal fleet dynamics may help assess the validity of control sites used in response ratios, support the direct inclusion of these fleet dynamics into statistical models, and allow managers to take into account potential negative interactions between MPAs and fleet dynamics such as those that may occur under constant-catch dynamics.
@@
 
-Non-equilibrium analyses also help set expectations for effect sizes over time [@nickols2019,@kaplan2019,Fig.\ref{did-plot}].
+Non-equilibrium analyses also help set expectations for effect sizes over time [@nickols2019,@kaplan2019,@ovando2016,Fig.\ref{did-plot}].
 Educating communities about the challenges of estimating the effects of MPAs can help ensure that a lack of a clear effect is not necessarily viewed as a failure of the program, or large positive result based solely on response ratios as a clear sign of success and subsequent relaxation of other fishery management strategies.
@@
 
+MPAs are an important part of the marine resource management toolbox.
 As the number and size of global MPA networks increase, it is critical that we both set appropriate expectations for their outcomes, and plan how we will monitor the performance of these protected areas over time.
@@
 Bio-economic modeling can help frame community expectations, reducing the potential for a reduction in support if unrealistic conservation or fishery expectations are not realized.
-Statistical approaches like difference-in-difference regression that explicitly address complications such as the spatial spillover effects of MPAs (such as the difference-in-difference approach used here) may give users an improved understanding of the performance of their MPAs, but even they may struggle when expected effect sizes are small.
+Statistical approaches like difference-in-difference regression that explicitly address complications such as the spatial spillover effects of MPAs may give users an improved understanding of the performance of their MPAs, but even they may struggle when expected effect sizes are small.
 Clearly communicating what we should expect, and what we can detect, from MPAs is critical to ensuring that MPAs play effective roles in fisheries management and marine conservation.

### 2020-07-09 make-pop-effects-of-mpas.R

@@ -68,7 +68,7 @@ validate_mpas <- FALSE
 
 process_results <- TRUE
 
-knit_paper <- FALSE
+knit_paper <- TRUE
 
 get_cdfw_catches <-  FALSE
 

### 2020-07-11 documents/ovando-et-al-pop-effects-of-mpas.Rmd
finally updated tables to be correct enough for WORD

@@
 
-* fix that mentino of inside and outside stocks
-
-* make clearer the idea that size of parallel trends violation exceeds effect size of MPA
-
 # Front Matter
@@
 For any one simulation we randomly pull a species and its associated life history traits from the `FishLife` [@thorson2017c] package.
-We pair these data with randomly selected values governing the characteristics of the simulation (Table.1).
+We pair these data with randomly selected values governing the characteristics of the simulation (Table.\@ref(tab:sim-table)).
 These include parameters governing recruitment (variation, autocorrelation, and timing), movement (magnitude of larval and adult dispersal and density dependent movement), fleet dynamics (degree of fishing pressure, distribution of fishing effort), and MPA design (size and placement, habitat characteristics, larval source).
@@
 Regression details and diagnostic plots are available in the online appendix.
-The full table of covariates and their posterior means and 89% credible interval can be found in Table.\@ref(tab:did-coefs).
+The full table of covariates and their posterior means and 89% credible interval can be found in Table.\@ref(tab:coef-table).
 
@@
 
-I need to clean these up
+```{r sim-table, results = "asis", echo=FALSE}
+variable <- c(
+"Scientific Name",
+"steepness",
+"Adult movement ($\\sigma_{s=a}$)",
+"Larval movement ($\\sigma_{s=l}$)",
+"Recruitment variation ($\\sigma_{r}$)",
+"Recruitment autocorrelation ($ac_{r}$)",
+"DD adult movement",
+"Density-dependence timing ",
+"% Patches in MPA",
+"Initial fishing relative to natural mortality",
+"Selectivity as a multiple of maturity length",
+"Fleet model",
+"Spatial effort model",
+"Years into simulation to start MPA",
+"MPA is Larval Source?",
+"Randomly place MPA?",
+"Fleet reaction to MPA",
+"Patchiness",
+"MPA habitat factor")
 
-Table.2 - Range of simulated variables
+distribution <- 
+c(
+"Drawn from all possible species in `FishLife` (@thorson2017c)",
+"uniform(0.6,0.95)",
+"uniform(0,0.25 * P)",
+"uniform(0,0.25 * P)",
+ "$\\in{0,0.05,.1,.2}$",
+ "$\\in{0,0.05,.1,.2}$",
+ "$\\in{0.25,1}$",
+ "$\\in{Local, Global,PostDispersal}$",
+ "uniform(0.01,1)",
+ "uniform(0.01,4)",
+ "uniform(0.1,1.25)",
+"$\\in{OpenAccess, ConstantEffort,ConstantCatch}$",
+ "$\\in{Uniform, Biomass,Profits}$",
+ "uniform(5,0.66)",
+"$\\in{TRUE,FALSE}$",
+"$\\in{TRUE,FALSE}$",
+"$\\in{Concentrate,Leave}$",
+"uniform(0.01,0.75)",
+"$\\in{1, 4}$"
+)
 
-| Variable | Distribution|
-|---------:|:-----------:|
-| Scientific Name | Drawn from all possible species in `FishLife` (@thorson2017c)|
-|steepness (h) | ~uniform(0.6,0.95)|
-|Adult movement ($\sigma_{s=a}$) | ~uniform(0,0.25 * P)|
-|Larval movement ($\sigma_{s=l}$) | ~uniform(0,0.25 * P)|
-|Recruitment variation ($\sigma_{r}$) |  $\in\{0,0.05,.1,.2\}$|
-|Recruitment autocorrelation ($ac_{r}$) |  $\in\{0,0.05,.1,.2\}$|
-|DD adult movement (dmod) | $\in\{0.25,1\}$|
-|Density-dependence timing | $\in\{local, global,post-dispersal\}$|
-|\% Patches in MPA | ~uniform(0.01,1)|
-|Initial fishing relative to natural mortality | ~uniform(0.01,4)|
-|Selectivity as a multiple of maturity length| ~uniform(0.1,1.25)|
-|Fleet model| $\in\{open-access, constant-effort,constant-catch\}$|
-|Spatial effort model| $\in\{uniform, biomass,profits\}$|
-|Years into simulation to start MPA| ~round(uniform(5,0.66 * T))|
-|MPA is sprinkler?| $\in\{TRUE,FALSE\}$|
-|Randomly place MPA?| $\in\{TRUE,FALSE\}$|
-|Fleet reaction to MPA| $\in\{concentrate, abandon-ship\}$|
-|Patchiness | ~uniform(0.01,0.75)|
-|MPA habitat factor| $\in\{1, 4\}$|
+equations <- tibble(Variable = variable, Distribution = distribution )
 
-```{r did-coefs, results = "asis",eval = TRUE}
-did_fit %>%
+knitr::kable(equations,
+             escape = FALSE,
+             caption = "Variables for MPA simulations. U mean uniform",
+             format = "pipe") %>% 
+  kableExtra::kable_styling()
+
+```
+
+```{r coef-table, results = "asis", echo=FALSE}
+
+coefs <- did_fit %>%
   broom::tidy(intervals = TRUE, prob = 0.89) %>%
@@
   rename(
-    mean = estimate,
-    lower_89th_ci = lower,
-    upper_89th_ci = upper
-  ) %>%
+    Mean = estimate,
+    `Lower 89th Percentile` = lower,
+    `Upper 89th Percentile` = upper
+  )
+
   knitr::kable(
+    coefs,
     digits = 2,
-    caption = "Posterior means and credible interval for key model coefficients",
-    longtable = TRUE,
-    booktabs = TRUE ) %>%
-  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
+    caption = "Posterior means and credible interval for key model coefficients.
+Conditional on the assumptions of the model, the targeted:year_bins coefficients reflect the effect of the MPAs on mean total biomass of targeted finfish in the Channel Islands.
+targeted indicates whether the species in question is targeted, 'var' refers to a covariate, where tex = technican experience, surge is mean reported surge, kelp is mean reported kelp cover, lag catch is the total reported catch for targeted species in the previous year, temp is the mean temperature at a site, temp_dev is the deviation of a temperature observation from the mean temperatuer for that island in question",
+    format = "pipe")
 

### 2020-07-17 documents/ovando-et-al-pop-effects-of-mpas-si.Rmd
starting final word sladh, need to lose 1600 :(

--- 2020-04-26 documents/ovando-et-al-pop-effects-of-mpas-si.Rmd
@@
 ---
-title: "Supporting Information"
+title: "Online Supporting Information"
 subtitle: "Assessing the Population-level Conservation Effects of Marine Protected Areas"
@@
   pdf_document: default
-bibliography: My Library.bib
+bibliography: references.bib
 header-includes:
@@
 library(Synth)
+library(here)
 
@@
 
+# (APPENDIX) Appendix {-}
+
 # Supporting Information (SI)
-
-# SI Text {#si-text .unnumbered}
 
@@
 \newpage
-```{r cartoon, fig.cap= "Cartoon illustration of the hierarchichal difference-in-difference estimator", fig.width=3}
-knitr::include_graphics("figs/estimation-cartoon.png")
+```{r cartoon, fig.cap= "Cartoon illustration of the hierarchichal difference-in-difference estimator", fig.width=8}
+knitr::include_graphics(here::here("documents","figs","estimation-cartoon.png"))
 ```

### 2020-07-17 documents/ovando-et-al-pop-effects-of-mpas.Rmd
starting final word sladh, need to lose 1600 :(

@@
 ---
-title: |
-  | Assessing the Population-level Conservation Effects of
-  | Marine Protected Areas
+title: Assessing the Population-level Conservation Effects of Marine Protected Areas
 author: 
@@
 Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level.
-MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present), and the size of many MPA networks suggests that effect sizes may often be smaller than 30%.
+We show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); the size of many MPA networks suggests that effect sizes may often be smaller than 30%.
 Our results provide a novel assessment of the population-level effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.
-bibliography: My Library.bib
+bibliography: references.bib
 csl: conservation-biology.csl
@@
     reference_docx: template.docx
-  bookdown::pdf_document2: default
+  bookdown::pdf_document2:
+    keep_tex: true
   bookdown::html_document2: default
@@
 ```{r, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
-                      cache = FALSE, fig.width = 3.42, fig.asp = 1.5, dev = "png")
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 300,
+                      cache = FALSE, fig.width = 3.42, fig.asp = 1.5, dev = "pdf")
 
@@
 
-# Front Matter
-
-## Author Contributions: 
-  D.O., S.G., and R.H.
-developed structure of simulation model.
-D.O., O.D., C.C., and J.C.
-developed estimation strategy.
-J.C.
-provided support on collecting and interpreting data.
-All simulations, statistics, and sensitivity analyses performed by D.O.
-and O.L.
-.
-All authors contributed to the manuscript.
-
-## Conflicts of Interest:
-  R.H's research program receives funding from environmental NGOs, foundations, fishing industry, governments and international agencies.
-All of these can be interpreted as a conflict of interest when evaluating fisheries policy.
-C.C., S.G., and O.D's research program includes funding from environmental NGOs and foundations with an interest in ocean conservation and food security.
-## Significance Statement
-  Healthy marine ecosystems are critical to the well-being of the planet.
-Marine protected areas (MPAs) are increasingly being used to conserve and manage these ecosystems.
-We pair theory with empirical methods to show that the population-level conservation effects of marine protected areas can be highly variable and challenging to detect.
-We find no statistically clear effect of a large and well protected network of MPAs in the Channel Island, California, on targeted fin-fish after 14 years of protection, which our simulation analysis suggests is in fact to be expected.
-MPA practitioners must clearly communicate expectations from MPA networks, and plan monitoring and adaptations strategies that directly address the challenges of estimating population-level conservation effects of protected areas.
-
-## Acknowledgements
-  Funding for the simulation model was provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship.
-Empirical data collection was funded primarily by the David and Lucille Packard Foundation in support of the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) with additional funding from the California Ocean Protection Council and California SeaGrant.
-This study would not be possible without the work provided by PISCO divers over the years, with special thanks to to Kathyrn Davis Koehn and Avrey Parsons-Field.
-We also thank Cody Szuwalski, Julia Lawson, André Punt, and two anonymous reviewers for helpful comments and technical support.
-
-## Keywords:
-  - Marine Protected Area Networks
-  - Conservation
-  - Bio-economic modeling
-  - Program Evaluation
-
 # Main Text
@@
 
-The empirical MPA literature has generally focused on assessing conservation effects within MPA borders [@lester2009;@halpern2003; @edgar2014; @gerber2005].As conservation benefits accrue inside MPAs, MPAs can affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
-These dynamics link the state of fishes inside MPAs to those outside.
-Several studies have documented empirical evidence for the existence of adult or larval fish spillover affecting both fish abundance and fisheries productivity [e.g.
+The empirical MPA literature has generally focused on assessing the ability of MPAs to provide conservation gains within their borders [@lester2009;@halpern2003; @edgar2014; @gerber2005].
+However, as conservation benefits accrue inside MPAs, MPAs can affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
+These dynamics result in MPAs affecting not only the organisms within their borders, but the broader population, the connected biological stock, of affected species.
+Theory states that numerous factors influence how MPAs might affect region-wide fish populations.
+These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], the strength, timing, and location of density dependence [@burgess2014], the degree of enforcement [@edgar2014], how overfished the population would be without the MPA, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006a;@walters2000a;@white2011,;@white2010c;@ovando2016;@moffitt2013].
+Even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through network effects [@gaines2010; @costello2010a].
+
+What empirical evidence is there for these population-level effects of MPAs?
+Conservation outcomes of MPAs are generally evaluated by versions of "response ratios", commonly measured as the ratio of biomass densities of species inside relative to outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015].
+These studies have by and large found clear evidence that well enforced and sufficiently sized MPAs are associated with high response ratios.
+Several studies have also documented empirical evidence for the existence of adult or larval fish spillover affecting both fish abundance and fisheries productivity [e.g.
 @goni2010; @halpern2009;@kay2012; @stobart2009; @mcclanahan2000; @russ1996; @thompson2017;@baetscher2019;@pelc2009].
-MPAs can also cause alterations in fishing effort [@costa2013;@mason2012;@murawski2005;@mcdermott2019].
-
-Numerous factors influence how MPAs affect region-wide fish populations.
-These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], the strength and timing of density dependence [@burgess2014], the degree of enforcement [@edgar2014], how overfished the population would be without the MPA, the time span under evaluation [@ovando2016;@moffitt2013], and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006a;@walters2000a;@white2011,;@white2010c].
-In addition, even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through network effects [@gaines2010; @costello2010a].
-
-The empirical evidence of the conservation outcomes of MPAs is made up almost entirely by versions of "response ratios", commonly measured as the ratio of biomass densities of species inside relative to outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015].
-These studies have by and large found clear evidence that well enforced and sufficiently sized MPAs are associated with high response ratios.
+MPAs have also caused alterations in fishing effort [@costa2013;@mason2012;@murawski2005;@mcdermott2019].
 Why does this large body of empirical work not serve as sufficient evidence for the population-level effects of MPAs?
 
+While spillover-based studies provide evidence that MPAs may have effects outside their borders, they are not sufficient evidence for the net effect of the MPAs.
+Detection of a biomass density or catch-per-unit-effort gradient decreasing with distance from an MPA border suggests that there may be spillover, but does not tell us the net effect of that spillover on the population or the fishery.
+Response ratios have shortcomings as measures of population-level effects as well.
 Control sites used in calculating response ratios are often selected based on habitat characteristics [@ferraro2018].
@@
 In theory, control sites far enough away to negate both biological spillover and concentration by the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs is challenging.
-As a result, response ratios can be a highly imprecise and biased measure of the population-level conservation effect of an MPA network [@ferraro2018].
-This means that our collective scientific understanding of the regional-scale impacts of MPAs is surprisingly limited.
+As a result, response ratios can be a highly imprecise and biased measure of the population-level conservation effect of an MPA network [@ferraro2018;@moffitt2013].
+As response ratios are our primary source of evidence for the conservation effects of MPAs, this means that our collective scientific understanding of the population-level impacts of MPAs is surprisingly limited.
 
@@
 The MPAs within the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) provide an opportunity to improve our understanding of population-level MPA effects.
-A network of protected areas covering approximately 20% of the Islands' waters was put in place in 2003 [see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of these MPAs].
-The network has been used as a model case study in protected area design around the world [@botsford2014 and references therein].
+A network of protected areas covering approximately 20% of the Islands' waters was put in place in 2003 (see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of these MPAs).
+The network has been used as a model case study in protected area design around the world (@botsford2014 and references therein).
 We use data from the first 14 years of protection to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of fin-fish species.
@@
 Our difference-in-difference model was fit using Stan [@carpenter2017] using the `rstanarm` package [@goodrich2020].
-All data and code needed to fully reproduce our results are publicly available at [github.com/DanOvando/population-effects-of-mpas](https://github.com/DanOvando/population-effects-of-mpas).
+All data and code needed to fully reproduce this manuscript are publicly available at [github.com/DanOvando/population-effects-of-mpas](https://github.com/DanOvando/population-effects-of-mpas).
+A detailed description of the simulation model structure, as well as sensitivity analyses of our estimation model, are available in the online supporting information (SI).
 
@@
 Readers can explore the functionality of the model using an online tool available at [danovando.shinyapps.io/simmpa/](https://danovando.shinyapps.io/simmpa/).
-Detailed methods can be found in the online appendix.
+Detailed methods can be found in the online supporting information.
 
@@
 The shape and scale parameters of the Gamma distribution are estimated as well.
-Regression details and diagnostic plots are available in the online appendix.
+Regression details and diagnostic plots are available in the SI.
 The full table of covariates and their posterior means and 89% credible interval can be found in Table.\@ref(tab:coef-table).
@@
 We compared this range of plausible outcomes to the empirical data.
-Previous research [@caselle2015] found a statistically significant increase in the response ratios of targeted species over time, and evidence that this increase is smaller in the non-targeted species.
+@caselle2015 found a statistically significant increase in the response ratios of targeted species over time, and evidence that this increase is smaller for non-targeted species.
 Updating the results of @caselle2015 through 2017 shows a continuation of the increasing trend in the response ratios of targeted species (Fig.\@ref(fig:rr-plot)).
@@
 
-While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (see SI Appendix Fig.S1) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
+While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (see SI Fig.S1) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
 However, some of the most common targeted species in the Channel Islands, such as blue rockfish and copper rockfish, have seen dramatic increases in reported catches in the Santa Barbara region since the year 2010 (although other such as California sheephead have seen declines relative to pre-MPA catch level).
@@
 We used convergent cross mapping (CCM), in the manner of @clark2015, to test for significant dynamic interactions between species and therefore the possibility of the trophic cascades biasing our results.
-Our results found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (SI Appendix Fig.S42).
+We found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (SI Fig.S42).
 However, the longer MPAs are in place, the greater the possibility that substantial species interactions that can affect use of non-targeted species as a control may arise.
@@
 <!-- Once simulation models have been used to help design an MPA, how can users evaluate whether it is achieving their objectives?  -->
-How then should stakeholders reconsider the population-level effects of MPAs?
+How then should stakeholders evaluate the population-level effects of MPAs?
 Response ratios are commonly used as evidence for conservation outcomes of MPAs; @lester2009 and @halpern2002 present meta-analyses of hundreds of such studies.
@@
 As @larsen2019 suggests, there are many potential alternatives for estimating the effects of MPAs that better account for the challenges of causal inference (though that may be more data-intensive).
-We applied one such approach here (a difference-in-difference estimator, see online appendix for alternative strategies such as synthetic controls [@abadie2010] that show similar results), and yet were still unable to reach robust conclusions as to the effect of MPAs on the total biomass density of targeted fin-fish in the Channel Islands, due to the likely small size of the true effect relative to the influence of environmental variability.
+We applied one such approach here (a difference-in-difference estimator, see SI for alternative strategies such as synthetic controls, @abadie2010, that show similar results), and yet were still unable to reach robust conclusions as to the effect of MPAs on the total biomass density of targeted fin-fish in the Channel Islands, due to the likely small size of the true effect relative to the influence of environmental variability.
 
@@
 
-Non-equilibrium analyses also help set expectations for effect sizes over time [@nickols2019,@kaplan2019,@ovando2016,Fig.\ref{did-plot}].
+Non-equilibrium analyses also help set expectations for effect sizes over time (@nickols2019;@kaplan2019;@ovando2016;Fig.\ref{fig:mpa-effect}).
 Educating communities about the challenges of estimating the effects of MPAs can help ensure that a lack of a clear effect is not necessarily viewed as a failure of the program, or large positive result based solely on response ratios as a clear sign of success and subsequent relaxation of other fishery management strategies.
 Rather, results and subsequent management actions must be considered in the context of reasonable expectations given the size, age, and degree of enforcement of the MPAs in question, together with the ecological and economic dynamics of a given system.
-While this paper has focused on the conservation outcomes of MPAs, future work must also address the challenge of predicting and estimating the fishery impacts of protected areas [in the manner of @smith2006a].
+While this paper has focused on the conservation outcomes of MPAs, future work must also address the challenge of predicting and estimating the fishery impacts of protected areas, an important question for which extremely little empirical evidence is available [@smith2006a].
 
@@
 
-# Supporting Information
+# Literature Cited
 
-A detailed description of the simulation model structure, as well as sensitivity analyses of our estimation model, are available in the online appendix.
+<div id="refs"></div>
 
@@
              escape = FALSE,
-             caption = "Variables for MPA simulations. U mean uniform",
+             caption = "Variables for MPA simulations",
              format = "pipe") %>% 
@@
 Simulated population-level (pop.) effect on biomass densities matched to empirical response ratios in red.
-For response ratios, a value of zero indicates that biomass densities of targeted species are identical inside and outside MPAs, a value of one that biomass densities of targeted species are 100% inside MPAs relative to outside.
+For response ratios, a value of zero indicates that biomass densities of targeted species are identical inside and outside MPAs, a value of one that biomass densities of targeted species are 100% greater inside MPAs relative to outside.
 For MPA population effect, a value of zero indicates that biomass densities are identical in the with- and without- MPA scenarios.
@@
 ```
-
-# References

### 2020-07-17 documents/pop-effects-of-mpas-front-matter.Rmd
starting final word sladh, need to lose 1600 :(

---
title: Assessing the Population-level Conservation Effects of Marine Protected Areas
subtitle: Cover Page
date: "`r Sys.Date()`"
output:
  bookdown::word_document2:
    template: uw.docx
  pdf_document: default
---

Submit your cover page as a separate document. It should NOT be part of the manuscript itself. The cover page must include the title of the paper; an article impact statement (see below); a running head (short title of 20 or fewer characters); a list of 5-8 keywords (do not repeat words in the title); word count (all text from the first word of the Abstract through the last word of the Literature Cited but not including table or figure legends or the body of tables); authors' complete mailing addresses (including street addresses and postal codes) at the time the work was conducted and present addresses if different; name and complete address (including email) of the person to whom correspondence should be sent; and text of your Acknowledgments section.

## Article Impact Statement

  Healthy marine ecosystems are critical to the well-being of the planet. Marine protected areas (MPAs) are increasingly being used to conserve and manage these ecosystems. We pair theory with empirical methods to show that the population-level conservation effects of marine protected areas can be highly variable and challenging to detect. We find no statistically clear effect of a large and well protected network of MPAs in the Channel Island, California, on targeted fin-fish after 14 years of protection, which our simulation analysis suggests is in fact to be expected. MPA practitioners must clearly communicate expectations from MPA networks, and plan monitoring and adaptations strategies that directly address the challenges of estimating population-level conservation effects of protected areas.

## Running Head

The Population Effects of MPAS

## Keywords:
  - Marine Protected Area Networks
  - Marine Conservation
  - Program Evaluation
  - Causal Inference
  - Bio-economic Modeling

## Word Count

7166

## Author Details

J

## Corresponding Author

## Acknowledgements

Funding for the simulation model was provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship. Empirical data collection was funded primarily by the David and Lucille Packard Foundation in support of the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) with additional funding from the California Ocean Protection Council and California SeaGrant. This study would not be possible without the work provided by PISCO divers over the years, with special thanks to to Kathyrn Davis Koehn and Avrey Parsons-Field. We also thank Cody Szuwalski, Julia Lawson, André Punt, and two anonymous reviewers for helpful comments and technical support.

### 2020-07-18 documents/ovando-et-al-pop-effects-of-mpas.Rmd
version submitted to conservation biology

@@
 title: Assessing the Population-level Conservation Effects of Marine Protected Areas
-author: 
-  - Daniel Ovando
-  - Jennifer E. Caselle
-  - Christopher Costello
-  - Olivier Deschenes
-  - Steven D. Gaines
-  - Ray Hilborn
-  - Owen Liu
 abstract: |
@@
 While numerous studies show that MPAs produce conservation benefits *inside* their borders, many MPAs are also justified on the grounds that they confer conservation benefits to the broader population *beyond* their borders.
-This paper examines the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and shows that even in cases where the population benefits are large, they are inherently difficult to detect empirically.
+We examine the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and show that even in cases where the population benefits are large, they are inherently difficult to detect empirically.
 A network of MPAs was put in place in The Channel Islands National Marine Sanctuary in 2003, with a goal of providing regional conservation and fishery benefits.
@@
 output:
+  bookdown::pdf_document2:
+    keep_tex: true
   bookdown::word_document2:
     reference_docx: template.docx
-  bookdown::pdf_document2:
-    keep_tex: true
   bookdown::html_document2: default
@@
 
-# Main Text
-
-## Introduction
+# Introduction {#intro .unnumbered}
 
@@
 Traditional cultures in Oceania utilized - often temporary - MPAs as "fish banks" for times of need [@johannes1978].
-Modern MPAs were first established primarily as marine analogs to the terrestrial protection of iconic landscapes [@iucn1976;@watson2014a].
+Modern MPAs were first established as marine analogs to the terrestrial protection of iconic landscapes [@iucn1976].
 Over time our goals and expectations for MPAs have evolved and become more ambitious.
@@
 
-The empirical MPA literature has generally focused on assessing the ability of MPAs to provide conservation gains within their borders [@lester2009;@halpern2003; @edgar2014; @gerber2005].
-However, as conservation benefits accrue inside MPAs, MPAs can affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
-These dynamics result in MPAs affecting not only the organisms within their borders, but the broader population, the connected biological stock, of affected species.
-Theory states that numerous factors influence how MPAs might affect region-wide fish populations.
-These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], the strength, timing, and location of density dependence [@burgess2014], the degree of enforcement [@edgar2014], how overfished the population would be without the MPA, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006a;@walters2000a;@white2011,;@white2010c;@ovando2016;@moffitt2013].
-Even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through network effects [@gaines2010; @costello2010a].
+The empirical MPA literature has focused on assessing the ability of MPAs to provide conservation gains within their borders [@lester2009; @edgar2014; @gerber2005].
+As conservation benefits accrue inside MPAs, MPAs can affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
+These dynamics result in MPAs affecting not only the organisms within their borders, but the broader connected population of affected species.
+Theory states that numerous factors influence how MPAs might affect fish populations.
+These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], the strength, timing, and location of density dependence [@burgess2014], the degree of enforcement [@edgar2014], how overfished the population would be without the MPA, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006a;@walters2000a;@white2011,;@white2010c;@ovando2016;@moffitt2013].
+The location and spacing of the MPAs within a network can also influence population-level impacts through network effects [@gaines2010; @costello2010a].
 
 What empirical evidence is there for these population-level effects of MPAs?
-Conservation outcomes of MPAs are generally evaluated by versions of "response ratios", commonly measured as the ratio of biomass densities of species inside relative to outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015].
-These studies have by and large found clear evidence that well enforced and sufficiently sized MPAs are associated with high response ratios.
-Several studies have also documented empirical evidence for the existence of adult or larval fish spillover affecting both fish abundance and fisheries productivity [e.g.
-@goni2010; @halpern2009;@kay2012; @stobart2009; @mcclanahan2000; @russ1996; @thompson2017;@baetscher2019;@pelc2009].
+Conservation outcomes of MPAs are generally evaluated by "response ratios", commonly measured as the ratio of biomass densities of species inside relative to outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015].
+These studies found clear evidence that well enforced and sufficiently sized MPAs are associated with high response ratios.
+Several studies have also documented empirical evidence for the existence of adult or larval fish spillover affecting fish abundance [@halpern2009;@kay2012; @stobart2009; @mcclanahan2000; @russ1996; @thompson2017;@baetscher2019;@pelc2009].
 MPAs have also caused alterations in fishing effort [@costa2013;@mason2012;@murawski2005;@mcdermott2019].
@@
 
-While spillover-based studies provide evidence that MPAs may have effects outside their borders, they are not sufficient evidence for the net effect of the MPAs.
-Detection of a biomass density or catch-per-unit-effort gradient decreasing with distance from an MPA border suggests that there may be spillover, but does not tell us the net effect of that spillover on the population or the fishery.
+Spillover-based studies provide evidence that MPAs may have effects outside their borders, but they are not sufficient evidence for the population effect of MPAs.
 Response ratios have shortcomings as measures of population-level effects as well.
@@
 In theory, control sites far enough away to negate both biological spillover and concentration by the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs is challenging.
-As a result, response ratios can be a highly imprecise and biased measure of the population-level conservation effect of an MPA network [@ferraro2018;@moffitt2013].
-As response ratios are our primary source of evidence for the conservation effects of MPAs, this means that our collective scientific understanding of the population-level impacts of MPAs is surprisingly limited.
+Response ratios can be a highly imprecise and biased measure of the population-level conservation effect of an MPA network [@ferraro2018;@moffitt2013].
+As response ratios are our primary source of evidence for the conservation effects of MPAs, this means that our collective understanding of the population-level impacts of MPAs is surprisingly limited.
 
@@
 The MPAs within the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) provide an opportunity to improve our understanding of population-level MPA effects.
-A network of protected areas covering approximately 20% of the Islands' waters was put in place in 2003 (see @osmond2010, @kirlin2013, @botsford2014, and @hilborn2012 for information on the creation of these MPAs).
+A network of protected areas covering approximately 20% of the Islands' waters was put in place in 2003 [@kirlin2013].
 The network has been used as a model case study in protected area design around the world (@botsford2014 and references therein).
@@
 
-## Methods
+# Methods {#methods .unnumbered}
 
 Our primary methods consist of a spatially-explicit bio-economic simulation model and a Bayesian difference-in-difference regression.
-We use the simulation model to provide provide theoretical grounding and context for our empirical results.
-The difference-in-difference regression provides our estimate of the population-level conservation effect of the Channel Islands MPAs.
+We use the simulation model to provide provide theoretical context for our empirical results.
+The regression provides our estimate of the population-level conservation effect of the Channel Islands MPAs.
 All analysis were conducted in R [@rcoreteam2019].
@@
 
-### Simulation Model
+## Simulation Model {#simmodel .unnumbered}
 
@@
 Readers can explore the functionality of the model using an online tool available at [danovando.shinyapps.io/simmpa/](https://danovando.shinyapps.io/simmpa/).
-Detailed methods can be found in the online supporting information.
 
@@
 We pair these data with randomly selected values governing the characteristics of the simulation (Table.\@ref(tab:sim-table)).
-These include parameters governing recruitment (variation, autocorrelation, and timing), movement (magnitude of larval and adult dispersal and density dependent movement), fleet dynamics (degree of fishing pressure, distribution of fishing effort), and MPA design (size and placement, habitat characteristics, larval source).
 The population begins at unfished equilibrium.
@@
 In each time step, we measure the difference in biomass densities between the scenario with and without the MPAs to calculate the population effect of the MPAs over time.
-Simulation results provide a library of plausible MPA effects for a range of biological and economic assumptions.
+Simulation results provide a library of plausible MPA effects for a range of biological and economic assumptions.One set of simulations is specifically designed to reflect the dynamics of the Channel Islands.
+We only include species of the same genus as those targeted by fishing in the Channel Islands, restrict fishing pressure to be mostly moderate to low, and cap the MPA size at 20% of the population's range (a likely upper end to the effective MPA size in question, @rassweiler2012).
 
-One set of simulations is specifically designed to reflect the dynamics of the Channel Islands.
-We only include species of the same genus as those targeted by fishing in the Channel Islands.
-The MPAs in this study cover 20% of the surface waters of the Channel Islands.
-Detailed dispersal studies are not available for all of the species covered by our study, but what information we have suggest that while adults of some of these species are likely to exhibit site fidelity [@love2011], larvae are likely distributed beyond the Channel Islands [@rassweiler2012].
-Therefore, if we assume that the population of the targeted species is at least equal to the extent of the Channel Islands, we can assume that at most the MPAs cover roughly 20% of the targeted populations' ranges.
-As such we cap the MPA size in the Channel Islands simulation to 20%.
-
-### Difference in Difference Regression
+## Difference in Difference Regression {#did .unnumbered}
 
@@
 The details of the monitoring program are described in @caselle2015.
+We define the population-level conservation effects of MPAs as the change in mean total biomass densities of targeted fin-fish both inside and outside of MPAs, relative to the mean total biomass densities of targeted fin-fish inside and outside of MPAs that would have occurred without the MPAs.
 
@@
 Targeted species in the Channel Islands include fin-fish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
-Each of these targeted species was the subject of prior bio-economic modeling related to the effects of MPAs in southern California [@costello2010a;@rassweiler2014].
-Our analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*), although our methods could similarly be applied to these species given data.
 Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*).
-We used a Bayesian difference-in-difference regression to estimate any difference in mean total biomass densities of fin-fish species targeted by fishing effort (i.e., those potentially affected by an MPA) and those species not targeted by fishing before and after MPA implementation [see @larsen2019 for an explanation of difference-in-difference estimators in ecological settings].
+We used a Bayesian difference-in-difference regression to estimate any difference in mean total biomass densities of fin-fish species targeted by fishing effort (i.e., those potentially affected by an MPA) and those species not targeted by fishing before and after MPA implementation.
 This identification strategy controls for unobserved environmental shocks to the system that are independent of the MPAs.
 Conditional on the assumptions of the model, this regression produces an estimate of the effect of the MPAs on the mean total biomass densities of targeted species throughout the Channel Islands.
+For example, consider an evenly distributed population that has 50% of its range protected by an MPA.
+If the MPA increases biomass densities inside its boundaries by 20%, and by 0% outside the reserve, the population effect of the MPA would be 10%.
 
-We define the population-level conservation effects of MPAs as the change in mean total biomass densities of targeted fin-fish both inside and outside of MPAs, relative to the mean total biomass densities of targeted fin-fish inside and outside of MPAs that would have occurred without the MPAs.
-For example, consider an evenly distributed population that has 50% of its range protected by an MPA.
-Suppose that the MPA increases biomass densities inside its boundaries by 20%, and by 0% outside the reserve.
-By our metric the population conservation effect of the MPA would be 10%.
-The regression amounts to estimating the pre-post MPA difference in the mean total biomass density of targeted species minus the same difference for non-targeted species in the Channel Islands.
+ <!-- Our analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*), although our methods could similarly be applied to these species given data.
+-->
+
+<!-- The regression amounts to estimating the pre-post MPA difference in the mean total biomass density of targeted species minus the same difference for non-targeted species in the Channel Islands.
+-->
 
@@
 $\beta_2MPA$ is the effect of the post-MPA period on the non-targeted species.
-The shape and scale parameters of the Gamma distribution are estimated as well.
-Regression details and diagnostic plots are available in the SI.
-The full table of covariates and their posterior means and 89% credible interval can be found in Table.\@ref(tab:coef-table).
+The shape and scale parameters of the Gamma distribution are estimated as well (Table.\@ref(tab:coef-table)).
 
@@
 
-## Results
-
-### Effects of the Channel Islands MPA Network
+# Results {#results .unnumbered}
 
 Our Channel Islands-specific simulation provides a set of outcomes that can feasibly bound our expectations for the empirical effects of the MPA network.
-The targeted species in our database span a range of life histories, but are largely made up of fishes in the perch and rockfish complexes, with a mean Von Bertalanffy growth coefficient of `r round(mean(used_data$k[used_data$targeted == 1]),2)`, and a median age at maturity of `r round(median(used_data$tm[used_data$targeted == 1]),0)` years.
-Although stock assessments are largely lacking for these species, what evidence we have suggests mostly moderate fishing pressure, with high fishing mortality for some species such as California sheephead (*Semicossyphus pulcher*) and blue rockfish (*Sebastes mystinus*).
-For many species, this likely means that biomass levels during the early 2000s were below target management levels [@alonzo2004;@dick2017].
-We integrated these network design features, life history traits, and exploitation histories into our simulation model, varying key unknowns such as the degree of pre-MPA depletion, species' movement rates, and fleet dynamics.
-While a wide range of MPA effects are plausible, from 0% to upwards of 150%, the median simulated effect size for this stylized version of the Channel Islands after over ten years of protection was roughly `r percent(median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected > 10]))` (Fig.\@ref(fig:mpa-effect)).
+The targeted species in our database span a range of life histories, but are largely made up of fishes in the perch and rockfish complexes, with a mean Von Bertalanffy growth coefficient of `r round(mean(used_data$k[used_data$targeted == 1]),2)`, and a median age at maturity of `r round(median(used_data$tm[used_data$targeted == 1]),0)` years, and a mix of exploitation histories, ranging from low to high (but mostly relatively low) [@alonzo2004;@dick2017].
+While a wide range of MPA effects are plausible, from 0% to upwards of 150%, the median simulated effect size for this stylized version of the Channel Islands after over ten years of protection was `r percent(median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected > 10]))` (Fig.\@ref(fig:mpa-effect)).
 
@@
 
-These response ratios cannot, however, be used as a definitive indicator of overall population effects of the MPAs.
+These response ratios cannot, however, be used as a definitive indicator of population effects of the MPAs.
 In the case of the Channel Islands MPAs, control sites are often located within a few kilometers of an MPA, suggesting that they are susceptible to both local biological spillover and by concentration of fishing effort excluded from the MPAs.
-Given these complications, the population-wide effects of the Channel Islands MPAs are difficult to determine with response ratios.
-
-To illustrate this problem, we paired our simulated Channel Islands response ratios to the estimated posterior probability distributions of the empirical Channel Island response ratios.
-For example, if the mean estimated response ratio in the year 2006 (year three after implementation) is one, we found all simulations that had response ratios near one after three years of protection, and then pulled out the "true" simulated population-level MPA effect from each of those simulations.
-This provides us with a distribution of simulated population-level MPA effects that could plausibly generate the types of response ratios actually observed in the Channel Islands.
-
+To illustrate this problem, we paired our simulated Channel Islands response ratios to the estimated posterior probability distributions of the empirical Channel Island response ratios by matching the patterns in the empirical response ratios.
 The response ratio trends we observe in the data could plausibly be produced by a wide range of population-level MPA effects (Fig.\@ref(fig:rr-plot)).
@@
 This can occur if for example fishing pressure is only moderate, adult movement is low, larval dispersal is high, and displaced fishing effort concentrates around the border of the MPAs.
+
+<!-- For example, if the mean estimated response ratio in the year 2006 (year three after implementation) is one, we found all simulations that had response ratios near one after three years of protection, and then pulled out the "true" simulated population-level MPA effect from each of those simulations.
+This provides us with a distribution of simulated population-level MPA effects that could plausibly generate the types of response ratios actually observed in the Channel Islands.
+-->
 
@@
 
-Our empirical difference-in-difference regression provides a new perspective on population-wide MPA effects.
-Consistent with MPA theory, over the first three years of implementation (2003-2006), the effects of the MPAs are unclear, with support for a small negative effect to a substantially positive effect, with much higher probability of a small positive effect (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`,
+Our empirical difference-in-difference regression provides a different perspective on population-wide MPA effects.
+Over the first three years of implementation (2003-2006), the effects of the MPAs are unclear, with support for a small negative effect to a substantially positive effect, with much higher probability of a small positive effect (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`,
 90\% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`
@@
 These empirical estimates are in line with the outcomes that our simulation model suggests are plausible.
-While this concordance between theory and empirics is not proof of the robustness of our results, it is an important line of evidence.
 However, in the subsequent years the trend reverses itself, and in 2015-2017 we once again see no clear effect of the MPAs (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`,
@@
 
-## Discussion {#discussion .unnumbered}
+# Discussion {#discussion .unnumbered}
 
@@
 The persistently high response ratios suggest that despite overall decreases in targeted biomass densities inside and outside MPAs relative to pre-MPA measurements, the MPAs may still be providing protection within their borders.
-But, as we have shown here these response ratios are not necessarily an indicator of broader population-level effects.
+But, these response ratios are not necessarily an indicator of broader population-level effects.
 The difference-in-difference strategy presents an alternative identification strategy.
@@
 However, these models generally simulate fleet dynamics through fishing mortality rates (e.g.
-concentration of fishing mortality @halpern2004).
-The implication of this is that total catches will scale with the size of the available population: if fishing effort moves outside the reserve, total catch will go down if the pool of fishable biomass available to the fleet is lower outside the reserve.
+concentration of fishing mortality, @halpern2004).
+The implication of this is that total catches will scale with the size of the available population.
 
@@
 
-Alternatively, though, fishers can pursue a  “constant-catch” rather than an effort-based strategy.
-Under the constant catch model, fishers have a catch objective, and exert as much (or little) effort as needed to achieve that objective.
-While a constant-catch greater than the maximum sustainable catch from the population is not possible over the long-term, over the short-term a constant-catch scenario is possible.
+Alternatively, fishers can pursue a  “constant-catch” strategy, where fishers have a catch objective and exert as much (or little) effort as needed to achieve that objective.
 Subsistence fisheries may use a constant-catch style policy over the short-term, as they seek to ensure that their food needs are met.
 Constant-catch dynamics might also occur in fisheries with constraining quotas that are not updated after the implementation of MPAs.
-Fishers pursuing constant catch under declining fish population size have to continually fish harder to achieve the same catch from a smaller part of the population, reducing the size structure of the population and subsequently causing net conservation loss under `r percent(1 - positive_outcomes$pp[positive_outcomes$fleet_model == "constant-catch"], accuracy = 1)` of our constant-catch simulations.
+Fishers pursuing constant catch in areas outside an MPA have to fish harder to achieve the same catch from a smaller part of the population, causing a population conservation loss under `r percent(1 - positive_outcomes$pp[positive_outcomes$fleet_model == "constant-catch"], accuracy = 1)` of our constant-catch simulations.
 This is an important and often overlooked possibility, especially as MPAs are increasingly implemented in quota-managed fisheries [@liu2018].
 
-While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (see SI Fig.S1) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
+While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (SI Fig.S1) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
 However, some of the most common targeted species in the Channel Islands, such as blue rockfish and copper rockfish, have seen dramatic increases in reported catches in the Santa Barbara region since the year 2010 (although other such as California sheephead have seen declines relative to pre-MPA catch level).
-To the extent that these increases in catches are due to external market forces, the disappearance of clear positive population-level MPA effects estimated by our model could be due to exogenous increases in fishing pressure overwhelming any population-level increases caused by the MPA network.
-Unfortunately, we lack a control that would allow us to estimate the MPA effects relative to a counterfactual in which the same market forces increase catches but no MPAs are in place.
+While we include catch histories in our regression, to the extent that these increases in catches are due to external market forces, the disappearance of clear positive population-level MPA effects estimated by our model could be due to exogenous increases in fishing pressure overwhelming any population-level increases caused by the MPA network.
 
-Another possibility for the observed decline in Channel Islands MPA effects is environmental disturbance.
-The Channel Islands region (and the entire West coast of the USA) experienced a dramatic ‘marine heatwave’ beginning in 2014 and persisting through 2016, resulting in part in extremely elevated water temperatures throughout the region [@gentemann2017].
+Another possibility for the estimated decline in Channel Islands MPA effects is environmental disturbance.
+The Channel Islands region experienced a dramatic ‘marine heatwave’ beginning in 2014 and persisting through 2016, resulting in part in extremely elevated water temperatures throughout the region [@gentemann2017].
 Many of the non-targeted species in the Channel Islands have warm thermal affinities and have increased in numbers since the heatwave [@freedman2019].
-The targeted group is made up mostly of fishes with cold-water affinities, such as members of the genus *Sebastes*.
-We hypothesize that the recent evidence for a decline in densities of targeted species is due to environmental conditions that disproportionately affected the targeted group.
+The targeted group is made up mostly of fishes with cold-water affinities.
+We hypothesize that the recent evidence for a decline in densities of targeted species is primarily due to environmental conditions that disproportionately affected the targeted group.
 The practical result of this is that in the presence of this marine heatwave the non-targeted species may no longer serve as an effective control for the evolution of biomass densities of targeted fin-fish in the absence of the MPAs, given the magnitude of the environmental shock relative to the size of the population-level MPA effect.
@@
 Our library of simulation results allows us to explore how MPA effects, and our ability to measure them, might vary with observation error and natural fish recruitment variation (including the possibility of recruitment regimes).
-Using a simple Bayesian difference-in-difference regression on these simulated data, we estimated the percent error between the posterior probability distribution of the estimated MPA effect from the regression and the true simulated MPA effect.
+Using a Bayesian difference-in-difference regression on these simulated data, we estimated the percent error between the posterior probability distribution of the estimated MPA effect from the regression and the true simulated MPA effect.
 
-While unbiased across simulations, the difference-in-difference model struggled severely when MPA effect sizes were less than 25% and the model was faced with both observation and process errors (Fig.\@ref(fig:val-plot)).
+The model struggled severely when MPA effect sizes were less than 25% and the model was faced with both observation and process errors (Fig.\@ref(fig:val-plot)).
 Even models fit to data generated from large effect sizes commonly mis-estimated the true MPA effect by 50% or more.
 Obtaining a mean absolute percent error (MAPE) of 25% or less across our simulated datasets required a population-level MPA effect of at least 30%.
-This is merely an illustrative exercise, omitting critical factors such as detection probability and sampling strategy.
+In the context of the Channel Islands, this finding suggests that we should not be surprised at our difficulty in precisely estimating the population-level effect of the MPAs.
+<!-- This is merely an illustrative exercise, omitting critical factors such as detection probability and sampling strategy.
 However, since nearly any omission which one can think of would make an MPA effect harder to detect, not easier, these results serve as a useful floor for the likely difficulty in estimating MPA effects.
-In the context of the Channel Islands, this finding unfortunately suggests that we should not be surprised at our difficulty in precisely estimating the population-level effect of the MPAs.
+-->
 
 Two of the most critical drivers of MPA effect size are the size of the MPA network and the degree of fishing pressure.
-Looking across these two variables, the MPA network must be large (25% or more of a species range) and the target species overfished (pre-MPA depletion greater than about 60%) to achieve an effect size with an MAPE of 25%.
+Looking across these two variables, based on our simulations the MPA network must be large (25% or more of a species range) and the target species overfished (pre-MPA depletion greater than about 60%) to achieve an effect size with a likely MAPE of 25% or less.
 Although recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, most MPA networks for near-shore commercial fin-fish are likely to cover areas more in line with the Channel Islands (20%) or smaller.
@@
 <!-- Once simulation models have been used to help design an MPA, how can users evaluate whether it is achieving their objectives?  -->
-How then should stakeholders evaluate the population-level effects of MPAs?
-Response ratios are commonly used as evidence for conservation outcomes of MPAs; @lester2009 and @halpern2002 present meta-analyses of hundreds of such studies.
-These results often find massively higher densities and biomass inside MPAs than outside [median biomass increases of  over 400%, @lester2009].
-But as suggested in @white2011, @ferraro2018, @larsen2019, @moffitt2013 and further demonstrated here, when MPAs affect nearby control sites used in response ratios through biological spillover or concentration of fishing effort, it is entirely possible to for MPA to produce massive response ratios while simultaneously having minimal effects on the entire population partially protected by the MPAs (and *vice versa*).
+@lester2009 found massively higher densities and biomass inside MPAs than outside [median biomass increases of  over 400%].
+But, when MPAs affect nearby control sites used in response ratios through biological spillover or concentration of fishing effort, it is entirely possible to for MPA to produce massive response ratios while simultaneously having minimal effects on the entire population partially protected by the MPAs (and *vice versa*).
 
-As @larsen2019 suggests, there are many potential alternatives for estimating the effects of MPAs that better account for the challenges of causal inference (though that may be more data-intensive).
-We applied one such approach here (a difference-in-difference estimator, see SI for alternative strategies such as synthetic controls, @abadie2010, that show similar results), and yet were still unable to reach robust conclusions as to the effect of MPAs on the total biomass density of targeted fin-fish in the Channel Islands, due to the likely small size of the true effect relative to the influence of environmental variability.
+As @larsen2019 suggests, there are many potential alternatives for estimating the population effects of MPAs that better account for the challenges of causal inference (though that may be more data-intensive).
+We applied one such approach here, and yet were still unable to reach robust conclusions as to the effect of MPAs on the total biomass density of targeted fin-fish in the Channel Islands, due to the likely small size of the true effect relative to the influence of environmental variability.
 
@@
 
-Stock assessment models [as described in @hilborn1992b] that account for the population within the MPAs [@field2006a] may be able to answer the management relevant question of whether fishing mortality rates and biomass levels from systems containing MPAs are in-line with management objectives [e.g.
+Stock assessment models that account for the population within the MPAs [@field2006a] may be able to answer the management relevant question of whether fishing mortality rates and biomass levels from systems containing MPAs are in-line with management objectives [e.g.
 @nickols2019].
@@
 MPAs are an important part of the marine resource management toolbox.
-As the number and size of global MPA networks increase, it is critical that we both set appropriate expectations for their outcomes, and plan how we will monitor the performance of these protected areas over time.
+As the number and size of global MPA networks increase, we must set appropriate expectations for their outcomes and design effective monitoring programs.
 While the history of MPA science has made important strides in helping us understand the dynamics of protected areas, the future of MPA science must directly tackle the challenge of evaluating the performance of these MPAs at the regional scale.
@@
 Commonly employed metrics such as response ratios may be applicable in some circumstances, but are vulnerable to inaccuracy or misuse as metrics of population-level effects.
-Dependence on unreliable estimators of MPA effects may lead to stakeholders incorrectly attributing negative environmental shocks as MPA failures, or interpreting data arising from scorched-earth fishing outside MPAs as a conservation success.
-Both of these scenarios would hinder the ability of MPAs to serve as effective marine resource management tools at scale.
 Bio-economic modeling can help frame community expectations, reducing the potential for a reduction in support if unrealistic conservation or fishery expectations are not realized.
@@
 
-# Literature Cited
+<!-- Dependence on unreliable estimators of MPA effects may lead to stakeholders incorrectly attributing negative environmental shocks as MPA failures, or interpreting data arising from scorched-earth fishing outside MPAs as a conservation success.
+Both of these scenarios would hinder the ability of MPAs to serve as effective marine resource management tools at scale.
+-->
+
+# Literature Cited {#lit .unnumbered}
 
@@
 
-# Tables
+# Tables {#tables .unnumbered}
 
@@
     digits = 2,
-    caption = "Posterior means and credible interval for key model coefficients.
+    caption = glue::glue("Posterior means and credible interval for key model coefficients.
 Conditional on the assumptions of the model, the targeted:year_bins coefficients reflect the effect of the MPAs on mean total biomass of targeted finfish in the Channel Islands.
-targeted indicates whether the species in question is targeted, 'var' refers to a covariate, where tex = technican experience, surge is mean reported surge, kelp is mean reported kelp cover, lag catch is the total reported catch for targeted species in the previous year, temp is the mean temperature at a site, temp_dev is the deviation of a temperature observation from the mean temperatuer for that island in question",
+targeted indicates whether the species in question is targeted, tex = technican experience, surge is mean reported surge, kelp is mean reported kelp cover, lag catch is the total reported catch for targeted species in the previous year, temp is the mean temperature at a site, temp_dev is the deviation of a temperature observation from the mean temperatuer for that island in question.
+Raw data contained {nrow(used_data)} individual observations, total number of targeted and non-targeted mean mean total biomass densities by site and year used to fit model was {nrow(did_fit$data)}"),
     format = "pipe")
@@
 
-# Figures
+# Figures {#figures .unnumbered}
 
@@
 
-```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="90% Posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis) in blue.
-Simulated population-level (pop.) effect on biomass densities matched to empirical response ratios in red.
+```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="90% Posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis) in dark grey.
+Simulated population-level (pop.) effect on biomass densities matched to empirical response ratios in light grey.
 For response ratios, a value of zero indicates that biomass densities of targeted species are identical inside and outside MPAs, a value of one that biomass densities of targeted species are 100% greater inside MPAs relative to outside.
@@
   theme_1col + 
-  theme(axis.title.x = element_textbox(size = 10))
+  theme(axis.title.x = element_textbox(size = 10)) + 
+  geom_vline(aes(xintercept = 0), color = "black", linetype = 2)
 
@@
 ```{r val-plot, include = TRUE,fig.width = width_1col, fig.asp=asp_1col, fig.cap = "Distribution of percent error in posterior estimates of population-level MPA effect (y-axis) plotted against true simulated MPA effect (x-axis).
-Color shows concentration of simulations.
+Shading shows concentration of simulations.
 Black line shows mean absolute percent error (MAPE) as a function of true simulated MPA effect."}
@@
   theme_1col + 
-  theme(legend.position = "top") + 
-  scale_fill_viridis(option = "C",
-                     name = "Sims(#)",
-                     guide = hgc <- guide_colorbar(frame.colour = "black",
-                                                   ticks.colour = "black",
-                                                   barwidth = ggplot2::unit(5,"lines")))
+  theme(legend.position = "top") #+ 
+  # scale_fill_viridis(option = "C",
+  #                    name = "Sims(#)",
+  #                    guide = hgc <- guide_colorbar(frame.colour = "black",
+  #                                                  ticks.colour = "black",
+  #                                                  barwidth = ggplot2::unit(5,"lines")))
 

### 2020-07-18 documents/pop-effects-of-mpas-front-matter.Rmd
version submitted to conservation biology

@@
 ---
-title: Assessing the Population-level Conservation Effects of Marine Protected Areas
+title: "Assessing the Population-level Conservation Effects of Marine Protected Areas"
 subtitle: Cover Page
@@
   bookdown::word_document2:
-    template: uw.docx
+    reference_docx: uw.docx
   pdf_document: default
+  word_document: default
 ---
-
-Submit your cover page as a separate document.
-It should NOT be part of the manuscript itself.
-The cover page must include the title of the paper; an article impact statement (see below); a running head (short title of 20 or fewer characters); a list of 5-8 keywords (do not repeat words in the title); word count (all text from the first word of the Abstract through the last word of the Literature Cited but not including table or figure legends or the body of tables); authors' complete mailing addresses (including street addresses and postal codes) at the time the work was conducted and present addresses if different; name and complete address (including email) of the person to whom correspondence should be sent; and text of your Acknowledgments section.
 
@@
 ## Keywords:
-  - Marine Protected Area Networks
-  - Marine Conservation
-  - Program Evaluation
-  - Causal Inference
-  - Bio-economic Modeling
+
+Marine Protected Area Networks
+
+Marine Conservation
+
+Program Evaluation
+
+Causal Inference
+
+Bio-economic Modeling
 
@@
 
-7166
+6000
 
@@
 
-J
+Jennifer E. Caselle
+UCSB Marine Science Institute Bldg 520 Rm 4002 Fl 4L 
+Santa Barbara, CA 93106-6150 USA
+caselle@msi.ucsb.edu
+
+Christopher Costello
+Bren Hall 4410 
+University of California, Santa Barbara
+Santa Barbara, CA 93106-6150 USA
+costello@bren.ucsb.edu
+
+Olivier Deschenes
+2127 North Hall
+University of California, Santa Barbara
+Santa Barbara, CA 93106-6150 USA
+olivier@econ.ucsb.edu
+
+Steven D. Gaines
+Bren Hall 2410 
+University of California, Santa Barbara
+Santa Barbara, CA 93106-6150 USA
+gaines@bren.ucsb.edu
+
+Ray Hilborn
+FSH 352B
+University of Washington
+1122 NE Boat Street
+Seattle, WA 98195-5020 USA
+
+Owen Liu
+Bren Hall 4330 
+University of California, Santa Barbara
+Santa Barbara, CA 93106-6150 United States
+owen.r.liu@gmail.com
 
 ## Corresponding Author
+
+Daniel Ovando
+Fishery Science Building
+University of Washington
+1122 NE Boat Street, Seattle, WA 98195-5020 USA
+danovan@uw.edu
+phone: (805) 284 6864
 
@@
 Empirical data collection was funded primarily by the David and Lucille Packard Foundation in support of the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) with additional funding from the California Ocean Protection Council and California SeaGrant.
-This study would not be possible without the work provided by PISCO divers over the years, with special thanks to to Kathyrn Davis Koehn and Avrey Parsons-Field.
-We also thank Cody Szuwalski, Julia Lawson, André Punt, and two anonymous reviewers for helpful comments and technical support.
+This study would not be possible without the work provided by PISCO divers over the years, with special thanks to to K.
+Davis Koehn and A.
+Parsons-Field.
+We also thank C.
+Szuwalski, J.
+Lawson, and A.
+Punt for helpful comments and technical support.

### 2020-07-18 make-pop-effects-of-mpas.R

@@ -3973,7 +3973,8 @@ total_trend_plot <- classcode_trends %>%
   ggplot() + 
   geom_line(aes(year, scaled_mbd, group = interaction(targeted, classcode), color = targeted == 0),alpha = 0.25) + 
   geom_line(data = targeted_trends, aes(year, scaled_mbd, color = targeted == 0), size = 2) + 
-  scale_color_npg(labels = c("Targeted", 'Non-Targeted'), name = "") + 
+  # scale_color_npg(labels = c("Targeted", 'Non-Targeted'), name = "") + 
+  scale_colour_grey(labels = c("Targeted", 'Non-Targeted'), name = "") + 
   scale_x_continuous(name = '') + 
   scale_y_continuous(name = "Scaled Mean Biomass Density") +
   theme(legend.position = "top",
@@ -3987,7 +3988,8 @@ total_trend_plot <- classcode_trends %>%
   mpa_trend_plot <- targeted_trends_by_mpa %>% 
     ggplot() + 
     geom_line( aes(year, scaled_mbd, color = targeted == 0), size = 2, show.legend = FALSE) + 
-    scale_color_npg(labels = c("Targeted", 'Non-Targeted'), name = "") + 
+    # scale_color_npg(labels = c("Targeted", 'Non-Targeted'), name = "") + 
+    scale_colour_grey(labels = c("Targeted", 'Non-Targeted'), name = "") + 
     scale_x_continuous(name = 'Year') + 
     scale_y_continuous(name = "") +
     facet_wrap(~eventual_mpa, labeller = labeller(eventual_mpa = targlab)) 
@@ -4072,7 +4074,7 @@ implications %>%
 # select only sebastes, perches, and wrasses, MPA size <= 25%, F/M <= 1.5
 
 
-title = "<span style = 'color:red;'> Paired Simulated Pop. Effect / <span style = 'color:blue;'>Empirical Response Ratio</span>"
+title = "<span style = 'color:grey;'> Paired Simulated Pop. Effect / <span style = 'color:black;'>Empirical Response Ratio</span>"
 
 response_ratio_plot <-   targ_rr %>%
   select(year, response_ratio) %>%
@@ -4097,7 +4099,7 @@ response_ratio_plot <-   targ_rr %>%
   scale_x_continuous(name = title, limits = c(NA, 2)) +
   scale_y_continuous(name = "Year", labels = seq(2003,2017, by = 3), breaks = seq(2003,2017, by = 3)) +
[…]

### 2020-07-18 documents/ovando-et-al-pop-effects-of-mpas-si.Rmd
Cons bio (#4)
* version submitted to conservation biology

Co-authored-by: Owen.Liu <1571646680@mil>

@@
 subtitle: "Assessing the Population-level Conservation Effects of Marine Protected Areas"
-author: 
-  - Daniel Ovando
-  - Jennifer E. Caselle
-  - Christopher Costello
-  - Olivier Deschenes
-  - Steven D. Gaines
-  - Ray Hilborn
-  - Owen Liu
 output: 

### 2020-07-18 make-pop-effects-of-mpas.R

@@ -76,7 +76,7 @@ sim_years <- 50
 
 num_patches <- 50
 
-n_cores <- 1
+n_cores <- 6
 
 # prepare run -------------------------------------------------------------
 
@@ -1813,9 +1813,9 @@ if (simulate_channel_islands == TRUE){
   
   num_patches <- 50
   
-  run_experiments <- FALSE
+  run_experiments <- TRUE
   
-  create_grid <- FALSE
+  create_grid <- TRUE
   
   save_experiment <- TRUE
   
@@ -1883,7 +1883,7 @@ if (simulate_channel_islands == TRUE){
             larval_movement = sample(0:round(.25 * num_patches), samps, replace = T),
             density_movement_modifier = sample(c(0.25, 1), samps, replace = T),
             density_dependence_form = sample(1:3, samps, replace = T),
-            mpa_size = runif(samps, min = .05, max = .4),
+            mpa_size = runif(samps, min = .05, max = .2),
             f_v_m = runif(samps, min = 0.5, max = 1.5),
             fleet_model = sample(
               c("open-access", "constant-effort"),
@@ -2369,6 +2369,8 @@ if (process_results == TRUE){
   
   processed_grid$adult_movement <- (2 * processed_grid$adult_movement) / num_patches
   
+
+
   
 #
 #   zissou_theme <-
@@ -2404,7 +2406,6 @@ if (process_results == TRUE){
 
   processed_grid <- processed_grid %>%
     filter(!experiment %in% unique(bad_sims$experiment))
-
   
   write_rds(processed_grid, file.path(run_dir,"filtered_processed_grid.rds"))
 
@@ -2466,7 +2467,7 @@ if (process_results == TRUE){
     mutate(pop_effect = pmin(1,(`with-mpa` - `no-mpa`) / b0))
 
   write_rds(outcomes, file.path(run_dir,"outcomes.rds"))
-
+  
   density_ratios <- processed_grid %>%
     select(-absolute_mpa_effect, fishery_effect) %>% 
     mutate(mpa_effect = map(mpa_effect,~.x %>% select(-year,-mpa_size))) %>% 
@@ -3026,7 +3027,7 @@ if (process_results == TRUE){
     scale_y_continuous(labels = percent, expand = expansion(0,0)) +
     scale_x_continuous(labels = percent,expand = expansion(0,0)) +
     theme(legend.position = "top",
-          legend.text = element_text(size = 6))
+          legend.text = element_text(size = 8))
[…]

### 2020-11-19 documents/ovando-et-al-pop-effects-of-mpas-si.Rmd
OK, path sketched to respond to reviewer comments

@@
 
-```
-
-### Transect Level Difference-in-difference Model
-
-Our base model aggregated the data to the level of total biomass densities of targeted and non-targeted species.
-This serves to average out some of the sampling noise from the individual transect level, and we would argue is the appropriate level of estimation for many key policy questions.
-However, this approach does oversimplify the data collection process, treating the total biomass densities at the year/site level as estimated without error.
-
-As an added robustness test, we include version of the model in which we operate at the level of individual transect observations.
-The raw data are estimated length compositions by fish species along a survey transect at a site.
-Lengths are converted to biomass per allometric relationships supplied by PISCO and supplemented by the `FishLife` [@thorson2017c] package in R where needed.
-We performed some minimal data filtering to reduce noise in the data.
-We only include species that were observed at least twice in each year of the dataset (2000-2017) somewhere in the core Channel Islands (Anacapa, Santa Cruz, Santa Rosa, San Miguel).
-While some data are available from 1999, per consultation with PISCO we omit those data due to changes in survey protocols.
-We assign species to targeted and non-targeted groups per the PISCO classifications.
-This filtering process results in `r filter_summary$no` non-targeted species and `r filter_summary$yes` targeted species remaining in the analysis.
-
-The simplified explanation of the estimation is a hierarchical model in which we first standardize the observed biomass densities into an abundance index of each species over time.
-The abundance indices in each year are assumed to be log-normally distributed with means and standard deviations for the targeted and non-targeted groups, giving an estimate of the mean densities of targeted and non-targeted species over time.
-We then calculate the difference between mean density of targeted species and the mean density of non-targeted species in each year.
-
-It is important to note that this model asks a slightly different question than our base model.
-The base model estimates the effect on total mean biomass of targeted finfish at the site level.
-This model estimates the effect on the mean biomass density of targeted finfish across all sites.
-While this model attempts to account for finer-scale observation error than our aggregated model, it also introduces a large amount of sampling variation that cannot be explained by the model.
-
-\newpage
-```{r cartoon, fig.cap= "Cartoon illustration of the hierarchichal difference-in-difference estimator", fig.width=8}
-knitr::include_graphics(here::here("documents","figs","estimation-cartoon.png"))
-```
-
-\pagebreak
-
-The first stage of the regression is a log-normal delta model.
-The model estimates two regressions, the first is a binomial generalized linear model (GLM) with a logit link estimating the probability of observing a given fish species at a observation *i* (transect at time *t*).
-The probability that a given species was observed *o* at a given observation is distributed
-
-\begin{equation}
-o_{s,i} \sim binomial(\frac{1}{1 +e^{-\beta^{o}{X}}})
-  (\#eq:binom)
-\end{equation}
-
- where $\beta^{o}$ are the estimated coefficients for the observation model and *X* is a matrix of covariates that include random effects for each year in the data (2000 to 2017).
-
-The expected density *d* of positive observations is modeled per a log-normal distribution
-
- \begin{equation}
-  log(d_{s,i}) \sim normal(\beta^{d}X, \sigma_s)
-   (\#eq:logd)
- \end{equation}
-
-where $\beta^{d}$ are the estimated coefficients for the expected density model and *X* is the same matrix of covariates as used in the observation portion of the model and $\sigma_s$ allows for each species *s* to have different standard deviations.
-
-Our covariate matrix *X* contains both fixed and random effects.
-Fixed effects include the depth level of the transect, the sampling site, the month of the observation, the estimated surge at the transect, visibility, the depth of the transect, and the experience (and experience squared) of the diver conducting the transect.
-We classify each species into one of two clusters based on the mean longitude the species was encountered at, breaking the species into two groups: those primarily found in the western end of the Channel Islands those found more in the eastern end.
-We then estimate random effects for each island for each cluster
-
-\begin{equation}
-\beta_{island,cluster} \sim normal(0,\sigma_{cluster})
-  (\#eq:island)
-\end{equation}
-
-This allows the mean effect of each island to differ for each cluster, e.g.
-allowing San Miguel, the easternmost island, to have a higher mean density for eastern species than for more western species (if the data suggest it).
-
-The second critical component of the covariate matrix *X* are random effects for each year for each species
-
-\begin{equation}
-  \beta_{year,species} \sim normal(0,\sigma_{species})
-  (\#eq:species)
-\end{equation}
-
-These $\beta_{year,species}$ represent our "standardized" estimate of observed abundance of each species in each time step, controlling for the included covariates.
-
-However, we still need to account for changes in the probability of detection over time.
-For that, we create a standard matrix of with rows equal to the number of years and columns corresponding to each of the columns in *X*, holding everything fixed at mean (or most frequently observed level for factors) levels for all variables in X except for the year and species interaction indices.
-Calling this standardized matrix $X^{standard}$, the probability of observing a given species in year *y* is then
-
-\begin{equation}
-p_{s,y} = (\frac{1}{1 +e^{-\beta^{o}{X^{standard}}}})
-  (\#eq:pdetect)
-\end{equation}
-
-In the same manner as described by @punt2000, The standardized index of abundance for species *s* in year *y* then is
-
-\begin{equation}
-I_{species,year} = p_{species,year}e^{\beta_{species,year}}
-  (\#eq:index)
-\end{equation}
-
-The next phase of the model requires us to estimate the mean abundance of targeted and non-targeted species over time.
-The concept here is that each $I_{species,year}$ can be modeled by a regression that contains random effects for each year for targeted and non-targeted fishes, the assumption then being that there is a mean density for targeted and non-target species, and $I_{species,year}$ represent deviations from that mean.
-
-\begin{equation}
-log(I_{species,year}) \sim normal(\beta^{effect}X^{effect}, \sigma_I)
-  (\#eq:indexprior)
-\end{equation}
-
-$X^{effect}$ contains both fixed and random effects. The fixed effects include an intercept and the temperature deviation for a given species in a year, where temperature deviation is
-
-\begin{equation}
-t_{s,y} = (t^{pref}_{s} -  \bar{t_{y}})^2
-  (\#eq:tempdev)
-\end{equation}
-
-where $t^{pref}$ is the preferred temperature for species *s* (drawn form `FishLife`, @thorson2017c), and $\bar{t_{y}}$ is the mean temperature encountered by that species in year *y*.
-We also include as variables in the model the mean kelp cover experienced by a given species in a given year, as well as the total fishery catches reported in the previous year for that species in the Santa Barbara region [drawn from the California Department of Fish and Wildlife database].
-We also include random intercepts for each species in $X^{effect}$.
-The most important random effects are year effects for targeted and non-targeted species
-
-\begin{equation}
-\beta_{year,targeted} \sim normal(0,\sigma_{targeted})
-  (\#eq:betatarg)
-\end{equation}
-
-$\beta_{year,targeted}$ is the mean log density of targeted species in year *y*, controlling for included covariates.
-Therefore, the final step in the model, the divergence in the standardized abundance trends of targeted and non-targeted species is
-
-\begin{equation}
-divergence_{year} =  \beta_{year,targeted = 1} - \beta_{year,targeted = 0}
-  (\#eq:mpaeffect)
-\end{equation}
-
-The model is fit in TMB to integrate the uncertainty across all levels of the model, with standard errors for each coefficient in the model estimated through the Laplace approximation.
-
-```{r tmb-fit, fig.cap="Difference-in-difference results from mixed effect model fit to the individual transect observation level of the data."}
-tmb_runs <- tmb_runs %>%
-  mutate(tmb_fit = map(tmb_fit,"result")) %>%
-  mutate(processed_fits = map(tmb_fit, process_fits)) %>%
-  mutate(did_plot = map(processed_fits, "did_plot"))
-
-tmb_runs$did_plot[[1]] + 
-scale_y_continuous(name = "Estimated MPA Effect")
-```
-
-A complete table of estimated coefficients can be seen in Table.S\@ref(tab:betatable).
-
-<!-- Stock status estimates are not available for many of the targeted species in the database.
-The California Department of Fish and Wildlife compiles status estimates for many nearshore finfish (https://www.wildlife.ca.gov/Conservation/Marine/Groundfish/Nearshore-Finfish), including recent stock assessments.
-For species not listed here we searched for available information on recent stock status from relevant stock assessments.
--->
-
-```{r betatable, results="asis"}
-
-fit_report <- tmb_runs$tmb_fit[[1]]$zissou_report
-
-observed_seen <- tmb_runs$tmb_fit[[1]]$zissou_data$log_density
-
-observed_seeing <- tmb_runs$tmb_fit[[1]]$zissou_data$any_seen
-
-seen_diagnostics <- data_frame(observed = observed_seen, predicted = as.numeric(fit_report$log_density_hat)) %>% 
-  mutate(residuals = predicted - observed)
-
-seeing_diagnostics <- data_frame(observed = observed_seeing, predicted = as.numeric(fit_report$prob_seeing))
-
-betas <- tmb_runs$processed_fits[[1]]$betas
-
-table <- betas %>% select(-std_error,-group)
-
-knitr::kable(table,
-             digits = 2,
-             caption = "Complete table of estimated coefficients from hierarchical difference-in-difference model",
-             longtable = TRUE,
-             booktabs = TRUE,
-             col.names = colnames(table)) %>% 
-  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
-
-```
-
-Figures S\@ref(fig:fe-plot):S\@ref(fig:targ-plot) present estimated effects for covariates included in the model, along with the raw estimated mean trends of the targeted and non-targeted species (while the difference between these trends is presented in our main results).
-
-\newpage
-
-```{r fe-plot, fig.cap = "Estimated coefficients for non-spatial fixed effects in observation model (seeing) and observed model (seen)"}
-
- non_nested_beta_plot <- betas %>%
-   filter(str_detect(group, "non_nested") & !str_detect(variable, "site_side")) %>%
-  mutate(group = str_remove_all(group,"(?=\\.).*" )) %>% 
-   ggplot() +
-   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
-   geom_pointrange(aes(x =variable,
-                       y = estimate,
-                       ymin = lower,
-                       ymax = upper)) +
-   facet_wrap(~group, labeller=labeller(group=c(
-     seeing_non_nested_betas="Observation Model\n(seeing)",
-     seen_non_nested_betas = "Observed Model\n(seen)"
-   ))) +
-   coord_flip() +
-   theme(axis.text.y = element_text(size = 10))
-
-non_nested_beta_plot
-
-```
-
-\pagebreak
-
-```{r site-plot, fig.cap = "Estimated coefficients for spatial random effects in observation model (seeing) and observed model (seen)", fig.height=10}
-
- non_nested_site_beta_plot <- betas %>%
-   filter(str_detect(group, "non_nested") & str_detect(variable, "site_side")) %>%
-    mutate(group = str_remove_all(group,"(?=\\.).*" )) %>% 
-  mutate(variable = tolower(variable)) %>% 
-  mutate(variable = str_replace(variable, "site_side","")) %>% 
-  mutate(island = str_replace(str_split(variable, '_', simplify = TRUE)[,1],'-','')
- ) %>% 
-   ggplot() +
-   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
-   geom_pointrange(aes(x = variable,
-                       y = estimate,
-                       ymin = lower,
-                       ymax = upper)) +
-   facet_grid(island~group, scales = "free",labeller=labeller(group=c(
-     seeing_non_nested_betas="Observation Model\n(seeing)",
-     seen_non_nested_betas = "Observed Model\n(seen)"),
-     island=c(anacapa="ANA",sci="SCI",smi="SMI",sri="SRI"))) +
-   coord_flip() +
-   theme(axis.text.y = element_text(size = 10)) + 
-  theme_minimal()
-
-non_nested_site_beta_plot + 
-  theme(axis.text.y = element_text(size = 6))
-
-```
-
-\pagebreak
-
-```{r region-plot, fig.cap="Random effects for region by cluster", eval = FALSE}
-
-  region_cluster_plots <- betas %>%
-   filter(str_detect(group, "region_cluster_betas")) %>%
-    mutate(group = str_remove_all(group,"(?=\\.).*" )) %>% 
-   mutate(cluster = str_replace_all(variable,"\\D","")) %>%
-   mutate(region = str_split(variable,'-', simplify = T)[,3]) %>%
-   ggplot() +
-    geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
-    geom_pointrange(aes(x = region,
-                        y = estimate,
-                        ymin = lower,
-                        ymax = upper, color = cluster)) +
-    facet_wrap(~group,labeller=labeller(group=c(
-     seeing_region_cluster_betas="Observation Model\n(seeing)",
-     seen_region_cluster_betas = "Observed Model\n(seen)")))
-
-region_cluster_plots
-
-```
-
-\pagebreak
-
-```{r targ-plot, fig.cap = "Trends in standardized mean abundance of targeted and non-targeted species"}
-
-report <- tmb_runs$tmb_fit[[1]]
-
-targeted_trends <- report$zissou_estimates %>% 
-  filter(str_detect(variable, "targeted_did")) %>% 
-  mutate(variable = str_remove_all(variable,"(?=\\.).*" )) %>% 
-  group_by(variable) %>% 
-  mutate(year =report$did_data$year %>% unique(),
-         labvar=ifelse(variable=="nontargeted_did_betas","Nontargeted","Targeted")) %>% 
-  ungroup() %>% 
-  ggplot() + 
-  geom_line(aes(year, estimate, color = labvar),alpha = 0.5) +
-  geom_pointrange(aes(year, estimate, ymin = lower, ymax = upper, color = labvar)) + 
-  scale_color_d3(name = "") +
-  theme(legend.position = c(0.8,0.9))
-
-targeted_trends
-
-```
-
-\pagebreak
-
-#### Transect Level Regression Diagnostics
-
-We include visual diagnostics of our estimation model. All coefficients passed convergence criteria for TMB.
-
-Looking first at the predictions of the model for the positive observations in the data (i.e.
-using the full model to predict biomass densities, and then comparing those predictions to cases where some positive biomass densities were observed), the model diagnostics show no clear problems.
-The $R^2$ of the model is 0.43.
-Residuals do not exhibit trends, though some grouping the residuals is evident.
-The quantile-quantile plot suggest that on the assumption of log-normal errors on the observed densities is reasonable, though the model appears to have some slight problems estimating the highest observed densities (Fig.S\@ref(fig:obs-v-pred-plot)).
-
-```{r, eval = FALSE}
-
-a = tmb_runs$tmb_fit[[1]]$zissou_data$x_seen_non_nested %>% 
-  as_data_frame() %>% 
-  select(contains("mean_")) %>% 
-  mutate(resid = seen_diagnostics$residuals) %>% 
-  gather(variable, value, -resid) %>% 
-  ggplot(aes(value, resid)) + 
-  geom_point() + 
-  facet_wrap(~variable)
-
-a
-
-```
-
-```{r obs-v-pred-plot, fig.cap="High level diagnostics for observed compontent of Delta-GLM: Observed vs predicted log densities (A), predicted log density vs residuals (B), and a normal qq-plot of the residuals (C)"}
-
-obs_v_pred_plot <- seen_diagnostics %>% 
-  ggplot(aes(observed, predicted)) + 
-  geom_point(alpha = 0.5) + 
-  geom_abline(aes(intercept = 0, slope = 1), linetype = 2, color = "red") + 
-  labs(caption = glue::glue("R^2 is {round(yardstick::rsq(data = seen_diagnostics, truth = observed, estimate = predicted)$.estimate,2)}"), title= "A") + 
-  theme(plot.margin = unit(rep(.1,4), units = "lines"))
-
-fit_v_resid_plot <- seen_diagnostics %>% 
-  ggplot(aes(predicted, residuals)) + 
-  geom_point() + 
-  geom_hline(aes(yintercept = 0), linetype = 2, color = "red") + 
-  labs(title = "B") +
-    theme(plot.margin = unit(rep(.1,4), units = "lines"))
-
-qq_plot <- ggplot(seen_diagnostics, aes(sample = residuals)) +
-  stat_qq() +
-  stat_qq_line() + 
-  labs(title = "C") + 
-    theme(plot.margin = unit(rep(.1,4), units = "lines"))
-
-obs_v_pred_plot/(fit_v_resid_plot + qq_plot)
-
-```
-
-In order to evaluate the ability of the model to estimate positive observations, we can compare the the binned predicted probability of a positive observation to the proportion of observations in that bin that recorded positive observations.
-If our model is working well, we would expect a group of fisheries that our model estimates on average should have a 50% probability of a positive observation, then we should expect about 50% of those observations to have positive observations.
-This is indeed what we see from the model (Fig.S\@ref(fig:obs-prob-plot)).
-
-```{r obs-prob-plot, fig.cap = "Binned mean predicted probability of detection provided by the first stage of the hurdle model vs observed proportion of positive detections"}
-
-seeing_diag_plot <- seeing_diagnostics %>% 
-  ungroup() %>% 
-  mutate(binned_prob = ntile(predicted, 20)) %>% 
-  arrange(binned_prob) %>% 
-  group_by(binned_prob) %>% 
-  summarise(mean_prob_seen = mean(predicted),
-            percent_observed = mean(observed))  %>% 
-  ggplot() +
-  geom_point(aes(mean_prob_seen, percent_observed)) + 
-  geom_abline(aes(intercept = 0, slope = 1), color = "red", linetype = 2) + 
-  labs(x = "Binned Predicted Probability of Detection", y = "Proportion of Positive Observations")
-
-seeing_diag_plot
-```
-
-We can also examine the receiver-operator-curve (ROC) to assess the performance of the observation component of the model.
-The area under the curve (AUC) value for the model is 0.84 (on a scale of 0.5 to 1), indicating the model is an overall good predictor of whether or not a given observation of biomass densities will be positive or not.
-
-```{r roc-plot, fig.cap = "Receiver operating characteristic curve of predictions of positive biomass densities"}
-library(plotROC)
-rocplot <- ggplot(seeing_diagnostics, aes(m = predicted, d = observed))+ geom_roc(n.cuts=20,labels=FALSE) + 
-  geom_abline(aes(slope = 1, intercept = 0))
-
-roc_plot <- rocplot + style_roc()
-
-roc_plot
-```
-
-#### Standardized Abundance Indices
-
-Overall most species showed consistent trends in biomass densities across the different islands at which they have been observed (Fig.S\@ref(fig:reg-pop)).
-
-```{r reg-pop, fig.cap = "Mean density by island by year for each fish species included in the analysis"}
-
-fitted_data <- used_data
-
-reg_pop_plot <- fitted_data %>% 
-  group_by(region, year, classcode) %>% 
-  summarise(mean_density = mean(exp(log_density))) %>% 
-  mutate(log_mean_density = log(mean_density)) %>% 
-  ungroup() %>% 
-  group_by(classcode, region) %>% 
-  mutate(cs_mean_density = (mean_density - mean(mean_density)) / sd(mean_density)) %>% 
-  ungroup() %>% 
-  ggplot(aes(year, cs_mean_density, color = region)) + 
-  scale_color_npg(name="Region")+
-  geom_line() + 
-  labs(y="Centered and Scaled Density",x="Year")+
-  facet_wrap(~classcode, scales = "free_y") + 
-  theme_minimal()+
-  theme(axis.text=element_text(size=6))
-
-reg_pop_plot
-
-```
-
-The standardized indices of abundance generally did not very substantially from the raw mean densities by species over time.
-However, for some species, such as blue rockfish, the standardized abundance index suggests much higher biomass densities in the pre-MPA period than those reported in the raw data.
-We suspect this is largely a function of changes in sampling sites over time, that the standardization is better able to account for (Fig.S\@ref(fig:raw-v-stand-plot)).
-
-```{r raw-v-stand-plot, fig.cap = "Raw (points) and standardized (lines) indices of abundance for each of the fishes included in the analysis"}
-
-life_history <- fitted_data %>%
-  select(classcode) %>%
-  unique() %>%
-  mutate(numeric_classcode = as.numeric(as.factor(classcode)))
-
-raw <- fitted_data %>%
-  group_by(year, classcode, targeted) %>%
-  summarise(md = mean(exp(log_density)),
-            mt = mean(temp_deviation)) %>%
-  group_by(classcode) %>%
-  mutate(smd = (md - mean(md))/(sd(md))) %>%
-  ungroup() %>%
-  left_join(life_history, by = "classcode")
-
-years <- report$did_data$year %>% unique()
-
-abundance_trends <-
-  data_frame(
-    abundance_hat = fit_report$abundance_hat,
-    classcode = rep(1:n_distinct(fitted_data$classcode), each = length(years))
-  ) %>%
-  mutate(log_abundance_hat = log(abundance_hat)) %>%
-  group_by(classcode) %>%
-  mutate(year = years) %>%
-  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
-  ungroup() %>%
-  rename(numeric_classcode = classcode) %>%
-  left_join(life_history, by = "numeric_classcode")
-
-# 
-# abundance_trends %>%
-#   ggplot(aes(year, log(abundance_hat), color = factor(classcode))) +
-#   geom_line(show.legend = F) +
-#   geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
-#   facet_wrap(~classcode) +
-#   theme_minimal()
-
-raw_v_std_plot <- abundance_trends %>%
-  filter(year %in% years) %>% 
-  ggplot(aes(year, scaled_abundance_hat, color = factor(classcode))) +
-  geom_line(show.legend = F) +
-  geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
-  geom_point(data = raw,aes(year, smd, color = factor(classcode)), show.legend = F)+
-  facet_wrap(~classcode) +
-  labs(y="Raw (dots) and Standardized (lines) Abundance",x="Year")+
-  theme_minimal()+
-  theme(axis.text=element_text(size=6))
-
-raw_v_std_plot
-
-```
-
-We include a variety of environmental, observation, and temporal indicators in our model.
-Inclusion of highly co-linear variables in a model can inflate standard errors and obscure "true" effects.
-To account for this we calculated the Pearson's correlation coefficients for all of the continuous data included in our model to ensure that none of the included variables had correlation coefficients greater than 0.7, a general rule of thumb for co-inclusion of variables.
-We did not find problematic levels of correlation among any of our included continuous variables.
-
-```{r numeric-cor, fig.cap="Pearson correlation coefficients of continuos data included in the regression model"}
-
-numeric_cor_plot <- used_data %>%
-  select_if(is.numeric) %>%
-  select(contains("mean_"), n_obs) %>%
-  na.omit() %>%
-  corrr::correlate() %>%
-  gather("colname", "correlation", -rowname) %>%
-  ggplot(aes(rowname, colname, fill = correlation)) +
-  geom_tile() +
-  geom_text(aes(rowname, colname, label = round(correlation, 2))) +
-  scale_fill_gradient2(
-    low = "tomato",
-    mid = "white",
-    high = "steelblue",
-    midpoint = 0,
-    limits = c(-1, 1)
-  ) +
-  theme_minimal() +
-  theme(
-    axis.text.x = element_text(
-      angle = 45,
-      vjust = 1,
-      hjust = 1,
-    size = 6),
-    axis.text.y = element_text(size = 6),
-    axis.title.x = element_blank(),
-    axis.title.y = element_blank()
-  )
-
-numeric_cor_plot
 ```

### 2020-11-19 documents/referee-replies.Rmd

---
title: "Referee Replies"
author: "Dan Ovando"
date: "11/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

# General Comments

# Reply to REGIONAL EDITOR'S COMMENTS

Comments to the Author:

Thank you for submitting your manuscript ‘Assessing the Population-level Conservation Effects of Marine Protected Areas’ to ‘Conservation Biology’.

The Handling Editor obtained three thoughtful and comprehensive reviews, which variously recommended ‘Minor Revision’ , ‘Moderate Revision’ and ‘Major Revision’.  The Handling Editor recommended ‘Moderate Revision’, a recommendation I support and recommend to the Editor-in Chief.

Identifying the conditions under which MPAs can provide conservation benefits beyond their borders is an important area of research that should be of wide interest to the generic readership of Conservation Biology. The Channel Islands National Marine Sanctuary is an appropriate and data-rich case study.

*Your chief challenge will be to construct a coherent and accessible story in the main text, without reliance on the supplementary material, which should be designed to house the material of interest to the specialist reader only.*

We thank you for the constructive comments, and hope that the revised version of the paper has sufficiently address the helpful points raised by each of the referees. In particular, we have sought to clarify the general contributions of this paper in the introduction. XX. We thank you for your consideration and look forward to hearing your response.

I look forward to a revised manuscript, which will be sent out for another round of review.

I hope you will continue to consider ‘Conservation Biology’ as your journal of choice for relevant manuscripts that are of potential interest to a wide international readership.

Sincerely
Helene Marsh
Regional Editor

# Reply to HANDLING EDITOR'S COMMENTS
Comments to the Author:

The Authors assess under which conditions MPAs can provide conservation benefits beyond MPA-borders. This is an important area of research – helping to guide expectations about how well MPAs protect species within and outside their borders, and whether these effects will be detectable. The Authors demonstrate their approach in The Channel Islands National Marine Sanctuary, a well-studied system with available data. The Authors are also to be commended for their efforts to provide fully open access code.

The Authors show that conservation benefits are very difficult to detect empirically—even when population-level benefits are large. MPAs are often criticised for their inability to provide spillover benefits to fishers who are negatively impacted by area closures. The current results provide insight into why evidence of spillover benefits are often lacking.

The three reviewers provided a comprehensive assessment of the manuscript. Their feedback was generally positive, but emphasized the need for *greater clarity in the presentation of the methods, and structuring of the material between results, introduction and discussion*. I must agree with this assessment, although I can appreciate that the Authors have had a difficult task clearly explaining an involved methodology with many results that add value to the interpretation and applicability of the findings. *The main elements that the Reviewers felt could be better explained (or moved from the supplementary to the main text) include details of the Bayesian Difference in Difference approach, formulas for the simulation model, and details regarding the model scenarios. I also felt that key results from the study were presented in the Discussion—namely results from simulation studies describing when MPA effect sizes were likely to be detectable (paragraphs beginning Line 220 and Line 228).*

Thank you for this clear synthesis of the helpful feedback provided by the referees. We have addressed each of their comments in the text, and outlined our responses below. In particular, we have included more details about the simulation model and Bayesian Difference in Difference in the methods, ensured that all results are in the results section, and made clearer the specific objectives of our study in the introduction.

# REVIEWER COMMENTS

## Reply to Reviewer: 1
Comments to the Author

I have read with interest the manuscript entitled: “Assessing the Population-level Conservation Effects of Marine Protected Areas”. This manuscript compiles very interesting approaches and outcomes. However, I believe the manuscript requires substantial modifications to make it clearer. Here are some suggestions.

Major comments:

Abstract:
- *Please indicate and describe briefly the method you used in your article*
  - We have clarrified in the abstract that we use a Bayesian difference-in-difference model

Introduction:
*- Clear definitions of the terms are needed (“population-level effect”, “alteration of fishing effort” etc…)*

We have sought to clarify these terms in the the introduction, thank you for pointing this out

*- A clearer explanation of the gap in the current literature that you are willing to fill is needed*

I truly have no idea here

*- Please be logic in the demonstration of the idea that you want to develop in each paragraph. Sometimes, it looks confusing to me. For instance, in the 3rd paragraph (line 43 to 52), the conclusion does not correspond to the literature cited*

Again, no idea

*- You simply indicated that control sites in empirical studies are not good controls but you don’t even mention BACI or BACIPS methods that try to take into account the temporal and spatial variations*

We apologize for the confusion here. BACI and BACIPS are both powerful tools for program evaluation in natural systems. However, both depend on the control being unaffacted by the treatment. If the treatment affects the control, e.g. if an MPA displaces fishing effort onto a control site, then any attempt to use the control sites as a control in a BACI will produce biased estimates. The key problem here then is not the temporal or spatial variation that BACI-style methods seek to address, but the likely influence of the treatment on the control sites in the context of MPA studies. We have endeavored to make this clearer in the text, on line XX>

  - GOD DAMN IT ALL TO HELL PEOPLE, SUTVA!!!!

*- Please cite and explain briefly which methods you are using in this article*

I really don't understand this at all, I guess maybe Snow 1855? Larsen cites Angrist & Pischke 2009, go with that

*- Please be clearer when presenting the objectives of the study*

Like seriously, what? I guess we do present the results in a more "modern" way, i.e. here are the results and then we explain why, and we could do a more traditional "we ask these questions somewhere in the intro".

Methods:

*- More description on the simulations is needed: how many replicates? How do MPAs were implemented? Are they randomly assigned? How many simulations for each chosen species in the bio-economic model? …?*

Given word limits we unfortunately had to leave many of the details of the simulation model in the supplementary materials. We have however added expanded descriptions of these key attributes in the main text. Line XX.

*- Please add clear log-ratio and confidence interval equations.*

We apologize but we are not clear what this comment is in reference to. Confidence intervals in a Bayesian regression do not have an equation per-say...

I don't know how to respond to this without sounding snarky.

*- Please explain at least quickly the method of Bayesian regression (or explain the choice of this methods)*

We chose to employ a Bayesian regression as we feel that Bayesian inference has several advantages for program evaluation. Specifically, it allows us to answer questions such as "what is the probability that the Channel Island MPAs had a greater than 0% effect on biomass densities of targeted finfish". For comparison, we have added a frequentist mixed effects version of our model (Fig.SXX), which profuces qualitatively the same results XX.

*- Please explain your fishing dynamics in the bio-economic model (constant catch? Fishing mortality?) because you are discussing this element a lot in the discussion section*

Thank you for pointing out that these terms are poorly described in the text, we have improved our descriptions (line XX) as well as expanded our definitions in the SI.

-       Please explain why do you choose to use two different methods to assess the effects of MPAs. Line 113-114: why this method has been chosen in particular? What does it offer compared to other assessment methods?

We take the referees comment to mean our choice of using both a difference-in-difference model and a simulation model? We apologize for the confusion. In this case, we have not used the simulation model as an explicit estimation strategy, but rather to frame expectations. However one could, in the manner of @nickols2018, tune the simulation model to the local system and use simulation projections as a counterfactual. Since our simulation model is not as empirically calibrated as the one used in @nickols2018, we do not use it as an identification strategy. As discussed on line XX, we feel that it is important to ground any empirical results in theoretical predictions, and that is why we present the simulation study along with the empirical assessment.

In regards to the difference-in-difference approach, we selected it as it fits the nature of the policy intervention in question: we have a treatment (targeted species) and a control (non-targeted species) observed before and after impact (MPA implementation). As we have sought to clarify on line XX, DiD is for all intents and purposes the same as BACI, which is a commonly used metric in MPA evaluations. We use the term DiD following the vein of @larsen2018, to more closely tie the program evaluation literature in ecology to the econometrics literature that more directly deals with causal inference in program evaluating. Our main methods difference then is in choosing to use targeted and non-targeted species as our treatment and control group, rather than inside and outside MPAs, since we believe there is strong a strong possibility that the MPAs affect the "control" sites outside their borders in the Channel Islands. In addition, because our goal is to estimate the population level effect of the MPAs, and for all of the species evalauted here the population stretches well beyond the borders of the MPAs, we need a control group that is ...

Given the nature of the study, other econometric techniques such as regression discontinuity or instrumental variables are not viable (we do not expect a discontinuous effect given the population dynamcis in play, and we have no valid instrument for the implementation of MPAs that is exogenous to the outcome). Synthetic controls [@abadie] do present another potential solution to our problem, which we use in SM.XX, and find no clear effect.

Results:
*-       Results are unclear and do not fully described what is presented in the figures. Some parts should be in the introduction or method or discussion sections. For instance: you cannot say that the effect is due to size and enforcement because you don’t have tested for that;*

We apologize that our results were unclear and have sought to clarify our findings. We are however unclear where the referees concern about "size and enforcement" come in. We mention size and enforcement on line 277 in the discussion in the context of setting expectations, which we feel is an important point to make. We do not mention enforcement in our results that we can find.

XX don't be snarky XX

*Line 146: all this paragraph should be in the introduction*

We have moved our discussion of the challenges of response ratios starting this paragraph to the introductin. However, we feel that the specific findings shown in Fig.3 are results, and as such have left them in the results section.

*- Please put figures in appropriate (figure 4 is cited before Fig. 3 and Fig. 2)*
*- Fig.2 is not described. Why?*

Thank you for pointing this out, we have corrected both of these errors.

Discussion:

*-       Please, do not cite a figure for the first time in the discussion that has a major result in it (Fig. 5)*

Thank you for pointing this out, we have moved all of this section to the results

Figures:
*-  Please be consistent with the terms used in figure titles, figure, and in the text*

Thank you for pointing this out, we have checked the text and figures for consistency.

*-  Please make the figures cleaners (position of axis titles, clear legends etc…)*

We apologies that our figures were not sufficiently clear. We have tried to clarify to the best of our abilities.

Minor comments:

*Line 25: “Recent international agreements” which one(s) are you referring to?*
Reference 30 by 30

*Line 27: a space is missing before the citation*

Corrected

*Line 37: “how overfished the population would be without the MPA” do you mean that the population should be overfished before the establishment of the MPAs so we can see an effect?*

More or less. The effect size, and hence detectability, of an MPA is a function of the degree of fishing pressure. An MPA on a lightly fished population may have an effect, but it will me small and hence hard to detect... etc.

*Line 42: please define what do you mean by “network effect”?*

Thank you for pointing this out. We have clarrified that by "network effect" we mean the idea that a network of MPAs can produce impacts greater than the sum of their individual parts, @XX

*Line 44: “commonly measured as the ratio of biomass densities of species” Please define what do you mean by “biomass densities”.*

We have clarified this on line XX, meaning the biomass per unit area XX.

*Line 48: please put citations in chronological order*

Citations are automatically formatted according to the citation style langauge file provided by Conservation Biology, so we have left the order as dictated by the software, unless the editors ask otherwise.

*Line 50: there is a problem with your reference: it is “Horta e costa et al., 2013” (same problem in the reference section)*

Thank you for pointing this out, we have corrected the reference

*Line 50: please precise what do you mean by “alteration in fishing effort”*

AHHHHHHHHH

*Line 51-52: “Why does this large body of empirical work not serve as sufficient evidence for the population-level effects of MPAs?” In my opinion, you did not define clearly why all the literature you cited before does not provide sufficient evidence for the population-level effects of MPAs. You just referred to publication but you did not explain why the results are debatable.*

Per the referees helpful suggestion, we have moved materials related to this from the results to the introduction (line XX), and broadly clarified that the primary issue is the violation of SUTVA, i.e. the fact that the MPAs affect nearby fished areas through biological spillover and concentration of the fishing fleet

*Line 57: reference needed here*

We have added @ferraro2018 and @larsen2018 as references here

*Line 58: Is “biological spillover” different from “spillover”? Please be consistent with the terms*

Thank you for pointing this out XX

*Line 58: “concentration by the fishing fleet” Shouldn’t it be “concentration of the fishing fleet”? Could you precise where does it concentrate (I assume this is next to MPA sites)*

We have corrected this and clarified the explanation (meaning that if there were 10 boats fishing inside the MPA and 10 boats outside the MPA, closure of the MPA may result in the 10 boats the used to used the MPA moving to the targeted area, so that there are now 20 boats fishing outside the MPA)

*Line 60: “Response ratios can be a highly imprecise and biased measure” You should explain why and what are precisely the problems you target with this method*

We have sought to clarify the specific problem of SUTVA violoations that these problems face and that our proposed strategy tries to resolve

*Line 96: “restrict fishing pressure to be mostly moderate to low” should be rephrased for a better understanding*

XX

**Line 136: please add standard deviation with means**
Thank you for pointing out this omission, these have been added XX

*Line 150: “estimated posterior probability distributions” please explain what it is in the method section*

Thank you for pointing out that this may not be a commonly used term XX. Bayes theorem states that the probability of x conditional on y, *p(x|y)*, is proportional to the probability of y conditional on x, *p(y|x)* times the prior probability of x, *p(x)*; *p(x|y) ~ p(y|x)p(x)*.

The posterior probability distribution is the distribution of *p(x|y)*, the outcome of any Bayesian analysis. See @gelmanXX.  So, in this case, we are reporting the posterior probability distribution of the MPA effect, conditional on the data (and the other covaraites etc). Given the word constraints in this paper we do not have space for an in-depth explanation of Bayesian methods, but have clarrified our choice of Bayesian methods and included references for readers interested in learning more

XX too snarky?XX

*Line 163: bracket missing*
Corrected

*Line 179-182: not only model-based papers find that size do matter for MPA effectiveness, also meta-analyses like Claudet et al., 2008*

Thank you for pointing this out, we have included this citation

*Table 1: “DD adult movement” please define DD*

Defined, thank you!

Fig. 3:
-       be consistent with the terms in the title and in the figure please
-       put the legend in the middle of y axis
-       Please change the legend for “Paired simulation Pop. Effect” and “Empirical Response Ratio” because it seems that “/” in between is a ratio

Thank you for pointing out the confusion in this figure, we hope the new version is more suitable.

# Reply to Reviewer 2

Comments to the Author

I enjoyed reading this paper and feel it makes an important contribution to the MPA literature. The focus on distinguishing seascape-wide MPA effects from local MPA effects is timely and important.

*Thank you, we are glad to hear the reviewer feels this is an important contribution.*

My main concerns on reading the paper are that the methods and results are not clearly enough described.

*1) Although the parameterisation of the simulation model is provided the equations should be included in supplementary. I did visit the github site and the simulator but neither allowed me to find formulations easily. So first off the simulation model needs to be explained in detail in supp methods including equations.*

 Thank you for your comment. We have moved the operating model equations from the SI (and clarrified that those equations are the simulation model) to Table.XX in the main methods, and clarrified in Section XX. the role of these equations in the simulation model.

*2) On a related note I didn't come away with a clear sense of the scenarios used to simulate the potential fish outcomes. For example I have no clear sense how you have simulated spatial patterns of fishing intensity. Similarly, it was not clear to me how you decide which model scenarios are most plausible when conditioning on observed response ratios. Yet conveying the ways in which observed RR could be derived would be useful. There is some discussion of this but not enough to detail. I believe that a good part of the discussion could be removed to make way for this.*

Given the word constraints of the paper, we felt that most of the simulation detail had to be left to Table.XX, Supplemental Information section XX. We agree though that readers would benefit from more clarity in the main text. Given the particularly critical nature of movement and fishing behavior in MPA models, we have included the following statement on line XX of the methods.

XX

3) The figures are generally not very clear; particularly 2 and 6. In the case of 2 it just isn't visually very clear what's going on and for 6 it requires a more interpretation in the caption to help someone follow it.

*We apologize for the confusion in the figures, we have... *

*4. Minor - some of the statements would benefit from being more nuanced. Line 187 states that fishers have to work harder to maintin constant catch when MPAs are in place. Well, potentially but not if they benefit from significant biological spillover.*

Thank you for pointing this out, we have restated this as "Fishers pursuing constant catch in areas outside an MPA may have to fish harder to achieve the same catch from a smaller part of the population"

# Reply to Reviewer 3

Comments to the Author

*This study fills an important gap in knowledge about the seascape wide effects of MPAs on fish populations. The introduction clearly lays out the challenge: we know a lot about how fish biomass increases inside MPAs, but very little about whether this affects seascape scale biomass overall. It uses a novel combination of simulation and empirical modelling to ask if the Channel Islands MPA system has seascape wide effects on fish populations, and under what conditions that might occur.*

We are glad that the reviewr found our contribution important and novel, thank you!

*I loved the open data/code approach to this paper, it goes the extra mile by having an online application. I had fun exploring that thanks.*

You are welcome, we are glad it helped

*The simulation model represents on its own a very complete synthesis of the fish biomass MPA modelling literature, and includes most of the processes and their variants, that have been shown over the past 20 years to affect MPA outcomes.*

We are glad you found the simulation model sufficiently comprehensive

*There are some minor comments/edits/clarifications below, but my main questions are about the ‘difference-in-difference model’, which seems to be a type of GLMM. I haven’t heard the term before, so I think it needs a bit more explanation. You might want to move some of the explanation from the SI which was really clear, over to the main text.*

We have moved some of the methods from the SI as suggested, which we hope clarifies what a difference-in-difference is (essentially the same thing as a BACI). We use the term difference-in-difference to tie our results more clearly to the causal inference literature in econometrics. We have also clarrified on line XX the clear link between DiD and BACI.

*After reading just the main text I was left with a few questions about the DiD: how the years were binned (and why those bins)? The year ‘bins’ are in the table and figure 4, but in the methods it says there was a just a pre and post MPA time-period.*

Thank you for pointing this out. We have clarified the pre and post time period discussion in the methods (line XX). We chose to bin the results for visual clarity. At the risk of further expanding the already extensive supplementary materials, we have also included a version with a binwidth of 1 (i.e. no binning) to make clear that our results are not sensitive to the choice of binwidth.

*Also, it seems like MPA status (so inside vs putative ‘control’ sites) should also be a factor in the model, but that wasn’t included? *

This is a very good point. We include hierarchical terms for each site in the data, clustered by the specific island in the Channel Islands a specific site was located at. This should serve to control for any intrinsic traits of the MPA sites themselves, e.g. if MPA sites were preferentially placed on better habitat than "control" sites. The site-specific intercept then represents the mean total biomass density of non-targeted species pre-MPA, conditional on the other covariates in the model.

Our reasoning for not including a specific interaction term for site, whether that site is an eventual MPA, and whether the MPA is in place in a given year, is that this would muddy the interpretation of our key variable of interest; namely the difference in mean total biomass density of targeted finfish throughout the Islands. Adding in a term representing the presence of an active MPA at a site would effectively mean that the our main coefficient of interest, the interaction of the targeted dummy variable with the year blocks, would represent the mean effect of the MPAs on mean total biomass density of targeted finfish controlling for any effect that those MPAs had inside the MPAs.

XX clarify this messXX

*If it is included, then the predictions for total density should be post-stratified by total area of MPAs in the seascape?*

This is correct, it was quite buried in the SI, but we do weight the MPA and non-MPA sites according to the total area of MPAs in the seascape (SI line XX), since as the referee points out the sampling regime is not random. We have mentioned this in the main methods as well now on line XX. We also include an unweighted version of our results, to demonstrate that our results are not sensitive to this weighting choice (Fig.SXX).

*The supplemental materials are very extensive because they include exploration of multiple different model formulations both models, and also a simulation study testing the difference in difference approach. I think you could cut them down a bit if they wanted, and save some of this for more complete exploration in another paper? I think most readers will never get this far, so it seems a bit of a waste to have all this interesting information in a supplemental. The parts I found most in the SI useful for this particular study were the explanation of the DiD, the simulation study and the documentation of the simulation model.*

Thank you for this point! To make things more streamlined, we have removed the transect-level DiD from the SM, as it is VERY long, very convoluted, and is by far the most difficult part of the code for others to replicate given the long run time.

Minor comments

*Introduction reads nicely. One comment: it might be worth stating that modelling studies are generally predicting population level effects, but empirical studies are measuring local effects. So another great example of a modelling-empirical disconnect. You’ve set a high expectation in the reader’s mind that you are going to fill this gap, which is then met through the novel combination of simulation and statistical modelling.*

Thank you for this excellent suggestion. We have reflected this suggestion on line XX.

*The references are pretty focused on California, and there is certainly a lot of great MPA literature from that part of the world, but perhaps you could expand the geography a bit to make the relevance to international readers clear?* See suggestions below.

We agree that inclusion of a wider set of references may broaden the impact of our paper and have included the recommended references, thank you for the suggestion.

*81: thanks for sharing the code. Note doing that does invalidate your double blind.*

We are glad you found it helpful. We did consider this, and decided that the extra effort needed to blind the code itself was not worthwhile. An interesting challenge for journals seeking to do double blind reviews to consider, it's hard to post code that is both open and reproducible and anonymous!

*89: Nils Kruek did a similar thing in his paper, it might be worth reading and relevant citing:
https://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.2000537*

Thank you for pointing this out XX dig into this one

*91: “equilibrium”. But most field studies won’t be measuring equilibrium, so herein lies an important difference?*

We have clarified that the simulation keeps track of outcomes at a variety of timesteps, including non-equilibrium. The "begins at" is just to clarify that we run the population out to equiblirum since there isn't an analytical solution to the biomass distribution in an age-structured spatial model when habitats vary.

*105: so just to be clear, its not one mean density for inside, one for outside, but one mean density for everywhere?*

That is correct, so if the MPAs cover 50% of the habitat, biomass density goes 20% inside, 10% outside, our model would estimate 15%. We do however include site-specific intercepts that control for different mean baseline densities at each site, as well as site-specific environmental covariates (e.g. kelp cover) that can also account for some site specific differences.

*Eqn 1: Don’t the gamma shape and scale depend on its mean? So maybe more correct to write this as:
Expectation(d) = exp(beta….) and then say observations were sampled from a gamma, where the dispersion parameter was free/estimated.*

Thank you for the suggestion, we have modified the text accordingly.

*It could be helpful to show an equation for the net DiD outcome (as explained at A.3). I think its something like:
(D[post, fished]-D[pre, fished]) – (D[post, unfished]-D[pre, unfished])*

We have included this section from the SM XX.

*Was temporal autocorrelation was dealt with/checked for?*

This is a good question. We considered treating the year effects as hierarchichal terms, which would help account for the autocorrelated nature of observations over time. We elected not to, as doing so would negate their benefits as an identification strategy, controlling for unobserved variables that are correlated with with the timing of MPA placement with the total biomass densities of targeted finfish. As discussed in @larsen2018, treating our year terms as hierarchical (or random effects in the context of that paper) means that we assume that the error distribution of the year terms is uncorrelated with the treatment, meaning that the difference-in-difference no longer acts to remove omitted variable bias, which is its main purpose here.

This may mean that we are underestimating the error in our estimated MPA effects, since the model treats observations from different years as conditionally independent. We feel that this would be a more important concern if we had estimated a for example positive but small effect of the MPAs, where perhaps slightly broader error bars might start to span zero. Given that one of our main results is that we are unable to detect a clear signal, we feel that the possible risk of underestimating the error distributions is outweighted by the potential gains of helping control for omitted variable bias.

*Fig 2: can you make the lines a bit narrower?*
Lines are narrower and XX

*Fig 4: indicate pre-MPA period.*

We have used XX to indicate the pre-MPA period

*189: I don’t think it is overlooked. E.g. see work done on the Great Barrier Reef, where quota management was coordinated with a major MPA rezonining in 2004. Papers by Little in this journal (2009 “Complementarity of No-Take Marine Reserves and Individual Transferable Catch Quotas for Managing the Line Fishery of the Great Barrier Reef”) and Hopf in Current Biology (2016 “Synergistic Effects of Marine Reserves and Harvest Controls on the Abundance and Catch Dynamics of a Coral Reef Fishery”)*

Thank you for pointing this out, and have rephrased this section accordingly (line XX)/

*225: I don’t think this simulation study was reported in the methods and results, but should be, or reference given to appendix?*
*228: this paragraph belongs in results.*

We have moved this section to the results

*Discussion: How would poaching in MPAs affect the interpretation of the difference-in-difference model?*

Interesting question! As we discuss on line XX, what evidence we have is that poaching is of limited concern in the Channel Islands (but clear can be a major issue in many MPAs). However, the presence of poaching should not bias the estimates of the DiD model in any way: it poaching were severe enough so as to render the MPAs completely ineffective, then in theory a DiD model such as the one implemented here would simply estimate that the MPA had no effect, as their would be no difference in the conditional biomass densities of targeted and non-targeted species.

*271: It would be good to read more comparison with the Nickols study, given the similarities to this study. In particular, they used a dynamic model to explore whether MPAs are achieving biomass recovery, also for Channel Islands. So how do their findings relate to what was found here?*

We briefly (due to word limit constraints) expanded our discussion of the synergies between our work and the Nickols study. Broadlyt we feel that the Nickols study further makes the case for...

One key difference is in the estimation strategy employed by our two studies. The Nickols approach might be considered a structural econometrics approach, in which we simulate the expected outcome in the absence of treatment, and then compare observed outcomes to the simulated. While they do not explicitly do so in their paper, their discussion of comparing the observed values to the expected baseline is in this vein. We on the other hand take an empirical approach, in which we construct the counterfactual "what would have happened without" the MPAs based on the observed trajectories of the non-targeted species (though in the same vein as Nickols et al. we simulate a counterfactual as well). Both of these approaches have strengths and weaknesses, and our studies share a common emphasis on the need for robust simulation modeling to ground expectations.

Chris Brown

### 2020-11-19 documents/ovando-et-al-pop-effects-of-mpas.Rmd

@@
 Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level.
-We show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); the size of many MPA networks suggests that effect sizes may often be smaller than 30%.
+Using a Bayesian difference-in-difference approach, we show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); the size of many MPA networks suggests that effect sizes may often be smaller than 30%.
 Our results provide a novel assessment of the population-level effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.
@@
 Over time our goals and expectations for MPAs have evolved and become more ambitious.
-Recent international agreements to expand MPAs are based on the assumption that well-designed MPAs will achieve benefits both within, and outside, their borders[@gaines2010].
+Recent international agreements to expand MPAs are based on the assumption that well-designed MPAs will achieve benefits both within, and outside, their borders [@gaines2010].
 
@@
 However, export of adults or larvae from the MPA to these "control" site affects their status as controls, as does displacement of fishing effort from MPAs to control sites.
-In theory, control sites far enough away to negate both biological spillover and concentration by the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs is challenging.
+In theory, control sites far enough away to negate both biological spillover and concentration of the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs is challenging.
 Response ratios can be a highly imprecise and biased measure of the population-level conservation effect of an MPA network [@ferraro2018;@moffitt2013].
@@
 We pair these data with randomly selected values governing the characteristics of the simulation (Table.\@ref(tab:sim-table)).
+These include
+
 The population begins at unfished equilibrium.
-After fishing effort is applied, each simulation is run to equilibrium with and without the selected MPA design (holding all else constant).
+After fishing effort is applied, each simulation is run to equilibrium with and without the selected MPA design (holding all else constant) and sampled throughout the time period, to reflect the fact that few empirical MPA studies are occuring in systems that have reached a theoretical stable equilibrium.
+
+Movement and fishing behavior are two of the key determinants of MPA effectivness.
+While we include many other variables in our simulation (Table.\@ref(tab:sim-table)), for movement we simulate a range of movement rates for both larvae and adult fish, as well as density dependent and independent movement.
+For fishing behavior, we allow total effort or total catch to remain fixed, with or without MPA mediated attrition, or effort to follow open-access dynamics.
+The effort dictated by the fleet model is distributed in space either uniformly, such that marginal catch is equal in space, or such that marginal profits are equal in space.
+
+XX clarify here the
+
 In each time step, we measure the difference in biomass densities between the scenario with and without the MPAs to calculate the population effect of the MPAs over time.
-Simulation results provide a library of plausible MPA effects for a range of biological and economic assumptions.One set of simulations is specifically designed to reflect the dynamics of the Channel Islands.
+Simulation results provide a library of plausible MPA effects for a range of biological and economic assumptions.
+One set of simulations is specifically designed to reflect the dynamics of the Channel Islands.
 We only include species of the same genus as those targeted by fishing in the Channel Islands, restrict fishing pressure to be mostly moderate to low, and cap the MPA size at 20% of the population's range (a likely upper end to the effective MPA size in question, @rassweiler2012).
@@
 Over the next six years the model estimates greater probabilities of an increasingly positive MPA effect, peaking in 2009-2011 with a median estimate of MPA effect of a `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)` increase in mean total biomass density of targeted species (90% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)`
-- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)` (Fig.\@ref(fig:mpa-effect)).
+- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)`) (Fig.\@ref(fig:mpa-effect)).
 These empirical estimates are in line with the outcomes that our simulation model suggests are plausible.
@@
 Constant-catch dynamics might also occur in fisheries with constraining quotas that are not updated after the implementation of MPAs.
-Fishers pursuing constant catch in areas outside an MPA have to fish harder to achieve the same catch from a smaller part of the population, causing a population conservation loss under `r percent(1 - positive_outcomes$pp[positive_outcomes$fleet_model == "constant-catch"], accuracy = 1)` of our constant-catch simulations.
-This is an important and often overlooked possibility, especially as MPAs are increasingly implemented in quota-managed fisheries [@liu2018].
+Fishers pursuing constant catch in areas outside an MPA may have to fish harder to achieve the same catch from a smaller part of the population, causing a population conservation loss under `r percent(1 - positive_outcomes$pp[positive_outcomes$fleet_model == "constant-catch"], accuracy = 1)` of our constant-catch simulations.
+This is an important scenario to consider (as done so in @little2011b) especially as MPAs are increasingly implemented in quota-managed fisheries [@liu2018].
 

### 2020-11-21 documents/referee-replies.Rmd
referee 1 responses mostly done

@@
 
+don't sell the model short, there's a lot in here that we just don't have space to explain in detail in the main text
+
 # REVIEWER COMMENTS
@@
 ## Reply to Reviewer: 1
+
+ADd summary paragraph, in our view these are the three main things, here are the main responses
+
 Comments to the Author
@@
 
-We have sought to clarify these terms in the the introduction, thank you for pointing this out
+We have sought to clarify these terms in the introduction, thank you for pointing this out.
 
@@
 
-I truly have no idea here
+We have provided a clearer statement of the gap in the current literature that our paper fills on line XX, and line XX
 
@@
 
-Again, no idea
+Thank you for pointing this out, we have clarified the paragraph in question and provided greater detail on why exactly this existing body of evidence does not neccesarily serve as a population-level measure of MPA effects (Line XX).
 
@@
 We apologize for the confusion here.
-BACI and BACIPS are both powerful tools for program evaluation in natural systems.
-However, both depend on the control being unaffacted by the treatment.
+BACI and BACIPS are both powerful tools for program evaluation in natural systems, and we now mention BACI explicitly in the text.
+However, both response ratios and BACI/BACIPS depend on the control being unaffacted by the treatment.
 If the treatment affects the control, e.g.
-if an MPA displaces fishing effort onto a control site, then any attempt to use the control sites as a control in a BACI will produce biased estimates.
+if an MPA displaces fishing effort onto a control site, then any attempt to use the control sites as a control in a response ratio BACI will produce biased estimates unless this spillover is somehow controlled for.
+
 The key problem here then is not the temporal or spatial variation that BACI-style methods seek to address, but the likely influence of the treatment on the control sites in the context of MPA studies.
@@
 
-  - GOD DAMN IT ALL TO HELL PEOPLE, SUTVA!!!!
-
 *- Please cite and explain briefly which methods you are using in this article*
 
-I really don't understand this at all, I guess maybe Snow 1855? Larsen cites Angrist & Pischke 2009, go with that
+We take this to mean a explanation and citation for difference-in-difference.
+We have included a summary citation for the method (line XX), as well as a brief description, given the length constraints of this paper, explaining the link between BACI and difference in difference.
 
@@
 
-Like seriously, what?
-I guess we do present the results in a more "modern" way, i.e.
-here are the results and then we explain why, and we could do a more traditional "we ask these questions somewhere in the intro".
+We have clarified the objectives of this study, and its place in the literature, on lines XX and XX
 
@@
 Given word limits we unfortunately had to leave many of the details of the simulation model in the supplementary materials.
-We have however added expanded descriptions of these key attributes in the main text.
-Line XX.
+We have however added expanded descriptions of these key attributes in the main text, in particular the nature of the variables requested here, as well as clarifying the role and nature of the simulation model.
+Line XX
 
@@
 
-We apologize but we are not clear what this comment is in reference to. Confidence intervals in a Bayesian regression do not have an equation per-say...
+We apologize but we are not clear what this comment is in reference to.
+We do not use log-ratios in our model, except in the sense that we use a Bayesian Gamma GLMER with a log link function to run our , so the estimated effect sizes are in log scales, roughly corresponding to percentage-level effects.
 
-I don't know how to respond to this without sounding snarky.
+Since this is a Bayesian regression, we provide credible intervals rather than confidence intervals [@gelman2014a].
+Similarly, there is no equation per-say for credible intervals, as they are essentially just an exercise in adding probabilities.
+Our regression provides a posterior probability distribution parameters, and we then select the values of the coefficients within a desired range.
+For example to construct a 95% credible interval, we would select the values of the coefficient from the posterior probability distribution that fall between the 2.5%th and 97.5th percentiles of the posterior probability distribution.
+
+See @gelman2014a, @mcelreath2020, @hobbs2015, and zuur2009 (for the GLMER terminology).
 
@@
 Specifically, it allows us to answer questions such as "what is the probability that the Channel Island MPAs had a greater than 0% effect on biomass densities of targeted finfish".
-For comparison, we have added a frequentist mixed effects version of our model (Fig.SXX), which profuces qualitatively the same results XX.
+For comparison, we have added a frequentist mixed effects version of our model (Fig.SXX), which produces qualitatively the same results XX.
+
+We have briefly addressed our choice of Bayesian methods on line XX.
 
@@
 
-Thank you for pointing out that these terms are poorly described in the text, we have improved our descriptions (line XX) as well as expanded our definitions in the SI.
+Thank you for pointing out that these terms are poorly described in the text, we have improved our descriptions (line XX).
 
--       Please explain why do you choose to use two different methods to assess the effects of MPAs.
+*-       Please explain why do you choose to use two different methods to assess the effects of MPAs.
 Line 113-114: why this method has been chosen in particular?
-What does it offer compared to other assessment methods?
+What does it offer compared to other assessment methods?*
+
+We did the simulation study purely to ground expectations, and then compared our simulation to the empirical estimation model.
 
@@
 
+We have more clearly explained that the DiD approach is more or less identical to a BACI on line XX, which we hope will make this particular model choice seem more familiar.
+
+XX made it here XX
 Results:
@@
 We are however unclear where the referees concern about "size and enforcement" come in.
-We mention size and enforcement on line 277 in the discussion in the context of setting expectations, which we feel is an important point to make.
+We do not claim to have tested for enforcement in our results.
+We mention size and enforcement on line 277XX in the discussion in the context of setting expectations, which we feel is an important point to make.
 We do not mention enforcement in our results that we can find.
-
-XX don't be snarky XX
+We do test for the effect of MPA size, which we make clearer in our methods description, on line XX.
 
@@
 
-We have moved our discussion of the challenges of response ratios starting this paragraph to the introductin.
-However, we feel that the specific findings shown in Fig.3 are results, and as such have left them in the results section.
+We respectfully disagree with this suggestion.
+Following referee #1's helpful suggestions, we have clarified the broader challenges with interpreting response ratios as population-level effects in the introduction on line XX, which we agree were poorly articulated.
+However, we feel that this section exploring the specific response ratios in the Channel Islands is a result of our simulation modeling, and so belongs in the results section.
+Moving this paragraph to the introduction would require explaining our simulation modeling in the introduction as well, which we feel is not ideal.
+
+XX Owen, thoughts on tone here?XX
 
@@
 
-Thank you for pointing this out, we have moved all of this section to the results
+Thank you for pointing this out, we have moved all of this section to the results (line XX).
 
@@
 
-We apologies that our figures were not sufficiently clear. We have tried to clarify to the best of our abilities.
+We apologies that our figures were not sufficiently clear.
+We have tried to clarify to the best of our abilities.
+We checked all the figures, and in particular cleaned up the presentation of Fig 2.
+and Fig.3
 
@@
 *Line 25: “Recent international agreements” which one(s) are you referring to?*
-Reference 30 by 30
+
+We have clarified specific agreements on line XX
 
@@
 
-Corrected
+TRemoved, thank you.
 
@@
 
-More or less.
-The effect size, and hence detectability, of an MPA is a function of the degree of fishing pressure.
-An MPA on a lightly fished population may have an effect, but it will me small and hence hard to detect...
-etc.
+We have clarified this statement on line XX.
 
@@
 
-Thank you for pointing this out. We have clarrified that by "network effect" we mean the idea that a network of MPAs can produce impacts greater than the sum of their individual parts, @XX
+Thank you for pointing this out.
+We have clarified that by "network effect" we mean the idea that a network of MPAs can produce impacts greater than the sum of their individual parts through synergistic action.
+Line XX.
 
@@
 
-We have clarified this on line XX, meaning the biomass per unit area XX.
+We have clarified on line XX that biomass ratios refer to the weight of organisms per unit area.
 
@@
 
-Citations are automatically formatted according to the citation style langauge file provided by Conservation Biology, so we have left the order as dictated by the software, unless the editors ask otherwise.
+Our citations are automatically formatted according to the citation style langauge file provided by Conservation Biology, so we have left the order as dictated by the software, unless the editors ask otherwise.
 
@@
 
-Thank you for pointing this out, we have corrected the reference
+Thank you for pointing this out.
+We have removed that line following the Referee #1's suggestion to ensure to continuity of paragraphs, as the displacement of fishing effort is not directly a population level effect, but rather a cause of population effect.
 
@@
 
-AHHHHHHHHH
+Thank you for pointing this out.
+We have removed that line following the Referee #1's suggestion to ensure to continuity of paragraphs, as the displacement of fishing effort is not directly a population level effect, but rather a cause of population effect.
 
@@
 
-Per the referees helpful suggestion, we have moved materials related to this from the results to the introduction (line XX), and broadly clarified that the primary issue is the violation of SUTVA, i.e.
-the fact that the MPAs affect nearby fished areas through biological spillover and concentration of the fishing fleet
+Per the referees helpful suggestion, we have moved materials related to this from the results to the introduction (line XX).
+The problem is that response ratio / spatial BACI style designs will still be biased if the treatment unit affects the control unit, a violation of the stable unit treatment value assumption (SUTVA).
+For example, suppose we employ a spatial BACI/BACIPS design on a system with two patches of equal size, one of which is put in a perfectly enforced MPA, the other left open to fishing.
+Suppose then that all the fishing effort that used to operate inside the MPA moves and now operates outside the MPA, doubling the amount of fishing pressure outside.
+A BACI design that assumes that the control site is a measure of what would have happened in the absence of an MPA will conclude that the MPAs had a large positive effect, since biomass inside the MPA will decline post-MPA, while biomass inside the MPA will increase.
+Convsersely, consider that same MPA but without effort displacement but with very strong larval and adult spillover, which results in the control site benefit greatly from the MPA.
+A BACI style design would conclude that the MPA had little to no effect in this case, since biomass levels would be similar inside and outside the MPA.
+
+XX Owen check if you think this is clearer now.
 
@@
 
-Thank you for pointing this out XX
+We have clarified that by spillover we mean spillover of fish or larvae.
 
@@
 
-We have sought to clarify the specific problem of SUTVA violoations that these problems face and that our proposed strategy tries to resolve
+We have sought to clarify the specific problem of SUTVA violoations (treated MPA sites affecting nearby control sites) that justify why response ratio studies do not necessarily serve as an index of population level effects, and in our methods clarified that it is not the DiD model itself that resolves this problem (which is akin to a BACI), but rather using non-targeted species as the control group, rather than relying on spatial controls.
 
@@
 
-XX
+We have rephrased this to reflect that the populations are likely moderately to lightly exploited in the Channel Islands.
 
 **Line 136: please add standard deviation with means**
-Thank you for pointing out this omission, these have been added XX
+
+Thank you for pointing out this omission, these have been added the standard deviation of the simulated values on line XX.
 
@@
 
-Thank you for pointing out that this may not be a commonly used term XX.
+we have clarified that the posterior probability distribution is the outcome of a Bayesian model (line XX.) Thank you for pointing out that this may not be a commonly used term.
 Bayes theorem states that the probability of x conditional on y, *p(x|y)*, is proportional to the probability of y conditional on x, *p(y|x)* times the prior probability of x, *p(x)*; *p(x|y) ~ p(y|x)p(x)*.
-
 The posterior probability distribution is the distribution of *p(x|y)*, the outcome of any Bayesian analysis.
 See @gelmanXX.
-So, in this case, we are reporting the posterior probability distribution of the MPA effect, conditional on the data (and the other covaraites etc).
-Given the word constraints in this paper we do not have space for an in-depth explanation of Bayesian methods, but have clarrified our choice of Bayesian methods and included references for readers interested in learning more
 
-XX too snarky?XX
+The posterior probability distribution is what allows us to interpret the say 95% credible interval of a parameter as "there is a 95% probability that conditional on the data, model, and priors the effect of Z was between X and Y".
+In contrast a frequentist 95% confidence interval tells us that there is a 95% change that if were to resample the data infinite times and refit the model, there is a 95% chance that the true parameter value would fall within that range.
+
+So, in this case, we are reporting the posterior probability distribution of the MPA effect, conditional on the data (and the other covariates etc, priors, and model structure).
+
+Given the space constraints of this paper, we do not feel that we are able to include a broader discussion of Bayesian inference, but have included a citation to a standard textbook on the topic in our introduction of our methods on line XX.
+
+XX Owen, too snarky? XX
 
 *Line 163: bracket missing*
+
 Corrected
@@
 
-Thank you for pointing this out, we have included this citation
+Thank you for pointing this out, we have included this citation.
 
@@
 
-Defined, thank you!
+Thank you for pointing that out, we have replaced DD with density-dependent
 
@@
 
+How did you decide which scenarios to run, given the results how did you interpret them. THis is how we interpret that request
+
+explain more the variety of ways that these results could have been observed
+
 Given the word constraints of the paper, we felt that most of the simulation detail had to be left to Table.XX, Supplemental Information section XX.
@@
 
-3) The figures are generally not very clear; particularly 2 and 6.
-In the case of 2 it just isn't visually very clear what's going on and for 6 it requires a more interpretation in the caption to help someone follow it.
+*3) The figures are generally not very clear; particularly 2 and 6.
+In the case of 2 it just isn't visually very clear what's going on and for 6 it requires a more interpretation in the caption to help someone follow it.*
 
-*We apologize for the confusion in the figures, we have... *
+Declarative sentance for what this figure is out. Results of ... Insight first, then terms at the end. warmer colors indicate larger effect
+
+Add MPA implementation line. Make clearer in the figure caption. Removing lindividial lines would artificially make it clearer
+
+We apologize for the confusion in the figures, we have...
 
@@
 *225: I don’t think this simulation study was reported in the methods and results, but should be, or reference given to appendix?*
+
 *228: this paragraph belongs in results.*

### 2020-11-21 documents/ovando-et-al-pop-effects-of-mpas.Rmd
referee 1 responses mostly done

@@
 Over time our goals and expectations for MPAs have evolved and become more ambitious.
-Recent international agreements to expand MPAs are based on the assumption that well-designed MPAs will achieve benefits both within, and outside, their borders [@gaines2010].
+Recent international efforts to expand MPAs,such as the Biodiversity Beyond National Jurisdiction (BBNJ) Agreement and the Convention on Biological Diversity’s Post-2020 Global Biodiversity Framework, are based in part on the assumption that well-designed MPAs will achieve benefits both within, and outside, their borders [@gaines2010].
 
@@
 As conservation benefits accrue inside MPAs, MPAs can affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
-These dynamics result in MPAs affecting not only the organisms within their borders, but the broader connected population of affected species.
+These dynamics result in MPAs affecting not only the organisms within their borders, but the broader population of affected species connected by movement and reproduction, creating both local (e.g.
+inside and outside MPA effects) and population-level effects XX Owen help!
+I seriously don't know how to define this any clearer XX.
 Theory states that numerous factors influence how MPAs might affect fish populations.
-These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], the strength, timing, and location of density dependence [@burgess2014], the degree of enforcement [@edgar2014], how overfished the population would be without the MPA, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006a;@walters2000a;@white2011,;@white2010c;@ovando2016;@moffitt2013].
-The location and spacing of the MPAs within a network can also influence population-level impacts through network effects [@gaines2010; @costello2010a].
+These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], the strength, timing, and location of density dependence [@burgess2014], the degree of enforcement [@edgar2014], the level of fishing pressure, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006a;@walters2000a;@white2011,;@white2010c;@ovando2016;@moffitt2013].
+The location and spacing of the MPAs within a network can also influence their population-level impacts through synergistic effects, for example acting as stepping stones for larvae [@gaines2010; @costello2010a].
 
-What empirical evidence is there for these population-level effects of MPAs?
-Conservation outcomes of MPAs are generally evaluated by "response ratios", commonly measured as the ratio of biomass densities of species inside relative to outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015].
+<!-- What empirical evidence is there for these population-level effects of MPAs?  -->
+
+These theoretical results are generally based on modeling of closed populations, where some fraction of the connected population is protected inside of MPAs.
+In contrast to this population paradigm used in MPA simulations, conservation outcomes of MPAs are generally evaluated at local scales using "response ratios", commonly measured as the ratio of biomass densities (weight of organisms per unit area) of species inside relative to selected areas outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015].
 These studies found clear evidence that well enforced and sufficiently sized MPAs are associated with high response ratios.
 Several studies have also documented empirical evidence for the existence of adult or larval fish spillover affecting fish abundance [@halpern2009;@kay2012; @stobart2009; @mcclanahan2000; @russ1996; @thompson2017;@baetscher2019;@pelc2009].
-MPAs have also caused alterations in fishing effort [@costa2013;@mason2012;@murawski2005;@mcdermott2019].
-Why does this large body of empirical work not serve as sufficient evidence for the population-level effects of MPAs?
+Where response ratios are available before and after MPA implementation, before-after-control-impact (BACI) style studies have shown similarly clear and positive results [@thiault2019a].
 
-Spillover-based studies provide evidence that MPAs may have effects outside their borders, but they are not sufficient evidence for the population effect of MPAs.
+While these studies have clearly demonstrated the ability of MPAs to create differences between fished and unfished areas, why does this large body of empirical work not serve as sufficient evidence for the population-level effects of MPAs?
+Spillover-based studies provide evidence that MPAs may have effects outside their borders, but they are not sufficient evidence for the population effect of MPAs, as on their own they lack any counterfactual.
 Response ratios have shortcomings as measures of population-level effects as well.
@@
 However, export of adults or larvae from the MPA to these "control" site affects their status as controls, as does displacement of fishing effort from MPAs to control sites.
-In theory, control sites far enough away to negate both biological spillover and concentration of the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs is challenging.
-Response ratios can be a highly imprecise and biased measure of the population-level conservation effect of an MPA network [@ferraro2018;@moffitt2013].
-As response ratios are our primary source of evidence for the conservation effects of MPAs, this means that our collective understanding of the population-level impacts of MPAs is surprisingly limited.
-
+These factors can act to bias estimators that require the assumption that the control sites are unaffected by the treatment, such as response ratios and BACI designs [@ferraro2018;@moffitt2013].
+In theory, control sites far enough away to negate both spillover of fish or larvae, and concentration of the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs is challenging.
+As variations of response ratios are our primary source of evidence for the conservation effects of MPAs, this means that our collective empirical understanding of the population-level impacts of MPAs is surprisingly limited.
+We show here that in contrast to the re
 <!-- Given the potential unreliability of response ratios for the task, how can we empirically measure the population-level effects of MPAs?
@@
 
-The MPAs within the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) provide an opportunity to improve our understanding of population-level MPA effects.
-A network of protected areas covering approximately 20% of the Islands' waters was put in place in 2003 [@kirlin2013].
+To fill this important gap in our understanding of protected area impacts, we present here an paired theoretical and empirical assessment demonstrating the challenges in assessing the population-level impacts of MPAs.The MPAs within the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) implemented a network of protected areas covering approximately 20% of the Islands' waters in 2003 [@kirlin2013].
 The network has been used as a model case study in protected area design around the world (@botsford2014 and references therein).
@@
 
-Our primary methods consist of a spatially-explicit bio-economic simulation model and a Bayesian difference-in-difference regression.
-We use the simulation model to provide provide theoretical context for our empirical results.
-The regression provides our estimate of the population-level conservation effect of the Channel Islands MPAs.
+Our primary methods consist of a spatially-explicit bio-economic simulation model and a Bayesian difference-in-difference regression [@angrist2009;@mcelreath2020].
+Difference-in-difference is a regression-based approach to BACI, assessing changes in control and treatment groups before and after treatment [@larsen2018].
+We use the term difference-in-difference here to more closely tie our results to the terminology in causal-inference literature.
+This difference-in-difference regression provides our estimate of the population-level conservation effect of the Channel Islands MPAs.
+We use a Bayesian regression since it allows us to interpret our estimated effects probabalistically [@mcelreath2020], e.g.
+stating the probability that conditional on the model, priors,and data the Channel Island MPAs increased or decreased biomass of targeted species.
+Being a Bayesian regression, our difference-in-difference model produces posterior probability distributions of our coefficients, from which we can construct Bayesian credible intervals.
+
+As we have discussed, theory shows that *a priori* we would expect the population-level effects of MPAs to be highly context dependent.
+We use the bio-economic simulation model to provide provide theoretical expectations for population-level effects of MPAs, which we then confront with the empirical results from our difference-in-difference model.
+
 All analysis were conducted in R [@rcoreteam2019].
@@
 
+The purpose of the simulation model is to both set expectations for our empirical results, and demonstrate more broadly the ways in which ecological and economic dynamics can interact to produce a wide range of population-level MPA effects.
+The full range of factors explored are shown in  (Table.\@ref(tab:sim-table)), and the equations of the simulation model can be viewed in the SI section XX.
+We used this model to generate XX simulated MPA outcomes across XX species.
+
 The simulation model consists of 50 patches with wrapped edges.
-For any one simulation we randomly pull a species and its associated life history traits from the `FishLife` [@thorson2017c] package.
+For each simulation we first randomly pull a species and its associated life history traits from the `FishLife` [@thorson2017c] package.
 We pair these data with randomly selected values governing the characteristics of the simulation (Table.\@ref(tab:sim-table)).
-These include
+We highlight here key choices available to the model governing fishing pressure and MPA design.
+For a given simulation, after selecting a species the model randomly selects a fleet model and effort allocation strategy.
+The fleet model can either be constant-catch (the fleet exerts as much effort as needed to maintain a fixed amount of catch), constant-effort (the fleet maintains a constant amount of effort over time), or open-access (fishing effort of the fleet expands and contracts in response to available profits).
+The total fishing effort exerted by the fleet can then be distributed in space uniformly, in proportion to catch per unit effort, or in proportion to profit per unit effort.
 
-The population begins at unfished equilibrium.
-After fishing effort is applied, each simulation is run to equilibrium with and without the selected MPA design (holding all else constant) and sampled throughout the time period, to reflect the fact that few empirical MPA studies are occuring in systems that have reached a theoretical stable equilibrium.
+The simulation then begins to apply the fleet model to the population, and in a randomly selected year implements an MPA network.
+The model samples a percentage of the population's range to place in MPAs, and randomly assigns patches to MPAs either across across the system or preferentially on higher-quality habitat.
+The model then randomly selects whether fishing effort that used to operate inside the MPAs is redistributed to the areas outside the MPAs, or leaves the fishery entirely.
+We then contine the simulations with the MPAs in place.
+Each simulation is paired with a simulation identical in every way except that MPAs are no longer put in place.
+We then calculate the effect of the MPAs on the population as the difference in relevant metrics in the simulation with MPAs relative to their values in the simulation without MPAs.
+We ran this process XX times, over which we simulate outcomes for XX different species.
 
-Movement and fishing behavior are two of the key determinants of MPA effectivness.
-While we include many other variables in our simulation (Table.\@ref(tab:sim-table)), for movement we simulate a range of movement rates for both larvae and adult fish, as well as density dependent and independent movement.
-For fishing behavior, we allow total effort or total catch to remain fixed, with or without MPA mediated attrition, or effort to follow open-access dynamics.
-The effort dictated by the fleet model is distributed in space either uniformly, such that marginal catch is equal in space, or such that marginal profits are equal in space.
-
-XX clarify here the
-
-In each time step, we measure the difference in biomass densities between the scenario with and without the MPAs to calculate the population effect of the MPAs over time.
-Simulation results provide a library of plausible MPA effects for a range of biological and economic assumptions.
+These simulation results provide a library of plausible MPA effects for a range of biological and economic assumptions.
 One set of simulations is specifically designed to reflect the dynamics of the Channel Islands.
-We only include species of the same genus as those targeted by fishing in the Channel Islands, restrict fishing pressure to be mostly moderate to low, and cap the MPA size at 20% of the population's range (a likely upper end to the effective MPA size in question, @rassweiler2012).
+We only include species of the same genus as those targeted by fishing in the Channel Islands, restrict fishing pressure such that the simulated popuilations are moderately to lightly exploited (given the best available knowledge of stock status for the targeted species in the Channel Islands), and cap the MPA size at 20% of the population's range (a likely upper end to the effective size of the Channel Island MPAs relative to the fish populations in question, @rassweiler2012).
 
@@
 The difference-in-difference model used empirical kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands.
-PISCO conducts visual SCUBA surveys at a large number of rocky reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\ref{fig:ci-map}).
+PISCO conducts visual SCUBA surveys at a large number of rocky reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\ref{fig:ci-map},Fig.\@ref{fig:raw-density}).
 The details of the monitoring program are described in @caselle2015.
@@
 
-Building off of the concepts explored in @caselle2015, we used an identification strategy utilizing biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 0])` species that are not directly targeted by fishing as our control group (non-targeted), and biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 1])` species targeted by fishing as our treatment group.
+Building off of the concepts explored in @caselle2015, we used an identification strategy utilizing biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 0])` species that are not directly targeted by fishing as our control group (non-targeted), and biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 1])` species targeted by fishing as our treatment group (see Fig.\@ref(fig:raw-density) for raw trends in targeted and non-targeted biomass densities).
 Targeted species in the Channel Islands include fin-fish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
@@
 We used a Bayesian difference-in-difference regression to estimate any difference in mean total biomass densities of fin-fish species targeted by fishing effort (i.e., those potentially affected by an MPA) and those species not targeted by fishing before and after MPA implementation.
+Note that this is akin to a BACI, but rather than relying on spatial control sites as our identification, we utilize non-targeted species as our control.
+
 This identification strategy controls for unobserved environmental shocks to the system that are independent of the MPAs.
@@
 Our Channel Islands-specific simulation provides a set of outcomes that can feasibly bound our expectations for the empirical effects of the MPA network.
-The targeted species in our database span a range of life histories, but are largely made up of fishes in the perch and rockfish complexes, with a mean Von Bertalanffy growth coefficient of `r round(mean(used_data$k[used_data$targeted == 1]),2)`, and a median age at maturity of `r round(median(used_data$tm[used_data$targeted == 1]),0)` years, and a mix of exploitation histories, ranging from low to high (but mostly relatively low) [@alonzo2004;@dick2017].
-While a wide range of MPA effects are plausible, from 0% to upwards of 150%, the median simulated effect size for this stylized version of the Channel Islands after over ten years of protection was `r percent(median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected > 10]))` (Fig.\@ref(fig:mpa-effect)).
+The targeted species in our database span a range of life histories, but are largely made up of fishes in the perch and rockfish complexes, with a mean Von Bertalanffy growth coefficient of `r round(mean(used_data$k[used_data$targeted == 1]),2)` (standard deviation `r round(sd(used_data$k[used_data$targeted == 1]),2)`), and a median age at maturity of `r round(median(used_data$tm[used_data$targeted == 1]),0)` years (standard deviation `r round(sd(used_data$tm[used_data$targeted == 1]),0)`), and a mix of exploitation histories, ranging from low to high (but mostly relatively low) [@alonzo2004;@dick2017].
+While a wide range of MPA effects are plausible, from 0% to upwards of 150%, the median simulated effect size for this stylized version of the Channel Islands after over ten years of protection was `r percent(median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected > 10]))`.
 
@@
 
+Given the natural variability of marine ecosystems, how large of an effect would an MPA network have to have in order to allow a difference-in-difference strategy such as this to be a reliable measure of MPA effects?
+Our library of simulation results allows us to explore how MPA effects, and our ability to measure them, might vary with observation error and natural fish recruitment variation (including the possibility of recruitment regimes).
+Using a Bayesian difference-in-difference regression on these simulated data, we estimated the percent error between the posterior probability distribution of the estimated MPA effect from the regression and the true simulated MPA effect.
+
+The model struggled severely when MPA effect sizes were less than 25% and the model was faced with both observation and process errors (Fig.\@ref(fig:val-plot)).
+Even models fit to data generated from large effect sizes commonly mis-estimated the true MPA effect by 50% or more.
+Obtaining a mean absolute percent error (MAPE) of 25% or less across our simulated datasets required a population-level MPA effect of at least 30%.
+In the context of the Channel Islands, this finding suggests that we should not be surprised at our difficulty in precisely estimating the population-level effect of the MPAs.
+<!-- This is merely an illustrative exercise, omitting critical factors such as detection probability and sampling strategy.
+However, since nearly any omission which one can think of would make an MPA effect harder to detect, not easier, these results serve as a useful floor for the likely difficulty in estimating MPA effects.
+-->
+
+Two of the most critical drivers of MPA effect size are the size of the MPA network and the degree of fishing pressure.
+Looking across these two variables, based on our simulations the MPA network must be large (25% or more of a species range) and the target species overfished (pre-MPA depletion greater than about 60%) to achieve an effect size with a likely MAPE of 25% or less.
+Although recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, most MPA networks for near-shore commercial fin-fish are likely to cover areas more in line with the Channel Islands (20%) or smaller.
+As such, many MPA networks are expected to have population-level effect sizes that are difficult to detect unless target species were experiencing extreme overfishing prior to MPA establishment (Fig.\@ref(fig:sim-effects)-A).
+
 # Discussion {#discussion .unnumbered}
@@
 Much of the theoretical literature on MPAs assumes that larger reserves produce larger conservation gains [@white2011 and references therein].
-However, these models generally simulate fleet dynamics through fishing mortality rates (e.g.
-concentration of fishing mortality, @halpern2004).
+However, these models generally simulate fleet dynamics through fishing mortality rates, the proportion of total mortality experienced by a population attributable to fishing pressure [e.g.
+@halpern2004].
 The implication of this is that total catches will scale with the size of the available population.
@@
 
-Our finding of small MPA effects in recent years is therefore likely a result of both changing fishing dynamics and environmental disturbance.
-Given the natural variability of marine ecosystems, how large of an effect would an MPA network have to have in order to allow a difference-in-difference strategy such as this to be a reliable measure of MPA effects?
-Our library of simulation results allows us to explore how MPA effects, and our ability to measure them, might vary with observation error and natural fish recruitment variation (including the possibility of recruitment regimes).
-Using a Bayesian difference-in-difference regression on these simulated data, we estimated the percent error between the posterior probability distribution of the estimated MPA effect from the regression and the true simulated MPA effect.
-
-The model struggled severely when MPA effect sizes were less than 25% and the model was faced with both observation and process errors (Fig.\@ref(fig:val-plot)).
-Even models fit to data generated from large effect sizes commonly mis-estimated the true MPA effect by 50% or more.
-Obtaining a mean absolute percent error (MAPE) of 25% or less across our simulated datasets required a population-level MPA effect of at least 30%.
-In the context of the Channel Islands, this finding suggests that we should not be surprised at our difficulty in precisely estimating the population-level effect of the MPAs.
-<!-- This is merely an illustrative exercise, omitting critical factors such as detection probability and sampling strategy.
-However, since nearly any omission which one can think of would make an MPA effect harder to detect, not easier, these results serve as a useful floor for the likely difficulty in estimating MPA effects.
--->
-
-Two of the most critical drivers of MPA effect size are the size of the MPA network and the degree of fishing pressure.
-Looking across these two variables, based on our simulations the MPA network must be large (25% or more of a species range) and the target species overfished (pre-MPA depletion greater than about 60%) to achieve an effect size with a likely MAPE of 25% or less.
-Although recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, most MPA networks for near-shore commercial fin-fish are likely to cover areas more in line with the Channel Islands (20%) or smaller.
-As such, many MPA networks are expected to have population-level effect sizes that are difficult to detect unless target species were experiencing extreme overfishing prior to MPA establishment (Fig.\@ref(fig:sim-effects)-A).
+<!-- Our finding of small MPA effects in recent years is therefore likely a result of both changing fishing dynamics and environmental disturbance. Given the  -->
 
@@
 "Recruitment autocorrelation ($ac_{r}$)",
-"DD adult movement",
+"Density-dependent adult movement",
 "Density-dependence timing ",

### 2020-11-22 documents/ovando-et-al-pop-effects-of-mpas-si.Rmd
fixed broom.mixed dependency for summary tables

@@
 output: 
+  bookdown::word_document2:
+    reference_docx: template.docx
   bookdown::pdf_document2:
@@
 
+\renewcommand{\thefigure}{S\arabic{figure}}
+
+\setcounter{figure}{0}
+
+\renewcommand{\thetable}{S\arabic{table}}
+
+\setcounter{table}{0}
+
+\renewcommand{\theequation}{S\arabic{equation}}
+
+\setcounter{equation}{0}
+
 # Supporting Information (SI)
@@
 All analysis were performed in `r R.version.string`.
-Package versions are shown in Table.S\@ref(tab:pkg-info).
+Package versions are shown in Table.\@ref(tab:pkg-info).
 
@@
 
-## PISCO Data
-
-All fish data used in the primary difference-in-difference model were collected by PISCO. The dive transect survey methods are described in @caselle2015, provided below for ease of reference
-
-"Fish assemblages were surveyed annually as part of a long-term monitoring program conducted by the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) using standard underwater visual belt survey methods (www.piscoweb.org).
-We analyzed data from 47 PISCO sites at the northern Channel Islands that were sampled annually from at least 2003 to 2012.
-We also excluded 3 sites on Anacapa where a much older MPA had already been established in 1978.
-MPAs on each island were sampled annually during June-October and we surveyed multiple sites inside and outside of any individual MPA.
-Details of MPA characteristics such as size and coastline extent are given in Hamilton et al.
-At each site, we conducted 8 to 12 fish transects that measured 30×2×2m at multiple levels in the water column: benthic, midwater, and kelp canopy (when present).
-Transects are laid out in a stratified random design, with multiple nonpermanent transects located in fixed strata (i.e., outer, middle, and inner edges of the reef).
-At each level in the water column, one SCUBA diver per transect counted and estimated the sizes of all fish to the nearest centimeter (total length), excluding small cryptic fishes"
-- @caselle2015
-
-We take a number of steps to translate the raw transect data into the total biomass densities used in this study.
-For the default run, we only include species that were observed at least once for at least 15 of the 18 years of available data.
-We also exclude "young of the year" observations due the challenges of correctly identifying and measuring these individuals.
-We omitted data from 1999 due to changes in the sampling procedures that occurred after 1999.
-Per recommendations from PISCO staff we omit observations from the canopy level of the transects (leaving the middle, bottom, and middle canopy levels).
-
-PISCO data report positive observations of fish, in the manner of number of individuals of a species seen within a particular size group.
-In order to use these data in our model we need to add in zeros for any transect that could have observed a given species of fish but did not.
-We assume that a fish could have been observed on a given transect if that species has ever been observed at that site in any time period in the data (PISCO data are organized by sites, with multiple transect at different locations within the borders of a site).
-If a species has never been observed at a site we assume that it does not occur at that site.
-
-Once zeros have been introduced to the database, we convert positive observations of fish from numbers to biomass.
-Each observation in the raw database lists the species, the number of individuals seen, and the size of those individuals (either as one value or as a minimum and maximum size for the group seen).
-PISCO staff compiled allometric information used to convert lengths to expected weights.
-For each observation then, we convert the observed lengths to weights per these relationships (accounting for variations in length types such as standard vs.
-total length).
-When minimum and maximum ranges were reported, we drew a number of samples equal to the number of observed fish in that group from a uniform distribution spanning the minimum and maximum reported size in that group.
-We assume all length-to-weight conversions are constant and deterministic.
-
-For each species at each transect, we calculate the biomass density of that transect as the sum of the observed biomass divided by the transect area.
-We then average the biomass densities for each species across all the transects at a given site, and lastly sum these mean species biomass densities to achieve the total mean species biomass at the site level.
-
-We include several additional sources of data in our regression analysis.
-Temperature readings are included from the PISCO data for each transect.
-We also include PISCO data on the estimated surge and visibility.
-We augmented these data with information on kelp cover over time from the Santa Barbara Channel Long Term Ecological Research Network [@lter2017].
-We used a k-nearest neighbors algorithm to fill in missing kelp observations, and matched the interpolated kelp data to the PISCO data at the resolution of year-month-site (Fig.\@ref(fig:kelp-plot)).
-We include a variable capturing the mean cumulative number of observations across all observers conducting transects, in an effort to control for evolving observer skill.
-
-<!-- Temperature data were augmented with data from `FishLife` [@thorson2017c] to include the estimated preferred temperature for a given species, so that we can include deviations from the preferred temperature envelope as a predictor in the model.
-This allows different temperatures to have different effects on each species (and is less computationally intensive than estimating species-temperature slopes) (Fig.\@ref(fig:tempdev-plot)).
--->
-
-We also included lagged catch totals in the Santa Barbara region for the commercially harvest species in the database, in an effort to control for changes in density caused by changes in fishing pressure.
-Catches were pulled from the CDFW website (https://www.wildlife.ca.gov/Fishing/Commercial/Landings), and extracted using the `tabulizer` package in R [@leeper2018] (Fig.\@ref(fig:catches-plot)).
-
-```{r catches-plot, fig.cap = "Total CDFW reported commercial catches in the Santa Barbara region"}
-
-cdfw_catches %>% 
-  group_by(classcode) %>% 
-  mutate(nc = sum(catch > 0)) %>% 
-  ungroup() %>% 
-  filter(classcode %in% unique(used_data$classcode), nc > 13) %>% 
-  ggplot(aes(year, catch)) + 
-    geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
-  geom_line() + 
-  facet_wrap(~commonname, scales = "free_y") + 
-  theme_minimal() + 
-  labs(x = "Year", y = "Catch (lbs)")
-
-```
-
-```{r kelp-plot, fig.cap="Mean kelp biomass by island over time from SBC LTER"}
-
-used_data %>% 
-  group_by(region,year) %>% 
-  summarise(mean_kelp = mean(interp_kelp, na.rm = TRUE)) %>% 
-  ggplot(aes(year, (mean_kelp), color = region)) + 
-  geom_line() + 
-  labs(x = "Year", y = "Mean Kelp Biomass")
-
-```
-
-```{r targ-trend-plot, fig.cap="Centered and scaled mean biomass densities of all targeted finfish in analysis before filtering"}
-
-targeted_trend_plot <- pisco_data %>% 
-  filter(targeted == 1) %>% 
-  group_by(year, classcode, targeted, commonname) %>% 
-  summarise(mean_biomass_density = mean(density_g_m2, na.rm = TRUE)) %>% 
-  group_by(classcode) %>% 
-  mutate(mean_biomass_density = scale(mean_biomass_density)) %>% 
-  ggplot(aes(year,mean_biomass_density)) + 
-  geom_hline(aes(yintercept = 0), linetype = 2) + 
-  geom_vline(aes(xintercept = 2003), color = "red") +
-  geom_line() + 
-  facet_wrap(~commonname) +
-  theme_minimal() + 
-  theme(strip.text = element_text(size = 6),
-        axis.text.x = element_text(size = 8)) + 
-  labs(x = "Year", y = "Centered and Scaled Mean Biomass Density")
-
-targeted_trend_plot
-```
-
-```{r nontarg-trend-plot, fig.cap="Centered and scaled mean biomass densities of all non-targeted finfish in analysis before filtering"}
-
-nontargeted_trend_plot <- pisco_data %>% 
-  filter(targeted == 0) %>% 
-  group_by(year, classcode, targeted, commonname) %>% 
-  summarise(mean_biomass_density = mean(density_g_m2, na.rm = TRUE)) %>% 
-  group_by(classcode) %>% 
-  mutate(mean_biomass_density = scale(mean_biomass_density)) %>% 
-  ggplot(aes(year,mean_biomass_density)) + 
-  geom_hline(aes(yintercept = 0), linetype = 2) + 
-  geom_vline(aes(xintercept = 2003), color = "red") +
-  geom_line() + 
-  facet_wrap(~commonname) +
-  theme_minimal() + 
-  theme(strip.text = element_text(size = 6),
-        axis.text.x = element_text(size = 8)) + 
-  labs(x = "Year", y = "Centered and Scaled Mean Biomass Density")
-
-nontargeted_trend_plot
-```
-
-```{r classcode, eval = FALSE}
-
-species_data <- used_data %>% 
-  select(classcode, commonname, taxa, targeted) %>% 
-  unique() %>% 
-  arrange(desc(targeted)) %>% 
-  mutate(targeted = targeted == 1) %>% 
-  rename(`Common Name` = commonname, 
-         `Scientific Name` = taxa,
-         `Targeted?` = targeted)
-
-status <- read_csv(here::here("data","stock-status-summaries.csv")) %>% 
-  select(classcode, `Stock Status`)
-
-species_data <- species_data %>% 
-  left_join(status, by = "classcode")
-
-knitr::kable(species_data,
-             digits = 2,
-             caption = "Species included in estimation model",
-             longtable = TRUE,
-             booktabs = TRUE,
-             col.names = colnames(species_data)) %>%
-  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
-
-```
-
-## Difference-in-Difference Model
-
-The difference-in-difference (DiD) regression amounts to estimating the pre-post MPA difference in the biomass densities of targeted species minus the same difference for non-targeted species in the Channel Islands.
-
-The simplified form of this model is
-
-\begin{equation}
-  d_{i} \sim Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + \mathbf{B^cX_i} + \mathbf{B^sS_i}},shape, scale)
-\label{eq:did}
-\end{equation}
-
-The full list of included coefficients and their estimated values can be seen in Table.\@ref(tab:did-coefs).
-
-To provide greater detail on this process, we first conduct the data pre-processing described earlier.
-From there, we aggregated data to the level of total mean biomass density of targeted and non-targeted species at each site each year.
-As such, our base model estimates the effect of the MPAs on total mean biomass density of targeted species, though we also explore the effect on alternative specifications such as mean biomass of targeted species.
-the model was then fit using the `rstanarm` package in R (5000 iterations, 2500 warmup, 4 chains, adapt_delta = 0.85).
-
-The intercept prior was was determined using the `rstanarm` `autoscale` function.
-For the non-intercept terms, we manually set a normal(0,2) prior.
-This implies that coefficients such as the MPA effect have a prior that provides support for an effect size centered on zero with a range of roughly -4 to 4.
-Since the covariates are centered and scaled, and the dependent variable is on the log-scale, this is an extremely diffuse prior (e.g.
-a one standard deviation change in an independent variable changes the dependent variable by 4 on the log scale).
-
-We weight each observation such that within MPA data are assigned a total weight of 0.2 and outside-MPA data a total weight of 0.8, in order to represent the proportion of the available habitat inside MPAs and outside.
-This makes the assumption that all areas throughout that Channel Islands are identical in terms of habitat / carrying capacity for the species in question.
-If detailed data on habitat are available, this method could be modified to weight each area by the available habitat in that area.
-See Fig.\@ref(fig:no-weight-plot) for a companion analysis without weighting (which implicitly assumes that the MPAs contain more of the available habitat than the outside-MPA areas, since the MPAs make up a disproportionate amount of the samples relative to their physcial area).
-
-The full table of covariates and their posterior means and 89% credible interval can be found in Table.S\@ref(tab:did-coefs).
-"var" refers to variable, "_2" indicates a squared term.
-tex is the mean cumulative number of observations of the observers collecting the raw data, surge is the mean reported surge across transects, kelp is the mean kelp coverage at the site, temp is the water temperature, regional_temp_dev is the water temperature at a given site scaled by the mean water temperature at that site.
-Targeted is a Boolean marker indicating whether the given observation is for targeted or non-targeted species.
-
-```{r did-coefs, results = "asis",eval = TRUE}
-did_fit %>%
-  broom::tidy(intervals = TRUE, prob = 0.89) %>%
-  select(term, estimate, lower, upper) %>%
-  rename(
-    mean = estimate,
-    lower_89th_ci = lower,
-    upper_89th_ci = upper
-  ) %>%
-  knitr::kable(
-    digits = 2,
-    caption = "Posterior means and credible interval for key model coefficients",
-    longtable = TRUE,
-    booktabs = TRUE ) %>%
-  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
-
-```
-
-```{r prior-plot, fig.cap="Prior and posterior distributions of the estimated MPA effect over time"}
-
-priors <- tibble(year = unique(did_results$year)) %>% 
-  mutate(prior = list(rnorm(500,0,2))) %>% 
-  unnest(cols = prior)
-
-did_results %>% 
-  filter(between(prank, 0.05, 0.95)) %>% 
-  ggplot() + 
-  geom_density(aes(did, fill = "Posterior", y = ..scaled..), color = "transparent", alpha = 0.75) + 
-  geom_density(data = priors, aes(prior, fill = "prior", y = ..scaled..), color = "transparent",
-               alpha = 0.5) +
-  facet_wrap(~year) + 
-  scale_fill_discrete(name = '') + 
-  scale_y_continuous(name = "Scaled Density")
-
-```
-
-We include a series of standard visual diagnostic plot for Bayesian models below, with brief descriptions of the key outcome of the diagnostic included in the figure captions.
-The model had 0 divergences or max tree depth saturations.
-
-```{r, echo = FALSE}
-
-rstan::check_hmc_diagnostics(did_fit$stanfit)
-
-```
-
-```{r rhat-plot, fig.cap="Histogram of potential scale reduction statistic Rhat. All values are below 1.05, indicating there is not evidence of chain convergence failure"}
-
-data.frame(rhat = bayesplot::rhat(did_fit)) %>% 
-             ggplot(aes(rhat)) + 
-             geom_histogram() + 
-             geom_vline(aes(xintercept = 1.05))
-
-```
-
-```{r energy-plot, fig.cap = "HMC energy diagnostic plots. Closely matching histograms shows there is no evidence of excessively fat tails"}
-
-mcmc_nuts_energy(nuts_params(did_fit))
-
-```
-
-```{r}
-y <- (did_fit$y)
-
-yrep <- (rstanarm::posterior_predict(did_fit, draws = 500))
-```
-
-```{r, fig.cap = "Posterior Predictive distribution of total biomass density (blue lines) and observed distribution of total biomass density (red line)"}
-
-ppc_dens_overlay(y = y,
-                 yrep = yrep) +
-  scale_x_continuous(name = "Total Biomass Density") +
-  scale_colour_discrete(labels = c("Observed", "Posterior Predictive"),
-                       name = '')
-
-```
-
-```{r mean_test_plot, fig.cap = "Posterior predictive mean (distribution) and empirical mean of total biomass density. Posterior predictive mean adequately covers empirical mean."}
-
-ppc_stat(y,  yrep, stat = "mean") 
-```
-
-```{r sum_test_plot, fig.cap = "Posterior predictive sum (distribution) and empirical sum of total biomass density. Posterior predictive sum adequately covers empirical sum."}
-
-ppc_stat(y,  yrep, stat = "sum") 
-```
-
-```{r sd_test_plot, fig.cap = "Posterior predictive standard deviation (distribution) and empirical standard deviation of total biomass density.
-Posterior predictive standard deviation adequately covers empirical standard deviation"}
-ppc_stat(y,  yrep, stat = "sd")
-
-```
-
-```{r min_test_plot, fig.cap = "Posterior predictive minimum (distribution) and empirical minimum of log total biomass density. Model slightly overpredicts minimum value in the data"}
-
-ppc_stat(y,  yrep, stat = "min") 
-```
-
-```{r max_test_plot, fig.cap = "Posterior predictive maximum (distribution) and empirical maximum of total biomass density. Model underestimates true maximum of the data."}
-ppc_stat(y,  yrep, stat = "max")
-
-```
-
-```{r resid-plot,fig.cap = "Posterior predictive total biomass densities plotted against mean posterior predictive residuals" }
-ppc_error_scatter_avg_vs_x(y, yrep, colMeans(yrep)) + 
-  scale_x_continuous(name = "Predicted total biomass density") + 
-  scale_y_continuous(name = "Mean Posterior Predictive Residuals") + 
-  geom_hline(yintercept = 0)
-```
-
-```{r resid-hist-plot, fig.cap = "Histograms of residuals for targeted and non-targeted obesrvations."}
-
-temp <- did_data %>% 
-  mutate(targ = ifelse(targeted == 1, "Targeted","Non-Targeted"))
-ppc_error_hist_grouped(y, yrep = yrep[1:2,], group = temp$targ) + 
-  geom_vline(aes(xintercept = 0)) + 
-  scale_x_continuous(name = "Residuals")
-```
-
-### Additional Difference-in-difference Runs
-
-We include a variety of additional model runs designed to explore the sensitivity of our key results to various assumptions and data processing steps included in our base results.
-Description and key message are provided inside figure captions.
-
-Our base run uses data collected by PISCO.
-As a robustness check to our main results, we repeated our analysis utilizing data provided by the [Kelp Forest Monitoring Program (KFM)](https://science.nature.nps.gov/im/units/medn/monitor/kelpforest.cfm) conducted in the Channel Islands.
-
-We include one run using only variables available in both datasets, to explore sensitivity of our results to selected covariates (Fig.\@ref(fig:kfm-data)).
-
-```{r kfm-data, fig.cap="Model fit to PISCO data but only using variables also available for KFM data", results="hide"}
-
-val_runs <- model_runs %>% 
-  mutate(did_results = map(did_fit,"did_results"))
-
-ggplot() +
- tidybayes::stat_halfeye(
-    data = val_runs$did_results[[2]],
-    aes(year, did, color = "Empirical Estimate"),
-    .width = c(0.5, 0.95),
-    slab_alpha = 0.95  ) +
-  scale_y_continuous(name = "MPA Effect") +
-  scale_x_discrete(name = element_blank()) +
-  theme(legend.position = "top") +
-  scale_fill_brewer(name = "Simulation Quantiles") + 
-  scale_color_manual(values = "red", name = '') + 
-  geom_hline(yintercept = 0)
-
-```
-
-We fit an alternative model using data from the KFM program to test whether a different dataset from the same region provides different results.
-KFM based results are qualitatively similar to those from PISCO data (Fig.\@ref(fig:kfm-fit))
-
-```{r kfm-fit, fig.cap="Model fit to KFM data. Results are more uncertain than using PISCO data, but follow the same trend estimated from the PISCO data."}
-
-ggplot() +
- tidybayes::stat_halfeye(
-    data = val_runs$did_results[[3]],
-    aes(year, did, color = "Empirical Estimate"),
-    .width = c(0.5, 0.95),
-    slab_alpha = 0.95  ) +
-  scale_y_continuous(name = "MPA Effect") +
-  scale_x_discrete(name = element_blank()) +
-  theme(legend.position = "top") +
-  scale_fill_brewer(name = "Simulation Quantiles") + 
-  scale_color_manual(values = "red", name = '') + 
-    geom_hline(yintercept = 0)
-
-```
-
-We ran a version of the model using data from only inside MPAs.
-This serves as a test of whether our results are being entirely driven by the outside-MPA data, and allows us to compare results from inside MPAs to those outside (Fig.\@ref(fig:mpa-fit-plot)).
-
-```{r mpa-fit-plot, fig.cap="Estimated MPA effects using PISCO data from inside MPAs only.
-The base model uses data from both inside and outside MPAs.
-These results show that the same general trend holds when only using data from inside the MPAs, though the model estimates greater probability of higher positive effects with only MPA data, compared to the model fit using all the data.",results = "hide"}
-
-ggplot() +
- tidybayes::stat_halfeye(
-    data = val_runs$did_results[[4]],
-    aes(year, did, color = "Empirical Estimate"),
-    .width = c(0.5, 0.95),
-    slab_alpha = 0.95  ) +
-  scale_y_continuous(name = "MPA Effect") +
-  scale_x_discrete(name = element_blank()) +
-  theme(legend.position = "top") +
-  scale_fill_brewer(name = "Simulation Quantiles") + 
-  scale_color_manual(values = "red", name = '') + 
-  geom_hline(yintercept = 0)
-
-```
-
-We ran a version of the model using data from only outside MPAs.
-This serves as a test of whether our results are being entirely driven by the inside-MPA data, and allows us to compare results from inside MPAs to those outside (Fig.\@ref(fig:fished-fit-plot)).
-
-```{r fished-fit-plot, fig.cap="Estimated MPA effects using PISCO data from outside MPAs only.
-The base model uses data from both inside and outside MPAs.
-These results show that the same general trend holds when only using data from outside the MPAs, though the model estimates less probability of high positive effects witout MPA data, compared to the model fit using all the data.",results = "hide"}
-
-ggplot() +
- tidybayes::stat_halfeye(
-    data = val_runs$did_results[[6]],
-    aes(year, did, color = "Empirical Estimate"),
-    .width = c(0.5, 0.95),
-    slab_alpha = 0.95  ) +
-  scale_y_continuous(name = "MPA Effect") +
-  scale_x_discrete(name = element_blank()) +
-  theme(legend.position = "top") +
-  scale_fill_brewer(name = "Simulation Quantiles") + 
-  scale_color_manual(values = "red", name = '') + 
-    geom_hline(yintercept = 0)
-
-```
-
-We include a model run using nearly all observed species, to test whether our results are being influenced by omission of rarely observed species (Fig.\@ref(fig:no-filter-plot)).
-
-```{r no-filter-plot, fig.cap="The base model only includes consistently observed species of finfish.
-For this run, we include species that have been seen in at least two years (down from fifteen).
-The results remain essentially unchanged.",results = "hide"}
-no_filter <- model_runs %>%
-  slice(1) %>%
-  left_join(abundance_data, by = "data_source") %>%
-  mutate(
-    did_fit = pmap(
-      list(
-        data = data,
-        data_to_use = data_to_use,
-        data_source = data_source
-      ),
-      estimate_did,
-      min_classcode_years = 2,
-      iter = 2500,
-      site_data = site_data,
-      cdfw_catches = cdfw_catches,
-      life_history_data = life_history_data
-    )
-  )
-
-ggplot() +
- tidybayes::stat_halfeye(
-    data = no_filter$did_fit[[1]]$did_results,
-    aes(year, did, color = "Empirical Estimate"),
-    .width = c(0.5, 0.95),
-    slab_alpha = 0.95  ) +
-  scale_y_continuous(name = "MPA Effect") +
-  scale_x_discrete(name = element_blank()) +
-  theme(legend.position = "top") +
-  scale_fill_brewer(name = "Simulation Quantiles") + 
-  scale_color_manual(values = "red", name = '') + 
-    geom_hline(yintercept = 0)
-
-```
-
-```{r no-weight-plot, fig.cap="Version of the model fit without reweighting samples in proportion to area",results = "hide"}
-no_weight <- model_runs %>%
-  slice(1) %>%
-  left_join(abundance_data, by = "data_source") %>%
-  mutate(
-    did_fit = pmap(
-      list(
-        data = data,
-        data_to_use = data_to_use,
-        data_source = data_source
-      ),
-      estimate_did,
-      iter = 2500,
-      site_data = site_data,
-      cdfw_catches = cdfw_catches,
-      life_history_data = life_history_data,
-      weight_samples = FALSE
-    )
-  )
-
-ggplot() +
- tidybayes::stat_halfeye(
-    data = no_weight$did_fit[[1]]$did_results,
-    aes(year, did, color = "Empirical Estimate"),
-    .width = c(0.5, 0.95),
-    slab_alpha = 0.95  ) +
-  scale_y_continuous(name = "MPA Effect") +
-  scale_x_discrete(name = element_blank()) +
-  theme(legend.position = "top") +
-  scale_fill_brewer(name = "Simulation Quantiles") + 
-  scale_color_manual(values = "red", name = '') + 
-    geom_hline(yintercept = 0)
-
-```
-
-For ease of interpretation and model convergence, our default model estimates MPA effects in three-year bins.
-For this run we estimate the effect annually, relative to the year 2000.
-Results show the same overall trend and magnitude reported in our base model (Fig.\@ref(fig:no-bin-plot)).
-
-```{r no-bin-plot, fig.cap="For ease of interpretation and model convergence, our default model estimates MPA effects in three-year bins.
-For this run we estimate the effect annually, relative to the year 2000.
-Results show the same overall trend and magnitude reported in our base model.", results = "hide", eval = TRUE}
-no_bin <- model_runs %>%
-  slice(1) %>%
-  left_join(abundance_data, by = "data_source") %>%
-  mutate(
-    did_fit = pmap(
-      list(
-        data = data,
-        data_to_use = data_to_use,
-        data_source = data_source
-      ),
-      estimate_did,
-      iter = 2500,
-      first_year = 2001,
-      bin_width = 1,
-      site_data = site_data,
-      cdfw_catches = cdfw_catches,
-      life_history_data = life_history_data
-    )
-  )
-
-ggplot() +
- tidybayes::stat_halfeye(
-    data = no_bin$did_fit[[1]]$did_results,
-    aes(year, did, color = "Empirical Estimate"),
-    .width = c(0.5, 0.95),
-    slab_alpha = 0.95  ) +
-  scale_y_continuous(name = "MPA Effect") +
-  scale_x_discrete(name = element_blank()) +
-  theme(legend.position = "top") +
-  scale_fill_brewer(name = "Simulation Quantiles") + 
-  scale_color_manual(values = "red", name = '') + 
-    scale_x_discrete(guide = guide_axis(n.dodge = 2))
-
-```
-
-Our base run estimates the effect of MPAs on total biomass densities of targeted finfish.
-For this run, we instead estimate the effect on mean biomass densities of targeted species.
-Results show the same pattern as our base run (Fig.\@ref(fig:mean-fit)).
-
-```{r mean-fit, fig.cap="Our base run estimates the effect of MPAs on total biomass densities of targeted finfish.
-For this run, we instead estimate the effect on mean biomass densities of targeted species.
-Results show the same pattern as our base run.", results="hide"}
-mean_fit <- model_runs %>%
-  slice(1) %>%
-  left_join(abundance_data, by = "data_source") %>%
-  mutate(
-    did_fit = pmap(
-      list(
-        data = data,
-        data_to_use = data_to_use,
-        data_source = data_source
-      ),
-      estimate_did,
-      fit_mean = TRUE,
-      iter = 2500,
-      site_data = site_data,
-      cdfw_catches = cdfw_catches,
-      life_history_data = life_history_data,
-      refresh = 1000
-    )
-  )
-
-ggplot() +
- tidybayes::stat_halfeye(
-    data = mean_fit$did_fit[[1]]$did_results,
-    aes(year, did, color = "Empirical Estimate"),
-    .width = c(0.5, 0.95),
-    slab_alpha = 0.95  ) +
-  scale_y_continuous(name = "MPA Effect") +
-  scale_x_discrete(name = element_blank()) +
-  theme(legend.position = "top") +
-  scale_fill_brewer(name = "Simulation Quantiles") + 
-  scale_color_manual(values = "red", name = '') + 
-    scale_x_discrete(guide = guide_axis(n.dodge = 2))
-
-```
-
-Appropriately addressing the problem of "missing" observations is a critical challenge in any field observation study.
-If no observations of a given fish species were recorded on a given transect, should the density of that species on that transect be marked as zero, and influence the estimate of the overall mean density accordingly?
-The obvious answer seems to be yes, but what if that species simply does not live in the environment covered by a particular transect, or was not present during the particular time of the diver's observation?
-For our base runs, we assign a value of zero density on a given transect for any fish species that has been observed at least once at a given site at any time in our data but was not observed on that particular transect.
-If that species was never observed at that site, we do not include a zero for that species.
-Our rationale for this is that given the shifting nature of the sampled sites, and the intensity of sampling at those sites, we do not want to skew density trends by changes in the amount of suitable habitat for a given species sampled.
-However, this is clearly a strong assumption.
-For example, perhaps the decreasing trend in mean densities from 2000 to 2004 is due to increased number of sites (and therefore zeros) included in the data.
-To assess the potential importance of this choice, we can compare the mean densities of targeted and non-targeted species over time with the added zeros to the mean densities using only positive observations (i.e.
-not including any zeros in the data, (Fig.S\@ref(fig:no-zero-fit)).
-The trends in the raw densities, and most importantly the mean trends of targeted and non-targeted fishes, are nearly identical whether or not zeros are added, providing strong evidence that our choice of how to incorporate missing observations into the data are not strongly influencing our overall results.
-
-```{r no-zero-fit, fig.cap="Our base run estimates the effect of MPAs on total biomass densities of targeted finfish.
-For this run, we instead estimate the effect on mean biomass densities of targeted species.
-Results show the same pattern as our base run.", results = "hide"}
-
-nozero_fit <- model_runs %>%
-  slice(1) %>%
-  left_join(abundance_data, by = "data_source")
-
-nozero_fit$data[[1]] <- nozero_fit$data[[1]] %>% filter(any_seen == TRUE)
-
-nozero_fit <-  nozero_fit%>%
-  mutate(
-    did_fit = pmap(
-      list(
-        data = data,
-        data_to_use = data_to_use,
-        data_source = data_source
-      ),
-      estimate_did,
-      iter = 2500,
-      site_data = site_data,
-      cdfw_catches = cdfw_catches,
-      life_history_data = life_history_data
-    )
-  )
-
-ggplot() +
- tidybayes::stat_halfeye(
-    data = nozero_fit$did_fit[[1]]$did_results,
-    aes(year, did, color = "Empirical Estimate"),
-    .width = c(0.5, 0.95),
-    slab_alpha = 0.95  ) +
-  scale_y_continuous(name = "MPA Effect") +
-  scale_x_discrete(name = element_blank()) +
-  theme(legend.position = "top") +
-  scale_fill_brewer(name = "Simulation Quantiles") + 
-  scale_color_manual(values = "red", name = '') + 
-    scale_x_discrete(guide = guide_axis(n.dodge = 2))
-
-```
-
-### Synthetic controls
-
-Synthetic controls are an alternative method for attempting to estimate the causal effect of a policy intervention [@abadie2010].
-A difference-in-difference approach assumes that some observable group serves as an adequate control for the state of the treated group in an untreated world.
-In our default case, we assume that the mean standardized index of non-targeted species are our control for the targeted species.
-Alternatively, synthetic controls use timeseries of treated and non-treated groups before and after treatment to construct a new "control" group built by weighting the pre-treatment timeseries of un-treated observations (together with covariates) such that the synthetic control group matches the trends in the treated group pre-treatment.
-
-We chose to present difference-in-difference as our main result since it better allows us to capture the uncertainty in the data generating process through our hierarchical model.
-However, we felt that it was worth exploring whether synthetic controls provided substantially different results than our default model.
-
-For the first synthetic control, we pulled our standardized mean index of abundance for targeted species as a whole from our difference-in-difference model as our treated group.
-We then pull the standardized indices of abundance for each of the non-targeted groups from the difference-in-difference to use as the candidate untreated components for the synthetic controls.
-A complete synthetic control analysis would require more extensive validation of the methods, but we use this approach simply to explore whether we observe substantially divergent results in the synthetic control versus the difference-in-difference model.
-
-We centered and scaled the candidate abundance indices to facilitate model convergence given the very few number of pre-treatment years available.
-The results of a synthetic control model are presented as the difference between the observed treatment outcome and the synthetic control (the difference in this case being in units of standard deviations).
-The model was not able to construct an adequate synthetic control at this level, as shown by the differences between the treated group and the synthetic control pre-treatment.
-However, we would note that the post-treatment results do show similarities with our main results, namely a lack of a clear divergence between the treatment and the control, and an upwards trend up through the earl 2010s followed by a decline (Fig.S\@ref(fig:total-synth)).
-
-```{r, include = FALSE}
-
-years <- unique(used_data$year)
-
-life_history <- used_data %>%
-  select(classcode) %>%
-  unique() %>%
-  mutate(numeric_classcode = as.numeric(as.factor(classcode)))
-
-nontargeted_index <- data_frame(abundance_hat = fit_report$abundance_hat,
-                               classcode = rep(1:n_distinct(used_data$classcode), each = length(years))) %>%
-  mutate(log_abundance_hat = log(abundance_hat)) %>%
-  group_by(classcode) %>%
-  mutate(year =years) %>%
-  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
-  ungroup() %>%
-  rename(numeric_classcode = classcode) %>%
-  left_join(life_history, by = "numeric_classcode" ) %>% 
-  left_join(unique(abundance_data$data[[1]] %>% select(classcode, targeted)), by = "classcode") %>% 
-  filter(targeted == 0) %>% 
-  select(log_abundance_hat, year, classcode, targeted)
-
-targeted_index <- tmb_runs$tmb_fit[[1]]$zissou_estimates %>%
-  filter(
-    stringr::str_detect(variable, "(targeted_did_betas)"),!str_detect(variable, "non")
-  ) %>%
-  mutate(
-    year = years,
-    classcode = "targeted",
-    targeted = 1
-  ) %>%
-  rename(log_abundance_hat = estimate) %>% 
-  select(colnames(nontargeted_index))
-
-  # filter(variable == "seeing_year_species_betas") %>%
-
-nontargeted_covariates <- used_data %>% 
-  select(classcode, contains("mean"),year, targeted) %>% 
-  filter(targeted == 0) %>% 
-  select(-mean_length) %>% 
-  gather(variable, value, -classcode,-year,-targeted) %>% 
-  group_by(classcode, variable, year) %>% 
-  summarise(mean_value = mean(value, na.rm = TRUE)) %>% 
-  spread(variable, mean_value)
-
-targeted_covariates <- used_data %>% 
-  select(classcode, contains("mean"),year, targeted) %>% 
-  filter(targeted == 1) %>% 
-  mutate(classcode = "targeted") %>% 
-  select(-mean_length) %>% 
-  gather(variable, value, -classcode,-year,-targeted) %>% 
-  group_by(classcode, variable, year) %>% 
-  summarise(mean_value = mean(value, na.rm = TRUE)) %>% 
-  spread(variable, mean_value)
-
-synth_covariates <- nontargeted_covariates %>% 
-  bind_rows(targeted_covariates)
-
-synth_data <- targeted_index %>% 
-  bind_rows(nontargeted_index) %>% 
-  left_join(synth_covariates, by = c("classcode","year")) %>% 
-    mutate(numeric_classcode = as.numeric(factor(classcode))) %>% 
-  group_by(classcode) %>% 
-  mutate(log_abundance_hat = scale(exp(log_abundance_hat))) %>% 
-  ungroup() %>% 
-  as.data.frame() %>% 
-  select(-rolling_mean)
-
-# synth_data %>%
-#   group_by(classcode) %>%
-#   mutate(index = (log_abundance_hat)) %>%
-#   ggplot(aes(
-#     year,
-#     index,
-#     color = factor(targeted),
-#     group = interaction(targeted, classcode)
-#   )) +
-#   geom_line()
-
-prepped_synth_data <-
-  Synth::dataprep(
-    foo = synth_data,
-    predictors = colnames(synth_data)[str_detect(colnames(synth_data),"mean")],
-    predictors.op = "mean",
-    dependent = "log_abundance_hat",
-    unit.variable = "numeric_classcode",
-    unit.names.variable = "classcode",
-    time.variable = "year",
-    treatment.identifier = "targeted",
-    controls.identifier = unique(synth_data$classcode)[unique(synth_data$classcode) != "targeted"],
-    time.predictors.prior = years[years <= 2003],
-    time.optimize.ssr = years[years <= 2003],
-    time.plot = years
-  )
-
-synth_fit <- Synth::synth(prepped_synth_data)
-
-gaps <-
-  rownames_to_column(
-    prepped_synth_data$Y1plot - (prepped_synth_data$Y0plot %*% synth_fit$solution.w) %>%
-      as.data.frame()
-  ) %>%
-  rename(gap = w.weight,
-         year = rowname) %>%
-  mutate(year = as.numeric(year)) %>%
-  mutate(synthetic = as.numeric(prepped_synth_data$Y0plot %*% synth_fit$solution.w)) %>% 
-  mutate(treatment = gap + synthetic)
-
-# path.plot(synth.res = synth_fit,
-#           dataprep.res = prepped_synth_data
-# )
-
-total_gap_plot <- gaps %>% 
-  ggplot(aes(year, gap)) + 
-  geom_hline(aes(yintercept = 0), linetype = 2) +
-  geom_vline(aes(xintercept = 2003), color = "red")+
-  geom_line() + 
-    labs(x = "Year", y = "Standard deviations from synthetic control")
-
-total_synth_plot <- gaps %>% 
-  select(year, synthetic, treatment) %>% 
-  gather(source, value, -year) %>% 
-  group_by(source) %>% 
-  # mutate(value = scale(value)) %>% 
-  ggplot(aes(year, value, color = source, linetype = source)) + 
-  geom_vline(aes(xintercept = 2003), color = "red")+
-  geom_line() + 
-    labs(x = "Year", y = "Standard deviations from synthetic control")
-
-```
-
-```{r  total-synth, fig.cap = "Difference in centered and scaled standardized targeted abundance and synthetic standardized targeted abundance"}
-total_gap_plot
-```
-
-```{r raw-synth, include = FALSE, eval = FALSE}
-
-  consistent_classcodes <- used_data %>%
-      group_by(classcode, region) %>%
-      filter(any_seen) %>%
-      summarise(n_years = n_distinct(year)) %>%
-      ungroup() %>%
-      filter(n_years == n_distinct(used_data$year))
-
-synth_data <- used_data %>% 
-  filter(classcode %in% consistent_classcodes$classcode,
-         region %in% consistent_classcodes$region) %>% 
-  mutate(classcode = ifelse(targeted == 1, "targeted", classcode),
-         region = ifelse(targeted == 1, "targeted", region)) %>% 
-      group_by(year,
-               site_side,
-               region,
-               zone,
-               transect,
-               eventual_mpa,
-               classcode,
-               targeted) %>%
-      summarise(
-        total_classcode_density = sum(exp(log_density)),
-        var_tex = sum(cumulative_n_obs),
-        var_vis = mean(mean_vis),
-        var_temp = mean(mean_temp),
-        var_depth = mean(mean_depth),
-        var_surge = mean(surge),
-        var_kelp = mean(interp_kelp)
-      ) %>%  # sum density across all levels of a transect
-      group_by(year, site_side, region, eventual_mpa, classcode, targeted) %>%
-      summarise(
-        md = mean(total_classcode_density),
-        var_tex = mean(var_tex),
-        var_vis = mean(var_vis),
-        var_temp = mean(var_temp),
-        var_depth = mean(var_depth),
-        var_surge = mean(var_surge),
-        var_kelp = mean(var_kelp)
-      ) %>% # calculate mean density per year site, side, species, averaging over zone, transect
-      ungroup()
-
- synth_data <- synth_data %>% 
-     group_by(year,region,classcode,targeted) %>%
-      summarise(
-        total_biomass_density = (sum(md) / 1e6) * 10000,
-        # calculate total and mean biomass densities across all species per year site side
-        mean_biomass_density = log((mean(md) / 1e6) * 10000 + 1e-3),
-        var_tex = mean(var_tex),
-        var_vis = mean(var_vis),
-        var_temp = mean(var_temp),
-        var_kelp = mean(var_kelp)
-      ) %>%
-   ungroup() %>% 
-   mutate(classcode = ifelse(targeted == 1, classcode, paste(classcode, region, sep = '_'))) %>% 
-      mutate(numeric_classcode = as.numeric(factor(classcode))) %>% 
-   as.data.frame()
-
-prepped_synth_data <-
-  Synth::dataprep(
-    foo = synth_data,
-    predictors = colnames(synth_data)[str_detect(colnames(synth_data),"var")],
-    predictors.op = "mean",
-    dependent = "mean_biomass_density",
-    unit.variable = "numeric_classcode",
-    unit.names.variable = "classcode",
-    time.variable = "year",
-    treatment.identifier = "targeted",
-    controls.identifier = unique(synth_data$classcode[synth_data$targeted == 0]),
-    time.predictors.prior = years[years <= 2003],
-    time.optimize.ssr = years[years <= 2003],
-    time.plot = years
-  )
-
-synth_fit <- Synth::synth(prepped_synth_data)
-
-gaps <-
-  rownames_to_column(
-    prepped_synth_data$Y1plot - (prepped_synth_data$Y0plot %*% synth_fit$solution.w) %>%
-      as.data.frame()
-  ) %>%
-  rename(gap = w.weight,
-         year = rowname) %>%
-  mutate(year = as.numeric(year)) %>%
-  mutate(synthetic = as.numeric(prepped_synth_data$Y0plot %*% synth_fit$solution.w)) %>% 
-  mutate(treatment = gap + synthetic)
-
-# path.plot(synth.res = synth_fit,
-#           dataprep.res = prepped_synth_data
-# )
-
-total_gap_plot <- gaps %>% 
-  ggplot(aes(year, gap)) + 
-  # geom_hline(aes(yintercept = 0), linetype = 2) +
-  geom_vline(aes(xintercept = 2003), color = "red")+
-  geom_line() + 
-    labs(x = "Year", y = "Standard deviations from synthetic control")
-
-total_synth_plot <- gaps %>% 
-  select(year, synthetic, treatment) %>% 
-  gather(source, value, -year) %>% 
-  group_by(source) %>% 
-  # mutate(value = scale(value)) %>% 
-  ggplot(aes(year, value, color = source, linetype = source)) + 
-  geom_vline(aes(xintercept = 2003), color = "red")+
-  geom_line() + 
-    labs(x = "Year", y = "Standard deviations from synthetic control")
-
-```
-
-As an extension, we repeated this process, but now treating each targeted species individually as the treated group, and the non-targeted species as the non-targeted.
-This is intended to explore whether we see clearer signals for individual species than we do for the targeted class as a whole.
-
-Overall we see similarly unclear results as the aggregate targeted synthetic control (and our main results).
-The synthetic control was better constructed for some individual species, but not clearly for any one, and most species showed some evidence of the upward-then-downward trend seen throughout our results (Fig.S\@ref(fig:synth-classcode-fits)).
-
-```{r, include = FALSE}
-
-synth_foo <-
-  function(treated_classcode,
-           synth_data,
-           targeted_classcodes) {
-    # treated_classcode <- targeted_classcodes[1]
-
-    temp_synth_data <- synth_data %>%
-      filter(classcode == treated_classcode |
-               !(classcode %in% targeted_classcodes))
-
-    prepped_synth_data <-
-      Synth::dataprep(
-        foo = temp_synth_data,
-        predictors = colnames(temp_synth_data)[str_detect(colnames(temp_synth_data), "mean")],
-        predictors.op = "mean",
-        dependent = "log_abundance_hat",
-        unit.variable = "numeric_classcode",
-        unit.names.variable = "classcode",
-        time.variable = "year",
-        treatment.identifier = treated_classcode,
-        controls.identifier = unique(temp_synth_data$classcode)[unique(temp_synth_data$classcode) != treated_classcode],
-        time.predictors.prior = c(2000:2003),
-        time.optimize.ssr = c(2000:2003),
-        time.plot = years
-      )
-
-    synth_fit <- Synth::synth(prepped_synth_data)
-
-    gaps <-
-      rownames_to_column(
-        prepped_synth_data$Y1plot - (prepped_synth_data$Y0plot %*% synth_fit$solution.w) %>%
-          as.data.frame()
-      ) %>%
-      rename(gap = w.weight,
-             year = rowname) %>%
-      mutate(year = as.numeric(year)) %>%
-      mutate(synthetic = as.numeric(prepped_synth_data$Y0plot %*% synth_fit$solution.w)) %>%
-      mutate(treatment = gap + synthetic)
-  }
-
-nontargeted_index <- data_frame(abundance_hat = fit_report$abundance_hat,
-                               classcode = rep(1:n_distinct(used_data$classcode), each = length(years))) %>%
-  mutate(log_abundance_hat = log(abundance_hat)) %>%
-  group_by(classcode) %>%
-  mutate(year = 1999 + 1:length(abundance_hat)) %>%
-  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
-  ungroup() %>%
-  rename(numeric_classcode = classcode) %>%
-  left_join(life_history, by = "numeric_classcode" ) %>% 
-  left_join(unique(abundance_data$data[[1]] %>% select(classcode, targeted)), by = "classcode") %>% 
-  select(log_abundance_hat, year, classcode, targeted)
-
-synth_covariates <- pisco_data %>% 
-  select(classcode, contains("mean"),year, targeted) %>% 
-  select(-mean_length) %>% 
-  gather(variable, value, -classcode,-year,-targeted) %>% 
-  group_by(classcode, variable, year) %>% 
-  summarise(mean_value = mean(value, na.rm = TRUE)) %>% 
-  spread(variable, mean_value)
-
-synth_data <- nontargeted_index %>% 
-  left_join(synth_covariates, by = c("classcode","year")) %>% 
-    mutate(numeric_classcode = as.numeric(factor(classcode))) %>% 
-  group_by(classcode) %>% 
-  mutate(log_abundance_hat = scale(exp(log_abundance_hat))) %>% 
-  ungroup() %>% 
-  as.data.frame()
-
-targeted_classcodes <- unique(synth_data$classcode[synth_data$targeted == 1])
-
-synth_classcode_fits <-
-  tibble(targeted_classcodes = targeted_classcodes) %>%
-  mutate(
-    synthetic_fit = map(
-      targeted_classcodes,
-      synth_foo,
-      synth_data = synth_data,
-      targeted_classcodes = targeted_classcodes
-    )
-  )
-
-synth_classcode_fits <- synth_classcode_fits %>% 
-  unnest()
-
-synth_classcode_fits_plot <- 
-synth_classcode_fits %>% 
-  ggplot(aes(year, gap)) + 
-  geom_hline(aes(yintercept = 0), linetype = 2) +
-  geom_vline(aes(xintercept = 2003), color = "red")+
-  geom_line() + 
-  facet_wrap(~targeted_classcodes) + 
-  labs(x = "Year", y = "Standard deviations from synthetic control")
-
-```
-
-```{r synth-classcode-fits, fig.cap="Synthetic control gaps for each targeted species"}
-synth_classcode_fits_plot + 
-  theme_minimal()
-```
-
-## Testing Model Assumptions
-
-### Simulation testing
-
-We state that a difference-in-difference model using targeted and non-targeted species is capable (conditional on assumptions) of estimating the causal effect of MPAs.
-We simulated MPA outcomes to test this claim.
-We first test our estimation strategy under idealized circumstances, where recruitment is deterministic and PISCO divers all have constant and perfect observer skills.
-We simulate five species that vary only in their maximum size and length at maturity.
-For each of these species, we set one version that is targeted by fishing and one that is not.
-We set a constant fishing mortality rate for each simulated targeted species, and then ran two matched simulations, one with MPAs and one without.
-We then have our simulated divers sample data from each of these scenarios, and then pass the sampled biomass densities to a simplified version of our difference-in-difference model (omitting the probability of detection step).
-We can then compare the difference-in-difference estimates of the MPA effect to the true simulated effect.
-The difference in difference model is able to capture the simulated MPA effect under these circumstances (Fig.S\@ref(fig:simple-did-test))
-
-```{r simple-did-test, fig.cap = "Simulated mean (red dashed line) and individual species (solid lines) MPA effects over time, along with difference-in-difference estimated MPA effects (mean with 95% confidence intervals)"}
-
-load(file = file.path(run_dir,'simulated_did.Rdata'))
-
-simple_performance$mixed_effect_did + 
-  labs(title = '', x = "Years with MPA protection",
-       y = "MPA Effect")
-```
-
-We then simulated a more complex example.
-We use the actual targeted and non-targeted species from our model.
-We assign species predominately seen in the western Channel Islands as "cold water" and those in the eastern Channel Islands as "warm water".
-We allow for stochasticity in recruitment.
-We use El Niño data as a simulated environmental recruitment driver, where we assume that El Niño events produce negative recruitment shocks for cold water species and *vice versa* for warm water species.
-We simulate three different divers each with different base skill levels, visual selectivities, and an evolving skill rate (such that observers get better over time).
-We hold fishing mortality rates constant for each species, although that fishing mortality affects each species differently because of intrinsic biological differences in maturity-at-age and steepness.
-We then test the ability of the difference-in-difference model to isolate the mean MPA effect across all of these targeted species, which our results show it is capable of doing (Fig.S\@ref(fig:complex-did-test)).
-
-```{r complex-did-test,  fig.cap = "Simulated mean (red dashed line) and individual species (solid lines) MPA effects over time, along with difference-in-difference estimated MPA effects (mean with 95% confidence intervals)"}
-
-pisco_performance$mixed_effect_did + 
-    labs(title = '', x = "Years with MPA protection",
-       y = "MPA Effect")
-
-```
-
-### Testing SUTVA with Convergent Cross Mapping
-
-The difference-in-difference model also assumes that the targeted and non-targeted fishes do not directly or indirectly affect each other.This assumption is clearly violated on some level: all the fishes in this analysis are part of the same ecosystem and therefore interact to some degree.
-For example, if the protection of targeted predatory fishes results in increased mortality of non-targeted fishes, the model would attribute that as an increased population effect (greater divergence between the abundance of targeted and non-targeted species).
-Given the time scale of analysis (15 years of protection), we do not feel that massive trophic cascades are likely to have developed yet, given both the pace and complexity of trophic cascade development [@babcock2010; @pershing2015a].
-A complete assessment of evidence for trophic cascades in the Channel Islands is beyond the scope of this study, but to address this question somewhat we utilized convergent cross mapping *sensu* @sugihara2012 to test for a significant causal signal between different broad trophic groups in the data, implemented in the `rEDM` package in R.
-
-Convergent cross mapping is a nonlinear forecasting method that uses observed time series data to test for significant causal links between variables.
-Following methods laid out in @clark2015 and @sugihara2012, we pool the abundance of each broad trophic group by region (Fig.S\@ref(fig:trophic-plot).
-This uses the data from the islands as "replicates", requiring the assumption that the islands are all part of the same dynamic system, but allowing us to take advantage of the extra information provided by each island to further resolve the reconstructed manifolds.
-Using these aggregations, we then test whether the variables can be properly embedded, i.e., if they have predictable manifold dynamics.
-We do this through a simplex forecasting test, using an individual timeseries' own lags to build a manifold.
-For each timeseries, the "best embedding dimension" is an approximation of the dimensionality of the dynamic system, in other words, the number of dimensions that define and predict the evolving states of the timeseries.
-This analysis shows that only the carnivore, piscivore, and planktivores groups show evidence of significant predictability (that is, that past dynamics of these species groups can predict future dynamics, Fig.S\@ref(fig:embed-plot)).
-
-```{r trophic-plot, fig.cap="Centered and scaled densities by broad trophic group and island over time"}
-cip_data <- used_data
-trophic_groups <- used_data %>% 
-  select(classcode, broadtrophic) %>% 
-  unique()
-
-# dat <- abundance_trends %>%
-#   left_join(trophic_groups, by = "classcode") %>%
-#   group_by(broadtrophic, year) %>%
-#   summarise(density = mean(scaled_abundance_hat)) %>%
-#   group_by(broadtrophic) %>%
-#   mutate(density = scale(density)) %>%
-#   ungroup() %>% 
-#   spread(broadtrophic, density) %>% 
-#   arrange(year) %>% 
-#   mutate(lag4_carnivore = lag(carnivore,4),
-#          lag4_piscovore = lag(piscivore,4)) %>% 
-#   na.omit() %>% 
-#   ungroup() %>% 
-#   filter(year > 1999) 
-# dat %>%
-#   ggplot(aes(year, density, color = broadtrophic)) +
-#   geom_line()
-
-dat <- cip_data %>%
-  group_by(broadtrophic, region, site, side, transect, year, eventual_mpa) %>%
-  summarise(density = sum(density_g_m2)) %>%
-  group_by(broadtrophic, region, site, year) %>%
-  summarise(mean_density = mean(density, na.rm = T)) %>%
-  group_by(broadtrophic,region, year) %>%
-  summarise(regional_density = mean(mean_density, na.rm = T)) %>% 
-  ungroup() %>% 
-  spread(broadtrophic, regional_density) %>% 
-  group_by(region) %>% 
-  arrange(year) %>% 
-  mutate(lag4_carnivore = lag(carnivore,4),
-         lag4_piscovore = lag(piscivore,4)) %>% 
-  na.omit() %>% 
-  ungroup() %>% 
-  filter(year > 1999)
-
- dat <- dat %>%
-  select(-lag4_carnivore,-lag4_piscovore)%>%
-  group_by(region)%>%
-  mutate_at(vars(carnivore:planktivore),funs(norm=(.-mean(.))/sd(.)))->dat
-
-group_trends_plot <- dat %>%
-  select(region,year,contains("norm"))%>%
-  gather("group","density",-region,-year)%>%
-  ggplot(aes(year,density,col=group))+
-  geom_line()+
-  geom_point() +
-  scale_color_locuszoom(name="Trophic Group",labels=c("carnivore","herbivore","piscivore","planktivore"))+
-  labs(x="year",y="Centered and Scaled Density")+
-  facet_wrap(~region)
-
-group_trends_plot
-```
-
-```{r embed-plot, fig.cap = "Predictive skill as a function of embedding dimensions"}
-
-## split data by region to analyze timeseries from each island separately
-ana.dat <- dat %>% filter(region=="ANA")
-sci.dat <- dat %>% filter(region=="SCI")
-smi.dat <- dat %>% filter(region=="SMI")
-sri.dat <- dat %>% filter(region=="SRI")
-
-datnest <- dat %>% group_by(region) %>% nest()
-
-datnorm <- dat %>% select(region,year,contains('norm')) %>% ungroup()
-
-# have to record the segments corresponding to each "replicate" so simplex algorithm does not try to make predictions crossing time barriers
-segs <- datnorm %>% mutate(ind=row_number()) %>% group_by(region) %>% summarise(first=first(ind),last=last(ind)) %>%
-  select(-region)
-
-var_names <- c("carnivore_norm","herbivore_norm","piscivore_norm","planktivore_norm")
-
-regions.combined.simp.list <- map(var_names,function(x){
-  temp <- datnorm %>% ungroup() %>% select(matches(x)) %>% as.data.frame()
-  out <- simplex(as.numeric(temp[,1]),E=1:10,lib=as.matrix(segs),silent=T) %>%
-    mutate(trophic=str_replace(x,"_norm",""))
-  out
-})
-
-embed_plot <- bind_rows(regions.combined.simp.list) %>%
-  ggplot(aes(E,rho,color=trophic))+
-  geom_line(size=2)+
-  facet_wrap(~trophic,nrow=2,scales="free_y")+
-  geom_hline(yintercept = 0,color="black")+
-  labs(x="Embedding Dimension (E)",y=expression(paste("Predictive Skill, ",rho)))+
-  scale_x_continuous(breaks=seq(0,12,by=2))+
-  scale_color_locuszoom()+
-  guides(color=F)
-
-embed_plot
-
-```
-
-Focusing on just these three groups then (removing herbivores), we can test for causal relationships between groups using convergent cross mapping and the logic of Takens' theorem of dynamic systems.
-Generalizations of Takens' theorem indicate that if two variables (in our case, species or physical variables) are part of the same dynamic system, their individual dynamics should reflect their relative causal influence [@sugihara2012].
-In other words, if one variable is causally forced by another, that forcing should leave a signature on the first time series.
-Convergent cross mapping (CCM) tests for causation by using the attractor/manifold built from the time series of one variable to predict another (hence the "cross-mapping").
-In simple terms, the *causal effect of A on B is determined by how well B cross-maps A*.
-
-There are two criteria for CCM to establish causality:  First, and most obviously, predictive cross-map skill using all available data should be significantly greater than zero.
-Second, that predictability should be convergent.
-Convergence means that cross-mapped estimates improve with library length (the number of state-space vectors used to build the attractor), because the attractor is more fully resolved and therefore estimation error should decline.
-Convergence is key to distinguishing causation from simple or spurious correlation.
-If two variables are spuriously correlated and not causally linked, CCM should fail to satisfy this second criterion.
-Based on these criteria, there is little evidence of significant dynamic interactions between trophic groups (Fig.S\@ref(fig:cross-map-cp)-\@ref(fig:cross-map-pis)).
-Cross-mapping produced positive predictive skill, but was non-convergent for all cross-mappings with the exception of carnivores cross-mapping planktivores (providing some evidence that planktivore dynamics may be a driver of carnivore dynamics).
-This analysis provides evidence that trophic cascades are unlikely to be a significant driver of our results.
-It is important to note though that this analysis does not mean that trophic cascades could not emerge in this system, rather that we do not detect them with these data at this time.
-
-```{r cross-map-fxn}
-# plots a cross mapping between two variables in the data, given their variable names (grp), their optimal embedding dimensions (from simplex forecasting), and their preferred names for plotting
-xmap <- function(grp1,grp2,E1,E2,name1,name2){
-    # cross-map group 1 to group 2
-
-  ccm1 <- ccm(datnorm,lib=as.matrix(segs),pred=as.matrix(segs),E=E1,lib_column= grp1,target_column = grp2,lib_sizes = c(10,25,50,75),num_samples=100,replace=T,silent=T,RNGseed = 41389)
-
-  # test for convergence (is predictability at max library size significantly greater than at minimum library size?)
-
-  rhomeans1 <- ccm1 %>% ccm_means()
-
-  conv_test1 <- t.test(ccm1$rho[ccm1$lib_size==10],ccm1$rho[ccm1$lib_size==75])$p.value
-  if((rhomeans1$rho[rhomeans1$lib_size==75]-rhomeans1$rho[rhomeans1$lib_size==10])<0.05) conv_test1 <- 1
-  conv_test1 <- ifelse(conv_test1<0.05,"Convergent","Non-convergent")
-
-  xmap1 <- ccm1 %>%
-    group_by(lib_size)%>%
-    summarise(rhomean=mean(rho,na.rm=T),upper=quantile(rho, 0.975),lower=quantile(rho, 0.025))%>%
-    ungroup()%>%
-    ggplot(aes(lib_size,rhomean))+
-    geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.3,fill="red")+
-    geom_line(color="darkorchid3")+
-    geom_hline(aes(yintercept = 0), linetype = 2) +
-    labs(x="Library Size",y=expression(paste(rho, " (predictive skill)")),title=paste0(name1," xmap ",name2,"\n","(",conv_test1,")"))+
-    theme(plot.title = element_text(size=6))
-
-  # cross map group 2 to group 1
-  # inspect the output of simplex from the previous step and use the best embedding dimension (highest rho) for the carnivore time series
-
-  ccm2 <- ccm(datnorm,lib=as.matrix(segs),pred=as.matrix(segs),E=E2,lib_column= grp2,target_column = grp1,lib_sizes = c(10,25,50,75),num_samples=100,replace=T,silent=T,RNGseed = 41389)
-
-  rhomeans2 <- ccm2 %>% ccm_means()
-
-  conv_test2 <- t.test(ccm2$rho[ccm2$lib_size==10],ccm2$rho[ccm2$lib_size==75])$p.value
-  if((rhomeans2$rho[rhomeans2$lib_size==75]-rhomeans2$rho[rhomeans2$lib_size==10])<0.05) conv_test2 <- 1
-  conv_test2 <- ifelse(conv_test2<0.05,"Convergent","Non-convergent")
-
-  xmap2 <- ccm2 %>%
-    group_by(lib_size)%>%
-    summarise(rhomean=mean(rho,na.rm=T),upper=quantile(rho, 0.975),lower=quantile(rho, 0.025))%>%
-    ungroup()%>%
-    ggplot(aes(lib_size,rhomean))+
-    geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.3,fill="red")+
-    geom_line(color="darkorchid3")+
-    geom_hline(aes(yintercept = 0), linetype = 2) +
-    labs(x="Library Size",y=expression(paste(rho, " (predictive skill)")),title=paste0(name2," xmap ",name1,"\n","(",conv_test2,")"))+
-    theme(plot.title = element_text(size=6))
-
-  cross_map_plot <- xmap1 + xmap2
-
-cross_map_plot
-}
-```
-
-```{r cross-map-cp, fig.cap = "Cross mapping of effect of piscivores on carnivores (A) and carnivores on piscivores (B) in the PISCO data from 2000 to 2017.
-Shaded region show 95% confidence interval"}
-
-car_pis<-xmap(grp1="carnivore_norm",grp2="piscivore_norm",E1=4,E2=4,name1="Carnivores",name2="Piscivores")
-car_pl <- xmap(grp1="carnivore_norm",grp2="planktivore_norm",E1=4,E2=4,name1="Carnivores",name2="Planktivores")
-pl_pis <- xmap(grp1="planktivore_norm",grp2="piscivore_norm",E1=4,E2=4,name1="Planktivores",name2="Piscivores")
-
-car_pis
-```
-
-```{r cross-map-pl, fig.cap = "Cross mapping of effect of planktivores on carnivores (A) and carnivores on planktivores (B) in the PISCO data from 2000 to 2017.
-Shaded region show 95% confidence interval"}
-
-car_pl
-
-```
-
-```{r cross-map-pis, fig.cap = "Cross mapping of effect of piscivores on carnivores (A) and carnivores on planktivores (B) in the PISCO data from 2000 to 2017.
-Shaded region show 95% confidence interval"}
-
- pl_pis 
-```
-
-```{r trophic-cor-plot, fig.cap="Estimated correlation coefficients between broad trophic groups at the island by year resolution", eval = FALSE}
-
-trophic_effects_plot <- cip_data %>%
-  group_by(broadtrophic, region, site, side, transect, year, eventual_mpa) %>%
-  summarise(density = sum(density_g_m2)) %>%
-  group_by(broadtrophic, region, site, year) %>%
-  summarise(mean_density = mean(density, na.rm = T)) %>%
-  group_by(broadtrophic,region, year) %>%
-  summarise(regional_density = mean(mean_density, na.rm = T)) %>% 
-  ungroup() %>% 
-  spread(broadtrophic, regional_density) %>% 
-  group_by(region) %>% 
-  arrange(year) %>% 
-  mutate(lag4_carnivore = lag(carnivore,4),
-         lag4_piscovore = lag(piscivore,4)) %>% 
-  na.omit() %>% 
-  ungroup() %>% 
-  filter(year > 1999) %>% 
-  select(-year,-region) %>% 
-  corrr::correlate() %>% 
-   gather("colname","correlation",-rowname) %>% 
-  ggplot(aes(rowname, colname,fill = correlation)) + 
-  geom_tile() + 
-  geom_text(aes(rowname, colname, label = round(correlation,2))) + 
-  scale_fill_gradient2(low = "tomato", mid = "white", high = "steelblue", midpoint = 0, limits = c(-1,1),
-                       guide = guide_colorbar(frame.colour = "black",barheight = 13)) + 
-  theme(axis.text.x = element_text(angle = 45, vjust = 1,
-                                   hjust = 1),
-        axis.title.x = element_blank(),
-        axis.title.y = element_blank())
-
-trophic_effects_plot
-
-```
-
-## Operating Model
+## Simulation Model
 
@@
 
-Table.S1 - Range of simulated variables
+Table.1 - Range of simulated variables
 
@@
 
+## PISCO Data
+
+All fish data used in the primary difference-in-difference model were collected by PISCO. The dive transect survey methods are described in @caselle2015, provided below for ease of reference
+
+"Fish assemblages were surveyed annually as part of a long-term monitoring program conducted by the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) using standard underwater visual belt survey methods (www.piscoweb.org).
+We analyzed data from 47 PISCO sites at the northern Channel Islands that were sampled annually from at least 2003 to 2012.
+We also excluded 3 sites on Anacapa where a much older MPA had already been established in 1978.
+MPAs on each island were sampled annually during June-October and we surveyed multiple sites inside and outside of any individual MPA.
+Details of MPA characteristics such as size and coastline extent are given in Hamilton et al.
+At each site, we conducted 8 to 12 fish transects that measured 30×2×2m at multiple levels in the water column: benthic, midwater, and kelp canopy (when present).
+Transects are laid out in a stratified random design, with multiple nonpermanent transects located in fixed strata (i.e., outer, middle, and inner edges of the reef).
+At each level in the water column, one SCUBA diver per transect counted and estimated the sizes of all fish to the nearest centimeter (total length), excluding small cryptic fishes"
+- @caselle2015
+
+We take a number of steps to translate the raw transect data into the total biomass densities used in this study.
+For the default run, we only include species that were observed at least once for at least 15 of the 18 years of available data.
+We also exclude "young of the year" observations due the challenges of correctly identifying and measuring these individuals.
+We omitted data from 1999 due to changes in the sampling procedures that occurred after 1999.
+Per recommendations from PISCO staff we omit observations from the canopy level of the transects (leaving the middle, bottom, and middle canopy levels).
+
+PISCO data report positive observations of fish, in the manner of number of individuals of a species seen within a particular size group.
+In order to use these data in our model we need to add in zeros for any transect that could have observed a given species of fish but did not.
+We assume that a fish could have been observed on a given transect if that species has ever been observed at that site in any time period in the data (PISCO data are organized by sites, with multiple transect at different locations within the borders of a site).
+If a species has never been observed at a site we assume that it does not occur at that site.
+
+Once zeros have been introduced to the database, we convert positive observations of fish from numbers to biomass.
+Each observation in the raw database lists the species, the number of individuals seen, and the size of those individuals (either as one value or as a minimum and maximum size for the group seen).
+PISCO staff compiled allometric information used to convert lengths to expected weights.
+For each observation then, we convert the observed lengths to weights per these relationships (accounting for variations in length types such as standard vs.
+total length).
+When minimum and maximum ranges were reported, we drew a number of samples equal to the number of observed fish in that group from a uniform distribution spanning the minimum and maximum reported size in that group.
+We assume all length-to-weight conversions are constant and deterministic.
+
+For each species at each transect, we calculate the biomass density of that transect as the sum of the observed biomass divided by the transect area.
+We then average the biomass densities for each species across all the transects at a given site, and lastly sum these mean species biomass densities to achieve the total mean species biomass at the site level.
+
+We include several additional sources of data in our regression analysis.
+Temperature readings are included from the PISCO data for each transect.
+We also include PISCO data on the estimated surge and visibility.
+We augmented these data with information on kelp cover over time from the Santa Barbara Channel Long Term Ecological Research Network [@lter2017].
+We used a k-nearest neighbors algorithm to fill in missing kelp observations, and matched the interpolated kelp data to the PISCO data at the resolution of year-month-site (Fig.\@ref(fig:kelp-plot)).
+We include a variable capturing the mean cumulative number of observations across all observers conducting transects, in an effort to control for evolving observer skill.
+
+<!-- Temperature data were augmented with data from `FishLife` [@thorson2017c] to include the estimated preferred temperature for a given species, so that we can include deviations from the preferred temperature envelope as a predictor in the model.
+This allows different temperatures to have different effects on each species (and is less computationally intensive than estimating species-temperature slopes) (Fig.\@ref(fig:tempdev-plot)).
+-->
+
+We also included lagged catch totals in the Santa Barbara region for the commercially harvest species in the database, in an effort to control for changes in density caused by changes in fishing pressure.
+Catches were pulled from the CDFW website (https://www.wildlife.ca.gov/Fishing/Commercial/Landings), and extracted using the `tabulizer` package in R [@leeper2018] (Fig.\@ref(fig:catches-plot)).
+
+```{r catches-plot, fig.cap = "Total CDFW reported commercial catches in the Santa Barbara region"}
+
+cdfw_catches %>% 
+  group_by(classcode) %>% 
+  mutate(nc = sum(catch > 0)) %>% 
+  ungroup() %>% 
+  filter(classcode %in% unique(used_data$classcode), nc > 13) %>% 
+  ggplot(aes(year, catch)) + 
+    geom_vline(aes(xintercept = 2003), linetype = 2, color = "red") +
+  geom_line() + 
+  facet_wrap(~commonname, scales = "free_y") + 
+  theme_minimal() + 
+  labs(x = "Year", y = "Catch (lbs)")
+
+```
+
+```{r kelp-plot, fig.cap="Mean kelp biomass by island over time from SBC LTER"}
+
+used_data %>% 
+  group_by(region,year) %>% 
+  summarise(mean_kelp = mean(interp_kelp, na.rm = TRUE)) %>% 
+  ggplot(aes(year, (mean_kelp), color = region)) + 
+  geom_line() + 
+  labs(x = "Year", y = "Mean Kelp Biomass")
+
+```
+
+```{r targ-trend-plot, fig.cap="Centered and scaled mean biomass densities of all targeted finfish in analysis before filtering"}
+
+targeted_trend_plot <- pisco_data %>% 
+  filter(targeted == 1) %>% 
+  group_by(year, classcode, targeted, commonname) %>% 
+  summarise(mean_biomass_density = mean(density_g_m2, na.rm = TRUE)) %>% 
+  group_by(classcode) %>% 
+  mutate(mean_biomass_density = scale(mean_biomass_density)) %>% 
+  ggplot(aes(year,mean_biomass_density)) + 
+  geom_hline(aes(yintercept = 0), linetype = 2) + 
+  geom_vline(aes(xintercept = 2003), color = "red") +
+  geom_line() + 
+  facet_wrap(~commonname) +
+  theme_minimal() + 
+  theme(strip.text = element_text(size = 6),
+        axis.text.x = element_text(size = 8)) + 
+  labs(x = "Year", y = "Centered and Scaled Mean Biomass Density")
+
+targeted_trend_plot
+```
+
+```{r nontarg-trend-plot, fig.cap="Centered and scaled mean biomass densities of all non-targeted finfish in analysis before filtering"}
+
+nontargeted_trend_plot <- pisco_data %>% 
+  filter(targeted == 0) %>% 
+  group_by(year, classcode, targeted, commonname) %>% 
+  summarise(mean_biomass_density = mean(density_g_m2, na.rm = TRUE)) %>% 
+  group_by(classcode) %>% 
+  mutate(mean_biomass_density = scale(mean_biomass_density)) %>% 
+  ggplot(aes(year,mean_biomass_density)) + 
+  geom_hline(aes(yintercept = 0), linetype = 2) + 
+  geom_vline(aes(xintercept = 2003), color = "red") +
+  geom_line() + 
+  facet_wrap(~commonname) +
+  theme_minimal() + 
+  theme(strip.text = element_text(size = 6),
+        axis.text.x = element_text(size = 8)) + 
+  labs(x = "Year", y = "Centered and Scaled Mean Biomass Density")
+
+nontargeted_trend_plot
+```
+
+```{r classcode, eval = FALSE}
+
+species_data <- used_data %>% 
+  select(classcode, commonname, taxa, targeted) %>% 
+  unique() %>% 
+  arrange(desc(targeted)) %>% 
+  mutate(targeted = targeted == 1) %>% 
+  rename(`Common Name` = commonname, 
+         `Scientific Name` = taxa,
+         `Targeted?` = targeted)
+
+status <- read_csv(here::here("data","stock-status-summaries.csv")) %>% 
+  select(classcode, `Stock Status`)
+
+species_data <- species_data %>% 
+  left_join(status, by = "classcode")
+
+knitr::kable(species_data,
+             digits = 2,
+             caption = "Species included in estimation model",
+             longtable = TRUE,
+             booktabs = TRUE,
+             col.names = colnames(species_data)) %>%
+  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
+
+```
+
+## Difference-in-Difference Model
+
+The difference-in-difference (DiD) regression amounts to estimating the pre-post MPA difference in the biomass densities of targeted species minus the same difference for non-targeted species in the Channel Islands.
+
+The simplified form of this model is
+
+\begin{equation}
+  d_{i} \sim Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + \mathbf{B^cX_i} + \mathbf{B^sS_i}},shape, scale)
+\label{eq:did}
+\end{equation}
+
+The full list of included coefficients and their estimated values can be seen in Table.\@ref(tab:did-coefs).
+
+To provide greater detail on this process, we first conduct the data pre-processing described earlier.
+From there, we aggregated data to the level of total mean biomass density of targeted and non-targeted species at each site each year.
+As such, our base model estimates the effect of the MPAs on total mean biomass density of targeted species, though we also explore the effect on alternative specifications such as mean biomass of targeted species.
+the model was then fit using the `rstanarm` package in R (5000 iterations, 2500 warmup, 4 chains, adapt_delta = 0.85).
+
+The intercept prior was was determined using the `rstanarm` `autoscale` function.
+For the non-intercept terms, we manually set a normal(0,2) prior.
+This implies that coefficients such as the MPA effect have a prior that provides support for an effect size centered on zero with a range of roughly -4 to 4.
+Since the covariates are centered and scaled, and the dependent variable is on the log-scale, this is an extremely diffuse prior (e.g.
+a one standard deviation change in an independent variable changes the dependent variable by 4 on the log scale).
+
+We weight each observation such that within MPA data are assigned a total weight of 0.2 and outside-MPA data a total weight of 0.8, in order to represent the proportion of the available habitat inside MPAs and outside.
+This makes the assumption that all areas throughout that Channel Islands are identical in terms of habitat / carrying capacity for the species in question.
+If detailed data on habitat are available, this method could be modified to weight each area by the available habitat in that area.
+See Fig.\@ref(fig:no-weight-plot) for a companion analysis without weighting (which implicitly assumes that the MPAs contain more of the available habitat than the outside-MPA areas, since the MPAs make up a disproportionate amount of the samples relative to their physcial area).
+
+The full table of covariates and their posterior means and 89% credible interval can be found in Table.\@ref(tab:did-coefs).
+"var" refers to variable, "_2" indicates a squared term.
+tex is the mean cumulative number of observations of the observers collecting the raw data, surge is the mean reported surge across transects, kelp is the mean kelp coverage at the site, temp is the water temperature, regional_temp_dev is the water temperature at a given site scaled by the mean water temperature at that site.
+Targeted is a Boolean marker indicating whether the given observation is for targeted or non-targeted species.
+
+```{r did-coefs, results = "asis",eval = TRUE}
+did_fit %>%
+  broom.mixed::tidy(conf.int = TRUE, prob = 0.89) %>%
+    rename(lower = conf.low,
+         upper = conf.high) %>% 
+  select(term, estimate, lower, upper) %>%
+  rename(
+    mean = estimate,
+    lower_89th_ci = lower,
+    upper_89th_ci = upper
+  ) %>%
+  knitr::kable(
+    digits = 2,
+    caption = "Posterior means and credible interval for key model coefficients",
+    longtable = TRUE,
+    booktabs = TRUE ) %>%
+  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
+
+```
+
+```{r prior-plot, fig.cap="Prior and posterior distributions of the estimated MPA effect over time"}
+
+priors <- tibble(year = unique(did_results$year)) %>% 
+  mutate(prior = list(rnorm(500,0,2))) %>% 
+  unnest(cols = prior)
+
+did_results %>% 
+  filter(between(prank, 0.05, 0.95)) %>% 
+  ggplot() + 
+  geom_density(aes(did, fill = "Posterior", y = ..scaled..), color = "transparent", alpha = 0.75) + 
+  geom_density(data = priors, aes(prior, fill = "prior", y = ..scaled..), color = "transparent",
+               alpha = 0.5) +
+  facet_wrap(~year) + 
+  scale_fill_discrete(name = '') + 
+  scale_y_continuous(name = "Scaled Density")
+
+```
+
+We include a series of standard visual diagnostic plot for Bayesian models below, with brief descriptions of the key outcome of the diagnostic included in the figure captions.
+The model had 0 divergences or max tree depth saturations.
+
+```{r, echo = FALSE}
+
+rstan::check_hmc_diagnostics(did_fit$stanfit)
+
+```
+
+```{r rhat-plot, fig.cap="Histogram of potential scale reduction statistic Rhat. All values are below 1.05, indicating there is not evidence of chain convergence failure"}
+
+data.frame(rhat = bayesplot::rhat(did_fit)) %>% 
+             ggplot(aes(rhat)) + 
+             geom_histogram() + 
+             geom_vline(aes(xintercept = 1.05))
+
+```
+
+```{r energy-plot, fig.cap = "HMC energy diagnostic plots. Closely matching histograms shows there is no evidence of excessively fat tails"}
+
+mcmc_nuts_energy(nuts_params(did_fit))
+
+```
+
+```{r}
+y <- (did_fit$y)
+
+yrep <- (rstanarm::posterior_predict(did_fit, draws = 500))
+```
+
+```{r, fig.cap = "Posterior Predictive distribution of total biomass density (blue lines) and observed distribution of total biomass density (red line)"}
+
+ppc_dens_overlay(y = y,
+                 yrep = yrep) +
+  scale_x_continuous(name = "Total Biomass Density") +
+  scale_colour_discrete(labels = c("Observed", "Posterior Predictive"),
+                       name = '')
+
+```
+
+```{r mean_test_plot, fig.cap = "Posterior predictive mean (distribution) and empirical mean of total biomass density. Posterior predictive mean adequately covers empirical mean."}
+
+ppc_stat(y,  yrep, stat = "mean") 
+```
+
+```{r sum_test_plot, fig.cap = "Posterior predictive sum (distribution) and empirical sum of total biomass density. Posterior predictive sum adequately covers empirical sum."}
+
+ppc_stat(y,  yrep, stat = "sum") 
+```
+
+```{r sd_test_plot, fig.cap = "Posterior predictive standard deviation (distribution) and empirical standard deviation of total biomass density.
+Posterior predictive standard deviation adequately covers empirical standard deviation"}
+ppc_stat(y,  yrep, stat = "sd")
+
+```
+
+```{r min_test_plot, fig.cap = "Posterior predictive minimum (distribution) and empirical minimum of log total biomass density. Model slightly overpredicts minimum value in the data"}
+
+ppc_stat(y,  yrep, stat = "min") 
+```
+
+```{r max_test_plot, fig.cap = "Posterior predictive maximum (distribution) and empirical maximum of total biomass density. Model underestimates true maximum of the data."}
+ppc_stat(y,  yrep, stat = "max")
+
+```
+
+```{r resid-plot,fig.cap = "Posterior predictive total biomass densities plotted against mean posterior predictive residuals" }
+ppc_error_scatter_avg_vs_x(y, yrep, colMeans(yrep)) + 
+  scale_x_continuous(name = "Predicted total biomass density") + 
+  scale_y_continuous(name = "Mean Posterior Predictive Residuals") + 
+  geom_hline(yintercept = 0)
+```
+
+```{r resid-hist-plot, fig.cap = "Histograms of residuals for targeted and non-targeted obesrvations."}
+
+temp <- did_data %>% 
+  mutate(targ = ifelse(targeted == 1, "Targeted","Non-Targeted"))
+ppc_error_hist_grouped(y, yrep = yrep[1:2,], group = temp$targ) + 
+  geom_vline(aes(xintercept = 0)) + 
+  scale_x_continuous(name = "Residuals")
+```
+
+### Additional Difference-in-difference Runs
+
+We include a variety of additional model runs designed to explore the sensitivity of our key results to various assumptions and data processing steps included in our base results.
+Description and key message are provided inside figure captions.
+
+Our base run uses data collected by PISCO.
+As a robustness check to our main results, we repeated our analysis utilizing data provided by the [Kelp Forest Monitoring Program (KFM)](https://science.nature.nps.gov/im/units/medn/monitor/kelpforest.cfm) conducted in the Channel Islands.
+
+We include one run using only variables available in both datasets, to explore sensitivity of our results to selected covariates (Fig.\@ref(fig:kfm-data)).
+
+```{r kfm-data, fig.cap="Model fit to PISCO data but only using variables also available for KFM data", results="hide"}
+
+val_runs <- model_runs %>% 
+  mutate(did_results = map(did_fit,"did_results"))
+
+ggplot() +
+ tidybayes::stat_halfeye(
+    data = val_runs$did_results[[2]],
+    aes(year, did, color = "Empirical Estimate"),
+    .width = c(0.5, 0.95),
+    slab_alpha = 0.95  ) +
+  scale_y_continuous(name = "MPA Effect") +
+  scale_x_discrete(name = element_blank()) +
+  theme(legend.position = "top") +
+  scale_fill_brewer(name = "Simulation Quantiles") + 
+  scale_color_manual(values = "red", name = '') + 
+  geom_hline(yintercept = 0)
+
+```
+
+We fit an alternative model using data from the KFM program to test whether a different dataset from the same region provides different results.
+KFM based results are qualitatively similar to those from PISCO data (Fig.\@ref(fig:kfm-fit))
+
+```{r kfm-fit, fig.cap="Model fit to KFM data. Results are more uncertain than using PISCO data, but follow the same trend estimated from the PISCO data."}
+
+ggplot() +
+ tidybayes::stat_halfeye(
+    data = val_runs$did_results[[3]],
+    aes(year, did, color = "Empirical Estimate"),
+    .width = c(0.5, 0.95),
+    slab_alpha = 0.95  ) +
+  scale_y_continuous(name = "MPA Effect") +
+  scale_x_discrete(name = element_blank()) +
+  theme(legend.position = "top") +
+  scale_fill_brewer(name = "Simulation Quantiles") + 
+  scale_color_manual(values = "red", name = '') + 
+    geom_hline(yintercept = 0)
+
+```
+
+We ran a version of the model using data from only inside MPAs.
+This serves as a test of whether our results are being entirely driven by the outside-MPA data, and allows us to compare results from inside MPAs to those outside (Fig.\@ref(fig:mpa-fit-plot)).
+
+```{r mpa-fit-plot, fig.cap="Estimated MPA effects using PISCO data from inside MPAs only.
+The base model uses data from both inside and outside MPAs.
+These results show that the same general trend holds when only using data from inside the MPAs, though the model estimates greater probability of higher positive effects with only MPA data, compared to the model fit using all the data.",results = "hide"}
+
+ggplot() +
+ tidybayes::stat_halfeye(
+    data = val_runs$did_results[[4]],
+    aes(year, did, color = "Empirical Estimate"),
+    .width = c(0.5, 0.95),
+    slab_alpha = 0.95  ) +
+  scale_y_continuous(name = "MPA Effect") +
+  scale_x_discrete(name = element_blank()) +
+  theme(legend.position = "top") +
+  scale_fill_brewer(name = "Simulation Quantiles") + 
+  scale_color_manual(values = "red", name = '') + 
+  geom_hline(yintercept = 0)
+
+```
+
+We ran a version of the model using data from only outside MPAs.
+This serves as a test of whether our results are being entirely driven by the inside-MPA data, and allows us to compare results from inside MPAs to those outside (Fig.\@ref(fig:fished-fit-plot)).
+
+```{r fished-fit-plot, fig.cap="Estimated MPA effects using PISCO data from outside MPAs only.
+The base model uses data from both inside and outside MPAs.
+These results show that the same general trend holds when only using data from outside the MPAs, though the model estimates less probability of high positive effects witout MPA data, compared to the model fit using all the data.",results = "hide"}
+
+ggplot() +
+ tidybayes::stat_halfeye(
+    data = val_runs$did_results[[6]],
+    aes(year, did, color = "Empirical Estimate"),
+    .width = c(0.5, 0.95),
+    slab_alpha = 0.95  ) +
+  scale_y_continuous(name = "MPA Effect") +
+  scale_x_discrete(name = element_blank()) +
+  theme(legend.position = "top") +
+  scale_fill_brewer(name = "Simulation Quantiles") + 
+  scale_color_manual(values = "red", name = '') + 
+    geom_hline(yintercept = 0)
+
+```
+
+We include a model run using nearly all observed species, to test whether our results are being influenced by omission of rarely observed species (Fig.\@ref(fig:no-filter-plot)).
+
+```{r no-filter-plot, fig.cap="The base model only includes consistently observed species of finfish.
+For this run, we include species that have been seen in at least two years (down from fifteen).
+The results remain essentially unchanged.",results = "hide"}
+no_filter <- model_runs %>%
+  slice(1) %>%
+  left_join(abundance_data, by = "data_source") %>%
+  mutate(
+    did_fit = pmap(
+      list(
+        data = data,
+        data_to_use = data_to_use,
+        data_source = data_source
+      ),
+      estimate_did,
+      min_classcode_years = 2,
+      iter = 2500,
+      site_data = site_data,
+      cdfw_catches = cdfw_catches,
+      life_history_data = life_history_data
+    )
+  )
+
+ggplot() +
+ tidybayes::stat_halfeye(
+    data = no_filter$did_fit[[1]]$did_results,
+    aes(year, did, color = "Empirical Estimate"),
+    .width = c(0.5, 0.95),
+    slab_alpha = 0.95  ) +
+  scale_y_continuous(name = "MPA Effect") +
+  scale_x_discrete(name = element_blank()) +
+  theme(legend.position = "top") +
+  scale_fill_brewer(name = "Simulation Quantiles") + 
+  scale_color_manual(values = "red", name = '') + 
+    geom_hline(yintercept = 0)
+
+```
+
+```{r no-weight-plot, fig.cap="Version of the model fit without reweighting samples in proportion to area",results = "hide"}
+no_weight <- model_runs %>%
+  slice(1) %>%
+  left_join(abundance_data, by = "data_source") %>%
+  mutate(
+    did_fit = pmap(
+      list(
+        data = data,
+        data_to_use = data_to_use,
+        data_source = data_source
+      ),
+      estimate_did,
+      iter = 2500,
+      site_data = site_data,
+      cdfw_catches = cdfw_catches,
+      life_history_data = life_history_data,
+      weight_samples = FALSE
+    )
+  )
+
+ggplot() +
+ tidybayes::stat_halfeye(
+    data = no_weight$did_fit[[1]]$did_results,
+    aes(year, did, color = "Empirical Estimate"),
+    .width = c(0.5, 0.95),
+    slab_alpha = 0.95  ) +
+  scale_y_continuous(name = "MPA Effect") +
+  scale_x_discrete(name = element_blank()) +
+  theme(legend.position = "top") +
+  scale_fill_brewer(name = "Simulation Quantiles") + 
+  scale_color_manual(values = "red", name = '') + 
+    geom_hline(yintercept = 0)
+
+```
+
+For ease of interpretation and model convergence, our default model estimates MPA effects in three-year bins.
+For this run we estimate the effect annually, relative to the year 2000.
+Results show the same overall trend and magnitude reported in our base model (Fig.\@ref(fig:no-bin-plot)).
+
+```{r no-bin-plot, fig.cap="For ease of interpretation and model convergence, our default model estimates MPA effects in three-year bins.
+For this run we estimate the effect annually, relative to the year 2000.
+Results show the same overall trend and magnitude reported in our base model.", results = "hide", eval = TRUE}
+no_bin <- model_runs %>%
+  slice(1) %>%
+  left_join(abundance_data, by = "data_source") %>%
+  mutate(
+    did_fit = pmap(
+      list(
+        data = data,
+        data_to_use = data_to_use,
+        data_source = data_source
+      ),
+      estimate_did,
+      iter = 2500,
+      first_year = 2001,
+      bin_width = 1,
+      site_data = site_data,
+      cdfw_catches = cdfw_catches,
+      life_history_data = life_history_data
+    )
+  )
+
+ggplot() +
+ tidybayes::stat_halfeye(
+    data = no_bin$did_fit[[1]]$did_results,
+    aes(year, did, color = "Empirical Estimate"),
+    .width = c(0.5, 0.95),
+    slab_alpha = 0.95  ) +
+  scale_y_continuous(name = "MPA Effect") +
+  scale_x_discrete(name = element_blank()) +
+  theme(legend.position = "top") +
+  scale_fill_brewer(name = "Simulation Quantiles") + 
+  scale_color_manual(values = "red", name = '') + 
+    scale_x_discrete(guide = guide_axis(n.dodge = 2))
+
+```
+
+Our base run estimates the effect of MPAs on total biomass densities of targeted finfish.
+For this run, we instead estimate the effect on mean biomass densities of targeted species.
+Results show the same pattern as our base run (Fig.\@ref(fig:mean-fit)).
+
+```{r mean-fit, fig.cap="Our base run estimates the effect of MPAs on total biomass densities of targeted finfish.
+For this run, we instead estimate the effect on mean biomass densities of targeted species.
+Results show the same pattern as our base run.", results="hide"}
+mean_fit <- model_runs %>%
+  slice(1) %>%
+  left_join(abundance_data, by = "data_source") %>%
+  mutate(
+    did_fit = pmap(
+      list(
+        data = data,
+        data_to_use = data_to_use,
+        data_source = data_source
+      ),
+      estimate_did,
+      fit_mean = TRUE,
+      iter = 2500,
+      site_data = site_data,
+      cdfw_catches = cdfw_catches,
+      life_history_data = life_history_data,
+      refresh = 1000
+    )
+  )
+
+ggplot() +
+ tidybayes::stat_halfeye(
+    data = mean_fit$did_fit[[1]]$did_results,
+    aes(year, did, color = "Empirical Estimate"),
+    .width = c(0.5, 0.95),
+    slab_alpha = 0.95  ) +
+  scale_y_continuous(name = "MPA Effect") +
+  scale_x_discrete(name = element_blank()) +
+  theme(legend.position = "top") +
+  scale_fill_brewer(name = "Simulation Quantiles") + 
+  scale_color_manual(values = "red", name = '') + 
+    scale_x_discrete(guide = guide_axis(n.dodge = 2))
+
+```
+
+Appropriately addressing the problem of "missing" observations is a critical challenge in any field observation study.
+If no observations of a given fish species were recorded on a given transect, should the density of that species on that transect be marked as zero, and influence the estimate of the overall mean density accordingly?
+The obvious answer seems to be yes, but what if that species simply does not live in the environment covered by a particular transect, or was not present during the particular time of the diver's observation?
+For our base runs, we assign a value of zero density on a given transect for any fish species that has been observed at least once at a given site at any time in our data but was not observed on that particular transect.
+If that species was never observed at that site, we do not include a zero for that species.
+Our rationale for this is that given the shifting nature of the sampled sites, and the intensity of sampling at those sites, we do not want to skew density trends by changes in the amount of suitable habitat for a given species sampled.
+However, this is clearly a strong assumption.
+For example, perhaps the decreasing trend in mean densities from 2000 to 2004 is due to increased number of sites (and therefore zeros) included in the data.
+To assess the potential importance of this choice, we can compare the mean densities of targeted and non-targeted species over time with the added zeros to the mean densities using only positive observations (i.e.
+not including any zeros in the data, (Fig.\@ref(fig:no-zero-fit)).
+The trends in the raw densities, and most importantly the mean trends of targeted and non-targeted fishes, are nearly identical whether or not zeros are added, providing strong evidence that our choice of how to incorporate missing observations into the data are not strongly influencing our overall results.
+
+```{r no-zero-fit, fig.cap="Our base run estimates the effect of MPAs on total biomass densities of targeted finfish.
+For this run, we instead estimate the effect on mean biomass densities of targeted species.
+Results show the same pattern as our base run.", results = "hide"}
+
+nozero_fit <- model_runs %>%
+  slice(1) %>%
+  left_join(abundance_data, by = "data_source")
+
+nozero_fit$data[[1]] <- nozero_fit$data[[1]] %>% filter(any_seen == TRUE)
+
+nozero_fit <-  nozero_fit%>%
+  mutate(
+    did_fit = pmap(
+      list(
+        data = data,
+        data_to_use = data_to_use,
+        data_source = data_source
+      ),
+      estimate_did,
+      iter = 2500,
+      site_data = site_data,
+      cdfw_catches = cdfw_catches,
+      life_history_data = life_history_data
+    )
+  )
+
+ggplot() +
+ tidybayes::stat_halfeye(
+    data = nozero_fit$did_fit[[1]]$did_results,
+    aes(year, did, color = "Empirical Estimate"),
+    .width = c(0.5, 0.95),
+    slab_alpha = 0.95  ) +
+  scale_y_continuous(name = "MPA Effect") +
+  scale_x_discrete(name = element_blank()) +
+  theme(legend.position = "top") +
+  scale_fill_brewer(name = "Simulation Quantiles") + 
+  scale_color_manual(values = "red", name = '') + 
+    scale_x_discrete(guide = guide_axis(n.dodge = 2))
+
+```
+
+### Synthetic controls
+
+Synthetic controls are an alternative method for attempting to estimate the causal effect of a policy intervention [@abadie2010].
+A difference-in-difference approach assumes that some observable group serves as an adequate control for the state of the treated group in an untreated world.
+In our default case, we assume that the mean standardized index of non-targeted species are our control for the targeted species.
+Alternatively, synthetic controls use timeseries of treated and non-treated groups before and after treatment to construct a new "control" group built by weighting the pre-treatment timeseries of un-treated observations (together with covariates) such that the synthetic control group matches the trends in the treated group pre-treatment.
+
+We chose to present difference-in-difference as our main result since it better allows us to capture the uncertainty in the data generating process through our hierarchical model.
+However, we felt that it was worth exploring whether synthetic controls provided substantially different results than our default model.
+
+For the first synthetic control, we pulled our standardized mean index of abundance for targeted species as a whole from our difference-in-difference model as our treated group.
+We then pull the standardized indices of abundance for each of the non-targeted groups from the difference-in-difference to use as the candidate untreated components for the synthetic controls.
+A complete synthetic control analysis would require more extensive validation of the methods, but we use this approach simply to explore whether we observe substantially divergent results in the synthetic control versus the difference-in-difference model.
+
+We centered and scaled the candidate abundance indices to facilitate model convergence given the very few number of pre-treatment years available.
+The results of a synthetic control model are presented as the difference between the observed treatment outcome and the synthetic control (the difference in this case being in units of standard deviations).
+The model was not able to construct an adequate synthetic control at this level, as shown by the differences between the treated group and the synthetic control pre-treatment.
+However, we would note that the post-treatment results do show similarities with our main results, namely a lack of a clear divergence between the treatment and the control, and an upwards trend up through the earl 2010s followed by a decline (Fig.\@ref(fig:total-synth)).
+
+```{r, include = FALSE, eval = FALSE}
+
+years <- unique(used_data$year)
+
+life_history <- used_data %>%
+  select(classcode) %>%
+  unique() %>%
+  mutate(numeric_classcode = as.numeric(as.factor(classcode)))
+
+nontargeted_index <- data_frame(abundance_hat = fit_report$abundance_hat,
+                               classcode = rep(1:n_distinct(used_data$classcode), each = length(years))) %>%
+  mutate(log_abundance_hat = log(abundance_hat)) %>%
+  group_by(classcode) %>%
+  mutate(year =years) %>%
+  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
+  ungroup() %>%
+  rename(numeric_classcode = classcode) %>%
+  left_join(life_history, by = "numeric_classcode" ) %>% 
+  left_join(unique(abundance_data$data[[1]] %>% select(classcode, targeted)), by = "classcode") %>% 
+  filter(targeted == 0) %>% 
+  select(log_abundance_hat, year, classcode, targeted)
+
+targeted_index <- tmb_runs$tmb_fit[[1]]$zissou_estimates %>%
+  filter(
+    stringr::str_detect(variable, "(targeted_did_betas)"),!str_detect(variable, "non")
+  ) %>%
+  mutate(
+    year = years,
+    classcode = "targeted",
+    targeted = 1
+  ) %>%
+  rename(log_abundance_hat = estimate) %>% 
+  select(colnames(nontargeted_index))
+
+  # filter(variable == "seeing_year_species_betas") %>%
+
+nontargeted_covariates <- used_data %>% 
+  select(classcode, contains("mean"),year, targeted) %>% 
+  filter(targeted == 0) %>% 
+  select(-mean_length) %>% 
+  gather(variable, value, -classcode,-year,-targeted) %>% 
+  group_by(classcode, variable, year) %>% 
+  summarise(mean_value = mean(value, na.rm = TRUE)) %>% 
+  spread(variable, mean_value)
+
+targeted_covariates <- used_data %>% 
+  select(classcode, contains("mean"),year, targeted) %>% 
+  filter(targeted == 1) %>% 
+  mutate(classcode = "targeted") %>% 
+  select(-mean_length) %>% 
+  gather(variable, value, -classcode,-year,-targeted) %>% 
+  group_by(classcode, variable, year) %>% 
+  summarise(mean_value = mean(value, na.rm = TRUE)) %>% 
+  spread(variable, mean_value)
+
+synth_covariates <- nontargeted_covariates %>% 
+  bind_rows(targeted_covariates)
+
+synth_data <- targeted_index %>% 
+  bind_rows(nontargeted_index) %>% 
+  left_join(synth_covariates, by = c("classcode","year")) %>% 
+    mutate(numeric_classcode = as.numeric(factor(classcode))) %>% 
+  group_by(classcode) %>% 
+  mutate(log_abundance_hat = scale(exp(log_abundance_hat))) %>% 
+  ungroup() %>% 
+  as.data.frame() %>% 
+  select(-rolling_mean)
+
+# synth_data %>%
+#   group_by(classcode) %>%
+#   mutate(index = (log_abundance_hat)) %>%
+#   ggplot(aes(
+#     year,
+#     index,
+#     color = factor(targeted),
+#     group = interaction(targeted, classcode)
+#   )) +
+#   geom_line()
+
+prepped_synth_data <-
+  Synth::dataprep(
+    foo = synth_data,
+    predictors = colnames(synth_data)[str_detect(colnames(synth_data),"mean")],
+    predictors.op = "mean",
+    dependent = "log_abundance_hat",
+    unit.variable = "numeric_classcode",
+    unit.names.variable = "classcode",
+    time.variable = "year",
+    treatment.identifier = "targeted",
+    controls.identifier = unique(synth_data$classcode)[unique(synth_data$classcode) != "targeted"],
+    time.predictors.prior = years[years <= 2003],
+    time.optimize.ssr = years[years <= 2003],
+    time.plot = years
+  )
+
+synth_fit <- Synth::synth(prepped_synth_data)
+
+gaps <-
+  rownames_to_column(
+    prepped_synth_data$Y1plot - (prepped_synth_data$Y0plot %*% synth_fit$solution.w) %>%
+      as.data.frame()
+  ) %>%
+  rename(gap = w.weight,
+         year = rowname) %>%
+  mutate(year = as.numeric(year)) %>%
+  mutate(synthetic = as.numeric(prepped_synth_data$Y0plot %*% synth_fit$solution.w)) %>% 
+  mutate(treatment = gap + synthetic)
+
+# path.plot(synth.res = synth_fit,
+#           dataprep.res = prepped_synth_data
+# )
+
+total_gap_plot <- gaps %>% 
+  ggplot(aes(year, gap)) + 
+  geom_hline(aes(yintercept = 0), linetype = 2) +
+  geom_vline(aes(xintercept = 2003), color = "red")+
+  geom_line() + 
+    labs(x = "Year", y = "Standard deviations from synthetic control")
+
+total_synth_plot <- gaps %>% 
+  select(year, synthetic, treatment) %>% 
+  gather(source, value, -year) %>% 
+  group_by(source) %>% 
+  # mutate(value = scale(value)) %>% 
+  ggplot(aes(year, value, color = source, linetype = source)) + 
+  geom_vline(aes(xintercept = 2003), color = "red")+
+  geom_line() + 
+    labs(x = "Year", y = "Standard deviations from synthetic control")
+
+```
+
+```{r  total-synth, fig.cap = "Difference in centered and scaled standardized targeted abundance and synthetic standardized targeted abundance", eval=FALSE}
+total_gap_plot
+```
+
+```{r raw-synth, include = FALSE, eval = FALSE}
+
+  consistent_classcodes <- used_data %>%
+      group_by(classcode, region) %>%
+      filter(any_seen) %>%
+      summarise(n_years = n_distinct(year)) %>%
+      ungroup() %>%
+      filter(n_years == n_distinct(used_data$year))
+
+synth_data <- used_data %>% 
+  filter(classcode %in% consistent_classcodes$classcode,
+         region %in% consistent_classcodes$region) %>% 
+  mutate(classcode = ifelse(targeted == 1, "targeted", classcode),
+         region = ifelse(targeted == 1, "targeted", region)) %>% 
+      group_by(year,
+               site_side,
+               region,
+               zone,
+               transect,
+               eventual_mpa,
+               classcode,
+               targeted) %>%
+      summarise(
+        total_classcode_density = sum(exp(log_density)),
+        var_tex = sum(cumulative_n_obs),
+        var_vis = mean(mean_vis),
+        var_temp = mean(mean_temp),
+        var_depth = mean(mean_depth),
+        var_surge = mean(surge),
+        var_kelp = mean(interp_kelp)
+      ) %>%  # sum density across all levels of a transect
+      group_by(year, site_side, region, eventual_mpa, classcode, targeted) %>%
+      summarise(
+        md = mean(total_classcode_density),
+        var_tex = mean(var_tex),
+        var_vis = mean(var_vis),
+        var_temp = mean(var_temp),
+        var_depth = mean(var_depth),
+        var_surge = mean(var_surge),
+        var_kelp = mean(var_kelp)
+      ) %>% # calculate mean density per year site, side, species, averaging over zone, transect
+      ungroup()
+
+ synth_data <- synth_data %>% 
+     group_by(year,region,classcode,targeted) %>%
+      summarise(
+        total_biomass_density = (sum(md) / 1e6) * 10000,
+        # calculate total and mean biomass densities across all species per year site side
+        mean_biomass_density = log((mean(md) / 1e6) * 10000 + 1e-3),
+        var_tex = mean(var_tex),
+        var_vis = mean(var_vis),
+        var_temp = mean(var_temp),
+        var_kelp = mean(var_kelp)
+      ) %>%
+   ungroup() %>% 
+   mutate(classcode = ifelse(targeted == 1, classcode, paste(classcode, region, sep = '_'))) %>% 
+      mutate(numeric_classcode = as.numeric(factor(classcode))) %>% 
+   as.data.frame()
+
+prepped_synth_data <-
+  Synth::dataprep(
+    foo = synth_data,
+    predictors = colnames(synth_data)[str_detect(colnames(synth_data),"var")],
+    predictors.op = "mean",
+    dependent = "mean_biomass_density",
+    unit.variable = "numeric_classcode",
+    unit.names.variable = "classcode",
+    time.variable = "year",
+    treatment.identifier = "targeted",
+    controls.identifier = unique(synth_data$classcode[synth_data$targeted == 0]),
+    time.predictors.prior = years[years <= 2003],
+    time.optimize.ssr = years[years <= 2003],
+    time.plot = years
+  )
+
+synth_fit <- Synth::synth(prepped_synth_data)
+
+gaps <-
+  rownames_to_column(
+    prepped_synth_data$Y1plot - (prepped_synth_data$Y0plot %*% synth_fit$solution.w) %>%
+      as.data.frame()
+  ) %>%
+  rename(gap = w.weight,
+         year = rowname) %>%
+  mutate(year = as.numeric(year)) %>%
+  mutate(synthetic = as.numeric(prepped_synth_data$Y0plot %*% synth_fit$solution.w)) %>% 
+  mutate(treatment = gap + synthetic)
+
+# path.plot(synth.res = synth_fit,
+#           dataprep.res = prepped_synth_data
+# )
+
+total_gap_plot <- gaps %>% 
+  ggplot(aes(year, gap)) + 
+  # geom_hline(aes(yintercept = 0), linetype = 2) +
+  geom_vline(aes(xintercept = 2003), color = "red")+
+  geom_line() + 
+    labs(x = "Year", y = "Standard deviations from synthetic control")
+
+total_synth_plot <- gaps %>% 
+  select(year, synthetic, treatment) %>% 
+  gather(source, value, -year) %>% 
+  group_by(source) %>% 
+  # mutate(value = scale(value)) %>% 
+  ggplot(aes(year, value, color = source, linetype = source)) + 
+  geom_vline(aes(xintercept = 2003), color = "red")+
+  geom_line() + 
+    labs(x = "Year", y = "Standard deviations from synthetic control")
+
+```
+
+As an extension, we repeated this process, but now treating each targeted species individually as the treated group, and the non-targeted species as the non-targeted.
+This is intended to explore whether we see clearer signals for individual species than we do for the targeted class as a whole.
+
+Overall we see similarly unclear results as the aggregate targeted synthetic control (and our main results).
+The synthetic control was better constructed for some individual species, but not clearly for any one, and most species showed some evidence of the upward-then-downward trend seen throughout our results (Fig.\@ref(fig:synth-classcode-fits)).
+
+```{r, include = FALSE, eval=FALSE}
+
+synth_foo <-
+  function(treated_classcode,
+           synth_data,
+           targeted_classcodes) {
+    # treated_classcode <- targeted_classcodes[1]
+
+    temp_synth_data <- synth_data %>%
+      filter(classcode == treated_classcode |
+               !(classcode %in% targeted_classcodes))
+
+    prepped_synth_data <-
+      Synth::dataprep(
+        foo = temp_synth_data,
+        predictors = colnames(temp_synth_data)[str_detect(colnames(temp_synth_data), "mean")],
+        predictors.op = "mean",
+        dependent = "log_abundance_hat",
+        unit.variable = "numeric_classcode",
+        unit.names.variable = "classcode",
+        time.variable = "year",
+        treatment.identifier = treated_classcode,
+        controls.identifier = unique(temp_synth_data$classcode)[unique(temp_synth_data$classcode) != treated_classcode],
+        time.predictors.prior = c(2000:2003),
+        time.optimize.ssr = c(2000:2003),
+        time.plot = years
+      )
+
+    synth_fit <- Synth::synth(prepped_synth_data)
+
+    gaps <-
+      rownames_to_column(
+        prepped_synth_data$Y1plot - (prepped_synth_data$Y0plot %*% synth_fit$solution.w) %>%
+          as.data.frame()
+      ) %>%
+      rename(gap = w.weight,
+             year = rowname) %>%
+      mutate(year = as.numeric(year)) %>%
+      mutate(synthetic = as.numeric(prepped_synth_data$Y0plot %*% synth_fit$solution.w)) %>%
+      mutate(treatment = gap + synthetic)
+  }
+
+nontargeted_index <- data_frame(abundance_hat = fit_report$abundance_hat,
+                               classcode = rep(1:n_distinct(used_data$classcode), each = length(years))) %>%
+  mutate(log_abundance_hat = log(abundance_hat)) %>%
+  group_by(classcode) %>%
+  mutate(year = 1999 + 1:length(abundance_hat)) %>%
+  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
+  ungroup() %>%
+  rename(numeric_classcode = classcode) %>%
+  left_join(life_history, by = "numeric_classcode" ) %>% 
+  left_join(unique(abundance_data$data[[1]] %>% select(classcode, targeted)), by = "classcode") %>% 
+  select(log_abundance_hat, year, classcode, targeted)
+
+synth_covariates <- pisco_data %>% 
+  select(classcode, contains("mean"),year, targeted) %>% 
+  select(-mean_length) %>% 
+  gather(variable, value, -classcode,-year,-targeted) %>% 
+  group_by(classcode, variable, year) %>% 
+  summarise(mean_value = mean(value, na.rm = TRUE)) %>% 
+  spread(variable, mean_value)
+
+synth_data <- nontargeted_index %>% 
+  left_join(synth_covariates, by = c("classcode","year")) %>% 
+    mutate(numeric_classcode = as.numeric(factor(classcode))) %>% 
+  group_by(classcode) %>% 
+  mutate(log_abundance_hat = scale(exp(log_abundance_hat))) %>% 
+  ungroup() %>% 
+  as.data.frame()
+
+targeted_classcodes <- unique(synth_data$classcode[synth_data$targeted == 1])
+
+synth_classcode_fits <-
+  tibble(targeted_classcodes = targeted_classcodes) %>%
+  mutate(
+    synthetic_fit = map(
+      targeted_classcodes,
+      synth_foo,
+      synth_data = synth_data,
+      targeted_classcodes = targeted_classcodes
+    )
+  )
+
+synth_classcode_fits <- synth_classcode_fits %>% 
+  unnest()
+
+synth_classcode_fits_plot <- 
+synth_classcode_fits %>% 
+  ggplot(aes(year, gap)) + 
+  geom_hline(aes(yintercept = 0), linetype = 2) +
+  geom_vline(aes(xintercept = 2003), color = "red")+
+  geom_line() + 
+  facet_wrap(~targeted_classcodes) + 
+  labs(x = "Year", y = "Standard deviations from synthetic control")
+
+```
+
+```{r synth-classcode-fits, fig.cap="Synthetic control gaps for each targeted species", eval=FALSE}
+synth_classcode_fits_plot + 
+  theme_minimal()
+```
+
+## Testing Model Assumptions
+
+### Simulation testing
+
+We state that a difference-in-difference model using targeted and non-targeted species is capable (conditional on assumptions) of estimating the causal effect of MPAs.
+We simulated MPA outcomes to test this claim.
+We first test our estimation strategy under idealized circumstances, where recruitment is deterministic and PISCO divers all have constant and perfect observer skills.
+We simulate five species that vary only in their maximum size and length at maturity.
+For each of these species, we set one version that is targeted by fishing and one that is not.
+We set a constant fishing mortality rate for each simulated targeted species, and then ran two matched simulations, one with MPAs and one without.
+We then have our simulated divers sample data from each of these scenarios, and then pass the sampled biomass densities to a simplified version of our difference-in-difference model (omitting the probability of detection step).
+We can then compare the difference-in-difference estimates of the MPA effect to the true simulated effect.
+The difference in difference model is able to capture the simulated MPA effect under these circumstances (Fig.\@ref(fig:simple-did-test))
+
+```{r simple-did-test, fig.cap = "Simulated mean (red dashed line) and individual species (solid lines) MPA effects over time, along with difference-in-difference estimated MPA effects (mean with 95% confidence intervals)"}
+
+load(file = file.path(run_dir,'simulated_did.Rdata'))
+
+simple_performance$mixed_effect_did + 
+  labs(title = '', x = "Years with MPA protection",
+       y = "MPA Effect")
+```
+
+We then simulated a more complex example.
+We use the actual targeted and non-targeted species from our model.
+We assign species predominately seen in the western Channel Islands as "cold water" and those in the eastern Channel Islands as "warm water".
+We allow for stochasticity in recruitment.
+We use El Niño data as a simulated environmental recruitment driver, where we assume that El Niño events produce negative recruitment shocks for cold water species and *vice versa* for warm water species.
+We simulate three different divers each with different base skill levels, visual selectivities, and an evolving skill rate (such that observers get better over time).
+We hold fishing mortality rates constant for each species, although that fishing mortality affects each species differently because of intrinsic biological differences in maturity-at-age and steepness.
+We then test the ability of the difference-in-difference model to isolate the mean MPA effect across all of these targeted species, which our results show it is capable of doing (Fig.\@ref(fig:complex-did-test)).
+
+```{r complex-did-test,  fig.cap = "Simulated mean (red dashed line) and individual species (solid lines) MPA effects over time, along with difference-in-difference estimated MPA effects (mean with 95% confidence intervals)"}
+
+pisco_performance$mixed_effect_did + 
+    labs(title = '', x = "Years with MPA protection",
+       y = "MPA Effect")
+
+```
+
+### Testing SUTVA with Convergent Cross Mapping
+
+The difference-in-difference model also assumes that the targeted and non-targeted fishes do not directly or indirectly affect each other.This assumption is clearly violated on some level: all the fishes in this analysis are part of the same ecosystem and therefore interact to some degree.
+For example, if the protection of targeted predatory fishes results in increased mortality of non-targeted fishes, the model would attribute that as an increased population effect (greater divergence between the abundance of targeted and non-targeted species).
+Given the time scale of analysis (15 years of protection), we do not feel that massive trophic cascades are likely to have developed yet, given both the pace and complexity of trophic cascade development [@babcock2010; @pershing2015a].
+A complete assessment of evidence for trophic cascades in the Channel Islands is beyond the scope of this study, but to address this question somewhat we utilized convergent cross mapping *sensu* @sugihara2012 to test for a significant causal signal between different broad trophic groups in the data, implemented in the `rEDM` package in R.
+
+Convergent cross mapping is a nonlinear forecasting method that uses observed time series data to test for significant causal links between variables.
+Following methods laid out in @clark2015 and @sugihara2012, we pool the abundance of each broad trophic group by region (Fig.\@ref(fig:trophic-plot).
+This uses the data from the islands as "replicates", requiring the assumption that the islands are all part of the same dynamic system, but allowing us to take advantage of the extra information provided by each island to further resolve the reconstructed manifolds.
+Using these aggregations, we then test whether the variables can be properly embedded, i.e., if they have predictable manifold dynamics.
+We do this through a simplex forecasting test, using an individual timeseries' own lags to build a manifold.
+For each timeseries, the "best embedding dimension" is an approximation of the dimensionality of the dynamic system, in other words, the number of dimensions that define and predict the evolving states of the timeseries.
+This analysis shows that only the carnivore, piscivore, and planktivores groups show evidence of significant predictability (that is, that past dynamics of these species groups can predict future dynamics, Fig.\@ref(fig:embed-plot)).
+
+```{r trophic-plot, fig.cap="Centered and scaled densities by broad trophic group and island over time"}
+cip_data <- used_data
+trophic_groups <- used_data %>% 
+  select(classcode, broadtrophic) %>% 
+  unique()
+
+# dat <- abundance_trends %>%
+#   left_join(trophic_groups, by = "classcode") %>%
+#   group_by(broadtrophic, year) %>%
+#   summarise(density = mean(scaled_abundance_hat)) %>%
+#   group_by(broadtrophic) %>%
+#   mutate(density = scale(density)) %>%
+#   ungroup() %>% 
+#   spread(broadtrophic, density) %>% 
+#   arrange(year) %>% 
+#   mutate(lag4_carnivore = lag(carnivore,4),
+#          lag4_piscovore = lag(piscivore,4)) %>% 
+#   na.omit() %>% 
+#   ungroup() %>% 
+#   filter(year > 1999) 
+# dat %>%
+#   ggplot(aes(year, density, color = broadtrophic)) +
+#   geom_line()
+
+dat <- cip_data %>%
+  group_by(broadtrophic, region, site, side, transect, year, eventual_mpa) %>%
+  summarise(density = sum(density_g_m2)) %>%
+  group_by(broadtrophic, region, site, year) %>%
+  summarise(mean_density = mean(density, na.rm = T)) %>%
+  group_by(broadtrophic,region, year) %>%
+  summarise(regional_density = mean(mean_density, na.rm = T)) %>% 
+  ungroup() %>% 
+  spread(broadtrophic, regional_density) %>% 
+  group_by(region) %>% 
+  arrange(year) %>% 
+  mutate(lag4_carnivore = lag(carnivore,4),
+         lag4_piscovore = lag(piscivore,4)) %>% 
+  na.omit() %>% 
+  ungroup() %>% 
+  filter(year > 1999)
+
+ dat <- dat %>%
+  select(-lag4_carnivore,-lag4_piscovore)%>%
+  group_by(region)%>%
+  mutate_at(vars(carnivore:planktivore),funs(norm=(.-mean(.))/sd(.)))->dat
+
+group_trends_plot <- dat %>%
+  select(region,year,contains("norm"))%>%
+  gather("group","density",-region,-year)%>%
+  ggplot(aes(year,density,col=group))+
+  geom_line()+
+  geom_point() +
+  scale_color_locuszoom(name="Trophic Group",labels=c("carnivore","herbivore","piscivore","planktivore"))+
+  labs(x="year",y="Centered and Scaled Density")+
+  facet_wrap(~region)
+
+group_trends_plot
+```
+
+```{r embed-plot, fig.cap = "Predictive skill as a function of embedding dimensions"}
+
+## split data by region to analyze timeseries from each island separately
+ana.dat <- dat %>% filter(region=="ANA")
+sci.dat <- dat %>% filter(region=="SCI")
+smi.dat <- dat %>% filter(region=="SMI")
+sri.dat <- dat %>% filter(region=="SRI")
+
+datnest <- dat %>% group_by(region) %>% nest()
+
+datnorm <- dat %>% select(region,year,contains('norm')) %>% ungroup()
+
+# have to record the segments corresponding to each "replicate" so simplex algorithm does not try to make predictions crossing time barriers
+segs <- datnorm %>% mutate(ind=row_number()) %>% group_by(region) %>% summarise(first=first(ind),last=last(ind)) %>%
+  select(-region)
+
+var_names <- c("carnivore_norm","herbivore_norm","piscivore_norm","planktivore_norm")
+
+regions.combined.simp.list <- map(var_names,function(x){
+  temp <- datnorm %>% ungroup() %>% select(matches(x)) %>% as.data.frame()
+  out <- simplex(as.numeric(temp[,1]),E=1:10,lib=as.matrix(segs),silent=T) %>%
+    mutate(trophic=str_replace(x,"_norm",""))
+  out
+})
+
+embed_plot <- bind_rows(regions.combined.simp.list) %>%
+  ggplot(aes(E,rho,color=trophic))+
+  geom_line(size=2)+
+  facet_wrap(~trophic,nrow=2,scales="free_y")+
+  geom_hline(yintercept = 0,color="black")+
+  labs(x="Embedding Dimension (E)",y=expression(paste("Predictive Skill, ",rho)))+
+  scale_x_continuous(breaks=seq(0,12,by=2))+
+  scale_color_locuszoom()+
+  guides(color=F)
+
+embed_plot
+
+```
+
+Focusing on just these three groups then (removing herbivores), we can test for causal relationships between groups using convergent cross mapping and the logic of Takens' theorem of dynamic systems.
+Generalizations of Takens' theorem indicate that if two variables (in our case, species or physical variables) are part of the same dynamic system, their individual dynamics should reflect their relative causal influence [@sugihara2012].
+In other words, if one variable is causally forced by another, that forcing should leave a signature on the first time series.
+Convergent cross mapping (CCM) tests for causation by using the attractor/manifold built from the time series of one variable to predict another (hence the "cross-mapping").
+In simple terms, the *causal effect of A on B is determined by how well B cross-maps A*.
+
+There are two criteria for CCM to establish causality:  First, and most obviously, predictive cross-map skill using all available data should be significantly greater than zero.
+Second, that predictability should be convergent.
+Convergence means that cross-mapped estimates improve with library length (the number of state-space vectors used to build the attractor), because the attractor is more fully resolved and therefore estimation error should decline.
+Convergence is key to distinguishing causation from simple or spurious correlation.
+If two variables are spuriously correlated and not causally linked, CCM should fail to satisfy this second criterion.
+Based on these criteria, there is little evidence of significant dynamic interactions between trophic groups (Fig.\@ref(fig:cross-map-cp)-\@ref(fig:cross-map-pis)).
+Cross-mapping produced positive predictive skill, but was non-convergent for all cross-mappings with the exception of carnivores cross-mapping planktivores (providing some evidence that planktivore dynamics may be a driver of carnivore dynamics).
+This analysis provides evidence that trophic cascades are unlikely to be a significant driver of our results.
+It is important to note though that this analysis does not mean that trophic cascades could not emerge in this system, rather that we do not detect them with these data at this time.
+
+```{r cross-map-fxn}
+# plots a cross mapping between two variables in the data, given their variable names (grp), their optimal embedding dimensions (from simplex forecasting), and their preferred names for plotting
+xmap <- function(grp1,grp2,E1,E2,name1,name2){
+    # cross-map group 1 to group 2
+
+  ccm1 <- ccm(datnorm,lib=as.matrix(segs),pred=as.matrix(segs),E=E1,lib_column= grp1,target_column = grp2,lib_sizes = c(10,25,50,75),num_samples=100,replace=T,silent=T,RNGseed = 41389)
+
+  # test for convergence (is predictability at max library size significantly greater than at minimum library size?)
+
+  rhomeans1 <- ccm1 %>% ccm_means()
+
+  conv_test1 <- t.test(ccm1$rho[ccm1$lib_size==10],ccm1$rho[ccm1$lib_size==75])$p.value
+  if((rhomeans1$rho[rhomeans1$lib_size==75]-rhomeans1$rho[rhomeans1$lib_size==10])<0.05) conv_test1 <- 1
+  conv_test1 <- ifelse(conv_test1<0.05,"Convergent","Non-convergent")
+
+  xmap1 <- ccm1 %>%
+    group_by(lib_size)%>%
+    summarise(rhomean=mean(rho,na.rm=T),upper=quantile(rho, 0.975),lower=quantile(rho, 0.025))%>%
+    ungroup()%>%
+    ggplot(aes(lib_size,rhomean))+
+    geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.3,fill="red")+
+    geom_line(color="darkorchid3")+
+    geom_hline(aes(yintercept = 0), linetype = 2) +
+    labs(x="Library Size",y=expression(paste(rho, " (predictive skill)")),title=paste0(name1," xmap ",name2,"\n","(",conv_test1,")"))+
+    theme(plot.title = element_text(size=6))
+
+  # cross map group 2 to group 1
+  # inspect the output of simplex from the previous step and use the best embedding dimension (highest rho) for the carnivore time series
+
+  ccm2 <- ccm(datnorm,lib=as.matrix(segs),pred=as.matrix(segs),E=E2,lib_column= grp2,target_column = grp1,lib_sizes = c(10,25,50,75),num_samples=100,replace=T,silent=T,RNGseed = 41389)
+
+  rhomeans2 <- ccm2 %>% ccm_means()
+
+  conv_test2 <- t.test(ccm2$rho[ccm2$lib_size==10],ccm2$rho[ccm2$lib_size==75])$p.value
+  if((rhomeans2$rho[rhomeans2$lib_size==75]-rhomeans2$rho[rhomeans2$lib_size==10])<0.05) conv_test2 <- 1
+  conv_test2 <- ifelse(conv_test2<0.05,"Convergent","Non-convergent")
+
+  xmap2 <- ccm2 %>%
+    group_by(lib_size)%>%
+    summarise(rhomean=mean(rho,na.rm=T),upper=quantile(rho, 0.975),lower=quantile(rho, 0.025))%>%
+    ungroup()%>%
+    ggplot(aes(lib_size,rhomean))+
+    geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.3,fill="red")+
+    geom_line(color="darkorchid3")+
+    geom_hline(aes(yintercept = 0), linetype = 2) +
+    labs(x="Library Size",y=expression(paste(rho, " (predictive skill)")),title=paste0(name2," xmap ",name1,"\n","(",conv_test2,")"))+
+    theme(plot.title = element_text(size=6))
+
+  cross_map_plot <- xmap1 + xmap2
+
+cross_map_plot
+}
+```
+
+```{r cross-map-cp, fig.cap = "Cross mapping of effect of piscivores on carnivores (A) and carnivores on piscivores (B) in the PISCO data from 2000 to 2017.
+Shaded region show 95% confidence interval"}
+
+car_pis<-xmap(grp1="carnivore_norm",grp2="piscivore_norm",E1=4,E2=4,name1="Carnivores",name2="Piscivores")
+car_pl <- xmap(grp1="carnivore_norm",grp2="planktivore_norm",E1=4,E2=4,name1="Carnivores",name2="Planktivores")
+pl_pis <- xmap(grp1="planktivore_norm",grp2="piscivore_norm",E1=4,E2=4,name1="Planktivores",name2="Piscivores")
+
+car_pis
+```
+
+```{r cross-map-pl, fig.cap = "Cross mapping of effect of planktivores on carnivores (A) and carnivores on planktivores (B) in the PISCO data from 2000 to 2017.
+Shaded region show 95% confidence interval"}
+
+car_pl
+
+```
+
+```{r cross-map-pis, fig.cap = "Cross mapping of effect of piscivores on carnivores (A) and carnivores on planktivores (B) in the PISCO data from 2000 to 2017.
+Shaded region show 95% confidence interval"}
+
+ pl_pis 
+```
+
+```{r trophic-cor-plot, fig.cap="Estimated correlation coefficients between broad trophic groups at the island by year resolution", eval = FALSE}
+
+trophic_effects_plot <- cip_data %>%
+  group_by(broadtrophic, region, site, side, transect, year, eventual_mpa) %>%
+  summarise(density = sum(density_g_m2)) %>%
+  group_by(broadtrophic, region, site, year) %>%
+  summarise(mean_density = mean(density, na.rm = T)) %>%
+  group_by(broadtrophic,region, year) %>%
+  summarise(regional_density = mean(mean_density, na.rm = T)) %>% 
+  ungroup() %>% 
+  spread(broadtrophic, regional_density) %>% 
+  group_by(region) %>% 
+  arrange(year) %>% 
+  mutate(lag4_carnivore = lag(carnivore,4),
+         lag4_piscovore = lag(piscivore,4)) %>% 
+  na.omit() %>% 
+  ungroup() %>% 
+  filter(year > 1999) %>% 
+  select(-year,-region) %>% 
+  corrr::correlate() %>% 
+   gather("colname","correlation",-rowname) %>% 
+  ggplot(aes(rowname, colname,fill = correlation)) + 
+  geom_tile() + 
+  geom_text(aes(rowname, colname, label = round(correlation,2))) + 
+  scale_fill_gradient2(low = "tomato", mid = "white", high = "steelblue", midpoint = 0, limits = c(-1,1),
+                       guide = guide_colorbar(frame.colour = "black",barheight = 13)) + 
+  theme(axis.text.x = element_text(angle = 45, vjust = 1,
+                                   hjust = 1),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank())
+
+trophic_effects_plot
+
+```
+
 ## References

### 2020-11-22 documents/ovando-et-al-pop-effects-of-mpas.Rmd
fixed broom.mixed dependency for summary tables

@@
 output:
+  bookdown::word_document2:
+    reference_docx: template.docx
   bookdown::pdf_document2:
     keep_tex: true
-  bookdown::word_document2:
-    reference_docx: template.docx
   bookdown::html_document2: default
@@
 knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 300,
-                      cache = FALSE, fig.width = 3.42, fig.asp = 1.5, dev = "pdf")
+                      cache = FALSE, fig.width = 3.42, fig.asp = 1.5)
 
@@
 library(here)
+library(broom.mixed)
 extrafont::loadfonts()
@@
          }
-did_coefs <- did_model %>% broom::tidy(intervals = TRUE,conf.level = 0.9)
+did_coefs <- did_model %>%  broom.mixed::tidy(conf.int = TRUE,conf.level = 0.9) %>% 
+  rename(lower = conf.low,
+         upper = conf.high)
 
@@
 coefs <- did_fit %>%
-  broom::tidy(intervals = TRUE, prob = 0.89) %>%
+  broom.mixed::tidy(conf.int = TRUE, prob = 0.9) %>%
+    rename(lower = conf.low,
+         upper = conf.high) %>% 
   select(term, estimate, lower, upper) %>%
@@
     Mean = estimate,
-    `Lower 89th Percentile` = lower,
-    `Upper 89th Percentile` = upper
+    `Lower 90th Percentile` = lower,
+    `Upper 90th Percentile` = upper
   )
@@
 
-```{r raw-density, fig.width = width_1col, fig.asp=asp_1col,fig.cap = "Centered and scaled trends in biomass densities of targeted and non-targeted fin-fish included in our study Top panel shows trends across all sites, with smaller background lines showing trends for each individual species.
+```{r raw-density, fig.width = width_1col, fig.asp=asp_1col,fig.cap = "Centered and scaled trends in biomass densities of targeted and non-targeted fin-fish included in our study.
+Top panel shows mean trends across all sites, with shaded areas shwoing 95% confidence intervals.
 Bottom two panels show aggregate biomass density trends outside and inside MPAs"}
@@
 temp_plot <-
-  (total_trend_plot / (mpa_trend_plot)) & theme(legend.position = "top", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), units = "lines"),
+  (total_trend_plot / (mpa_trend_plot)) & theme( plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), units = "lines"),
                                                 panel.grid = element_blank()
@@
   theme_1col + 
-  theme(axis.title.x = element_textbox(size = 10)) + 
+  theme(axis.title.x = element_textbox(size = 10),
+        legend.position = "top") + 
   geom_vline(aes(xintercept = 0), color = "black", linetype = 2)

### 2020-11-22 make-pop-effects-of-mpas.R

@@ -4316,41 +4316,84 @@ classcode_trends <- base_run$did_fit[[1]]$classcode_level_data %>%
 
 targeted_trends <- base_run$did_fit[[1]]$did_data %>%
   group_by(year, targeted) %>%
-  summarise(mbd = mean(total_biomass_density)) %>%
+  summarise(mbd = mean(total_biomass_density),
+            sdbd = sd(total_biomass_density)) %>%
   group_by(targeted) %>%
   mutate(scaled_mbd = scale(mbd))
 
 targeted_trends_by_mpa <- base_run$did_fit[[1]]$did_data %>%
   group_by(year, targeted, eventual_mpa) %>%
-  summarise(mbd = mean(total_biomass_density)) %>%
+  summarise(mbd = mean(total_biomass_density),
+            sdbd = sd(total_biomass_density)) %>%
   group_by(targeted, eventual_mpa) %>%
   mutate(scaled_mbd = scale(mbd)) %>%
   ungroup()
 
+targeted_trends_by_mpa %>% 
+  ggplot(aes(year, scaled_mbd, color = targeted == 1)) + 
+  geom_ribbon(aes(year, ymin = scaled_mbd - 1.96, ymax = scaled_mbd + 1.96, fill = targeted == 1),alpha = 0.25) +
+  geom_line(size = 1.5) + 
+  facet_wrap(~eventual_mpa)
 
-total_trend_plot <- classcode_trends %>%
+
+total_trend_plot <- targeted_trends %>%
   ggplot() +
-  geom_line(aes(
+  geom_line(aes(year, scaled_mbd, color = targeted == 0, linetype = targeted == 0),
+            size = 1.25) +
+  geom_ribbon(aes(
     year,
-    scaled_mbd,
-    group = interaction(targeted, classcode),
-    color = targeted == 0
+    ymin = scaled_mbd - 1.96,
+    ymax = scaled_mbd + 1.96,
+    fill = targeted == 0
   ),
-  alpha = 0.25) +
-  geom_line(data = targeted_trends,
-            aes(year, scaled_mbd, color = targeted == 0),
-            size = 2) +
-  # scale_color_npg(labels = c("Targeted", 'Non-Targeted'), name = "") +
-  scale_colour_grey(labels = c("Targeted", 'Non-Targeted'),
-                    name = "") +
+  alpha = 0.5) +
+  scale_color_manual(
+    values = c("black", "grey50"),
+    labels = c("Targeted", 'Non-Targeted'),
[…]

### 2020-11-22 documents/referee-replies.Rmd
better plots

@@
 
- Thank you for your comment.
-We have moved the operating model equations from the SI (and clarrified that those equations are the simulation model) to Table.XX in the main methods, and clarrified in Section XX.
-the role of these equations in the simulation model.
+Thank you for pointing out the lack of sufficient detail on the equations for the simulation model.
+We have restructured the SI so that the simulation model equations are front  and center (SI Line XX), and clarified our explanations of the equations.
+We have also expanded our description of the simulation model in the text.
+We hope this clarifies the structure behind the simulation model.
 
 *2) On a related note I didn't come away with a clear sense of the scenarios used to simulate the potential fish outcomes.
-For example I have no clear sense how you have simulated spatial patterns of fishing intensity.
-Similarly, it was not clear to me how you decide which model scenarios are most plausible when conditioning on observed response ratios.
+For example I have no clear sense how you have simulated spatial patterns of fishing intensity.*
+
+We apologize for the confusion in the simulation scenarios.
+Given the wide array of scenarios employed, and the word constraints of the paper, we unfortunately could not include a thorough description of every scenario in the main text.
+We have however included a clearer summary of the simulation in the methods section, highlighting among other key attributes the ways in which we simulate spatial patterns of fishing intensity, which as Referee #2 has noted is a critical variable to consider.
+
+*Similarly, it was not clear to me how you decide which model scenarios are most plausible when conditioning on observed response ratios.
 Yet conveying the ways in which observed RR could be derived would be useful.
@@
 
-Declarative sentance for what this figure is out. Results of ... Insight first, then terms at the end. warmer colors indicate larger effect
+Thank you for pointing this out.
+We have redesigned Figure 2 to make clearer that the raw data a) support the parallel trends assumption between the targeted and non-targeted populations pre-MPA and b) that the raw data support the statistical finding of no clear MPA effect.
+We have rewritten the caption to make Figure 6 clearer: we acknowledge that it is a complicated figure but we feel that it is important to convey the range of possible outcomes produced by our simulation model.
 
-Add MPA implementation line. Make clearer in the figure caption. Removing lindividial lines would artificially make it clearer
-
-We apologize for the confusion in the figures, we have...
+Declarative sentance for what this figure is out. Results of ... Insight first, then terms at the end. warmer colors indicate larger effect XX
 
@@
 We include hierarchical terms for each site in the data, clustered by the specific island in the Channel Islands a specific site was located at.
-This should serve to control for any intrinsic traits of the MPA sites themselves, e.g.
+This serves to control for any intrinsic traits of the MPA sites themselves, e.g.
 if MPA sites were preferentially placed on better habitat than "control" sites.
@@
 *Fig 2: can you make the lines a bit narrower?*
-Lines are narrower and XX
+
+We have cleaned up Fig.2 substantially, and hope that the message that the raw data a) support the parallel trends assumption between the targeted and non-targeted populations pre-MPA and b) that the raw data support the statistical finding of no clear MPA effect.
+We have rewritten the caption to make Figure 6 clearer: we acknowledge that it is a complicated figure but we feel that it is important to convey the range of possible outcomes produced by our simulation model.
 
@@
 
-We have used XX to indicate the pre-MPA period
+We have clarified that all the year bins are post-MPA, since the MPAs went in place in 2003, in the figure caption.
+We worry that given the large amount going on in Figure 4 already adding another shaded area might be confusing.
 
@@
 
-Thank you for pointing this out, and have rephrased this section accordingly (line XX)/
+Thank you for pointing this out, and have rephrased this section accordingly (line XX) to reflect that this is something that stakeholders should be aware of, as demonstrated in Little et al. 2009.
 
@@
 
+We believe there is some confusion, this is the same simulation model as presented in the methods (line XX) and used elsewhere in the paper.
+This clearly was not evident in the text though, we have clarified that this is the same model on line XX.
+
 *228: this paragraph belongs in results.*
 
-We have moved this section to the results
+We have moved this section to the results.
 
@@
 As we discuss on line XX, what evidence we have is that poaching is of limited concern in the Channel Islands (but clear can be a major issue in many MPAs).
-However, the presence of poaching should not bias the estimates of the DiD model in any way: it poaching were severe enough so as to render the MPAs completely ineffective, then in theory a DiD model such as the one implemented here would simply estimate that the MPA had no effect, as their would be no difference in the conditional biomass densities of targeted and non-targeted species.
+However, the presence of poaching should not bias the estimates of the DiD model in any way: if poaching were severe enough so as to render the MPAs completely ineffective, then in theory a DiD model such as the one implemented here would simply estimate that the MPA had no effect, as their would be no difference in the conditional biomass densities of targeted and non-targeted species.
+
+However, poaching might certainly influence ones interpretation of the results.
+If poaching is rampant but stakeholders are largely unaware of it, a model like this might estimate the the MPAs had no effect (which in that case they might not).
+Stakeholders might then interpret that as a problem of MPA design, a misunderstanding of the movement dynamics or fishing pressure in the fishery, or any other number of variables, when in fact the problem was simply that the MPA was not actually acting as an MPA.
 

### 2020-11-22 documents/ovando-et-al-pop-effects-of-mpas.Rmd
better plots

@@
 Constant-catch dynamics might also occur in fisheries with constraining quotas that are not updated after the implementation of MPAs.
-Fishers pursuing constant catch in areas outside an MPA may have to fish harder to achieve the same catch from a smaller part of the population, causing a population conservation loss under `r percent(1 - positive_outcomes$pp[positive_outcomes$fleet_model == "constant-catch"], accuracy = 1)` of our constant-catch simulations.
-This is an important scenario to consider (as done so in @little2011b) especially as MPAs are increasingly implemented in quota-managed fisheries [@liu2018].
+Fishers pursuing constant-catch in areas outside an MPA may have to fish harder to achieve the same catch from a smaller part of the population, causing a population conservation loss under `r percent(1 - positive_outcomes$pp[positive_outcomes$fleet_model == "constant-catch"], accuracy = 1)` of our constant-catch simulations.
+This potential negative interaction between constant-catch and MPAs is important risk to consider (as done so in @little2011b) especially as MPAs are increasingly implemented in quota-managed fisheries [@liu2018].
 

### 2020-11-22 make-pop-effects-of-mpas.R

@@ -4338,15 +4338,19 @@ targeted_trends_by_mpa %>%
 
 total_trend_plot <- targeted_trends %>%
   ggplot() +
+  geom_vline(aes(xintercept = 2003), linetype = "dotted", size = 1) +
+  geom_segment(x = 2006, xend = 2003.25, y = 4, yend = 4, arrow =arrow(length = unit(0.2,"cm"))) +
+  geom_text(x = 2008.5, y = 4, label = "MPAs Implemented", nudge_x = 5, size = 4) +
   geom_line(aes(year, scaled_mbd, color = targeted == 0, linetype = targeted == 0),
-            size = 1.25) +
+            size = 1) +
   geom_ribbon(aes(
     year,
     ymin = scaled_mbd - 1.96,
     ymax = scaled_mbd + 1.96,
-    fill = targeted == 0
-  ),
-  alpha = 0.5) +
+    fill = targeted == 0),
+    outline.type = "both",
+    color = "black",
+    alpha = 0.5) +
   scale_color_manual(
     values = c("black", "grey50"),
     labels = c("Targeted", 'Non-Targeted'),
@@ -4369,14 +4373,16 @@ targlab <- c(`TRUE` = "Inside MPAs",
 
 mpa_trend_plot <- targeted_trends_by_mpa %>%
   ggplot() +
+  geom_vline(aes(xintercept = 2003), linetype = "dotted", size = 1) +
   geom_line(aes(year, scaled_mbd, color = targeted == 0, linetype = targeted == 0),
-            size = 1.25) +
+            size = 1) +
   geom_ribbon(aes(
     year,
     ymin = scaled_mbd - 1.96,
     ymax = scaled_mbd + 1.96,
-    fill = targeted == 0
-  ),
+    fill = targeted == 0),
+  outline.type = "both",
+  color = "black",
   alpha = 0.5) +
   scale_color_manual(
     values = c("black", "grey50"),
@@ -4384,7 +4390,7 @@ mpa_trend_plot <- targeted_trends_by_mpa %>%
     name = ""
   ) +
   scale_fill_manual(
-    values = c("black", "lightgrey"),
+    values = c("black", "grey50"),
     labels = c("Targeted", 'Non-Targeted'),
     name = ""
   ) +
@@ -4392,20 +4398,6 @@ mpa_trend_plot <- targeted_trends_by_mpa %>%
   scale_x_continuous(name = '') +
   scale_y_continuous(name = "") +
[…]

### 2020-11-25 documents/referee-replies.Rmd
updated draft

@@
 date: "11/11/2020"
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 ---
@@
 
+XX Throughout introduction XX
+
 *- A clearer explanation of the gap in the current literature that you are willing to fill is needed*
@@
 We have provided a clearer statement of the gap in the current literature that our paper fills on line XX, and line XX
+
+XX Paragraph in intro starting with "While these studies have clearly demonstrated the ability" XX
 
@@
 Thank you for pointing this out, we have clarified the paragraph in question and provided greater detail on why exactly this existing body of evidence does not neccesarily serve as a population-level measure of MPA effects (Line XX).
+
+XX Paragraph in intro starting with "While these studies have clearly demonstrated the ability" XX
 
@@
 
+XX Paragraph in intro starting with "While these studies have clearly demonstrated the ability" XX
+
 *- Please cite and explain briefly which methods you are using in this article*
@@
 We take this to mean a explanation and citation for difference-in-difference.
-We have included a summary citation for the method (line XX), as well as a brief description, given the length constraints of this paper, explaining the link between BACI and difference in difference.
+We have included a summary citation for the method (line XX, start of methods), as well as a brief description, given the length constraints of this paper, explaining the link between BACI and difference in difference.
+
+XX paragraph in intro starting with "To fill this important gap"
 
@@
 We have clarified the objectives of this study, and its place in the literature, on lines XX and XX
+
+XX basically the whole introduction XX
 
@@
 Line XX
+XX Simulation Model part of the methods section XX
 
@@
 We apologize but we are not clear what this comment is in reference to.
-We do not use log-ratios in our model, except in the sense that we use a Bayesian Gamma GLMER with a log link function to run our , so the estimated effect sizes are in log scales, roughly corresponding to percentage-level effects.
+We do not use log-ratios in our model, except in the sense that we use a Bayesian Gamma GLMER with a log link function to run our difference-in-diffecence (basically BACI) , so the estimated effect sizes are in log scales, roughly corresponding to percentage-level effects.
+
+We have however included an equation to clarify the BACI-like nature of our regression, on line XX
+
+XX methods section starting with 
+For readers more familiar with BACI methodXX
 
@@
 
-We have briefly addressed our choice of Bayesian methods on line XX.
+We have briefly addressed our choice of Bayesian methods on line XX
+
+XX end of methods section on diff-in-diff XX
 
@@
 Thank you for pointing out that these terms are poorly described in the text, we have improved our descriptions (line XX).
+
+XX methods simulation model section XX
 
@@
 
-XX made it here XX
+XX intro to methods, end of diff-in-diff section XX
+
 Results:
@@
 We do test for the effect of MPA size, which we make clearer in our methods description, on line XX.
+
+XX methods section XX
 
@@
 
+XX The location and spacing of the MPAs within ... XX
+
 *Line 44: “commonly measured as the ratio of biomass densities of species” Please define what do you mean by “biomass densities”.*
@@
 
-XX Owen check if you think this is clearer now.
+XX Owen check if you think this is clearer now. basically most fo the intro XX
 
@@
 
-We have sought to clarify the specific problem of SUTVA violoations (treated MPA sites affecting nearby control sites) that justify why response ratio studies do not necessarily serve as an index of population level effects, and in our methods clarified that it is not the DiD model itself that resolves this problem (which is akin to a BACI), but rather using non-targeted species as the control group, rather than relying on spatial controls.
+We have sought to clarify the specific problem of SUTVA violations (treated MPA sites affecting nearby control sites) that justify why response ratio studies do not necessarily serve as an index of population level effects, and in our methods clarified that it is not the DiD model itself that resolves this problem (which is akin to a BACI), but rather using non-targeted species as the control group, rather than relying on spatial controls.
+We have also modified the language in that sentence to be more nuanced.
 
@@
 
+XX Simulation model part of the methods XX
+
 *2) On a related note I didn't come away with a clear sense of the scenarios used to simulate the potential fish outcomes.
@@
 We apologize for the confusion in the simulation scenarios.
-Given the wide array of scenarios employed, and the word constraints of the paper, we unfortunately could not include a thorough description of every scenario in the main text.
+Given the wide array of scenarios employed, and the word constraints of the paper, we unfortunately could not include a thorough description of every scenario in the main text (line XX).
 We have however included a clearer summary of the simulation in the methods section, highlighting among other key attributes the ways in which we simulate spatial patterns of fishing intensity, which as Referee #2 has noted is a critical variable to consider.
+
+XX Simulation model part of the methods XX
 
@@
 
-How did you decide which scenarios to run, given the results how did you interpret them. THis is how we interpret that request
+We have expanded our description of the simulation model in an effort to explain this more clearly, and in particular expanded our matching process to address this comment (line XX.).
+Our Channel Islands simulations were designed to match the genera of species observed in the Channel Islands, our best sense of the exploitation history of those species, and the potential size of the MPAs for those species.
+For example, we know that all of the finfish species in this analysis are part of populations that span at least the range of the Channel Islands.
+Since the Channel Islands MPAs make up roughly 20% of the Islands, we know that the MPAs are likely to be no more than 20% of a species range, which we incorporate into the simulations.
 
-explain more the variety of ways that these results could have been observed
+For each of these simulations intended to reflect the broad context of the Channel Islands, we sampled the simulated biomass density response ratios.
+We then matched these simulated response ratios to the empirical response ratios observed in the Channel Islands.
+So, for example, if we observe a response ratio of 2 after 4 years of protection in the Channel Islands, we matched that results to all simulated results that also had a response ratio of 2 after 4 yearse of MPA protection.
 
-Given the word constraints of the paper, we felt that most of the simulation detail had to be left to Table.XX, Supplemental Information section XX.
-We agree though that readers would benefit from more clarity in the main text.
-Given the particularly critical nature of movement and fishing behavior in MPA models, we have included the following statement on line XX of the methods.
-
-XX
+XX see end of simulation model part of methods XX
 
@@
 
-We are glad that the reviewr found our contribution important and novel, thank you!
+We are glad that the reviewer found our contribution important and novel, thank you!
 
@@
 We use the term difference-in-difference to tie our results more clearly to the causal inference literature in econometrics.
-We have also clarrified on line XX the clear link between DiD and BACI.
+We have also clarified on line XX the clear link between DiD and BACI.
+
+XX DiD methods section XX
 
@@
 no binning) to make clear that our results are not sensitive to the choice of binwidth.
+
+XX end of diff in diff methods XX
 
@@
 
-This is correct, it was quite buried in the SI, but we do weight the MPA and non-MPA sites according to the total area of MPAs in the seascape (SI line XX), since as the referee points out the sampling regime is not random.
-We have mentioned this in the main methods as well now on line XX.
+This is correct.
+it was quite buried in the SI, but we do weight the MPA and non-MPA sites according to the total area of MPAs in the seascape (SI line XX), since as the referee points out the sampling regime is not random.
+Otherwise, if the MPAs caused an increase of 50% inside, and 0% outside, given that the PISCO sampling regime has roughly equal sampling inside and outside MPAs, the model would estimate a net effect of 25%, even though MPAs in fact make up roughly 20% of the area.
+
+We have clarified that we weighted the regressions to account for this non-uniform sampling regime on line XX.
 We also include an unweighted version of our results, to demonstrate that our results are not sensitive to this weighting choice (Fig.SXX).
+
+XX diff in diff methods XX
 
@@
 Thank you for this point!
-To make things more streamlined, we have removed the transect-level DiD from the SM, as it is VERY long, very convoluted, and is by far the most difficult part of the code for others to replicate given the long run time.
+We reorganized the supplemental materials for ease of reading.
+*However, we have not removed sections at this time.
+Each of the analyses included in the SM (beyond that basic explanations of the methods) were a result of a question that we received from prior reviewers or peers.
+As such, we feel that it is better to include these in there for the odd person that really wants to see the results via synthetic controls.
+Given that the referees primary concern here was our not wasting useful analyses (which we appreciate!), and that we do not intent do dive further into the empirics of this particular question in the Channel Islands at this time, we have elected to go with the conservative approach of a perhaps excessively thorough SM.
 
@@
 
-Thank you for this excellent suggestion. We have reflected this suggestion on line XX.
+Thank you for this excellent suggestion that we have sought to reflect on line XX.
+
+ XX Introduction XX
 
@@
 
-We agree that inclusion of a wider set of references may broaden the impact of our paper and have included the recommended references, thank you for the suggestion.
+Thank you for pointing this out.
+We only see one suggested reference (Krueck et al.
+2017), which we have incorporated on line XX.
+We have also cited @fulton2015 in order to link our mork more explicitly to the broader literature on MPA simulation model.
 
@@
 
-*89: Nils Kruek did a similar thing in his paper, it might be worth reading and relevant citing:
-https://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.2000537*
+*89: Nils Kruek did a similar thing in his paper, it might be worth reading and relevant citing:https://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.2000537*
 
-Thank you for pointing this out XX dig into this one
+Thank you for pointing this out, we have included this citation, as well as @fulton2015, to place are results more clearly into a broader literature of MPA simulation (Line XX)
 
@@
 We have clarified that the simulation keeps track of outcomes at a variety of timesteps, including non-equilibrium.
-The "begins at" is just to clarify that we run the population out to equiblirum since there isn't an analytical solution to the biomass distribution in an age-structured spatial model when habitats vary.
+The "begins at" is just to clarify that we run the population out to equilibrium to start the simulation since there isn't an analytical solution to the biomass distribution in an age-structured spatial model when habitats vary.
 
@@
 
-That is correct, so if the MPAs cover 50% of the habitat, biomass density goes 20% inside, 10% outside, our model would estimate 15%.
+That is correct, so if the MPAs cover 50% of the habitat, MPAs cause biomass density goes 20% inside, 10% outside, our model would estimate 15%.
 We do however include site-specific intercepts that control for different mean baseline densities at each site, as well as site-specific environmental covariates (e.g.
@@
 
-*Eqn 1: Don’t the gamma shape and scale depend on its mean? So maybe more correct to write this as:
-Expectation(d) = exp(beta….) and then say observations were sampled from a gamma, where the dispersion parameter was free/estimated.*
+*Eqn 1: Don’t the gamma shape and scale depend on its mean?
+So maybe more correct to write this as: Expectation(d) = exp(beta….) and then say observations were sampled from a gamma, where the dispersion parameter was free/estimated.*
 
-Thank you for the suggestion, we have modified the text accordingly.
+Thank you for the suggestion, in keeping with the likelihood notation here, we have kept the shape (dispersion) parameter in the equation, since the Gamma likelihood does depend on it, but you are right that the scale is a transformation of the mean and the shape and so should be (and has been) omitted.
 
-*It could be helpful to show an equation for the net DiD outcome (as explained at A.3). I think its something like:
-(D[post, fished]-D[pre, fished]) – (D[post, unfished]-D[pre, unfished])*
+*It could be helpful to show an equation for the net DiD outcome (as explained at A.3). I think its something like: (D[post, fished]-D[pre, fished]) – (D[post, unfished]-D[pre, unfished])*
 
-We have included this section from the SM XX.
+We agree this would help readers understand the model and have included this suggestion on line XX of the methods.
 
@@
 Given that one of our main results is that we are unable to detect a clear signal, we feel that the possible risk of underestimating the error distributions is outweighted by the potential gains of helping control for omitted variable bias.
+
+XX Owen I'll work on this while you check my edits XX
 
@@
 
+XX paragraph starting with "Stock assessment models that account for the population within the MPAs [@" XX
+
 Chris Brown

### 2020-11-25 documents/ovando-et-al-pop-effects-of-mpas-si.Rmd
updated draft

@@
 output: 
+  bookdown::pdf_document2:
+    keep_tex: true
   bookdown::word_document2:
     reference_docx: template.docx
-  bookdown::pdf_document2:
-    keep_tex: true
   pdf_document: default
@@
 
-For the first synthetic control, we pulled our standardized mean index of abundance for targeted species as a whole from our difference-in-difference model as our treated group.
-We then pull the standardized indices of abundance for each of the non-targeted groups from the difference-in-difference to use as the candidate untreated components for the synthetic controls.
+For the first synthetic control, we first calculated the mean total biomass density of targeted species across the Channel Islands, our outcome variable of interest.
+We then calculated the mean total biomass dentisy of each individual non-targeted species in our database as a set of donor data to construct our synthetic control.
 A complete synthetic control analysis would require more extensive validation of the methods, but we use this approach simply to explore whether we observe substantially divergent results in the synthetic control versus the difference-in-difference model.
@@
 
-```{r, include = FALSE, eval = FALSE}
+```{r, include = FALSE, eval = TRUE}
 
@@
 
-nontargeted_index <- data_frame(abundance_hat = fit_report$abundance_hat,
-                               classcode = rep(1:n_distinct(used_data$classcode), each = length(years))) %>%
-  mutate(log_abundance_hat = log(abundance_hat)) %>%
-  group_by(classcode) %>%
-  mutate(year =years) %>%
-  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
+targeted_index <- base_run$did_fit[[1]]$did_data %>%
+  group_by(year, targeted) %>%
+  summarise(mtbd = mean(total_biomass_density)) %>%
+  group_by(targeted) %>%
+  filter(targeted == 1) %>% 
+  mutate(classcode = "targeted")
+
+nontargeted_index <- abundance_data$data[[which(abundance_data$data_source == "pisco")]] %>% 
+  group_by(year, targeted, classcode, site_side) %>% 
+  summarise(tb = sum(total_biomass_g)) %>% 
+  group_by(year, targeted, classcode) %>% 
+  summarise(mtbd = mean(tb)) %>% 
+  filter(targeted == 0) %>% 
   ungroup() %>% 
-  rename(numeric_classcode = classcode) %>%
-  left_join(life_history, by = "numeric_classcode" ) %>% 
-  left_join(unique(abundance_data$data[[1]] %>% select(classcode, targeted)), by = "classcode") %>% 
-  filter(targeted == 0) %>% 
-  select(log_abundance_hat, year, classcode, targeted)
-
-targeted_index <- tmb_runs$tmb_fit[[1]]$zissou_estimates %>%
-  filter(
-    stringr::str_detect(variable, "(targeted_did_betas)"),!str_detect(variable, "non")
-  ) %>%
-  mutate(
-    year = years,
-    classcode = "targeted",
-    targeted = 1
-  ) %>%
-  rename(log_abundance_hat = estimate) %>% 
-  select(colnames(nontargeted_index))
-
-  # filter(variable == "seeing_year_species_betas") %>%
+  select(colnames(targeted_index))
 
@@
   left_join(synth_covariates, by = c("classcode","year")) %>% 
+  ungroup() %>% 
   mutate(numeric_classcode = as.numeric(factor(classcode))) %>% 
   group_by(classcode) %>% 
-  mutate(log_abundance_hat = scale(exp(log_abundance_hat))) %>% 
+  mutate(scaled_mtbd = as.numeric(scale(mtbd))) %>% 
   ungroup() %>% 
@@
 
+synth_data %>% 
+  group_by(numeric_classcode, year) %>% 
+  count()
+
 prepped_synth_data <-
@@
     predictors.op = "mean",
-    dependent = "log_abundance_hat",
+    dependent = "scaled_mtbd",
     unit.variable = "numeric_classcode",
@@
 
-```{r  total-synth, fig.cap = "Difference in centered and scaled standardized targeted abundance and synthetic standardized targeted abundance", eval=FALSE}
+```{r  total-synth, fig.cap = "Difference in centered and scaled standardized targeted abundance and synthetic standardized targeted abundance", eval=TRUE}
 total_gap_plot
-```
-
-```{r raw-synth, include = FALSE, eval = FALSE}
-
-  consistent_classcodes <- used_data %>%
-      group_by(classcode, region) %>%
-      filter(any_seen) %>%
-      summarise(n_years = n_distinct(year)) %>%
-      ungroup() %>%
-      filter(n_years == n_distinct(used_data$year))
-
-synth_data <- used_data %>% 
-  filter(classcode %in% consistent_classcodes$classcode,
-         region %in% consistent_classcodes$region) %>% 
-  mutate(classcode = ifelse(targeted == 1, "targeted", classcode),
-         region = ifelse(targeted == 1, "targeted", region)) %>% 
-      group_by(year,
-               site_side,
-               region,
-               zone,
-               transect,
-               eventual_mpa,
-               classcode,
-               targeted) %>%
-      summarise(
-        total_classcode_density = sum(exp(log_density)),
-        var_tex = sum(cumulative_n_obs),
-        var_vis = mean(mean_vis),
-        var_temp = mean(mean_temp),
-        var_depth = mean(mean_depth),
-        var_surge = mean(surge),
-        var_kelp = mean(interp_kelp)
-      ) %>%  # sum density across all levels of a transect
-      group_by(year, site_side, region, eventual_mpa, classcode, targeted) %>%
-      summarise(
-        md = mean(total_classcode_density),
-        var_tex = mean(var_tex),
-        var_vis = mean(var_vis),
-        var_temp = mean(var_temp),
-        var_depth = mean(var_depth),
-        var_surge = mean(var_surge),
-        var_kelp = mean(var_kelp)
-      ) %>% # calculate mean density per year site, side, species, averaging over zone, transect
-      ungroup()
-
- synth_data <- synth_data %>% 
-     group_by(year,region,classcode,targeted) %>%
-      summarise(
-        total_biomass_density = (sum(md) / 1e6) * 10000,
-        # calculate total and mean biomass densities across all species per year site side
-        mean_biomass_density = log((mean(md) / 1e6) * 10000 + 1e-3),
-        var_tex = mean(var_tex),
-        var_vis = mean(var_vis),
-        var_temp = mean(var_temp),
-        var_kelp = mean(var_kelp)
-      ) %>%
-   ungroup() %>% 
-   mutate(classcode = ifelse(targeted == 1, classcode, paste(classcode, region, sep = '_'))) %>% 
-      mutate(numeric_classcode = as.numeric(factor(classcode))) %>% 
-   as.data.frame()
-
-prepped_synth_data <-
-  Synth::dataprep(
-    foo = synth_data,
-    predictors = colnames(synth_data)[str_detect(colnames(synth_data),"var")],
-    predictors.op = "mean",
-    dependent = "mean_biomass_density",
-    unit.variable = "numeric_classcode",
-    unit.names.variable = "classcode",
-    time.variable = "year",
-    treatment.identifier = "targeted",
-    controls.identifier = unique(synth_data$classcode[synth_data$targeted == 0]),
-    time.predictors.prior = years[years <= 2003],
-    time.optimize.ssr = years[years <= 2003],
-    time.plot = years
-  )
-
-synth_fit <- Synth::synth(prepped_synth_data)
-
-gaps <-
-  rownames_to_column(
-    prepped_synth_data$Y1plot - (prepped_synth_data$Y0plot %*% synth_fit$solution.w) %>%
-      as.data.frame()
-  ) %>%
-  rename(gap = w.weight,
-         year = rowname) %>%
-  mutate(year = as.numeric(year)) %>%
-  mutate(synthetic = as.numeric(prepped_synth_data$Y0plot %*% synth_fit$solution.w)) %>% 
-  mutate(treatment = gap + synthetic)
-
-# path.plot(synth.res = synth_fit,
-#           dataprep.res = prepped_synth_data
-# )
-
-total_gap_plot <- gaps %>% 
-  ggplot(aes(year, gap)) + 
-  # geom_hline(aes(yintercept = 0), linetype = 2) +
-  geom_vline(aes(xintercept = 2003), color = "red")+
-  geom_line() + 
-    labs(x = "Year", y = "Standard deviations from synthetic control")
-
-total_synth_plot <- gaps %>% 
-  select(year, synthetic, treatment) %>% 
-  gather(source, value, -year) %>% 
-  group_by(source) %>% 
-  # mutate(value = scale(value)) %>% 
-  ggplot(aes(year, value, color = source, linetype = source)) + 
-  geom_vline(aes(xintercept = 2003), color = "red")+
-  geom_line() + 
-    labs(x = "Year", y = "Standard deviations from synthetic control")
-
 ```
@@
 
-```{r, include = FALSE, eval=FALSE}
+```{r, include = FALSE, eval=TRUE}
 
@@
         predictors.op = "mean",
-        dependent = "log_abundance_hat",
+        dependent = "scaled_mtbd",
         unit.variable = "numeric_classcode",
@@
 
-nontargeted_index <- data_frame(abundance_hat = fit_report$abundance_hat,
-                               classcode = rep(1:n_distinct(used_data$classcode), each = length(years))) %>%
-  mutate(log_abundance_hat = log(abundance_hat)) %>%
-  group_by(classcode) %>%
-  mutate(year = 1999 + 1:length(abundance_hat)) %>%
-  mutate(scaled_abundance_hat = (abundance_hat - mean(abundance_hat)) / sd(abundance_hat)) %>%
+nontargeted_index <- abundance_data$data[[which(abundance_data$data_source == "pisco")]] %>% 
+  group_by(year, targeted, classcode, site_side) %>% 
+  summarise(tb = sum(total_biomass_g)) %>% 
+  group_by(year, targeted, classcode) %>% 
+  summarise(mtbd = mean(tb)) %>% 
   ungroup() %>% 
-  rename(numeric_classcode = classcode) %>%
-  left_join(life_history, by = "numeric_classcode" ) %>% 
-  left_join(unique(abundance_data$data[[1]] %>% select(classcode, targeted)), by = "classcode") %>% 
-  select(log_abundance_hat, year, classcode, targeted)
+  select(colnames(targeted_index))
 
@@
   group_by(classcode) %>% 
-  mutate(log_abundance_hat = scale(exp(log_abundance_hat))) %>% 
+  mutate(scaled_mtbd = scale(mtbd)) %>% 
   ungroup() %>% 
@@
 
-```{r synth-classcode-fits, fig.cap="Synthetic control gaps for each targeted species", eval=FALSE}
+```{r synth-classcode-fits, fig.cap="Synthetic control gaps for each targeted species", eval=TRUE}
 synth_classcode_fits_plot + 

### 2020-11-25 documents/ovando-et-al-pop-effects-of-mpas.Rmd
updated draft

@@
 ---
-title: Assessing the Population-level Conservation Effects of Marine Protected Areas
+title: Assessing the Population-Level Conservation Effects of Marine Protected Areas
 abstract: |
@@
 Our results provide a novel assessment of the population-level effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.
-bibliography: references.bib
+bibliography: references.yaml
 csl: conservation-biology.csl
@@
   - \usepackage{lineno}\linenumbers
+  - \usepackage{draftwatermark}
+  - \SetWatermarkText{DRAFT}
 always_allow_html: true
@@
 The network has been used as a model case study in protected area design around the world (@botsford2014 and references therein).
-We use data from the first 14 years of protection to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of fin-fish species.
+We use data from the first 14 years of protection in a difference-in-difference mode [see @angrist2009 and references therein] to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of fin-fish species.
 In contrast to clear differences in biomass densities observed inside and outside of well-protected MPAs both globally [@lester2009] and in the Channel Islands [@caselle2015] we are unable to detect a clear population effect from the Channel Islands MPAs.
@@
 We use the term difference-in-difference here to more closely tie our results to the terminology in causal-inference literature.
-This difference-in-difference regression provides our estimate of the population-level conservation effect of the Channel Islands MPAs.
-We use a Bayesian regression since it allows us to interpret our estimated effects probabalistically [@mcelreath2020], e.g.
-stating the probability that conditional on the model, priors,and data the Channel Island MPAs increased or decreased biomass of targeted species.
-Being a Bayesian regression, our difference-in-difference model produces posterior probability distributions of our coefficients, from which we can construct Bayesian credible intervals.
-
-As we have discussed, theory shows that *a priori* we would expect the population-level effects of MPAs to be highly context dependent.
-We use the bio-economic simulation model to provide provide theoretical expectations for population-level effects of MPAs, which we then confront with the empirical results from our difference-in-difference model.
+The population-level effects of MPAs will depend on the social-ecological contexts in which they occur.
+We use our bio-economic simulation model to provide theoretical expectations for population-level effects of MPAs, which we then confront with the empirical results from our difference-in-difference model.
 
@@
 Readers can explore the functionality of the model using an online tool available at [danovando.shinyapps.io/simmpa/](https://danovando.shinyapps.io/simmpa/).
-
 The purpose of the simulation model is to both set expectations for our empirical results, and demonstrate more broadly the ways in which ecological and economic dynamics can interact to produce a wide range of population-level MPA effects.
 The full range of factors explored are shown in  (Table.\@ref(tab:sim-table)), and the equations of the simulation model can be viewed in the SI section XX.
-We used this model to generate XX simulated MPA outcomes across XX species.
+We used this model to generate `r n_distinct(outcomes$experiment)` simulated MPA outcomes across `n_distinct(outcomes$scientific_name)` species.
 
+Many papers have presented simulation analyses of MPA outcomes [@fulton2015].
+Our model incorporates a broad range of core ecological and economic drivers of MPA performance assessed by individual papers into a cohesive model, similar in spirit to @krueck2017.
 The simulation model consists of 50 patches with wrapped edges.
@@
 The simulation then begins to apply the fleet model to the population, and in a randomly selected year implements an MPA network.
-The model samples a percentage of the population's range to place in MPAs, and randomly assigns patches to MPAs either across across the system or preferentially on higher-quality habitat.
+The model samples a percentage of the population's range to place in MPAs, and randomly assigns patches to MPAs either across across a uniform system or preferentially on higher-quality habitat.
 The model then randomly selects whether fishing effort that used to operate inside the MPAs is redistributed to the areas outside the MPAs, or leaves the fishery entirely.
-We then contine the simulations with the MPAs in place.
+We then continue the simulations with the MPAs in place.
 Each simulation is paired with a simulation identical in every way except that MPAs are no longer put in place.
 We then calculate the effect of the MPAs on the population as the difference in relevant metrics in the simulation with MPAs relative to their values in the simulation without MPAs.
-We ran this process XX times, over which we simulate outcomes for XX different species.
 
@@
 One set of simulations is specifically designed to reflect the dynamics of the Channel Islands.
-We only include species of the same genus as those targeted by fishing in the Channel Islands, restrict fishing pressure such that the simulated popuilations are moderately to lightly exploited (given the best available knowledge of stock status for the targeted species in the Channel Islands), and cap the MPA size at 20% of the population's range (a likely upper end to the effective size of the Channel Island MPAs relative to the fish populations in question, @rassweiler2012).
+We only include species of the same genus as those targeted by fishing in the Channel Islands, restrict fishing pressure such that the simulated populations are moderately to lightly exploited (given the best available knowledge of stock status for the targeted species in the Channel Islands), and cap the MPA size at 20% of the population's range (a likely upper end to the effective size of the Channel Island MPAs relative to the fish populations in question, @rassweiler2012).
+For each of these Channel Island simulations, we calculated the biomass density response ratios for each simulation.
+We then binned both the simulated and empirical response ratios, and joined our simulated values to the empirical values based on matches of these bins.
+This provides a library of simulated population-level MPA effects that could have produced the types of empirical response ratios seen in the Channel Islands.
 
-## Difference in Difference Regression {#did .unnumbered}
+## Difference-in-Difference Regression {#did .unnumbered}
 
@@
 Note that this is akin to a BACI, but rather than relying on spatial control sites as our identification, we utilize non-targeted species as our control.
+To account for the fact that sampling locations are not uniformly distributed across the islands, we weight the samples in our regression in proportion to the area inside and outside of MPAs in the Channel Islands.
 
@@
 If the MPA increases biomass densities inside its boundaries by 20%, and by 0% outside the reserve, the population effect of the MPA would be 10%.
+
+For readers more familiar with BACI methods, the difference-in-difference (DiD) regression amounts to estimating the pre-post MPA difference in the biomass densities of targeted species minus the same difference for non-targeted species in the Channel Islands.
+
+$$(log(D_{MPA=1, T = 1})-log(D_{MPA=0, T = 1}))-(log(D_{MPA = 1,T = 0})-log(D_{MPA =0, T = 0}))$$
+
+Where *T* indicates targeted species, *MPA* indicates whether MPAs are in place, and *D* is the observed mean total biomass density across all sites.
 
@@
 
-The simplified form of this model is
+The simplified regression form of this model is
 
 \begin{equation}
-  d_{i} \sim Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + \mathbf{B^cX_i} + \mathbf{B^sS_i}},shape, scale)
+  d_{i} \sim Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + \mathbf{B^cX_i} + \mathbf{B^sS_i}},shape)
 \label{eq:did}
@@
 The shape and scale parameters of the Gamma distribution are estimated as well (Table.\@ref(tab:coef-table)).
+
+This difference-in-difference regression provides our estimate of the population-level conservation effect of the Channel Islands MPAs.
+We use a Bayesian regression since it allows us to interpret our estimated effects probabalistically [@mcelreath2020], e.g.
+stating the probability that conditional on the model, priors,and data the Channel Island MPAs increased or decreased biomass of targeted species.
+Being a Bayesian regression, our difference-in-difference model produces posterior probability distributions of our coefficients, from which we can construct Bayesian credible intervals.
+Since the effects of MPAs will evolve over time, we use this model to estimate a series of population-level effects of the Channel Islands MPAs in three-year intervals from 2003 to 2017.
 
@@
 
-Stock assessment models that account for the population within the MPAs [@field2006a] may be able to answer the management relevant question of whether fishing mortality rates and biomass levels from systems containing MPAs are in-line with management objectives [e.g.
-@nickols2019].
-However, such an approach does not necessarily shed light on whether the MPAs themselves caused the estimated state of the population, and of course are highly data intensive, potentially restricting our ability to provide stock-assessment based inference of MPA outcomes for broad arrays of targeted species.
+Stock assessment models that account for the population within the MPAs [@field2006a] may be able to answer the management relevant question of whether fishing mortality rates and biomass levels from systems containing MPAs are in-line with management objectives.However, such an approach does not necessarily shed light on whether the MPAs themselves caused the estimated state of the population, and of course are highly data intensive, potentially restricting our ability to provide stock-assessment based inference of MPA outcomes for broad arrays of targeted species.
+@nickols2019 and @white2011 provide excellent examples of using models fit to local data to set expectations for the outcomes of MPAs.
+In the manner of @abadie2010], models such as these that have been demonstrated to be reliable predictors of the true dynamics of the system in the pre-MPA period can be used as an estimate for plausible outcomes in a world without MPAs in the post-MPA period and subsequently as a strategy for estimating the population-level effects of MPAs.
+In contrast, we use our simulation model to set expectations but use an empirical difference-in-difference approach to construct our counterfactual of likely outcomes of the Channel Islands in the absence of MPAs.
+Both approaches can help address the problem of assessing the population-level effects of MPAs.
 

### 2020-11-25 make-pop-effects-of-mpas.R

@@ -4268,7 +4268,7 @@ sample_location_plot <-  ggplot() +
     data = channel_islands_mpas,
     fill = "lightgrey",
     alpha = 0.25,
-    color = "tomato"
+    color = "black"
   ) +
   geom_sf(data = california, fill = "darkgrey") +
   geom_point(data = cities, aes(x = long, y = lat)) +
@@ -4297,12 +4297,15 @@ sample_location_plot <-  ggplot() +
   ) +
   #   coord_sf(xlim = c(bbox['xmin'] - .3, bbox['xmax'] + .06),
   # ylim = c(bbox['ymin'] - .1, bbox['ymax'] + .1)) +
-  scale_fill_binned(
-    type = "viridis",
-    name = "Sampling Events",
-    guide = hgc,
-    option = "E"
-  ) +
+  scale_fill_steps2(low = "grey75", high = "black",
+                    name = "Sampling Events",
+                    guide = hgc) +
+  # scale_fill_binned(
+  #   type = "viridis",
+  #   name = "Sampling Events",
+  #   guide = hgc,
+  #   option = "E"
+  # ) +
   theme_bw() +
   theme(legend.position = "top",
         panel.grid = element_blank())
@@ -4338,26 +4341,24 @@ targeted_trends_by_mpa %>%
 
 total_trend_plot <- targeted_trends %>%
   ggplot() +
-  geom_vline(aes(xintercept = 2003), linetype = "dotted", size = 1) +
+  geom_vline(aes(xintercept = 2003), linetype = "dotted", size = 1.1) +
   geom_segment(x = 2006, xend = 2003.25, y = 4, yend = 4, arrow =arrow(length = unit(0.2,"cm"))) +
-  geom_text(x = 2008.5, y = 4, label = "MPAs Implemented", nudge_x = 5, size = 4) +
-  geom_line(aes(year, scaled_mbd, color = targeted == 0, linetype = targeted == 0),
-            size = 1) +
   geom_ribbon(aes(
     year,
     ymin = scaled_mbd - 1.96,
     ymax = scaled_mbd + 1.96,
     fill = targeted == 0),
-    outline.type = "both",
-    color = "black",
-    alpha = 0.5) +
+    alpha = 0.25) +
   scale_color_manual(
-    values = c("black", "grey50"),
+    values = c("black", "grey40"),
     labels = c("Targeted", 'Non-Targeted'),
     name = ""
   ) +
[…]

### 2020-11-28 documents/ovando-et-al-pop-effects-of-mpas.Rmd
shortening and cleaning up

@@
 No-take Marine Protected Areas (MPAs), spatial regions of the ocean in which fishing is prohibited, have a long history in the management of marine resources.
-Traditional cultures in Oceania utilized - often temporary - MPAs as "fish banks" for times of need [@johannes1978].
 Modern MPAs were first established as marine analogs to the terrestrial protection of iconic landscapes [@iucn1976].
 Over time our goals and expectations for MPAs have evolved and become more ambitious.
-Recent international efforts to expand MPAs,such as the Biodiversity Beyond National Jurisdiction (BBNJ) Agreement and the Convention on Biological Diversity’s Post-2020 Global Biodiversity Framework, are based in part on the assumption that well-designed MPAs will achieve benefits both within, and outside, their borders [@gaines2010].
+Recent international efforts to expand MPAs, such as the Biodiversity Beyond National Jurisdiction (BBNJ) Agreement and the Convention on Biological Diversity’s Post-2020 Global Biodiversity Framework, are based in part on the assumption that well-designed MPAs will not only provide conservation benefits inside their borders, but will have broader conservation effects on unprotected areas surrounding the MPAs[@gaines2010].
 
 The empirical MPA literature has focused on assessing the ability of MPAs to provide conservation gains within their borders [@lester2009; @edgar2014; @gerber2005].
-As conservation benefits accrue inside MPAs, MPAs can affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
-These dynamics result in MPAs affecting not only the organisms within their borders, but the broader population of affected species connected by movement and reproduction, creating both local (e.g.
-inside and outside MPA effects) and population-level effects XX Owen help!
-I seriously don't know how to define this any clearer XX.
+However, as conservation benefits accrue inside MPAs, MPAs also affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
+Therefore, MPAs contribute to both local and population-level effects (i.e., inside-MPA effects and broader regional effects).
 Theory states that numerous factors influence how MPAs might affect fish populations.
@@
 <!-- What empirical evidence is there for these population-level effects of MPAs?  -->
-
 These theoretical results are generally based on modeling of closed populations, where some fraction of the connected population is protected inside of MPAs.
-In contrast to this population paradigm used in MPA simulations, conservation outcomes of MPAs are generally evaluated at local scales using "response ratios", commonly measured as the ratio of biomass densities (weight of organisms per unit area) of species inside relative to selected areas outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015].
+In contrast to this population paradigm used in MPA simulations, MPAs are often evaluated empirically at local scales using "response ratios", commonly measured as the ratio of biomass densities (weight of organisms per unit area) of species inside relative to selected control sites outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015].
 These studies found clear evidence that well enforced and sufficiently sized MPAs are associated with high response ratios.
@@
 
-While these studies have clearly demonstrated the ability of MPAs to create differences between fished and unfished areas, why does this large body of empirical work not serve as sufficient evidence for the population-level effects of MPAs?
-Spillover-based studies provide evidence that MPAs may have effects outside their borders, but they are not sufficient evidence for the population effect of MPAs, as on their own they lack any counterfactual.
-Response ratios have shortcomings as measures of population-level effects as well.
+While these studies have clearly demonstrated the ability of MPAs to create differences between local fished and unfished areas, this body of empirical work does not serve as sufficient evidence for the broader, population-level effects of MPAs.
+Spillover-based studies provide evidence that MPAs may affect areas outside their boundaries, but these studies often lack a counterfactual, and so cannot determine the magnitude of population-level effects.
+Response ratios have shortcomings as measures of population-level MPA effects as well.
 Control sites used in calculating response ratios are often selected based on habitat characteristics [@ferraro2018].
-However, export of adults or larvae from the MPA to these "control" site affects their status as controls, as does displacement of fishing effort from MPAs to control sites.
-These factors can act to bias estimators that require the assumption that the control sites are unaffected by the treatment, such as response ratios and BACI designs [@ferraro2018;@moffitt2013].
+However, export of adults or larvae from the MPA to these "control" sites affects their status as controls, as does displacement of fishing effort from MPAs to control sites.
+These factors can act to bias estimators that require the assumption that the control sites are unaffected by the treatment (a required assumption of response ratios and BACI designs [@ferraro2018;@moffitt2013]).
 In theory, control sites far enough away to negate both spillover of fish or larvae, and concentration of the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs is challenging.
 As variations of response ratios are our primary source of evidence for the conservation effects of MPAs, this means that our collective empirical understanding of the population-level impacts of MPAs is surprisingly limited.
-We show here that in contrast to the re
+
 <!-- Given the potential unreliability of response ratios for the task, how can we empirically measure the population-level effects of MPAs?
@@
 
-To fill this important gap in our understanding of protected area impacts, we present here an paired theoretical and empirical assessment demonstrating the challenges in assessing the population-level impacts of MPAs.The MPAs within the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) implemented a network of protected areas covering approximately 20% of the Islands' waters in 2003 [@kirlin2013].
+We present here an paired theoretical and empirical assessment demonstrating the challenges in assessing the population-level impacts of MPAs.
+In 2003, a network of MPAs was established in the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) covering approximately 20% of the Islands' waters [@kirlin2013].
 The network has been used as a model case study in protected area design around the world (@botsford2014 and references therein).
-We use data from the first 14 years of protection in a difference-in-difference mode [see @angrist2009 and references therein] to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of fin-fish species.
+We use data from the first 14 years of protection in a difference-in-difference model [see @angrist2009 and references therein] to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of fin-fish species.
 In contrast to clear differences in biomass densities observed inside and outside of well-protected MPAs both globally [@lester2009] and in the Channel Islands [@caselle2015] we are unable to detect a clear population effect from the Channel Islands MPAs.
@@
 The population-level effects of MPAs will depend on the social-ecological contexts in which they occur.
-We use our bio-economic simulation model to provide theoretical expectations for population-level effects of MPAs, which we then confront with the empirical results from our difference-in-difference model.
+We use a bio-economic simulation model to provide theoretical bounds on potential population-level effects of MPAs, which we then confront with the empirical results from our difference-in-difference model.
 
@@
 We pair these data with randomly selected values governing the characteristics of the simulation (Table.\@ref(tab:sim-table)).
-We highlight here key choices available to the model governing fishing pressure and MPA design.
-For a given simulation, after selecting a species the model randomly selects a fleet model and effort allocation strategy.
+Key choices available to the model include parameters governing fishing pressure and MPA design.
+For a given simulation, the model randomly selects a fleet model and fishing effort allocation strategy.
 The fleet model can either be constant-catch (the fleet exerts as much effort as needed to maintain a fixed amount of catch), constant-effort (the fleet maintains a constant amount of effort over time), or open-access (fishing effort of the fleet expands and contracts in response to available profits).
@@
 We then continue the simulations with the MPAs in place.
-Each simulation is paired with a simulation identical in every way except that MPAs are no longer put in place.
+Each simulation is paired with a simulation identical in every way except that MPAs are not implemented (i.e., a simulated control).
 We then calculate the effect of the MPAs on the population as the difference in relevant metrics in the simulation with MPAs relative to their values in the simulation without MPAs.
@@
 One set of simulations is specifically designed to reflect the dynamics of the Channel Islands.
-We only include species of the same genus as those targeted by fishing in the Channel Islands, restrict fishing pressure such that the simulated populations are moderately to lightly exploited (given the best available knowledge of stock status for the targeted species in the Channel Islands), and cap the MPA size at 20% of the population's range (a likely upper end to the effective size of the Channel Island MPAs relative to the fish populations in question, @rassweiler2012).
-For each of these Channel Island simulations, we calculated the biomass density response ratios for each simulation.
-We then binned both the simulated and empirical response ratios, and joined our simulated values to the empirical values based on matches of these bins.
-This provides a library of simulated population-level MPA effects that could have produced the types of empirical response ratios seen in the Channel Islands.
+We only include species of the same genus as those targeted by fishing in the Channel Islands, restrict fishing pressure such that the simulated populations are moderately to lightly exploited (given the best available knowledge of stock status for the targeted species in the Channel Islands), and cap the MPA size at 20% of the population's range [@rassweiler2012].
+For each of these Channel Island simulations, we calculated biomass density response ratios.
 
@@
 
-For readers more familiar with BACI methods, the difference-in-difference (DiD) regression amounts to estimating the pre-post MPA difference in the biomass densities of targeted species minus the same difference for non-targeted species in the Channel Islands.
+For readers more familiar with BACI methods, the difference-in-difference (DiD) regression amounts to estimating the pre-post MPA difference in the biomass densities of targeted species, minus the same difference for non-targeted species in the Channel Islands.
 
@@
 This difference-in-difference regression provides our estimate of the population-level conservation effect of the Channel Islands MPAs.
-We use a Bayesian regression since it allows us to interpret our estimated effects probabalistically [@mcelreath2020], e.g.
+We use a Bayesian regression since it allows us to interpret our estimated effects probabilistically [@mcelreath2020], e.g.
 stating the probability that conditional on the model, priors,and data the Channel Island MPAs increased or decreased biomass of targeted species.
@@
 Since the effects of MPAs will evolve over time, we use this model to estimate a series of population-level effects of the Channel Islands MPAs in three-year intervals from 2003 to 2017.
+
+To compare response ratios from the Channel Islands simulations to our empirical results, we organized simulated and empirical response ratios into bins.
+Because each simulation included measures of both response ratios and population-level effects, this provided a library of simulations (and their associated attributes) that could have produced the types of empirical response ratios measured in the Channel Islands.
 
@@
 In the case of the Channel Islands MPAs, control sites are often located within a few kilometers of an MPA, suggesting that they are susceptible to both local biological spillover and by concentration of fishing effort excluded from the MPAs.
-To illustrate this problem, we paired our simulated Channel Islands response ratios to the estimated posterior probability distributions of the empirical Channel Island response ratios by matching the patterns in the empirical response ratios.
+To illustrate this problem, we paired our simulated response ratios to the estimated posterior probability distributions of the empirical response ratios by matching the patterns in the empirical response ratios.
 The response ratio trends we observe in the data could plausibly be produced by a wide range of population-level MPA effects (Fig.\@ref(fig:rr-plot)).
@@
 
-Given the natural variability of marine ecosystems, how large of an effect would an MPA network have to have in order to allow a difference-in-difference strategy such as this to be a reliable measure of MPA effects?
+Given natural variability in ecosystems and uncertainty in measurements, MPA effects may be difficult to detect with precision.
 Our library of simulation results allows us to explore how MPA effects, and our ability to measure them, might vary with observation error and natural fish recruitment variation (including the possibility of recruitment regimes).
@@
 Even models fit to data generated from large effect sizes commonly mis-estimated the true MPA effect by 50% or more.
-Obtaining a mean absolute percent error (MAPE) of 25% or less across our simulated datasets required a population-level MPA effect of at least 30%.
+Obtaining a mean absolute percent error (MAPE) of 25% or less across our simulated datasets required a "true" population-level MPA effect of at least 30%.
 In the context of the Channel Islands, this finding suggests that we should not be surprised at our difficulty in precisely estimating the population-level effect of the MPAs.
@@
 Two of the most critical drivers of MPA effect size are the size of the MPA network and the degree of fishing pressure.
-Looking across these two variables, based on our simulations the MPA network must be large (25% or more of a species range) and the target species overfished (pre-MPA depletion greater than about 60%) to achieve an effect size with a likely MAPE of 25% or less.
+Looking across these two variables, based on our simulations the MPA network must be large (25% or more of a species' range) and the target species overfished (pre-MPA depletion greater than about 60%) to achieve an effect size with a likely MAPE of 25% or less.
 Although recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, most MPA networks for near-shore commercial fin-fish are likely to cover areas more in line with the Channel Islands (20%) or smaller.
@@
 Containing a carefully designed, well-enforced, and well-studied MPA network, the Channel Islands seems to be an ideal location to study the population-level effects of protected areas.
-The persistently high response ratios suggest that despite overall decreases in targeted biomass densities inside and outside MPAs relative to pre-MPA measurements, the MPAs may still be providing protection within their borders.
+Persistently high response ratios suggest that despite overall decreases in targeted biomass densities inside and outside MPAs relative to pre-MPA measurements (Fig.\@ref(fig:raw-density)), the MPAs may still be providing protection within their borders.
 But, these response ratios are not necessarily an indicator of broader population-level effects.
@@
 
-How can we explain this result?
+Uncertain fishing dynamics may be one factor contributing to a lack of strong MPA network effects.
 Much of the theoretical literature on MPAs assumes that larger reserves produce larger conservation gains [@white2011 and references therein].
@@
 
-Alternatively, fishers can pursue a  “constant-catch” strategy, where fishers have a catch objective and exert as much (or little) effort as needed to achieve that objective.
+Alternatively, under a “constant-catch” strategy, fishers have a catch objective and exert as much (or little) effort as needed to achieve that objective.
 Subsistence fisheries may use a constant-catch style policy over the short-term, as they seek to ensure that their food needs are met.
@@
 Fishers pursuing constant-catch in areas outside an MPA may have to fish harder to achieve the same catch from a smaller part of the population, causing a population conservation loss under `r percent(1 - positive_outcomes$pp[positive_outcomes$fleet_model == "constant-catch"], accuracy = 1)` of our constant-catch simulations.
-This potential negative interaction between constant-catch and MPAs is important risk to consider (as done so in @little2011b) especially as MPAs are increasingly implemented in quota-managed fisheries [@liu2018].
+This potential negative interaction between constant-catch and MPAs is an important risk to consider (as done so in @little2011b) especially as MPAs are increasingly implemented in quota-managed fisheries [@liu2018].
 
 While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (SI Fig.S1) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
-However, some of the most common targeted species in the Channel Islands, such as blue rockfish and copper rockfish, have seen dramatic increases in reported catches in the Santa Barbara region since the year 2010 (although other such as California sheephead have seen declines relative to pre-MPA catch level).
+However, some of the most common targeted species in the Channel Islands, such as blue and copper rockfish, have seen dramatic increases in reported catches in the Santa Barbara region since the year 2010 (although other such as California sheephead have seen declines relative to pre-MPA catch level).
 While we include catch histories in our regression, to the extent that these increases in catches are due to external market forces, the disappearance of clear positive population-level MPA effects estimated by our model could be due to exogenous increases in fishing pressure overwhelming any population-level increases caused by the MPA network.
@@
 The targeted group is made up mostly of fishes with cold-water affinities.
-We hypothesize that the recent evidence for a decline in densities of targeted species is primarily due to environmental conditions that disproportionately affected the targeted group.
+We hypothesize that the recent evidence for a decline in densities of targeted species is primarily due to altered environmental conditions that disproportionately affected the targeted group.
 The practical result of this is that in the presence of this marine heatwave the non-targeted species may no longer serve as an effective control for the evolution of biomass densities of targeted fin-fish in the absence of the MPAs, given the magnitude of the environmental shock relative to the size of the population-level MPA effect.
@@
 <!-- Once simulation models have been used to help design an MPA, how can users evaluate whether it is achieving their objectives?  -->
-@lester2009 found massively higher densities and biomass inside MPAs than outside [median biomass increases of  over 400%].
+
+<!-- @lester2009 found massively higher densities and biomass inside MPAs than outside [median biomass increases of  over 400%].
 But, when MPAs affect nearby control sites used in response ratios through biological spillover or concentration of fishing effort, it is entirely possible to for MPA to produce massive response ratios while simultaneously having minimal effects on the entire population partially protected by the MPAs (and *vice versa*).
+-->
 
-As @larsen2019 suggests, there are many potential alternatives for estimating the population effects of MPAs that better account for the challenges of causal inference (though that may be more data-intensive).
-We applied one such approach here, and yet were still unable to reach robust conclusions as to the effect of MPAs on the total biomass density of targeted fin-fish in the Channel Islands, due to the likely small size of the true effect relative to the influence of environmental variability.
-
-While this does not mean that all MPAs will face similar challenges in estimating their effects, our results in the relatively large, well-enforced, and rigorously studied (though also relatively lightly fished) Channel Islands Marine Protected Area network make clear that in many instances empirically detecting a clear effect of MPAs on total biomass densities of targeted fin-fish at the population level may not be possible.
+MPAs are an important part of the marine resource management toolbox.
+As the number and size of global MPA networks increase, we must set appropriate expectations for their outcomes on both local and regional scales.
+Our results for the relatively large, well-enforced, and rigorously studied (though also relatively lightly fished) Channel Islands Marine Protected Area network make clear that in many instances empirically detecting a clear effect of MPAs on total biomass densities of targeted fin-fish at the population level may not be possible.
 In the absence of clear empirical findings, simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can be tuned to focus on the species groups that have the highest chance of a detectable effect size over the early years of the reserve.
@@
 
-Stock assessment models that account for the population within the MPAs [@field2006a] may be able to answer the management relevant question of whether fishing mortality rates and biomass levels from systems containing MPAs are in-line with management objectives.However, such an approach does not necessarily shed light on whether the MPAs themselves caused the estimated state of the population, and of course are highly data intensive, potentially restricting our ability to provide stock-assessment based inference of MPA outcomes for broad arrays of targeted species.
+As @larsen2019 suggests, there are many potential alternatives to empirical response ratios for estimating the population effects of MPAs that better account for the challenges of causal inference (though that may be more data-intensive).
+We applied one such approach here, and yet were still unable to reach robust conclusions as to the effect of MPAs on the total biomass density of targeted fin-fish in the Channel Islands, due to the likely small size of the true effect relative to the influence of environmental variability.
+There are other promising statistical approaches to setting expectations for MPA effects, including using models fitted to local data to set population-level expectations [@nickols2019; @white2011].
+
+However, as the scientific community continues to develop appropriate measures of population-level MPA effects, we must also effectively communicate the challenges of estimating those effects.
+Lack of a clear population-level MPA effect should not necessarily be viewed as a failure of a conservation program, and likewise, a large positive effect based solely on response ratios should not be the sole justification for the relaxation of other conservation measures.
+Rather, results and subsequent management actions must be considered in the context of reasonable expectations given the size, age, and degree of enforcement of the MPAs in question, together with the ecological and economic dynamics of a given system.
+
+<!--Stock assessment models that account for the population within the MPAs [@field2006a] may be able to answer the management relevant question of whether fishing mortality rates and biomass levels from systems containing MPAs are in-line with management objectives.However, such an approach does not necessarily shed light on whether the MPAs themselves caused the estimated state of the population, and of course are highly data intensive, potentially restricting our ability to provide stock-assessment based inference of MPA outcomes for broad arrays of targeted species.
 @nickols2019 and @white2011 provide excellent examples of using models fit to local data to set expectations for the outcomes of MPAs.
@@
 In contrast, we use our simulation model to set expectations but use an empirical difference-in-difference approach to construct our counterfactual of likely outcomes of the Channel Islands in the absence of MPAs.
-Both approaches can help address the problem of assessing the population-level effects of MPAs.
+Both approaches can help address the problem of assessing the population-level effects of MPAs.-->
 
-Non-equilibrium analyses also help set expectations for effect sizes over time (@nickols2019;@kaplan2019;@ovando2016;Fig.\ref{fig:mpa-effect}).
+<!--Non-equilibrium analyses also help set expectations for effect sizes over time (@nickols2019;@kaplan2019;@ovando2016;Fig.\ref{fig:mpa-effect}).
 Educating communities about the challenges of estimating the effects of MPAs can help ensure that a lack of a clear effect is not necessarily viewed as a failure of the program, or large positive result based solely on response ratios as a clear sign of success and subsequent relaxation of other fishery management strategies.
 Rather, results and subsequent management actions must be considered in the context of reasonable expectations given the size, age, and degree of enforcement of the MPAs in question, together with the ecological and economic dynamics of a given system.
-While this paper has focused on the conservation outcomes of MPAs, future work must also address the challenge of predicting and estimating the fishery impacts of protected areas, an important question for which extremely little empirical evidence is available [@smith2006a].
+While this paper has focused on the conservation outcomes of MPAs, future work must also address the challenge of predicting and estimating the fishery impacts of protected areas, an important question for which extremely little empirical evidence is available [@smith2006a].-->
 
-MPAs are an important part of the marine resource management toolbox.
-As the number and size of global MPA networks increase, we must set appropriate expectations for their outcomes and design effective monitoring programs.
-While the history of MPA science has made important strides in helping us understand the dynamics of protected areas, the future of MPA science must directly tackle the challenge of evaluating the performance of these MPAs at the regional scale.
-This is particularly true if communities are depending on MPAs as their primary marine resource management tool.
+As advocacy for broad networks of MPAs is growing globally, future MPA science must directly tackle the challenge of evaluating the performance of MPAs at the regional scale.
 Commonly employed metrics such as response ratios may be applicable in some circumstances, but are vulnerable to inaccuracy or misuse as metrics of population-level effects.

### 2020-11-30 documents/ovando-et-al-pop-effects-of-mpas-si.Rmd
updating SI

@@
 library(here)
+library(modelsummary)
+library(sandwich)
 
@@
 
+We also ran a frequentist version of our results, as further demonstration that our results are not driven by our priors or Bayesian framework in general.
+For this version we applied a Newey-West correction to the standard errors to account for the clustering of the observations in space and time (Fig.\@ref(fig:freq-did)).
+
+```{r freq-did, fig.cap="Frequentist version of core difference-in-difference model, with standard errors adjusted by Newey-West correction to reflect clustering in space and time.
+Points are mean estimate, error bars 95% confidence interval."}
+
+did_dat  <- model_runs$did_fit[[which(model_runs$data_source == "pisco" & model_runs$var_names == "pisco_a" & model_runs$data_to_use == "all")]]$did_data
+
+did_reg <-
+  lm(
+    log(total_biomass_density) ~ targeted * year_bins + site_side - 1 + var_tex + var_tex_2
+    + var_surge + var_kelp + var_lag_catch + var_temp +
+      regional_temp_dev + regional_temp_dev_2,
+    weights = did_data$weight,
+    data = did_data
+  )
+
+a = vcovPL(did_reg, cluster = ~  region +year,lag = "NW1994")
+
+freq_did <- broom.mixed::tidy(did_reg) %>% 
+  mutate(se = sqrt(diag(a))) %>% 
+  mutate(effect = exp(estimate + se^2 / 2) - 1)
+
+freq_did %>% 
+  filter(str_detect(term, "targeted:")) %>% 
+  mutate(term = str_remove_all(term, "targeted:year_bins")) %>% 
+  ggplot(aes(term, estimate)) + 
+  geom_hline(aes(yintercept = 0)) +
+  geom_pointrange(aes(ymin = estimate - 1.96 * se, ymax = estimate + 1.96 * se)) + 
+    scale_y_continuous(name = "MPA Effect") +
+  scale_x_discrete(name = element_blank())
+
+```
+
 ### Synthetic controls

### 2020-11-30 documents/referee-replies.Rmd

@@
 XX Owen, thoughts on tone here?XX
+XX Looks great to me!--O XX
 
@@
 Suppose then that all the fishing effort that used to operate inside the MPA moves and now operates outside the MPA, doubling the amount of fishing pressure outside.
-A BACI design that assumes that the control site is a measure of what would have happened in the absence of an MPA will conclude that the MPAs had a large positive effect, since biomass inside the MPA will decline post-MPA, while biomass inside the MPA will increase.
-Convsersely, consider that same MPA but without effort displacement but with very strong larval and adult spillover, which results in the control site benefit greatly from the MPA.
+A BACI design that assumes that the control site is a measure of what would have happened in the absence of an MPA will conclude that the MPAs had a large positive effect, since biomass outside the MPA will decline post-MPA, while biomass inside the MPA will increase.
+Conversely, consider that same MPA but without effort displacement but with very strong larval and adult spillover, which results in the control site benefit greatly from the MPA.
 A BACI style design would conclude that the MPA had little to no effect in this case, since biomass levels would be similar inside and outside the MPA.
@@
 XX Owen check if you think this is clearer now. basically most fo the intro XX
+XX This explanation is good and I have made minor edits in the intro XX
 
@@
 XX Owen, too snarky? XX
+XX Thisf response is fine, and I will check the brief description in the main text as well XX
 

### 2020-11-30 documents/ovando-et-al-pop-effects-of-mpas.Rmd
pre-chop-back-commit

@@
 Over time our goals and expectations for MPAs have evolved and become more ambitious.
-Recent international efforts to expand MPAs, such as the Biodiversity Beyond National Jurisdiction (BBNJ) Agreement and the Convention on Biological Diversity’s Post-2020 Global Biodiversity Framework, are based in part on the assumption that well-designed MPAs will not only provide conservation benefits inside their borders, but will have broader conservation effects on unprotected areas surrounding the MPAs[@gaines2010].
+Recent international efforts to expand MPAs, such as the Biodiversity Beyond National Jurisdiction (BBNJ) Agreement and the Convention on Biological Diversity’s Post-2020 Global Biodiversity Framework, are based in part on the assumption that well-designed MPAs will not only provide conservation benefits inside their borders, but also have broader conservation effects on unprotected areas surrounding the MPAs [@gaines2010].
 
@@
 Therefore, MPAs contribute to both local and population-level effects (i.e., inside-MPA effects and broader regional effects).
-Theory states that numerous factors influence how MPAs might affect fish populations.
-These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], the strength, timing, and location of density dependence [@burgess2014], the degree of enforcement [@edgar2014], the level of fishing pressure, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006a;@walters2000a;@white2011,;@white2010c;@ovando2016;@moffitt2013].
-The location and spacing of the MPAs within a network can also influence their population-level impacts through synergistic effects, for example acting as stepping stones for larvae [@gaines2010; @costello2010a].
+Numerous factors influence how MPAs affect fish populations.
+These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], the strength, timing, and location of density dependence [@burgess2014], the design of the network [@gaines2010; @costello2010a],the degree of enforcement [@edgar2014], the level of fishing pressure, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006a;@walters2000a;@white2011;@white2010c;@ovando2016;@moffitt2013].
 
 <!-- What empirical evidence is there for these population-level effects of MPAs?  -->
-These theoretical results are generally based on modeling of closed populations, where some fraction of the connected population is protected inside of MPAs.
-In contrast to this population paradigm used in MPA simulations, MPAs are often evaluated empirically at local scales using "response ratios", commonly measured as the ratio of biomass densities (weight of organisms per unit area) of species inside relative to selected control sites outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015].
+These theoretical results are generally based on modeling of closed populations, where some fraction of a population is protected inside of MPAs.
+In contrast to this population paradigm used in MPA simulations, MPAs are often evaluated empirically at local scales using spatial "response ratios", commonly measured as the ratio of biomass densities (weight of organisms per unit area) of species inside relative to selected control sites outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015].
 These studies found clear evidence that well enforced and sufficiently sized MPAs are associated with high response ratios.
 Several studies have also documented empirical evidence for the existence of adult or larval fish spillover affecting fish abundance [@halpern2009;@kay2012; @stobart2009; @mcclanahan2000; @russ1996; @thompson2017;@baetscher2019;@pelc2009].
-Where response ratios are available before and after MPA implementation, before-after-control-impact (BACI) style studies have shown similarly clear and positive results [@thiault2019a].
+Where response ratios are available before and after MPA implementation, spatial before-after-control-impact (BACI) style studies have shown similarly clear and positive results [@thiault2019a].
 
-While these studies have clearly demonstrated the ability of MPAs to create differences between local fished and unfished areas, this body of empirical work does not serve as sufficient evidence for the broader, population-level effects of MPAs.
-Spillover-based studies provide evidence that MPAs may affect areas outside their boundaries, but these studies often lack a counterfactual, and so cannot determine the magnitude of population-level effects.
-Response ratios have shortcomings as measures of population-level MPA effects as well.
-Control sites used in calculating response ratios are often selected based on habitat characteristics [@ferraro2018].
-However, export of adults or larvae from the MPA to these "control" sites affects their status as controls, as does displacement of fishing effort from MPAs to control sites.
-These factors can act to bias estimators that require the assumption that the control sites are unaffected by the treatment (a required assumption of response ratios and BACI designs [@ferraro2018;@moffitt2013]).
-In theory, control sites far enough away to negate both spillover of fish or larvae, and concentration of the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs is challenging.
-As variations of response ratios are our primary source of evidence for the conservation effects of MPAs, this means that our collective empirical understanding of the population-level impacts of MPAs is surprisingly limited.
+While these studies have clearly demonstrated the ability of MPAs to create differences between local fished and unfished areas, this body of empirical work does not necessarily serve as sufficient evidence for the population-level effects of MPAs.
+This is because these studies rely on assuming that selected control sites serve as a measure of what would have happened in the absence of MPAs.
+Control sites used in calculating response ratios are often selected based on habitat characteristics to justify their use as counterfactuals [@ferraro2018].
+However, the very spillover effects MPAs are increasingly hoped to produce can negate the ability of spatial response ratio / BACI designs to accurately estimate the effects of MPAs.
+Export of adults or larvae from MPAs to "control" sites can affect their status as controls, as does displacement of fishing effort from MPAs to control sites.
+These factors can bias estimators of population-level effects that require the assumption that the control sites are unaffected by the treatment (a required assumption of response ratios and BACI designs [@ferraro2018;@moffitt2013]).
+Control sites sufficiently far from MPAs to negate both spillover of fish or larvae and concentration of the fishing fleet could be selected, but finding suitably distant sites that are also appropriate proxies for the ecological and economic context of the MPAs is challenging.
+As variations of spatial response ratios and BACI studies are our primary source of evidence for the conservation effects of MPAs, this means that our empirical understanding of the population-level impacts of MPAs is surprisingly limited.
 
@@
 We use data from the first 14 years of protection in a difference-in-difference model [see @angrist2009 and references therein] to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of fin-fish species.
+Rather than relying on spatial controls as is typically done in MPA studies, we use groups of species targeted and not-targeted by fishing pressure as our treatment and control groups.
 In contrast to clear differences in biomass densities observed inside and outside of well-protected MPAs both globally [@lester2009] and in the Channel Islands [@caselle2015] we are unable to detect a clear population effect from the Channel Islands MPAs.
@@
 We only include species of the same genus as those targeted by fishing in the Channel Islands, restrict fishing pressure such that the simulated populations are moderately to lightly exploited (given the best available knowledge of stock status for the targeted species in the Channel Islands), and cap the MPA size at 20% of the population's range [@rassweiler2012].
-For each of these Channel Island simulations, we calculated biomass density response ratios.
+For these Channel Island-esque simulations, we calculated 1) the population-level response for each species (the 'true' population-wide difference in biomass between the simulations with and without the MPAs) and 2) the measured response ratio of biomass densities inside and outside the simulated MPAs.
+To compare response ratios from the Channel Islands simulations to our empirical results, we organized simulated and empirical response ratios into bins.
+Because each simulation included measures of both response ratios and population-level effects, this provided a library of simulations (and their associated attributes) that could have produced the types of empirical response ratios measured in the Channel Islands.
 
@@
 
-For readers more familiar with BACI methods, the difference-in-difference (DiD) regression amounts to estimating the pre-post MPA difference in the biomass densities of targeted species, minus the same difference for non-targeted species in the Channel Islands.
+The difference-in-difference (DiD) regression amounts to estimating the pre-post MPA difference in the biomass densities of targeted species, minus the same difference for non-targeted species in the Channel Islands.
 
@@
 
-This difference-in-difference regression provides our estimate of the population-level conservation effect of the Channel Islands MPAs.
+This difference-in-difference regression provides our estimate of the population-level effect of the Channel Islands MPAs.
 We use a Bayesian regression since it allows us to interpret our estimated effects probabilistically [@mcelreath2020], e.g.
@@
 Since the effects of MPAs will evolve over time, we use this model to estimate a series of population-level effects of the Channel Islands MPAs in three-year intervals from 2003 to 2017.
-
-To compare response ratios from the Channel Islands simulations to our empirical results, we organized simulated and empirical response ratios into bins.
-Because each simulation included measures of both response ratios and population-level effects, this provided a library of simulations (and their associated attributes) that could have produced the types of empirical response ratios measured in the Channel Islands.
 
@@
 In the case of the Channel Islands MPAs, control sites are often located within a few kilometers of an MPA, suggesting that they are susceptible to both local biological spillover and by concentration of fishing effort excluded from the MPAs.
-To illustrate this problem, we paired our simulated response ratios to the estimated posterior probability distributions of the empirical response ratios by matching the patterns in the empirical response ratios.
-The response ratio trends we observe in the data could plausibly be produced by a wide range of population-level MPA effects (Fig.\@ref(fig:rr-plot)).
+Our simulation results show that the response ratio trends we observe in the data could plausibly be produced by a wide range of population-level MPA effects (Fig.\@ref(fig:rr-plot)).
 Response ratios well over one were associated with "true" population-level MPA effects generally less than 25%, and, importantly, many simulations produced large response ratios but population-level MPA effects close to 0%.
@@
 
-Our empirical difference-in-difference regression provides a different perspective on population-wide MPA effects.
+Our difference-in-difference regression provides a different perspective on population-wide MPA effects.
 Over the first three years of implementation (2003-2006), the effects of the MPAs are unclear, with support for a small negative effect to a substantially positive effect, with much higher probability of a small positive effect (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`,
@@
 Using a Bayesian difference-in-difference regression on these simulated data, we estimated the percent error between the posterior probability distribution of the estimated MPA effect from the regression and the true simulated MPA effect.
-
 The model struggled severely when MPA effect sizes were less than 25% and the model was faced with both observation and process errors (Fig.\@ref(fig:val-plot)).
@@
 But, these response ratios are not necessarily an indicator of broader population-level effects.
-The difference-in-difference strategy presents an alternative identification strategy.
-It is not without its own strict caveats, but provides some potential improvements over response-ratios as a means of estimating population-level effects.
+The targeted vs.
+non-targeted difference-in-difference strategy presents an alternative means of estimating the population-level effects of MPAs.
+While having its own strict caveats, this strategy provides some potential improvements over spatial response-ratios as a means of estimating population-level effects.
 While we estimate an uncertain but overall positive effect of the MPA network in its first few years, we are unable to detect a robust positive signal from 2012-2017.
-After 14 years of MPA protection we are left without a clear picture of the effect of the Channel Island MPA network on biomass densities of targeted fin-fish species from either response ratios or the difference-in-difference model.
+After 14 years of MPA protection we are left without a clear picture of the population-level effect of the Channel Island MPA network on biomass densities of targeted fin-fish species.
 
-Uncertain fishing dynamics may be one factor contributing to a lack of strong MPA network effects.
+Fishing dynamics may be one factor contributing to a lack of strong MPA network effects.
 Much of the theoretical literature on MPAs assumes that larger reserves produce larger conservation gains [@white2011 and references therein].
-However, these models generally simulate fleet dynamics through fishing mortality rates, the proportion of total mortality experienced by a population attributable to fishing pressure [e.g.
+However, these models generally simulate fleet dynamics through fishing mortality rates; the proportion of total mortality experienced by a population attributable to fishing pressure [e.g.
 @halpern2004].
@@
 
-While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (SI Fig.S1) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
+While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (SI Fig.S1 XX ) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
 However, some of the most common targeted species in the Channel Islands, such as blue and copper rockfish, have seen dramatic increases in reported catches in the Santa Barbara region since the year 2010 (although other such as California sheephead have seen declines relative to pre-MPA catch level).
@@
 
-MPAs are an important part of the marine resource management toolbox.
+<!-- MPAs are an important part of the marine resource management toolbox.
 As the number and size of global MPA networks increase, we must set appropriate expectations for their outcomes on both local and regional scales.
@@
 Whenever possible monitoring programs should be implemented prior to MPA implementation to provide a pre-treatment benchmark.
+-->
 

### 2020-12-01 documents/referee-replies.Rmd
Edits almost done

@@
 ---
-title: "Referee Replies"
-author: "Dan Ovando"
-date: "11/11/2020"
+title: Assessing the Population-Level Conservation Effects of Marine Protected Areas
+subtitle: "Reviewer Replies"
 output:
-  pdf_document: default
-  html_document: default
+  bookdown::pdf_document2: default
 ---
@@
 
+We thank the editors and referees for their helpful comments.
+The most pressing requests were in our mind to a) specify the specific gap in the literature filled by our work and b) clarify the methods with regards to both our simulation model and the difference-in-difference equations.
+In order to address these concerts, we have
+
+1. Clarified the specific gap that our paper fills in the literature, and how the empirical approach here builds off of methods such as before-after-control-impact (BACI) designs
+
+2. Included a clearer description of our simulation and estimation methods, and reorganized the Supporting Information to allow readers to more clearer find detailed equations.
+
+While we have endeavored to include every request made by the referees, we do note that the manuscript is currently at the word limit of 6000 words (including literature cited) stipulated by Conservation Biology.
+As such we have not included as in-depth responses in the text of the manuscript as we might otherwise, but we hope our brief responses address the issues identified by the referees.
+However, we understand that it is our responsibility to ensure that we address all comments are sufficiently addressed within the word limits, so we will of course work to accommodate any outstanding issues.
+
+We provide detailed responses to each of the editors' and referees' comments below. Editor's or referee's comments are *italicized*, and our responses are in plain text.
+
 # Reply to REGIONAL EDITOR'S COMMENTS
 
-Comments to the Author:
+*Comments to the Author:*
 
-Thank you for submitting your manuscript ‘Assessing the Population-level Conservation Effects of Marine Protected Areas’ to ‘Conservation Biology’.
+*Thank you for submitting your manuscript ‘Assessing the Population-level Conservation Effects of Marine Protected Areas’ to ‘Conservation Biology’.*
 
-The Handling Editor obtained three thoughtful and comprehensive reviews, which variously recommended ‘Minor Revision’ , ‘Moderate Revision’ and ‘Major Revision’.
-The Handling Editor recommended ‘Moderate Revision’, a recommendation I support and recommend to the Editor-in Chief.
+*The Handling Editor obtained three thoughtful and comprehensive reviews, which variously recommended ‘Minor Revision’ , ‘Moderate Revision’ and ‘Major Revision’.
+The Handling Editor recommended ‘Moderate Revision’, a recommendation I support and recommend to the Editor-in Chief.*
 
-Identifying the conditions under which MPAs can provide conservation benefits beyond their borders is an important area of research that should be of wide interest to the generic readership of Conservation Biology.
-The Channel Islands National Marine Sanctuary is an appropriate and data-rich case study.
+*Identifying the conditions under which MPAs can provide conservation benefits beyond their borders is an important area of research that should be of wide interest to the generic readership of Conservation Biology.
+The Channel Islands National Marine Sanctuary is an appropriate and data-rich case study.*
 
@@
 In particular, we have sought to clarify the general contributions of this paper in the introduction.
-XX.
 We thank you for your consideration and look forward to hearing your response.
 
-I look forward to a revised manuscript, which will be sent out for another round of review.
+# Reply to HANDLING EDITOR'S COMMENTS
 
-I hope you will continue to consider ‘Conservation Biology’ as your journal of choice for relevant manuscripts that are of potential interest to a wide international readership.
+*Comments to the Author:*
 
-Sincerely
-Helene Marsh
-Regional Editor
-
-# Reply to HANDLING EDITOR'S COMMENTS
-Comments to the Author:
-
-The Authors assess under which conditions MPAs can provide conservation benefits beyond MPA-borders.
+*The Authors assess under which conditions MPAs can provide conservation benefits beyond MPA-borders.
 This is an important area of research – helping to guide expectations about how well MPAs protect species within and outside their borders, and whether these effects will be detectable.
 The Authors demonstrate their approach in The Channel Islands National Marine Sanctuary, a well-studied system with available data.
-The Authors are also to be commended for their efforts to provide fully open access code.
+The Authors are also to be commended for their efforts to provide fully open access code.*
 
-The Authors show that conservation benefits are very difficult to detect empirically—even when population-level benefits are large.
-MPAs are often criticised for their inability to provide spillover benefits to fishers who are negatively impacted by area closures.
-The current results provide insight into why evidence of spillover benefits are often lacking.
+*The Authors show that conservation benefits are very difficult to detect empirically—even when population-level benefits are large.
+MPAs are often criticized for their inability to provide spillover benefits to fishers who are negatively impacted by area closures.
+The current results provide insight into why evidence of spillover benefits are often lacking.*
 
-The three reviewers provided a comprehensive assessment of the manuscript.
-Their feedback was generally positive, but emphasized the need for *greater clarity in the presentation of the methods, and structuring of the material between results, introduction and discussion*.
+*The three reviewers provided a comprehensive assessment of the manuscript.
+Their feedback was generally positive, but emphasized the need for greater clarity in the presentation of the methods, and structuring of the material between results, introduction and discussion.
 I must agree with this assessment, although I can appreciate that the Authors have had a difficult task clearly explaining an involved methodology with many results that add value to the interpretation and applicability of the findings.
-*The main elements that the Reviewers felt could be better explained (or moved from the supplementary to the main text) include details of the Bayesian Difference in Difference approach, formulas for the simulation model, and details regarding the model scenarios.
+The main elements that the Reviewers felt could be better explained (or moved from the supplementary to the main text) include details of the Bayesian Difference in Difference approach, formulas for the simulation model, and details regarding the model scenarios.
 I also felt that key results from the study were presented in the Discussion—namely results from simulation studies describing when MPA effect sizes were likely to be detectable (paragraphs beginning Line 220 and Line 228).*
@@
 We have addressed each of their comments in the text, and outlined our responses below.
-In particular, we have included more details about the simulation model and Bayesian Difference in Difference in the methods, ensured that all results are in the results section, and made clearer the specific objectives of our study in the introduction.
-
-don't sell the model short, there's a lot in here that we just don't have space to explain in detail in the main text
+In particular, we have included more details about the simulation model and Bayesian Difference-in-Difference in the methods, ensured that all results are in the results section, and made clearer the specific objectives of our study in the introduction.
 
@@
 
-I have read with interest the manuscript entitled: “Assessing the Population-level Conservation Effects of Marine Protected Areas”.
+*I have read with interest the manuscript entitled: “Assessing the Population-level Conservation Effects of Marine Protected Areas”.
 This manuscript compiles very interesting approaches and outcomes.
 However, I believe the manuscript requires substantial modifications to make it clearer.
-Here are some suggestions.
+Here are some suggestions.*
+
+We thank Reviewer #1 for their through comments and helpful suggestions.
+The reviewer's main concern appears to be that we 1) did not clearly articulate the gap in the literature that our paper fills and b) a lack of clarity in the details of the difference-in-difference model, its relationship to BACI designs, and how our methods attempt to resolve some of the challenges with spatial identification strategies discussed in our paper.
+
+To address these concerns we
+
+- Clarified in the introduction that the gap filled by our method is not in the statistical method per say (a difference-in-difference is essentially a BACI by another name), but in the use of targeted and non-targeted species as our treatment and control group, as opposed to inside and outside MPAs.
+See line XX
+
+- Expanded the description of our methods on line XX
 
@@
 - *Please indicate and describe briefly the method you used in your article*
-  - We have clarrified in the abstract that we use a Bayesian difference-in-difference model
+  - We have clarified in the abstract that we use a Bayesian difference-in-difference model
 
@@
 
-XX Throughout introduction XX
-
 *- A clearer explanation of the gap in the current literature that you are willing to fill is needed*
 
-We have provided a clearer statement of the gap in the current literature that our paper fills on line XX, and line XX
+We have clarified that the reason that the large body of inside-outside MPA style studies does not necessarily serve as a measure of population-level effects is due to the potential of the MPA itself to affect the control sites.
+This is a problem that BACI style analyses are not designed to address.
 
-XX Paragraph in intro starting with "While these studies have clearly demonstrated the ability" XX
+On line XX we state
+
+"XX"
+
+We synthesize the gap filled by our paper on line XX
+
+"XX"
 
@@
 
-Thank you for pointing this out, we have clarified the paragraph in question and provided greater detail on why exactly this existing body of evidence does not neccesarily serve as a population-level measure of MPA effects (Line XX).
+Thank you for pointing this out, we have clarified the paragraph in question and provided greater detail on why exactly this existing body of evidence does not necessarily serve as a population-level measure of MPA effects (Line XX).
 
@@
 BACI and BACIPS are both powerful tools for program evaluation in natural systems, and we now mention BACI explicitly in the text.
-However, both response ratios and BACI/BACIPS depend on the control being unaffacted by the treatment.
+However, both response ratios and BACI/BACIPS depend on the control being unaffected by the treatment.
 If the treatment affects the control, e.g.
@@
 We apologize but we are not clear what this comment is in reference to.
-We do not use log-ratios in our model, except in the sense that we use a Bayesian Gamma GLMER with a log link function to run our difference-in-diffecence (basically BACI) , so the estimated effect sizes are in log scales, roughly corresponding to percentage-level effects.
+We do not use log-ratios in our model, except in the sense that we use a Bayesian Gamma GLMER with a log link function to run our difference-in-difference (basically BACI) , so the estimated effect sizes are in log scales, roughly corresponding to percentage-level effects.
 
@@
 Since this is a Bayesian regression, we provide credible intervals rather than confidence intervals [@gelman2014a].
-Similarly, there is no equation per-say for credible intervals, as they are essentially just an exercise in adding probabilities.
+Similarly, there is no equation per-say for credible intervals, as they are essentially just an exercise in manipulating probabilities.
 Our regression provides a posterior probability distribution parameters, and we then select the values of the coefficients within a desired range.
@@
 Specifically, it allows us to answer questions such as "what is the probability that the Channel Island MPAs had a greater than 0% effect on biomass densities of targeted finfish".
-For comparison, we have added a frequentist mixed effects version of our model (Fig.SXX), which produces qualitatively the same results XX.
+For comparison, we have added a frequentist mixed effects version of our model (Supplemental Information Fig.S26), which produces qualitatively the same results.
 
@@
 
-We did the simulation study purely to ground expectations, and then compared our simulation to the empirical estimation model.
-
 We take the referees comment to mean our choice of using both a difference-in-difference model and a simulation model?
@@
 Our main methods difference then is in choosing to use targeted and non-targeted species as our treatment and control group, rather than inside and outside MPAs, since we believe there is strong a strong possibility that the MPAs affect the "control" sites outside their borders in the Channel Islands.
-In addition, because our goal is to estimate the population level effect of the MPAs, and for all of the species evalauted here the population stretches well beyond the borders of the MPAs, we need a control group that is ...
+In addition, because our goal is to estimate the population level effect of the MPAs, and for all of the species evaluated here the population stretches well beyond the borders of the MPAs, we need a control group that reflects this dynamics.
+We feel that the targeted vs non-targeted identification strategy employed here better accomplishes this than spatial controls within the Channel Islands.
 
-Given the nature of the study, other econometric techniques such as regression discontinuity or instrumental variables are not viable (we do not expect a discontinuous effect given the population dynamcis in play, and we have no valid instrument for the implementation of MPAs that is exogenous to the outcome).
-Synthetic controls [@abadie] do present another potential solution to our problem, which we use in SM.XX, and find no clear effect.
+Given the nature of the study, other econometric techniques such as regression discontinuity or instrumental variables are not viable (we do not expect a discontinuous effect given the population dynamics in play, and we have no valid instrument for the implementation of MPAs that is exogenous to the outcome).
+Synthetic controls [@abadie] do present another potential solution to our problem, which we use in the Supplementary Information Section 5.1 and find no clear effect, similar to our difference-in-difference based approach.
 
@@
 We apologize that our results were unclear and have sought to clarify our findings.
-We are however unclear where the referees concern about "size and enforcement" come in.
+We are however unclear about the referees concern about "size and enforcement".
 We do not claim to have tested for enforcement in our results.
@@
 We do not mention enforcement in our results that we can find.
-We do test for the effect of MPA size, which we make clearer in our methods description, on line XX.
+We do simulate the effect of MPA size, which we make clearer in our methods description, on line XX.
 
@@
 Moving this paragraph to the introduction would require explaining our simulation modeling in the introduction as well, which we feel is not ideal.
-
-XX Owen, thoughts on tone here?XX
-XX Looks great to me!--O XX
 
@@
 
-TRemoved, thank you.
+Removed, thank you.
 
@@
 
-Thank you for pointing this out.
-We have clarified that by "network effect" we mean the idea that a network of MPAs can produce impacts greater than the sum of their individual parts through synergistic action.
-Line XX.
-
-XX The location and spacing of the MPAs within ... XX
+To make room for other material in the paper we have removed this specific discussion
 
@@
 
-Our citations are automatically formatted according to the citation style langauge file provided by Conservation Biology, so we have left the order as dictated by the software, unless the editors ask otherwise.
+Our citations are automatically formatted according to the citation style language file provided by Conservation Biology, so we have left the order as dictated by the software, unless the editors ask otherwise.
 
@@
 A BACI style design would conclude that the MPA had little to no effect in this case, since biomass levels would be similar inside and outside the MPA.
-
-XX Owen check if you think this is clearer now. basically most fo the intro XX
-XX This explanation is good and I have made minor edits in the intro XX
 
@@
 The posterior probability distribution is what allows us to interpret the say 95% credible interval of a parameter as "there is a 95% probability that conditional on the data, model, and priors the effect of Z was between X and Y".
-In contrast a frequentist 95% confidence interval tells us that there is a 95% change that if were to resample the data infinite times and refit the model, there is a 95% chance that the true parameter value would fall within that range.
+In contrast a frequentist 95% confidence interval tells us that there is a 95% change that if were to resample the data infinite times and refit the model each time, there is a 95% chance that the true parameter value would fall within that range.
 
@@
 Given the space constraints of this paper, we do not feel that we are able to include a broader discussion of Bayesian inference, but have included a citation to a standard textbook on the topic in our introduction of our methods on line XX.
-
-XX Owen, too snarky? XX
-XX Thisf response is fine, and I will check the brief description in the main text as well XX
 
@@
 
-# Reply to Reviewer 2
+## Reply to Reviewer 2
 
@@
 
-My main concerns on reading the paper are that the methods and results are not clearly enough described.
+*My main concerns on reading the paper are that the methods and results are not clearly enough described.*
+
+Thank you for pointing this out, we have reorganized and clarified our methods and results.
 
@@
 Thank you for pointing out the lack of sufficient detail on the equations for the simulation model.
-We have restructured the SI so that the simulation model equations are front  and center (SI Line XX), and clarified our explanations of the equations.
+We have restructured the SI so that the simulation model equations are front  and center (SI Section 2), and clarified our explanations of the equations.
 We have also expanded our description of the simulation model in the text.
 We hope this clarifies the structure behind the simulation model.
-
-XX Simulation model part of the methods XX
 
@@
 We have however included a clearer summary of the simulation in the methods section, highlighting among other key attributes the ways in which we simulate spatial patterns of fishing intensity, which as Referee #2 has noted is a critical variable to consider.
-
-XX Simulation model part of the methods XX
 
@@
 We then matched these simulated response ratios to the empirical response ratios observed in the Channel Islands.
-So, for example, if we observe a response ratio of 2 after 4 years of protection in the Channel Islands, we matched that results to all simulated results that also had a response ratio of 2 after 4 yearse of MPA protection.
+So, for example, if we observe a response ratio of 2 after 4 years of protection in the Channel Islands, we matched that results to all simulated results that also had a response ratio of 2 after 4 years of MPA protection.
 
@@
 
-Declarative sentance for what this figure is out. Results of ... Insight first, then terms at the end. warmer colors indicate larger effect XX
+Declarative sentence for what this figure is out. Results of ... Insight first, then terms at the end. warmer colors indicate larger effect XX
 
@@
 Minor - some of the statements would benefit from being more nuanced.
-Line 187 states that fishers have to work harder to maintin constant catch when MPAs are in place.
+Line 187 states that fishers have to work harder to maintain constant catch when MPAs are in place.
 Well, potentially but not if they benefit from significant biological spillover.*
@@
 
-# Reply to Reviewer 3
+## Reply to Reviewer 3
 
@@
 The introduction clearly lays out the challenge: we know a lot about how fish biomass increases inside MPAs, but very little about whether this affects seascape scale biomass overall.
-It uses a novel combination of simulation and empirical modelling to ask if the Channel Islands MPA system has seascape wide effects on fish populations, and under what conditions that might occur.*
+It uses a novel combination of simulation and empirical modeling to ask if the Channel Islands MPA system has seascape wide effects on fish populations, and under what conditions that might occur.*
 
@@
 
-*The simulation model represents on its own a very complete synthesis of the fish biomass MPA modelling literature, and includes most of the processes and their variants, that have been shown over the past 20 years to affect MPA outcomes.*
+*The simulation model represents on its own a very complete synthesis of the fish biomass MPA modeling literature, and includes most of the processes and their variants, that have been shown over the past 20 years to affect MPA outcomes.*
 
@@
 At the risk of further expanding the already extensive supplementary materials, we have also included a version with a binwidth of 1 (i.e.
-no binning) to make clear that our results are not sensitive to the choice of binwidth.
+no binning) to make clear that our results are not sensitive to the choice of binwidth (Fig.S23)
 
@@
 Adding in a term representing the presence of an active MPA at a site would effectively mean that the our main coefficient of interest, the interaction of the targeted dummy variable with the year blocks, would represent the mean effect of the MPAs on mean total biomass density of targeted finfish controlling for any effect that those MPAs had inside the MPAs.
-
-XX clarify this messXX
 
@@
 *Introduction reads nicely.
-One comment: it might be worth stating that modelling studies are generally predicting population level effects, but empirical studies are measuring local effects.
-So another great example of a modelling-empirical disconnect.
-You’ve set a high expectation in the reader’s mind that you are going to fill this gap, which is then met through the novel combination of simulation and statistical modelling.*
+One comment: it might be worth stating that modeling studies are generally predicting population level effects, but empirical studies are measuring local effects.
+So another great example of a modeling-empirical disconnect.
+You’ve set a high expectation in the reader’s mind that you are going to fill this gap, which is then met through the novel combination of simulation and statistical modeling.*
 
@@
 2017), which we have incorporated on line XX.
-We have also cited @fulton2015 in order to link our mork more explicitly to the broader literature on MPA simulation model.
+We have also cited @fulton2015 in order to link our more more explicitly to the broader literature on MPA simulation model.
 
@@
 We feel that this would be a more important concern if we had estimated a for example positive but small effect of the MPAs, where perhaps slightly broader error bars might start to span zero.
-Given that one of our main results is that we are unable to detect a clear signal, we feel that the possible risk of underestimating the error distributions is outweighted by the potential gains of helping control for omitted variable bias.
+Given that one of our main results is that we are unable to detect a clear signal, we feel that the possible risk of underestimating the error distributions is outweighed by the potential gains of helping control for omitted variable bias.
+
+We have included a frequentist version of our model in the Supporting Information Fig.S26, which uses the Newey-West correction to account for clustering of the observations in space and time, and the results are qualitatively identical to our main Bayesian difference-in-difference results.
 
@@
 We have cleaned up Fig.2 substantially, and hope that the message that the raw data a) support the parallel trends assumption between the targeted and non-targeted populations pre-MPA and b) that the raw data support the statistical finding of no clear MPA effect.
-We have rewritten the caption to make Figure 6 clearer: we acknowledge that it is a complicated figure but we feel that it is important to convey the range of possible outcomes produced by our simulation model.
 
@@
 We worry that given the large amount going on in Figure 4 already adding another shaded area might be confusing.
+The "pre" MPA period is effectively reflected by the intercept term and its relevant interactions.
 
@@
 
+Given the word constraints, we have not included this discussion in the main text. However, we are happy to find a way to do so if the reviewer feels it needs to be included.
+
 *271: It would be good to read more comparison with the Nickols study, given the similarities to this study.
@@
 
-We briefly (due to word limit constraints) expanded our discussion of the synergies between our work and the Nickols study. Broadlyt we feel that the Nickols study further makes the case for...
+We briefly (due to word limit constraints) expanded our discussion of the synergies between our work and the Nickols study, in particular how the sorts of locally-fit population models used there can be used as a synthetic counterfactual for MPA evaluation, which is hinted at in Nickols et al.
+2019, but not explicitly done.
 

### 2020-12-01 documents/ovando-et-al-pop-effects-of-mpas-si.Rmd
Edits almost done

@@
 ---
-title: "Online Supporting Information"
+title: "Appendix Supporting Information 1"
 subtitle: "Assessing the Population-level Conservation Effects of Marine Protected Areas"
@@
 
-# (APPENDIX) Appendix {-}
-
 \renewcommand{\thefigure}{S\arabic{figure}}
@@
 
-# Supporting Information (SI)
-
-## Computing environment
+# Computing environment
 
@@
 
-## Simulation Model
+# Simulation Model
 
@@
 
-### Recruitment
+## Recruitment
 
@@
 
-### Dispersal
+## Dispersal
 
@@
 
-### Fleet Dynamics
+## Fleet Dynamics
 
@@
 
-### Spatial Fleet Distribution
+## Spatial Fleet Distribution
 
@@
 
-### MPA Design
+## MPA Design
 
@@
 
-##### Filtering Simulations
+### Filtering Simulations
 
@@
 
-#### Additional Simulation Results
+### Additional Simulation Results
 
@@
 
-## PISCO Data
+# PISCO Data
 
@@
 
-## Difference-in-Difference Model
+# Difference-in-Difference Model
 
@@
 
-### Additional Difference-in-difference Runs
+# Additional Difference-in-difference Runs
 
@@
 
-### Synthetic controls
+## Synthetic controls
 
@@
 
-## Testing Model Assumptions
-
-### Simulation testing
+# Testing Model Assumptions
+
+## Simulation testing
 
@@
 
-### Testing SUTVA with Convergent Cross Mapping
+## Testing SUTVA with Convergent Cross Mapping
 
@@
 
-## References
+# Literature Cited

### 2020-12-01 documents/ovando-et-al-pop-effects-of-mpas.Rmd
Edits almost done

@@
 abstract: |
-   Marine Protected Areas (MPAs) cover 3-7% of the world’s ocean, up from less than 1% in the year 2000, and international commitments call for 10%, 30%, and even 50% coverage.
+   Marine Protected Areas (MPAs) cover 3-7% of the world’s ocean, up from less than 1% in the year 2000, and international organizations call for 30% by 2030.
 The premise underlying MPA expansion is that they conserve biodiversity, habitats, and fished populations.
@@
 A network of MPAs was put in place in The Channel Islands National Marine Sanctuary in 2003, with a goal of providing regional conservation and fishery benefits.
-Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level.
-Using a Bayesian difference-in-difference approach, we show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); the size of many MPA networks suggests that effect sizes may often be smaller than 30%.
+Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level using a Bayesian difference-in-difference appreoach.
+We show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); the size of many MPA networks suggests that effect sizes may often be smaller than 30%.
 Our results provide a novel assessment of the population-level effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.
@@
 Over time our goals and expectations for MPAs have evolved and become more ambitious.
-Recent international efforts to expand MPAs, such as the Biodiversity Beyond National Jurisdiction (BBNJ) Agreement and the Convention on Biological Diversity’s Post-2020 Global Biodiversity Framework, are based in part on the assumption that well-designed MPAs will not only provide conservation benefits inside their borders, but also have broader conservation effects on unprotected areas surrounding the MPAs [@gaines2010].
+Recent international efforts to expand MPAs, such as The International Union for Conservation of Nature's 30% by 2030 MPA targets are based in part on the assumption that well-designed MPAs will not only provide conservation benefits inside their borders, but also have broader conservation effects on unprotected areas surrounding the MPAs [@gaines2010].
 
-The empirical MPA literature has focused on assessing the ability of MPAs to provide conservation gains within their borders [@lester2009; @edgar2014; @gerber2005].
+The empirical MPA literature has focused on assessing the ability of MPAs to provide conservation gains within their borders [@lester2009; @edgar2014].
 However, as conservation benefits accrue inside MPAs, MPAs also affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
@@
 Numerous factors influence how MPAs affect fish populations.
-These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], the strength, timing, and location of density dependence [@burgess2014], the design of the network [@gaines2010; @costello2010a],the degree of enforcement [@edgar2014], the level of fishing pressure, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006a;@walters2000a;@white2011;@white2010c;@ovando2016;@moffitt2013].
+These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003; @botsford2008], the strength, timing, and location of density dependence [@burgess2014], the design of the network [@gaines2010; @costello2010a],the degree of enforcement [@edgar2014], the level of fishing pressure, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@walters2000a;@white2011;@ovando2016;@moffitt2013].
 
 <!-- What empirical evidence is there for these population-level effects of MPAs?  -->
-These theoretical results are generally based on modeling of closed populations, where some fraction of a population is protected inside of MPAs.
+These largely theoretical results are generally based on modeling of closed populations with some fraction protected inside of MPAs.
 In contrast to this population paradigm used in MPA simulations, MPAs are often evaluated empirically at local scales using spatial "response ratios", commonly measured as the ratio of biomass densities (weight of organisms per unit area) of species inside relative to selected control sites outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015].
 These studies found clear evidence that well enforced and sufficiently sized MPAs are associated with high response ratios.
-Several studies have also documented empirical evidence for the existence of adult or larval fish spillover affecting fish abundance [@halpern2009;@kay2012; @stobart2009; @mcclanahan2000; @russ1996; @thompson2017;@baetscher2019;@pelc2009].
+Several studies have also documented empirical evidence for the existence of adult or larval fish spillover affecting fish abundance [e.g.
+@halpern2009;@kay2012; @mcclanahan2000; @russ1996].
 Where response ratios are available before and after MPA implementation, spatial before-after-control-impact (BACI) style studies have shown similarly clear and positive results [@thiault2019a].
@@
 While these studies have clearly demonstrated the ability of MPAs to create differences between local fished and unfished areas, this body of empirical work does not necessarily serve as sufficient evidence for the population-level effects of MPAs.
-This is because these studies rely on assuming that selected control sites serve as a measure of what would have happened in the absence of MPAs.
+These spatial studies rely on assuming that selected control sites serve as a measure of what would have happened in the absence of MPAs.
 Control sites used in calculating response ratios are often selected based on habitat characteristics to justify their use as counterfactuals [@ferraro2018].
-However, the very spillover effects MPAs are increasingly hoped to produce can negate the ability of spatial response ratio / BACI designs to accurately estimate the effects of MPAs.
+However, beyond habitat differences, the very spillover effects MPAs are increasingly hoped to produce can negate the ability of spatial response ratio or BACI designs to accurately estimate the effects of MPAs.
 Export of adults or larvae from MPAs to "control" sites can affect their status as controls, as does displacement of fishing effort from MPAs to control sites.
-These factors can bias estimators of population-level effects that require the assumption that the control sites are unaffected by the treatment (a required assumption of response ratios and BACI designs [@ferraro2018;@moffitt2013]).
+These factors can bias estimators of population-level effects like response ratios that require the assumption that the control sites are unaffected by the treatment [@ferraro2018;@moffitt2013].
 Control sites sufficiently far from MPAs to negate both spillover of fish or larvae and concentration of the fishing fleet could be selected, but finding suitably distant sites that are also appropriate proxies for the ecological and economic context of the MPAs is challenging.
@@
 We present here an paired theoretical and empirical assessment demonstrating the challenges in assessing the population-level impacts of MPAs.
-In 2003, a network of MPAs was established in the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) covering approximately 20% of the Islands' waters [@kirlin2013].
-The network has been used as a model case study in protected area design around the world (@botsford2014 and references therein).
-We use data from the first 14 years of protection in a difference-in-difference model [see @angrist2009 and references therein] to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of fin-fish species.
+In 2003, a network of MPAs was established in the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) covering approximately 20% of the Islands' waters.
+The network has been used as a model case study in protected area design around the world [@botsford2014 and references therein].
+We use data from the first 14 years of protection in a difference-in-difference (DiD) model [see @angrist2009 and references therein] to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of fin-fish species.
 Rather than relying on spatial controls as is typically done in MPA studies, we use groups of species targeted and not-targeted by fishing pressure as our treatment and control groups.
@@
 
-Our primary methods consist of a spatially-explicit bio-economic simulation model and a Bayesian difference-in-difference regression [@angrist2009;@mcelreath2020].
-Difference-in-difference is a regression-based approach to BACI, assessing changes in control and treatment groups before and after treatment [@larsen2018].
-We use the term difference-in-difference here to more closely tie our results to the terminology in causal-inference literature.
-The population-level effects of MPAs will depend on the social-ecological contexts in which they occur.
-We use a bio-economic simulation model to provide theoretical bounds on potential population-level effects of MPAs, which we then confront with the empirical results from our difference-in-difference model.
+Our methods consist of a spatially-explicit bio-economic simulation model and a Bayesian DiD regression [@angrist2009;@mcelreath2020].
+DiD is a regression-based approach to BACI, assessing changes in control and treatment groups before and after treatment [@larsen2018].
+We use the term DiD here to more closely tie our results to the terminology in causal-inference literature.
+We used a bio-economic simulation model to provide theoretical bounds on potential population-level effects of MPAs, which we then confront with the empirical results from our DiD model.
 
 All analysis were conducted in R [@rcoreteam2019].
-Our difference-in-difference model was fit using Stan [@carpenter2017] using the `rstanarm` package [@goodrich2020].
+Our DiD model was fit using Stan [@carpenter2017] using the rstanarm package [@goodrich2020].
 All data and code needed to fully reproduce this manuscript are publicly available at [github.com/DanOvando/population-effects-of-mpas](https://github.com/DanOvando/population-effects-of-mpas).
-A detailed description of the simulation model structure, as well as sensitivity analyses of our estimation model, are available in the online supporting information (SI).
+A detailed description of the simulation model structure, as well as sensitivity analyses of our estimation model, are available in the online Appendix Supporting Information 1.
 
@@
 Readers can explore the functionality of the model using an online tool available at [danovando.shinyapps.io/simmpa/](https://danovando.shinyapps.io/simmpa/).
-The purpose of the simulation model is to both set expectations for our empirical results, and demonstrate more broadly the ways in which ecological and economic dynamics can interact to produce a wide range of population-level MPA effects.
-The full range of factors explored are shown in  (Table.\@ref(tab:sim-table)), and the equations of the simulation model can be viewed in the SI section XX.
-We used this model to generate `r n_distinct(outcomes$experiment)` simulated MPA outcomes across `n_distinct(outcomes$scientific_name)` species.
+The purpose of the simulation model is to both set expectations for our empirical results, and demonstrate the ways in which ecological and economic dynamics can interact to produce a wide range of population-level MPA effects.
+The full range of factors explored are shown in  Table.\@ref(tab:sim-table), and the equations of the simulation model can be viewed in the Appendix S1 Section 2.
+We used this model to generate `r n_distinct(outcomes$experiment)` simulated MPA outcomes across `r n_distinct(outcomes$scientific_name)` species.
 
 Many papers have presented simulation analyses of MPA outcomes [@fulton2015].
-Our model incorporates a broad range of core ecological and economic drivers of MPA performance assessed by individual papers into a cohesive model, similar in spirit to @krueck2017.
+Our model incorporates a broad range of core ecological and economic drivers of MPA performance assessed by these individual papers into a cohesive model, similar in spirit to @krueck2017.
 The simulation model consists of 50 patches with wrapped edges.
@@
 The fleet model can either be constant-catch (the fleet exerts as much effort as needed to maintain a fixed amount of catch), constant-effort (the fleet maintains a constant amount of effort over time), or open-access (fishing effort of the fleet expands and contracts in response to available profits).
-The total fishing effort exerted by the fleet can then be distributed in space uniformly, in proportion to catch per unit effort, or in proportion to profit per unit effort.
+The total fishing effort exerted by the fleet can then be distributed in space uniformly, in proportion to spatial catch per unit effort, or in proportion to spatial profit per unit effort.
 
@@
 Each simulation is paired with a simulation identical in every way except that MPAs are not implemented (i.e., a simulated control).
-We then calculate the effect of the MPAs on the population as the difference in relevant metrics in the simulation with MPAs relative to their values in the simulation without MPAs.
+We then calculate the effect of the MPAs on the population as the difference in biomass densities in the simulation with MPAs relative to biomass densities in the simulation without MPAs.
 
@@
 One set of simulations is specifically designed to reflect the dynamics of the Channel Islands.
-We only include species of the same genus as those targeted by fishing in the Channel Islands, restrict fishing pressure such that the simulated populations are moderately to lightly exploited (given the best available knowledge of stock status for the targeted species in the Channel Islands), and cap the MPA size at 20% of the population's range [@rassweiler2012].
-For these Channel Island-esque simulations, we calculated 1) the population-level response for each species (the 'true' population-wide difference in biomass between the simulations with and without the MPAs) and 2) the measured response ratio of biomass densities inside and outside the simulated MPAs.
-To compare response ratios from the Channel Islands simulations to our empirical results, we organized simulated and empirical response ratios into bins.
-Because each simulation included measures of both response ratios and population-level effects, this provided a library of simulations (and their associated attributes) that could have produced the types of empirical response ratios measured in the Channel Islands.
+For this set we only include species of the same genus as those targeted by fishing in the Channel Islands, restrict fishing pressure such that the simulated populations are moderately to lightly exploited, and cap the MPA size at 20% of the population's range [@rassweiler2012].
+For each of these Channel Island style simulations, we calculated 1) the 'true' population-wide difference in biomass between the simulations with and without the MPAs and 2) the response ratio of biomass densities inside and outside the simulated MPAs.
+We then calculated the empirical response ratios observed in the real Channel Islands, and matched these empirical results with simulations that produced similar response ratios after the same number of years of MPA protection.
+Because each simulation included measures of both response ratios and population-level effects, this process provides a library of simulations (and their associated attributes) that could have produced the types of empirical response ratios measured in the Channel Islands.
 
@@
 
-The difference-in-difference model used empirical kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands.
+The DiD model used empirical kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands.
 PISCO conducts visual SCUBA surveys at a large number of rocky reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\ref{fig:ci-map},Fig.\@ref{fig:raw-density}).
@@
 Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*).
-We used a Bayesian difference-in-difference regression to estimate any difference in mean total biomass densities of fin-fish species targeted by fishing effort (i.e., those potentially affected by an MPA) and those species not targeted by fishing before and after MPA implementation.
-Note that this is akin to a BACI, but rather than relying on spatial control sites as our identification, we utilize non-targeted species as our control.
-To account for the fact that sampling locations are not uniformly distributed across the islands, we weight the samples in our regression in proportion to the area inside and outside of MPAs in the Channel Islands.
+Our regression estimates any difference in mean total biomass densities of fin-fish species targeted by fishing effort (i.e., those potentially affected by an MPA) and those species not targeted by fishing before and after MPA implementation.
+To account for the fact that sampling locations are not uniformly distributed across the islands, we weight the samples in our regression in proportion to the total area inside and outside of MPAs in the Channel Islands.
 
-This identification strategy controls for unobserved environmental shocks to the system that are independent of the MPAs.
+This identification strategy attempts to control for unobserved environmental shocks to the system that are independent of the MPAs.
 Conditional on the assumptions of the model, this regression produces an estimate of the effect of the MPAs on the mean total biomass densities of targeted species throughout the Channel Islands.
 For example, consider an evenly distributed population that has 50% of its range protected by an MPA.
-If the MPA increases biomass densities inside its boundaries by 20%, and by 0% outside the reserve, the population effect of the MPA would be 10%.
+If the MPA increases biomass densities inside its boundaries by 20%, and by 0% outside the reserve, the population effect of the MPA estimated by our DiD would be 10%.
 
-The difference-in-difference (DiD) regression amounts to estimating the pre-post MPA difference in the biomass densities of targeted species, minus the same difference for non-targeted species in the Channel Islands.
+The DiD regression amounts to estimating the pre-post MPA difference in the biomass densities of targeted species, minus the same difference for non-targeted species in the Channel Islands.
 
@@
 where $d_i$ is the biomass density at observation *i*, *T* indicates whether the observation *i* is for a targeted ($T = 1$) or non-targeted ($T = 0$) species, and *MPA* marks whether observation *i* is in a pre MPA ($MPA = 0$) or post MPA ($MPA = 1$) state.
-**B^c^** is a vector of coefficients for additional control variables in matrix *X* such as water visibility and observer experience.
+To account for the fact that MPA effects will evolve over time, we estimate a vector of MPA effects in three-year blocks for the all years after the MPAs were implemented in 2003.
+**B^c^** is a vector of coefficients for additional control variables in matrix *X* such as water temperature and observer experience.
 **B^s^** is a vector of hierarchical coefficients for each sampling location **S** , clustered by island.
@@
 $\beta_0$ is the mean total biomass of non-targeted species pre-MPA, and $\beta_1T$ is the mean total biomass of targeted species pre-MPA.
-$\beta_2MPA$ is the effect of the post-MPA period on the non-targeted species.
-The shape and scale parameters of the Gamma distribution are estimated as well (Table.\@ref(tab:coef-table)).
-
-This difference-in-difference regression provides our estimate of the population-level effect of the Channel Islands MPAs.
-We use a Bayesian regression since it allows us to interpret our estimated effects probabilistically [@mcelreath2020], e.g.
-stating the probability that conditional on the model, priors,and data the Channel Island MPAs increased or decreased biomass of targeted species.
-Being a Bayesian regression, our difference-in-difference model produces posterior probability distributions of our coefficients, from which we can construct Bayesian credible intervals.
-Since the effects of MPAs will evolve over time, we use this model to estimate a series of population-level effects of the Channel Islands MPAs in three-year intervals from 2003 to 2017.
+$\beta_2MPA$ is the effect of the post-MPA period on the non-targeted species (Table.\@ref(tab:coef-table)).
+We use a Bayesian regression since it allows us to interpret our estimated effects probabilistically [@mcelreath2020].
+Being a Bayesian regression, our DiD model produces posterior probability distributions (the probability distribution of our coefficients conditional on the data, priors, and model assumptions) of our coefficients, from which we can construct Bayesian credible intervals [@gelman2014a].
 
@@
 
-Our Channel Islands-specific simulation provides a set of outcomes that can feasibly bound our expectations for the empirical effects of the MPA network.
+<!-- Our Channel Islands-specific simulation provides a set of outcomes that can feasibly bound our expectations for the empirical effects of the MPA network.
 The targeted species in our database span a range of life histories, but are largely made up of fishes in the perch and rockfish complexes, with a mean Von Bertalanffy growth coefficient of `r round(mean(used_data$k[used_data$targeted == 1]),2)` (standard deviation `r round(sd(used_data$k[used_data$targeted == 1]),2)`), and a median age at maturity of `r round(median(used_data$tm[used_data$targeted == 1]),0)` years (standard deviation `r round(sd(used_data$tm[used_data$targeted == 1]),0)`), and a mix of exploitation histories, ranging from low to high (but mostly relatively low) [@alonzo2004;@dick2017].
-While a wide range of MPA effects are plausible, from 0% to upwards of 150%, the median simulated effect size for this stylized version of the Channel Islands after over ten years of protection was `r percent(median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected > 10]))`.
+-->
 
-We compared this range of plausible outcomes to the empirical data.
 @caselle2015 found a statistically significant increase in the response ratios of targeted species over time, and evidence that this increase is smaller for non-targeted species.
@@
 
-These response ratios cannot, however, be used as a definitive indicator of population effects of the MPAs.
+These response ratios cannot, however, be used as a definitive indicator of population effects of these MPAs.
 In the case of the Channel Islands MPAs, control sites are often located within a few kilometers of an MPA, suggesting that they are susceptible to both local biological spillover and by concentration of fishing effort excluded from the MPAs.
@@
 
-Our difference-in-difference regression provides a different perspective on population-wide MPA effects.
+Our targeted vs.
+non-targeted DiD regression provides an alternative approach to spatial controls for estimating population-level MPA effects.
 Over the first three years of implementation (2003-2006), the effects of the MPAs are unclear, with support for a small negative effect to a substantially positive effect, with much higher probability of a small positive effect (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`,
@@
 These empirical estimates are in line with the outcomes that our simulation model suggests are plausible.
+While a wide range of MPA effects are plausible, from 0% to upwards of 150%, the median simulated effect size for this stylized version of the Channel Islands after over ten years of protection was `r percent(median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected > 10]))`.
 However, in the subsequent years the trend reverses itself, and in 2015-2017 we once again see no clear effect of the MPAs (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`,
@@
 
-Given natural variability in ecosystems and uncertainty in measurements, MPA effects may be difficult to detect with precision.
-Our library of simulation results allows us to explore how MPA effects, and our ability to measure them, might vary with observation error and natural fish recruitment variation (including the possibility of recruitment regimes).
-Using a Bayesian difference-in-difference regression on these simulated data, we estimated the percent error between the posterior probability distribution of the estimated MPA effect from the regression and the true simulated MPA effect.
+Our library of simulation results allows us to explore how MPA effects, and our ability to measure them, might vary with observation error and natural fish recruitment variation.
+Using a Bayesian DiD regression on these simulated data, we estimated the percent error between the posterior probability distribution of the estimated MPA effect from the regression and the true simulated MPA effect.
 The model struggled severely when MPA effect sizes were less than 25% and the model was faced with both observation and process errors (Fig.\@ref(fig:val-plot)).
@@
 Containing a carefully designed, well-enforced, and well-studied MPA network, the Channel Islands seems to be an ideal location to study the population-level effects of protected areas.
-Persistently high response ratios suggest that despite overall decreases in targeted biomass densities inside and outside MPAs relative to pre-MPA measurements (Fig.\@ref(fig:raw-density)), the MPAs may still be providing protection within their borders.
-But, these response ratios are not necessarily an indicator of broader population-level effects.
+High response ratios suggest that despite overall decreases in targeted biomass densities inside and outside MPAs relative to pre-MPA measurements (Fig.\@ref(fig:raw-density)), the MPAs may still be providing protection within their borders.
+But, these spatial response ratios are not necessarily an indicator of broader population-level effects.
 The targeted vs.
-non-targeted difference-in-difference strategy presents an alternative means of estimating the population-level effects of MPAs.
+non-targeted DiD strategy presents an alternative means of estimating the population-level effects of MPAs.
 While having its own strict caveats, this strategy provides some potential improvements over spatial response-ratios as a means of estimating population-level effects.
-While we estimate an uncertain but overall positive effect of the MPA network in its first few years, we are unable to detect a robust positive signal from 2012-2017.
+While we estimate an uncertain but overall positive effect of the MPA network in its first few years, we are unable to detect a robust signal from 2012-2017.
 After 14 years of MPA protection we are left without a clear picture of the population-level effect of the Channel Island MPA network on biomass densities of targeted fin-fish species.
@@
 Fishers pursuing constant-catch in areas outside an MPA may have to fish harder to achieve the same catch from a smaller part of the population, causing a population conservation loss under `r percent(1 - positive_outcomes$pp[positive_outcomes$fleet_model == "constant-catch"], accuracy = 1)` of our constant-catch simulations.
-This potential negative interaction between constant-catch and MPAs is an important risk to consider (as done so in @little2011b) especially as MPAs are increasingly implemented in quota-managed fisheries [@liu2018].
+This potential negative interaction between constant-catch and MPAs is an important risk to consider (as done in @little2011b) especially as MPAs are increasingly implemented in quota-managed fisheries [@liu2018].
 
-While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (SI Fig.S1 XX ) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
+While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (Appendix S1 Fig.S2) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
 However, some of the most common targeted species in the Channel Islands, such as blue and copper rockfish, have seen dramatic increases in reported catches in the Santa Barbara region since the year 2010 (although other such as California sheephead have seen declines relative to pre-MPA catch level).
@@
 
-<!-- MPAs are an important part of the marine resource management toolbox.
+MPAs are an important part of the marine resource management toolbox.
 As the number and size of global MPA networks increase, we must set appropriate expectations for their outcomes on both local and regional scales.
 Our results for the relatively large, well-enforced, and rigorously studied (though also relatively lightly fished) Channel Islands Marine Protected Area network make clear that in many instances empirically detecting a clear effect of MPAs on total biomass densities of targeted fin-fish at the population level may not be possible.
-In the absence of clear empirical findings, simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can be tuned to focus on the species groups that have the highest chance of a detectable effect size over the early years of the reserve.
+In the absence of clear empirical findings, simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can be tuned to focus on the species groups that have the highest chance of a detectable effect size over the early years of the reserve [@nickols2019].
 Expanding data collection to include robust monitoring of spatio-temporal fleet dynamics may help assess the validity of control sites used in response ratios, support the direct inclusion of these fleet dynamics into statistical models, and allow managers to take into account potential negative interactions between MPAs and fleet dynamics such as those that may occur under constant-catch dynamics.
 Whenever possible monitoring programs should be implemented prior to MPA implementation to provide a pre-treatment benchmark.
--->
 
-As @larsen2019 suggests, there are many potential alternatives to empirical response ratios for estimating the population effects of MPAs that better account for the challenges of causal inference (though that may be more data-intensive).
+There are many potential alternatives to spatial response ratios for estimating the population effects of MPAs that better account for the challenges of causal inference (though that may be more data-intensive) [@larsen2019].
 We applied one such approach here, and yet were still unable to reach robust conclusions as to the effect of MPAs on the total biomass density of targeted fin-fish in the Channel Islands, due to the likely small size of the true effect relative to the influence of environmental variability.
-There are other promising statistical approaches to setting expectations for MPA effects, including using models fitted to local data to set population-level expectations [@nickols2019; @white2011].
+There are other promising statistical approaches to setting expectations for MPA effects, including using models fitted to local data to set population-level expectations and create synthetic counterfactuals [@nickols2019; @white2011].
 
 However, as the scientific community continues to develop appropriate measures of population-level MPA effects, we must also effectively communicate the challenges of estimating those effects.
-Lack of a clear population-level MPA effect should not necessarily be viewed as a failure of a conservation program, and likewise, a large positive effect based solely on response ratios should not be the sole justification for the relaxation of other conservation measures.
+In light of these results, lack of a clear population-level MPA effect should not necessarily be viewed as a failure of a conservation program, and likewise, large response ratios should not be automatically taken as evidence of a population-level conservation success.
 Rather, results and subsequent management actions must be considered in the context of reasonable expectations given the size, age, and degree of enforcement of the MPAs in question, together with the ecological and economic dynamics of a given system.
@@
 
-As advocacy for broad networks of MPAs is growing globally, future MPA science must directly tackle the challenge of evaluating the performance of MPAs at the regional scale.
-Commonly employed metrics such as response ratios may be applicable in some circumstances, but are vulnerable to inaccuracy or misuse as metrics of population-level effects.
+As advocacy for broad networks of MPAs is growing globally, MPA science must directly tackle the challenge of evaluating the performance of MPAs at the population scale.
+Commonly employed metrics such as spatial response ratios may be applicable in some circumstances, but are vulnerable to inaccuracy or misuse as metrics of population-level effects.
 Bio-economic modeling can help frame community expectations, reducing the potential for a reduction in support if unrealistic conservation or fishery expectations are not realized.
-Statistical approaches like difference-in-difference regression that explicitly address complications such as the spatial spillover effects of MPAs may give users an improved understanding of the performance of their MPAs, but even they may struggle when expected effect sizes are small.
+Statistical approaches that explicitly address complications such as the spatial spillover effects of MPAs may give users an improved understanding of the performance of their MPAs, but even they may struggle when expected effect sizes are small.
 Clearly communicating what we should expect, and what we can detect, from MPAs is critical to ensuring that MPAs play effective roles in fisheries management and marine conservation.
@@
 
+# Supporting Information  {#si .unnumbered}
+
+Detailed descriptions of the simulation model and DiD estimator (Appendix S1) are available online.
+The authors are solely responsible for the content and functionality of these materials.
+Queries (other than absence of the material) should be directed to the corresponding author.
+
 # Literature Cited {#lit .unnumbered}
@@
 
+\newpage
+
 # Tables {#tables .unnumbered}
 
-```{r sim-table, results = "asis", echo=FALSE}
+```{r sim-table, echo=FALSE}
 variable <- c(
@@
              caption = "Variables for MPA simulations",
-             format = "pipe") %>% 
+             format = "pipe",
+             align = "l") %>% 
   kableExtra::kable_styling()
@@
 
-```{r coef-table, results = "asis", echo=FALSE}
+\newpage
+
+```{r coef-table, echo=FALSE}
 
@@
     Mean = estimate,
-    `Lower 90th Percentile` = lower,
-    `Upper 90th Percentile` = upper
+    `5th Percentile` = lower,
+    `95th Percentile` = upper
   )
+
+coefs$term <- str_remove_all(coefs$term, "var_")
+
+coefs$term <- str_replace_all(coefs$term, "targeted","Targeted by Fishing")
+
+coefs$term <- str_replace_all(coefs$term, "year_bins","Year Block ")
+
+coefs$term <- str_replace_all(coefs$term, ":","$\\\\times$")
+
+coefs <- coefs %>% 
+  mutate(term = case_when(term == "tex" ~ "Observer Experience",
+                          term == "tex_2" ~ "(Observer Experience)$^2$",
+                          term == "temp" ~ "Water Temperature",
+                          term == "regional_temp_dev" ~ "Deviation from regional mean temperature",
+                          term == "regional_temp_dev_2" ~ "(Deviation from regional mean temperature)$^2$",
+                          term == "lag_catch" ~ "Lagged Commercial Catch",
+                          term == "targeted" ~ "Targeted by Fishing",
+                          term == "(Intercept)" ~ "Intercept",
+                          term == "surge" ~ "Wave Surge",
+                          term == "kelp" ~ "Kelp Cover",
+                           TRUE ~ term))  %>% 
+  rename(Variable = term)
 
@@
     caption = glue::glue("Posterior means and credible interval for key model coefficients.
-Conditional on the assumptions of the model, the targeted:year_bins coefficients reflect the effect of the MPAs on mean total biomass of targeted finfish in the Channel Islands.
-targeted indicates whether the species in question is targeted, tex = technican experience, surge is mean reported surge, kelp is mean reported kelp cover, lag catch is the total reported catch for targeted species in the previous year, temp is the mean temperature at a site, temp_dev is the deviation of a temperature observation from the mean temperatuer for that island in question.
-Raw data contained {nrow(used_data)} individual observations, total number of targeted and non-targeted mean mean total biomass densities by site and year used to fit model was {nrow(did_fit$data)}"),
-    format = "pipe")
+Raw data contained {nrow(used_data)} individual observations, total number of targeted and non-targeted mean mean total biomass densities by site and year used to fit model was {nrow(did_fit$data)}.
+Site-level terms omitted for clarity."),
+    format = "pipe", 
+    escape = FALSE,
+    align = "l")
 
 ```
+
+\newpage
 
@@
 ```{r ci-map, include = TRUE,fig.width=width_1col, fig.asp=1.5, fig.cap="Map of study region; the Northern Channel Islands, California, USA.
-Colors show binned number of PISCO sampling events across the time period of our study.", include=TRUE}
+Shading shows binned number of PISCO sampling events across the time period of our study.", include=TRUE}
 
@@
 
+\newpage
+
 ```{r raw-density, fig.width = width_1col, fig.asp=asp_1col,fig.cap = "Centered and scaled trends in biomass densities of targeted and non-targeted fin-fish included in our study.
 Top panel shows mean trends across all sites, with shaded areas shwoing 95% confidence intervals.
-Bottom two panels show aggregate biomass density trends outside and inside MPAs"}
+Bottom two panels show same trends but only including sites outside and inside MPAs"}
 
@@
 
-```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="90% Posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis) in dark grey.
+\newpage
+
+```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="Simulated population effects (grey) that could have produced observed response ratios (black) in the Channel Islands.
+Black histograms show 90% Posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis).
 Simulated population-level (pop.) effect on biomass densities matched to empirical response ratios in light grey.
@@
 
-```{r mpa-effect, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Results of difference-in-difference regression estimating the population effect of the Channel Island MPAs on mean total biomass densities of targeted species (difference in mean total biomass density of targeted species over time relative to expected levels using non-targeted species as a control).
-Grey distributions show posterior probability distribution of estimated MPA effect; red point is median estimated effect, thicker red section 50% credible interval, thinner red line 90% credible interval).
+\newpage
+
+```{r mpa-effect, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Results of difference-in-difference regression estimating the population effect of the Channel Island MPAs on mean total biomass densities of targeted species.
+Grey distributions show posterior probability distribution of estimated MPA effect; red point is median estimated effect, thicker red section 50% credible interval, thinner red line 90% credible interval.
 Blue distributions in background show range of MPA effects produced by simulation model tuned to reflect the dynamics of the Channel Island MPAs (black dashed line is median simulated value).
@@
 ```
+
+\newpage
 
@@
 
-```{r sim-effects, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Simulated population-level (pop.) MPA effect sizes as a function of percent of species' range inside MPA (x-axis), and pre-MPA depletion (y-axis).
+\newpage
+
+```{r sim-effects, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Simulated population-level effects of MPAs.
+Panel a) shows median simulated population-level (pop.) MPA effect sizes as a function of percent of species' range inside MPA (x-axis), and pre-MPA depletion (y-axis).
 Pre-MPA depletion is a measure of fishing pressure, where 0 means that the population is unfished, and 1 means that the population is extinct in the time period immediately prior to MPA implementation.
-Panel A) shows median MPA effects across range in MPA and pre-MPA depletion Panel B) shows distribution of simulations across range of MPA size and pre-MPA depletion separately."}
+Panel b) shows distribution of simulations across range of MPA size and pre-MPA depletion separately."}
 

### 2020-12-01 documents/test-tables.Rmd
Edits almost done

--- 2020-12-01 documents/ovando-et-al-pop-effects-of-mpas.Rmd
@@
 ---
-title: Assessing the Population-Level Conservation Effects of Marine Protected Areas
-abstract: |
-   Marine Protected Areas (MPAs) cover 3-7% of the world’s ocean, up from less than 1% in the year 2000, and international organizations call for 30% by 2030.
-The premise underlying MPA expansion is that they conserve biodiversity, habitats, and fished populations.
-While numerous studies show that MPAs produce conservation benefits *inside* their borders, many MPAs are also justified on the grounds that they confer conservation benefits to the broader population *beyond* their borders.
-We examine the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and show that even in cases where the population benefits are large, they are inherently difficult to detect empirically.
-A network of MPAs was put in place in The Channel Islands National Marine Sanctuary in 2003, with a goal of providing regional conservation and fishery benefits.
-Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level using a Bayesian difference-in-difference appreoach.
-We show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); the size of many MPA networks suggests that effect sizes may often be smaller than 30%.
-Our results provide a novel assessment of the population-level effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.
-bibliography: references.yaml
-csl: conservation-biology.csl
+title: "test-table"
+author: "Dan Ovando"
+date: "11/30/2020"
 output:
@@
     reference_docx: template.docx
-  bookdown::pdf_document2:
-    keep_tex: true
-  bookdown::html_document2: default
-params:
-  run_name: ["v6.0"]
-  sim_years: [50]
-  burn_years: [20]
-  num_patches: [25]
-linkcolor: blue
-toc: false
-header-includes:
-  - \usepackage{setspace}\doublespacing
-  - \usepackage{lineno}\linenumbers
-  - \usepackage{draftwatermark}
-  - \SetWatermarkText{DRAFT}
-always_allow_html: true
 ---
 
-```{r, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 300,
-                      cache = FALSE, fig.width = 3.42, fig.asp = 1.5)
-
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+library(tidyverse)
+library(flextable)
 ```
-
-```{r, include=FALSE}
-library(sf)
-library(ggmap)
-library(viridis)
-library(hrbrthemes)
-library(scales)
-library(patchwork)
-library(rpart)
-library(extrafont)
-library(caret)
-library(ggsci)
-library(rEDM)
-library(tidyverse)
-library(ggtext)
-library(here)
-library(broom.mixed)
-extrafont::loadfonts()
-functions <- list.files(here::here("functions"))
-
-walk(functions, ~ here::here("functions", .x) %>% source()) # load local functions
-save_plots <- TRUE
-
-# sim_years <- 50x
-# 
-# burn_years <- 20
-# 
-# num_patches <- 25
-
-# run_name <- "v5.0"
-
-width_1col <- 3.42 # inches
-
-asp_1col <- 1.25
-
-width_2col <- 7 # inches
-
-asp_2col <-  .75
-
-dpi <- 900
-
-theme_1col <- hrbrthemes::theme_ipsum(base_size = 11,
-                                      axis_title_size = 12)
-
-theme_2col <- hrbrthemes::theme_ipsum(base_size = 12,
-                                        axis_title_size = 14)
-
-run_name <- params$run_name
-
-sim_years <- params$sim_years
-
-burn_years <- params$burn_years
-
-num_patches <- params$num_patches
-
-run_dir <- here::here("results", run_name)
-
-experiment_dir <- here::here("results",run_name, "experiments")
-
-load(file = here::here("results",run_name, "rawish_zissou_data.Rdata"))
-
-load(file.path(run_dir, 'abundance_data.Rdata'))
-
-model_runs <- read_rds( file.path(run_dir,"did_fits.rds"))
-
-# load(file = here::here("results",run_name, "model_runs.Rdata"))
-
-processed_grid <-  read_rds(file.path(run_dir,"filtered_processed_grid.rds"))
-
-load(file = file.path(run_dir ,"sim_grid.Rdata"))
-
-load(file = file.path(run_dir,"plots.RData"))
-
-valgrid <- read_rds(file.path(run_dir, "valgrid.rds"))
-
-cdfw_catches <-
-  read_csv(file = here::here("data", 'cdfw-catches.csv')) %>%
-  group_by(sci_name, year) %>%
-  summarise(catch = sum(pounds_caught, na.rm = T)) %>%
-  left_join(
-    life_history_data %>% select(taxa, classcode, commonname) %>% mutate(taxa = tolower(taxa)),
-    by = c("sci_name" = "taxa")
-  ) %>%
-  filter(!is.na(classcode))
-
-# cdfw_catches %>%
-#   group_by(commonname, classcode, sci_name) %>%
-#   summarise(catch =sum(catch, na.rm = TRUE)) %>%
-#   mutate(in_analysis = classcode %in% unique(cip_data$classcode)) %>%
-#   View()
-
-density_ratios <- readRDS(file.path(run_dir,"density_ratios.rds"))
-
-outcomes <- readRDS(file.path(run_dir,"outcomes.rds"))
-
-outcomes <- outcomes %>% 
-  filter(!is.na(mpa_effect))
-
-density_ratios <- density_ratios %>% 
-  filter(!is.na(biased_density_ratio))
-
-   eqo <- outcomes %>%
-     filter(year == max(year))
-
-facet_labels <- c(
-  mpa_size = "Range in MPA",
-  depletion = "Depletion"
-)
-
-testplot <- ggplot()
-
-panel_height = unit(1,"npc") - sum(ggplotGrob(testplot)[["heights"]][-3]) - unit(4,"line")
-
-panel_width = unit(1,"npc") - sum(ggplotGrob(testplot)[["widths"]][-3]) - unit(4,"line")
-
-gc <- guide_colorbar(frame.colour = "black",
-                     ticks.colour = "white",
-                     barheight = panel_height)
-
-  hgc <- guide_colorbar(frame.colour = "black",
-                       ticks.colour = "black",
-                       barwidth = 10)
-
-    wide_hgc <- guide_colorbar(frame.colour = "black",
-                       ticks.colour = "black",
-                       barwidth = 25)
-
-theme_set(theme_1col)
-
-# models_worked <- model_runs$tmb_fit %>% map("error") %>% map_lgl(is_null)
-
-# model_runs <- model_runs %>%
-#   filter(models_worked) %>%
-#   mutate(tmb_fit = map(tmb_fit,"result")) %>%
-#   mutate(processed_fits = map(tmb_fit, process_fits)) %>%
-#   mutate(did_plot = map(processed_fits, "did_plot"))
-
-base_run <- model_runs %>%
-filter(var_names == "pisco_a", data_to_use == "all", center_scale == TRUE)
-
-mpa_run <- model_runs %>%
-  filter(data_source == "pisco",
-         var_names == "pisco_a",
-         data_to_use == "mpa_only")
-
-fishsed_run <- model_runs %>%
-  filter(data_source == "pisco",
-         var_names == "pisco_a",
-         data_to_use == "fished_only")
-
-kfm_run <- model_runs %>%
-  filter(data_source == "kfm")
-
-used_data <- abundance_data$data[abundance_data$data_source == "pisco"][[1]]
-
-tmdata <- used_data %>% 
-  filter(targeted == 1) %>% 
-  select(tm, everything())
-
-median_tm = round(median(tmdata$tm))
-
-site_data <- read_csv(here::here("data",'Final_Site_Table_UCSB.csv')) %>%
-  magrittr::set_colnames(., tolower(colnames(.))) %>%
-  select(
-    site,
-    side,
-    mpagroup,
-    mpa_status,
-    reserve,
-    region,
-    year_mpa,
-    mpaareanm2,
-    lat_wgs84,
-    lon_wgs84
-  ) %>%
-  rename(lat_wgs84 = lon_wgs84,
-         lon_wgs84 = lat_wgs84) %>%
-  unique() %>%
-  mutate(eventual_mpa = (year_mpa > 0))
-
-top_species <- used_data$classcode %>% unique()
-
-life_summaries <- life_history_data %>% 
-  select(classcode, tm, vbgf.k, vbgf.linf) %>% 
-  rename(age_mature = tm, vbk = vbgf.k, linf = vbgf.linf) %>% 
-  unique() %>% 
-  filter(classcode %in% top_species) %>% 
-  select(-classcode) %>% 
-  gather(variable,value) %>% 
-  group_by(variable) %>% 
-  summarise(min_val = min(value, na.rm = TRUE),
-         max_val = max(value, na.rm = TRUE))
-
- sim_life_history <- sim_grid %>%
-    mutate(
-      age_mature = map_dbl(fish, "age_mature"),
-      vbk = map_dbl(fish, "vbk"),
-      linf = map_dbl(fish, "linf")
-    ) %>%
-    select(scientific_name, age_mature, vbk, linf) %>%
-    unique() %>% 
-   filter(age_mature > life_summaries$min_val[life_summaries$variable == "age_mature"],
-          age_mature < life_summaries$max_val[life_summaries$variable == "age_mature"],
-          vbk > life_summaries$min_val[life_summaries$variable == "vbk"],
-          vbk < life_summaries$max_val[life_summaries$variable == "vbk"],
-          linf > life_summaries$min_val[life_summaries$variable == "linf"],
-          linf < life_summaries$max_val[life_summaries$variable == "linf"])
-
- year_bins <-
-  seq(2003, max(pisco_data$year) + 1, by = 3)
-
- sim_mpa_effects <- outcomes %>% 
-  filter(set == "ci",
-         final_depletion > 0) %>% 
-  mutate(year = years_protected + 2002) %>% 
-  filter(year <= max(pisco_data$year)) %>% 
-  mutate(binned_year = cut(year, year_bins, include.lowest = FALSE))  %>% 
-  filter(year >= 2003,
-         !is.na(binned_year)) %>% 
-  mutate(mpa_effect = pmin(mpa_effect,3))
-
- did_fit <- model_runs$did_fit[model_runs$data_source == "pisco" &
-                                 model_runs$var_names == "pisco_a" &
-                                 model_runs$data_to_use == "all"][[1]]$did_reg
-
- if (!dir.exists(here("documents", "figs"))){
-
-   dir.create(here("documents", "figs"))
-
- }
-```
-
-<!-- \AddToShipoutPictureFG{ -->
-<!--   \AtPageCenter{% or \AtTextCenter -->
-<!--     \makebox[0pt]{\rotatebox[origin=c]{45}{% -->
-<!--       \scalebox{5}{\texttransparent{0.3}{DRAFT}}% -->
-<!--     }} -->
-<!--   } -->
-<!-- } -->
-
-# Introduction {#intro .unnumbered}
-
-No-take Marine Protected Areas (MPAs), spatial regions of the ocean in which fishing is prohibited, have a long history in the management of marine resources.
-Modern MPAs were first established as marine analogs to the terrestrial protection of iconic landscapes [@iucn1976].
-Over time our goals and expectations for MPAs have evolved and become more ambitious.
-Recent international efforts to expand MPAs, such as The International Union for Conservation of Nature's 30% by 2030 MPA targets are based in part on the assumption that well-designed MPAs will not only provide conservation benefits inside their borders, but also have broader conservation effects on unprotected areas surrounding the MPAs [@gaines2010].
-
-The empirical MPA literature has focused on assessing the ability of MPAs to provide conservation gains within their borders [@lester2009; @edgar2014].
-However, as conservation benefits accrue inside MPAs, MPAs also affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
-Therefore, MPAs contribute to both local and population-level effects (i.e., inside-MPA effects and broader regional effects).
-Numerous factors influence how MPAs affect fish populations.
-These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003; @botsford2008], the strength, timing, and location of density dependence [@burgess2014], the design of the network [@gaines2010; @costello2010a],the degree of enforcement [@edgar2014], the level of fishing pressure, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@walters2000a;@white2011;@ovando2016;@moffitt2013].
-
-<!-- What empirical evidence is there for these population-level effects of MPAs?  -->
-These largely theoretical results are generally based on modeling of closed populations with some fraction protected inside of MPAs.
-In contrast to this population paradigm used in MPA simulations, MPAs are often evaluated empirically at local scales using spatial "response ratios", commonly measured as the ratio of biomass densities (weight of organisms per unit area) of species inside relative to selected control sites outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015].
-These studies found clear evidence that well enforced and sufficiently sized MPAs are associated with high response ratios.
-Several studies have also documented empirical evidence for the existence of adult or larval fish spillover affecting fish abundance [e.g.
-@halpern2009;@kay2012; @mcclanahan2000; @russ1996].
-Where response ratios are available before and after MPA implementation, spatial before-after-control-impact (BACI) style studies have shown similarly clear and positive results [@thiault2019a].
-
-While these studies have clearly demonstrated the ability of MPAs to create differences between local fished and unfished areas, this body of empirical work does not necessarily serve as sufficient evidence for the population-level effects of MPAs.
-These spatial studies rely on assuming that selected control sites serve as a measure of what would have happened in the absence of MPAs.
-Control sites used in calculating response ratios are often selected based on habitat characteristics to justify their use as counterfactuals [@ferraro2018].
-However, beyond habitat differences, the very spillover effects MPAs are increasingly hoped to produce can negate the ability of spatial response ratio or BACI designs to accurately estimate the effects of MPAs.
-Export of adults or larvae from MPAs to "control" sites can affect their status as controls, as does displacement of fishing effort from MPAs to control sites.
-These factors can bias estimators of population-level effects like response ratios that require the assumption that the control sites are unaffected by the treatment [@ferraro2018;@moffitt2013].
-Control sites sufficiently far from MPAs to negate both spillover of fish or larvae and concentration of the fishing fleet could be selected, but finding suitably distant sites that are also appropriate proxies for the ecological and economic context of the MPAs is challenging.
-As variations of spatial response ratios and BACI studies are our primary source of evidence for the conservation effects of MPAs, this means that our empirical understanding of the population-level impacts of MPAs is surprisingly limited.
-
-<!-- Given the potential unreliability of response ratios for the task, how can we empirically measure the population-level effects of MPAs?
-The perfect experiment would involve two parallel worlds that were identical, except for the implementation of an MPA.
-In world “A”, no MPA would be implemented, and in the facsimile world “B”, the MPA would be implemented.
-Both worlds would be tracked before and after MPA implementation, and the biomass densities would be compared after treatment.
-Instead of two parallel worlds, a similar experimental design would involve random placement of MPAs.
-Unfortunately, neither of these experimental designs has been implemented, because parallel worlds do not exist and MPAs are to our knowledge never placed at random.
-Careful causal inference requires some other means of controlling for biases introduced by factors such as the unobserved environmental shocks, the MPA siting process, biological spillover, and concentration of fishing effort  [@ferraro2018;@larsen2019].
--->
-
-<!-- As stakeholders around the world seek to use MPAs in marine resource management portfolios, it is critical that we develop a better understanding of the magnitude and drivers of population-level MPA effects.
-This is particularly true in the context of a changing climate, in which "set it and forget it" management strategies are increasingly untenable; we need to be able to monitor the performance of MPAs in order to adapt them as dictated by a shifting environment.
-From a fishery perspective, are spillover benefits sufficient to offset losses in fishing grounds and changes in responses of displaced fishers caused by an MPA?
-From a conservation perspective, how much does the buildup of fish inside an MPA increase biomass outside the protected area?
-Overall, what are the population-level effects of MPAs?
--->
-
-We present here an paired theoretical and empirical assessment demonstrating the challenges in assessing the population-level impacts of MPAs.
-In 2003, a network of MPAs was established in the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) covering approximately 20% of the Islands' waters.
-The network has been used as a model case study in protected area design around the world [@botsford2014 and references therein].
-We use data from the first 14 years of protection in a difference-in-difference (DiD) model [see @angrist2009 and references therein] to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of fin-fish species.
-Rather than relying on spatial controls as is typically done in MPA studies, we use groups of species targeted and not-targeted by fishing pressure as our treatment and control groups.
-In contrast to clear differences in biomass densities observed inside and outside of well-protected MPAs both globally [@lester2009] and in the Channel Islands [@caselle2015] we are unable to detect a clear population effect from the Channel Islands MPAs.
-We build off of existing MPA theory to consider why this might be, and provide guidance for scientists and managers as to when and how we might expect to estimate the population-level conservation effects of MPAs.
-
-# Methods {#methods .unnumbered}
-
-Our methods consist of a spatially-explicit bio-economic simulation model and a Bayesian DiD regression [@angrist2009;@mcelreath2020].
-DiD is a regression-based approach to BACI, assessing changes in control and treatment groups before and after treatment [@larsen2018].
-We use the term DiD here to more closely tie our results to the terminology in causal-inference literature.
-We used a bio-economic simulation model to provide theoretical bounds on potential population-level effects of MPAs, which we then confront with the empirical results from our DiD model.
-
-All analysis were conducted in R [@rcoreteam2019].
-Our DiD model was fit using Stan [@carpenter2017] using the rstanarm package [@goodrich2020].
-All data and code needed to fully reproduce this manuscript are publicly available at [github.com/DanOvando/population-effects-of-mpas](https://github.com/DanOvando/population-effects-of-mpas).
-A detailed description of the simulation model structure, as well as sensitivity analyses of our estimation model, are available in the online Appendix Supporting Information 1.
-
-## Simulation Model {#simmodel .unnumbered}
-
-Our bio-economic model simulates the effect of MPAs on a spatially explicit age-structured representation of a fish population.
-Readers can explore the functionality of the model using an online tool available at [danovando.shinyapps.io/simmpa/](https://danovando.shinyapps.io/simmpa/).
-The purpose of the simulation model is to both set expectations for our empirical results, and demonstrate the ways in which ecological and economic dynamics can interact to produce a wide range of population-level MPA effects.
-The full range of factors explored are shown in  Table.\@ref(tab:sim-table), and the equations of the simulation model can be viewed in the Appendix S1 Section 2.
-We used this model to generate `r n_distinct(outcomes$experiment)` simulated MPA outcomes across `r n_distinct(outcomes$scientific_name)` species.
-
-Many papers have presented simulation analyses of MPA outcomes [@fulton2015].
-Our model incorporates a broad range of core ecological and economic drivers of MPA performance assessed by these individual papers into a cohesive model, similar in spirit to @krueck2017.
-The simulation model consists of 50 patches with wrapped edges.
-For each simulation we first randomly pull a species and its associated life history traits from the `FishLife` [@thorson2017c] package.
-We pair these data with randomly selected values governing the characteristics of the simulation (Table.\@ref(tab:sim-table)).
-Key choices available to the model include parameters governing fishing pressure and MPA design.
-For a given simulation, the model randomly selects a fleet model and fishing effort allocation strategy.
-The fleet model can either be constant-catch (the fleet exerts as much effort as needed to maintain a fixed amount of catch), constant-effort (the fleet maintains a constant amount of effort over time), or open-access (fishing effort of the fleet expands and contracts in response to available profits).
-The total fishing effort exerted by the fleet can then be distributed in space uniformly, in proportion to spatial catch per unit effort, or in proportion to spatial profit per unit effort.
-
-The simulation then begins to apply the fleet model to the population, and in a randomly selected year implements an MPA network.
-The model samples a percentage of the population's range to place in MPAs, and randomly assigns patches to MPAs either across across a uniform system or preferentially on higher-quality habitat.
-The model then randomly selects whether fishing effort that used to operate inside the MPAs is redistributed to the areas outside the MPAs, or leaves the fishery entirely.
-We then continue the simulations with the MPAs in place.
-Each simulation is paired with a simulation identical in every way except that MPAs are not implemented (i.e., a simulated control).
-We then calculate the effect of the MPAs on the population as the difference in biomass densities in the simulation with MPAs relative to biomass densities in the simulation without MPAs.
-
-These simulation results provide a library of plausible MPA effects for a range of biological and economic assumptions.
-One set of simulations is specifically designed to reflect the dynamics of the Channel Islands.
-For this set we only include species of the same genus as those targeted by fishing in the Channel Islands, restrict fishing pressure such that the simulated populations are moderately to lightly exploited, and cap the MPA size at 20% of the population's range [@rassweiler2012].
-For each of these Channel Island style simulations, we calculated 1) the 'true' population-wide difference in biomass between the simulations with and without the MPAs and 2) the response ratio of biomass densities inside and outside the simulated MPAs.
-We then calculated the empirical response ratios observed in the real Channel Islands, and matched these empirical results with simulations that produced similar response ratios after the same number of years of MPA protection.
-Because each simulation included measures of both response ratios and population-level effects, this process provides a library of simulations (and their associated attributes) that could have produced the types of empirical response ratios measured in the Channel Islands.
-
-## Difference-in-Difference Regression {#did .unnumbered}
-
-The DiD model used empirical kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands.
-PISCO conducts visual SCUBA surveys at a large number of rocky reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\ref{fig:ci-map},Fig.\@ref{fig:raw-density}).
-The details of the monitoring program are described in @caselle2015.
-We define the population-level conservation effects of MPAs as the change in mean total biomass densities of targeted fin-fish both inside and outside of MPAs, relative to the mean total biomass densities of targeted fin-fish inside and outside of MPAs that would have occurred without the MPAs.
-
-Building off of the concepts explored in @caselle2015, we used an identification strategy utilizing biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 0])` species that are not directly targeted by fishing as our control group (non-targeted), and biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 1])` species targeted by fishing as our treatment group (see Fig.\@ref(fig:raw-density) for raw trends in targeted and non-targeted biomass densities).
-Targeted species in the Channel Islands include fin-fish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
-Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*).
-Our regression estimates any difference in mean total biomass densities of fin-fish species targeted by fishing effort (i.e., those potentially affected by an MPA) and those species not targeted by fishing before and after MPA implementation.
-To account for the fact that sampling locations are not uniformly distributed across the islands, we weight the samples in our regression in proportion to the total area inside and outside of MPAs in the Channel Islands.
-
-This identification strategy attempts to control for unobserved environmental shocks to the system that are independent of the MPAs.
-Conditional on the assumptions of the model, this regression produces an estimate of the effect of the MPAs on the mean total biomass densities of targeted species throughout the Channel Islands.
-For example, consider an evenly distributed population that has 50% of its range protected by an MPA.
-If the MPA increases biomass densities inside its boundaries by 20%, and by 0% outside the reserve, the population effect of the MPA estimated by our DiD would be 10%.
-
-The DiD regression amounts to estimating the pre-post MPA difference in the biomass densities of targeted species, minus the same difference for non-targeted species in the Channel Islands.
-
-$$(log(D_{MPA=1, T = 1})-log(D_{MPA=0, T = 1}))-(log(D_{MPA = 1,T = 0})-log(D_{MPA =0, T = 0}))$$
-
-Where *T* indicates targeted species, *MPA* indicates whether MPAs are in place, and *D* is the observed mean total biomass density across all sites.
-
- <!-- Our analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*), although our methods could similarly be applied to these species given data.
--->
-
-<!-- The regression amounts to estimating the pre-post MPA difference in the mean total biomass density of targeted species minus the same difference for non-targeted species in the Channel Islands.
--->
-
-The simplified regression form of this model is
-
-\begin{equation}
-  d_{i} \sim Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + \mathbf{B^cX_i} + \mathbf{B^sS_i}},shape)
-\label{eq:did}
-\end{equation}
-
-where $d_i$ is the biomass density at observation *i*, *T* indicates whether the observation *i* is for a targeted ($T = 1$) or non-targeted ($T = 0$) species, and *MPA* marks whether observation *i* is in a pre MPA ($MPA = 0$) or post MPA ($MPA = 1$) state.
-To account for the fact that MPA effects will evolve over time, we estimate a vector of MPA effects in three-year blocks for the all years after the MPAs were implemented in 2003.
-**B^c^** is a vector of coefficients for additional control variables in matrix *X* such as water temperature and observer experience.
-**B^s^** is a vector of hierarchical coefficients for each sampling location **S** , clustered by island.
-Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment (*MPA*) on the treated (targeted species).
-$\beta_0$ is the mean total biomass of non-targeted species pre-MPA, and $\beta_1T$ is the mean total biomass of targeted species pre-MPA.
-$\beta_2MPA$ is the effect of the post-MPA period on the non-targeted species (Table.\@ref(tab:coef-table)).
-We use a Bayesian regression since it allows us to interpret our estimated effects probabilistically [@mcelreath2020].
-Being a Bayesian regression, our DiD model produces posterior probability distributions (the probability distribution of our coefficients conditional on the data, priors, and model assumptions) of our coefficients, from which we can construct Bayesian credible intervals [@gelman2014a].
-
-<!-- r percent(mean(used_data$eventual_mpa))` of the samples in the raw data come from within MPAs, while the MPAs themselves cover 20% of the surface waters of the Channel Islands.
-As such, we weight each observation such that within MPA data are assigned a total weight of 0.2 and outside-MPA data a total weight of 0.8.
-The intercept prior was was determined using the `rstanarm` `autoscale` function.
-For the non-intercept terms, we manually set a prior as a normal distribution with a mean of 0 and a standard deviation of 2.
-This implies that coefficients such as the MPA effect have a prior that provides support for an effect size centered on zero with a range of roughly -4 to 4.
-Since the covariates are centered and scaled, and the dependent variable is on the log-scale, this is an extremely diffuse prior (e.g.
-a one standard deviation change in an independent variable changes the dependent variable by 4 on the log scale) -->
-
-# Results {#results .unnumbered}
-
-<!-- Our Channel Islands-specific simulation provides a set of outcomes that can feasibly bound our expectations for the empirical effects of the MPA network.
-The targeted species in our database span a range of life histories, but are largely made up of fishes in the perch and rockfish complexes, with a mean Von Bertalanffy growth coefficient of `r round(mean(used_data$k[used_data$targeted == 1]),2)` (standard deviation `r round(sd(used_data$k[used_data$targeted == 1]),2)`), and a median age at maturity of `r round(median(used_data$tm[used_data$targeted == 1]),0)` years (standard deviation `r round(sd(used_data$tm[used_data$targeted == 1]),0)`), and a mix of exploitation histories, ranging from low to high (but mostly relatively low) [@alonzo2004;@dick2017].
--->
-
-@caselle2015 found a statistically significant increase in the response ratios of targeted species over time, and evidence that this increase is smaller for non-targeted species.
-Updating the results of @caselle2015 through 2017 shows a continuation of the increasing trend in the response ratios of targeted species (Fig.\@ref(fig:rr-plot)).
-This provides evidence that the Channel Islands MPAs are large enough and sufficiently well-enforced as to provide meaningful protection within their borders.
-
-These response ratios cannot, however, be used as a definitive indicator of population effects of these MPAs.
-In the case of the Channel Islands MPAs, control sites are often located within a few kilometers of an MPA, suggesting that they are susceptible to both local biological spillover and by concentration of fishing effort excluded from the MPAs.
-Our simulation results show that the response ratio trends we observe in the data could plausibly be produced by a wide range of population-level MPA effects (Fig.\@ref(fig:rr-plot)).
-Response ratios well over one were associated with "true" population-level MPA effects generally less than 25%, and, importantly, many simulations produced large response ratios but population-level MPA effects close to 0%.
-This can occur if for example fishing pressure is only moderate, adult movement is low, larval dispersal is high, and displaced fishing effort concentrates around the border of the MPAs.
-
-<!-- For example, if the mean estimated response ratio in the year 2006 (year three after implementation) is one, we found all simulations that had response ratios near one after three years of protection, and then pulled out the "true" simulated population-level MPA effect from each of those simulations.
-This provides us with a distribution of simulated population-level MPA effects that could plausibly generate the types of response ratios actually observed in the Channel Islands.
--->
-
-```{r}
-did_model <- model_runs %>%
-  filter(data_source == "pisco",
-         var_names == "pisco_a",
-         data_to_use == "all") %>% {
-           .$did_fit[[1]]$did_reg
-         }
-did_coefs <- did_model %>%  broom.mixed::tidy(conf.int = TRUE,conf.level = 0.9) %>% 
-  rename(lower = conf.low,
-         upper = conf.high)
-
-```
-
-Our targeted vs.
-non-targeted DiD regression provides an alternative approach to spatial controls for estimating population-level MPA effects.
-Over the first three years of implementation (2003-2006), the effects of the MPAs are unclear, with support for a small negative effect to a substantially positive effect, with much higher probability of a small positive effect (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`,
-90\% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`
-- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`).
-Over the next six years the model estimates greater probabilities of an increasingly positive MPA effect, peaking in 2009-2011 with a median estimate of MPA effect of a `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)` increase in mean total biomass density of targeted species (90% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)`
-- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)`) (Fig.\@ref(fig:mpa-effect)).
-These empirical estimates are in line with the outcomes that our simulation model suggests are plausible.
-While a wide range of MPA effects are plausible, from 0% to upwards of 150%, the median simulated effect size for this stylized version of the Channel Islands after over ten years of protection was `r percent(median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected > 10]))`.
-However, in the subsequent years the trend reverses itself, and in 2015-2017 we once again see no clear effect of the MPAs (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`,
-90\% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`
-- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`.) Fig.\@ref(fig:mpa-effect).
-
-Our library of simulation results allows us to explore how MPA effects, and our ability to measure them, might vary with observation error and natural fish recruitment variation.
-Using a Bayesian DiD regression on these simulated data, we estimated the percent error between the posterior probability distribution of the estimated MPA effect from the regression and the true simulated MPA effect.
-The model struggled severely when MPA effect sizes were less than 25% and the model was faced with both observation and process errors (Fig.\@ref(fig:val-plot)).
-Even models fit to data generated from large effect sizes commonly mis-estimated the true MPA effect by 50% or more.
-Obtaining a mean absolute percent error (MAPE) of 25% or less across our simulated datasets required a "true" population-level MPA effect of at least 30%.
-In the context of the Channel Islands, this finding suggests that we should not be surprised at our difficulty in precisely estimating the population-level effect of the MPAs.
-<!-- This is merely an illustrative exercise, omitting critical factors such as detection probability and sampling strategy.
-However, since nearly any omission which one can think of would make an MPA effect harder to detect, not easier, these results serve as a useful floor for the likely difficulty in estimating MPA effects.
--->
-
-Two of the most critical drivers of MPA effect size are the size of the MPA network and the degree of fishing pressure.
-Looking across these two variables, based on our simulations the MPA network must be large (25% or more of a species' range) and the target species overfished (pre-MPA depletion greater than about 60%) to achieve an effect size with a likely MAPE of 25% or less.
-Although recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, most MPA networks for near-shore commercial fin-fish are likely to cover areas more in line with the Channel Islands (20%) or smaller.
-As such, many MPA networks are expected to have population-level effect sizes that are difficult to detect unless target species were experiencing extreme overfishing prior to MPA establishment (Fig.\@ref(fig:sim-effects)-A).
-
-# Discussion {#discussion .unnumbered}
-
-Containing a carefully designed, well-enforced, and well-studied MPA network, the Channel Islands seems to be an ideal location to study the population-level effects of protected areas.
-High response ratios suggest that despite overall decreases in targeted biomass densities inside and outside MPAs relative to pre-MPA measurements (Fig.\@ref(fig:raw-density)), the MPAs may still be providing protection within their borders.
-But, these spatial response ratios are not necessarily an indicator of broader population-level effects.
-The targeted vs.
-non-targeted DiD strategy presents an alternative means of estimating the population-level effects of MPAs.
-While having its own strict caveats, this strategy provides some potential improvements over spatial response-ratios as a means of estimating population-level effects.
-While we estimate an uncertain but overall positive effect of the MPA network in its first few years, we are unable to detect a robust signal from 2012-2017.
-After 14 years of MPA protection we are left without a clear picture of the population-level effect of the Channel Island MPA network on biomass densities of targeted fin-fish species.
-
-Fishing dynamics may be one factor contributing to a lack of strong MPA network effects.
-Much of the theoretical literature on MPAs assumes that larger reserves produce larger conservation gains [@white2011 and references therein].
-However, these models generally simulate fleet dynamics through fishing mortality rates; the proportion of total mortality experienced by a population attributable to fishing pressure [e.g.
-@halpern2004].
-The implication of this is that total catches will scale with the size of the available population.
-
-```{r}
-positive_outcomes <- outcomes %>% 
-  filter(years_protected == 15) %>% 
-  group_by(fleet_model) %>% 
-  summarise(pp = mean(mpa_effect > 0))
-
-```
-
-Alternatively, under a “constant-catch” strategy, fishers have a catch objective and exert as much (or little) effort as needed to achieve that objective.
-Subsistence fisheries may use a constant-catch style policy over the short-term, as they seek to ensure that their food needs are met.
-Constant-catch dynamics might also occur in fisheries with constraining quotas that are not updated after the implementation of MPAs.
-Fishers pursuing constant-catch in areas outside an MPA may have to fish harder to achieve the same catch from a smaller part of the population, causing a population conservation loss under `r percent(1 - positive_outcomes$pp[positive_outcomes$fleet_model == "constant-catch"], accuracy = 1)` of our constant-catch simulations.
-This potential negative interaction between constant-catch and MPAs is an important risk to consider (as done in @little2011b) especially as MPAs are increasingly implemented in quota-managed fisheries [@liu2018].
-
-While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (Appendix S1 Fig.S2) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
-However, some of the most common targeted species in the Channel Islands, such as blue and copper rockfish, have seen dramatic increases in reported catches in the Santa Barbara region since the year 2010 (although other such as California sheephead have seen declines relative to pre-MPA catch level).
-While we include catch histories in our regression, to the extent that these increases in catches are due to external market forces, the disappearance of clear positive population-level MPA effects estimated by our model could be due to exogenous increases in fishing pressure overwhelming any population-level increases caused by the MPA network.
-
-Another possibility for the estimated decline in Channel Islands MPA effects is environmental disturbance.
-The Channel Islands region experienced a dramatic ‘marine heatwave’ beginning in 2014 and persisting through 2016, resulting in part in extremely elevated water temperatures throughout the region [@gentemann2017].
-Many of the non-targeted species in the Channel Islands have warm thermal affinities and have increased in numbers since the heatwave [@freedman2019].
-The targeted group is made up mostly of fishes with cold-water affinities.
-We hypothesize that the recent evidence for a decline in densities of targeted species is primarily due to altered environmental conditions that disproportionately affected the targeted group.
-The practical result of this is that in the presence of this marine heatwave the non-targeted species may no longer serve as an effective control for the evolution of biomass densities of targeted fin-fish in the absence of the MPAs, given the magnitude of the environmental shock relative to the size of the population-level MPA effect.
-
-<!-- This hypothesis is supported by the fact that we estimate a similar decline in the estimated MPA effect both inside and outside of the MPAs (SI Appendix Fig.S18) (as well as observe the decline in the raw data, Fig.\@ref(fig:raw-density)).
-If the cause of recent declines was due to increases in fishing pressure we would expect to see substantial declines only in the fished areas.
-While we see evidence for the validity of the parallel trends assumption in the early years of our data (Fig.\@ref(fig:mpa-effect), SI Appendix Fig.S26), it is likely that the non-targeted species reacted in a different way to the marine heatwave than the targeted species did, causing a violation of the parallel trends assumption in the post 2014 data.
--->
-
-<!-- Simulation analysis suggests that we should not be surprised by this result.
-The Channel Islands MPAs cover 20% of the surface waters in the Channel Islands, and while formal stock assessments are not available for many of the targeted species in our analysis, what evidence we have suggests that, as a group, these fish are on average not heavily overfished.
-Some species, such as California sheephead and blue rockfish were likely below target levels during the period [@alonzo2004;@dick2017], but projections based upon the overall average response across all species will likely suggest modest benefits even if a subset of species could experience much larger population gains.
-Under such circumstances, the average percentage difference in densities of targeted species with and without MPAs is likely to be modest (Fig.\ref{fig:mpa-effect}).
-Effects of this size will often be challenging to detect empirically given the large natural variation of marine ecosystems (especially temperate reefs) and the observation error inherent in infrequent (e.g.
-annual) common in marine systems.
--->
-
-<!-- Our simulations and difference-in-difference model examine percentage changes in biomass densities (as opposed to changes in total population size).
-An increase in biomass densities from 0.02kg/m^2^ to 0.04kg/m^2^ would be measured as a 100% increase.
-While a large percentage effect, this is a small change in biomass densities relative to the variance in the observation process itself.
-While an MPA providing 25% protection might be expected to produce a 30% increase in biomass densities, expressed as a percentage of unfished biomass, this gain is on the order of 15% (SI Appendix Fig.S45).
-In addition, the median estimated age at sexual maturity for the targeted species included in this study is `r median_tm` years, meaning that the span of this analysis represents less than three generations of MPA protection for half of the measured species.
-Ongoing monitoring may yet reveal clearer effects, especially if the time series extends beyond the periods of environmental shocks.
-Analysis of more rapidly growing and maturing species, e.g.
-spiny lobster, may also reveal clearer signals.
--->
-
-Trophic cascades are another possible explanation for our equivocal results.
-All of the species in this empirical analysis may affect each other through mechanisms such as predation, competition, and habitat modification.
-We used convergent cross mapping (CCM), in the manner of @clark2015, to test for significant dynamic interactions between species and therefore the possibility of the trophic cascades biasing our results.
-We found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (SI Fig.S42).
-However, the longer MPAs are in place, the greater the possibility that substantial species interactions that can affect use of non-targeted species as a control may arise.
-
-<!-- Our finding of small MPA effects in recent years is therefore likely a result of both changing fishing dynamics and environmental disturbance. Given the  -->
-
-<!-- The MPA literature highlights a large number of variables beyond simply size and fishing pressure that can affect performance.
-We examined the variability in expected MPA effects across each of the two major axes (pre-MPA depletion and MPA size).
-As both MPA size and pre-MPA depletion increase, the potential for a large MPA effect increases.
-However, for both variables even at extremely large or extremely small values a wide range of MPA effects were possible (though as we might expect the effect of pre-MPA depletion, in other words fishing pressure, was much clearer than MPA size alone) (Fig.\@ref(fig:sim-effects)-B).
--->
-
-<!-- MPAs are an important part of the marine resource management toolbox.
-Under ideal circumstances they can protect individual species and ecosystem linkages, while supporting local economies through tourism and fishing opportunities.
-One rationale for the expansion of MPAs is that they will deliver net conservation benefits both inside *and outside* their borders.
-To date, this assumption is insufficiently tested, and this is the focus of our paper.
-Our results show that population-level conservation benefits of MPAs are highly context dependent and in many circumstances, are likely to be small enough that they are nearly impossible to detect empirically.
-This is exactly what we found in our empirical case study from the relatively large and well-studied network of MPAs in the Channel Islands, California, USA.
--->
-
-<!-- A large theoretical literature has pointed out the importance of a wide range of economic and environmental factors in determining the effectiveness of MPAs [e.g.
-@hilborn2004; @smith2006a; @white2010c].
-Our simulation model allows users to place this collective theory into context, clarifying under what states of the world we might expect to see particular effect sizes, and demonstrating what effect sizes we might reasonably expect to be detectable in the real world.
-In the case of the Channel Islands, simulation modeling suggests for the network in question, an effect size on the order of 30% or smaller is likely, which our analysis in turn demonstrates may be challenging to detect even with a comprehensive monitoring program.
--->
-
-<!-- Once simulation models have been used to help design an MPA, how can users evaluate whether it is achieving their objectives?  -->
-
-<!-- @lester2009 found massively higher densities and biomass inside MPAs than outside [median biomass increases of  over 400%].
-But, when MPAs affect nearby control sites used in response ratios through biological spillover or concentration of fishing effort, it is entirely possible to for MPA to produce massive response ratios while simultaneously having minimal effects on the entire population partially protected by the MPAs (and *vice versa*).
--->
-
-MPAs are an important part of the marine resource management toolbox.
-As the number and size of global MPA networks increase, we must set appropriate expectations for their outcomes on both local and regional scales.
-Our results for the relatively large, well-enforced, and rigorously studied (though also relatively lightly fished) Channel Islands Marine Protected Area network make clear that in many instances empirically detecting a clear effect of MPAs on total biomass densities of targeted fin-fish at the population level may not be possible.
-In the absence of clear empirical findings, simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can be tuned to focus on the species groups that have the highest chance of a detectable effect size over the early years of the reserve [@nickols2019].
-Expanding data collection to include robust monitoring of spatio-temporal fleet dynamics may help assess the validity of control sites used in response ratios, support the direct inclusion of these fleet dynamics into statistical models, and allow managers to take into account potential negative interactions between MPAs and fleet dynamics such as those that may occur under constant-catch dynamics.
-Whenever possible monitoring programs should be implemented prior to MPA implementation to provide a pre-treatment benchmark.
-
-There are many potential alternatives to spatial response ratios for estimating the population effects of MPAs that better account for the challenges of causal inference (though that may be more data-intensive) [@larsen2019].
-We applied one such approach here, and yet were still unable to reach robust conclusions as to the effect of MPAs on the total biomass density of targeted fin-fish in the Channel Islands, due to the likely small size of the true effect relative to the influence of environmental variability.
-There are other promising statistical approaches to setting expectations for MPA effects, including using models fitted to local data to set population-level expectations and create synthetic counterfactuals [@nickols2019; @white2011].
-
-However, as the scientific community continues to develop appropriate measures of population-level MPA effects, we must also effectively communicate the challenges of estimating those effects.
-In light of these results, lack of a clear population-level MPA effect should not necessarily be viewed as a failure of a conservation program, and likewise, large response ratios should not be automatically taken as evidence of a population-level conservation success.
-Rather, results and subsequent management actions must be considered in the context of reasonable expectations given the size, age, and degree of enforcement of the MPAs in question, together with the ecological and economic dynamics of a given system.
-
-<!--Stock assessment models that account for the population within the MPAs [@field2006a] may be able to answer the management relevant question of whether fishing mortality rates and biomass levels from systems containing MPAs are in-line with management objectives.However, such an approach does not necessarily shed light on whether the MPAs themselves caused the estimated state of the population, and of course are highly data intensive, potentially restricting our ability to provide stock-assessment based inference of MPA outcomes for broad arrays of targeted species.
-@nickols2019 and @white2011 provide excellent examples of using models fit to local data to set expectations for the outcomes of MPAs.
-In the manner of @abadie2010], models such as these that have been demonstrated to be reliable predictors of the true dynamics of the system in the pre-MPA period can be used as an estimate for plausible outcomes in a world without MPAs in the post-MPA period and subsequently as a strategy for estimating the population-level effects of MPAs.
-In contrast, we use our simulation model to set expectations but use an empirical difference-in-difference approach to construct our counterfactual of likely outcomes of the Channel Islands in the absence of MPAs.
-Both approaches can help address the problem of assessing the population-level effects of MPAs.-->
-
-<!--Non-equilibrium analyses also help set expectations for effect sizes over time (@nickols2019;@kaplan2019;@ovando2016;Fig.\ref{fig:mpa-effect}).
-Educating communities about the challenges of estimating the effects of MPAs can help ensure that a lack of a clear effect is not necessarily viewed as a failure of the program, or large positive result based solely on response ratios as a clear sign of success and subsequent relaxation of other fishery management strategies.
-Rather, results and subsequent management actions must be considered in the context of reasonable expectations given the size, age, and degree of enforcement of the MPAs in question, together with the ecological and economic dynamics of a given system.
-While this paper has focused on the conservation outcomes of MPAs, future work must also address the challenge of predicting and estimating the fishery impacts of protected areas, an important question for which extremely little empirical evidence is available [@smith2006a].-->
-
-As advocacy for broad networks of MPAs is growing globally, MPA science must directly tackle the challenge of evaluating the performance of MPAs at the population scale.
-Commonly employed metrics such as spatial response ratios may be applicable in some circumstances, but are vulnerable to inaccuracy or misuse as metrics of population-level effects.
-Bio-economic modeling can help frame community expectations, reducing the potential for a reduction in support if unrealistic conservation or fishery expectations are not realized.
-Statistical approaches that explicitly address complications such as the spatial spillover effects of MPAs may give users an improved understanding of the performance of their MPAs, but even they may struggle when expected effect sizes are small.
-Clearly communicating what we should expect, and what we can detect, from MPAs is critical to ensuring that MPAs play effective roles in fisheries management and marine conservation.
-
-<!-- Dependence on unreliable estimators of MPA effects may lead to stakeholders incorrectly attributing negative environmental shocks as MPA failures, or interpreting data arising from scorched-earth fishing outside MPAs as a conservation success.
-Both of these scenarios would hinder the ability of MPAs to serve as effective marine resource management tools at scale.
--->
-
-# Supporting Information  {#si .unnumbered}
-
-Detailed descriptions of the simulation model and DiD estimator (Appendix S1) are available online.
-The authors are solely responsible for the content and functionality of these materials.
-Queries (other than absence of the material) should be directed to the corresponding author.
-
-# Literature Cited {#lit .unnumbered}
-
-<div id="refs"></div>
-
-\newpage
-
-# Tables {#tables .unnumbered}
 
@@
 
-knitr::kable(equations,
-             escape = FALSE,
-             caption = "Variables for MPA simulations",
-             format = "pipe",
-             align = "l") %>% 
-  kableExtra::kable_styling()
+a = flextable(equations) %>% 
+  set_caption("Test")
+
+  autofit(a)
+
+# knitr::kable(equations,
+#              escape = FALSE,
+#              caption = "Variables for MPA simulations",
+#              format = "simple", 
+#              align = "c")
 
 ```
-
-\newpage
-
-```{r coef-table, echo=FALSE}
-
-coefs <- did_fit %>%
-  broom.mixed::tidy(conf.int = TRUE, prob = 0.9) %>%
-    rename(lower = conf.low,
-         upper = conf.high) %>% 
-  select(term, estimate, lower, upper) %>%
-  rename(
-    Mean = estimate,
-    `5th Percentile` = lower,
-    `95th Percentile` = upper
-  )
-
-coefs$term <- str_remove_all(coefs$term, "var_")
-
-coefs$term <- str_replace_all(coefs$term, "targeted","Targeted by Fishing")
-
-coefs$term <- str_replace_all(coefs$term, "year_bins","Year Block ")
-
-coefs$term <- str_replace_all(coefs$term, ":","$\\\\times$")
-
-coefs <- coefs %>% 
-  mutate(term = case_when(term == "tex" ~ "Observer Experience",
-                          term == "tex_2" ~ "(Observer Experience)$^2$",
-                          term == "temp" ~ "Water Temperature",
-                          term == "regional_temp_dev" ~ "Deviation from regional mean temperature",
-                          term == "regional_temp_dev_2" ~ "(Deviation from regional mean temperature)$^2$",
-                          term == "lag_catch" ~ "Lagged Commercial Catch",
-                          term == "targeted" ~ "Targeted by Fishing",
-                          term == "(Intercept)" ~ "Intercept",
-                          term == "surge" ~ "Wave Surge",
-                          term == "kelp" ~ "Kelp Cover",
-                           TRUE ~ term))  %>% 
-  rename(Variable = term)
-
-  knitr::kable(
-    coefs,
-    digits = 2,
-    caption = glue::glue("Posterior means and credible interval for key model coefficients.
-Raw data contained {nrow(used_data)} individual observations, total number of targeted and non-targeted mean mean total biomass densities by site and year used to fit model was {nrow(did_fit$data)}.
-Site-level terms omitted for clarity."),
-    format = "pipe", 
-    escape = FALSE,
-    align = "l")
-
-```
-
-\newpage
-
-# Figures {#figures .unnumbered}
-
-```{r ci-map, include = TRUE,fig.width=width_1col, fig.asp=1.5, fig.cap="Map of study region; the Northern Channel Islands, California, USA.
-Shading shows binned number of PISCO sampling events across the time period of our study.", include=TRUE}
-
-temp_plot <- sample_location_plot + 
-  theme(axis.text = element_blank())
-
-temp_plot
-ggsave(
-  temp_plot,
-  filename = here::here("documents","figs","sample_location_plot.pdf"),
-  width = width_1col,
-  height = width_1col * asp_1col,
-  device = cairo_pdf
-)
-
-```
-
-\newpage
-
-```{r raw-density, fig.width = width_1col, fig.asp=asp_1col,fig.cap = "Centered and scaled trends in biomass densities of targeted and non-targeted fin-fish included in our study.
-Top panel shows mean trends across all sites, with shaded areas shwoing 95% confidence intervals.
-Bottom two panels show same trends but only including sites outside and inside MPAs"}
-
-temp_plot <-
-  (total_trend_plot / (mpa_trend_plot)) & theme( plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), units = "lines"),
-                                                panel.grid = element_blank()
-  )
-
-temp_plot
-
-ggsave(
-  temp_plot,
-  filename = here::here("documents","figs","raw_biomass_density_plot.pdf"),
-  width = width_1col,
-  height = width_1col * asp_1col,
-  device = cairo_pdf
-)
-
-```
-
-\newpage
-
-```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="Simulated population effects (grey) that could have produced observed response ratios (black) in the Channel Islands.
-Black histograms show 90% Posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis).
-Simulated population-level (pop.) effect on biomass densities matched to empirical response ratios in light grey.
-For response ratios, a value of zero indicates that biomass densities of targeted species are identical inside and outside MPAs, a value of one that biomass densities of targeted species are 100% greater inside MPAs relative to outside.
-For MPA population effect, a value of zero indicates that biomass densities are identical in the with- and without- MPA scenarios.
-A value of 1 indicates that biomass densities are 100% greater in the scenario with MPAs than the scenario without MPAs."}
-
-temp_plot <- response_ratio_plot + 
-  theme_1col + 
-  theme(axis.title.x = element_textbox(size = 10),
-        legend.position = "top") + 
-  geom_vline(aes(xintercept = 0), color = "black", linetype = 2)
-
-ggsave(
-  temp_plot,
-  filename = here::here("documents","figs","response_ratio_plot.pdf"),
-  device = cairo_pdf,
-  width = width_1col,
-  height = width_1col * asp_1col,
-  units = "in",
-)
-
-temp_plot
-# pdf(file =  "figs/density-ratio.png",width = 8, height = 5, useDingbats = TRUE)
-# print(dr_plot)
-# dev.off()
-
-```
-
-\newpage
-
-```{r mpa-effect, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Results of difference-in-difference regression estimating the population effect of the Channel Island MPAs on mean total biomass densities of targeted species.
-Grey distributions show posterior probability distribution of estimated MPA effect; red point is median estimated effect, thicker red section 50% credible interval, thinner red line 90% credible interval.
-Blue distributions in background show range of MPA effects produced by simulation model tuned to reflect the dynamics of the Channel Island MPAs (black dashed line is median simulated value).
-Results are estimated in blocks of three years, including years greater than or equal to left-hand value and less than right-hand value."}
-
-mpa_effect_plot
-
-ggsave(
-  mpa_effect_plot,
-  filename = here::here("documents","figs","mpa_effect_plot.pdf"),
-  width = width_2col,
-  height = width_2col * asp_2col,
-  device = cairo_pdf
-)
-
-```
-
-\newpage
-
-```{r val-plot, include = TRUE,fig.width = width_1col, fig.asp=asp_1col, fig.cap = "Distribution of percent error in posterior estimates of population-level MPA effect (y-axis) plotted against true simulated MPA effect (x-axis).
-Shading shows concentration of simulations.
-Black line shows mean absolute percent error (MAPE) as a function of true simulated MPA effect."}
-
-temp_plot <- validation_plot + 
-  theme_1col + 
-  theme(legend.position = "top") #+ 
-  # scale_fill_viridis(option = "C",
-  #                    name = "Sims(#)",
-  #                    guide = hgc <- guide_colorbar(frame.colour = "black",
-  #                                                  ticks.colour = "black",
-  #                                                  barwidth = ggplot2::unit(5,"lines")))
-
-temp_plot
-ggsave(
-  temp_plot,
-  filename = here::here("documents","figs","validation_plot.pdf"),
-  width = width_1col,
-  height = width_1col * asp_1col,
-  device = cairo_pdf
-)
-
-```
-
-\newpage
-
-```{r sim-effects, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Simulated population-level effects of MPAs.
-Panel a) shows median simulated population-level (pop.) MPA effect sizes as a function of percent of species' range inside MPA (x-axis), and pre-MPA depletion (y-axis).
-Pre-MPA depletion is a measure of fishing pressure, where 0 means that the population is unfished, and 1 means that the population is extinct in the time period immediately prior to MPA implementation.
-Panel b) shows distribution of simulations across range of MPA size and pre-MPA depletion separately."}
-
-expected_mpa_effect_plot
-ggsave(
-  expected_mpa_effect_plot,
-  filename = here::here("documents","figs","expected_mpa_effect.pdf"),
-  width = width_2col,
-  height = width_2col * asp_2col,
-  device = cairo_pdf
-
-)
-
-```

### 2020-12-01 make-pop-effects-of-mpas.R

@@ -3248,7 +3248,7 @@ pop_depletion_plot <- outcomes %>%
                      expand = expansion(0, 0))
 
 expected_mpa_effect_plot <-
-  (pop_depletion_and_size_plot + labs(title = "A")) + ((pop_size_plot + labs(title = "B")) / pop_depletion_plot)  + plot_layout(widths = c(1.5, 1)) &
+  (pop_depletion_and_size_plot + labs(title = "a)")) + ((pop_size_plot + labs(title = "b)")) / pop_depletion_plot)  + plot_layout(widths = c(1.5, 1)) &
   theme(
     plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), units = "lines"),
     axis.text.x = element_text(size = 8),
@@ -4488,7 +4488,7 @@ implications %>%
 
 
 # title = "<span style = 'color:grey;'> Paired Simulated Pop. Effect / <span style = 'color:black;'>Empirical Response Ratio</span>"
-
+browser()
 response_ratio_plot <-   targ_rr %>%
   select(year, response_ratio) %>%
   rename(mpa_effect = response_ratio) %>%
@@ -4523,9 +4523,10 @@ response_ratio_plot <-   targ_rr %>%
     labels = c("Simulated Pop. Effect", "Empirical Response Ratio"),
     name = element_blank()
   ) +
-  theme(legend.position = "top",
-        axis.title.x = element_textbox_simple(valign = 1.75,
-                                              halign = 1))
+  theme(legend.position = "top")
+  
+              # axis.title.x = element_textbox_simple(valign = 1.75,
+              #                                 halign = 1))
 
 response_ratio_plot
 

### 2020-12-01 documents/ovando-et-al-pop-effects-of-mpas.Rmd
a couple of tiny edits

@@
 A network of MPAs was put in place in The Channel Islands National Marine Sanctuary in 2003, with a goal of providing regional conservation and fishery benefits.
-Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level using a Bayesian difference-in-difference appreoach.
+Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level using a Bayesian difference-in-difference approach.
 We show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); the size of many MPA networks suggests that effect sizes may often be smaller than 30%.
@@
 The simulation then begins to apply the fleet model to the population, and in a randomly selected year implements an MPA network.
-The model samples a percentage of the population's range to place in MPAs, and randomly assigns patches to MPAs either across across a uniform system or preferentially on higher-quality habitat.
+The model samples a percentage of the population's range to place in MPAs, and randomly assigns patches to MPAs either across a uniform system or preferentially on higher-quality habitat.
 The model then randomly selects whether fishing effort that used to operate inside the MPAs is redistributed to the areas outside the MPAs, or leaves the fishery entirely.
@@
 The DiD model used empirical kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands.
-PISCO conducts visual SCUBA surveys at a large number of rocky reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\ref{fig:ci-map},Fig.\@ref{fig:raw-density}).
+PISCO conducts visual SCUBA surveys at a large number of rocky reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\@ref{fig:ci-map},Fig.\@ref{fig:raw-density}).
 The details of the monitoring program are described in @caselle2015.

### 2020-12-01 documents/ovando-et-al-pop-effects-of-mpas.Rmd
mergin in owen edits

--- 2020-12-01 documents/ovando-et-al-pop-effects-of-mpas.Rmd
@@
 
-theme_1col <- hrbrthemes::theme_ipsum(base_size = 11,
-                                      axis_title_size = 12)
+theme_1col <- hrbrthemes::theme_ipsum(base_size = 8,
+                                      axis_title_size = 6,
+                                      plot_title_size = 10,
+                                      strip_text_size = 8)
 
-theme_2col <- hrbrthemes::theme_ipsum(base_size = 12,
-                                        axis_title_size = 14)
+theme_2col <- hrbrthemes::theme_ipsum(base_size = 10,
+                                        axis_title_size = 12,
+                                      plot_title_size = 12)
 
@@
 Our methods consist of a spatially-explicit bio-economic simulation model and a Bayesian DiD regression [@angrist2009;@mcelreath2020].
-DiD is a regression-based approach to BACI, assessing changes in control and treatment groups before and after treatment [@larsen2018].
+DiD is a regression-based approach to BACI, assessing changes in control and treatment groups before and after treatment [@larsen2019].
 We use the term DiD here to more closely tie our results to the terminology in causal-inference literature.
@@
 temp_plot <-
-  (total_trend_plot / (mpa_trend_plot)) & theme( plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), units = "lines"),
-                                                panel.grid = element_blank()
-  )
+  ((total_trend_plot + labs(subtitle = "a) Channel Islands") + theme_1col) / (mpa_trend_plot + theme_1col + guides(color = FALSE, fill = FALSE, linetype = FALSE))) & theme(plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), units = "lines"),
+                         panel.grid = element_blank(),
+                         legend.position = "top",
+                         panel.spacing = unit(.5, "lines"))
 
@@
   theme(axis.title.x = element_textbox(size = 10),
-        legend.position = "top") + 
+        legend.position = "top",
+        panel.spacing.y = unit(1, "lines")) + 
   geom_vline(aes(xintercept = 0), color = "black", linetype = 2)
@@
 
-mpa_effect_plot
+mpa_effect_plot + theme_2col + theme(legend.position = "top")
 

### 2020-12-01 make-pop-effects-of-mpas.R

@@ -4368,8 +4368,8 @@ total_trend_plot <- targeted_trends %>%
   theme(legend.position = "top")
 
 
-targlab <- c(`TRUE` = "Inside MPAs",
-             `FALSE` = 'Outside MPAs')
+targlab <- c(`TRUE` = "c) Inside MPAs",
+             `FALSE` = 'b) Outside MPAs')
 
 
 mpa_trend_plot <- targeted_trends_by_mpa %>%
@@ -4488,7 +4488,12 @@ implications %>%
 
 
 # title = "<span style = 'color:grey;'> Paired Simulated Pop. Effect / <span style = 'color:black;'>Empirical Response Ratio</span>"
-browser()
+
+lab <- c(
+  biased = "b) Paired Simulated Population-Level MPA Effect",
+  `Response Ratio` = "a) Empirical Response Ratio"
+)
+
 response_ratio_plot <-   targ_rr %>%
   select(year, response_ratio) %>%
   rename(mpa_effect = response_ratio) %>%
@@ -4509,7 +4514,7 @@ response_ratio_plot <-   targ_rr %>%
     stat = "binline",
     color = "black",
     size = .1,
-    show.legend = TRUE,
+    show.legend = FALSE,
     bins = 20
   ) +
   scale_x_continuous(name = "Percent Difference", limits = c(NA, 2), labels = percent) +
@@ -4523,7 +4528,10 @@ response_ratio_plot <-   targ_rr %>%
     labels = c("Simulated Pop. Effect", "Empirical Response Ratio"),
     name = element_blank()
   ) +
-  theme(legend.position = "top")
+  coord_flip() + 
+  facet_wrap(~source, nrow = 2, labeller =  labeller(source = lab), as.table = FALSE) + 
+  theme(legend.position = "top", panel.spacing.y = unit(.1, "lines"))
+  
   
               # axis.title.x = element_textbox_simple(valign = 1.75,
               #                                 halign = 1))

### 2020-12-02 documents/referee-replies.Rmd
fixed that damn conservation-letters.csl error

@@
 
-Our citations are automatically formatted according to the citation style language file provided by Conservation Biology, so we have left the order as dictated by the software, unless the editors ask otherwise.
+Thank you for catching this, there was an error in the conservation biology citation style language file we were using, which we have corrected. Citations are now in chronological order.
 

### 2020-12-02 documents/ovando-et-al-pop-effects-of-mpas-si.Rmd
fixed that damn conservation-letters.csl error

@@
   pdf_document: default
-bibliography: references.bib
+bibliography: references.yaml
+csl: conservation-biology.csl
 header-includes:

### 2020-12-02 documents/ovando-et-al-pop-effects-of-mpas.Rmd
fixed that damn conservation-letters.csl error

@@
 Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level using a Bayesian difference-in-difference approach.
-We show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); the size of many MPA networks suggests that effect sizes may often be smaller than 30%.
+We show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); smaller effect sizes (which we find are more common) are even harder to detect.
 Our results provide a novel assessment of the population-level effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.
@@
 output:
+  bookdown::word_document2:
+    reference_docx: template.docx
   bookdown::pdf_document2:
     keep_tex: true
-  bookdown::word_document2:
-    reference_docx: template.docx
+    latex_engine: xelatex
   bookdown::html_document2: default
@@
 Numerous factors influence how MPAs affect fish populations.
-These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003; @botsford2008], the strength, timing, and location of density dependence [@burgess2014], the design of the network [@gaines2010; @costello2010a],the degree of enforcement [@edgar2014], the level of fishing pressure, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@walters2000a;@white2011;@ovando2016;@moffitt2013].
+These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003; @botsford2008], the strength, timing, and location of density dependence [@burgess2014], the design of the network [@gaines2010;@costello2010a],the degree of enforcement [@edgar2014], the level of fishing pressure, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@hastings2003;@gerber2003;@hilborn2004a;@gaines2010;@botsford2003;@smith2003a;@walters2000a;@white2011;@ovando2016;@moffitt2013].
 
@@
 These largely theoretical results are generally based on modeling of closed populations with some fraction protected inside of MPAs.
-In contrast to this population paradigm used in MPA simulations, MPAs are often evaluated empirically at local scales using spatial "response ratios", commonly measured as the ratio of biomass densities (weight of organisms per unit area) of species inside relative to selected control sites outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015].
+In contrast to this population paradigm used in MPA simulations, MPAs are often evaluated empirically at local scales using spatial "response ratios", commonly measured as the ratio of biomass densities (weight of organisms per unit area) of species inside relative to selected control sites outside of MPAs [@lester2009;@halpern2003a;@edgar2014;@caselle2015].
 These studies found clear evidence that well enforced and sufficiently sized MPAs are associated with high response ratios.
@@
 The DiD model used empirical kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands.
-PISCO conducts visual SCUBA surveys at a large number of rocky reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\@ref{fig:ci-map},Fig.\@ref{fig:raw-density}).
+PISCO conducts visual SCUBA surveys at a large number of rocky reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\@ref(fig:ci-map),Fig.\@ref(fig:raw-density)).
 The details of the monitoring program are described in @caselle2015.
@@
 Using a Bayesian DiD regression on these simulated data, we estimated the percent error between the posterior probability distribution of the estimated MPA effect from the regression and the true simulated MPA effect.
-The model struggled severely when MPA effect sizes were less than 25% and the model was faced with both observation and process errors (Fig.\@ref(fig:val-plot)).
+The error in the DiD model's estimate of the population-level MPA effect was extremely high when MPA effect sizes were less than 25% and the model was faced with both observation and process errors (Fig.\@ref(fig:val-plot)).
 Even models fit to data generated from large effect sizes commonly mis-estimated the true MPA effect by 50% or more.
@@
 temp_plot <-
-  ((total_trend_plot + labs(subtitle = "a) Channel Islands") + theme_1col) / (mpa_trend_plot + theme_1col + guides(color = FALSE, fill = FALSE, linetype = FALSE))) & theme(plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), units = "lines"),
+  ((total_trend_plot + labs(subtitle = "a) Channel Islands") + theme_1col + theme(plot.subtitle = element_text(size = 8))) / (mpa_trend_plot + theme_1col + guides(color = FALSE, fill = FALSE, linetype = FALSE))) & theme(plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), units = "lines"),
                          panel.grid = element_blank(),
@@
         legend.position = "top",
-        panel.spacing.y = unit(1, "lines")) + 
-  geom_vline(aes(xintercept = 0), color = "black", linetype = 2)
+        panel.spacing.y = unit(1, "lines"))
 

### 2020-12-02 documents/test-tables.Rmd
fixed that damn conservation-letters.csl error

@@
 ---
-title: "test-table"
-author: "Dan Ovando"
-date: "11/30/2020"
+title: Assessing the Population-Level Conservation Effects of Marine Protected Areas
+abstract: |
+  Marine Protected Areas (MPAs) cover 3-7% of the world’s ocean, up from less than 1% in the year 2000, and international organizations call for 30% by 2030.
+The premise underlying MPA expansion is that they conserve biodiversity, habitats, and fished populations.
+While numerous studies show that MPAs produce conservation benefits *inside* their borders, many MPAs are also justified on the grounds that they confer conservation benefits to the broader population *beyond* their borders.
+We examine the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and show that even in cases where the population benefits are large, they are inherently difficult to detect empirically.
+A network of MPAs was put in place in The Channel Islands National Marine Sanctuary in 2003, with a goal of providing regional conservation and fishery benefits.
+Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level using a Bayesian difference-in-difference approach.
+We show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); smaller effect sizes (which we find are more common) are even harder to detect.
+Our results provide a novel assessment of the population-level effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.
+bibliography: references.yaml
+csl: conservation-biology.csl
 output:
+  bookdown::html_document2: default
   bookdown::word_document2:
     reference_docx: template.docx
+  bookdown::pdf_document2:
+    keep_tex: true
+    latex_engine: xelatex
+params:
+  run_name: ["v6.0"]
+  sim_years: [50]
+  burn_years: [20]
+  num_patches: [25]
+linkcolor: blue
+toc: false
+header-includes:
+  - \usepackage{setspace}\doublespacing
+  - \usepackage{lineno}\linenumbers
+  - \usepackage{draftwatermark}
+  - \SetWatermarkText{DRAFT}
+always_allow_html: true
 ---
@@
 
-```{r sim-table, echo=FALSE}
-variable <- c(
-"Scientific Name",
-"steepness",
-"Adult movement ($\\sigma_{s=a}$)",
-"Larval movement ($\\sigma_{s=l}$)",
-"Recruitment variation ($\\sigma_{r}$)",
-"Recruitment autocorrelation ($ac_{r}$)",
-"Density-dependent adult movement",
-"Density-dependence timing ",
-"% Patches in MPA",
-"Initial fishing relative to natural mortality",
-"Selectivity as a multiple of maturity length",
-"Fleet model",
-"Spatial effort model",
-"Years into simulation to start MPA",
-"MPA is Larval Source?",
-"Randomly place MPA?",
-"Fleet reaction to MPA",
-"Patchiness",
-"MPA habitat factor")
-
-distribution <- 
-c(
-"Drawn from all possible species in `FishLife` (@thorson2017c)",
-"uniform(0.6,0.95)",
-"uniform(0,0.25 * P)",
-"uniform(0,0.25 * P)",
- "$\\in{0,0.05,.1,.2}$",
- "$\\in{0,0.05,.1,.2}$",
- "$\\in{0.25,1}$",
- "$\\in{Local, Global,PostDispersal}$",
- "uniform(0.01,1)",
- "uniform(0.01,4)",
- "uniform(0.1,1.25)",
-"$\\in{OpenAccess, ConstantEffort,ConstantCatch}$",
- "$\\in{Uniform, Biomass,Profits}$",
- "uniform(5,0.66)",
-"$\\in{TRUE,FALSE}$",
-"$\\in{TRUE,FALSE}$",
-"$\\in{Concentrate,Leave}$",
-"uniform(0.01,0.75)",
-"$\\in{1, 4}$"
-)
-
-equations <- tibble(Variable = variable, Distribution = distribution )
-
-a = flextable(equations) %>% 
-  set_caption("Test")
-
-  autofit(a)
-
-# knitr::kable(equations,
-#              escape = FALSE,
-#              caption = "Variables for MPA simulations",
-#              format = "simple", 
-#              align = "c")
-
-```
+Hmmm [@hastings2003;@gerber2003;@botsford2003;@smith2003a;@hilborn2004a;@gaines2010;@white2011;@moffitt2013;@ovando2016;@walters2000a]

### 2020-12-02 make-pop-effects-of-mpas.R

@@ -33,7 +33,6 @@ library(tidyverse)
 library(ggtext)
 library(ggrepel)
 library(rnaturalearth)
-library(Cairo)
 options(dplyr.summarise.inform = FALSE)
 
 # library(rnaturalearthhires)
@@ -4339,10 +4338,12 @@ targeted_trends_by_mpa %>%
   facet_wrap(~eventual_mpa)
 
 
+
+
 total_trend_plot <- targeted_trends %>%
   ggplot() +
   geom_vline(aes(xintercept = 2003), linetype = "dotted", size = 1.1) +
-  geom_segment(x = 2006, xend = 2003.25, y = 4, yend = 4, arrow =arrow(length = unit(0.2,"cm"))) +
+  geom_segment(data = data.frame(x = 2006, xend = 2003.25, y = 4, yend = 4), aes(x = x, xend = xend, y = y, yend = yend), arrow =arrow(length = unit(0.2,"cm"))) +
   geom_ribbon(aes(
     year,
     ymin = scaled_mbd - 1.96,
@@ -4354,7 +4355,7 @@ total_trend_plot <- targeted_trends %>%
     labels = c("Targeted", 'Non-Targeted'),
     name = ""
   ) +
-  geom_text(x = 2008.5, y = 4, label = "MPAs Implemented", nudge_x = 5, size = 2) +
+  geom_text(data = data.frame(x = 2008, y = 4), aes(x = x, y = y), label = "MPAs Implemented", nudge_x = 1, size = 3) +
   geom_line(aes(year, scaled_mbd, color = targeted == 0, linetype = targeted == 0),
             size = 1) +
   scale_fill_manual(
@@ -4499,10 +4500,11 @@ response_ratio_plot <-   targ_rr %>%
   rename(mpa_effect = response_ratio) %>%
   mutate(source = "Response Ratio",
          mpa_effect = mpa_effect - 1) %>%
-  bind_rows(biased_implication) %>%
+  bind_rows(biased_implication %>% mutate(mpa_effect = pmin(0.5, mpa_effect))) %>%
+  # filter(source == "biased") %>% 
   # filter((year - 2000) %% 3 == 0) %>%
   ggplot() +
-  geom_vline(aes(xintercept = 0), color = "red", linetype = 2) +
+  geom_vline(aes(xintercept = 0), color = "black", linetype = 2) +
   ggridges::geom_density_ridges(
     aes(
       mpa_effect,
@@ -4510,14 +4512,14 @@ response_ratio_plot <-   targ_rr %>%
[…]

### 2020-12-03 documents/referee-replies.Rmd
just about ready

@@
 subtitle: "Reviewer Replies"
+bibliography: references.yaml
+csl: conservation-biology.csl
 output:
-  bookdown::pdf_document2: default
+  bookdown::pdf_document2:
+    keep_tex: true
+  bookdown::word_document2:
+    reference_docx: template.docx
+  bookdown::html_document2: default
 ---
@@
 
-We thank the editors and referees for their helpful comments.
-The most pressing requests were in our mind to a) specify the specific gap in the literature filled by our work and b) clarify the methods with regards to both our simulation model and the difference-in-difference equations.
+We thank the editors and reviewers for their helpful comments.
+The most pressing requests were in our mind to a) more clearly specify the gap in the literature filled by our work and b) clarify the methods with regards to both our simulation model and the difference-in-difference equations.
 In order to address these concerts, we have
 
-1. Clarified the specific gap that our paper fills in the literature, and how the empirical approach here builds off of methods such as before-after-control-impact (BACI) designs
+1. Clarified the specific gap that our paper fills in the literature, and how the empirical approach here builds off of methods such as before-after-control-impact (BACI) designs.
 
-2. Included a clearer description of our simulation and estimation methods, and reorganized the Supporting Information to allow readers to more clearer find detailed equations.
+2. Included a clearer description of our simulation and estimation methods, and reorganized the Supporting Information to help readers find our simulation model equations.
 
@@
 As such we have not included as in-depth responses in the text of the manuscript as we might otherwise, but we hope our brief responses address the issues identified by the referees.
-However, we understand that it is our responsibility to ensure that we address all comments are sufficiently addressed within the word limits, so we will of course work to accommodate any outstanding issues.
+However, we understand that it is our responsibility to ensure that all comments are sufficiently addressed within the word limits, so we will of course work to accommodate any outstanding issues.
 
-We provide detailed responses to each of the editors' and referees' comments below. Editor's or referee's comments are *italicized*, and our responses are in plain text.
+We provide detailed responses to each of the editors' and referees' comments below. Editor or reviewer comments are *italicized*, and our responses are in plain text.
 
@@
 
-We thank you for the constructive comments, and hope that the revised version of the paper has sufficiently address the helpful points raised by each of the referees.
-In particular, we have sought to clarify the general contributions of this paper in the introduction.
+We thank you for the constructive comments, and hope that the revised version of the paper has sufficiently address the helpful points raised by each of the reviewers.
+In particular, we have sought to clarify the general contributions of this paper in the introduction and discussion.
 We thank you for your consideration and look forward to hearing your response.
@@
 
-Thank you for this clear synthesis of the helpful feedback provided by the referees.
-We have addressed each of their comments in the text, and outlined our responses below.
+Thank you for this clear synthesis of the helpful feedback provided by the reviewers We have addressed each of their comments in the text, and outlined our responses below.
 In particular, we have included more details about the simulation model and Bayesian Difference-in-Difference in the methods, ensured that all results are in the results section, and made clearer the specific objectives of our study in the introduction.
@@
 ## Reply to Reviewer: 1
-
-ADd summary paragraph, in our view these are the three main things, here are the main responses
 
@@
 - Clarified in the introduction that the gap filled by our method is not in the statistical method per say (a difference-in-difference is essentially a BACI by another name), but in the use of targeted and non-targeted species as our treatment and control group, as opposed to inside and outside MPAs.
-See line XX
+  - See paragraph in introduction starting with "While these studies have clearly demonstrated the ability of MPAs to create differences between local fished and unfished areas, this body of empirical work does not necessarily serve as sufficient evidence for the population-level effects of MPAs.
+These spatial studies rely on assuming that selected control sites serve as a measure of what would have happened in the absence of MPAs"
 
-- Expanded the description of our methods on line XX
+- Expanded the description of our methods, in particular the structure of the difference-in-difference model in the Methods section, and its relationship to BACI designs
 
-Major comments:
+- We also more clearly explain the gap filled by our method in the opening paragraph of the Discussion
+
+### Major comments:
 
 Abstract:
+
 - *Please indicate and describe briefly the method you used in your article*
-  - We have clarified in the abstract that we use a Bayesian difference-in-difference model
+
+We have clarified in the abstract that we use a Bayesian difference-in-difference model
 
 Introduction:
+
 *- Clear definitions of the terms are needed (“population-level effect”, “alteration of fishing effort” etc…)*
@@
 This is a problem that BACI style analyses are not designed to address.
-
-On line XX we state
-
-"XX"
-
-We synthesize the gap filled by our paper on line XX
-
-"XX"
+We make this point much more explicitly in the Introduction, Methods, and Discussion sections.
 
@@
 
-Thank you for pointing this out, we have clarified the paragraph in question and provided greater detail on why exactly this existing body of evidence does not necessarily serve as a population-level measure of MPA effects (Line XX).
-
-XX Paragraph in intro starting with "While these studies have clearly demonstrated the ability" XX
+Thank you for pointing this out, we have clarified the paragraph in question and provided greater detail on why exactly this existing body of evidence does not necessarily serve as a population-level measure of MPA effects in the Introduction.
 
@@
 We apologize for the confusion here.
-BACI and BACIPS are both powerful tools for program evaluation in natural systems, and we now mention BACI explicitly in the text.
+BACI and BACIPS are both powerful tools for program evaluation in natural systems, and we now mention BACI explicitly in the text (while referencing @thiault2019a and its discussion of BACIPS).
 However, both response ratios and BACI/BACIPS depend on the control being unaffected by the treatment.
@@
 
-The key problem here then is not the temporal or spatial variation that BACI-style methods seek to address, but the likely influence of the treatment on the control sites in the context of MPA studies.
-We have endeavored to make this clearer in the text, on line XX>
-
-XX Paragraph in intro starting with "While these studies have clearly demonstrated the ability" XX
+The key problem here then is not the temporal or spatial variation that BACI-style methods seek to address (e.g.
+a BACI can control for the possibility that MPAs are placed in better habitat), but the likely influence of the treatment on the control sites in the context of MPA studies.
+We have clarified in the introduction that the key to our strategy is not new statistical procedure, difference-in-difference being a form of BACI, but in using species targeted and non-targeted by fishing as the treatment and control groups.
+While this strategy comes with a host of different assumptions that we discuss and test for, for the purposes of assessing population level effects we feel this approach is an important change as it allows us to consider the effects of MPAs across seascapes when their effects stretch beyond their borders.
 
@@
 We take this to mean a explanation and citation for difference-in-difference.
-We have included a summary citation for the method (line XX, start of methods), as well as a brief description, given the length constraints of this paper, explaining the link between BACI and difference in difference.
-
-XX paragraph in intro starting with "To fill this important gap"
+We have included a summary citation for the method [@angrist2009], as well as a brief description, given the length constraints of this paper, explaining the link between BACI and difference in difference in our Methods.
 
@@
 
-We have clarified the objectives of this study, and its place in the literature, on lines XX and XX
-
-XX basically the whole introduction XX
+We hope that the concluding paragraph of our Introduction now makes the objectives of our study clearer.
 
@@
 Given word limits we unfortunately had to leave many of the details of the simulation model in the supplementary materials.
-We have however added expanded descriptions of these key attributes in the main text, in particular the nature of the variables requested here, as well as clarifying the role and nature of the simulation model.
-Line XX
-XX Simulation Model part of the methods section XX
+We have however added expanded descriptions of key attributes of the simulation model in the Simulation Model section of our Methods, in particular the nature of the variables requested here, as well as clarifying the role and nature of the simulation model.
 
@@
 
-We have however included an equation to clarify the BACI-like nature of our regression, on line XX
-
-XX methods section starting with 
-For readers more familiar with BACI methodXX
+We have however included an equation to clarify the BACI-like nature of our regression in the Difference-in-Difference section of our Methods, which is perhaps what the reviewer was wanting to see?
 
 Since this is a Bayesian regression, we provide credible intervals rather than confidence intervals [@gelman2014a].
-Similarly, there is no equation per-say for credible intervals, as they are essentially just an exercise in manipulating probabilities.
+As such, there is no equation per-say for credible intervals, as they are essentially just an exercise in manipulating probabilities.
 Our regression provides a posterior probability distribution parameters, and we then select the values of the coefficients within a desired range.
 For example to construct a 95% credible interval, we would select the values of the coefficient from the posterior probability distribution that fall between the 2.5%th and 97.5th percentiles of the posterior probability distribution.
+We also provide a very brief clarification of the term "posterior probability distribution" at the end of the Difference-in-Difference portion of the Methods section.
 
@@
 
-We have briefly addressed our choice of Bayesian methods on line XX
-
-XX end of methods section on diff-in-diff XX
+We briefly elaborate on our choice of a Bayesian model at the end of the Difference-in-Difference portion of the Methods section.
 
@@
 
-Thank you for pointing out that these terms are poorly described in the text, we have improved our descriptions (line XX).
-
-XX methods simulation model section XX
+Thank you for pointing out that these terms are poorly described in the text, we have improved our descriptions in the Simulation Model, and sought to use more general language in our Introduction.
 
@@
 
-We take the referees comment to mean our choice of using both a difference-in-difference model and a simulation model?
+We take the referees comment to mean our choice of using both a difference-in-difference model and a simulation model.
 We apologize for the confusion.
 In this case, we have not used the simulation model as an explicit estimation strategy, but rather to frame expectations.
-However one could, in the manner of @nickols2018, tune the simulation model to the local system and use simulation projections as a counterfactual.
-Since our simulation model is not as empirically calibrated as the one used in @nickols2018, we do not use it as an identification strategy.
-As discussed on line XX, we feel that it is important to ground any empirical results in theoretical predictions, and that is why we present the simulation study along with the empirical assessment.
+However one could, in the manner of @nickols2019, tune the simulation model to the local system and use simulation projections as a counterfactual.
+Since our simulation model is not as empirically calibrated as the one used in @nickols2019, we do not use it as an identification strategy.
+As stated in our Discussion, we feel that it is important to ground any empirical results in theoretical predictions, and that is why we present the simulation study along with the empirical assessment.
 
 In regards to the difference-in-difference approach, we selected it as it fits the nature of the policy intervention in question: we have a treatment (targeted species) and a control (non-targeted species) observed before and after impact (MPA implementation).
-As we have sought to clarify on line XX, DiD is for all intents and purposes the same as BACI, which is a commonly used metric in MPA evaluations.
-We use the term DiD following the vein of @larsen2018, to more closely tie the program evaluation literature in ecology to the econometrics literature that more directly deals with causal inference in program evaluating.
+As we have sought to clarify in the Methods, DiD is for all intents and purposes the same as BACI, which is a commonly used metric in MPA evaluations.
+We use the term DiD following the vein of @larsen2019, to more closely tie the program evaluation literature in ecology to the econometrics literature that more directly deals with causal inference in program evaluating.
+
 Our main methods difference then is in choosing to use targeted and non-targeted species as our treatment and control group, rather than inside and outside MPAs, since we believe there is strong a strong possibility that the MPAs affect the "control" sites outside their borders in the Channel Islands.
@@
 
-We have more clearly explained that the DiD approach is more or less identical to a BACI on line XX, which we hope will make this particular model choice seem more familiar.
-
-XX intro to methods, end of diff-in-diff section XX
+We have more clearly explained that the DiD approach is more or less identical to a BACI in our Methods, which we hope will make this particular model choice seem more familiar.
 
@@
 We do not claim to have tested for enforcement in our results.
-We mention size and enforcement on line 277XX in the discussion in the context of setting expectations, which we feel is an important point to make.
+We mention size and enforcement in the Introduction and Discussion in the context of setting expectations, which we feel is an important point to make.
 We do not mention enforcement in our results that we can find.
-We do simulate the effect of MPA size, which we make clearer in our methods description, on line XX.
-
-XX methods section XX
+We do simulate the effect of MPA size, which we make clearer in our Simulation Model Methods description.
 
@@
 We respectfully disagree with this suggestion.
-Following referee #1's helpful suggestions, we have clarified the broader challenges with interpreting response ratios as population-level effects in the introduction on line XX, which we agree were poorly articulated.
+Following referee #1's helpful suggestions, we have clarified the broader challenges with interpreting response ratios as population-level effects in the Introduction and Discussion, which we agree were poorly articulated.
 However, we feel that this section exploring the specific response ratios in the Channel Islands is a result of our simulation modeling, and so belongs in the results section.
@@
 
-*-       Please, do not cite a figure for the first time in the discussion that has a major result in it (Fig. 5)*
+*-Please, do not cite a figure for the first time in the discussion that has a major result in it (Fig. 5)*
 
-Thank you for pointing this out, we have moved all of this section to the results (line XX).
+Thank you for pointing this out, we have moved all of this section to the Results section.
 
@@
 
-Minor comments:
+### Minor comments:
 
@@
 
-We have clarified specific agreements on line XX
+We have clarified specific agreements in our Introduction.
 
@@
 
-We have clarified this statement on line XX.
+We have clarified this statement.
 
@@
 
-To make room for other material in the paper we have removed this specific discussion
+To make room for other material in the paper we have removed this specific discussion.
 
@@
 
-We have clarified on line XX that biomass ratios refer to the weight of organisms per unit area.
+We have clarified that biomass densities refer to the weight of organisms per unit area.
 
@@
 
-Thank you for catching this, there was an error in the conservation biology citation style language file we were using, which we have corrected. Citations are now in chronological order.
+Thank you for catching this, citations are now in chronological order.
 
@@
 
-Thank you for pointing this out.
 We have removed that line following the Referee #1's suggestion to ensure to continuity of paragraphs, as the displacement of fishing effort is not directly a population level effect, but rather a cause of population effect.
@@
 
-Per the referees helpful suggestion, we have moved materials related to this from the results to the introduction (line XX).
+Per the referees helpful suggestion, we have moved materials related to this from the results to the introduction.
 The problem is that response ratio / spatial BACI style designs will still be biased if the treatment unit affects the control unit, a violation of the stable unit treatment value assumption (SUTVA).
@@
 A BACI style design would conclude that the MPA had little to no effect in this case, since biomass levels would be similar inside and outside the MPA.
+See @ferraro2018 for a specific discussion of this problem in the context of MPAs.
 
@@
 
-We have added @ferraro2018 and @larsen2018 as references here
+We have added @ferraro2018 and @larsen2019 as references here
 
@@
 
-Thank you for pointing out this omission, these have been added the standard deviation of the simulated values on line XX.
+Thank you for pointing out this omission, these have been added the standard deviation of the simulated values in our Simulation Model section of the Methods.
 
@@
 
-we have clarified that the posterior probability distribution is the outcome of a Bayesian model (line XX.) Thank you for pointing out that this may not be a commonly used term.
+We have clarified that the posterior probability distribution is the outcome of a Bayesian model at the end of our Difference-in-Difference Methods section.
+Thank you for pointing out that this may not be a commonly used term.
 Bayes theorem states that the probability of x conditional on y, *p(x|y)*, is proportional to the probability of y conditional on x, *p(y|x)* times the prior probability of x, *p(x)*; *p(x|y) ~ p(y|x)p(x)*.
-The posterior probability distribution is the distribution of *p(x|y)*, the outcome of any Bayesian analysis.
-See @gelmanXX.
+The posterior probability distribution is the distribution of *p(x|y)*, the outcome of any Bayesian analysis, see @gelman2014a.
 
@@
 
-Given the space constraints of this paper, we do not feel that we are able to include a broader discussion of Bayesian inference, but have included a citation to a standard textbook on the topic in our introduction of our methods on line XX.
+Given the space constraints of this paper, we do not feel that we are able to include a broader discussion of Bayesian inference, but have included a citation to a standard textbook on the topic in our introduction of our Methods.
 
@@
 
-Fig. 3:
--       be consistent with the terms in the title and in the figure please
--       put the legend in the middle of y axis
--       Please change the legend for “Paired simulation Pop. Effect” and “Empirical Response Ratio” because it seems that “/” in between is a ratio
+*Fig.
+3: be consistent with the terms in the title and in the figure please put the legend in the middle of y axis.
+Please change the legend for “Paired simulation Pop.
+Effect” and “Empirical Response Ratio” because it seems that “/” in between is a ratio*
 
@@
 
-I enjoyed reading this paper and feel it makes an important contribution to the MPA literature. The focus on distinguishing seascape-wide MPA effects from local MPA effects is timely and important.
+*I enjoyed reading this paper and feel it makes an important contribution to the MPA literature. The focus on distinguishing seascape-wide MPA effects from local MPA effects is timely and important.*
 
-*Thank you, we are glad to hear the reviewer feels this is an important contribution.*
+Thank you, we are glad to hear the reviewer feels this is an important contribution.
 
@@
 We have restructured the SI so that the simulation model equations are front and center (SI Section 2), and clarified our explanations of the equations.
-We have also expanded our description of the simulation model in the text.
+We have also expanded our description of the simulation model in the Methods section of the main text.
 We hope this clarifies the structure behind the simulation model.
@@
 We apologize for the confusion in the simulation scenarios.
-Given the wide array of scenarios employed, and the word constraints of the paper, we unfortunately could not include a thorough description of every scenario in the main text (line XX).
-We have however included a clearer summary of the simulation in the methods section, highlighting among other key attributes the ways in which we simulate spatial patterns of fishing intensity, which as Referee #2 has noted is a critical variable to consider.
+Given the wide array of scenarios employed, and the word constraints of the paper, we unfortunately could not include a thorough description of every scenario in the main text.
+We have however included a clearer summary of the simulation in the methods section, highlighting among other key attributes the ways in which we simulate spatial patterns of fishing intensity, which as Reviewer #2 has noted is a critical variable to consider.
 
@@
 
-We have expanded our description of the simulation model in an effort to explain this more clearly, and in particular expanded our matching process to address this comment (line XX.).
+We have expanded our description of the simulation model in an effort to explain this more clearly, and in particular expanded our matching process to address this comment (end of the Simulation Model Methods section).
 Our Channel Islands simulations were designed to match the genera of species observed in the Channel Islands, our best sense of the exploitation history of those species, and the potential size of the MPAs for those species.
@@
 
-XX see end of simulation model part of methods XX
-
 *3) The figures are generally not very clear; particularly 2 and 6.
@@
 We have rewritten the caption to make Figure 6 clearer: we acknowledge that it is a complicated figure but we feel that it is important to convey the range of possible outcomes produced by our simulation model.
-
-Declarative sentence for what this figure is out. Results of ... Insight first, then terms at the end. warmer colors indicate larger effect XX
 
@@
 
-Comments to the Author
-
 *This study fills an important gap in knowledge about the seascape wide effects of MPAs on fish populations.
@@
 
-You are welcome, we are glad it helped
+You are welcome, we are glad it helped.
 
@@
 
-We are glad you found the simulation model sufficiently comprehensive
+We are glad you found the simulation model sufficiently comprehensive.
 
@@
 
-We have moved some of the methods from the SI as suggested, which we hope clarifies what a difference-in-difference is (essentially the same thing as a BACI).
+We have moved some of the methods from the SI as suggested, which we hope clarifies the connection between difference-in-difference and BACI.
 We use the term difference-in-difference to tie our results more clearly to the causal inference literature in econometrics.
-We have also clarified on line XX the clear link between DiD and BACI.
-
-XX DiD methods section XX
+WE have also clarified that use use a GLMM-style model (we refer to it as a hierarchichal generalized linear model since the term mixed effects doesn't quite apply in a Bayesian context) model to account for the spatial nature of our data.
 
@@
 Thank you for pointing this out.
-We have clarified the pre and post time period discussion in the methods (line XX).
+We have clarified the pre and post time period discussion in the Difference-in-Difference section of our Methods.
 We chose to bin the results for visual clarity.
@@
 no binning) to make clear that our results are not sensitive to the choice of binwidth (Fig.S23)
-
-XX end of diff in diff methods XX
 
@@
 This is correct.
-it was quite buried in the SI, but we do weight the MPA and non-MPA sites according to the total area of MPAs in the seascape (SI line XX), since as the referee points out the sampling regime is not random.
+It was quite buried in the SI, but we do weight the MPA and non-MPA sites according to the total area of MPAs in the seascape, since as the referee points out the sampling regime is not random.
+We note this weighting in our Methods section now.
 Otherwise, if the MPAs caused an increase of 50% inside, and 0% outside, given that the PISCO sampling regime has roughly equal sampling inside and outside MPAs, the model would estimate a net effect of 25%, even though MPAs in fact make up roughly 20% of the area.
 
-We have clarified that we weighted the regressions to account for this non-uniform sampling regime on line XX.
-We also include an unweighted version of our results, to demonstrate that our results are not sensitive to this weighting choice (Fig.SXX).
-
-XX diff in diff methods XX
+We also include an unweighted version of our results, to demonstrate that our results are not sensitive to this weighting choice (Fig.S22).
 
@@
 
-Thank you for this point!
+Thank you for this point.
 We reorganized the supplemental materials for ease of reading.
-*However, we have not removed sections at this time.
-Each of the analyses included in the SM (beyond that basic explanations of the methods) were a result of a question that we received from prior reviewers or peers.
+However, we have not removed sections at this time.
+Each of the analyses included in the SI (beyond that basic explanations of the methods) were a result of a question that we received from prior reviewers or peers.
 As such, we feel that it is better to include these in there for the odd person that really wants to see the results via synthetic controls.
-Given that the referees primary concern here was our not wasting useful analyses (which we appreciate!), and that we do not intent do dive further into the empirics of this particular question in the Channel Islands at this time, we have elected to go with the conservative approach of a perhaps excessively thorough SM.
+Given that the referees primary concern here was our not wasting useful analyses (which we appreciate!), and that we do not intent do dive further into the empirics of this particular question in the Channel Islands at this time, we have elected to go with the conservative approach of a perhaps excessively thorough SI.
 
@@
 
-Thank you for this excellent suggestion that we have sought to reflect on line XX.
+Thank you for this excellent suggestion that we have sought to reflect in our Introduction.
 
- XX Introduction XX
-
-*The references are pretty focused on California, and there is certainly a lot of great MPA literature from that part of the world, but perhaps you could expand the geography a bit to make the relevance to international readers clear?* See suggestions below.
+*The references are pretty focused on California, and there is certainly a lot of great MPA literature from that part of the world, but perhaps you could expand the geography a bit to make the relevance to international readers clear?
+See suggestions below.*
 
 Thank you for pointing this out.
-We only see one suggested reference (Krueck et al.
-2017), which we have incorporated on line XX.
+We only see one suggested reference (@krueck2017), which we have incorporated.
 We have also cited @fulton2015 in order to link our more more explicitly to the broader literature on MPA simulation model.
@@
 We did consider this, and decided that the extra effort needed to blind the code itself was not worthwhile.
-An interesting challenge for journals seeking to do double blind reviews to consider, it's hard to post code that is both open and reproducible and anonymous!
+An interesting challenge for journals seeking to do double blind reviews to consider; it's hard to post code that is both open and reproducible and anonymous!
 
@@
 
-Thank you for pointing this out, we have included this citation, as well as @fulton2015, to place are results more clearly into a broader literature of MPA simulation (Line XX)
+Thank you for pointing this out, we have included this citation, as well as @fulton2015, to place are results more clearly into a broader literature of MPA simulation.
 
@@
 
-We agree this would help readers understand the model and have included this suggestion on line XX of the methods.
+We agree this would help readers understand the model and have included this suggestion on in our Methods section.
 
@@
 We elected not to, as doing so would negate their benefits as an identification strategy, controlling for unobserved variables that are correlated with with the timing of MPA placement with the total biomass densities of targeted finfish.
-As discussed in @larsen2018, treating our year terms as hierarchical (or random effects in the context of that paper) means that we assume that the error distribution of the year terms is uncorrelated with the treatment, meaning that the difference-in-difference no longer acts to remove omitted variable bias, which is its main purpose here.
+As discussed in @larsen2019, treating our year terms as hierarchical (or random effects in the context of that paper) means that we assume that the error distribution of the year terms is uncorrelated with the treatment, meaning that the difference-in-difference no longer acts to remove omitted variable bias, which is its main purpose here.
 
@@
 We have included a frequentist version of our model in the Supporting Information Fig.S26, which uses the Newey-West correction to account for clustering of the observations in space and time, and the results are qualitatively identical to our main Bayesian difference-in-difference results.
-
-XX Owen I'll work on this while you check my edits XX
 
@@
 
-Thank you for pointing this out, and have rephrased this section accordingly (line XX) to reflect that this is something that stakeholders should be aware of, as demonstrated in Little et al. 2009.
+Thank you for pointing this out, and have rephrased this section accordingly to reflect that this is something that stakeholders should be aware of, as demonstrated in Little et al. 2009.
 
@@
 
-We believe there is some confusion, this is the same simulation model as presented in the methods (line XX) and used elsewhere in the paper.
-This clearly was not evident in the text though, we have clarified that this is the same model on line XX.
+We believe there is some confusion, this is the same simulation model as presented in the methods and used elsewhere in the paper.
+This clearly was not evident in the text though, we have clarified that this is the Simulation Model section of the Methods.
 
@@
 
-Interesting question!
-As we discuss on line XX, what evidence we have is that poaching is of limited concern in the Channel Islands (but clear can be a major issue in many MPAs).
+Interesting question.
+What evidence we have is that poaching is of limited concern in the Channel Islands (but clear can be a major issue in many MPAs).
 However, the presence of poaching should not bias the estimates of the DiD model in any way: if poaching were severe enough so as to render the MPAs completely ineffective, then in theory a DiD model such as the one implemented here would simply estimate that the MPA had no effect, as their would be no difference in the conditional biomass densities of targeted and non-targeted species.
 
-However, poaching might certainly influence ones interpretation of the results.
+However, poaching might certainly influence stakeholder interpretation of the results.
 If poaching is rampant but stakeholders are largely unaware of it, a model like this might estimate the the MPAs had no effect (which in that case they might not).
@@
 
-XX paragraph starting with "Stock assessment models that account for the population within the MPAs [@" XX
+Chris Brown
 
-Chris Brown
+# Literature Cited

### 2020-12-03 documents/ovando-et-al-pop-effects-of-mpas.Rmd
just about ready

@@
   Marine Protected Areas (MPAs) cover 3-7% of the world’s ocean, up from less than 1% in the year 2000, and international organizations call for 30% by 2030.
-The premise underlying MPA expansion is that they conserve biodiversity, habitats, and fished populations.
 While numerous studies show that MPAs produce conservation benefits *inside* their borders, many MPAs are also justified on the grounds that they confer conservation benefits to the broader population *beyond* their borders.
@@
 Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level using a Bayesian difference-in-difference approach.
-We show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); smaller effect sizes (which we find are more common) are even harder to detect.
+We show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); smaller effect sizes (which we find are common) are even harder to detect.
 Our results provide a novel assessment of the population-level effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.
@@
 Numerous factors influence how MPAs affect fish populations.
-These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003; @botsford2008], the strength, timing, and location of density dependence [@burgess2014], the design of the network [@gaines2010;@costello2010a],the degree of enforcement [@edgar2014], the level of fishing pressure, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@hastings2003;@gerber2003;@hilborn2004a;@gaines2010;@botsford2003;@smith2003a;@walters2000a;@white2011;@ovando2016;@moffitt2013].
+These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003; @botsford2008], the strength, timing, and location of density dependence [@burgess2014], the design of the network [@gaines2010;@rassweiler2014],the degree of enforcement [@edgar2014], the level of fishing pressure, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@gerber2003;@hilborn2004a;@gaines2010;@botsford2003;@smith2003a;@walters2000a;@white2011;@ovando2016;@moffitt2013].
 
@@
 These studies found clear evidence that well enforced and sufficiently sized MPAs are associated with high response ratios.
-Several studies have also documented empirical evidence for the existence of adult or larval fish spillover affecting fish abundance [e.g.
-@halpern2009;@kay2012; @mcclanahan2000; @russ1996].
+Several studies have also documented empirical evidence for the existence of adult or larval fish spillover affecting fish abundance [@halpern2009;@kay2012; @mcclanahan2000; @russ1996].
 Where response ratios are available before and after MPA implementation, spatial before-after-control-impact (BACI) style studies have shown similarly clear and positive results [@thiault2019a].
@@
 
-We present here an paired theoretical and empirical assessment demonstrating the challenges in assessing the population-level impacts of MPAs.
+In this paper, we present a paired theoretical and empirical assessment demonstrating the challenges in assessing the population-level impacts of MPAs.
 In 2003, a network of MPAs was established in the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) covering approximately 20% of the Islands' waters.
 The network has been used as a model case study in protected area design around the world [@botsford2014 and references therein].
-We use data from the first 14 years of protection in a difference-in-difference (DiD) model [see @angrist2009 and references therein] to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of fin-fish species.
+We use data from the first 14 years of protection in a difference-in-difference (DiD) model [@angrist2009] to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of fin-fish species.
 Rather than relying on spatial controls as is typically done in MPA studies, we use groups of species targeted and not-targeted by fishing pressure as our treatment and control groups.
@@
 We use the term DiD here to more closely tie our results to the terminology in causal-inference literature.
-We used a bio-economic simulation model to provide theoretical bounds on potential population-level effects of MPAs, which we then confront with the empirical results from our DiD model.
+We used a bio-economic simulation model to provide theoretical bounds on potential population-level effects of MPAs, which we then confront with the empirical results from our DiD regression.
 
 All analysis were conducted in R [@rcoreteam2019].
-Our DiD model was fit using Stan [@carpenter2017] using the rstanarm package [@goodrich2020].
+Our DiD regression was fit using Stan [@carpenter2017] using the `stan_glmer` function from the `rstanarm` package [@goodrich2020].
 All data and code needed to fully reproduce this manuscript are publicly available at [github.com/DanOvando/population-effects-of-mpas](https://github.com/DanOvando/population-effects-of-mpas).
@@
 
-The DiD model used empirical kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands.
+The DiD analysis used kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands.
 PISCO conducts visual SCUBA surveys at a large number of rocky reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\@ref(fig:ci-map),Fig.\@ref(fig:raw-density)).
@@
 
-The simplified regression form of this model is
+The simplified form of this regression model is
 
@@
 
+\begin{equation}
+  \mathbf{B^s} \sim Normal(\beta_r,\sigma_r) 
+\label{eq:did}
+\end{equation}
+
 where $d_i$ is the biomass density at observation *i*, *T* indicates whether the observation *i* is for a targeted ($T = 1$) or non-targeted ($T = 0$) species, and *MPA* marks whether observation *i* is in a pre MPA ($MPA = 0$) or post MPA ($MPA = 1$) state.
@@
 **B^c^** is a vector of coefficients for additional control variables in matrix *X* such as water temperature and observer experience.
-**B^s^** is a vector of hierarchical coefficients for each sampling location **S** , clustered by island.
+**B^s^** is a vector of hierarchical coefficients for each sampling location **S** , clustered by island $\beta_r$ with variance $\sigma_r^2$.
 Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment (*MPA*) on the treated (targeted species).
@@
 $\beta_2MPA$ is the effect of the post-MPA period on the non-targeted species (Table.\@ref(tab:coef-table)).
-We use a Bayesian regression since it allows us to interpret our estimated effects probabilistically [@mcelreath2020].
-Being a Bayesian regression, our DiD model produces posterior probability distributions (the probability distribution of our coefficients conditional on the data, priors, and model assumptions) of our coefficients, from which we can construct Bayesian credible intervals [@gelman2014a].
+We use a Bayesian hierarchical generalized linear model since it allows us to interpret our estimated effects probabilistically [@mcelreath2020].
+Being a Bayesian regression, our DiD analysis produces posterior probability distributions (the probability distribution of our coefficients conditional on the data, priors, and model assumptions) of our coefficients, from which we can construct Bayesian credible intervals [@gelman2014a].
 
@@
 Using a Bayesian DiD regression on these simulated data, we estimated the percent error between the posterior probability distribution of the estimated MPA effect from the regression and the true simulated MPA effect.
-The error in the DiD model's estimate of the population-level MPA effect was extremely high when MPA effect sizes were less than 25% and the model was faced with both observation and process errors (Fig.\@ref(fig:val-plot)).
+The error in the DiD regression's estimate of the population-level MPA effect was extremely high when MPA effect sizes were less than 25% and the model was faced with both observation and process errors (Fig.\@ref(fig:val-plot)).
 Even models fit to data generated from large effect sizes commonly mis-estimated the true MPA effect by 50% or more.
@@
 High response ratios suggest that despite overall decreases in targeted biomass densities inside and outside MPAs relative to pre-MPA measurements (Fig.\@ref(fig:raw-density)), the MPAs may still be providing protection within their borders.
-But, these spatial response ratios are not necessarily an indicator of broader population-level effects.
+But, these spatial response ratios are not necessarily an indicator of broader population-level effects due to the likelihood of MPAs affecting areas outside their borders.
 The targeted vs.
@@
 Fishing dynamics may be one factor contributing to a lack of strong MPA network effects.
-Much of the theoretical literature on MPAs assumes that larger reserves produce larger conservation gains [@white2011 and references therein].
-However, these models generally simulate fleet dynamics through fishing mortality rates; the proportion of total mortality experienced by a population attributable to fishing pressure [e.g.
-@halpern2004].
+Much of the theoretical literature on MPAs assumes that larger reserves produce larger conservation gains (@white2011 and references therein).
+However, these models generally simulate fleet dynamics through fishing mortality rates; the proportion of total mortality experienced by a population attributable to fishing pressure (e.g.
+@halpern2004).
 The implication of this is that total catches will scale with the size of the available population.
@@
 ```{r raw-density, fig.width = width_1col, fig.asp=asp_1col,fig.cap = "Centered and scaled trends in biomass densities of targeted and non-targeted fin-fish included in our study.
-Top panel shows mean trends across all sites, with shaded areas shwoing 95% confidence intervals.
-Bottom two panels show same trends but only including sites outside and inside MPAs"}
+Panel a) shows mean trends across all sites, with shaded areas shwoing 95% confidence intervals.
+Bottom two panels show same trends but only including sites b) outside and c) inside MPAs"}
 
@@
 
-```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="Simulated population effects (grey) that could have produced observed response ratios (black) in the Channel Islands.
-Black histograms show 90% Posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis).
-Simulated population-level (pop.) effect on biomass densities matched to empirical response ratios in light grey.
-For response ratios, a value of zero indicates that biomass densities of targeted species are identical inside and outside MPAs, a value of one that biomass densities of targeted species are 100% greater inside MPAs relative to outside.
-For MPA population effect, a value of zero indicates that biomass densities are identical in the with- and without- MPA scenarios.
-A value of 1 indicates that biomass densities are 100% greater in the scenario with MPAs than the scenario without MPAs."}
+```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="Simulated population effects (panel a) in grey) that could have produced observed response ratios (panel B) in black) in the Channel Islands.
+Panel a) histograms show 90% posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis).
+Panel b) Simulated population-level (pop.) effect on biomass densities matched to empirical response ratios in a).
+For response ratios, a percent difference value of zero indicates that biomass densities of targeted species are identical inside and outside MPAs, a percent difference value of one that biomass densities of targeted species are 100% greater inside MPAs relative to outside.
+For MPA population effect, a percent difference value of zero indicates that biomass densities are identical in the with- and without- MPA scenarios.
+A percent difference value of 1 indicates that biomass densities are 100% greater in the scenario with MPAs than the scenario without MPAs."}
 
@@
 
-```{r mpa-effect, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Results of difference-in-difference regression estimating the population effect of the Channel Island MPAs on mean total biomass densities of targeted species.
+```{r mpa-effect, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Results of Difference-in-Difference regression estimating the population-level effect of the Channel Island MPAs on mean total biomass densities of targeted species.
 Grey distributions show posterior probability distribution of estimated MPA effect; red point is median estimated effect, thicker red section 50% credible interval, thinner red line 90% credible interval.
 Blue distributions in background show range of MPA effects produced by simulation model tuned to reflect the dynamics of the Channel Island MPAs (black dashed line is median simulated value).
-Results are estimated in blocks of three years, including years greater than or equal to left-hand value and less than right-hand value."}
+Results are estimated in blocks of three years, including years greater than or equal to left-hand value and less than right-hand value.
+MPAs were implemented in 2003."}
 

### 2020-12-03 make-pop-effects-of-mpas.R

@@ -4677,7 +4677,7 @@ validation_plot <- valplot %>%
   scale_y_percent(name = "% Error",
                   labels = ylabs,
                   breaks = seq(-2.5, 2.5, by = .5)) +
-  scale_x_percent(name = "Population-Level Effect") +
+  scale_x_percent(name = "Population-Level MPA Effect") +
   scale_fill_gradient(
     low = "lightgrey",
     high = "#454545",

### 2020-12-04 documents/ovando-et-al-pop-effects-of-mpas-si.Rmd
ready to resubmit

@@
 
+The list of species included in our base results is available in Table.\@ref(tab:fish-tab).
+
+```{r fish-tab}
+a = life_history_data %>% 
+  filter(classcode %in% unique(used_data$classcode)) %>% 
+  select(taxa, commonname, targeted) %>% 
+  mutate(targeted = case_when(targeted == 1 ~ "Targeted",
+                              TRUE ~ "Non-Targeted")) %>% 
+  rename("Scientific Name" = taxa,
+         "Common Name" = commonname,
+         "Fishery Status" = targeted) %>% 
+  arrange(`Fishery Status`)
+
+  knitr::kable(
+    a,
+    digits = 2,
+    caption = "Fin-fish species available in PISCO data used in primary difference-in-difference regression.",
+    longtable = TRUE,
+    booktabs = TRUE ) %>%
+  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
+```
+
 The full table of covariates and their posterior means and 89% credible interval can be found in Table.\@ref(tab:did-coefs).

### 2020-12-04 documents/ovando-et-al-pop-effects-of-mpas.Rmd
ready to resubmit

@@
 Modern MPAs were first established as marine analogs to the terrestrial protection of iconic landscapes [@iucn1976].
-Over time our goals and expectations for MPAs have evolved and become more ambitious.
-Recent international efforts to expand MPAs, such as The International Union for Conservation of Nature's 30% by 2030 MPA targets are based in part on the assumption that well-designed MPAs will not only provide conservation benefits inside their borders, but also have broader conservation effects on unprotected areas surrounding the MPAs [@gaines2010].
+Recent international efforts to expand MPAs, such as The International Union for Conservation of Nature's 30% by 2030 MPA targets are based in part on the assumption, critical to MPAs designed explicitly for conservation, fishery benefits, or both, that well-designed MPAs will not only provide conservation benefits inside their borders, but also have broader conservation effects on unprotected areas surrounding the MPAs [@gaines2010].
 
@@
 However, as conservation benefits accrue inside MPAs, MPAs also affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
-Therefore, MPAs contribute to both local and population-level effects (i.e., inside-MPA effects and broader regional effects).
+Therefore, MPAs contribute to both local and population-level effects.
 Numerous factors influence how MPAs affect fish populations.
-These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003; @botsford2008], the strength, timing, and location of density dependence [@burgess2014], the design of the network [@gaines2010;@rassweiler2014],the degree of enforcement [@edgar2014], the level of fishing pressure, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@gerber2003;@hilborn2004a;@gaines2010;@botsford2003;@smith2003a;@walters2000a;@white2011;@ovando2016;@moffitt2013].
+These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003], the strength, timing, and location of density dependence [@burgess2014], the design of the network [@gaines2010;@rassweiler2014],the degree of enforcement [@edgar2014], the level of fishing pressure, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@gerber2003;@hilborn2004a;@gaines2010;@botsford2003;@smith2003a;@walters2000a;@white2011;@ovando2016;@moffitt2013].
 
@@
 These largely theoretical results are generally based on modeling of closed populations with some fraction protected inside of MPAs.
-In contrast to this population paradigm used in MPA simulations, MPAs are often evaluated empirically at local scales using spatial "response ratios", commonly measured as the ratio of biomass densities (weight of organisms per unit area) of species inside relative to selected control sites outside of MPAs [@lester2009;@halpern2003a;@edgar2014;@caselle2015].
+In contrast to this population paradigm used in MPA simulations, MPAs are often evaluated empirically at local scales using spatial response ratios, commonly measured as the ratio of biomass densities (weight of organisms per unit area) of species inside relative to selected control sites outside of MPAs [@lester2009;@halpern2003a;@edgar2014;@caselle2015].
 These studies found clear evidence that well enforced and sufficiently sized MPAs are associated with high response ratios.
@@
 Control sites sufficiently far from MPAs to negate both spillover of fish or larvae and concentration of the fishing fleet could be selected, but finding suitably distant sites that are also appropriate proxies for the ecological and economic context of the MPAs is challenging.
-As variations of spatial response ratios and BACI studies are our primary source of evidence for the conservation effects of MPAs, this means that our empirical understanding of the population-level impacts of MPAs is surprisingly limited.
+As variations of spatial response ratios and BACI studies have been a primary source of evidence for the conservation effects of MPAs, this means that our empirical understanding of the population-level impacts of MPAs is surprisingly limited.
 
@@
 Our methods consist of a spatially-explicit bio-economic simulation model and a Bayesian DiD regression [@angrist2009;@mcelreath2020].
-DiD is a regression-based approach to BACI, assessing changes in control and treatment groups before and after treatment [@larsen2019].
+DiD is a regression-based approach akin to BACI, assessing changes in control and treatment groups before and after treatment [@larsen2019].
 We use the term DiD here to more closely tie our results to the terminology in causal-inference literature.
@@
 Readers can explore the functionality of the model using an online tool available at [danovando.shinyapps.io/simmpa/](https://danovando.shinyapps.io/simmpa/).
-The purpose of the simulation model is to both set expectations for our empirical results, and demonstrate the ways in which ecological and economic dynamics can interact to produce a wide range of population-level MPA effects.
+The purpose of the simulation model is to both set expectations for our empirical results, and to demonstrate the ways in which ecological and economic dynamics can interact to produce a wide range of population-level MPA effects.
 The full range of factors explored are shown in  Table.\@ref(tab:sim-table), and the equations of the simulation model can be viewed in the Appendix S1 Section 2.
@@
 These simulation results provide a library of plausible MPA effects for a range of biological and economic assumptions.
-One set of simulations is specifically designed to reflect the dynamics of the Channel Islands.
-For this set we only include species of the same genus as those targeted by fishing in the Channel Islands, restrict fishing pressure such that the simulated populations are moderately to lightly exploited, and cap the MPA size at 20% of the population's range [@rassweiler2012].
+One set of simulations was specifically designed to reflect the dynamics of the subset of species available in the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring data from the Channel Islands used in this study.
+For this set we only include species of the same genus as those targeted by fishing in the PISCO data, restrict fishing pressure such that the simulated populations are moderately to lightly exploited (since the PISCO data used here excludes deeper-water species such as boccacio *Sebastes paucispinis* which were overexploited at the inception of the Channel Island MPAs, and threatened invertebrates such as red abalone *Haliotis rufescens*), and cap the MPA size at 20% of the population's range [@rassweiler2012].
 For each of these Channel Island style simulations, we calculated 1) the 'true' population-wide difference in biomass between the simulations with and without the MPAs and 2) the response ratio of biomass densities inside and outside the simulated MPAs.
-We then calculated the empirical response ratios observed in the real Channel Islands, and matched these empirical results with simulations that produced similar response ratios after the same number of years of MPA protection.
+We then calculated the response ratios observed in the PISCO survey data from the Channel Islands, and matched these empirical results with simulations that produced similar response ratios after the same number of years of MPA protection.
 Because each simulation included measures of both response ratios and population-level effects, this process provides a library of simulations (and their associated attributes) that could have produced the types of empirical response ratios measured in the Channel Islands.
@@
 
-The DiD analysis used kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands.
+The DiD analysis used kelp forest survey data PISCO surveys in the Channel Islands.
 PISCO conducts visual SCUBA surveys at a large number of rocky reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\@ref(fig:ci-map),Fig.\@ref(fig:raw-density)).
@@
 Building off of the concepts explored in @caselle2015, we used an identification strategy utilizing biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 0])` species that are not directly targeted by fishing as our control group (non-targeted), and biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 1])` species targeted by fishing as our treatment group (see Fig.\@ref(fig:raw-density) for raw trends in targeted and non-targeted biomass densities).
-Targeted species in the Channel Islands include fin-fish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
+Targeted  fin-fish species in the Channel Islands available to this study include California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
 Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*).
@@
 @caselle2015 found a statistically significant increase in the response ratios of targeted species over time, and evidence that this increase is smaller for non-targeted species.
-Updating the results of @caselle2015 through 2017 shows a continuation of the increasing trend in the response ratios of targeted species (Fig.\@ref(fig:rr-plot)).
-This provides evidence that the Channel Islands MPAs are large enough and sufficiently well-enforced as to provide meaningful protection within their borders.
+Updating the results of @caselle2015 using data collected through 2017 shows a continuation of the increasing trend in the response ratios of targeted species (Fig.\@ref(fig:rr-plot)).
+This provides evidence that the Channel Islands MPAs are large enough and sufficiently well-enforced as to provide meaningful protection within their borders [@white2020].
 
@@
 
+<!-- High response ratios suggest that despite overall decreases in targeted biomass densities inside and outside MPAs relative to pre-MPA measurements (Fig.\@ref(fig:raw-density)), the MPAs may still be providing protection within their borders.
+But, these spatial response ratios are not necessarily an indicator of broader population-level effects due to the likelihood of MPAs affecting areas outside their borders.
+-->
+
 Containing a carefully designed, well-enforced, and well-studied MPA network, the Channel Islands seems to be an ideal location to study the population-level effects of protected areas.
-High response ratios suggest that despite overall decreases in targeted biomass densities inside and outside MPAs relative to pre-MPA measurements (Fig.\@ref(fig:raw-density)), the MPAs may still be providing protection within their borders.
-But, these spatial response ratios are not necessarily an indicator of broader population-level effects due to the likelihood of MPAs affecting areas outside their borders.
 The targeted vs.
-non-targeted DiD strategy presents an alternative means of estimating the population-level effects of MPAs.
-While having its own strict caveats, this strategy provides some potential improvements over spatial response-ratios as a means of estimating population-level effects.
+non-targeted DiD strategy used here presents an alternative means of estimating the population-level effects of MPAs.
+Despite having its own strict caveats, this strategy provides some potential improvements over spatial response-ratios as a means of estimating population-level effects.
 While we estimate an uncertain but overall positive effect of the MPA network in its first few years, we are unable to detect a robust signal from 2012-2017.
-After 14 years of MPA protection we are left without a clear picture of the population-level effect of the Channel Island MPA network on biomass densities of targeted fin-fish species.
-
-Fishing dynamics may be one factor contributing to a lack of strong MPA network effects.
-Much of the theoretical literature on MPAs assumes that larger reserves produce larger conservation gains (@white2011 and references therein).
-However, these models generally simulate fleet dynamics through fishing mortality rates; the proportion of total mortality experienced by a population attributable to fishing pressure (e.g.
-@halpern2004).
-The implication of this is that total catches will scale with the size of the available population.
+We show that under the dynamics of the Channel Islands, particularly given the lack of heavily exploited species in the available data, this result is to be expected.
+After 14 years of MPA protection we are left without a clear picture of the population-level effect of the Channel Island MPA network on biomass densities of targeted fin-fish.
 
@@
 
+Fishing dynamics may be one factor contributing to a lack of strong MPA network effects.
+Much of the theoretical literature on MPAs assumes that larger reserves produce larger conservation gains (@white2011 and references therein).
+However, these models generally simulate fleet dynamics through fishing mortality rates; the proportion of total mortality experienced by a population attributable to fishing pressure (e.g.
+@halpern2004).
 Alternatively, under a “constant-catch” strategy, fishers have a catch objective and exert as much (or little) effort as needed to achieve that objective.
@@
 This potential negative interaction between constant-catch and MPAs is an important risk to consider (as done in @little2011b) especially as MPAs are increasingly implemented in quota-managed fisheries [@liu2018].
+While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (Appendix S1 Fig.S2) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
 
-While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (Appendix S1 Fig.S2) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
-However, some of the most common targeted species in the Channel Islands, such as blue and copper rockfish, have seen dramatic increases in reported catches in the Santa Barbara region since the year 2010 (although other such as California sheephead have seen declines relative to pre-MPA catch level).
+<!-- However, some of the most common targeted species in the Channel Islands, such as blue and copper rockfish, have seen dramatic increases in reported catches in the Santa Barbara region since the year 2010 (although other such as California sheephead have seen declines relative to pre-MPA catch level).
 While we include catch histories in our regression, to the extent that these increases in catches are due to external market forces, the disappearance of clear positive population-level MPA effects estimated by our model could be due to exogenous increases in fishing pressure overwhelming any population-level increases caused by the MPA network.
+-->
+
+<!-- We hypothesize that the recent evidence for a decline in densities of targeted species is primarily due to altered environmental conditions that may have disproportionately affected the targeted group.
+The practical result of this is tha -->
 
@@
 The Channel Islands region experienced a dramatic ‘marine heatwave’ beginning in 2014 and persisting through 2016, resulting in part in extremely elevated water temperatures throughout the region [@gentemann2017].
-Many of the non-targeted species in the Channel Islands have warm thermal affinities and have increased in numbers since the heatwave [@freedman2019].
+Many of the non-targeted species in the Channel Islands have warm thermal affinities and have increased in numbers since the heatwave [@freedman2020].
 The targeted group is made up mostly of fishes with cold-water affinities.
-We hypothesize that the recent evidence for a decline in densities of targeted species is primarily due to altered environmental conditions that disproportionately affected the targeted group.
-The practical result of this is that in the presence of this marine heatwave the non-targeted species may no longer serve as an effective control for the evolution of biomass densities of targeted fin-fish in the absence of the MPAs, given the magnitude of the environmental shock relative to the size of the population-level MPA effect.
+In the presence of this marine heatwave the non-targeted species may no longer serve as an effective control for the evolution of biomass densities of targeted fin-fish in the absence of the MPAs, given the magnitude of the environmental shock relative to the size of the population-level MPA effect.
 
@@
 
-Trophic cascades are another possible explanation for our equivocal results.
 All of the species in this empirical analysis may affect each other through mechanisms such as predation, competition, and habitat modification.
@@
 
-MPAs are an important part of the marine resource management toolbox.
 As the number and size of global MPA networks increase, we must set appropriate expectations for their outcomes on both local and regional scales.
-Our results for the relatively large, well-enforced, and rigorously studied (though also relatively lightly fished) Channel Islands Marine Protected Area network make clear that in many instances empirically detecting a clear effect of MPAs on total biomass densities of targeted fin-fish at the population level may not be possible.
-In the absence of clear empirical findings, simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can be tuned to focus on the species groups that have the highest chance of a detectable effect size over the early years of the reserve [@nickols2019].
+Simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can be tuned to focus on the species groups that have the highest chance of a detectable effect size over the early years of the reserve [@nickols2019].
 Expanding data collection to include robust monitoring of spatio-temporal fleet dynamics may help assess the validity of control sites used in response ratios, support the direct inclusion of these fleet dynamics into statistical models, and allow managers to take into account potential negative interactions between MPAs and fleet dynamics such as those that may occur under constant-catch dynamics.
@@
 
-```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="Simulated population effects (panel a) in grey) that could have produced observed response ratios (panel B) in black) in the Channel Islands.
-Panel a) histograms show 90% posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis).
+```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="Simulated population effects (panel *a* in grey) that could have produced observed response ratios (panel *b*) in black) in the Channel Islands.
+Panel *a*) histograms show 90% posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis).
 Panel b) Simulated population-level (pop.) effect on biomass densities matched to empirical response ratios in a).
@@
 For MPA population effect, a percent difference value of zero indicates that biomass densities are identical in the with- and without- MPA scenarios.
-A percent difference value of 1 indicates that biomass densities are 100% greater in the scenario with MPAs than the scenario without MPAs."}
+A percent difference value of one indicates that biomass densities are 100% greater in the scenario with MPAs than the scenario without MPAs."}
 
@@
 Shading shows concentration of simulations.
-Black line shows mean absolute percent error (MAPE) as a function of true simulated MPA effect."}
+Black line shows mean absolute percent error (MAPE) as a function of simulated population-level MPA effect."}
 

### 2020-12-04 documents/pop-effects-of-mpas-front-matter.Rmd
Edits almost done
ready to resubmit

@@
 We pair theory with empirical methods to show that the population-level conservation effects of marine protected areas can be highly variable and challenging to detect.
-We find no statistically clear effect of a large and well protected network of MPAs in the Channel Island, California, on targeted fin-fish after 14 years of protection, which our simulation analysis suggests is in fact to be expected.
+We find no statistically clear effect of a large and well protected network of MPAs in the Channel Island, California, on targeted fin-fish after 14 years of protection, which we show is in fact to be expected.
 MPA practitioners must clearly communicate expectations from MPA networks, and plan monitoring and adaptations strategies that directly address the challenges of estimating population-level conservation effects of protected areas.
@@
 Szuwalski, J.
-Lawson, and A.
-Punt for helpful comments and technical support.
+Lawson, A.
+Punt, Chris Brown, and two anonymous reviewers for helpful comments and technical support.

### 2020-12-07 documents/pop-effects-of-mpas-front-matter.Rmd
Cb edits (#5)
* edits for revisions at Conservation Biology
* clean up and merge back to main

Co-authored-by: Owen Liu <owenrliu@gmail.com>

@@
 ---
-title: "Assessing the Population-level Conservation Effects of Marine Protected Areas"
-subtitle: Cover Page
-date: "`r Sys.Date()`"
 output:
   bookdown::word_document2:
-    reference_docx: uw.docx
+    reference_docx: template.docx
   pdf_document: default
@@
 
-## Article Impact Statement
+## Title  {.unnumbered}
+Assessing the Population-level Conservation Effects of Marine Protected Areas
 
-Healthy marine ecosystems are critical to the well-being of the planet.
-Marine protected areas (MPAs) are increasingly being used to conserve and manage these ecosystems.
-We pair theory with empirical methods to show that the population-level conservation effects of marine protected areas can be highly variable and challenging to detect.
-We find no statistically clear effect of a large and well protected network of MPAs in the Channel Island, California, on targeted fin-fish after 14 years of protection, which we show is in fact to be expected.
-MPA practitioners must clearly communicate expectations from MPA networks, and plan monitoring and adaptations strategies that directly address the challenges of estimating population-level conservation effects of protected areas.
+## Article Impact Statement {.unnumbered}
 
-## Running Head
+The population-level conservation effects of Marine Protected Areas are likely to be extremely challenging to detect
 
-The Population Effects of MPAS
+## Running Head {.unnumbered}
 
-## Keywords:
+The Population-Level Effects of MPAS
+
+## Keywords: {.unnumbered}
 
@@
 
-## Word Count
+## Word Count {.unnumbered}
 
-6000
+5976
 
-## Author Details
+## Author Details {.unnumbered}
 
@@
 
-## Corresponding Author
+## Corresponding Author  {.unnumbered}
 
@@
 danovan@uw.edu
-phone: (805) 284 6864
 
-## Acknowledgements
+## Acknowledgements  {.unnumbered}
 

### 2020-12-07 documents/ovando-et-al-pop-effects-of-mpas.Rmd

@@
 abstract: |
-  Marine Protected Areas (MPAs) cover 3-7% of the world’s ocean, up from less than 1% in the year 2000, and international organizations call for 30% by 2030.
+  Marine Protected Areas (MPAs) cover 3-7% of the world’s ocean, and international organizations call for 30% by 2030.
 While numerous studies show that MPAs produce conservation benefits *inside* their borders, many MPAs are also justified on the grounds that they confer conservation benefits to the broader population *beyond* their borders.
@@
 
-No-take Marine Protected Areas (MPAs), spatial regions of the ocean in which fishing is prohibited, have a long history in the management of marine resources.
+No-take Marine Protected Areas (MPAs), spatial regions of the ocean in which fishing is prohibited, have a long history in the management of marine resources [@johannes1978].
 Modern MPAs were first established as marine analogs to the terrestrial protection of iconic landscapes [@iucn1976].
-Recent international efforts to expand MPAs, such as The International Union for Conservation of Nature's 30% by 2030 MPA targets are based in part on the assumption, critical to MPAs designed explicitly for conservation, fishery benefits, or both, that well-designed MPAs will not only provide conservation benefits inside their borders, but also have broader conservation effects on unprotected areas surrounding the MPAs [@gaines2010].
+Recent international efforts to expand MPAs, such as The International Union for Conservation of Nature's 30% by 2030 MPA targets are based in part on the assumption that well-designed MPAs will not only provide conservation benefits inside their borders, but also have broader conservation effects on unprotected areas surrounding the MPAs, whether MPAs are designed explicitly for conservation, fishery benefits, or both [@gaines2010].
 
@@
 However, as conservation benefits accrue inside MPAs, MPAs also affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort.
-Therefore, MPAs contribute to both local and population-level effects.
+Therefore, MPAs contribute to both local and regional population-level effects.
 Numerous factors influence how MPAs affect fish populations.
-These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003], the strength, timing, and location of density dependence [@burgess2014], the design of the network [@gaines2010;@rassweiler2014],the degree of enforcement [@edgar2014], the level of fishing pressure, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@gerber2003;@hilborn2004a;@gaines2010;@botsford2003;@smith2003a;@walters2000a;@white2011;@ovando2016;@moffitt2013].
+These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003], the strength, timing, and location of density dependence [@burgess2014], the design of the network [@gaines2010;@rassweiler2014],the degree of enforcement [@edgar2014], the level of fishing pressure, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@gerber2003;@hilborn2004a;@gaines2010;@botsford2003;@smith2003a;@walters2000a;@white2011;@ovando2016;@moffitt2013;@jaco2020].
 
 <!-- What empirical evidence is there for these population-level effects of MPAs?  -->
-These largely theoretical results are generally based on modeling of closed populations with some fraction protected inside of MPAs.
+This largely theoretical literature is generally based on modeling of closed populations with some fraction protected inside of MPAs.
 In contrast to this population paradigm used in MPA simulations, MPAs are often evaluated empirically at local scales using spatial response ratios, commonly measured as the ratio of biomass densities (weight of organisms per unit area) of species inside relative to selected control sites outside of MPAs [@lester2009;@halpern2003a;@edgar2014;@caselle2015].
@@
 Export of adults or larvae from MPAs to "control" sites can affect their status as controls, as does displacement of fishing effort from MPAs to control sites.
-These factors can bias estimators of population-level effects like response ratios that require the assumption that the control sites are unaffected by the treatment [@ferraro2018;@moffitt2013].
+These factors can bias estimators of population-level effects like response ratios that require that the control sites are unaffected by the treatment [@ferraro2018;@moffitt2013].
 Control sites sufficiently far from MPAs to negate both spillover of fish or larvae and concentration of the fishing fleet could be selected, but finding suitably distant sites that are also appropriate proxies for the ecological and economic context of the MPAs is challenging.
@@
 In 2003, a network of MPAs was established in the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) covering approximately 20% of the Islands' waters.
-The network has been used as a model case study in protected area design around the world [@botsford2014 and references therein].
+The network has been used as a model case study in protected area design around the world [@botsford2014].
 We use data from the first 14 years of protection in a difference-in-difference (DiD) model [@angrist2009] to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of fin-fish species.
@@
 DiD is a regression-based approach akin to BACI, assessing changes in control and treatment groups before and after treatment [@larsen2019].
-We use the term DiD here to more closely tie our results to the terminology in causal-inference literature.
-We used a bio-economic simulation model to provide theoretical bounds on potential population-level effects of MPAs, which we then confront with the empirical results from our DiD regression.
+We used a bio-economic simulation model to provide theoretical expectations of population-level effects of MPAs, which we then confront with the empirical results from our DiD regression.
 
 All analysis were conducted in R [@rcoreteam2019].
-Our DiD regression was fit using Stan [@carpenter2017] using the `stan_glmer` function from the `rstanarm` package [@goodrich2020].
+Our DiD regression was fit using Stan [@carpenter2017] through the `rstanarm` package [@goodrich2020].
 All data and code needed to fully reproduce this manuscript are publicly available at [github.com/DanOvando/population-effects-of-mpas](https://github.com/DanOvando/population-effects-of-mpas).
@@
 Many papers have presented simulation analyses of MPA outcomes [@fulton2015].
-Our model incorporates a broad range of core ecological and economic drivers of MPA performance assessed by these individual papers into a cohesive model, similar in spirit to @krueck2017.
+Our model incorporates core ecological and economic drivers of MPA performance assessed by these individual papers into a cohesive model, similar in spirit to @krueck2017.
 The simulation model consists of 50 patches with wrapped edges.
-For each simulation we first randomly pull a species and its associated life history traits from the `FishLife` [@thorson2017c] package.
+For each simulation we first randomly pull a species and its associated life history traits from the FishLife [@thorson2017c] package.
 We pair these data with randomly selected values governing the characteristics of the simulation (Table.\@ref(tab:sim-table)).
@@
 One set of simulations was specifically designed to reflect the dynamics of the subset of species available in the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring data from the Channel Islands used in this study.
-For this set we only include species of the same genus as those targeted by fishing in the PISCO data, restrict fishing pressure such that the simulated populations are moderately to lightly exploited (since the PISCO data used here excludes deeper-water species such as boccacio *Sebastes paucispinis* which were overexploited at the inception of the Channel Island MPAs, and threatened invertebrates such as red abalone *Haliotis rufescens*), and cap the MPA size at 20% of the population's range [@rassweiler2012].
+For this set we only include species of the same genus as those targeted by fishing in the PISCO data, restrict fishing pressure such that the simulated populations are moderately to lightly exploited (since the PISCO data used here excludes deeper-water species such as boccacio, *Sebastes paucispinis*, which were overexploited at the inception of the Channel Island MPAs, and threatened invertebrates such as red abalone *Haliotis rufescens*), and cap the MPA size at 20% of the population's range [@rassweiler2012].
 For each of these Channel Island style simulations, we calculated 1) the 'true' population-wide difference in biomass between the simulations with and without the MPAs and 2) the response ratio of biomass densities inside and outside the simulated MPAs.
@@
 
-The DiD analysis used kelp forest survey data PISCO surveys in the Channel Islands.
+The DiD analysis used kelp forest survey data from the PISCO surveys in the Channel Islands.
 PISCO conducts visual SCUBA surveys at a large number of rocky reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\@ref(fig:ci-map),Fig.\@ref(fig:raw-density)).
@@
 
-Building off of the concepts explored in @caselle2015, we used an identification strategy utilizing biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 0])` species that are not directly targeted by fishing as our control group (non-targeted), and biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 1])` species targeted by fishing as our treatment group (see Fig.\@ref(fig:raw-density) for raw trends in targeted and non-targeted biomass densities).
+Building off of @caselle2015, we used an identification strategy utilizing biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 0])` species that are not directly targeted by fishing as our control group (non-targeted), and biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 1])` species targeted by fishing as our treatment group (Fig.\@ref(fig:raw-density)).
 Targeted  fin-fish species in the Channel Islands available to this study include California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
@@
 **B^s^** is a vector of hierarchical coefficients for each sampling location **S** , clustered by island $\beta_r$ with variance $\sigma_r^2$.
-Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment (*MPA*) on the treated (targeted species).
-$\beta_0$ is the mean total biomass of non-targeted species pre-MPA, and $\beta_1T$ is the mean total biomass of targeted species pre-MPA.
-$\beta_2MPA$ is the effect of the post-MPA period on the non-targeted species (Table.\@ref(tab:coef-table)).
+Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment (*MPA*) on the treated (targeted species) (Table.\@ref(tab:coef-table)).
 We use a Bayesian hierarchical generalized linear model since it allows us to interpret our estimated effects probabilistically [@mcelreath2020].
@@
 
-These response ratios cannot, however, be used as a definitive indicator of population effects of these MPAs.
-In the case of the Channel Islands MPAs, control sites are often located within a few kilometers of an MPA, suggesting that they are susceptible to both local biological spillover and by concentration of fishing effort excluded from the MPAs.
+These response ratios cannot, however, be used as a definitive indicator of population-level effects of these MPAs.
+In the case of the Channel Islands MPAs, control sites are often located within a few kilometers of an MPA, making them susceptible to both biological spillover and concentration of fishing effort excluded from the MPAs.
 Our simulation results show that the response ratio trends we observe in the data could plausibly be produced by a wide range of population-level MPA effects (Fig.\@ref(fig:rr-plot)).
 Response ratios well over one were associated with "true" population-level MPA effects generally less than 25%, and, importantly, many simulations produced large response ratios but population-level MPA effects close to 0%.
-This can occur if for example fishing pressure is only moderate, adult movement is low, larval dispersal is high, and displaced fishing effort concentrates around the border of the MPAs.
+This can occur if for example fishing pressure is moderate, adult movement is low, larval dispersal is high, and displaced fishing effort concentrates around the border of the MPAs.
 
@@
 These empirical estimates are in line with the outcomes that our simulation model suggests are plausible.
-While a wide range of MPA effects are plausible, from 0% to upwards of 150%, the median simulated effect size for this stylized version of the Channel Islands after over ten years of protection was `r percent(median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected > 10]))`.
 However, in the subsequent years the trend reverses itself, and in 2015-2017 we once again see no clear effect of the MPAs (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`,
 90\% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`
-- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`.) Fig.\@ref(fig:mpa-effect).
+- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`.) (Fig.\@ref(fig:mpa-effect)).
 
-Our library of simulation results allows us to explore how MPA effects, and our ability to measure them, might vary with observation error and natural fish recruitment variation.
-Using a Bayesian DiD regression on these simulated data, we estimated the percent error between the posterior probability distribution of the estimated MPA effect from the regression and the true simulated MPA effect.
+<!-- While a wide range of MPA effects are plausible, from 0% to upwards of 150%, the median simulated effect size for this stylized version of the Channel Islands after over ten years of protection was `r percent(median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected > 10]))` -->
+
+Our library of simulation results allows us to explore how observation error and natural fish recruitment variation might explain this lack of a clear result in the Channel Islands MPAs.
+Using a Bayesian DiD regression on these simulated data, we estimated the percent error between the posterior probability distribution of the estimated MPA effect from the regression and the simulated MPA effect.
 The error in the DiD regression's estimate of the population-level MPA effect was extremely high when MPA effect sizes were less than 25% and the model was faced with both observation and process errors (Fig.\@ref(fig:val-plot)).
@@
 Obtaining a mean absolute percent error (MAPE) of 25% or less across our simulated datasets required a "true" population-level MPA effect of at least 30%.
-In the context of the Channel Islands, this finding suggests that we should not be surprised at our difficulty in precisely estimating the population-level effect of the MPAs.
+In the context of the moderately exploited species in the Channel Islands PISCO data evaluated here, this finding suggests that we should not be surprised at our difficulty in precisely estimating the population-level effect of the MPAs.
 <!-- This is merely an illustrative exercise, omitting critical factors such as detection probability and sampling strategy.
@@
 Although recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, most MPA networks for near-shore commercial fin-fish are likely to cover areas more in line with the Channel Islands (20%) or smaller.
-As such, many MPA networks are expected to have population-level effect sizes that are difficult to detect unless target species were experiencing extreme overfishing prior to MPA establishment (Fig.\@ref(fig:sim-effects)-A).
+As such, many MPA networks are expected to have population-level effect sizes that are difficult to detect unless target species were experiencing extreme overfishing prior to MPA establishment, or in the hypothetical world without MPAs (Fig.\@ref(fig:sim-effects)-A).
 
@@
 While we estimate an uncertain but overall positive effect of the MPA network in its first few years, we are unable to detect a robust signal from 2012-2017.
-We show that under the dynamics of the Channel Islands, particularly given the lack of heavily exploited species in the available data, this result is to be expected.
+We show that under the dynamics of the Channel Islands, particularly given the lack of heavily exploited species such as abalone and deep-water rockfish that helped motivate the Channel Island MPAs in the available data, this result is to be expected.
 After 14 years of MPA protection we are left without a clear picture of the population-level effect of the Channel Island MPA network on biomass densities of targeted fin-fish.
@@
 We used convergent cross mapping (CCM), in the manner of @clark2015, to test for significant dynamic interactions between species and therefore the possibility of the trophic cascades biasing our results.
-We found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (SI Fig.S42).
+We found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (Appendix S1 Fig.S33-S35).
 However, the longer MPAs are in place, the greater the possibility that substantial species interactions that can affect use of non-targeted species as a control may arise.
@@
 
-However, as the scientific community continues to develop appropriate measures of population-level MPA effects, we must also effectively communicate the challenges of estimating those effects.
-In light of these results, lack of a clear population-level MPA effect should not necessarily be viewed as a failure of a conservation program, and likewise, large response ratios should not be automatically taken as evidence of a population-level conservation success.
+The scientific communicate must effectively communicate the challenges of estimating the population-level effects of MPAs.
+Lack of a clear population-level MPA effect should not necessarily be viewed as a failure of a conservation program, and likewise, large response ratios should not be automatically taken as evidence of a population-level conservation success.
 Rather, results and subsequent management actions must be considered in the context of reasonable expectations given the size, age, and degree of enforcement of the MPAs in question, together with the ecological and economic dynamics of a given system.

### 2020-12-07 README.Rmd
updating README with renv instructions

@@
 
-Users will need to install the simulation model that can be found [here](https://github.com/DanOvando/spasm), and installed using `remotes::install_github("danovando/spasm")`
+This project is set up with [`renv`](https://rstudio.github.io/renv/articles/renv.html) to manage package dependencies.
+Inside R (and with your working directory set correctly) run `renv::restore()`.
+Follow all prompts.
+This will install the correct versions of all the packages needed to replicate our results.
+Packages are installed in a stand-alone project library for this paper, and will not affect your installed R packages anywhere else.
 
->>  Marine Protected Areas (MPAs) cover 3-7% of the world’s ocean, up from less than 1% in the year 2000, and international commitments call for 10%, 30%, and even 50% coverage.
-The premise underlying MPA expansion is that they conserve biodiversity, habitats, and fished populations.
+>>  Marine Protected Areas (MPAs) cover 3-7% of the world’s ocean, and international organizations call for 30% by 2030.
 While numerous studies show that MPAs produce conservation benefits *inside* their borders, many MPAs are also justified on the grounds that they confer conservation benefits to the broader population *beyond* their borders.
-This paper examines the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and shows that even in cases where the conservation benefits are large, they are inherently difficult to detect empirically.
+We examine the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and show that even in cases where the population benefits are large, they are inherently difficult to detect empirically.
 A network of MPAs was put in place in The Channel Islands National Marine Sanctuary in 2003, with a goal of providing regional conservation and fishery benefits.
-Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level.
-MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present), and the size of many MPA networks suggests that effect sizes may often be smaller than 30%.
-Our results provide a novel assessment of the population-level effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.
+Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level using a Bayesian difference-in-difference approach.
+We show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); smaller effect sizes (which we find are common) are even harder to detect.
+Our results provide a novel assessment of the population-level effects of a large and iconic Marine Protected Area

### 2021-03-07 documents/referee-replies-2.Rmd
pushing for final review

---
title: Assessing the Population-Level Conservation Effects of Marine Protected Areas
subtitle: "Reviewer Replies 2"
bibliography: references.yaml
csl: conservation-biology.csl
output:
  bookdown::pdf_document2:
    keep_tex: true
  bookdown::word_document2:
    reference_docx: template.docx
  bookdown::html_document2: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

## General Comments

We are pleased to hear that the reviewers and editors found our last round of edits helpful and sufficient, and appreciate the comments provided for this round of minor revisions. We agree with all of Reviewer #3's suggestions and have addressed them in the updated manuscript. 
Since Dr. Marsh indicated that the manuscript should not need to go out for further review we have not provided detailed responses to each of Reviewer #3...

Thanks for comprehensively addressing my comments from the last review. Below I just have some minor editorial suggestions. I hope you can find the time to do further studies studies exploring the many interesting ideas raised in the supplemental material in more depth.

Comment: It would be helpful context to have the 20% MPA coverage for Channel Islands in the Abstract (if you can squeeze it in).

Response: included

Comment: 15: Can you be more specific with the implications? e.g. something from your final paragraph about how MPA performance should be communicated to communities?

Response: Included

Comment: 52: The aims of the study hinge on this sentence. Can you make the key problem statement standout more? (I recognize though this is a style choice). e.g. "What we do not know is whether MPAs have population level effects". Then move first half of the sentence up to prior paragraph.

Response: Modified to be clearer.

76: I always think doing science well is more important than being first. So you could reword to remove the priority statement and make this sentence more straightforward. The following sentences are sufficient to clearly distinguish this study from previous ones.

Response: Priority statement removed

168, 179: Thanks for adding these.

Comment: 
180: Is D observed or predicted biomass? I thought it was the mean predicted by the model? If it is observed, should it be lower case d, (or is D the sum of all the di?)
1

80 and 186-188: You could combine these variable definitions to just explain once and save words. e.g. at line 180 change to "MPA marks whether observation i is in a pre MPA (MPA =0) or ...." Then remove same text below.

Response: Condensed

Comment: 
198: Are Gelman and McElreath refs redundant? you could move one to save space if needed?
Response: Removed.

206: Really key paragraph, marine scientists need to do more simulation studies like this to test the robustness of their empirical conclusions!

233: This is a pretty significant finding.

Comment: 
247: "or in the hypothetical world..." check grammar, this sentence doesn't make sense. You could delete 'or in the hypothetical world...' and it would make more sense.

Response: Very true! Reworded for grammar and clarity.

Comment: 
253: Delete these words: "Despite having its own strict caveats".
There are quite a few sentences like 253 that start with a qualification, caveat or exception. E.g. like:
 "While these studies demonstrated XXX, they do not necessarily do YYY". I think the writing could be tightened by editing these sentences to get straight to the point and also by removing qualifiers and adverbs, e.g. "Previous studies have not done YYYY" or "We do not know if YYYY".

 Reply: Corrected, and edits done throughout to address this

### 2021-03-07 documents/ovando-et-al-pop-effects-of-mpas.Rmd
pushing for final review

@@
 We examine the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and show that even in cases where the population benefits are large, they are inherently difficult to detect empirically.
-A network of MPAs was put in place in The Channel Islands National Marine Sanctuary in 2003, with a goal of providing regional conservation and fishery benefits.
+A network of MPAs covering roughly 20% of the Channel Islands National Marine Sanctuary was put in place in 2003, with a goal of providing regional conservation and fishery benefits.
 Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level using a Bayesian difference-in-difference approach.
 We show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); smaller effect sizes (which we find are common) are even harder to detect.
+Clearly communicating expectations and uncertainties around MPAs is critical to ensuring that MPAs play effective roles marine conservation.
 Our results provide a novel assessment of the population-level effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.
@@
 ```{r, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 300,
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
                       cache = FALSE, fig.width = 3.42, fig.asp = 1.5)
@@
 No-take Marine Protected Areas (MPAs), spatial regions of the ocean in which fishing is prohibited, have a long history in the management of marine resources [@johannes1978].
-Modern MPAs were first established as marine analogs to the terrestrial protection of iconic landscapes [@iucn1976].
+Modern MPAs were first established largely as marine analogs to the terrestrial protection of iconic landscapes [@iucn1976].
 Recent international efforts to expand MPAs, such as The International Union for Conservation of Nature's 30% by 2030 MPA targets are based in part on the assumption that well-designed MPAs will not only provide conservation benefits inside their borders, but also have broader conservation effects on unprotected areas surrounding the MPAs, whether MPAs are designed explicitly for conservation, fishery benefits, or both [@gaines2010].
@@
 Therefore, MPAs contribute to both local and regional population-level effects.
-Numerous factors influence how MPAs affect fish populations.
+Numerous factors influence the population-level effects of MPAs.
 These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003], the strength, timing, and location of density dependence [@burgess2014], the design of the network [@gaines2010;@rassweiler2014],the degree of enforcement [@edgar2014], the level of fishing pressure, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@gerber2003;@hilborn2004a;@gaines2010;@botsford2003;@smith2003a;@walters2000a;@white2011;@ovando2016;@moffitt2013;@jaco2020].
@@
 Where response ratios are available before and after MPA implementation, spatial before-after-control-impact (BACI) style studies have shown similarly clear and positive results [@thiault2019a].
+These studies demonstrate the ability of MPAs to create differences between local fished and unfished areas,
 
-While these studies have clearly demonstrated the ability of MPAs to create differences between local fished and unfished areas, this body of empirical work does not necessarily serve as sufficient evidence for the population-level effects of MPAs.
-These spatial studies rely on assuming that selected control sites serve as a measure of what would have happened in the absence of MPAs.
-Control sites used in calculating response ratios are often selected based on habitat characteristics to justify their use as counterfactuals [@ferraro2018].
-However, beyond habitat differences, the very spillover effects MPAs are increasingly hoped to produce can negate the ability of spatial response ratio or BACI designs to accurately estimate the effects of MPAs.
-Export of adults or larvae from MPAs to "control" sites can affect their status as controls, as does displacement of fishing effort from MPAs to control sites.
-These factors can bias estimators of population-level effects like response ratios that require that the control sites are unaffected by the treatment [@ferraro2018;@moffitt2013].
+What we do not have is clear evidence for the population-level effects of MPAs.
+Spatial "inside-vs-outside" studies rely on assuming that selected control sites serve as a measure of what would have happened in the absence of MPAs.
+Habitat characteristics are often used to justify the selection of particular fished sites as as counterfactuals (controls) in response ratios [@ferraro2018].
+However, beyond habitat differences, the very spillover effects MPAs are often hoped to produce can negate the ability of spatial response ratio or BACI designs to accurately estimate the effects of MPAs, since these methods require that the control sites are conditionally unaffected by the treatment [@ferraro2018;@moffitt2013].
+Spillover of adults or larvae from MPAs to control sites can mask conservation benefits, while displacement of fishing effort from MPAs to control sites (which is rarely addressed directly, @ferraro2018) can lead to overestimates of conservation gains caused by MPAs measured by spatial response ratios.
 Control sites sufficiently far from MPAs to negate both spillover of fish or larvae and concentration of the fishing fleet could be selected, but finding suitably distant sites that are also appropriate proxies for the ecological and economic context of the MPAs is challenging.
-As variations of spatial response ratios and BACI studies have been a primary source of evidence for the conservation effects of MPAs, this means that our empirical understanding of the population-level impacts of MPAs is surprisingly limited.
+Since variations of spatial response ratios and BACI studies have been a primary source of evidence for the conservation effects of MPAs, this means that our empirical understanding of the population-level impacts of MPAs is surprisingly limited.
 
@@
 In this paper, we present a paired theoretical and empirical assessment demonstrating the challenges in assessing the population-level impacts of MPAs.
-In 2003, a network of MPAs was established in the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) covering approximately 20% of the Islands' waters.
+In 2003, a network of MPAs was established in the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) covering approximately 20% of the Islands' waters (an area of over 800 KM^2^).
 The network has been used as a model case study in protected area design around the world [@botsford2014].
-We use data from the first 14 years of protection in a difference-in-difference (DiD) model [@angrist2009] to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of fin-fish species.
-Rather than relying on spatial controls as is typically done in MPA studies, we use groups of species targeted and not-targeted by fishing pressure as our treatment and control groups.
+We use data from the first 14 years of protection in a difference-in-difference (DiD) model [@angrist2009] to assess the population-level effect of a large  MPA network on a wide array of fin-fish species.
+Rather than relying on spatial controls, we use groups of species targeted and not-targeted by fishing pressure as our treatment and control groups.
 In contrast to clear differences in biomass densities observed inside and outside of well-protected MPAs both globally [@lester2009] and in the Channel Islands [@caselle2015] we are unable to detect a clear population effect from the Channel Islands MPAs.
@@
 
-Where *T* indicates targeted species, *MPA* indicates whether MPAs are in place, and *D* is the observed mean total biomass density across all sites.
+Where *T* stands for targeted ($T = 1$) or non-targeted ($T = 0$) by fishing, *MPA* marks whether the data are from the pre-MPA ($MPA = 0$) or post-MPA ($MPA = 1$) period, and *D* is the observed mean total biomass density across all observations of the appropriate group.
 
@@
 
-The simplified form of this regression model is
+The expanded form of the DiD regression is
 
@@
 
-where $d_i$ is the biomass density at observation *i*, *T* indicates whether the observation *i* is for a targeted ($T = 1$) or non-targeted ($T = 0$) species, and *MPA* marks whether observation *i* is in a pre MPA ($MPA = 0$) or post MPA ($MPA = 1$) state.
+where $d_i$ is the biomass density at observation *i*.
 To account for the fact that MPA effects will evolve over time, we estimate a vector of MPA effects in three-year blocks for the all years after the MPAs were implemented in 2003.
@@
 **B^s^** is a vector of hierarchical coefficients for each sampling location **S** , clustered by island $\beta_r$ with variance $\sigma_r^2$.
-Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment (*MPA*) on the treated (targeted species) (Table.\@ref(tab:coef-table)).
-We use a Bayesian hierarchical generalized linear model since it allows us to interpret our estimated effects probabilistically [@mcelreath2020].
+Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment ($MPA=1$) on the treated targeted group ($T=1$) (Table.\@ref(tab:coef-table)).
+We use a Bayesian hierarchical generalized linear model since it allows us to interpret our estimated effects probabilistically.
 Being a Bayesian regression, our DiD analysis produces posterior probability distributions (the probability distribution of our coefficients conditional on the data, priors, and model assumptions) of our coefficients, from which we can construct Bayesian credible intervals [@gelman2014a].
@@
 Although recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, most MPA networks for near-shore commercial fin-fish are likely to cover areas more in line with the Channel Islands (20%) or smaller.
-As such, many MPA networks are expected to have population-level effect sizes that are difficult to detect unless target species were experiencing extreme overfishing prior to MPA establishment, or in the hypothetical world without MPAs (Fig.\@ref(fig:sim-effects)-A).
+As such, many MPA networks are expected to have population-level effect sizes that are difficult to detect unless the target species would have been extremely overfished without the protection of MPAs (Fig.\@ref(fig:sim-effects)-A).
 
@@
 The targeted vs.
-non-targeted DiD strategy used here presents an alternative means of estimating the population-level effects of MPAs.
-Despite having its own strict caveats, this strategy provides some potential improvements over spatial response-ratios as a means of estimating population-level effects.
-While we estimate an uncertain but overall positive effect of the MPA network in its first few years, we are unable to detect a robust signal from 2012-2017.
+non-targeted DiD strategy used here presents an alternative means of estimating the population-level effects of MPAs, that does not rely on the assumption that MPAs do not affect control sites required of spatial response-ratios.
+While we estimate an uncertain but overall positive effect of the MPA network in its first few years of existence, we are unable to detect a robust signal from 2012-2017.
 We show that under the dynamics of the Channel Islands, particularly given the lack of heavily exploited species such as abalone and deep-water rockfish that helped motivate the Channel Island MPAs in the available data, this result is to be expected.
@@
 This potential negative interaction between constant-catch and MPAs is an important risk to consider (as done in @little2011b) especially as MPAs are increasingly implemented in quota-managed fisheries [@liu2018].
-While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (Appendix S1 Fig.S2) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
+We do not have access to fine scale fishing data from the Channel Islands alone, but reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (Appendix S1 Fig.S2) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
 
@@
 
-Another possibility for the estimated decline in Channel Islands MPA effects is environmental disturbance.
+Environmental disturbance is another possible explanation for the decline in the population-level effects of Channel Islands MPA estimated by our model.
 The Channel Islands region experienced a dramatic ‘marine heatwave’ beginning in 2014 and persisting through 2016, resulting in part in extremely elevated water temperatures throughout the region [@gentemann2017].
@@
 
-The scientific communicate must effectively communicate the challenges of estimating the population-level effects of MPAs.
+The scientific community must effectively communicate the challenges of estimating the population-level effects of MPAs.
 Lack of a clear population-level MPA effect should not necessarily be viewed as a failure of a conservation program, and likewise, large response ratios should not be automatically taken as evidence of a population-level conservation success.
@@
 
-As advocacy for broad networks of MPAs is growing globally, MPA science must directly tackle the challenge of evaluating the performance of MPAs at the population scale.
+As advocacy for large networks of MPAs grows around the world, MPA science must directly tackle the challenge of evaluating the performance of MPAs at the population scale.
 Commonly employed metrics such as spatial response ratios may be applicable in some circumstances, but are vulnerable to inaccuracy or misuse as metrics of population-level effects.
@@
              escape = FALSE,
-             caption = "Variables for MPA simulations",
+             caption = "Distributions of variables for MPA simulations",
              format = "pipe",
@@
     caption = glue::glue("Posterior means and credible interval for key model coefficients.
-Raw data contained {nrow(used_data)} individual observations, total number of targeted and non-targeted mean mean total biomass densities by site and year used to fit model was {nrow(did_fit$data)}.
-Site-level terms omitted for clarity."),
+Raw data contained {nrow(used_data)} individual observations.
+Total number of targeted and non-targeted mean total biomass densities by site and year used to fit model was {nrow(did_fit$data)}.
+Site-level terms omitted from table for visual clarity."),
     format = "pipe", 
@@
 
-```{r raw-density, fig.width = width_1col, fig.asp=asp_1col,fig.cap = "Centered and scaled trends in biomass densities of targeted and non-targeted fin-fish included in our study.
-Panel a) shows mean trends across all sites, with shaded areas shwoing 95% confidence intervals.
-Bottom two panels show same trends but only including sites b) outside and c) inside MPAs"}
+```{r raw-density, fig.width = width_1col, fig.asp=asp_1col,fig.cap = "Centered and scaled trends in biomass densities of targeted (solid darker lines) and non-targeted (dashed lighter lines) fin-fish included in our study of the Channel Islands MPA network.
+Panel (a) shows mean trends across all sites, with shaded areas shwoing 95% confidence intervals.
+Bottom two panels show same trends but only including sites (b) outside and (c) inside MPAs.
+Black vertical dotted line shows MPA implementation in 2003."}
 
 temp_plot <-
-  ((total_trend_plot + labs(subtitle = "a) Channel Islands") + theme_1col + theme(plot.subtitle = element_text(size = 8))) / (mpa_trend_plot + theme_1col + guides(color = FALSE, fill = FALSE, linetype = FALSE))) & theme(plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), units = "lines"),
+  ((total_trend_plot + labs(subtitle = "(a) Channel Islands") + theme_1col + theme(plot.subtitle = element_text(size = 8))) / (mpa_trend_plot + theme_1col + guides(color = FALSE, fill = FALSE, linetype = FALSE))) & theme(plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), units = "lines"),
                          panel.grid = element_blank(),
@@
 
-```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="Simulated population effects (panel *a* in grey) that could have produced observed response ratios (panel *b*) in black) in the Channel Islands.
-Panel *a*) histograms show 90% posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis).
-Panel b) Simulated population-level (pop.) effect on biomass densities matched to empirical response ratios in a).
+```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="Simulated population effects (panel (a) in grey) that could have produced observed response ratios (panel (b) in black) in the Channel Islands.
+Panel (a) histograms of 90% posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis).
+Panel (b) Histograms of simulated population-level (pop.) effect on biomass densities matched to empirical response ratios in Panel (a).
 For response ratios, a percent difference value of zero indicates that biomass densities of targeted species are identical inside and outside MPAs, a percent difference value of one that biomass densities of targeted species are 100% greater inside MPAs relative to outside.
@@
 ```{r sim-effects, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Simulated population-level effects of MPAs.
-Panel a) shows median simulated population-level (pop.) MPA effect sizes as a function of percent of species' range inside MPA (x-axis), and pre-MPA depletion (y-axis).
+Panel (a) shows median simulated population-level (pop.) MPA effect sizes as a function of percent of species' range inside MPA (x-axis), and pre-MPA depletion (y-axis).
 Pre-MPA depletion is a measure of fishing pressure, where 0 means that the population is unfished, and 1 means that the population is extinct in the time period immediately prior to MPA implementation.
-Panel b) shows distribution of simulations across range of MPA size and pre-MPA depletion separately."}
+Panel (b) shows distribution of simulations across range of MPA size and pre-MPA depletion separately."}
 

### 2021-03-07 make-pop-effects-of-mpas.R

@@ -3171,7 +3171,7 @@ pop_depletion_and_size_plot <- outcomes %>%
   summarise(median_mpa_effect = median(mpa_effect)) %>%
   ggplot(aes(depletion, mpa_size, fill = median_mpa_effect)) +
   geom_tile(alpha = 0.9) +
-  geom_vline(aes(xintercept = .6), linetype = 2) +
+  # geom_vline(aes(xintercept = .6), linetype = 2) +
   # geom_contour(aes(z = median_mpa_effect)) +
   scale_fill_binned(
     type = "viridis",
@@ -3247,7 +3247,7 @@ pop_depletion_plot <- outcomes %>%
                      expand = expansion(0, 0))
 
 expected_mpa_effect_plot <-
-  (pop_depletion_and_size_plot + labs(title = "a)")) + ((pop_size_plot + labs(title = "b)")) / pop_depletion_plot)  + plot_layout(widths = c(1.5, 1)) &
+  (pop_depletion_and_size_plot + labs(title = "(a)")) + ((pop_size_plot + labs(title = "(b)")) / pop_depletion_plot)  + plot_layout(widths = c(1.5, 1)) &
   theme(
     plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), units = "lines"),
     axis.text.x = element_text(size = 8),
@@ -4369,8 +4369,8 @@ total_trend_plot <- targeted_trends %>%
   theme(legend.position = "top")
 
 
-targlab <- c(`TRUE` = "c) Inside MPAs",
-             `FALSE` = 'b) Outside MPAs')
+targlab <- c(`TRUE` = "(c) Inside MPAs",
+             `FALSE` = '(b) Outside MPAs')
 
 
 mpa_trend_plot <- targeted_trends_by_mpa %>%
@@ -4491,8 +4491,8 @@ implications %>%
 # title = "<span style = 'color:grey;'> Paired Simulated Pop. Effect / <span style = 'color:black;'>Empirical Response Ratio</span>"
 
 lab <- c(
-  biased = "b) Paired Simulated Population-Level MPA Effect",
-  `Response Ratio` = "a) Empirical Response Ratio"
+  biased = "(b) Paired Simulated Population-Level MPA Effect",
+  `Response Ratio` = "(a) Empirical Response Ratio"
 )
 
 response_ratio_plot <-   targ_rr %>%

### 2021-03-08 documents/ovando-et-al-pop-effects-of-mpas.Rmd
small typo editing for final draft

@@
 Where response ratios are available before and after MPA implementation, spatial before-after-control-impact (BACI) style studies have shown similarly clear and positive results [@thiault2019a].
-These studies demonstrate the ability of MPAs to create differences between local fished and unfished areas,
+These studies demonstrate the ability of MPAs to create differences between local fished and unfished areas.
 
 What we do not have is clear evidence for the population-level effects of MPAs.
-Spatial "inside-vs-outside" studies rely on assuming that selected control sites serve as a measure of what would have happened in the absence of MPAs.
+Spatial "inside-vs-outside" studies rely on an assumption that selected control sites serve as a measure of what would have happened in the absence of MPAs.
 Habitat characteristics are often used to justify the selection of particular fished sites as as counterfactuals (controls) in response ratios [@ferraro2018].
@@
 In this paper, we present a paired theoretical and empirical assessment demonstrating the challenges in assessing the population-level impacts of MPAs.
-In 2003, a network of MPAs was established in the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) covering approximately 20% of the Islands' waters (an area of over 800 KM^2^).
+In 2003, a network of MPAs was established in the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) covering approximately 20% of the Islands' waters (an area of over 800 km^2^).
 The network has been used as a model case study in protected area design around the world [@botsford2014].
@@
 The purpose of the simulation model is to both set expectations for our empirical results, and to demonstrate the ways in which ecological and economic dynamics can interact to produce a wide range of population-level MPA effects.
-The full range of factors explored are shown in  Table.\@ref(tab:sim-table), and the equations of the simulation model can be viewed in the Appendix S1 Section 2.
+The full range of factors explored are shown in  Table \@ref(tab:sim-table), and the equations of the simulation model can be viewed in the Appendix S1 Section 2.
 We used this model to generate `r n_distinct(outcomes$experiment)` simulated MPA outcomes across `r n_distinct(outcomes$scientific_name)` species.
@@
 For each simulation we first randomly pull a species and its associated life history traits from the FishLife [@thorson2017c] package.
-We pair these data with randomly selected values governing the characteristics of the simulation (Table.\@ref(tab:sim-table)).
+We pair these data with randomly selected values governing the characteristics of the simulation (Table \@ref(tab:sim-table)).
 Key choices available to the model include parameters governing fishing pressure and MPA design.
@@
 Each simulation is paired with a simulation identical in every way except that MPAs are not implemented (i.e., a simulated control).
-We then calculate the effect of the MPAs on the population as the difference in biomass densities in the simulation with MPAs relative to biomass densities in the simulation without MPAs.
+Using these paired simulations, we calculate the effect of the MPAs on the population as the difference in biomass densities in the simulation with MPAs relative to biomass densities in the simulation without MPAs.
 
@@
 One set of simulations was specifically designed to reflect the dynamics of the subset of species available in the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring data from the Channel Islands used in this study.
-For this set we only include species of the same genus as those targeted by fishing in the PISCO data, restrict fishing pressure such that the simulated populations are moderately to lightly exploited (since the PISCO data used here excludes deeper-water species such as boccacio, *Sebastes paucispinis*, which were overexploited at the inception of the Channel Island MPAs, and threatened invertebrates such as red abalone *Haliotis rufescens*), and cap the MPA size at 20% of the population's range [@rassweiler2012].
-For each of these Channel Island style simulations, we calculated 1) the 'true' population-wide difference in biomass between the simulations with and without the MPAs and 2) the response ratio of biomass densities inside and outside the simulated MPAs.
+For this set we only include species of the same genus as those targeted by fishing in the PISCO data, restrict fishing pressure such that the simulated populations are moderately to lightly exploited (since the PISCO data used here excludes deeper-water species such as boccacio, *Sebastes paucispinis*, which were overexploited at the inception of the Channel Islands MPAs, and threatened invertebrates such as red abalone *Haliotis rufescens*), and cap the MPA size at 20% of the population's range [@rassweiler2012].
+For each of these Channel Islands-style simulations, we calculated 1) the 'true' population-wide difference in biomass between the simulations with and without the MPAs and 2) the response ratio of biomass densities inside and outside the simulated MPAs.
 We then calculated the response ratios observed in the PISCO survey data from the Channel Islands, and matched these empirical results with simulations that produced similar response ratios after the same number of years of MPA protection.
@@
 Building off of @caselle2015, we used an identification strategy utilizing biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 0])` species that are not directly targeted by fishing as our control group (non-targeted), and biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 1])` species targeted by fishing as our treatment group (Fig.\@ref(fig:raw-density)).
-Targeted  fin-fish species in the Channel Islands available to this study include California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish.
+Targeted  fin-fish species in the Channel Islands available to this study include California sheephead (*Semicossyphus pulcher*), and copper and blue rockfish (*Sebastes caurinus* and *Sebastes mystinus*).
 Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*).
@@
 **B^s^** is a vector of hierarchical coefficients for each sampling location **S** , clustered by island $\beta_r$ with variance $\sigma_r^2$.
-Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment ($MPA=1$) on the treated targeted group ($T=1$) (Table.\@ref(tab:coef-table)).
+Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment ($MPA=1$) on the treated targeted group ($T=1$) (Table\@ref(tab:coef-table)).
 We use a Bayesian hierarchical generalized linear model since it allows us to interpret our estimated effects probabilistically.
@@
 These empirical estimates are in line with the outcomes that our simulation model suggests are plausible.
-However, in the subsequent years the trend reverses itself, and in 2015-2017 we once again see no clear effect of the MPAs (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`,
+However, in the subsequent years the trend reverses, and in 2015-2017 we once again see no clear effect of the MPAs (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`,
 90\% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`
@@
 
-```{r ci-map, include = TRUE,fig.width=width_1col, fig.asp=1.5, fig.cap="Map of study region; the Northern Channel Islands, California, USA.
+```{r ci-map, include = TRUE,fig.width=width_1col, fig.asp=1.5, fig.cap="Map of the study region in the Northern Channel Islands, California, USA.
 Shading shows binned number of PISCO sampling events across the time period of our study.", include=TRUE}
@@
 ```{r raw-density, fig.width = width_1col, fig.asp=asp_1col,fig.cap = "Centered and scaled trends in biomass densities of targeted (solid darker lines) and non-targeted (dashed lighter lines) fin-fish included in our study of the Channel Islands MPA network.
-Panel (a) shows mean trends across all sites, with shaded areas shwoing 95% confidence intervals.
+Panel (a) shows mean trends across all sites, with shaded areas showing 95% confidence intervals.
 Bottom two panels show same trends but only including sites (b) outside and (c) inside MPAs.
@@
 
-```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="Simulated population effects (panel (a) in grey) that could have produced observed response ratios (panel (b) in black) in the Channel Islands.
+```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="Simulated population effects (panel (b) in grey) that could have produced empirically-observed response ratios (panel (a) in black) in the Channel Islands.
 Panel (a) histograms of 90% posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis).
-Panel (b) Histograms of simulated population-level (pop.) effect on biomass densities matched to empirical response ratios in Panel (a).
+Panel (b) Histograms of simulated population-level effects on biomass densities matched to empirical response ratios in Panel (a).
 For response ratios, a percent difference value of zero indicates that biomass densities of targeted species are identical inside and outside MPAs, a percent difference value of one that biomass densities of targeted species are 100% greater inside MPAs relative to outside.

### 2021-03-08 documents/ovando-et-al-pop-effects-of-mpas.Rmd
addressing Owen's minor edits and figure improvements

@@
 
-```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="Simulated population effects (panel (b) in grey) that could have produced empirically-observed response ratios (panel (a) in black) in the Channel Islands.
-Panel (a) histograms of 90% posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis).
-Panel (b) Histograms of simulated population-level effects on biomass densities matched to empirical response ratios in Panel (a).
-For response ratios, a percent difference value of zero indicates that biomass densities of targeted species are identical inside and outside MPAs, a percent difference value of one that biomass densities of targeted species are 100% greater inside MPAs relative to outside.
-For MPA population effect, a percent difference value of zero indicates that biomass densities are identical in the with- and without- MPA scenarios.
-A percent difference value of one indicates that biomass densities are 100% greater in the scenario with MPAs than the scenario without MPAs."}
+```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,fig.cap=" Empirically-observed response ratios (panel (a) in black) in the Channel Islands and simulated population effects (panel (b) in grey) that could have produced observed response ratios in panel (a).
+Panel (a) shows histograms of 90% posterior probability distributions of response ratios for targeted species (y-axis) over time (x-axis).
+Panel (b) shows histograms of simulated population-level effects on biomass densities matched to empirical response ratios in Panel (a).
+For response ratios, a percent difference value of 0% indicates that biomass densities of targeted species are identical inside and outside MPAs, a percent difference value of 100% that biomass densities of targeted species are 100% greater inside MPAs relative to outside.
+For MPA population effect, a percent difference value of 0% indicates that biomass densities are identical in the with- and without- MPA scenarios.
+A percent difference value of 100% indicates that biomass densities are 100% greater in the scenario with MPAs than the scenario without MPAs."}
 
@@
 ```{r sim-effects, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Simulated population-level effects of MPAs.
-Panel (a) shows median simulated population-level (pop.) MPA effect sizes as a function of percent of species' range inside MPA (x-axis), and pre-MPA depletion (y-axis).
-Pre-MPA depletion is a measure of fishing pressure, where 0 means that the population is unfished, and 1 means that the population is extinct in the time period immediately prior to MPA implementation.
+Panel (a) shows median simulated population-level (pop.) MPA effect sizes as a function of percent of species' range inside MPA (y-axis), and pre-MPA depletion (x-axis).
+Pre-MPA depletion is a measure of fishing pressure, where 0% means that the population is unfished, and 100% means that the population is extinct in the time period immediately prior to MPA implementation.
 Panel (b) shows distribution of simulations across range of MPA size and pre-MPA depletion separately."}

### 2021-03-08 make-pop-effects-of-mpas.R

@@ -3196,7 +3196,7 @@ pop_depletion_and_size_plot <- outcomes %>%
   scale_y_continuous(labels = percent, expand = expansion(0, 0)) +
   scale_x_continuous(labels = percent, expand = expansion(0, 0)) +
   theme(legend.position = "top",
-        legend.text = element_text(size = 8))
+        legend.text = element_text(size = 7))
 
 
 
@@ -3224,8 +3224,10 @@ pop_size_plot <- outcomes %>%
                      expand = expansion(0, 0)) +
   scale_y_continuous(labels = percent,
                      name = "Pop. Effect",
-                     expand = expansion(0, 0)) +
-  theme(legend.position = "top")
+                     n.breaks = 6) +
+  theme(legend.position = "top",
+        axis.text.x = element_text(size = 6),
+        axis.text.y = element_text(size = 6))
 
 pop_depletion_plot <- outcomes %>%
   filter(years_protected == short_frame) %>%
@@ -3244,17 +3246,17 @@ pop_depletion_plot <- outcomes %>%
                      expand = expansion(0, 0)) +
   scale_y_continuous(labels = percent,
                      name = "",
-                     expand = expansion(0, 0))
+                     n.breaks =6) +
+  theme(axis.text.x = element_text(size = 6),
+        axis.text.y = element_text(size = 6))
 
 expected_mpa_effect_plot <-
   (pop_depletion_and_size_plot + labs(title = "(a)")) + ((pop_size_plot + labs(title = "(b)")) / pop_depletion_plot)  + plot_layout(widths = c(1.5, 1)) &
   theme(
-    plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), units = "lines"),
-    axis.text.x = element_text(size = 8),
+    plot.margin = unit(c(0.2, 0.4, 0.2, 0.4), units = "lines"),
     legend.box.margin = unit(c(0, 0, 0, 0), units = "lines"),
-    axis.text.y = element_text(size = 10)
+    axis.title = element_text(size = 12)
   )
-
[…]

### 2021-03-14 documents/ovando-et-al-pop-effects-of-mpas.Rmd
Minor revisions (#6)
* addressing Owen's minor edits and figure improvements
* merging back to main

Co-authored-by: Owen R. Liu <owenrliu@gmail.com>

@@
   Marine Protected Areas (MPAs) cover 3-7% of the world’s ocean, and international organizations call for 30% by 2030.
-While numerous studies show that MPAs produce conservation benefits *inside* their borders, many MPAs are also justified on the grounds that they confer conservation benefits to the broader population *beyond* their borders.
+While numerous studies show that MPAs produce conservation benefits *inside* their borders, many MPAs are also justified on the grounds that they confer conservation benefits to the broader populations that span *beyond* their borders.
 We examine the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and show that even in cases where the population benefits are large, they are inherently difficult to detect empirically.
@@
 theme_2col <- hrbrthemes::theme_ipsum(base_size = 10,
-                                        axis_title_size = 12,
+                                        axis_title_size = 14,
                                       plot_title_size = 12)
@@
 We use a Bayesian hierarchical generalized linear model since it allows us to interpret our estimated effects probabilistically.
-Being a Bayesian regression, our DiD analysis produces posterior probability distributions (the probability distribution of our coefficients conditional on the data, priors, and model assumptions) of our coefficients, from which we can construct Bayesian credible intervals [@gelman2014a].
+Being a Bayesian regression, our DiD analysis produces posterior probability distributions (the probability distribution of our coefficients conditional on the data, priors, and model assumptions) of our coefficients, from which we can construct Bayesian credible intervals (also termed "uncertainty" or "compatibility" intervals) [@gelman2014a;@mcelreath2020];.
 
@@
 Pre-MPA depletion is a measure of fishing pressure, where 0% means that the population is unfished, and 100% means that the population is extinct in the time period immediately prior to MPA implementation.
-Panel (b) shows distribution of simulations across range of MPA size and pre-MPA depletion separately."}
+Panel (b) shows distribution of simulations across range of MPA sizes and pre-MPA depletions separately."}
 

### 2021-03-14 documents/pop-effects-of-mpas-front-matter.Rmd
Minor revisions (#6)
* addressing Owen's minor edits and figure improvements
* merging back to main

Co-authored-by: Owen R. Liu <owenrliu@gmail.com>

@@
 
-The population-level conservation effects of Marine Protected Areas are likely to be extremely challenging to detect
+The population-level conservation effects of Marine Protected Areas are likely to be extremely challenging to measure and detect
 
@@
 
+Channel Islands National Marine Sanctuary
+
 ## Word Count {.unnumbered}
 
-5976
+5994
 
@@
 
-Funding for the simulation model was provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship.
+Funding for the simulation model was provided by the National Marine Fisheries Service-SeaGrant Population and Ecosystem Dynamics Fellowship.
 Empirical data collection was funded primarily by the David and Lucille Packard Foundation in support of the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) with additional funding from the California Ocean Protection Council and California SeaGrant.
-This study would not be possible without the work provided by PISCO divers over the years, with special thanks to to K.
+This study would not have been possible without the work provided by PISCO divers over the years, with special thanks to K.
 Davis Koehn and A.
@@
 Lawson, A.
-Punt, Chris Brown, and two anonymous reviewers for helpful comments and technical support.
+Punt, C.
+Brown, and two anonymous reviewers for helpful comments and technical support.

### 2021-03-14 make-pop-effects-of-mpas.R

@@ -3171,7 +3171,7 @@ pop_depletion_and_size_plot <- outcomes %>%
   summarise(median_mpa_effect = median(mpa_effect)) %>%
   ggplot(aes(depletion, mpa_size, fill = median_mpa_effect)) +
   geom_tile(alpha = 0.9) +
-  geom_vline(aes(xintercept = .6), linetype = 2) +
+  # geom_vline(aes(xintercept = .6), linetype = 2) +
   # geom_contour(aes(z = median_mpa_effect)) +
   scale_fill_binned(
     type = "viridis",
@@ -3196,7 +3196,9 @@ pop_depletion_and_size_plot <- outcomes %>%
   scale_y_continuous(labels = percent, expand = expansion(0, 0)) +
   scale_x_continuous(labels = percent, expand = expansion(0, 0)) +
   theme(legend.position = "top",
-        legend.text = element_text(size = 8))
+        legend.text = element_text(size = 7),
+        axis.text.x = element_text(size = 12),
+        axis.text.y = element_text(size = 12))
 
 
 
@@ -3224,8 +3226,10 @@ pop_size_plot <- outcomes %>%
                      expand = expansion(0, 0)) +
   scale_y_continuous(labels = percent,
                      name = "Pop. Effect",
-                     expand = expansion(0, 0)) +
-  theme(legend.position = "top")
+                     n.breaks = 6) +
+  theme(legend.position = "top",
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 7))
 
 pop_depletion_plot <- outcomes %>%
   filter(years_protected == short_frame) %>%
@@ -3244,17 +3248,18 @@ pop_depletion_plot <- outcomes %>%
                      expand = expansion(0, 0)) +
   scale_y_continuous(labels = percent,
                      name = "",
-                     expand = expansion(0, 0))
+                     n.breaks =6) +
+  theme(axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 7))
 
 expected_mpa_effect_plot <-
-  (pop_depletion_and_size_plot + labs(title = "a)")) + ((pop_size_plot + labs(title = "b)")) / pop_depletion_plot)  + plot_layout(widths = c(1.5, 1)) &
[…]

### 2021-05-02 documents/ovando-et-al-pop-effects-of-mpas-si.Rmd
pre-print cleanup

@@
 ---
-title: "Appendix Supporting Information 1"
+title: "Supporting Information"
 subtitle: "Assessing the Population-level Conservation Effects of Marine Protected Areas"
+author: 
+  - Daniel Ovando
+  - Jennifer E. Caselle
+  - Christopher Costello
+  - Olivier Deschenes
+  - Steven D. Gaines
+  - Ray Hilborn
+  - Owen Liu
 output: 
@@
 always_allow_html: true
+toc: false
 ---
@@
 
+\renewcommand{\thesection}{Appendix S\arabic{section}}
+
 \renewcommand{\thefigure}{S\arabic{figure}}
@@
 
-All code needed to reproduce our main results and manuscript can be found at https://github.com/DanOvando/population-effects-of-mpas.
-All analysis were performed in `r R.version.string`.
-Package versions are shown in Table.\@ref(tab:pkg-info).
+All data and code needed to fully reproduce replicate our study this manuscript are publicly available at from github.com/DanOvando/population-effects-of-mpas
+
+All analysis were performed in `r R.version.string`. Package versions are shown in Table.\@ref(tab:pkg-info).
 
@@
 
+```{r sim-table, echo=FALSE}
+variable <- c(
+"Scientific Name",
+"steepness",
+"Adult movement ($\\sigma_{s=a}$)",
+"Larval movement ($\\sigma_{s=l}$)",
+"Recruitment variation ($\\sigma_{r}$)",
+"Recruitment autocorrelation ($ac_{r}$)",
+"Density-dependent adult movement",
+"Density-dependence timing ",
+"% Patches in MPA",
+"Initial fishing relative to natural mortality",
+"Selectivity as a multiple of maturity length",
+"Fleet model",
+"Spatial effort model",
+"Years into simulation to start MPA",
+"MPA is Larval Source?",
+"Randomly place MPA?",
+"Fleet reaction to MPA",
+"Patchiness",
+"MPA habitat factor")
+
+distribution <- 
+c(
+"Drawn from all possible species in `FishLife` (@thorson2017c)",
+"uniform(0.6,0.95)",
+"uniform(0,0.25 * P)",
+"uniform(0,0.25 * P)",
+ "$\\in{0,0.05,.1,.2}$",
+ "$\\in{0,0.05,.1,.2}$",
+ "$\\in{0.25,1}$",
+ "$\\in{Local, Global,PostDispersal}$",
+ "uniform(0.01,1)",
+ "uniform(0.01,4)",
+ "uniform(0.1,1.25)",
+"$\\in{OpenAccess, ConstantEffort,ConstantCatch}$",
+ "$\\in{Uniform, Biomass,Profits}$",
+ "uniform(5,0.66)",
+"$\\in{TRUE,FALSE}$",
+"$\\in{TRUE,FALSE}$",
+"$\\in{Concentrate,Leave}$",
+"uniform(0.01,0.75)",
+"$\\in{1, 4}$"
+)
+
+equations <- tibble(Variable = variable, Distribution = distribution )
+
+knitr::kable(equations,
+             escape = FALSE,
+             caption = "Distributions of variables used in MPA simulations.
+Simulations combine random combinations of the variables included here, and using those sampled states simulate a scenario with and a scenario without MPAs.
+The difference between these two scenarios is then used to calcualte the effect of the MPA in that simulated scenario.",
+             format = "pipe",
+             align = "l") %>% 
+  kableExtra::kable_styling()
+
+```
+
 <!--   1. Local density dependence: Density dependence occurs independently in each patch, and recruits then disperse to nearby patches -->
@@
 
-## Recruitment
+## Recruitment {-}
 
@@
 
-## Dispersal
+## Dispersal {-}
 
@@
 
-## Fleet Dynamics
+## Fleet Dynamics {-}
 
@@
 
-## Spatial Fleet Distribution
+## Spatial Fleet Distribution {-}
 
@@
 
-## MPA Design
+## MPA Design {-}
 
@@
 
-## Simulations
+## Simulations{-}
 
@@
 
-### Filtering Simulations
+### Filtering Simulations {-}
 
@@
 
-### Additional Simulation Results
+### Additional Simulation Results {-}
 
@@
 
-The full table of covariates and their posterior means and 89% credible interval can be found in Table.\@ref(tab:did-coefs).
+The full table of covariates and their posterior means and 89% credible interval can be found in Table.\@ref(tab:coef-table) and Table.\@ref(tab:did-coefs).
 "var" refers to variable, "_2" indicates a squared term.
@@
 Targeted is a Boolean marker indicating whether the given observation is for targeted or non-targeted species.
+
+```{r coef-table, echo=FALSE}
+
+coefs <- did_fit %>%
+  broom.mixed::tidy(conf.int = TRUE, prob = 0.9) %>%
+    rename(lower = conf.low,
+         upper = conf.high) %>% 
+  select(term, estimate, lower, upper) %>%
+  rename(
+    Mean = estimate,
+    `5th Percentile` = lower,
+    `95th Percentile` = upper
+  )
+
+coefs$term <- str_remove_all(coefs$term, "var_")
+
+coefs$term <- str_replace_all(coefs$term, "targeted","Targeted by Fishing")
+
+coefs$term <- str_replace_all(coefs$term, "year_bins","Year Block ")
+
+coefs$term <- str_replace_all(coefs$term, ":","$\\\\times$")
+
+coefs <- coefs %>% 
+  mutate(term = case_when(term == "tex" ~ "Observer Experience",
+                          term == "tex_2" ~ "(Observer Experience)$^2$",
+                          term == "temp" ~ "Water Temperature",
+                          term == "regional_temp_dev" ~ "Deviation from regional mean temperature",
+                          term == "regional_temp_dev_2" ~ "(Deviation from regional mean temperature)$^2$",
+                          term == "lag_catch" ~ "Lagged Commercial Catch",
+                          term == "targeted" ~ "Targeted by Fishing",
+                          term == "(Intercept)" ~ "Intercept",
+                          term == "surge" ~ "Wave Surge",
+                          term == "kelp" ~ "Kelp Cover",
+                           TRUE ~ term))  %>% 
+  rename(Variable = term)
+
+  knitr::kable(
+    coefs,
+    digits = 2,
+    caption = glue::glue("Posterior means and credible interval for key model coefficients.
+Raw data contained {nrow(used_data)} individual observations.
+Total number of targeted and non-targeted mean total biomass densities by site and year used to fit model was {nrow(did_fit$data)}.
+Site-level terms omitted from table for visual clarity."),
+    format = "pipe", 
+    escape = FALSE,
+    align = "l")
+
+```
 
@@
 
-## Synthetic controls
+## Synthetic controls {-}
 
@@
 
-# Testing Model Assumptions
-
-## Simulation testing
+# Simulation Testing
 
@@
 
-## Testing SUTVA with Convergent Cross Mapping
+# Testing SUTVA with Convergent Cross Mapping
 
@@
 
-# Literature Cited
+# Appendix S8 Literature Cited

### 2021-05-02 documents/ovando-et-al-pop-effects-of-mpas.Rmd
pre-print cleanup

@@
                                       plot_title_size = 10,
-                                      strip_text_size = 8)
+                                      strip_text_size = 8) + 
+    theme(axis.title.x = element_text(hjust = 0.5),
+        axis.title.y = element_text(hjust = 0.5))
 
@@
                                         axis_title_size = 14,
-                                      plot_title_size = 12)
+                                      plot_title_size = 12) + 
+    theme(axis.title.x = element_text(hjust = 0.5),
+        axis.title.y = element_text(hjust = 0.5))
 
@@
 All data and code needed to fully reproduce this manuscript are publicly available at [github.com/DanOvando/population-effects-of-mpas](https://github.com/DanOvando/population-effects-of-mpas).
-A detailed description of the simulation model structure, as well as sensitivity analyses of our estimation model, are available in the online Appendix Supporting Information 1.
+A detailed description of the simulation model structure, as well as sensitivity analyses of our estimation model, are available in the online Appendix S2.
 
@@
 The purpose of the simulation model is to both set expectations for our empirical results, and to demonstrate the ways in which ecological and economic dynamics can interact to produce a wide range of population-level MPA effects.
-The full range of factors explored are shown in  Table \@ref(tab:sim-table), and the equations of the simulation model can be viewed in the Appendix S1 Section 2.
+The full range of factors explored are shown in  Table \@ref(tab:sim-table), and the equations of the simulation model can be viewed in the Appendix S2.
 We used this model to generate `r n_distinct(outcomes$experiment)` simulated MPA outcomes across `r n_distinct(outcomes$scientific_name)` species.
@@
 This potential negative interaction between constant-catch and MPAs is an important risk to consider (as done in @little2011b) especially as MPAs are increasingly implemented in quota-managed fisheries [@liu2018].
-We do not have access to fine scale fishing data from the Channel Islands alone, but reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (Appendix S1 Fig.S2) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
+We do not have access to fine scale fishing data from the Channel Islands alone, but reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (Appendix S3) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely.
 
@@
 We used convergent cross mapping (CCM), in the manner of @clark2015, to test for significant dynamic interactions between species and therefore the possibility of the trophic cascades biasing our results.
-We found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (Appendix S1 Fig.S33-S35).
+We found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (Appendix S7).
 However, the longer MPAs are in place, the greater the possibility that substantial species interactions that can affect use of non-targeted species as a control may arise.
@@
 
-Detailed descriptions of the simulation model and DiD estimator (Appendix S1) are available online.
-The authors are solely responsible for the content and functionality of these materials.
+Detailed descriptions of the simulation model and DiD estimator are available in the online supporting information The authors are solely responsible for the content and functionality of these materials.
 Queries (other than absence of the material) should be directed to the corresponding author.
@@
 <div id="refs"></div>
-
-\newpage
-
-# Tables {#tables .unnumbered}
-
-```{r sim-table, echo=FALSE}
-variable <- c(
-"Scientific Name",
-"steepness",
-"Adult movement ($\\sigma_{s=a}$)",
-"Larval movement ($\\sigma_{s=l}$)",
-"Recruitment variation ($\\sigma_{r}$)",
-"Recruitment autocorrelation ($ac_{r}$)",
-"Density-dependent adult movement",
-"Density-dependence timing ",
-"% Patches in MPA",
-"Initial fishing relative to natural mortality",
-"Selectivity as a multiple of maturity length",
-"Fleet model",
-"Spatial effort model",
-"Years into simulation to start MPA",
-"MPA is Larval Source?",
-"Randomly place MPA?",
-"Fleet reaction to MPA",
-"Patchiness",
-"MPA habitat factor")
-
-distribution <- 
-c(
-"Drawn from all possible species in `FishLife` (@thorson2017c)",
-"uniform(0.6,0.95)",
-"uniform(0,0.25 * P)",
-"uniform(0,0.25 * P)",
- "$\\in{0,0.05,.1,.2}$",
- "$\\in{0,0.05,.1,.2}$",
- "$\\in{0.25,1}$",
- "$\\in{Local, Global,PostDispersal}$",
- "uniform(0.01,1)",
- "uniform(0.01,4)",
- "uniform(0.1,1.25)",
-"$\\in{OpenAccess, ConstantEffort,ConstantCatch}$",
- "$\\in{Uniform, Biomass,Profits}$",
- "uniform(5,0.66)",
-"$\\in{TRUE,FALSE}$",
-"$\\in{TRUE,FALSE}$",
-"$\\in{Concentrate,Leave}$",
-"uniform(0.01,0.75)",
-"$\\in{1, 4}$"
-)
-
-equations <- tibble(Variable = variable, Distribution = distribution )
-
-knitr::kable(equations,
-             escape = FALSE,
-             caption = "Distributions of variables for MPA simulations",
-             format = "pipe",
-             align = "l") %>% 
-  kableExtra::kable_styling()
-
-```
-
-\newpage
-
-```{r coef-table, echo=FALSE}
-
-coefs <- did_fit %>%
-  broom.mixed::tidy(conf.int = TRUE, prob = 0.9) %>%
-    rename(lower = conf.low,
-         upper = conf.high) %>% 
-  select(term, estimate, lower, upper) %>%
-  rename(
-    Mean = estimate,
-    `5th Percentile` = lower,
-    `95th Percentile` = upper
-  )
-
-coefs$term <- str_remove_all(coefs$term, "var_")
-
-coefs$term <- str_replace_all(coefs$term, "targeted","Targeted by Fishing")
-
-coefs$term <- str_replace_all(coefs$term, "year_bins","Year Block ")
-
-coefs$term <- str_replace_all(coefs$term, ":","$\\\\times$")
-
-coefs <- coefs %>% 
-  mutate(term = case_when(term == "tex" ~ "Observer Experience",
-                          term == "tex_2" ~ "(Observer Experience)$^2$",
-                          term == "temp" ~ "Water Temperature",
-                          term == "regional_temp_dev" ~ "Deviation from regional mean temperature",
-                          term == "regional_temp_dev_2" ~ "(Deviation from regional mean temperature)$^2$",
-                          term == "lag_catch" ~ "Lagged Commercial Catch",
-                          term == "targeted" ~ "Targeted by Fishing",
-                          term == "(Intercept)" ~ "Intercept",
-                          term == "surge" ~ "Wave Surge",
-                          term == "kelp" ~ "Kelp Cover",
-                           TRUE ~ term))  %>% 
-  rename(Variable = term)
-
-  knitr::kable(
-    coefs,
-    digits = 2,
-    caption = glue::glue("Posterior means and credible interval for key model coefficients.
-Raw data contained {nrow(used_data)} individual observations.
-Total number of targeted and non-targeted mean total biomass densities by site and year used to fit model was {nrow(did_fit$data)}.
-Site-level terms omitted from table for visual clarity."),
-    format = "pipe", 
-    escape = FALSE,
-    align = "l")
-
-```
 
@@
   theme_1col + 
-  theme(axis.title.x = element_textbox(size = 10),
+  theme(
         legend.position = "top",
@@
 ggsave(
-  mpa_effect_plot,
+  mpa_effect_plot + theme_2col + theme(legend.position = "top"),
   filename = here::here("documents","figs","mpa_effect_plot.pdf"),

### 2021-05-02 documents/pop-effects-of-mpas-front-matter.Rmd
pre-print cleanup

@@
 
+Daniel Ovando*
+School of Aquatic and Fishery Sciences
+University of Washington, Seattle, WA 98195-5020 USA
+danovan@uw.edu
+
 Jennifer E. Caselle
-UCSB Marine Science Institute Bldg 520 Rm 4002 Fl 4L 
-Santa Barbara, CA 93106-6150 USA
-caselle@msi.ucsb.edu
+Marine Science Institute, University of California Santa Barbara, Santa Barbara, CA, USA
 
 Christopher Costello
-Bren Hall 4410 
-University of California, Santa Barbara
-Santa Barbara, CA 93106-6150 USA
-costello@bren.ucsb.edu
+Environmental Market Solutions Lab, University of California, Santa Barbara, CA, USA
 
 Olivier Deschenes
-2127 North Hall
-University of California, Santa Barbara
-Santa Barbara, CA 93106-6150 USA
-olivier@econ.ucsb.edu
+Environmental Market Solutions Lab, University of California, Santa Barbara, CA, USA 
+Department of Economics, University of California, Santa Barbara, CA, USA
 
 Steven D. Gaines
-Bren Hall 2410 
-University of California, Santa Barbara
-Santa Barbara, CA 93106-6150 USA
-gaines@bren.ucsb.edu
+Bren School of Environmental Science and Management, University of California, Santa Barbara, CA 93117
 
 Ray Hilborn
-FSH 352B
-University of Washington
-1122 NE Boat Street
-Seattle, WA 98195-5020 USA
+School of Aquatic and Fishery Sciences
+University of Washington, Seattle, WA 98195-5020 USA
 
 Owen Liu
-Bren Hall 4330 
-University of California, Santa Barbara
-Santa Barbara, CA 93106-6150 United States
-owen.r.liu@gmail.com
+Bren School of Environmental Science and Management, University of California, Santa Barbara, CA 93117
 
@@
 Empirical data collection was funded primarily by the David and Lucille Packard Foundation in support of the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) with additional funding from the California Ocean Protection Council and California SeaGrant.
-This study would not have been possible without the work provided by PISCO divers over the years, with special thanks to K.
-Davis Koehn and A.
+This study would not have been possible without the work provided by PISCO divers over the years.
+We especially thank K.
+Davis Koehn, and A.
 Parsons-Field.
@@
 Punt, C.
-Brown, and two anonymous reviewers for helpful comments and technical support.
+Brown, and 2 anonymous reviewers for helpful comments and technical support.

### 2021-05-02 make-pop-effects-of-mpas.R

@@ -95,7 +95,9 @@ write(run_description,
       file = paste(run_dir, 'RUN_DESCRIPTION.txt', sep = '/'))
 
 plot_theme <- hrbrthemes::theme_ipsum(base_size = 12,
-                                      axis_title_size = 14)
+                                      axis_title_size = 14) + 
+  theme(axis.title.x = element_text(hjust = 0.5),
+        axis.title.y = element_text(hjust = 0.5))
 
 
 theme_set(plot_theme)
@@ -2916,8 +2918,8 @@ fishery_eqo <- fishery_outcomes %>%
   filter(year == max(year))
 
 
-facet_labels <- c(mpa_size = "Range in MPA",
-                  depletion = "Pre-MPA Depletion")
+facet_labels <- c(mpa_size = "Range in MPA (%)",
+                  depletion = "Pre-MPA Depletion (%)")
 
 
 
@@ -3169,15 +3171,14 @@ pop_depletion_and_size_plot <- outcomes %>%
   filter(years_protected == short_frame, depletion < 1) %>%
   group_by(depletion, mpa_size) %>%
   summarise(median_mpa_effect = median(mpa_effect)) %>%
-  ggplot(aes(depletion, mpa_size, fill = median_mpa_effect)) +
+  ggplot(aes(depletion * 100, mpa_size * 100, fill = median_mpa_effect * 100)) +
   geom_tile(alpha = 0.9) +
   # geom_vline(aes(xintercept = .6), linetype = 2) +
   # geom_contour(aes(z = median_mpa_effect)) +
   scale_fill_binned(
     type = "viridis",
-    labels = label_percent(accuracy = 1),
-    name = "Pop. Effect",
-    breaks = seq(0, 2, by = .25),
+    name = "Pop. Effect (%)",
+    breaks = seq(0, 200, by = 25),
     guide = guide_bins(
       keywidth = ggplot2::unit(1.25, "lines"),
       reverse = FALSE,
@@ -3185,16 +3186,9 @@ pop_depletion_and_size_plot <- outcomes %>%
       axis.linewidth = 1
     )
   ) +
-  # scale_fill_gradient2(low = "tomato", high = "steelblue", mid = "white", midpoint = .25,
-  #                      labels = percent,
-  #                      guide = hgc,
-  #                      name = "Median Effect") +
-  # scale_fill_viridis(labels = percent,
-  #                    guide = hgc,
-  #                    name = "Median Effect") +
-  labs(x = "Pre-MPA Depletion", y = "Range in MPA") +
[…]

### 2021-07-01 README.Rmd
adding DOI to readme

@@
 
-This repo contains code and data needed to reproduce "Assessing the Population-level Conservation Effects of Marine Protected Areas" by Ovando et al. In Prep.
+This respository contains code and data needed to reproduce "Assessing the Population-level Conservation Effects of Marine Protected Areas" by Ovando et al.
+2021.
+The manuscript in *Conservation Biology* can be accessed at https://conbio.onlinelibrary.wiley.com/doi/10.1111/cobi.13782
+
+## Abstract
+
+>>  Marine Protected Areas (MPAs) cover 3-7% of the world’s ocean, and international organizations call for 30% by 2030.
+While numerous studies show that MPAs produce conservation benefits *inside* their borders, many MPAs are also justified on the grounds that they confer conservation benefits to the broader population *beyond* their borders.
+We examine the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and show that even in cases where the population benefits are large, they are inherently difficult to detect empirically.
+A network of MPAs was put in place in The Channel Islands National Marine Sanctuary in 2003, with a goal of providing regional conservation and fishery benefits.
+Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level using a Bayesian difference-in-difference approach.
+We show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); smaller effect sizes (which we find are common) are even harder to detect.
+Our results provide a novel assessment of the population-level effects of a large and iconic Marine Protected Area
+
+## Reproducing Results
 
@@
 Packages are installed in a stand-alone project library for this paper, and will not affect your installed R packages anywhere else.
-
->>  Marine Protected Areas (MPAs) cover 3-7% of the world’s ocean, and international organizations call for 30% by 2030.
-While numerous studies show that MPAs produce conservation benefits *inside* their borders, many MPAs are also justified on the grounds that they confer conservation benefits to the broader population *beyond* their borders.
-We examine the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and show that even in cases where the population benefits are large, they are inherently difficult to detect empirically.
-A network of MPAs was put in place in The Channel Islands National Marine Sanctuary in 2003, with a goal of providing regional conservation and fishery benefits.
-Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level using a Bayesian difference-in-difference approach.
-We show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); smaller effect sizes (which we find are common) are even harder to detect.
-Our results provide a novel assessment of the population-level effects of a large and iconic Marine Protected Area
